### 2013-05-13 Writing/paper.tex
Uploading the rough disfunctional current code/data.

%Use this template for comments in the margin: \com{Note!}

%\date{December 2010}                           % Activate to display a given date or no date

%
%  From Riggle:
%
%----------------------------------------------------------------------------%
%                    Basic page layout parameters                            %
%----------------------------------------------------------------------------%

% \documentclass[11pt,leqno,tbtags]{article}
\documentclass[12pt,leqno,tbtags,twoside]{article}

    % leqno = equations at the left
    % tbtags (+leqno) = equation numbers at the top (rather than centered)

\usepackage[letterpaper,hmargin=1in,vmargin=1in]{geometry} 
\usepackage[bottom]{footmisc}
\usepackage{fancyhdr}

\fancyhead{} % clear all header fields

\fancyhead[LO]{Samuel R. Bowman}
\fancyhead[RE]{Logical Reasoning in Compositional Vector Space Semantics}

\fancyfoot{} % clear all footer fields
\fancyfoot[CE, CO]{\thepage}

\renewcommand{\headrulewidth}{0.4pt}

\renewcommand{\footrulewidth}{0.4pt}

\pagestyle{fancy}
\setlength{\headheight}{15pt}

\usepackage{setspace}
\spacing{1}

\usepackage{enumitem}
\setlist{noitemsep}

\usepackage{float}

\usepackage[round]{natbib}
\bibpunct[,~]{(}{)}{,}{a}{,}{,}
\setlength{\bibsep}{0pt}

%----------------------------------------------------------------------------%
%                    Title, author, and affiliation                          %
%----------------------------------------------------------------------------%

\title{\Large Logical Reasoning in Compositional Vector Space Semantics}
\author{Samuel R. Bowman\\\textsc{Ling 236}}
%\\\\Qualifying Paper \textit{for}\\The Department of Linguistics, Stanford University\\\\Committee:\\Paul Kiparsky (chair), Arto Anttila, and Christopher Manning
% \date{Edinburgh P-Workshop, May 2013}
% \author{Max Bane and Jason Riggle}
% \date{draft -- \today}

%----------------------------------------------------------------------------%
%                    Fonts and encoding                                      %
%----------------------------------------------------------------------------%

\usepackage[dirty,tipa]{ucs}	% Utf-8
\usepackage[utf8x]{inputenc}	% fonts
\usepackage[T1]{fontenc}		% in Tex

\usepackage{tipa}\let\ipa\textipa
\usepackage{vowel}

\usepackage{pifont}
\newcommand{\BlankCell}{}

%\usepackage{palatino}           % body font 
\renewcommand{\ttdefault}{lmtt} % tt set to latinmodern

%----------------------------------------------------------------------------%
%                    Graphics, tables, and figures                           %
%----------------------------------------------------------------------------%
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{multido}
\usepackage{ot-tableau}
\usepackage{pstricks}
\usepackage{pst-slpe}

\psset{gridcolor=green, subgridcolor=white, dimen=middle,
	   linewidth=1pt, subgriddiv=1, gridlabels=4pt, gridwidth=0.2pt}
\psset{cornersize=absolute,linearc=0.06}

\newpsstyle{final}{linewidth=0.5pt,fillstyle=solid,fillcolor=gray95,%
				   doubleline=true,doublesep=0.8pt,dimen=outer}

%----------------------------------------------------------------------------%
%                    Algorithms and Theorems                                 %
%----------------------------------------------------------------------------%

\usepackage[ruled,noend,linesnumbered]{algorithm2e}

\usepackage[leqno,fleqn]{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{wasysym}

\usepackage{stmaryrd}

\newtheoremstyle%
{dotless}%      name
{\topsep}%      Space above
{\topsep}%      Space below
{\normalfont}%  Body font
{}%             Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}%    Thm head font
{}%             Punctuation after thm head
{\newline}%     Space after head: " " = interword space, \newline = linebreak
{}%             Thm head spec (can be left empty, meaning `normal')
%               alternative spec: {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}

\theoremstyle{dotless}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}

\newtheorem{exe}[thm]{Example}

%----------------------------------------------------------------------------%
%                    Hyperlinks                                              %
%----------------------------------------------------------------------------%
\newrgbcolor{lnkClr}{0.0 0.0 0.33}

\usepackage{hyperref}			% Clickable links in pdfs
\hypersetup{colorlinks,			% nice for large docs and 
			breaklinks,			% for links to web pages
            linkcolor=lnkClr,
			urlcolor=lnkClr,
            anchorcolor=lnkClr,
			citecolor=lnkClr}

%----------------------------------------------------------------------------%
%  Example environment
%----------------------------------------------------------------------------%

\newenvironment{example}
{\refstepcounter{equation}\begin{list}{(\theequation)}{\leftmargin=7mm \labelwidth=5mm}\item}%
{\end{list}}

\def\citeapos#1{\citeauthor{#1}'s (\citeyear{#1})}

%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%
%./\..../\..../\..../\..../\..../\..../\..../\..../\..../\..../\..../\..../\.%
%/  \../  \../  \../  \../  \../  \../  \../  \../  \../  \../  \../  \../  \%
%    \/    \/    \/    \/    \/    \/    \/    \/    \/    \/    \/    \/    %
%%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%
\usepackage{gb4e}

\begin{document}
\maketitle
% \addtocounter{section}{-1}

\def\Y{\textbabygamma}
\def\P{\textceltpal}
\def\Rs{\textsubwedge} % restore this
\def\E{\tipaloweraccent[.1ex]{8}{e}} %restore this
\def\A{\sublptr{\"{a}}}

\maketitle

%TODO: See leichturm margins

%TODO: See Ken Shan slides

\section{Introduction}

Vector space models of semantic meaning have  substantially advanced the state of the art in applied natural language processing tasks like parsing, paraphrase detection, and sentiment analysis. These models have benefited in particular from two properties: The ability to learn representations of individual words that are \textit{optimized for the task at hand}, and their ability to compose those word representations into higher-order constituent representations \textit{in the same semantic vector space}. Together, these properties allow the model to capture nuanced compositional representations of sentence meaning without the need for any independent lexical semantic model.

There is not at this time any general understanding of what \textit{kind} of information these models are capable of learning and encoding in the vector space. It is clear that they are able to learn various useful facets of meaning well, but it is possible that many of the properties that are necessary for a full account of linguistic meaning like quantification and monotonicity are missed: while these models achieve state-of-the-art performance, the models that they supplant are largely models based on simple expert-designed features with substantial known limitations.

In this study, I propose to probe the representational capacities of these systems. In particular, I propose to build a model which learns task-specific compositional vector representations---in the style of Socher et al.---for the task of recognizing textual entailment in a regimented formal language. This model should shed light on the abilities of these schemes to capture reasoning patterns involving quantification, monotonicity, intersective and nonintersective modification, and other phenomena.

\section{The Pilot Experiment: Reasoning with idealized data}

If the pilot model works: Distributional system with X composition function gets quantification. Publish and push on how best to learn it in broader domain applications.

If it fails: Can we say that the system isn't powerful enough? Squib and return to even more supervised approaches.

\section{Background}

\citet{baroni2012entailment} present a real world version of the entailment reasoning problem, and have the current state of the art on that task. [Ken Shan is the last author on this.]

\begin{quote}
Interestingly, on the QN entailment task, neither our classifier trained on AN-N pairs nor the balAPinc method beat baseline methods. This suggests that our successful QN classifiers tap into vector properties beyond such relations as feature inclusion that those methods for nominal entailment rely upon.
\end{quote}

\begin{quote}
The FS definition of entailment has been modified
by taking common sense into account. Instead of
a relation from the truth of the consequent to the
truth of the antecedent in any circumstance, the
applied view looks at entailment in terms of plausibility \\\\... if a human who reads (and trusts) would most likely infer that is also true.
\end{quote}

\citet{erk2009representing} discusses notions of subsumption from distributional vectors. His results would be discouraging if we were trying to use pure distributional vectors rather than task-learned ones.

\begin{quote}
\textit{Baroni et al on Erk:} Erk (2009) suggests that it
may not be possible to induce lexical entailment
directly from a vector space representation, but it
is possible to encode the relation in this space after it has been derived through other means.\\\\
...run implies move in some contexts, but not all, for example not in the context computer runs.
\end{quote}

\citet{baroni2011we} present a corpus of word pairs manually annotated with relations like \textit{hypernym}, \textit{hyponym}, \textit{meronym}, \textit{associated event}, and \textit{common attribute}, as well as negative example relations.

\begin{quote}
\begin{verbatim}
missile-n	weapon	hyper	rocket-n
missile-n	weapon	hyper	vehicle-n
missile-n	weapon	hyper	weapon-n
missile-n	weapon	mero	engine-n
missile-n	weapon	mero	explosive-n
\end{verbatim}
\end{quote}

This may be useful as a fringe test case, but the lack of proper entailment/cover/equiv relationships may make it more difficult to test the logical capacity of the model by these means.

\citet{lenci2012identifying} have more of the same: distributional inclusion does capture some entailment information.

\begin{quote}
We introduce BLESS, a data set specifically
designed for the evaluation of distributional
semantic models. BLESS contains a set of tuples instantiating different, explicitly typed semantic relations, plus a number of controlled
random tuples. It is thus possible to assess the
ability of a model to detect truly related word
pairs, as well as to perform in-depth analyses of the types of semantic relations that a
model favors. We discuss the motivations for
BLESS, describe its construction and structure, and present examples of its usage in the
evaluation of distributional semantic models.
\end{quote}

In my notes for some reason: LDC's ACE08 Local Relation Detection and Recognition (LRDR) task

\citet{goller1996learning} present the core of the machine learning algorithm needed for the task-specific representations used in this project. I can't make heads or tails of their writing.

\citeapos{widdows2004geometry} book primarily serves as an primer on vector space semantics for those without sufficient mathematical background, but it does briefly address some aspects of the logical capacities of such systems. In particular, he presents a system for constructing information-retrieval targeted query vectors, and builds a practical quasi-logic around it which is capable of conjunction (\textit{rock} AND \textit{blues}) using vector addition, and subtraction (\textit{rock} NOT \textit{blues}) through the the operation of finding the orthogonal vector, expressed as:

\begin{equation}
a \text{~NOT~} b = a - (a \cdot b)b
\end{equation}

\citet{maccartney2009extended} \citep[from Bill's dissertation][]{maccartney2009natural} presents the framing of entailment and monotonicity used here, and shows that they can (with a lot of hacks) be used to do the Recognizing Textual Entailment task on real language. No vectors here.

\citet{icard2012inclusion} presents an extension to the NLI/MacCartney and Manning logic involving an elaborated notion of monotonicity. He uses the same seven basic entailment relations as MacCartney. The added detail relates to the model described in this paper as a formal description of what the model should ideally be able to learn, and not as a description of a formal property that should be externally introduced into the model.

\citet{socher2011semi} is the primary introduction to the idea of learning a set of quasi-distributional vectors and a composition function on those vectors, all guided to optimize performance on some task.

\citet{socher2011dynamic} extends this to paraphrase detection, an (easy, by our standards) task of comparing two learned sentence representations.

Socher et al. (in prep; ACL 2013) and \citet{chen2013learning} define a more powerful version of the composition function based on a third-order tensor. I'm using this layer as the top node that builds the comparison vector from which relation types are extracted.

Socher and co. have a few other recent and in-progress papers setting various task-performance records. Useful for motivation, but not that informative.

% Distributional Models of the Representation and Acquisition of Natural Language Categories, Lapata and Fountain: Likely worth ignoring

Basile et al: Encoding syntactic dependencies by vector permutation. Slides are a bit vague. Give it a closer look!

\bibliography{../../Bibs/MLSemantics}{}
\bibliographystyle{plainnat}

\end{document}

### 2013-05-13 Sigmoid.m

@@ -1,0 +1,5 @@
+function outVec = Sigmoid (inVec)
+
+outVec = tanh(inVec);
+
+end

### 2013-05-13 SigmoidDeriv.m

@@ -1,0 +1,5 @@
+function outVec = SigmoidDeriv (inVec)
+
+outVec = sech(inVec).^2;
+
+end

### 2013-05-13 Tree.m

@@ -1,0 +1,80 @@
+classdef Tree
+    
+    % Represents a single binary branching syntactic tree with three representations at each node:
+    % - The index with which the feature vector can be looked up - IF LEAF
+    % - The text with which the tree can be displayed.
+    % - The features at the node.
+    
+    properties (Hidden) %TODO: Make all private.
+        daughters = []; % 2 x 1 vector of trees
+        text = 'NULL';
+        features = []; % DIM x 1 vector
+        wordIndex = -1; % -1 => Not a lexical item node.
+    end
+    
+    methods(Static)
+        % TODO: Constructors for multiword trees?
+        
+        function t = makeLeaf(iText)
+            global wordMap;
+            global wordFeatures;
+            t = Tree();
+            t.text = iText;
+            t.wordIndex = wordMap(t.text);
+            t.features = wordFeatures(t.wordIndex, :);
+        end
+        
+    end
+    methods
+        
+        function resp = isLeaf(obj)
+            resp = (length(obj.daughters) == 0); % TODO: Fill in for undefined.
+        end
+        
+        function ld = getLeftDaughter(obj)
+            if (length(obj.daughters) > 0)
+                ld = obj.daughters(1);
+            else
+                ld = 0;
+            end
+        end
+        
+        function rd = getRightDaughter(obj)
+            if (length(obj.daughters) > 1)
+                rd = obj.daughters(2);
+            else
+                rd = 0;
+            end
+        end
+        
+        function t = getText(obj)
+            if isLeaf(obj)
+                t = obj.text;
+            else
+                t = [obj.getLeftDaughter().getText(), ' ', obj.getRightDaughter().getText()];
+            end
+        end
+        
+        function f = getFeatures(obj)
+            f = obj.features;
+        end
+        
+        function i = getWordIndex(obj)
+            i = obj.wordIndex;
+        end
+        
+        function updateFeatures(obj)
+            global wordFeatures;
+            
+            if (~isempty(obj.daughters))
+                for (daughterIndex = 1:length(obj.daughters))
+                    obj.daughters(daughterIndex).updateFeatures();
+                end
+                % TODO: APPLY COMPOSITION FUNCTION.
+            else
+                % In this case, we are a leaf.
+                features = wordFeatures(:, obj.wordIndex);
+            end
+        end
+    end
+end

### 2013-10-15 Writing/paper.tex
Update after a long absence… still not quite done

@@
 
+%TOOD: Mention Potts ministudy in 236 handout
+
+%TODO: Replicate Baroni et al 2012
+
 \section{Introduction}
@@
 
-\bibliography{../../Bibs/MLSemantics}{}
+\section{Methods: The pilot model}
+
+The initial model is built upon three functions: The composition function, which takes individual word representations and composes them into higher-order constituent representations, the comparison function, which takes pairs of word representations and compares them to yield a relation representation, and the softmax function, which turns this abstract relation representation into a probability distribution over the seven relation types.
+Each of these has numerous parameters which we learn from training data, and when composed appropriately, they form a relation classifier for pairs of sentences of arbitrary length.
+
+In this section, I lay out the definitions of these three functions, how they are applied, and how their parameters can be learned.
+
+\subsection{Forward computation: Identifying a single relation}
+
+The figure shows an example of the structure over which information flows in a forward (classifying) run of the model.
+The leaves of the tree are the vector representations of individual words.
+The left input, ``dog'', is comprised of only a single word, and can be fed directly into the comparison function.
+``very big cat'', however, contains three of these representations, which need to be reduced to a single vector.
+This is done by applying the composition function twice, yielding a vector of the same length as the words themselves at each application.
+The order of application of the composition function is determined by a constituency structure (unlabeled binary parse tree) provided with the input text.
+This structure is annotated manually for these inputs, but could in other cases be derived from the outputs of any of a large number of extant tree parsers.
+
+Once the input has been reduced to two vectors of equal length, these are fed into the comparison layer, and subsequently into the softmax layer.
+The output of the softmax layer yields a distribution over possible relations between the two constituents, and choosing the maximum at this point yields the model's hypothesized relation.
+
+\subsubsection{The comparison function}
+
+The next layer is the comparison layer.
+This takes in the two $D$-dimensional representations of each constituent, $\vec{x^1}$ and $\vec{x^2}$, and transform them into a $D'$-dimensional vector $y$ representing the relation between the two.
+While this layer is not intended to be the same function as the composition function that builds up the individual constituents, I parameterize it in the same way, as an RNTN layer with three parameters, one a third-order tensor (in other words, a vector of matrices, marked $\mathbf{A^{[1...D']}}$), one a matrix  ($\mathbf{B}$), and one a vector ($\vec{c}$).
+
+Each element in the output vector is computed using a single matrix drawn from the third order tensor parameter, a single row of the matrix parameter, and a single entry of the vector parameter, as follows.
+
+\begin{equation}
+y_i = f(\vec{x^1}^T \mathbf{A^{[1...D']}} \vec{x^2} + \vec{B_{i,:}} [\vec{x^1}; \vec{x^2}] + c_i)
+\end{equation}
+
+\begin{equation}
+Cpr(\vec{x^1}, \vec{x^2}) = [y_1; y_2; ...; y_D']
+\end{equation}
+
+The result of each of these computations is fed into a sigmoid nonlinearity $f$. In my implementation, I use $f(x) = \tanh(x)$.
+
+\subsubsection{The composition function}
+
+As with the comparison layer, the composition layer is an RNTN layer with three parameters, one a third-order tensor ($\mathbf{K^{[1...D']}}$), one a matrix  ($\mathbf{L}$), and one a vector ($\vec{m}$).
+As above, the output of the layer is computed as follows.
+
+\begin{equation}
+p_i = f(\vec{x^1}^T \mathbf{K^{[1...D]}} \vec{x^2} + \vec{L_{i,:}} [\vec{x^1}; \vec{x^2}] + m_i)
+\end{equation}
+
+\begin{equation}
+Mrg(\vec{x^1}, \vec{x^2}) = [y_1; y_2; ...; y_D']
+\end{equation}
+
+\subsubsection{The softmax layer}
+
+TODO: Cite UFDL
+
+The top layer of the model is a softmax classifier, which takes the model output and transforms it into a multinomial probability distribution over the seven possible relation types.
+
+The softmax function takes the input from the layer below, denoted $\vec{y}$, appends a 1 to yield a bias term, multiplies the result by a matrix $\mathbf{S}$ and exponentiates it elementwise to yield an unnormalized probability measure.
+
+\begin{equation}
+\vec{\mu}(\vec{y}) = exp(\mathbf{S}\cdot[1; \vec{y}])
+\end{equation}
+
+From this, we can normalize to get a probability function by dividing each entry in the vector by the sum of the vector:
+\begin{equation}
+\vec{\pi}(\vec{y}) = \frac{\vec{\mu}(\vec{y})_r}{\sum_{r' = 1}^R \mu(\vec{y})_{r'}}
+\end{equation}
+
+\begin{equation}
+P(rel(x_1, x_2)=r) = \pi(\vec{y})_r
+\end{equation}
+
+$\mathbf{S}$ above is a $D' \times R$ matrix.
+
+\subsection{Backpropagation: Learning from a single relation}
+
+The parameters of the model---both the function parameters and the word parameters---are learned using the L-BFGS algorithm TODO: CITE. L-BFGS is a TODO: CHECK
+
+These gradients for each data point are computed using the backpropagation through structure algorithm \cite{goller1996learning}.
+In this algorithm, the model is first run forward in order to compute the predicted distribution, which is in turn used to compute the objective function.
+
+\subsubsection{The Objective}
+
+For an objective function, I use the negative logarithm of the probability of the true relation to capture correctness, and the sum of the squared parameter values (L2 normalization) to minimize overfitting.
+I present it here as a function of the two top-level vectors of the inputs, and the single correct relation specification.
+
+\begin{equation}
+E(\vec{w_1}, \vec{w_2}, \vec{r}, \theta) = - \sum_{i=1}^N \log(\pi(Cpr(\vec{w_1}, \vec{w_2}))_{r_i}) + \frac{\lambda}{2}\sum_j\theta_j^2
+\end{equation}
+
+Where $\theta$ is a single vector containing all of the entries of all of the model parameters:
+\begin{equation}
+\theta = \text{linearize}([\mathbf{A^{[1...D']}}, \mathbf{B}, \vec{c}, \mathbf{K^{[1...D]}}, \mathbf{L}, \vec{m}, \mathbf{S}])
+\end{equation}
+
+\subsection{Gradients for S}
+
+The simplest gradient to compute is that of the last parameter to be applied, the matrix $\mathbf{S}$ in the probability function. Here I do so, assuming for simplicity that $N = 1$.
+
+\begin{equation}
+\frac{\partial E}{\partial S_{j,k}} = \lambda S_{j,k} - \frac{\partial E}{\partial \pi(\vec{\mu}(\vec{y}))_{r_i}}\frac{\partial \pi(\vec{\mu}(\vec{y}))_{r_i}}{\partial S_{j,k}} = 
+\end{equation}
+
+Where $\pi(\vec{\mu}(\vec{y}))_{r_i}$ is the probability of the true relation.
+
+\begin{equation}
+\frac{\partial E}{\partial \pi(\vec{\mu}(\vec{y}))_{r_i}} = -\frac{ 1 }{\pi(\vec{\mu}(\vec{y}))_{r_i}}
+\end{equation}
+
+\begin{equation}
+\vec{\mu}(\vec{y}) = exp(\mathbf{S}\cdot[1; \vec{y}])
+\end{equation}
+
+\begin{equation}
+\vec{\pi}(\vec{y})_r = \frac{\vec{\mu}(\vec{y})_r}{\sum_{r' = 1}^R \mu(\vec{y})_{r'}}
+\end{equation}
+
+if $r \neq j$
+
+%f = \vec{\mu}(\vec{y})_r
+%g = \sum_{r' = 1}^R \mu(\vec{y})_{r'}
+%f' = 0
+%g' = [1; \vec{y}]_{k}
+%
+%if r \neq j
+
+\begin{equation}
+\frac{\partial \pi(\vec{y})_{r_i}}{\partial S_{j,k}} = \frac{-\vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
+\end{equation}
+
+\begin{equation}
+\frac{\partial E}{\partial S_{j,k}} = \lambda S_{j,k} - \frac{ }{\pi(\vec{y})_{r_i}} \frac{\vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
+\end{equation}
+
+%f = \vec{\mu}(\vec{y})_r
+%g = \sum_{r' = 1}^R \mu(\vec{y})_{r'}
+%f' = [1; \vec{y}]_{k}
+%g' = [1; \vec{y}]_{k}
+%
+
+if eq
+
+\begin{equation}
+\frac{\partial \pi(\vec{y})_{r_i}}{\partial S_{j,k}} = \frac{[1; \vec{y}]_{k}\cdot \sum_{r' = 1}^R \mu(\vec{y})_{r'} - \vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
+\end{equation}
+
+\begin{equation}
+\frac{\partial E}{\partial S_{j,k}} = \lambda S_{j,k}  + \frac{ 1 }{\pi(\vec{\mu}(\vec{y}))_{r_i}} \frac{[1; \vec{y}]_{k}\cdot \sum_{r' = 1}^R \mu(\vec{y})_{r'} - \vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
+\end{equation}
+
+%d pi_ri  / d mu
+
+%f = \vec{\mu}(\vec{y})_r
+%g =
+
+%\vec{\pi}(\vec{y}) = \frac{\vec{\mu}(\vec{y})_r}{\sum_{r' = 1}^R \mu(\vec{y})_{r'}}
+
+%\log(\pi(Cpr(\vec{w_1}, \vec{w_2}))_{r_i})
+
+\subsection{The comparison function}
+
+Since we use $\tanh(x)$ for the sigmoid function $f$, we have that $\frac{df}{dx} = sech^2(x)$.
+
+\bibliography{../../../Bibs/MLSemantics}{}
 \bibliographystyle{plainnat}

### 2013-10-15 Writing/paper_cogsci.tex
Update after a long absence… still not quite done

--- 2013-10-15 Writing/paper.tex
@@
-%Use this template for comments in the margin: \com{Note!}
+% 
+% Annual Cognitive Science Conference
 
-%\date{December 2010}                           % Activate to display a given date or no date
+% Modified: Niels Taatgen (taatgen@cmu.edu) 10/24/2006
 
-%
-%  From Riggle:
-%
-%----------------------------------------------------------------------------%
-%                    Basic page layout parameters                            %
-%----------------------------------------------------------------------------%
+%% Change ``a4paper'' in the following line to ``letterpaper'' if you are
+%% producing a letter-format document.
 
-% \documentclass[11pt,leqno,tbtags]{article}
-\documentclass[12pt,leqno,tbtags,twoside]{article}
+\documentclass[10pt,letterpaper]{article}
 
-    % leqno = equations at the left
-    % tbtags (+leqno) = equation numbers at the top (rather than centered)
+\usepackage{cogsci}
+\usepackage{pslatex}
+\usepackage{apacite}
+\usepackage{amsmath}
+\usepackage{qtree}
 
-\usepackage[letterpaper,hmargin=1in,vmargin=1in]{geometry} 
-\usepackage[bottom]{footmisc}
-\usepackage{fancyhdr}
+\title{Logical Reasoning in Compositional Vector Space Semantics}
 
-\fancyhead{} % clear all header fields
-
-\fancyhead[LO]{Samuel R. Bowman}
-\fancyhead[RE]{Logical Reasoning in Compositional Vector Space Semantics}
-
-\fancyfoot{} % clear all footer fields
-\fancyfoot[CE, CO]{\thepage}
-
-\renewcommand{\headrulewidth}{0.4pt}
-
-\renewcommand{\footrulewidth}{0.4pt}
-
-\pagestyle{fancy}
-\setlength{\headheight}{15pt}
-
-\usepackage{setspace}
-\spacing{1}
-
-\usepackage{enumitem}
-\setlist{noitemsep}
-
-\usepackage{float}
-
-\usepackage[round]{natbib}
-\bibpunct[,~]{(}{)}{,}{a}{,}{,}
-\setlength{\bibsep}{0pt}
-
-%----------------------------------------------------------------------------%
-%                    Title, author, and affiliation                          %
-%----------------------------------------------------------------------------%
-
-\title{\Large Logical Reasoning in Compositional Vector Space Semantics}
-\author{Samuel R. Bowman\\\textsc{Ling 236}}
-%\\\\Qualifying Paper \textit{for}\\The Department of Linguistics, Stanford University\\\\Committee:\\Paul Kiparsky (chair), Arto Anttila, and Christopher Manning
-% \date{Edinburgh P-Workshop, May 2013}
-% \author{Max Bane and Jason Riggle}
-% \date{draft -- \today}
-
-%----------------------------------------------------------------------------%
-%                    Fonts and encoding                                      %
-%----------------------------------------------------------------------------%
-
-\usepackage[dirty,tipa]{ucs}	% Utf-8
-\usepackage[utf8x]{inputenc}	% fonts
-\usepackage[T1]{fontenc}		% in Tex
-
-\usepackage{tipa}\let\ipa\textipa
-\usepackage{vowel}
-
-\usepackage{pifont}
-\newcommand{\BlankCell}{}
-
-%\usepackage{palatino}           % body font 
-\renewcommand{\ttdefault}{lmtt} % tt set to latinmodern
-
-%----------------------------------------------------------------------------%
-%                    Graphics, tables, and figures                           %
-%----------------------------------------------------------------------------%
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{graphicx}
-\usepackage{multido}
-\usepackage{ot-tableau}
-\usepackage{pstricks}
-\usepackage{pst-slpe}
-
-\psset{gridcolor=green, subgridcolor=white, dimen=middle,
-	   linewidth=1pt, subgriddiv=1, gridlabels=4pt, gridwidth=0.2pt}
-\psset{cornersize=absolute,linearc=0.06}
-
-\newpsstyle{final}{linewidth=0.5pt,fillstyle=solid,fillcolor=gray95,%
-				   doubleline=true,doublesep=0.8pt,dimen=outer}
-
-%----------------------------------------------------------------------------%
-%                    Algorithms and Theorems                                 %
-%----------------------------------------------------------------------------%
-
-\usepackage[ruled,noend,linesnumbered]{algorithm2e}
-
-\usepackage[leqno,fleqn]{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{wasysym}
-
-\usepackage{stmaryrd}
-
-\newtheoremstyle%
-{dotless}%      name
-{\topsep}%      Space above
-{\topsep}%      Space below
-{\normalfont}%  Body font
-{}%             Indent amount (empty = no indent, \parindent = para indent)
-{\bfseries}%    Thm head font
-{}%             Punctuation after thm head
-{\newline}%     Space after head: " " = interword space, \newline = linebreak
-{}%             Thm head spec (can be left empty, meaning `normal')
-%               alternative spec: {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}
-
-\theoremstyle{dotless}
-
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{lem}[thm]{Lemma}
-
-\newtheorem{exe}[thm]{Example}
-
-%----------------------------------------------------------------------------%
-%                    Hyperlinks                                              %
-%----------------------------------------------------------------------------%
-\newrgbcolor{lnkClr}{0.0 0.0 0.33}
-
-\usepackage{hyperref}			% Clickable links in pdfs
-\hypersetup{colorlinks,			% nice for large docs and 
-			breaklinks,			% for links to web pages
-            linkcolor=lnkClr,
-			urlcolor=lnkClr,
-            anchorcolor=lnkClr,
-			citecolor=lnkClr}
-
-%----------------------------------------------------------------------------%
-%  Example environment
-%----------------------------------------------------------------------------%
-
-\newenvironment{example}
-{\refstepcounter{equation}\begin{list}{(\theequation)}{\leftmargin=7mm \labelwidth=5mm}\item}%
-{\end{list}}
-
-\def\citeapos#1{\citeauthor{#1}'s (\citeyear{#1})}
-
-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%
-%./\..../\..../\..../\..../\..../\..../\..../\..../\..../\..../\..../\..../\.%
-%/  \../  \../  \../  \../  \../  \../  \../  \../  \../  \../  \../  \../  \%
-%    \/    \/    \/    \/    \/    \/    \/    \/    \/    \/    \/    \/    %
-%%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%
-\usepackage{gb4e}
+\author{{\large \bf Samuel R. Bowman (sbowman@stanford.edu)} \\
+  Department of Linguistics, 450 Serra Mall \\
+  Stanford, CA 94305-2150
+  \AND {\large \bf Sharon J. Derry (SDJ@Macc.Wisc.Edu)} \\
+  Department of Educational Psychology, 1025 W. Johnson Street \\
+  Madison, WI 53706 USA}
 
 \begin{document}
-\maketitle
-% \addtocounter{section}{-1}
-
-\def\Y{\textbabygamma}
-\def\P{\textceltpal}
-\def\Rs{\textsubwedge} % restore this
-\def\E{\tipaloweraccent[.1ex]{8}{e}} %restore this
-\def\A{\sublptr{\"{a}}}
 
@@
 
-%TODO: See leichturm margins
+\begin{abstract}
+The abstract should be one paragraph, indented 1/8~inch on both sides,
+in 9~point font with single spacing. The heading {\bf Abstract} should
+be 10~point, bold, centered, with one line space below it. This
+one-paragraph abstract section is required only for standard spoken
+papers and standard posters (i.e., those presentations that will be
+represented by six page papers in the Proceedings).
 
-%TODO: See Ken Shan slides
+\textbf{Keywords:} 
+vector space models, computational semantics\end{abstract}
+
+%\section{Formalities, Footnotes, and Floats}
+
+%Use standard APA citation format. Citations within the text should
+%include the author's last name and year. If the authors' names are
+%included in the sentence, place only the year in parentheses, as in
+%\citeA{NewellSimon1972a}, but otherwise place the entire reference in
+%parentheses with the authors and year separated by a comma
+%\citeA{NewellSimon1972a}. List multiple references alphabetically and
+%separate them by semicolons
+%\citeA{ChalnickBillman1988a,NewellSimon1972a}. Use the
+%et~al. construction only after listing all the authors to a
+%publication in an earlier reference and for citations with four or
+%more authors.
+
+%\subsection{Footnotes}
+
+%Indicate footnotes with a number\footnote{Sample of the first
+%  footnote.} in the text. Place the footnotes in 9~point type at the
+%bottom of the page on which they appear. Precede the footnote with a
+%horizontal rule.\footnote{Sample of the second footnote.}
+
+%\subsection{Tables}
+
+%Number tables consecutively; place the table number and title (in
+%10~point) above the table with one line space above the caption and
+%one line space below it, as in Table~\ref{sample-table}. You may float
+%tables to the top or bottom of a column, set wide tables across both
+%columns.
+
+%\begin{table}[!ht]
+%\begin{center} 
+%\caption{Sample table title.} 
+%\label{sample-table} 
+%\vskip 0.12in
+%\begin{tabular}{ll} 
+%\hline
+%Error type    &  Example \\
+%\hline
+%Take smaller        &   63 - 44 = 21 \\
+%Always borrow~~~~   &   96 - 42 = 34 \\
+%0 - N = N           &   70 - 47 = 37 \\
+%0 - N = 0           &   70 - 47 = 30 \\
+%\hline
+%\end{tabular} 
+%\end{center} 
+%\end{table}
+
+%\subsection{Figures}
+
+%All artwork must be very dark for purposes of reproduction and should
+%not be hand drawn. Number figures sequentially, placing the figure
+%number and caption, in 10~point, after the figure with one line space
+%above the caption and one line space below it, as in
+%Figure~\ref{sample-figure}. If necessary, leave extra white space at
+%the bottom of the page to avoid splitting the figure and figure
+%caption. You may float figures to the top or bottom of a column, or
+%set wide figures across both columns.
+
+%\begin{figure}[ht]
+%\begin{center}
+%\fbox{CoGNiTiVe ScIeNcE}
+%\end{center}
+%\caption{This is a figure.} 
+%\label{sample-figure}
+%\end{figure}
 
@@
 
-%TODO: Replicate Baroni et al 2012
+%TODO: Replicate Baroni et al 2012?
 
@@
 
-\citet{baroni2012entailment} present a real world version of the entailment reasoning problem, and have the current state of the art on that task. [Ken Shan is the last author on this.]
+\citeA{baroni2012entailment} present a real world version of the entailment reasoning problem, and have the current state of the art on that task. [Ken Shan is the last author on this.]
 
@@
 
-\citet{erk2009representing} discusses notions of subsumption from distributional vectors.
+\citeA{erk2009representing} discusses notions of subsumption from distributional vectors.
 His results would be discouraging if we were trying to use pure distributional vectors rather than task-learned ones.
@@
 
-\citet{baroni2011we} present a corpus of word pairs manually annotated with relations like \textit{hypernym}, \textit{hyponym}, \textit{meronym}, \textit{associated event}, and \textit{common attribute}, as well as negative example relations.
+\citeA{baroni2011we} present a corpus of word pairs manually annotated with relations like \textit{hypernym}, \textit{hyponym}, \textit{meronym}, \textit{associated event}, and \textit{common attribute}, as well as negative example relations.
 
@@
 
-\citet{lenci2012identifying} have more of the same: distributional inclusion does capture some entailment information.
+\citeA{lenci2012identifying} have more of the same: distributional inclusion does capture some entailment information.
 
@@
 
-\citet{goller1996learning} present the core of the machine learning algorithm needed for the task-specific representations used in this project. I can't make heads or tails of their writing.
+\citeA{goller1996learning} present the core of the machine learning algorithm needed for the task-specific representations used in this project. I can't make heads or tails of their writing.
 
-\citeapos{widdows2004geometry} book primarily serves as an primer on vector space semantics for those without sufficient mathematical background, but it does briefly address some aspects of the logical capacities of such systems.
+\citeA{widdows2004geometry} primarily serves as an primer on vector space semantics for those without sufficient mathematical background, but it does briefly address some aspects of the logical capacities of such systems.
 In particular, he presents a system for constructing information-retrieval targeted query vectors, and builds a practical quasi-logic around it which is capable of conjunction (\textit{rock} AND \textit{blues}) using vector addition, and subtraction (\textit{rock} NOT \textit{blues}) through the the operation of finding the orthogonal vector, expressed as:
@@
 
-\citet{maccartney2009extended} \citep[from Bill's dissertation][]{maccartney2009natural} presents the framing of entailment and monotonicity used here, and shows that they can (with a lot of hacks) be used to do the Recognizing Textual Entailment task on real language.
+\citeA{maccartney2009extended} \cite{maccartney2009natural}  presents the framing of entailment and monotonicity used here, and shows that they can (with a lot of hacks) be used to do the Recognizing Textual Entailment task on real language.
 No vectors here.
 
-\citet{icard2012inclusion} presents an extension to the NLI/MacCartney and Manning logic involving an elaborated notion of monotonicity.
+\citeA{icard2012inclusion} presents an extension to the NLI/MacCartney and Manning logic involving an elaborated notion of monotonicity.
 He uses the same seven basic entailment relations as MacCartney.
@@
 
-\citet{socher2011semi} is the primary introduction to the idea of learning a set of quasi-distributional vectors and a composition function on those vectors, all guided to optimize performance on some task.
+\citeA{socher2011semi} is the primary introduction to the idea of learning a set of quasi-distributional vectors and a composition function on those vectors, all guided to optimize performance on some task.
 
-\citet{socher2011dynamic} extends this to paraphrase detection, an (easy, by our standards) task of comparing two learned sentence representations.
+\citeA{socher2011dynamic} extends this to paraphrase detection, an (easy, by our standards) task of comparing two learned sentence representations.
 
 Socher et al.
-(in prep; ACL 2013) and \citet{chen2013learning} define a more powerful version of the composition function based on a third-order tensor.
+(in prep; ACL 2013) and \citeA{chen2013learning} define a more powerful version of the composition function based on a third-order tensor.
 I'm using this layer as the top node that builds the comparison vector from which relation types are extracted.
@@
 Basile et al: Encoding syntactic dependencies by vector permutation. Slides are a bit vague. Give it a closer look!
+
+--------
+Worth naming at least one of the Clarke/Grefenstette/... papers:
+http://www.aclweb.org/anthology/W/W11/W11-0114.pdf
+
+http://arxiv.org/pdf/1003.4394v1.pdf
+http://www.aclweb.org/anthology-new/S/S13/S13-1001.pdf
+- Pure theory - no experiment yet.
+- Formalism treats prepositions as fifth-order tensors - not meant to be viable in a machine learning setting as presented.
+
+http://www.aclweb.org/anthology-new/S/S13/S13-1001.pdf
+- Might be the most overt presentation of the tensor approach
+- overfitting issues
+
+Work on similarity
+- http://www.aclweb.org/anthology-new/D/D11/D11-1129.pdf
+
+Most recent
+- http://www.cl.cam.ac.uk/~sc609/pubs/mol13.pdf
+
+\section{Methods}
 
@@
 \subsection{Forward computation: Identifying a single relation}
+
+\begin{figure}[ht]
+\begin{center}
+\fbox{\Tree [.{$\underset{r}{\operatorname{argmax}}(\pi_{\operatorname{index}(r)}) = \text{``}|\text{''}$\\Most probable relation.} [.{$\vec{\pi}$\\Probabilities of each relation.} [.{[Softmax layer]} [.{$\vec{y}$\\World pair representation.} [.{[Comparison layer]} [.{$x_1$\\``dog'' vector} ] [.{$x_6$\\``very big cat'' vector} [.{[Composition layer]} [.{$x_5$\\``very big'' vector} [.{[Composition layer]} [.{$x_2$\\``very'' vector} ] [.{$x_3$\\``big'' vector} ] ] ] [.{$x_4$\\``cat'' vector} ] ] ] ] ] ] ] ] }
+\end{center}
+\caption{Example model structure for comparing ``dog'' and ``very big cat''. When running the model forward to compute the predicted relation, information flows up the tree.} 
+\label{sample-figure}
+\end{figure}
 
@@
 
-The parameters of the model---both the function parameters and the word parameters---are learned using the L-BFGS algorithm TODO: CITE. L-BFGS is a TODO: CHECK
+The parameters of the model---both the function parameters and the word parameters---are learned using the minFunc package for multivariate optimization \citeA{schmidt2012minfunc}, configured to use L-BFGS.
+The update rule by which this algorithm optimizes the parameters requires a closed-form computation of the gradient at each iteration: the partial derivatives for each parameter with respect to the objective function.
 
-These gradients for each data point are computed using the backpropagation through structure algorithm \cite{goller1996learning}.
+This gradient is computed using the sum of the gradients for each item of the training data set, combined with an L2 regularization term.
+The gradients for each data point are computed using the backpropagation through structure algorithm \cite{goller1996learning}.
 In this algorithm, the model is first run forward in order to compute the predicted distribution, which is in turn used to compute the objective function.
@@
 
+The gradient of this function can be decomposed as follows:
+
+\begin{equation}
+\frac{\partial E(\vec{w_1}, \vec{w_2}, \vec{r}, \theta)}{\partial \theta_i} = - \partial{\sum_{i=1}^N \log(\pi(Cpr(\vec{w_1}, \vec{w_2}))_{r_i})}{\partial \theat + \frac{\lambda}{2}\sum_j\theta_j^2
+\end{equation}
+
 \subsection{Gradients for S}
@@
 
-\begin{equation}
-\frac{\partial E}{\partial \pi(\vec{\mu}(\vec{y}))_{r_i}} = -\frac{ 1 }{\pi(\vec{\mu}(\vec{y}))_{r_i}}
-\end{equation}
-
-\begin{equation}
-\vec{\mu}(\vec{y}) = exp(\mathbf{S}\cdot[1; \vec{y}])
-\end{equation}
-
-\begin{equation}
-\vec{\pi}(\vec{y})_r = \frac{\vec{\mu}(\vec{y})_r}{\sum_{r' = 1}^R \mu(\vec{y})_{r'}}
-\end{equation}
-
-if $r \neq j$
-
-%f = \vec{\mu}(\vec{y})_r
-%g = \sum_{r' = 1}^R \mu(\vec{y})_{r'}
-%f' = 0
-%g' = [1; \vec{y}]_{k}
+%\begin{equation}
+%\frac{\partial E}{\partial \pi(\vec{\mu}(\vec{y}))_{r_i}} = -\frac{ 1 }{\pi(\vec{\mu}(\vec{y}))_{r_i}}
+%\end{equation}
 %
-%if r \neq j
-
-\begin{equation}
-\frac{\partial \pi(\vec{y})_{r_i}}{\partial S_{j,k}} = \frac{-\vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
-\end{equation}
-
-\begin{equation}
-\frac{\partial E}{\partial S_{j,k}} = \lambda S_{j,k} - \frac{ }{\pi(\vec{y})_{r_i}} \frac{\vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
-\end{equation}
-
-%f = \vec{\mu}(\vec{y})_r
-%g = \sum_{r' = 1}^R \mu(\vec{y})_{r'}
-%f' = [1; \vec{y}]_{k}
-%g' = [1; \vec{y}]_{k}
+%\begin{equation}
+%\vec{\mu}(\vec{y}) = exp(\mathbf{S}\cdot[1; \vec{y}])
+%\end{equation}
 %
-
-if eq
-
-\begin{equation}
-\frac{\partial \pi(\vec{y})_{r_i}}{\partial S_{j,k}} = \frac{[1; \vec{y}]_{k}\cdot \sum_{r' = 1}^R \mu(\vec{y})_{r'} - \vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
-\end{equation}
-
-\begin{equation}
-\frac{\partial E}{\partial S_{j,k}} = \lambda S_{j,k}  + \frac{ 1 }{\pi(\vec{\mu}(\vec{y}))_{r_i}} \frac{[1; \vec{y}]_{k}\cdot \sum_{r' = 1}^R \mu(\vec{y})_{r'} - \vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
-\end{equation}
+%\begin{equation}
+%\vec{\pi}(\vec{y})_r = \frac{\vec{\mu}(\vec{y})_r}{\sum_{r' = 1}^R \mu(\vec{y})_{r'}}
+%\end{equation}
+%
+%if $r \neq j$
+%
+%
+%%f = \vec{\mu}(\vec{y})_r
+%%g = \sum_{r' = 1}^R \mu(\vec{y})_{r'}
+%%f' = 0
+%%g' = [1; \vec{y}]_{k}
+%%
+%%if r \neq j
+%
+%\begin{equation}
+%\frac{\partial \pi(\vec{y})_{r_i}}{\partial S_{j,k}} = \frac{-\vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
+%\end{equation}
+%
+%
+%\begin{equation}
+%\frac{\partial E}{\partial S_{j,k}} = \lambda S_{j,k} - \frac{ }{\pi(\vec{y})_{r_i}} \frac{\vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
+%\end{equation}
+%
+%
+%%f = \vec{\mu}(\vec{y})_r
+%%g = \sum_{r' = 1}^R \mu(\vec{y})_{r'}
+%%f' = [1; \vec{y}]_{k}
+%%g' = [1; \vec{y}]_{k}
+%%
+%
+%if eq
+%
+%
+%\begin{equation}
+%\frac{\partial \pi(\vec{y})_{r_i}}{\partial S_{j,k}} = \frac{[1; \vec{y}]_{k}\cdot \sum_{r' = 1}^R \mu(\vec{y})_{r'} - \vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
+%\end{equation}
+%
+%
+%\begin{equation}
+%\frac{\partial E}{\partial S_{j,k}} = \lambda S_{j,k}  + \frac{ 1 }{\pi(\vec{\mu}(\vec{y}))_{r_i}} \frac{[1; \vec{y}]_{k}\cdot \sum_{r' = 1}^R \mu(\vec{y})_{r'} - \vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
+%\end{equation}
 
@@
 
-\subsection{The comparison function}
+% \subsection{The comparison function}
 
-Since we use $\tanh(x)$ for the sigmoid function $f$, we have that $\frac{df}{dx} = sech^2(x)$.
+% Since we use $\tanh(x)$ for the sigmoid function $f$, we have that $\frac{df}{dx} = sech^2(x)$.
 
-\bibliography{../../../Bibs/MLSemantics}{}
-\bibliographystyle{plainnat}
+\section{Current status, results and discussion}
+
+We have not yet fully built the necessary gradient computation functions to learn the parameters of the composition layer or to propagate information past the composition layer down to the words.
+Aside from that...
+
+% implemented in matlab
+
+% learning overall distribution
+
+% tuning needed
+
+% data set cut needed
+
+\bibliographystyle{apacite}
+
+\setlength{\bibleftmargin}{.125in}
+\setlength{\bibindent}{-\bibleftmargin}
+
+\bibliography{MLSemantics}
 

### 2013-10-15 ?LoadConstitTrainingData.m

@@ -1,0 +1,81 @@
+function [ data, wordMap, relationMap ] = LoadConstitTrainingData(filename)
+%[ dataset ] = LoadTrainingData()
+
+% 'constitpairs-v1.tsv'
+
+% Import data
+fid = fopen(filename);
+C = textscan(fid,'%s','delimiter',sprintf('\n'));
+fclose(fid);
+
+
+% TRUNCATE
+%C{1} = C{1}(1:5)
+%C{2} = C{2}(1:5)
+%C{3} = C{3}(1:5)
+
+% Load the file
+
+% Initialize the data array
+rawData = repmat(struct('relation', 0, 'leftText', '', 'rightText', ''), ...
+    length(C{1}), 1);
+wordList = cell(length(C{1}), 1);
+
+% Establish (manually specified) relations
+relations = {'#', '=', '>', '<', '|', '^', 'v'};
+relationMap = containers.Map(relations,1:length(relations));
+
+% Parse the file
+itemNo = 1;
+wordNo = 1;
+maxLine = length(C{1});
+% maxLine = 6; % Truncate data.
+
+for line = 1:maxLine;
+    if ~isempty(C{1}{line}) 
+        splitLine = textscan(C{1}{line}, '%s', 'delimiter', '\t');
+        splitLine = splitLine{1};
+        
+        if ~(length(splitLine{1}) ~= 1 || splitLine{1} == '%')
+            % Skip lines that are blank or have a multicharacter first chunk
+            rawData(itemNo).relation = relationMap(splitLine{1});
+            rawData(itemNo).leftText = splitLine{2};
+            rawData(itemNo).rightText = splitLine{3};
+
+            % Add to wordList
+            lWords = textscan(splitLine{1}, '%s', 'delimiter', ' ');
+            rWords = textscan(splitLine{2}, '%s', 'delimiter', ' ');
+            words = unique([lWords{1}, rWords{1}]);
+            wordList(wordNo:wordNo + (length(words) - 1)) = cellstr(words);
+            wordNo = wordNo + length(words);
+
+            itemNo = itemNo + 1;
+        end
+    end
+end
+
+rawData = rawData(1:itemNo - 1);
+
+% Learn vocabulary
+wordList = wordList(1:wordNo - 1);
+vocabulary = unique(wordList);
+
+% Remove syntactic symbols from vocabulary
+vocabulary = setdiff(vocabulary, {'(', ')'});
+
+% Build word map
[…]

### 2013-10-15 ?TrainConstitModel.m

@@ -1,0 +1,73 @@
+[data, wordMap, relationMap] = LoadConstitTrainingData();
+
+V = size(wordMap, 1); % Num of words in vocab.
+[theta, thetaDecoder, hyperParams] = InitializeModel(V);
+
+data = InitializeWordFeatures(data, theta, thetaDecoder);
+
+addpath('../minFunc/minFunc/')
+addpath('../minFunc/minFunc/compiled/')
+addpath('../minFunc/minFunc/mex/')
+addpath('../minFunc/autoDif/')
+
+% minfunc options (not tuned)
+options.Method = 'lbfgs';
+options.MaxIter = 200;
+options.TolX = 1e-4;
+options.DerivativeCheck = 'off';
+options.Display = 'excessive';
+options.numDiff = 0;
+options.LS_init = '2' % Attempt to minimize evaluations per step...
+
+% %%% PRETEST
+% for i = 1:3
+%     [classifierMatrices, classifierMatrix, classifierBias, ...
+%         classifierParameters, wordFeatures] = stack2param(theta, thetaDecoder);
+%     leftTree = data(i).leftTree;
+%     rightTree = data(i).rightTree;
+%     leftTree.updateFeatures(wordFeatures);
+%     rightTree.updateFeatures(wordFeatures);
+%     leftFeatures = leftTree.getFeatures();
+%     rightFeatures = rightTree.getFeatures();
+%     trueRelation = data(i).relation;
+%     disp(leftTree.getText)
+%     disp(rightTree.getText)
+%     disp(trueRelation)
+%     tensorInnerOutput = ComputeInnerTensorLayer(leftFeatures, rightFeatures, classifierMatrices, classifierMatrix, classifierBias);
+%     tensorOutput = Sigmoid(tensorInnerOutput);
+%     tensorDeriv = SigmoidDeriv(tensorInnerOutput);
+%     relationProbs = ComputeSoftmaxProbabilities(tensorOutput, classifierParameters)
+% end
+% wordFeatures
+% %%% END
+
+theta = minFunc(@ComputeFullCostAndGrad, theta, options, thetaDecoder, data, hyperParams);
+
+[cost, grad, acc, confusion] = ComputeFullCostAndGrad(theta, thetaDecoder, data, hyperParams);
+
+acc
+confusion
+
+%  minFunc
+
+%%% TEST
+% for i = 1:3
+%     [classifierMatrices, classifierMatrix, classifierBias, ...
[…]

### 2013-10-15 ComputeCost.m

@@ -1,0 +1,36 @@
+function [ cost, correct ] = ComputeCost( theta, decoder, dataPoint, hyperParams )
+%function [ cost, grad ] = ComputeCostAndGrad( theta, decoder, dataPoint, hyperParams )
+%   Detailed explanation goes here
+
+% Unpack theta.
+[classifierMatrices, classifierMatrix, classifierBias, ...
+    classifierParameters, wordFeatures] = stack2param(theta, decoder);
+
+% Unpack hyperparams.
+NUM_RELATIONS = hyperParams.numRelations;
+PENULT_DIM = hyperParams.penultDim;
+DIM = hyperParams.dim;
+
+leftTree = dataPoint.leftTree;
+rightTree = dataPoint.rightTree;
+trueRelation = dataPoint.relation;
+
+% Make sure word features are current.
+leftTree.updateFeatures(wordFeatures);
+rightTree.updateFeatures(wordFeatures);
+
+leftFeatures = leftTree.getFeatures();
+rightFeatures = rightTree.getFeatures();
+
+% Use the tensor layer to build classifier input:
+tensorInnerOutput = ComputeInnerTensorLayer(leftFeatures, rightFeatures, classifierMatrices, classifierMatrix, classifierBias);
+tensorOutput = Sigmoid(tensorInnerOutput);
+tensorDeriv = SigmoidDeriv(tensorInnerOutput);
+
+relationProbs = ComputeSoftmaxProbabilities(tensorOutput, classifierParameters);
+
+% Increment local error
+cost = Objective(trueRelation, relationProbs);
+
+end
+

### 2013-10-15 ComputeFullCostAndGrad.m

@@ -1,0 +1,80 @@
+function [ cost, grad, trainingError, confusion ] = ComputeFullCostAndGrad( theta, decoder, data, hyperParams )
+%function [ cost, grad ] = ComputeFullCostAndGrad( theta, decoder, data, hyperParams )
+%   Compute cost and gradient over a full dataset for some parameters.
+
+N = length(data);
+
+argout = nargout;
+if nargout > 3
+    confusions = zeros(N, 2);
+end
+
+accumulatedCost = 0;
+accumulatedSuccess = 0;
+if nargout > 1
+    accumulatedGrad = zeros(length(theta), 1);
+end
+
+% Parallelize
+if matlabpool('size') == 0 % checking to see if my pool is already open
+    matlabpool;
+end
+
+if nargout > 2
+    parfor i = 1:N
+        [localCost, localGrad, localPred] = ...
+            ComputeCostAndGrad(theta, decoder, data(i), hyperParams);
+        accumulatedCost = accumulatedCost + localCost;
+        accumulatedGrad = accumulatedGrad + localGrad;
+        
+        localCorrect = localPred == data(i).relation;
+        if ~localCorrect
+            disp('error. correct:')
+            disp(data(i).leftTree.getText)
+            disp(data(i).rightTree.getText)
+            disp(data(i).relation)
+            disp('predicted:')
+            disp(localPred)
+        end
+        
+        if argout > 3
+            confusions(i,:) = [localPred, data(i).relation]
+        end
+             
+        accumulatedSuccess = accumulatedSuccess + localCorrect;
+    end
+    
+    if nargout > 3
+        confusion = zeros(hyperParams.numRelations);
+        for i = 1:N
+           confusion(confusions(i,1), confusions(i,2)) = ...
+               confusion(confusions(i,1), confusions(i,2)) + 1;
+        end
+    end
+elseif nargout > 1
+    parfor i = 1:N
+        [localCost, localGrad] = ...
+            ComputeCostAndGrad(theta, decoder, data(i), hyperParams);
+        accumulatedCost = accumulatedCost + localCost;
+        accumulatedGrad = accumulatedGrad + localGrad;
+    end
+else
+    parfor i = 1:N
+        localCost = ...
+            ComputeCostAndGrad(theta, decoder, data(i), hyperParams);
+        accumulatedCost = accumulatedCost + localCost;
+    end
+end
+
[…]

### 2013-10-15 TestGradientInTensor.m

@@ -1,21 +1,24 @@
-function gradient_error = TestGradientInTensor (i, j, delta, classifierMatrices, classifierParameters, combinedFeatures, trueRelation)
-% TestGradientInTensor (i, j, delta, classifierMatrices, classifierParameters, combinedFeatures, trueRelation)
-% Find out how close a gradinent for one entry in classifierMatrices is to
+function gradient_error = TestGradientInTensor (i, grad, params, decoder, a, b, trueRelation)
+% TestGradientInTensor (i, delta, feats, decoder, a, b, trueRelation)
+% Find out how close a gradinent for one entry in tensor parameters is to
 % the local slope.
+% Use: feats = param2stack(classifierMatrices, classifierMatrix, classifierBias)
 
-EPSILON = 0.1;
+DELTA = 0.0001;
 
-originalValue = classifierMatrices(i, j);
+originalValue = params(i);
 objs = [];
 for polarity = -1:2:1
-    classifierMatrices(i,j) = originalValue + polarity * EPSILON;
-    tensorOutput = ComputeTensorLayer(combinedFeatures, classifierMatrices);    
+    params(i) = originalValue + polarity * DELTA;
+    [classifierMatrices, classifierMatrix, classifierBias, classifierParameters] = stack2param(params, decoder);
+    tensorOutput = ComputeTensorLayer(a, b, classifierMatrices, classifierMatrix, classifierBias);    
     probs = ComputeSoftmaxProbabilities(tensorOutput, classifierParameters);
-    objs = [objs Objective(trueRelation, probs)];
+    objs = [objs Objective(trueRelation, probs, params)];
 end
-    
-localSlope = (objs(2) - objs(1)) / (2 * EPSILON);
-gradient_error = (localSlope / delta);
+
+localSlope = (objs(2) - objs(1)) / (2 * DELTA);
+
+gradient_error = (grad / localSlope);
     
 % if localSlope > 0.1
 %     % For histogram of error:

### 2013-10-15 Tree.m

@@ -1,4 +1,4 @@
-classdef Tree
+classdef Tree < handle
     
     % Represents a single binary branching syntactic tree with three representations at each node:
     % - The index with which the feature vector can be looked up - IF LEAF
@@ -15,24 +15,90 @@ classdef Tree
     methods(Static)
         % TODO: Constructors for multiword trees?
         
-        function t = makeLeaf(iText)
-            global wordMap;
-            global wordFeatures;
+        function t = makeTree(iText, wordMap)
+            % Parsing strategy:          
+            % ( a b ) ( c d )
+            % (
+            %  cache - a
+            %  cache - a b
+            % )
+            % cache -> ab - merge last two nodes
+            % (
+            %  cache ab c
+            %  cache ab c d
+            % )
+            % cache ab cd
+            % cache abcd
+            % 
+            % 
+            % ( a ( ( b c ) d ) )
+            % (
+            % cache a
+            % (
+            % (
+            % cache a b
+            % cache a b c
+            % )
+            % cache a bc
+            % cache a bc d
+            % )
+            % cache a bcd
+            % )
+            % cache abcd
+            
+            C = textscan(iText, '%s', 'delimiter', ' ');
+            C = C{1};
+             
+            stack = cell(length(C));
+            stackTop = 0;
+            
+            for i = 1:length(C)
+                if ~strcmp(C{i}, '(') && ~strcmp(C{i}, ')')
+                    % Turn words into leaf nodes
+                    stack{stackTop + 1} = Tree.makeLeaf(C{i}, wordMap);
+                    stackTop = stackTop + 1;
+                elseif strcmp(C{i}, ')')
+                    % Merge at the ends of constituents
+                    r = stack{stackTop};
+                    l = stack{stackTop - 1};
+                    stack{stackTop - 1} = Tree.mergeTrees(l, r);
+                    stackTop = stackTop - 1;
+                end
+            end
+            
+            t = stack{stackTop};
+            stackTop = stackTop - 1;
+            
+            % Merge in any residual words to the left
+            while stackTop > 0
+                p = stack{stackTop};
+                stackTop = stackTop - 1;
+                
+                t = Tree.mergeTrees(p, t);
+            end            
+        end
+        
+        function t = makeLeaf(iText, wordMap)
             t = Tree();
             t.text = iText;
             t.wordIndex = wordMap(t.text);
-            t.features = wordFeatures(t.wordIndex, :);
+        end
+        
+        function t = mergeTrees(l, r)
+            t = Tree();
[…]

### 2014-05-30 writing/S'14 paper/abstract.tex
Adding paper draft, cleaning up data dirs.

\begin{abstract}
Supervised recursive neural network models (RNNs) for sentence meaning
have been successful in a wide array of sophisticated language tasks,
but it remains an open question whether they can achieve the same
results as compositional semantic grammars based in logical forms. We
address this question directly by using a logical grammar to generate
controlled data sets encoding the relationships (entailment, synonymy,
contradiction) between pairs of expressions and evaluating whether
each of two classes of neural model --- plain RNNs and recursive
neural tensor networks (RNTNs) --- can learn those relationships
correctly. Our first experiment confirms that both models can learn
the basic algebra of logical relations involved. Our second and third
experiments expand on this result to cover complex recursive
structures and sentences involving quantification. We find that the
plain RNN achieves only mixed results on the second and third
experiments, but that the stronger RNTN models generalized well in
every case.
\end{abstract}

### 2014-05-30 writing/S'14 paper/discussion.tex
Adding paper draft, cleaning up data dirs.

\subsection{General discussion}  % TODO

1 Para

- capable of a geometric encoding of fundamentals of logic

- capable of capturing recursive structure up to a scale comparable to what's used in language, and of true recursion

- capable of learning functional word meanings given adequate data, frontiers: more complex cases

### 2014-05-30 writing/S'14 paper/intro.tex
Adding paper draft, cleaning up data dirs.

\section{Introduction}
% TODO: cite \cite{watanabe2012latent} for past work on ML for McC&M-style NLI

Deep learning methods in NLP which learn vector representations for words have seen successful uses in recent years on increasingly sophisticated tasks \cite{collobert2011natural, socher2011semi, socher2013acl1, chen2013learning}. Given the still modest performance of semantically rich NLP systems in many domains---question answering and machine translation, for instance---it is worth exploring the degree to which learned vectors can serve as general purpose semantic representations. 
Much of the work to date analyzing vector representations for words (see \cite{baroni2013frege}) has focused on lexical semantic behaviors---like the similarity between words like \ii{Paris} and \ii{France}. Good similarity functions are valuable for many NLP tasks, but there are real use cases for which it is necessary to know how words relate to one another or to some extrinsic representation, and to model the ways in which word meanings combine to form phrase, sentence, or document meanings. This paper explores the ability of linguistic representations developed using supervised deep learning techniques to support interpretation and reasoning.

% TODO: Name RNTNs

There are two broad family of tasks that could be used to test the ability of a model to develop general purpose meaning representations. In an interpretation task, sentences are mapped onto some denotation, such as  \ii{true} or \ii{false} for statements, or a factual answer for questions. There has been preliminary work in developing distributed models for interpretation \cite{grefenstette2013towards, rocktaschellow}, but developing a representation of world knowledge that corresponds accurately to the content expressed in language introduces considerable design challenges. I approach the problem by way of an inference task instead. Inferring the truth of one sentence from another does not require any preexisting knowledge representations, but nonetheless requires a precise representation of sentence meaning. I borrow the structure of the task from MacCartney and Manning  \cite{maccartney2009extended}. In it, the model is presented with a pair of sentences, and made to label the logical relation between the sentences as equivalence, entailment, or any of five other classes, as here:

\begin{quote}
\begin{enumerate}\small
\item Many smartphone users avoid high bills overseas by disabling data service.
\item Not everyone uses their smartphones for email when traveling abroad.
\end{enumerate} 
$\Rightarrow$ Entailment
\end{quote}

In this paper, we test the ability of recursive models to on three simple tasks, each of which is meant to capture a property that is necessary in representing natural language meaning in the setting of inference. I begin with an overview of MacCartney and Manning's \cite{maccartney2009extended} framework for inference, and of the recursive neural networks that I study. by showing that these models can learn to correctly represent entailment representations between sentences. I then show that these models can capture the meanings of recursive structurers accurately up to a sufficient depth. I finally close with a brief demonstration of the ability of these models to reason over short natural language sentences involving quantifiers.

% TODO: Cite ICLR paper, emphasize new work since

\subsection{The task: natural language inference}

% Condense?
In the standard formulation of the inference task (and the one used in the RTE datasets \cite{dagan2006pascal}), the goal is to determine whether a reasonable human would infer a hypothesis from a premise.
MacCartney formalizes a method of inferring entailment relations, and moves past two way entailment/non-entailment classification, proposing the set $\mathfrak{B}$ of seven labels meant to describe all of the possible non-trivial relations that might hold between pairs of statements, shown in Table \ref{b-table}.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|} \hline
name & symbol & set-theoretic definition & example \\ \hline \hline
entailment & $x \sqsubset y$ & $x \subset y$ & \ii{crow, bird}  \\ \hline
reverse entailment & $x \sqsupset y$ & $x \supset y$ & \ii{Asian, Thai}  \\ \hline
equivalence & $x \equiv y$ & $x = y$ & \ii{couch, sofa} \\ \hline
alteration & $x$ $|$ $y$ & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{cat, dog} \\ \hline
negation & $x \natneg y$ & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{able, unable} \\ \hline
cover & $x \smallsmile y$ & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-ape} \\ \hline
independence & $x$ \# $y$ & (else) & \ii{hungry, hippo}\\ \hline
\end{tabular}
\caption{The entailment relations in  $\mathfrak{B}$. $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, and the relation \# applies whenever none of the other six do, including when there is insufficient knowledge to choose a label.}
\label{b-table}
\end{center}
\end{table}

In order to minimize this possibility, I define the task of strict unambiguous NLI (SU-NLI). In this task, only entailments that are licensed by a strictly literal interpretation of the provided sentences are considered valid, and several constraints are applied to the language to minimize ambiguity:
\begin{itemize}
\ex A small, unambiguous vocabulary is used.
\ex All strings are given an explicit unlabeled tree structure parse.
\ex Statements involving the hard-to-formalize generic senses of nouns---i.e. \ii{dogs bark} as opposed to the non-generic \ii{all dogs bark}---are excluded.
\ex The sentences do not contain context dependent elements. This includes any reference to time or any tense morphology, and all pronouns.
\ex The morphology is dramatically simplified: the copula is not used (\ii{some puppy is French} is simplified to \ii{some puppy French}, to make it more directly comparable to sentences like \ii{some puppy bark}), and agreement marking (\ii{they walk} vs. \ii{she walks}) is omitted.
\end{itemize}

The key to success at this task is to learn a set of representations and functions that can mimic the logical structure underlying the data. There is limited precedent that deterministic logical behavior can be learned in supervised deep learning models: \citet{socher2012semantic} show in an aside that a Boolean logic with negation and conjunction can be learned in a minimal recursive neural network model with one-dimensional (scalar) representations for words. Modeling the logical behavior that underlies linguistic reasoning, though, is a substantially more difficult challenge, even in modular hand-built models.

The natural logic engine at the core of MacCartney's \cite{maccartney2009natural} NLI system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures. These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
for dozens of different cases of insertion, deletion and substitution of different types of lexical item. For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences. Here, the substitution of \ii{no} for \ii{some}  yields $\natneg$ and the substitution of \ii{dogs} for \ii{animals} yields $\sqsupset$. It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
\begin{quote}

1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}

\end{quote}

This study is the first that I am aware of to attempt to build an inference engine based on learned vector representations, but two recent projects have attempted to introduce vector representations into inference systems in other ways: 
\citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations. They propose a deterministic inference engine (similar to MacCartney's) which is augmented by a probabilistic component that evaluates individual lexical substitution steps in the derivation using vector representations, though again these representations are not learned, and no evaluations of this system have been published to date.
\label{sec2}

### 2014-05-30 writing/S'14 paper/join.tex
Adding paper draft, cleaning up data dirs.

\section{Relation composition}

In order for a model to be able to learn to mimic the behavior of a relational logic like the one presented here from a finite amount data, it must be able to learn to deduce new relations from seen relations. The simplest such deductions involve facts about the relations themselves that do not involve considering the internal structure of the things being compared. For example, given that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that $a\sqsupset c$ by the transitivity of $\sqsupset$, even without understanding $a$, $b$, or $c$. These seven relations support more than just transitivity: MacCartney and Manning's \cite{maccartney2009extended} join table defines 32 valid inferences that can be made on the basis of pairs of relations of the form $a R b$ and $b R' c$, including several less intuitive ones such as that if $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.

\begin{figure}[t]
\begin{center}
\begin{tabular}{ll|lll}
$a := \{1, 3, 4, 5, 6\}	$&~~~&~~~& $a \equiv a$	    	& $a~\#~c$	\\
$b := \{0, 3, 5, 6\}	$&~~~&~~~& $e \sqsupset c$	&$b \smallsmile c$		\\
$c := \{1, 2, 3, 4\}	$&~~~&~~~& $d \smallsmile e$	& $b~\#~d$		\\
$d := \{0, 4\}		$&~~~&~~~& $a \sqsubset e$	& $ d \equiv d$	\\
$e := \{1, 2, 3, 4, 5, 6\}$&~~~&~~~& $a~\#~b$		& ... 	\\
\end{tabular}
\end{center}

\caption{Some sample randomly generated sets, and some of the relations defined between them.  \label{lattice-figure}} 
\end{figure}

I test the model's ability to learn this behavior by creating an artificial dataset of terms which represent sets of numbers. Since MacCartney and Manning's set of relations hold between sets as well as between sentences, I can use the underlying set structure to generate the all of the relations that hold between any pair of these terms, as in Figure \ref{lattice-figure}. I train the model defined above on a subset of these relations, but rather then presenting the model with a pair of tree-structured sentences as inputs, simply present it with two single terms, each of which corresponds to a single vector in the (randomly initialized) vocabulary matrix $V$, ensuring that the model has no information about the terms being compared except the relations between them.

I generate 80 randomly generated sets drawing from the same seven elements, and create a dataset consisting of the relations between every pair of sets, yielding 6400 pairs. 3200 of these pairs were then chosen as a test dataset, and that test dataset was further split into the 2960 examples that can be provably derived from the test data using MacCartney and Manning's join table (or by the symmetry of the relations in about half of the cases) and the 240 that cannot. % TODO: Say more about symmetry?

 A TODO-dimensional RNTN model tuned for the task was able to correctly label TODO \% of the derivable test examples, and TODO \% of the remaining examples. The simpler RNN model with the same number of dimensions was able to derive only TODO \% and TODO \%, respectively.

### 2014-05-30 writing/S'14 paper/methods.tex
Adding paper draft, cleaning up data dirs.

\section{Methods}

The model that I study is a recursive neural network model \citet{socher2011dynamic}, and is thus centered on a recursively applied composition function which is meant to mimic the recursive construction of meanings in formal models of semantics. In this scheme, pairs of words are merged into phrase representations by a function that maps from representations of length $2N$ to representations of length $N$. These phrase representations are then further merged with other words and phrases until the entire phrase or sentence being evaluated is represented in a single vector. This vector is then used as the input to a classifier and used in a supervised learning task.

Borrowing a model directly from the existing literature for this task is impossible since none has been proposed to detect asymmetric relations between phrases. Instead, I build a combination model, depicted in Figure \ref{sample-figure}. The two phrases being compared are built up separately on each side of the tree using the same composition function until they have each been reduced to single vectors. Then, the two phrase vectors are fed into a separate comparison layer that is meant to generate a feature vector capturing the relation between the two phrases. The output of this layer is then given to a softmax classifier, which in turn produces a hypothesized distribution over the seven relations.

For a composition layer, I experiment with both a plain neural network layer function (eq. \ref{rnn}) and the RNTN layer function proposed in \citet{chen2013learning} (eq. \ref{rntn}). A sigmoid nonlinearity (elementwise $\tanh$) is applied to the output of either layer function, following Socher et al.

 \label{rnn}\label{rntn} % TODO: Fix labeling
\begin{equation}
y_{RNN} = f(M [\vec{x}^{(l)}; \vec{x}^{(r)}] + b_i)\\
y_{RNTN}^i = f(\vec{x}^{(l)T} \mathbf{A^{[i]}} \vec{x}^{(r)} + \vec{B_{i,:}} [\vec{x}^{(l)}; \vec{x}^{(r)}] + c_i)
\end{equation}

The comparison layer uses the same type of function with different parameters and a different nonlinearity function wrapped around it. Rather than use a $\tanh$ nonlinearity here, I found a substantial improvement in performance by using a rectified linear function for $f_{b}$. In particular, I use the leaky rectified linear function \cite{maasrectifier}: $f_{b}(\vec{x})=\max(\vec{x}, 0) + 0.01\min(\vec{x}, 0)$,  applied elementwise.

To run the model forward and label a pair of phrases, the structure of the lower layers of the network is assembled so as to mirror the tree structures provided for each phrase. The word vectors are then looked up from the vocabulary matrix $V$ (one of the model parameters), and the composition and comparison functions are used to pass information up the tree and into the classifier at the top. The model is trained using backpropagation through structure  \cite{goller1996learning}, wherein the negative log of the probability assigned to the correct label is taken as a cost function, and the gradient of each parameter with respect to that cost function is computed at each node, with information passing down the tree and into both the function parameters and the vocabulary matrix. Gradients for different instances of the composition function or different instances of a word in the same tree are summed to produce at most a single gradient update per parameter.

\textit{Optimization:} I train the model with stochastic gradient descent (SGD), with gradients pooled from randomly chosen minibatches of 32 training examples, and learning rates computed using AdaGrad \cite{duchi2011adaptive} from a starting rate of 0.2. The parameters (including the vocabulary) are initialized randomly using a uniform distribution over $[-0.1, 0.1]$. % TODO: Update

\textit{Model size:} The dimensionalities of the model were tuned extensively. The performance of the model is approximately optimal with the dimensionality of the word vectors set to 16, and the dimensionality of the feature vector produced by the comparison layer set to 45. % TODO: Later

\begin{figure}
\begin{center}
\Tree [.{\sc softmax classifier}  [.{\sc comparison layer} [.{\sc \ii{dog} vector from $V$} ] [.{\sc composition layer} [.{\sc composition layer} [.{\sc \ii{very} vector from $V$} ] [.{\sc \ii{big} vector from $V$} ] ] [.{\sc \ii{cat} vector from $V$} ] ] ] ]
\end{center}
\caption{The model structure used to compare \ii{dog} and \ii{(very big) cat}. \label{sample-figure}} 
\end{figure} % TODO: Create a more compact figure using something better than qtree.

\ii{Source code and data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?

### 2014-05-30 writing/S'14 paper/quantifiers.tex
Adding paper draft, cleaning up data dirs.

\section{Reasoning with natural language quantifiers}

% Condense condense condense

Interest in NLI in the NLP community is ongoing, and a version of it has been included in the 2014 SemEval challenge specially targeted towards the evaluation of distributional models. Neither this dataset, nor any other existing RTE and NLI datasets are appropriate for the task at hand however. I intend for the present experiment to evaluate the ability of a class of models to learn certain types of inference behavior, and need a dataset that precisely tests these phenomena. With existing datasets that use unrestricted natural language, there is the risk that a model could successfully capture monotonicity inference, but fail to accurately label test data due to problems with, for example, lexical ambiguity, syntactic ambiguity, coreference resolution, or pragmatic enrichment.

\subsection{Data}

The dataset is built on a small purpose-built vocabulary, which is intended to be large and diverse enough to permit a wide variety of experiments on different semantic phenomena beyond those shown here, but which is still small enough to allow the relations between the lexical items themselves to be exhaustively manually labeled when needed. The vocabulary contains 41 predicates (nouns, adjectives, and intransitive verbs; see Appendix B for more on the design of the wordlist), six quantifiers, the logical conjunctions \emph{and} and \emph{or}, and the unary operator \emph{not}.

The data take the form of 12k pairs of short sentences, each annotated with a relation label, which I divide into about 200 smaller datasets. These datasets contain tightly constrained variants of a specific valid reasoning pattern, each differing in at most one lexical item, and all sharing the same relation label. The remainder of this section describes the composition of these 200 datasets. Examples from four of the datasets are provided in Table \ref{examplesofdata}.

\begin{table}\small\centering
\begin{tabular}{|ll|l|}
\hline
\textbf{(some $x$) mobile $\sqsupset$ (some $y$) mobile [s.t. $x\sqsupset y$]}&~&\textbf{(all $x$) French $\sqsubset$ (most $x$) European}		
\\\hline
(some dog) mobile	 $\sqsupset$  (some puppy) mobile	&&
(all puppy) French $\sqsubset$ (most puppy) European
\\
(some animal) mobile $\sqsupset$ (some cat) mobile&&
(all cat) French $\sqsubset$ (most cat) European 		\\

(some Asian) mobile $\sqsupset$ (some Thai) mobile&&(all hippo) French $\sqsubset$ (most hippo) European \\ ...&&...\\\hline
\end{tabular}
\caption{Sample datasets from two different templates.\label{examplesofdata}}
\end{table}

% TODO: Mention all seven relations seen, but some rare

% TODO: Condense the below to a paragraph

\noindent\textit{Basic monotonicity:} This set of datasets contains reasoning patterns like those seen in example 1  in Table \ref{examplesofdata}, where one of the predicates on one side entails the predicate in the corresponding position on the other side. In some of the datasets (as in 1), this alternation is in the first argument position, in some the second argument position, and in some both. For the alternating first argument subclasses, I have every lexical entailment pair in the data in the first argument position, the terms \ii{bark}, \ii{mobile}, and \ii{European} in the second argument position, and every quantifier. 
For alternating second argument datasets, I have all predicates in the first argument position, and the pairs \ii{bark--animate}, \ii{French--European}, and \ii{Parisian--French} in the second argument position. The datasets in which there is an alternation in both positions fall into two categories. In some, the entailment relations between the arguments work in the same direction (\ii{some dog French} $\sqsubset$ \ii{some animal European}), in others, they work in opposite directions (\ii{some dog European} \# \ii{some animal French}).

\noindent\textit{Quantifier substitution:} These datasets, exemplified in 2 above, show the interactions between each possible pair of different quantifiers, with the same predicates on both sides. Datasets exist with \ii{bark}, \ii{mobile}, and \ii{European} in the second argument position, and within each dataset every possible predicate is shown in first argument position.

\noindent\textit{Monotonicity with quantifier substitution:} These datasets show the more complex monotonicity interactions that emerge when reasoning between phrases with differing arguments and differing quantifiers, as in  3. 
Exhaustively creating datasets of this kind involves too large an amount of data to readily verify manually, so I sampled  from the extensive set of possible combinations of the six quantifiers and the fixed argument fillers used in the monotonicity datasets. Each possible relation but `$\equiv$' (which does not apply sentences like these unless the predicates on both sides are identical) is expressed, as is every possible pair of quantifiers.

\noindent\textit{Negation:}
To show the interaction of negation and monotonicity, I included variants of many of the datasets described above in which one of the four argument positions is systematically negated, as demonstrated in example 4.

\section{Experiments and results}

% TODO: Update with newer experiments

In the simplest experimental setting, which I label \textsc{all-split}, I randomly sample 85\% the data---making sure to sample 85\% of each of the individual datasets---train the model on that portion, and evaluate on the remaining 15\% of the data. This setting is meant to test whether the model is able to correctly generalize the individual reasoning patterns represented by each of the datasets.

Performance on this setting is perfect: the model quickly converges to 100\% accuracy on the test data, showing that it is capable of accurately learning to capture all of the reasoning patterns in the data. The remaining experimental settings serve to determine whether what is learned captures the underlying logical structure of the data in a way that allows it to accurately label unseen kinds of reasoning pattern. In each of them, I choose one of three arbitrarily chosen target datasets, all involving quantifier substitution, to test on. I then then hold out that dataset and---depending on the experimental setting---other similar datasets from the training data in an attempt to discover just how different a test example can be from anything seen in training and still be classified accurately. Table \ref{patterntable} shows what information is included in the training data for each of the four settings for one of the three target datasets.

\begin{table}\small\centering
\begin{tabular}{|l|ll|l|}\hline
Target dataset & Data evaluated   &  \textsc{subcl.-out} & \textsc{pair-out} \\\hline
\ii{(most $x$) bark $|$ (no $x$)} bark&target dataset only		&100\%&	93.6\%\\
\ii{(two $x$) bark \# (all $x$) bark}&target dataset only  			&100\%&	94.7\%\\
 \ii{(some $x$) bark $\natneg$ (no $x$) bark}&target dataset only 	&0\%&	0\%\\\hline
\end{tabular}
\caption{Test set accuracy.\label{resultstable}}
\end{table} % TODO: Replace

% TODO: Replace

\textsc{set-out:} In these experiments, I hold out the target dataset, training on none of it and testing on all of it, and additionally split each remaining dataset as in \textsc{all-split}. This setting is meant to test whether the model can generalize a broader reasoning pattern from one dataset to another---the model will still have seen other similar quantifier substitution datasets that differ from the target dataset only in which filler word is placed in the second argument position, as in row 5 in Table \ref{patterntable} above.

\textsc{pair-out:} In these experiments, I hold out all of the datasets that demonstrate the interaction between a particular pair of quantifiers. For the target dataset \ii{(most $x$) bark $|$ (no $x$) bark} in this setting, all examples with \ii{most} on one side and \ii{no} on the other, in either order, with or without negation or differing predicates, are held out from training. The remaining datasets are split as above.

These last three experiments have multiple sources of test data: some from the 15\% test portions of the datasets that were split, some from the target datasets, and some from any other similar datasets that were held out. In  Table \ref{resultstable}, I report results for each experiment on the entire test dataset, on the single target dataset, and on all of the held out datasets (i.e., the test data excluding the 15\% portions from the training datasets) in aggregate. This third figure is identical to the second for the \textsc{set-out} experiments, since the only held out dataset is the target dataset.

%\begin{table}\small\centering
%\begin{tabular}{|l|l|lll|}\hline
%Target dataset & Data evaluated   & \textsc{set-out} & \textsc{subcl.-out} & \textsc{pair-out} \\\hline
%\ii{(most $x$) bark $|$ (no $x$)} bark&target dataset only &100\%&100\%&0\%\\
%& all held out datasets                                 &(100\%) & 100\%& 49.4\%\\
%& all test data                                  &100\% & 99.9\%&85.7\%\\\hline
%\ii{(two $x$) bark \# (all $x$) bark}&target dataset only  &94.7\%&0\%&0\%\\
%& all held out datasets                                 &(94.7\%) & 70.0\%& 36.0\%\\
%& all test data                                  &98.9\% & 97.5\%&88.0\%\\\hline
% \ii{(some $x$) bark $\natneg$ (no $x$) bark}&target dataset only &2.6\%&0\%&0\%\\
%& all held out datasets                                 &(2.6\%) & 0\%& 17.6\%\\
%& all test data                                   &97.7\% & 93.7\%&83.9\%\\\hline
%\end{tabular}
%\caption{Test accuracy after 64 passes through the training data.\label{resultstable}}
%\end{table}

% The model did not converge well for any of these experiments: convergence can take hundreds or thousands of passes through the data, and the performance of the model at convergence  on test data was generally worse than its performance during the first hundred or so iterations. To sidestep this problem somewhat, I report results here for the models learned after 64 passes through the data. Even so, results are somewhat difficult to interpret: the performance on the test data fluctuates substantially between iterations, so a model that performs very poorly at iteration 64 may perform well at iteration 70. It is possible to stabilize the model by manipulating the hyperparameters away from their present tuned settings (raising minibatch size, shrinking model size, and increasing regularization, for example), but doing so consistently hurts performance.

Performance was perfect for at least some held out datasets in both the \textsc{set-out} and \textsc{subclass-out} settings, and near perfect for some in \textsc{pair-out}. Performance for one of the three target datasets---\ii{(some $x$) bark $\natneg$ (no $x$) bark}---was consistently poor across training settings. This poor performance is consistent across random initializations, and should be a target of further investigation. % TODO

\subsection{Discussion}

% TODO: Revise

The model learns to generalize well to novel data in most but not all of the training configurations. This inconsistent performance suggests that there is room to improve the optimization techniques used on the model, but the fact that it is able to perform well in these settings even some of the time suggests that the structure of the model is basically capable of learning meaning representations that support inference.

The results in the \textsc{all-split} condition show two important behaviors. The model is able to learn to identify the difference between two unseen sentences and consistently return the label that corresponds to that difference. In addition, the model can learn lexical relations from the training data, such as \ii{dog $\sqsubset$ animal} from \ii{(no dog) bark $\sqsupset$ (no animal) bark}, and it can then use these learned lexical relations to compute the relation for a sentence pair like \ii{(some dog) bark $\sqsubset$ (all animal) bark}. The results from the other three experimental settings show that the model is able to learn general purpose representations for quantifiers which, at least in many cases, allow it to perform inference when a crucial difference between two sentences---the  substitution of one specific quantifier for another---has not been seen. These results serve to confirm that the representations learned are capturing some of the underlying logic, rather than than just supporting the memorization of fixed reasoning patterns.

%  There has also been some research, most recently including \citet{grefenstette2013towards}, into models which represent quantifiers and other function words as higher order tensors rather than vectors, motivated in part by the hypothesis that their compositional behavior cannot be captured in strictly vector-based representatons. A successful inference engine that used those representations would not be a strict rebuttal of this assumption, but it could provide insight into precisely what circumstances would warrant the use of these richer tensor representation models in practice.  % TODO: Proof

These results leave open the question of how much information is minimally needed to learn general purpose representations for quantifiers in this setting. There are two lines of experimental work that could help to clarify this. Including longer sentences and constructions involving conjunctions (i.e. \ii{and, or}), transitive verbs (i.e. \ii{eats, kicks}) or other constructions in the training and test data could further test what kinds of behavior can be learned from a small training set, as could further experiments on this data involving even smaller training sets than those shown here or differently structured configurations of train and test sets.

%The representations that the model learns are not truly general-purpose, though. Human speakers of English who know the meanings of all of the words in a sentence are able to reason about whether that sentence entails any other sentence of English even if they have never seen either sentence before, or have never used the particular pattern of inference before. This broad ability to generalize does not appear to have been learned in these experiments. Pessimistically, the model might be primarily memorizing the reasoning patterns themselves, with any generalization beyond the \textsc{set-out} setting being erratic and case-specific. Optimistically, though, the model might be learning to handle logic in a way more like MacCartney's formal approach, in which the model must know dozens of different possible inference steps involving specific quantifiers and entailment patterns to succeed: if the model hasn't seen the substitution of \ii{most} with \ii{no}, it doesn't have the tools to reason about that substitution. 
%An incomplete ability to generalize is not necessarily detrimental to a model's ability to support reasoning in practice, especially in this later interpretation, since the body of knowledge used to do inference in MacCartney's approach is small enough that it could realistically be learned.

%It remains an open research question whether generalization of the type seen in the \textsc{pair-out} case is possible for any model. A model that could successfully generalize in this way could be potentially valuable both as a more versatile reasoning tool, and as a source of information about the ways in which functional types like quantifiers can be represented in vector space models.
%There are two ways that this problem can be addressed, one practical and one formal.

% An obvious direction for future work is to test for this behavior in more powerful models, or to see whether a dataset containing a more diverse array of reasoning patterns (beyond those possible just with quantification and negation) could encourage generalization. Running experiments like the ones presented above on models with matrices or higher order tensors as word representations \cite{coecke2010mathematical, grefenstette2013towards} might be a promising start.
%It is also not impossible that even an RNTN like the one studied could generalize perfectly: the dramatic instability of the model over the course of training suggests that better approaches to learning could improve performance.

### 2014-05-30 writing/S'14 paper/recursion.tex
Adding paper draft, cleaning up data dirs.

\section{Recursive structure}

\begin{figure}[t]
\begin{center}
\begin{tabular}{lll}
$a\equiv a$		&~~~&	$(c~(and~(not~d)))~\#~f$\\
$b~\#~c$			&~~~&	$(not~(c~(or~b)))~\sqsubset~(not~c)$\\
$d\natneg(not~d)$	&~~~&	$f~\#~((c~(or~(not~d)))~(and~a))$\\
$(c~(and~d))\sqsubset d$&~~~&$d\sqsupset((d~(or~d))~(and~(not~b)))$\\
\end{tabular}
\end{center}

\caption{Some sample randomly generated pairs of propositional logic statements.  \label{prop-figure}} 
\end{figure}

% TODO: Cite Chomsky/Hauser/Fitch?

Recursive structure is a prominent feature of natural language. Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice said hello}, and \ii{Carl thinks that Bob said that Alice said hello}. Overt recursion of this kind is easy to find, and theoretical accounts of natural language syntax and semantics rely heavily on recursive structures.
In order for a model to be able to accurately learn natural language meanings, then, we expect that it would need to be able to learn to represent the meanings of function words in a such a way that they are able to behave correctly when taking their own outputs as isput.

We again test this phenomenon within the framework of MacCartney and Manning-style entailment reasoning, but we replace the unanalyzed symbols from the previous experiment with expressions that involve recursive structure. To define these expressions, we turn to propositional logic, a relatively simple logic in which each variable represents either \ii{true} or \ii{false}. We generate data of the form seen in Figure \ref{prop-figure}: strings of arbitrary length consisting of six elementary proposition symbols and the operators \ii{and}, \ii{or}, and \ii{not}, arranged in pairs with the logical relations between them specified. It should be noted that we preserve the same  The data are generated randomly and deduplicated, and consist of about 248k training examples and 44k test examples. Each of the six symbols and each of the three operators is treated as a word for the purposes of our model, and is represented by a randomly initialized vector representation.

% TODO: Worth explicitly calling this project theorem proving?
Socher et al. \cite{socher2012semantic} have previously demonstrated the learning of a logic in a matrix-vector RNN model somewhat similar to our own, but the logic discussed here is substantially stronger, and a much better approximation of the kind of structure that is needed for natural language. The logic learned in that experiment is boolean, wherein the atomic symbols are simply the values 0 and 1, rather than variables over those values. While learning the operators of that logic is not trivial, the ouptuts of each operator can be represented accurately by a single bit. The statements of propositional logic learned here describe conditions on the truth values of propositions where those truth values are not known. As opposed to the two-way contrasts seen in \cite{socher2012semantic}, this logic distinguishes between 64 (2^6) possible assignments of truth values, and expressions of this logic define arbitrary conditions on these possible assignments, for a total of 2^{64} ($\approx 10^{20}$) possible statements that the recursive model needs to be able to distinguish.

...

% TODO: Add figure for results

train up to depth 5
test up to depth 12

% Describe data

### 2014-05-30 TrainModel.m

@@ -1,5 +1,5 @@
 % Want to distribute this code? Have other questions? -> sbowman@stanford.edu
-function TrainModel(dataflag, pretrainingFilename, expName, mbs, dim, lr, tot)
+function TrainModel(dataflag, pretrainingFilename, expName, mbs, dim, lr, lambda, tot)
 % The main training and testing script. The first arguments to the function
 % have been tweaked quite a few times depending on what is being tuned.
 
@@ -9,7 +9,7 @@ else
     expName = '.';
 end
 
-if ~strcmp(dataflag, 'and-or')
+if ~strcmp(dataflag, 'and-or') &&  ~strcmp(dataflag, 'and-or-deep')
     [wordMap, relationMap, relations] = ...
         LoadTrainingData('./wordpairs-v2.tsv');
 else
@@ -36,7 +36,7 @@ hyperParams.topDepth = 1;
 hyperParams.penultDim = 75;
 
 % Regularization coefficient.
-hyperParams.lambda = 0.002;
+hyperParams.lambda = lambda %0.002;
 
 % A vector of text relation labels.
 hyperParams.relations = relations;
@@ -62,7 +62,7 @@ hyperParams.datasetsPortion = 1;
 hyperParams.dataPortion = 1;
 
 hyperParams.useThirdOrder = tot;
-hyperParams.useThirdOrderComparison = 1;
+hyperParams.useThirdOrderComparison = tot;
 
 % Nonlinearities.
 hyperParams.compNL = @Sigmoid;
@@ -129,10 +129,38 @@ disp(options)
 
 % Choose which files to load in each category.
 listing = dir('data-4/*.tsv');
+listing5 = dir('data-5/*.tsv');
+
 splitFilenames = {listing.name};
 trainFilenames = {};
 testFilenames = {};
-if strcmp(dataflag, 'one-mn')
+
+if strcmp(dataflag, 'sub-tn')
+    testFilenames = {'MQ-three-no-bark.tsv', 'MQ-three-no-European.tsv', ...
+        'MQ-three-no-mobile.tsv'};
+    splitFilenames = {listing5.name};
+elseif strcmp(dataflag, 'pair-tn')
+    listing = [dir('data-5/*three-no*'); dir('data-5/*no-three*')];
+    testFilenames = {'MQ-three-no-bark.tsv', listing.name};
+    splitFilenames = {listing5.name};
[…]

### 2014-05-31 writing/S'14 paper/intro.tex
Sketch of alternative intro

@@
+\section{Introduction}\label{sec:intro}
+
+Supervised recursive neural network models (RNNs) for sentence meaning
+have been successful in a wide array of sophisticated language tasks,
+including sentiment analysis, analogy completion, image description,
+and paraphase detection. These results are encouraging about the
+ability of these models to learn compositional semantic grammars, but
+it remains an open question whether they can achieve the same results
+as grammars based in logical forms when it comes to core semantic
+concepts like quantification, entailment, and contradiction defined
+over complex linguistic structures. These concepts are central to
+robust natural language understanding, so it is essential that our
+computational models be able to capture them.
+
+We address this question directly by using a logical grammar to
+generate controlled data sets encoding the semantic relationships
+between pairs of expressions and evaluating whether each of two
+classes of neural model --- plain RNNs and recursive neural tensor
+networks (RNTNs) --- can learn those relationships correctly. Our
+logical grammar is a version of the natural logic developed by
+\cite{maccartney2009extended}, which defines seven core relations of
+synonymy, entailment, contradiction, and mutual consistency. Natural
+logic is well-suited to our purposes because of it defines a rich set
+of relations between words, phrases, and sentences, and because its
+logical expressions are derived directly from surface forms, making it
+amenable to large-scale natural language processing.
+
+In our first experiment, we show that both of these models can learn
+the core relational algebra of natural logic from data. Our second
+experiment builds on this result to cover relations between complex
+recursive structures like `(A or B)' and `not(not A and not B)', and
+our third experiment involves relations between quantified statements
+like `every reptile walks' and `every turtle moves'. We find that the
+plain RNN achieves only mixed results on the second and third
+experiments, but that the stronger RNTN models generalized well in
+every case, suggesting that it has in fact learned, or at least
+learned to simulate, our target logical concepts.
+
 \section{Introduction}

### 2014-06-01 writing/S'14 paper/quantifiers.tex
Bracketing bug fix, wrote up quantifier examples.

@@
 \begin{table}\small\centering
-\begin{tabular}{|ll|l|}
+\begin{tabular}{|l|}
 \hline
-\textbf{(some $x$) mobile $\sqsupset$ (some $y$) mobile [s.t. $x\sqsupset y$]}&~&\textbf{(all $x$) French $\sqsubset$ (most $x$) European}		
-\\\hline
-(some dog) mobile	 $\sqsupset$  (some puppy) mobile	&&
-(all puppy) French $\sqsubset$ (most puppy) European
-\\
-(some animal) mobile $\sqsupset$ (some cat) mobile&&
-(all cat) French $\sqsubset$ (most cat) European 		\\
-
-(some Asian) mobile $\sqsupset$ (some Thai) mobile&&(all hippo) French $\sqsubset$ (most hippo) European \\ ...&&...\\\hline
+(most warthogs) walk $\natneg$ (not-most warthogs) walk\\
+(most mammals) move \# (not-most (not turtles)) move\\
+(most (not pets)) (not swim) $\sqsupset$ (not-most (not pets)) move\\
+\hline
+(no turtles) (not growl) $\|$ (no turtles) (not swim)\\
+(no warthogs) swim $\sqsupset$ (no warthogs) move\\
+(no warthogs) move $\sqsubset$ (no (not reptiles)) swim\\
+\hline
 \end{tabular}
-\caption{Sample datasets from two different templates.\label{examplesofdata}}
+\caption{Sample data from two different quantifier pairs.\label{examplesofdata}}
 \end{table}
@@
 \begin{table}\small\centering
-\begin{tabular}{|l|ll|l|}\hline
-Target dataset & Data evaluated   &  \textsc{subcl.-out} & \textsc{pair-out} \\\hline
-\ii{(most $x$) bark $|$ (no $x$)} bark&target dataset only		&100\%&	93.6\%\\
-\ii{(two $x$) bark \# (all $x$) bark}&target dataset only  			&100\%&	94.7\%\\
- \ii{(some $x$) bark $\natneg$ (no $x$) bark}&target dataset only 	&0\%&	0\%\\\hline
+\begin{tabular}{|l|l|}\hline
+\textbf{Training configuration} & \textbf{Test accuracy}\\\hline
+\textsc{all split}	&	TODO\% \\\hline
+\textsc{pair no/no}	&	TODO\% \\
+\textsc{pair two/$<$two}	&	TODO\% \\
+\textsc{pair not-all/not-most}	&	TODO\% \\
+\textsc{pair all/some}	&	TODO\% \\\hline
 \end{tabular}
-\caption{Test set accuracy.\label{resultstable}}
+
+\caption{Quantifier experiment performance.\label{resultstable}}
 \end{table} % TODO: Replace
 
-% TODO: Replace
-
-\textsc{set-out:} In these experiments, I hold out the target dataset, training on none of it and testing on all of it, and additionally split each remaining dataset as in \textsc{all-split}.
-This setting is meant to test whether the model can generalize a broader reasoning pattern from one dataset to another---the model will still have seen other similar quantifier substitution datasets that differ from the target dataset only in which filler word is placed in the second argument position, as in row 5 in Table \ref{patterntable} above.
-
-\textsc{pair-out:} In these experiments, I hold out all of the datasets that demonstrate the interaction between a particular pair of quantifiers.
-For the target dataset \ii{(most $x$) bark $|$ (no $x$) bark} in this setting, all examples with \ii{most} on one side and \ii{no} on the other, in either order, with or without negation or differing predicates, are held out from training.
-The remaining datasets are split as above.
-
-These last three experiments have multiple sources of test data: some from the 15\% test portions of the datasets that were split, some from the target datasets, and some from any other similar datasets that were held out.
-In  Table \ref{resultstable}, I report results for each experiment on the entire test dataset, on the single target dataset, and on all of the held out datasets (i.e., the test data excluding the 15\% portions from the training datasets) in aggregate.
-This third figure is identical to the second for the \textsc{set-out} experiments, since the only held out dataset is the target dataset.
-
-%\begin{table}\small\centering
-%\begin{tabular}{|l|l|lll|}\hline
-%Target dataset & Data evaluated   & \textsc{set-out} & \textsc{subcl.-out} & \textsc{pair-out} \\\hline
-%\ii{(most $x$) bark $|$ (no $x$)} bark&target dataset only &100\%&100\%&0\%\\
-%& all held out datasets                                 &(100\%) & 100\%& 49.4\%\\
-%& all test data                                  &100\% & 99.9\%&85.7\%\\\hline
-%\ii{(two $x$) bark \# (all $x$) bark}&target dataset only  &94.7\%&0\%&0\%\\
-%& all held out datasets                                 &(94.7\%) & 70.0\%& 36.0\%\\
-%& all test data                                  &98.9\% & 97.5\%&88.0\%\\\hline
-% \ii{(some $x$) bark $\natneg$ (no $x$) bark}&target dataset only &2.6\%&0\%&0\%\\
-%& all held out datasets                                 &(2.6\%) & 0\%& 17.6\%\\
-%& all test data                                   &97.7\% & 93.7\%&83.9\%\\\hline
-%\end{tabular}
-%\caption{Test accuracy after 64 passes through the training data.\label{resultstable}}
-%\end{table}
-
-% The model did not converge well for any of these experiments: convergence can take hundreds or thousands of passes through the data, and the performance of the model at convergence  on test data was generally worse than its performance during the first hundred or so iterations.
-To sidestep this problem somewhat, I report results here for the models learned after 64 passes through the data.
-Even so, results are somewhat difficult to interpret: the performance on the test data fluctuates substantially between iterations, so a model that performs very poorly at iteration 64 may perform well at iteration 70.
-It is possible to stabilize the model by manipulating the hyperparameters away from their present tuned settings (raising minibatch size, shrinking model size, and increasing regularization, for example), but doing so consistently hurts performance.
-
-Performance was perfect for at least some held out datasets in both the \textsc{set-out} and \textsc{subclass-out} settings, and near perfect for some in \textsc{pair-out}.
-Performance for one of the three target datasets---\ii{(some $x$) bark $\natneg$ (no $x$) bark}---was consistently poor across training settings.
-This poor performance is consistent across random initializations, and should be a target of further investigation.
-% TODO
+% TODO: Define experimental setups
 

### 2014-06-01 writing/S'14 paper/intro.tex
Quantifier section update. Not done yet.

@@
 
-Deep learning methods in NLP which learn vector representations for words have seen successful uses in recent years on increasingly sophisticated tasks \cite{collobert2011natural, socher2011semi, socher2013acl1, chen2013learning}.
+% Deep learning methods in NLP which learn vector representations for words have seen successful uses in recent years on increasingly sophisticated tasks \cite{collobert2011natural, socher2011semi, socher2013acl1, chen2013learning}.
 Given the still modest performance of semantically rich NLP systems in many domains---question answering and machine translation, for instance---it is worth exploring the degree to which learned vectors can serve as general purpose semantic representations.
@@
 
-There are two broad family of tasks that could be used to test the ability of a model to develop general purpose meaning representations.
+% There are two broad family of tasks that could be used to test the ability of a model to develop general purpose meaning representations.
 In an interpretation task, sentences are mapped onto some denotation, such as  \ii{true} or \ii{false} for statements, or a factual answer for questions.
@@
 
-\begin{quote}
-\begin{enumerate}\small
-\item Many smartphone users avoid high bills overseas by disabling data service.
-\item Not everyone uses their smartphones for email when traveling abroad.
-\end{enumerate} 
-$\Rightarrow$ Entailment
-\end{quote}
+%\begin{quote}
+%\begin{enumerate}\small
+%\item Many smartphone users avoid high bills overseas by disabling data service.
+%\item Not everyone uses their smartphones for email when traveling abroad.
+%\end{enumerate} 
+%$\Rightarrow$ Entailment
+%\end{quote}
 
-In this paper, we test the ability of recursive models to on three simple tasks, each of which is meant to capture a property that is necessary in representing natural language meaning in the setting of inference.
+%In this paper, we test the ability of recursive models to on three simple tasks, each of which is meant to capture a property that is necessary in representing natural language meaning in the setting of inference.
 I begin with an overview of MacCartney and Manning's \cite{maccartney2009extended} framework for inference, and of the recursive neural networks that I study.

### 2014-06-01 writing/S'14 paper/quantifiers.tex
Quantifier section update. Not done yet.

@@
 
-% Condense condense condense
-
-Interest in NLI in the NLP community is ongoing, and a version of it has been included in the 2014 SemEval challenge specially targeted towards the evaluation of distributional models.
-Neither this dataset, nor any other existing RTE and NLI datasets are appropriate for the task at hand however.
-I intend for the present experiment to evaluate the ability of a class of models to learn certain types of inference behavior, and need a dataset that precisely tests these phenomena.
-With existing datasets that use unrestricted natural language, there is the risk that a model could successfully capture monotonicity inference, but fail to accurately label test data due to problems with, for example, lexical ambiguity, syntactic ambiguity, coreference resolution, or pragmatic enrichment.
+% Introduce the task
 
@@
 
-The dataset is built on a small purpose-built vocabulary, which is intended to be large and diverse enough to permit a wide variety of experiments on different semantic phenomena beyond those shown here, but which is still small enough to allow the relations between the lexical items themselves to be exhaustively manually labeled when needed.
-The vocabulary contains 41 predicates (nouns, adjectives, and intransitive verbs; see Appendix B for more on the design of the wordlist), six quantifiers, the logical conjunctions \emph{and} and \emph{or}, and the unary operator \emph{not}.
+Our data consists of pairs of sentences generated from a small artificial grammar.
+Each sentence contains a quantifier, a noun, which may be negated, and an intransitive verb which may be negated.
+We use the basic quantifiers \ii{some}, \ii{most}, \ii{all}, \ii{two}, and \ii{three}, and each of their duals over negation \ii{no}, \ii{not-all}, \ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}.
+We also include five nouns, four intransitive verbs, and the negation symbol \ii{not}.
+In order to be able to define relations between sentences with differing lexical items, we define the lexical relations between each noun--noun pair, each verb--verb pair, and each quantifier--quantifier pair.
 
-The data take the form of 12k pairs of short sentences, each annotated with a relation label, which I divide into about 200 smaller datasets.
-These datasets contain tightly constrained variants of a specific valid reasoning pattern, each differing in at most one lexical item, and all sharing the same relation label.
-The remainder of this section describes the composition of these 200 datasets.
-Examples from four of the datasets are provided in Table \ref{examplesofdata}.
+%nouns = ['warthogs', 'turtles', 'mammals', 'reptiles', 'pets']
+%verbs = ['walk', 'move', 'swim', 'growl']
+%dets = ['all', 'not_all', 'some', 'no', 'most', 'not_most', 'two', 'lt_two', 'three', 'lt_three']
+%adverbs = ['', 'not']
+
+To assign relation labels to sentence pairs, we built a small task-specific implemenation of MacCartney's logic that can accurately label sentences of this restricted language.
+The logic is not able to derive all intuitively true relations of this language, and fails to derive a single unique relation for certain types of statement, including De Morgean's laws (e.g.
+\ii{(all pets) growl $\natneg$ (some pet) (not growl)}), and we simply discard these examples.
+Exhaustively generating the valid sentences under this grammar and choosing those to which a relation label can be assigned
+yields 66k sentence pairs. Some examples of these data are provided in Table \ref{examplesofdata}.
 
@@
 \hline
-(most warthogs) walk $\natneg$ (not-most warthogs) walk\\
-(most mammals) move \# (not-most (not turtles)) move\\
-(most (not pets)) (not swim) $\sqsupset$ (not-most (not pets)) move\\
+\ii{(most warthogs) walk $\natneg$ (not-most warthogs) walk}\\
+\ii{(most mammals) move $\#$ (not-most (not turtles)) move}\\
+\ii{(most (not pets)) (not swim) $\sqsupset$ (not-most (not pets)) move}\\
 \hline
-(no turtles) (not growl) $\|$ (no turtles) (not swim)\\
-(no warthogs) swim $\sqsupset$ (no warthogs) move\\
-(no warthogs) move $\sqsubset$ (no (not reptiles)) swim\\
+\ii{(no turtles) (not growl) $\|$ (no turtles) (not swim)}\\
+\ii{(no warthogs) swim $\sqsupset$ (no warthogs) move}\\
+\ii{(no warthogs) move $\sqsubset$ (no (not reptiles)) swim}\\
 \hline
 \end{tabular}
-\caption{Sample data from two different quantifier pairs.\label{examplesofdata}}
+\caption{Sample data involving two different quantifier pairs.\label{examplesofdata}}
 \end{table}
@@
 % TODO: Mention all seven relations seen, but some rare
-
-% TODO: Condense the below to a paragraph
-
-\noindent\textit{Basic monotonicity:} This set of datasets contains reasoning patterns like those seen in example 1  in Table \ref{examplesofdata}, where one of the predicates on one side entails the predicate in the corresponding position on the other side.
-In some of the datasets (as in 1), this alternation is in the first argument position, in some the second argument position, and in some both.
-For the alternating first argument subclasses, I have every lexical entailment pair in the data in the first argument position, the terms \ii{bark}, \ii{mobile}, and \ii{European} in the second argument position, and every quantifier.
-For alternating second argument datasets, I have all predicates in the first argument position, and the pairs \ii{bark--animate}, \ii{French--European}, and \ii{Parisian--French} in the second argument position.
-The datasets in which there is an alternation in both positions fall into two categories.
-In some, the entailment relations between the arguments work in the same direction (\ii{some dog French} $\sqsubset$ \ii{some animal European}), in others, they work in opposite directions (\ii{some dog European} \# \ii{some animal French}).
-
-\noindent\textit{Quantifier substitution:} These datasets, exemplified in 2 above, show the interactions between each possible pair of different quantifiers, with the same predicates on both sides.
-Datasets exist with \ii{bark}, \ii{mobile}, and \ii{European} in the second argument position, and within each dataset every possible predicate is shown in first argument position.
-
-\noindent\textit{Monotonicity with quantifier substitution:} These datasets show the more complex monotonicity interactions that emerge when reasoning between phrases with differing arguments and differing quantifiers, as in  3.
-Exhaustively creating datasets of this kind involves too large an amount of data to readily verify manually, so I sampled  from the extensive set of possible combinations of the six quantifiers and the fixed argument fillers used in the monotonicity datasets.
-Each possible relation but `$\equiv$' (which does not apply sentences like these unless the predicates on both sides are identical) is expressed, as is every possible pair of quantifiers.
-
-\noindent\textit{Negation:}
-To show the interaction of negation and monotonicity, I included variants of many of the datasets described above in which one of the four argument positions is systematically negated, as demonstrated in example 4.
 
@@
 \textsc{pair no/no}	&	TODO\% \\
-\textsc{pair two/$<$two}	&	TODO\% \\
+\textsc{pair two/less-than-two}	&	TODO\% \\
 \textsc{pair not-all/not-most}	&	TODO\% \\

### 2014-06-02 writing/S'14 paper/abstract.tex
Incremental paper update.

@@
 structures and sentences involving quantification. We find that the
-plain RNN achieves only mixed results on the second and third
+plain RNN achieves only mixed results on the first and second
 experiments, but that the stronger RNTN models generalized well in
-every case.
+every case. % TODO: Update to reflect results
 \end{abstract}

### 2014-06-02 writing/S'14 paper/methods.tex
Incremental paper update.

@@
-\section{Methods}
+\section{Recursive neural network models}
 
-The model that I study is a recursive neural network model \citet{socher2011dynamic}, and is thus centered on a recursively applied composition function which is meant to mimic the recursive construction of meanings in formal models of semantics.
+The models that we study is are centered aronund a recursively applied composition function which is meant to mimic the recursive construction of meanings in formal semantics.
 In this scheme, pairs of words are merged into phrase representations by a function that maps from representations of length $2N$ to representations of length $N$.
@@
 Borrowing a model directly from the existing literature for this task is impossible since none has been proposed to detect asymmetric relations between phrases.
-Instead, I build a combination model, depicted in Figure \ref{sample-figure}.
+Instead, we build a combination model, depicted in Figure \ref{sample-figure}.
 The two phrases being compared are built up separately on each side of the tree using the same composition function until they have each been reduced to single vectors.
@@
 
-For a composition layer, I experiment with both a plain neural network layer function (eq.
+For a composition layer, we experiment with both a plain neural network layer function (eq.
 \ref{rnn}) and the RNTN layer function proposed in \citet{chen2013learning} (eq.
@@
 A sigmoid nonlinearity (elementwise $\tanh$) is applied to the output of either layer function, following Socher et al.
-
- \label{rnn}\label{rntn} % TODO: Fix labeling
-\begin{equation}
-y_{RNN} = f(M [\vec{x}^{(l)}; \vec{x}^{(r)}] + b_i)\\
-y_{RNTN}^i = f(\vec{x}^{(l)T} \mathbf{A^{[i]}} \vec{x}^{(r)} + \vec{B_{i,:}} [\vec{x}^{(l)}; \vec{x}^{(r)}] + c_i)
-\end{equation}
+\begin{gather} \label{rnn}
+y_{RNN} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + b_i)\\ % TODO: Add column vectors?
+\label{rntn}
+y_{RNTN} = f(\vec{x}^{(l)T} \mathbf{A}^{[1...N]} \vec{x}^{(r)} + \mathbf{B} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
+\end{gather} % TODO: Explain third order tensor parameter?
 
@@
 
-\textit{Model size:} The dimensionalities of the model were tuned extensively.
-The performance of the model is approximately optimal with the dimensionality of the word vectors set to 16, and the dimensionality of the feature vector produced by the comparison layer set to 45.
-% TODO: Later
-
 \begin{figure}
 \begin{center}
-\Tree [.{\sc softmax classifier}  [.{\sc comparison layer} [.{\sc \ii{dog} vector from $V$} ] [.{\sc composition layer} [.{\sc composition layer} [.{\sc \ii{very} vector from $V$} ] [.{\sc \ii{big} vector from $V$} ] ] [.{\sc \ii{cat} vector from $V$} ] ] ] ]
+% \Tree [.{\sc softmax classifier}  [.{\sc comparison layer} [.{\sc \ii{dog} vector from $V$} ] [.{\sc composition layer} [.{\sc composition layer} [.{\sc \ii{very} vector from $V$} ] [.{\sc \ii{big} vector from $V$} ] ] [.{\sc \ii{cat} vector from $V$} ] ] ] ]
+\includegraphics[scale=0.36]{model.eps}
 \end{center}
-\caption{The model structure used to compare \ii{dog} and \ii{(very big) cat}. \label{sample-figure}} 
+\caption{The model structure used to compare \ii{no dog} and \ii{(not all) dog}. \label{sample-figure}} 
 \end{figure} % TODO: Create a more compact figure using something better than qtree.

### 2014-06-02 writing/S'14 paper/quantifiers.tex
Incremental paper update.

@@
 
-% Introduce the task
+Even to the extent that RNTN models can handle functional meanings in the form of the operators of propositional logic, this is not a guarantee that these models can handle functional meanings of the forms seen in natural language.
+As a first step towards investigating the latter, we attempt to directly measure the degree to which RNNs are able to develop representations for natural language quantifiers like \ii{some} and \ii{all} that are adequate for inference.
+Quantification is far from the only place in natural language where complex functional meanings are found, but it is a natural starting point, since it can be tested in setences whose structures are otherwise quite simple, and since it has formed a standard case study in prior formal work on natural language inference.
 
-\subsection{Data}
+% \subsection{Data}
+
+This experiment replicates similar work described in \citet{bowman2013can}, but replaces the partially manually annotated data in that paper with data that is generated directly using the logical system that we hope to model, yielding results that we believe to be substantially more straightforward to interpret.
 
@@
 
-% TODO: Mention all seven relations seen, but some rare
+% \subsection{Experiments and results}
 
-\section{Experiments and results}
+We evaluate the model using two experimental settings.
+In the simpler setting we randomly sample 85\% of the data and evaluate on the remaining 15\%.
+In this setting, the model is being asked to learn a complete reasoning system for the limited language and logic presented in the training data, but it is not being asked to generalize to test examples that are substantially different from those it was trained on.
+Crucially though, to succeed on this task, the model must be able to recognize all of the lexical relations between the nouns, verbs, and quantifiers and how they interact.
 
-% TODO: Update with newer experiments
-
-In the simplest experimental setting, which I label \textsc{all-split}, I randomly sample 85\% the data---making sure to sample 85\% of each of the individual datasets---train the model on that portion, and evaluate on the remaining 15\% of the data.
-This setting is meant to test whether the model is able to correctly generalize the individual reasoning patterns represented by each of the datasets.
-
-Performance on this setting is perfect: the model quickly converges to 100\% accuracy on the test data, showing that it is capable of accurately learning to capture all of the reasoning patterns in the data.
-The remaining experimental settings serve to determine whether what is learned captures the underlying logical structure of the data in a way that allows it to accurately label unseen kinds of reasoning pattern.
-In each of them, I choose one of three arbitrarily chosen target datasets, all involving quantifier substitution, to test on.
-I then then hold out that dataset and---depending on the experimental setting---other similar datasets from the training data in an attempt to discover just how different a test example can be from anything seen in training and still be classified accurately.
-Table \ref{patterntable} shows what information is included in the training data for each of the four settings for one of the three target datasets.
+While our primary interest is in discovering the extent to which the model can learn to encode the logic given an arbitrary amount of data, we are also interested in the degree to which the model can infer a correct representation from the logic from more constrained training data.
+In particular, we propose to segment the sentence pairs according to which quantifiers appear in each pair, and then hold out one such pair for testing.
+We hypothesize that a model that can efficiently learn to represent a logic should be able to construct an accurate representation of each held out quantifier from the way that it interacts with the other nine quantifiers which are not held out.
+Since running this experiment requires choosing a pair of quantifiers to hold out before training, the resource demans of training prevent us from testing each of the 55 possible possible pairs of quantifiers, and we choose only four pairs to test on.
+We attempt choose three pairs of differing quantifiers (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and \ii{all}/\ii{some}) to represent six different quantifiers in the test data, choosing pairs that maximize the diversity of relation labels appearing in the test data, and additionally choose one pair in which a quantifier is paired with itself (\ii{no}/\ii{no}).
 
@@
 \textsc{all split}	&	TODO\% \\\hline
-\textsc{pair no/no}	&	TODO\% \\
 \textsc{pair two/less-than-two}	&	TODO\% \\
 \textsc{pair not-all/not-most}	&	TODO\% \\
-\textsc{pair all/some}	&	TODO\% \\\hline
+\textsc{pair all/some}	&	TODO\% \\
+\textsc{pair no/no}	&	TODO\% \\
+\hline
 \end{tabular}
@@
 
-% TODO: Define experimental setups
+% \subsection{Discussion}
 
-\subsection{Discussion}
+The RNN model was approximately optimal with N dimensional word representations and an M dimensional comparison layer. The RNTN was approximately optimal with N and M dimensions, respectively.
+% TODO: Update dimensionality
+
+To be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...
 
 % TODO: Revise
-
-The model learns to generalize well to novel data in most but not all of the training configurations.
-This inconsistent performance suggests that there is room to improve the optimization techniques used on the model, but the fact that it is able to perform well in these settings even some of the time suggests that the structure of the model is basically capable of learning meaning representations that support inference.
-
-The results in the \textsc{all-split} condition show two important behaviors.
-The model is able to learn to identify the difference between two unseen sentences and consistently return the label that corresponds to that difference.
-In addition, the model can learn lexical relations from the training data, such as \ii{dog $\sqsubset$ animal} from \ii{(no dog) bark $\sqsupset$ (no animal) bark}, and it can then use these learned lexical relations to compute the relation for a sentence pair like \ii{(some dog) bark $\sqsubset$ (all animal) bark}.
-The results from the other three experimental settings show that the model is able to learn general purpose representations for quantifiers which, at least in many cases, allow it to perform inference when a crucial difference between two sentences---the  substitution of one specific quantifier for another---has not been seen.
-These results serve to confirm that the representations learned are capturing some of the underlying logic, rather than than just supporting the memorization of fixed reasoning patterns.
-
-%  There has also been some research, most recently including \citet{grefenstette2013towards}, into models which represent quantifiers and other function words as higher order tensors rather than vectors, motivated in part by the hypothesis that their compositional behavior cannot be captured in strictly vector-based representatons.
-A successful inference engine that used those representations would not be a strict rebuttal of this assumption, but it could provide insight into precisely what circumstances would warrant the use of these richer tensor representation models in practice.
-% TODO: Proof
-
-These results leave open the question of how much information is minimally needed to learn general purpose representations for quantifiers in this setting.
-There are two lines of experimental work that could help to clarify this.
-Including longer sentences and constructions involving conjunctions (i.e.
-\ii{and, or}), transitive verbs (i.e.
-\ii{eats, kicks}) or other constructions in the training and test data could further test what kinds of behavior can be learned from a small training set, as could further experiments on this data involving even smaller training sets than those shown here or differently structured configurations of train and test sets.
-
-%The representations that the model learns are not truly general-purpose, though.
-Human speakers of English who know the meanings of all of the words in a sentence are able to reason about whether that sentence entails any other sentence of English even if they have never seen either sentence before, or have never used the particular pattern of inference before.
-This broad ability to generalize does not appear to have been learned in these experiments.
-Pessimistically, the model might be primarily memorizing the reasoning patterns themselves, with any generalization beyond the \textsc{set-out} setting being erratic and case-specific.
-Optimistically, though, the model might be learning to handle logic in a way more like MacCartney's formal approach, in which the model must know dozens of different possible inference steps involving specific quantifiers and entailment patterns to succeed: if the model hasn't seen the substitution of \ii{most} with \ii{no}, it doesn't have the tools to reason about that substitution.
-%An incomplete ability to generalize is not necessarily detrimental to a model's ability to support reasoning in practice, especially in this later interpretation, since the body of knowledge used to do inference in MacCartney's approach is small enough that it could realistically be learned.
-
-%It remains an open research question whether generalization of the type seen in the \textsc{pair-out} case is possible for any model.
-A model that could successfully generalize in this way could be potentially valuable both as a more versatile reasoning tool, and as a source of information about the ways in which functional types like quantifiers can be represented in vector space models.
-%There are two ways that this problem can be addressed, one practical and one formal.
-
-% An obvious direction for future work is to test for this behavior in more powerful models, or to see whether a dataset containing a more diverse array of reasoning patterns (beyond those possible just with quantification and negation) could encourage generalization.
-Running experiments like the ones presented above on models with matrices or higher order tensors as word representations \cite{coecke2010mathematical, grefenstette2013towards} might be a promising start.
-%It is also not impossible that even an RNTN like the one studied could generalize perfectly: the dramatic instability of the model over the course of training suggests that better approaches to learning could improve performance.

### 2014-06-02 writing/S'14 paper/recursion.tex
Incremental paper update.

@@
 Overt recursion of this kind is easy to find, and theoretical accounts of natural language syntax and semantics rely heavily on recursive structures.
-In order for a model to be able to accurately learn natural language meanings, then, we expect that it would need to be able to learn to represent the meanings of function words in a such a way that they are able to behave correctly when taking their own outputs as isput.
+In order for a model to be able to accurately learn natural language meanings, then, we expect that it would need to be able to learn to represent the meanings of function words in a such a way that they are able to behave correctly when taking their own outputs as input.
+
+% Goal: Deep recursion!
 
@@
 To define these expressions, we turn to propositional logic, a relatively simple logic in which each variable represents either \ii{true} or \ii{false}.
-We generate data of the form seen in Figure \ref{prop-figure}: strings of arbitrary length consisting of six elementary proposition symbols and the operators \ii{and}, \ii{or}, and \ii{not}, arranged in pairs with the logical relations between them specified.
-It should be noted that we preserve the same  The data are generated randomly and deduplicated, and consist of about 248k training examples and 44k test examples.
-Each of the six symbols and each of the three operators is treated as a word for the purposes of our model, and is represented by a randomly initialized vector representation.
+We generate data of the form seen in Figure \ref{prop-figure}: strings of arbitrary length consisting of six elementary proposition variables and the operators \ii{and}, \ii{or}, and \ii{not}, arranged in pairs with the logical relations between them specified.
 
-% TODO: Worth explicitly calling this project theorem proving?
+% TODO: Worth explicitly calling this project theorem proving? Yes! There was some confusion at CSLI.
+
 Socher et al.
@@
 As opposed to the two-way contrasts seen in \cite{socher2012semantic}, this logic distinguishes between 64 (2^6) possible assignments of truth values, and expressions of this logic define arbitrary conditions on these possible assignments, for a total of 2^{64} ($\approx 10^{20}$) possible statements that the recursive model needs to be able to distinguish.
+To frame this distinction in another way, the relational statements in our data our theorems about the relations between statements in the logic tested in \cite{socher2012semantic}.
+
+We randomly generate pairs of parentheses-bracketed statements of the logic and then randomly divide the results into training and test data sets.
+To compute the relation between each pair of statements, we exhaustively enumerate the sets of assignments of truth values to proposition variables that would satisfy each of the statements and then convert the set-theoretic relation between those assignments into one of the seven relations.
+If we do not implement any constraint that the two statements being compared are similar in any way, the generated data consists in large part of statements in which the two refer to largely separate subsets of the six variables, and to which we will nearly always assign the \# relation.
+In an effort to balance the distribution of relation labels without departing from the basic task of modeling propositional logic, we disallow individual pairs of statements from referring to more than four of the six proposition variables.
+
+We deduplicate the data and discard pairs in which either statement is a tautology or contradiction (a statement that is true of either all or no possible assignments), for which none of the seven relation labels can accurately apply.
+We then divide the generated pairs into size bins based on the number of logical operators (\ii{and}, \ii{or}, or \ii{not}) in the larger of the two pairs being compared, and discard examples of size greater than twelve by this measure.
+Finally, we randomly sample 15\% of each bin for a held out test set.
+...
+65k train
+
+ compute the relaitons between them, discard  , and consist of about 248k training examples and 44k test examples.
+
+ Each of the six symbols and each of the three operators is treated as a word for the purposes of our model, and is represented by a randomly initialized vector representation.
 
@@
 % Describe data
+
+\begin{figure}[t]
+\begin{center}
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+\end{center}
+
+\caption{Model performance by expression size. The vertical dashed line indicates the size of the largest expressions included in the training data.  \label{prop-figure}} 
+\end{figure}
+
+The RNN model was approximately optimal with N dimensional word representations and an M dimensional comparison layer. The RNTN was approximately optimal with N and M dimensions, respectively.
+
+% TODO: Include dimensionality

### 2014-06-02 writing/S'14 paper/discussion.tex
Incremental paper update.
Adding Monday morning results to paper.

@@
 
-1 Para
+Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...
 
-- capable of a geometric encoding of fundamentals of logic
-
-- capable of capturing recursive structure up to a scale comparable to what's used in language, and of true recursion
-
-- capable of learning functional word meanings given adequate data, frontiers: more complex cases
+% TODO: One para of closing discussion.
+% - Results strongly suggest that RNTNs can encode the basics of this logic. First such result.
+% -- Decay over size in recursion results not ideal, but this kind of behavior may not be problematic since there is a practical bound on the length of natural language sentences.
+% - Future work: 
+% -- Harder problems
+% --- Sentences with more types of structure: Transitive verbs, relative clauses, etc.
+% -- More models
+% --- BP-RNNs, conv-RNNs

### 2014-06-02 writing/S'14 paper/join.tex
Incremental paper update.
Adding Monday morning results to paper.

@@
 
- A TODO-dimensional RNTN model tuned for the task was able to correctly label TODO \% of the derivable test examples, and TODO \% of the remaining examples.
-The simpler RNN model with the same number of dimensions was able to derive only TODO \% and TODO \%, respectively.
+We found that the RNTN model worked best with 11 dimensional vector representations for the 80 sets and a 90 dimensional feature vector for the classifier.
+This model was able to correctly label 99.3\% of the derivable test examples, and 99.1\% of the remaining examples.
+The simpler RNN model worked best with 11 and 75 dimensions, respectively, but was able to achieve accuracies of only 90.0\% and 87.\%, respectively.
+
+Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...
+
+% Points for discussion: 
+% - Fairly simple interpretation: RNTN generalized, RNN didn't
+% - Future work: Can this generalize to models with huge numbers of entities, or does this result depend on the examples reflecting a small model in any way?
+Can't exhaustively generate large models without getting \#s - need sampling.

### 2014-06-02 writing/S'14 paper/quantifiers.tex
Adding Monday morning results to paper.

@@
 
-This experiment replicates similar work described in \citet{bowman2013can}, but replaces the partially manually annotated data in that paper with data that is generated directly using the logical system that we hope to model, yielding results that we believe to be substantially more straightforward to interpret.
+This experiment replicates similar work described in \citet{bowman2013can}, which found that RNTNs can learn to reason well with quantifier meanings given sufficient training data.
+This paper replaces the partially manually annotated data in that paper with data that is generated directly using the logical system that we hope to model, yielding results that we believe to be substantially more straightforward to interpret.
 
@@
 
-% \subsection{Experiments and results}
-
 We evaluate the model using two experimental settings.
@@
 \begin{table}\small\centering
-\begin{tabular}{|l|l|}\hline
-\textbf{Training configuration} & \textbf{Test accuracy}\\\hline
-\textsc{all split}	&	TODO\% \\\hline
-\textsc{pair two/less-than-two}	&	TODO\% \\
-\textsc{pair not-all/not-most}	&	TODO\% \\
-\textsc{pair all/some}	&	TODO\% \\
-\textsc{pair no/no}	&	TODO\% \\
+\begin{tabular}{|l|lll|}\hline
+\textbf{Data} & \textbf{Most frequent class*} & \textbf{TODO dim RNN} & \textbf{16 dim RNTN}\\\hline
+\textsc{all split}	& 35.4\% &	TODO\%&	88.1\% \\\hline
+\textsc{pair two/less-than-two}	& 59.8\% &	TODO\% &	92.4\% \\
+\textsc{pair not-all/not-most}	&0\% &	   TODO\%  &	82.1\% \\
+\textsc{pair all/some}	& 0\%& TODO\%  &	80.4\% \\
+\textsc{pair no/no}	& 30.8\% &	TODO\% &	100\% \\
 \hline
@@
 
-\caption{Quantifier experiment performance.\label{resultstable}}
+\caption{Quantifier experiment performance. *Most frequent class accuracy is measured against the most frequent class in the training data, \#.\label{resultstable}}
 \end{table} % TODO: Replace
-
-% \subsection{Discussion}
 
@@
 
-To be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...
+Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...
+
+% Notes for discussion:
+% - First effort to learn a logic, results somewhat unclear.
+% -- No straightforward way to prove that there is enough information in the data to learn quantifier projecitivies in this setting
+% - Perfect performance on no-no promising that the model is learning at least some structure
 

### 2014-06-02 writing/S'14 paper/recursion.tex
Adding Monday morning results to paper.

@@
 In order for a model to be able to accurately learn natural language meanings, then, we expect that it would need to be able to learn to represent the meanings of function words in a such a way that they are able to behave correctly when taking their own outputs as input.
-
-% Goal: Deep recursion!
+In evaluating the model, we take advantage of the fact that recursive structures of this kind define potentially infinite languages by testing the model on strings that are longer and more complex than any seen in testing.
 
@@
 
-% TODO: Worth explicitly calling this project theorem proving? Yes! There was some confusion at CSLI.
+% NOTE: Worth explicitly calling this project theorem proving? Yes! There was some confusion at CSLI.
 
@@
 Finally, we randomly sample 15\% of each bin for a held out test set.
-...
-65k train
 
- compute the relaitons between them, discard  , and consist of about 248k training examples and 44k test examples.
+We trained both the RNN and RNTN models on the data of size four or less (65k pairs), and tested it on examples of up to size 12 (44k pairs).
+We initialized the model parameters randomly, including the vector representations of the six variables.
 
- Each of the six symbols and each of the three operators is treated as a word for the purposes of our model, and is represented by a randomly initialized vector representation.
-
-...
-
-% TODO: Add figure for results
-
-train up to depth 5
-test up to depth 12
-
-% Describe data
+The results are shown in Figure \ref{prop-results}.
+In tuning, we found that the RNN model was approximately optimal with 45 dimensional vector representations, and the RNTN model was approximately optimal with 25 dimensions.
+We fixed the size of the feature vector for the classifier at 75 dimensions.
+We found that the RNTN model was able to perform almost perfectly on unseen small test examples, with accuracy above 99\% below size four.
+After depth four, performance gradually falls with increasing size.
+The RNN model did not perform well, reaching only 88.2\% accuracy on the smallest test examples, and declining from there to near-baseline performance at size 12.
 
@@
 \begin{center}
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
+\includegraphics[width=5.5in]{recursion\string_results.eps}
 \end{center}
 
-\caption{Model performance by expression size. The vertical dashed line indicates the size of the largest expressions included in the training data.  \label{prop-figure}} 
+\caption{Model performance by expression size.  \label{prop-results}} 
 \end{figure}
 
-The RNN model was approximately optimal with N dimensional word representations and an M dimensional comparison layer. The RNTN was approximately optimal with N and M dimensions, respectively.
+Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...
 
-% TODO: Include dimensionality
+% TODO:
+% Points for discussion
+% - The RNTN has learned to approximate the correct funtionn
+% -- The error introduced by the approximation likely grows with depth, but is sufficiently small at the training example sizes
+
+% - The RNN has plausibly learned an approximation as well, but one which is quite noisy even without recursion
+% -- Possible but unlikely that further tuning and different types or amounts of training data could get the RNN to work
+
+% - Future work:
+% -- Generate much larger examples to tease apart the ability of the model to approximate a representable function from data, and the ability of the model to represent

### 2014-06-03 writing/S'14 paper/discussion.tex
Updating paper.

@@
-\subsection{General discussion}  % TODO
+\section{General discussion}
 
-Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...
+These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators.
+This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic.
+The positive results that we have fonud, however, are extremely promising for the future of learned representation models in the applied modeling of meaning.
+We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators.
+We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.
+
+There is ample room to build on these results.
+In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses.
+In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.
 

### 2014-06-03 writing/S'14 paper/intro.tex
Incremental paper update.
Adding Monday morning results to paper.
Updating paper.

@@
 have been successful in a wide array of sophisticated language tasks,
-including sentiment analysis, analogy completion, image description,
-and paraphase detection. These results are encouraging about the
+including sentiment analysis \cite{socher2011semi}, analogy completion, image description,
+and paraphase detection . These results are encouraging about the
 ability of these models to learn compositional semantic grammars, but
@@
 robust natural language understanding, so it is essential that our
-computational models be able to capture them.
+computational models be able to capture them. % TODO: Insert more citations for past RNN work.
 
-We address this question directly by using a logical grammar to
-generate controlled data sets encoding the semantic relationships
-between pairs of expressions and evaluating whether each of two
-classes of neural model --- plain RNNs and recursive neural tensor
-networks (RNTNs) --- can learn those relationships correctly. Our
-logical grammar is a version of the natural logic developed by
-\cite{maccartney2009extended}, which defines seven core relations of
-synonymy, entailment, contradiction, and mutual consistency. Natural
-logic is well-suited to our purposes because of it defines a rich set
-of relations between words, phrases, and sentences, and because its
-logical expressions are derived directly from surface forms, making it
-amenable to large-scale natural language processing.
+We address this question directly by using a logical grammar togenerate controlled data sets encoding the semantic relationships between pairs of expressions and evaluating whether each of twoclasses of neural model --- plain RNNs and recursive neural tensor networks (RNTNs, \cite{socher2013acl1}) --- can learn those relationships correctly.
+Following \citet{bowman2013can}, we use the natural logic developed by\cite{maccartney2009extended} as our formal model, which defines seven core relations ofsynonymy, entailment, contradiction, and mutual consistency.
+Natural logic is well-suited to our purposes because of it defines a rich setof relations between words, phrases, and sentences, and because its logical expressions are derived directly from surface forms, making itamenable to large-scale natural language processing.
+% Is this the right place for the first self citation?
 
@@
 every case, suggesting that it has in fact learned, or at least
-learned to simulate, our target logical concepts.
+learned to simulate, our target logical concepts. % TODO: Update with current results
 
-\section{Introduction}
-% TODO: cite \cite{watanabe2012latent} for past work on ML for McC&M-style NLI
+Citations to additional past work to be added.\\...\\...\\...\\...\\...
+
+% TODO: Mention earlier work on NN interpretation - \cite{Garcez-etal:2001}
 
@@
 This paper explores the ability of linguistic representations developed using supervised deep learning techniques to support interpretation and reasoning.
-
-% TODO: Name RNTNs
 
@@
 
-% TODO: Cite ICLR paper, emphasize new work since
+\subsection{Natural language inference and natural logic}
 
-\subsection{The task: natural language inference}
+In its simplest form, \ii{natural language inference} (also known as \ii{recognizing textual entailment}, as in \cite{dagan2006pascal}) is the task of determine whether one sentence entails another.
+Much of the theoretical work on this task (and some successful implemented models \cite{maccartney2009natural, watanabe2012latent}) involve \ii{natural logic}, formal systems that define sound rules of inference from one sentence of natural language to another without the need for intermediate representations in some other logic.
+The most powerful model that we are aware of for natural logic is due to MacCartney and Manning \cite{maccartney2009extended} and Icard \cite{icard2012inclusion}, and is centered around the definition of a set of seven logical relations which can hold between sentences, shown in Table \ref{b-table}.
 
-% Condense?
-In the standard formulation of the inference task (and the one used in the RTE datasets \cite{dagan2006pascal}), the goal is to determine whether a reasonable human would infer a hypothesis from a premise.
-MacCartney formalizes a method of inferring entailment relations, and moves past two way entailment/non-entailment classification, proposing the set $\mathfrak{B}$ of seven labels meant to describe all of the possible non-trivial relations that might hold between pairs of statements, shown in Table \ref{b-table}.
+This approach to natural logic can capture much of the complexity of natural language meaning within a well understood framework, and is also fairly straightforward to implement in a machine learning setting since it can be reduced to a seven-way classification problem on sentence pairs.
+Our goal in this paper is to learn recursive neural network models which are able to mimic key behaviors of this system.
 
@@
 
-In order to minimize this possibility, I define the task of strict unambiguous NLI (SU-NLI).
-In this task, only entailments that are licensed by a strictly literal interpretation of the provided sentences are considered valid, and several constraints are applied to the language to minimize ambiguity:
-\begin{itemize}
-\ex A small, unambiguous vocabulary is used.
-\ex All strings are given an explicit unlabeled tree structure parse.
-\ex Statements involving the hard-to-formalize generic senses of nouns---i.e. \ii{dogs bark} as opposed to the non-generic \ii{all dogs bark}---are excluded.
-\ex The sentences do not contain context dependent elements. This includes any reference to time or any tense morphology, and all pronouns.
-\ex The morphology is dramatically simplified: the copula is not used (\ii{some puppy is French} is simplified to \ii{some puppy French}, to make it more directly comparable to sentences like \ii{some puppy bark}), and agreement marking (\ii{they walk} vs.
-\ii{she walks}) is omitted.
-\end{itemize}
+The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes.
+It should be noted that this is not a balanced classification problem.
+If arbitrary pairs of statements are chosen for comparison in almost any domain of natural langauge, the minimally informative \# relation will apply between them.
+% TODO: How much should we say about this?
 
-The key to success at this task is to learn a set of representations and functions that can mimic the logical structure underlying the data.
-There is limited precedent that deterministic logical behavior can be learned in supervised deep learning models: \citet{socher2012semantic} show in an aside that a Boolean logic with negation and conjunction can be learned in a minimal recursive neural network model with one-dimensional (scalar) representations for words.
-Modeling the logical behavior that underlies linguistic reasoning, though, is a substantially more difficult challenge, even in modular hand-built models.
-
-The natural logic engine at the core of MacCartney's \cite{maccartney2009natural} NLI system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures.
+% The natural logic engine at the core of MacCartney and Manning's system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures.
 These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
-for dozens of different cases of insertion, deletion and substitution of different types of lexical item.
+%for dozens of different cases of insertion, deletion and substitution of different types of lexical item.
 For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences.
@@
 It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
-\begin{quote}
+%\begin{quote}
 
-1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
-2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
-3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}
+%1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
+%2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
+%3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}
 
-\end{quote}
+%\end{quote}
 
-This study is the first that I am aware of to attempt to build an inference engine based on learned vector representations, but two recent projects have attempted to introduce vector representations into inference systems in other ways:
+Work to date on inference in neural network models is quite limited.
 \citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.

### 2014-06-03 writing/S'14 paper/join.tex
Updating paper.

@@
 
-I test the model's ability to learn this behavior by creating an artificial dataset of terms which represent sets of numbers.
-Since MacCartney and Manning's set of relations hold between sets as well as between sentences, I can use the underlying set structure to generate the all of the relations that hold between any pair of these terms, as in Figure \ref{lattice-figure}.
-I train the model defined above on a subset of these relations, but rather then presenting the model with a pair of tree-structured sentences as inputs, simply present it with two single terms, each of which corresponds to a single vector in the (randomly initialized) vocabulary matrix $V$, ensuring that the model has no information about the terms being compared except the relations between them.
+We test the model's ability to learn this behavior by creating an artificial dataset of terms which represent sets of numbers.
+Since MacCartney and Manning's set of relations hold between sets as well as between sentences, we can use the underlying set structure to generate the all of the relations that hold between any pair of these terms, as in Figure \ref{lattice-figure}.
+We train the model defined above on a subset of these relations, but rather then presenting the model with a pair of tree-structured sentences as inputs, simply present it with two single terms, each of which corresponds to a single vector in the (randomly initialized) vocabulary matrix $V$, ensuring that the model has no information about the terms being compared except the relations between them.
 
-I generate 80 randomly generated sets drawing from the same seven elements, and create a dataset consisting of the relations between every pair of sets, yielding 6400 pairs.
+We generate 80 randomly generated sets drawing from the same seven elements, and create a dataset consisting of the relations between every pair of sets, yielding 6400 pairs.
 3200 of these pairs were then chosen as a test dataset, and that test dataset was further split into the 2960 examples that can be provably derived from the test data using MacCartney and Manning's join table (or by the symmetry of the relations in about half of the cases) and the 240 that cannot.
@@
 
+We tested a version of both the RNN model and the RNTN model on these data.
+In both cases, the models were implemented exactly as described in Section \ref{methods}, but since the items being compared are single terms rather than full tree structures, the composition layer was not used, and the two models differed only in which layer function was used for the comparison layer.
 We found that the RNTN model worked best with 11 dimensional vector representations for the 80 sets and a 90 dimensional feature vector for the classifier.
@@
 
-Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...
+% TODO: T-sne the RNTN model's embeddings.
 
-% Points for discussion: 
-% - Fairly simple interpretation: RNTN generalized, RNN didn't
-% - Future work: Can this generalize to models with huge numbers of entities, or does this result depend on the examples reflecting a small model in any way?
-Can't exhaustively generate large models without getting \#s - need sampling.
+These results are fairly simple to interpret.
+The RNTN model was able to accurately encode the relations between the terms in the geometric relations between their vectors, and was able to then use that information to recover relations that were not overtly included in the training data.
+The RNN model was able to approximate this behavior only incompletely.
+It is possible but not likely that it could be made to find a good solution with further optimization on different learning algorithms, or that it would do better on a larger universe of sets for which there would be a larger set of training data to learn from.

### 2014-06-03 writing/S'14 paper/methods.tex
Adding Monday morning results to paper.
Updating paper.

@@
-\section{Recursive neural network models}
+\section{Recursive neural network models} \label{methods}
 
-The models that we study is are centered aronund a recursively applied composition function which is meant to mimic the recursive construction of meanings in formal semantics.
+The models that we study are are centered aronund a recursively applied composition function which is meant to mimic the recursive construction of meanings in formal semantics.
 In this scheme, pairs of words are merged into phrase representations by a function that maps from representations of length $2N$ to representations of length $N$.
@@
 
-Borrowing a model directly from the existing literature for this task is impossible since none has been proposed to detect asymmetric relations between phrases.
-Instead, we build a combination model, depicted in Figure \ref{sample-figure}.
+We use the model architecture proposed in \citet{bowman2013can} and depicted in Figure \ref{sample-figure}.
 The two phrases being compared are built up separately on each side of the tree using the same composition function until they have each been reduced to single vectors.
@@
 
-For a composition layer, we experiment with both a plain neural network layer function (eq.
+For a composition layer, we evaluate models with both the plain neural network layer function (eq.
 \ref{rnn}) and the RNTN layer function proposed in \citet{chen2013learning} (eq.
@@
 
-The comparison layer uses the same type of function with different parameters and a different nonlinearity function wrapped around it.
-Rather than use a $\tanh$ nonlinearity here, I found a substantial improvement in performance by using a rectified linear function for $f_{b}$.
-In particular, I use the leaky rectified linear function \cite{maasrectifier}: $f_{b}(\vec{x})=\max(\vec{x}, 0) + 0.01\min(\vec{x}, 0)$,  applied elementwise.
+The comparison layer uses the same type function template as the composition function (either an NN layer or an NTN layer) with independantly learned parameters and a separate nonlinearity function.
+Rather than use a $\tanh$ nonlinearity here, we found a substantial improvement in performance by using a rectified linear function for $f_{b}$.
+In particular, we use the leaky rectified linear function \cite{maasrectifier}: $f_{b}(\vec{x})=\max(\vec{x}, 0) + 0.01\min(\vec{x}, 0)$,  applied elementwise.
 
@@
 The word vectors are then looked up from the vocabulary matrix $V$ (one of the model parameters), and the composition and comparison functions are used to pass information up the tree and into the classifier at the top.
-The model is trained using backpropagation through structure  \cite{goller1996learning}, wherein the negative log of the probability assigned to the correct label is taken as a cost function, and the gradient of each parameter with respect to that cost function is computed at each node, with information passing down the tree and into both the function parameters and the vocabulary matrix.
-Gradients for different instances of the composition function or different instances of a word in the same tree are summed to produce at most a single gradient update per parameter.
+For an objective function, we use the negative log of the probability assigned to the correct label.
 
-\textit{Optimization:} I train the model with stochastic gradient descent (SGD), with gradients pooled from randomly chosen minibatches of 32 training examples, and learning rates computed using AdaGrad \cite{duchi2011adaptive} from a starting rate of 0.2.
+We train the model with minibatch stochastic gradient descent (SGD) with learning rates computed using AdaGrad \cite{duchi2011adaptive} from a starting rate of 0.2.
 The parameters (including the vocabulary) are initialized randomly using a uniform distribution over $[-0.1, 0.1]$.
@@
 \begin{center}
-% \Tree [.{\sc softmax classifier}  [.{\sc comparison layer} [.{\sc \ii{dog} vector from $V$} ] [.{\sc composition layer} [.{\sc composition layer} [.{\sc \ii{very} vector from $V$} ] [.{\sc \ii{big} vector from $V$} ] ] [.{\sc \ii{cat} vector from $V$} ] ] ] ]
-\includegraphics[scale=0.36]{model.eps}
+\includegraphics[scale=0.35]{model.eps}
 \end{center}
-\caption{The model structure used to compare \ii{no dog} and \ii{(not all) dog}. \label{sample-figure}} 
-\end{figure} % TODO: Create a more compact figure using something better than qtree.
+\caption{The model structure used to compare \ii{no dog} and \ii{(not all) dog}. The same structure is used for both the RNN and RNTN layer functions. \label{sample-figure}} 
+\end{figure}
 
-\ii{Source code and data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?
+\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?

### 2014-06-03 writing/S'14 paper/quantifiers.tex
Updating paper.

@@
 \caption{Quantifier experiment performance. *Most frequent class accuracy is measured against the most frequent class in the training data, \#.\label{resultstable}}
-\end{table} % TODO: Replace
+\end{table}
 
@@
 
-Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...
+ (Note: Models still running.
+Results might improve and this para is likely to change.) The strong performance by the RNN on the \textsc{pair no/no} experiment suggests that it is at least plausible that this model is able to handle quantifiers correctly given sufficient training data, but the sub 90\% results on most of the training settings suggest that the data provided in these experiments are not generally sufficient, and that the model may not be able to learn to reason with quantifiers with high accuracy.
+However, the fact that both models perform far above baseline is promising, and the question of how much data is necessary to accurately capture quantifier behavior in a na\"ive model remains open.
 
-% Notes for discussion:
-% - First effort to learn a logic, results somewhat unclear.
-% -- No straightforward way to prove that there is enough information in the data to learn quantifier projecitivies in this setting
-% - Perfect performance on no-no promising that the model is learning at least some structure
-
-% TODO: Revise
+Discussion of RNN results (and updated discussion of better RNTN results) to be filled in as results arrive.\\...\\... % TODO

### 2014-06-03 writing/S'14 paper/recursion.tex
Updating paper.

@@
 
-% NOTE: Worth explicitly calling this project theorem proving? Yes! There was some confusion at CSLI.
-
+% NOTE: There was some confusion at CSLI and the 224U lecture about whether this was theorem proving or some kind of model checking. Hopefully this is explicit enough.
 Socher et al.
@@
 
-Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...
-
-% TODO:
-% Points for discussion
-% - The RNTN has learned to approximate the correct funtionn
-% -- The error introduced by the approximation likely grows with depth, but is sufficiently small at the training example sizes
-
-% - The RNN has plausibly learned an approximation as well, but one which is quite noisy even without recursion
-% -- Possible but unlikely that further tuning and different types or amounts of training data could get the RNN to work
-
-% - Future work:
-% -- Generate much larger examples to tease apart the ability of the model to approximate a representable function from data, and the ability of the model to represent
+The performance of the RNTN model on small unseen test examples indicates that it learned a correct approximation of the underlying logic.
+It appears that this approximation is accurate enough to yield correct answers when the composition layer is only applied a small number of times, but that the error in the approximation grows with depth, resulting in the gradually dropping performance.
+This is not necessarily a significant flaw in the model, since it remains possible that a less lossy approximation of the same logic would do no better on the training objective, since that objective only considers short strings and includes a regularization term.
+Whether a more accurate encoding could be learned from a larger set of training data remains a quention for future work.
+In contrast, the RNN model did not appear to be able to learn a correct approximation of the logic.

### 2014-06-03 writing/S'14 paper/intro.tex
Revised, more compact intro, and new refs.

@@
 have been successful in a wide array of sophisticated language tasks,
-including sentiment analysis \cite{socher2011semi}, analogy completion, image description,
-and paraphase detection . These results are encouraging about the
-ability of these models to learn compositional semantic grammars, but
-it remains an open question whether they can achieve the same results
-as grammars based in logical forms when it comes to core semantic
-concepts like quantification, entailment, and contradiction defined
-over complex linguistic structures. These concepts are central to
-robust natural language understanding, so it is essential that our
-computational models be able to capture them. % TODO: Insert more citations for past RNN work.
+including sentiment analysis \cite{socher2011semi,socher2013acl1},
+image description \cite{sochergrounded}, and paraphase detection
+\cite{Socher-etal:2011:Paraphrase}. These results are encouraging
+about the ability of these models to learn compositional semantic
+grammars, but it remains an open question whether they can achieve the
+same results as grammars based in logical forms when it comes to core
+semantic concepts like quantification, entailment and contradiction.
+To date, experimental investigations of these concepts using
+distributed representations have been largely confined to short
+phrases
+\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
+For robust natural language understanding, we must be able to model
+these phenomena in full generality on complex linguistic structures.
 
-We address this question directly by using a logical grammar togenerate controlled data sets encoding the semantic relationships between pairs of expressions and evaluating whether each of twoclasses of neural model --- plain RNNs and recursive neural tensor networks (RNTNs, \cite{socher2013acl1}) --- can learn those relationships correctly.
-Following \citet{bowman2013can}, we use the natural logic developed by\cite{maccartney2009extended} as our formal model, which defines seven core relations ofsynonymy, entailment, contradiction, and mutual consistency.
-Natural logic is well-suited to our purposes because of it defines a rich setof relations between words, phrases, and sentences, and because its logical expressions are derived directly from surface forms, making itamenable to large-scale natural language processing.
-% Is this the right place for the first self citation?
+We address this question in the context of \ii{natural language
+  inference} (also known as \ii{recognizing textual entailment};
+\cite{dagan2006pascal}), in which the goal is to determine the core
+inferential relationship between two sentences. Much of the
+theoretical work on this task (and some successful implemented models
+\cite{maccartney2009natural,watanabe2012latent}) involves \ii{natural
+  logics}, which are formal systems that define sound rules of
+inference between words, phrases, and sentences of language without
+the need of intermediate representations in an artificial logical
+language. Following \cite{bowman2013can}, we use the natural logic
+developed by \cite{maccartney2009extended} as our formal model. This
+logic defines seven core relations of synonymy, entailment,
+contradiction, and mutual consistency, and it provides rules of
+semantic combination for projecting these relations from the lexicon
+up to complex phrases. The formal properties and inferential strength
+of this system are now well-understood
+\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
 
-In our first experiment, we show that both of these models can learn
-the core relational algebra of natural logic from data. Our second
-experiment builds on this result to cover relations between complex
-recursive structures like `(A or B)' and `not(not A and not B)', and
-our third experiment involves relations between quantified statements
-like `every reptile walks' and `every turtle moves'. We find that the
-plain RNN achieves only mixed results on the second and third
-experiments, but that the stronger RNTN models generalized well in
-every case, suggesting that it has in fact learned, or at least
-learned to simulate, our target logical concepts. % TODO: Update with current results
+In our experiments, we use this pre-specified logical grammar to
+generate controlled data sets encoding the semantic relationships
+between pairs of expressions and evaluating whether each of two
+classes of neural model --- plain RNNs and recursive neural tensor
+networks (RNTNs, \cite{socher2013acl1}) --- can learn those
+relationships correctly. In our first experiment, we evaluate the
+ability of these models to learn the core relational algebra of
+natural logic from data. Our second experiment extends this evaluation
+to cover relations between complex recursive structures like `(A or
+B)' and `not(not A and not B)', and our third experiment involves
+relations between quantified statements like `every reptile walks' and
+`every turtle moves'. We find that the plain RNN achieves only mixed
+results on all three experiments, whereas the stronger RNTN models
+generalized well in every case, suggesting that it has in fact
+learned, or at least learned to simulate, our target logical
+concepts. % TODO: Update with current results
 
-Citations to additional past work to be added.\\...\\...\\...\\...\\...
+% Citations to additional past work to be added.\\...\\...\\...\\...\\...
 
@@
 
-\subsection{Natural language inference and natural logic}
+% \subsection{Natural language inference and natural logic}
 
-In its simplest form, \ii{natural language inference} (also known as \ii{recognizing textual entailment}, as in \cite{dagan2006pascal}) is the task of determine whether one sentence entails another.
+% In its simplest form, \ii{natural language inference} (also known as \ii{recognizing textual entailment}, as in \cite{dagan2006pascal}) is the task of determine whether one sentence entails another.
 Much of the theoretical work on this task (and some successful implemented models \cite{maccartney2009natural, watanabe2012latent}) involve \ii{natural logic}, formal systems that define sound rules of inference from one sentence of natural language to another without the need for intermediate representations in some other logic.
@@
 
-This approach to natural logic can capture much of the complexity of natural language meaning within a well understood framework, and is also fairly straightforward to implement in a machine learning setting since it can be reduced to a seven-way classification problem on sentence pairs.
+% This approach to natural logic can capture much of the complexity of natural language meaning within a well understood framework, and is also fairly straightforward to implement in a machine learning setting since it can be reduced to a seven-way classification problem on sentence pairs.
 Our goal in this paper is to learn recursive neural network models which are able to mimic key behaviors of this system.
@@
 
-The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes.
+% The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes.
 It should be noted that this is not a balanced classification problem.
@@
 
-Work to date on inference in neural network models is quite limited.
-\citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
-In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations.
+% Work to date on inference in neural network models is quite limited.
+% \citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
+% In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations.
 They propose a deterministic inference engine (similar to MacCartney's) which is augmented by a probabilistic component that evaluates individual lexical substitution steps in the derivation using vector representations, though again these representations are not learned, and no evaluations of this system have been published to date.
-\label{sec2}
+% \label{sec2}

### 2014-06-03 writing/S'14 paper/intro.tex
Improved join table table

@@
 
-\begin{table}
-\begin{center}
-\begin{tabular}{|c|c|c|c|} \hline
-name & symbol & set-theoretic definition & example \\ \hline \hline
-entailment & $x \sqsubset y$ & $x \subset y$ & \ii{crow, bird}  \\ \hline
-reverse entailment & $x \sqsupset y$ & $x \supset y$ & \ii{Asian, Thai}  \\ \hline
-equivalence & $x \equiv y$ & $x = y$ & \ii{couch, sofa} \\ \hline
-alteration & $x$ $|$ $y$ & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{cat, dog} \\ \hline
-negation & $x \natneg y$ & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{able, unable} \\ \hline
-cover & $x \smallsmile y$ & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-ape} \\ \hline
-independence & $x$ \# $y$ & (else) & \ii{hungry, hippo}\\ \hline
-\end{tabular}
-\caption{The entailment relations in  $\mathfrak{B}$.
+% \begin{table}
+% \begin{center}
+% \begin{tabular}{|c|c|c|c|} \hline
+% name & symbol & set-theoretic definition & example \\ \hline \hline
+% entailment & $x \sqsubset y$ & $x \subset y$ & \ii{crow, bird}  \\ \hline
+% reverse entailment & $x \sqsupset y$ & $x \supset y$ & \ii{Asian, Thai}  \\ \hline
+% equivalence & $x \equiv y$ & $x = y$ & \ii{couch, sofa} \\ \hline
+% alteration & $x$ $|$ $y$ & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{cat, dog} \\ \hline
+% negation & $x \natneg y$ & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{able, unable} \\ \hline
+% cover & $x \smallsmile y$ & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-ape} \\ \hline
+% independence & $x$ \# $y$ & (else) & \ii{hungry, hippo}\\ \hline
+% \end{tabular}
+% \caption{The entailment relations in  $\mathfrak{B}$.
 $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, and the relation \# applies whenever none of the other six do, including when there is insufficient knowledge to choose a label.}
-\label{b-table}
-\end{center}
-\end{table}
+% \label{b-table}
+% \end{center}
+% \end{table}
 

### 2014-06-03 writing/S'14 paper/join.tex
Improved join table table

@@
+\begin{table}[htp]
+  \centering
+  \setlength{\tabcolsep}{15pt}
+  \renewcommand{\arraystretch}{1.1}
+  \begin{tabular}{l c l l} 
+    \toprule
+    Name & Symbol & Set-theoretic definition & Example \\ 
+    \midrule
+    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
+    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
+    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
+    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
+    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
+    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
+    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
+    \bottomrule
+  \end{tabular}
+  \caption{The entailment relations in $\mathfrak{B}$. 
+    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
+    and the relation $\natind$ applies whenever none of the other six do, including when there 
+    is insufficient knowledge to choose a label.}
+  \label{b-table}
+\end{table}
+
 \section{Relation composition}

### 2014-06-03 writing/S'14 paper/abstract.tex
Revisions focused on the join table section

--- 2014-05-30 writing/S'14 paper/abstract.tex
@@
 \begin{abstract}
-Supervised recursive neural network models (RNNs) for sentence meaning
-have been successful in a wide array of sophisticated language tasks,
-but it remains an open question whether they can achieve the same
-results as compositional semantic grammars based in logical forms. We
-address this question directly by using a logical grammar to generate
-controlled data sets encoding the relationships (entailment, synonymy,
-contradiction) between pairs of expressions and evaluating whether
-each of two classes of neural model --- plain RNNs and recursive
-neural tensor networks (RNTNs) --- can learn those relationships
-correctly. Our first experiment confirms that both models can learn
-the basic algebra of logical relations involved. Our second and third
-experiments expand on this result to cover complex recursive
-structures and sentences involving quantification. We find that the
-plain RNN achieves only mixed results on the second and third
-experiments, but that the stronger RNTN models generalized well in
-every case.
+  Supervised recursive neural network models (RNNs) for sentence
+  meaning have been successful in a wide array of sophisticated
+  language tasks, but it remains an open question whether they can
+  achieve the same results as compositional semantic grammars based in
+  logical forms. We address this question directly by using a logical
+  grammar to generate controlled data sets encoding the relationships
+  (e.g., entailment, contradiction) between pairs of expressions and
+  evaluating whether each of two classes of neural model --- plain
+  RNNs and recursive neural tensor networks (RNTNs) --- can learn
+  those relationships correctly. Our first experiment evaluates
+  whether these models can learn the basic algebra of logical
+  relations involved. Our second and third experiments extend this
+  evaluation to complex recursive structures and sentences involving
+  quantification.  We find that the plain RNN achieves only mixed
+  results on all three experiments, whereas the stronger RNTN models
+  generalizes well in every setting.
 \end{abstract}

### 2014-06-03 writing/S'14 paper/intro.tex
Revisions focused on the join table section

@@
 \cite{maccartney2009natural,watanabe2012latent}) involves \ii{natural
-  logics}, which are formal systems that define sound rules of
-inference between words, phrases, and sentences of language without
-the need of intermediate representations in an artificial logical
+  logics}, which are formal systems that define rules of inference
+between natural language words, phrases, and sentences without the
+need of intermediate representations in an artificial logical
 language. Following \cite{bowman2013can}, we use the natural logic
@@
 logic defines seven core relations of synonymy, entailment,
-contradiction, and mutual consistency, and it provides rules of
-semantic combination for projecting these relations from the lexicon
-up to complex phrases. The formal properties and inferential strength
-of this system are now well-understood
-\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
+contradiction, and mutual consistency, as summarized in
+Table~\ref{b-table}, and it provides rules of semantic combination for
+projecting these relations from the lexicon up to complex phrases. The
+formal properties and inferential strength of this system are now
+well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
 
@@
 concepts. % TODO: Update with current results
+
+\begin{table}[htp]
+  \centering
+  \setlength{\tabcolsep}{15pt}
+  \renewcommand{\arraystretch}{1.1}
+  \begin{tabular}{l c l l} 
+    \toprule
+    Name & Symbol & Set-theoretic definition & Example \\ 
+    \midrule
+    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
+    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
+    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
+    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
+    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
+    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
+    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
+    \bottomrule
+  \end{tabular}
+  \caption{The entailment relations in $\mathfrak{B}$. 
+    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
+    and the relation $\natind$ applies whenever none of the other six do, including when there 
+    is insufficient knowledge to choose a label.}
+  \label{b-table}
+\end{table}
 

### 2014-06-03 writing/S'14 paper/join.tex
Revisions focused on the join table section

--- 2014-06-02 writing/S'14 paper/join.tex
@@
 
-In order for a model to be able to learn to mimic the behavior of a relational logic like the one presented here from a finite amount data, it must be able to learn to deduce new relations from seen relations.
-The simplest such deductions involve facts about the relations themselves that do not involve considering the internal structure of the things being compared.
-For example, given that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that $a\sqsupset c$ by the transitivity of $\sqsupset$, even without understanding $a$, $b$, or $c$.
-These seven relations support more than just transitivity: MacCartney and Manning's \cite{maccartney2009extended} join table defines 32 valid inferences that can be made on the basis of pairs of relations of the form $a R b$ and $b R' c$, including several less intuitive ones such as that if $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
+If a model is to learn the behavior of a relational logic like the one
+presented here from a finite amount data, it must be able to learn to
+deduce new relations from seen relations in a sound and complete
+manner. The simplest such deductions involve atomic statements using
+the relations in Table~\ref{b-table}. For instance, given that $a
+\natrev b$ and $b \natrev c$, one can conclude that $a \natrev c$, by
+basic set-theoretic reasoning (transitivity of $\natrev$). Similarly,
+from $a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.
+The full set of sound inferences of this form is summarized in
+Table~\ref{tab:jointable}.
 
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{ll|lll}
-$a := \{1, 3, 4, 5, 6\}	$&~~~&~~~& $a \equiv a$	    	& $a~\#~c$	\\
-$b := \{0, 3, 5, 6\}	$&~~~&~~~& $e \sqsupset c$	&$b \smallsmile c$		\\
-$c := \{1, 2, 3, 4\}	$&~~~&~~~& $d \smallsmile e$	& $b~\#~d$		\\
-$d := \{0, 4\}		$&~~~&~~~& $a \sqsubset e$	& $ d \equiv d$	\\
-$e := \{1, 2, 3, 4, 5, 6\}$&~~~&~~~& $a~\#~b$		& ... 	\\
+% about the relations themselves that do not depend on the
+% internal structure of the things being compared. For example, given
+% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
+% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
+% understanding $a$, $b$, or $c$. These seven relations support more
+% than just transitivity: MacCartney and Manning's
+% \cite{maccartney2009extended} join table defines 32 valid inferences
+% that can be made on the basis of pairs of relations of the form $a R
+% b$ and $b R' c$, including several less intuitive ones such as that if
+% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
+
+\begin{table}[htp]
+  \centering  
+  \setlength{\arraycolsep}{12pt}
+  \renewcommand{\arraystretch}{1.1}
+  %\newcommand{\rcg}{\cellcolor[gray]{.9}}
+  \newcommand{\rcg}{}
+  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
+    %\hline
+    \multicolumn{1}{c}{}
+             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
+    \cline{2-8}
+    \nateq   & \rcg\nateq & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \natind\\
+    \natfor  & \natfor    & \rcg\natfor & \natind     & \natalt    & \natalt     & \natind     & \natind\\
+    \natrev  & \natrev    & \natind     & \rcg\natrev & \natcov    & \natind     & \natcov     & \natind\\
+    \natneg  & \natneg    & \natcov     & \natalt     & \rcg\nateq & \natrev     & \natfor     & \natind\\
+    \natalt  & \natalt    & \natind     & \natalt     & \natfor    & \rcg\natind & \natfor     & \natind\\
+    \natcov  & \natcov    & \natcov     & \natind     & \natrev    & \natrev     & \rcg\natind & \natind\\
+    \natind  & \natind    & \natind     & \natind     & \natind    & \natind     & \natind     & \rcg\natind \\
+    \cline{2-8}
+  \end{array}$
+  \caption{Full set of inferences between natural logic relations.
+    Given $a \mathbin{R} b$ and $b \mathbin{S} c$, where $R$ and $S$
+    are the row and column relations, respectively, and $a$, $b$, and
+    $c$ are arbitrary formulae, the table provides the relation
+    $a \mathbin{T} c$.}
+  \label{tab:jointable}
+\end{table}
+
+To test our models' ability to learn this relational structure, we
+create small models for our logic in which terms denote sets of
+entities from a single domain of seven entities (integers).
+Figure~\ref{lattice-figure} depicts a small model of this form. The
+lattice structure gives the full model, in which all the statements on
+the right can be proved. The highlighted nodes represent a
+hypothetical training set drawn from this larger space, which allows
+us to prove only a subset of the full set of true statements.
+
+\begin{figure}[htp]
+  \centering
+  \begin{subfigure}[t]{0.4\textwidth}
+    \centering
+    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
+    \setlength{\unitlength}{1cm}
+    \begin{picture}(5,5.5)
+      \labelednode{2.50}{5}{a}{1,2,3}
+
+      \put(0.75,4){\line(3,1){1.5}}
+      \put(2.5,4){\line(0,1){0.5}}
+      \put(4.25,4){\line(-3,1){1.5}}
+
+      \labelednode{0.75}{3.5}{b}{1,2}
+      \labelednode{2.50}{3.5}{c}{1,3}
+      \labelednode{4.25}{3.5}{d}{2,3}
+
+      \put(0.75,2.5){\line(0,1){0.5}}
+      \put(0.75,2.5){\line(3,1){1.5}}
+
+      \put(2.5,2.5){\line(-3,1){1.5}}
+      \put(2.5,2.5){\line(3,1){1.5}}
+
+      \put(4.25,2.5){\line(0,1){0.5}}
+      \put(4.25,2.5){\line(-3,1){1.5}}
+
+      \labelednode{0.75}{2}{e}{1}
+      \labelednode{2.50}{2}{f}{2}
+      \labelednode{4.25}{2}{g}{3}
+
+      \put(2.5,1){\line(-3,1){1.5}}
+      \put(2.5,1){\line(0,1){0.5}}
+      \put(2.5,1){\line(3,1){1.5}}
+
+      \labelednode{2.5}{0.5}{h}{}
+    \end{picture}
+    \caption{Simple boolean model.}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[t]{0.5\textwidth}
+    \centering
+    \setlength{\tabcolsep}{12pt}
+    \begin{tabular}[b]{c  c}
+      \toprule
+      Train & Test \\
+      \midrule
+                    & $b \nateq b$ \\
+      $b \natcov c$ &               \\
+                    & $b \natcov d$ \\
+                    & $b \natrev e$ \\
+      $c \natcov d$ &               \\
+      $c \natrev e$ &               \\
+                    & \strikeout{$c \natneg f$} \\
+      $c \natrev g$ &               \\ 
+                    & $e \natfor b$ \\
+      $e \natfor c$ &               \\[-1ex]
+      $\vdots$      & $\vdots$ \\
+      \bottomrule
     \end{tabular}
-\end{center}
 
-\caption{Some sample randomly generated sets, and some of the relations defined between them.  \label{lattice-figure}} 
+    \caption{An example train/test split from the full set
+      of statements one can make about the model.
+      Statements not provable from the test data are crossed out.}
+  \end{subfigure}  
+  % \begin{tabular}{ll|lll}
+  %   $a := \{1, 3, 4, 5, 6\}	$&~~~&~~~& $a \equiv a$	    	& $a~\#~c$	\\
+  %   $b := \{0, 3, 5, 6\}	$&~~~&~~~& $e \sqsupset c$	&$b \smallsmile c$		\\
+  %   $c := \{1, 2, 3, 4\}	$&~~~&~~~& $d \smallsmile e$	& $b~\#~d$		\\
+  %   $d := \{0, 4\}		$&~~~&~~~& $a \sqsubset e$	& $ d \equiv d$	\\
+  %   $e := \{1, 2, 3, 4, 5, 6\}$&~~~&~~~& $a~\#~b$		& ... 	\\
+  % \end{tabular}
+  \caption{Some sample randomly generated sets, and some of the relations defined between them.}
+  \label{lattice-figure}
 \end{figure}
 
-I test the model's ability to learn this behavior by creating an artificial dataset of terms which represent sets of numbers.
-Since MacCartney and Manning's set of relations hold between sets as well as between sentences, I can use the underlying set structure to generate the all of the relations that hold between any pair of these terms, as in Figure \ref{lattice-figure}.
-I train the model defined above on a subset of these relations, but rather then presenting the model with a pair of tree-structured sentences as inputs, simply present it with two single terms, each of which corresponds to a single vector in the (randomly initialized) vocabulary matrix $V$, ensuring that the model has no information about the terms being compared except the relations between them.
+% We test the model's ability to learn this behavior by creating
+% artificial data sets of terms which represent sets of numbers. Since
+% MacCartney and Manning's set of relations hold between sets as well as
+% between sentences, we can use the underlying set structure to generate
+% the all of the relations that hold between any pair of these terms, as
+% in Figure \ref{lattice-figure}. We train the model defined above on a
+% subset of these relations, but rather then presenting the model with a
+% pair of tree-structured sentences as inputs, simply present it with
+% two single terms, each of which corresponds to a single vector in the
+% (randomly initialized) vocabulary matrix $V$, ensuring that the model
+% has no information about the terms being compared except the relations
+% between them.
 
-I generate 80 randomly generated sets drawing from the same seven elements, and create a dataset consisting of the relations between every pair of sets, yielding 6400 pairs.
-3200 of these pairs were then chosen as a test dataset, and that test dataset was further split into the 2960 examples that can be provably derived from the test data using MacCartney and Manning's join table (or by the symmetry of the relations in about half of the cases) and the 240 that cannot.
+In our experiments, we create 80 randomly generated sets drawing from
+the same domain of seven element. This yields a data set consisting of
+6400 statements about pairs of entities. 3200 of these pairs are
+chosen as a test set, and that test set is further reduced to the 2960
+examples that can be provably derived from the test data using
+Table~\ref{tab:jointable}.
+
 % TODO: Say more about symmetry?
+We trained versions of both the RNN model and the RNTN model on these
+data sets. In both cases, the models were implemented exactly as
+described in Section~\ref{methods}, but since the items being compared
+are single terms rather than full tree structures, the composition
+layer was not involved, and the two models differed only in which
+layer function was used for the comparison layer. We simply present
+the models with two single terms, each of which corresponds to a
+single vector in the (randomly initialized) vocabulary matrix $V$,
+ensuring that the model has no information about the terms being
+compared except the relations between them. We found that the RNTN
+model worked best with 11-dimensional vector representations for the
+80 sets and a 90-dimensional feature vector for the classifier. This
+model was able to correctly label 99.3\% of the derivable test
+examples, and 99.1\% of the remaining examples. The simpler RNN model
+worked best with 11 and 75 dimensions, respectively, but was able to
+achieve accuracies of only 90.0\% and 87.\%, respectively.
 
-We found that the RNTN model worked best with 11 dimensional vector representations for the 80 sets and a 90 dimensional feature vector for the classifier.
-This model was able to correctly label 99.3\% of the derivable test examples, and 99.1\% of the remaining examples.
-The simpler RNN model worked best with 11 and 75 dimensions, respectively, but was able to achieve accuracies of only 90.0\% and 87.\%, respectively.
+% and create a dataset consisting of the relations between every pair of
+% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
+% test dataset, and that test dataset was further split into the 2960
+% examples that can be provably derived from the test data using
+% MacCartney and Manning's join table (or by the symmetry of the
+% relations in about half of the cases) and the 240 that
+% cannot. % TODO: Say more about symmetry?
 
-Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...
+% We tested a version of both the RNN model and the RNTN model on these
+% data. In both cases, the models were implemented exactly as described
+% in Section~\ref{methods}, but since the items being compared are
+% single terms rather than full tree structures, the composition layer
+% was not used, and the two models differed only in which layer function
+% was used for the comparison layer. We found that the RNTN model worked
+% best with 11 dimensional vector representations for the 80 sets and a
+% 90 dimensional feature vector for the classifier. This model was able
+% to correctly label 99.3\% of the derivable test examples, and 99.1\%
+% of the remaining examples. The simpler RNN model worked best with 11
+% and 75 dimensions, respectively, but was able to achieve accuracies of
+% only 90.0\% and 87.\%, respectively.
 
-% Points for discussion: 
-% - Fairly simple interpretation: RNTN generalized, RNN didn't
-% - Future work: Can this generalize to models with huge numbers of entities, or does this result depend on the examples reflecting a small model in any way?
-Can't exhaustively generate large models without getting \#s - need sampling.
+% TODO: T-sne the RNTN model's embeddings.
+
+These results are fairly straightforward to interpret. The RNTN model
+was able to accurately encode the relations between the terms in the
+geometric relations between their vectors, and was able to then use
+that information to recover relations that were not overtly included
+in the training data. In contrast, the RNN model was able to
+approximate this behavior only incompletely. It is possible but not
+likely that it could be made to find a good solution with further
+optimization on different learning algorithms, or that it would do
+better on a larger universe of sets for which there would be a larger
+set of training data to learn from, but the RNTN is able to achieve
+these effects in the well-studied framework of Section~\ref{methods}.

### 2014-06-03 writing/S'14 paper/methods.tex
Revisions focused on the join table section

@@
 
-The models that we study are are centered aronund a recursively applied composition function which is meant to mimic the recursive construction of meanings in formal semantics.
+We study neural models that adhere to the linguistic \ii{principle of
+  compositionality}, which says that the meanings for complex
+expressions are derived from the meanings of their constituent parts
+via specific combination functions \cite{Partee84,Janssen97}. In our
+distributional model, word meanings are vectors of length $N$. The
+combination function concatenates pairs of them to form vectors of
+length $2N$ and maps those larger vectors back into length $N$
+vectors, which can then be merged again to create more complex
+phrases. Once the entire sentence-level representation has been
+derived, it fed to a classifier trained in a supervised manner.
+
+% The models that we study are centered around a recursively applied composition function which is meant to mimic the recursive construction of meanings in formal semantics.
 In this scheme, pairs of words are merged into phrase representations by a function that maps from representations of length $2N$ to representations of length $N$.
@@
 
-We use the model architecture proposed in \citet{bowman2013can} and depicted in Figure \ref{sample-figure}.
-The two phrases being compared are built up separately on each side of the tree using the same composition function until they have each been reduced to single vectors.
-Then, the two phrase vectors are fed into a separate comparison layer that is meant to generate a feature vector capturing the relation between the two phrases.
-The output of this layer is then given to a softmax classifier, which in turn produces a hypothesized distribution over the seven relations.
+We use the model architecture proposed in \cite{bowman2013can} and
+depicted in Figure~\ref{sample-figure}. The two phrases being compared
+are built up separately on each side of the tree, using the same
+composition function, until they have each been reduced to single
+vectors. The resulting vectors are fed into a separate comparison
+layer that is meant to generate a feature vector capturing the
+relation between the two phrases. The output of this layer is then
+given to a softmax classifier, which in turn produces a hypothesized
+distribution over the seven relations represented in Table~\ref{b-table}.
 
-For a composition layer, we evaluate models with both the plain neural network layer function (eq.
-\ref{rnn}) and the RNTN layer function proposed in \citet{chen2013learning} (eq.
-\ref{rntn}).
-A sigmoid nonlinearity (elementwise $\tanh$) is applied to the output of either layer function, following Socher et al.
+For a composition layer, we evaluate models with both the plain neural
+network layer function (eq.~\ref{rnn}) and the RNTN layer function
+proposed in \citet{chen2013learning} (eq.~\ref{rntn}). A sigmoid
+nonlinearity (element-wise $\tanh$) is applied to the output of either
+layer function, following \cite{socher2013acl1}.
+%
 \begin{gather} \label{rnn}
-y_{RNN} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + b_i)\\ % TODO: Add column vectors?
+y_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + b_i)\\ % TODO: Add column vectors?
 \label{rntn}
-y_{RNTN} = f(\vec{x}^{(l)T} \mathbf{A}^{[1...N]} \vec{x}^{(r)} + \mathbf{B} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
+y_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{A}^{[1 \ldots N]} \vec{x}^{(r)} + \mathbf{B} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
 \end{gather} % TODO: Explain third order tensor parameter?
+%
+Here, the $l$ and $r$ superscripts indicate the vector representations
+for the left and right children of the local structure interpreted as
+$y$. The RNN concatenates them, multiplies them by an $N \times 2N$
+matrix of learned weights, and applies the element-wise non-linearity
+to the resulting vector (plus a bias term). The RNTN has the same
+basic structure, but with the addition of a tensor $\mathbf{A}$,
+dimension $2N \times N \times N$, modeling interactions between the
+child vectors.
 
-The comparison layer uses the same type function template as the composition function (either an NN layer or an NTN layer) with independantly learned parameters and a separate nonlinearity function.
-Rather than use a $\tanh$ nonlinearity here, we found a substantial improvement in performance by using a rectified linear function for $f_{b}$.
-In particular, we use the leaky rectified linear function \cite{maasrectifier}: $f_{b}(\vec{x})=\max(\vec{x}, 0) + 0.01\min(\vec{x}, 0)$,  applied elementwise.
+The comparison layer uses the same type function template as the
+composition function (either an NN layer or an NTN layer) with
+independantly learned parameters and a separate nonlinearity function.
+Rather than use a $\tanh$ nonlinearity here, we found a substantial
+improvement in performance by using a rectified linear function for
+$f_{b}$. In particular, we use the leaky rectified linear function
+\cite{maasrectifier}: $f_{b}(\vec{x})=\max(\vec{x}, 0) +
+0.01\min(\vec{x}, 0)$, applied element-wise.
 
-To run the model forward and label a pair of phrases, the structure of the lower layers of the network is assembled so as to mirror the tree structures provided for each phrase.
-The word vectors are then looked up from the vocabulary matrix $V$ (one of the model parameters), and the composition and comparison functions are used to pass information up the tree and into the classifier at the top.
-For an objective function, we use the negative log of the probability assigned to the correct label.
+To run the model forward and label a pair of phrases, the structure of
+the lower layers of the network is assembled so as to mirror the tree
+structures provided for each phrase. The word vectors are then looked
+up from the vocabulary matrix $V$ (one of the model parameters), and
+the composition and comparison functions are used to pass information
+up the tree and into the classifier at the top. For an objective
+function, we use the negative log of the probability assigned to the
+correct label.
 
-We train the model with minibatch stochastic gradient descent (SGD) with learning rates computed using AdaGrad \cite{duchi2011adaptive} from a starting rate of 0.2.
-The parameters (including the vocabulary) are initialized randomly using a uniform distribution over $[-0.1, 0.1]$.
-% TODO: Update
+We train the model with minibatch stochastic gradient descent (SGD)
+with learning rates computed using AdaGrad \cite{duchi2011adaptive}
+from a starting rate of 0.2. The parameters (including the vocabulary)
+are initialized randomly using a uniform distribution over $[-0.1,
+0.1]$. % TODO: Update
+Source code and generated data will be released after the conclusion
+of the review period.
 
-\begin{figure}
-\begin{center}
+\begin{figure}[htp]
+  \centering
   \includegraphics[scale=0.35]{model.eps}
-\end{center}
-\caption{The model structure used to compare \ii{no dog} and \ii{(not all) dog}. The same structure is used for both the RNN and RNTN layer functions. \label{sample-figure}} 
+  \caption{The model structure used to compare \ii{no dog} and \ii{(not all) dog}. 
+    The same structure is used for both the RNN and RNTN layer functions.} 
+  \label{sample-figure}
 \end{figure}
 
-\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?
+%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?

### 2014-06-03 writing/S'14 paper/intro.tex
Revised recursion section

@@
 grammars, but it remains an open question whether they can achieve the
-same results as grammars based in logical forms when it comes to core
-semantic concepts like quantification, entailment and contradiction.
-To date, experimental investigations of these concepts using
-distributed representations have been largely confined to short
-phrases
+same results as grammars based in logical forms
+\cite{ZetCol:2005,LiangJordan:2013} when it comes to core semantic
+concepts like quantification, entailment and contradiction. To date,
+experimental investigations of these concepts using distributed
+representations have been largely confined to short phrases
 \cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.

### 2014-06-03 writing/S'14 paper/recursion.tex
Revisions focused on the join table section
Revised recursion section

@@
 
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{lll}
-$a\equiv a$		&~~~&	$(c~(and~(not~d)))~\#~f$\\
-$b~\#~c$			&~~~&	$(not~(c~(or~b)))~\sqsubset~(not~c)$\\
-$d\natneg(not~d)$	&~~~&	$f~\#~((c~(or~(not~d)))~(and~a))$\\
-$(c~(and~d))\sqsubset d$&~~~&$d\sqsupset((d~(or~d))~(and~(not~b)))$\\
+% TODO: Cite Chomsky/Hauser/Fitch? -- no, not appropriate here --- CP.
+
+Recursive structure is an prominent feature of natural languages and
+an important part of their expressivity, as it allows people to use
+and interpret complex structures they have never encountered before.
+Theoretical accounts of natural language syntax and semantics
+therefore rely heavily on recursive structures, and we expect that
+accurate computational models will have to come to grips with it as
+well. The present section pursues this question for our two classes of
+RNN. In evaluating the model, we take advantage of the fact that
+recursive structures of this kind define potentially infinite
+languages, by testing the model on strings that are longer and more
+complex than any seen in testing.
+
+% Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice
+%   said hello}, and \ii{Carl thinks that Bob said that Alice said
+%   hello}. Overt recursion of this kind is easy to find, and
+% theoretical accounts of natural language syntax and semantics rely
+% heavily on recursive structures. In order for a model to be able to
+% accurately learn natural language meanings, then, we expect that it
+% would need to be able to learn to represent the meanings of function
+% words in a such a way that they are able to behave correctly when
+% taking their own outputs as input. In evaluating the model, we take
+% advantage of the fact that recursive structures of this kind define
+% potentially infinite languages by testing the model on strings that
+% are longer and more complex than any seen in testing.
+
+As in Section~\ref{sec:join}, we test this phenomenon within the
+framework of natural logic, but we now replace the unanalyzed symbols
+from that experiment with expressions that involve recursive
+structure. The result is a truth-functionally complete classical
+propositional logic: each atomic variable denotes either $\True$ or
+$\False$, and the only operators are truth-functional ones.
+Table~\ref{tab:pl} defines this logic, and Table~\ref{tab:plexs} gives
+some examples of natural logic relational statements that we can make
+in these terms. To compute these relations between statements, we
+exhaustively enumerate the sets of assignments of truth values to
+proposition variables that would satisfy each of the statements and
+then convert the set-theoretic relation between those assignments into
+one of the seven relations in Table~\ref{b-table}. This is equivalent
+to proving theorems about these statements model-theoretically.
+
+\newcommand{\plneg}{\mathop{\textit{not}}}
+\newcommand{\pland}{\mathbin{\textit{and}}}
+\newcommand{\plor}{\mathbin{\textit{or}}}
+
+\begin{table}[htp]
+  \centering
+  \begin{subtable}[t]{0.45\textwidth}
+    \centering
+    \begin{tabular}[t]{l l}
+      \toprule
+      Formula     & Interpretation \\
+      \midrule
+      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
+      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
+      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
+      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
+      \bottomrule
     \end{tabular}    
-\end{center}
+    \caption{Well-formed formulae. $\varphi$ and $\psi$
+      range over all well-formed formulae, and $\sem{\cdot}$ is
+      interpretation function mapping formulae into $\{\True,
+      \False\}$.}\label{tab:pl}
+  \end{subtable}
+  \quad
+  \begin{subtable}[t]{0.45\textwidth}
+    \centering
+    \begin{tabular}[t]{r c l}
+      \toprule
+      $\plneg a$        & $\natneg$ & $a$ \\
+      $\plneg \plneg a$ & $\nateq$  & $a$ \\
+      $a$               & $\natfor$ & $(a \plor b)$ \\
+      $a$               & $\natrev$ & $(a \pland b)$ \\
+      $(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
+      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
+      \bottomrule
+    \end{tabular}
+    \caption{Examples of statements about relations between
+      well-formed formulae, defined in terms of sets of satisfying
+      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
+  \end{subtable}
+  \caption{Propositional logic with natural logic relations.}  
+  \label{prop-figure}
+\end{table}
 
-\caption{Some sample randomly generated pairs of propositional logic statements.  \label{prop-figure}} 
+% NOTE: There was some confusion at CSLI and the 224U lecture about
+% whether this was theorem proving or some kind of model checking.
+% Hopefully this is explicit enough.
+Socher et al.~\cite{socher2012semantic} demonstrate the learning of a
+logic in a matrix-vector RNN model somewhat similar to our own, but
+the logic discussed here is substantially stronger, and a much better
+approximation of the kind of structure that is needed for natural
+language. The logic learned in that experiment is boolean, wherein the
+atomic symbols are simply the values $0$ and $1$, rather than
+variables over those values. While learning the operators of that
+logic is not trivial, the outputs of each operator can be represented
+accurately by a single bit. The statements of propositional logic
+learned here describe conditions on the truth values of propositions
+where those truth values are not known. As opposed to the two-way
+contrasts seen in \cite{socher2012semantic}, this logic distinguishes
+between $2^{6} = 64$ possible assignments of truth values, and
+expressions of this logic define arbitrary conditions on these
+possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
+possible statements that the recursive model needs to be able to
+distinguish.
+
+For our experiments, we randomly generate large sets of syntactically
+unique formulae and compute the natural logic relation between them.
+We discard pairs in which either statement is a tautology or
+contradiction (a statement that is true of either all or no possible
+assignments), for which none of the seven relation labels in
+Table~\ref{b-table} can hold. The resulting set of formula pairs is
+then partitioned into bins based on the number of logical operators in
+the longer of the two formulae. We then randomly sample 15\% of each
+bin for a held out test set.
+
+If we do not implement any constraint that the two statements
+being compared are similar in any way, the generated data consists in
+large part of statements in which the two formulae refer to largely
+separate subsets of the six variables, and to which we will nearly
+always assign the $\natind$ relation. In an effort to balance the
+distribution of relation labels without departing from the basic task
+of modeling propositional logic, we disallow individual pairs of
+statements from referring to more than four of the six proposition
+variables.
+
+% We deduplicate the data and discard pairs in which either statement is
+% a tautology or contradiction (a statement that is true of either all
+% or no possible assignments), for which none of the seven relation
+% labels can accurately apply. We then divide the generated pairs into
+% size bins based on the number of logical operators (\ii{and}, \ii{or},
+% or \ii{not}) in the larger of the two pairs being compared, and
+% discard examples of size greater than twelve by this measure. Finally,
+% we randomly sample 15\% of each bin for a held out test set.
+
+We trained both the RNN and RNTN models on the data of size four or
+less (65k pairs), and tested it on examples of up to size 12 (44k
+pairs). We initialized the model parameters randomly, including the
+vector representations of the six variables. The results are shown in
+Figure~\ref{prop-results}. In tuning, we found that the RNN model was
+approximately optimal with 45-dimensional vector representations, and
+the RNTN model was approximately optimal with 25 dimensions. We fixed
+the size of the feature vector for the classifier at 75 dimensions. We
+found that the RNTN model was able to perform almost perfectly on
+unseen small test examples, with accuracy above 99\% below size four.
+After depth four, performance gradually falls with increasing size.
+The RNN model did not perform well, reaching only 88.2\% accuracy on
+the smallest test examples, and declining from there to near-baseline
+performance at size 12.
+
+\begin{figure}[htp]
+  \centering
+  \includegraphics[width=5.5in]{recursion\string_results.eps}
+  \caption{Model performance by expression size.}  
+  \label{prop-results}
 \end{figure}
 
-% TODO: Cite Chomsky/Hauser/Fitch?
-
-Recursive structure is a prominent  feature of natural language.
-Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice said hello}, and \ii{Carl thinks that Bob said that Alice said hello}.
-Overt recursion of this kind is easy to find, and theoretical accounts of natural language syntax and semantics rely heavily on recursive structures.
-In order for a model to be able to accurately learn natural language meanings, then, we expect that it would need to be able to learn to represent the meanings of function words in a such a way that they are able to behave correctly when taking their own outputs as input.
-In evaluating the model, we take advantage of the fact that recursive structures of this kind define potentially infinite languages by testing the model on strings that are longer and more complex than any seen in testing.
-
-We again test this phenomenon within the framework of MacCartney and Manning-style entailment reasoning, but we replace the unanalyzed symbols from the previous experiment with expressions that involve recursive structure.
-To define these expressions, we turn to propositional logic, a relatively simple logic in which each variable represents either \ii{true} or \ii{false}.
-We generate data of the form seen in Figure \ref{prop-figure}: strings of arbitrary length consisting of six elementary proposition variables and the operators \ii{and}, \ii{or}, and \ii{not}, arranged in pairs with the logical relations between them specified.
-
-% NOTE: There was some confusion at CSLI and the 224U lecture about whether this was theorem proving or some kind of model checking. Hopefully this is explicit enough.
-Socher et al.
-\cite{socher2012semantic} have previously demonstrated the learning of a logic in a matrix-vector RNN model somewhat similar to our own, but the logic discussed here is substantially stronger, and a much better approximation of the kind of structure that is needed for natural language.
-The logic learned in that experiment is boolean, wherein the atomic symbols are simply the values 0 and 1, rather than variables over those values.
-While learning the operators of that logic is not trivial, the ouptuts of each operator can be represented accurately by a single bit.
-The statements of propositional logic learned here describe conditions on the truth values of propositions where those truth values are not known.
-As opposed to the two-way contrasts seen in \cite{socher2012semantic}, this logic distinguishes between 64 (2^6) possible assignments of truth values, and expressions of this logic define arbitrary conditions on these possible assignments, for a total of 2^{64} ($\approx 10^{20}$) possible statements that the recursive model needs to be able to distinguish.
-To frame this distinction in another way, the relational statements in our data our theorems about the relations between statements in the logic tested in \cite{socher2012semantic}.
-
-We randomly generate pairs of parentheses-bracketed statements of the logic and then randomly divide the results into training and test data sets.
-To compute the relation between each pair of statements, we exhaustively enumerate the sets of assignments of truth values to proposition variables that would satisfy each of the statements and then convert the set-theoretic relation between those assignments into one of the seven relations.
-If we do not implement any constraint that the two statements being compared are similar in any way, the generated data consists in large part of statements in which the two refer to largely separate subsets of the six variables, and to which we will nearly always assign the \# relation.
-In an effort to balance the distribution of relation labels without departing from the basic task of modeling propositional logic, we disallow individual pairs of statements from referring to more than four of the six proposition variables.
-
-We deduplicate the data and discard pairs in which either statement is a tautology or contradiction (a statement that is true of either all or no possible assignments), for which none of the seven relation labels can accurately apply.
-We then divide the generated pairs into size bins based on the number of logical operators (\ii{and}, \ii{or}, or \ii{not}) in the larger of the two pairs being compared, and discard examples of size greater than twelve by this measure.
-Finally, we randomly sample 15\% of each bin for a held out test set.
-
-We trained both the RNN and RNTN models on the data of size four or less (65k pairs), and tested it on examples of up to size 12 (44k pairs).
-We initialized the model parameters randomly, including the vector representations of the six variables.
-
-The results are shown in Figure \ref{prop-results}.
-In tuning, we found that the RNN model was approximately optimal with 45 dimensional vector representations, and the RNTN model was approximately optimal with 25 dimensions.
-We fixed the size of the feature vector for the classifier at 75 dimensions.
-We found that the RNTN model was able to perform almost perfectly on unseen small test examples, with accuracy above 99\% below size four.
-After depth four, performance gradually falls with increasing size.
-The RNN model did not perform well, reaching only 88.2\% accuracy on the smallest test examples, and declining from there to near-baseline performance at size 12.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=5.5in]{recursion\string_results.eps}
-\end{center}
-
-\caption{Model performance by expression size.  \label{prop-results}} 
-\end{figure}
-
-The performance of the RNTN model on small unseen test examples indicates that it learned a correct approximation of the underlying logic.
-It appears that this approximation is accurate enough to yield correct answers when the composition layer is only applied a small number of times, but that the error in the approximation grows with depth, resulting in the gradually dropping performance.
-This is not necessarily a significant flaw in the model, since it remains possible that a less lossy approximation of the same logic would do no better on the training objective, since that objective only considers short strings and includes a regularization term.
-Whether a more accurate encoding could be learned from a larger set of training data remains a quention for future work.
-In contrast, the RNN model did not appear to be able to learn a correct approximation of the logic.
+The performance of the RNTN model on small unseen test examples
+indicates that it learned a correct approximation of the underlying
+logic. It appears that this approximation is accurate enough to yield
+correct answers when the composition layer is only applied a small
+number of times, but that the error in the approximation grows with
+depth, resulting in the gradually dropping performance. This is not
+necessarily a significant flaw in the model, since it remains possible
+that a less lossy approximation of the same logic would do no better
+on the training objective, since that objective only considers short
+strings and includes a regularization term. Whether a more accurate
+encoding could be learned from a larger set of training data remains a
+quention for future work. In contrast, the RNN model did not appear to
+be able to learn a correct approximation of the logic.

### 2014-06-03 writing/S'14 paper/quantifiers.tex
Revision to quantifiers section

@@
-\section{Reasoning with natural language quantifiers}
+\section{Reasoning with natural language quantifiers}\label{sec:quantifiers}
 
-Even to the extent that RNTN models can handle functional meanings in the form of the operators of propositional logic, this is not a guarantee that these models can handle functional meanings of the forms seen in natural language.
-As a first step towards investigating the latter, we attempt to directly measure the degree to which RNNs are able to develop representations for natural language quantifiers like \ii{some} and \ii{all} that are adequate for inference.
-Quantification is far from the only place in natural language where complex functional meanings are found, but it is a natural starting point, since it can be tested in setences whose structures are otherwise quite simple, and since it has formed a standard case study in prior formal work on natural language inference.
+Even to the extent that RNTN models can handle functional meanings in
+the form of the operators of propositional logic, this is not a
+guarantee that these models can handle functional meanings of the form
+seen in natural language. As a first step towards investigating this
+issue, we attempt to directly measure the degree to which RNNs are
+able to develop suitable representations for natural language
+quantifiers like \ii{some} and \ii{all}. Quantification is far from
+the only place in natural language where complex functional meanings
+are found, but it is a natural starting point, since it can be tested
+in sentences whose structures are otherwise quite simple, and since it
+has formed a standard case study in prior formal work on natural
+language inference.
 
@@
 
-This experiment replicates similar work described in \citet{bowman2013can}, which found that RNTNs can learn to reason well with quantifier meanings given sufficient training data.
-This paper replaces the partially manually annotated data in that paper with data that is generated directly using the logical system that we hope to model, yielding results that we believe to be substantially more straightforward to interpret.
+This experiment replicates similar work described in
+\cite{bowman2013can}, which found that RNTNs can learn to reason well
+with quantifier meanings given sufficient training data. This paper
+replaces the partially manually annotated data in that paper with data
+that is generated directly using the logical system that we hope to
+model, yielding results that we believe to be substantially more
+straightforward to interpret.
 
-Our data consists of pairs of sentences generated from a small artificial grammar.
-Each sentence contains a quantifier, a noun, which may be negated, and an intransitive verb which may be negated.
-We use the basic quantifiers \ii{some}, \ii{most}, \ii{all}, \ii{two}, and \ii{three}, and each of their duals over negation \ii{no}, \ii{not-all}, \ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}.
-We also include five nouns, four intransitive verbs, and the negation symbol \ii{not}.
-In order to be able to define relations between sentences with differing lexical items, we define the lexical relations between each noun--noun pair, each verb--verb pair, and each quantifier--quantifier pair.
+Our data consists of pairs of sentences generated from a small
+artificial grammar. Each sentence contains a quantifier, a noun, which
+may be negated, and an intransitive verb which may be negated. We use
+the basic quantifiers \ii{some}, \ii{most}, \ii{all}, \ii{two}, and
+\ii{three}, and each of their duals over negation \ii{no},
+\ii{not-all}, \ii{not-most}, \ii{less-than-two}, and
+\ii{less-than-three}. We also include five nouns, four intransitive
+verbs, and the negation symbol \ii{not}. In order to be able to define
+relations between sentences with differing lexical items, we define
+the lexical relations between each noun--noun pair, each verb--verb
+pair, and each quantifier--quantifier pair. The grammar accepts
+aligned pairs of sentences of this form and calculates the natural
+logic relationship between them.  Some examples of these data are
+provided in Table~\ref{examplesofdata}.  As in previous sections, the
+goal of learning is then to assign these relational labels accurately
+on unseen cases.
 
@@
 
-To assign relation labels to sentence pairs, we built a small task-specific implemenation of MacCartney's logic that can accurately label sentences of this restricted language.
-The logic is not able to derive all intuitively true relations of this language, and fails to derive a single unique relation for certain types of statement, including De Morgean's laws (e.g.
-\ii{(all pets) growl $\natneg$ (some pet) (not growl)}), and we simply discard these examples.
-Exhaustively generating the valid sentences under this grammar and choosing those to which a relation label can be assigned
-yields 66k sentence pairs. Some examples of these data are provided in Table \ref{examplesofdata}.
+% To assign relation labels to sentence pairs, we built a small
+% task-specific implemenation of MacCartney's logic that can
+% accurately label sentences of this restricted language. The logic is
+% not able to derive all intuitively true relations of this language,
+% and fails to derive a single unique relation for certain types of
+% statement, including De Morgean's laws (e.g. \ii{(all pets) growl
+% $\natneg$ (some pet) (not growl)}), and we simply discard these
+% examples. Exhaustively generating the valid sentences under this
+% grammar and choosing those to which a relation label can be assigned
+% yields 66k sentence pairs. Some examples of these data are provided
+% in Table~\ref{examplesofdata}.
 
-\begin{table}\small\centering
-\begin{tabular}{|l|}
-\hline
-\ii{(most warthogs) walk $\natneg$ (not-most warthogs) walk}\\
-\ii{(most mammals) move $\#$ (not-most (not turtles)) move}\\
-\ii{(most (not pets)) (not swim) $\sqsupset$ (not-most (not pets)) move}\\
-\hline
-\ii{(no turtles) (not growl) $\|$ (no turtles) (not swim)}\\
-\ii{(no warthogs) swim $\sqsupset$ (no warthogs) move}\\
-\ii{(no warthogs) move $\sqsubset$ (no (not reptiles)) swim}\\
-\hline
+\begin{table}[htp]
+  \centering
+  \begin{tabular}{l c l}
+    \toprule
+    (most warthogs) walk         & $\natneg$ & (not-most warthogs) walk\\
+    (most mammals) move          & $\natind$ &  (not-most (not turtles)) move\\
+    (most (not pets)) (not swim) & $\natrev$ & (not-most (not pets)) move 
+    \\[2ex]    
+    (no turtles) (not growl)     & $\natalt$ & (no turtles) (not swim)\\
+    (no warthogs) swim           & $\natrev$ & (no warthogs) move\\
+    (no warthogs) move           & $\natfor$ & (no (not reptiles)) swim\\
+    \bottomrule
   \end{tabular}
-\caption{Sample data involving two different quantifier pairs.\label{examplesofdata}}
+  \caption{Sample data involving two different quantifier pairs.}
+  \label{examplesofdata}
 \end{table}
 
-We evaluate the model using two experimental settings.
-In the simpler setting we randomly sample 85\% of the data and evaluate on the remaining 15\%.
-In this setting, the model is being asked to learn a complete reasoning system for the limited language and logic presented in the training data, but it is not being asked to generalize to test examples that are substantially different from those it was trained on.
-Crucially though, to succeed on this task, the model must be able to recognize all of the lexical relations between the nouns, verbs, and quantifiers and how they interact.
+We evaluate the model using two experimental settings. In the simpler
+setting, we randomly sample 85\% of the data and evaluate on the
+remaining 15\%. In this setting, the model is being asked to learn a
+complete reasoning system for the limited language and logic presented
+in the training data, but it is not being asked to generalize to test
+examples that are substantially different from those it was trained
+on. Crucially though, to succeed on this task, the model must be able
+to recognize all of the lexical relations between the nouns, verbs,
+and quantifiers and how they interact. For instance, it might see
+\eqref{p1} and \eqref{p2} in training and, from that information,
+determine \eqref{p3}.
+%
+\begin{gather}
+  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
+  \\
+  \text{(all lizard) reptile} \natrev  \text{(some lizard) animal}\label{p2}
+  \\
+  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
+\end{gather}
 
-While our primary interest is in discovering the extent to which the model can learn to encode the logic given an arbitrary amount of data, we are also interested in the degree to which the model can infer a correct representation from the logic from more constrained training data.
-In particular, we propose to segment the sentence pairs according to which quantifiers appear in each pair, and then hold out one such pair for testing.
-We hypothesize that a model that can efficiently learn to represent a logic should be able to construct an accurate representation of each held out quantifier from the way that it interacts with the other nine quantifiers which are not held out.
-Since running this experiment requires choosing a pair of quantifiers to hold out before training, the resource demans of training prevent us from testing each of the 55 possible possible pairs of quantifiers, and we choose only four pairs to test on.
-We attempt choose three pairs of differing quantifiers (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and \ii{all}/\ii{some}) to represent six different quantifiers in the test data, choosing pairs that maximize the diversity of relation labels appearing in the test data, and additionally choose one pair in which a quantifier is paired with itself (\ii{no}/\ii{no}).
+\mynote{Discuss results of this experiment}
 
-\begin{table}\small\centering
-\begin{tabular}{|l|lll|}\hline
-\textbf{Data} & \textbf{Most frequent class*} & \textbf{TODO dim RNN} & \textbf{16 dim RNTN}\\\hline
-\textsc{all split}	& 35.4\% &	TODO\%&	88.1\% \\\hline
-\textsc{pair two/less-than-two}	& 59.8\% &	TODO\% &	92.4\% \\
-\textsc{pair not-all/not-most}	&0\% &	   TODO\%  &	82.1\% \\
-\textsc{pair all/some}	& 0\%& TODO\%  &	80.4\% \\
-\textsc{pair no/no}	& 30.8\% &	TODO\% &	100\% \\
-\hline
+While our primary interest is in discovering the extent to which our
+models' can learn to encode the logic given an arbitrary amount of
+data, we are also interested in the degree to which they can infer a
+correct representation from the logic from more constrained training
+data. To this end, we segment the sentence pairs according to which
+quantifiers appear in each pair, and then hold out one such pair for
+testing. We hypothesize that a model that can efficiently learn to
+represent a logic should be able to construct an accurate
+representation of each held out quantifier from the way that it
+interacts with the other nine quantifiers which are not held
+out. Since running this experiment requires choosing a pair of
+quantifiers to hold out before training, the resource demands of
+training prevent us from testing each of the 55 possible possible
+pairs of quantifiers, and we choose only four pairs to test on.  Three
+of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
+\ii{all}/\ii{some}) were chosen because they determine the most
+diverse set of logical relationship in our data, thereby providing the
+most complex training and testing scenario possible. The fourth is a
+self-pair (\ii{no}/\ii{no}), as a check to make sure the model does
+perfectly on maximally easy instances.
+
+\begin{table}[htp]
+  \centering
+  \begin{tabular}{ l rrr }
+    \toprule
+    Data & Most frequent class & TODO dim RNN (\%) & 16 dim RNTN (\%)\\
+    \midrule
+    \textsc{all split}	& 35.4 &	TODO&	88.1 
+    \\[1ex]    
+    \textsc{pair two/less-than-two}	& 59.8 &	TODO &	92.4 \\
+    \textsc{pair not-all/not-most}	&0 &	   TODO  &	82.1 \\
+    \textsc{pair all/some}	& 0& TODO  &	80.4 \\
+    \textsc{pair no/no}	& 30.8 &	TODO &	100 \\
+    \bottomrule
   \end{tabular}
-
-\caption{Quantifier experiment performance. *Most frequent class accuracy is measured against the most frequent class in the training data, \#.\label{resultstable}}
+  \caption{Quantifier experiment performance. 
+    Most frequent class accuracy is measured against the most frequent class in the training data, $\natind$.}
+  \label{resultstable}
 \end{table}
 
+\mynote{
 The RNN model was approximately optimal with N dimensional word representations and an M dimensional comparison layer. The RNTN was approximately optimal with N and M dimensions, respectively.
 % TODO: Update dimensionality
-
- (Note: Models still running.
-Results might improve and this para is likely to change.) The strong performance by the RNN on the \textsc{pair no/no} experiment suggests that it is at least plausible that this model is able to handle quantifiers correctly given sufficient training data, but the sub 90\% results on most of the training settings suggest that the data provided in these experiments are not generally sufficient, and that the model may not be able to learn to reason with quantifiers with high accuracy.
-However, the fact that both models perform far above baseline is promising, and the question of how much data is necessary to accurately capture quantifier behavior in a na\"ive model remains open.
+(Note: Models still running. Results might improve and this para is
+likely to change.) The strong performance by the RNN on the
+\textsc{pair no/no} experiment suggests that it is at least plausible
+that this model is able to handle quantifiers correctly given
+sufficient training data, but the sub 90\% results on most of the
+training settings suggest that the data provided in these experiments
+are not generally sufficient, and that the model may not be able to
+learn to reason with quantifiers with high accuracy. However, the fact
+that both models perform far above baseline is promising, and the
+question of how much data is necessary to accurately capture
+quantifier behavior in a na\"ive model remains open.
 
 Discussion of RNN results (and updated discussion of better RNTN results) to be filled in as results arrive.\\...\\... % TODO
+}

### 2014-06-03 writing/S'14 paper/discussion.tex
Revised closing discussion

--- 2014-06-02 writing/S'14 paper/discussion.tex
@@
-\subsection{General discussion}  % TODO
+\section{General discussion}\label{sec:discussion}
 
-Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...
+This paper evaluated two RNNs in a series of three increasingly
+challenging interpretive tasks involving natural language inference:
+the core relational algebra of natural logic with entailment and
+exclusion; recursive propositional logic structures; and statements
+involving quantification and negation. The results suggest that RNTNs,
+but not plain RNNs, have the capacity to meet the challenges of these
+tasks with reasonably-sized training sets. These positive results are
+promising for the future of learned representation models in the
+applied modeling of meaning.
+
+Of course, challenges remain. In terms of our experimental data, even
+the RNTN falls short of perfection in our more complex tasks, with
+performance falling off steadily as the depth of recursion grows. It
+remains to be seen whether these deficiences can be overcome with
+improvements to the model, the optimization procedures, or the
+linguistic representations
+\cite{sochergrounded,kalchbrenner2014convolutional}. In addition,
+there remain subtle questions about how to fairly assess whether these
+models have truly generalized in the way we want them to --- what
+kinds of training and testing data will give them a chance to simulate
+the target logical functions without revealing the answer to them in a
+way that encourages overfitting? Finally, we have only scratched the
+surface of the logical complexity of natural language; in future
+experiments, we hope to test sentences with embedded quantifiers,
+multiple interacting quantifiers, relative clauses, and other kinds of
+recursive structure. Nonetheless, the rapid progress we have seen on
+these models in recent years provides ample reason to be optimistic
+that they can rise to the challenges of natural language semantics.
+
+% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators.
+This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic.
+The positive results that we have fonud, however, are extremely promising for the future of learned representation models in the applied modeling of meaning.
+We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators.
+We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.
+
+% There is ample room to build on these results.
+In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses.
+In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.
 

### 2014-06-03 writing/S'14 paper/join.tex
Revised recursion section
Tweaks

@@
-\section{Relation composition}
+\section{Relation composition}\label{sec:join}
 
@@
 The full set of sound inferences of this form is summarized in
-Table~\ref{tab:jointable}.
+Table~\ref{tab:jointable}; blank cells correspond to pairs of relations
+for which no valid inference can be drawn in our logic.
 
@@
     \begin{picture}(5,5.5)
-      \labelednode{2.50}{5}{a}{1,2,3}
+      \labelednode{2.50}{5}{}{1,2,3}
 
@@
 
-      \labelednode{0.75}{3.5}{b}{1,2}
+      \labelednode{0.75}{3.5}{a,b}{1,2}
       \labelednode{2.50}{3.5}{c}{1,3}
@@
 
-      \labelednode{0.75}{2}{e}{1}
-      \labelednode{2.50}{2}{f}{2}
-      \labelednode{4.25}{2}{g}{3}
+      \labelednode{0.75}{2}{e,f}{1}
+      \labelednode{2.50}{2}{}{2}
+      \labelednode{4.25}{2}{g,h}{3}
 
@@
 
-      \labelednode{2.5}{0.5}{h}{}
+      \labelednode{2.5}{0.5}{}{}
     \end{picture}
-    \caption{Simple boolean model.}
+    \caption{Simple boolean model. The letters name the sets. Not all sets have names, and
+    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
   \end{subfigure}
@@
                     & $b \natcov d$ \\
-                    & $b \natrev e$ \\
+                    & \strikeout{$b \natrev e$} \\
       $c \natcov d$ &               \\
       $c \natrev e$ &               \\
-                    & \strikeout{$c \natneg f$} \\
+                    & \strikeout{$c \nateq f$} \\
       $c \natrev g$ &               \\ 

### 2014-06-03 writing/S'14 paper/join.tex
Revised join table

@@
 The full set of sound inferences of this form is summarized in
-Table~\ref{tab:jointable}; blank cells correspond to pairs of relations
-for which no valid inference can be drawn in our logic.
+Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
+of relations for which no valid inference can be drawn in our logic.
 
@@
   \centering  
-  \setlength{\arraycolsep}{12pt}
+  \setlength{\arraycolsep}{8pt}
   \renewcommand{\arraystretch}{1.1}
-  %\newcommand{\rcg}{\cellcolor[gray]{.9}}
-  \newcommand{\rcg}{}
+  \newcommand{\UNK}{\cdot}  
   $\begin{array}[t]{c@{ \ }|*{7}{c}|}
@@
     \cline{2-8}
-    \nateq   & \rcg\nateq & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \natind\\
-    \natfor  & \natfor    & \rcg\natfor & \natind     & \natalt    & \natalt     & \natind     & \natind\\
-    \natrev  & \natrev    & \natind     & \rcg\natrev & \natcov    & \natind     & \natcov     & \natind\\
-    \natneg  & \natneg    & \natcov     & \natalt     & \rcg\nateq & \natrev     & \natfor     & \natind\\
-    \natalt  & \natalt    & \natind     & \natalt     & \natfor    & \rcg\natind & \natfor     & \natind\\
-    \natcov  & \natcov    & \natcov     & \natind     & \natrev    & \natrev     & \rcg\natind & \natind\\
-    \natind  & \natind    & \natind     & \natind     & \natind    & \natind     & \natind     & \rcg\natind \\
+    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
+    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
+    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
+    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
+    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
+    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
+    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
     \cline{2-8}

### 2014-06-04 writing/S'14 paper/recursion.tex
Revision to quantifiers section
Intermediate paper update.

@@
-\section{Recursive structure}
+\section{Recursive structure}\label{sec:recursion}
 
@@
 therefore rely heavily on recursive structures, and we expect that
-accurate computational models will have to come to grips with it as
+accurate computational models will have to come to grips with them as
 well. The present section pursues this question for our two classes of
@@
 from that experiment with expressions that involve recursive
-structure. The result is a truth-functionally complete classical
+structure. These expressions represent a truth-functionally complete classical
 propositional logic: each atomic variable denotes either $\True$ or
@@
 Table~\ref{tab:pl} defines this logic, and Table~\ref{tab:plexs} gives
-some examples of natural logic relational statements that we can make
+some examples of relational statements that we can make
 in these terms. To compute these relations between statements, we
@@
 then convert the set-theoretic relation between those assignments into
-one of the seven relations in Table~\ref{b-table}. This is equivalent
-to proving theorems about these statements model-theoretically.
-
-\newcommand{\plneg}{\mathop{\textit{not}}}
-\newcommand{\pland}{\mathbin{\textit{and}}}
-\newcommand{\plor}{\mathbin{\textit{or}}}
+one of the seven relations in Table~\ref{b-table}. As a result, each relational
+statement represents a valid theorem of the propositional logic.
 
@@
 possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
-possible statements that the recursive model needs to be able to
-distinguish.
+possible statements that the intermediate vector representations need 
+to be able to distinguish.
 
-For our experiments, we randomly generate large sets of syntactically
-unique formulae and compute the natural logic relation between them.
+For our experiments, we randomly generate a large set of  unique pairs 
+of formulae and compute the relation that holds for each pair.
 We discard pairs in which either statement is a tautology or
@@
 variables.
-
-% We deduplicate the data and discard pairs in which either statement is
-% a tautology or contradiction (a statement that is true of either all
-% or no possible assignments), for which none of the seven relation
-% labels can accurately apply. We then divide the generated pairs into
-% size bins based on the number of logical operators (\ii{and}, \ii{or},
-% or \ii{not}) in the larger of the two pairs being compared, and
-% discard examples of size greater than twelve by this measure. Finally,
-% we randomly sample 15\% of each bin for a held out test set.
 
@@
   \includegraphics[width=5.5in]{recursion\string_results.eps}
-  \caption{Model performance by expression size.}  
+  \caption{Model performance on propositional logic, by expression size.}  
   \label{prop-results}

### 2014-06-04 writing/S'14 paper/intro.tex
Intro updates

@@
 including sentiment analysis \cite{socher2011semi,socher2013acl1},
-image description \cite{sochergrounded}, and paraphase detection
+image description \cite{sochergrounded}, and paraphrase detection
 \cite{Socher-etal:2011:Paraphrase}. These results are encouraging
@@
 same results as grammars based in logical forms
-\cite{ZetCol:2005,LiangJordan:2013} when it comes to core semantic
+\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013} when it comes to core semantic
 concepts like quantification, entailment and contradiction. To date,
@@
 networks (RNTNs, \cite{socher2013acl1}) --- can learn those
-relationships correctly. In our first experiment, we evaluate the
-ability of these models to learn the core relational algebra of
-natural logic from data. Our second experiment extends this evaluation
-to cover relations between complex recursive structures like `(A or
-B)' and `not(not A and not B)', and our third experiment involves
-relations between quantified statements like `every reptile walks' and
-`every turtle moves'. We find that the plain RNN achieves only mixed
-results on all three experiments, whereas the stronger RNTN models
-generalized well in every case, suggesting that it has in fact
-learned, or at least learned to simulate, our target logical
+relationships correctly. In our first experiment
+(Section~\ref{sec:join}), we evaluate the ability of these models to
+learn the core relational algebra of natural logic from data. Our
+second experiment (Section~\ref{sec:recursion}) extends this
+evaluation to cover relations between complex recursive structures
+like $(a \plor b)$ and $\plneg(\plneg a \pland \plneg b)$, and our
+third experiment (Section~\ref{sec:quantifiers}) involves relations
+between quantified statements like `every reptile walks' and `every
+turtle moves'. We find that the plain RNN achieves only mixed results
+in all three experiments, whereas the stronger RNTN models generalized
+well in every case, suggesting that it has in fact learned, or at
+least learned to simulate, our target logical
 concepts. % TODO: Update with current results
@@
 by showing that these models can learn to correctly represent entailment representations between sentences.
-I then show that these models can capture the meanings of recursive structurers accurately up to a sufficient depth.
+I then show that these models can capture the meanings of recursive structures accurately up to a sufficient depth.
 I finally close with a brief demonstration of the ability of these models to reason over short natural language sentences involving quantifiers.
@@
 It should be noted that this is not a balanced classification problem.
-If arbitrary pairs of statements are chosen for comparison in almost any domain of natural langauge, the minimally informative \# relation will apply between them.
+If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them.
 % TODO: How much should we say about this?

### 2014-06-04 writing/S'14 paper/recursion.tex
Proofed sec. 4

@@
 \section{Recursive structure}\label{sec:recursion}
-
-% TODO: Cite Chomsky/Hauser/Fitch? -- no, not appropriate here --- CP.
 
@@
 number of times, but that the error in the approximation grows with
-depth, resulting in the gradually dropping performance. This is not
+increasing depth (and with the increasing complexity of the expressions),
+ resulting in the gradually dropping performance. This is not
 necessarily a significant flaw in the model, since it remains possible

### 2014-06-04 writing/S'14 paper/intro.tex

@@
 \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013} when it comes to core semantic
-concepts like quantification, entailment and contradiction. To date,
+concepts like quantification, entailment, and contradiction. To date,
 experimental investigations of these concepts using distributed
@@
 For robust natural language understanding, we must be able to model
-these phenomena in full generality on complex linguistic structures.
+these phenomena in their full generality on complex linguistic structures.
 

### 2014-06-04 writing/S'14 paper/intro.tex
First proof of intro and methods

@@
 third experiment (Section~\ref{sec:quantifiers}) involves relations
-between quantified statements like `every reptile walks' and `every
-turtle moves'. We find that the plain RNN achieves only mixed results
+between quantified statements like \ii{every reptile walks} and \ii{every
+turtle moves}. We find that the plain RNN achieves only mixed results
 in all three experiments, whereas the stronger RNTN models generalized

### 2014-06-04 writing/S'14 paper/methods.tex
First proof of intro and methods

@@
 via specific combination functions \cite{Partee84,Janssen97}. In our
-distributional model, word meanings are vectors of length $N$. The
+distributed setting, word meanings are vectors of length $N$. The
 combination function concatenates pairs of them to form vectors of
@@
 For a composition layer, we evaluate models with both the plain neural
-network layer function (eq.~\ref{rnn}) and the RNTN layer function
-proposed in \citet{chen2013learning} (eq.~\ref{rntn}). A sigmoid
+network layer function \eqref{rnn} and the RNTN layer function
+proposed in \citet{chen2013learning} \eqref{rntn}. A sigmoid
 nonlinearity (element-wise $\tanh$) is applied to the output of either
@@
 \begin{gather} \label{rnn}
-y_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + b_i)\\ % TODO: Add column vectors?
+y_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
 \label{rntn}
@@
 basic structure, but with the addition of a tensor $\mathbf{A}$,
-dimension $2N \times N \times N$, modeling interactions between the
+dimension $N \times N \times N$, modeling interactions between the
 child vectors.
 
-The comparison layer uses the same type function template as the
+The comparison layer uses the same kind of function template as the
 composition function (either an NN layer or an NTN layer) with
-independantly learned parameters and a separate nonlinearity function.
+independently learned parameters and a separate nonlinearity function.
 Rather than use a $\tanh$ nonlinearity here, we found a substantial

### 2014-06-04 writing/S'14 paper/discussion.tex
Revised conclusion

@@
 promising for the future of learned representation models in the
-applied modeling of meaning.
+applied modeling of compositional semantics.
 
@@
 performance falling off steadily as the depth of recursion grows. It
-remains to be seen whether these deficiences can be overcome with
+remains to be seen whether these deficiencies can be overcome with
 improvements to the model, the optimization procedures, or the
@@
 there remain subtle questions about how to fairly assess whether these
-models have truly generalized in the way we want them to --- what
-kinds of training and testing data will give them a chance to simulate
-the target logical functions without revealing the answer to them in a
-way that encourages overfitting? Finally, we have only scratched the
-surface of the logical complexity of natural language; in future
-experiments, we hope to test sentences with embedded quantifiers,
-multiple interacting quantifiers, relative clauses, and other kinds of
-recursive structure. Nonetheless, the rapid progress we have seen on
-these models in recent years provides ample reason to be optimistic
-that they can rise to the challenges of natural language semantics.
+models have truly generalized in the way we want them to. There is a
+constant tension between giving the models training data that gives
+them a chance to learn the target logical functions and revealing the
+answer to them in a way that leads to overfitting. The underlying
+logical theories provide only limited guidance on this point. Finally,
+we have only scratched the surface of the logical complexity of
+natural language; in future experiments, we hope to test sentences
+with embedded quantifiers, multiple interacting quantifiers, relative
+clauses, and other kinds of recursive structure. Nonetheless, the
+rapid progress the field has made with these models in recent years
+provides ample reason to be optimistic that they can be trained to
+meet the challenges of natural language semantics.
 
@@
 This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic.
-The positive results that we have fonud, however, are extremely promising for the future of learned representation models in the applied modeling of meaning.
+The positive results that we have found, however, are extremely promising for the future of learned representation models in the applied modeling of meaning.
 We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators.

### 2014-06-04 writing/S'14 paper/quantifiers.tex
Proofed sec. 5

@@
 Our data consists of pairs of sentences generated from a small
-artificial grammar. Each sentence contains a quantifier, a noun, which
+artificial grammar. Each sentence contains a quantifier, a noun which
 may be negated, and an intransitive verb which may be negated. We use
 the basic quantifiers \ii{some}, \ii{most}, \ii{all}, \ii{two}, and
-\ii{three}, and each of their duals over negation \ii{no},
+\ii{three}, and their negations \ii{no},
 \ii{not-all}, \ii{not-most}, \ii{less-than-two}, and
@@
 goal of learning is then to assign these relational labels accurately
-on unseen cases.
+to unseen pairs of sentences.
 
@@
 We evaluate the model using two experimental settings. In the simpler
-setting, we randomly sample 85\% of the data and evaluate on the
+setting, \textsc{all split}, we randomly sample 85\% of the data and evaluate on the
 remaining 15\%. In this setting, the model is being asked to learn a
@@
     \midrule
-    \textsc{all split}	& 35.4 &	TODO&	88.1 
+    \textsc{all split}	& 35.4 &	60.4&	94.4
     \\[1ex]    
-    \textsc{pair two/less-than-two}	& 59.8 &	TODO &	92.4 \\
-    \textsc{pair not-all/not-most}	&0 &	   TODO  &	82.1 \\
-    \textsc{pair all/some}	& 0& TODO  &	80.4 \\
-    \textsc{pair no/no}	& 30.8 &	TODO &	100 \\
+    \textsc{pair two/less-than-two}	& 59.8 &	76.4 &	97.1 \\
+    \textsc{pair not-all/not-most}	&0 &	   65.6  &	79.2 \\
+    \textsc{pair all/some}	& 0& 60.7  &	79.5 \\
+    \textsc{pair no/no}	& 30.8 &	65.8 &	100 \\
     \bottomrule

### 2014-06-04 writing/S'14 paper/quantifiers.tex
Proofed sec. 5

@@
   \\
-  \text{(all lizard) reptile} \natrev  \text{(some lizard) animal}\label{p2}
+  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
   \\
@@
 most complex training and testing scenario possible. The fourth is a
-self-pair (\ii{no}/\ii{no}), as a check to make sure the model does
-perfectly on maximally easy instances.
+self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
+handles equality.
+
+% Further explanation for motivation for no/no?
 
@@
     \midrule
-    \textsc{all split}	& 35.4 &	60.4&	94.4
+    \textsc{all split}	& 35.4 &	60.4&	\textbf{94.4}
     \\[1ex]    
-    \textsc{pair two/less-than-two}	& 59.8 &	76.4 &	97.1 \\
-    \textsc{pair not-all/not-most}	&0 &	   65.6  &	79.2 \\
-    \textsc{pair all/some}	& 0& 60.7  &	79.5 \\
-    \textsc{pair no/no}	& 30.8 &	65.8 &	100 \\
+    \textsc{pair two/less-than-two}	& 59.8 &	76.4 &	\textbf{97.1} \\
+    \textsc{pair not-all/not-most}	&0 &	   65.6  &	\textbf{79.2} \\
+    \textsc{pair all/some}	& 0& 60.7  &	\textbf{79.5} \\
+    \textsc{pair no/no}	& 30.8 &	65.8 &	\textbf{100} \\
     \bottomrule
@@
   \caption{Quantifier experiment performance. 
-    Most frequent class accuracy is measured against the most frequent class in the training data, $\natind$.}
+    The most frequent class baseline accuracy reflects the most frequent class in the training data, $\natind$.}
   \label{resultstable}
@@
 (Note: Models still running. Results might improve and this para is
-likely to change.) The strong performance by the RNN on the
-\textsc{pair no/no} experiment suggests that it is at least plausible
+likely to change.) The perfect performance by the RNTN on the
+\textsc{pair no/no} experiment and its strong performance on
+\textsc{pair two/less-than-two} suggests that it is at least plausible
 that this model is able to handle quantifiers correctly given
-sufficient training data, but the sub 90\% results on most of the
-training settings suggest that the data provided in these experiments
-are not generally sufficient, and that the model may not be able to
-learn to reason with quantifiers with high accuracy. However, the fact
+sufficient training data, but the sub 90\% results on two of the
+training settings suggest that the model may not be able to
+generalize from data this impoverished in general. However, the fact
 that both models perform far above baseline is promising, and the

### 2014-06-04 writing/S'14 paper/join.tex
Proofed sec. 3

@@
 presented here from a finite amount data, it must be able to learn to
-deduce new relations from seen relations in a sound and complete
+deduce new relations from seen relations in a sound
 manner. The simplest such deductions involve atomic statements using
@@
 Figure~\ref{lattice-figure} depicts a small model of this form. The
-lattice structure gives the full model, in which all the statements on
-the right can be proved. The highlighted nodes represent a
-hypothetical training set drawn from this larger space, which allows
-us to prove only a subset of the full set of true statements.
+lattice structure gives the full model, for which all the statements on
+the right are valid. We divide these statements evenly into training and
+test sets, and remove from the test set those statements which cannot be 
+proven from the training statements using the logic depicted in 
+Figure~\ref{lattice-figure}.
 
@@
   \end{subfigure}  
-  % \begin{tabular}{ll|lll}
-  %   $a := \{1, 3, 4, 5, 6\}	$&~~~&~~~& $a \equiv a$	    	& $a~\#~c$	\\
-  %   $b := \{0, 3, 5, 6\}	$&~~~&~~~& $e \sqsupset c$	&$b \smallsmile c$		\\
-  %   $c := \{1, 2, 3, 4\}	$&~~~&~~~& $d \smallsmile e$	& $b~\#~d$		\\
-  %   $d := \{0, 4\}		$&~~~&~~~& $a \sqsubset e$	& $ d \equiv d$	\\
-  %   $e := \{1, 2, 3, 4, 5, 6\}$&~~~&~~~& $a~\#~b$		& ... 	\\
-  % \end{tabular}
   \caption{Some sample randomly generated sets, and some of the relations defined between them.}
@@
 \end{figure} 
+% TODO: Recheck strikethrough distribution.
 
@@
 In our experiments, we create 80 randomly generated sets drawing from
-the same domain of seven element. This yields a data set consisting of
+a domain of seven elements. This yields a data set consisting of
 6400 statements about pairs of entities. 3200 of these pairs are
 chosen as a test set, and that test set is further reduced to the 2960
-examples that can be provably derived from the test data using
-Table~\ref{tab:jointable}.
+examples that can be provably derived from the test data.
 
-% TODO: Say more about symmetry?
 We trained versions of both the RNN model and the RNTN model on these
@@
 worked best with 11 and 75 dimensions, respectively, but was able to
-achieve accuracies of only 90.0\% and 87.\%, respectively.
+achieve accuracies of only 90.0\% and 87.0\%, respectively.
 
@@
 better on a larger universe of sets for which there would be a larger
-set of training data to learn from, but the RNTN is able to achieve
-these effects in the well-studied framework of Section~\ref{methods}.
+set of training data to learn from, but the RNTN is readily able to achieve
+these effects in the setting discussed here.

### 2014-06-04 writing/S'14 paper/intro.tex
First minor tweaks from the review session.

@@
 generate controlled data sets encoding the semantic relationships
-between pairs of expressions and evaluating whether each of two
+between pairs of expressions and evaluate whether each of two
 classes of neural model --- plain RNNs and recursive neural tensor

### 2014-06-04 writing/S'14 paper/methods.tex
First minor tweaks from the review session.

@@
 phrases. Once the entire sentence-level representation has been
-derived, it fed to a classifier trained in a supervised manner.
+derived, it is fed to a classifier trained in a supervised manner.
 
@@
 \begin{gather} \label{rnn}
-y_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
+\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
 \label{rntn}
-y_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{A}^{[1 \ldots N]} \vec{x}^{(r)} + \mathbf{B} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
+\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots N]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
 \end{gather} % TODO: Explain third order tensor parameter?
 %
-Here, the $l$ and $r$ superscripts indicate the vector representations
-for the left and right children of the local structure interpreted as
-$y$. The RNN concatenates them, multiplies them by an $N \times 2N$
+Here, the $\vec{x}^{(l)}$ and $\vec{x}^{(l)}$ are the column vector representations
+for the left and right children of the node, and $\vec{y}$ is the node's output.
+The RNN concatenates them, multiplies them by an $N \times 2N$
 matrix of learned weights, and applies the element-wise non-linearity
 to the resulting vector (plus a bias term). The RNTN has the same
-basic structure, but with the addition of a tensor $\mathbf{A}$,
-dimension $N \times N \times N$, modeling interactions between the
-child vectors.
+basic structure, but with the addition of a learned third-order tensor $\mathbf{T}$,
+dimension $N \times N \times N$, modeling multiplicative interactions between the
+child vectors. Both models include a learned bias vector, represented $\vec{b}$.
 

### 2014-06-04 writing/S'14 paper/quantifiers.tex
First minor tweaks from the review session.

@@
 issue, we attempt to directly measure the degree to which RNNs are
-able to develop suitable representations for natural language
+able to develop suitable representations for the semantics of natural language
 quantifiers like \ii{some} and \ii{all}. Quantification is far from
@@
 
-\mynote{Discuss results of this experiment}
-
 While our primary interest is in discovering the extent to which our
-models' can learn to encode the logic given an arbitrary amount of
+models can learn to encode the logic given an arbitrary amount of
 data, we are also interested in the degree to which they can infer a
-correct representation from the logic from more constrained training
+correct representation for the logic from more constrained training
 data. To this end, we segment the sentence pairs according to which
@@
 of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
-\ii{all}/\ii{some}) were chosen because they determine the most
-diverse set of logical relationship in our data, thereby providing the
-most complex training and testing scenario possible. The fourth is a
+\ii{all}/\ii{some}) were chosen because they provide allow for the most
+different class labels at in the training data. The fourth is a
 self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
@@
   \end{tabular}
-  \caption{Quantifier experiment performance. 
-    The most frequent class baseline accuracy reflects the most frequent class in the training data, $\natind$.}
+  \caption{Quantifier experiment performance. }
   \label{resultstable}
 \end{table}
+
+The results for these experiments are shown at the top of Figure \ref{resultstable}.
+We compare the results to a most frequent class baseline, which reflects the frequency in the test data of most frequent class in the training data, $\natind$.
 
@@
 (Note: Models still running. Results might improve and this para is
-likely to change.) The perfect performance by the RNTN on the
+likely to change.)
+
+The perfect performance by the RNTN on the
 \textsc{pair no/no} experiment and its strong performance on

### 2014-06-04 writing/S'14 paper/recursion.tex
First minor tweaks from the review session.

@@
 
-Recursive structure is an prominent feature of natural languages and
+Recursive structure is a prominent feature of natural languages and
 an important part of their expressivity, as it allows people to use
@@
 encoding could be learned from a larger set of training data remains a
-quention for future work. In contrast, the RNN model did not appear to
+question for future work. In contrast, the RNN model did not appear to
 be able to learn a correct approximation of the logic.

### 2014-06-04 writing/S'14 paper/methods.tex
Adding new Figure 1, updating results on quantifiers.

@@
 
-\begin{figure}[htp]
+\begin{figure}[tp]
   \centering
-  \includegraphics[scale=0.35]{model.eps}
-  \caption{The model structure used to compare \ii{no dog} and \ii{(not all) dog}. 
+ % \includegraphics[scale=0.35]{model.eps}
+
+   \scriptsize
+\begin{picture}(20,7.5)
+
+  \labeledtreenode[2]{9,7}{$P(\sqsubset) = 0.8$}{softmaxcolor}
+  \labeledtreenode[2]{12.5,7}{Softmax classifier}{white}
+  \put(10.25,5.98){\vector(0,1){0.88}}
+
+  \labeledtreenode[7]{6.75,5.5}{all reptiles walk \emph{vs.}~some turtles move}{comparisoncolor}
+  \labeledtreenode[2]{16,5.5}{Comparison N(T)N layer}{white}
+
+  \put(5.5,4.48){\vector(4,1){3.25}}
+  \labeledtreenode{4,4}{all reptiles walk}{compositioncolor}
+
+  \labeledtreenode[2]{9.5,4}{Composition RN(T)N layers}{white}
+
+  \put(3.5,2.98){\vector(2,1){1.7}}
+  \labeledtreenode{2,2.5}{all reptiles}{compositioncolor}
+
+  \put(1.5,1.48){\vector(2,1){1.7}}
+  \labeledtreenode{0,1}{all}{lexcolor}
+
+  \put(5.5,1.48){\vector(-2,1){1.7}}
+  \labeledtreenode{4,1}{reptiles}{lexcolor}
+
+  \put(7.5,2.98){\vector(-2,1){1.7}}
+  \labeledtreenode{6,2.5}{walk}{lexcolor}
+
+  \put(15.5,4.48){\vector(-4,1){3.25}}
+  \labeledtreenode{14,4}{some turtles move}{compositioncolor}
+
+  \put(13.5,2.98){\vector(2,1){1.7}}
+  \labeledtreenode{12,2.5}{some turtles}{compositioncolor}
+
+  \put(11.5,1.48){\vector(2,1){1.7}}
+  \labeledtreenode{10,1}{some}{lexcolor}
+
+  \put(15.5,1.48){\vector(-2,1){1.7}}
+  \labeledtreenode{14,1}{turtles}{lexcolor}
+
+  \put(17.5,2.98){\vector(-2,1){1.7}}
+  \labeledtreenode{16,2.5}{move}{lexcolor}
+
+  \labeledtreenode{5,0.25}{Learned, randomly initialized word vectors}{white}
+\end{picture}
+
+  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
     The same structure is used for both the RNN and RNTN layer functions.} 

### 2014-06-04 writing/S'14 paper/abstract.tex
Interim update.

@@
   meaning have been successful in a wide array of sophisticated
-  language tasks, but it remains an open question whether they can
-  achieve the same results as compositional semantic grammars based in
-  logical forms. We address this question directly by using a logical
+  language tasks, but it remains an open question whether succeed in
+  learning compositional semantic grammars that allow logical deduction. 
+  We address this question directly by using a logical
   grammar to generate controlled data sets encoding the relationships
@@
   results on all three experiments, whereas the stronger RNTN models
-  generalizes well in every setting.
+  generalizes well in every setting. This serves as compelling evidence that
+  RNTNs provide a satisfactory representation of meaning for
+  natural language reasoning.
 \end{abstract}

### 2014-06-04 writing/S'14 paper/intro.tex
Interim update.

@@
 \cite{Socher-etal:2011:Paraphrase}. These results are encouraging
-about the ability of these models to learn compositional semantic
+for the ability of these models to learn compositional semantic
 grammars, but it remains an open question whether they can achieve the

### 2014-06-04 writing/S'14 paper/join.tex
First minor tweaks from the review session.
Interim update.

@@
 
-To test our models' ability to learn this relational structure, we
+To test our systems's ability to learn this relational structure, we
 create small models for our logic in which terms denote sets of
@@
 chosen as a test set, and that test set is further reduced to the 2960
-examples that can be provably derived from the test data.
+examples that can be provably derived from the training data.
 
@@
 ensuring that the model has no information about the terms being
-compared except the relations between them. We found that the RNTN
-model worked best with 11-dimensional vector representations for the
+compared except the relations between them.
+
+We found that the RNTN model worked best with 11-dimensional vector representations for the
 80 sets and a 90-dimensional feature vector for the classifier. This
-model was able to correctly label 99.3\% of the derivable test
-examples, and 99.1\% of the remaining examples. The simpler RNN model
+model was able to correctly label 99.3\% of the provably derivable test
+examples, and 99.1\% of the remaining test examples. The simpler RNN model
 worked best with 11 and 75 dimensions, respectively, but was able to
 achieve accuracies of only 90.0\% and 87.0\%, respectively.
+
+% RNTN log: tue-j-11-6x80-hip.txt
+% Train PER: 0.0021875
+
+% RNN log: tue-j-11-6x80-r-75.txt
+% Train PER: 0.047188
 

### 2014-06-04 writing/S'14 paper/methods.tex
Adding new Figure 1, updating results on quantifiers.
Revised figure 1
Interim update.

--- 2014-06-04 writing/S'14 paper/methods.tex
@@
 \label{rntn}
-\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots N]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
+\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots N]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
 \end{gather} % TODO: Explain third order tensor parameter?
 %
-Here, the $\vec{x}^{(l)}$ and $\vec{x}^{(l)}$ are the column vector representations
+Here, the $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector representations
 for the left and right children of the node, and $\vec{y}$ is the node's output.
@@
 
-\begin{figure}[htp]
+\begin{figure}[tp]
   \centering
-  \includegraphics[scale=0.35]{model.eps}
-  \caption{The model structure used to compare \ii{no dog} and \ii{(not all) dog}. 
+ % \includegraphics[scale=0.35]{model.eps}
+  \input{figure1}
+  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
     The same structure is used for both the RNN and RNTN layer functions.} 

### 2014-06-04 writing/S'14 paper/quantifiers.tex
Adding new Figure 1, updating results on quantifiers.
Interim update.

@@
 of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
-\ii{all}/\ii{some}) were chosen because they provide allow for the most
+\ii{all}/\ii{some}) were chosen because they allow for the most
 different class labels at in the training data. The fourth is a
@@
     \midrule
-    \textsc{all split}	& 35.4 &	60.4&	\textbf{94.4}
+    \textsc{all split}	& 35.4 &	63.9&	\textbf{93.4}
     \\[1ex]    
-    \textsc{pair two/less-than-two}	& 59.8 &	76.4 &	\textbf{97.1} \\
-    \textsc{pair not-all/not-most}	&0 &	   65.6  &	\textbf{79.2} \\
-    \textsc{pair all/some}	& 0& 60.7  &	\textbf{79.5} \\
-    \textsc{pair no/no}	& 30.8 &	65.8 &	\textbf{100} \\
+    \textsc{pair two/less-than-two}	& 59.8 &	79.5 &	\textbf{98.3} \\
+    \textsc{pair not-all/not-most}	&0 &	   62.2  &	\textbf{83.3} \\
+    \textsc{pair all/some}	& 0& 63.6  &	\textbf{81.0} \\
+    \textsc{pair no/no}	& 30.8 &	66.0 &	\textbf{100} \\
     \bottomrule
@@
 
-The results for these experiments are shown at the top of Figure \ref{resultstable}.
+The results for these experiments are shown in Figure \ref{resultstable}.
 We compare the results to a most frequent class baseline, which reflects the frequency in the test data of most frequent class in the training data, $\natind$.

### 2014-06-04 writing/S'14 paper/recursion.tex
Interim update.

@@
 languages, by testing the model on strings that are longer and more
-complex than any seen in testing.
+complex than any seen in training.
 

### 2014-06-05 writing/S'14 paper/join.tex
Few more assorted paper edits.

@@
   \end{array}$
-  \caption{Full set of inferences between natural logic relations.
-    Given $a \mathbin{R} b$ and $b \mathbin{S} c$, where $R$ and $S$
-    are the row and column relations, respectively, and $a$, $b$, and
-    $c$ are arbitrary formulae, the table provides the relation
-    $a \mathbin{T} c$.}
+  \caption{Inference path from premises $a R S$ (row) and $b S c$ (column) to the relation that holds between $a$ and $c$, if any.
+These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
+We assess our model's ability to reproduce such inferential paths.}
   \label{tab:jointable}
@@
 
-We found that the RNTN model worked best with 11-dimensional vector representations for the
+\paragraph{Results.} We found that the RNTN model worked best with 11-dimensional vector representations for the
 80 sets and a 90-dimensional feature vector for the classifier. This

### 2014-06-05 writing/S'14 paper/quantifiers.tex
Few more assorted paper edits.

@@
-\section{Reasoning with natural language quantifiers}\label{sec:quantifiers}
+\section{Reasoning with natural language quantifiers and negation}\label{sec:quantifiers}
 
@@
 
-The results for these experiments are shown in Figure \ref{resultstable}.
+\paragraph{Results.} The results for these experiments are shown in Figure \ref{resultstable}.
 We compare the results to a most frequent class baseline, which reflects the frequency in the test data of most frequent class in the training data, $\natind$.

### 2014-06-05 writing/S'14 paper/recursion.tex
Few more assorted paper edits.

@@
-\section{Recursive structure}\label{sec:recursion}
+\section{Recursive structure in propositional logic}\label{sec:recursion}
 
@@
 
-We trained both the RNN and RNTN models on the data of size four or
+\paragraph{Results.} We trained both the RNN and RNTN models on the data of size four or
 less (65k pairs), and tested it on examples of up to size 12 (44k

### 2014-06-05 writing/S'14 paper/join.tex
Two more items of the paper to do list.

@@
 \paragraph{Results.} We found that the RNTN model worked best with 11-dimensional vector representations for the
-80 sets and a 90-dimensional feature vector for the classifier. This
+80 sets and a 90-dimensional feature vector for the classifier, though the performance of the model was
+not highly sensitive to either dimensionality parameter in any of the experiments discussed here. This
 model was able to correctly label 99.3\% of the provably derivable test

### 2014-06-05 writing/S'14 paper/quantifiers.tex
Two more items of the paper to do list.

@@
 
-% Further explanation for motivation for no/no?
-
 \begin{table}[htp]
@@
 We compare the results to a most frequent class baseline, which reflects the frequency in the test data of most frequent class in the training data, $\natind$.
-
-\mynote{
 The RNN model was approximately optimal with N dimensional word representations and an M dimensional comparison layer. The RNTN was approximately optimal with N and M dimensions, respectively.
 % TODO: Update dimensionality
-(Note: Models still running. Results might improve and this para is
-likely to change.)
 
@@
 quantifier behavior in a na\"ive model remains open.
-
-Discussion of RNN results (and updated discussion of better RNTN results) to be filled in as results arrive.\\...\\... % TODO
-}

### 2014-06-05 writing/S'14 paper/recursion.tex
RNN falloff fig.

@@
   \centering
-  \includegraphics[width=5.5in]{recursion\string_results.eps}
+  \includegraphics[width=5.5in]{recursion\string_results\string_updated.eps}
   \caption{Model performance on propositional logic, by expression size.}  
@@
 increasing depth (and with the increasing complexity of the expressions),
- resulting in the gradually dropping performance. This is not
-necessarily a significant flaw in the model, since it remains possible
+ resulting in the gradually dropping performance. In contrast, the RNN model 
+ did not appear to be able to learn a correct approximation of the logic for
+ statements of any length, but it appears that the faulty logic that the RNN did
+ learn decayed similarly with expression size. 
+ This is not necessarily a significant flaw in the model, since it remains possible
 that a less lossy approximation of the same logic would do no better
@@
 be able to learn a correct approximation of the logic.
+
+\begin{figure}[htp]
+  \centering
+  \includegraphics[width=5.5in]{recursion\string_pairwise.eps}
+  \caption{...}  
+  \label{prop-results}
+\end{figure}

### 2014-06-05 writing/S'14 paper/discussion.tex
First minor tweaks from the review session.
Intermediate paper update

@@
 
-This paper evaluated two RNNs in a series of three increasingly
+This paper evaluated two recursive models in a series of three increasingly
 challenging interpretive tasks involving natural language inference:
@@
 answer to them in a way that leads to overfitting. The underlying
-logical theories provide only limited guidance on this point. Finally,
-we have only scratched the surface of the logical complexity of
+logical theories provide only limited guidance on this point, and the fact that 
+there is a finite universe of possible expressions makes this an unavoidable issue. 
+Finally, we have only scratched the surface of the logical complexity of
 natural language; in future experiments, we hope to test sentences

### 2014-06-05 writing/S'14 paper/recursion.tex
Intermediate paper update

@@
   \includegraphics[width=5.5in]{recursion\string_results\string_updated.eps}
-  \caption{Model performance on propositional logic, by expression size.}  
+  \caption{Model performance on propositional logic, by expression size. The dashed line indicates that only expressions of size four or less appeared in the training data.}  
   \label{prop-results}
@@
  statements of any length, but it appears that the faulty logic that the RNN did
- learn decayed similarly with expression size. 
- This is not necessarily a significant flaw in the model, since it remains possible
-that a less lossy approximation of the same logic would do no better
+ learn decayed similarly with expression size.  This is not necessarily a significant flaw in the model, since it remains possible that a less lossy approximation of the same logic would do no better
 on the training objective, since that objective only considers short
@@
 
+In an attempt to better understand the limited success of the RNN model, we discovered that this model looses information about in longer statements in a particularly problematic way.
+It appears that that model is unable to distinguish between two sentences when the only difference between those sentences is embedded within a very deep structure.
+We evaluated both models on sentences that differed in only one term, but for which that one term was embedded under a large number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and b)))}.
+We then measured the Euclidean distance between the vector representations of the two sentences in each pair.
+Our findings are shown in Figure~\ref{prop-falloff}, and show that while the RNTN can distinguish the two sentences well at every size that we test, the RNN fails after a depth of approximately six.
+
 \begin{figure}[htp]
@@
   \includegraphics[width=5.5in]{recursion\string_pairwise.eps}
-  \caption{...}  
-  \label{prop-results}
+  \caption{How well each model can distinguish to expressions that differ by one term, measured for terms embedded under different numbers of tokens of \ii{and}.}  
+  \label{prop-falloff}
 \end{figure}

### 2014-06-05 writing/S'14 paper/join.tex
Adding first two batches of new results.

@@
 not highly sensitive to either dimensionality parameter in any of the experiments discussed here. This
-model was able to correctly label 99.3\% of the provably derivable test
-examples, and 99.1\% of the remaining test examples. The simpler RNN model
+model was able to correctly label 99.4\% (94.3 macroaveraged F1) of the provably derivable test
+examples, and 95.4\% (92.2) of the remaining test examples. The simpler RNN model
 worked best with 11 and 75 dimensions, respectively, but was able to
-achieve accuracies of only 90.0\% and 87.0\%, respectively.
+achieve accuracies of only 89.0\% (80.9)  and 62.5\% (40.1), respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for the RNN.
 
@@
 
-% TODO: T-sne the RNTN model's embeddings.
-
 These results are fairly straightforward to interpret. The RNTN model

### 2014-06-05 writing/S'14 paper/recursion.tex
Adding first two batches of new results.

@@
 found that the RNTN model was able to perform almost perfectly on
-unseen small test examples, with accuracy above 99\% below size four.
-After depth four, performance gradually falls with increasing size.
-The RNN model did not perform well, reaching only 88.2\% accuracy on
+unseen small test examples, with accuracy above 99\% below size four (Macroaveraged F1 was above 98.9).
+Starting at size four, performance gradually falls with increasing size.
+The RNN model did not perform well, reaching only 88.2\% (67.2) accuracy on
 the smallest test examples, and declining from there to near-baseline
-performance at size 12.
+performance at size 12. Training accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.
+
+% thu-and-or-deep-25-32b.txt iter 102
+% mon-and-or-deep-45r-32.txt iter 480
 
@@
   \centering
-  \includegraphics[width=5.5in]{recursion\string_results\string_updated.eps}
+  \includegraphics[width=5.3in]{recursion\string_results\string_updated.eps}
   \caption{Model performance on propositional logic, by expression size. The dashed line indicates that only expressions of size four or less appeared in the training data.}  
@@
   \centering
-  \includegraphics[width=5.5in]{recursion\string_pairwise.eps}
+  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
   \caption{How well each model can distinguish to expressions that differ by one term, measured for terms embedded under different numbers of tokens of \ii{and}.}  

### 2014-06-05 writing/S'14 paper/quantifiers.tex
All results updated.

@@
     \toprule
-    Data & Most frequent class & TODO dim RNN (\%) & 16 dim RNTN (\%)\\
+    Data & Most frequent class & 20 dim RNN (\%) & 20 dim RNTN (\%)\\
     \midrule
-    \textsc{all split}	& 35.4 &	63.9&	\textbf{93.4}
+    \textsc{all split}	& 35.4 &	65.3&	\textbf{99.0}
     \\[1ex]    
-    \textsc{pair two/less-than-two}	& 59.8 &	79.5 &	\textbf{98.3} \\
-    \textsc{pair not-all/not-most}	&0 &	   62.2  &	\textbf{83.3} \\
-    \textsc{pair all/some}	& 0& 63.6  &	\textbf{81.0} \\
-    \textsc{pair no/no}	& 30.8 &	66.0 &	\textbf{100} \\
+    \textsc{pair two/less-than-two}	& 59.8 &	77.6 &	\textbf{100} \\
+    \textsc{pair not-all/not-most}	&0 & 60.8 &	\textbf{82.0} \\
+    \textsc{pair all/some}	& 0& 52.5  &	\textbf{85.1} \\
+    \textsc{pair no/no}	& 30.8 &	68.9 &	\textbf{99.9} \\
     \bottomrule
@@
 We compare the results to a most frequent class baseline, which reflects the frequency in the test data of most frequent class in the training data, $\natind$.
-The RNN model was approximately optimal with N dimensional word representations and an M dimensional comparison layer. The RNTN was approximately optimal with N and M dimensions, respectively.
+After some cross-validation, we chose 20-dimensional word and phrase representations, and 75 dimensional feature vectors for the classifier.
 % TODO: Update dimensionality
 
-The perfect performance by the RNTN on the
+The RNN performed poorly at this task,
+even though the sentences used in these examples are short enough
+to avoid the pathology shown in Figure~\ref{prop-falloff}.
+However, the perfect performance by the RNTN on the
 \textsc{pair no/no} experiment and its strong performance on
-\textsc{pair two/less-than-two} suggests that it is at least plausible
-that this model is able to handle quantifiers correctly given
-sufficient training data, but the sub 90\% results on two of the
+\textsc{all split} and \textsc{pair two/less-than-two} 
+suggests that it is at least plausible
+that this stronger model is able to handle quantifiers correctly given
+sufficient training data, but the sub-90\% results on two of the
 training settings suggest that the model may not be able to

### 2014-06-06 writing/S'14 paper/abstract.tex
Abstract update.
Proofreading etc.

--- 2014-06-03 writing/S'14 paper/abstract.tex
@@
   Supervised recursive neural network models (RNNs) for sentence
-  meaning have been successful in a wide array of sophisticated
-  language tasks, but it remains an open question whether they can
-  achieve the same results as compositional semantic grammars based in
-  logical forms. We address this question directly by using a logical
-  grammar to generate controlled data sets encoding the relationships
-  (e.g., entailment, contradiction) between pairs of expressions and
-  evaluating whether each of two classes of neural model --- plain
-  RNNs and recursive neural tensor networks (RNTNs) --- can learn
-  those relationships correctly. Our first experiment evaluates
-  whether these models can learn the basic algebra of logical
-  relations involved. Our second and third experiments extend this
-  evaluation to complex recursive structures and sentences involving
-  quantification.  We find that the plain RNN achieves only mixed
-  results on all three experiments, whereas the stronger RNTN models
-  generalizes well in every setting.
+  meaning have been successful in an array of sophisticated language
+  tasks, but it remains an open question whether they can learn
+  compositional semantic grammars that support logical deduction.  We
+  address this question directly by for the first time evaluating
+  whether each of two classes of neural model --- plain RNNs and
+  recursive neural tensor networks (RNTNs) --- can correctly learn
+  relationships such as entailment and contradiction between pairs of
+  sentences, where we have generated controlled data sets of sentences
+  from a logical grammar.  Our first experiment evaluates whether
+  these models can learn the basic algebra of logical relations
+  involved. Our second and third experiments extend this evaluation to
+  complex recursive structures and sentences involving quantification.
+  We find that the plain RNN achieves only mixed results on all three
+  experiments, whereas the stronger RNTN models generalizes well in
+  every setting and appears capable of learning suitable
+  representations for natural language logical inference.
 \end{abstract}
+
+% We address this question directly by for the first time evaluating whether each of two classes of neural model — plain RNNs and recursive neural tensor networks (RNTNs) — can correctly learn relationships such as entailment and contradiction between pairs of sentences, where we have generated controlled data sets of sentences from a logical grammar.

### 2014-06-06 writing/S'14 paper/intro.tex
Proofreading etc.

@@
 Supervised recursive neural network models (RNNs) for sentence meaning
-have been successful in a wide array of sophisticated language tasks,
+have been successful in an array of sophisticated language tasks,
 including sentiment analysis \cite{socher2011semi,socher2013acl1},
 image description \cite{sochergrounded}, and paraphrase detection
-\cite{Socher-etal:2011:Paraphrase}. These results are encouraging
-for the ability of these models to learn compositional semantic
-grammars, but it remains an open question whether they can achieve the
-same results as grammars based in logical forms
-\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013} when it comes to core semantic
-concepts like quantification, entailment, and contradiction. To date,
+\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
+the ability of these models to learn compositional semantic grammars,
+but it remains an open question whether they can achieve the same
+results as grammars based in logical forms
+\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
+when it comes to the core semantic concepts of quantification,
+entailment, and contradiction needed to identify the relationships
+between sentences like \ii{No child failed a language exam}, \ii{No
+  girl failed a French exam}, and \ii{A boy failed French}. To date,
 experimental investigations of these concepts using distributed
@@
 \cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
-For robust natural language understanding, we must be able to model
-these phenomena in their full generality on complex linguistic structures.
+For robust natural language understanding, it is essential to model
+these phenomena in their full generality on complex linguistic
+structures.
 
@@
 formal properties and inferential strength of this system are now
-well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
+well-understood \cite{Icard:Moss:2013:LILT}.
 
@@
 generate controlled data sets encoding the semantic relationships
-between pairs of expressions and evaluate whether each of two
-classes of neural model --- plain RNNs and recursive neural tensor
-networks (RNTNs, \cite{socher2013acl1}) --- can learn those
-relationships correctly. In our first experiment
-(Section~\ref{sec:join}), we evaluate the ability of these models to
-learn the core relational algebra of natural logic from data. Our
-second experiment (Section~\ref{sec:recursion}) extends this
-evaluation to cover relations between complex recursive structures
-like $(a \plor b)$ and $\plneg(\plneg a \pland \plneg b)$, and our
-third experiment (Section~\ref{sec:quantifiers}) involves relations
-between quantified statements like \ii{every reptile walks} and \ii{every
-turtle moves}. We find that the plain RNN achieves only mixed results
-in all three experiments, whereas the stronger RNTN models generalized
-well in every case, suggesting that it has in fact learned, or at
-least learned to simulate, our target logical
+between pairs of expressions and evaluate whether each of two classes
+of neural model --- plain RNNs and recursive neural tensor networks
+(RNTNs, \cite{socher2013acl1}) --- can learn those relationships
+correctly. In our first experiment (Section~\ref{sec:join}), we
+evaluate the ability of these models to learn the core relational
+algebra of natural logic from data. Our second experiment
+(Section~\ref{sec:recursion}) extends this evaluation to relations
+between complex recursive structures like $(a \plor b)$ and
+$\plneg(\plneg a \pland \plneg b)$, and our third experiment
+(Section~\ref{sec:quantifiers}) involves relations between quantified
+statements like \ii{every reptile walks} and \ii{every turtle moves}.
+We find that the plain RNN achieves only mixed results in all three
+experiments, whereas the stronger RNTN model generalizes well in every
+case, suggesting that it has in fact learned, or at least learned to
+simulate, our target logical
 concepts. % TODO: Update with current results
 
-\begin{table}[htp]
+\begin{table}[tp]
   \centering
@@
     $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do, including when there 
-    is insufficient knowledge to choose a label.}
+    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
+    %is insufficient knowledge to choose a label.}
   \label{b-table}

### 2014-06-06 writing/S'14 paper/join.tex
Dispatching a few more little TODOs.
Proofreading etc.

@@
-\section{Relation composition}\label{sec:join}
+\section{Reasoning about semanic relations}\label{sec:join}
 
 If a model is to learn the behavior of a relational logic like the one
-presented here from a finite amount data, it must be able to learn to
-deduce new relations from seen relations in a sound
-manner. The simplest such deductions involve atomic statements using
-the relations in Table~\ref{b-table}. For instance, given that $a
-\natrev b$ and $b \natrev c$, one can conclude that $a \natrev c$, by
-basic set-theoretic reasoning (transitivity of $\natrev$). Similarly,
-from $a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.
-The full set of sound inferences of this form is summarized in
+presented here from a finite amount data, it must learn to deduce new
+relations from seen relations in a sound manner. The simplest such
+deductions involve atomic statements using the relations in
+Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
+\natrev c$, one can conclude that $a \natrev c$, by basic
+set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
+$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
+full set of sound inferences of this form is summarized in
 Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
@@
 
-To test our systems's ability to learn this relational structure, we
-create small models for our logic in which terms denote sets of
-entities from a single domain of seven entities (integers).
-Figure~\ref{lattice-figure} depicts a small model of this form. The
-lattice structure gives the full model, for which all the statements on
-the right are valid. We divide these statements evenly into training and
-test sets, and remove from the test set those statements which cannot be 
-proven from the training statements using the logic depicted in 
-Figure~\ref{lattice-figure}.
+\paragraph{Experiments}
+To test our models' ability to learn these inferential patterns, we
+create small boolean structures for our logic in which terms denote
+sets of entities from a single domain of seven entities (integers).
+Figure~\ref{lattice-figure} depicts a small structure of this
+form. The lattice gives the full model, for which all the statements
+on the right are valid. We divide these statements evenly into
+training and test sets, and remove from the test set those statements
+which cannot be proven from the training statements.
+% using the logic depicted in Figure~\ref{lattice-figure}.
 
@@
   \centering
-  \begin{subfigure}[t]{0.4\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
     \centering
@@
     \end{picture}
-    \caption{Simple boolean model. The letters name the sets. Not all sets have names, and
+    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
     some sets have multiple names, so that learning $\nateq$ is non-trivial.}
@@
   \qquad
-  \begin{subfigure}[t]{0.5\textwidth}
+  \begin{subfigure}[t]{0.43\textwidth}
     \centering
@@
 
-    \caption{An example train/test split from the full set
-      of statements one can make about the model.
-      Statements not provable from the test data are crossed out.}
+    \caption{A train/test split of the atomic statements about the
+      model.  Test statements not provable from the training data are
+      crossed out.}
   \end{subfigure}  
-  \caption{Some sample randomly generated sets, and some of the relations defined between them.}
+  \caption{Small example structure and data for learning relation composition.}
   \label{lattice-figure}
@@
 
-\paragraph{Results.} We found that the RNTN model worked best with 11-dimensional vector representations for the
-80 sets and a 90-dimensional feature vector for the classifier, though the performance of the model was
-not highly sensitive to either dimensionality parameter in any of the experiments discussed here. This
-model was able to correctly label 99.4\% (94.3 macroaveraged F1) of the provably derivable test
-examples, and 95.4\% (92.2) of the remaining test examples. The simpler RNN model
-worked best with 11 and 75 dimensions, respectively, but was able to
-achieve accuracies of only 89.0\% (80.9)  and 62.5\% (40.1), respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for the RNN.
+\paragraph{Results} 
+We found that the RNTN model worked best with 11-dimensional vector
+representations for the 80 sets and a 90-dimensional feature vector
+for the classifier, though the performance of the model was not highly
+sensitive to either dimensionality parameter in any of the experiments
+discussed here. This model was able to correctly label 99.4\% (94.3
+macroaveraged F1) of the provably derivable test examples, and 95.4\%
+(92.2) of the remaining test examples. The simpler RNN model worked
+best with 11 and 75 dimensions, respectively, but was able to achieve
+accuracies of only 89.0\% (80.9) and 62.5\% (40.1),
+respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
+the RNN.
 
@@
 % relations in about half of the cases) and the 240 that
-% cannot. % TODO: Say more about symmetry?
+% cannot.
 

### 2014-06-06 writing/S'14 paper/methods.tex
Proofreading etc.

@@
 %
-Here, the $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector representations
-for the left and right children of the node, and $\vec{y}$ is the node's output.
-The RNN concatenates them, multiplies them by an $N \times 2N$
-matrix of learned weights, and applies the element-wise non-linearity
-to the resulting vector (plus a bias term). The RNTN has the same
-basic structure, but with the addition of a learned third-order tensor $\mathbf{T}$,
-dimension $N \times N \times N$, modeling multiplicative interactions between the
-child vectors. Both models include a learned bias vector, represented $\vec{b}$.
+Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
+representations for the left and right children of the node, and
+$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
+them by an $N \times 2N$ matrix of learned weights, and applies the
+element-wise non-linearity to the resulting vector. The RNTN has the
+same basic structure, but with the addition of a learned third-order
+tensor $\mathbf{T}$, dimension $N \times N \times N$, modeling
+multiplicative interactions between the child vectors. Both models
+include a learned bias vector, represented as $\vec{b}$.
 

### 2014-06-06 writing/S'14 paper/recursion.tex
Proofreading etc.

@@
 
-Recursive structure is a prominent feature of natural languages and
-an important part of their expressivity, as it allows people to use
-and interpret complex structures they have never encountered before.
+Recursive structure is a prominent feature of natural languages and an
+important part of their expressivity, as it allows people to use and
+interpret complex structures they have never encountered before.
 Theoretical accounts of natural language syntax and semantics
@@
 well. The present section pursues this question for our two classes of
-RNN. In evaluating the model, we take advantage of the fact that
-recursive structures of this kind define potentially infinite
-languages, by testing the model on strings that are longer and more
+RNN. In our evaluations, we take exploit the fact that our logical
+language is infinite by testing on strings that are longer and more
 complex than any seen in training.
@@
 
+\paragraph{Experiments}
 As in Section~\ref{sec:join}, we test this phenomenon within the
 framework of natural logic, but we now replace the unanalyzed symbols
-from that experiment with expressions that involve recursive
-structure. These expressions represent a truth-functionally complete classical
-propositional logic: each atomic variable denotes either $\True$ or
-$\False$, and the only operators are truth-functional ones.
-Table~\ref{tab:pl} defines this logic, and Table~\ref{tab:plexs} gives
-some examples of relational statements that we can make
-in these terms. To compute these relations between statements, we
-exhaustively enumerate the sets of assignments of truth values to
-proposition variables that would satisfy each of the statements and
-then convert the set-theoretic relation between those assignments into
-one of the seven relations in Table~\ref{b-table}. As a result, each relational
-statement represents a valid theorem of the propositional logic.
+from that experiment with complex formulae. These formulae
+represent a truth-functionally complete classical propositional logic:
+each atomic variable denotes either $\True$ or $\False$, and the only
+operators are truth-functional ones.  Table~\ref{tab:pl} defines this
+logic, and Table~\ref{tab:plexs} gives some examples of relational
+statements that we can make in these terms. To compute these relations
+between statements, we exhaustively enumerate the sets of assignments
+of truth values to proposition variables that would satisfy each of
+the statements and then convert the set-theoretic relation between
+those assignments into one of the seven relations in
+Table~\ref{b-table}. As a result, each relational statement represents
+a valid theorem of the propositional logic.
 
-\begin{table}[htp]
+\begin{table}[tp]
   \centering
@@
       $a$               & $\natrev$ & $(a \pland b)$ \\
-      $(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
+      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
       $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
@@
 % Hopefully this is explicit enough.
-Socher et al.~\cite{socher2012semantic} demonstrate the learning of a
-logic in a matrix-vector RNN model somewhat similar to our own, but
-the logic discussed here is substantially stronger, and a much better
-approximation of the kind of structure that is needed for natural
+Socher et al.~\cite{socher2012semantic} show that a matrix-vector RNN
+model somewhat similar to our own can learn a logic, but the logic
+discussed here is substantially stronger, and a much better
+approximation of the kind of expressivity needed for natural
 language. The logic learned in that experiment is boolean, wherein the
@@
 
-If we do not implement any constraint that the two statements
-being compared are similar in any way, the generated data consists in
-large part of statements in which the two formulae refer to largely
-separate subsets of the six variables, and to which we will nearly
-always assign the $\natind$ relation. In an effort to balance the
-distribution of relation labels without departing from the basic task
-of modeling propositional logic, we disallow individual pairs of
-statements from referring to more than four of the six proposition
-variables.
+If we do not implement any constraint that the two statements being
+compared are similar in any way, then the generated data are dominated
+by statements in which the two formulae refer to largely separate
+subsets of the six variables, which means that the $\natind$ relation
+is almost always correct.  In an effort to balance the distribution of
+relation labels without departing from the basic task of modeling
+propositional logic, we disallow individual pairs of statements from
+referring to more than four of the six proposition variables.
 
-\paragraph{Results.} We trained both the RNN and RNTN models on the data of size four or
+\paragraph{Results} 
+We trained both the RNN and RNTN models on the data of size four or
 less (65k pairs), and tested it on examples of up to size 12 (44k
@@
 found that the RNTN model was able to perform almost perfectly on
-unseen small test examples, with accuracy above 99\% below size four (Macroaveraged F1 was above 98.9).
-Starting at size four, performance gradually falls with increasing size.
-The RNN model did not perform well, reaching only 88.2\% (67.2) accuracy on
-the smallest test examples, and declining from there to near-baseline
-performance at size 12. Training accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.
+unseen small test examples, with accuracy above 99\% below size four
+(Macroaveraged F1 was above 98.9).  Starting at size four, performance
+gradually falls with increasing size.  The RNN model did not perform
+well, reaching only 88.2\% (67.2) accuracy on the smallest test
+examples, and declining from there to near-baseline performance at
+size 12. Training accuracy was 99.6\% for the RNTN and 87.6\% for the
+RNN.
 
@@
 
-\begin{figure}[htp]
+\begin{figure}[tp]
   \centering
   \includegraphics[width=5.3in]{recursion\string_results\string_updated.eps}
-  \caption{Model performance on propositional logic, by expression size. The dashed line indicates that only expressions of size four or less appeared in the training data.}  
+  \caption{Model performance on propositional logic, by expression size. 
+    The dashed line indicates that only expressions of size four or less appeared in the training data.}  
   \label{prop-results}

### 2014-06-06 writing/S'14 paper/recursion.tex
More editing

@@
 
-In an attempt to better understand the limited success of the RNN model, we discovered that this model looses information about in longer statements in a particularly problematic way.
-It appears that that model is unable to distinguish between two sentences when the only difference between those sentences is embedded within a very deep structure.
-We evaluated both models on sentences that differed in only one term, but for which that one term was embedded under a large number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and b)))}.
-We then measured the Euclidean distance between the vector representations of the two sentences in each pair.
-Our findings are shown in Figure~\ref{prop-falloff}, and show that while the RNTN can distinguish the two sentences well at every size that we test, the RNN fails after a depth of approximately six.
+In an effort to better understand why the RNTN outperforms the plain
+RNN, we evaluated both models on pairs of long formulae involving
+binary connectives, to assess the extent to which they achieve
+substantially different representations for semantically distinct
+formulae and substantially similar representations for synonymous
+formulae. We found that both models are able to identify synonymous
+pairs like $(a \pland a)$ and $(a \pland (a \pland (a \pland a)))$,
+and similarly for disjunctions, but only the RNTN is able to learn
+substantially different representations for formulae like $(a \pland
+a)$ and $(a \pland (a \pland (a \pland
+b)))$. Figure~\ref{prop-falloff} summarizes the result for pairs of
+distinct formulae, showing that the RNN basically cannot distinguish
+them after depth~2.
+
+% we discovered that this model looses information about in longer
+% statements in a particularly problematic way. It appears that that
+% model is unable to distinguish between two sentences when the only
+% difference between those sentences is embedded within a very deep
+% structure. We evaluated both models on sentences that differed in only
+% one term, but for which that one term was embedded under a large
+% number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and
+%   b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and
+%   b)))}. We then measured the Euclidean distance between the vector
+% representations of the two sentences in each pair. Our findings are
+% shown in Figure~\ref{prop-falloff}, and show that while the RNTN can
+% distinguish the two sentences well at every size that we test, the RNN
+% fails after a depth of approximately six.
 
@@
   \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-  \caption{How well each model can distinguish to expressions that differ by one term, measured for terms embedded under different numbers of tokens of \ii{and}.}  
+  \caption{Truth-conditionally distinct formulae should have different
+    representations. As measured by Euclidean distance, only the RNTN
+    achieves this for formulae containing more than two
+    connectives. The RNN quickly collapses the representations,
+    failing to capture the meaning contrasts.}
   \label{prop-falloff}

### 2014-06-06 writing/S'14 paper/discussion.tex
Dispatching a few more little TODOs.
More edits

@@
 models have truly generalized in the way we want them to. There is a
-constant tension between giving the models training data that gives
+constant tension between showing the models training data that gives
 them a chance to learn the target logical functions and revealing the
 answer to them in a way that leads to overfitting. The underlying
-logical theories provide only limited guidance on this point, and the fact that 
-there is a finite universe of possible expressions makes this an unavoidable issue. 
-Finally, we have only scratched the surface of the logical complexity of
-natural language; in future experiments, we hope to test sentences
+logical theories provide only limited guidance on this point.
+%
+%, and the fact that there is a finite universe of possible
+%  expressions makes his an unavoidable issue. 
+%
+%  CP: I don't understand the above. It is false if taken literally;
+%  our PL generates an infinte number of formulae.
+%
+Finally, we have only scratched the surface of the logical complexity
+of natural language; in future experiments, we hope to test sentences
 with embedded quantifiers, multiple interacting quantifiers, relative
@@
 In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.
-
-% TODO: One para of closing discussion.
-% - Results strongly suggest that RNTNs can encode the basics of this logic. First such result.
-% -- Decay over size in recursion results not ideal, but this kind of behavior may not be problematic since there is a practical bound on the length of natural language sentences.
-% - Future work: 
-% -- Harder problems
-% --- Sentences with more types of structure: Transitive verbs, relative clauses, etc.
-% -- More models
-% --- BP-RNNs, conv-RNNs

### 2014-06-06 writing/S'14 paper/intro.tex
More edits

@@
 entailment, and contradiction needed to identify the relationships
-between sentences like \ii{No child failed a language exam}, \ii{No
-  girl failed a French exam}, and \ii{A boy failed French}. To date,
-experimental investigations of these concepts using distributed
-representations have been largely confined to short phrases
+between sentences like \ii{every animal walks}, \ii{every turtle
+  moves}, and \ii{most reptiles don't move}. To date, experimental
+investigations of these concepts using distributed representations
+have been largely confined to short phrases
 \cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
@@
 theoretical work on this task (and some successful implemented models
-\cite{maccartney2009natural,watanabe2012latent}) involves \ii{natural
+\cite{maccartney2009extended,watanabe2012latent}) involves \ii{natural
   logics}, which are formal systems that define rules of inference
@@
 formal properties and inferential strength of this system are now
-well-understood \cite{Icard:Moss:2013:LILT}.
+well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
 
@@
 (Section~\ref{sec:quantifiers}) involves relations between quantified
-statements like \ii{every reptile walks} and \ii{every turtle moves}.
+statements like \ii{every reptile walks} and \ii{no turtle moves}.
 We find that the plain RNN achieves only mixed results in all three

### 2014-06-06 writing/S'14 paper/join.tex
More edits

@@
 create small boolean structures for our logic in which terms denote
-sets of entities from a single domain of seven entities (integers).
-Figure~\ref{lattice-figure} depicts a small structure of this
-form. The lattice gives the full model, for which all the statements
-on the right are valid. We divide these statements evenly into
-training and test sets, and remove from the test set those statements
-which cannot be proven from the training statements.
+sets of entities from a small domain.  Figure~\ref{lattice-figure}
+depicts a structure of this form. The lattice gives the full model,
+for which all the statements on the right are valid. We divide these
+statements evenly into training and test sets, and remove from the
+test set those statements which cannot be proven from the training
+statements.
 % using the logic depicted in Figure~\ref{lattice-figure}.
@@
 single vector in the (randomly initialized) vocabulary matrix $V$,
-ensuring that the model has no information about the terms being
-compared except the relations between them.
+thereby ensuring that the model has no information about the terms
+being compared except the relations between them.
 
@@
 macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2) of the remaining test examples. The simpler RNN model worked
+(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
 best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.9) and 62.5\% (40.1),
+accuracies of only 89.0\% (80.9 macro-F1) and 62.5\% (40.1 macro-F1),
 respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for

### 2014-06-06 writing/S'14 paper/methods.tex
More edits

@@
 network layer function \eqref{rnn} and the RNTN layer function
-proposed in \citet{chen2013learning} \eqref{rntn}. A sigmoid
+\eqref{rntn} proposed in \citet{chen2013learning}. A sigmoid
 nonlinearity (element-wise $\tanh$) is applied to the output of either
@@
 \label{rntn}
-\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots N]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
+\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
 \end{gather} % TODO: Explain third order tensor parameter?
@@
 $\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
-them by an $N \times 2N$ matrix of learned weights, and applies the
+them by an $n \times 2n$ matrix of learned weights, and applies the
 element-wise non-linearity to the resulting vector. The RNTN has the
 same basic structure, but with the addition of a learned third-order
-tensor $\mathbf{T}$, dimension $N \times N \times N$, modeling
+tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
 multiplicative interactions between the child vectors. Both models
-include a learned bias vector, represented as $\vec{b}$.
+include a bias vector, represented as $\vec{b}$.
 
@@
 Rather than use a $\tanh$ nonlinearity here, we found a substantial
-improvement in performance by using a rectified linear function for
-$f_{b}$. In particular, we use the leaky rectified linear function
+improvement in performance by using a rectified linear function. In
+particular, we use the leaky rectified linear function
 \cite{maasrectifier}: $f_{b}(\vec{x})=\max(\vec{x}, 0) +

### 2014-06-06 writing/S'14 paper/quantifiers.tex
Dispatching a few more little TODOs.
More edits

@@
 
-Even to the extent that RNTN models can handle functional meanings in
-the form of the operators of propositional logic, this is not a
-guarantee that these models can handle functional meanings of the form
-seen in natural language. As a first step towards investigating this
-issue, we attempt to directly measure the degree to which RNNs are
-able to develop suitable representations for the semantics of natural language
+We have seen that the RNTN can learn an approximation of propositional
+logic.  However, natural languages can express functional meanings of
+considerably greater complexity than this.  As a first step towards
+investigating whether our models can capture this complexity, we now
+attempt to directly measure the degree to which RNNs are able to
+develop suitable representations for the semantics of natural language
 quantifiers like \ii{some} and \ii{all}. Quantification is far from
@@
 
-Our data consists of pairs of sentences generated from a small
-artificial grammar. Each sentence contains a quantifier, a noun which
-may be negated, and an intransitive verb which may be negated. We use
-the basic quantifiers \ii{some}, \ii{most}, \ii{all}, \ii{two}, and
-\ii{three}, and their negations \ii{no},
-\ii{not-all}, \ii{not-most}, \ii{less-than-two}, and
-\ii{less-than-three}. We also include five nouns, four intransitive
-verbs, and the negation symbol \ii{not}. In order to be able to define
-relations between sentences with differing lexical items, we define
-the lexical relations between each noun--noun pair, each verb--verb
-pair, and each quantifier--quantifier pair. The grammar accepts
-aligned pairs of sentences of this form and calculates the natural
-logic relationship between them.  Some examples of these data are
-provided in Table~\ref{examplesofdata}.  As in previous sections, the
-goal of learning is then to assign these relational labels accurately
-to unseen pairs of sentences.
+\paragraph{Experiments}
+Our experimental data consist of pairs of sentences generated from a
+small artificial grammar. Each sentence contains a quantifier, a noun
+which may be negated, and an intransitive verb which may be
+negated. We use the basic quantifiers \ii{some}, \ii{most}, \ii{all},
+\ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
+\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}. We also
+include five nouns, four intransitive verbs, and the negation symbol
+\ii{not}. In order to be able to define relations between sentences
+with differing lexical items, we define the lexical relations between
+each noun--noun pair, each verb--verb pair, and each
+quantifier--quantifier pair. The grammar accepts aligned pairs of
+sentences of this form and calculates the natural logic relationship
+between them.  Some examples of these data are provided in
+Table~\ref{examplesofdata}.  As in previous sections, the goal of
+learning is then to assign these relational labels accurately to
+unseen pairs of sentences.
 
@@
 %
+% do not allow a blank line --- adds too much space
+%
 \begin{gather}
@@
 \end{gather}
-
+%
+% do not allow a blank line --- adds too much space
+%
 While our primary interest is in discovering the extent to which our
@@
 
-\begin{table}[htp]
+\begin{table}[tp]
   \centering
@@
     \\[1ex]    
-    \textsc{pair two/less-than-two}	& 59.8 &	77.6 &	\textbf{100} \\
+    \textsc{pair two/less-than-two}	& 59.8 &	77.6 &	\textbf{100.0} \\
     \textsc{pair not-all/not-most}	&0 & 60.8 &	\textbf{82.0} \\
@@
   \end{tabular}
-  \caption{Quantifier experiment performance. }
+  \caption{Performance on the quantifier experiments. }
   \label{resultstable}
@@
 
-\paragraph{Results.} The results for these experiments are shown in Figure \ref{resultstable}.
-We compare the results to a most frequent class baseline, which reflects the frequency in the test data of most frequent class in the training data, $\natind$.
-After some cross-validation, we chose 20-dimensional word and phrase representations, and 75 dimensional feature vectors for the classifier.
-% TODO: Update dimensionality
+\paragraph{Results} 
+The results for these experiments are shown in
+Figure~\ref{resultstable}. We compare the results to a most frequent
+class baseline, which reflects the frequency in the test data of most
+frequent class in the training data, $\natind$.  After some
+cross-validation, we chose 20-dimensional word and phrase
+representations, and 75 dimensional feature vectors for the
+classifier.
 
-The RNN performed poorly at this task,
-even though the sentences used in these examples are short enough
-to avoid the pathology shown in Figure~\ref{prop-falloff}.
-However, the perfect performance by the RNTN on the
-\textsc{pair no/no} experiment and its strong performance on
-\textsc{all split} and \textsc{pair two/less-than-two} 
-suggests that it is at least plausible
-that this stronger model is able to handle quantifiers correctly given
-sufficient training data, but the sub-90\% results on two of the
-training settings suggest that the model may not be able to
-generalize from data this impoverished in general. However, the fact
-that both models perform far above baseline is promising, and the
-question of how much data is necessary to accurately capture
-quantifier behavior in a na\"ive model remains open.
+The RNN performed poorly at this task, even though the sentences used
+in these examples are short enough to avoid the pathology shown in
+Figure~\ref{prop-falloff}.  However, the perfect performance by the
+RNTN on the \textsc{pair no/no} experiment and its strong performance
+on \textsc{all split} and \textsc{pair two/less-than-two} suggests
+that that this stronger model is able to handle quantifiers correctly
+given sufficient training data, but the sub-90\% results on two of the
+training settings suggest that it may not be able to generalize from
+data this impoverished in general. However, though the question of how
+much data is necessary to accurately capture quantifier behavior in a
+na\"ive model remains open, the fact that both models perform far
+above baseline is promising.

### 2014-06-06 writing/S'14 paper/recursion.tex
More edits

@@
 well. The present section pursues this question for our two classes of
-RNN. In our evaluations, we take exploit the fact that our logical
-language is infinite by testing on strings that are longer and more
-complex than any seen in training.
+RNN. In our evaluations, we exploit the fact that our logical language
+is infinite by testing on strings that are longer and more complex
+than any seen in training.
 
@@
 Socher et al.~\cite{socher2012semantic} show that a matrix-vector RNN
-model somewhat similar to our own can learn a logic, but the logic
-discussed here is substantially stronger, and a much better
-approximation of the kind of expressivity needed for natural
-language. The logic learned in that experiment is boolean, wherein the
-atomic symbols are simply the values $0$ and $1$, rather than
-variables over those values. While learning the operators of that
-logic is not trivial, the outputs of each operator can be represented
-accurately by a single bit. The statements of propositional logic
-learned here describe conditions on the truth values of propositions
-where those truth values are not known. As opposed to the two-way
+model somewhat similar to our RNTN can learn a logic, but the logic
+discussed here is a much better approximation of the kind of
+expressivity needed for natural language. The logic learned in that
+experiment is boolean, wherein the atomic symbols are simply the
+values $0$ and $1$, rather than variables over those values. While
+learning the operators of that logic is not trivial, the outputs of
+each operator can be represented accurately by a single bit. The
+statements of propositional logic learned here describe conditions on
+unknown truth values of propositions. As opposed to the two-way
 contrasts seen in \cite{socher2012semantic}, this logic distinguishes
@@
 possible statements that the intermediate vector representations need
-to be able to distinguish.
+to be able to distinguish if it is to learn this logic system in its
+full generality.
 
@@
 the longer of the two formulae. We then randomly sample 15\% of each
-bin for a held out test set.
+bin for a held-out test set.
 
@@
 \paragraph{Results} 
-We trained both the RNN and RNTN models on the data of size four or
-less (65k pairs), and tested it on examples of up to size 12 (44k
-pairs). We initialized the model parameters randomly, including the
-vector representations of the six variables. The results are shown in
-Figure~\ref{prop-results}. In tuning, we found that the RNN model was
-approximately optimal with 45-dimensional vector representations, and
-the RNTN model was approximately optimal with 25 dimensions. We fixed
-the size of the feature vector for the classifier at 75 dimensions. We
-found that the RNTN model was able to perform almost perfectly on
-unseen small test examples, with accuracy above 99\% below size four
-(Macroaveraged F1 was above 98.9).  Starting at size four, performance
-gradually falls with increasing size.  The RNN model did not perform
-well, reaching only 88.2\% (67.2) accuracy on the smallest test
-examples, and declining from there to near-baseline performance at
-size 12. Training accuracy was 99.6\% for the RNTN and 87.6\% for the
-RNN.
+We trained both the RNN and RNTN models on data with up to 4
+connectives (65k pairs) and tested it on examples of up to 12
+connectives (44k pairs). We initialized the model parameters randomly,
+including the vector representations of the six variables. The results
+are shown in Figure~\ref{prop-results}. In tuning, we found that the
+RNN model was approximately optimal with 45-dimensional vector
+representations, and the RNTN model was approximately optimal with 25
+dimensions. We fixed the size of the feature vector for the classifier
+at 75 dimensions. We found that the RNTN model was able to perform
+almost perfectly on unseen small test examples, with accuracy above
+99\% below size four (macro-F1 was above 98.9).  Starting at size
+four, performance gradually falls with increasing size.  The RNN model
+did not perform well, reaching only 88.2\% (67.2 micro-F1) accuracy on
+the smallest test examples, and declining from there to near-baseline
+performance at size 12. Training accuracy was 99.6\% for the RNTN and
+87.6\% for the RNN.
 
@@
 number of times, but that the error in the approximation grows with
-increasing depth (and with the increasing complexity of the expressions),
- resulting in the gradually dropping performance. In contrast, the RNN model 
- did not appear to be able to learn a correct approximation of the logic for
- statements of any length, but it appears that the faulty logic that the RNN did
- learn decayed similarly with expression size.  This is not necessarily a significant flaw in the model, since it remains possible that a less lossy approximation of the same logic would do no better
-on the training objective, since that objective only considers short
-strings and includes a regularization term. Whether a more accurate
-encoding could be learned from a larger set of training data remains a
-question for future work. In contrast, the RNN model did not appear to
-be able to learn a correct approximation of the logic.
+increasing depth (and with the increasing complexity of the
+expressions), resulting in gradually dropping performance. In
+contrast, the RNN model did not appear to be able to learn a correct
+approximation of the logic for statements of any length, but it
+appears that the faulty logic that the RNN did learn decayed similarly
+with expression size.  This is not necessarily a significant flaw in
+the model, since it remains possible that a less lossy approximation
+of the same logic would do no better on the training objective, since
+that objective only considers short strings and includes a
+regularization term. Whether a more accurate encoding could be learned
+from a larger set of training data remains a question for future
+work.
 
-In an effort to better understand why the RNTN outperforms the plain
+In an effort to better understand why the RNTN outperformed the plain
 RNN, we evaluated both models on pairs of long formulae involving

### 2014-06-06 writing/S'14 paper/recursion.tex
Typos fixed

@@
 possible statements that the intermediate vector representations need
-to be able to distinguish if it is to learn this logic system in its
+to be able to distinguish if it is to learn this logical system in its
 full generality.

### 2014-06-06 writing/S'14 paper/join.tex
Typo correction
Temp

@@
-\section{Reasoning about semanic relations}\label{sec:join}
+\section{Reasoning about semantic relations}\label{sec:join}
 
@@
 best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.9 macro-F1) and 62.5\% (40.1 macro-F1),
+accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
 respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for

### 2014-06-06 writing/S'14 paper/methods.tex
Temp

@@
 We train the model with minibatch stochastic gradient descent (SGD)
-with learning rates computed using AdaGrad \cite{duchi2011adaptive}
-from a starting rate of 0.2. The parameters (including the vocabulary)
-are initialized randomly using a uniform distribution over $[-0.1,
-0.1]$. % TODO: Update
-Source code and generated data will be released after the conclusion
-of the review period.
+with learning rates computed using AdaGrad \cite{duchi2011adaptive}.
+The parameters (including the vocabulary) are initialized randomly 
+using a uniform distribution over $[-0.1, 0.1]$.
+Because the seven classes are not balanced in general, we report performance
+using both accuracy and macroaveraged F1, since the latter is sensitive
+to performance on infrequent classes. We compute macroaveraged F1 
+as the harmonic mean of mean precision and mean recall, both computed
+for all classes for which there is test data, setting precision to 0 
+where it is not defined. Source code and generated data will be released
+after the review period.
 

### 2014-06-06 writing/S'14 paper/recursion.tex
Temp

@@
 almost perfectly on unseen small test examples, with accuracy above
-99\% below size four (macro-F1 was above 98.9).  Starting at size
+98.7\% below size four (macro-F1 was above 95.5).  Starting at size
 four, performance gradually falls with increasing size.  The RNN model
-did not perform well, reaching only 88.2\% (67.2 micro-F1) accuracy on
+did not perform well, reaching only 90.5\% (81.4 micro-F1) accuracy on
 the smallest test examples, and declining from there to near-baseline
@@
 
-% thu-and-or-deep-25-32b.txt iter 102
+% thu-and-or-deep-25-32b.txt iter 136
 % mon-and-or-deep-45r-32.txt iter 480
@@
   \centering
-  \includegraphics[width=5.3in]{recursion\string_results\string_updated.eps}
+  \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
   \caption{Model performance on propositional logic, by expression size. 

### 2014-06-06 writing/S'14 paper/quantifiers.tex
Temp
Final results in!

@@
     \toprule
-    Data & Most frequent class & 20 dim RNN (\%) & 20 dim RNTN (\%)\\
+    Data & Most frequent class & 16 dim RNN  & 20 dim RNTN\\
     \midrule
-    \textsc{all split}	& 35.4 &	65.3&	\textbf{99.0}
+    \textsc{all split}	& 35.4 (7.5) &	67.4 (56.5)&	\textbf{100.0 (100.0)}
     \\[1ex]    
-    \textsc{pair two/less-than-two}	& 59.8 &	77.6 &	\textbf{100.0} \\
-    \textsc{pair not-all/not-most}	&0 & 60.8 &	\textbf{82.0} \\
-    \textsc{pair all/some}	& 0& 52.5  &	\textbf{85.1} \\
-    \textsc{pair no/no}	& 30.8 &	68.9 &	\textbf{99.9} \\
+    \textsc{pair two/less-than-two}	& 59.8 (18.4) &	77.2 (57.4) &	\textbf{100.0 (100.0)} \\
+    \textsc{pair not-all/not-most}	&0 (0) & 66.0 (54.3) &	\textbf{93.8 (91.5)} \\
+    \textsc{pair all/some}	& 0 (0) & 62.7 (65.9)  &	\textbf{78.4 (84.7)} \\
+    \textsc{pair no/no}	& 30.8 (10.2) &	67.8 (61.8) &	\textbf{99.9 (99.7)} \\
     \bottomrule
   \end{tabular}
-  \caption{Performance on the quantifier experiments. }
+  \caption{Performance on the quantifier experiments. Results are reported as accuracy scores followed by macroaveraged F1 scores in parentheses.}
   \label{resultstable}
@@
 frequent class in the training data, $\natind$.  After some
-cross-validation, we chose 20-dimensional word and phrase
-representations, and 75 dimensional feature vectors for the
-classifier.
+cross-validation, we chose 16 and 20-dimensional
+representations for the RNN and RNTN respectively, and 75 dimensional 
+feature vectors for the classifier.
 
@@
 Figure~\ref{prop-falloff}.  However, the perfect performance by the
-RNTN on the \textsc{pair no/no} experiment and its strong performance
-on \textsc{all split} and \textsc{pair two/less-than-two} suggests
+RNTN on the \textsc{all split} and \textsc{pair two/less-than-two} 
+experiment and its strong performance on \textsc{pair no/no} suggests
 that that this stronger model is able to handle quantifiers correctly
-given sufficient training data, but the sub-90\% results on two of the
+given sufficient training data, but the weaker results on two of the
 training settings suggest that it may not be able to generalize from

### 2014-06-06 writing/S'14 paper/methods.tex
Typo

@@
 \vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
-\end{gather} % TODO: Explain third order tensor parameter?
+\end{gather} 
 %
@@
 $\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
-them by an $n \times 2n$ matrix of learned weights, and applies the
+them by an $N \times 2N$ matrix of learned weights, and applies the
 element-wise non-linearity to the resulting vector. The RNTN has the
 same basic structure, but with the addition of a learned third-order
-tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
+tensor $\mathbf{T}$, dimension $N \times N \times N$, modeling
 multiplicative interactions between the child vectors. Both models

### 2014-06-06 writing/S'14 paper/methods.tex
Methods error fix.

@@
 distributed setting, word meanings are vectors of length $N$. The
-combination function concatenates pairs of them to form vectors of
-length $2N$ and maps those larger vectors back into length $N$
+combination function maps pairs of them to single vectors of length $N$
 vectors, which can then be merged again to create more complex

### 2014-06-06 writing/S'14 paper/methods.tex
A few methods clarifications.

@@
 expressions are derived from the meanings of their constituent parts
-via specific combination functions \cite{Partee84,Janssen97}. In our
+via specific composition functions \cite{Partee84,Janssen97}. In our
 distributed setting, word meanings are vectors of length $N$. The
-combination function maps pairs of them to single vectors of length $N$
-vectors, which can then be merged again to create more complex
+composition function maps pairs of them to single vectors of length $N$, 
+which can then be merged again to create more complex
 phrases. Once the entire sentence-level representation has been
@@
 The comparison layer uses the same kind of function template as the
-composition function (either an NN layer or an NTN layer) with
+composition layers (either an NN layer or an NTN layer) with
 independently learned parameters and a separate nonlinearity function.
@@
 
-We train the model with minibatch stochastic gradient descent (SGD)
+% Removed 'minibatch' -> For two of the experiments, we use minibatches
+% of size 1, which doesn't really count.
+We train the model using stochastic gradient descent (SGD)
 with learning rates computed using AdaGrad \cite{duchi2011adaptive}.

### 2014-06-06 writing/S'14 paper/join.tex
Typo
Proofreading

@@
 
-\begin{figure}[htp]
+\begin{figure}[tp]
   \centering
@@
 a domain of seven elements. This yields a data set consisting of
-6400 statements about pairs of entities. 3200 of these pairs are
+6400 statements about pairs of formulae. 3200 of these pairs are
 chosen as a test set, and that test set is further reduced to the 2960
@@
 that information to recover relations that were not overtly included
-in the training data. In contrast, the RNN model was able to
-approximate this behavior only incompletely. It is possible but not
-likely that it could be made to find a good solution with further
-optimization on different learning algorithms, or that it would do
-better on a larger universe of sets for which there would be a larger
-set of training data to learn from, but the RNTN is readily able to achieve
-these effects in the setting discussed here.
+in the training data. In contrast, the RNN model was able to achieve
+this behavior only incompletely. It is possible but not likely that it
+could be made to find a good solution with further optimization on
+different learning algorithms, or that it would do better on a larger
+universe of sets, for which there would be a larger set of training
+data to learn from, but the RNTN is readily able to achieve these
+effects in the setting discussed here.

### 2014-06-06 writing/S'14 paper/methods.tex
Proofreading

--- 2014-06-06 writing/S'14 paper/methods.tex
@@
 via specific combination functions \cite{Partee84,Janssen97}. In our
-distributed setting, word meanings are vectors of length $N$. The
-combination function maps pairs of them to single vectors of length $N$
+distributed setting, word meanings are vectors of length $n$. The
+combination function concatenates pairs of them to form vectors of
+length $2n$ and maps those larger vectors back into length $n$
 vectors, which can then be merged again to create more complex
@@
 \vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
-\end{gather} 
+\end{gather} % TODO: Explain third order tensor parameter?
 %
@@
 $\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
-them by an $N \times 2N$ matrix of learned weights, and applies the
+them by an $n \times 2n$ matrix of learned weights, and applies the
 element-wise non-linearity to the resulting vector. The RNTN has the
 same basic structure, but with the addition of a learned third-order
-tensor $\mathbf{T}$, dimension $N \times N \times N$, modeling
+tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
 multiplicative interactions between the child vectors. Both models

### 2014-06-06 writing/S'14 paper/quantifiers.tex
Proofreading

--- 2014-06-06 writing/S'14 paper/quantifiers.tex
@@
 investigating whether our models can capture this complexity, we now
-attempt to directly measure the degree to which RNNs are able to
+attempt to directly measure the degree to which they are able to
 develop suitable representations for the semantics of natural language
@@
 has formed a standard case study in prior formal work on natural
-language inference.
+language inference \cite{Icard:Moss:2013:LILT}.
 
@@
 that is generated directly using the logical system that we hope to
-model, yielding results that we believe to be substantially more
+learn, yielding results that we believe to be substantially more
 straightforward to interpret.
@@
 examples that are substantially different from those it was trained
-on. Crucially though, to succeed on this task, the model must be able
+on. Crucially, though, to succeed on this task, the model must be able
 to recognize all of the lexical relations between the nouns, verbs,
@@
 quantifiers to hold out before training, the resource demands of
-training prevent us from testing each of the 55 possible possible
-pairs of quantifiers, and we choose only four pairs to test on.  Three
+training prevent us from testing each of the 55 possible 
+pairs of quantifiers, so we choose only four pairs to test on.  Three
 of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
 \ii{all}/\ii{some}) were chosen because they allow for the most
-different class labels at in the training data. The fourth is a
+different class labels in the training data. The fourth is a
 self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
@@
   \centering
-  \begin{tabular}{ l rrr }
+  \setlength{\tabcolsep}{10pt}
+  \begin{tabular}{ l r r@{ \ }r r@{ \ }r }
     \toprule
-    Data & Most frequent class & 20 dim RNN (\%) & 20 dim RNTN (\%)\\
+    Data & Most frequent class & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
     \midrule
-    \textsc{all split}	& 35.4 &	65.3&	\textbf{99.0}
+    \textsc{all split}	            & 35.4 (7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}
     \\[1ex]    
-    \textsc{pair two/less-than-two}	& 59.8 &	77.6 &	\textbf{100.0} \\
-    \textsc{pair not-all/not-most}	&0 & 60.8 &	\textbf{82.0} \\
-    \textsc{pair all/some}	& 0& 52.5  &	\textbf{85.1} \\
-    \textsc{pair no/no}	& 30.8 &	68.9 &	\textbf{99.9} \\
+    \textsc{pair two/less-than-two} & 59.8 (18.4) & 77.2 &(57.4) &	\textbf{100.0} &\textbf{(100.0)} \\
+    \textsc{pair not-all/not-most}  &    0 (0)    & 66.0 &(54.3) &	\textbf{93.8}  &\textbf{(91.5)} \\
+    \textsc{pair all/some}	    &    0 (0)    & 62.7 &(65.9)  &	\textbf{78.4}  &\textbf{(84.7)} \\
+    \textsc{pair no/no}	            & 30.8  (10.2) & 67.8 & (61.8) &	\textbf{99.9}  &\textbf{(99.7)} \\
     \bottomrule
   \end{tabular}
-  \caption{Performance on the quantifier experiments. }
+  \caption{Performance on the quantifier experiments. Results are reported as accuracy scores followed by macroaveraged F1 scores in parentheses.}
   \label{resultstable}
@@
 Figure~\ref{resultstable}. We compare the results to a most frequent
-class baseline, which reflects the frequency in the test data of most
-frequent class in the training data, $\natind$.  After some
-cross-validation, we chose 20-dimensional word and phrase
-representations, and 75 dimensional feature vectors for the
-classifier.
+class baseline, which reflects the frequency in the test data of the
+most frequent class in the training data, $\natind$.  After some
+cross-validation, we chose 16- and 20-dimensional representations for
+the RNN and RNTN, respectively, and 75-dimensional feature vectors for
+the classifier.
 
@@
 Figure~\ref{prop-falloff}.  However, the perfect performance by the
-RNTN on the \textsc{pair no/no} experiment and its strong performance
-on \textsc{all split} and \textsc{pair two/less-than-two} suggests
-that that this stronger model is able to handle quantifiers correctly
-given sufficient training data, but the sub-90\% results on two of the
-training settings suggest that it may not be able to generalize from
-data this impoverished in general. However, though the question of how
-much data is necessary to accurately capture quantifier behavior in a
-na\"ive model remains open, the fact that both models perform far
-above baseline is promising.
+RNTN on the \textsc{all split} and \textsc{pair two/less-than-two}
+experiments, and its strong performance on the \textsc{pair no/no}
+experiment, suggest that this stronger model is able to handle
+quantifiers correctly given sufficient training data, but the weaker
+results on two of the training settings suggest that it may not be
+able to generalize from data in general. However, though the question
+of how much data is necessary to accurately capture quantifier
+behavior in a na\"ive model remains open, the fact that both models
+perform far above the baseline is promising.

### 2014-06-06 writing/S'14 paper/recursion.tex
Proofreading

@@
 between statements, we exhaustively enumerate the sets of assignments
-of truth values to proposition variables that would satisfy each of
-the statements and then convert the set-theoretic relation between
+of truth values to propositional variables that would satisfy each of
+the statements, and then we convert the set-theoretic relation between
 those assignments into one of the seven relations in
@@
 possible statements that the intermediate vector representations need
-to be able to distinguish if it is to learn this logical system in its
+to be able to distinguish if we are to learn this logical system in its
 full generality.
@@
 propositional logic, we disallow individual pairs of statements from
-referring to more than four of the six proposition variables.
+referring to more than four of the six propositional variables.
 
@@
 almost perfectly on unseen small test examples, with accuracy above
-98.7\% below size four (macro-F1 was above 95.5).  Starting at size
-four, performance gradually falls with increasing size.  The RNN model
-did not perform well, reaching only 90.5\% (81.4 micro-F1) accuracy on
-the smallest test examples, and declining from there to near-baseline
-performance at size 12. Training accuracy was 99.6\% for the RNTN and
-87.6\% for the RNN.
+98.7\% on formulae below length four (macro-F1 was above 95.5).
+Starting at size four, performance gradually falls with increasing
+size.  The RNN model did not perform well, reaching only 90.5\% (81.4
+micro-F1) accuracy on the smallest test examples, and declining from
+there to near-baseline performance on formulae of length 12. Training
+accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.
 

### 2014-06-06 writing/S'14 paper/join.tex
Typo in join

--- 2014-06-06 writing/S'14 paper/join.tex
@@
 If a model is to learn the behavior of a relational logic like the one
-presented here from a finite amount data, it must learn to deduce new
+presented here from a finite amount of data, it must learn to deduce new
 relations from seen relations in a sound manner. The simplest such
@@
   \end{array}$
-  \caption{Inference path from premises $a R S$ (row) and $b S c$ (column) to the relation that holds between $a$ and $c$, if any.
+  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.
 These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
@@
 \end{figure} 
-% TODO: Recheck strikethrough distribution.
+% Note: None of these test examples is derivable from the shown training data, 
+% but we suggest that there is additional training data, so we can cross lines
+% out willy-nilly without fear.
 

### 2014-06-06 writing/S'14 paper/recursion.tex
Some results tweak-

@@
 appears that the faulty logic that the RNN did learn decayed similarly
-with expression size.  This is not necessarily a significant flaw in
-the model, since it remains possible that a less lossy approximation
-of the same logic would do no better on the training objective, since
-that objective only considers short strings and includes a
-regularization term. Whether a more accurate encoding could be learned
-from a larger set of training data remains a question for future
-work.
+with expression size.
+This is not necessarily a significant flaw in the model, since it remains possible that  different training regimes would allow it to learn accurate representations of the larger expressions.
 

### 2014-06-06 writing/S'14 paper/methods.tex
Caltrain edits.

@@
 expressions are derived from the meanings of their constituent parts
-via specific combination functions \cite{Partee84,Janssen97}. In our
+via specific composition functions \cite{Partee84,Janssen97}. In our
 distributed setting, word meanings are vectors of length $n$. The
-combination function concatenates pairs of them to form vectors of
-length $2n$ and maps those larger vectors back into length $n$
-vectors, which can then be merged again to create more complex
+composition function maps pairs of them to single vectors of length $n$, 
+which can then be merged again to represent more complex
 phrases. Once the entire sentence-level representation has been
-derived, it is fed to a classifier trained in a supervised manner.
-
-% The models that we study are centered around a recursively applied composition function which is meant to mimic the recursive construction of meanings in formal semantics.
-In this scheme, pairs of words are merged into phrase representations by a function that maps from representations of length $2N$ to representations of length $N$.
-These phrase representations are then further merged with other words and phrases until the entire phrase or sentence being evaluated is represented in a single vector.
-This vector is then used as the input to a classifier and used in a supervised learning task.
+derived, it serves as a fixed-length input for some subsequent function.
 
@@
 independently learned parameters and a separate nonlinearity function.
-Rather than use a $\tanh$ nonlinearity here, we found a substantial
-improvement in performance by using a rectified linear function. In
+Rather than use a $\tanh$ nonlinearity here, we follow \cite{bowman2013can} 
+in using a rectified linear function. In
 particular, we use the leaky rectified linear function
-\cite{maasrectifier}: $f_{b}(\vec{x})=\max(\vec{x}, 0) +
+\cite{maasrectifier}: $f(\vec{x})=\max(\vec{x}, 0) +
 0.01\min(\vec{x}, 0)$, applied element-wise.
@@
 Because the seven classes are not balanced in general, we report performance
-using both accuracy and macroaveraged F1, since the latter is sensitive
-to performance on infrequent classes. We compute macroaveraged F1 
+using both accuracy and macroaveraged F1, since the latter emphasizes
+ performance on infrequent classes. We compute macroaveraged F1 
 as the harmonic mean of mean precision and mean recall, both computed

### 2014-06-06 writing/S'14 paper/quantifiers.tex
Some results tweak-
Caltrain edits.

@@
 represent a logic should be able to construct an accurate
-representation of each held out quantifier from the way that it
-interacts with the other nine quantifiers which are not held
-out. Since running this experiment requires choosing a pair of
+representation of each of the two quantifiers in the pair from the way that it
+interacts with the other nine quantifiers with which it is seen in training. 
+Since running this experiment requires choosing a pair of
 quantifiers to hold out before training, the resource demands of
@@
 \ii{all}/\ii{some}) were chosen because they allow for the most
-different class labels in the training data. The fourth is a
+different class labels in the test data. The fourth is a
 self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
@@
   \setlength{\tabcolsep}{10pt}
-  \begin{tabular}{ l r r@{ \ }r r@{ \ }r }
+  \begin{tabular}{ l rr@{ \ }r r@{ \ }r r@{ \ }r }
     \toprule
-    Data & Most frequent class & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
+    Data & \multicolumn{3}{c}{Most frequent class} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
     \midrule
-    \textsc{all split}	            & 35.4 (7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}
+    \textsc{all split}	            &~~~~& 35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}
     \\[1ex]    
-    \textsc{pair two/less-than-two} & 59.8 (18.4) & 77.2 &(57.4) &	\textbf{100.0} &\textbf{(100.0)} \\
-    \textsc{pair not-all/not-most}  &    0 (0)    & 66.0 &(54.3) &	\textbf{93.8}  &\textbf{(91.5)} \\
-    \textsc{pair all/some}	    &    0 (0)    & 62.7 &(65.9)  &	\textbf{78.4}  &\textbf{(84.7)} \\
-    \textsc{pair no/no}	            & 30.8  (10.2) & 67.8 & (61.8) &	\textbf{99.9}  &\textbf{(99.7)} \\
+    \textsc{pair two/less-than-two} && 59.8 &(18.4) & 77.2 &(57.4) &	\textbf{100.0} &\textbf{(100.0)} \\
+    \textsc{pair not-all/not-most}  &&    0 &(0)    & 66.0 &(54.3) &	\textbf{93.8}  &\textbf{(91.5)} \\
+    \textsc{pair all/some}	    &&    0& (0)    & 62.7 &(65.9)  &	\textbf{78.4}  &\textbf{(84.7)} \\
+    \textsc{pair no/no}	            && 30.8  &(10.2) & 67.8 & (61.8) &	\textbf{99.9}  &\textbf{(99.7)} \\
     \bottomrule
@@
 the RNN and RNTN, respectively, and 75-dimensional feature vectors for
-the classifier.
+the classifier. Training
+accuracy was near 80\% for the each of the five RNN experiments, and 
+near 98\% for each of the five RNTN experiments.
+
+%train rnn 0.20302, 0.017195
 
@@
 results on two of the training settings suggest that it may not be
-able to generalize from data in general. However, though the question
+able to generalize from this kind of restricted training data in general. However, though the question
 of how much data is necessary to accurately capture quantifier
-behavior in a na\"ive model remains open, the fact that both models
-perform far above the baseline is promising.
+behavior in a na\"ive model remains open, the fact that this model is able to perform so well is promising.

### 2014-06-06 writing/S'14 paper/recursion.tex
Caltrain edits.

@@
       range over all well-formed formulae, and $\sem{\cdot}$ is
-      interpretation function mapping formulae into $\{\True,
+      the interpretation function mapping formulae into $\{\True,
       \False\}$.}\label{tab:pl}
@@
 pairs like $(a \pland a)$ and $(a \pland (a \pland (a \pland a)))$,
-and similarly for disjunctions, but only the RNTN is able to learn
-substantially different representations for formulae like $(a \pland
-a)$ and $(a \pland (a \pland (a \pland
-b)))$. Figure~\ref{prop-falloff} summarizes the result for pairs of
-distinct formulae, showing that the RNN basically cannot distinguish
-them after depth~2.
+with both \ii{and} and \ii{or}, but only the RNTN is able to learn
+substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
+ and $(a \pland (a \pland b))$ when the difference between the two is embedded under multiple operators. 
+ Figure~\ref{prop-falloff} summarizes this result.
 
@@
   \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-  \caption{Truth-conditionally distinct formulae should have different
+  \caption{Semantically distinct formulae should have different
     representations. As measured by Euclidean distance, only the RNTN
-    achieves this for formulae containing more than two
-    connectives. The RNN quickly collapses the representations,
+    achieves this for formulae containing more than a small number of
+    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
     failing to capture the meaning contrasts.}

### 2014-06-06 writing/S'14 paper/discussion.tex
Add intro sentence.

@@
 
-This paper evaluated two recursive models in a series of three increasingly
+This paper evaluated two recursive models on a series of three increasingly
 challenging interpretive tasks involving natural language inference:

### 2014-08-25 writing/BayLearn abstract/intro.tex
Draft abstract.

--- 2014-06-06 writing/S'14 paper/intro.tex
@@
 simulate, our target logical
-concepts. % TODO: Update with current results
+concepts. These experiments differentiate the increased power of RNTNs
+better than previous work and provide the most convincing
+demonstration to date of the ability of neural networks to model
+semantic inferences in complex natural language sentences.
 
@@
   \end{tabular}
-  \caption{The entailment relations in $\mathfrak{B}$. 
+  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
     $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
@@
 % Citations to additional past work to be added.\\...\\...\\...\\...\\...
-
-% TODO: Mention earlier work on NN interpretation - \cite{Garcez-etal:2001}
 

### 2014-08-25 writing/BayLearn abstract/long_abstract.tex
Draft abstract.

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage[leqno, fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{qtree}
\usepackage[numbers]{natbib}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}

\definecolor{mylinkcolor}{rgb}{0,0,0} % black
\hypersetup{colorlinks, linkcolor=mylinkcolor, urlcolor=mylinkcolor, citecolor=mylinkcolor}

\newcommand{\nateq}{\equiv}
\newcommand{\natind}{\mathbin{\#}}
%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
\newcommand{\natneg}{\mathbin{^{\wedge}}}
\newcommand{\natfor}{\sqsubset}
\newcommand{\natrev}{\sqsupset}
\newcommand{\natalt}{\mathbin{|}}
\newcommand{\natcov}{\mathbin{\smallsmile}}

\newcommand{\plneg}{\mathop{\textit{not}}}
\newcommand{\pland}{\mathbin{\textit{and}}}
\newcommand{\plor}{\mathbin{\textit{or}}}

% Strikeout
\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}

\newcommand{\True}{\texttt{T}}
\newcommand{\False}{\texttt{F}}
\usepackage{stmaryrd}
\newcommand{\sem}[1]{\ensuremath{\llbracket#1\rrbracket}}

\renewcommand{\bibsection}{\subsubsection*{References}}

\usepackage{gb4e}

\def\ii#1{\textit{#1}}

\newcommand{\mynote}[1]{{\color{red}\framebox{\begin{tabular}{p{0.9\textwidth}}\footnotesize#1 \end{tabular}}}}

\title{Recursive Neural Networks for Learning Logical Semantics}

\author{
Samuel R.\ Bowman$^{\ast\dag}$ \\
\texttt{sbowman@stanford.edu} \\[2ex]
$^{\ast}$Stanford Linguistics \\
\And
Christopher Potts$^{\ast}$\\
\texttt{cgpotts@stanford.edu} \\[2ex]
$^{\dag}$Stanford NLP Group
\And
Christopher D.\ Manning$^{\ast\dag\ddag}$\\
\texttt{manning@stanford.edu}\\[2ex]
$^{\ddag}$Stanford Computer Science
}

% \author{
% Samuel R.\ Bowman \\
% NLP Group, Dept.\ of Linguistics\\
% Stanford University\\
% Stanford, CA 94305-2150 \\
% \texttt{sbowman@stanford.edu}
%  \And
%  Christopher Potts \\
% Dept.\ of Linguistics\\
% Stanford University\\
% Stanford, CA 94305-2150 \\
% \texttt{cgpotts@stanford.edu}
%  \And
% Christopher D.\ Manning \\
% NLP Group,  Depts.\ of Computer Science and Linguistics\\
% Stanford University\\
% Stanford, CA 94305-2150 \\
% \texttt{manning@stanford.edu}
% }

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle

% Compare with Grefenstette etc?

  Supervised recursive neural network models (RNNs) for sentence
  meaning have been successful in an array of sophisticated language
  tasks, but it remains an open question whether they can learn
  compositional semantic grammars that support logical deduction.  We
  address this question directly by for the first time evaluating
  whether each of two classes of neural model --- plain RNNs and
  recursive neural tensor networks (RNTNs) --- can correctly learn
  relationships such as entailment and contradiction between pairs of
  sentences, where we have generated controlled data sets of sentences
  from a logical grammar.  Our first experiment evaluates whether
  these models can learn the basic algebra of logical relations
  involved. Our second and third experiments extend this evaluation to
  complex recursive structures and sentences involving quantification.
  We find that the plain RNN achieves only mixed results on all three
  experiments, whereas the stronger RNTN model generalizes well in
  every setting and appears capable of learning suitable
  representations for natural language logical inference.

\subsection*{Recursive neural network models}

\begin{figure}[hp]
  \centering
  \input{figure1}
  % \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
  %  The same structure is used for both the RNN and RNTN layer functions.} 
  \label{sample-figure}
\end{figure}

In our experiments, we train pairs of recursive (tree structured) neural network models \cite{socher2013acl1} which are joined together with a shared top layer that generates features for a classifier. The classifier predicts the logical relation that holds between the sentences represented by the two trees. For an activation function, we use either a plain NN layer or a tensor combination layer.

% TODO gold parse structures

\subsection*{Reasoning about semantic relations}

\begin{table}[h]
  \centering
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{l c l l} 
    \toprule
    Name & Symbol & Set-theoretic definition & Example \\ 
    \midrule
    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
    \bottomrule
  \end{tabular}
  \label{b-table}
\end{table}

Our models classify into the seven relations from \cite{maccartney2009extended}, shown in the table below, which define possible relationships between pairs of terms or sentences of natural language in terms of their set-theoretic denotations.
If any model is to learn the behavior of a relational logic like the one
presented here from a finite amount of data, it must learn to deduce new
relations from already seen relations. Our first experiment evaluates the ability of our models to do this over pairs of atomic symbols in a large corpus of artificial data. The model is trained on examples like \{$a \natfor b$, $b \natneg c$\}, and tested on examples that follow from them, like \{$a \natalt c$\}.

\subsection*{Recursive structure in propositional logic}\label{sec:recursion}

Our second experiment introduces compositionality to our examples, training on short statements of propositional logic, like $\plneg a\natrev(a \pland b)$. We train our models on only pairs of statements with up to four symbols, but observe that the RNN performs reasonably both on those and on much longer test pairs.

\begin{figure}[h]
  \centering
  \includegraphics[width=4in]{recursion\string_results\string_final.eps}
  \label{prop-results}
\end{figure}

\subsection*{Reasoning with natural language quantifiers and negation}\label{sec:quantifiers}

For our third experiment, we generate pairs of sentences in which each sentence contains one quantifier, and any of a small set of common nouns, as in the example \textit{(no warthogs) move $\natfor$ (no (not reptiles)) swim}. The parentheses indicate the tree structure for each sentence as it will be used by the model. We defined several different types of train--test split for this experiment. A tuned RNTN model performed either well ($>85\%$ accuracy) or perfectly on all of them, while a tuned RNN did not break 80\% in any setting.

\bibliographystyle{unsrtnat}

\small % Note: Explicitly allowed in style guide
\bibliography{MLSemantics}

\end{document}

### 2014-08-25 writing/BayLearn abstract/paper.tex
Draft abstract.

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage[leqno, fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{qtree}
\usepackage[numbers]{natbib}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}

\definecolor{mylinkcolor}{rgb}{0,0,0} % black
\hypersetup{colorlinks, linkcolor=mylinkcolor, urlcolor=mylinkcolor, citecolor=mylinkcolor}

\newcommand{\nateq}{\equiv}
\newcommand{\natind}{\mathbin{\#}}
%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
\newcommand{\natneg}{\mathbin{^{\wedge}}}
\newcommand{\natfor}{\sqsubset}
\newcommand{\natrev}{\sqsupset}
\newcommand{\natalt}{\mathbin{|}}
\newcommand{\natcov}{\mathbin{\smallsmile}}

\newcommand{\plneg}{\mathop{\textit{not}}}
\newcommand{\pland}{\mathbin{\textit{and}}}
\newcommand{\plor}{\mathbin{\textit{or}}}

% Strikeout
\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}

\newcommand{\True}{\texttt{T}}
\newcommand{\False}{\texttt{F}}
\usepackage{stmaryrd}
\newcommand{\sem}[1]{\ensuremath{\llbracket#1\rrbracket}}

\renewcommand{\bibsection}{\subsubsection*{References}}

\usepackage{gb4e}

\def\ii#1{\textit{#1}}

\newcommand{\mynote}[1]{{\color{red}\framebox{\begin{tabular}{p{0.9\textwidth}}\footnotesize#1 \end{tabular}}}}

\title{Recursive Neural Networks for Learning Logical Semantics}

\author{
Samuel R.\ Bowman$^{\ast\dag}$ \\
\texttt{sbowman@stanford.edu} \\[2ex]
$^{\ast}$Stanford Linguistics \\
\And
Christopher Potts$^{\ast}$\\
\texttt{cgpotts@stanford.edu} \\[2ex]
$^{\dag}$Stanford NLP Group
\And
Christopher D.\ Manning$^{\ast\dag\ddag}$\\
\texttt{manning@stanford.edu}\\[2ex]
$^{\ddag}$Stanford Computer Science
}

% \author{
% Samuel R.\ Bowman \\
% NLP Group, Dept.\ of Linguistics\\
% Stanford University\\
% Stanford, CA 94305-2150 \\
% \texttt{sbowman@stanford.edu}
%  \And
%  Christopher Potts \\
% Dept.\ of Linguistics\\
% Stanford University\\
% Stanford, CA 94305-2150 \\
% \texttt{cgpotts@stanford.edu}
%  \And
% Christopher D.\ Manning \\
% NLP Group,  Depts.\ of Computer Science and Linguistics\\
% Stanford University\\
% Stanford, CA 94305-2150 \\
% \texttt{manning@stanford.edu}
% }

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle

  Supervised recursive neural network models (RNNs) for sentence
  meaning have been successful in an array of sophisticated language
  tasks, but it remains an open question whether they can learn
  compositional semantic grammars that support logical deduction.  We
  address this question directly by for the first time evaluating
  whether each of two classes of neural model --- plain RNNs and
  recursive neural tensor networks (RNTNs) --- can correctly learn
  relationships such as entailment and contradiction between pairs of
  sentences, where we have generated controlled data sets of sentences
  from a logical grammar.  Our first experiment evaluates whether
  these models can learn the basic algebra of logical relations
  involved. Our second and third experiments extend this evaluation to
  complex recursive structures and sentences involving quantification.
  We find that the plain RNN achieves only mixed results on all three
  experiments, whereas the stronger RNTN model generalizes well in
  every setting and appears capable of learning suitable
  representations for natural language logical inference.

  \section{Introduction}\label{sec:intro}

Supervised recursive neural network models (RNNs) for sentence meaning
have been successful in an array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,socher2013acl1},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
the ability of these models to learn compositional semantic grammars,
but it remains an open question whether they can achieve the same
results as grammars based in logical forms
\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
when it comes to the core semantic concepts of quantification,
entailment, and contradiction needed to identify the relationships
between sentences like \ii{every animal walks}, \ii{every turtle
  moves}, and \ii{most reptiles don't move}. To date, experimental
investigations of these concepts using distributed representations
have been largely confined to short phrases
\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
For robust natural language understanding, it is essential to model
these phenomena in their full generality on complex linguistic
structures.

We address this question in the context of \ii{natural language
  inference} (also known as \ii{recognizing textual entailment};
\cite{dagan2006pascal}), in which the goal is to determine the core
inferential relationship between two sentences. Much of the
theoretical work on this task (and some successful implemented models
\cite{maccartney2009extended,watanabe2012latent}) involves \ii{natural
  logics}, which are formal systems that define rules of inference
between natural language words, phrases, and sentences without the
need of intermediate representations in an artificial logical
language. Following \cite{bowman2013can}, we use the natural logic
developed by \cite{maccartney2009extended} as our formal model. This
logic defines seven core relations of synonymy, entailment,
contradiction, and mutual consistency, as summarized in
Table~\ref{b-table}, and it provides rules of semantic combination for
projecting these relations from the lexicon up to complex phrases. The
formal properties and inferential strength of this system are now
well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.

In our experiments, we use this pre-specified logical grammar to
generate controlled data sets encoding the semantic relationships
between pairs of expressions and evaluate whether each of two classes
of neural model --- plain RNNs and recursive neural tensor networks
(RNTNs, \cite{socher2013acl1}) --- can learn those relationships
correctly. In our first experiment (Section~\ref{sec:join}), we
evaluate the ability of these models to learn the core relational
algebra of natural logic from data. Our second experiment
(Section~\ref{sec:recursion}) extends this evaluation to relations
between complex recursive structures like $(a \plor b)$ and
$\plneg(\plneg a \pland \plneg b)$, and our third experiment
(Section~\ref{sec:quantifiers}) involves relations between quantified
statements like \ii{every reptile walks} and \ii{no turtle moves}.
We find that the plain RNN achieves only mixed results in all three
experiments, whereas the stronger RNTN model generalizes well in every
case, suggesting that it has in fact learned, or at least learned to
simulate, our target logical
concepts. These experiments differentiate the increased power of RNTNs
better than previous work and provide the most convincing
demonstration to date of the ability of neural networks to model
semantic inferences in complex natural language sentences.

\begin{table}[tp]
  \centering
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{l c l l} 
    \toprule
    Name & Symbol & Set-theoretic definition & Example \\ 
    \midrule
    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
    \bottomrule
  \end{tabular}
  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
    %is insufficient knowledge to choose a label.}
  \label{b-table}
\end{table}

% Citations to additional past work to be added.\\...\\...\\...\\...\\...

% Deep learning methods in NLP which learn vector representations for words have seen successful uses in recent years on increasingly sophisticated tasks \cite{collobert2011natural, socher2011semi, socher2013acl1, chen2013learning}. Given the still modest performance of semantically rich NLP systems in many domains---question answering and machine translation, for instance---it is worth exploring the degree to which learned vectors can serve as general purpose semantic representations. Much of the work to date analyzing vector representations for words (see \cite{baroni2013frege}) has focused on lexical semantic behaviors---like the similarity between words like \ii{Paris} and \ii{France}. Good similarity functions are valuable for many NLP tasks, but there are real use cases for which it is necessary to know how words relate to one another or to some extrinsic representation, and to model the ways in which word meanings combine to form phrase, sentence, or document meanings. This paper explores the ability of linguistic representations developed using supervised deep learning techniques to support interpretation and reasoning.

% There are two broad family of tasks that could be used to test the ability of a model to develop general purpose meaning representations. In an interpretation task, sentences are mapped onto some denotation, such as  \ii{true} or \ii{false} for statements, or a factual answer for questions. There has been preliminary work in developing distributed models for interpretation \cite{grefenstette2013towards, rocktaschellow}, but developing a representation of world knowledge that corresponds accurately to the content expressed in language introduces considerable design challenges. I approach the problem by way of an inference task instead. Inferring the truth of one sentence from another does not require any preexisting knowledge representations, but nonetheless requires a precise representation of sentence meaning. I borrow the structure of the task from MacCartney and Manning  \cite{maccartney2009extended}. In it, the model is presented with a pair of sentences, and made to label the logical relation between the sentences as equivalence, entailment, or any of five other classes, as here:

%\begin{quote}
%\begin{enumerate}\small
%\item Many smartphone users avoid high bills overseas by disabling data service.
%\item Not everyone uses their smartphones for email when traveling abroad.
%\end{enumerate} 
%$\Rightarrow$ Entailment
%\end{quote}

%In this paper, we test the ability of recursive models to on three simple tasks, each of which is meant to capture a property that is necessary in representing natural language meaning in the setting of inference. I begin with an overview of MacCartney and Manning's \cite{maccartney2009extended} framework for inference, and of the recursive neural networks that I study. by showing that these models can learn to correctly represent entailment representations between sentences. I then show that these models can capture the meanings of recursive structures accurately up to a sufficient depth. I finally close with a brief demonstration of the ability of these models to reason over short natural language sentences involving quantifiers.

% \subsection{Natural language inference and natural logic}

% In its simplest form, \ii{natural language inference} (also known as \ii{recognizing textual entailment}, as in \cite{dagan2006pascal}) is the task of determine whether one sentence entails another. Much of the theoretical work on this task (and some successful implemented models \cite{maccartney2009natural, watanabe2012latent}) involve \ii{natural logic}, formal systems that define sound rules of inference from one sentence of natural language to another without the need for intermediate representations in some other logic. The most powerful model that we are aware of for natural logic is due to MacCartney and Manning \cite{maccartney2009extended} and Icard \cite{icard2012inclusion}, and is centered around the definition of a set of seven logical relations which can hold between sentences, shown in Table \ref{b-table}.

% This approach to natural logic can capture much of the complexity of natural language meaning within a well understood framework, and is also fairly straightforward to implement in a machine learning setting since it can be reduced to a seven-way classification problem on sentence pairs. Our goal in this paper is to learn recursive neural network models which are able to mimic key behaviors of this system.

% \begin{table}
% \begin{center}
% \begin{tabular}{|c|c|c|c|} \hline
% name & symbol & set-theoretic definition & example \\ \hline \hline
% entailment & $x \sqsubset y$ & $x \subset y$ & \ii{crow, bird}  \\ \hline
% reverse entailment & $x \sqsupset y$ & $x \supset y$ & \ii{Asian, Thai}  \\ \hline
% equivalence & $x \equiv y$ & $x = y$ & \ii{couch, sofa} \\ \hline
% alteration & $x$ $|$ $y$ & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{cat, dog} \\ \hline
% negation & $x \natneg y$ & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{able, unable} \\ \hline
% cover & $x \smallsmile y$ & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-ape} \\ \hline
% independence & $x$ \# $y$ & (else) & \ii{hungry, hippo}\\ \hline
% \end{tabular}
% \caption{The entailment relations in  $\mathfrak{B}$. $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, and the relation \# applies whenever none of the other six do, including when there is insufficient knowledge to choose a label.}
% \label{b-table}
% \end{center}
% \end{table}

% The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes. It should be noted that this is not a balanced classification problem. If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them. % TODO: How much should we say about this?

% The natural logic engine at the core of MacCartney and Manning's system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures. These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
%for dozens of different cases of insertion, deletion and substitution of different types of lexical item. For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences. Here, the substitution of \ii{no} for \ii{some}  yields $\natneg$ and the substitution of \ii{dogs} for \ii{animals} yields $\sqsupset$. It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
%\begin{quote}

%1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
%2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
%3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}

%\end{quote}

% Work to date on inference in neural network models is quite limited.
% \citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
% In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations. They propose a deterministic inference engine (similar to MacCartney's) which is augmented by a probabilistic component that evaluates individual lexical substitution steps in the derivation using vector representations, though again these representations are not learned, and no evaluations of this system have been published to date.
% \label{sec2}

\section{Recursive neural network models} \label{methods}

We study neural models that adhere to the linguistic \ii{principle of
 compositionality}, which says that the meanings for complex
expressions are derived from the meanings of their constituent parts
via specific composition functions \cite{Partee84,Janssen97}. In our
distributed setting, word meanings are vectors of length $n$. The
composition function maps pairs of them to single vectors of length $n$, 
which can then be merged again to represent more complex
phrases. Once the entire sentence-level representation has been
derived, it serves as a fixed-length input for some subsequent function.

We use the model architecture proposed in \cite{bowman2013can} and
depicted in Figure~\ref{sample-figure}. The two phrases being compared
are built up separately on each side of the tree, using the same
composition function, until they have each been reduced to single
vectors. The resulting vectors are fed into a separate comparison
layer that is meant to generate a feature vector capturing the
relation between the two phrases. The output of this layer is then
given to a softmax classifier, which in turn produces a hypothesized
distribution over the seven relations represented in Table~\ref{b-table}.

For a composition layer, we evaluate models with both the plain neural
network layer function \eqref{rnn} and the RNTN layer function
\eqref{rntn} proposed in \citet{chen2013learning}. A sigmoid
nonlinearity (element-wise $\tanh$) is applied to the output of either
layer function, following \cite{socher2013acl1}.
%
\begin{gather} \label{rnn}
\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
\label{rntn}
\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
\end{gather} 
%
Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
representations for the left and right children of the node, and
$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
them by an $n \times 2n$ matrix of learned weights, and applies the
element-wise non-linearity to the resulting vector. The RNTN has the
same basic structure, but with the addition of a learned third-order
tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
multiplicative interactions between the child vectors. Both models
include a bias vector~$\vec{b}$.

The comparison layer uses the same kind of function template as the
composition layers (either an NN layer or an NTN layer) with
independently learned parameters and a separate nonlinearity function.
Rather than use a $\tanh$ nonlinearity here, we follow \cite{bowman2013can} 
in using a rectified linear function. In
particular, we use the leaky rectified linear function
\cite{maasrectifier}: $f(\vec{x})=\max(\vec{x}, 0) +
0.01\min(\vec{x}, 0)$, applied element-wise.

To run the model forward and label a pair of phrases, the structure of
the lower layers of the network is assembled so as to mirror the tree
structures provided for each phrase. The word vectors are then looked
up from the vocabulary matrix $V$ (one of the model parameters), and
the composition and comparison functions are used to pass information
up the tree and into the classifier at the top. For an objective
function, we use the negative log of the probability assigned to the
correct label.

% Removed 'minibatch' -> For two of the experiments, we use minibatches
% of size 1, which doesn't really count.
We train the model using stochastic gradient descent (SGD)
with learning rates computed using AdaGrad \cite{duchi2011adaptive}.
The parameters (including the vocabulary) are initialized randomly 
using a uniform distribution over $[-0.1, 0.1]$.
Because the seven classes are not balanced in general, we report performance
using both accuracy and macroaveraged F1, since the latter emphasizes
 performance on infrequent classes. We compute macroaveraged F1 
as the harmonic mean of mean precision and mean recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined. Source code and generated data will be released
after the review period.

%\begin{figure}[tp]
%  \centering
% % \includegraphics[scale=0.35]{model.eps}
%  \input{figure1}
%  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
%    The same structure is used for both the RNN and RNTN layer functions.} 
%  \label{sample-figure}
%\end{figure}

%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?

\section{Reasoning about semantic relations}\label{sec:join}

If a model is to learn the behavior of a relational logic like the one
presented here from a finite amount of data, it must learn to deduce new
relations from seen relations in a sound manner. The simplest such
deductions involve atomic statements using the relations in
Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
\natrev c$, one can conclude that $a \natrev c$, by basic
set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
full set of sound inferences of this form is summarized in
Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
of relations for which no valid inference can be drawn in our logic.

% about the relations themselves that do not depend on the
% internal structure of the things being compared. For example, given
% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
% understanding $a$, $b$, or $c$. These seven relations support more
% than just transitivity: MacCartney and Manning's
% \cite{maccartney2009extended} join table defines 32 valid inferences
% that can be made on the basis of pairs of relations of the form $a R
% b$ and $b R' c$, including several less intuitive ones such as that if
% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
%
%\begin{table}[htp]
%  \centering  
%  \setlength{\arraycolsep}{8pt}
%  \renewcommand{\arraystretch}{1.1}
%  \newcommand{\UNK}{\cdot}  
%  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
%    %\hline
%    \multicolumn{1}{c}{}
%             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
%    \cline{2-8}
%    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
%    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
%    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
%    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
%    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
%    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
%    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
%    \cline{2-8}
%  \end{array}$
%  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.  These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}. We assess our models' ability to reproduce such inferential paths.}
%  \label{tab:jointable}
%\end{table}

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create small boolean structures for our logic in which terms denote
sets of entities from a small domain.  Figure~\ref{lattice-figure}
depicts a structure of this form. The lattice gives the full model,
for which all the statements on the right are valid. We divide these
statements evenly into training and test sets, and remove from the
test set those statements which cannot be proven from the training
statements.
% using the logic depicted in Figure~\ref{lattice-figure}.

% We test the model's ability to learn this behavior by creating
% artificial data sets of terms which represent sets of numbers. Since
% MacCartney and Manning's set of relations hold between sets as well as
% between sentences, we can use the underlying set structure to generate
% the all of the relations that hold between any pair of these terms, as
% in Figure \ref{lattice-figure}. We train the model defined above on a
% subset of these relations, but rather then presenting the model with a
% pair of tree-structured sentences as inputs, simply present it with
% two single terms, each of which corresponds to a single vector in the
% (randomly initialized) vocabulary matrix $V$, ensuring that the model
% has no information about the terms being compared except the relations
% between them.

In our experiments, we create 80 randomly generated sets drawing from
a domain of seven elements. This yields a data set consisting of
6400 statements about pairs of formulae. 3200 of these pairs are
chosen as a test set, and that test set is further reduced to the 2960
examples that can be provably derived from the training data.

We trained versions of both the RNN model and the RNTN model on these
data sets. In both cases, the models were implemented exactly as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer was not involved, and the two models differed only in which
layer function was used for the comparison layer. We simply present
the models with two single terms, each of which corresponds to a
single vector in the (randomly initialized) vocabulary matrix $V$,
thereby ensuring that the model has no information about the terms
being compared except the relations between them.

\paragraph{Results} 
We found that the RNTN model worked best with 11-dimensional vector
representations for the 80 sets and a 90-dimensional feature vector
for the classifier, though the performance of the model was not highly
sensitive to either dimensionality parameter in any of the experiments
discussed here. This model was able to correctly label 99.4\% (94.3
macroaveraged F1) of the provably derivable test examples, and 95.4\%
(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
best with 11 and 75 dimensions, respectively, but was able to achieve
accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
the RNN.

\begin{figure}[tp]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
    \setlength{\unitlength}{1cm}
    \begin{picture}(5,5.5)
      \labelednode{2.50}{5}{}{1,2,3}

      \put(0.75,4){\line(3,1){1.5}}
      \put(2.5,4){\line(0,1){0.5}}
      \put(4.25,4){\line(-3,1){1.5}}

      \labelednode{0.75}{3.5}{a,b}{1,2}
      \labelednode{2.50}{3.5}{c}{1,3}
      \labelednode{4.25}{3.5}{d}{2,3}

      \put(0.75,2.5){\line(0,1){0.5}}
      \put(0.75,2.5){\line(3,1){1.5}}

      \put(2.5,2.5){\line(-3,1){1.5}}
      \put(2.5,2.5){\line(3,1){1.5}}

      \put(4.25,2.5){\line(0,1){0.5}}
      \put(4.25,2.5){\line(-3,1){1.5}}

      \labelednode{0.75}{2}{e,f}{1}
      \labelednode{2.50}{2}{}{2}
      \labelednode{4.25}{2}{g,h}{3}

      \put(2.5,1){\line(-3,1){1.5}}
      \put(2.5,1){\line(0,1){0.5}}
      \put(2.5,1){\line(3,1){1.5}}

      \labelednode{2.5}{0.5}{}{}
    \end{picture}
    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
  \end{subfigure}
  \qquad
  \begin{subfigure}[t]{0.43\textwidth}
    \centering
    \setlength{\tabcolsep}{12pt}
    \begin{tabular}[b]{c  c}
      \toprule
      Train & Test \\
      \midrule
                    & $b \nateq b$ \\
      $b \natcov c$ &               \\
                    & $b \natcov d$ \\
                    & \strikeout{$b \natrev e$} \\
      $c \natcov d$ &               \\
      $c \natrev e$ &               \\
                    & \strikeout{$c \nateq f$} \\
      $c \natrev g$ &               \\ 
                    & $e \natfor b$ \\
      $e \natfor c$ &               \\[-1ex]
      $\vdots$      & $\vdots$ \\
      \bottomrule
    \end{tabular}

    \caption{A train/test split of the atomic statements about the
      model.  Test statements not provable from the training data are
      crossed out.}
  \end{subfigure}  
  \caption{Small example structure and data for learning relation composition.}
  \label{lattice-figure}
\end{figure} 
% Note: None of these test examples is derivable from the shown training data, 
% but we suggest that there is additional training data, so we can cross lines
% out willy-nilly without fear.

% RNTN log: tue-j-11-6x80-hip.txt
% Train PER: 0.0021875

% RNN log: tue-j-11-6x80-r-75.txt
% Train PER: 0.047188

% and create a dataset consisting of the relations between every pair of
% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
% test dataset, and that test dataset was further split into the 2960
% examples that can be provably derived from the test data using
% MacCartney and Manning's join table (or by the symmetry of the
% relations in about half of the cases) and the 240 that
% cannot.

% We tested a version of both the RNN model and the RNTN model on these
% data. In both cases, the models were implemented exactly as described
% in Section~\ref{methods}, but since the items being compared are
% single terms rather than full tree structures, the composition layer
% was not used, and the two models differed only in which layer function
% was used for the comparison layer. We found that the RNTN model worked
% best with 11 dimensional vector representations for the 80 sets and a
% 90 dimensional feature vector for the classifier. This model was able
% to correctly label 99.3\% of the derivable test examples, and 99.1\%
% of the remaining examples. The simpler RNN model worked best with 11
% and 75 dimensions, respectively, but was able to achieve accuracies of
% only 90.0\% and 87.\%, respectively.

These results are fairly straightforward to interpret. The RNTN model
was able to accurately encode the relations between the terms in the
geometric relations between their vectors, and was able to then use
that information to recover relations that were not overtly included
in the training data. In contrast, the RNN model was able to achieve
this behavior only incompletely. It is possible but not likely that it
could be made to find a good solution with further optimization on
different learning algorithms, or that it would do better on a larger
universe of sets, for which there would be a larger set of training
data to learn from, but the RNTN is readily able to achieve these
effects in the setting discussed here.

\section{Recursive structure in propositional logic}\label{sec:recursion}

Recursive structure is a prominent feature of natural languages and an
important part of their expressivity, as it allows people to use and
interpret complex structures they have never encountered before.
Theoretical accounts of natural language syntax and semantics
therefore rely heavily on recursive structures, and we expect that
accurate computational models will have to come to grips with them as
well. The present section pursues this question for our two classes of
RNN. In our evaluations, we exploit the fact that our logical language
is infinite by testing on strings that are longer and more complex
than any seen in training.

% Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice
%   said hello}, and \ii{Carl thinks that Bob said that Alice said
%   hello}. Overt recursion of this kind is easy to find, and
% theoretical accounts of natural language syntax and semantics rely
% heavily on recursive structures. In order for a model to be able to
% accurately learn natural language meanings, then, we expect that it
% would need to be able to learn to represent the meanings of function
% words in a such a way that they are able to behave correctly when
% taking their own outputs as input. In evaluating the model, we take
% advantage of the fact that recursive structures of this kind define
% potentially infinite languages by testing the model on strings that
% are longer and more complex than any seen in testing.

\paragraph{Experiments}
As in Section~\ref{sec:join}, we test this phenomenon within the
framework of natural logic, but we now replace the unanalyzed symbols
from that experiment with complex formulae. These formulae
represent a truth-functionally complete classical propositional logic:
each atomic variable denotes either $\True$ or $\False$, and the only
operators are truth-functional ones.  Table~\ref{tab:pl} defines this
logic, and Table~\ref{tab:plexs} gives some examples of relational
statements that we can make in these terms. To compute these relations
between statements, we exhaustively enumerate the sets of assignments
of truth values to propositional variables that would satisfy each of
the statements, and then we convert the set-theoretic relation between
those assignments into one of the seven relations in
Table~\ref{b-table}. As a result, each relational statement represents
a valid theorem of the propositional logic.
%
%\begin{table}[tp]
%  \centering
%  \begin{subtable}[t]{0.45\textwidth}
%    \centering
%    \begin{tabular}[t]{l l}
%      \toprule
%      Formula     & Interpretation \\
%      \midrule
%      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
%      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
%      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
%      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
%      \bottomrule
%    \end{tabular}    
%    \caption{Well-formed formulae. $\varphi$ and $\psi$
%      range over all well-formed formulae, and $\sem{\cdot}$ is
%      the interpretation function mapping formulae into $\{\True,
%      \False\}$.}\label{tab:pl}
%  \end{subtable}
%  \quad
%  \begin{subtable}[t]{0.45\textwidth}
%    \centering
%    \begin{tabular}[t]{r c l}
%      \toprule
%      $\plneg a$        & $\natneg$ & $a$ \\
%      $\plneg \plneg a$ & $\nateq$  & $a$ \\
%      $a$               & $\natfor$ & $(a \plor b)$ \\
%      $a$               & $\natrev$ & $(a \pland b)$ \\
%      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
%      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
%      \bottomrule
%    \end{tabular}
%    \caption{Examples of statements about relations between
%      well-formed formulae, defined in terms of sets of satisfying
%      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
%  \end{subtable}
%  \caption{Propositional logic with natural logic relations.}  
%  \label{prop-figure}
%\end{table}

% NOTE: There was some confusion at CSLI and the 224U lecture about
% whether this was theorem proving or some kind of model checking.
% Hopefully this is explicit enough.
Socher et al.~\cite{socher2012semantic} show that a matrix-vector RNN
model somewhat similar to our RNTN can learn a logic, but the logic
discussed here is a much better approximation of the kind of
expressivity needed for natural language. The logic learned in that
experiment is boolean, wherein the atomic symbols are simply the
values $0$ and $1$, rather than variables over those values. While
learning the operators of that logic is not trivial, the outputs of
each operator can be represented accurately by a single bit. The
statements of propositional logic learned here describe conditions on
unknown truth values of propositions. As opposed to the two-way
contrasts seen in \cite{socher2012semantic}, this logic distinguishes
between $2^{6} = 64$ possible assignments of truth values, and
expressions of this logic define arbitrary conditions on these
possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
possible statements that the intermediate vector representations need
to be able to distinguish if we are to learn this logical system in its
full generality.

For our experiments, we randomly generate a large set of  unique pairs 
of formulae and compute the relation that holds for each pair.
We discard pairs in which either statement is a tautology or
contradiction (a statement that is true of either all or no possible
assignments), for which none of the seven relation labels in
Table~\ref{b-table} can hold. The resulting set of formula pairs is
then partitioned into bins based on the number of logical operators in
the longer of the two formulae. We then randomly sample 15\% of each
bin for a held-out test set.

If we do not implement any constraint that the two statements being
compared are similar in any way, then the generated data are dominated
by statements in which the two formulae refer to largely separate
subsets of the six variables, which means that the $\natind$ relation
is almost always correct.  In an effort to balance the distribution of
relation labels without departing from the basic task of modeling
propositional logic, we disallow individual pairs of statements from
referring to more than four of the six propositional variables.

\paragraph{Results} 
We trained both the RNN and RNTN models on data with up to 4
connectives (65k pairs) and tested it on examples of up to 12
connectives (44k pairs). We initialized the model parameters randomly,
including the vector representations of the six variables. The results
are shown in Figure~\ref{prop-results}. In tuning, we found that the
RNN model was approximately optimal with 45-dimensional vector
representations, and the RNTN model was approximately optimal with 25
dimensions. We fixed the size of the feature vector for the classifier
at 75 dimensions. We found that the RNTN model was able to perform
almost perfectly on unseen small test examples, with accuracy above
98.7\% on formulae below length four (macro-F1 was above 95.5).
Starting at size four, performance gradually falls with increasing
size.  The RNN model did not perform well, reaching only 90.5\% (81.4
micro-F1) accuracy on the smallest test examples, and declining from
there to near-baseline performance on formulae of length 12. Training
accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.

% thu-and-or-deep-25-32b.txt iter 136
% mon-and-or-deep-45r-32.txt iter 480

\begin{figure}[tp]
  \centering
  \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
  \caption{Model performance on propositional logic, by expression size. 
    The dashed line indicates that only expressions of size four or less appeared in the training data.}  
  \label{prop-results}
\end{figure}

The performance of the RNTN model on small unseen test examples
indicates that it learned a correct approximation of the underlying
logic. It appears that this approximation is accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing depth (and with the increasing complexity of the
expressions), resulting in gradually dropping performance. In
contrast, the RNN model did not appear to be able to learn a correct
approximation of the logic for statements of any length, but it
appears that the faulty logic that the RNN did learn decayed similarly
with expression size.  This is not necessarily a significant flaw in the model, since it remains possible that  different training regimes would allow it to learn accurate representations of the larger expressions.

In an effort to better understand why the RNTN outperformed the plain
RNN, we evaluated both models on pairs of long formulae involving
binary connectives, to assess the extent to which they achieve
substantially different representations for semantically distinct
formulae and substantially similar representations for synonymous
formulae. We found that both models are able to identify synonymous
pairs like $(a \pland a)$ and $(a \pland (a \pland (a \pland a)))$,
with both \ii{and} and \ii{or}, but only the RNTN is able to learn
substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
 and $(a \pland (a \pland b))$ when the difference between the two is embedded under multiple operators. 
 Figure~\ref{prop-falloff} summarizes this result.

% we discovered that this model looses information about in longer
% statements in a particularly problematic way. It appears that that
% model is unable to distinguish between two sentences when the only
% difference between those sentences is embedded within a very deep
% structure. We evaluated both models on sentences that differed in only
% one term, but for which that one term was embedded under a large
% number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and
%   b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and
%   b)))}. We then measured the Euclidean distance between the vector
% representations of the two sentences in each pair. Our findings are
% shown in Figure~\ref{prop-falloff}, and show that while the RNTN can
% distinguish the two sentences well at every size that we test, the RNN
% fails after a depth of approximately six.

%\begin{figure}[htp]
%  \centering
%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
%  \caption{Semantically distinct formulae should have different
%    representations. As measured by Euclidean distance, only the RNTN
%    achieves this for formulae containing more than a small number of
%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
%    failing to capture the meaning contrasts.}
%  \label{prop-falloff}
%\end{figure}

\section{Reasoning with natural language quantifiers and negation}\label{sec:quantifiers}

We have seen that the RNTN can learn an approximation of propositional
logic.  However, natural languages can express functional meanings of
considerably greater complexity than this.  As a first step towards
investigating whether our models can capture this complexity, we now
attempt to directly measure the degree to which they are able to
develop suitable representations for the semantics of natural language
quantifiers like \ii{some} and \ii{all}. Quantification is far from
the only place in natural language where complex functional meanings
are found, but it is a natural starting point, since it can be tested
in sentences whose structures are otherwise quite simple, and since it
has formed a standard case study in prior formal work on natural
language inference \cite{Icard:Moss:2013:LILT}.

% \subsection{Data}

This experiment replicates similar work described in
\cite{bowman2013can}, which found that RNTNs can learn to reason well
with quantifier meanings given sufficient training data. This paper
replaces the partially manually annotated data in that paper with data
that is generated directly using the logical system that we hope to
learn, yielding results that we believe to be substantially more
straightforward to interpret.

\paragraph{Experiments}
Our experimental data consist of pairs of sentences generated from a
small artificial grammar. Each sentence contains a quantifier, a noun
which may be negated, and an intransitive verb which may be
negated. We use the basic quantifiers \ii{some}, \ii{most}, \ii{all},
\ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}. We also
include five nouns, four intransitive verbs, and the negation symbol
\ii{not}. In order to be able to define relations between sentences
with differing lexical items, we define the lexical relations between
each noun--noun pair, each verb--verb pair, and each
quantifier--quantifier pair. The grammar accepts aligned pairs of
sentences of this form and calculates the natural logic relationship
between them.  Some examples of these data are provided in
Table~\ref{examplesofdata}.  As in previous sections, the goal of
learning is then to assign these relational labels accurately to
unseen pairs of sentences.

%nouns = ['warthogs', 'turtles', 'mammals', 'reptiles', 'pets']
%verbs = ['walk', 'move', 'swim', 'growl']
%dets = ['all', 'not_all', 'some', 'no', 'most', 'not_most', 'two', 'lt_two', 'three', 'lt_three']
%adverbs = ['', 'not']

% To assign relation labels to sentence pairs, we built a small
% task-specific implemenation of MacCartney's logic that can
% accurately label sentences of this restricted language. The logic is
% not able to derive all intuitively true relations of this language,
% and fails to derive a single unique relation for certain types of
% statement, including De Morgean's laws (e.g. \ii{(all pets) growl
% $\natneg$ (some pet) (not growl)}), and we simply discard these
% examples. Exhaustively generating the valid sentences under this
% grammar and choosing those to which a relation label can be assigned
% yields 66k sentence pairs. Some examples of these data are provided
% in Table~\ref{examplesofdata}.

%\begin{table}[htp]
%  \centering
%  \begin{tabular}{l c l}
%    \toprule
%    (most warthogs) walk         & $\natneg$ & (not-most warthogs) walk\\
%    (most mammals) move          & $\natind$ &  (not-most (not turtles)) move\\
%    (most (not pets)) (not swim) & $\natrev$ & (not-most (not pets)) move 
%    \\[2ex]    
%    (no turtles) (not growl)     & $\natalt$ & (no turtles) (not swim)\\
%    (no warthogs) swim           & $\natrev$ & (no warthogs) move\\
%    (no warthogs) move           & $\natfor$ & (no (not reptiles)) swim\\
%    \bottomrule
%  \end{tabular}
%  \caption{Sample data involving two different quantifier pairs.}
%  \label{examplesofdata}
%\end{table}

We evaluate the model using two experimental settings. In the simpler
setting, \textsc{all split}, we randomly sample 85\% of the data and evaluate on the
remaining 15\%. In this setting, the model is being asked to learn a
complete reasoning system for the limited language and logic presented
in the training data, but it is not being asked to generalize to test
examples that are substantially different from those it was trained
on. Crucially, though, to succeed on this task, the model must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
\eqref{p1} and \eqref{p2} in training and, from that information,
determine \eqref{p3}.
%
% do not allow a blank line --- adds too much space
%
\begin{gather}
  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
  \\
  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
  \\
  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
\end{gather}
%
% do not allow a blank line --- adds too much space
%
While our primary interest is in discovering the extent to which our
models can learn to encode the logic given an arbitrary amount of
data, we are also interested in the degree to which they can infer a
correct representation for the logic from more constrained training
data. To this end, we segment the sentence pairs according to which
quantifiers appear in each pair, and then hold out one such pair for
testing. We hypothesize that a model that can efficiently learn to
represent a logic should be able to construct an accurate
representation of each of the two quantifiers in the pair from the way that it
interacts with the other nine quantifiers with which it is seen in training. 
Since running this experiment requires choosing a pair of
quantifiers to hold out before training, the resource demands of
training prevent us from testing each of the 55 possible 
pairs of quantifiers, so we choose only four pairs to test on.  Three
of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
\ii{all}/\ii{some}) were chosen because they allow for the most
different class labels in the test data. The fourth is a
self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
handles equality.

\begin{table}[tp]
  \centering
  \setlength{\tabcolsep}{10pt}
  \begin{tabular}{ l rr@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    Data & \multicolumn{3}{c}{Most frequent class} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
    \midrule
    \textsc{all split}	            &~~~~& 35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}
    \\[1ex]    
    \textsc{pair two/less-than-two} && 59.8 &(18.4) & 77.2 &(57.4) &	\textbf{100.0} &\textbf{(100.0)} \\
    \textsc{pair not-all/not-most}  &&    0 &(0)    & 66.0 &(54.3) &	\textbf{93.8}  &\textbf{(91.5)} \\
    \textsc{pair all/some}	    &&    0& (0)    & 62.7 &(65.9)  &	\textbf{78.4}  &\textbf{(84.7)} \\
    \textsc{pair no/no}	            && 30.8  &(10.2) & 67.8 & (61.8) &	\textbf{99.9}  &\textbf{(99.7)} \\
    \bottomrule
  \end{tabular}
  \caption{Performance on the quantifier experiments. Results are reported as accuracy scores followed by macroaveraged F1 scores in parentheses.}
  \label{resultstable}
\end{table}

\paragraph{Results} 
The results for these experiments are shown in
Figure~\ref{resultstable}. We compare the results to a most frequent
class baseline, which reflects the frequency in the test data of the
most frequent class in the training data, $\natind$.  After some
cross-validation, we chose 16- and 20-dimensional representations for
the RNN and RNTN, respectively, and 75-dimensional feature vectors for
the classifier. Training
accuracy was near 80\% for the each of the five RNN experiments, and 
near 98\% for each of the five RNTN experiments.

%train rnn 0.20302, 0.017195

The RNN performed poorly at this task, even though the sentences used
in these examples are short enough to avoid the pathology shown in
Figure~\ref{prop-falloff}.  However, the perfect performance by the
RNTN on the \textsc{all split} and \textsc{pair two/less-than-two}
experiments, and its strong performance on the \textsc{pair no/no}
experiment, suggest that this stronger model is able to handle
quantifiers correctly given sufficient training data, but the weaker
results on two of the training settings suggest that it may not be
able to generalize from this kind of restricted training data in general. However, though the question
of how much data is necessary to accurately capture quantifier
behavior in a na\"ive model remains open, the fact that this model is able to perform so well is promising.

\section{General discussion}\label{sec:discussion}

This paper evaluated two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference:
the core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. The results suggest that RNTNs,
but not plain RNNs, have the capacity to meet the challenges of these
tasks with reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

Of course, challenges remain. In terms of our experimental data, even
the RNTN falls short of perfection in our more complex tasks, with
performance falling off steadily as the depth of recursion grows. It
remains to be seen whether these deficiencies can be overcome with
improvements to the model, the optimization procedures, or the
linguistic representations
\cite{sochergrounded,kalchbrenner2014convolutional}. In addition,
there remain subtle questions about how to fairly assess whether these
models have truly generalized in the way we want them to. There is a
constant tension between showing the models training data that gives
them a chance to learn the target logical functions and revealing the
answer to them in a way that leads to overfitting. The underlying
logical theories provide only limited guidance on this point.
%
%, and the fact that there is a finite universe of possible
%  expressions makes his an unavoidable issue. 
%
%  CP: I don't understand the above. It is false if taken literally;
%  our PL generates an infinte number of formulae.
%
Finally, we have only scratched the surface of the logical complexity
of natural language; in future experiments, we hope to test sentences
with embedded quantifiers, multiple interacting quantifiers, relative
clauses, and other kinds of recursive structure. Nonetheless, the
rapid progress the field has made with these models in recent years
provides ample reason to be optimistic that they can be trained to
meet the challenges of natural language semantics.

% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators. This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic. The positive results that we have found, however, are extremely promising for the future of learned representation models in the applied modeling of meaning. We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators. We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.

% There is ample room to build on these results. In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses. In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.

% \textit{Withheld for review.}\\ \\
% Do later: Fill in, cite grants, internal reviewers

\bibliographystyle{unsrtnat}

\small % Note: Explicitly allowed in style guide
\bibliography{MLSemantics}

\end{document}

### 2014-08-25 writing/BayLearn abstract/long_abstract.tex
Updated draft abstract.

@@
 \begin{figure}[hp]
-  \centering
-  \input{figure1}
+  \centering\resizebox{4.5in}{!}{
+  \input{figure1}}
   % \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
@@
 
-\subsection*{Reasoning about semantic relations}
+\subsection*{Inference and the semantic relations}
 
-\begin{table}[h]
-  \centering
+Our aim is to evaluate the ability of learned recusive models to represent semantic structure.
+We pursue this through an inference task, where the models must learn to choose the logical relation that applies between a pair of statements.
+The possible relations are the seven below from \cite{maccartney2009extended}.
+This is distinct from the limited prior work on learning neural network models for formal semantics \cite{grefenstette2013towards, rocktaschellow}, which use an interpretation task in which the model evaluates sentences as \textit{true} or \textit{false} with respect to some representation of the world.
+Our approach allows us to sidestep serious open problems involved in the representation of a complete set of knowledge about the world.
+
+\begin{table}[h]\small
+\centering\resizebox{4.5in}{!}{
   \setlength{\tabcolsep}{15pt}
@@
     \bottomrule
-  \end{tabular}
+  \end{tabular}}
   \label{b-table}
@@
 
-Our models classify into the seven relations from \cite{maccartney2009extended}, shown in the table below, which define possible relationships between pairs of terms or sentences of natural language in terms of their set-theoretic denotations.
-If any model is to learn the behavior of a relational logic like the one
-presented here from a finite amount of data, it must learn to deduce new
-relations from already seen relations.
-Our first experiment evaluates the ability of our models to do this over pairs of atomic symbols in a large corpus of artificial data.
-The model is trained on examples like \{$a \natfor b$, $b \natneg c$\}, and tested on examples that follow from them, like \{$a \natalt c$\}.
+\subsection*{Reasoning with atomic symbols}
+
+If any model is to learn the behavior of a relational logic like the one presented here from a finite amount of data, it must learn to deduce new relations from already seen relations.
+Our first experiment evaluates the ability of our models to do this over pairs of atomic symbols.
+The model is trained on a randomly generated set of consistent statements like \{$a \natfor b$, $b \natneg c$, $c \nateq d$\}, and tested on novel examples that follow from the statements seen in training, like \{$b \natneg d$\}.
+A tuned RNTN reached greater than 99\% test accuracy, but no plain RNN surpassed 89\%.
 
@@
 
-Our second experiment introduces compositionality to our examples, training on short statements of propositional logic, like $\plneg a\natrev(a \pland b)$.
-We train our models on only pairs of statements with up to four symbols, but observe that the RNN performs reasonably both on those and on much longer test pairs.
+Our second experiment introduces compositional structure to our examples, replacing the atomic statements above with pairs of short statements of propositional logic, like $\plneg a\natalt((a~(\pland~ b))~(\pland~c))$.
+We train our models on only pairs of statements with up to four symbols, but observe that the RNTN performs reasonably both on pairs of that length and on much longer test pairs, as shown below.
 
@@
 We defined several different types of train--test split for this experiment.
-A tuned RNTN model performed either well ($>85\%$ accuracy) or perfectly on all of them, while a tuned RNN did not break 80\% in any setting.
+RNTNs performed either well ($>85\%$ accuracy) or perfectly on all of them, while plain RNNs did not break 80\% in any setting.
 

### 2014-10-08 writing/AAAI SSS '15/intro.tex
Create AAAI paper directory

--- 2014-08-25 writing/BayLearn abstract/paper.tex
@@
-\documentclass{article} % For LaTeX2e
-\usepackage{nips14submit_e,times}
-\usepackage{hyperref}
-\usepackage{url}
-\usepackage[leqno, fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{qtree}
-\usepackage[numbers]{natbib}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{colortbl}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{xcolor}
-
-\definecolor{mylinkcolor}{rgb}{0,0,0} % black
-\hypersetup{colorlinks, linkcolor=mylinkcolor, urlcolor=mylinkcolor, citecolor=mylinkcolor}
-
-\newcommand{\nateq}{\equiv}
-\newcommand{\natind}{\mathbin{\#}}
-%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
-\newcommand{\natneg}{\mathbin{^{\wedge}}}
-\newcommand{\natfor}{\sqsubset}
-\newcommand{\natrev}{\sqsupset}
-\newcommand{\natalt}{\mathbin{|}}
-\newcommand{\natcov}{\mathbin{\smallsmile}}
-
-\newcommand{\plneg}{\mathop{\textit{not}}}
-\newcommand{\pland}{\mathbin{\textit{and}}}
-\newcommand{\plor}{\mathbin{\textit{or}}}
-
-% Strikeout
-\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
-\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}
-
-\newcommand{\True}{\texttt{T}}
-\newcommand{\False}{\texttt{F}}
-\usepackage{stmaryrd}
-\newcommand{\sem}[1]{\ensuremath{\llbracket#1\rrbracket}}
-
-\renewcommand{\bibsection}{\subsubsection*{References}}
-
-\usepackage{gb4e}
-
-\def\ii#1{\textit{#1}}
-
-\newcommand{\mynote}[1]{{\color{red}\framebox{\begin{tabular}{p{0.9\textwidth}}\footnotesize#1 \end{tabular}}}}
-
-\title{Recursive Neural Networks for Learning Logical Semantics}
-
-\author{
-Samuel R.\ Bowman$^{\ast\dag}$ \\
-\texttt{sbowman@stanford.edu} \\[2ex]
-$^{\ast}$Stanford Linguistics \\
-\And
-Christopher Potts$^{\ast}$\\
-\texttt{cgpotts@stanford.edu} \\[2ex]
-$^{\dag}$Stanford NLP Group
-\And
-Christopher D.\ Manning$^{\ast\dag\ddag}$\\
-\texttt{manning@stanford.edu}\\[2ex]
-$^{\ddag}$Stanford Computer Science
-}
-
-% \author{
-% Samuel R.\ Bowman \\
-% NLP Group, Dept.\ of Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{sbowman@stanford.edu}
-%  \And
-%  Christopher Potts \\
-% Dept.\ of Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{cgpotts@stanford.edu}
-%  \And
-% Christopher D.\ Manning \\
-% NLP Group,  Depts.\ of Computer Science and Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{manning@stanford.edu}
-% }
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-
-\nipsfinalcopy % Uncomment for camera-ready version
-
-\begin{document}
-\maketitle
-
-  Supervised recursive neural network models (RNNs) for sentence
-  meaning have been successful in an array of sophisticated language
-  tasks, but it remains an open question whether they can learn
-  compositional semantic grammars that support logical deduction.  We
-  address this question directly by for the first time evaluating
-  whether each of two classes of neural model --- plain RNNs and
-  recursive neural tensor networks (RNTNs) --- can correctly learn
-  relationships such as entailment and contradiction between pairs of
-  sentences, where we have generated controlled data sets of sentences
-  from a logical grammar.  Our first experiment evaluates whether
-  these models can learn the basic algebra of logical relations
-  involved. Our second and third experiments extend this evaluation to
-  complex recursive structures and sentences involving quantification.
-  We find that the plain RNN achieves only mixed results on all three
-  experiments, whereas the stronger RNTN model generalizes well in
-  every setting and appears capable of learning suitable
-  representations for natural language logical inference.
-
-  \section{Introduction}\label{sec:intro}
+\subsection*{Introduction}\label{sec:intro}
+  % TODO: Trim and revise
+  % Less focus on RNTNs, more on embedding spaces
 
@@
 % \label{sec2}
-
-\section{Recursive neural network models} \label{methods}
-
-We study neural models that adhere to the linguistic \ii{principle of
- compositionality}, which says that the meanings for complex
-expressions are derived from the meanings of their constituent parts
-via specific composition functions \cite{Partee84,Janssen97}. In our
-distributed setting, word meanings are vectors of length $n$. The
-composition function maps pairs of them to single vectors of length $n$, 
-which can then be merged again to represent more complex
-phrases. Once the entire sentence-level representation has been
-derived, it serves as a fixed-length input for some subsequent function.
-
-We use the model architecture proposed in \cite{bowman2013can} and
-depicted in Figure~\ref{sample-figure}. The two phrases being compared
-are built up separately on each side of the tree, using the same
-composition function, until they have each been reduced to single
-vectors. The resulting vectors are fed into a separate comparison
-layer that is meant to generate a feature vector capturing the
-relation between the two phrases. The output of this layer is then
-given to a softmax classifier, which in turn produces a hypothesized
-distribution over the seven relations represented in Table~\ref{b-table}.
-
-For a composition layer, we evaluate models with both the plain neural
-network layer function \eqref{rnn} and the RNTN layer function
-\eqref{rntn} proposed in \citet{chen2013learning}. A sigmoid
-nonlinearity (element-wise $\tanh$) is applied to the output of either
-layer function, following \cite{socher2013acl1}.
-%
-\begin{gather} \label{rnn}
-\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
-\label{rntn}
-\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
-\end{gather} 
-%
-Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
-representations for the left and right children of the node, and
-$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
-them by an $n \times 2n$ matrix of learned weights, and applies the
-element-wise non-linearity to the resulting vector. The RNTN has the
-same basic structure, but with the addition of a learned third-order
-tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
-multiplicative interactions between the child vectors. Both models
-include a bias vector~$\vec{b}$.
-
-The comparison layer uses the same kind of function template as the
-composition layers (either an NN layer or an NTN layer) with
-independently learned parameters and a separate nonlinearity function.
-Rather than use a $\tanh$ nonlinearity here, we follow \cite{bowman2013can} 
-in using a rectified linear function. In
-particular, we use the leaky rectified linear function
-\cite{maasrectifier}: $f(\vec{x})=\max(\vec{x}, 0) +
-0.01\min(\vec{x}, 0)$, applied element-wise.
-
-To run the model forward and label a pair of phrases, the structure of
-the lower layers of the network is assembled so as to mirror the tree
-structures provided for each phrase. The word vectors are then looked
-up from the vocabulary matrix $V$ (one of the model parameters), and
-the composition and comparison functions are used to pass information
-up the tree and into the classifier at the top. For an objective
-function, we use the negative log of the probability assigned to the
-correct label.
-
-% Removed 'minibatch' -> For two of the experiments, we use minibatches
-% of size 1, which doesn't really count.
-We train the model using stochastic gradient descent (SGD)
-with learning rates computed using AdaGrad \cite{duchi2011adaptive}.
-The parameters (including the vocabulary) are initialized randomly 
-using a uniform distribution over $[-0.1, 0.1]$.
-Because the seven classes are not balanced in general, we report performance
-using both accuracy and macroaveraged F1, since the latter emphasizes
- performance on infrequent classes. We compute macroaveraged F1 
-as the harmonic mean of mean precision and mean recall, both computed
-for all classes for which there is test data, setting precision to 0 
-where it is not defined. Source code and generated data will be released
-after the review period.
-
-%\begin{figure}[tp]
-%  \centering
-% % \includegraphics[scale=0.35]{model.eps}
-%  \input{figure1}
-%  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
-%    The same structure is used for both the RNN and RNTN layer functions.} 
-%  \label{sample-figure}
-%\end{figure}
-
-%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?
-
-\section{Reasoning about semantic relations}\label{sec:join}
-
-If a model is to learn the behavior of a relational logic like the one
-presented here from a finite amount of data, it must learn to deduce new
-relations from seen relations in a sound manner. The simplest such
-deductions involve atomic statements using the relations in
-Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
-\natrev c$, one can conclude that $a \natrev c$, by basic
-set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
-full set of sound inferences of this form is summarized in
-Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
-of relations for which no valid inference can be drawn in our logic.
-
-% about the relations themselves that do not depend on the
-% internal structure of the things being compared. For example, given
-% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
-% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven relations support more
-% than just transitivity: MacCartney and Manning's
-% \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of relations of the form $a R
-% b$ and $b R' c$, including several less intuitive ones such as that if
-% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-%
-%\begin{table}[htp]
-%  \centering  
-%  \setlength{\arraycolsep}{8pt}
-%  \renewcommand{\arraystretch}{1.1}
-%  \newcommand{\UNK}{\cdot}  
-%  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
-%    %\hline
-%    \multicolumn{1}{c}{}
-%             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
-%    \cline{2-8}
-%    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
-%    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
-%    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
-%    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
-%    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
-%    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
-%    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
-%    \cline{2-8}
-%  \end{array}$
-%  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.
-These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
-We assess our models' ability to reproduce such inferential paths.}
-%  \label{tab:jointable}
-%\end{table}
-
-\paragraph{Experiments}
-To test our models' ability to learn these inferential patterns, we
-create small boolean structures for our logic in which terms denote
-sets of entities from a small domain.  Figure~\ref{lattice-figure}
-depicts a structure of this form. The lattice gives the full model,
-for which all the statements on the right are valid. We divide these
-statements evenly into training and test sets, and remove from the
-test set those statements which cannot be proven from the training
-statements.
-% using the logic depicted in Figure~\ref{lattice-figure}.
-
-% We test the model's ability to learn this behavior by creating
-% artificial data sets of terms which represent sets of numbers. Since
-% MacCartney and Manning's set of relations hold between sets as well as
-% between sentences, we can use the underlying set structure to generate
-% the all of the relations that hold between any pair of these terms, as
-% in Figure \ref{lattice-figure}. We train the model defined above on a
-% subset of these relations, but rather then presenting the model with a
-% pair of tree-structured sentences as inputs, simply present it with
-% two single terms, each of which corresponds to a single vector in the
-% (randomly initialized) vocabulary matrix $V$, ensuring that the model
-% has no information about the terms being compared except the relations
-% between them.
-
-In our experiments, we create 80 randomly generated sets drawing from
-a domain of seven elements. This yields a data set consisting of
-6400 statements about pairs of formulae. 3200 of these pairs are
-chosen as a test set, and that test set is further reduced to the 2960
-examples that can be provably derived from the training data.
-
-We trained versions of both the RNN model and the RNTN model on these
-data sets. In both cases, the models were implemented exactly as
-described in Section~\ref{methods}, but since the items being compared
-are single terms rather than full tree structures, the composition
-layer was not involved, and the two models differed only in which
-layer function was used for the comparison layer. We simply present
-the models with two single terms, each of which corresponds to a
-single vector in the (randomly initialized) vocabulary matrix $V$,
-thereby ensuring that the model has no information about the terms
-being compared except the relations between them.
-
-\paragraph{Results} 
-We found that the RNTN model worked best with 11-dimensional vector
-representations for the 80 sets and a 90-dimensional feature vector
-for the classifier, though the performance of the model was not highly
-sensitive to either dimensionality parameter in any of the experiments
-discussed here. This model was able to correctly label 99.4\% (94.3
-macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
-best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
-respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
-the RNN.
-
-\begin{figure}[tp]
-  \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
-    \setlength{\unitlength}{1cm}
-    \begin{picture}(5,5.5)
-      \labelednode{2.50}{5}{}{1,2,3}
-
-      \put(0.75,4){\line(3,1){1.5}}
-      \put(2.5,4){\line(0,1){0.5}}
-      \put(4.25,4){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{3.5}{a,b}{1,2}
-      \labelednode{2.50}{3.5}{c}{1,3}
-      \labelednode{4.25}{3.5}{d}{2,3}
-
-      \put(0.75,2.5){\line(0,1){0.5}}
-      \put(0.75,2.5){\line(3,1){1.5}}
-
-      \put(2.5,2.5){\line(-3,1){1.5}}
-      \put(2.5,2.5){\line(3,1){1.5}}
-
-      \put(4.25,2.5){\line(0,1){0.5}}
-      \put(4.25,2.5){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{2}{e,f}{1}
-      \labelednode{2.50}{2}{}{2}
-      \labelednode{4.25}{2}{g,h}{3}
-
-      \put(2.5,1){\line(-3,1){1.5}}
-      \put(2.5,1){\line(0,1){0.5}}
-      \put(2.5,1){\line(3,1){1.5}}
-
-      \labelednode{2.5}{0.5}{}{}
-    \end{picture}
-    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
-    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}[t]{0.43\textwidth}
-    \centering
-    \setlength{\tabcolsep}{12pt}
-    \begin{tabular}[b]{c  c}
-      \toprule
-      Train & Test \\
-      \midrule
-                    & $b \nateq b$ \\
-      $b \natcov c$ &               \\
-                    & $b \natcov d$ \\
-                    & \strikeout{$b \natrev e$} \\
-      $c \natcov d$ &               \\
-      $c \natrev e$ &               \\
-                    & \strikeout{$c \nateq f$} \\
-      $c \natrev g$ &               \\ 
-                    & $e \natfor b$ \\
-      $e \natfor c$ &               \\[-1ex]
-      $\vdots$      & $\vdots$ \\
-      \bottomrule
-    \end{tabular}
-
-    \caption{A train/test split of the atomic statements about the
-      model.  Test statements not provable from the training data are
-      crossed out.}
-  \end{subfigure}  
-  \caption{Small example structure and data for learning relation composition.}
-  \label{lattice-figure}
-\end{figure} 
-% Note: None of these test examples is derivable from the shown training data, 
-% but we suggest that there is additional training data, so we can cross lines
-% out willy-nilly without fear.
-
-% RNTN log: tue-j-11-6x80-hip.txt
-% Train PER: 0.0021875
-
-% RNN log: tue-j-11-6x80-r-75.txt
-% Train PER: 0.047188
-
-% and create a dataset consisting of the relations between every pair of
-% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
-% test dataset, and that test dataset was further split into the 2960
-% examples that can be provably derived from the test data using
-% MacCartney and Manning's join table (or by the symmetry of the
-% relations in about half of the cases) and the 240 that
-% cannot.
-
-% We tested a version of both the RNN model and the RNTN model on these
-% data. In both cases, the models were implemented exactly as described
-% in Section~\ref{methods}, but since the items being compared are
-% single terms rather than full tree structures, the composition layer
-% was not used, and the two models differed only in which layer function
-% was used for the comparison layer. We found that the RNTN model worked
-% best with 11 dimensional vector representations for the 80 sets and a
-% 90 dimensional feature vector for the classifier. This model was able
-% to correctly label 99.3\% of the derivable test examples, and 99.1\%
-% of the remaining examples. The simpler RNN model worked best with 11
-% and 75 dimensions, respectively, but was able to achieve accuracies of
-% only 90.0\% and 87.\%, respectively.
-
-These results are fairly straightforward to interpret. The RNTN model
-was able to accurately encode the relations between the terms in the
-geometric relations between their vectors, and was able to then use
-that information to recover relations that were not overtly included
-in the training data. In contrast, the RNN model was able to achieve
-this behavior only incompletely. It is possible but not likely that it
-could be made to find a good solution with further optimization on
-different learning algorithms, or that it would do better on a larger
-universe of sets, for which there would be a larger set of training
-data to learn from, but the RNTN is readily able to achieve these
-effects in the setting discussed here.
-
-\section{Recursive structure in propositional logic}\label{sec:recursion}
-
-Recursive structure is a prominent feature of natural languages and an
-important part of their expressivity, as it allows people to use and
-interpret complex structures they have never encountered before.
-Theoretical accounts of natural language syntax and semantics
-therefore rely heavily on recursive structures, and we expect that
-accurate computational models will have to come to grips with them as
-well. The present section pursues this question for our two classes of
-RNN. In our evaluations, we exploit the fact that our logical language
-is infinite by testing on strings that are longer and more complex
-than any seen in training.
-
-% Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice
-%   said hello}, and \ii{Carl thinks that Bob said that Alice said
-%   hello}. Overt recursion of this kind is easy to find, and
-% theoretical accounts of natural language syntax and semantics rely
-% heavily on recursive structures. In order for a model to be able to
-% accurately learn natural language meanings, then, we expect that it
-% would need to be able to learn to represent the meanings of function
-% words in a such a way that they are able to behave correctly when
-% taking their own outputs as input. In evaluating the model, we take
-% advantage of the fact that recursive structures of this kind define
-% potentially infinite languages by testing the model on strings that
-% are longer and more complex than any seen in testing.
-
-\paragraph{Experiments}
-As in Section~\ref{sec:join}, we test this phenomenon within the
-framework of natural logic, but we now replace the unanalyzed symbols
-from that experiment with complex formulae. These formulae
-represent a truth-functionally complete classical propositional logic:
-each atomic variable denotes either $\True$ or $\False$, and the only
-operators are truth-functional ones.  Table~\ref{tab:pl} defines this
-logic, and Table~\ref{tab:plexs} gives some examples of relational
-statements that we can make in these terms. To compute these relations
-between statements, we exhaustively enumerate the sets of assignments
-of truth values to propositional variables that would satisfy each of
-the statements, and then we convert the set-theoretic relation between
-those assignments into one of the seven relations in
-Table~\ref{b-table}. As a result, each relational statement represents
-a valid theorem of the propositional logic.
-%
-%\begin{table}[tp]
-%  \centering
-%  \begin{subtable}[t]{0.45\textwidth}
-%    \centering
-%    \begin{tabular}[t]{l l}
-%      \toprule
-%      Formula     & Interpretation \\
-%      \midrule
-%      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
-%      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
-%      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
-%      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
-%      \bottomrule
-%    \end{tabular}    
-%    \caption{Well-formed formulae. $\varphi$ and $\psi$
-%      range over all well-formed formulae, and $\sem{\cdot}$ is
-%      the interpretation function mapping formulae into $\{\True,
-%      \False\}$.}\label{tab:pl}
-%  \end{subtable}
-%  \quad
-%  \begin{subtable}[t]{0.45\textwidth}
-%    \centering
-%    \begin{tabular}[t]{r c l}
-%      \toprule
-%      $\plneg a$        & $\natneg$ & $a$ \\
-%      $\plneg \plneg a$ & $\nateq$  & $a$ \\
-%      $a$               & $\natfor$ & $(a \plor b)$ \\
-%      $a$               & $\natrev$ & $(a \pland b)$ \\
-%      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
-%      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
-%      \bottomrule
-%    \end{tabular}
-%    \caption{Examples of statements about relations between
-%      well-formed formulae, defined in terms of sets of satisfying
-%      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
-%  \end{subtable}
-%  \caption{Propositional logic with natural logic relations.}  
-%  \label{prop-figure}
-%\end{table}
-
-% NOTE: There was some confusion at CSLI and the 224U lecture about
-% whether this was theorem proving or some kind of model checking.
-% Hopefully this is explicit enough.
-Socher et al.~\cite{socher2012semantic} show that a matrix-vector RNN
-model somewhat similar to our RNTN can learn a logic, but the logic
-discussed here is a much better approximation of the kind of
-expressivity needed for natural language. The logic learned in that
-experiment is boolean, wherein the atomic symbols are simply the
-values $0$ and $1$, rather than variables over those values. While
-learning the operators of that logic is not trivial, the outputs of
-each operator can be represented accurately by a single bit. The
-statements of propositional logic learned here describe conditions on
-unknown truth values of propositions. As opposed to the two-way
-contrasts seen in \cite{socher2012semantic}, this logic distinguishes
-between $2^{6} = 64$ possible assignments of truth values, and
-expressions of this logic define arbitrary conditions on these
-possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
-possible statements that the intermediate vector representations need
-to be able to distinguish if we are to learn this logical system in its
-full generality.
-
-For our experiments, we randomly generate a large set of  unique pairs 
-of formulae and compute the relation that holds for each pair.
-We discard pairs in which either statement is a tautology or
-contradiction (a statement that is true of either all or no possible
-assignments), for which none of the seven relation labels in
-Table~\ref{b-table} can hold. The resulting set of formula pairs is
-then partitioned into bins based on the number of logical operators in
-the longer of the two formulae. We then randomly sample 15\% of each
-bin for a held-out test set.
-
-If we do not implement any constraint that the two statements being
-compared are similar in any way, then the generated data are dominated
-by statements in which the two formulae refer to largely separate
-subsets of the six variables, which means that the $\natind$ relation
-is almost always correct.  In an effort to balance the distribution of
-relation labels without departing from the basic task of modeling
-propositional logic, we disallow individual pairs of statements from
-referring to more than four of the six propositional variables.
-
-\paragraph{Results} 
-We trained both the RNN and RNTN models on data with up to 4
-connectives (65k pairs) and tested it on examples of up to 12
-connectives (44k pairs). We initialized the model parameters randomly,
-including the vector representations of the six variables. The results
-are shown in Figure~\ref{prop-results}. In tuning, we found that the
-RNN model was approximately optimal with 45-dimensional vector
-representations, and the RNTN model was approximately optimal with 25
-dimensions. We fixed the size of the feature vector for the classifier
-at 75 dimensions. We found that the RNTN model was able to perform
-almost perfectly on unseen small test examples, with accuracy above
-98.7\% on formulae below length four (macro-F1 was above 95.5).
-Starting at size four, performance gradually falls with increasing
-size.  The RNN model did not perform well, reaching only 90.5\% (81.4
-micro-F1) accuracy on the smallest test examples, and declining from
-there to near-baseline performance on formulae of length 12. Training
-accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.
-
-% thu-and-or-deep-25-32b.txt iter 136
-% mon-and-or-deep-45r-32.txt iter 480
-
-\begin{figure}[tp]
-  \centering
-  \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
-  \caption{Model performance on propositional logic, by expression size. 
-    The dashed line indicates that only expressions of size four or less appeared in the training data.}  
-  \label{prop-results}
-\end{figure}
-
-The performance of the RNTN model on small unseen test examples
-indicates that it learned a correct approximation of the underlying
-logic. It appears that this approximation is accurate enough to yield
-correct answers when the composition layer is only applied a small
-number of times, but that the error in the approximation grows with
-increasing depth (and with the increasing complexity of the
-expressions), resulting in gradually dropping performance. In
-contrast, the RNN model did not appear to be able to learn a correct
-approximation of the logic for statements of any length, but it
-appears that the faulty logic that the RNN did learn decayed similarly
-with expression size.
-This is not necessarily a significant flaw in the model, since it remains possible that  different training regimes would allow it to learn accurate representations of the larger expressions.
-
-In an effort to better understand why the RNTN outperformed the plain
-RNN, we evaluated both models on pairs of long formulae involving
-binary connectives, to assess the extent to which they achieve
-substantially different representations for semantically distinct
-formulae and substantially similar representations for synonymous
-formulae. We found that both models are able to identify synonymous
-pairs like $(a \pland a)$ and $(a \pland (a \pland (a \pland a)))$,
-with both \ii{and} and \ii{or}, but only the RNTN is able to learn
-substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
- and $(a \pland (a \pland b))$ when the difference between the two is embedded under multiple operators. 
- Figure~\ref{prop-falloff} summarizes this result.
-
-% we discovered that this model looses information about in longer
-% statements in a particularly problematic way. It appears that that
-% model is unable to distinguish between two sentences when the only
-% difference between those sentences is embedded within a very deep
-% structure. We evaluated both models on sentences that differed in only
-% one term, but for which that one term was embedded under a large
-% number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and
-%   b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and
-%   b)))}. We then measured the Euclidean distance between the vector
-% representations of the two sentences in each pair. Our findings are
-% shown in Figure~\ref{prop-falloff}, and show that while the RNTN can
-% distinguish the two sentences well at every size that we test, the RNN
-% fails after a depth of approximately six.
-
-%\begin{figure}[htp]
-%  \centering
-%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-%  \caption{Semantically distinct formulae should have different
-%    representations. As measured by Euclidean distance, only the RNTN
-%    achieves this for formulae containing more than a small number of
-%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
-%    failing to capture the meaning contrasts.}
-%  \label{prop-falloff}
-%\end{figure}
-
-\section{Reasoning with natural language quantifiers and negation}\label{sec:quantifiers}
-
-We have seen that the RNTN can learn an approximation of propositional
-logic.  However, natural languages can express functional meanings of
-considerably greater complexity than this.  As a first step towards
-investigating whether our models can capture this complexity, we now
-attempt to directly measure the degree to which they are able to
-develop suitable representations for the semantics of natural language
-quantifiers like \ii{some} and \ii{all}. Quantification is far from
-the only place in natural language where complex functional meanings
-are found, but it is a natural starting point, since it can be tested
-in sentences whose structures are otherwise quite simple, and since it
-has formed a standard case study in prior formal work on natural
-language inference \cite{Icard:Moss:2013:LILT}.
-
-% \subsection{Data}
-
-This experiment replicates similar work described in
-\cite{bowman2013can}, which found that RNTNs can learn to reason well
-with quantifier meanings given sufficient training data. This paper
-replaces the partially manually annotated data in that paper with data
-that is generated directly using the logical system that we hope to
-learn, yielding results that we believe to be substantially more
-straightforward to interpret.
-
-\paragraph{Experiments}
-Our experimental data consist of pairs of sentences generated from a
-small artificial grammar. Each sentence contains a quantifier, a noun
-which may be negated, and an intransitive verb which may be
-negated. We use the basic quantifiers \ii{some}, \ii{most}, \ii{all},
-\ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
-\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}. We also
-include five nouns, four intransitive verbs, and the negation symbol
-\ii{not}. In order to be able to define relations between sentences
-with differing lexical items, we define the lexical relations between
-each noun--noun pair, each verb--verb pair, and each
-quantifier--quantifier pair. The grammar accepts aligned pairs of
-sentences of this form and calculates the natural logic relationship
-between them.  Some examples of these data are provided in
-Table~\ref{examplesofdata}.  As in previous sections, the goal of
-learning is then to assign these relational labels accurately to
-unseen pairs of sentences.
-
-%nouns = ['warthogs', 'turtles', 'mammals', 'reptiles', 'pets']
-%verbs = ['walk', 'move', 'swim', 'growl']
-%dets = ['all', 'not_all', 'some', 'no', 'most', 'not_most', 'two', 'lt_two', 'three', 'lt_three']
-%adverbs = ['', 'not']
-
-% To assign relation labels to sentence pairs, we built a small
-% task-specific implemenation of MacCartney's logic that can
-% accurately label sentences of this restricted language. The logic is
-% not able to derive all intuitively true relations of this language,
-% and fails to derive a single unique relation for certain types of
-% statement, including De Morgean's laws (e.g. \ii{(all pets) growl
-% $\natneg$ (some pet) (not growl)}), and we simply discard these
-% examples. Exhaustively generating the valid sentences under this
-% grammar and choosing those to which a relation label can be assigned
-% yields 66k sentence pairs. Some examples of these data are provided
-% in Table~\ref{examplesofdata}.
-
-%\begin{table}[htp]
-%  \centering
-%  \begin{tabular}{l c l}
-%    \toprule
-%    (most warthogs) walk         & $\natneg$ & (not-most warthogs) walk\\
-%    (most mammals) move          & $\natind$ &  (not-most (not turtles)) move\\
-%    (most (not pets)) (not swim) & $\natrev$ & (not-most (not pets)) move 
-%    \\[2ex]    
-%    (no turtles) (not growl)     & $\natalt$ & (no turtles) (not swim)\\
-%    (no warthogs) swim           & $\natrev$ & (no warthogs) move\\
-%    (no warthogs) move           & $\natfor$ & (no (not reptiles)) swim\\
-%    \bottomrule
-%  \end{tabular}
-%  \caption{Sample data involving two different quantifier pairs.}
-%  \label{examplesofdata}
-%\end{table}
-
-We evaluate the model using two experimental settings. In the simpler
-setting, \textsc{all split}, we randomly sample 85\% of the data and evaluate on the
-remaining 15\%. In this setting, the model is being asked to learn a
-complete reasoning system for the limited language and logic presented
-in the training data, but it is not being asked to generalize to test
-examples that are substantially different from those it was trained
-on. Crucially, though, to succeed on this task, the model must be able
-to recognize all of the lexical relations between the nouns, verbs,
-and quantifiers and how they interact. For instance, it might see
-\eqref{p1} and \eqref{p2} in training and, from that information,
-determine \eqref{p3}.
-%
-% do not allow a blank line --- adds too much space
-%
-\begin{gather}
-  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
-  \\
-  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
-  \\
-  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
-\end{gather}
-%
-% do not allow a blank line --- adds too much space
-%
-While our primary interest is in discovering the extent to which our
-models can learn to encode the logic given an arbitrary amount of
-data, we are also interested in the degree to which they can infer a
-correct representation for the logic from more constrained training
-data. To this end, we segment the sentence pairs according to which
-quantifiers appear in each pair, and then hold out one such pair for
-testing. We hypothesize that a model that can efficiently learn to
-represent a logic should be able to construct an accurate
-representation of each of the two quantifiers in the pair from the way that it
-interacts with the other nine quantifiers with which it is seen in training. 
-Since running this experiment requires choosing a pair of
-quantifiers to hold out before training, the resource demands of
-training prevent us from testing each of the 55 possible 
-pairs of quantifiers, so we choose only four pairs to test on.  Three
-of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
-\ii{all}/\ii{some}) were chosen because they allow for the most
-different class labels in the test data. The fourth is a
-self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
-handles equality.
-
-\begin{table}[tp]
-  \centering
-  \setlength{\tabcolsep}{10pt}
-  \begin{tabular}{ l rr@{ \ }r r@{ \ }r r@{ \ }r }
-    \toprule
-    Data & \multicolumn{3}{c}{Most frequent class} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
-    \midrule
-    \textsc{all split}	            &~~~~& 35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}
-    \\[1ex]    
-    \textsc{pair two/less-than-two} && 59.8 &(18.4) & 77.2 &(57.4) &	\textbf{100.0} &\textbf{(100.0)} \\
-    \textsc{pair not-all/not-most}  &&    0 &(0)    & 66.0 &(54.3) &	\textbf{93.8}  &\textbf{(91.5)} \\
-    \textsc{pair all/some}	    &&    0& (0)    & 62.7 &(65.9)  &	\textbf{78.4}  &\textbf{(84.7)} \\
-    \textsc{pair no/no}	            && 30.8  &(10.2) & 67.8 & (61.8) &	\textbf{99.9}  &\textbf{(99.7)} \\
-    \bottomrule
-  \end{tabular}
-  \caption{Performance on the quantifier experiments. Results are reported as accuracy scores followed by macroaveraged F1 scores in parentheses.}
-  \label{resultstable}
-\end{table}
-
-\paragraph{Results} 
-The results for these experiments are shown in
-Figure~\ref{resultstable}. We compare the results to a most frequent
-class baseline, which reflects the frequency in the test data of the
-most frequent class in the training data, $\natind$.  After some
-cross-validation, we chose 16- and 20-dimensional representations for
-the RNN and RNTN, respectively, and 75-dimensional feature vectors for
-the classifier. Training
-accuracy was near 80\% for the each of the five RNN experiments, and 
-near 98\% for each of the five RNTN experiments.
-
-%train rnn 0.20302, 0.017195
-
-The RNN performed poorly at this task, even though the sentences used
-in these examples are short enough to avoid the pathology shown in
-Figure~\ref{prop-falloff}.  However, the perfect performance by the
-RNTN on the \textsc{all split} and \textsc{pair two/less-than-two}
-experiments, and its strong performance on the \textsc{pair no/no}
-experiment, suggest that this stronger model is able to handle
-quantifiers correctly given sufficient training data, but the weaker
-results on two of the training settings suggest that it may not be
-able to generalize from this kind of restricted training data in general. However, though the question
-of how much data is necessary to accurately capture quantifier
-behavior in a na\"ive model remains open, the fact that this model is able to perform so well is promising.
-
-\section{General discussion}\label{sec:discussion}
-
-This paper evaluated two recursive models on a series of three increasingly
-challenging interpretive tasks involving natural language inference:
-the core relational algebra of natural logic with entailment and
-exclusion; recursive propositional logic structures; and statements
-involving quantification and negation. The results suggest that RNTNs,
-but not plain RNNs, have the capacity to meet the challenges of these
-tasks with reasonably-sized training sets. These positive results are
-promising for the future of learned representation models in the
-applied modeling of compositional semantics.
-
-Of course, challenges remain. In terms of our experimental data, even
-the RNTN falls short of perfection in our more complex tasks, with
-performance falling off steadily as the depth of recursion grows. It
-remains to be seen whether these deficiencies can be overcome with
-improvements to the model, the optimization procedures, or the
-linguistic representations
-\cite{sochergrounded,kalchbrenner2014convolutional}. In addition,
-there remain subtle questions about how to fairly assess whether these
-models have truly generalized in the way we want them to. There is a
-constant tension between showing the models training data that gives
-them a chance to learn the target logical functions and revealing the
-answer to them in a way that leads to overfitting. The underlying
-logical theories provide only limited guidance on this point.
-%
-%, and the fact that there is a finite universe of possible
-%  expressions makes his an unavoidable issue. 
-%
-%  CP: I don't understand the above. It is false if taken literally;
-%  our PL generates an infinte number of formulae.
-%
-Finally, we have only scratched the surface of the logical complexity
-of natural language; in future experiments, we hope to test sentences
-with embedded quantifiers, multiple interacting quantifiers, relative
-clauses, and other kinds of recursive structure. Nonetheless, the
-rapid progress the field has made with these models in recent years
-provides ample reason to be optimistic that they can be trained to
-meet the challenges of natural language semantics.
-
-% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators.
-This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic.
-The positive results that we have found, however, are extremely promising for the future of learned representation models in the applied modeling of meaning.
-We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators.
-We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.
-
-% There is ample room to build on these results.
-In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses.
-In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.
-
-% \textit{Withheld for review.}\\ \\
-% Do later: Fill in, cite grants, internal reviewers
-
-\bibliographystyle{unsrtnat}
-
-\small % Note: Explicitly allowed in style guide
-\bibliography{MLSemantics}
-
-\end{document}

### 2014-10-08 writing/AAAI SSS '15/join.tex
Create AAAI paper directory

--- 2014-06-06 writing/S'14 paper/join.tex
@@
-\section{Reasoning about semantic relations}\label{sec:join}
+\subsection*{Reasoning about semantic relations}\label{sec:join}
+
+\begin{table}[tp]
+  \centering
+  \setlength{\tabcolsep}{15pt}
+  \renewcommand{\arraystretch}{1.1}
+  \begin{tabular}{l c l l} 
+    \toprule
+    Name & Symbol & Set-theoretic definition & Example \\ 
+    \midrule
+    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
+    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
+    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
+    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
+    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
+    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
+    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
+    \bottomrule
+  \end{tabular}
+  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
+    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
+    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
+    %is insufficient knowledge to choose a label.}
+  \label{b-table}
+\end{table}
 
@@
 These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
-We assess our model's ability to reproduce such inferential paths.}
+We assess our models' ability to reproduce such inferential paths.}
   \label{tab:jointable}
@@
 % using the logic depicted in Figure~\ref{lattice-figure}.
+
+% We test the model's ability to learn this behavior by creating
+% artificial data sets of terms which represent sets of numbers. Since
+% MacCartney and Manning's set of relations hold between sets as well as
+% between sentences, we can use the underlying set structure to generate
+% the all of the relations that hold between any pair of these terms, as
+% in Figure \ref{lattice-figure}. We train the model defined above on a
+% subset of these relations, but rather then presenting the model with a
+% pair of tree-structured sentences as inputs, simply present it with
+% two single terms, each of which corresponds to a single vector in the
+% (randomly initialized) vocabulary matrix $V$, ensuring that the model
+% has no information about the terms being compared except the relations
+% between them.
+
+In our experiments, we create 80 randomly generated sets drawing from
+a domain of seven elements. This yields a data set consisting of
+6400 statements about pairs of formulae. 3200 of these pairs are
+chosen as a test set, and that test set is further reduced to the 2960
+examples that can be provably derived from the training data.
+
+We trained versions of both the RNN model and the RNTN model on these
+data sets. In both cases, the models were implemented exactly as
+described in Section~\ref{methods}, but since the items being compared
+are single terms rather than full tree structures, the composition
+layer was not involved, and the two models differed only in which
+layer function was used for the comparison layer. We simply present
+the models with two single terms, each of which corresponds to a
+single vector in the (randomly initialized) vocabulary matrix $V$,
+thereby ensuring that the model has no information about the terms
+being compared except the relations between them.
+
+\paragraph{Results} 
+We found that the RNTN model worked best with 11-dimensional vector
+representations for the 80 sets and a 90-dimensional feature vector
+for the classifier, though the performance of the model was not highly
+sensitive to either dimensionality parameter in any of the experiments
+discussed here. This model was able to correctly label 99.4\% (94.3
+macroaveraged F1) of the provably derivable test examples, and 95.4\%
+(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
+best with 11 and 75 dimensions, respectively, but was able to achieve
+accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
+respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
+the RNN.
 
@@
 
-% We test the model's ability to learn this behavior by creating
-% artificial data sets of terms which represent sets of numbers. Since
-% MacCartney and Manning's set of relations hold between sets as well as
-% between sentences, we can use the underlying set structure to generate
-% the all of the relations that hold between any pair of these terms, as
-% in Figure \ref{lattice-figure}. We train the model defined above on a
-% subset of these relations, but rather then presenting the model with a
-% pair of tree-structured sentences as inputs, simply present it with
-% two single terms, each of which corresponds to a single vector in the
-% (randomly initialized) vocabulary matrix $V$, ensuring that the model
-% has no information about the terms being compared except the relations
-% between them.
-
-In our experiments, we create 80 randomly generated sets drawing from
-a domain of seven elements. This yields a data set consisting of
-6400 statements about pairs of formulae. 3200 of these pairs are
-chosen as a test set, and that test set is further reduced to the 2960
-examples that can be provably derived from the training data.
-
-We trained versions of both the RNN model and the RNTN model on these
-data sets. In both cases, the models were implemented exactly as
-described in Section~\ref{methods}, but since the items being compared
-are single terms rather than full tree structures, the composition
-layer was not involved, and the two models differed only in which
-layer function was used for the comparison layer. We simply present
-the models with two single terms, each of which corresponds to a
-single vector in the (randomly initialized) vocabulary matrix $V$,
-thereby ensuring that the model has no information about the terms
-being compared except the relations between them.
-
-\paragraph{Results} 
-We found that the RNTN model worked best with 11-dimensional vector
-representations for the 80 sets and a 90-dimensional feature vector
-for the classifier, though the performance of the model was not highly
-sensitive to either dimensionality parameter in any of the experiments
-discussed here. This model was able to correctly label 99.4\% (94.3
-macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
-best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
-respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
-the RNN.
-
 % RNTN log: tue-j-11-6x80-hip.txt

### 2014-10-08 writing/AAAI SSS '15/wordnet.tex
Create AAAI paper directory

\subsection*{Reasoning about lexical relations in WordNet}\label{sec:wordnet}

% TODO: Fill out

% Can a model to do this with something that has the scale and structure of a real vocabulary?

% Wordnet! Subsumption lattice for words!

% Three relations: Hyper, hypo, coordinate
% Explain extraction and rationale for each

% Data creation:
% - Enumerate filtering strategy

% Experiment:
% - Explain train test split

% Reults:
% RNN / RNTN x Data volume x Accuracy vs. MacroF1

### 2014-10-09 writing/AAAI SSS '15/abstract.tex

--- 2014-08-25 writing/BayLearn abstract/paper.tex
@@
-\documentclass{article} % For LaTeX2e
-\usepackage{nips14submit_e,times}
-\usepackage{hyperref}
-\usepackage{url}
-\usepackage[leqno, fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{qtree}
-\usepackage[numbers]{natbib}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{colortbl}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{xcolor}
-
-\definecolor{mylinkcolor}{rgb}{0,0,0} % black
-\hypersetup{colorlinks, linkcolor=mylinkcolor, urlcolor=mylinkcolor, citecolor=mylinkcolor}
-
-\newcommand{\nateq}{\equiv}
-\newcommand{\natind}{\mathbin{\#}}
-%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
-\newcommand{\natneg}{\mathbin{^{\wedge}}}
-\newcommand{\natfor}{\sqsubset}
-\newcommand{\natrev}{\sqsupset}
-\newcommand{\natalt}{\mathbin{|}}
-\newcommand{\natcov}{\mathbin{\smallsmile}}
-
-\newcommand{\plneg}{\mathop{\textit{not}}}
-\newcommand{\pland}{\mathbin{\textit{and}}}
-\newcommand{\plor}{\mathbin{\textit{or}}}
-
-% Strikeout
-\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
-\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}
-
-\newcommand{\True}{\texttt{T}}
-\newcommand{\False}{\texttt{F}}
-\usepackage{stmaryrd}
-\newcommand{\sem}[1]{\ensuremath{\llbracket#1\rrbracket}}
-
-\renewcommand{\bibsection}{\subsubsection*{References}}
-
-\usepackage{gb4e}
-
-\def\ii#1{\textit{#1}}
-
-\newcommand{\mynote}[1]{{\color{red}\framebox{\begin{tabular}{p{0.9\textwidth}}\footnotesize#1 \end{tabular}}}}
-
-\title{Recursive Neural Networks for Learning Logical Semantics}
-
-\author{
-Samuel R.\ Bowman$^{\ast\dag}$ \\
-\texttt{sbowman@stanford.edu} \\[2ex]
-$^{\ast}$Stanford Linguistics \\
-\And
-Christopher Potts$^{\ast}$\\
-\texttt{cgpotts@stanford.edu} \\[2ex]
-$^{\dag}$Stanford NLP Group
-\And
-Christopher D.\ Manning$^{\ast\dag\ddag}$\\
-\texttt{manning@stanford.edu}\\[2ex]
-$^{\ddag}$Stanford Computer Science
-}
-
-% \author{
-% Samuel R.\ Bowman \\
-% NLP Group, Dept.\ of Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{sbowman@stanford.edu}
-%  \And
-%  Christopher Potts \\
-% Dept.\ of Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{cgpotts@stanford.edu}
-%  \And
-% Christopher D.\ Manning \\
-% NLP Group,  Depts.\ of Computer Science and Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{manning@stanford.edu}
-% }
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-
-\nipsfinalcopy % Uncomment for camera-ready version
-
-\begin{document}
-\maketitle
+\begin{abstract}
+  % TODO: Rewrite
 
@@
   representations for natural language logical inference.
+\end{abstract}
 
-  \section{Introduction}\label{sec:intro}
-
-Supervised recursive neural network models (RNNs) for sentence meaning
-have been successful in an array of sophisticated language tasks,
-including sentiment analysis \cite{socher2011semi,socher2013acl1},
-image description \cite{sochergrounded}, and paraphrase detection
-\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
-the ability of these models to learn compositional semantic grammars,
-but it remains an open question whether they can achieve the same
-results as grammars based in logical forms
-\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
-when it comes to the core semantic concepts of quantification,
-entailment, and contradiction needed to identify the relationships
-between sentences like \ii{every animal walks}, \ii{every turtle
-  moves}, and \ii{most reptiles don't move}. To date, experimental
-investigations of these concepts using distributed representations
-have been largely confined to short phrases
-\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
-For robust natural language understanding, it is essential to model
-these phenomena in their full generality on complex linguistic
-structures.
-
-We address this question in the context of \ii{natural language
-  inference} (also known as \ii{recognizing textual entailment};
-\cite{dagan2006pascal}), in which the goal is to determine the core
-inferential relationship between two sentences. Much of the
-theoretical work on this task (and some successful implemented models
-\cite{maccartney2009extended,watanabe2012latent}) involves \ii{natural
-  logics}, which are formal systems that define rules of inference
-between natural language words, phrases, and sentences without the
-need of intermediate representations in an artificial logical
-language. Following \cite{bowman2013can}, we use the natural logic
-developed by \cite{maccartney2009extended} as our formal model. This
-logic defines seven core relations of synonymy, entailment,
-contradiction, and mutual consistency, as summarized in
-Table~\ref{b-table}, and it provides rules of semantic combination for
-projecting these relations from the lexicon up to complex phrases. The
-formal properties and inferential strength of this system are now
-well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
-
-In our experiments, we use this pre-specified logical grammar to
-generate controlled data sets encoding the semantic relationships
-between pairs of expressions and evaluate whether each of two classes
-of neural model --- plain RNNs and recursive neural tensor networks
-(RNTNs, \cite{socher2013acl1}) --- can learn those relationships
-correctly. In our first experiment (Section~\ref{sec:join}), we
-evaluate the ability of these models to learn the core relational
-algebra of natural logic from data. Our second experiment
-(Section~\ref{sec:recursion}) extends this evaluation to relations
-between complex recursive structures like $(a \plor b)$ and
-$\plneg(\plneg a \pland \plneg b)$, and our third experiment
-(Section~\ref{sec:quantifiers}) involves relations between quantified
-statements like \ii{every reptile walks} and \ii{no turtle moves}.
-We find that the plain RNN achieves only mixed results in all three
-experiments, whereas the stronger RNTN model generalizes well in every
-case, suggesting that it has in fact learned, or at least learned to
-simulate, our target logical
-concepts. These experiments differentiate the increased power of RNTNs
-better than previous work and provide the most convincing
-demonstration to date of the ability of neural networks to model
-semantic inferences in complex natural language sentences.
-
-\begin{table}[tp]
-  \centering
-  \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{l c l l} 
-    \toprule
-    Name & Symbol & Set-theoretic definition & Example \\ 
-    \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
-    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
-    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
-    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
-    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
-    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
-    \bottomrule
-  \end{tabular}
-  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
-    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
-    %is insufficient knowledge to choose a label.}
-  \label{b-table}
-\end{table}
-
-% Citations to additional past work to be added.\\...\\...\\...\\...\\...
-
-% Deep learning methods in NLP which learn vector representations for words have seen successful uses in recent years on increasingly sophisticated tasks \cite{collobert2011natural, socher2011semi, socher2013acl1, chen2013learning}.
-Given the still modest performance of semantically rich NLP systems in many domains---question answering and machine translation, for instance---it is worth exploring the degree to which learned vectors can serve as general purpose semantic representations.
-Much of the work to date analyzing vector representations for words (see \cite{baroni2013frege}) has focused on lexical semantic behaviors---like the similarity between words like \ii{Paris} and \ii{France}.
-Good similarity functions are valuable for many NLP tasks, but there are real use cases for which it is necessary to know how words relate to one another or to some extrinsic representation, and to model the ways in which word meanings combine to form phrase, sentence, or document meanings.
-This paper explores the ability of linguistic representations developed using supervised deep learning techniques to support interpretation and reasoning.
-
-% There are two broad family of tasks that could be used to test the ability of a model to develop general purpose meaning representations.
-In an interpretation task, sentences are mapped onto some denotation, such as  \ii{true} or \ii{false} for statements, or a factual answer for questions.
-There has been preliminary work in developing distributed models for interpretation \cite{grefenstette2013towards, rocktaschellow}, but developing a representation of world knowledge that corresponds accurately to the content expressed in language introduces considerable design challenges.
-I approach the problem by way of an inference task instead.
-Inferring the truth of one sentence from another does not require any preexisting knowledge representations, but nonetheless requires a precise representation of sentence meaning.
-I borrow the structure of the task from MacCartney and Manning  \cite{maccartney2009extended}.
-In it, the model is presented with a pair of sentences, and made to label the logical relation between the sentences as equivalence, entailment, or any of five other classes, as here:
-
-%\begin{quote}
-%\begin{enumerate}\small
-%\item Many smartphone users avoid high bills overseas by disabling data service.
-%\item Not everyone uses their smartphones for email when traveling abroad.
-%\end{enumerate} 
-%$\Rightarrow$ Entailment
-%\end{quote}
-
-%In this paper, we test the ability of recursive models to on three simple tasks, each of which is meant to capture a property that is necessary in representing natural language meaning in the setting of inference.
-I begin with an overview of MacCartney and Manning's \cite{maccartney2009extended} framework for inference, and of the recursive neural networks that I study.
-by showing that these models can learn to correctly represent entailment representations between sentences.
-I then show that these models can capture the meanings of recursive structures accurately up to a sufficient depth.
-I finally close with a brief demonstration of the ability of these models to reason over short natural language sentences involving quantifiers.
-
-% \subsection{Natural language inference and natural logic}
-
-% In its simplest form, \ii{natural language inference} (also known as \ii{recognizing textual entailment}, as in \cite{dagan2006pascal}) is the task of determine whether one sentence entails another.
-Much of the theoretical work on this task (and some successful implemented models \cite{maccartney2009natural, watanabe2012latent}) involve \ii{natural logic}, formal systems that define sound rules of inference from one sentence of natural language to another without the need for intermediate representations in some other logic.
-The most powerful model that we are aware of for natural logic is due to MacCartney and Manning \cite{maccartney2009extended} and Icard \cite{icard2012inclusion}, and is centered around the definition of a set of seven logical relations which can hold between sentences, shown in Table \ref{b-table}.
-
-% This approach to natural logic can capture much of the complexity of natural language meaning within a well understood framework, and is also fairly straightforward to implement in a machine learning setting since it can be reduced to a seven-way classification problem on sentence pairs.
-Our goal in this paper is to learn recursive neural network models which are able to mimic key behaviors of this system.
-
-% \begin{table}
-% \begin{center}
-% \begin{tabular}{|c|c|c|c|} \hline
-% name & symbol & set-theoretic definition & example \\ \hline \hline
-% entailment & $x \sqsubset y$ & $x \subset y$ & \ii{crow, bird}  \\ \hline
-% reverse entailment & $x \sqsupset y$ & $x \supset y$ & \ii{Asian, Thai}  \\ \hline
-% equivalence & $x \equiv y$ & $x = y$ & \ii{couch, sofa} \\ \hline
-% alteration & $x$ $|$ $y$ & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{cat, dog} \\ \hline
-% negation & $x \natneg y$ & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{able, unable} \\ \hline
-% cover & $x \smallsmile y$ & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-ape} \\ \hline
-% independence & $x$ \# $y$ & (else) & \ii{hungry, hippo}\\ \hline
-% \end{tabular}
-% \caption{The entailment relations in  $\mathfrak{B}$.
-$\mathcal{D}$ is the universe of possible objects of the same type as those being compared, and the relation \# applies whenever none of the other six do, including when there is insufficient knowledge to choose a label.}
-% \label{b-table}
-% \end{center}
-% \end{table}
-
-% The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes.
-It should be noted that this is not a balanced classification problem.
-If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them.
-% TODO: How much should we say about this?
-
-% The natural logic engine at the core of MacCartney and Manning's system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures.
-These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
-%for dozens of different cases of insertion, deletion and substitution of different types of lexical item.
-For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences.
-Here, the substitution of \ii{no} for \ii{some}  yields $\natneg$ and the substitution of \ii{dogs} for \ii{animals} yields $\sqsupset$.
-It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
-%\begin{quote}
-
-%1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
-%2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
-%3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}
-
-%\end{quote}
-
-% Work to date on inference in neural network models is quite limited.
-% \citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
-% In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations.
-They propose a deterministic inference engine (similar to MacCartney's) which is augmented by a probabilistic component that evaluates individual lexical substitution steps in the derivation using vector representations, though again these representations are not learned, and no evaluations of this system have been published to date.
-% \label{sec2}
-
-\section{Recursive neural network models} \label{methods}
-
-We study neural models that adhere to the linguistic \ii{principle of
- compositionality}, which says that the meanings for complex
-expressions are derived from the meanings of their constituent parts
-via specific composition functions \cite{Partee84,Janssen97}. In our
-distributed setting, word meanings are vectors of length $n$. The
-composition function maps pairs of them to single vectors of length $n$, 
-which can then be merged again to represent more complex
-phrases. Once the entire sentence-level representation has been
-derived, it serves as a fixed-length input for some subsequent function.
-
-We use the model architecture proposed in \cite{bowman2013can} and
-depicted in Figure~\ref{sample-figure}. The two phrases being compared
-are built up separately on each side of the tree, using the same
-composition function, until they have each been reduced to single
-vectors. The resulting vectors are fed into a separate comparison
-layer that is meant to generate a feature vector capturing the
-relation between the two phrases. The output of this layer is then
-given to a softmax classifier, which in turn produces a hypothesized
-distribution over the seven relations represented in Table~\ref{b-table}.
-
-For a composition layer, we evaluate models with both the plain neural
-network layer function \eqref{rnn} and the RNTN layer function
-\eqref{rntn} proposed in \citet{chen2013learning}. A sigmoid
-nonlinearity (element-wise $\tanh$) is applied to the output of either
-layer function, following \cite{socher2013acl1}.
-%
-\begin{gather} \label{rnn}
-\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
-\label{rntn}
-\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
-\end{gather} 
-%
-Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
-representations for the left and right children of the node, and
-$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
-them by an $n \times 2n$ matrix of learned weights, and applies the
-element-wise non-linearity to the resulting vector. The RNTN has the
-same basic structure, but with the addition of a learned third-order
-tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
-multiplicative interactions between the child vectors. Both models
-include a bias vector~$\vec{b}$.
-
-The comparison layer uses the same kind of function template as the
-composition layers (either an NN layer or an NTN layer) with
-independently learned parameters and a separate nonlinearity function.
-Rather than use a $\tanh$ nonlinearity here, we follow \cite{bowman2013can} 
-in using a rectified linear function. In
-particular, we use the leaky rectified linear function
-\cite{maasrectifier}: $f(\vec{x})=\max(\vec{x}, 0) +
-0.01\min(\vec{x}, 0)$, applied element-wise.
-
-To run the model forward and label a pair of phrases, the structure of
-the lower layers of the network is assembled so as to mirror the tree
-structures provided for each phrase. The word vectors are then looked
-up from the vocabulary matrix $V$ (one of the model parameters), and
-the composition and comparison functions are used to pass information
-up the tree and into the classifier at the top. For an objective
-function, we use the negative log of the probability assigned to the
-correct label.
-
-% Removed 'minibatch' -> For two of the experiments, we use minibatches
-% of size 1, which doesn't really count.
-We train the model using stochastic gradient descent (SGD)
-with learning rates computed using AdaGrad \cite{duchi2011adaptive}.
-The parameters (including the vocabulary) are initialized randomly 
-using a uniform distribution over $[-0.1, 0.1]$.
-Because the seven classes are not balanced in general, we report performance
-using both accuracy and macroaveraged F1, since the latter emphasizes
- performance on infrequent classes. We compute macroaveraged F1 
-as the harmonic mean of mean precision and mean recall, both computed
-for all classes for which there is test data, setting precision to 0 
-where it is not defined. Source code and generated data will be released
-after the review period.
-
-%\begin{figure}[tp]
-%  \centering
-% % \includegraphics[scale=0.35]{model.eps}
-%  \input{figure1}
-%  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
-%    The same structure is used for both the RNN and RNTN layer functions.} 
-%  \label{sample-figure}
-%\end{figure}
-
-%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?
-
-\section{Reasoning about semantic relations}\label{sec:join}
-
-If a model is to learn the behavior of a relational logic like the one
-presented here from a finite amount of data, it must learn to deduce new
-relations from seen relations in a sound manner. The simplest such
-deductions involve atomic statements using the relations in
-Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
-\natrev c$, one can conclude that $a \natrev c$, by basic
-set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
-full set of sound inferences of this form is summarized in
-Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
-of relations for which no valid inference can be drawn in our logic.
-
-% about the relations themselves that do not depend on the
-% internal structure of the things being compared. For example, given
-% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
-% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven relations support more
-% than just transitivity: MacCartney and Manning's
-% \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of relations of the form $a R
-% b$ and $b R' c$, including several less intuitive ones such as that if
-% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-%
-%\begin{table}[htp]
-%  \centering  
-%  \setlength{\arraycolsep}{8pt}
-%  \renewcommand{\arraystretch}{1.1}
-%  \newcommand{\UNK}{\cdot}  
-%  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
-%    %\hline
-%    \multicolumn{1}{c}{}
-%             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
-%    \cline{2-8}
-%    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
-%    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
-%    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
-%    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
-%    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
-%    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
-%    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
-%    \cline{2-8}
-%  \end{array}$
-%  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.
-These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
-We assess our models' ability to reproduce such inferential paths.}
-%  \label{tab:jointable}
-%\end{table}
-
-\paragraph{Experiments}
-To test our models' ability to learn these inferential patterns, we
-create small boolean structures for our logic in which terms denote
-sets of entities from a small domain.  Figure~\ref{lattice-figure}
-depicts a structure of this form. The lattice gives the full model,
-for which all the statements on the right are valid. We divide these
-statements evenly into training and test sets, and remove from the
-test set those statements which cannot be proven from the training
-statements.
-% using the logic depicted in Figure~\ref{lattice-figure}.
-
-% We test the model's ability to learn this behavior by creating
-% artificial data sets of terms which represent sets of numbers. Since
-% MacCartney and Manning's set of relations hold between sets as well as
-% between sentences, we can use the underlying set structure to generate
-% the all of the relations that hold between any pair of these terms, as
-% in Figure \ref{lattice-figure}. We train the model defined above on a
-% subset of these relations, but rather then presenting the model with a
-% pair of tree-structured sentences as inputs, simply present it with
-% two single terms, each of which corresponds to a single vector in the
-% (randomly initialized) vocabulary matrix $V$, ensuring that the model
-% has no information about the terms being compared except the relations
-% between them.
-
-In our experiments, we create 80 randomly generated sets drawing from
-a domain of seven elements. This yields a data set consisting of
-6400 statements about pairs of formulae. 3200 of these pairs are
-chosen as a test set, and that test set is further reduced to the 2960
-examples that can be provably derived from the training data.
-
-We trained versions of both the RNN model and the RNTN model on these
-data sets. In both cases, the models were implemented exactly as
-described in Section~\ref{methods}, but since the items being compared
-are single terms rather than full tree structures, the composition
-layer was not involved, and the two models differed only in which
-layer function was used for the comparison layer. We simply present
-the models with two single terms, each of which corresponds to a
-single vector in the (randomly initialized) vocabulary matrix $V$,
-thereby ensuring that the model has no information about the terms
-being compared except the relations between them.
-
-\paragraph{Results} 
-We found that the RNTN model worked best with 11-dimensional vector
-representations for the 80 sets and a 90-dimensional feature vector
-for the classifier, though the performance of the model was not highly
-sensitive to either dimensionality parameter in any of the experiments
-discussed here. This model was able to correctly label 99.4\% (94.3
-macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
-best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
-respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
-the RNN.
-
-\begin{figure}[tp]
-  \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
-    \setlength{\unitlength}{1cm}
-    \begin{picture}(5,5.5)
-      \labelednode{2.50}{5}{}{1,2,3}
-
-      \put(0.75,4){\line(3,1){1.5}}
-      \put(2.5,4){\line(0,1){0.5}}
-      \put(4.25,4){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{3.5}{a,b}{1,2}
-      \labelednode{2.50}{3.5}{c}{1,3}
-      \labelednode{4.25}{3.5}{d}{2,3}
-
-      \put(0.75,2.5){\line(0,1){0.5}}
-      \put(0.75,2.5){\line(3,1){1.5}}
-
-      \put(2.5,2.5){\line(-3,1){1.5}}
-      \put(2.5,2.5){\line(3,1){1.5}}
-
-      \put(4.25,2.5){\line(0,1){0.5}}
-      \put(4.25,2.5){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{2}{e,f}{1}
-      \labelednode{2.50}{2}{}{2}
-      \labelednode{4.25}{2}{g,h}{3}
-
-      \put(2.5,1){\line(-3,1){1.5}}
-      \put(2.5,1){\line(0,1){0.5}}
-      \put(2.5,1){\line(3,1){1.5}}
-
-      \labelednode{2.5}{0.5}{}{}
-    \end{picture}
-    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
-    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}[t]{0.43\textwidth}
-    \centering
-    \setlength{\tabcolsep}{12pt}
-    \begin{tabular}[b]{c  c}
-      \toprule
-      Train & Test \\
-      \midrule
-                    & $b \nateq b$ \\
-      $b \natcov c$ &               \\
-                    & $b \natcov d$ \\
-                    & \strikeout{$b \natrev e$} \\
-      $c \natcov d$ &               \\
-      $c \natrev e$ &               \\
-                    & \strikeout{$c \nateq f$} \\
-      $c \natrev g$ &               \\ 
-                    & $e \natfor b$ \\
-      $e \natfor c$ &               \\[-1ex]
-      $\vdots$      & $\vdots$ \\
-      \bottomrule
-    \end{tabular}
-
-    \caption{A train/test split of the atomic statements about the
-      model.  Test statements not provable from the training data are
-      crossed out.}
-  \end{subfigure}  
-  \caption{Small example structure and data for learning relation composition.}
-  \label{lattice-figure}
-\end{figure} 
-% Note: None of these test examples is derivable from the shown training data, 
-% but we suggest that there is additional training data, so we can cross lines
-% out willy-nilly without fear.
-
-% RNTN log: tue-j-11-6x80-hip.txt
-% Train PER: 0.0021875
-
-% RNN log: tue-j-11-6x80-r-75.txt
-% Train PER: 0.047188
-
-% and create a dataset consisting of the relations between every pair of
-% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
-% test dataset, and that test dataset was further split into the 2960
-% examples that can be provably derived from the test data using
-% MacCartney and Manning's join table (or by the symmetry of the
-% relations in about half of the cases) and the 240 that
-% cannot.
-
-% We tested a version of both the RNN model and the RNTN model on these
-% data. In both cases, the models were implemented exactly as described
-% in Section~\ref{methods}, but since the items being compared are
-% single terms rather than full tree structures, the composition layer
-% was not used, and the two models differed only in which layer function
-% was used for the comparison layer. We found that the RNTN model worked
-% best with 11 dimensional vector representations for the 80 sets and a
-% 90 dimensional feature vector for the classifier. This model was able
-% to correctly label 99.3\% of the derivable test examples, and 99.1\%
-% of the remaining examples. The simpler RNN model worked best with 11
-% and 75 dimensions, respectively, but was able to achieve accuracies of
-% only 90.0\% and 87.\%, respectively.
-
-These results are fairly straightforward to interpret. The RNTN model
-was able to accurately encode the relations between the terms in the
-geometric relations between their vectors, and was able to then use
-that information to recover relations that were not overtly included
-in the training data. In contrast, the RNN model was able to achieve
-this behavior only incompletely. It is possible but not likely that it
-could be made to find a good solution with further optimization on
-different learning algorithms, or that it would do better on a larger
-universe of sets, for which there would be a larger set of training
-data to learn from, but the RNTN is readily able to achieve these
-effects in the setting discussed here.
-
-\section{Recursive structure in propositional logic}\label{sec:recursion}
-
-Recursive structure is a prominent feature of natural languages and an
-important part of their expressivity, as it allows people to use and
-interpret complex structures they have never encountered before.
-Theoretical accounts of natural language syntax and semantics
-therefore rely heavily on recursive structures, and we expect that
-accurate computational models will have to come to grips with them as
-well. The present section pursues this question for our two classes of
-RNN. In our evaluations, we exploit the fact that our logical language
-is infinite by testing on strings that are longer and more complex
-than any seen in training.
-
-% Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice
-%   said hello}, and \ii{Carl thinks that Bob said that Alice said
-%   hello}. Overt recursion of this kind is easy to find, and
-% theoretical accounts of natural language syntax and semantics rely
-% heavily on recursive structures. In order for a model to be able to
-% accurately learn natural language meanings, then, we expect that it
-% would need to be able to learn to represent the meanings of function
-% words in a such a way that they are able to behave correctly when
-% taking their own outputs as input. In evaluating the model, we take
-% advantage of the fact that recursive structures of this kind define
-% potentially infinite languages by testing the model on strings that
-% are longer and more complex than any seen in testing.
-
-\paragraph{Experiments}
-As in Section~\ref{sec:join}, we test this phenomenon within the
-framework of natural logic, but we now replace the unanalyzed symbols
-from that experiment with complex formulae. These formulae
-represent a truth-functionally complete classical propositional logic:
-each atomic variable denotes either $\True$ or $\False$, and the only
-operators are truth-functional ones.  Table~\ref{tab:pl} defines this
-logic, and Table~\ref{tab:plexs} gives some examples of relational
-statements that we can make in these terms. To compute these relations
-between statements, we exhaustively enumerate the sets of assignments
-of truth values to propositional variables that would satisfy each of
-the statements, and then we convert the set-theoretic relation between
-those assignments into one of the seven relations in
-Table~\ref{b-table}. As a result, each relational statement represents
-a valid theorem of the propositional logic.
-%
-%\begin{table}[tp]
-%  \centering
-%  \begin{subtable}[t]{0.45\textwidth}
-%    \centering
-%    \begin{tabular}[t]{l l}
-%      \toprule
-%      Formula     & Interpretation \\
-%      \midrule
-%      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
-%      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
-%      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
-%      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
-%      \bottomrule
-%    \end{tabular}    
-%    \caption{Well-formed formulae. $\varphi$ and $\psi$
-%      range over all well-formed formulae, and $\sem{\cdot}$ is
-%      the interpretation function mapping formulae into $\{\True,
-%      \False\}$.}\label{tab:pl}
-%  \end{subtable}
-%  \quad
-%  \begin{subtable}[t]{0.45\textwidth}
-%    \centering
-%    \begin{tabular}[t]{r c l}
-%      \toprule
-%      $\plneg a$        & $\natneg$ & $a$ \\
-%      $\plneg \plneg a$ & $\nateq$  & $a$ \\
-%      $a$               & $\natfor$ & $(a \plor b)$ \\
-%      $a$               & $\natrev$ & $(a \pland b)$ \\
-%      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
-%      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
-%      \bottomrule
-%    \end{tabular}
-%    \caption{Examples of statements about relations between
-%      well-formed formulae, defined in terms of sets of satisfying
-%      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
-%  \end{subtable}
-%  \caption{Propositional logic with natural logic relations.}  
-%  \label{prop-figure}
-%\end{table}
-
-% NOTE: There was some confusion at CSLI and the 224U lecture about
-% whether this was theorem proving or some kind of model checking.
-% Hopefully this is explicit enough.
-Socher et al.~\cite{socher2012semantic} show that a matrix-vector RNN
-model somewhat similar to our RNTN can learn a logic, but the logic
-discussed here is a much better approximation of the kind of
-expressivity needed for natural language. The logic learned in that
-experiment is boolean, wherein the atomic symbols are simply the
-values $0$ and $1$, rather than variables over those values. While
-learning the operators of that logic is not trivial, the outputs of
-each operator can be represented accurately by a single bit. The
-statements of propositional logic learned here describe conditions on
-unknown truth values of propositions. As opposed to the two-way
-contrasts seen in \cite{socher2012semantic}, this logic distinguishes
-between $2^{6} = 64$ possible assignments of truth values, and
-expressions of this logic define arbitrary conditions on these
-possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
-possible statements that the intermediate vector representations need
-to be able to distinguish if we are to learn this logical system in its
-full generality.
-
-For our experiments, we randomly generate a large set of  unique pairs 
-of formulae and compute the relation that holds for each pair.
-We discard pairs in which either statement is a tautology or
-contradiction (a statement that is true of either all or no possible
-assignments), for which none of the seven relation labels in
-Table~\ref{b-table} can hold. The resulting set of formula pairs is
-then partitioned into bins based on the number of logical operators in
-the longer of the two formulae. We then randomly sample 15\% of each
-bin for a held-out test set.
-
-If we do not implement any constraint that the two statements being
-compared are similar in any way, then the generated data are dominated
-by statements in which the two formulae refer to largely separate
-subsets of the six variables, which means that the $\natind$ relation
-is almost always correct.  In an effort to balance the distribution of
-relation labels without departing from the basic task of modeling
-propositional logic, we disallow individual pairs of statements from
-referring to more than four of the six propositional variables.
-
-\paragraph{Results} 
-We trained both the RNN and RNTN models on data with up to 4
-connectives (65k pairs) and tested it on examples of up to 12
-connectives (44k pairs). We initialized the model parameters randomly,
-including the vector representations of the six variables. The results
-are shown in Figure~\ref{prop-results}. In tuning, we found that the
-RNN model was approximately optimal with 45-dimensional vector
-representations, and the RNTN model was approximately optimal with 25
-dimensions. We fixed the size of the feature vector for the classifier
-at 75 dimensions. We found that the RNTN model was able to perform
-almost perfectly on unseen small test examples, with accuracy above
-98.7\% on formulae below length four (macro-F1 was above 95.5).
-Starting at size four, performance gradually falls with increasing
-size.  The RNN model did not perform well, reaching only 90.5\% (81.4
-micro-F1) accuracy on the smallest test examples, and declining from
-there to near-baseline performance on formulae of length 12. Training
-accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.
-
-% thu-and-or-deep-25-32b.txt iter 136
-% mon-and-or-deep-45r-32.txt iter 480
-
-\begin{figure}[tp]
-  \centering
-  \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
-  \caption{Model performance on propositional logic, by expression size. 
-    The dashed line indicates that only expressions of size four or less appeared in the training data.}  
-  \label{prop-results}
-\end{figure}
-
-The performance of the RNTN model on small unseen test examples
-indicates that it learned a correct approximation of the underlying
-logic. It appears that this approximation is accurate enough to yield
-correct answers when the composition layer is only applied a small
-number of times, but that the error in the approximation grows with
-increasing depth (and with the increasing complexity of the
-expressions), resulting in gradually dropping performance. In
-contrast, the RNN model did not appear to be able to learn a correct
-approximation of the logic for statements of any length, but it
-appears that the faulty logic that the RNN did learn decayed similarly
-with expression size.
-This is not necessarily a significant flaw in the model, since it remains possible that  different training regimes would allow it to learn accurate representations of the larger expressions.
-
-In an effort to better understand why the RNTN outperformed the plain
-RNN, we evaluated both models on pairs of long formulae involving
-binary connectives, to assess the extent to which they achieve
-substantially different representations for semantically distinct
-formulae and substantially similar representations for synonymous
-formulae. We found that both models are able to identify synonymous
-pairs like $(a \pland a)$ and $(a \pland (a \pland (a \pland a)))$,
-with both \ii{and} and \ii{or}, but only the RNTN is able to learn
-substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
- and $(a \pland (a \pland b))$ when the difference between the two is embedded under multiple operators. 
- Figure~\ref{prop-falloff} summarizes this result.
-
-% we discovered that this model looses information about in longer
-% statements in a particularly problematic way. It appears that that
-% model is unable to distinguish between two sentences when the only
-% difference between those sentences is embedded within a very deep
-% structure. We evaluated both models on sentences that differed in only
-% one term, but for which that one term was embedded under a large
-% number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and
-%   b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and
-%   b)))}. We then measured the Euclidean distance between the vector
-% representations of the two sentences in each pair. Our findings are
-% shown in Figure~\ref{prop-falloff}, and show that while the RNTN can
-% distinguish the two sentences well at every size that we test, the RNN
-% fails after a depth of approximately six.
-
-%\begin{figure}[htp]
-%  \centering
-%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-%  \caption{Semantically distinct formulae should have different
-%    representations. As measured by Euclidean distance, only the RNTN
-%    achieves this for formulae containing more than a small number of
-%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
-%    failing to capture the meaning contrasts.}
-%  \label{prop-falloff}
-%\end{figure}
-
-\section{Reasoning with natural language quantifiers and negation}\label{sec:quantifiers}
-
-We have seen that the RNTN can learn an approximation of propositional
-logic.  However, natural languages can express functional meanings of
-considerably greater complexity than this.  As a first step towards
-investigating whether our models can capture this complexity, we now
-attempt to directly measure the degree to which they are able to
-develop suitable representations for the semantics of natural language
-quantifiers like \ii{some} and \ii{all}. Quantification is far from
-the only place in natural language where complex functional meanings
-are found, but it is a natural starting point, since it can be tested
-in sentences whose structures are otherwise quite simple, and since it
-has formed a standard case study in prior formal work on natural
-language inference \cite{Icard:Moss:2013:LILT}.
-
-% \subsection{Data}
-
-This experiment replicates similar work described in
-\cite{bowman2013can}, which found that RNTNs can learn to reason well
-with quantifier meanings given sufficient training data. This paper
-replaces the partially manually annotated data in that paper with data
-that is generated directly using the logical system that we hope to
-learn, yielding results that we believe to be substantially more
-straightforward to interpret.
-
-\paragraph{Experiments}
-Our experimental data consist of pairs of sentences generated from a
-small artificial grammar. Each sentence contains a quantifier, a noun
-which may be negated, and an intransitive verb which may be
-negated. We use the basic quantifiers \ii{some}, \ii{most}, \ii{all},
-\ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
-\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}. We also
-include five nouns, four intransitive verbs, and the negation symbol
-\ii{not}. In order to be able to define relations between sentences
-with differing lexical items, we define the lexical relations between
-each noun--noun pair, each verb--verb pair, and each
-quantifier--quantifier pair. The grammar accepts aligned pairs of
-sentences of this form and calculates the natural logic relationship
-between them.  Some examples of these data are provided in
-Table~\ref{examplesofdata}.  As in previous sections, the goal of
-learning is then to assign these relational labels accurately to
-unseen pairs of sentences.
-
-%nouns = ['warthogs', 'turtles', 'mammals', 'reptiles', 'pets']
-%verbs = ['walk', 'move', 'swim', 'growl']
-%dets = ['all', 'not_all', 'some', 'no', 'most', 'not_most', 'two', 'lt_two', 'three', 'lt_three']
-%adverbs = ['', 'not']
-
-% To assign relation labels to sentence pairs, we built a small
-% task-specific implemenation of MacCartney's logic that can
-% accurately label sentences of this restricted language. The logic is
-% not able to derive all intuitively true relations of this language,
-% and fails to derive a single unique relation for certain types of
-% statement, including De Morgean's laws (e.g. \ii{(all pets) growl
-% $\natneg$ (some pet) (not growl)}), and we simply discard these
-% examples. Exhaustively generating the valid sentences under this
-% grammar and choosing those to which a relation label can be assigned
-% yields 66k sentence pairs. Some examples of these data are provided
-% in Table~\ref{examplesofdata}.
-
-%\begin{table}[htp]
-%  \centering
-%  \begin{tabular}{l c l}
-%    \toprule
-%    (most warthogs) walk         & $\natneg$ & (not-most warthogs) walk\\
-%    (most mammals) move          & $\natind$ &  (not-most (not turtles)) move\\
-%    (most (not pets)) (not swim) & $\natrev$ & (not-most (not pets)) move 
-%    \\[2ex]    
-%    (no turtles) (not growl)     & $\natalt$ & (no turtles) (not swim)\\
-%    (no warthogs) swim           & $\natrev$ & (no warthogs) move\\
-%    (no warthogs) move           & $\natfor$ & (no (not reptiles)) swim\\
-%    \bottomrule
-%  \end{tabular}
-%  \caption{Sample data involving two different quantifier pairs.}
-%  \label{examplesofdata}
-%\end{table}
-
-We evaluate the model using two experimental settings. In the simpler
-setting, \textsc{all split}, we randomly sample 85\% of the data and evaluate on the
-remaining 15\%. In this setting, the model is being asked to learn a
-complete reasoning system for the limited language and logic presented
-in the training data, but it is not being asked to generalize to test
-examples that are substantially different from those it was trained
-on. Crucially, though, to succeed on this task, the model must be able
-to recognize all of the lexical relations between the nouns, verbs,
-and quantifiers and how they interact. For instance, it might see
-\eqref{p1} and \eqref{p2} in training and, from that information,
-determine \eqref{p3}.
-%
-% do not allow a blank line --- adds too much space
-%
-\begin{gather}
-  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
-  \\
-  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
-  \\
-  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
-\end{gather}
-%
-% do not allow a blank line --- adds too much space
-%
-While our primary interest is in discovering the extent to which our
-models can learn to encode the logic given an arbitrary amount of
-data, we are also interested in the degree to which they can infer a
-correct representation for the logic from more constrained training
-data. To this end, we segment the sentence pairs according to which
-quantifiers appear in each pair, and then hold out one such pair for
-testing. We hypothesize that a model that can efficiently learn to
-represent a logic should be able to construct an accurate
-representation of each of the two quantifiers in the pair from the way that it
-interacts with the other nine quantifiers with which it is seen in training. 
-Since running this experiment requires choosing a pair of
-quantifiers to hold out before training, the resource demands of
-training prevent us from testing each of the 55 possible 
-pairs of quantifiers, so we choose only four pairs to test on.  Three
-of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
-\ii{all}/\ii{some}) were chosen because they allow for the most
-different class labels in the test data. The fourth is a
-self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
-handles equality.
-
-\begin{table}[tp]
-  \centering
-  \setlength{\tabcolsep}{10pt}
-  \begin{tabular}{ l rr@{ \ }r r@{ \ }r r@{ \ }r }
-    \toprule
-    Data & \multicolumn{3}{c}{Most frequent class} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
-    \midrule
-    \textsc{all split}	            &~~~~& 35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}
-    \\[1ex]    
-    \textsc{pair two/less-than-two} && 59.8 &(18.4) & 77.2 &(57.4) &	\textbf{100.0} &\textbf{(100.0)} \\
-    \textsc{pair not-all/not-most}  &&    0 &(0)    & 66.0 &(54.3) &	\textbf{93.8}  &\textbf{(91.5)} \\
-    \textsc{pair all/some}	    &&    0& (0)    & 62.7 &(65.9)  &	\textbf{78.4}  &\textbf{(84.7)} \\
-    \textsc{pair no/no}	            && 30.8  &(10.2) & 67.8 & (61.8) &	\textbf{99.9}  &\textbf{(99.7)} \\
-    \bottomrule
-  \end{tabular}
-  \caption{Performance on the quantifier experiments. Results are reported as accuracy scores followed by macroaveraged F1 scores in parentheses.}
-  \label{resultstable}
-\end{table}
-
-\paragraph{Results} 
-The results for these experiments are shown in
-Figure~\ref{resultstable}. We compare the results to a most frequent
-class baseline, which reflects the frequency in the test data of the
-most frequent class in the training data, $\natind$.  After some
-cross-validation, we chose 16- and 20-dimensional representations for
-the RNN and RNTN, respectively, and 75-dimensional feature vectors for
-the classifier. Training
-accuracy was near 80\% for the each of the five RNN experiments, and 
-near 98\% for each of the five RNTN experiments.
-
-%train rnn 0.20302, 0.017195
-
-The RNN performed poorly at this task, even though the sentences used
-in these examples are short enough to avoid the pathology shown in
-Figure~\ref{prop-falloff}.  However, the perfect performance by the
-RNTN on the \textsc{all split} and \textsc{pair two/less-than-two}
-experiments, and its strong performance on the \textsc{pair no/no}
-experiment, suggest that this stronger model is able to handle
-quantifiers correctly given sufficient training data, but the weaker
-results on two of the training settings suggest that it may not be
-able to generalize from this kind of restricted training data in general. However, though the question
-of how much data is necessary to accurately capture quantifier
-behavior in a na\"ive model remains open, the fact that this model is able to perform so well is promising.
-
-\section{General discussion}\label{sec:discussion}
-
-This paper evaluated two recursive models on a series of three increasingly
-challenging interpretive tasks involving natural language inference:
-the core relational algebra of natural logic with entailment and
-exclusion; recursive propositional logic structures; and statements
-involving quantification and negation. The results suggest that RNTNs,
-but not plain RNNs, have the capacity to meet the challenges of these
-tasks with reasonably-sized training sets. These positive results are
-promising for the future of learned representation models in the
-applied modeling of compositional semantics.
-
-Of course, challenges remain. In terms of our experimental data, even
-the RNTN falls short of perfection in our more complex tasks, with
-performance falling off steadily as the depth of recursion grows. It
-remains to be seen whether these deficiencies can be overcome with
-improvements to the model, the optimization procedures, or the
-linguistic representations
-\cite{sochergrounded,kalchbrenner2014convolutional}. In addition,
-there remain subtle questions about how to fairly assess whether these
-models have truly generalized in the way we want them to. There is a
-constant tension between showing the models training data that gives
-them a chance to learn the target logical functions and revealing the
-answer to them in a way that leads to overfitting. The underlying
-logical theories provide only limited guidance on this point.
-%
-%, and the fact that there is a finite universe of possible
-%  expressions makes his an unavoidable issue. 
-%
-%  CP: I don't understand the above. It is false if taken literally;
-%  our PL generates an infinte number of formulae.
-%
-Finally, we have only scratched the surface of the logical complexity
-of natural language; in future experiments, we hope to test sentences
-with embedded quantifiers, multiple interacting quantifiers, relative
-clauses, and other kinds of recursive structure. Nonetheless, the
-rapid progress the field has made with these models in recent years
-provides ample reason to be optimistic that they can be trained to
-meet the challenges of natural language semantics.
-
-% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators.
-This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic.
-The positive results that we have found, however, are extremely promising for the future of learned representation models in the applied modeling of meaning.
-We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators.
-We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.
-
-% There is ample room to build on these results.
-In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses.
-In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.
-
-% \textit{Withheld for review.}\\ \\
-% Do later: Fill in, cite grants, internal reviewers
-
-\bibliographystyle{unsrtnat}
-
-\small % Note: Explicitly allowed in style guide
-\bibliography{MLSemantics}
-
-\end{document}
+% We address this question directly by for the first time evaluating whether each of two classes of neural model — plain RNNs and recursive neural tensor networks (RNTNs) — can correctly learn relationships such as entailment and contradiction between pairs of sentences, where we have generated controlled data sets of sentences from a logical grammar.

### 2014-10-09 writing/AAAI SSS '15/wordnet.tex

@@
 
-% Can a model to do this with something that has the scale and structure of a real vocabulary?
+We have shown that a model can learn to reason using relations over a small artificial dataset, 
+but it the question of whether the same model can show this behavior over the type and number of
+relations seen in a real natural language vocabulary.
+
+Unfortunately, we are aware of no reliable source of relation data that covers a large natural 
+language vocabulary using the natural logic relations discussed above. However, WordNet 
+\cite{fellbaum2010wordnet} provides a different type of lattice structure which is compatible with
+our learning paradigm.
+
+We extract three types of relation from WordNet. \ii{Hypernym} and \ii{hyponym} can be are represented
+directly in the WordNet graph structure, and correspond closely to the \natrev and \natfor relations from
+natural logic. As in natural logic, these relations are mirror images of one another: if \ii{dog} is a
+hyponym of \ii{animal}, then \ii{animal} must be a hypernym of \ii{dog}.
 

### 2014-10-09 writing/AAAI SSS '15/wordnet.tex
Add draft of WordNet section

@@
 \subsection*{Reasoning about lexical relations in WordNet}\label{sec:wordnet}
-
-% TODO: Fill out
 
@@
 We extract three types of relation from WordNet. \ii{Hypernym} and \ii{hyponym} can be are represented
-directly in the WordNet graph structure, and correspond closely to the \natrev and \natfor relations from
+directly in the WordNet graph structure, and correspond closely to the $\natrev$ and $\natfor$ relations from
 natural logic. As in natural logic, these relations are mirror images of one another: if \ii{dog} is a
-hyponym of \ii{animal}, then \ii{animal} must be a hypernym of \ii{dog}.
+hyponym of \ii{animal} (perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then \ii{animal} must be a 
+hypernym of \ii{dog}. A third informative relation can 
+be readily derived from the graph, which we call \ii{coordinate}. Coordinate terms are those which share
+a direct hypernym, like \ii{dalmation}, \ii{pug}, and \ii{puppy}, which are all direct hyponyms of \ii{dog}. 
+Coordinate terms tend to exclude one another (as with $\natalt$ above) \cite{Hurford:1974}, but this tendancy
+is not universal. WordNet defines hypernymy and hyponymy over sets of synonyms, rather than over individual 
+terms, so we do not include a \ii{synonym} or \ii{equivalent} relation, but rather simply consider only one
+member of each set of synonyms.
 
-% Wordnet! Subsumption lattice for words!
+To limit the size of the vocabulary without otherwise simplifying the learning problem, we extract all of the
+instances of these three relations for those terms in WordNet that are hyponyms of the node \texttt{organism.n.01}.
+In order to balance the distribution of the classes, we slightly downsample instances of the \ii{coordinate} relation,
+yielding 36,772 relations among 3,217 terms. We randomly select 15\% of these relations as test data, and train on
+some or all of the remaining data.
 
-% Three relations: Hyper, hypo, coordinate
-% Explain extraction and rationale for each
+We find that with only a small amount of hyperparamater tuning, our model can perform almost perfectly, and that performance remains fairly good even when the fraction of the training data used is decreased.
 
-% Data creation:
-% - Enumerate filtering strategy
+% TODO: Describe use of GloVe once there are clearer results. Should we omit mention of it entirely?
 
-% Experiment:
-% - Explain train test split
-
-% Reults:
-% RNN / RNTN x Data volume x Accuracy vs. MacroF1
+\begin{table}[h]
+\centering\resizebox{4in}{!}{\small
+  \setlength{\tabcolsep}{15pt}
+  \renewcommand{\arraystretch}{1.1}
+  \begin{tabular}{l c l l} 
+    \toprule
+    Comparison layer & \% training data & w/ GloVe & w/o GloVe \\
+    NN       & 100\%   & TODO & TODO \\
+    NN       & 75\%   & 98.6 (100.0) & TODO \\
+    NN       & 33\%   & TODO & 37.0 (37.0) \\ % TODO: See if this improves
+    NN       & 16\%   & TODO & 92.0 (100.0)\\
+    NTN       & 100\%   & 99.6 (100.0) & 99.4 (99.9) \\
+    NTN       & 75\%   & 98.6 (100.0) & TODO\\
+    NTN       & 33\%   & 95.1 (99.8) & 95.2 (100.0) \\
+    NTN       & 16\%   & 94.6 (100.0) & 92.1 (100.0) \\
+    \bottomrule
+  \end{tabular}}
+ \caption{Test (train) accuracy figures on the WordNet data.\label{b-table}}  
+\end{table}

### 2014-10-09 writing/AAAI SSS '15/methods.tex
Create AAAI paper directory
Minor update

--- 2014-06-06 writing/S'14 paper/methods.tex
@@
-\section{Recursive neural network models} \label{methods}
+\subsection*{Neural network models for relation classification} \label{methods}
 
-We study neural models that adhere to the linguistic \ii{principle of
- compositionality}, which says that the meanings for complex
-expressions are derived from the meanings of their constituent parts
-via specific composition functions \cite{Partee84,Janssen97}. In our
-distributed setting, word meanings are vectors of length $n$. The
-composition function maps pairs of them to single vectors of length $n$, 
-which can then be merged again to represent more complex
-phrases. Once the entire sentence-level representation has been
-derived, it serves as a fixed-length input for some subsequent function.
+% TODO: Should we cite our NIPS manuscript?
 
-We use the model architecture proposed in \cite{bowman2013can} and
-depicted in Figure~\ref{sample-figure}. The two phrases being compared
+We follow the approach to learning semantically meaningful embeddings 
+proposed in \cite{bowman2013can}, which is centered on the problem of
+labeling a pair of words or sentences with one of a small set of logical
+relations. The architecture of the model that we use is depicted in
+Figure~\ref{sample-figure}. The model represents the two
+
+The two phrases being compared
 are built up separately on each side of the tree, using the same
@@
 multiplicative interactions between the child vectors. Both models
-include a bias vector, represented as $\vec{b}$.
+include a bias vector~$\vec{b}$.
 
@@
   \centering
- % \includegraphics[scale=0.35]{model.eps}
   \input{figure1}
-  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
+  \caption{The model structure used to compare \ii{turtle} and \ii{animal}. 
     The same structure is used for both the RNN and RNTN layer functions.} 

### 2014-10-09 writing/AAAI SSS '15/wordnet.tex
Minor update

@@
 
-% TODO: Describe use of GloVe once there are clearer results. Should we omit mention of it entirely?
+% TODO: Describe use of GloVe once there are clearer results. Since it doesn't look like it makes a huge difference,
+% should we perhaps just omit mention of it entirely to save space?
 
@@
     Comparison layer & \% training data & w/ GloVe & w/o GloVe \\
+    Most frequent class & -- & 37.9 (--) & 37.9 (--) \\
     NN       & 100\%   & TODO & TODO \\

### 2014-10-09 writing/AAAI SSS '15/abstract.tex
Draft abstract and intro

@@
 \begin{abstract}
-  % TODO: Rewrite
-
-  Supervised recursive neural network models (RNNs) for sentence
-  meaning have been successful in an array of sophisticated language
-  tasks, but it remains an open question whether they can learn
-  compositional semantic grammars that support logical deduction.  We
-  address this question directly by for the first time evaluating
-  whether each of two classes of neural model --- plain RNNs and
-  recursive neural tensor networks (RNTNs) --- can correctly learn
-  relationships such as entailment and contradiction between pairs of
-  sentences, where we have generated controlled data sets of sentences
-  from a logical grammar.  Our first experiment evaluates whether
-  these models can learn the basic algebra of logical relations
-  involved. Our second and third experiments extend this evaluation to
-  complex recursive structures and sentences involving quantification.
-  We find that the plain RNN achieves only mixed results on all three
-  experiments, whereas the stronger RNTN model generalizes well in
-  every setting and appears capable of learning suitable
+  Natural logic offers a powerful relational conception of meaning
+  that fits well with recent distributed approaches to semantic
+  representation. Such representations have proven valuable in a wide
+  range of sophisticated language tasks, but it remains an open
+  question whether it is possible to train them to support the rich,
+  diverse logical reasoning captured by natural logic. We address this
+  question for two classes of neural model: plain neural networks and
+  neural tensor networks.  Our first experiment evaluates these
+  models' ability to learn the basic algebra of natural logic
+  relations from simulated data, and our second experiment evaluates
+  them on the WordNet noun graph.  We find that the plain RNN achieves
+  only mixed results, whereas the stronger NTN generalizes well in
+  both settings and appears capable of learning suitable
   representations for natural language logical inference.
 \end{abstract}
-
-% We address this question directly by for the first time evaluating whether each of two classes of neural model — plain RNNs and recursive neural tensor networks (RNTNs) — can correctly learn relationships such as entailment and contradiction between pairs of sentences, where we have generated controlled data sets of sentences from a logical grammar.

### 2014-10-09 writing/AAAI SSS '15/intro.tex
Clean up methods section a bit
Draft abstract and intro

@@
 
-Supervised recursive neural network models (RNNs) for sentence meaning
-have been successful in an array of sophisticated language tasks,
-including sentiment analysis \cite{socher2011semi,socher2013acl1},
-image description \cite{sochergrounded}, and paraphrase detection
-\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
-the ability of these models to learn compositional semantic grammars,
-but it remains an open question whether they can achieve the same
-results as grammars based in logical forms
-\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
-when it comes to the core semantic concepts of quantification,
-entailment, and contradiction needed to identify the relationships
-between sentences like \ii{every animal walks}, \ii{every turtle
-  moves}, and \ii{most reptiles don't move}. To date, experimental
-investigations of these concepts using distributed representations
-have been largely confined to short phrases
-\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
-For robust natural language understanding, it is essential to model
-these phenomena in their full generality on complex linguistic
-structures.
+Natural logic offers a powerful \emph{relational} conception of
+semantics: the meanings for expressions are given, at least in part,
+by their inferential connections with other expressions
+\cite{vanBenthem08NATLOG,maccartney2009extended}. For instance,
+\word{turtle} is analyzed, not primarily by its extension in the
+world, but rather by its lexical network: it entails \word{reptile},
+excludes \word{chair}, is entailed by \word{sea
+  turtle}, and so forth. With generalized notions of entailment and
+contradiction, these relationships can be defined for all lexical
+categories as well as complex phrases, sentences, and even texts. The
+resulting theories of meaning offer valuable new analytic tools for
+tasks involving database inference, relation extraction, and textual
+entailment.
 
-We address this question in the context of \ii{natural language
-  inference} (also known as \ii{recognizing textual entailment};
-\cite{dagan2006pascal}), in which the goal is to determine the core
-inferential relationship between two sentences. Much of the
-theoretical work on this task (and some successful implemented models
-\cite{maccartney2009extended,watanabe2012latent}) involves \ii{natural
-  logics}, which are formal systems that define rules of inference
-between natural language words, phrases, and sentences without the
-need of intermediate representations in an artificial logical
-language. Following \cite{bowman2013can}, we use the natural logic
-developed by \cite{maccartney2009extended} as our formal model. This
-logic defines seven core relations of synonymy, entailment,
-contradiction, and mutual consistency, as summarized in
-Table~\ref{b-table}, and it provides rules of semantic combination for
-projecting these relations from the lexicon up to complex phrases. The
-formal properties and inferential strength of this system are now
-well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
+These approaches are naturally paired with theories using distributed
+(e.g., vector) representations, where meaning is also naturally
+modeled as relational. Distributed representations have been used
+successfully in a wide array of sophisticated language tasks,
+including sentiment analysis, analogy completion, relation extraction,
+and named entity recognition. However, it remains an open question
+whether it is possible to train such representations to support the
+rich, diverse logical reasoning captured by natural logic; while such
+representations excel at synonymy (similarity), the results are more
+mixed for entailment, contradiction, and mutual consistency.
 
-In our experiments, we use this pre-specified logical grammar to
-generate controlled data sets encoding the semantic relationships
-between pairs of expressions and evaluate whether each of two classes
-of neural model --- plain RNNs and recursive neural tensor networks
-(RNTNs, \cite{socher2013acl1}) --- can learn those relationships
-correctly. In our first experiment (Section~\ref{sec:join}), we
-evaluate the ability of these models to learn the core relational
-algebra of natural logic from data. Our second experiment
-(Section~\ref{sec:recursion}) extends this evaluation to relations
-between complex recursive structures like $(a \plor b)$ and
-$\plneg(\plneg a \pland \plneg b)$, and our third experiment
-(Section~\ref{sec:quantifiers}) involves relations between quantified
-statements like \ii{every reptile walks} and \ii{no turtle moves}.
-We find that the plain RNN achieves only mixed results in all three
-experiments, whereas the stronger RNTN model generalizes well in every
-case, suggesting that it has in fact learned, or at least learned to
-simulate, our target logical
-concepts. These experiments differentiate the increased power of RNTNs
-better than previous work and provide the most convincing
-demonstration to date of the ability of neural networks to model
-semantic inferences in complex natural language sentences.
+Using the natural logic of \cite{maccartney2009extended} as our formal
+model, we address this open question for two classes of neural model:
+plain neural networks and neural tensor networks (NTNs;
+\cite{socher2013acl1}). The natural logic is built from the seven
+relations defined in Table~\ref{b-table}. Its formal properties and
+inferential strength of the natural logic are now well-understood
+\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}, so it provides a rigorous
+and challenging set of goals for our distributed representations. To
+keep the discussion managable, we limit attention to experiments
+involving the lexicon; for a more extended treatment involving complex
+expressions involving logical connectives and quantifiers, see
+\cite{Bowman:Potts:Manning:2014}.
+
+In our first experiment, we use the pre-specified logical grammar to
+generate controlled data sets and assses the ability of the two
+classes of neural network to learn the core underlying relational
+algebra from this simulated data. In our second experiment, we define
+a simple modification of the logical grammar so that it aligns well
+with the WordNet noun graph, and we again assess our two classes of
+RNN on this data. The controlled data of our first experiment helps us
+to achieve analytic insights into what the models are able to learn,
+and the large, diverse lexicon of the second experiment shows us how
+they fare in the real world. We find that the plain RNN achieves only
+mixed results in these experiments, whereas the stronger RNTN model
+generalizes well in both cases, suggesting that it has in fact
+learned, or at least learned to simulate, our target logical concepts.
+% These experiments differentiate the increased power of RNTNs better
+% than previous work and provide the most convincing demonstration to
+% date of the ability of neural networks to model semantic inferences
+% in complex natural language sentences.
 
@@
 \end{table}
-
-% Citations to additional past work to be added.\\...\\...\\...\\...\\...
-
-% Deep learning methods in NLP which learn vector representations for words have seen successful uses in recent years on increasingly sophisticated tasks \cite{collobert2011natural, socher2011semi, socher2013acl1, chen2013learning}.
-Given the still modest performance of semantically rich NLP systems in many domains---question answering and machine translation, for instance---it is worth exploring the degree to which learned vectors can serve as general purpose semantic representations.
-Much of the work to date analyzing vector representations for words (see \cite{baroni2013frege}) has focused on lexical semantic behaviors---like the similarity between words like \ii{Paris} and \ii{France}.
-Good similarity functions are valuable for many NLP tasks, but there are real use cases for which it is necessary to know how words relate to one another or to some extrinsic representation, and to model the ways in which word meanings combine to form phrase, sentence, or document meanings.
-This paper explores the ability of linguistic representations developed using supervised deep learning techniques to support interpretation and reasoning.
-
-% There are two broad family of tasks that could be used to test the ability of a model to develop general purpose meaning representations.
-In an interpretation task, sentences are mapped onto some denotation, such as  \ii{true} or \ii{false} for statements, or a factual answer for questions.
-There has been preliminary work in developing distributed models for interpretation \cite{grefenstette2013towards, rocktaschellow}, but developing a representation of world knowledge that corresponds accurately to the content expressed in language introduces considerable design challenges.
-I approach the problem by way of an inference task instead.
-Inferring the truth of one sentence from another does not require any preexisting knowledge representations, but nonetheless requires a precise representation of sentence meaning.
-I borrow the structure of the task from MacCartney and Manning  \cite{maccartney2009extended}.
-In it, the model is presented with a pair of sentences, and made to label the logical relation between the sentences as equivalence, entailment, or any of five other classes, as here:
-
-%\begin{quote}
-%\begin{enumerate}\small
-%\item Many smartphone users avoid high bills overseas by disabling data service.
-%\item Not everyone uses their smartphones for email when traveling abroad.
-%\end{enumerate} 
-%$\Rightarrow$ Entailment
-%\end{quote}
-
-%In this paper, we test the ability of recursive models to on three simple tasks, each of which is meant to capture a property that is necessary in representing natural language meaning in the setting of inference.
-I begin with an overview of MacCartney and Manning's \cite{maccartney2009extended} framework for inference, and of the recursive neural networks that I study.
-by showing that these models can learn to correctly represent entailment representations between sentences.
-I then show that these models can capture the meanings of recursive structures accurately up to a sufficient depth.
-I finally close with a brief demonstration of the ability of these models to reason over short natural language sentences involving quantifiers.
-
-% \subsection{Natural language inference and natural logic}
-
-% In its simplest form, \ii{natural language inference} (also known as \ii{recognizing textual entailment}, as in \cite{dagan2006pascal}) is the task of determine whether one sentence entails another.
-Much of the theoretical work on this task (and some successful implemented models \cite{maccartney2009natural, watanabe2012latent}) involve \ii{natural logic}, formal systems that define sound rules of inference from one sentence of natural language to another without the need for intermediate representations in some other logic.
-The most powerful model that we are aware of for natural logic is due to MacCartney and Manning \cite{maccartney2009extended} and Icard \cite{icard2012inclusion}, and is centered around the definition of a set of seven logical relations which can hold between sentences, shown in Table \ref{b-table}.
-
-% This approach to natural logic can capture much of the complexity of natural language meaning within a well understood framework, and is also fairly straightforward to implement in a machine learning setting since it can be reduced to a seven-way classification problem on sentence pairs.
-Our goal in this paper is to learn recursive neural network models which are able to mimic key behaviors of this system.
-
-% \begin{table}
-% \begin{center}
-% \begin{tabular}{|c|c|c|c|} \hline
-% name & symbol & set-theoretic definition & example \\ \hline \hline
-% entailment & $x \sqsubset y$ & $x \subset y$ & \ii{crow, bird}  \\ \hline
-% reverse entailment & $x \sqsupset y$ & $x \supset y$ & \ii{Asian, Thai}  \\ \hline
-% equivalence & $x \equiv y$ & $x = y$ & \ii{couch, sofa} \\ \hline
-% alteration & $x$ $|$ $y$ & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{cat, dog} \\ \hline
-% negation & $x \natneg y$ & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{able, unable} \\ \hline
-% cover & $x \smallsmile y$ & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-ape} \\ \hline
-% independence & $x$ \# $y$ & (else) & \ii{hungry, hippo}\\ \hline
-% \end{tabular}
-% \caption{The entailment relations in  $\mathfrak{B}$.
-$\mathcal{D}$ is the universe of possible objects of the same type as those being compared, and the relation \# applies whenever none of the other six do, including when there is insufficient knowledge to choose a label.}
-% \label{b-table}
-% \end{center}
-% \end{table}
-
-% The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes.
-It should be noted that this is not a balanced classification problem.
-If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them.
-% TODO: How much should we say about this?
-
-% The natural logic engine at the core of MacCartney and Manning's system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures.
-These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
-%for dozens of different cases of insertion, deletion and substitution of different types of lexical item.
-For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences.
-Here, the substitution of \ii{no} for \ii{some}  yields $\natneg$ and the substitution of \ii{dogs} for \ii{animals} yields $\sqsupset$.
-It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
-%\begin{quote}
-
-%1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
-%2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
-%3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}
-
-%\end{quote}
-
-% Work to date on inference in neural network models is quite limited.
-% \citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
-% In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations.
-They propose a deterministic inference engine (similar to MacCartney's) which is augmented by a probabilistic component that evaluates individual lexical substitution steps in the derivation using vector representations, though again these representations are not learned, and no evaluations of this system have been published to date.
-% \label{sec2}

### 2014-10-09 writing/AAAI SSS '15/wordnet.tex
Draft abstract and intro

--- 2014-10-09 writing/AAAI SSS '15/wordnet.tex
@@
 We extract three types of relation from WordNet. \ii{Hypernym} and \ii{hyponym} can be are represented
-directly in the WordNet graph structure, and correspond closely to the \natrev and \natfor relations from
+directly in the WordNet graph structure, and correspond closely to the $\natrev$ and $\natfor$ relations from
 natural logic. As in natural logic, these relations are mirror images of one another: if \ii{dog} is a

### 2014-10-09 writing/AAAI SSS '15/methods.tex

@@
 We follow the approach to learning semantically meaningful embeddings 
-proposed in \cite{bowman2013can}, which is centered on the problem of
+proposed in \citet{bowman2013can}, which is centered on the problem of
 labeling a pair of words or sentences with one of a small set of logical
-relations. The architecture of the model that we use is depicted in
-Figure~\ref{sample-figure}. The model represents the two
+relations. The architecture of the model that we use, which is limited
+to only pairs of single symbols (such as words), is depicted in
+Figure~\ref{sample-figure}. The model represents the two input symbols
+as embeddings, which are fed into a comparison function based on one
+of two types of neural network layer functions to produce a representation
+for the relationship between the two symbols. This representation is then
+fed into a simple softmax classifier which outputs a distribution over
+possible labels. The entire network, including the embeddings, are trained
+through backpropagation with AdaGrad \cite{duchi2011adaptive}.
 
-The two phrases being compared
-are built up separately on each side of the tree, using the same
-composition function, until they have each been reduced to single
-vectors. The resulting vectors are fed into a separate comparison
-layer that is meant to generate a feature vector capturing the
-relation between the two phrases. The output of this layer is then
-given to a softmax classifier, which in turn produces a hypothesized
-distribution over the seven relations represented in Table~\ref{b-table}.
+\begin{figure}[tp]
+  \centering
+  \input{figure1}
+  \caption{The model structure used to compare \ii{turtle} and \ii{animal}. 
+    The same structure is used for both the RNN and RNTN layer functions.} 
+  \label{sample-figure}
+\end{figure}
 
-For a composition layer, we evaluate models with both the plain neural
-network layer function \eqref{rnn} and the RNTN layer function
-\eqref{rntn} proposed in \citet{chen2013learning}. A sigmoid
-nonlinearity (element-wise $\tanh$) is applied to the output of either
-layer function, following \cite{socher2013acl1}.
+For a comparison funtion, we evaluate versions of the model with both a plain neural
+network (NN) layer function and a neural tensor network (NTN) layer function
+\eqref{rntn} proposed in \citet{chen2013learning}. A leaky ReLU
+nonlinearity \cite{maasrectifier} is applied to the output of either
+layer function.
 %
 \begin{gather} \label{rnn}
-\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
+\vec{y}_{\textit{NN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
 \label{rntn}
-\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
+\vec{y}_{\textit{NTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
 \end{gather} 
@@
 
-The comparison layer uses the same kind of function template as the
-composition layers (either an NN layer or an NTN layer) with
-independently learned parameters and a separate nonlinearity function.
-Rather than use a $\tanh$ nonlinearity here, we follow \cite{bowman2013can} 
-in using a rectified linear function. In
-particular, we use the leaky rectified linear function
-\cite{maasrectifier}: $f(\vec{x})=\max(\vec{x}, 0) +
-0.01\min(\vec{x}, 0)$, applied element-wise.
-
-To run the model forward and label a pair of phrases, the structure of
-the lower layers of the network is assembled so as to mirror the tree
-structures provided for each phrase. The word vectors are then looked
-up from the vocabulary matrix $V$ (one of the model parameters), and
-the composition and comparison functions are used to pass information
-up the tree and into the classifier at the top. For an objective
-function, we use the negative log of the probability assigned to the
-correct label.
-
-% Removed 'minibatch' -> For two of the experiments, we use minibatches
-% of size 1, which doesn't really count.
-We train the model using stochastic gradient descent (SGD)
-with learning rates computed using AdaGrad \cite{duchi2011adaptive}.
-The parameters (including the vocabulary) are initialized randomly 
-using a uniform distribution over $[-0.1, 0.1]$.
-Because the seven classes are not balanced in general, we report performance
-using both accuracy and macroaveraged F1, since the latter emphasizes
- performance on infrequent classes. We compute macroaveraged F1 
-as the harmonic mean of mean precision and mean recall, both computed
-for all classes for which there is test data, setting precision to 0 
-where it is not defined. Source code and generated data will be released
-after the review period.
-
-\begin{figure}[tp]
-  \centering
-  \input{figure1}
-  \caption{The model structure used to compare \ii{turtle} and \ii{animal}. 
-    The same structure is used for both the RNN and RNTN layer functions.} 
-  \label{sample-figure}
-\end{figure}
+This model differs from that of \citet{bowman2013can} in two ways. Because 
+the inputs are single symbols, there is no need for the composition functions
+which are used in that (and prior) work. Also, for our second experiment on 
+WordNet data, we introduce a new neural network layer between the embedding input
+and the comparison function, which is meant to facilitate initializing the embeddings
+from an outside source, and was found to help peformance in that setting.
 

### 2014-10-09 writing/AAAI SSS '15/wordnet.tex

--- 2014-10-09 writing/AAAI SSS '15/wordnet.tex
@@
 relations seen in a real natural language vocabulary.
-
 Unfortunately, we are aware of no reliable source of relation data that covers a large natural 
@@
 
+\paragraph{Results} 
 We find that with only a small amount of hyperparamater tuning, our model can perform almost perfectly, and that performance remains fairly good even when the fraction of the training data used is decreased.

### 2014-10-09 writing/AAAI SSS '15/wordnet.tex
Incomplete results update

@@
 We find that with only a small amount of hyperparamater tuning, our model can perform almost perfectly, and that performance remains fairly good even when the fraction of the training data used is decreased.
+Neither model has reached perfect performance in any of our experiments, though we are
+continuing to tune both models. We find that the NN and NTN both capture the required
+generalizations extremely well.
 
@@
   \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{l c l l} 
+  \begin{tabular}{l c l l l l} 
     \toprule
-    Comparison layer & \% training data & w/ GloVe & w/o GloVe \\
-    Most frequent class & -- & 37.9 (--) & 37.9 (--) \\
-    NN       & 100\%   & TODO & TODO \\
-    NN       & 75\%   & 98.6 (100.0) & TODO \\
-    NN       & 33\%   & TODO & 37.0 (37.0) \\ % TODO: See if this improves
-    NN       & 16\%   & TODO & 92.0 (100.0)\\
-    NTN       & 100\%   & 99.6 (100.0) & 99.4 (99.9) \\
-    NTN       & 75\%   & 98.6 (100.0) & TODO\\
-    NTN       & 33\%   & 95.1 (99.8) & 95.2 (100.0) \\
-    NTN       & 16\%   & 94.6 (100.0) & 92.1 (100.0) \\
+    Comparison layer & \% training data & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} \\
+     &  & NN & NTN & NN & NTN \\
+    \midrule
+    Baseline & -- & 37.9 (--) & 37.9 (--) & 37.9 (--) & 37.9 (--) \\
+    \midrule
+    NN       & 100\%   & 99.5 (100.0) & 98.9 (99.7) \\
+    NN       & 75\%   & 98.9 (100.0) & 37.9 (37.0) \\
+    NN       & 33\%   & 96.0 (100.0) & 37.0 (37.0) \\ % TODO: Rerunning. Got stuck predicting the most frequent class.
+    NN       & 16\%   & 91.8 (100.0) & 91.9 (100.0)\\
+    \midrule
+    NTN       & 100\%   & \textbf{99.6} (100.0) & 99.2 (100.0) \\
+    NTN       & 75\%   & 98.6 (100.0) & 98.4 (100.0)\\
+    NTN       & 33\%   & 95.4 (100.0) & 95.3 (100.0) \\
+    NTN       & 16\%   & 92.1 (99.7) & 92.2 (100.0) \\
     \bottomrule
   \end{tabular}}
- \caption{Test (train) accuracy figures on the WordNet data.\label{b-table}}  
+ \caption{Test (train) accuracy figures on the WordNet data. The baseline figure is simply the frequency of the most frequent class, \ii{coordinate}.\label{b-table}}  
 \end{table}

### 2014-10-09 RunExperiments.m

@@ -155,4 +155,7 @@ echo "cd quant; mbs = 64; lambda = 0.0005; ed = 1; loadWords = 0; penult = 80, d
 echo "cd quant; mbs = 64; lambda = 0.0005; ed = 1; loadWords = 1; penult = 80, dp = 0.16, name=['orss-l', num2str(lambda), '-ed', num2str(ed), '-tr', num2str(loadWords), '-pen', num2str(penult), '-mbs', num2str(mbs), '-dp', num2str(dp)]; disp(name); TrainModel('synset-relations', '', name, mbs,25, penult, 0.1, lambda, 0, ed, loadWords, dp);" | /afs/cs/software/bin/matlab_r2012b
 echo "cd quant; mbs = 64; lambda = 0.0005; ed = 1; loadWords = 0; penult = 80, dp = 0.75, name=['oss-l', num2str(lambda), '-ed', num2str(ed), '-tr', num2str(loadWords), '-pen', num2str(penult), '-mbs', num2str(mbs), '-dp', num2str(dp)]; disp(name); TrainModel('synset-relations', '', name, mbs,25, penult, 0.1, lambda, 1, ed, loadWords, dp);" | /afs/cs/software/bin/matlab_r2012b
 
+echo "cd quant; mbs = 64; lambda = 0.0005; ed = 1; loadWords = 0; penult = 80, dp = 0.75, name=['orbss-l', num2str(lambda), '-ed', num2str(ed), '-tr', num2str(loadWords), '-pen', num2str(penult), '-mbs', num2str(mbs), '-dp', num2str(dp)]; disp(name); TrainModel('synset-relations', '', name, mbs,25, penult, 0.1, lambda, 0, ed, loadWords, dp);" | /afs/cs/software/bin/matlab_r2012b
+echo "cd quant; mbs = 64; lambda = 0.0005; ed = 1; loadWords = 0; penult = 80, dp = 0.33, name=['orbss-l', num2str(lambda), '-ed', num2str(ed), '-tr', num2str(loadWords), '-pen', num2str(penult), '-mbs', num2str(mbs), '-dp', num2str(dp)]; disp(name); TrainModel('synset-relations', '', name, mbs,25, penult, 0.1, lambda, 0, ed, loadWords, dp);" | /afs/cs/software/bin/matlab_r2012b
[…]

### 2014-10-09 writing/AAAI SSS '15/intro.tex
Attempting to resolve conflicts, shrunk figs

@@
 
-Natural logic offers a powerful \emph{relational} conception of
-semantics: the meanings for expressions are given, at least in part,
-by their inferential connections with other expressions
-\cite{vanBenthem08NATLOG,maccartney2009extended}. For instance,
-\word{turtle} is analyzed, not primarily by its extension in the
-world, but rather by its lexical network: it entails \word{reptile},
-excludes \word{chair}, is entailed by \word{sea
-  turtle}, and so forth. With generalized notions of entailment and
-contradiction, these relationships can be defined for all lexical
-categories as well as complex phrases, sentences, and even texts. The
-resulting theories of meaning offer valuable new analytic tools for
-tasks involving database inference, relation extraction, and textual
-entailment.
+Supervised recursive neural network models (RNNs) for sentence meaning
+have been successful in an array of sophisticated language tasks,
+including sentiment analysis \cite{socher2011semi,socher2013acl1},
+image description \cite{sochergrounded}, and paraphrase detection
+\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
+the ability of these models to learn compositional semantic grammars,
+but it remains an open question whether they can achieve the same
+results as grammars based in logical forms
+\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
+when it comes to the core semantic concepts of quantification,
+entailment, and contradiction needed to identify the relationships
+between sentences like \\resizebox{5.25in}{!}{{every animal walks}, \ii{every turtle
+  moves}, and \ii{most reptiles don't move}. To date, experimental
+investigations of these concepts using distributed representations
+have been largely confined to short phrases
+\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
+For robust natural language understanding, it is essential to model
+these phenomena in their full generality on complex linguistic
+structures.
 
-These approaches are naturally paired with theories using distributed
-(e.g., vector) representations, where meaning is also naturally
-modeled as relational. Distributed representations have been used
-successfully in a wide array of sophisticated language tasks,
-including sentiment analysis, analogy completion, relation extraction,
-and named entity recognition. However, it remains an open question
-whether it is possible to train such representations to support the
-rich, diverse logical reasoning captured by natural logic; while such
-representations excel at synonymy (similarity), the results are more
-mixed for entailment, contradiction, and mutual consistency.
+We address this question in the context of \ii{natural language
+  inference} (also known as \ii{recognizing textual entailment};
+\cite{dagan2006pascal}), in which the goal is to determine the core
+inferential relationship between two sentences. Much of the
+theoretical work on this task (and some successful implemented models
+\cite{maccartney2009extended,watanabe2012latent}) involves \ii{natural
+  logics}, which are formal systems that define rules of inference
+between natural language words, phrases, and sentences without the
+need of intermediate representations in an artificial logical
+language. Following \cite{bowman2013can}, we use the natural logic
+developed by \cite{maccartney2009extended} as our formal model. This
+logic defines seven core relations of synonymy, entailment,
+contradiction, and mutual consistency, as summarized in
+Table~\ref{b-table}, and it provides rules of semantic combination for
+projecting these relations from the lexicon up to complex phrases. The
+formal properties and inferential strength of this system are now
+well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
 
-Using the natural logic of \cite{maccartney2009extended} as our formal
-model, we address this open question for two classes of neural model:
-plain neural networks and neural tensor networks (NTNs;
-\cite{socher2013acl1}). The natural logic is built from the seven
-relations defined in Table~\ref{b-table}. Its formal properties and
-inferential strength of the natural logic are now well-understood
-\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}, so it provides a rigorous
-and challenging set of goals for our distributed representations. To
-keep the discussion managable, we limit attention to experiments
-involving the lexicon; for a more extended treatment involving complex
-expressions involving logical connectives and quantifiers, see
-\cite{Bowman:Potts:Manning:2014}.
-
-In our first experiment, we use the pre-specified logical grammar to
-generate controlled data sets and assses the ability of the two
-classes of neural network to learn the core underlying relational
-algebra from this simulated data. In our second experiment, we define
-a simple modification of the logical grammar so that it aligns well
-with the WordNet noun graph, and we again assess our two classes of
-RNN on this data. The controlled data of our first experiment helps us
-to achieve analytic insights into what the models are able to learn,
-and the large, diverse lexicon of the second experiment shows us how
-they fare in the real world. We find that the plain RNN achieves only
-mixed results in these experiments, whereas the stronger RNTN model
-generalizes well in both cases, suggesting that it has in fact
-learned, or at least learned to simulate, our target logical concepts.
-% These experiments differentiate the increased power of RNTNs better
-% than previous work and provide the most convincing demonstration to
-% date of the ability of neural networks to model semantic inferences
-% in complex natural language sentences.
+In our experiments, we use this pre-specified logical grammar to
+generate controlled data sets encoding the semantic relationships
+between pairs of expressions and evaluate whether each of two classes
+of neural model --- plain RNNs and recursive neural tensor networks
+(RNTNs, \cite{socher2013acl1}) --- can learn those relationships
+correctly. In our first experiment (Section~\ref{sec:join}), we
+evaluate the ability of these models to learn the core relational
+algebra of natural logic from data. Our second experiment
+(Section~\ref{sec:recursion}) extends this evaluation to relations
+between complex recursive structures like $(a \plor b)$ and
+$\plneg(\plneg a \pland \plneg b)$, and our third experiment
+(Section~\ref{sec:quantifiers}) involves relations between quantified
+statements like \ii{every reptile walks} and \ii{no turtle moves}.
+We find that the plain RNN achieves only mixed results in all three
+experiments, whereas the stronger RNTN model generalizes well in every
+case, suggesting that it has in fact learned, or at least learned to
+simulate, our target logical
+concepts. These experiments differentiate the increased power of RNTNs
+better than previous work and provide the most convincing
+demonstration to date of the ability of neural networks to model
+semantic inferences in complex natural language sentences.
 
 \begin{table}[tp]
-  \centering
+  \centering\small
   \setlength{\tabcolsep}{15pt}
@@
 \end{table}
+
+% TODO: Add citations on related work in structuring relations in embedding spaces:
+% - Representing part–whole relations in conceptual spaces
+% - Something McCallum/Universal Schema?

### 2014-10-09 writing/AAAI SSS '15/wordnet.tex
Attempting to resolve conflicts, shrunk figs

@@
 In order to balance the distribution of the classes, we slightly downsample instances of the \ii{coordinate} relation,
-yielding 36,772 relations among 3,217 terms. We randomly select 15\% of these relations as test data, and train on
-some or all of the remaining data.
+yielding 36,772 relations among 3,217 terms. We randomly select 15\% of these relations as test data. Unlike in the
+previous experiment, it is not straightforward here to determine in advance how much training data should be required
+to learn an accurate model, so we performed training runs with various fractions of the training portion of the 
+dataset.
 
@@
 We find that with only a small amount of hyperparamater tuning, our model can perform almost perfectly, and that performance remains fairly good even when the fraction of the training data used is decreased.
-Neither model has reached perfect performance in any of our experiments, though we are
-continuing to tune both models. We find that the NN and NTN both capture the required
-generalizations extremely well.
+Neither model has reached perfect performance in any of our experiments, though we are continuing to tune both models.
+We find that the NN and NTN both capture the required generalizations very well.
 
@@
 \begin{table}[h]
-\centering\resizebox{4in}{!}{\small
+\centering\resizebox{5.5in}{!}{
   \setlength{\tabcolsep}{15pt}
   \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{l c l l l l} 
+  \begin{tabular}{l l l l l l} 
     \toprule
-    Comparison layer & \% training data & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} \\
-     &  & NN & NTN & NN & NTN \\
+     Portion of training data & \multicolumn{2}{c}{NN} & \multicolumn{2}{c}{NTN} & Baseline\\
+       & \multicolumn{1}{c}{w/ GloVe} & \multicolumn{1}{c}{w/o GloVe} & \multicolumn{1}{c}{w/ GloVe} & \multicolumn{1}{c}{w/o GloVe} \\
     \midrule
-    Baseline & -- & 37.9 (--) & 37.9 (--) & 37.9 (--) & 37.9 (--) \\
-    \midrule
-    NN       & 100\%   & 99.5 (100.0) & 98.9 (99.7) \\
-    NN       & 75\%   & 98.9 (100.0) & 37.9 (37.0) \\
-    NN       & 33\%   & 96.0 (100.0) & 37.0 (37.0) \\ % TODO: Rerunning. Got stuck predicting the most frequent class.
-    NN       & 16\%   & 91.8 (100.0) & 91.9 (100.0)\\
-    \midrule
-    NTN       & 100\%   & \textbf{99.6} (100.0) & 99.2 (100.0) \\
-    NTN       & 75\%   & 98.6 (100.0) & 98.4 (100.0)\\
-    NTN       & 33\%   & 95.4 (100.0) & 95.3 (100.0) \\
-    NTN       & 16\%   & 92.1 (99.7) & 92.2 (100.0) \\
+     100\%   & 99.5 (100.0) & 98.9 (99.7) & \textbf{99.6} (100.0) & 99.2 (100.0) & 37.9 (--)\\
+     75\%   & 98.9 (100.0) & 37.9 (37.0) & 98.6 (100.0) & 98.4 (100.0) & 37.9 (--)\\
+     33\%   & 96.0 (100.0) & 37.0 (37.0) & 95.4 (100.0) & 95.3 (100.0) & 37.9 (--)\\ % TODO: Rerunning. Got stuck predicting the most frequent class.
+     16\%   & 91.8 (100.0) & 91.9 (100.0) & 92.1 (99.7) & 92.2 (100.0) & 37.9 (--)\\
     \bottomrule

### 2014-10-09 writing/AAAI SSS '15/intro.tex
Conflict repair
Small methods cleanup, more squishing

--- 2014-10-09 writing/AAAI SSS '15/intro.tex
@@
 
-\begin{table}[tp]
-  \centering
-  \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{l c l l} 
-    \toprule
-    Name & Symbol & Set-theoretic definition & Example \\ 
-    \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
-    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
-    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
-    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
-    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
-    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
-    \bottomrule
-  \end{tabular}
-  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
-    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
-    %is insufficient knowledge to choose a label.}
-  \label{b-table}
-\end{table}
+% TODO: Add citations on related work in structuring relations in embedding spaces:
+% - Representing part–whole relations in conceptual spaces
+% - Something McCallum/Universal Schema?

### 2014-10-09 writing/AAAI SSS '15/methods.tex
Small methods cleanup, more squishing

@@
 We follow the approach to learning semantically meaningful embeddings 
-proposed in \citet{bowman2013can}, which is centered on the problem of
+of \citet{Bowman:Potts:Manning:2014}, which is centered on the problem of
 labeling a pair of words or sentences with one of a small set of logical
@@
 network (NN) layer function and a neural tensor network (NTN) layer function
-\eqref{rntn} proposed in \citet{chen2013learning}. A leaky ReLU
+proposed in \citet{chen2013learning}, which adds a third order tensor parameter
+to mediate multiplicative interactions between the two inputs. A leaky ReLU
 nonlinearity \cite{maasrectifier} is applied to the output of either
@@
 \begin{gather} \label{rnn}
-\vec{y}_{\textit{NN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
-\label{rntn}
-\vec{y}_{\textit{NTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
+\vec{y}_{\textit{NN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) ~~~~~~~~~~~~~~~ \vec{y}_{\textit{NTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
 \end{gather} 
 %
-Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
-representations for the left and right children of the node, and
-$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
-them by an $n \times 2n$ matrix of learned weights, and applies the
-element-wise non-linearity to the resulting vector. The RNTN has the
-same basic structure, but with the addition of a learned third-order
-tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
-multiplicative interactions between the child vectors. Both models
-include a bias vector~$\vec{b}$.
 
-This model differs from that of \citet{bowman2013can} in two ways. Because 
+This model differs from that of \citet{Bowman:Potts:Manning:2014} in two ways. Because 
 the inputs are single symbols, there is no need for the composition functions

### 2014-10-09 writing/AAAI SSS '15/methods.tex
Updates to methods/wordnet.

@@
 
-We follow the approach to learning semantically meaningful embeddings 
-of \citet{Bowman:Potts:Manning:2014}, which is centered on the problem of
+We build a model to learn embeddings using the 
+method of \citet{Bowman:Potts:Manning:2014}, which is centered on the problem of
 labeling a pair of words or sentences with one of a small set of logical
@@
   \input{figure1}
-  \caption{The model structure used to compare \ii{turtle} and \ii{animal}. 
-    The same structure is used for both the RNN and RNTN layer functions.} 
+  \caption{The model structure used to compare \ii{turtle} and \ii{animal}.} 
   \label{sample-figure}
@@
 the inputs are single symbols, there is no need for the composition functions
-which are used in that (and prior) work. Also, for our second experiment on 
-WordNet data, we introduce a new neural network layer between the embedding input
+which are used in that (and prior) work. Also, for our experiment on 
+WordNet data, we introduce an additional neural network layer between the embedding input
 and the comparison function, which is meant to facilitate initializing the embeddings
-from an outside source, and was found to help peformance in that setting.
+themselves from pretrained vectors, and was found to help peformance in that setting.
 

### 2014-10-09 writing/AAAI SSS '15/wordnet.tex
Updates to methods/wordnet.

@@
 hyponym of \ii{animal} (perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then \ii{animal} must be a 
-hypernym of \ii{dog}. A third informative relation can 
-be readily derived from the graph, which we call \ii{coordinate}. Coordinate terms are those which share
+hypernym of \ii{dog}. We also extract \ii{coordinate} terms, which share
 a direct hypernym, like \ii{dalmation}, \ii{pug}, and \ii{puppy}, which are all direct hyponyms of \ii{dog}. 
@@
 To limit the size of the vocabulary without otherwise simplifying the learning problem, we extract all of the
-instances of these three relations for those terms in WordNet that are hyponyms of the node \texttt{organism.n.01}.
-In order to balance the distribution of the classes, we slightly downsample instances of the \ii{coordinate} relation,
-yielding 36,772 relations among 3,217 terms. We randomly select 15\% of these relations as test data. Unlike in the
-previous experiment, it is not straightforward here to determine in advance how much training data should be required
-to learn an accurate model, so we performed training runs with various fractions of the training portion of the 
-dataset.
+instances of these three relations for single word nouns in WordNet that are hyponyms of the node 
+\texttt{organism.n.01}. In order to balance the distribution of the classes, we slightly downsample instances 
+of the \ii{coordinate} relation, yielding 36,772 relations among 3,217 terms. We randomly select 15\% of these 
+relations as test data. Unlike in the previous experiment, it is not straightforward here to determine in advance
+ how much training data should be required to learn an accurate model, so we performed training runs with 
+ various fractions of the training portion of the 
+dataset. Embeddings were fixed at 25 dimensions and were initialized either randomly or using distributional 
+vectors from GloVe \cite{pennington2014glove}.
 
 \paragraph{Results} 
-We find that with only a small amount of hyperparamater tuning, our model can perform almost perfectly, and that performance remains fairly good even when the fraction of the training data used is decreased.
-Neither model has reached perfect performance in any of our experiments, though we are continuing to tune both models.
-We find that the NN and NTN both capture the required generalizations very well.
-
-% TODO: Describe use of GloVe once there are clearer results. Since it doesn't look like it makes a huge difference,
-% should we perhaps just omit mention of it entirely to save space?
+We find that our model can perform almost perfectly, and that performance remains fairly good even when the model 
+is trained only on a fraction of the ground truth relations. Neither model has reached perfect performance in any
+of our experiments, though we are continuing to tune both models. We also find that the two comparison functions
 

### 2014-10-09 writing/AAAI SSS '15/abstract.tex
Editing

@@
   Natural logic offers a powerful relational conception of meaning
-  that fits well with recent distributed approaches to semantic
-  representation. Such representations have proven valuable in a wide
-  range of sophisticated language tasks, but it remains an open
-  question whether it is possible to train them to support the rich,
-  diverse logical reasoning captured by natural logic. We address this
-  question for two classes of neural model: plain neural networks and
-  neural tensor networks.  Our first experiment evaluates these
-  models' ability to learn the basic algebra of natural logic
-  relations from simulated data, and our second experiment evaluates
-  them on the WordNet noun graph.  We find that the plain RNN achieves
-  only mixed results, whereas the stronger NTN generalizes well in
-  both settings and appears capable of learning suitable
-  representations for natural language logical inference.
+  that is a natural counterpart to distributed semantic
+  representations, which have proven valuable in a wide range of
+  sophisticated language tasks. However, it remains an open question
+  whether it is possible to train distributed representations to
+  support the rich, diverse logical reasoning captured by natural
+  logic. We address this question for two classes of neural model:
+  plain neural networks and neural tensor networks (NTNs).  Our
+  experiments evaluate these models' ability to learn the basic
+  algebra of natural logic relations from simulated data and from the
+  WordNet noun graph.  We find that the plain RNN achieves only mixed
+  results, whereas the stronger NTN generalizes well in both settings
+  and appears capable of learning suitable representations for natural
+  language logical inference.
 \end{abstract}

### 2014-10-09 writing/AAAI SSS '15/intro.tex
Editing

@@
 
-These approaches are naturally paired with theories using distributed
-(e.g., vector) representations, where meaning is also naturally
-modeled as relational. Distributed representations have been used
-successfully in a wide array of sophisticated language tasks,
-including sentiment analysis, analogy completion, relation extraction,
-and named entity recognition. However, it remains an open question
-whether it is possible to train such representations to support the
-rich, diverse logical reasoning captured by natural logic; while such
-representations excel at synonymy (similarity), the results are more
-mixed for entailment, contradiction, and mutual consistency.
+Natural logic aligns well with distributed (e.g., vector)
+representations, which also naturally model meaning relationally.
+Distributed representations have been used successfully in a wide
+array of sophisticated language tasks. %,
+% including sentiment analysis, analogy completion, relation
+% extraction, and named entity recognition.
+However, it remains an open question whether it is possible to train
+such representations to support the rich, diverse logical reasoning
+captured by natural logic; while they excel at
+synonymy (similarity), the results are more mixed for entailment,
+contradiction, and mutual consistency.  Using the natural logic of
+\cite{maccartney2009extended} as our formal model, we address this
+open question for two classes of neural model: plain neural networks
+and neural tensor networks (NTNs; \cite{socher2013acl1}). The natural
+logic is built from the seven relations defined in
+Table~\ref{b-table}. Its formal properties are now well-understood
+\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}, so it provides a rigorous
+set of goals for our neural models. To keep the discussion managable,
+we limit attention to experiments involving the lexicon; for a more
+extended treatment of complex expressions involving logical
+connectives and quantifiers, see \citet{Bowman:Potts:Manning:2014}.
 
-Using the natural logic of \cite{maccartney2009extended} as our formal
-model, we address this open question for two classes of neural model:
-plain neural networks and neural tensor networks (NTNs;
-\cite{socher2013acl1}). The natural logic is built from the seven
-relations defined in Table~\ref{b-table}. Its formal properties and
-inferential strength of the natural logic are now well-understood
-\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}, so it provides a rigorous
-and challenging set of goals for our distributed representations. To
-keep the discussion managable, we limit attention to experiments
-involving the lexicon; for a more extended treatment involving complex
-expressions involving logical connectives and quantifiers, see
-\cite{Bowman:Potts:Manning:2014}.
-
-In our first experiment, we use the pre-specified logical grammar to
-generate controlled data sets and assses the ability of the two
-classes of neural network to learn the core underlying relational
-algebra from this simulated data. In our second experiment, we define
-a simple modification of the logical grammar so that it aligns well
-with the WordNet noun graph, and we again assess our two classes of
-RNN on this data. The controlled data of our first experiment helps us
-to achieve analytic insights into what the models are able to learn,
-and the large, diverse lexicon of the second experiment shows us how
-they fare in the real world. We find that the plain RNN achieves only
+In our experiments, we evaluate these models' ability to learn the
+basic algebra of natural logic relations from simulated data and from
+the WordNet noun graph. The simulated data help us to achieve analytic
+insights into what the models learn, and WordNet data show how they
+fare in the real world.  We find that the plain RNN achieves only
 mixed results in these experiments, whereas the stronger RNTN model
@@
 learned, or at least learned to simulate, our target logical concepts.
+
+% In our first experiment, we use the pre-specified logical grammar to
+% generate controlled data sets and assess the ability of the two
+% classes of neural network to learn the core underlying relational
+% algebra from this simulated data. In our second experiment, we define
+% a simple modification of the logical grammar so that it aligns with
+% the WordNet noun graph, and we again assess our two classes of RNN on
+% this data. The controlled data of our first experiment helps us to
+% achieve analytic insights into what the models are able to learn, and
+% the large, diverse lexicon of the second experiment shows us how they
+% fare in the real world.
+
 % These experiments differentiate the increased power of RNTNs better

### 2014-10-09 writing/AAAI SSS '15/methods.tex
Editing

--- 2014-10-09 writing/AAAI SSS '15/methods.tex
@@
 We follow the approach to learning semantically meaningful embeddings 
-of \citet{Bowman:Potts:Manning:2014}, which is centered on the problem of
+of \cite{Bowman:Potts:Manning:2014}, which is centered on the problem of
 labeling a pair of words or sentences with one of a small set of logical

### 2014-10-09 writing/AAAI SSS '15/methods.tex

--- 2014-10-09 writing/AAAI SSS '15/methods.tex
@@
 
-We build a model to learn embeddings using the 
-method of \citet{Bowman:Potts:Manning:2014}, which is centered on the problem of
-labeling a pair of words or sentences with one of a small set of logical
-relations. The architecture of the model that we use, which is limited
-to only pairs of single symbols (such as words), is depicted in
-Figure~\ref{sample-figure}. The model represents the two input symbols
-as embeddings, which are fed into a comparison function based on one
-of two types of neural network layer functions to produce a representation
-for the relationship between the two symbols. This representation is then
-fed into a simple softmax classifier which outputs a distribution over
-possible labels. The entire network, including the embeddings, are trained
-through backpropagation with AdaGrad \cite{duchi2011adaptive}.
+We build a model to learn embeddings using the method of
+\citet{Bowman:Potts:Manning:2014}, which is centered on the problem of
+labeling a pair of words or sentences with one of a small set of
+logical relations. The architecture of the model that we use, which is
+limited to only pairs of single symbols (such as words), is depicted
+in Figure~\ref{sample-figure}. The model represents the two input
+symbols as embeddings, which are fed into a comparison function based
+on one of two types of neural network layer functions to produce a
+representation for the relationship between the two symbols. This
+representation is then fed into a simple softmax classifier which
+outputs a distribution over possible labels. The entire network,
+including the embeddings, are trained through backpropagation with
+AdaGrad \cite{duchi2011adaptive}.
 

### 2014-10-09 writing/AAAI SSS '15/join.tex
Attempting to resolve conflicts, shrunk figs
More compact join section

@@
 \begin{table}[tp]
-  \centering
+  \centering\small
   \setlength{\tabcolsep}{15pt}
@@
 full set of sound inferences of this form is summarized in
-Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
+Figure~\ref{tab:jointable}; cells containing a dot correspond to pairs
 of relations for which no valid inference can be drawn in our logic.
@@
 
-\begin{table}[htp]
-  \centering  
-  \setlength{\arraycolsep}{8pt}
-  \renewcommand{\arraystretch}{1.1}
-  \newcommand{\UNK}{\cdot}  
-  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
-    %\hline
-    \multicolumn{1}{c}{}
-             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
-    \cline{2-8}
-    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
-    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
-    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
-    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
-    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
-    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
-    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
-    \cline{2-8}
-  \end{array}$
-  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.
-These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
-We assess our models' ability to reproduce such inferential paths.}
-  \label{tab:jointable}
-\end{table}
-
 \paragraph{Experiments}
 To test our models' ability to learn these inferential patterns, we
-create small boolean structures for our logic in which terms denote
-sets of entities from a small domain.  Figure~\ref{lattice-figure}
-depicts a structure of this form. The lattice gives the full model,
-for which all the statements on the right are valid. We divide these
-statements evenly into training and test sets, and remove from the
-test set those statements which cannot be proven from the training
-statements.
+create boolean structures in which terms denote sets of entities from
+a small domain (e.g., Figure~\ref{lattice-figure}), employ logical
+deduction to identify the valid statements, divide those into train
+and test sets, and remove from the test set those statements which
+cannot be proven from the training statements
+(Figure~\ref{unprovable}).
 % using the logic depicted in Figure~\ref{lattice-figure}.
-
+%
+%
 % We test the model's ability to learn this behavior by creating
@@
 % between them.
-
-In our experiments, we create 80 randomly generated sets drawing from
-a domain of seven elements. This yields a data set consisting of
-6400 statements about pairs of formulae. 3200 of these pairs are
-chosen as a test set, and that test set is further reduced to the 2960
-examples that can be provably derived from the training data.
-
-We trained versions of both the RNN model and the RNTN model on these
-data sets. In both cases, the models were implemented exactly as
-described in Section~\ref{methods}, but since the items being compared
-are single terms rather than full tree structures, the composition
-layer was not involved, and the two models differed only in which
-layer function was used for the comparison layer. We simply present
-the models with two single terms, each of which corresponds to a
-single vector in the (randomly initialized) vocabulary matrix $V$,
-thereby ensuring that the model has no information about the terms
-being compared except the relations between them.
+%
+In our experiments, we create 80 randomly generated sets drawn from a
+domain of seven elements. This yields 6400 statements about pairs of
+formulae, of which 3200 are chosen as a test set, and that test set is
+further reduced to the 2960 examples that can be provably derived from
+the training data. We trained both the NN model and the NTN model on
+these data
+sets% . In both cases, the models were implemented exactly as
+% described in Section~\ref{methods}, but since the items being compared
+% are single terms rather than full tree structures, the composition
+% layer was not involved, and the two models differed only in which
+% layer function was used for the comparison layer. We simply present
+% the models with two single terms, each of which corresponds to a
+% single vector in the (randomly initialized) vocabulary matrix $V$,
+% thereby ensuring that the model has no information about the terms
+% being compared except the relations between them.
 
 \paragraph{Results} 
-We found that the RNTN model worked best with 11-dimensional vector
+We found that the NTN model worked best with 11-dimensional vector
 representations for the 80 sets and a 90-dimensional feature vector
@@
 macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
+(92.2 macro-F1) of the remaining test examples. The simpler NN worked
 best with 11 and 75 dimensions, respectively, but was able to achieve
 accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
-respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
-the RNN.
+respectively. Training accuracy was 99.8\% for the NTN and 95.4\% for
+the NN.
 
@@
   \centering
-  \begin{subfigure}[t]{0.45\textwidth}
+  \begin{subfigure}[b]{0.42\textwidth}
+    \centering  
+    \setlength{\arraycolsep}{8pt}
+    \renewcommand{\arraystretch}{1.1}
+    \newcommand{\UNK}{\cdot}  
+    \resizebox{2.2in}{!}{$\begin{array}[c]{c@{ \ }|*{7}{c}|}
+        % \hline
+        \multicolumn{1}{c}{}
+        & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
+        \cline{2-8}
+        \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
+        \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
+        \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
+        \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
+        \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
+        \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
+        \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
+        \cline{2-8}
+      \end{array}$}
+    \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.
+These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
+We assess our models' ability to reproduce such inferential paths.}
+    \label{tab:jointable}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.3\textwidth}
     \centering
@@
     \setlength{\unitlength}{1cm}
-    \begin{picture}(5,5.5)
+    \resizebox{1.5in}{!}{\begin{picture}(5,5.5)
       \labelednode{2.50}{5}{}{1,2,3}
@@
       \labelednode{2.5}{0.5}{}{}
-    \end{picture}
+    \end{picture}}
     \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
-    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
+    some sets have multiple names, so that learning $\nateq$ is non-trivial.}\label{lattice-figure}
   \end{subfigure}
-  \qquad
-  \begin{subfigure}[t]{0.43\textwidth}
+  \hfill
+  \begin{subfigure}[b]{0.2\textwidth}
     \centering
     \setlength{\tabcolsep}{12pt}
-    \begin{tabular}[b]{c  c}
+    \resizebox{1.05in}{!}{\begin{tabular}[c]{c  c}
       \toprule
@@
       \bottomrule
-    \end{tabular}
+    \end{tabular}}
 
@@
       model.  Test statements not provable from the training data are
-      crossed out.}
+      crossed out.}\label{unprovable}
   \end{subfigure}  
   \caption{Small example structure and data for learning relation composition.}
-  \label{lattice-figure}
+  \label{exp1}
 \end{figure} 
@@
 
-These results are fairly straightforward to interpret. The RNTN model
+These results are fairly straightforward to interpret. The NTN model
 was able to accurately encode the relations between the terms in the
@@
 that information to recover relations that were not overtly included
-in the training data. In contrast, the RNN model was able to achieve
-this behavior only incompletely. It is possible but not likely that it
+in the training data. In contrast, the NN was able to achieve this
+behavior only incompletely. It is possible but not likely that it
 could be made to find a good solution with further optimization on
@@
 universe of sets, for which there would be a larger set of training
-data to learn from, but the RNTN is readily able to achieve these
+data to learn from, but the NTN is readily able to achieve these
 effects in the setting discussed here.

### 2014-10-09 writing/AAAI SSS '15/methods.tex
More compact join section

@@
 representation for the relationship between the two symbols. This
-representation is then fed into a simple softmax classifier which
-outputs a distribution over possible labels. The entire network,
-including the embeddings, are trained through backpropagation with
-AdaGrad \cite{duchi2011adaptive}.
+representation is then fed into a softmax classifier, which outputs a
+distribution over possible labels. The entire network, including the
+embeddings, are trained through backpropagation with AdaGrad
+\cite{duchi2011adaptive}.
 
@@
 
-For a comparison funtion, we evaluate versions of the model with both a plain neural
-network (NN) layer function and a neural tensor network (NTN) layer function
-proposed in \citet{chen2013learning}, which adds a third order tensor parameter
-to mediate multiplicative interactions between the two inputs. A leaky ReLU
-nonlinearity \cite{maasrectifier} is applied to the output of either
-layer function.
+For a comparison funtion, we evaluate versions of the model with both
+a plain NN and the NTN layer function proposed in
+\citet{chen2013learning}, which adds a third order tensor parameter to
+mediate multiplicative interactions between the two inputs. A leaky
+ReLU nonlinearity \cite{maasrectifier} is applied to the output of
+either layer function.
 %
@@
 
-This model differs from that of \citet{Bowman:Potts:Manning:2014} in two ways. Because 
-the inputs are single symbols, there is no need for the composition functions
-which are used in that (and prior) work. Also, for our experiment on 
-WordNet data, we introduce an additional neural network layer between the embedding input
-and the comparison function, which is meant to facilitate initializing the embeddings
-themselves from pretrained vectors, and was found to help peformance in that setting.
+This model differs from that of \citet{Bowman:Potts:Manning:2014} in
+two ways. Because the inputs are single symbols, there is no need for
+the composition functions used in that (and prior) work. Also, for our
+experiment on WordNet data, we introduce an additional neural network
+layer between the embedding input and the comparison function, which
+facilitates initializing the embeddings themselves from pretrained
+vectors and was found to help peformance in that setting.
 

### 2014-10-09 writing/AAAI SSS '15/abstract.tex
Abstract update, added citation

@@
   support the rich, diverse logical reasoning captured by natural
-  logic. We address this question for two classes of neural model:
-  plain neural networks and neural tensor networks (NTNs).  Our
-  experiments evaluate these models' ability to learn the basic
-  algebra of natural logic relations from simulated data and from the
-  WordNet noun graph.  We find that the plain RNN achieves only mixed
-  results, whereas the stronger NTN generalizes well in both settings
-  and appears capable of learning suitable representations for natural
-  language logical inference.
+  logic. We address this question a simple neural network-based
+  model for learning embeddings.  Our experiments evaluate the models' 
+  ability to learn the basic algebra of natural logic relations from 
+  simulated data and from the WordNet noun graph.  We find that the model
+  can learn to do inference over relations with near perfect precision
+  from moderate amounts of training data.
 \end{abstract}

### 2014-10-09 writing/AAAI SSS '15/intro.tex
Adding results summary, cutting an Icard paper.

@@
 \cite{maccartney2009extended} as our formal model, we address this
-open question for two classes of neural model: plain neural networks
-and neural tensor networks (NTNs; \cite{socher2013acl1}). The natural
-logic is built from the seven relations defined in
+open question for a simple neural network-based classification model. 
+The natural logic is built from the seven relations defined in
 Table~\ref{b-table}. Its formal properties are now well-understood
-\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}, so it provides a rigorous
+\cite{Icard:Moss:2013:LILT}, so it provides a rigorous
 set of goals for our neural models. To keep the discussion managable,
@@
 insights into what the models learn, and WordNet data show how they
-fare in the real world.  We find that the plain RNN achieves only
-mixed results in these experiments, whereas the stronger RNTN model
-generalizes well in both cases, suggesting that it has in fact
-learned, or at least learned to simulate, our target logical concepts.
+fare in the real world.  We find that even our model is able to learn
+natural logic-style relational reasoning with near-perfect accuracy
+in both settings.
 

### 2014-10-09 writing/AAAI SSS '15/wordnet.tex
Revised wordnet section

@@
 
-We have shown that a model can learn to reason using relations over a small artificial dataset, 
-but it the question of whether the same model can show this behavior over the type and number of
-relations seen in a real natural language vocabulary.
-Unfortunately, we are aware of no reliable source of relation data that covers a large natural 
-language vocabulary using the natural logic relations discussed above. However, WordNet 
-\cite{fellbaum2010wordnet} provides a different type of lattice structure which is compatible with
-our learning paradigm.
+Using simulated data as above is reassuring about what the models
+learn and why, but we also want to know how they perform with a real
+natural language vocabulary. Unfortunately, as far as we are aware,
+there are no available resources labeling such a vocabulary with the
+relations from Table~\ref{b-table}. However, the relations in WordNet
+\cite{fellbaum2010wordnet} come close and pose the same substantive
+challenges.
 
-We extract three types of relation from WordNet. \ii{Hypernym} and \ii{hyponym} can be are represented
-directly in the WordNet graph structure, and correspond closely to the $\natrev$ and $\natfor$ relations from
-natural logic. As in natural logic, these relations are mirror images of one another: if \ii{dog} is a
-hyponym of \ii{animal} (perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then \ii{animal} must be a 
-hypernym of \ii{dog}. We also extract \ii{coordinate} terms, which share
-a direct hypernym, like \ii{dalmation}, \ii{pug}, and \ii{puppy}, which are all direct hyponyms of \ii{dog}. 
-Coordinate terms tend to exclude one another (as with $\natalt$ above) \cite{Hurford:1974}, but this tendancy
-is not universal. WordNet defines hypernymy and hyponymy over sets of synonyms, rather than over individual 
-terms, so we do not include a \ii{synonym} or \ii{equivalent} relation, but rather simply consider only one
-member of each set of synonyms.
+% WordNet \cite{fellbaum2010wordnet} provides a different type of
+% lattice structure which is compatible with our learning paradigm.
+
+We extract three types of relation from WordNet. \ii{Hypernym} and
+\ii{hyponym} can be are represented directly in the WordNet graph
+structure, and correspond closely to the $\natrev$ and $\natfor$
+relations from natural logic. As in natural logic, these relations are
+mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
+(perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then
+\ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
+terms, which share a direct hypernym, like \ii{dalmation}, \ii{pug},
+and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
+terms tend to exclude one another, thereby providing a reasonable
+approximation of exclusion. WordNet defines hypernymy and hyponymy over
+sets of synonyms, rather than over individual terms, so we do not
+include a \ii{synonym} or \ii{equivalent} relation, but rather
+consider only one member of each set of synonyms.
 
@@
 \paragraph{Results} 
-We find that our model can perform almost perfectly, and that performance remains fairly good even when the model 
-is trained only on a fraction of the ground truth relations. Neither model has reached perfect performance in any
-of our experiments, though we are continuing to tune both models. We also find that the two comparison functions
+We find that the NTN performs almost perfectly even when trained only
+on a fraction of the ground truth relations. The plain NN is much less
+successful. Neither model has reached perfect performance in any of
+our experiments, though we are continuing to tune both models.
 
@@
   \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{l l l l l l} 
+  \begin{tabular}{r r@{ }r r@{ }r r@{ }r r@{ }r r@{ }r} 
     \toprule
-     Portion of training data & \multicolumn{2}{c}{NN} & \multicolumn{2}{c}{NTN} & Baseline\\
-       & \multicolumn{1}{c}{w/ GloVe} & \multicolumn{1}{c}{w/o GloVe} & \multicolumn{1}{c}{w/ GloVe} & \multicolumn{1}{c}{w/o GloVe} \\
+     Portion of & \multicolumn{4}{c}{NN} & \multicolumn{4}{c}{NTN} & \multicolumn{2}{c}{Baseline}\\
+     training data  & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} \\
     \midrule
-     100\%   & 99.5 (100.0) & 98.9 (99.7) & \textbf{99.6} (100.0) & 99.2 (100.0) & 37.9 (--)\\
-     75\%   & 98.9 (100.0) & 37.9 (37.0) & 98.6 (100.0) & 98.4 (100.0) & 37.9 (--)\\
-     33\%   & 96.0 (100.0) & 37.0 (37.0) & 95.4 (100.0) & 95.3 (100.0) & 37.9 (--)\\ % TODO: Rerunning. Got stuck predicting the most frequent class.
-     16\%   & 91.8 (100.0) & 91.9 (100.0) & 92.1 (99.7) & 92.2 (100.0) & 37.9 (--)\\
+     100\%   & 99.5 &(100.0) & 98.9 &(99.7)    & \textbf{99.6} &(100.0) & 99.2 &(100.0) & 37.9 &(--)\\
+     75\%    & 98.9 &(100.0) & 37.9 &(37.0)    & 98.6 &(100.0)          & 98.4 &(100.0) & 37.9 &(--)\\
+     33\%    & 96.0 &(100.0) & 37.0 &(37.0)    & 95.4 &(100.0)          & 95.3 &(100.0) & 37.9 &(--)\\ % TODO: Rerunning. Got stuck predicting the most frequent class.
+     16\%    & 91.8 &(100.0) & 91.9 &(100.0)   & 92.1 &(99.7)           & 92.2 &(100.0) & 37.9 &(--)\\
     \bottomrule

### 2014-10-09 writing/AAAI SSS '15/methods.tex
Revised wordnet section
Trimming methods.

@@
 \subsection*{Neural network models for relation classification} \label{methods}
-
-% TODO: Should we cite our NIPS manuscript?
 
@@
 
+The simpler version of the comparison simply concatenates the two input
+vectors before feeding them into a standard neural network (NN) layer.
+The more powerful neural tensor network (NTN) version uses an additional 
+third-order tensor parameter to allow for multiplicative
+interactions between the two inputs. For more details on the implementation and
+training of the layer functions, see \cite{Bowman:Potts:Manning:2014}.
+
 \begin{figure}[tp]
@@
 
-For a comparison funtion, we evaluate versions of the model with both
-a plain NN and the NTN layer function proposed in
-\citet{chen2013learning}, which adds a third order tensor parameter to
-mediate multiplicative interactions between the two inputs. A leaky
-ReLU nonlinearity \cite{maasrectifier} is applied to the output of
-either layer function.
-%
-\begin{gather} \label{rnn}
-\vec{y}_{\textit{NN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) ~~~~~~~~~~~~~~~ \vec{y}_{\textit{NTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
-\end{gather} 
-%
-
-This model differs from that of \citet{Bowman:Potts:Manning:2014} in
-two ways. Because the inputs are single symbols, there is no need for
-the composition functions used in that (and prior) work. Also, for our
-experiment on WordNet data, we introduce an additional neural network
-layer between the embedding input and the comparison function, which
-facilitates initializing the embeddings themselves from pretrained
-vectors and was found to help peformance in that setting.
+This model used here differs from the one described in that work in
+two ways. Because the inputs are single symbols, we do not use
+a composition function here. Also, for our experiment on WordNet data, 
+we introduce an additional neural network layer between the embedding 
+input and the comparison function, which facilitates initializing the 
+embeddings themselves from pretrained vectors and was found to help 
+peformance in that setting.
 

### 2014-10-09 writing/AAAI SSS '15/methods.tex
One more citation cut.
One more methods tweak.

@@
 
-We build a model to learn embeddings using the method of
-\citet{Bowman:Potts:Manning:2014}, which is centered on the problem of
+We build an embedding based model using the method of
+\cite{Bowman:Potts:Manning:2014}, which is centered on the task of
 labeling a pair of words or sentences with one of a small set of
@@
 distribution over possible labels. The entire network, including the
-embeddings, are trained through backpropagation with AdaGrad
-\cite{duchi2011adaptive}.
+embeddings, are trained using backpropagation.
 

### 2014-10-10 writing/AAAI SSS '15/join.tex
Caption update
Tweaks.

@@
 
-If a model is to learn the behavior of a relational logic like the one
-presented here from a finite amount of data, it must learn to deduce new
-relations from seen relations in a sound manner. The simplest such
-deductions involve atomic statements using the relations in
-Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
+If a model is to learn the behavior of a relational logic like the natural
+logic presented here over some domain from a finite amount of data, it 
+must learn to deduce new relational statements from seen ones in a sound manner.
+The simplest such deductions, and the only ones that can be made using relations
+between single terms, involve the set of inference rules schematized in 
+Figure~\ref{tab:jointable}. For instance, given that $a \natrev b$ and $b
 \natrev c$, one can conclude that $a \natrev c$, by basic
 set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
-full set of sound inferences of this form is summarized in
-Figure~\ref{tab:jointable}; cells containing a dot correspond to pairs
-of relations for which no valid inference can be drawn in our logic.
+$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$. Cells in the
+table containing a dot correspond to pairs of relations for which no valid 
+inference can be drawn in our logic.
 
@@
 To test our models' ability to learn these inferential patterns, we
-create boolean structures in which terms denote sets of entities from
+create artificial boolean structures in which terms denote sets of entities from
 a small domain (e.g., Figure~\ref{lattice-figure}), employ logical
@@
   \end{subfigure}  
-  \caption{Small example structure and data for learning relation composition.}
+  \caption{Experimental goal and set-up for reasoning about semantic relations.}
   \label{exp1}

### 2014-10-10 writing/AAAI SSS '15/wordnet.tex
Tweaks.

@@
  various fractions of the training portion of the 
-dataset. Embeddings were fixed at 25 dimensions and were initialized either randomly or using distributional 
-vectors from GloVe \cite{pennington2014glove}.
+dataset. Embeddings were fixed at 25 dimensions and were initialized randomly or using distributional 
+vectors from GloVe \cite{pennington2014glove}. The feature vector produced by the comparison layer was fixed at 
+80 dimensions.
 
@@
 We find that the NTN performs almost perfectly even when trained only
-on a fraction of the ground truth relations. The plain NN is much less
+on a fraction of the ground truth relations. The plain NN is somewhat less
 successful. Neither model has reached perfect performance in any of

### 2014-10-10 writing/AAAI SSS '15/intro.tex
Minor edits

@@
-\subsection*{Introduction}\label{sec:intro}
+\section{Introduction}\label{sec:intro}
   % TODO: Trim and revise
@@
 Distributed representations have been used successfully in a wide
-array of sophisticated language tasks. %,
+array of sophisticated language tasks \cite{collobert2011natural}. %,
 % including sentiment analysis, analogy completion, relation
@@
 \cite{Icard:Moss:2013:LILT}, so it provides a rigorous
-set of goals for our neural models. To keep the discussion managable,
+set of goals for our neural models. To keep the discussion manageable,
 we limit attention to experiments involving the lexicon; for a more
@@
 insights into what the models learn, and WordNet data show how they
-fare in the real world.  We find that even our model is able to learn
-natural logic-style relational reasoning with near-perfect accuracy
-in both settings.
+fare in the real world.  We find that only the NTN is able to fully
+learn the underlying algebra, but that both models excel in the 
+WordNet experiment.
 

### 2014-10-10 writing/AAAI SSS '15/methods.tex
Minor edits

@@
-\subsection*{Neural network models for relation classification} \label{methods}
+\section{Neural network models for relation classification} \label{methods}
 
-We build an embedding based model using the method of
+We build an embedding-based model using the method of
 \cite{Bowman:Potts:Manning:2014}, which is centered on the task of
@@
 
-The simpler version of the comparison simply concatenates the two input
-vectors before feeding them into a standard neural network (NN) layer.
-The more powerful neural tensor network (NTN) version uses an additional 
-third-order tensor parameter to allow for multiplicative
-interactions between the two inputs. For more details on the implementation and
-training of the layer functions, see \cite{Bowman:Potts:Manning:2014}.
+The simpler version of the comparison simply concatenates the two
+input vectors before feeding them into a standard neural network (NN)
+layer.  The more powerful neural tensor network (NTN) version uses an
+additional third-order tensor parameter to allow for multiplicative
+interactions between the two inputs \cite{chen2013learning}. For more
+details on the implementation and training of the layer functions, see
+\cite{Bowman:Potts:Manning:2014}.
 
@@
 embeddings themselves from pretrained vectors and was found to help 
-peformance in that setting.
+performance in that setting.
 

### 2014-10-10 writing/AAAI SSS '15/wordnet.tex
Minor edits

@@
-\subsection*{Reasoning about lexical relations in WordNet}\label{sec:wordnet}
+\section{Reasoning about lexical relations in WordNet}\label{sec:wordnet}
 
@@
 \ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
-terms, which share a direct hypernym, like \ii{dalmation}, \ii{pug},
+terms, which share a direct hypernym, like \ii{dalmatian}, \ii{pug},
 and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate

### 2014-10-10 writing/AAAI SSS '15/join.tex
Minor edits
Misc clarifications.

@@
-\subsection*{Reasoning about semantic relations}\label{sec:join}
+\section{Reasoning about semantic relations}\label{sec:join}
 
@@
 
-If a model is to learn the behavior of a relational logic like the natural
-logic presented here over some domain from a finite amount of data, it 
-must learn to deduce new relational statements from seen ones in a sound manner.
-The simplest such deductions, and the only ones that can be made using relations
-between single terms, involve the set of inference rules schematized in 
+In this experiment, we test our model's ability to learn and use
+the natural logic inference rules schematized in 
 Figure~\ref{tab:jointable}. For instance, given that $a \natrev b$ and $b
@@
 for the classifier, though the performance of the model was not highly
-sensitive to either dimensionality parameter in any of the experiments
-discussed here. This model was able to correctly label 99.4\% (94.3
+dependant on either dimensionality setting. 
+This model was able to correctly label 99.4\% (94.3
 macroaveraged F1) of the provably derivable test examples, and 95.4\%

### 2014-10-10 writing/AAAI SSS '15/wordnet.tex
Misc clarifications.

@@
 \cite{fellbaum2010wordnet} come close and pose the same substantive
-challenges.
+challenges within a somewhat easier classification problem.
 
@@
 and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
-terms tend to exclude one another, thereby providing a reasonable
-approximation of exclusion. WordNet defines hypernymy and hyponymy over
+terms tend to exclude one another, thereby providing a loose approximation 
+of the natural logic exclusion relation $\natneg$. 
+WordNet defines hypernymy and hyponymy over
 sets of synonyms, rather than over individual terms, so we do not
 include a \ii{synonym} or \ii{equivalent} relation, but rather
-consider only one member of each set of synonyms.
+consider only one member of each set of synonyms. Word pairs which do
+not fall into these three relations are not included in the dataset.
 
@@
 \texttt{organism.n.01}. In order to balance the distribution of the classes, we slightly downsample instances 
-of the \ii{coordinate} relation, yielding 36,772 relations among 3,217 terms. We randomly select 15\% of these 
-relations as test data. Unlike in the previous experiment, it is not straightforward here to determine in advance
- how much training data should be required to learn an accurate model, so we performed training runs with 
- various fractions of the training portion of the 
-dataset. Embeddings were fixed at 25 dimensions and were initialized randomly or using distributional 
-vectors from GloVe \cite{pennington2014glove}. The feature vector produced by the comparison layer was fixed at 
-80 dimensions.
+of the \ii{coordinate} relation, yielding a total of 36,772 relations among 3,217 terms.
+We report results below using crossvalidation, choosing a disjoin 10\% test sample for each of five runs.
+Unlike in the previous experiment,
+it is not straightforward here to determine in advance
+how much data should be required to train an accurate model, so we performed training runs with 
+various fractions of the remaining data. Embeddings were fixed at 25 dimensions and were initialized 
+randomly or using distributional vectors from GloVe \cite{pennington2014glove}. The feature vector 
+produced by the comparison layer was fixed at 80 dimensions.
 
@@
      100\%   & 99.5 &(100.0) & 98.9 &(99.7)    & \textbf{99.6} &(100.0) & 99.2 &(100.0) & 37.9 &(--)\\
-     75\%    & 98.9 &(100.0) & 37.9 &(37.0)    & 98.6 &(100.0)          & 98.4 &(100.0) & 37.9 &(--)\\
-     33\%    & 96.0 &(100.0) & 37.0 &(37.0)    & 95.4 &(100.0)          & 95.3 &(100.0) & 37.9 &(--)\\ % TODO: Rerunning. Got stuck predicting the most frequent class.
-     16\%    & 91.8 &(100.0) & 91.9 &(100.0)   & 92.1 &(99.7)           & 92.2 &(100.0) & 37.9 &(--)\\
+     33\%    & 96.0 &(100.0) & 37.0 &(37.0)    & 95.4 &(100.0)          & 95.3 &(100.0) & 37.9 &(--)\\
+     11\%    & 91.8 &(100.0) & 91.9 &(100.0)   & 92.1 &(99.7)           & 92.2 &(100.0) & 37.9 &(--)\\
+     % 2\%    & 86.0 &(100.0) & ... &(100.0)   & 86.2 &(100.0)           & 76.4 &(87.3) & 37.9 &(--)\\ % Accidental run with slightly different train/test split.
     \bottomrule

### 2014-10-10 RunExperiments.m

@@ -178,4 +178,8 @@ echo "cd quant; mbs = 64; lambda = 0.0005; ed = 1; loadWords = 1; penult = 80, d
 echo "cd quant; mbs = 64; lambda = 0.0005; ed = 1; loadWords = 1; penult = 80, dp = .33, name=['otss-l', num2str(lambda), '-ed', num2str(ed), '-tr', num2str(loadWords), '-pen', num2str(penult), '-mbs', num2str(mbs), '-dp', num2str(dp)]; disp(name); Crossvalidate('synset-relations', '', name, mbs, 25, penult, 0.1, lambda, 1, ed, loadWords, dp);" | /afs/cs/software/bin/matlab_r2012b
 echo "cd quant; mbs = 64; lambda = 0.0005; ed = 1; loadWords = 1; penult = 80, dp = .11, name=['otss-l', num2str(lambda), '-ed', num2str(ed), '-tr', num2str(loadWords), '-pen', num2str(penult), '-mbs', num2str(mbs), '-dp', num2str(dp)]; disp(name); Crossvalidate('synset-relations', '', name, mbs, 25, penult, 0.1, lambda, 1, ed, loadWords, dp);" | /afs/cs/software/bin/matlab_r2012b
 
+# Join
+
+echo "cd quant; mbs = 64; lambda = 0.0005; ed = 1; penult = 80, name=['join-l', num2str(lambda), '-ed', num2str(ed), '-pen', num2str(penult), '-mbs', num2str(mbs)]; disp(name);  TrainJoinModel(name, mbs, dim, 1, lambda, 90, 0);" | /afs/cs/software/bin/matlab_r2012b
+
 

### 2014-10-12 writing/AAAI SSS '15/join.tex
Add final results to paper; related configs

@@
 dependant on either dimensionality setting. 
-This model was able to correctly label 99.4\% (94.3
-macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2 macro-F1) of the remaining test examples. The simpler NN worked
+Averaging over five randomly generated datasets, the model was able to correctly label 98.1\% (standard deviation $\sigma = 1.5\%$) of the provably derivable test examples, and 87.7\%
+($\sigma = 8.0\%$) of the remaining test examples. The simpler NN worked
 best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
-respectively. Training accuracy was 99.8\% for the NTN and 95.4\% for
+accuracies of only 84.8\% ($\sigma = 1.9\%$) and 68.7\% ($\sigma = 3.5\%$),
+respectively. Training accuracy was 99.8\% ($\sigma = 0.1\%$) for the NTN and 94.7\% ($\sigma = 2.0\%$) for
 the NN.

### 2014-10-12 writing/AAAI SSS '15/methods.tex
Add final results to paper; related configs

@@
 logical relations. The architecture of the model that we use, which is
-limited to only pairs of single symbols (such as words), is depicted
+limited to only pairs of single terms (such as words), is depicted
 in Figure~\ref{sample-figure}. The model represents the two input
-symbols as embeddings, which are fed into a comparison function based
+terms as embeddings, which are fed into a comparison function based
 on one of two types of neural network layer functions to produce a
-representation for the relationship between the two symbols. This
+representation for the relationship between the two terms. This
 representation is then fed into a softmax classifier, which outputs a
@@
 This model used here differs from the one described in that work in
-two ways. Because the inputs are single symbols, we do not use
+two ways. Because the inputs are single terms, we do not use
 a composition function here. Also, for our experiment on WordNet data, 

### 2014-10-12 writing/AAAI SSS '15/wordnet.tex
Add final results to paper; related configs

@@
 \paragraph{Results} 
-We find that the NTN performs almost perfectly even when trained only
-on a fraction of the ground truth relations. The plain NN is somewhat less
-successful. Neither model has reached perfect performance in any of
-our experiments, though we are continuing to tune both models.
+We find that the NTN performs perfectly with random initialization, and that the plain NN performs almost as well.
+We also find that initialization with GloVe is helpful in allowing the models to maintain fair performance
+with smaller amounts of training data. Some of the randomly initialized model runs failed to learn
+usable representations at all and labeled all examples with the most frequent labels.
+We excluded these runs from the statistics, but marked settings for which this occurred with the symbol $\dagger$.
+For all of the remaining runs, training accuracy was consistently above 99\%.
 
@@
     \midrule
-     100\%   & 99.5 &(100.0) & 98.9 &(99.7)    & \textbf{99.6} &(100.0) & 99.2 &(100.0) & 37.9 &(--)\\
-     33\%    & 96.0 &(100.0) & 37.0 &(37.0)    & 95.4 &(100.0)          & 95.3 &(100.0) & 37.9 &(--)\\
-     11\%    & 91.8 &(100.0) & 91.9 &(100.0)   & 92.1 &(99.7)           & 92.2 &(100.0) & 37.9 &(--)\\
+     100\%   & 99.7 &(0.0) & 99.4$^\dagger$ &(0.3)    & 99.6 &(0.0) & \textbf{100.0} &(0.1) & 37.9 &(--)\\
+     33\%    & 96.0 &(0.5) & 95.3 &(0.2)     & 95.8 &(0.8)          & 95.5$^\dagger$ &(0.5) & 37.9 &(--)\\
+     11\%    & 91.1 &(0.5) & 90.8$^\dagger$ &(0.3)    & 91.3 &(0.6)          & 90.9$^\dagger$ &(0.2) & 37.9 &(--)\\
      % 2\%    & 86.0 &(100.0) & ... &(100.0)   & 86.2 &(100.0)           & 76.4 &(87.3) & 37.9 &(--)\\ % Accidental run with slightly different train/test split.
@@
   \end{tabular}}
- \caption{Test (train) accuracy figures on the WordNet data. The baseline figure is simply the frequency of the most frequent class, \ii{coordinate}.\label{b-table}}  
+ \caption{Mean test \% accuracy scores (with standard deviations) on the WordNet data over five-fold crossvalidation.
+The baseline figure is simply the frequency of the most frequent class, \ii{coordinate}.\label{b-table}}
 \end{table}

### 2014-10-12 writing/AAAI SSS '15/abstract.tex
Adding results summary, cutting an Icard paper.
Edits

@@
   support the rich, diverse logical reasoning captured by natural
-  logic. We address this question a simple neural network-based
-  model for learning embeddings.  Our experiments evaluate the models' 
-  ability to learn the basic algebra of natural logic relations from 
-  simulated data and from the WordNet noun graph.  We find that the model
-  can learn to do inference over relations with near perfect precision
-  from moderate amounts of training data.
+  logic. We address this question using two neural network-based
+  models for learning embeddings: plain neural networks and neural
+  tensor networks. Our experiments evaluate the models' ability to
+  learn the basic algebra of natural logic relations from simulated
+  data and from the WordNet noun graph.  The overall positive results
+  are promising for the future of learned distributed representations in
+  the applied modeling of logical semantics.
 \end{abstract}

### 2014-10-12 writing/AAAI SSS '15/intro.tex
Edits

@@
 Distributed representations have been used successfully in a wide
-array of sophisticated language tasks \cite{collobert2011natural}. %,
+array of sophisticated language tasks (e.g., \cite{collobert2011natural}). %,
 % including sentiment analysis, analogy completion, relation
@@
 the WordNet noun graph. The simulated data help us to achieve analytic
-insights into what the models learn, and WordNet data show how they
-fare in the real world.  We find that only the NTN is able to fully
+insights into what the models learn, and the WordNet data show how they
+fare with a real natural language vocabulary.  We find that only the NTN is able to fully
 learn the underlying algebra, but that both models excel in the 

### 2014-10-12 writing/AAAI SSS '15/join.tex
Edits

@@
 these data
-sets% . In both cases, the models were implemented exactly as
+sets.% . In both cases, the models were implemented exactly as
 % described in Section~\ref{methods}, but since the items being compared
@@
 dependant on either dimensionality setting. 
-Averaging over five randomly generated datasets, the model was able to correctly label 98.1\% (standard deviation $\sigma = 1.5\%$) of the provably derivable test examples, and 87.7\%
+Averaging over five randomly generated data sets, the model was able to correctly label 98.1\% (standard deviation $\sigma = 1.5\%$) of the provably derivable test examples, and 87.7\%
 ($\sigma = 8.0\%$) of the remaining test examples. The simpler NN worked

### 2014-10-12 writing/AAAI SSS '15/methods.tex
Edits

@@
 
-We build an embedding-based model using the method of
+We build embedding-based models using the method of
 \cite{Bowman:Potts:Manning:2014}, which is centered on the task of
 labeling a pair of words or sentences with one of a small set of
-logical relations. The architecture of the model that we use, which is
-limited to only pairs of single terms (such as words), is depicted
-in Figure~\ref{sample-figure}. The model represents the two input
-terms as embeddings, which are fed into a comparison function based
-on one of two types of neural network layer functions to produce a
+logical relations. The architecture that we use, which is limited to
+only pairs of single terms (such as words), is depicted in
+Figure~\ref{sample-figure}. The model represents the two input terms
+as embeddings, which are fed into a comparison function based on one
+of two types of neural network layer functions to produce a
 representation for the relationship between the two terms. This
@@
 distribution over possible labels. The entire network, including the
-embeddings, are trained using backpropagation.
+embeddings, is trained using backpropagation.
 
-The simpler version of the comparison simply concatenates the two
+The simpler version of the comparison concatenates the two
 input vectors before feeding them into a standard neural network (NN)
@@
 This model used here differs from the one described in that work in
-two ways. Because the inputs are single terms, we do not use
-a composition function here. Also, for our experiment on WordNet data, 
+two ways. First, because the inputs are single terms, we do not use
+a composition function here. Second, for our experiment on WordNet data, 
 we introduce an additional neural network layer between the embedding 

### 2014-10-12 writing/AAAI SSS '15/wordnet.tex
Edits

@@
 We extract three types of relation from WordNet. \ii{Hypernym} and
-\ii{hyponym} can be are represented directly in the WordNet graph
+\ii{hyponym} can be represented directly in the WordNet graph
 structure, and correspond closely to the $\natrev$ and $\natfor$
@@
 terms tend to exclude one another, thereby providing a loose approximation 
-of the natural logic exclusion relation $\natneg$. 
+of the natural logic exclusion relation $\natalt$. 
 WordNet defines hypernymy and hyponymy over
@@
 consider only one member of each set of synonyms. Word pairs which do
-not fall into these three relations are not included in the dataset.
+not fall into these three relations are not included in the data set.
 
@@
 of the \ii{coordinate} relation, yielding a total of 36,772 relations among 3,217 terms.
-We report results below using crossvalidation, choosing a disjoin 10\% test sample for each of five runs.
+We report results below using crossvalidation, choosing a disjoint 10\% test sample for each of five runs.
 Unlike in the previous experiment,
@@
 \paragraph{Results} 
-We find that the NTN performs perfectly with random initialization, and that the plain NN performs almost as well.
-We also find that initialization with GloVe is helpful in allowing the models to maintain fair performance
+We find that the NTN performs perfectly with random initialization, and that the plain NN performs almost as well,
+a point of contrast with the results of Section~\ref{sec:join}. We also find that initialization with GloVe is helpful in allowing the models to maintain fair performance
 with smaller amounts of training data. Some of the randomly initialized model runs failed to learn
@@
 For all of the remaining runs, training accuracy was consistently above 99\%.
+
+\newcommand{\nodagger}{\phantom{$^\dagger$}}
 
@@
     \midrule
-     100\%   & 99.7 &(0.0) & 99.4$^\dagger$ &(0.3)    & 99.6 &(0.0) & \textbf{100.0} &(0.1) & 37.9 &(--)\\
-     33\%    & 96.0 &(0.5) & 95.3 &(0.2)     & 95.8 &(0.8)          & 95.5$^\dagger$ &(0.5) & 37.9 &(--)\\
+     100\%   & 99.7 &(0.0) & 99.4$^\dagger$ &(0.3)    & 99.6 &(0.0) & \textbf{100.0}\nodagger &(0.1) & 37.9 &(--)\\
+     33\%    & 96.0 &(0.5) & 95.3\nodagger &(0.2)     & 95.8 &(0.8)          & 95.5$^\dagger$ &(0.5) & 37.9 &(--)\\
      11\%    & 91.1 &(0.5) & 90.8$^\dagger$ &(0.3)    & 91.3 &(0.6)          & 90.9$^\dagger$ &(0.2) & 37.9 &(--)\\
@@
  \caption{Mean test \% accuracy scores (with standard deviations) on the WordNet data over five-fold crossvalidation.
-The baseline figure is simply the frequency of the most frequent class, \ii{coordinate}.\label{b-table}}
+The baseline figure is the frequency of the most frequent class, \ii{coordinate}.\label{wordnet-table}}
 \end{table}

### 2014-10-12 writing/AAAI SSS '15/join.tex
Tweaks to stats reporting.

--- 2014-10-10 writing/AAAI SSS '15/join.tex
@@
 dependant on either dimensionality setting. 
-This model was able to correctly label 99.4\% (94.3
-macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2 macro-F1) of the remaining test examples. The simpler NN worked
+Averaging over five randomly generated datasets, the model was able to correctly label 98.1\% (standard error $0.67\%$) of the provably derivable test examples, and 87.7\%
+($SE = 3.59\%$) of the remaining test examples. The simpler NN worked
 best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
-respectively. Training accuracy was 99.8\% for the NTN and 95.4\% for
+accuracies of only 84.8\% ($SE = 0.84\%$) and 68.7\% ($SE = 1.58\%$),
+respectively. Training accuracy was 99.8\% ($SE = 0.04\%$) for the NTN and 94.7\% ($SE= 0.89\%$) for
 the NN.

### 2014-10-12 writing/AAAI SSS '15/join.tex
Post-merge cleanup.

@@
 dependant on either dimensionality setting. 
-Averaging over five randomly generated datasets, the model was able to correctly label 98.1\% (standard error $0.67\%$) of the provably derivable test examples, and 87.7\%
+Averaging over five randomly generated data sets, the model was able to correctly label 98.1\% (standard error $0.67\%$) of the provably derivable test examples, and 87.7\%
 ($SE = 3.59\%$) of the remaining test examples. The simpler NN worked

### 2014-10-12 writing/AAAI SSS '15/wordnet.tex
Tweaks to stats reporting.
Post-merge cleanup.

@@
     \midrule
-     100\%   & 99.7 &(0.0) & 99.4$^\dagger$ &(0.3)    & 99.6 &(0.0) & \textbf{100.0}\nodagger &(0.1) & 37.9 &(--)\\
-     33\%    & 96.0 &(0.5) & 95.3\nodagger &(0.2)     & 95.8 &(0.8)          & 95.5$^\dagger$ &(0.5) & 37.9 &(--)\\
-     11\%    & 91.1 &(0.5) & 90.8$^\dagger$ &(0.3)    & 91.3 &(0.6)          & 90.9$^\dagger$ &(0.2) & 37.9 &(--)\\
+     100\%   & 99.73 &(0.04) & 99.37$^\dagger$ &(0.14)    & 99.61 &(0.02) & \textbf{99.95}\nodagger &(0.03) & 37.05 &(--)\\
+     33\%    & 95.96 &(0.20) & 95.30\nodagger &(0.12)                & 95.83 &(0.35)          & 95.45$^\dagger$ &(0.31) & 37.05 &(--)\\
+     11\%    & 91.11 &(0.24) & 90.81$^\dagger$ &(0.20)    & 91.27 &(0.27)          & 90.90$^\dagger$ &(0.13) & 37.05 &(--)\\
      % 2\%    & 86.0 &(100.0) & ... &(100.0)   & 86.2 &(100.0)           & 76.4 &(87.3) & 37.9 &(--)\\ % Accidental run with slightly different train/test split.
@@
   \end{tabular}}
- \caption{Mean test \% accuracy scores (with standard deviations) on the WordNet data over five-fold crossvalidation.
-The baseline figure is the frequency of the most frequent class, \ii{coordinate}.\label{wordnet-table}}
+ \caption{Mean test \% accuracy scores (with standard error) on the WordNet data over five-fold crossvalidation.
+The baseline figure is the frequency of the most frequent class, \ii{hypernym}.\label{b-table}}
 \end{table}

### 2014-10-12 writing/AAAI SSS '15/intro.tex
Ref fix and figure positioning.

@@
 % - Something McCallum/Universal Schema?
+
+\begin{table}[tp]
+  \centering\small
+  \setlength{\tabcolsep}{15pt}
+  \renewcommand{\arraystretch}{1.1}
+  \begin{tabular}{l c l l} 
+    \toprule
+    Name & Symbol & Set-theoretic definition & Example \\ 
+    \midrule
+    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
+    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
+    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
+    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
+    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
+    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
+    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
+    \bottomrule
+  \end{tabular}
+  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
+    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
+    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
+    %is insufficient knowledge to choose a label.}
+  \label{b-table}
+\end{table}

### 2014-10-12 writing/AAAI SSS '15/join.tex
Ref fix and figure positioning.

@@
 \section{Reasoning about semantic relations}\label{sec:join}
-
-\begin{table}[tp]
-  \centering\small
-  \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{l c l l} 
-    \toprule
-    Name & Symbol & Set-theoretic definition & Example \\ 
-    \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
-    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
-    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
-    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
-    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
-    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
-    \bottomrule
-  \end{tabular}
-  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
-    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
-    %is insufficient knowledge to choose a label.}
-  \label{b-table}
-\end{table}
 

### 2014-10-13 writing/AAAI SSS '15/join.tex
Corrected math mode typing

@@
 Averaging over five randomly generated data sets, the model was able to correctly label 98.1\% (standard error $0.67\%$) of the provably derivable test examples, and 87.7\%
-($SE = 3.59\%$) of the remaining test examples. The simpler NN worked
+($\textit{SE} = 3.59\%$) of the remaining test examples. The simpler NN worked
 best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 84.8\% ($SE = 0.84\%$) and 68.7\% ($SE = 1.58\%$),
-respectively. Training accuracy was 99.8\% ($SE = 0.04\%$) for the NTN and 94.7\% ($SE= 0.89\%$) for
+accuracies of only 84.8\% ($\textit{SE} = 0.84\%$) and 68.7\% ($\textit{SE} = 1.58\%$),
+respectively. Training accuracy was 99.8\% ($\textit{SE} = 0.04\%$) for the NTN and 94.7\% ($\textit{SE} = 0.89\%$) for
 the NN.

### 2014-10-13 writing/AAAI SSS '15/intro.tex
Tweaks related to model count

@@
 such representations to support the rich, diverse logical reasoning
-captured by natural logic; while they excel at
-synonymy (similarity), the results are more mixed for entailment,
-contradiction, and mutual consistency.  Using the natural logic of
-\cite{maccartney2009extended} as our formal model, we address this
-open question for a simple neural network-based classification model. 
-The natural logic is built from the seven relations defined in
-Table~\ref{b-table}. Its formal properties are now well-understood
-\cite{Icard:Moss:2013:LILT}, so it provides a rigorous
-set of goals for our neural models. To keep the discussion manageable,
-we limit attention to experiments involving the lexicon; for a more
-extended treatment of complex expressions involving logical
-connectives and quantifiers, see \citet{Bowman:Potts:Manning:2014}.
+captured by natural logic; while they excel at synonymy (similarity),
+the results are more mixed for entailment, contradiction, and mutual
+consistency.  Using the natural logic of \cite{maccartney2009extended}
+as our formal model, we address this open question using two neural
+network-based models for learning embeddings: plain neural networks
+and neural tensor networks (NTNs).  The natural logic is built from
+the seven relations defined in Table~\ref{b-table}. Its formal
+properties are now well-understood \cite{Icard:Moss:2013:LILT}, so it
+provides a rigorous set of goals for our neural models. To keep the
+discussion manageable, we limit attention to experiments involving the
+lexicon; for a more extended treatment of complex expressions
+involving logical connectives and quantifiers, see
+\citet{Bowman:Potts:Manning:2014}.
 

### 2014-10-20 writing/AAAI SSS '15/intro.tex
Misc fixes around fragmented training.

@@
     \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
-    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
-    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
-    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
-    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
-    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
+    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle $\natfor$ reptile}  \\ 
+    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile $\natrev$ turtle}  \\ 
+    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch $\nateq$ sofa} \\ 
+    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle $\natalt$ warthog} \\ 
+    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able $\natneg$ unable} \\
+    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal $\natcov$ non-turtle} \\ 
+    independence       & $x \natind y$   & (else) & \ii{turtle $\natind$ pet}\\
     \bottomrule

### 2014-10-20 config/Sick.m

@@ -1,4 +1,4 @@
-function [ hyperParams, options, wordMap, relationMap ] = Sick(dataflag, transDepth, penult, lambda, tot, mbs, lr, trainwords, frag)
+function [ hyperParams, options, wordMap, relationMap ] = Sick(dataflag, transDepth, penult, lambda, tot, mbs, lr, trainwords, frag, loadwords)
 % Configuration for experiments involving the SemEval SICK challenge and ImageFlickr 30k. 
 
 [hyperParams, options] = Defaults();
@@ -23,7 +23,7 @@ hyperParams.lambda = lambda; % 0.002 works?;
 hyperParams.useThirdOrder = tot;
 hyperParams.useThirdOrderComparison = tot;
 
-hyperParams.loadWords = true;
+hyperParams.loadWords = loadwords;
 hyperParams.trainWords = trainwords;
 
 % How many examples to run before taking a parameter update step on the accumulated gradients.
@@ -75,14 +75,14 @@ elseif findstr(dataflag, 'sick-plus')
     hyperParams.fragmentData = true;
 elseif strcmp(dataflag, 'imageflickr')
     % The number of relations.
-    hyperParams.numRelations = 2; 
+    hyperParams.numRelations = 4; 
 
-    hyperParams.relations = {{'ENTAILMENT', 'NONENTAILMENT'}};
+    hyperParams.relations = {{'ENTAILMENT', 'na', 'na2', 'NONENTAILMENT'}};
 	relationMap = cell(1, 1);
 	relationMap{1} = containers.Map(hyperParams.relations{1}, 1:length(hyperParams.relations{1}));
 
     wordMap = InitializeMaps('sick_data/flickr_words_t4.txt');
-    hyperParams.vocabName = 'spt4-2cl';
+    hyperParams.vocabName = 'spt4b';
 
     hyperParams.trainFilenames = {'/scr/nlp/data/ImageFlickrEntailments/clean_parsed_entailment_pairs.tsv'};
     hyperParams.testFilenames = {'/scr/nlp/data/ImageFlickrEntailments/clean_parsed_entailment_pairs_first500.tsv', ...

### 2014-11-23 writing/AAAI SSS '15/intro.tex
Paper first push; init; naming; new data; PBS

@@
 
+% TODO: Better Gref citation; cite Oriol/Ilya
+
 % In our first experiment, we use the pre-specified logical grammar to
@@
   \end{tabular}
-  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
+  \protect\caption{\protect\label{b-table}The seven natural logic relations of \protect\cite{maccartney2009extended}. 
     $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
+    and the relation $\natind$ applies whenever none of the other six do.55} %, including when there 
     %is insufficient knowledge to choose a label.}
-  \label{b-table}
 \end{table}

### 2014-11-23 writing/F'14 paper/abstract.tex
Paper first push; init; naming; new data; PBS

\begin{abstract}
Recursive neural network models for sentence meaning have been successful in many tasks, but it remains an open question whether any such model can learn compositional semantic representations that support logical deduction. We pursue this question directly by evaluating whether two such models can correctly learn to identify logical relationships such as entailment and contradiction between sentences. We first describe a recursive neural tensor network model for entailment and demonstrate it on the SICK challenge. We then use artificial data from a logical grammar to test the model's ability to learn to handle basic relational reasoning, recursive functions, and quantification. The model performs competitively on the SICK data and generalizes well in all three experiments on simulated data, suggesting that it can learn suitable representations for logical inference in natural language.
\end{abstract}

% TODO: Is there anything I should add that would be more specific than 'many tasks'? 'Tasks like paraphrase detection'?

### 2014-11-23 writing/F'14 paper/discussion.tex
Paper first push; init; naming; new data; PBS

--- 2014-08-25 writing/BayLearn abstract/paper.tex
@@
-\documentclass{article} % For LaTeX2e
-\usepackage{nips14submit_e,times}
-\usepackage{hyperref}
-\usepackage{url}
-\usepackage[leqno, fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{qtree}
-\usepackage[numbers]{natbib}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{colortbl}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{xcolor}
-
-\definecolor{mylinkcolor}{rgb}{0,0,0} % black
-\hypersetup{colorlinks, linkcolor=mylinkcolor, urlcolor=mylinkcolor, citecolor=mylinkcolor}
-
-\newcommand{\nateq}{\equiv}
-\newcommand{\natind}{\mathbin{\#}}
-%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
-\newcommand{\natneg}{\mathbin{^{\wedge}}}
-\newcommand{\natfor}{\sqsubset}
-\newcommand{\natrev}{\sqsupset}
-\newcommand{\natalt}{\mathbin{|}}
-\newcommand{\natcov}{\mathbin{\smallsmile}}
-
-\newcommand{\plneg}{\mathop{\textit{not}}}
-\newcommand{\pland}{\mathbin{\textit{and}}}
-\newcommand{\plor}{\mathbin{\textit{or}}}
-
-% Strikeout
-\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
-\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}
-
-\newcommand{\True}{\texttt{T}}
-\newcommand{\False}{\texttt{F}}
-\usepackage{stmaryrd}
-\newcommand{\sem}[1]{\ensuremath{\llbracket#1\rrbracket}}
-
-\renewcommand{\bibsection}{\subsubsection*{References}}
-
-\usepackage{gb4e}
-
-\def\ii#1{\textit{#1}}
-
-\newcommand{\mynote}[1]{{\color{red}\framebox{\begin{tabular}{p{0.9\textwidth}}\footnotesize#1 \end{tabular}}}}
-
-\title{Recursive Neural Networks for Learning Logical Semantics}
-
-\author{
-Samuel R.\ Bowman$^{\ast\dag}$ \\
-\texttt{sbowman@stanford.edu} \\[2ex]
-$^{\ast}$Stanford Linguistics \\
-\And
-Christopher Potts$^{\ast}$\\
-\texttt{cgpotts@stanford.edu} \\[2ex]
-$^{\dag}$Stanford NLP Group
-\And
-Christopher D.\ Manning$^{\ast\dag\ddag}$\\
-\texttt{manning@stanford.edu}\\[2ex]
-$^{\ddag}$Stanford Computer Science
-}
-
-% \author{
-% Samuel R.\ Bowman \\
-% NLP Group, Dept.\ of Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{sbowman@stanford.edu}
-%  \And
-%  Christopher Potts \\
-% Dept.\ of Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{cgpotts@stanford.edu}
-%  \And
-% Christopher D.\ Manning \\
-% NLP Group,  Depts.\ of Computer Science and Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{manning@stanford.edu}
-% }
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-
-\nipsfinalcopy % Uncomment for camera-ready version
-
-\begin{document}
-\maketitle
-
-  Supervised recursive neural network models (RNNs) for sentence
-  meaning have been successful in an array of sophisticated language
-  tasks, but it remains an open question whether they can learn
-  compositional semantic grammars that support logical deduction.  We
-  address this question directly by for the first time evaluating
-  whether each of two classes of neural model --- plain RNNs and
-  recursive neural tensor networks (RNTNs) --- can correctly learn
-  relationships such as entailment and contradiction between pairs of
-  sentences, where we have generated controlled data sets of sentences
-  from a logical grammar.  Our first experiment evaluates whether
-  these models can learn the basic algebra of logical relations
-  involved. Our second and third experiments extend this evaluation to
-  complex recursive structures and sentences involving quantification.
-  We find that the plain RNN achieves only mixed results on all three
-  experiments, whereas the stronger RNTN model generalizes well in
-  every setting and appears capable of learning suitable
-  representations for natural language logical inference.
-
-  \section{Introduction}\label{sec:intro}
-
-Supervised recursive neural network models (RNNs) for sentence meaning
-have been successful in an array of sophisticated language tasks,
-including sentiment analysis \cite{socher2011semi,socher2013acl1},
-image description \cite{sochergrounded}, and paraphrase detection
-\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
-the ability of these models to learn compositional semantic grammars,
-but it remains an open question whether they can achieve the same
-results as grammars based in logical forms
-\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
-when it comes to the core semantic concepts of quantification,
-entailment, and contradiction needed to identify the relationships
-between sentences like \ii{every animal walks}, \ii{every turtle
-  moves}, and \ii{most reptiles don't move}. To date, experimental
-investigations of these concepts using distributed representations
-have been largely confined to short phrases
-\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
-For robust natural language understanding, it is essential to model
-these phenomena in their full generality on complex linguistic
-structures.
-
-We address this question in the context of \ii{natural language
-  inference} (also known as \ii{recognizing textual entailment};
-\cite{dagan2006pascal}), in which the goal is to determine the core
-inferential relationship between two sentences. Much of the
-theoretical work on this task (and some successful implemented models
-\cite{maccartney2009extended,watanabe2012latent}) involves \ii{natural
-  logics}, which are formal systems that define rules of inference
-between natural language words, phrases, and sentences without the
-need of intermediate representations in an artificial logical
-language. Following \cite{bowman2013can}, we use the natural logic
-developed by \cite{maccartney2009extended} as our formal model. This
-logic defines seven core relations of synonymy, entailment,
-contradiction, and mutual consistency, as summarized in
-Table~\ref{b-table}, and it provides rules of semantic combination for
-projecting these relations from the lexicon up to complex phrases. The
-formal properties and inferential strength of this system are now
-well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
-
-In our experiments, we use this pre-specified logical grammar to
-generate controlled data sets encoding the semantic relationships
-between pairs of expressions and evaluate whether each of two classes
-of neural model --- plain RNNs and recursive neural tensor networks
-(RNTNs, \cite{socher2013acl1}) --- can learn those relationships
-correctly. In our first experiment (Section~\ref{sec:join}), we
-evaluate the ability of these models to learn the core relational
-algebra of natural logic from data. Our second experiment
-(Section~\ref{sec:recursion}) extends this evaluation to relations
-between complex recursive structures like $(a \plor b)$ and
-$\plneg(\plneg a \pland \plneg b)$, and our third experiment
-(Section~\ref{sec:quantifiers}) involves relations between quantified
-statements like \ii{every reptile walks} and \ii{no turtle moves}.
-We find that the plain RNN achieves only mixed results in all three
-experiments, whereas the stronger RNTN model generalizes well in every
-case, suggesting that it has in fact learned, or at least learned to
-simulate, our target logical
-concepts. These experiments differentiate the increased power of RNTNs
-better than previous work and provide the most convincing
-demonstration to date of the ability of neural networks to model
-semantic inferences in complex natural language sentences.
-
-\begin{table}[tp]
-  \centering
-  \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{l c l l} 
-    \toprule
-    Name & Symbol & Set-theoretic definition & Example \\ 
-    \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
-    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
-    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
-    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
-    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
-    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
-    \bottomrule
-  \end{tabular}
-  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
-    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
-    %is insufficient knowledge to choose a label.}
-  \label{b-table}
-\end{table}
-
-% Citations to additional past work to be added.\\...\\...\\...\\...\\...
-
-% Deep learning methods in NLP which learn vector representations for words have seen successful uses in recent years on increasingly sophisticated tasks \cite{collobert2011natural, socher2011semi, socher2013acl1, chen2013learning}.
-Given the still modest performance of semantically rich NLP systems in many domains---question answering and machine translation, for instance---it is worth exploring the degree to which learned vectors can serve as general purpose semantic representations.
-Much of the work to date analyzing vector representations for words (see \cite{baroni2013frege}) has focused on lexical semantic behaviors---like the similarity between words like \ii{Paris} and \ii{France}.
-Good similarity functions are valuable for many NLP tasks, but there are real use cases for which it is necessary to know how words relate to one another or to some extrinsic representation, and to model the ways in which word meanings combine to form phrase, sentence, or document meanings.
-This paper explores the ability of linguistic representations developed using supervised deep learning techniques to support interpretation and reasoning.
-
-% There are two broad family of tasks that could be used to test the ability of a model to develop general purpose meaning representations.
-In an interpretation task, sentences are mapped onto some denotation, such as  \ii{true} or \ii{false} for statements, or a factual answer for questions.
-There has been preliminary work in developing distributed models for interpretation \cite{grefenstette2013towards, rocktaschellow}, but developing a representation of world knowledge that corresponds accurately to the content expressed in language introduces considerable design challenges.
-I approach the problem by way of an inference task instead.
-Inferring the truth of one sentence from another does not require any preexisting knowledge representations, but nonetheless requires a precise representation of sentence meaning.
-I borrow the structure of the task from MacCartney and Manning  \cite{maccartney2009extended}.
-In it, the model is presented with a pair of sentences, and made to label the logical relation between the sentences as equivalence, entailment, or any of five other classes, as here:
-
-%\begin{quote}
-%\begin{enumerate}\small
-%\item Many smartphone users avoid high bills overseas by disabling data service.
-%\item Not everyone uses their smartphones for email when traveling abroad.
-%\end{enumerate} 
-%$\Rightarrow$ Entailment
-%\end{quote}
-
-%In this paper, we test the ability of recursive models to on three simple tasks, each of which is meant to capture a property that is necessary in representing natural language meaning in the setting of inference.
-I begin with an overview of MacCartney and Manning's \cite{maccartney2009extended} framework for inference, and of the recursive neural networks that I study.
-by showing that these models can learn to correctly represent entailment representations between sentences.
-I then show that these models can capture the meanings of recursive structures accurately up to a sufficient depth.
-I finally close with a brief demonstration of the ability of these models to reason over short natural language sentences involving quantifiers.
-
-% \subsection{Natural language inference and natural logic}
-
-% In its simplest form, \ii{natural language inference} (also known as \ii{recognizing textual entailment}, as in \cite{dagan2006pascal}) is the task of determine whether one sentence entails another.
-Much of the theoretical work on this task (and some successful implemented models \cite{maccartney2009natural, watanabe2012latent}) involve \ii{natural logic}, formal systems that define sound rules of inference from one sentence of natural language to another without the need for intermediate representations in some other logic.
-The most powerful model that we are aware of for natural logic is due to MacCartney and Manning \cite{maccartney2009extended} and Icard \cite{icard2012inclusion}, and is centered around the definition of a set of seven logical relations which can hold between sentences, shown in Table \ref{b-table}.
-
-% This approach to natural logic can capture much of the complexity of natural language meaning within a well understood framework, and is also fairly straightforward to implement in a machine learning setting since it can be reduced to a seven-way classification problem on sentence pairs.
-Our goal in this paper is to learn recursive neural network models which are able to mimic key behaviors of this system.
-
-% \begin{table}
-% \begin{center}
-% \begin{tabular}{|c|c|c|c|} \hline
-% name & symbol & set-theoretic definition & example \\ \hline \hline
-% entailment & $x \sqsubset y$ & $x \subset y$ & \ii{crow, bird}  \\ \hline
-% reverse entailment & $x \sqsupset y$ & $x \supset y$ & \ii{Asian, Thai}  \\ \hline
-% equivalence & $x \equiv y$ & $x = y$ & \ii{couch, sofa} \\ \hline
-% alteration & $x$ $|$ $y$ & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{cat, dog} \\ \hline
-% negation & $x \natneg y$ & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{able, unable} \\ \hline
-% cover & $x \smallsmile y$ & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-ape} \\ \hline
-% independence & $x$ \# $y$ & (else) & \ii{hungry, hippo}\\ \hline
-% \end{tabular}
-% \caption{The entailment relations in  $\mathfrak{B}$.
-$\mathcal{D}$ is the universe of possible objects of the same type as those being compared, and the relation \# applies whenever none of the other six do, including when there is insufficient knowledge to choose a label.}
-% \label{b-table}
-% \end{center}
-% \end{table}
-
-% The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes.
-It should be noted that this is not a balanced classification problem.
-If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them.
-% TODO: How much should we say about this?
-
-% The natural logic engine at the core of MacCartney and Manning's system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures.
-These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
-%for dozens of different cases of insertion, deletion and substitution of different types of lexical item.
-For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences.
-Here, the substitution of \ii{no} for \ii{some}  yields $\natneg$ and the substitution of \ii{dogs} for \ii{animals} yields $\sqsupset$.
-It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
-%\begin{quote}
-
-%1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
-%2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
-%3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}
-
-%\end{quote}
-
-% Work to date on inference in neural network models is quite limited.
-% \citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
-% In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations.
-They propose a deterministic inference engine (similar to MacCartney's) which is augmented by a probabilistic component that evaluates individual lexical substitution steps in the derivation using vector representations, though again these representations are not learned, and no evaluations of this system have been published to date.
-% \label{sec2}
-
-\section{Recursive neural network models} \label{methods}
-
-We study neural models that adhere to the linguistic \ii{principle of
- compositionality}, which says that the meanings for complex
-expressions are derived from the meanings of their constituent parts
-via specific composition functions \cite{Partee84,Janssen97}. In our
-distributed setting, word meanings are vectors of length $n$. The
-composition function maps pairs of them to single vectors of length $n$, 
-which can then be merged again to represent more complex
-phrases. Once the entire sentence-level representation has been
-derived, it serves as a fixed-length input for some subsequent function.
-
-We use the model architecture proposed in \cite{bowman2013can} and
-depicted in Figure~\ref{sample-figure}. The two phrases being compared
-are built up separately on each side of the tree, using the same
-composition function, until they have each been reduced to single
-vectors. The resulting vectors are fed into a separate comparison
-layer that is meant to generate a feature vector capturing the
-relation between the two phrases. The output of this layer is then
-given to a softmax classifier, which in turn produces a hypothesized
-distribution over the seven relations represented in Table~\ref{b-table}.
-
-For a composition layer, we evaluate models with both the plain neural
-network layer function \eqref{rnn} and the RNTN layer function
-\eqref{rntn} proposed in \citet{chen2013learning}. A sigmoid
-nonlinearity (element-wise $\tanh$) is applied to the output of either
-layer function, following \cite{socher2013acl1}.
-%
-\begin{gather} \label{rnn}
-\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
-\label{rntn}
-\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
-\end{gather} 
-%
-Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
-representations for the left and right children of the node, and
-$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
-them by an $n \times 2n$ matrix of learned weights, and applies the
-element-wise non-linearity to the resulting vector. The RNTN has the
-same basic structure, but with the addition of a learned third-order
-tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
-multiplicative interactions between the child vectors. Both models
-include a bias vector~$\vec{b}$.
-
-The comparison layer uses the same kind of function template as the
-composition layers (either an NN layer or an NTN layer) with
-independently learned parameters and a separate nonlinearity function.
-Rather than use a $\tanh$ nonlinearity here, we follow \cite{bowman2013can} 
-in using a rectified linear function. In
-particular, we use the leaky rectified linear function
-\cite{maasrectifier}: $f(\vec{x})=\max(\vec{x}, 0) +
-0.01\min(\vec{x}, 0)$, applied element-wise.
-
-To run the model forward and label a pair of phrases, the structure of
-the lower layers of the network is assembled so as to mirror the tree
-structures provided for each phrase. The word vectors are then looked
-up from the vocabulary matrix $V$ (one of the model parameters), and
-the composition and comparison functions are used to pass information
-up the tree and into the classifier at the top. For an objective
-function, we use the negative log of the probability assigned to the
-correct label.
-
-% Removed 'minibatch' -> For two of the experiments, we use minibatches
-% of size 1, which doesn't really count.
-We train the model using stochastic gradient descent (SGD)
-with learning rates computed using AdaGrad \cite{duchi2011adaptive}.
-The parameters (including the vocabulary) are initialized randomly 
-using a uniform distribution over $[-0.1, 0.1]$.
-Because the seven classes are not balanced in general, we report performance
-using both accuracy and macroaveraged F1, since the latter emphasizes
- performance on infrequent classes. We compute macroaveraged F1 
-as the harmonic mean of mean precision and mean recall, both computed
-for all classes for which there is test data, setting precision to 0 
-where it is not defined. Source code and generated data will be released
-after the review period.
-
-%\begin{figure}[tp]
-%  \centering
-% % \includegraphics[scale=0.35]{model.eps}
-%  \input{figure1}
-%  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
-%    The same structure is used for both the RNN and RNTN layer functions.} 
-%  \label{sample-figure}
-%\end{figure}
-
-%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?
-
-\section{Reasoning about semantic relations}\label{sec:join}
-
-If a model is to learn the behavior of a relational logic like the one
-presented here from a finite amount of data, it must learn to deduce new
-relations from seen relations in a sound manner. The simplest such
-deductions involve atomic statements using the relations in
-Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
-\natrev c$, one can conclude that $a \natrev c$, by basic
-set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
-full set of sound inferences of this form is summarized in
-Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
-of relations for which no valid inference can be drawn in our logic.
-
-% about the relations themselves that do not depend on the
-% internal structure of the things being compared. For example, given
-% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
-% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven relations support more
-% than just transitivity: MacCartney and Manning's
-% \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of relations of the form $a R
-% b$ and $b R' c$, including several less intuitive ones such as that if
-% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-%
-%\begin{table}[htp]
-%  \centering  
-%  \setlength{\arraycolsep}{8pt}
-%  \renewcommand{\arraystretch}{1.1}
-%  \newcommand{\UNK}{\cdot}  
-%  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
-%    %\hline
-%    \multicolumn{1}{c}{}
-%             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
-%    \cline{2-8}
-%    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
-%    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
-%    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
-%    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
-%    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
-%    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
-%    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
-%    \cline{2-8}
-%  \end{array}$
-%  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.
-These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
-We assess our models' ability to reproduce such inferential paths.}
-%  \label{tab:jointable}
-%\end{table}
-
-\paragraph{Experiments}
-To test our models' ability to learn these inferential patterns, we
-create small boolean structures for our logic in which terms denote
-sets of entities from a small domain.  Figure~\ref{lattice-figure}
-depicts a structure of this form. The lattice gives the full model,
-for which all the statements on the right are valid. We divide these
-statements evenly into training and test sets, and remove from the
-test set those statements which cannot be proven from the training
-statements.
-% using the logic depicted in Figure~\ref{lattice-figure}.
-
-% We test the model's ability to learn this behavior by creating
-% artificial data sets of terms which represent sets of numbers. Since
-% MacCartney and Manning's set of relations hold between sets as well as
-% between sentences, we can use the underlying set structure to generate
-% the all of the relations that hold between any pair of these terms, as
-% in Figure \ref{lattice-figure}. We train the model defined above on a
-% subset of these relations, but rather then presenting the model with a
-% pair of tree-structured sentences as inputs, simply present it with
-% two single terms, each of which corresponds to a single vector in the
-% (randomly initialized) vocabulary matrix $V$, ensuring that the model
-% has no information about the terms being compared except the relations
-% between them.
-
-In our experiments, we create 80 randomly generated sets drawing from
-a domain of seven elements. This yields a data set consisting of
-6400 statements about pairs of formulae. 3200 of these pairs are
-chosen as a test set, and that test set is further reduced to the 2960
-examples that can be provably derived from the training data.
-
-We trained versions of both the RNN model and the RNTN model on these
-data sets. In both cases, the models were implemented exactly as
-described in Section~\ref{methods}, but since the items being compared
-are single terms rather than full tree structures, the composition
-layer was not involved, and the two models differed only in which
-layer function was used for the comparison layer. We simply present
-the models with two single terms, each of which corresponds to a
-single vector in the (randomly initialized) vocabulary matrix $V$,
-thereby ensuring that the model has no information about the terms
-being compared except the relations between them.
-
-\paragraph{Results} 
-We found that the RNTN model worked best with 11-dimensional vector
-representations for the 80 sets and a 90-dimensional feature vector
-for the classifier, though the performance of the model was not highly
-sensitive to either dimensionality parameter in any of the experiments
-discussed here. This model was able to correctly label 99.4\% (94.3
-macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
-best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
-respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
-the RNN.
-
-\begin{figure}[tp]
-  \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
-    \setlength{\unitlength}{1cm}
-    \begin{picture}(5,5.5)
-      \labelednode{2.50}{5}{}{1,2,3}
-
-      \put(0.75,4){\line(3,1){1.5}}
-      \put(2.5,4){\line(0,1){0.5}}
-      \put(4.25,4){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{3.5}{a,b}{1,2}
-      \labelednode{2.50}{3.5}{c}{1,3}
-      \labelednode{4.25}{3.5}{d}{2,3}
-
-      \put(0.75,2.5){\line(0,1){0.5}}
-      \put(0.75,2.5){\line(3,1){1.5}}
-
-      \put(2.5,2.5){\line(-3,1){1.5}}
-      \put(2.5,2.5){\line(3,1){1.5}}
-
-      \put(4.25,2.5){\line(0,1){0.5}}
-      \put(4.25,2.5){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{2}{e,f}{1}
-      \labelednode{2.50}{2}{}{2}
-      \labelednode{4.25}{2}{g,h}{3}
-
-      \put(2.5,1){\line(-3,1){1.5}}
-      \put(2.5,1){\line(0,1){0.5}}
-      \put(2.5,1){\line(3,1){1.5}}
-
-      \labelednode{2.5}{0.5}{}{}
-    \end{picture}
-    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
-    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}[t]{0.43\textwidth}
-    \centering
-    \setlength{\tabcolsep}{12pt}
-    \begin{tabular}[b]{c  c}
-      \toprule
-      Train & Test \\
-      \midrule
-                    & $b \nateq b$ \\
-      $b \natcov c$ &               \\
-                    & $b \natcov d$ \\
-                    & \strikeout{$b \natrev e$} \\
-      $c \natcov d$ &               \\
-      $c \natrev e$ &               \\
-                    & \strikeout{$c \nateq f$} \\
-      $c \natrev g$ &               \\ 
-                    & $e \natfor b$ \\
-      $e \natfor c$ &               \\[-1ex]
-      $\vdots$      & $\vdots$ \\
-      \bottomrule
-    \end{tabular}
-
-    \caption{A train/test split of the atomic statements about the
-      model.  Test statements not provable from the training data are
-      crossed out.}
-  \end{subfigure}  
-  \caption{Small example structure and data for learning relation composition.}
-  \label{lattice-figure}
-\end{figure} 
-% Note: None of these test examples is derivable from the shown training data, 
-% but we suggest that there is additional training data, so we can cross lines
-% out willy-nilly without fear.
-
-% RNTN log: tue-j-11-6x80-hip.txt
-% Train PER: 0.0021875
-
-% RNN log: tue-j-11-6x80-r-75.txt
-% Train PER: 0.047188
-
-% and create a dataset consisting of the relations between every pair of
-% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
-% test dataset, and that test dataset was further split into the 2960
-% examples that can be provably derived from the test data using
-% MacCartney and Manning's join table (or by the symmetry of the
-% relations in about half of the cases) and the 240 that
-% cannot.
-
-% We tested a version of both the RNN model and the RNTN model on these
-% data. In both cases, the models were implemented exactly as described
-% in Section~\ref{methods}, but since the items being compared are
-% single terms rather than full tree structures, the composition layer
-% was not used, and the two models differed only in which layer function
-% was used for the comparison layer. We found that the RNTN model worked
-% best with 11 dimensional vector representations for the 80 sets and a
-% 90 dimensional feature vector for the classifier. This model was able
-% to correctly label 99.3\% of the derivable test examples, and 99.1\%
-% of the remaining examples. The simpler RNN model worked best with 11
-% and 75 dimensions, respectively, but was able to achieve accuracies of
-% only 90.0\% and 87.\%, respectively.
-
-These results are fairly straightforward to interpret. The RNTN model
-was able to accurately encode the relations between the terms in the
-geometric relations between their vectors, and was able to then use
-that information to recover relations that were not overtly included
-in the training data. In contrast, the RNN model was able to achieve
-this behavior only incompletely. It is possible but not likely that it
-could be made to find a good solution with further optimization on
-different learning algorithms, or that it would do better on a larger
-universe of sets, for which there would be a larger set of training
-data to learn from, but the RNTN is readily able to achieve these
-effects in the setting discussed here.
-
-\section{Recursive structure in propositional logic}\label{sec:recursion}
-
-Recursive structure is a prominent feature of natural languages and an
-important part of their expressivity, as it allows people to use and
-interpret complex structures they have never encountered before.
-Theoretical accounts of natural language syntax and semantics
-therefore rely heavily on recursive structures, and we expect that
-accurate computational models will have to come to grips with them as
-well. The present section pursues this question for our two classes of
-RNN. In our evaluations, we exploit the fact that our logical language
-is infinite by testing on strings that are longer and more complex
-than any seen in training.
-
-% Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice
-%   said hello}, and \ii{Carl thinks that Bob said that Alice said
-%   hello}. Overt recursion of this kind is easy to find, and
-% theoretical accounts of natural language syntax and semantics rely
-% heavily on recursive structures. In order for a model to be able to
-% accurately learn natural language meanings, then, we expect that it
-% would need to be able to learn to represent the meanings of function
-% words in a such a way that they are able to behave correctly when
-% taking their own outputs as input. In evaluating the model, we take
-% advantage of the fact that recursive structures of this kind define
-% potentially infinite languages by testing the model on strings that
-% are longer and more complex than any seen in testing.
-
-\paragraph{Experiments}
-As in Section~\ref{sec:join}, we test this phenomenon within the
-framework of natural logic, but we now replace the unanalyzed symbols
-from that experiment with complex formulae. These formulae
-represent a truth-functionally complete classical propositional logic:
-each atomic variable denotes either $\True$ or $\False$, and the only
-operators are truth-functional ones.  Table~\ref{tab:pl} defines this
-logic, and Table~\ref{tab:plexs} gives some examples of relational
-statements that we can make in these terms. To compute these relations
-between statements, we exhaustively enumerate the sets of assignments
-of truth values to propositional variables that would satisfy each of
-the statements, and then we convert the set-theoretic relation between
-those assignments into one of the seven relations in
-Table~\ref{b-table}. As a result, each relational statement represents
-a valid theorem of the propositional logic.
-%
-%\begin{table}[tp]
-%  \centering
-%  \begin{subtable}[t]{0.45\textwidth}
-%    \centering
-%    \begin{tabular}[t]{l l}
-%      \toprule
-%      Formula     & Interpretation \\
-%      \midrule
-%      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
-%      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
-%      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
-%      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
-%      \bottomrule
-%    \end{tabular}    
-%    \caption{Well-formed formulae. $\varphi$ and $\psi$
-%      range over all well-formed formulae, and $\sem{\cdot}$ is
-%      the interpretation function mapping formulae into $\{\True,
-%      \False\}$.}\label{tab:pl}
-%  \end{subtable}
-%  \quad
-%  \begin{subtable}[t]{0.45\textwidth}
-%    \centering
-%    \begin{tabular}[t]{r c l}
-%      \toprule
-%      $\plneg a$        & $\natneg$ & $a$ \\
-%      $\plneg \plneg a$ & $\nateq$  & $a$ \\
-%      $a$               & $\natfor$ & $(a \plor b)$ \\
-%      $a$               & $\natrev$ & $(a \pland b)$ \\
-%      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
-%      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
-%      \bottomrule
-%    \end{tabular}
-%    \caption{Examples of statements about relations between
-%      well-formed formulae, defined in terms of sets of satisfying
-%      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
-%  \end{subtable}
-%  \caption{Propositional logic with natural logic relations.}  
-%  \label{prop-figure}
-%\end{table}
-
-% NOTE: There was some confusion at CSLI and the 224U lecture about
-% whether this was theorem proving or some kind of model checking.
-% Hopefully this is explicit enough.
-Socher et al.~\cite{socher2012semantic} show that a matrix-vector RNN
-model somewhat similar to our RNTN can learn a logic, but the logic
-discussed here is a much better approximation of the kind of
-expressivity needed for natural language. The logic learned in that
-experiment is boolean, wherein the atomic symbols are simply the
-values $0$ and $1$, rather than variables over those values. While
-learning the operators of that logic is not trivial, the outputs of
-each operator can be represented accurately by a single bit. The
-statements of propositional logic learned here describe conditions on
-unknown truth values of propositions. As opposed to the two-way
-contrasts seen in \cite{socher2012semantic}, this logic distinguishes
-between $2^{6} = 64$ possible assignments of truth values, and
-expressions of this logic define arbitrary conditions on these
-possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
-possible statements that the intermediate vector representations need
-to be able to distinguish if we are to learn this logical system in its
-full generality.
-
-For our experiments, we randomly generate a large set of  unique pairs 
-of formulae and compute the relation that holds for each pair.
-We discard pairs in which either statement is a tautology or
-contradiction (a statement that is true of either all or no possible
-assignments), for which none of the seven relation labels in
-Table~\ref{b-table} can hold. The resulting set of formula pairs is
-then partitioned into bins based on the number of logical operators in
-the longer of the two formulae. We then randomly sample 15\% of each
-bin for a held-out test set.
-
-If we do not implement any constraint that the two statements being
-compared are similar in any way, then the generated data are dominated
-by statements in which the two formulae refer to largely separate
-subsets of the six variables, which means that the $\natind$ relation
-is almost always correct.  In an effort to balance the distribution of
-relation labels without departing from the basic task of modeling
-propositional logic, we disallow individual pairs of statements from
-referring to more than four of the six propositional variables.
-
-\paragraph{Results} 
-We trained both the RNN and RNTN models on data with up to 4
-connectives (65k pairs) and tested it on examples of up to 12
-connectives (44k pairs). We initialized the model parameters randomly,
-including the vector representations of the six variables. The results
-are shown in Figure~\ref{prop-results}. In tuning, we found that the
-RNN model was approximately optimal with 45-dimensional vector
-representations, and the RNTN model was approximately optimal with 25
-dimensions. We fixed the size of the feature vector for the classifier
-at 75 dimensions. We found that the RNTN model was able to perform
-almost perfectly on unseen small test examples, with accuracy above
-98.7\% on formulae below length four (macro-F1 was above 95.5).
-Starting at size four, performance gradually falls with increasing
-size.  The RNN model did not perform well, reaching only 90.5\% (81.4
-micro-F1) accuracy on the smallest test examples, and declining from
-there to near-baseline performance on formulae of length 12. Training
-accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.
-
-% thu-and-or-deep-25-32b.txt iter 136
-% mon-and-or-deep-45r-32.txt iter 480
-
-\begin{figure}[tp]
-  \centering
-  \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
-  \caption{Model performance on propositional logic, by expression size. 
-    The dashed line indicates that only expressions of size four or less appeared in the training data.}  
-  \label{prop-results}
-\end{figure}
-
-The performance of the RNTN model on small unseen test examples
-indicates that it learned a correct approximation of the underlying
-logic. It appears that this approximation is accurate enough to yield
-correct answers when the composition layer is only applied a small
-number of times, but that the error in the approximation grows with
-increasing depth (and with the increasing complexity of the
-expressions), resulting in gradually dropping performance. In
-contrast, the RNN model did not appear to be able to learn a correct
-approximation of the logic for statements of any length, but it
-appears that the faulty logic that the RNN did learn decayed similarly
-with expression size.
-This is not necessarily a significant flaw in the model, since it remains possible that  different training regimes would allow it to learn accurate representations of the larger expressions.
-
-In an effort to better understand why the RNTN outperformed the plain
-RNN, we evaluated both models on pairs of long formulae involving
-binary connectives, to assess the extent to which they achieve
-substantially different representations for semantically distinct
-formulae and substantially similar representations for synonymous
-formulae. We found that both models are able to identify synonymous
-pairs like $(a \pland a)$ and $(a \pland (a \pland (a \pland a)))$,
-with both \ii{and} and \ii{or}, but only the RNTN is able to learn
-substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
- and $(a \pland (a \pland b))$ when the difference between the two is embedded under multiple operators. 
- Figure~\ref{prop-falloff} summarizes this result.
-
-% we discovered that this model looses information about in longer
-% statements in a particularly problematic way. It appears that that
-% model is unable to distinguish between two sentences when the only
-% difference between those sentences is embedded within a very deep
-% structure. We evaluated both models on sentences that differed in only
-% one term, but for which that one term was embedded under a large
-% number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and
-%   b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and
-%   b)))}. We then measured the Euclidean distance between the vector
-% representations of the two sentences in each pair. Our findings are
-% shown in Figure~\ref{prop-falloff}, and show that while the RNTN can
-% distinguish the two sentences well at every size that we test, the RNN
-% fails after a depth of approximately six.
-
-%\begin{figure}[htp]
-%  \centering
-%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-%  \caption{Semantically distinct formulae should have different
-%    representations. As measured by Euclidean distance, only the RNTN
-%    achieves this for formulae containing more than a small number of
-%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
-%    failing to capture the meaning contrasts.}
-%  \label{prop-falloff}
-%\end{figure}
-
-\section{Reasoning with natural language quantifiers and negation}\label{sec:quantifiers}
-
-We have seen that the RNTN can learn an approximation of propositional
-logic.  However, natural languages can express functional meanings of
-considerably greater complexity than this.  As a first step towards
-investigating whether our models can capture this complexity, we now
-attempt to directly measure the degree to which they are able to
-develop suitable representations for the semantics of natural language
-quantifiers like \ii{some} and \ii{all}. Quantification is far from
-the only place in natural language where complex functional meanings
-are found, but it is a natural starting point, since it can be tested
-in sentences whose structures are otherwise quite simple, and since it
-has formed a standard case study in prior formal work on natural
-language inference \cite{Icard:Moss:2013:LILT}.
-
-% \subsection{Data}
-
-This experiment replicates similar work described in
-\cite{bowman2013can}, which found that RNTNs can learn to reason well
-with quantifier meanings given sufficient training data. This paper
-replaces the partially manually annotated data in that paper with data
-that is generated directly using the logical system that we hope to
-learn, yielding results that we believe to be substantially more
-straightforward to interpret.
-
-\paragraph{Experiments}
-Our experimental data consist of pairs of sentences generated from a
-small artificial grammar. Each sentence contains a quantifier, a noun
-which may be negated, and an intransitive verb which may be
-negated. We use the basic quantifiers \ii{some}, \ii{most}, \ii{all},
-\ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
-\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}. We also
-include five nouns, four intransitive verbs, and the negation symbol
-\ii{not}. In order to be able to define relations between sentences
-with differing lexical items, we define the lexical relations between
-each noun--noun pair, each verb--verb pair, and each
-quantifier--quantifier pair. The grammar accepts aligned pairs of
-sentences of this form and calculates the natural logic relationship
-between them.  Some examples of these data are provided in
-Table~\ref{examplesofdata}.  As in previous sections, the goal of
-learning is then to assign these relational labels accurately to
-unseen pairs of sentences.
-
-%nouns = ['warthogs', 'turtles', 'mammals', 'reptiles', 'pets']
-%verbs = ['walk', 'move', 'swim', 'growl']
-%dets = ['all', 'not_all', 'some', 'no', 'most', 'not_most', 'two', 'lt_two', 'three', 'lt_three']
-%adverbs = ['', 'not']
-
-% To assign relation labels to sentence pairs, we built a small
-% task-specific implemenation of MacCartney's logic that can
-% accurately label sentences of this restricted language. The logic is
-% not able to derive all intuitively true relations of this language,
-% and fails to derive a single unique relation for certain types of
-% statement, including De Morgean's laws (e.g. \ii{(all pets) growl
-% $\natneg$ (some pet) (not growl)}), and we simply discard these
-% examples. Exhaustively generating the valid sentences under this
-% grammar and choosing those to which a relation label can be assigned
-% yields 66k sentence pairs. Some examples of these data are provided
-% in Table~\ref{examplesofdata}.
-
-%\begin{table}[htp]
-%  \centering
-%  \begin{tabular}{l c l}
-%    \toprule
-%    (most warthogs) walk         & $\natneg$ & (not-most warthogs) walk\\
-%    (most mammals) move          & $\natind$ &  (not-most (not turtles)) move\\
-%    (most (not pets)) (not swim) & $\natrev$ & (not-most (not pets)) move 
-%    \\[2ex]    
-%    (no turtles) (not growl)     & $\natalt$ & (no turtles) (not swim)\\
-%    (no warthogs) swim           & $\natrev$ & (no warthogs) move\\
-%    (no warthogs) move           & $\natfor$ & (no (not reptiles)) swim\\
-%    \bottomrule
-%  \end{tabular}
-%  \caption{Sample data involving two different quantifier pairs.}
-%  \label{examplesofdata}
-%\end{table}
-
-We evaluate the model using two experimental settings. In the simpler
-setting, \textsc{all split}, we randomly sample 85\% of the data and evaluate on the
-remaining 15\%. In this setting, the model is being asked to learn a
-complete reasoning system for the limited language and logic presented
-in the training data, but it is not being asked to generalize to test
-examples that are substantially different from those it was trained
-on. Crucially, though, to succeed on this task, the model must be able
-to recognize all of the lexical relations between the nouns, verbs,
-and quantifiers and how they interact. For instance, it might see
-\eqref{p1} and \eqref{p2} in training and, from that information,
-determine \eqref{p3}.
-%
-% do not allow a blank line --- adds too much space
-%
-\begin{gather}
-  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
-  \\
-  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
-  \\
-  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
-\end{gather}
-%
-% do not allow a blank line --- adds too much space
-%
-While our primary interest is in discovering the extent to which our
-models can learn to encode the logic given an arbitrary amount of
-data, we are also interested in the degree to which they can infer a
-correct representation for the logic from more constrained training
-data. To this end, we segment the sentence pairs according to which
-quantifiers appear in each pair, and then hold out one such pair for
-testing. We hypothesize that a model that can efficiently learn to
-represent a logic should be able to construct an accurate
-representation of each of the two quantifiers in the pair from the way that it
-interacts with the other nine quantifiers with which it is seen in training. 
-Since running this experiment requires choosing a pair of
-quantifiers to hold out before training, the resource demands of
-training prevent us from testing each of the 55 possible 
-pairs of quantifiers, so we choose only four pairs to test on.  Three
-of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
-\ii{all}/\ii{some}) were chosen because they allow for the most
-different class labels in the test data. The fourth is a
-self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
-handles equality.
-
-\begin{table}[tp]
-  \centering
-  \setlength{\tabcolsep}{10pt}
-  \begin{tabular}{ l rr@{ \ }r r@{ \ }r r@{ \ }r }
-    \toprule
-    Data & \multicolumn{3}{c}{Most frequent class} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
-    \midrule
-    \textsc{all split}	            &~~~~& 35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}
-    \\[1ex]    
-    \textsc{pair two/less-than-two} && 59.8 &(18.4) & 77.2 &(57.4) &	\textbf{100.0} &\textbf{(100.0)} \\
-    \textsc{pair not-all/not-most}  &&    0 &(0)    & 66.0 &(54.3) &	\textbf{93.8}  &\textbf{(91.5)} \\
-    \textsc{pair all/some}	    &&    0& (0)    & 62.7 &(65.9)  &	\textbf{78.4}  &\textbf{(84.7)} \\
-    \textsc{pair no/no}	            && 30.8  &(10.2) & 67.8 & (61.8) &	\textbf{99.9}  &\textbf{(99.7)} \\
-    \bottomrule
-  \end{tabular}
-  \caption{Performance on the quantifier experiments. Results are reported as accuracy scores followed by macroaveraged F1 scores in parentheses.}
-  \label{resultstable}
-\end{table}
-
-\paragraph{Results} 
-The results for these experiments are shown in
-Figure~\ref{resultstable}. We compare the results to a most frequent
-class baseline, which reflects the frequency in the test data of the
-most frequent class in the training data, $\natind$.  After some
-cross-validation, we chose 16- and 20-dimensional representations for
-the RNN and RNTN, respectively, and 75-dimensional feature vectors for
-the classifier. Training
-accuracy was near 80\% for the each of the five RNN experiments, and 
-near 98\% for each of the five RNTN experiments.
-
-%train rnn 0.20302, 0.017195
-
-The RNN performed poorly at this task, even though the sentences used
-in these examples are short enough to avoid the pathology shown in
-Figure~\ref{prop-falloff}.  However, the perfect performance by the
-RNTN on the \textsc{all split} and \textsc{pair two/less-than-two}
-experiments, and its strong performance on the \textsc{pair no/no}
-experiment, suggest that this stronger model is able to handle
-quantifiers correctly given sufficient training data, but the weaker
-results on two of the training settings suggest that it may not be
-able to generalize from this kind of restricted training data in general. However, though the question
-of how much data is necessary to accurately capture quantifier
-behavior in a na\"ive model remains open, the fact that this model is able to perform so well is promising.
-
 \section{General discussion}\label{sec:discussion}
@@
 
+% TODO: Cite deep RNNs
+% TODO: Cite LSTMs
+
 Of course, challenges remain. In terms of our experimental data, even
@@
 linguistic representations
-\cite{sochergrounded,kalchbrenner2014convolutional}. In addition,
+\cite{sochergrounded,kalchbrenner2014convolutional,}. In addition,
 there remain subtle questions about how to fairly assess whether these
@@
 In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.
-
-% \textit{Withheld for review.}\\ \\
-% Do later: Fill in, cite grants, internal reviewers
-
-\bibliographystyle{unsrtnat}
-
-\small % Note: Explicitly allowed in style guide
-\bibliography{MLSemantics}
-
-\end{document}

### 2014-11-23 writing/F'14 paper/intro.tex
Paper first push; init; naming; new data; PBS

--- 2014-08-25 writing/BayLearn abstract/paper.tex
@@
-\documentclass{article} % For LaTeX2e
-\usepackage{nips14submit_e,times}
-\usepackage{hyperref}
-\usepackage{url}
-\usepackage[leqno, fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{qtree}
-\usepackage[numbers]{natbib}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{colortbl}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{xcolor}
-
-\definecolor{mylinkcolor}{rgb}{0,0,0} % black
-\hypersetup{colorlinks, linkcolor=mylinkcolor, urlcolor=mylinkcolor, citecolor=mylinkcolor}
-
-\newcommand{\nateq}{\equiv}
-\newcommand{\natind}{\mathbin{\#}}
-%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
-\newcommand{\natneg}{\mathbin{^{\wedge}}}
-\newcommand{\natfor}{\sqsubset}
-\newcommand{\natrev}{\sqsupset}
-\newcommand{\natalt}{\mathbin{|}}
-\newcommand{\natcov}{\mathbin{\smallsmile}}
-
-\newcommand{\plneg}{\mathop{\textit{not}}}
-\newcommand{\pland}{\mathbin{\textit{and}}}
-\newcommand{\plor}{\mathbin{\textit{or}}}
-
-% Strikeout
-\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
-\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}
-
-\newcommand{\True}{\texttt{T}}
-\newcommand{\False}{\texttt{F}}
-\usepackage{stmaryrd}
-\newcommand{\sem}[1]{\ensuremath{\llbracket#1\rrbracket}}
-
-\renewcommand{\bibsection}{\subsubsection*{References}}
-
-\usepackage{gb4e}
-
-\def\ii#1{\textit{#1}}
-
-\newcommand{\mynote}[1]{{\color{red}\framebox{\begin{tabular}{p{0.9\textwidth}}\footnotesize#1 \end{tabular}}}}
-
-\title{Recursive Neural Networks for Learning Logical Semantics}
-
-\author{
-Samuel R.\ Bowman$^{\ast\dag}$ \\
-\texttt{sbowman@stanford.edu} \\[2ex]
-$^{\ast}$Stanford Linguistics \\
-\And
-Christopher Potts$^{\ast}$\\
-\texttt{cgpotts@stanford.edu} \\[2ex]
-$^{\dag}$Stanford NLP Group
-\And
-Christopher D.\ Manning$^{\ast\dag\ddag}$\\
-\texttt{manning@stanford.edu}\\[2ex]
-$^{\ddag}$Stanford Computer Science
-}
-
-% \author{
-% Samuel R.\ Bowman \\
-% NLP Group, Dept.\ of Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{sbowman@stanford.edu}
-%  \And
-%  Christopher Potts \\
-% Dept.\ of Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{cgpotts@stanford.edu}
-%  \And
-% Christopher D.\ Manning \\
-% NLP Group,  Depts.\ of Computer Science and Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{manning@stanford.edu}
-% }
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-
-\nipsfinalcopy % Uncomment for camera-ready version
-
-\begin{document}
-\maketitle
-
-  Supervised recursive neural network models (RNNs) for sentence
-  meaning have been successful in an array of sophisticated language
-  tasks, but it remains an open question whether they can learn
-  compositional semantic grammars that support logical deduction.  We
-  address this question directly by for the first time evaluating
-  whether each of two classes of neural model --- plain RNNs and
-  recursive neural tensor networks (RNTNs) --- can correctly learn
-  relationships such as entailment and contradiction between pairs of
-  sentences, where we have generated controlled data sets of sentences
-  from a logical grammar.  Our first experiment evaluates whether
-  these models can learn the basic algebra of logical relations
-  involved. Our second and third experiments extend this evaluation to
-  complex recursive structures and sentences involving quantification.
-  We find that the plain RNN achieves only mixed results on all three
-  experiments, whereas the stronger RNTN model generalizes well in
-  every setting and appears capable of learning suitable
-  representations for natural language logical inference.
-
 \section{Introduction}\label{sec:intro}
@@
 between sentences like \ii{every animal walks}, \ii{every turtle
-  moves}, and \ii{most reptiles don't move}. To date, experimental
-investigations of these concepts using distributed representations
-have been largely confined to short phrases
-\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
+  moves}, and \ii{most reptiles don't move}.
+
+To date, experimental investigations of these concepts using 
+distributed representations have been largely confined to short phrases
+\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment}.
 For robust natural language understanding, it is essential to model
 these phenomena in their full generality on complex linguistic
-structures.
+structures. Other recent work \cite{grefenstette2013towards,rocktaschellow} has
+developed well-understood frameworks that can compute the meanings of 
+certain fragments of natural language compositionally from vector or
+tensor representations. These frameworks are too structured to allow broad
+coverage meaning representations to be learned in
+supervised natural language tasks, and as such do not provide direct insight
+into the question of what existing applied machine learning models are 
+capable of.
 
-We address this question in the context of \ii{natural language
-  inference} (also known as \ii{recognizing textual entailment};
+% TODO: Reskim Gref paper.
+
+% TODO: Any way to avoid nested parens?
+
+Our research uses the task of \ii{natural language
+  inference} (also known as \ii{recognizing textual entailment}
 \cite{dagan2006pascal}), in which the goal is to determine the core
@@
 need of intermediate representations in an artificial logical
-language. Following \cite{bowman2013can}, we use the natural logic
-developed by \cite{maccartney2009extended} as our formal model. This
-logic defines seven core relations of synonymy, entailment,
-contradiction, and mutual consistency, as summarized in
+language. We use the natural logic developed by MacCartney and Manning 
+\shortcite{maccartney2009extended} 
+as our formal model. This logic defines seven core relations of synonymy, 
+entailment, contradiction, and mutual consistency, as summarized in
 Table~\ref{b-table}, and it provides rules of semantic combination for
@@
 
-In our experiments, we use this pre-specified logical grammar to
+% TODO: Do we need both of these Icard/Moss citations?
+
+Since no existing model architecture can be adopted without modification
+to handle entailment classification, we describe a novel method for this task
+in the spirit of prior work on recursive neural networks (\ref{methods}).
+In three experiments, we use this pre-specified logical grammar to
 generate controlled data sets encoding the semantic relationships
-between pairs of expressions and evaluate whether each of two classes
-of neural model --- plain RNNs and recursive neural tensor networks
-(RNTNs, \cite{socher2013acl1}) --- can learn those relationships
-correctly. In our first experiment (Section~\ref{sec:join}), we
+between pairs of expressions and evaluate whether this model can 
+learn to correctly identify the relationships between novel pairs.
+In our first experiment (Section~\ref{sec:join}), we
 evaluate the ability of these models to learn the core relational
@@
 between complex recursive structures like $(a \plor b)$ and
-$\plneg(\plneg a \pland \plneg b)$, and our third experiment
+$\plneg (\plneg a \pland \plneg b)$, and our third experiment
 (Section~\ref{sec:quantifiers}) involves relations between quantified
@@
 case, suggesting that it has in fact learned, or at least learned to
-simulate, our target logical
-concepts. These experiments differentiate the increased power of RNTNs
-better than previous work and provide the most convincing
-demonstration to date of the ability of neural networks to model
+simulate, our target logical concepts. These experiments provide the 
+most convincing demonstration to date of the ability of neural networks to model
 semantic inferences in complex natural language sentences.
 
-\begin{table}[tp]
+Finally, we perform a simple additional experiment to verify that our novel
+model structure is able to show similar reasoning behavior on real natural language data.
+The small size of existing entailment data precludes a state-of-the-art result, but we show
+that our model is able to perform competitively on the SICK entailment challenge, suggesting
+that the representational abilities that we observe in the previous sections are not limited to 
+narrowly defined artificial tasks.
+
+\begin{table*}[tp]
   \centering
@@
   \end{tabular}
-  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
+  \caption{\label{b-table}The seven natural logic relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}. 
     $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
@@
     %is insufficient knowledge to choose a label.}
-  \label{b-table}
-\end{table}
+
+\end{table*}
 

### 2014-11-23 writing/F'14 paper/join.tex
Paper first push; init; naming; new data; PBS

--- 2014-06-06 writing/S'14 paper/join.tex
@@
 % using the logic depicted in Figure~\ref{lattice-figure}.
-
-\begin{figure}[tp]
-  \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
-    \setlength{\unitlength}{1cm}
-    \begin{picture}(5,5.5)
-      \labelednode{2.50}{5}{}{1,2,3}
-
-      \put(0.75,4){\line(3,1){1.5}}
-      \put(2.5,4){\line(0,1){0.5}}
-      \put(4.25,4){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{3.5}{a,b}{1,2}
-      \labelednode{2.50}{3.5}{c}{1,3}
-      \labelednode{4.25}{3.5}{d}{2,3}
-
-      \put(0.75,2.5){\line(0,1){0.5}}
-      \put(0.75,2.5){\line(3,1){1.5}}
-
-      \put(2.5,2.5){\line(-3,1){1.5}}
-      \put(2.5,2.5){\line(3,1){1.5}}
-
-      \put(4.25,2.5){\line(0,1){0.5}}
-      \put(4.25,2.5){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{2}{e,f}{1}
-      \labelednode{2.50}{2}{}{2}
-      \labelednode{4.25}{2}{g,h}{3}
-
-      \put(2.5,1){\line(-3,1){1.5}}
-      \put(2.5,1){\line(0,1){0.5}}
-      \put(2.5,1){\line(3,1){1.5}}
-
-      \labelednode{2.5}{0.5}{}{}
-    \end{picture}
-    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
-    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}[t]{0.43\textwidth}
-    \centering
-    \setlength{\tabcolsep}{12pt}
-    \begin{tabular}[b]{c  c}
-      \toprule
-      Train & Test \\
-      \midrule
-                    & $b \nateq b$ \\
-      $b \natcov c$ &               \\
-                    & $b \natcov d$ \\
-                    & \strikeout{$b \natrev e$} \\
-      $c \natcov d$ &               \\
-      $c \natrev e$ &               \\
-                    & \strikeout{$c \nateq f$} \\
-      $c \natrev g$ &               \\ 
-                    & $e \natfor b$ \\
-      $e \natfor c$ &               \\[-1ex]
-      $\vdots$      & $\vdots$ \\
-      \bottomrule
-    \end{tabular}
-
-    \caption{A train/test split of the atomic statements about the
-      model.  Test statements not provable from the training data are
-      crossed out.}
-  \end{subfigure}  
-  \caption{Small example structure and data for learning relation composition.}
-  \label{lattice-figure}
-\end{figure} 
-% Note: None of these test examples is derivable from the shown training data, 
-% but we suggest that there is additional training data, so we can cross lines
-% out willy-nilly without fear.
 
@@
 \paragraph{Results} 
-We found that the RNTN model worked best with 11-dimensional vector
-representations for the 80 sets and a 90-dimensional feature vector
-for the classifier, though the performance of the model was not highly
-sensitive to either dimensionality parameter in any of the experiments
-discussed here. This model was able to correctly label 99.4\% (94.3
-macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
-best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
-respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
-the RNN.
+TODO: Explain Table \ref{joinresultstable}.
+
+\begin{table}[tp]
+  \centering
+  \setlength{\tabcolsep}{10pt}
+  \begin{tabular}{ r@{ \ }r r@{ \ }r r@{ \ }r }
+    \toprule
+    \multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
+    \midrule
+    35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}\\
+    \bottomrule
+  \end{tabular}
+  \caption{Performance on the TODO experiments. Results are reported as accuracy scores followed by macroaveraged F1 scores in parentheses.}
+  \label{joinresultstable}
+\end{table} % TODO: Is this worth the fair amount of space it takes up?
+
+\begin{figure}[tp]
+  \centering
+%  \begin{subfigure}[t]{0.45\textwidth}
+%    \centering
+%    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
+%    \setlength{\unitlength}{1cm}
+%    \begin{picture}(5,5.5)
+%      \labelednode{2.50}{5}{}{1,2,3}
+%      
+%      \put(0.75,4){\line(3,1){1.5}}
+%      \put(2.5,4){\line(0,1){0.5}}
+%      \put(4.25,4){\line(-3,1){1.5}}
+%      
+%      \labelednode{0.75}{3.5}{a,b}{1,2}
+%      \labelednode{2.50}{3.5}{c}{1,3}
+%      \labelednode{4.25}{3.5}{d}{2,3}
+%      
+%      \put(0.75,2.5){\line(0,1){0.5}}
+%      \put(0.75,2.5){\line(3,1){1.5}}
+%      
+%      \put(2.5,2.5){\line(-3,1){1.5}}
+%      \put(2.5,2.5){\line(3,1){1.5}}
+%      
+%      \put(4.25,2.5){\line(0,1){0.5}}
+%      \put(4.25,2.5){\line(-3,1){1.5}}
+%      
+%
+%      \labelednode{0.75}{2}{e,f}{1}
+%      \labelednode{2.50}{2}{}{2}
+%      \labelednode{4.25}{2}{g,h}{3}
+%      
+%      \put(2.5,1){\line(-3,1){1.5}}
+%      \put(2.5,1){\line(0,1){0.5}}
+%      \put(2.5,1){\line(3,1){1.5}}
+%      
+%      \labelednode{2.5}{0.5}{}{}
+%    \end{picture}
+%    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
+%    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
+%  \end{subfigure}
+%  \qquad
+%  \begin{subfigure}[t]{0.43\textwidth}
+    \centering
+    \setlength{\tabcolsep}{12pt}
+    \begin{tabular}[b]{c  c}
+      \toprule
+      Train & Test \\
+      \midrule
+      $b \natcov c$              & $b \nateq b$ \\
+      $c \natcov d$              & $b \natcov d$ \\
+      $c \natrev e$              & \strikeout{$b \natrev e$} \\
+      $c \natrev g$               & \strikeout{$c \nateq f$} \\
+      $e \natfor c$              & $e \natfor b$ \\
+      \bottomrule
+    \end{tabular}
+    \caption{A train/test split of the atomic statements about the
+      model.  Test statements not provable from the training data are
+      crossed out.}
+ % \end{subfigure}  
+ %  \caption{Small example structure and data for learning relation composition.}
+  \label{lattice-figure}
+\end{figure} 
+% Note: None of these test examples is derivable from the shown training data, 
+% but we suggest that there is additional training data, so we can cross lines
+% out willy-nilly without fear.
 
@@
 
+% TODO: Write out results.
+% TODO: Do we still care about the unprovable test examples?
+
 These results are fairly straightforward to interpret. The RNTN model

### 2014-11-23 writing/F'14 paper/methods.tex
Paper first push; init; naming; new data; PBS

--- 2014-06-06 writing/S'14 paper/methods.tex
@@
 
-We use the model architecture proposed in \cite{bowman2013can} and
-depicted in Figure~\ref{sample-figure}. The two phrases being compared
-are built up separately on each side of the tree, using the same
-composition function, until they have each been reduced to single
-vectors. The resulting vectors are fed into a separate comparison
+We use the model architecture depicted in Figure~\ref{sample-figure}. 
+The two phrases being compared are processed separately using a pair 
+of recursive networks that share a single set of parameters. 
+The resulting vectors are fed into a separate comparison
 layer that is meant to generate a feature vector capturing the
@@
 For a composition layer, we evaluate models with both the plain neural
-network layer function \eqref{rnn} and the RNTN layer function
-\eqref{rntn} proposed in \citet{chen2013learning}. A sigmoid
-nonlinearity (element-wise $\tanh$) is applied to the output of either
-layer function, following \cite{socher2013acl1}.
+network layer function \eqref{rnn} and a more powerful extension, the RNTN layer function
+\eqref{rntn} proposed in \newcite{chen2013learning}. A $\tanh$ nonlinearity is applied to the output of the layer function.
 %
@@
 \label{rntn}
-\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
+\vec{y}_{\textit{RNTN}} = \vec{y}_{\textit{RNN}} + f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
 \end{gather} 
@@
 $\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
-them by an $n \times 2n$ matrix of learned weights, and applies the
-element-wise non-linearity to the resulting vector. The RNTN has the
-same basic structure, but with the addition of a learned third-order
-tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
-multiplicative interactions between the child vectors. Both models
-include a bias vector, represented as $\vec{b}$.
+them by an $n \times 2n$ matrix of learned weights, adds a bias $\vec{b}$, and applies the
+element-wise non-linearity to the resulting vector. The RNTN adds a learned third-order tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
+multiplicative interactions between the child vectors.
 
-The comparison layer uses the same kind of function template as the
+% TODO: Kill ReLU?
+
+The comparison layer uses the same layer function as the
 composition layers (either an NN layer or an NTN layer) with
 independently learned parameters and a separate nonlinearity function.
-Rather than use a $\tanh$ nonlinearity here, we follow \cite{bowman2013can} 
-in using a rectified linear function. In
+Rather than use a $\tanh$ nonlinearity here, we use a rectified linear function. In
 particular, we use the leaky rectified linear function
-\cite{maasrectifier}: $f(\vec{x})=\max(\vec{x}, 0) +
+\shortcite{maasrectifier}: $f(\vec{x})=\max(\vec{x}, 0) +
 0.01\min(\vec{x}, 0)$, applied element-wise.
@@
 with learning rates computed using AdaGrad \cite{duchi2011adaptive}.
-The parameters (including the vocabulary) are initialized randomly 
-using a uniform distribution over $[-0.1, 0.1]$.
-Because the seven classes are not balanced in general, we report performance
+The parameters (including the vocabulary) are initialized randomly.
+Because the classes are not balanced in general, we report performance
 using both accuracy and macroaveraged F1, since the latter emphasizes
@@
 
-\begin{figure}[tp]
+% TODO: Describe mean-of-five scheme for results reporting.
+
+\begin{figure*}[tp]
   \centering
@@
   \label{sample-figure}
-\end{figure}
-
-%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?
+\end{figure*}

### 2014-11-23 writing/F'14 paper/quantifiers.tex
Paper first push; init; naming; new data; PBS

--- 2014-08-25 writing/BayLearn abstract/paper.tex
@@
-\documentclass{article} % For LaTeX2e
-\usepackage{nips14submit_e,times}
-\usepackage{hyperref}
-\usepackage{url}
-\usepackage[leqno, fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{qtree}
-\usepackage[numbers]{natbib}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{colortbl}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{xcolor}
-
-\definecolor{mylinkcolor}{rgb}{0,0,0} % black
-\hypersetup{colorlinks, linkcolor=mylinkcolor, urlcolor=mylinkcolor, citecolor=mylinkcolor}
-
-\newcommand{\nateq}{\equiv}
-\newcommand{\natind}{\mathbin{\#}}
-%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
-\newcommand{\natneg}{\mathbin{^{\wedge}}}
-\newcommand{\natfor}{\sqsubset}
-\newcommand{\natrev}{\sqsupset}
-\newcommand{\natalt}{\mathbin{|}}
-\newcommand{\natcov}{\mathbin{\smallsmile}}
-
-\newcommand{\plneg}{\mathop{\textit{not}}}
-\newcommand{\pland}{\mathbin{\textit{and}}}
-\newcommand{\plor}{\mathbin{\textit{or}}}
-
-% Strikeout
-\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
-\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}
-
-\newcommand{\True}{\texttt{T}}
-\newcommand{\False}{\texttt{F}}
-\usepackage{stmaryrd}
-\newcommand{\sem}[1]{\ensuremath{\llbracket#1\rrbracket}}
-
-\renewcommand{\bibsection}{\subsubsection*{References}}
-
-\usepackage{gb4e}
-
-\def\ii#1{\textit{#1}}
-
-\newcommand{\mynote}[1]{{\color{red}\framebox{\begin{tabular}{p{0.9\textwidth}}\footnotesize#1 \end{tabular}}}}
-
-\title{Recursive Neural Networks for Learning Logical Semantics}
-
-\author{
-Samuel R.\ Bowman$^{\ast\dag}$ \\
-\texttt{sbowman@stanford.edu} \\[2ex]
-$^{\ast}$Stanford Linguistics \\
-\And
-Christopher Potts$^{\ast}$\\
-\texttt{cgpotts@stanford.edu} \\[2ex]
-$^{\dag}$Stanford NLP Group
-\And
-Christopher D.\ Manning$^{\ast\dag\ddag}$\\
-\texttt{manning@stanford.edu}\\[2ex]
-$^{\ddag}$Stanford Computer Science
-}
-
-% \author{
-% Samuel R.\ Bowman \\
-% NLP Group, Dept.\ of Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{sbowman@stanford.edu}
-%  \And
-%  Christopher Potts \\
-% Dept.\ of Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{cgpotts@stanford.edu}
-%  \And
-% Christopher D.\ Manning \\
-% NLP Group,  Depts.\ of Computer Science and Linguistics\\
-% Stanford University\\
-% Stanford, CA 94305-2150 \\
-% \texttt{manning@stanford.edu}
-% }
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-
-\nipsfinalcopy % Uncomment for camera-ready version
-
-\begin{document}
-\maketitle
-
-  Supervised recursive neural network models (RNNs) for sentence
-  meaning have been successful in an array of sophisticated language
-  tasks, but it remains an open question whether they can learn
-  compositional semantic grammars that support logical deduction.  We
-  address this question directly by for the first time evaluating
-  whether each of two classes of neural model --- plain RNNs and
-  recursive neural tensor networks (RNTNs) --- can correctly learn
-  relationships such as entailment and contradiction between pairs of
-  sentences, where we have generated controlled data sets of sentences
-  from a logical grammar.  Our first experiment evaluates whether
-  these models can learn the basic algebra of logical relations
-  involved. Our second and third experiments extend this evaluation to
-  complex recursive structures and sentences involving quantification.
-  We find that the plain RNN achieves only mixed results on all three
-  experiments, whereas the stronger RNTN model generalizes well in
-  every setting and appears capable of learning suitable
-  representations for natural language logical inference.
-
-  \section{Introduction}\label{sec:intro}
-
-Supervised recursive neural network models (RNNs) for sentence meaning
-have been successful in an array of sophisticated language tasks,
-including sentiment analysis \cite{socher2011semi,socher2013acl1},
-image description \cite{sochergrounded}, and paraphrase detection
-\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
-the ability of these models to learn compositional semantic grammars,
-but it remains an open question whether they can achieve the same
-results as grammars based in logical forms
-\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
-when it comes to the core semantic concepts of quantification,
-entailment, and contradiction needed to identify the relationships
-between sentences like \ii{every animal walks}, \ii{every turtle
-  moves}, and \ii{most reptiles don't move}. To date, experimental
-investigations of these concepts using distributed representations
-have been largely confined to short phrases
-\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment,kalchbrenner2014convolutional}.
-For robust natural language understanding, it is essential to model
-these phenomena in their full generality on complex linguistic
-structures.
-
-We address this question in the context of \ii{natural language
-  inference} (also known as \ii{recognizing textual entailment};
-\cite{dagan2006pascal}), in which the goal is to determine the core
-inferential relationship between two sentences. Much of the
-theoretical work on this task (and some successful implemented models
-\cite{maccartney2009extended,watanabe2012latent}) involves \ii{natural
-  logics}, which are formal systems that define rules of inference
-between natural language words, phrases, and sentences without the
-need of intermediate representations in an artificial logical
-language. Following \cite{bowman2013can}, we use the natural logic
-developed by \cite{maccartney2009extended} as our formal model. This
-logic defines seven core relations of synonymy, entailment,
-contradiction, and mutual consistency, as summarized in
-Table~\ref{b-table}, and it provides rules of semantic combination for
-projecting these relations from the lexicon up to complex phrases. The
-formal properties and inferential strength of this system are now
-well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
-
-In our experiments, we use this pre-specified logical grammar to
-generate controlled data sets encoding the semantic relationships
-between pairs of expressions and evaluate whether each of two classes
-of neural model --- plain RNNs and recursive neural tensor networks
-(RNTNs, \cite{socher2013acl1}) --- can learn those relationships
-correctly. In our first experiment (Section~\ref{sec:join}), we
-evaluate the ability of these models to learn the core relational
-algebra of natural logic from data. Our second experiment
-(Section~\ref{sec:recursion}) extends this evaluation to relations
-between complex recursive structures like $(a \plor b)$ and
-$\plneg(\plneg a \pland \plneg b)$, and our third experiment
-(Section~\ref{sec:quantifiers}) involves relations between quantified
-statements like \ii{every reptile walks} and \ii{no turtle moves}.
-We find that the plain RNN achieves only mixed results in all three
-experiments, whereas the stronger RNTN model generalizes well in every
-case, suggesting that it has in fact learned, or at least learned to
-simulate, our target logical
-concepts. These experiments differentiate the increased power of RNTNs
-better than previous work and provide the most convincing
-demonstration to date of the ability of neural networks to model
-semantic inferences in complex natural language sentences.
-
-\begin{table}[tp]
-  \centering
-  \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{l c l l} 
-    \toprule
-    Name & Symbol & Set-theoretic definition & Example \\ 
-    \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
-    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
-    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
-    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
-    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
-    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
-    \bottomrule
-  \end{tabular}
-  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
-    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
-    %is insufficient knowledge to choose a label.}
-  \label{b-table}
-\end{table}
-
-% Citations to additional past work to be added.\\...\\...\\...\\...\\...
-
-% Deep learning methods in NLP which learn vector representations for words have seen successful uses in recent years on increasingly sophisticated tasks \cite{collobert2011natural, socher2011semi, socher2013acl1, chen2013learning}.
-Given the still modest performance of semantically rich NLP systems in many domains---question answering and machine translation, for instance---it is worth exploring the degree to which learned vectors can serve as general purpose semantic representations.
-Much of the work to date analyzing vector representations for words (see \cite{baroni2013frege}) has focused on lexical semantic behaviors---like the similarity between words like \ii{Paris} and \ii{France}.
-Good similarity functions are valuable for many NLP tasks, but there are real use cases for which it is necessary to know how words relate to one another or to some extrinsic representation, and to model the ways in which word meanings combine to form phrase, sentence, or document meanings.
-This paper explores the ability of linguistic representations developed using supervised deep learning techniques to support interpretation and reasoning.
-
-% There are two broad family of tasks that could be used to test the ability of a model to develop general purpose meaning representations.
-In an interpretation task, sentences are mapped onto some denotation, such as  \ii{true} or \ii{false} for statements, or a factual answer for questions.
-There has been preliminary work in developing distributed models for interpretation \cite{grefenstette2013towards, rocktaschellow}, but developing a representation of world knowledge that corresponds accurately to the content expressed in language introduces considerable design challenges.
-I approach the problem by way of an inference task instead.
-Inferring the truth of one sentence from another does not require any preexisting knowledge representations, but nonetheless requires a precise representation of sentence meaning.
-I borrow the structure of the task from MacCartney and Manning  \cite{maccartney2009extended}.
-In it, the model is presented with a pair of sentences, and made to label the logical relation between the sentences as equivalence, entailment, or any of five other classes, as here:
-
-%\begin{quote}
-%\begin{enumerate}\small
-%\item Many smartphone users avoid high bills overseas by disabling data service.
-%\item Not everyone uses their smartphones for email when traveling abroad.
-%\end{enumerate} 
-%$\Rightarrow$ Entailment
-%\end{quote}
-
-%In this paper, we test the ability of recursive models to on three simple tasks, each of which is meant to capture a property that is necessary in representing natural language meaning in the setting of inference.
-I begin with an overview of MacCartney and Manning's \cite{maccartney2009extended} framework for inference, and of the recursive neural networks that I study.
-by showing that these models can learn to correctly represent entailment representations between sentences.
-I then show that these models can capture the meanings of recursive structures accurately up to a sufficient depth.
-I finally close with a brief demonstration of the ability of these models to reason over short natural language sentences involving quantifiers.
-
-% \subsection{Natural language inference and natural logic}
-
-% In its simplest form, \ii{natural language inference} (also known as \ii{recognizing textual entailment}, as in \cite{dagan2006pascal}) is the task of determine whether one sentence entails another.
-Much of the theoretical work on this task (and some successful implemented models \cite{maccartney2009natural, watanabe2012latent}) involve \ii{natural logic}, formal systems that define sound rules of inference from one sentence of natural language to another without the need for intermediate representations in some other logic.
-The most powerful model that we are aware of for natural logic is due to MacCartney and Manning \cite{maccartney2009extended} and Icard \cite{icard2012inclusion}, and is centered around the definition of a set of seven logical relations which can hold between sentences, shown in Table \ref{b-table}.
-
-% This approach to natural logic can capture much of the complexity of natural language meaning within a well understood framework, and is also fairly straightforward to implement in a machine learning setting since it can be reduced to a seven-way classification problem on sentence pairs.
-Our goal in this paper is to learn recursive neural network models which are able to mimic key behaviors of this system.
-
-% \begin{table}
-% \begin{center}
-% \begin{tabular}{|c|c|c|c|} \hline
-% name & symbol & set-theoretic definition & example \\ \hline \hline
-% entailment & $x \sqsubset y$ & $x \subset y$ & \ii{crow, bird}  \\ \hline
-% reverse entailment & $x \sqsupset y$ & $x \supset y$ & \ii{Asian, Thai}  \\ \hline
-% equivalence & $x \equiv y$ & $x = y$ & \ii{couch, sofa} \\ \hline
-% alteration & $x$ $|$ $y$ & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{cat, dog} \\ \hline
-% negation & $x \natneg y$ & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{able, unable} \\ \hline
-% cover & $x \smallsmile y$ & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-ape} \\ \hline
-% independence & $x$ \# $y$ & (else) & \ii{hungry, hippo}\\ \hline
-% \end{tabular}
-% \caption{The entailment relations in  $\mathfrak{B}$.
-$\mathcal{D}$ is the universe of possible objects of the same type as those being compared, and the relation \# applies whenever none of the other six do, including when there is insufficient knowledge to choose a label.}
-% \label{b-table}
-% \end{center}
-% \end{table}
-
-% The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes.
-It should be noted that this is not a balanced classification problem.
-If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them.
-% TODO: How much should we say about this?
-
-% The natural logic engine at the core of MacCartney and Manning's system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures.
-These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
-%for dozens of different cases of insertion, deletion and substitution of different types of lexical item.
-For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences.
-Here, the substitution of \ii{no} for \ii{some}  yields $\natneg$ and the substitution of \ii{dogs} for \ii{animals} yields $\sqsupset$.
-It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
-%\begin{quote}
-
-%1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
-%2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
-%3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}
-
-%\end{quote}
-
-% Work to date on inference in neural network models is quite limited.
-% \citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
-% In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations.
-They propose a deterministic inference engine (similar to MacCartney's) which is augmented by a probabilistic component that evaluates individual lexical substitution steps in the derivation using vector representations, though again these representations are not learned, and no evaluations of this system have been published to date.
-% \label{sec2}
-
-\section{Recursive neural network models} \label{methods}
-
-We study neural models that adhere to the linguistic \ii{principle of
- compositionality}, which says that the meanings for complex
-expressions are derived from the meanings of their constituent parts
-via specific composition functions \cite{Partee84,Janssen97}. In our
-distributed setting, word meanings are vectors of length $n$. The
-composition function maps pairs of them to single vectors of length $n$, 
-which can then be merged again to represent more complex
-phrases. Once the entire sentence-level representation has been
-derived, it serves as a fixed-length input for some subsequent function.
-
-We use the model architecture proposed in \cite{bowman2013can} and
-depicted in Figure~\ref{sample-figure}. The two phrases being compared
-are built up separately on each side of the tree, using the same
-composition function, until they have each been reduced to single
-vectors. The resulting vectors are fed into a separate comparison
-layer that is meant to generate a feature vector capturing the
-relation between the two phrases. The output of this layer is then
-given to a softmax classifier, which in turn produces a hypothesized
-distribution over the seven relations represented in Table~\ref{b-table}.
-
-For a composition layer, we evaluate models with both the plain neural
-network layer function \eqref{rnn} and the RNTN layer function
-\eqref{rntn} proposed in \citet{chen2013learning}. A sigmoid
-nonlinearity (element-wise $\tanh$) is applied to the output of either
-layer function, following \cite{socher2013acl1}.
-%
-\begin{gather} \label{rnn}
-\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
-\label{rntn}
-\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
-\end{gather} 
-%
-Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
-representations for the left and right children of the node, and
-$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
-them by an $n \times 2n$ matrix of learned weights, and applies the
-element-wise non-linearity to the resulting vector. The RNTN has the
-same basic structure, but with the addition of a learned third-order
-tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
-multiplicative interactions between the child vectors. Both models
-include a bias vector~$\vec{b}$.
-
-The comparison layer uses the same kind of function template as the
-composition layers (either an NN layer or an NTN layer) with
-independently learned parameters and a separate nonlinearity function.
-Rather than use a $\tanh$ nonlinearity here, we follow \cite{bowman2013can} 
-in using a rectified linear function. In
-particular, we use the leaky rectified linear function
-\cite{maasrectifier}: $f(\vec{x})=\max(\vec{x}, 0) +
-0.01\min(\vec{x}, 0)$, applied element-wise.
-
-To run the model forward and label a pair of phrases, the structure of
-the lower layers of the network is assembled so as to mirror the tree
-structures provided for each phrase. The word vectors are then looked
-up from the vocabulary matrix $V$ (one of the model parameters), and
-the composition and comparison functions are used to pass information
-up the tree and into the classifier at the top. For an objective
-function, we use the negative log of the probability assigned to the
-correct label.
-
-% Removed 'minibatch' -> For two of the experiments, we use minibatches
-% of size 1, which doesn't really count.
-We train the model using stochastic gradient descent (SGD)
-with learning rates computed using AdaGrad \cite{duchi2011adaptive}.
-The parameters (including the vocabulary) are initialized randomly 
-using a uniform distribution over $[-0.1, 0.1]$.
-Because the seven classes are not balanced in general, we report performance
-using both accuracy and macroaveraged F1, since the latter emphasizes
- performance on infrequent classes. We compute macroaveraged F1 
-as the harmonic mean of mean precision and mean recall, both computed
-for all classes for which there is test data, setting precision to 0 
-where it is not defined. Source code and generated data will be released
-after the review period.
-
-%\begin{figure}[tp]
-%  \centering
-% % \includegraphics[scale=0.35]{model.eps}
-%  \input{figure1}
-%  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
-%    The same structure is used for both the RNN and RNTN layer functions.} 
-%  \label{sample-figure}
-%\end{figure}
-
-%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?
-
-\section{Reasoning about semantic relations}\label{sec:join}
-
-If a model is to learn the behavior of a relational logic like the one
-presented here from a finite amount of data, it must learn to deduce new
-relations from seen relations in a sound manner. The simplest such
-deductions involve atomic statements using the relations in
-Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
-\natrev c$, one can conclude that $a \natrev c$, by basic
-set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
-full set of sound inferences of this form is summarized in
-Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
-of relations for which no valid inference can be drawn in our logic.
-
-% about the relations themselves that do not depend on the
-% internal structure of the things being compared. For example, given
-% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
-% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven relations support more
-% than just transitivity: MacCartney and Manning's
-% \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of relations of the form $a R
-% b$ and $b R' c$, including several less intuitive ones such as that if
-% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-%
-%\begin{table}[htp]
-%  \centering  
-%  \setlength{\arraycolsep}{8pt}
-%  \renewcommand{\arraystretch}{1.1}
-%  \newcommand{\UNK}{\cdot}  
-%  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
-%    %\hline
-%    \multicolumn{1}{c}{}
-%             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
-%    \cline{2-8}
-%    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
-%    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
-%    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
-%    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
-%    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
-%    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
-%    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
-%    \cline{2-8}
-%  \end{array}$
-%  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.
-These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
-We assess our models' ability to reproduce such inferential paths.}
-%  \label{tab:jointable}
-%\end{table}
-
-\paragraph{Experiments}
-To test our models' ability to learn these inferential patterns, we
-create small boolean structures for our logic in which terms denote
-sets of entities from a small domain.  Figure~\ref{lattice-figure}
-depicts a structure of this form. The lattice gives the full model,
-for which all the statements on the right are valid. We divide these
-statements evenly into training and test sets, and remove from the
-test set those statements which cannot be proven from the training
-statements.
-% using the logic depicted in Figure~\ref{lattice-figure}.
-
-% We test the model's ability to learn this behavior by creating
-% artificial data sets of terms which represent sets of numbers. Since
-% MacCartney and Manning's set of relations hold between sets as well as
-% between sentences, we can use the underlying set structure to generate
-% the all of the relations that hold between any pair of these terms, as
-% in Figure \ref{lattice-figure}. We train the model defined above on a
-% subset of these relations, but rather then presenting the model with a
-% pair of tree-structured sentences as inputs, simply present it with
-% two single terms, each of which corresponds to a single vector in the
-% (randomly initialized) vocabulary matrix $V$, ensuring that the model
-% has no information about the terms being compared except the relations
-% between them.
-
-In our experiments, we create 80 randomly generated sets drawing from
-a domain of seven elements. This yields a data set consisting of
-6400 statements about pairs of formulae. 3200 of these pairs are
-chosen as a test set, and that test set is further reduced to the 2960
-examples that can be provably derived from the training data.
-
-We trained versions of both the RNN model and the RNTN model on these
-data sets. In both cases, the models were implemented exactly as
-described in Section~\ref{methods}, but since the items being compared
-are single terms rather than full tree structures, the composition
-layer was not involved, and the two models differed only in which
-layer function was used for the comparison layer. We simply present
-the models with two single terms, each of which corresponds to a
-single vector in the (randomly initialized) vocabulary matrix $V$,
-thereby ensuring that the model has no information about the terms
-being compared except the relations between them.
-
-\paragraph{Results} 
-We found that the RNTN model worked best with 11-dimensional vector
-representations for the 80 sets and a 90-dimensional feature vector
-for the classifier, though the performance of the model was not highly
-sensitive to either dimensionality parameter in any of the experiments
-discussed here. This model was able to correctly label 99.4\% (94.3
-macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
-best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
-respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
-the RNN.
-
-\begin{figure}[tp]
-  \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
-    \setlength{\unitlength}{1cm}
-    \begin{picture}(5,5.5)
-      \labelednode{2.50}{5}{}{1,2,3}
-
-      \put(0.75,4){\line(3,1){1.5}}
-      \put(2.5,4){\line(0,1){0.5}}
-      \put(4.25,4){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{3.5}{a,b}{1,2}
-      \labelednode{2.50}{3.5}{c}{1,3}
-      \labelednode{4.25}{3.5}{d}{2,3}
-
-      \put(0.75,2.5){\line(0,1){0.5}}
-      \put(0.75,2.5){\line(3,1){1.5}}
-
-      \put(2.5,2.5){\line(-3,1){1.5}}
-      \put(2.5,2.5){\line(3,1){1.5}}
-
-      \put(4.25,2.5){\line(0,1){0.5}}
-      \put(4.25,2.5){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{2}{e,f}{1}
-      \labelednode{2.50}{2}{}{2}
-      \labelednode{4.25}{2}{g,h}{3}
-
-      \put(2.5,1){\line(-3,1){1.5}}
-      \put(2.5,1){\line(0,1){0.5}}
-      \put(2.5,1){\line(3,1){1.5}}
-
-      \labelednode{2.5}{0.5}{}{}
-    \end{picture}
-    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
-    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}[t]{0.43\textwidth}
-    \centering
-    \setlength{\tabcolsep}{12pt}
-    \begin{tabular}[b]{c  c}
-      \toprule
-      Train & Test \\
-      \midrule
-                    & $b \nateq b$ \\
-      $b \natcov c$ &               \\
-                    & $b \natcov d$ \\
-                    & \strikeout{$b \natrev e$} \\
-      $c \natcov d$ &               \\
-      $c \natrev e$ &               \\
-                    & \strikeout{$c \nateq f$} \\
-      $c \natrev g$ &               \\ 
-                    & $e \natfor b$ \\
-      $e \natfor c$ &               \\[-1ex]
-      $\vdots$      & $\vdots$ \\
-      \bottomrule
-    \end{tabular}
-
-    \caption{A train/test split of the atomic statements about the
-      model.  Test statements not provable from the training data are
-      crossed out.}
-  \end{subfigure}  
-  \caption{Small example structure and data for learning relation composition.}
-  \label{lattice-figure}
-\end{figure} 
-% Note: None of these test examples is derivable from the shown training data, 
-% but we suggest that there is additional training data, so we can cross lines
-% out willy-nilly without fear.
-
-% RNTN log: tue-j-11-6x80-hip.txt
-% Train PER: 0.0021875
-
-% RNN log: tue-j-11-6x80-r-75.txt
-% Train PER: 0.047188
-
-% and create a dataset consisting of the relations between every pair of
-% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
-% test dataset, and that test dataset was further split into the 2960
-% examples that can be provably derived from the test data using
-% MacCartney and Manning's join table (or by the symmetry of the
-% relations in about half of the cases) and the 240 that
-% cannot.
-
-% We tested a version of both the RNN model and the RNTN model on these
-% data. In both cases, the models were implemented exactly as described
-% in Section~\ref{methods}, but since the items being compared are
-% single terms rather than full tree structures, the composition layer
-% was not used, and the two models differed only in which layer function
-% was used for the comparison layer. We found that the RNTN model worked
-% best with 11 dimensional vector representations for the 80 sets and a
-% 90 dimensional feature vector for the classifier. This model was able
-% to correctly label 99.3\% of the derivable test examples, and 99.1\%
-% of the remaining examples. The simpler RNN model worked best with 11
-% and 75 dimensions, respectively, but was able to achieve accuracies of
-% only 90.0\% and 87.\%, respectively.
-
-These results are fairly straightforward to interpret. The RNTN model
-was able to accurately encode the relations between the terms in the
-geometric relations between their vectors, and was able to then use
-that information to recover relations that were not overtly included
-in the training data. In contrast, the RNN model was able to achieve
-this behavior only incompletely. It is possible but not likely that it
-could be made to find a good solution with further optimization on
-different learning algorithms, or that it would do better on a larger
-universe of sets, for which there would be a larger set of training
-data to learn from, but the RNTN is readily able to achieve these
-effects in the setting discussed here.
-
-\section{Recursive structure in propositional logic}\label{sec:recursion}
-
-Recursive structure is a prominent feature of natural languages and an
-important part of their expressivity, as it allows people to use and
-interpret complex structures they have never encountered before.
-Theoretical accounts of natural language syntax and semantics
-therefore rely heavily on recursive structures, and we expect that
-accurate computational models will have to come to grips with them as
-well. The present section pursues this question for our two classes of
-RNN. In our evaluations, we exploit the fact that our logical language
-is infinite by testing on strings that are longer and more complex
-than any seen in training.
-
-% Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice
-%   said hello}, and \ii{Carl thinks that Bob said that Alice said
-%   hello}. Overt recursion of this kind is easy to find, and
-% theoretical accounts of natural language syntax and semantics rely
-% heavily on recursive structures. In order for a model to be able to
-% accurately learn natural language meanings, then, we expect that it
-% would need to be able to learn to represent the meanings of function
-% words in a such a way that they are able to behave correctly when
-% taking their own outputs as input. In evaluating the model, we take
-% advantage of the fact that recursive structures of this kind define
-% potentially infinite languages by testing the model on strings that
-% are longer and more complex than any seen in testing.
-
-\paragraph{Experiments}
-As in Section~\ref{sec:join}, we test this phenomenon within the
-framework of natural logic, but we now replace the unanalyzed symbols
-from that experiment with complex formulae. These formulae
-represent a truth-functionally complete classical propositional logic:
-each atomic variable denotes either $\True$ or $\False$, and the only
-operators are truth-functional ones.  Table~\ref{tab:pl} defines this
-logic, and Table~\ref{tab:plexs} gives some examples of relational
-statements that we can make in these terms. To compute these relations
-between statements, we exhaustively enumerate the sets of assignments
-of truth values to propositional variables that would satisfy each of
-the statements, and then we convert the set-theoretic relation between
-those assignments into one of the seven relations in
-Table~\ref{b-table}. As a result, each relational statement represents
-a valid theorem of the propositional logic.
-%
-%\begin{table}[tp]
-%  \centering
-%  \begin{subtable}[t]{0.45\textwidth}
-%    \centering
-%    \begin{tabular}[t]{l l}
-%      \toprule
-%      Formula     & Interpretation \\
-%      \midrule
-%      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
-%      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
-%      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
-%      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
-%      \bottomrule
-%    \end{tabular}    
-%    \caption{Well-formed formulae. $\varphi$ and $\psi$
-%      range over all well-formed formulae, and $\sem{\cdot}$ is
-%      the interpretation function mapping formulae into $\{\True,
-%      \False\}$.}\label{tab:pl}
-%  \end{subtable}
-%  \quad
-%  \begin{subtable}[t]{0.45\textwidth}
-%    \centering
-%    \begin{tabular}[t]{r c l}
-%      \toprule
-%      $\plneg a$        & $\natneg$ & $a$ \\
-%      $\plneg \plneg a$ & $\nateq$  & $a$ \\
-%      $a$               & $\natfor$ & $(a \plor b)$ \\
-%      $a$               & $\natrev$ & $(a \pland b)$ \\
-%      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
-%      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
-%      \bottomrule
-%    \end{tabular}
-%    \caption{Examples of statements about relations between
-%      well-formed formulae, defined in terms of sets of satisfying
-%      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
-%  \end{subtable}
-%  \caption{Propositional logic with natural logic relations.}  
-%  \label{prop-figure}
-%\end{table}
-
-% NOTE: There was some confusion at CSLI and the 224U lecture about
-% whether this was theorem proving or some kind of model checking.
-% Hopefully this is explicit enough.
-Socher et al.~\cite{socher2012semantic} show that a matrix-vector RNN
-model somewhat similar to our RNTN can learn a logic, but the logic
-discussed here is a much better approximation of the kind of
-expressivity needed for natural language. The logic learned in that
-experiment is boolean, wherein the atomic symbols are simply the
-values $0$ and $1$, rather than variables over those values. While
-learning the operators of that logic is not trivial, the outputs of
-each operator can be represented accurately by a single bit. The
-statements of propositional logic learned here describe conditions on
-unknown truth values of propositions. As opposed to the two-way
-contrasts seen in \cite{socher2012semantic}, this logic distinguishes
-between $2^{6} = 64$ possible assignments of truth values, and
-expressions of this logic define arbitrary conditions on these
-possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
-possible statements that the intermediate vector representations need
-to be able to distinguish if we are to learn this logical system in its
-full generality.
-
-For our experiments, we randomly generate a large set of  unique pairs 
-of formulae and compute the relation that holds for each pair.
-We discard pairs in which either statement is a tautology or
-contradiction (a statement that is true of either all or no possible
-assignments), for which none of the seven relation labels in
-Table~\ref{b-table} can hold. The resulting set of formula pairs is
-then partitioned into bins based on the number of logical operators in
-the longer of the two formulae. We then randomly sample 15\% of each
-bin for a held-out test set.
-
-If we do not implement any constraint that the two statements being
-compared are similar in any way, then the generated data are dominated
-by statements in which the two formulae refer to largely separate
-subsets of the six variables, which means that the $\natind$ relation
-is almost always correct.  In an effort to balance the distribution of
-relation labels without departing from the basic task of modeling
-propositional logic, we disallow individual pairs of statements from
-referring to more than four of the six propositional variables.
-
-\paragraph{Results} 
-We trained both the RNN and RNTN models on data with up to 4
-connectives (65k pairs) and tested it on examples of up to 12
-connectives (44k pairs). We initialized the model parameters randomly,
-including the vector representations of the six variables. The results
-are shown in Figure~\ref{prop-results}. In tuning, we found that the
-RNN model was approximately optimal with 45-dimensional vector
-representations, and the RNTN model was approximately optimal with 25
-dimensions. We fixed the size of the feature vector for the classifier
-at 75 dimensions. We found that the RNTN model was able to perform
-almost perfectly on unseen small test examples, with accuracy above
-98.7\% on formulae below length four (macro-F1 was above 95.5).
-Starting at size four, performance gradually falls with increasing
-size.  The RNN model did not perform well, reaching only 90.5\% (81.4
-micro-F1) accuracy on the smallest test examples, and declining from
-there to near-baseline performance on formulae of length 12. Training
-accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.
-
-% thu-and-or-deep-25-32b.txt iter 136
-% mon-and-or-deep-45r-32.txt iter 480
-
-\begin{figure}[tp]
-  \centering
-  \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
-  \caption{Model performance on propositional logic, by expression size. 
-    The dashed line indicates that only expressions of size four or less appeared in the training data.}  
-  \label{prop-results}
-\end{figure}
-
-The performance of the RNTN model on small unseen test examples
-indicates that it learned a correct approximation of the underlying
-logic. It appears that this approximation is accurate enough to yield
-correct answers when the composition layer is only applied a small
-number of times, but that the error in the approximation grows with
-increasing depth (and with the increasing complexity of the
-expressions), resulting in gradually dropping performance. In
-contrast, the RNN model did not appear to be able to learn a correct
-approximation of the logic for statements of any length, but it
-appears that the faulty logic that the RNN did learn decayed similarly
-with expression size.
-This is not necessarily a significant flaw in the model, since it remains possible that  different training regimes would allow it to learn accurate representations of the larger expressions.
-
-In an effort to better understand why the RNTN outperformed the plain
-RNN, we evaluated both models on pairs of long formulae involving
-binary connectives, to assess the extent to which they achieve
-substantially different representations for semantically distinct
-formulae and substantially similar representations for synonymous
-formulae. We found that both models are able to identify synonymous
-pairs like $(a \pland a)$ and $(a \pland (a \pland (a \pland a)))$,
-with both \ii{and} and \ii{or}, but only the RNTN is able to learn
-substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
- and $(a \pland (a \pland b))$ when the difference between the two is embedded under multiple operators. 
- Figure~\ref{prop-falloff} summarizes this result.
-
-% we discovered that this model looses information about in longer
-% statements in a particularly problematic way. It appears that that
-% model is unable to distinguish between two sentences when the only
-% difference between those sentences is embedded within a very deep
-% structure. We evaluated both models on sentences that differed in only
-% one term, but for which that one term was embedded under a large
-% number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and
-%   b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and
-%   b)))}. We then measured the Euclidean distance between the vector
-% representations of the two sentences in each pair. Our findings are
-% shown in Figure~\ref{prop-falloff}, and show that while the RNTN can
-% distinguish the two sentences well at every size that we test, the RNN
-% fails after a depth of approximately six.
-
-%\begin{figure}[htp]
-%  \centering
-%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-%  \caption{Semantically distinct formulae should have different
-%    representations. As measured by Euclidean distance, only the RNTN
-%    achieves this for formulae containing more than a small number of
-%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
-%    failing to capture the meaning contrasts.}
-%  \label{prop-falloff}
-%\end{figure}
-
 \section{Reasoning with natural language quantifiers and negation}\label{sec:quantifiers}
@@
 
-% \subsection{Data}
-
-This experiment replicates similar work described in
-\cite{bowman2013can}, which found that RNTNs can learn to reason well
-with quantifier meanings given sufficient training data. This paper
-replaces the partially manually annotated data in that paper with data
-that is generated directly using the logical system that we hope to
-learn, yielding results that we believe to be substantially more
-straightforward to interpret.
+% TODO: We generate clean data from a grammar.
 
@@
 
-%\begin{table}[htp]
-%  \centering
-%  \begin{tabular}{l c l}
-%    \toprule
-%    (most warthogs) walk         & $\natneg$ & (not-most warthogs) walk\\
-%    (most mammals) move          & $\natind$ &  (not-most (not turtles)) move\\
-%    (most (not pets)) (not swim) & $\natrev$ & (not-most (not pets)) move 
-%    \\[2ex]    
-%    (no turtles) (not growl)     & $\natalt$ & (no turtles) (not swim)\\
-%    (no warthogs) swim           & $\natrev$ & (no warthogs) move\\
-%    (no warthogs) move           & $\natfor$ & (no (not reptiles)) swim\\
-%    \bottomrule
-%  \end{tabular}
-%  \caption{Sample data involving two different quantifier pairs.}
-%  \label{examplesofdata}
-%\end{table}
+\begin{table*}[htp]
+  \centering
+  \begin{tabular}{lcl}
+    \toprule
+    (all turtles) walk      &    $\natind$  & (less-than-three turtles) walk\\
+    (three warthogs) move  &        $\natneg$ & (less-than-three warthogs) move\\
+    (all mammals) (not move) &         $\natalt$ & (some warthogs) swim\\
+    (some (not pets)) swim     &     $\natcov$  & (all (not pets)) (not move)\\
+    \bottomrule
+  \end{tabular}
+  \caption{\label{examplesofdata}Examples of pairs of artificial setences with quantifiers.}
 
+\end{table*}
+
+% TODO: Rewrite experimental setup.
 We evaluate the model using two experimental settings. In the simpler
@@
 %
-While our primary interest is in discovering the extent to which our
-models can learn to encode the logic given an arbitrary amount of
-data, we are also interested in the degree to which they can infer a
-correct representation for the logic from more constrained training
-data. To this end, we segment the sentence pairs according to which
-quantifiers appear in each pair, and then hold out one such pair for
-testing. We hypothesize that a model that can efficiently learn to
-represent a logic should be able to construct an accurate
-representation of each of the two quantifiers in the pair from the way that it
-interacts with the other nine quantifiers with which it is seen in training. 
-Since running this experiment requires choosing a pair of
-quantifiers to hold out before training, the resource demands of
-training prevent us from testing each of the 55 possible 
-pairs of quantifiers, so we choose only four pairs to test on.  Three
-of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
-\ii{all}/\ii{some}) were chosen because they allow for the most
-different class labels in the test data. The fourth is a
-self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
-handles equality.
+In this configuration the model ...
 
@@
   \setlength{\tabcolsep}{10pt}
-  \begin{tabular}{ l rr@{ \ }r r@{ \ }r r@{ \ }r }
+  \begin{tabular}{ r@{ \ }r r@{ \ }r r@{ \ }r }
     \toprule
-    Data & \multicolumn{3}{c}{Most frequent class} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
+    \multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
     \midrule
-    \textsc{all split}	            &~~~~& 35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}
-    \\[1ex]    
-    \textsc{pair two/less-than-two} && 59.8 &(18.4) & 77.2 &(57.4) &	\textbf{100.0} &\textbf{(100.0)} \\
-    \textsc{pair not-all/not-most}  &&    0 &(0)    & 66.0 &(54.3) &	\textbf{93.8}  &\textbf{(91.5)} \\
-    \textsc{pair all/some}	    &&    0& (0)    & 62.7 &(65.9)  &	\textbf{78.4}  &\textbf{(84.7)} \\
-    \textsc{pair no/no}	            && 30.8  &(10.2) & 67.8 & (61.8) &	\textbf{99.9}  &\textbf{(99.7)} \\
+    35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}\\
     \bottomrule
   \end{tabular}
-  \caption{Performance on the quantifier experiments. Results are reported as accuracy scores followed by macroaveraged F1 scores in parentheses.}
-  \label{resultstable}
+  \caption{Performance on the quantifier experiments.}
+  \label{qresultstable}
 \end{table}
 
-\paragraph{Results} 
+\paragraph{Results} % TODO: Rewrite!
 The results for these experiments are shown in
-Figure~\ref{resultstable}. We compare the results to a most frequent
+Table~\ref{qresultstable}. We compare the results to a most frequent
 class baseline, which reflects the frequency in the test data of the
@@
 
-%train rnn 0.20302, 0.017195
-
-The RNN performed poorly at this task, even though the sentences used
-in these examples are short enough to avoid the pathology shown in
-Figure~\ref{prop-falloff}.  However, the perfect performance by the
-RNTN on the \textsc{all split} and \textsc{pair two/less-than-two}
-experiments, and its strong performance on the \textsc{pair no/no}
-experiment, suggest that this stronger model is able to handle
-quantifiers correctly given sufficient training data, but the weaker
-results on two of the training settings suggest that it may not be
-able to generalize from this kind of restricted training data in general. However, though the question
-of how much data is necessary to accurately capture quantifier
-behavior in a na\"ive model remains open, the fact that this model is able to perform so well is promising.
-
-\section{General discussion}\label{sec:discussion}
-
-This paper evaluated two recursive models on a series of three increasingly
-challenging interpretive tasks involving natural language inference:
-the core relational algebra of natural logic with entailment and
-exclusion; recursive propositional logic structures; and statements
-involving quantification and negation. The results suggest that RNTNs,
-but not plain RNNs, have the capacity to meet the challenges of these
-tasks with reasonably-sized training sets. These positive results are
-promising for the future of learned representation models in the
-applied modeling of compositional semantics.
-
-Of course, challenges remain. In terms of our experimental data, even
-the RNTN falls short of perfection in our more complex tasks, with
-performance falling off steadily as the depth of recursion grows. It
-remains to be seen whether these deficiencies can be overcome with
-improvements to the model, the optimization procedures, or the
-linguistic representations
-\cite{sochergrounded,kalchbrenner2014convolutional}. In addition,
-there remain subtle questions about how to fairly assess whether these
-models have truly generalized in the way we want them to. There is a
-constant tension between showing the models training data that gives
-them a chance to learn the target logical functions and revealing the
-answer to them in a way that leads to overfitting. The underlying
-logical theories provide only limited guidance on this point.
-%
-%, and the fact that there is a finite universe of possible
-%  expressions makes his an unavoidable issue. 
-%
-%  CP: I don't understand the above. It is false if taken literally;
-%  our PL generates an infinte number of formulae.
-%
-Finally, we have only scratched the surface of the logical complexity
-of natural language; in future experiments, we hope to test sentences
-with embedded quantifiers, multiple interacting quantifiers, relative
-clauses, and other kinds of recursive structure. Nonetheless, the
-rapid progress the field has made with these models in recent years
-provides ample reason to be optimistic that they can be trained to
-meet the challenges of natural language semantics.
-
-% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators.
-This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic.
-The positive results that we have found, however, are extremely promising for the future of learned representation models in the applied modeling of meaning.
-We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators.
-We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.
-
-% There is ample room to build on these results.
-In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses.
-In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.
-
-% \textit{Withheld for review.}\\ \\
-% Do later: Fill in, cite grants, internal reviewers
-
-\bibliographystyle{unsrtnat}
-
-\small % Note: Explicitly allowed in style guide
-\bibliography{MLSemantics}
-
-\end{document}
+Discussion discussion. 
+Discussion discussion. 
+Discussion discussion. 
+Discussion discussion. 
+Discussion discussion. 
+Discussion discussion. 
+Discussion discussion. 
+Discussion discussion. 
+Discussion discussion.

### 2014-11-23 writing/F'14 paper/recursion.tex
Paper first push; init; naming; new data; PBS

--- 2014-06-06 writing/S'14 paper/recursion.tex
@@
 represent a truth-functionally complete classical propositional logic:
-each atomic variable denotes either $\True$ or $\False$, and the only
+each atomic symbol is a variable over the domain \{$\True$, $\False$\}, and the only
 operators are truth-functional ones.  Table~\ref{tab:pl} defines this
@@
   \end{subtable}
-  \quad
   \begin{subtable}[t]{0.45\textwidth}
-    \centering
+    \centering\vspace{4mm} % TODO: Is there a better command to ensure padding above the table?
     \begin{tabular}[t]{r c l}
@@
   \end{subtable}
-  \caption{Propositional logic with natural logic relations.}  
+  \caption{Natural logic relations over sentences of propositional logic.}  
+  % TODO: Regenerate figure from final model.
   \label{prop-figure}
@@
 % Hopefully this is explicit enough.
-Socher et al.~\cite{socher2012semantic} show that a matrix-vector RNN
+Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector RNN
 model somewhat similar to our RNTN can learn a logic, but the logic
@@
 unknown truth values of propositions. As opposed to the two-way
-contrasts seen in \cite{socher2012semantic}, this logic distinguishes
+contrasts seen in \newcite{socher2012semantic}, this logic distinguishes
 between $2^{6} = 64$ possible assignments of truth values, and
@@
 
-\begin{figure}[tp]
+\begin{figure*}[t]
   \centering
   \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
+%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
   \caption{Model performance on propositional logic, by expression size. 
-    The dashed line indicates that only expressions of size four or less appeared in the training data.}  
+%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
+%    representations. As measured by Euclidean distance, only the RNTN
+%    achieves this for formulae containing more than a small number of
+%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
+%    failing to capture the meaning contrasts.
+    }  
   \label{prop-results}
-\end{figure}
+  % TODO: Regenerate figure from final model.
+\end{figure*}
 
@@
 substantially different representations for semantically distinct
-formulae and substantially similar representations for synonymous
-formulae. We found that both models are able to identify synonymous
-pairs like $(a \pland a)$ and $(a \pland (a \pland (a \pland a)))$,
-with both \ii{and} and \ii{or}, but only the RNTN is able to learn
+formulae. We found that only the RNTN is able to learn
 substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
- and $(a \pland (a \pland b))$ when the difference between the two is embedded under multiple operators. 
- Figure~\ref{prop-falloff} summarizes this result.
+ and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
+The RNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed at under three operators, and quickly approaches zero as depth increases.
+ The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.
 
@@
 % fails after a depth of approximately six.
-
-\begin{figure}[htp]
-  \centering
-  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-  \caption{Semantically distinct formulae should have different
-    representations. As measured by Euclidean distance, only the RNTN
-    achieves this for formulae containing more than a small number of
-    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
-    failing to capture the meaning contrasts.}
-  \label{prop-falloff}
-\end{figure}

### 2014-11-23 writing/F'14 paper/sick.tex
Paper first push; init; naming; new data; PBS

\section{The SICK textual entailment challenge}

What SICK is. \cite{marelli2014sick}
\\ \\ \\ \\

\begin{table*}[htp]
  \centering
  \begin{tabular}{lcl}
    \toprule
    (all turtles) walk      &    $\natind$  & (less-than-three turtles) walk\\
    (three warthogs) move  &        $\natneg$ & (less-than-three warthogs) move\\
    (all mammals) (not move) &         $\natalt$ & (some warthogs) swim\\
    (some (not pets)) swim     &     $\natcov$  & (all (not pets)) (not move)\\
    \bottomrule
  \end{tabular}
  \caption{\label{examplesofsickdata}Examples of successful classifications on SICK.}
\end{table*}

We initialized our word embeddings using precomputed TODO dimensional vectors from 
(TODO: Where?). Before any embedding is used as an input to a recursive layer, 
it is passed through an additional $\tanh$ neural network layer that transforms
the embedding from the embedding
dimension to the dimension of the recursive layer.

We supplemented the roughly 5000 SICK training examples\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with training data from (CITE CITE CITE CITE CITE CITE CITE CITE CITE CITE CITE). We trained a single model on data from all sources, but used a separate set of softmax parameters for classifying into the labels from each source.

\begin{table}[tp]
  \centering
  \setlength{\tabcolsep}{10pt}
  \begin{tabular}{ r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    \multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
    \midrule
    56.7 &(...) &	...&(...)& \textbf{...} & \textbf{(...)}\\
    \bottomrule
  \end{tabular}
  \caption{Performance on the SICK.}
  \label{sresultstable}
\end{table}

Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}.

% TODO: Last runs, train on dev.

### 2014-11-23 InitializeModel.m

@@ -16,52 +16,62 @@ else
     NUMCOMP = 3;
 end
 
-SCALE = hyperParams.initScale;
-OFFSET = 2 * hyperParams.initScale;
+SCALE = 0.01 * (30 / DIM);
+TSCALE = hyperParams.initScale * SCALE;
 
 % Randomly initialize softmax layer
-classifierParameters = rand(sum(hyperParams.numRelations), PENULT + 1) .* OFFSET - SCALE;
+% SCALE = 1 / sqrt(PENULT + 1);
+classifierParameters = rand(sum(hyperParams.numRelations), PENULT + 1) .* (2 * SCALE) - SCALE;
 
 % Randomly initialize tensor parameters
+% SCALE = (1 / (2 * sqrt(DIM))) * hyperParams.initScale;
 if hyperParams.useThirdOrderComparison
-    classifierMatrices = rand(DIM, DIM, PENULT) .* OFFSET - SCALE;
+    classifierMatrices = rand(DIM, DIM, PENULT) .* (2 * TSCALE) - TSCALE;
 else
     classifierMatrices = rand(0, 0, PENULT);
 end
-classifierMatrix = rand(PENULT, DIM * 2) .* OFFSET - SCALE;
+% SCALE = 1 / (2 * sqrt(2 * DIM));
+classifierMatrix = rand(PENULT, DIM * 2) .* (2 * SCALE) - SCALE;
 classifierBias = zeros(PENULT, 1);
 
+% SCALE = (1 / (2 * sqrt(DIM))) * hyperParams.initScale;
 if hyperParams.useThirdOrder
-    compositionMatrices = rand(DIM, DIM, DIM, NUMCOMP) .* OFFSET - SCALE;
+    compositionMatrices = rand(DIM, DIM, DIM, NUMCOMP) .* (2 * TSCALE) - TSCALE;
 else
     compositionMatrices = zeros(0, 0, 0, NUMCOMP);
 end
-compositionMatrix = rand(DIM, DIM * 2, NUMCOMP) .* OFFSET - SCALE;
-for i = 1:NUMCOMP
-  compositionMatrix(:, :, i) = compositionMatrix(:, :, i) + [eye(DIM) eye(DIM)];
+% SCALE = 1 / (2 * sqrt(2 * DIM));
+compositionMatrix = rand(DIM, DIM * 2, NUMCOMP) .* (2 * SCALE) - SCALE;
+if hyperParams.useEyes
+  for i = 1:NUMCOMP
[…]

### 2014-11-23 RunExperiments.m

@@ -501,3 +501,12 @@ echo "cd quant; lambda = 0.0003; dim = 30; ed = 200; td = 1; train = 1; penult =
 echo "cd quant; lambda = 0.0003; dim = 75; ed = 200; td = 1; train = 1; penult = 90; lr = 0.020; loadw = 1; dropout = [0.75, 0.9]; mult = 45; collo = 1; scale = 0.1; mbs = 32; dataflag='sick-plus-100k'; name=[dataflag, num2str(lambda), 'r-dim', num2str(dim), '-ed', num2str(ed), '-td', num2str(td), '-tr', num2str(train), '-pen', num2str(penult), '-lr', num2str(lr), '-loadw', num2str(loadw), '-do', num2str(dropout(1)), ':', num2str(dropout(2)), '-co', num2str(collo), '-m', num2str(mult), '-sc', num2str(scale), '-mb', num2str(mbs)]; disp(name); TrainModel('', name, 1, @Sick, dataflag, ed, dim, td, penult, lambda, 0, 0, mbs, lr, train, loadw, dropout(1), dropout(2), mult, collo, scale);" | /afs/cs/software/bin/matlab_r2012b
 
 echo "cd quant; lambda = 0.0003; dim = 30; ed = 200; td = 1; train = 1; penult = 75; lr = 0.020; loadw = 1; dropout = [0.75, 0.9]; mult = 270; collo = 1; scale = 0.1; mbs = 32; dataflag='sick-plus-600k'; name=[dataflag, num2str(lambda), '-dim', num2str(dim), '-ed', num2str(ed), '-td', num2str(td), '-tr', num2str(train), '-pen', num2str(penult), '-lr', num2str(lr), '-loadw', num2str(loadw), '-do', num2str(dropout(1)), ':', num2str(dropout(2)), '-co', num2str(collo), '-m', num2str(mult), '-sc', num2str(scale), '-mb', num2str(mbs)]; disp(name); TrainModel('', name, 1, @Sick, dataflag, ed, dim, td, penult, lambda, 1, 0, mbs, lr, train, loadw, dropout(1), dropout(2), mult, collo, scale);" | /afs/cs/software/bin/matlab_r2012b
+
[…]

### 2014-11-23 run.sh

@@ -1,34 +1,13 @@
-#!/bin/sh
-### Script to launch an experiment on PBS.
-### Work in progress.
+#/bin/bash
+#PBS -l nodes=1:ppn=4 	### Request at least 4 cores
+#PBS -l walltime=12:00:00	### Die after 12h
+#PBS -l mem=6000MB
+#PBS -W x=NACCESSPOLICY:SINGLEJOB	### Use all cores on node
+#PBS -n	### Use all cores on node (alternate specification)
+#PBS -m n
+#PBS -q nlp
 
-### Job name
-#PBS -N sbowman_pbs_test
-### Output files
-###PBS -o hello_world_job.stdout
-###PBS -e hello_world_job.stderr
-### Queue name 
-###PBS -q dqueue
-### Number of nodes 
-#PBS -l nodes=4:compute#shared
-# Print the default PBS server
-echo PBS default server is $PBS_DEFAULT
-# Print the job's working directory and enter it.
-echo Working directory is $PBS_O_WORKDIR
-cd $PBS_O_WORKDIR
-# Print some other environment information
-echo Running on host `hostname`
-echo Time is `date`
-echo Directory is `pwd`
-echo This jobs runs on the following processors:
-NODES=`cat $PBS_NODEFILE`
-echo $NODES
-# Compute the number of processors
-NPROCS=`wc -l < $PBS_NODEFILE`
-echo This job has allocated $NPROCS nodes
-# Run hello_world
-for NODE in $NODES; do
- ssh $NODE cat < "ABCDE" &
-done
-# Wait for background jobs to complete.
-wait 
+# [Don't] change to the submission directory
+cd $PBS_O_WORKDIR 
+echo $MATLABCMD 
+echo $MATLABCMD | /afs/cs/software/bin/matlab_r2012b -nodisplay

### 2014-11-24 writing/F'14 paper/discussion.tex
Provisional AdaDelta support; cleanup.

@@
 
-% TODO: Cite deep RNNs
-% TODO: Cite LSTMs
-
 Of course, challenges remain. In terms of our experimental data, even
@@
 remains to be seen whether these deficiencies can be overcome with
-improvements to the model, the optimization procedures, or the
-linguistic representations
-\cite{sochergrounded,kalchbrenner2014convolutional,}. In addition,
+stronger models---either extended recursive models 
+\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep} or recurrent
+ones \cite{sutskever2014sequence}---or the optimization procedures. In addition,
 there remain subtle questions about how to fairly assess whether these
@@
 In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses.
-In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.
+In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (-), or convolutional RNNs \cite{kalchbrenner2014convolutional}.

### 2014-11-24 AdaDeltaUpdate.m

@@ -1,0 +1,40 @@
+function modelState = AdaDeltaUpdate(modelState, options, grad, embGrad)
+
+if modelState.step == 0
+    modelState.sumSqGrad = zeros(size(modelState.theta));
+    modelState.sumSqDelta = zeros(size(modelState.theta));
+    if length(embGrad) > 0
+        % Set up a separate SubSqGrad tracker for the embeddings.
+        modelState.sumSqEmbGrad = zeros(size(modelState.separateWordFeatures));
+        modelState.sumSqEmbDelta = zeros(size(modelState.separateWordFeatures));
+    end
+end
+
+% Do an AdaGrad-scaled parameter update
+modelState.sumSqGrad = modelState.sumSqGrad .* options.adaDeltaRho + (grad.^2) .* (1 - options.adaDeltaRho);
+
+rmsLastDelta = sqrt(modelState.sumSqDelta + options.adaDeltaEps);
+rmsGrad = sqrt(modelState.sumSqGrad + options.adaDeltaEps);
+
+delta = -1 * (rmsLastDelta ./ rmsGrad) .* grad;
+modelState.sumSqDelta = modelState.sumSqDelta .* options.adaDeltaRho + (delta.^2) .* (1 - options.adaDeltaRho);
+
+modelState.theta = modelState.theta + delta;
+
+assert(sum(isnan(modelState.theta)) == 0, 'NaNs in theta.');
+assert(sum(isinf(modelState.theta)) == 0, 'Infs in theta.');
+
+% Do an AdaGrad-scaled parameter update to the separate word features
+if length(embGrad) > 0
+    modelState.sumSqEmbGrad = modelState.sumSqEmbGrad .* options.adaDeltaRho + (embGrad.^2) .* (1 - options.adaDeltaRho);
+
+    rmsLastEmbDelta = sqrt(modelState.sumSqEmbDelta + options.adaDeltaEps);
+    rmsEmbGrad = sqrt(modelState.sumSqEmbGrad + options.adaDeltaEps);
+
+    embDelta = -1 * (rmsLastEmbDelta ./ rmsEmbGrad) .* embGrad;
+    modelState.sumSqEmbDelta = modelState.sumSqEmbDelta .* options.adaDeltaRho + (embDelta.^2) .* (1 - options.adaDeltaRho);
+
+    modelState.separateWordFeatures = modelState.separateWordFeatures + embDelta;
+end
+
+end

### 2014-11-24 Log.m

@@ -10,7 +10,6 @@ else
 	try
 		fprintf(file, '%s\n', [datestr(clock, 0) ': ' message]);
 	catch err
-		% TODO: Fix logging from inside parfor
 	end
 end
 

### 2014-11-24 RunExperiments.m

@@ -61,8 +61,6 @@ echo "cd quant; lambda = 0.005; lr = 0.1; ed = 0; mbs = 64; name=['wr-tfrozen-',
 
 echo "cd quant; lambda = 0.000001; lr = 0.1; mbs = 16; trainwords = true; ed = 0; name=['sick-plus-demo-', num2str(lr), '-', num2str(lambda), '-', num2str(mbs), '-', num2str(ed), '-', num2str(trainwords)]; disp(name); TrainModel('sick-plus', '', name, mbs, 25, 80, lr, lambda, 1, ed, trainwords);" | /afs/cs/software/bin/matlab_r2012b
 
-# TODO: Turn on word loading
-# TODO: try not training words?
 function TrainModel(dataflag, pretrainingFilename, expName, mbs, dim, penult, lr, lambda, tot, transDepth, trainwords)
 
 echo "cd quant; lambda = 0.0005; ed = 0; train = 0; penult = 80, name=['wr-tfrozen-l', num2str(lambda), '-ed', num2str(ed), '-tr', num2str(train), '-pen', num2str(penult)]; disp(name); TrainModel('word-relations', '', name, 64, 25, penult, 0.1, lambda, 1, ed, train);" | /afs/cs/software/bin/matlab_r2012b

### 2014-11-24 RunExperiments.sh

@@ -576,6 +576,87 @@ export MATLABCMD="cd quant; lambda = 0.0003; dim = 35; ed = 50; td = 1; train = 
 export MATLABCMD="cd quant; lambda = 0.0003; dim = 35; ed = 50; td = 1; train = 1; penult = 75; lr = 0.02; loadw = 1; dropout = [0.75\, 0.9]; rtemult = 5; nlimult = 100; mult = 150; collo = 2; tscale = 0.1; wscale = 0.01; eyes = 1; mbs = 64; dataflag='sick-plus-100k'; name='sc'; TrainModel(''\, 1\, @Sick\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, 1\, 0\, mbs\, lr\, train\, loadw\, dropout(1)\, dropout(2)\, mult\, rtemult\, nlimult\, collo\, tscale\, wscale\, eyes);"; qsub -v MATLABCMD quant/run.sh -q john
 export MATLABCMD="cd quant; lambda = 0.0003; dim = 35; ed = 50; td = 1; train = 1; penult = 75; lr = 0.02; loadw = 1; dropout = [0.75\, 0.9]; rtemult = 30; nlimult = 80; mult = 50; collo = 2; tscale = 1; wscale = 0.05; eyes = 1; mbs = 64; dataflag='sick-plus-100k'; name='sc'; TrainModel(''\, 1\, @Sick\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, 1\, 0\, mbs\, lr\, train\, loadw\, dropout(1)\, dropout(2)\, mult\, rtemult\, nlimult\, collo\, tscale\, wscale\, eyes);"; qsub -v MATLABCMD quant/run.sh -q john
 
-echo "cd quant;TrainModel('', 1, @GradCheck, 1, 1, 1, 0, 1, 0);" | /afs/cs/software/bin/matlab_r2012b
[…]

### 2014-11-24 TrainModel.m

@@ -105,7 +105,7 @@ if hyperParams.minFunc
     modelState.theta = minFunc(@ComputeFullCostAndGrad, modelState.theta, options, ...
         modelState.thetaDecoder, trainDataset, modelState.separateWordFeatures, hyperParams, testDatasets);
 else
-    modelState.theta = AdaGradSGD(@ComputeFullCostAndGrad, modelState, options, ...
+    modelState.theta = TrainSGD(@ComputeFullCostAndGrad, modelState, options, ...
         trainDataset, hyperParams, testDatasets);
 end
     

### 2014-12-01 writing/F'14 paper/intro.tex
Provisional AdaDelta support; cleanup.
Big paper push; SICK data updates; hacks

@@
 
-% TODO: Reskim Gref paper.
-
 % TODO: Any way to avoid nested parens?
@@
 generate controlled data sets encoding the semantic relationships
-between pairs of expressions and evaluate whether this model can 
+between pairs of expressions and evaluate whether models with this architecture can 
 learn to correctly identify the relationships between novel pairs.
@@
 statements like \ii{every reptile walks} and \ii{no turtle moves}.
-We find that the plain RNN achieves only mixed results in all three
+We find that the plain recursive NN achieves only mixed results in all three
 experiments, whereas the stronger RNTN model generalizes well in every
@@
 If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them.
-% TODO: How much should we say about this?
+%  How much should we say about this?
 

### 2014-12-01 writing/F'14 paper/join.tex
Big paper push; SICK data updates; hacks

@@
 
-% We test the model's ability to learn this behavior by creating
-% artificial data sets of terms which represent sets of numbers. Since
-% MacCartney and Manning's set of relations hold between sets as well as
-% between sentences, we can use the underlying set structure to generate
-% the all of the relations that hold between any pair of these terms, as
-% in Figure \ref{lattice-figure}. We train the model defined above on a
-% subset of these relations, but rather then presenting the model with a
-% pair of tree-structured sentences as inputs, simply present it with
-% two single terms, each of which corresponds to a single vector in the
-% (randomly initialized) vocabulary matrix $V$, ensuring that the model
-% has no information about the terms being compared except the relations
-% between them.
-
 In our experiments, we create 80 randomly generated sets drawing from
@@
 
-\paragraph{Results} 
-TODO: Explain Table \ref{joinresultstable}.
+\begin{table}[tp]
+  \centering \small
+  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
+    \toprule
+    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{??d RNN}  & \multicolumn{2}{c}{??d RNTN}\\
+    \midrule
+    Provable test &52.0 &(10.3) &	-&(-)& \textbf{-} & \textbf{(-)}\\
+    All test &53.8 &(10.5) &	-&(-)& \textbf{-} & \textbf{(-)}\\
+    Train &53.8 &(10.5)    & -&(-)& \textbf{-} & \textbf{(-)}\\
 
-\begin{table}[tp]
-  \centering
-  \setlength{\tabcolsep}{10pt}
-  \begin{tabular}{ r@{ \ }r r@{ \ }r r@{ \ }r }
-    \toprule
-    \multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
-    \midrule
-    35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}\\
     \bottomrule
   \end{tabular}
-  \caption{Performance on the TODO experiments. Results are reported as accuracy scores followed by macroaveraged F1 scores in parentheses.}
+  \caption{Performance on the semantic relation experiments.
+These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
   \label{joinresultstable}
-\end{table} % TODO: Is this worth the fair amount of space it takes up?
+\end{table}
 
@@
 
-% RNTN log: tue-j-11-6x80-hip.txt
-% Train PER: 0.0021875
-
-% RNN log: tue-j-11-6x80-r-75.txt
-% Train PER: 0.047188
-
-% and create a dataset consisting of the relations between every pair of
-% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
-% test dataset, and that test dataset was further split into the 2960
-% examples that can be provably derived from the test data using
-% MacCartney and Manning's join table (or by the symmetry of the
-% relations in about half of the cases) and the 240 that
-% cannot.
-
-% We tested a version of both the RNN model and the RNTN model on these
-% data. In both cases, the models were implemented exactly as described
-% in Section~\ref{methods}, but since the items being compared are
-% single terms rather than full tree structures, the composition layer
-% was not used, and the two models differed only in which layer function
-% was used for the comparison layer. We found that the RNTN model worked
-% best with 11 dimensional vector representations for the 80 sets and a
-% 90 dimensional feature vector for the classifier. This model was able
-% to correctly label 99.3\% of the derivable test examples, and 99.1\%
-% of the remaining examples. The simpler RNN model worked best with 11
-% and 75 dimensions, respectively, but was able to achieve accuracies of
-% only 90.0\% and 87.\%, respectively.
-
-% TODO: Write out results.
 % TODO: Do we still care about the unprovable test examples?
 
-These results are fairly straightforward to interpret. The RNTN model
+\paragraph{Results} 
+Table \ref{joinresultstable} shows the results, which are fairly straightforward 
+to interpret. The RNTN model
 was able to accurately encode the relations between the terms in the

### 2014-12-01 writing/F'14 paper/methods.tex
Big paper push; SICK data updates; hacks

@@
 The resulting vectors are fed into a separate comparison
-layer that is meant to generate a feature vector capturing the
+layer (fixed at 75 dimensions) that is meant to generate a feature vector capturing the
 relation between the two phrases. The output of this layer is then
@@
 \begin{gather} \label{rnn}
-\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
+\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\
 \label{rntn}
@@
 them by an $n \times 2n$ matrix of learned weights, adds a bias $\vec{b}$, and applies the
-element-wise non-linearity to the resulting vector. The RNTN adds a learned third-order tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
+element-wise non-linearity to the resulting vector. The RNTN adds a learned third-order tensor 
+$\mathbf{T}$, dimension $n \times n \times n$, modeling
 multiplicative interactions between the child vectors.
-
-% TODO: Kill ReLU?
 
@@
 We train the model using stochastic gradient descent (SGD)
-with learning rates computed using AdaGrad \cite{duchi2011adaptive}.
+with learning rates computed using ADADELTA \cite{zeiler2012adadelta}.
 The parameters (including the vocabulary) are initialized randomly.
@@
 
-% TODO: Describe mean-of-five scheme for results reporting.
-
 \begin{figure*}[tp]

### 2014-12-01 writing/F'14 paper/quantifiers.tex
Provisional AdaDelta support; cleanup.
Big paper push; SICK data updates; hacks

@@
 language inference \cite{Icard:Moss:2013:LILT}.
-
-% TODO: We generate clean data from a grammar.
 
@@
 % TODO: Rewrite experimental setup.
-We evaluate the model using two experimental settings. In the simpler
-setting, \textsc{all split}, we randomly sample 85\% of the data and evaluate on the
-remaining 15\%. In this setting, the model is being asked to learn a
+We evaluate the model by randomly sampling 20\% of the data to
+test on and training on the remainder. In this setting, the model is being asked to learn a
 complete reasoning system for the limited language and logic presented
@@
 %
-In this configuration the model ...
-
 \begin{table}[tp]
-  \centering
-  \setlength{\tabcolsep}{10pt}
-  \begin{tabular}{ r@{ \ }r r@{ \ }r r@{ \ }r }
+  \centering \small
+  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
     \toprule
-    \multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
+    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{??d RNN}  & \multicolumn{2}{c}{??d RNTN}\\
     \midrule
-    35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}\\
+    Test & 35.4 &(7.5) &	99.1&(99.2)& \textbf{-} & \textbf{(-)}\\
+    Train & 35.4 &(7.5) &	-&	-&	-&	-\\	
     \bottomrule
@@
 \end{table} 
+%
+% do not allow a blank line --- adds too much space
+%
+\paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic well.
 
-\paragraph{Results} % TODO: Rewrite!
-The results for these experiments are shown in
-Table~\ref{qresultstable}. We compare the results to a most frequent
-class baseline, which reflects the frequency in the test data of the
-most frequent class in the training data, $\natind$.  After some
-cross-validation, we chose 16- and 20-dimensional representations for
-the RNN and RNTN, respectively, and 75-dimensional feature vectors for
-the classifier. Training
-accuracy was near 80\% for the each of the five RNN experiments, and 
-near 98\% for each of the five RNTN experiments.
-
-Discussion discussion. 
-Discussion discussion. 
-Discussion discussion. 
-Discussion discussion. 
-Discussion discussion. 
-Discussion discussion. 
-Discussion discussion. 
-Discussion discussion. 
-Discussion discussion.
+% TODO: More to say?

### 2014-12-01 writing/F'14 paper/sick.tex
Big paper push; SICK data updates; hacks

@@
 
-What SICK is. \cite{marelli2014sick}
-\\ \\ \\ \\
+The tree-pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference over real natural language.
+To ensure that this is the case, we also train a version of our model on an existing competition dataset, the ICK textual entailment challenge \cite{marelli2014sick}.
+The corpus consists of about 10k natural language sentence pairs (generated using data expansion from a smaller set), labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}.
+At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest hand labeled entailment corpus that we are aware of.
 
@@
   \centering
-  \begin{tabular}{lcl}
+  \begin{tabular}{l}
     \toprule
-    (all turtles) walk      &    $\natind$  & (less-than-three turtles) walk\\
-    (three warthogs) move  &        $\natneg$ & (less-than-three warthogs) move\\
-    (all mammals) (not move) &         $\natalt$ & (some warthogs) swim\\
-    (some (not pets)) swim     &     $\natcov$  & (all (not pets)) (not move)\\
+example example example example example example example example example example example \\
+example example example example example example example example example example example \\
+ENTAILMENT A skilled person is riding a bicycle on one wheel	A person is riding the bicycle on one wheel\\
+CONTRADICTION The player is missing the basket and a crowd is in background	The player is dunking the basketball into the net and a crowd is in background\\
     \bottomrule
@@
 
-We initialized our word embeddings using precomputed TODO dimensional vectors from 
-(TODO: Where?). Before any embedding is used as an input to a recursive layer, 
+We initialized our word embeddings using the TODO dimensional vectors provided with 
+GloVe \cite{pennington2014glove}. Before any embedding is used as an input to a recursive layer, 
 it is passed through an additional $\tanh$ neural network layer that transforms
-the embedding from the embedding
-dimension to the dimension of the recursive layer.
+the embedding from the embedding dimension to the dimension of the recursive layer.
 
-We supplemented the roughly 5000 SICK training examples\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with training data from (CITE CITE CITE CITE CITE CITE CITE CITE CITE CITE CITE).
-We trained a single model on data from all sources, but used a separate set of softmax parameters for classifying into the labels from each source.
+We supplemented the roughly 5000 SICK training examples\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with TODO examples of approximate entailment data from the Denotation Graph project \cite{hodoshimage}.
+We parsed all data with the Stanford PCFG Parser v.
+3.3.1 \cite{klein2003accurate}.
+To accelerate convergence, we collapsed subtrees that were identical across both sentences in a pair down to a single head word.
+In order to improve regularization on the noisier data, we used dropout \cite{hinton2012improving} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).
+We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source.
 
 \begin{table}[tp]
-  \centering
-  \setlength{\tabcolsep}{10pt}
-  \begin{tabular}{ r@{ \ }r r@{ \ }r r@{ \ }r }
+  \centering \small
+  \begin{tabular}{ l r@{ \ } r@{ \ } r@{ \ } }
     \toprule
-    \multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
+    ~&\multicolumn{1}{c}{$\natind$ only} & \multicolumn{1}{c}{??d RNN}  & \multicolumn{1}{c}{??d RNTN}\\
     \midrule
-    56.7 &(...) &	...&(...)& \textbf{...} & \textbf{(...)}\\
+    Test & 56.7 &	...& \textbf{...} \\
+    Train &- &- &-  \\
     \bottomrule
   \end{tabular}
-  \caption{Performance on the SICK.}
+  \caption{Performance on SICK.}
   \label{sresultstable}
@@
 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}. 
-Explain Figure \ref{examplesofsickdata}.
+Despite the small amount of high quality training data available, it is possible to train our model to perform competitively on textual entailment.
+Our performance did not reach the state of the art (84.6\%), but exceeded the median submission to the competition (77.06\%) in a field including many models which used sophisticated hand-engineered features specific to the specific version of the entailment task at hand.
 
-% TODO: Last runs, train on dev.
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}. 
+Explain Table \ref{examplesofsickdata}.

### 2014-12-01 RunAndOrExperiments.sh

@@ -4,3 +4,44 @@ export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.0003; dim = 40;
 export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.0003; dim = 40; td = 2; penult = 75; dropout = 0.9; tot = 1; name='ta'; relu = 0; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout);" ; qsub -v MATLABCMD quant/run.sh -l nodes=1:ppn=8
 export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.0003; dim = 40; td = 1; penult = 75; dropout = 1; tot = 1; name='ta'; relu = 1; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout);" ; qsub -v MATLABCMD quant/run.sh -l nodes=1:ppn=8
 export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.0003; dim = 40; td = 1; penult = 75; dropout = 0.9; tot = 1; name='ta'; relu = 0; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout);" ; qsub -v MATLABCMD quant/run.sh -l nodes=1:ppn=8
+
+export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.0003; dim = 36; td = 1; penult = 75; dropout = 1; tot = 1; name='ta'; relu = 1; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout);" ; qsub -v MATLABCMD quant/run.sh -q john
+export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.0003; dim = 40; td = 1; penult = 75; dropout = 1; tot = 1; name='ta'; relu = 1; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout);" ; qsub -v MATLABCMD quant/run.sh -q john
[…]

### 2014-12-01 writing/F'14 paper/sick.tex
Paper update, bug fixes.

@@
 \begin{table*}[htp]
-  \centering
-  \begin{tabular}{l}
+  \centering\small
+  \begin{tabular}{lcl}
     \toprule
-example example example example example example example example example example example \\
-example example example example example example example example example example example \\
-ENTAILMENT A skilled person is riding a bicycle on one wheel	A person is riding the bicycle on one wheel\\
-CONTRADICTION The player is missing the basket and a crowd is in background	The player is dunking the basketball into the net and a crowd is in background\\
+A woman is not boiling shrimps.& \ii{contradiction}&	A woman is boiling shrimps.\\
+A wild deer is jumping a fence. &\ii{entailment}	&A deer is jumping a fence.\\
+A potato is being sliced by a woman. &\ii{entailment}	&A woman is slicing a potato.\\
+The plane, which is south African, is flying in a blue sky.& \ii{entailment}&	An airplane is flying through the air.\\
     \bottomrule
@@
     \midrule
-    Test & 56.7 &	...& \textbf{...} \\
+    Test & 56.7 &	...& \textbf{75.5} \\
     Train &- &- &-  \\
@@
 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}. 
-Explain Table \ref{examplesofsickdata}.
+Table \ref{examplesofsickdata} shows a few typical examples of correct classifications.
+The model appears to be able to successfully recognize negation and the insertion and deletion of words, as demonstrated in the first two examples.
+The latter two show that the model still recognize entailments and contradictions between sentences which are not superficially similar, such as passive--active pairs and sentences with low lexical overlap.

### 2014-12-01 RunAndOrExperiments.sh

@@ -36,11 +36,11 @@ export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.001; dim = 45; 
 export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.001; dim = 45; td = 1; penult = 75; dropout = 1; tot = 0; name='taf'; relu = 1; TrainModel(''\, 4\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh 
 export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.001; dim = 45; td = 1; penult = 75; dropout = 1; tot = 0; name='taf'; relu = 1; TrainModel(''\, 5\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh 
 
-export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.0003; dim = 40; td = 1; penult = 75; dropout = 1; tot = 1; name='taf'; relu = 1; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh -q john
-export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.0003; dim = 40; td = 1; penult = 75; dropout = 1; tot = 1; name='taf'; relu = 1; TrainModel(''\, 2\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh -q john
-export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.0003; dim = 40; td = 1; penult = 75; dropout = 1; tot = 1; name='taf'; relu = 1; TrainModel(''\, 3\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh -q john
[…]

### 2014-12-01 writing/F'14 paper/sick.tex
SICK commentary update.

@@
 
-Despite the small amount of high quality training data available, it is possible to train our model to perform competitively on textual entailment.
+Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment.
 Our performance did not reach the state of the art (84.6\%), but exceeded the median submission to the competition (77.06\%) in a field including many models which used sophisticated hand-engineered features specific to the specific version of the entailment task at hand.

### 2014-12-01 writing/F'14 paper/abstract.tex
Revised abstract

@@
-\begin{abstract}
-Recursive neural network models for sentence meaning have been successful in many tasks, but it remains an open question whether any such model can learn compositional semantic representations that support logical deduction.
+% \begin{abstract}
+% Recursive neural network models for sentence meaning have been successful in many tasks, but it remains an open question whether any such model can learn compositional semantic representations that support logical deduction.
 We pursue this question directly by evaluating whether two such models can correctly learn to identify logical relationships such as entailment and contradiction between sentences.
@@
 The model performs competitively on the SICK data and generalizes well in all three experiments on simulated data, suggesting that it can learn suitable representations for logical inference in natural language.
+% \end{abstract}
+
+%%%% POTTS REVISION
+\begin{abstract} 
+  Recursive neural networks (RNNs) for sentence meaning have been
+  successful in many tasks, but it remains an open question whether
+  any such model can learn compositional semantic representations that
+  support logical deduction. We pursue this question by
+  evaluating whether two such models---RNNs and recursive neural
+  tensor networks (RNTNs)---can correctly learn to identify logical
+  relationships such as entailment and contradiction. % between sentences.
+  In our first set of experiments, we generate artificial data from a
+  logical grammar and use it to evaluate the models' ability to learn
+  to handle basic relational reasoning, recursive structures, and
+  quantification. We then evaluate the models on the more naturalistic
+  SICK challenge data. Both models perform competitively on the SICK
+  data and generalize well in all three experiments on simulated data,
+  suggesting that they can learn suitable representations for logical
+  inference in natural language. 
 \end{abstract}

### 2014-12-02 writing/F'14 paper/intro.tex
Paper update, bug fixes.
Revised intro

@@
 To date, experimental investigations of these concepts using
-distributed representations have been largely confined to short phrases
+distributed representations have been largely confined to short
+phrases
 \cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment}.
-For robust natural language understanding, it is essential to model
-these phenomena in their full generality on complex linguistic
-structures. Other recent work \cite{grefenstette2013towards,rocktaschellow} has
-developed well-understood frameworks that can compute the meanings of 
-certain fragments of natural language compositionally from vector or
-tensor representations. These frameworks are too structured to allow broad
-coverage meaning representations to be learned in
-supervised natural language tasks, and as such do not provide direct insight
-into the question of what existing applied machine learning models are 
-capable of.
+However, for robust natural language understanding, it is essential to
+model these phenomena in their full generality on complex linguistic
+structures. Other recent work
+\cite{grefenstette2013towards,rocktaschellow} has developed rich
+frameworks for computing the meanings of complex fragments of natural
+language compositionally from vector or tensor representations, but
+these approaches do not confront the question of how to learn their
+highly structured representations from data using existing applied
+machine learning models.
 
-% TODO: Any way to avoid nested parens?
+% TODO: Any way to avoid nested parens? Yes, use \citealt ---CP
 
 Our research uses the task of \ii{natural language
-  inference} (also known as \ii{recognizing textual entailment}
-\cite{dagan2006pascal}), in which the goal is to determine the core
+  inference} (also known as \ii{recognizing textual entailment}; 
+\citealt{dagan2006pascal}), in which the goal is to determine the core
 inferential relationship between two sentences. Much of the
-theoretical work on this task (and some successful implemented models
-\cite{maccartney2009extended,watanabe2012latent}) involves \ii{natural
+theoretical work on this task (and some successful implemented models; 
+\citealt{maccartney2009extended,watanabe2012latent}) involves \ii{natural
   logics}, which are formal systems that define rules of inference
@@
 need of intermediate representations in an artificial logical
-language. We use the natural logic developed by MacCartney and Manning 
-\shortcite{maccartney2009extended} 
-as our formal model. This logic defines seven core relations of synonymy, 
-entailment, contradiction, and mutual consistency, as summarized in
-Table~\ref{b-table}, and it provides rules of semantic combination for
-projecting these relations from the lexicon up to complex phrases. The
-formal properties and inferential strength of this system are now
-well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
+language. We use the natural logic developed by
+\newcite{maccartney2009extended} as our formal model. This logic
+defines seven core relations of synonymy, entailment, contradiction,
+and mutual consistency, as summarized in Table~\ref{b-table}, and it
+provides rules of semantic combination for projecting these relations
+from the lexicon up to complex phrases. The formal properties and
+inferential strength of this system are now well-understood
+\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
 
-% TODO: Do we need both of these Icard/Moss citations?
+% TODO: Do we need both of these Icard/Moss citations? -- I'd prefer
+% to keep them ---CP.
 
-Since no existing model architecture can be adopted without modification
-to handle entailment classification, we describe a novel method for this task
-in the spirit of prior work on recursive neural networks (\ref{methods}).
-In three experiments, we use this pre-specified logical grammar to
-generate controlled data sets encoding the semantic relationships
-between pairs of expressions and evaluate whether models with this architecture can 
-learn to correctly identify the relationships between novel pairs.
-In our first experiment (Section~\ref{sec:join}), we
-evaluate the ability of these models to learn the core relational
-algebra of natural logic from data. Our second experiment
-(Section~\ref{sec:recursion}) extends this evaluation to relations
-between complex recursive structures like $(a \plor b)$ and
-$\plneg (\plneg a \pland \plneg b)$, and our third experiment
-(Section~\ref{sec:quantifiers}) involves relations between quantified
-statements like \ii{every reptile walks} and \ii{no turtle moves}.
-We find that the plain recursive NN achieves only mixed results in all three
-experiments, whereas the stronger RNTN model generalizes well in every
-case, suggesting that it has in fact learned, or at least learned to
-simulate, our target logical concepts. These experiments provide the 
-most convincing demonstration to date of the ability of neural networks to model
-semantic inferences in complex natural language sentences.
+% Since no existing model architecture can be adopted without modification
+% to handle entailment classification, we describe a novel method for this task
+% in the spirit of prior work on recursive neural networks (\ref{methods}).
+In our first set of experiments, we use this pre-specified logical
+grammar to generate controlled data sets encoding the semantic
+relationships between pairs of expressions and evaluate whether each
+of two classes of neural model---plain RNNs and recursive neural
+tensor networks (RNTNs; \cite{socher2013acl1})---can learn those
+relationships correctly. In our first experiment
+(Section~\ref{sec:join}), we evaluate the ability of these models to
+learn the core relational algebra of natural logic from data. Our
+second experiment (Section~\ref{sec:recursion}) extends this
+evaluation to relations between complex recursive structures like
+$(a \plor b)$ and $\plneg (\plneg a \pland \plneg b)$, and our third
+experiment (Section~\ref{sec:quantifiers}) involves relations between
+quantified statements like \ii{every reptile walks} and \ii{no turtle
+  moves}. \tbchecked{We find that the plain recursive NN achieves only
+  mixed results in all three experiments, whereas the stronger RNTN
+  model generalizes well in every case, suggesting that it has in fact
+  learned, or at least learned to simulate, our target logical
+  concepts.}
 
-Finally, we perform a simple additional experiment to verify that our novel
-model structure is able to show similar reasoning behavior on real natural language data.
-The small size of existing entailment data precludes a state-of-the-art result, but we show
-that our model is able to perform competitively on the SICK entailment challenge, suggesting
-that the representational abilities that we observe in the previous sections are not limited to 
-narrowly defined artificial tasks.
+The experiments with simulated data provide a convincing demonstration
+of the ability of neural networks to learn to encode semantic
+inferences in complex natural language sentences from reasonably-sized
+training sets. However, we are also interested in the more practical
+question of whether they can learn these representations from more
+naturalistic text. To address this question, we apply our models to
+the SICK entailment challenge data. The small size of this corpus puts
+data-intensive neural models like ours at a disadvantage, but we are
+nonetheless able to achieve competitive performance on it
+\mynote{CONSIDER ADDING A CONCRETE DETAIL --- E.G., ACCURACY
+  AND ITS RELATION TO THE COMPETITION MEDIAN}, suggesting that the
+representational abilities that we observe in the previous sections
+are not limited to carefully circumscribed tasks.
 
 \begin{table*}[tp]
-  \centering
+  \centering\small
   \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.1}
+  \renewcommand{\arraystretch}{1}
   \begin{tabular}{l c l l} 

### 2014-12-02 writing/F'14 paper/intro.tex
citealt simulation

@@
     $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
+    and the relation $\natind$ applies where none of the other six do.} %, including when there 
     %is insufficient knowledge to choose a label.}

### 2014-12-02 writing/F'14 paper/discussion.tex
Paper update; mat file loading.

@@
 This paper evaluated two recursive models on a series of three increasingly
-challenging interpretive tasks involving natural language inference:
+challenging interpretive tasks involving natural language inference---
 the core relational algebra of natural logic with entailment and
 exclusion; recursive propositional logic structures; and statements
-involving quantification and negation. The results suggest that RNTNs,
-but not plain RNNs, have the capacity to meet the challenges of these
-tasks with reasonably-sized training sets. These positive results are
+involving quantification and negation---as well as on an entailment task on
+natural language image captions. The results suggest that RNTNs,
+but not necessarily plain RNNs, have the capacity to meet the 
+challenges of these tasks with reasonably-sized training sets. These positive results are
 promising for the future of learned representation models in the
@@
 Of course, challenges remain. In terms of our experimental data, even
-the RNTN falls short of perfection in our more complex tasks, with
-performance falling off steadily as the depth of recursion grows. It
+the RNTN falls short of perfection in the recursion experiment, with
+performance falling off steadily as the size of the expressions grows. It
 remains to be seen whether these deficiencies can be overcome with
-stronger models---either extended recursive models 
-\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep} or recurrent
-ones \cite{sutskever2014sequence}---or the optimization procedures. In addition,
+stronger models or the optimization procedures. The space of possible stronger
+models is sizable, including extensions of the recursive models used here
+\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, recurrent
+ones \cite{sutskever2014sequence}, and even learning-oriented variants 
+of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
+In addition,
 there remain subtle questions about how to fairly assess whether these
@@
 logical theories provide only limited guidance on this point.
-%
-%, and the fact that there is a finite universe of possible
-%  expressions makes his an unavoidable issue. 
-%
-%  CP: I don't understand the above. It is false if taken literally;
-%  our PL generates an infinte number of formulae.
-%
-Finally, we have only scratched the surface of the logical complexity
-of natural language; in future experiments, we hope to test sentences
+
+Finally, our artificial data experiments have only scratched the surface 
+of the logical complexity of natural language; in future experiments, we hope to test sentences
 with embedded quantifiers, multiple interacting quantifiers, relative

### 2014-12-02 writing/F'14 paper/join.tex
Paper update, bug fixes.
Paper update; mat file loading.

@@
 full set of sound inferences of this form is summarized in
-Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
-of relations for which no valid inference can be drawn in our logic.
+Table~\ref{tab:jointable}.
 
@@
 \begin{table}[htp]
-  \centering  
+  \centering  \small
   \setlength{\arraycolsep}{8pt}
@@
 These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
-We assess our models' ability to reproduce such inferential paths.}
+We assess our models' ability to reproduce such inferential paths.
+Cells containing a dot correspond to pairs
+of relations for which no valid inference can be drawn.}
   \label{tab:jointable}
@@
     \midrule
-    Provable test &52.0 &(10.3) &	-&(-)& \textbf{-} & \textbf{(-)}\\
+    Provable test &52.0 &(10.3) &	95...&(-)& \textbf{96.8...} & \textbf{(-)}\\
     All test &53.8 &(10.5) &	-&(-)& \textbf{-} & \textbf{(-)}\\
@@
 Table \ref{joinresultstable} shows the results, which are fairly straightforward 
-to interpret. The RNTN model
-was able to accurately encode the relations between the terms in the
-geometric relations between their vectors, and was able to then use
-that information to recover relations that were not overtly included
-in the training data. In contrast, the RNN model was able to achieve
-this behavior only incompletely. It is possible but not likely that it
-could be made to find a good solution with further optimization on
-different learning algorithms, or that it would do better on a larger
-universe of sets, for which there would be a larger set of training
-data to learn from, but the RNTN is readily able to achieve these
-effects in the setting discussed here.
+to interpret. Both models were able to accurately encode the relations 
+between the terms in the geometric relations between their vectors, 
+and was able to then use that information to recover relations that 
+were not overtly included in the training data.  The model's slightly 
+poorer performance on the unprovable test examples shows that it
+ has not simply encoded an approximation the entire underlying 
+ model, but is rather taking advantage of the specific statements seen in the test data.

### 2014-12-02 writing/F'14 paper/methods.tex
Version of figure 1 that fits in one column
Paper update; mat file loading.

@@
 via specific composition functions \cite{Partee84,Janssen97}. In our
-distributed setting, word meanings are vectors of length $n$. The
-composition function maps pairs of them to single vectors of length $n$, 
+distributed setting, word meanings are vectors of dimension $n$. The
+composition function maps pairs of them to single vectors of dimension $n$, 
 which can then be merged again to represent more complex
 phrases. Once the entire sentence-level representation has been
-derived, it serves as a fixed-length input for some subsequent function.
+derived, it serves as a fixed-dimensional input for some subsequent function.
 
@@
 For a composition layer, we evaluate models with both the plain neural
-network layer function \eqref{rnn} and a more powerful extension, the RNTN layer function
-\eqref{rntn} proposed in \newcite{chen2013learning}. A $\tanh$ nonlinearity is applied to the output of the layer function.
+network layer function \eqref{rnn} and the more powerful RNTN layer function
+\eqref{rntn} proposed in \newcite{chen2013learning}. The nonlinearity $f(x) = \tanh(x)$ is applied to the output of the layer function.
 %
 \begin{gather} \label{rnn}
-\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\
+\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}\,) \\
 \label{rntn}
@@
 element-wise non-linearity to the resulting vector. The RNTN adds a learned third-order tensor 
-$\mathbf{T}$, dimension $n \times n \times n$, modeling
+$\mathbf{T}$, of dimension $n \times n \times n$, modeling
 multiplicative interactions between the child vectors.
@@
 independently learned parameters and a separate nonlinearity function.
-Rather than use a $\tanh$ nonlinearity here, we use a rectified linear function. In
+Rather than use a $\tanh$ nonlinearity here, we found better results with a rectified linear function. In
 particular, we use the leaky rectified linear function
-\shortcite{maasrectifier}: $f(\vec{x})=\max(\vec{x}, 0) +
-0.01\min(\vec{x}, 0)$, applied element-wise.
+\shortcite{maasrectifier}: $f(x)=\max(x, 0) +
+0.01\min(x, 0)$.
 
@@
 We train the model using stochastic gradient descent (SGD)
-with learning rates computed using ADADELTA \cite{zeiler2012adadelta}.
+with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
 The parameters (including the vocabulary) are initialized randomly.
@@
 
-\begin{figure*}[tp]
+\begin{figure}[tp]
   \centering
@@
   \label{sample-figure}
-\end{figure*}
+\end{figure}

### 2014-12-02 writing/F'14 paper/quantifiers.tex
Paper update, bug fixes.
Paper update; mat file loading.

@@
 
-We have seen that the RNTN can learn an approximation of propositional
+% TODO: Compress?
+
+We have seen that recursive NNs can learn an approximation of propositional
 logic.  However, natural languages can express functional meanings of
@@
 sentences of this form and calculates the natural logic relationship
-between them.  Some examples of these data are provided in
-Table~\ref{examplesofdata}.  As in previous sections, the goal of
+between them.  Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.
+  As in previous sections, the goal of
 learning is then to assign these relational labels accurately to
@@
 
-\begin{table*}[htp]
-  \centering
-  \begin{tabular}{lcl}
-    \toprule
-    (all turtles) walk      &    $\natind$  & (less-than-three turtles) walk\\
-    (three warthogs) move  &        $\natneg$ & (less-than-three warthogs) move\\
-    (all mammals) (not move) &         $\natalt$ & (some warthogs) swim\\
-    (some (not pets)) swim     &     $\natcov$  & (all (not pets)) (not move)\\
-    \bottomrule
-  \end{tabular}
-  \caption{\label{examplesofdata}Examples of pairs of artificial setences with quantifiers.}
-
-\end{table*}
-
 % TODO: Rewrite experimental setup.
-We evaluate the model by randomly sampling 20\% of the data to
-test on and training on the remainder. In this setting, the model is being asked to learn a
+We evaluate the models by randomly sampling 20\% of the data to
+test on and training on the remainder. In this setting, the models are being asked to learn a
 complete reasoning system for the limited language and logic presented
@@
 examples that are substantially different from those it was trained
-on. Crucially, though, to succeed on this task, the model must be able
+on. Crucially, though, to succeed on this task, the models must be able
 to recognize all of the lexical relations between the nouns, verbs,
@@
 determine \eqref{p3}.
-%
-% do not allow a blank line --- adds too much space
-%
+
+\vspace{-0.6cm}
 \begin{gather}
@@
   \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
-\end{gather}
-%
-% do not allow a blank line --- adds too much space
-%
+\end{gather}\vspace{-0.8cm}
+
 \begin{table}[tp]
@@
     \midrule
-    Test & 35.4 &(7.5) &	99.1&(99.2)& \textbf{-} & \textbf{(-)}\\
+    Test & 35.4 &(7.5) &	99.1&(99.2)& \textbf{99.4...} & \textbf{(99.2)}\\
     Train & 35.4 &(7.5) &	-&	-&	-&	-\\	

### 2014-12-02 writing/F'14 paper/recursion.tex
Paper update, bug fixes.
Paper update; mat file loading.

@@
-\section{Recursive structure in propositional logic}\label{sec:recursion}
+\section{Recursive structure}\label{sec:recursion}
 
@@
 important part of their expressivity, as it allows people to use and
-interpret complex structures they have never encountered before.
+interpret complex compositional structures they have never encountered before.
 Theoretical accounts of natural language syntax and semantics
@@
 \begin{table}[tp]
-  \centering
+  \centering\small
   \begin{subtable}[t]{0.45\textwidth}
@@
 Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector RNN
-model somewhat similar to our RNTN can learn a logic, but the logic
+model somewhat similar to our RNTN can learn a boolean logic, but the logic
 discussed here is a much better approximation of the kind of
-expressivity needed for natural language. The logic learned in that
-experiment is boolean, wherein the atomic symbols are simply the
+expressivity needed for natural language. In a boolean logic
+ the atomic symbols are simply the
 values $0$ and $1$, rather than variables over those values. While
@@
 
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
+%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
+  \caption{Model performance on propositional logic, by expression size. 
+%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
+%    representations. As measured by Euclidean distance, only the RNTN
+%    achieves this for formulae containing more than a small number of
+%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
+%    failing to capture the meaning contrasts.
+    }  
+  \label{prop-results}
+  % TODO: Regenerate figure from final model.
+\end{figure*}
+
 \paragraph{Results} 
@@
 
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
-%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-  \caption{Model performance on propositional logic, by expression size. 
-%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
-%    representations. As measured by Euclidean distance, only the RNTN
-%    achieves this for formulae containing more than a small number of
-%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
-%    failing to capture the meaning contrasts.
-    }  
-  \label{prop-results}
-  % TODO: Regenerate figure from final model.
-\end{figure*}
-
 The performance of the RNTN model on small unseen test examples
@@
 number of times, but that the error in the approximation grows with
-increasing depth (and with the increasing complexity of the
-expressions), resulting in gradually dropping performance. In
-contrast, the RNN model did not appear to be able to learn a correct
-approximation of the logic for statements of any length, but it
-appears that the faulty logic that the RNN did learn decayed similarly
+increasing expression size, resulting in gradually dropping performance. In
+contrast, the RNN model did not appear to be able to accurately handle statements of any length, but it
+appears that the approximation that the RNN did learn decayed similarly
 with expression size.
@@
 RNN, we evaluated both models on pairs of long formulae involving
-binary connectives, to assess the extent to which they achieve
-substantially different representations for semantically distinct
+binary connectives, to assess how well they distinguish representations for semantically distinct
 formulae. We found that only the RNTN is able to learn

### 2014-12-02 writing/F'14 paper/sick.tex
Paper update; mat file loading.

@@
 
-The tree-pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference over real natural language.
-To ensure that this is the case, we also train a version of our model on an existing competition dataset, the ICK textual entailment challenge \cite{marelli2014sick}.
+The tree-pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and diverse range of linguistic structures seen in typical natural language data.
+To ensure that this is the case, we also train a version of our model on an existing competition dataset, the SICK textual entailment challenge \cite{marelli2014sick}.
 The corpus consists of about 10k natural language sentence pairs (generated using data expansion from a smaller set), labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}.
@@
 
-We initialized our word embeddings using the TODO dimensional vectors provided with 
-GloVe \cite{pennington2014glove}. Before any embedding is used as an input to a recursive layer, 
-it is passed through an additional $\tanh$ neural network layer that transforms
-the embedding from the embedding dimension to the dimension of the recursive layer.
+Adapting to this task required us to make a few additions to the techniques discussed in Section \ref{methods}.
+In order to better handle rare words, we initialized our word embeddings using the 200 dimensional vectors trained with
+GloVe \cite{pennington2014glove} on data from Wikipedia.
+Since 200 dimensional vectors are too large to be practical in an RNTN, especially on a small dataset, a new embedding transformation layer is needed.
+Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer.
 
-We supplemented the roughly 5000 SICK training examples\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with TODO examples of approximate entailment data from the Denotation Graph project \cite{hodoshimage}.
-We parsed all data with the Stanford PCFG Parser v.
+We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with TODO examples of entailment data from the Denotation Graph project \cite{hodoshimage}, a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn.
+We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source.
+We parsed the data from both sources with the Stanford PCFG Parser v.
 3.3.1 \cite{klein2003accurate}.
-To accelerate convergence, we collapsed subtrees that were identical across both sentences in a pair down to a single head word.
-In order to improve regularization on the noisier data, we used dropout \cite{hinton2012improving} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).
-We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source.
+We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word.
+Finally, in order to improve regularization on the noisier data, we used dropout \cite{hinton2012improving} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).
 

### 2014-12-02 writing/F'14 paper/intro.tex
Paper TODO tweaks

@@
 
-% TODO: Do we need both of these Icard/Moss citations? -- I'd prefer
-% to keep them ---CP.
-
-% Since no existing model architecture can be adopted without modification
-% to handle entailment classification, we describe a novel method for this task
-% in the spirit of prior work on recursive neural networks (\ref{methods}).
 In our first set of experiments, we use this pre-specified logical
@@
 relationships between pairs of expressions and evaluate whether each
-of two classes of neural model---plain RNNs and recursive neural
-tensor networks (RNTNs; \cite{socher2013acl1})---can learn those
+of two classes of neural model---plain recursive NNs and recursive neural
+tensor networks (RNTNs; \citealt{socher2013acl1})---can learn those
 relationships correctly. In our first experiment

### 2014-12-02 writing/F'14 paper/join.tex
Paper TODO tweaks

@@
   \end{tabular}
-  \caption{Performance on the semantic relation experiments.
-These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
+  \caption{\mynote{Performance on the semantic relation experiments.} These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
   \label{joinresultstable}
@@
 and was able to then use that information to recover relations that 
-were not overtly included in the training data.  The model's slightly 
+were not overtly included in the training data.  \tbchecked{The model's slightly 
 poorer performance on the unprovable test examples shows that it
  has not simply encoded an approximation the entire underlying 
- model, but is rather taking advantage of the specific statements seen in the test data.
+ model, but is rather taking advantage of the specific statements seen in the test data.}

### 2014-12-02 writing/F'14 paper/sick.tex
Paper TODO tweaks
One more TODO.

@@
 
-We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with TODO examples of entailment data from the Denotation Graph project \cite{hodoshimage}, a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn.
+We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with \mynote{TODO} examples of entailment data from the Denotation Graph project \cite{hodoshimage}, a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn.
 We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source.
@@
   \end{tabular}
-  \caption{Performance on SICK.}
+  \caption{\mynote{Performance on SICK.}}
   \label{sresultstable}

### 2014-12-02 writing/F'14 paper/recursion.tex
Paper TODO tweaks
Minor paper updates.

@@
 
-Recursive structure is a prominent feature of natural languages and an
-important part of their expressivity, as it allows people to use and
-interpret complex compositional structures they have never encountered before.
+Complex embedding is a prominent feature of natural languages. 
+The principle of compositionality explains how the meanings of 
+such structures are derived systematically from their parts, thereby
+ enabling people to use novel phrases and sentences.
 Theoretical accounts of natural language syntax and semantics
@@
 %  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-  \caption{Model performance on propositional logic, by expression size. 
+  \caption{\mynote{Model performance on propositional logic, by expression size. }
 %    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
@@
 including the vector representations of the six variables. The results
-are shown in Figure~\ref{prop-results}. In tuning, we found that the
+are shown in Figure~\ref{prop-results}. \mynote{In tuning, we found that the
 RNN model was approximately optimal with 45-dimensional vector
@@
 there to near-baseline performance on formulae of length 12. Training
-accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.
-
-% thu-and-or-deep-25-32b.txt iter 136
-% mon-and-or-deep-45r-32.txt iter 480
+accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.}
 

### 2014-12-02 writing/F'14 paper/abstract.tex
Paper update.

@@
   Recursive neural networks (RNNs) for sentence meaning have been
-  successful in many tasks, but it remains an open question whether
+  successful for many tasks, but it remains an open question whether
   any such model can learn compositional semantic representations that
@@
   to handle basic relational reasoning, recursive structures, and
-  quantification. We then evaluate the models on the more naturalistic
+  quantification. We then evaluate the models on the more natural
   SICK challenge data. Both models perform competitively on the SICK
@@
 \end{abstract}
-
-% TODO: Is there anything I should add that would be more specific than 'many tasks'? 'Tasks like paraphrase detection'?

### 2014-12-02 writing/F'14 paper/discussion.tex
Paper update.

@@
-\section{General discussion}\label{sec:discussion}
+\section{Discussion and conclusion}\label{sec:discussion}
 
-This paper evaluated two recursive models on a series of three increasingly
-challenging interpretive tasks involving natural language inference---
-the core relational algebra of natural logic with entailment and
+This paper first evaluates two recursive models on a series of three increasingly
+challenging interpretive tasks involving natural language inference---the 
+core relational algebra of natural logic with entailment and
 exclusion; recursive propositional logic structures; and statements
-involving quantification and negation---as well as on an entailment task on
-natural language image captions. The results suggest that RNTNs,
-but not necessarily plain RNNs, have the capacity to meet the 
-challenges of these tasks with reasonably-sized training sets. These positive results are
+involving quantification and negation. We then showed that the same models can learn to
+perform an entailment task on natural language image captions. \mynote{The results suggest that RNTNs,
+but not necessarily plain RNNs}, have the capacity to model these tasks with 
+reasonably-sized training sets. These positive results are
 promising for the future of learned representation models in the
@@
 remains to be seen whether these deficiencies can be overcome with
-stronger models or the optimization procedures. The space of possible stronger
+stronger models or optimization procedures. The space of possible stronger
 models is sizable, including extensions of the recursive models used here
 \cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, recurrent
-ones \cite{sutskever2014sequence}, and even learning-oriented variants 
+models \cite{sutskever2014sequence}, and even learning-enabled variants 
 of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
@@
 
-Finally, our artificial data experiments have only scratched the surface 
+Our artificial data experiments have only scratched the surface 
 of the logical complexity of natural language; in future experiments, we hope to test sentences
@@
 provides ample reason to be optimistic that they can be trained to
-meet the challenges of natural language semantics.
+meet all the challenges of natural language semantics.
 

### 2014-12-02 writing/F'14 paper/intro.tex
Paper update.

@@
 language compositionally from vector or tensor representations, but
-these approaches do not confront the question of how to learn their
+these approaches have not yet attempted to learn their
 highly structured representations from data using existing applied
 machine learning models.
-
-% TODO: Any way to avoid nested parens? Yes, use \citealt ---CP
 
@@
 representational abilities that we observe in the previous sections
-are not limited to carefully circumscribed tasks.
+are not limited to carefully circumscribed tasks. We conclude that RNTN models
+are adequate for typical cases of natural logic inference, and that there is not yet
+any clear level of inferential complexity for which other approaches work and NN models fail.
+
+% TODO: Clean up and condense to column length?
 

### 2014-12-02 writing/F'14 paper/join.tex
Minor paper updates.
Paper update.

@@
 statements evenly into training and test sets, and remove from the
-test set those statements which cannot be proven from the training
+test set statements which cannot be proven from the training
 statements.
@@
 We trained versions of both the RNN model and the RNTN model on these
-data sets. In both cases, the models were implemented exactly as
+data sets. In both cases, the models are implemented exactly as
 described in Section~\ref{methods}, but since the items being compared
 are single terms rather than full tree structures, the composition
-layer was not involved, and the two models differed only in which
-layer function was used for the comparison layer. We simply present
+layer is not involved, and the two models differed only in which
+layer function is used for the comparison layer. We simply present
 the models with two single terms, each of which corresponds to a
@@
     \midrule
-    Provable test &52.0 &(10.3) &	95...&(-)& \textbf{96.8...} & \textbf{(-)}\\
-    All test &53.8 &(10.5) &	-&(-)& \textbf{-} & \textbf{(-)}\\
+    Test &52.0 &(10.3) &	95...&(-)& \textbf{96.8...} & \textbf{(-)}\\
     Train &53.8 &(10.5)    & -&(-)& \textbf{-} & \textbf{(-)}\\
@@
 
-% TODO: Do we still care about the unprovable test examples?
-
 \paragraph{Results} 
-Table \ref{joinresultstable} shows the results, which are fairly straightforward 
-to interpret. Both models were able to accurately encode the relations 
+Table \ref{joinresultstable} shows the results, which are straightforward 
+to interpret. Both models are able to accurately encode the relations 
 between the terms in the geometric relations between their vectors, 
-and was able to then use that information to recover relations that 
-were not overtly included in the training data.  \tbchecked{The model's slightly 
-poorer performance on the unprovable test examples shows that it
- has not simply encoded an approximation the entire underlying 
- model, but is rather taking advantage of the specific statements seen in the test data.}
+and is able to then use that information to recover relations that 
+are not overtly included in the training data.

### 2014-12-02 writing/F'14 paper/methods.tex
Paper update.

@@
  compositionality}, which says that the meanings for complex
-expressions are derived from the meanings of their constituent parts
+expressions are derived from the meanings of their parts
 via specific composition functions \cite{Partee84,Janssen97}. In our
@@
 
-We use the model architecture depicted in Figure~\ref{sample-figure}. 
+We use the tree pair model architecture depicted in Figure~\ref{sample-figure}, which is novel 
+but closely modeled after past work on recursive NNs. 
 The two phrases being compared are processed separately using a pair 
@@
 particular, we use the leaky rectified linear function
-\shortcite{maasrectifier}: $f(x)=\max(x, 0) +
+\cite{maasrectifier}: $f(x)=\max(x, 0) +
 0.01\min(x, 0)$.
@@
  performance on infrequent classes. We compute macroaveraged F1 
-as the harmonic mean of mean precision and mean recall, both computed
+as the harmonic mean of average precision and average recall, both computed
 for all classes for which there is test data, setting precision to 0 

### 2014-12-02 writing/F'14 paper/quantifiers.tex
Paper TODO tweaks
Minor paper updates.
Paper update.

@@
 \section{Reasoning with natural language quantifiers and negation}\label{sec:quantifiers}
-
-% TODO: Compress?
 
@@
 logic.  However, natural languages can express functional meanings of
-considerably greater complexity than this.  As a first step towards
+considerably greater complexity than this.  As a step towards
 investigating whether our models can capture this complexity, we now
-attempt to directly measure the degree to which they are able to
+directly measure the degree to which they are able to
 develop suitable representations for the semantics of natural language
-quantifiers like \ii{some} and \ii{all}. Quantification is far from
+quantifiers like \ii{most} and \ii{all}. Quantification is far from
 the only place in natural language where complex functional meanings
@@
 
-% TODO: Rewrite experimental setup.
 We evaluate the models by randomly sampling 20\% of the data to
-test on and training on the remainder. In this setting, the models are being asked to learn a
+test on and training on the remainder. In this setting, each model being asked to learn a
 complete reasoning system for the limited language and logic presented
-in the training data, but it is not being asked to generalize to test
-examples that are substantially different from those it was trained
-on. Crucially, though, to succeed on this task, the models must be able
+in the training data. To succeed on this task, the model must be able
 to recognize all of the lexical relations between the nouns, verbs,
@@
     \toprule
-    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{??d RNN}  & \multicolumn{2}{c}{??d RNTN}\\
+    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{25d RNN}  & \multicolumn{2}{c}{25d RNTN}\\
     \midrule
-    Test & 35.4 &(7.5) &	99.1&(99.2)& \textbf{99.4...} & \textbf{(99.2)}\\
-    Train & 35.4 &(7.5) &	-&	-&	-&	-\\	
+    Test & 35.4 &(7.5) &	99.1&(99.2)& \textbf{99.5} & \textbf{(99.3)}\\
+    Train & 35.4 &(7.5) &	100&		(100)&	100&	(100)\\	
     \bottomrule
   \end{tabular}
-  \caption{Performance on the quantifier experiments.}
+  \caption{\mynote{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}}
   \label{qresultstable}
@@
 \paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic well.
-
-% TODO: More to say?

### 2014-12-02 writing/F'14 paper/recursion.tex
Paper update.

@@
 Theoretical accounts of natural language syntax and semantics
-therefore rely heavily on recursive structures, and we expect that
-accurate computational models will have to come to grips with them as
-well. The present section pursues this question for our two classes of
-RNN. In our evaluations, we exploit the fact that our logical language
+therefore rely heavily on recursive structures. This section examines the 
+question of whether our models are able to learn a 
+compositional semantics over such recursive structures.
+In our evaluations, we exploit the fact that our logical language
 is infinite by testing on strings that are longer and more complex
@@
   \begin{subtable}[t]{0.45\textwidth}
-    \centering\vspace{4mm} % TODO: Is there a better command to ensure padding above the table?
+    \centering\vspace{4mm}
     \begin{tabular}[t]{r c l}
@@
   \caption{Natural logic relations over sentences of propositional logic.}  
-  % TODO: Regenerate figure from final model.
   \label{prop-figure}
@@
 
-% NOTE: There was some confusion at CSLI and the 224U lecture about
-% whether this was theorem proving or some kind of model checking.
-% Hopefully this is explicit enough.
 Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector RNN
-model somewhat similar to our RNTN can learn a boolean logic, but the logic
-discussed here is a much better approximation of the kind of
-expressivity needed for natural language. In a boolean logic
- the atomic symbols are simply the
-values $0$ and $1$, rather than variables over those values. While
-learning the operators of that logic is not trivial, the outputs of
-each operator can be represented accurately by a single bit. The
-statements of propositional logic learned here describe conditions on
-unknown truth values of propositions. As opposed to the two-way
-contrasts seen in \newcite{socher2012semantic}, this logic distinguishes
+model somewhat similar to our RNTN can learn boolean logic, 
+a logic where the atomic symbols are simply the
+values $0$ and $1$. While learning the operators of that logic is not trivial, the outputs of
+each operator can be represented accurately by a single bit.
+In the propositional logic used here, the atomic symbols are variables over those values. 
+Statements of this logic thus describe conditions on
+unknown truth values of propositions. This logic distinguishes
 between $2^{6} = 64$ possible assignments of truth values, and
@@
 size.  The RNN model did not perform well, reaching only 90.5\% (81.4
-micro-F1) accuracy on the smallest test examples, and declining from
+macro-F1) accuracy on the smallest test examples, and declining from
 there to near-baseline performance on formulae of length 12. Training
@@
 with expression size.
-This is not necessarily a significant flaw in the model, since it remains possible that  different training regimes would allow it to learn accurate representations of the larger expressions.
 
-In an effort to better understand why the RNTN outperformed the plain
+\mynote{In an effort to better understand why the RNTN outperformed the plain
 RNN, we evaluated both models on pairs of long formulae involving
@@
  and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
-The RNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed at under three operators, and quickly approaches zero as depth increases.
- The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.
+The RNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed under three operators, and quickly approaches zero as depth increases.
+ The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}
+
+ % TODO: If results are still bad tomorrow, write something about why this is not fatal to the program.
 

### 2014-12-02 writing/F'14 paper/sick.tex
Minor paper updates.
Paper update.

@@
 
-The tree-pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and diverse range of linguistic structures seen in typical natural language data.
-To ensure that this is the case, we also train a version of our model on an existing competition dataset, the SICK textual entailment challenge \cite{marelli2014sick}.
+The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data are not alone sufficient to show this.
+To investigate the ability of the model to learn on natural data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}.
 The corpus consists of about 10k natural language sentence pairs (generated using data expansion from a smaller set), labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}.
@@
 
+% TODO: JMLR version of hinton paper.
+% TODO: Add Karpathy to TACL paper
+
 \begin{table}[tp]
@@
     \toprule
-    ~&\multicolumn{1}{c}{$\natind$ only} & \multicolumn{1}{c}{??d RNN}  & \multicolumn{1}{c}{??d RNTN}\\
+    ~&\multicolumn{1}{c}{$\natind$ only} & \multicolumn{1}{c}{50d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
     \midrule
@@
   \end{tabular}
-  \caption{\mynote{Performance on SICK.}}
+  \caption{\mynote{Performance on SICK. Training figures include performance on the Denotation Graph data.}}
   \label{sresultstable}
@@
 Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment.
-Our performance did not reach the state of the art (84.6\%), but exceeded the median submission to the competition (77.06\%) in a field including many models which used sophisticated hand-engineered features specific to the specific version of the entailment task at hand.
+Our performance did not reach the state of the art (84.6\%), but exceeded the median submission to the competition (77.06\%) in a field including many models which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
 
 Table \ref{examplesofsickdata} shows a few typical examples of correct classifications.
-The model appears to be able to successfully recognize negation and the insertion and deletion of words, as demonstrated in the first two examples.
-The latter two show that the model still recognize entailments and contradictions between sentences which are not superficially similar, such as passive--active pairs and sentences with low lexical overlap.
+The model is able to successfully recognize negation and the insertion and deletion of words, as demonstrated in the first two examples.
+The latter two show that the model can recognize entailments and contradictions between sentences which are not superficially similar, such as passive--active pairs and sentences with low lexical overlap.

### 2014-12-02 RunQuantifierExperiments.sh

@@ -35,8 +35,11 @@ export MATLABCMD="cd quant; lambda = 0.0001; dim = 25; td = 1; penult = 75; drop
 export MATLABCMD="cd quant; lambda = 0.0001; dim = 25; td = 1; penult = 75; dropout = 1; tot = 1; name='tq'; relu = 1; TrainModel(''\, 4\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh
 export MATLABCMD="cd quant; lambda = 0.0001; dim = 25; td = 1; penult = 75; dropout = 1; tot = 1; name='tq'; relu = 1; TrainModel(''\, 5\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh
 
-export MATLABCMD="cd quant; lambda = 0.0001; dim = 25; td = 1; penult = 75; dropout = 1; tot = 1; name='tq'; relu = 1; TrainModel(''\, 5\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh
-export MATLABCMD="cd quant; lambda = 0.0001; dim = 25; td = 1; penult = 75; dropout = 1; tot = 0; name='tq'; relu = 1; TrainModel(''\, 5\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh
+export MATLABCMD="cd quant; lambda = 0.0001; dim = 15; td = 1; penult = 75; dropout = 1; tot = 1; name='tq'; relu = 1; TrainModel(''\, 5\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh
+export MATLABCMD="cd quant; lambda = 0.0001; dim = 15; td = 1; penult = 75; dropout = 1; tot = 0; name='tq'; relu = 1; TrainModel(''\, 5\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh
[…]

### 2014-12-03 writing/F'14 paper/intro.tex
Faster testing; test data MacroF1

@@
 \cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
-the ability of these models to learn compositional semantic grammars,
-but it remains an open question whether they can achieve the same
-results as grammars based in logical forms
+the ability of these models to learn compositional semantic grammars.
+However, it remains an open question whether they can achieve the
+high-fidelity distributed representations of recent algebraic
+approaches \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,rocktaschellow},
+and whether they can match the performance of grammars based in logical forms
 \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
@@
 distributed representations have been largely confined to short
-phrases
-\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment}.
+phrases \cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment}.
 However, for robust natural language understanding, it is essential to
 model these phenomena in their full generality on complex linguistic
-structures. Other recent work
-\cite{grefenstette2013towards,rocktaschellow} has developed rich
-frameworks for computing the meanings of complex fragments of natural
+structures. Recent work on the algebraic approach of
+\newcite{ClarkCoeckeSadrzadeh2011} has yielded rich frameworks 
+for computing the meanings of complex fragments of natural
 language compositionally from vector or tensor representations, but
-these approaches have not yet attempted to learn their
-highly structured representations from data using existing applied
-machine learning models.
+how to learn these representations entirely from data in a typical
+machine learning setting remains an open question.
 
-Our research uses the task of \ii{natural language
-  inference} (also known as \ii{recognizing textual entailment};
+Our research uses the task of \ii{natural language inference} 
+(also known as \ii{recognizing textual entailment};
 \citealt{dagan2006pascal}), in which the goal is to determine the core

### 2014-12-03 writing/F'14 paper/join.tex
Faster testing; test data MacroF1

@@
 
-\begin{table}[tp]
-  \centering \small
-  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
-    \toprule
-    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{??d RNN}  & \multicolumn{2}{c}{??d RNTN}\\
-    \midrule
-    Test &52.0 &(10.3) &	95...&(-)& \textbf{96.8...} & \textbf{(-)}\\
-    Train &53.8 &(10.5)    & -&(-)& \textbf{-} & \textbf{(-)}\\
-
-    \bottomrule
-  \end{tabular}
-  \caption{\mynote{Performance on the semantic relation experiments.} These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
-  \label{joinresultstable}
-\end{table}
-
 \begin{figure}[tp]
@@
 
+\begin{table}[tp]
+  \centering \small
+  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
+    \toprule
+    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{15d RNN}  & \multicolumn{2}{c}{15d RNTN}\\
+    \midrule
+    Test &52.0 &(10.3) &	94.0&(87.0)& \textbf{99.6} & \textbf{(95.5)}\\
+    Train &53.8 &(10.5)    & 99.8&	(99.0) & \textbf{100} & \textbf{(100)}\\
+
+    \bottomrule
+  \end{tabular}
+  \caption{Performance on the semantic relation experiments.
+These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
+  \label{joinresultstable}
+\end{table}
+
 \paragraph{Results} 
 Table \ref{joinresultstable} shows the results, which are straightforward 
-to interpret. Both models are able to accurately encode the relations 
+to interpret. The RNTN is able to accurately encode the relations 
 between the terms in the geometric relations between their vectors, 
 and is able to then use that information to recover relations that 
-are not overtly included in the training data.
+are not overtly included in the training data. The RNN, in contrast, was not able to learn
+to do this well in this setting.

### 2014-12-03 writing/F'14 paper/quantifiers.tex
Faster testing; test data MacroF1

@@
   \end{tabular}
-  \caption{\mynote{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}}
+  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}
   \label{qresultstable}
@@
 \paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic well.
+We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment.

### 2014-12-03 TestModel.m

@@ -11,7 +11,7 @@ heldOutConfusion = zeros(hyperParams.numRelations(1));
 targetConfusion = zeros(hyperParams.numRelations(1));
 
 for i = 1:length(testDatasets{1})
-    [~, ~, ~, err, confusion] = CostGradFunc(theta, thetaDecoder, testDatasets{2}{i}, separateWordFeatures, hyperParams);
+    [~, ~, ~, err, confusion] = CostGradFunc(theta, thetaDecoder, testDatasets{2}{i}, separateWordFeatures, hyperParams, 0);
     if i == 1
         targetConfusion = confusion;
     end

### 2014-12-03 writing/F'14 paper/intro.tex
Intro update and misc paper cleanup

@@
 language compositionally from vector or tensor representations, but 
-how to learn these representations entirely from data in a typical
-machine learning setting remains an open question.
+has not yet yielded effective methods for learning these representations
+entirely from data in a typical machine learning setting.
 
@@
 quantified statements like \ii{every reptile walks} and \ii{no turtle
-  moves}. \tbchecked{We find that the plain recursive NN achieves only
+  moves}. We find that the plain recursive NN achieves only
   mixed results in all three experiments, whereas the stronger RNTN
@@
   learned, or at least learned to simulate, our target logical
-  concepts.}
+  concepts.
 
@@
 any clear level of inferential complexity for which other approaches work and NN models fail.
-
-% TODO: Clean up and condense to column length?
 

### 2014-12-03 writing/F'14 paper/join.tex
Intro update and misc paper cleanup

@@
 
-In our experiments, we create 80 randomly generated sets drawing from
+In each experimental run, we create 80 randomly generated sets drawing from
 a domain of seven elements. This yields a data set consisting of
@@
 are single terms rather than full tree structures, the composition
-layer is not involved, and the two models differed only in which
+layer is not involved, and the two models differ only in which
 layer function is used for the comparison layer. We simply present
@@
       \midrule
-      $b \natcov c$              & $b \nateq b$ \\
-      $c \natcov d$              & $b \natcov d$ \\
-      $c \natrev e$              & \strikeout{$b \natrev e$} \\
-      $c \natrev g$               & \strikeout{$c \nateq f$} \\
-      $e \natfor c$              & $e \natfor b$ \\
+      $a \natfor b$              & $c \natrev a$ \\
+      $b \natfor c$              & $a \natfor d$ \\
+      $b \nateq d$              & \strikeout{$a \natfor e$} \\
+      $d \natalt e$               & \strikeout{$e \natrev a$} \\
+      $e \natalt b$              & $e \nateq e$ \\
+%      a animal
+%     b mammal
+%      c dog
+ %     d mammal'
+ %     e nonmammal OR reptile
+ %     b nonmammal' OR reptile'
       \bottomrule
@@
     \midrule
+    Train &53.8 &(10.5)    & 99.8&	(99.0) & \textbf{100} & \textbf{(100)}\\
     Test &52.0 &(10.3) &	94.0&(87.0)& \textbf{99.6} & \textbf{(95.5)}\\
-    Train &53.8 &(10.5)    & 99.8&	(99.0) & \textbf{100} & \textbf{(100)}\\
-
     \bottomrule
@@
 are not overtly included in the training data. The RNN, in contrast, was not able to learn
-to do this well in this setting.
+to do this well under these conditions, though we cannot rule
+out the possibility that different optimization techniques or further hyperparameter 
+tuning could lead an RNN model to succeed here.

### 2014-12-03 writing/F'14 paper/methods.tex
Intro update and misc paper cleanup

@@
 
-We use the tree pair model architecture depicted in Figure~\ref{sample-figure}, which is novel 
-but closely modeled after past work on recursive NNs. 
-The two phrases being compared are processed separately using a pair 
+To apply recursive models to this inference class, we propose the tree 
+pair model architecture depicted in Figure~\ref{sample-figure}, which is novel but
+closely modeled after past work on recursive NNs. 
+In it, the two phrases being compared are processed separately using a pair 
 of recursive networks that share a single set of parameters. 
 The resulting vectors are fed into a separate comparison
-layer (fixed at 75 dimensions) that is meant to generate a feature vector capturing the
+layer that is meant to generate a feature vector (fixed at 75 dimensions) capturing the
 relation between the two phrases. The output of this layer is then
@@
 them by an $n \times 2n$ matrix of learned weights, adds a bias $\vec{b}$, and applies the
-element-wise non-linearity to the resulting vector. The RNTN adds a learned third-order tensor 
+non-linearity element-wise to the resulting vector. The RNTN adds a learned third-order tensor 
 $\mathbf{T}$, of dimension $n \times n \times n$, modeling

### 2014-12-03 writing/F'14 paper/recursion.tex
Intro update and misc paper cleanup

@@
 The principle of compositionality explains how the meanings of 
-such structures are derived systematically from their parts, thereby
+such structures are derived systematically from their parts,
  enabling people to use novel phrases and sentences.
@@
 
-For our experiments, we randomly generate a large set of  unique pairs 
+In our experiments, we randomly generate a large set of unique pairs 
 of formulae and compute the relation that holds for each pair.
@@
 then partitioned into bins based on the number of logical operators in
-the longer of the two formulae. We then randomly sample 15\% of each
+the longer of the two formulae. We then randomly sample 20\% of each
 bin for a held-out test set.
@@
 We trained both the RNN and RNTN models on data with up to 4
-connectives (65k pairs) and tested it on examples of up to 12
-connectives (44k pairs). We initialized the model parameters randomly,
-including the vector representations of the six variables. The results
-are shown in Figure~\ref{prop-results}. \mynote{In tuning, we found that the
+connectives (60k pairs) and tested it on a separate set of examples of up to 12
+connectives (21k pairs). We initialized the model parameters randomly,
+including the vector representations of the six variables, and tested the model
+in five fold cross-validation, discarding all but 20\% of the examples with 
+more than four connectives. The results are shown in Figure~\ref{prop-results}. 
+\mynote{In tuning, we found that the
 RNN model was approximately optimal with 45-dimensional vector

### 2014-12-03 TrainModel.m

@@ -53,7 +53,7 @@ hyperParams = FlushLogs(hyperParams);
 % hyperParams.splitFilenames = setdiff(hyperParams.splitFilenames, hyperParams.testFilenames);
 
 % TODO
-hyperParams.firstSplit = length(hyperParams.testFilenames) + 1;
+hyperParams.firstSplit = 1;
 
 % Trim out data files if needed
 if hyperParams.datasetsPortion < 1

### 2014-12-03 writing/F'14 paper/discussion.tex
Tentative results in paper

@@
 perform an entailment task on natural language image captions. \mynote{The results suggest that RNTNs,
-but not necessarily plain RNNs}, have the capacity to model these tasks with 
+and potentially also RNNs}, have the capacity to model these tasks with 
 reasonably-sized training sets. These positive results are
@@
 
-Of course, challenges remain. In terms of our experimental data, even
+Of course, challenges remain. Even
 the RNTN falls short of perfection in the recursion experiment, with

### 2014-12-03 writing/F'14 paper/intro.tex
Tentative results in paper

@@
 quantified statements like \ii{every reptile walks} and \ii{no turtle
-  moves}. We find that the plain recursive NN achieves only
-  mixed results in all three experiments, whereas the stronger RNTN
+  moves}. \mynote{Though the performance of the plain RNN model 
+ is somewhat poor in one experiment, we find that the stronger RNTN
   model generalizes well in every case, suggesting that it has in fact
   learned, or at least learned to simulate, our target logical
-  concepts.
+  concepts.}
 
@@
 data-intensive neural models like ours at a disadvantage, but we are
-nonetheless able to achieve competitive performance on it
-\mynote{CONSIDER ADDING A CONCRETE DETAIL --- E.G., ACCURACY
-  AND ITS RELATION TO THE COMPETITION MEDIAN}, suggesting that the
-representational abilities that we observe in the previous sections
+nonetheless able to achieve competitive performance on it,
+\mynote{surpassing several models with significant hand engineered task specific features.}
+This suggests that the representational abilities that we observe in the previous sections
 are not limited to carefully circumscribed tasks. We conclude that RNTN models

### 2014-12-03 writing/F'14 paper/quantifiers.tex
Intro update and misc paper cleanup
Tentative results in paper

@@
 We evaluate the models by randomly sampling 20\% of the data to
-test on and training on the remainder. In this setting, each model being asked to learn a
+test on and training on the remainder in five fold cross-validation.
+In this setting, each model being asked to learn a
 complete reasoning system for the limited language and logic presented
@@
     \midrule
-    Test & 35.4 &(7.5) &	99.1&(99.2)& \textbf{99.5} & \textbf{(99.3)}\\
     Train & 35.4 &(7.5) &	100&		(100)&	100&	(100)\\	
+    Test & 35.4 &(7.5) &	99.4&(99.4)& \textbf{99.5} & \textbf{(99.4)}\\
     \bottomrule
@@
 %
-\paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic well.
-We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment.
+\paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic satisfactorily.
+We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

### 2014-12-03 writing/F'14 paper/recursion.tex
Tentative results in paper

@@
   \centering
-  \includegraphics[width=5.3in]{recursion\string_results\string_final.eps}
+  \includegraphics[width=5.3in]{decayfig.eps}
 %  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
@@
 more than four connectives. The results are shown in Figure~\ref{prop-results}. 
-\mynote{In tuning, we found that the
+In tuning, we found that the
 RNN model was approximately optimal with 45-dimensional vector
-representations, and the RNTN model was approximately optimal with 25
-dimensions. We fixed the size of the feature vector for the classifier
-at 75 dimensions. We found that the RNTN model was able to perform
-almost perfectly on unseen small test examples, with accuracy above
-98.7\% on formulae below length four (macro-F1 was above 95.5).
+representations, \mynote{and the RNTN model was approximately optimal with 25
+dimensions. We found that the both models model were able to perform
+well on unseen small test examples, with RNN accuracy above
+97\% and RNTN accuracy above 98\% on formulae below length five.
 Starting at size four, performance gradually falls with increasing
-size.  The RNN model did not perform well, reaching only 90.5\% (81.4
-macro-F1) accuracy on the smallest test examples, and declining from
-there to near-baseline performance on formulae of length 12. Training
-accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.}
+size. Training accuracy was 99\% for the RNN and 99.5\% for the RNTN.}
 
-The performance of the RNTN model on small unseen test examples
+The performance of these model on small unseen test examples
 indicates that it learned a correct approximation of the underlying
@@
 number of times, but that the error in the approximation grows with
-increasing expression size, resulting in gradually dropping performance. In
-contrast, the RNN model did not appear to be able to accurately handle statements of any length, but it
-appears that the approximation that the RNN did learn decayed similarly
-with expression size.
+increasing expression size, resulting in gradually dropping performance.
 
-\mynote{In an effort to better understand why the RNTN outperformed the plain
-RNN, we evaluated both models on pairs of long formulae involving
+\mynote{In an effort to better understand how the models are decaying, 
+we evaluated both models on pairs of long formulae involving
 binary connectives, to assess how well they distinguish representations for semantically distinct
@@
  The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}
-
- % TODO: If results are still bad tomorrow, write something about why this is not fatal to the program.
 

### 2014-12-03 writing/F'14 paper/sick.tex
Intro update and misc paper cleanup
Tentative results in paper

@@
 
-We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with \mynote{TODO} examples of entailment data from the Denotation Graph project \cite{hodoshimage}, a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn.
+We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.
+We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}), a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn.
 We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source.
@@
 We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word.
-Finally, in order to improve regularization on the noisier data, we used dropout \cite{hinton2012improving} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).
-
-% TODO: JMLR version of hinton paper.
-% TODO: Add Karpathy to TACL paper
+Finally, in order to improve regularization on the noisier data, we used dropout \cite{srivastava2014dropout} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).
 
@@
     \toprule
-    ~&\multicolumn{1}{c}{$\natind$ only} & \multicolumn{1}{c}{50d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
+    ~&\multicolumn{1}{c}{\ii{neut.} only} & \multicolumn{1}{c}{30d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
     \midrule
-    Test & 56.7 &	...& \textbf{75.5} \\
-    Train &- &- &-  \\
+    SICK Train &  56.7 & \textbf{95.6} &  95.0  \\
+    DG Train &  50.0 & \textbf{68.0} & 66.0  \\
+    Test & 56.7 & 75.3 & \textbf{75.7}  \\
     \bottomrule
   \end{tabular}
-  \caption{\mynote{Performance on SICK. Training figures include performance on the Denotation Graph data.}}
+  \caption{\mynote{Performance on SICK.}}
   \label{sresultstable}
@@
 
+The results are shown in Table \ref{sresultstable}.
 Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment.
-Our performance did not reach the state of the art (84.6\%), but exceeded the median submission to the competition (77.06\%) in a field including many models which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
+Our performance did not reach the state of the art (84.6\%), but exceeded the performance of eight out of 18 submitted systems including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
 

### 2014-12-04 writing/F'14 paper/intro.tex
Paper update

@@
 have been successful in an array of sophisticated language tasks,
-including sentiment analysis \cite{socher2011semi,socher2013acl1},
+including sentiment analysis \cite{socher2011semi,irsoydeep},
 image description \cite{sochergrounded}, and paraphrase detection
@@
 high-fidelity distributed representations of recent algebraic
-approaches \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,rocktaschellow},
+approaches \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
 and whether they can match the performance of grammars based in logical forms
@@
 experiment (Section~\ref{sec:quantifiers}) involves relations between
-quantified statements like \ii{every reptile walks} and \ii{no turtle
+statements with quantification and negation like \ii{every reptile walks} and \ii{no turtle
  moves}. \mynote{Though the performance of the plain RNN model 
- is somewhat poor in one experiment, we find that the stronger RNTN
+ is somewhat poor in our first experiment, we find that the stronger RNTN
   model generalizes well in every case, suggesting that it has in fact
@@
 training sets. However, we are also interested in the more practical
-question of whether they can learn these representations from more
+question of whether they can learn these representations from
 naturalistic text. To address this question, we apply our models to
@@
 nonetheless able to achieve competitive performance on it,
-\mynote{surpassing several models with significant hand engineered task specific features.}
+\mynote{surpassing several models with significant hand-engineered task-specific features.}
 This suggests that the representational abilities that we observe in the previous sections

### 2014-12-04 writing/F'14 paper/join.tex
Tentative results in paper
Paper update

@@
 depicts a structure of this form. The lattice gives the full model,
-for which all the statements on the right are valid. We divide these
+for which all the statements on below are valid. We divide these
 statements evenly into training and test sets, and remove from the
@@
 
-\begin{figure}[tp]
+\begin{figure}[t]
   \centering
-%  \begin{subfigure}[t]{0.45\textwidth}
-%    \centering
-%    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
-%    \setlength{\unitlength}{1cm}
-%    \begin{picture}(5,5.5)
-%      \labelednode{2.50}{5}{}{1,2,3}
-%      
-%      \put(0.75,4){\line(3,1){1.5}}
-%      \put(2.5,4){\line(0,1){0.5}}
-%      \put(4.25,4){\line(-3,1){1.5}}
-%      
-%      \labelednode{0.75}{3.5}{a,b}{1,2}
-%      \labelednode{2.50}{3.5}{c}{1,3}
-%      \labelednode{4.25}{3.5}{d}{2,3}
-%      
-%      \put(0.75,2.5){\line(0,1){0.5}}
-%      \put(0.75,2.5){\line(3,1){1.5}}
-%      
-%      \put(2.5,2.5){\line(-3,1){1.5}}
-%      \put(2.5,2.5){\line(3,1){1.5}}
-%      
-%      \put(4.25,2.5){\line(0,1){0.5}}
-%      \put(4.25,2.5){\line(-3,1){1.5}}
-%      
-%
-%      \labelednode{0.75}{2}{e,f}{1}
-%      \labelednode{2.50}{2}{}{2}
-%      \labelednode{4.25}{2}{g,h}{3}
-%      
-%      \put(2.5,1){\line(-3,1){1.5}}
-%      \put(2.5,1){\line(0,1){0.5}}
-%      \put(2.5,1){\line(3,1){1.5}}
-%      
-%      \labelednode{2.5}{0.5}{}{}
-%    \end{picture}
-%    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
-%    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
-%  \end{subfigure}
-%  \qquad
-%  \begin{subfigure}[t]{0.43\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
     \centering
+    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
+    \setlength{\unitlength}{1cm}\scalebox{0.7}{
+    \begin{picture}(5,5.5)
+      \labelednode{2.50}{5}{}{1,2,3}
+
+      \put(0.75,4){\line(3,1){1.5}}
+      \put(2.5,4){\line(0,1){0.5}}
+      \put(4.25,4){\line(-3,1){1.5}}
+
+      \labelednode{0.75}{3.5}{a,b}{1,2}
+      \labelednode{2.50}{3.5}{c}{1,3}
+      \labelednode{4.25}{3.5}{d}{2,3}
+
+      \put(0.75,2.5){\line(0,1){0.5}}
+      \put(0.75,2.5){\line(3,1){1.5}}
+
+      \put(2.5,2.5){\line(-3,1){1.5}}
+      \put(2.5,2.5){\line(3,1){1.5}}
+
+      \put(4.25,2.5){\line(0,1){0.5}}
+      \put(4.25,2.5){\line(-3,1){1.5}}
+
+      \labelednode{0.75}{2}{e,f}{1}
+      \labelednode{2.50}{2}{}{2}
+      \labelednode{4.25}{2}{g,h}{3}
+
+      \put(2.5,1){\line(-3,1){1.5}}
+      \put(2.5,1){\line(0,1){0.5}}
+      \put(2.5,1){\line(3,1){1.5}}
+
+      \labelednode{2.5}{0.5}{}{}
+    \end{picture}}
+    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
+    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
+  \end{subfigure}
+  \qquad
+    \begin{subfigure}[t]{0.43\textwidth}
+    \centering \vspace{0.4cm}
     \setlength{\tabcolsep}{12pt}
@@
       \midrule
-      $a \natfor b$              & $c \natrev a$ \\
-      $b \natfor c$              & $a \natfor d$ \\
-      $b \nateq d$              & \strikeout{$a \natfor e$} \\
-      $d \natalt e$               & \strikeout{$e \natrev a$} \\
-      $e \natalt b$              & $e \nateq e$ \\
-%      a animal
-%     b mammal
-%      c dog
- %     d mammal'
- %     e nonmammal OR reptile
- %     b nonmammal' OR reptile'
+
+      $a \nateq b$              & $b \nateq b$ \\
+      $a \natfor e$              & $b \natfor e$ \\
+      $a \natneg g$              & \strikeout{$e \nateq f$} \\
+      $d \natfor h$              & \strikeout{$g \natrev d$} \\
+      $e \natalt g$            & $h \natrev d$ \\
+
       \bottomrule
     \end{tabular}
-    \caption{A train/test split of the atomic statements about the
-      model.  Test statements not provable from the training data are
+
+    \caption{A few examples of atomic statements about the
+      model.  Test statements that are not provable from the training data shown are
       crossed out.}
- % \end{subfigure}  
- %  \caption{Small example structure and data for learning relation composition.}
+  \end{subfigure}  
+  \caption{Small example structure and data for learning relation composition.}
   \label{lattice-figure}
 \end{figure}
+
+\paragraph{Results} 
+Table \ref{joinresultstable} shows the results, which are straightforward 
+to interpret. The RNTN is able to accurately encode the relations 
+between the terms in the geometric relations between their vectors, 
+and is able to then use that information to recover relations that 
+are not overtly included in the training data. The RNN, in contrast, was not able to learn
+to do this well under these conditions, though we cannot rule
+out the possibility that different optimization techniques or further hyperparameter 
+tuning could lead an RNN model to succeed here.
+
 % Note: None of these test examples is derivable from the shown training data, 
@@
 % out willy-nilly without fear.
-
 \begin{table}[tp]
@@
 \end{table}
-
-\paragraph{Results} 
-Table \ref{joinresultstable} shows the results, which are straightforward 
-to interpret. The RNTN is able to accurately encode the relations 
-between the terms in the geometric relations between their vectors, 
-and is able to then use that information to recover relations that 
-are not overtly included in the training data. The RNN, in contrast, was not able to learn
-to do this well under these conditions, though we cannot rule
-out the possibility that different optimization techniques or further hyperparameter 
-tuning could lead an RNN model to succeed here.

### 2014-12-04 writing/F'14 paper/methods.tex
Paper update

@@
 function, we use the negative log of the probability assigned to the
-correct label.
+correct label, plus L2 regularization.
 
@@
 with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
-The parameters (including the vocabulary) are initialized randomly.
-Because the classes are not balanced in general, we report performance
+Training times on CPUs vary from hours to days across experiments, and
+the hyperparameters of AdaDelta, the dimensions of the representations,
+and the initialization strategy are all tuned manually.
+Because the classes are not balanced in the artificial data experiments, we report performance
 using both accuracy and macroaveraged F1, since the latter emphasizes
@@
 for all classes for which there is test data, setting precision to 0 
-where it is not defined. Source code and generated data will be released
-after the review period.
+where it is not defined. On the artificial data experiments, we report mean
+results over five fold cross-validation, though differences across runs are small.
+Source code and generated data will be released after the review period.
 

### 2014-12-04 writing/F'14 paper/quantifiers.tex
Paper update

@@
-\section{Reasoning with natural language quantifiers and negation}\label{sec:quantifiers}
+\section{Reasoning with quantifiers and negation}\label{sec:quantifiers}
 
@@
 develop suitable representations for the semantics of natural language
-quantifiers like \ii{most} and \ii{all}. Quantification is far from
+quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantifiers 
+and negation are far from
 the only place in natural language where complex functional meanings
-are found, but it is a natural starting point, since it can be tested
-in sentences whose structures are otherwise quite simple, and since it
-has formed a standard case study in prior formal work on natural
+are found, but they are natural starting point, since they can be tested
+in sentences whose structures are otherwise quite simple, and since they have
+formed a standard case study in prior formal work on natural
 language inference \cite{Icard:Moss:2013:LILT}.
@@
 
-We evaluate the models by randomly sampling 20\% of the data to
-test on and training on the remainder in five fold cross-validation.
-In this setting, each model being asked to learn a
+We evaluate the models by sampling 20\% of the data to test on and 
+training on the remainder. In this setting, each model must learn a
 complete reasoning system for the limited language and logic presented
@@
 and quantifiers and how they interact. For instance, it might see
-\eqref{p1} and \eqref{p2} in training and, from that information,
-determine \eqref{p3}.
+\eqref{p1} and \eqref{p2} in training and be required to then label \eqref{p3}.
 
@@
     \midrule
-    Train & 35.4 &(7.5) &	100&		(100)&	100&	(100)\\	
-    Test & 35.4 &(7.5) &	99.4&(99.4)& \textbf{99.5} & \textbf{(99.4)}\\
+    Train & 35.4 &(7.5) &	\textbf{100}&		\textbf{(100)}&	99.5&	(99.3)\\	
+    Test & 35.4 &(7.5) &	\textbf{99.4}&\textbf{(99.5)}& \textbf{99.4} & (99.2)\\
     \bottomrule

### 2014-12-04 writing/F'14 paper/recursion.tex
Paper update

@@
 Theoretical accounts of natural language syntax and semantics
-therefore rely heavily on recursive structures. This section examines the 
+in turn rely heavily on recursive structures. This section examines the 
 question of whether our models are able to learn a 
@@
 possible statements that the intermediate vector representations need
-to be able to distinguish if we are to learn this logical system in its
-full generality.
+to be able to distinguish if we are to learn this logical system in full.
 
@@
   \centering
-  \includegraphics[width=5.3in]{decayfig.eps}
+  \includegraphics[width=5in]{decayfig.eps}
 %  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
@@
 connectives (21k pairs). We initialized the model parameters randomly,
-including the vector representations of the six variables, and tested the model
-in five fold cross-validation, discarding all but 20\% of the examples with 
-more than four connectives. The results are shown in Figure~\ref{prop-results}. 
-In tuning, we found that the
-RNN model was approximately optimal with 45-dimensional vector
-representations, \mynote{and the RNTN model was approximately optimal with 25
-dimensions. We found that the both models model were able to perform
-well on unseen small test examples, with RNN accuracy above
-97\% and RNTN accuracy above 98\% on formulae below length five.
+including the vector representations of the six variables.
+Figure~\ref{prop-results} shows test the relationship
+between test accuracy and statement length. \mynote{We found that both 
+models were able to perform well on unseen small test examples, 
+with RNN accuracy above
+97\% and RNTN accuracy above 99\% on formulae below length five.
 Starting at size four, performance gradually falls with increasing
-size. Training accuracy was 99\% for the RNN and 99.5\% for the RNTN.}
+size. Training accuracy was 99.4\% for the RNN and 99.9\% for the RNTN.}
 
-The performance of these model on small unseen test examples
-indicates that it learned a correct approximation of the underlying
-logic. It appears that this approximation is accurate enough to yield
+The performance of these models on small unseen test examples
+indicates that they learned correct approximations of the underlying
+logic. It appears that these approximations are accurate enough to yield
 correct answers when the composition layer is only applied a small
@@
 increasing expression size, resulting in gradually dropping performance. 
+It appears that the RNTN model overfit somewhat relative to the plain RNN,
+performing better on the training data and short examples, but worse on longer
+examples. Two factors are likely involved in this: even with the lower dimension,
+the RNTN composition function has about eight times as many parameters as the
+RNN, and this much more capacity to memorize a large number of narrow 
+generalizations that work on the test data but don't generalize as well. In addition,
+the RNTN used here used slightly less L2 regularization than the 
+RNN ($\lambda_{RNN} = 0.001$, $\lambda_{RNTN} = 0.0003$), as had
+yielded better performance earlier in tuning.
 
-\mynote{In an effort to better understand how the models are decaying, 
-we evaluated both models on pairs of long formulae involving
-binary connectives, to assess how well they distinguish representations for semantically distinct
-formulae. We found that only the RNTN is able to learn
-substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
- and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
-The RNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed under three operators, and quickly approaches zero as depth increases.
- The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}
+% \mynote{In an effort to better understand why model performance decays, 
+%we evaluated both models on pairs of long formulae involving
+%binary connectives, to assess how well they distinguish representations for semantically distinct
+%formulae. We found that only the RNTN is able to learn
+%substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
+% and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
+%The RNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed under three operators, and quickly approaches zero as depth increases.
+% The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}
 

### 2014-12-04 writing/F'14 paper/sick.tex
Paper update

@@
 
-The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data are not alone sufficient to show this.
+The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data are not alone sufficient to address this issue.
 To investigate the ability of the model to learn on natural data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}.
@@
 GloVe \cite{pennington2014glove} on data from Wikipedia.
-Since 200 dimensional vectors are too large to be practical in an RNTN, especially on a small dataset, a new embedding transformation layer is needed.
-Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer.
+Since 200 dimensional vectors are too large to be practical in an RNTN on a small dataset, a new embedding transformation layer is needed.
+Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer which aggregates any usable information from the embedding vectors into a more compact working representation.
 
@@
 We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word.
+The test data on which we report performance are collapsed in this way, and both collapsed and uncollapsed copies of the training data are used in training.
 Finally, in order to improve regularization on the noisier data, we used dropout \cite{srivastava2014dropout} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).
@@
     \midrule
-    SICK Train &  56.7 & \textbf{95.6} &  95.0  \\
-    DG Train &  50.0 & \textbf{68.0} & 66.0  \\
-    Test & 56.7 & 75.3 & \textbf{75.7}  \\
+    SICK Train &  56.7 & 95.6 &  \textbf{97.8}  \\
+    DG Train &  50.0 & 68.0 & \textbf{74.0}  \\
+    Test & 56.7 & 75.3 & \textbf{76.9}  \\
     \bottomrule
@@
 Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment.
-Our performance did not reach the state of the art (84.6\%), but exceeded the performance of eight out of 18 submitted systems including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
+Our performance did not reach that of the winning system (84.6\%), but exceeded the performance of eight out of 18 submitted systems including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
 
@@
 The model is able to successfully recognize negation and the insertion and deletion of words, as demonstrated in the first two examples.
-The latter two show that the model can recognize entailments and contradictions between sentences which are not superficially similar, such as passive--active pairs and sentences with low lexical overlap.
+The latter two show that the model can recognize entailments and contradictions between superficially different sentences, such as passive--active pairs and sentences with low lexical overlap.

### 2014-12-05 writing/F'14 paper/discussion.tex
Paper update
Last paper update for now.

@@
 exclusion; recursive propositional logic structures; and statements
-involving quantification and negation. We then showed that the same models can learn to
-perform an entailment task on natural language image captions. \mynote{The results suggest that RNTNs,
-and potentially also RNNs}, have the capacity to model these tasks with 
+involving quantification and negation. We then show that the same models can learn to
+perform an entailment task on natural language image captions. The results suggest that RNTNs,
+and potentially also RNNs, have the capacity to learn to model these tasks with 
 reasonably-sized training sets. These positive results are
@@
 with embedded quantifiers, multiple interacting quantifiers, relative
-clauses, and other kinds of recursive structure. Nonetheless, the
+clauses, and other kinds of recursive structure. Our SICK experiments 
+similarly only begin to reveal the potential of these models to learn to 
+perform complex semantic inferences from corpora. Nonetheless, the
 rapid progress the field has made with these models in recent years

### 2014-12-05 writing/F'14 paper/intro.tex
Last paper update for now.

@@
 \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
-when it comes to the core semantic concepts of quantification,
+on the core semantic issues of quantification,
 entailment, and contradiction needed to identify the relationships
@@
 evaluation to relations between complex recursive structures like
-$(a \plor b)$ and $\plneg (\plneg a \pland \plneg b)$, and our third
-experiment (Section~\ref{sec:quantifiers}) involves relations between
+$(a \plor b)$ and $\plneg~(\plneg a \pland \plneg b)$, and our third
+experiment (Section~\ref{sec:quantifiers}) covers relations between
 statements with quantification and negation like \ii{every reptile walks} and \ii{no turtle
- moves}. \mynote{Though the performance of the plain RNN model 
+ moves}. Though the performance of the plain RNN model 
  is somewhat poor in our first experiment, we find that the stronger RNTN
@@
 nonetheless able to achieve competitive performance on it,
-\mynote{surpassing several models with significant hand-engineered task-specific features.}
+surpassing several models with significant hand-engineered task-specific features.
 This suggests that the representational abilities that we observe in the previous sections

### 2014-12-05 writing/F'14 paper/join.tex
Last paper update for now.

@@
-\section{Reasoning about semantic relations}\label{sec:join}
-
-If a model is to learn the behavior of a relational logic like the one
-presented here from a finite amount of data, it must learn to deduce new
-relations from seen relations in a sound manner. The simplest such
-deductions involve atomic statements using the relations in
-Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
-\natrev c$, one can conclude that $a \natrev c$, by basic
-set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
-full set of sound inferences of this form is summarized in
-Table~\ref{tab:jointable}.
-
-% about the relations themselves that do not depend on the
-% internal structure of the things being compared. For example, given
-% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
-% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven relations support more
-% than just transitivity: MacCartney and Manning's
-% \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of relations of the form $a R
-% b$ and $b R' c$, including several less intuitive ones such as that if
-% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-
 \begin{table}[htp]
@@
 
+\section{Reasoning about semantic relations}\label{sec:join}
+
+If a model is to learn the behavior of a relational logic like the one
+presented here from a finite amount of data, it must learn to deduce new
+relations from seen relations in a sound manner. The simplest such
+deductions involve atomic statements using the relations in
+Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
+\natrev c$, one can conclude that $a \natrev c$, by basic
+set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
+$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
+full set of sound such inferences is summarized in
+Table~\ref{tab:jointable}.
+
+% about the relations themselves that do not depend on the
+% internal structure of the things being compared. For example, given
+% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
+% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
+% understanding $a$, $b$, or $c$. These seven relations support more
+% than just transitivity: MacCartney and Manning's
+% \cite{maccartney2009extended} join table defines 32 valid inferences
+% that can be made on the basis of pairs of relations of the form $a R
+% b$ and $b R' c$, including several less intuitive ones such as that if
+% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
+
 \paragraph{Experiments}
@@
 depicts a structure of this form. The lattice gives the full model,
-for which all the statements on below are valid. We divide these
+for which all the statements below it are valid. We divide these
 statements evenly into training and test sets, and remove from the
@@
 We trained versions of both the RNN model and the RNTN model on these
-data sets. In both cases, the models are implemented exactly as
+data sets. In both cases, the models are implemented as
 described in Section~\ref{methods}, but since the items being compared
@@
 layer function is used for the comparison layer. We simply present
-the models with two single terms, each of which corresponds to a
-single vector in the (randomly initialized) vocabulary matrix $V$,
-thereby ensuring that the model has no information about the terms
-being compared except the relations between them.
+the models with the (randomly initialized) embedding vectors for each
+of two terms, ensuring that the model has no information about the terms
+being compared except for the relations between them that appear in training.
 
@@
     \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
-    \setlength{\unitlength}{1cm}\scalebox{0.7}{
+    \setlength{\unitlength}{1cm}\scalebox{0.8}{
     \begin{picture}(5,5.5)
@@
 
-\paragraph{Results} 
-Table \ref{joinresultstable} shows the results, which are straightforward 
-to interpret. The RNTN is able to accurately encode the relations 
-between the terms in the geometric relations between their vectors, 
-and is able to then use that information to recover relations that 
-are not overtly included in the training data. The RNN, in contrast, was not able to learn
-to do this well under these conditions, though we cannot rule
-out the possibility that different optimization techniques or further hyperparameter 
-tuning could lead an RNN model to succeed here.
-
 % Note: None of these test examples is derivable from the shown training data, 
@@
 \end{table}
+
+\paragraph{Results} 
+Table \ref{joinresultstable} shows the results. 
+The RNTN is able to accurately encode the relations 
+between the terms in the geometric relations between their vectors, 
+and is able to then use that information to recover relations that 
+are not overtly included in the training data. The RNN also generalizes fairly well, 
+but makes enough errors that it remains an open question whether 
+it is capable of learning representations with these properties. 
+We cannot rule out the possibility that different optimization techniques or
+further hyperparameter tuning could lead an RNN model to succeed here.

### 2014-12-05 writing/F'14 paper/methods.tex
Last paper update for now.

@@
 
-To apply recursive models to this inference class, we propose the tree 
+\begin{figure}[tp]
+  \centering
+ % \includegraphics[scale=0.35]{model.eps}
+  \input{figure1}
+  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
+    The same structure is used for both the RNN and RNTN layer functions.} 
+  \label{sample-figure}
+\end{figure}
+
+To apply these recursive models to our task, we propose the tree 
 pair model architecture depicted in Figure~\ref{sample-figure}, which is novel but
@@
 where it is not defined. On the artificial data experiments, we report mean
-results over five fold cross-validation, though differences across runs are small.
+results over five fold cross-validation. Variance across runs is typically no
+more than two percentage points.
 Source code and generated data will be released after the review period.
-
-\begin{figure}[tp]
-  \centering
- % \includegraphics[scale=0.35]{model.eps}
-  \input{figure1}
-  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
-    The same structure is used for both the RNN and RNTN layer functions.} 
-  \label{sample-figure}
-\end{figure}

### 2014-12-05 writing/F'14 paper/quantifiers.tex
Last paper update for now.

@@
 develop suitable representations for the semantics of natural language
-quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantifiers 
+quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantification 
 and negation are far from
@@
 
-We evaluate the models by sampling 20\% of the data to test on and 
-training on the remainder. In this setting, each model must learn a
+In each run, we sample 20\% of the 60k generated pairs to test on and 
+train on the remainder. In this setting, each model must learn a
 complete reasoning system for the limited language and logic presented
-in the training data. To succeed on this task, the model must be able
+in the training data, and must be able
 to recognize all of the lexical relations between the nouns, verbs,
 and quantifiers and how they interact. For instance, it might see
-\eqref{p1} and \eqref{p2} in training and be required to then label \eqref{p3}.
+pairs like \eqref{p1} and \eqref{p2} in training and be required to 
+then label \eqref{p3}.
 

### 2014-12-05 writing/F'14 paper/recursion.tex
Last paper update for now.

@@
       %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
-      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
+      $\plneg\, (\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
       \bottomrule
@@
   \centering
-  \includegraphics[width=5in]{decayfig.eps}
+  \includegraphics[width=5.75in]{decayfig.eps}
 %  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-  \caption{\mynote{Model performance on propositional logic, by expression size. }
+  \caption{Model performance on propositional logic, by expression size.}
 %    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
@@
 %    failing to capture the meaning contrasts.
-    }  
+
   \label{prop-results}
-  % TODO: Regenerate figure from final model.
+
 \end{figure*}
@@
 Figure~\ref{prop-results} shows test the relationship
-between test accuracy and statement length. \mynote{We found that both 
+between test accuracy and statement length. We found that both 
 models were able to perform well on unseen small test examples, 
 with RNN accuracy above
-97\% and RNTN accuracy above 99\% on formulae below length five.
+98\% and RNTN accuracy above 99\% on formulae below length five.
 Starting at size four, performance gradually falls with increasing
-size. Training accuracy was 99.4\% for the RNN and 99.9\% for the RNTN.}
+size. Training accuracy was 99.4\% for the RNN and 99.8\% for the RNTN.
 
@@
 indicates that they learned correct approximations of the underlying
-logic. It appears that these approximations are accurate enough to yield
+logic. These approximations are accurate enough to yield
 correct answers when the composition layer is only applied a small
@@
 the RNTN composition function has about eight times as many parameters as the
-RNN, and this much more capacity to memorize a large number of narrow 
-generalizations that work on the test data but don't generalize as well. In addition,
-the RNTN used here used slightly less L2 regularization than the 
-RNN ($\lambda_{RNN} = 0.001$, $\lambda_{RNTN} = 0.0003$), as had
-yielded better performance earlier in tuning.
+RNN, and thus more capacity to find a solution involving a large number of narrow 
+generalizations that work on the training data but don't generalize well. In addition,
+the RNTN used here has slightly weaker L2 regularization than the 
+RNN ($\lambda_{\textit{RNN}} = 0.001$, $\lambda_{\textit{RNTN}} = 0.0003$). This
+yielded better average performance, but may have facilitated this effect.
 

### 2014-12-05 writing/F'14 paper/sick.tex
Last paper update for now.

@@
 To investigate the ability of the model to learn on natural data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}.
-The corpus consists of about 10k natural language sentence pairs (generated using data expansion from a smaller set), labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}.
-At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest hand labeled entailment corpus that we are aware of.
+The corpus consists of about 10k natural language sentence pairs, labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}.
+At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of.
 
@@
   \end{tabular}
-  \caption{\label{examplesofsickdata}Examples of successful classifications on SICK.}
+  \caption{\label{examplesofsickdata}Examples of correct RNTN classifications on SICK.}
 \end{table*}
@@
 Adapting to this task required us to make a few additions to the techniques discussed in Section \ref{methods}.
-In order to better handle rare words, we initialized our word embeddings using the 200 dimensional vectors trained with
+In order to better handle rare words, we initialized our word embeddings using 200 dimensional vectors trained with
 GloVe \cite{pennington2014glove} on data from Wikipedia.
 Since 200 dimensional vectors are too large to be practical in an RNTN on a small dataset, a new embedding transformation layer is needed.
-Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer which aggregates any usable information from the embedding vectors into a more compact working representation.
+Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer.
+This new layer aggregates any usable information from the embedding vectors into a more compact working representation.
 
@@
 We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word.
-The test data on which we report performance are collapsed in this way, and both collapsed and uncollapsed copies of the training data are used in training.
+The training and test data on which we report performance are collapsed in this way, and both collapsed and uncollapsed copies of the training data are used in training.
 Finally, in order to improve regularization on the noisier data, we used dropout \cite{srivastava2014dropout} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).
@@
     \midrule
-    SICK Train &  56.7 & 95.6 &  \textbf{97.8}  \\
-    DG Train &  50.0 & 68.0 & \textbf{74.0}  \\
-    Test & 56.7 & 75.3 & \textbf{76.9}  \\
+    SICK Train &  56.7 & 95.4 &  \textbf{97.8}  \\
+    DG (Train) &  50.0 & 67.0 & \textbf{74.0}  \\
+    SICK Test & 56.7 & 74.9 & \textbf{76.9}  \\
     \bottomrule
   \end{tabular}
-  \caption{\mynote{Performance on SICK.}}
+  \caption{Performance on collapsed SICK data.}
   \label{sresultstable}
@@
 Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment.
-Our performance did not reach that of the winning system (84.6\%), but exceeded the performance of eight out of 18 submitted systems including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
+Our RNTN did not reach the performance of the winning system (84.6\%), but it did exceed that of eight out of 18 submitted systems, including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
 
 Table \ref{examplesofsickdata} shows a few typical examples of correct classifications.
-The model is able to successfully recognize negation and the insertion and deletion of words, as demonstrated in the first two examples.
-The latter two show that the model can recognize entailments and contradictions between superficially different sentences, such as passive--active pairs and sentences with low lexical overlap.
+Our qualitative evaluation suggests that the model is able to successfully recognize negation and the insertion or deletion of words, as in the first two examples.
+The latter two show that the model can also identify the relationships between superficially different sentences, such as passive--active pairs and sentences with low lexical overlap.

### 2014-12-05 config/AndOr.m

@@ -24,7 +24,7 @@ hyperParams.topDropout = tdrop;
 hyperParams.topDepth = top;
 
 hyperParams.specialAndOrMode = 1;
-hyperParams.truncateTest = 1;
+hyperParams.truncateTest = 0;
 
 if relu
   hyperParams.classNL = @LReLU;
@@ -59,14 +59,14 @@ end
 
 
 % How often (in steps) to report cost.
-options.costFreq = 500; 'TODO'
+options.costFreq = 1; 'TODO'
 
 % How often (in steps) to run on test data.
-options.testFreq = 500; 'TODO'
+options.testFreq = 1; 'TODO'
 
 % How often to report confusion matrices. 
 % Should be a multiple of testFreq.
-options.confusionFreq = 500; 'TODO'
+options.confusionFreq = 1; 'TODO'
 
 % How often to display which items are misclassified.
 % Should be a multiple of testFreq.

### 2015-01-07 writing/AAAI SSS '15/BowmanPottsManning.tex
Update AAAI SSS KRR paper

\def\year{2015}
%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\usepackage[leqno, fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{qtree}
\usepackage[numbers]{natbib}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}

\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title Learning Distributed Word Representations for Natural Logic Reasoning
/Author Samuel R. Bowman, Christopher Potts, Christopher D. Manning
/Keywords distributional semantics, vector space models, natural logic, natural language processing, neural networks}
\setcounter{secnumdepth}{1}

\newcommand{\nateq}{\equiv}
\newcommand{\natind}{\mathbin{\#}}
%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
\newcommand{\natneg}{\mathbin{^{\wedge}}}
\newcommand{\natfor}{\sqsubset}
\newcommand{\natrev}{\sqsupset}
\newcommand{\natalt}{\mathbin{|}}
\newcommand{\natcov}{\mathbin{\smallsmile}}

\newcommand{\plneg}{\mathop{\textit{not}}}
\newcommand{\pland}{\mathbin{\textit{and}}}
\newcommand{\plor}{\mathbin{\textit{or}}}

% Strikeout
\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}

\newcommand{\word}[1]{\emph{#1}}
\def\ii#1{\textit{#1}}

 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Learning Distributed Word Representations for Natural Logic Reasoning}
\author{Samuel R. Bowman, Christopher Potts, and Christopher D. Manning\\
  Stanford University\\
  450 Serra Mall \\
  Stanford, CA 94305-2150\\
  \texttt{\{sbowman,cgpotts,manning\}@stanford.edu}} %TODO: Three affiliations?
\maketitle
\begin{abstract}
\begin{quote}
  Natural logic offers a powerful relational conception of meaning
  that is a natural counterpart to distributed semantic
  representations, which have proven valuable in a wide range of
  sophisticated language tasks. However, it remains an open question
  whether it is possible to train distributed representations to
  support the rich, diverse logical reasoning captured by natural
  logic. We address this question using two neural network-based
  models for learning embeddings: plain neural networks and neural
  tensor networks. Our experiments evaluate the models' ability to
  learn the basic algebra of natural logic relations from simulated
  data and from the WordNet noun graph.  The overall positive results
  are promising for the future of learned distributed representations in
  the applied modeling of logical semantics.
\end{quote}
\end{abstract}

Natural logic offers a powerful \emph{relational} conception of
semantics: the meanings for expressions are given, at least in part,
by their inferential connections with other expressions
\cite{vanBenthem08NATLOG,maccartney2009extended}. For instance,
\word{turtle} is analyzed, not primarily by its extension in the
world, but rather by its lexical network: it entails \word{reptile},
excludes \word{chair}, is entailed by \word{sea
  turtle}, and so forth. With generalized notions of entailment and
contradiction, these relationships can be defined for all lexical
categories as well as complex phrases, sentences, and even texts. The
resulting theories of meaning offer valuable new analytic tools for
tasks involving database inference, relation extraction, and textual
entailment.

Natural logic aligns well with distributed (e.g., vector)
representations, which also naturally model meaning relationally.
Distributed representations have been used successfully in a wide
array of sophisticated language tasks (e.g., \cite{collobert2011natural}). %,
% including sentiment analysis, analogy completion, relation
% extraction, and named entity recognition.
However, it remains an open question whether it is possible to train
such representations to support the rich, diverse logical reasoning
captured by natural logic; while they excel at synonymy (similarity),
the results are more mixed for entailment, contradiction, and mutual
consistency.  Using the natural logic of \cite{maccartney2009extended}
as our formal model, we address this open question using two neural
network-based models for learning embeddings: plain neural networks
and neural tensor networks (NTNs).  The natural logic is built from
the seven relations defined in Table~\ref{b-table}. Its formal
properties are now well-understood \cite{Icard:Moss:2013:LILT}, so it
provides a rigorous set of goals for our neural models. To keep the
discussion manageable, we limit attention to experiments involving the
lexicon; for a more extended treatment of complex expressions
involving logical connectives and quantifiers, see
\cite{Bowman:Potts:Manning:2014}. % CITET

In our experiments, we evaluate these models' ability to learn the
basic algebra of natural logic relations from simulated data and from
the WordNet noun graph. The simulated data help us to achieve analytic
insights into what the models learn, and the WordNet data show how they
fare with a real natural language vocabulary.  We find that only the NTN is able to fully
learn the underlying algebra, but that both models excel in the 
WordNet experiment.

% TODO: Better Gref citation; cite Oriol/Ilya

% In our first experiment, we use the pre-specified logical grammar to
% generate controlled data sets and assess the ability of the two
% classes of neural network to learn the core underlying relational
% algebra from this simulated data. In our second experiment, we define
% a simple modification of the logical grammar so that it aligns with
% the WordNet noun graph, and we again assess our two classes of RNN on
% this data. The controlled data of our first experiment helps us to
% achieve analytic insights into what the models are able to learn, and
% the large, diverse lexicon of the second experiment shows us how they
% fare in the real world.

% These experiments differentiate the increased power of RNTNs better
% than previous work and provide the most convincing demonstration to
% date of the ability of neural networks to model semantic inferences
% in complex natural language sentences.

% TODO: Add citations on related work in structuring relations in embedding spaces:
% - Representing part–whole relations in conceptual spaces
% - Something McCallum/Universal Schema?

\begin{table*}[tp]
  \centering\small
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{l c l l} 
    \toprule
    Name & Symbol & Set-theoretic definition & Example \\ 
    \midrule
    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle $\natfor$ reptile}  \\ 
    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile $\natrev$ turtle}  \\ 
    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch $\nateq$ sofa} \\ 
    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle $\natalt$ warthog} \\ 
    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able $\natneg$ unable} \\
    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal $\natcov$ non-turtle} \\ 
    independence       & $x \natind y$   & (else) & \ii{turtle $\natind$ pet}\\
    \bottomrule
  \end{tabular}
  \protect\caption{\protect\label{b-table}The seven natural logic relations of \protect\cite{maccartney2009extended}. 
    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
    %is insufficient knowledge to choose a label.}
\end{table*}

\section{Neural network models for relation classification} \label{methods}

We build embedding-based models using the method of
\cite{Bowman:Potts:Manning:2014}, which is centered on the task of
labeling a pair of words or sentences with one of a small set of
logical relations. The architecture that we use, which is limited to
only pairs of single terms (such as words), is depicted in
Figure~\ref{sample-figure}. The model represents the two input terms
as embeddings, which are fed into a comparison function based on one
of two types of neural network layer functions to produce a
representation for the relationship between the two terms. This
representation is then fed into a softmax classifier, which outputs a
distribution over possible labels. The entire network, including the
embeddings, is trained using backpropagation with AdaGrad \cite{duchi2011adaptive}.

The simpler version of the comparison concatenates the two
input vectors before feeding them into a standard neural network (NN)
layer.  The more powerful neural tensor network (NTN) version uses an
additional third-order tensor parameter to allow for multiplicative
interactions between the two inputs \cite{chen2013learning}. For more
details on the implementation and training of the layer functions, see
\cite{Bowman:Potts:Manning:2014}.

\begin{figure}[tp]
  \centering
    \footnotesize

	\newcommand{\labeledtreenode}[4][3]{\put(#2){\makebox(0,0){{\fcolorbox{black}{#4}{\makebox(#1,0.3){#3}}}}}}

	\newcommand{\textlabel}[4][3.5]{\put(#2){\makebox(0,0){{\makebox(#1,0.3){#3}}}}}

	\definecolor{lexcolor}{HTML}{F5F7C4}
	\definecolor{compositioncolor}{HTML}{BBEBFF}
	\definecolor{comparisoncolor}{HTML}{FFC895}
	\definecolor{softmaxcolor}{HTML}{A5FF8A}

	\setlength{\unitlength}{0.61cm}

	\resizebox{3in}{!}{
	\begin{picture}(13.15, 5.5)

	  \labeledtreenode[2.4]{6.5,5}{$P(\sqsubset) = 0.8$}{softmaxcolor}  
	  \put(6.5,3.7){\vector(0,1){1}}  
	  \labeledtreenode[5]{6.5,3.4}{turtle \emph{vs.}~animal}{comparisoncolor}

	  \textlabel{2.75,5}{Softmax classifier}{black}
	  \textlabel{2.0,3.4}{\parbox{2cm}{Comparison N(T)N layer}}{black}

	  \textlabel{6.5,2.1}{Optional embedding}{black}
	  \textlabel{6.5,1.5}{transformation NN layers}{black}

	  \textlabel{6.5,0.5}{Learned word vectors}{black}

	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	  \put(1.5,0.75){\vector(0,1){0.85}}
	  \labeledtreenode{1.5,0.5}{turtle}{lexcolor}

	  \put(1.5,2.25){\vector(4,1){3.25}}
	  \labeledtreenode{1.5,1.9}{turtle}{compositioncolor}

	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	  \put(11.5,0.75){\vector(0,1){0.85}}
	  \labeledtreenode{11.5,0.5}{animal}{lexcolor}

	  \put(11.5,2.25){\vector(-4,1){3.25}}
	  \labeledtreenode{11.5,1.9}{animal}{compositioncolor}

	\end{picture}
	}
  \caption{The model structure used to compare \ii{turtle} and \ii{animal}.} 
  \label{sample-figure}
\end{figure}

This model used here differs from the one described in that work in
two ways. First, because the inputs are single terms, we do not use
a composition function here. Second, for our experiment on WordNet data, 
we introduce an additional neural network layer between the embedding 
input and the comparison function, which facilitates initializing the 
embeddings themselves from pretrained vectors and was found to help 
performance in that setting.

%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?

\section{Reasoning about semantic relations}\label{sec:join}

In this experiment, we test our model's ability to learn and use
the natural logic inference rules schematized in 
Figure~\ref{tab:jointable}. For instance, given that $a \natrev b$ and $b
\natrev c$, one can conclude that $a \natrev c$, by basic
set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$. Cells in the
table containing a dot correspond to pairs of relations for which no valid 
inference can be drawn in our logic.

% about the relations themselves that do not depend on the
% internal structure of the things being compared. For example, given
% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
% understanding $a$, $b$, or $c$. These seven relations support more
% than just transitivity: MacCartney and Manning's
% \cite{maccartney2009extended} join table defines 32 valid inferences
% that can be made on the basis of pairs of relations of the form $a R
% b$ and $b R' c$, including several less intuitive ones such as that if
% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create artificial boolean structures in which terms denote sets of entities from
a small domain (e.g., Figure~\ref{lattice-figure}), employ logical
deduction to identify the valid statements, divide those into train
and test sets, and remove from the test set those statements which
cannot be proven from the training statements
(Figure~\ref{unprovable}).
% using the logic depicted in Figure~\ref{lattice-figure}.
%
%
% We test the model's ability to learn this behavior by creating
% artificial data sets of terms which represent sets of numbers. Since
% MacCartney and Manning's set of relations hold between sets as well as
% between sentences, we can use the underlying set structure to generate
% the all of the relations that hold between any pair of these terms, as
% in Figure \ref{lattice-figure}. We train the model defined above on a
% subset of these relations, but rather then presenting the model with a
% pair of tree-structured sentences as inputs, simply present it with
% two single terms, each of which corresponds to a single vector in the
% (randomly initialized) vocabulary matrix $V$, ensuring that the model
% has no information about the terms being compared except the relations
% between them.
%
In our experiments, we create 80 randomly generated sets drawn from a
domain of seven elements. This yields 6400 statements about pairs of
formulae, of which 3200 are chosen as a test set, and that test set is
further reduced to the 2960 examples that can be provably derived from
the training data. We trained both the NN model and the NTN model on
these data
sets.% . In both cases, the models were implemented exactly as
% described in Section~\ref{methods}, but since the items being compared
% are single terms rather than full tree structures, the composition
% layer was not involved, and the two models differed only in which
% layer function was used for the comparison layer. We simply present
% the models with two single terms, each of which corresponds to a
% single vector in the (randomly initialized) vocabulary matrix $V$,
% thereby ensuring that the model has no information about the terms
% being compared except the relations between them.

\paragraph{Results} 
We found that the NTN model worked best with 11-dimensional vector
representations for the 80 sets and a 90-dimensional feature vector
for the classifier, though the performance of the model was not highly
dependant on either dimensionality setting. 
Averaging over five randomly generated data sets, the model was able to correctly label 98.1\% (standard error $0.67\%$) of the provably derivable test examples, and 87.7\%
($\textit{SE} = 3.59\%$) of the remaining test examples. The simpler NN worked
best with 11 and 75 dimensions, respectively, but was able to achieve
accuracies of only 84.8\% ($\textit{SE} = 0.84\%$) and 68.7\% ($\textit{SE} = 1.58\%$),
respectively. Training accuracy was 99.8\% ($\textit{SE} = 0.04\%$) for the NTN and 94.7\% ($\textit{SE} = 0.89\%$) for
the NN.

\begin{figure*}[tp]
  \centering
  \begin{subfigure}[b]{0.42\textwidth}
    \centering  
    \setlength{\arraycolsep}{8pt}
    \renewcommand{\arraystretch}{1.1}
    \newcommand{\UNK}{\cdot}  
    \resizebox{2.4in}{!}{$\begin{array}[c]{c@{ \ }|*{7}{c}|}
        % \hline
        \multicolumn{1}{c}{}
        & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
        \cline{2-8}
        \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
        \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
        \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
        \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
        \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
        \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
        \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
        \cline{2-8}
      \end{array}$}
    \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.  These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}. We assess our models' ability to reproduce such inferential paths.}
    \label{tab:jointable}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
    \setlength{\unitlength}{1cm}
    \resizebox{1.5in}{!}{\begin{picture}(5,5.5)
      \labelednode{2.50}{5}{}{1,2,3}

      \put(0.75,4){\line(3,1){1.5}}
      \put(2.5,4){\line(0,1){0.5}}
      \put(4.25,4){\line(-3,1){1.5}}

      \labelednode{0.75}{3.5}{a,b}{1,2}
      \labelednode{2.50}{3.5}{c}{1,3}
      \labelednode{4.25}{3.5}{d}{2,3}

      \put(0.75,2.5){\line(0,1){0.5}}
      \put(0.75,2.5){\line(3,1){1.5}}

      \put(2.5,2.5){\line(-3,1){1.5}}
      \put(2.5,2.5){\line(3,1){1.5}}

      \put(4.25,2.5){\line(0,1){0.5}}
      \put(4.25,2.5){\line(-3,1){1.5}}

      \labelednode{0.75}{2}{e,f}{1}
      \labelednode{2.50}{2}{}{2}
      \labelednode{4.25}{2}{g,h}{3}

      \put(2.5,1){\line(-3,1){1.5}}
      \put(2.5,1){\line(0,1){0.5}}
      \put(2.5,1){\line(3,1){1.5}}

      \labelednode{2.5}{0.5}{}{}
    \end{picture}}
    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
    some sets have multiple names, so that learning $\nateq$ is non-trivial.}\label{lattice-figure}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.2\textwidth}
    \centering
    \setlength{\tabcolsep}{12pt}
    \begin{tabular}[b]{c  c}
      \toprule
      Train & Test \\
      \midrule

      $a \nateq b$              & $b \nateq b$ \\
      $a \natfor e$              & $b \natfor e$ \\
      $a \natneg g$              & \strikeout{$e \nateq f$} \\
      $d \natfor h$              & \strikeout{$g \natrev d$} \\
      $e \natalt g$            & $h \natrev d$ \\ 
      $\vdots$		& $\vdots$	\\
      \bottomrule
    \end{tabular}

    \caption{A train/test split of the atomic statements about the
      model.  Test statements not provable from the training data are
      crossed out.}\label{unprovable}
  \end{subfigure}  
  \caption{Experimental goal and set-up for reasoning about semantic relations.}
  \label{exp1}
\end{figure*} 
% Note: None of these test examples is derivable from the shown training data, 
% but we suggest that there is additional training data, so we can cross lines
% out willy-nilly without fear.

% RNTN log: tue-j-11-6x80-hip.txt
% Train PER: 0.0021875

% RNN log: tue-j-11-6x80-r-75.txt
% Train PER: 0.047188

% and create a dataset consisting of the relations between every pair of
% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
% test dataset, and that test dataset was further split into the 2960
% examples that can be provably derived from the test data using
% MacCartney and Manning's join table (or by the symmetry of the
% relations in about half of the cases) and the 240 that
% cannot.

% We tested a version of both the RNN model and the RNTN model on these
% data. In both cases, the models were implemented exactly as described
% in Section~\ref{methods}, but since the items being compared are
% single terms rather than full tree structures, the composition layer
% was not used, and the two models differed only in which layer function
% was used for the comparison layer. We found that the RNTN model worked
% best with 11 dimensional vector representations for the 80 sets and a
% 90 dimensional feature vector for the classifier. This model was able
% to correctly label 99.3\% of the derivable test examples, and 99.1\%
% of the remaining examples. The simpler RNN model worked best with 11
% and 75 dimensions, respectively, but was able to achieve accuracies of
% only 90.0\% and 87.\%, respectively.

These results are fairly straightforward to interpret. The NTN model
was able to accurately encode the relations between the terms in the
geometric relations between their vectors, and was able to then use
that information to recover relations that were not overtly included
in the training data. In contrast, the NN was able to achieve this
behavior only incompletely. It is possible but not likely that it
could be made to find a good solution with further optimization on
different learning algorithms, or that it would do better on a larger
universe of sets, for which there would be a larger set of training
data to learn from, but the NTN is readily able to achieve these
effects in the setting discussed here. The fact that both models performed
well above chance on the unprovable examples indicates that both were able
to identify statistical cues to the properties of the underlying boolean structure,
but that information of this kind is not sufficient to yield the near-perfect
accuracy that we saw on the provable data.

\section{Reasoning about lexical relations in WordNet}\label{sec:wordnet}

Using simulated data as above is reassuring about what the models
learn and why, but we also want to know how they perform with a real
natural language vocabulary. Unfortunately, as far as we are aware,
there are no available resources labeling such a vocabulary with the
relations from Table~\ref{b-table}. However, the relations in WordNet
\cite{fellbaum2010wordnet} come close and pose the same substantive
challenges within a somewhat easier classification problem.

% WordNet \cite{fellbaum2010wordnet} provides a different type of
% lattice structure which is compatible with our learning paradigm.

\newcommand{\nodagger}{\phantom{$^\dagger$}}

\begin{table*}
\centering\resizebox{5.5in}{!}{
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{r r@{ }r r@{ }r r@{ }r r@{ }r r@{ }r} 
    \toprule
     Portion of & \multicolumn{4}{c}{NN} & \multicolumn{4}{c}{NTN} & \multicolumn{2}{c}{Baseline}\\
     training data  & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} \\
    \midrule
     100\%   & 99.73 &(0.04) & 99.37$^\dagger$ &(0.14)    & 99.61 &(0.02) & \textbf{99.95}\nodagger &(0.03) & 37.05 &(--)\\
     33\%    & 95.96 &(0.20) & 95.30\nodagger &(0.12)                & 95.83 &(0.35)          & 95.45$^\dagger$ &(0.31) & 37.05 &(--)\\
     11\%    & 91.11 &(0.24) & 90.81$^\dagger$ &(0.20)    & 91.27 &(0.27)          & 90.90$^\dagger$ &(0.13) & 37.05 &(--)\\
     % 2\%    & 86.0 &(100.0) & ... &(100.0)   & 86.2 &(100.0)           & 76.4 &(87.3) & 37.9 &(--)\\ % Accidental run with slightly different train/test split.
    \bottomrule
  \end{tabular}}
 \caption{Mean test \% accuracy scores (with standard error) on the WordNet data over five-fold crossvalidation. The baseline figure is the frequency of the most frequent class, \ii{hypernym}.\label{wn-table}} 
\end{table*}

We extract three types of relation from WordNet. \ii{Hypernym} and
\ii{hyponym} can be represented directly in the WordNet graph
structure, and correspond closely to the $\natrev$ and $\natfor$
relations from natural logic. As in natural logic, these relations are
mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
(perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then
\ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
terms, which share a direct hypernym, like \ii{dalmatian}, \ii{pug},
and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
terms tend to exclude one another, thereby providing a loose approximation 
of the natural logic exclusion relation $\natalt$. WordNet antonyms are both too narrowly defined
and too rare to use for this purpose. WordNet defines its relations over
sets of synonyms, rather than over individual terms, so we do not
include a \ii{synonym} or \ii{equivalent} relation, but rather
consider only one member of each set of synonyms. Word pairs which do
not fall into these three relations are not included in the data set.

To limit the size of the vocabulary without otherwise simplifying the learning problem, we extract all of the
instances of these three relations for single word nouns in WordNet that are hyponyms of the node 
\texttt{organism.n.01}. In order to balance the distribution of the classes, we slightly downsample instances 
of the \ii{coordinate} relation, yielding a total of 36,772 relations among 3,217 terms. We report results below using crossvalidation, choosing a disjoint 10\% test sample for each of five runs. Unlike in the previous experiment,
it is not straightforward here to determine in advance
how much data should be required to train an accurate model, so we performed training runs with 
various fractions of the remaining data. Embeddings were fixed at 25 dimensions and were initialized 
randomly or using distributional vectors from GloVe \cite{pennington2014glove}. The feature vector 
produced by the comparison layer was fixed at 80 dimensions.

\paragraph{Results} 
The structured WordNet data contains substantially more redundancy than the random artificial data above, and we find that the NTN performs perfectly with random initialization. The plain NN performs almost as well, providing
a point of contrast with the results of Section~\ref{sec:join}. We also find that initialization with GloVe is helpful in allowing the models to maintain fair performance
with smaller amounts of training data. Some of the randomly initialized model runs failed to learn
usable representations at all and labeled all examples with the most frequent labels. We excluded these runs from the statistics, but marked settings for which this occurred with the symbol $\dagger$. For all of the remaining runs, training accuracy was consistently above 99\%.

\section{Conclusion}\label{sec:discussion}

Ours is not the first investigation into logical behavior in vector space models: 
 there is extensive prior work on constructing compositional word meanings in vector spaces \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
 extracting entailment information from distributional vector space models \cite{baroni2012entailment,rei2014looking},
 and transferring information between logic programs and neural networks in 
 constrained settings \cite{holldobler1999approximating,hitzler2004logic}. However, this investigation
  is the first that we are aware of on logical behavior in representations learned within conventionally
   optimized neural network models.

This paper evaluated two neural models on the task of learning 
natural logic relations between distributed word representations. The
results suggest that at least the neural tensor network
has the capacity to
meet this challenge with reasonably-sized training sets, learning both 
to embed a vocabulary in a way that encodes a diverse
set of relations, and to subsequently use those embeddings to infer new
relations. In
\cite{Bowman:Potts:Manning:2014}, we extend these results to include
complex expressions involving logical connectives and quantifiers, with similar
conclusions about (tree structured recursive versions of) these models. These findings
are promising for the future of learned distributed representations in the
applied modeling of logical semantics.

\section{Acknowledgments}

We thank Jeffrey Pennington, Richard Socher, and audiences at CSLI, Nuance, and BayLearn.

\bibliography{MLSemantics} 
\bibliographystyle{aaai}

\end{document}

### 2015-01-07 writing/AAAI SSS '15/BowmanPottsManning.tex
AAAI paper updates

@@
 Distributed representations have been used successfully in a wide
-array of sophisticated language tasks (e.g., \cite{collobert2011natural}). %,
-% including sentiment analysis, analogy completion, relation
-% extraction, and named entity recognition.
+array of sophisticated language tasks, including part of speach tagging,
+\cite{collobert2011natural}, semantic role labeling \cite{collobert2011natural},
+sentiment analysis \cite{socher2011semi}, and translation \cite{sutskever2014sequence} 
+among many others.
 However, it remains an open question whether it is possible to train
@@
 
-% TODO: Better Gref citation; cite Oriol/Ilya
-
-% In our first experiment, we use the pre-specified logical grammar to
-% generate controlled data sets and assess the ability of the two
-% classes of neural network to learn the core underlying relational
-% algebra from this simulated data. In our second experiment, we define
-% a simple modification of the logical grammar so that it aligns with
-% the WordNet noun graph, and we again assess our two classes of RNN on
-% this data. The controlled data of our first experiment helps us to
-% achieve analytic insights into what the models are able to learn, and
-% the large, diverse lexicon of the second experiment shows us how they
-% fare in the real world.
-
-% These experiments differentiate the increased power of RNTNs better
-% than previous work and provide the most convincing demonstration to
-% date of the ability of neural networks to model semantic inferences
-% in complex natural language sentences.
-
-% TODO: Add citations on related work in structuring relations in embedding spaces:
-% - Representing part–whole relations in conceptual spaces
-% - Something McCallum/Universal Schema?
-
 \begin{table*}[tp]
@@
     $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
-    %is insufficient knowledge to choose a label.}
+    and the relation $\natind$ applies whenever none of the other six do.} 
 \end{table*}
@@
 labeling a pair of words or sentences with one of a small set of
-logical relations. The architecture that we use, which is limited to
-only pairs of single terms (such as words), is depicted in
+logical relations. Unlike in a classical inference
+setting, the model is not given specific premises to reason with at test time, but rather
+is expected to encode all of the information that it learns about each term (word) at training
+time in the term vectors, and to then use only the information encoded in the relevant pair of
+term vectors at test time. 
+The architecture that we use, which is limited to
+only pairs of single terms, is depicted in
 Figure~\ref{sample-figure}. The model represents the two input terms
@@
 distribution over possible labels. The entire network, including the
-embeddings, is trained using backpropagation with AdaGrad \cite{duchi2011adaptive}.
+embeddings, is trained using backpropagation with AdaGrad \cite{duchi2011adaptive}
+with L2 regularization.
 
 The simpler version of the comparison concatenates the two
-input vectors before feeding them into a standard neural network (NN)
-layer.  The more powerful neural tensor network (NTN) version uses an
+input vectors before feeding them into a LReLU \cite{maasrectifier} neural network (NN)
+layer.  The more powerful neural tensor network (NTN) version adds an
 additional third-order tensor parameter to allow for multiplicative
@@
 a composition function here. Second, for our experiment on WordNet data, 
-we introduce an additional neural network layer between the embedding 
+we introduce an additional $\tanh$ neural network layer between the embedding 
 input and the comparison function, which facilitates initializing the 
@@
 domain of seven elements. This yields 6400 statements about pairs of
-formulae, of which 3200 are chosen as a test set, and that test set is
-further reduced to the 2960 examples that can be provably derived from
-the training data. We trained both the NN model and the NTN model on
-these data
-sets.% . In both cases, the models were implemented exactly as
-% described in Section~\ref{methods}, but since the items being compared
-% are single terms rather than full tree structures, the composition
-% layer was not involved, and the two models differed only in which
-% layer function was used for the comparison layer. We simply present
-% the models with two single terms, each of which corresponds to a
-% single vector in the (randomly initialized) vocabulary matrix $V$,
-% thereby ensuring that the model has no information about the terms
-% being compared except the relations between them.
+formulae, of which 3200 are chosen as a test set. For 240 of the test examples, 
+there is not sufficient evidence in the training data to support the choice of any 
+one relation, and we  exclude these from our primary test set. 
+We trained both the NN model and the 
+NTN model on these data sets.
 
@@
 to identify statistical cues to the properties of the underlying boolean structure,
-but that information of this kind is not sufficient to yield the near-perfect
+thereby going beyond what one would expect from a logical proof system. However, 
+information of this kind is not sufficient to yield the near-perfect
 accuracy that we saw on the provable data.
-
-\section{Reasoning about lexical relations in WordNet}\label{sec:wordnet}
-
-Using simulated data as above is reassuring about what the models
-learn and why, but we also want to know how they perform with a real
-natural language vocabulary. Unfortunately, as far as we are aware,
-there are no available resources labeling such a vocabulary with the
-relations from Table~\ref{b-table}. However, the relations in WordNet
-\cite{fellbaum2010wordnet} come close and pose the same substantive
-challenges within a somewhat easier classification problem.
-
-% WordNet \cite{fellbaum2010wordnet} provides a different type of
-% lattice structure which is compatible with our learning paradigm.
 
@@
 
+\section{Reasoning about lexical relations in WordNet}\label{sec:wordnet}
+
+Using simulated data as above is reassuring about what the models
+learn and why, but we also want to know how they perform with a real
+natural language vocabulary. Unfortunately, as far as we are aware,
+there are no available resources labeling such a vocabulary with the
+relations from Table~\ref{b-table}. However, the relations in WordNet
+\cite{fellbaum2010wordnet} come close and pose the same substantive
+challenges within a somewhat easier classification problem.
+
+% WordNet \cite{fellbaum2010wordnet} provides a different type of
+% lattice structure which is compatible with our learning paradigm.
+
 We extract three types of relation from WordNet. \ii{Hypernym} and
@@
 terms tend to exclude one another, thereby providing a loose approximation 
-of the natural logic exclusion relation $\natalt$. WordNet antonyms are both too narrowly defined
+of the natural logic exclusion relation $\natalt$. WordNet's antonym relation, however, is both too narrowly defined
 and too rare to use for this purpose. WordNet defines its relations over

### 2015-01-08 writing/AAAI SSS '15/BowmanPottsManning.tex
Some tweaks

@@
 the seven relations defined in Table~\ref{b-table}. Its formal
-properties are now well-understood \cite{Icard:Moss:2013:LILT}, so it
+properties are now well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}, so it
 provides a rigorous set of goals for our neural models. To keep the
@@
 \begin{table*}[tp]
-  \centering\small
+  \centering%\small
   \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.1}
+  \renewcommand{\arraystretch}{1.2}
   \begin{tabular}{l c l l} 
@@
 
-	\resizebox{3in}{!}{
+	%\resizebox{3in}{!}{
 	\begin{picture}(13.15, 5.5)
@@
 	\end{picture}
-	}
-  \caption{The model structure used to compare \ii{turtle} and \ii{animal}.} 
+	%}
+        \caption{The model structure used to compare \ii{turtle} and
+          \ii{animal}.  learned word representations are fed through
+          either an NN or NTN comparison network and then to a softmax
+        classifier over the seven relations in Table~\ref{b-table}.}
   \label{sample-figure}
@@
 
-In this experiment, we test our model's ability to learn and use
-the natural logic inference rules schematized in 
-Figure~\ref{tab:jointable}. For instance, given that $a \natrev b$ and $b
-\natrev c$, one can conclude that $a \natrev c$, by basic
+In this experiment, we test our model's ability to learn and use the
+natural logic inference rules schematized in
+Figure~\ref{tab:jointable}. For instance, given that $a \natrev b$ and
+$b \natrev c$, one can conclude that $a \natrev c$, by basic
 set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$. Cells in the
-table containing a dot correspond to pairs of relations for which no valid 
+$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  There
+are 32 valid inferences that can be made on the basis of pairs of
+relations of the form $a R b$ and $b R' c$. Cells in the table
+containing a dot correspond to pairs of relations for which no valid
 inference can be drawn in our logic.
@@
 data to learn from, but the NTN is readily able to achieve these
-effects in the setting discussed here. The fact that both models performed
-well above chance on the unprovable examples indicates that both were able
-to identify statistical cues to the properties of the underlying boolean structure,
-thereby going beyond what one would expect from a logical proof system. However, 
-information of this kind is not sufficient to yield the near-perfect
-accuracy that we saw on the provable data.
+effects in the setting discussed here.
+
+It's noteworthy that both models performed well above chance on the
+unprovable examples. This indicates that both were able to identify
+statistical cues to the properties of the underlying boolean
+structure, thereby going beyond what one could obtain from the natural
+logic. Information of this kind is not sufficient to yield the
+near-perfect accuracy that we saw on the provable data, but it does
+highlight one of the ways in which trained inference models can turn
+out to be more powerful that rigidly defined proof systems.
 
@@
 \begin{table*}
-\centering\resizebox{5.5in}{!}{
+\centering%\resizebox{5.5in}{!}
+{
   \setlength{\tabcolsep}{15pt}
@@
 
-Ours is not the first investigation into logical behavior in vector space models: 
- there is extensive prior work on constructing compositional word meanings in vector spaces \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
- extracting entailment information from distributional vector space models \cite{baroni2012entailment,rei2014looking},
- and transferring information between logic programs and neural networks in 
- constrained settings \cite{holldobler1999approximating,hitzler2004logic}. However, this investigation
-  is the first that we are aware of on logical behavior in representations learned within conventionally
-   optimized neural network models.
+This investigation builds upon extensive prior research on
+constructing compositional word meanings in vector spaces
+\cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
+extracting entailment information from distributional vector space
+models \cite{baroni2012entailment,rei2014looking}, and transferring
+information between logic programs and neural networks in constrained
+settings \cite{holldobler1999approximating,hitzler2004logic}. Our
+primary contribution is to establish a tight connection between these
+approaches and formal developments within natural logic.  More
+specifically, we showed that neural network models, optimized using
+established techniques, can learn distributed representations that
+closely approximate the behavior one would expect from a natural logic
+inference algorithm.
+
+% Ours is not the first investigation into logical behavior in vector space models: 
+%  there is extensive prior work on constructing compositional word meanings in vector spaces \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
+%  extracting entailment information from distributional vector space models \cite{baroni2012entailment,rei2014looking},
+%  and transferring information between logic programs and neural networks in 
+%  constrained settings \cite{holldobler1999approximating,hitzler2004logic}. However, this investigation
+%   is the first that we are aware of on logical behavior in representations learned within conventionally
+%    optimized neural network models.
 
@@
 complex expressions involving logical connectives and quantifiers, with similar
-conclusions about (tree structured recursive versions of) these models. These findings
+conclusions about (tree-structured recursive versions of) these models. These findings
 are promising for the future of learned distributed representations in the

### 2015-01-08 writing/AAAI SSS '15/BowmanPottsManning.tex
A couple of new refs

@@
 by their inferential connections with other expressions
-\cite{vanBenthem08NATLOG,maccartney2009extended}. For instance,
+\cite{Katz72,maccartney2009extended,vanBenthem08NATLOG}. For instance,
 \word{turtle} is analyzed, not primarily by its extension in the
@@
         \caption{The model structure used to compare \ii{turtle} and
-          \ii{animal}.  learned word representations are fed through
-          either an NN or NTN comparison network and then to a softmax
+          \ii{animal}.  Learned word representations are fed through
+          either an NN or NTN comparison layer and then to a softmax
         classifier over the seven relations in Table~\ref{b-table}.}
@@
 relations from Table~\ref{b-table}. However, the relations in WordNet
-\cite{fellbaum2010wordnet} come close and pose the same substantive
+\cite{fellbaum2010wordnet,WordNet95} come close and pose the same substantive
 challenges within a somewhat easier classification problem.
@@
 information between logic programs and neural networks in constrained
-settings \cite{holldobler1999approximating,hitzler2004logic}. Our
+settings \cite{hitzler2004logic,holldobler1999approximating}. Our
 primary contribution is to establish a tight connection between these

### 2015-01-09 writing/AAAI SSS '15/BowmanPottsManning/BowmanPottsManning.tex
AAAI formatted SSS submission

--- 2015-01-08 writing/AAAI SSS '15/BowmanPottsManning.tex
@@
 is expected to encode all of the information that it learns about each term (word) at training
-time in the term vectors, and to then use only the information encoded in the relevant pair of
+time in the term embedding vectors, and to then use only the information encoded in the relevant pair of
 term vectors at test time. 
-The architecture that we use, which is limited to
-only pairs of single terms, is depicted in
+The architecture that we use is depicted in
 Figure~\ref{sample-figure}. The model represents the two input terms
-as embeddings, which are fed into a comparison function based on one
-of two types of neural network layer functions to produce a
+as single embedding vectors, which are fed into a comparison function based on one
+of two types of neural network layer function to produce a
 representation for the relationship between the two terms. This
@@
 embeddings, is trained using backpropagation with AdaGrad \cite{duchi2011adaptive}
-with L2 regularization.
+and L2 regularization.
 
@@
 
-	  \textlabel{6.5,0.5}{Learned word vectors}{black}
+	  \textlabel{6.5,0.5}{Learned term vectors}{black}
 
@@
         \caption{The model structure used to compare \ii{turtle} and
-          \ii{animal}.  Learned word representations are fed through
+          \ii{animal}.  Learned term representations are fed through
           either an NN or NTN comparison layer and then to a softmax
@@
 domain of seven elements. This yields 6400 statements about pairs of
-formulae, of which 3200 are chosen as a test set. For 240 of the test examples, 
-there is not sufficient evidence in the training data to support the choice of any 
-one relation, and we  exclude these from our primary test set. 
-We trained both the NN model and the 
-NTN model on these data sets.
+formulae, of which 3200 are chosen as a test set. 240 of the test examples 
+were excluded from the primary test set for lack of evidence in the training 
+data to support the choice of any one relation. 
+We then trained and evaluated both the NN model and the NTN model on these data sets.
 
@@
 
-\bibliography{MLSemantics} 
+\begin{thebibliography}{}
+
+\bibitem[\protect\citeauthoryear{Baroni \bgroup et al\mbox.\egroup
+  }{2012}]{baroni2012entailment}
+Baroni, M.; Bernardi, R.; Do, N.-Q.; and Shan, C.-c.
+\newblock 2012.
+\newblock Entailment above the word level in distributional semantics.
+\newblock In {\em Proc. EACL}.
+
+\bibitem[\protect\citeauthoryear{Bowman, Potts, and
+  Manning}{2014}]{Bowman:Potts:Manning:2014}
+Bowman, S.; Potts, C.; and Manning, C.
+\newblock 2014.
+\newblock Recursive neural networks can learn logical semantics.
+\newblock {arXiv} manuscript 1406.1827.
+
+\bibitem[\protect\citeauthoryear{Chen \bgroup et al\mbox.\egroup
+  }{2013}]{chen2013learning}
+Chen, D.; Socher, R.; Manning, C.; and Ng, A.
+\newblock 2013.
+\newblock Learning new facts from knowledge bases with neural tensor networks
+  and semantic word vectors.
+\newblock In {\em Proc. {ICLR}}.
+
+\bibitem[\protect\citeauthoryear{Clark, Coecke, and
+  Sadrzadeh}{2011}]{ClarkCoeckeSadrzadeh2011}
+Clark, S.; Coecke, B.; and Sadrzadeh, M.
+\newblock 2011.
+\newblock Mathematical foundations for a compositional distributed model of
+  meaning.
+\newblock {\em Linguistic Analysis} 36(1--4):345--384.
+
+\bibitem[\protect\citeauthoryear{Collobert \bgroup et al\mbox.\egroup
+  }{2011}]{collobert2011natural}
+Collobert, R.; Weston, J.; Bottou, L.; Karlen, M.; Kavukcuoglu, K.; and Kuksa,
+  P.
+\newblock 2011.
+\newblock Natural language processing (almost) from scratch.
+\newblock {\em JLMR} 12.
+
+\bibitem[\protect\citeauthoryear{Duchi, Hazan, and
+  Singer}{2011}]{duchi2011adaptive}
+Duchi, J.; Hazan, E.; and Singer, Y.
+\newblock 2011.
+\newblock Adaptive subgradient methods for online learning and stochastic
+  optimization.
+\newblock {\em JMLR}.
+
+\bibitem[\protect\citeauthoryear{Fellbaum}{2010}]{fellbaum2010wordnet}
+Fellbaum, C.
+\newblock 2010.
+\newblock Word{N}et.
+\newblock In {\em Theory and Applications of Ontology: Computer Applications}.
+  Springer.
+
+\bibitem[\protect\citeauthoryear{Grefenstette}{2013}]{grefenstette2013towards}
+Grefenstette, E.
+\newblock 2013.
+\newblock Towards a formal distributional semantics: Simulating logical calculi
+  with tensors.
+\newblock arXiv manuscript 1304.5823.
+
+\bibitem[\protect\citeauthoryear{Hermann, Grefenstette, and
+  Blunsom}{2013}]{Hermann-etal:2013}
+Hermann, K.; Grefenstette, E.; and Blunsom, P.
+\newblock 2013.
+\newblock ``{N}ot not bad'' is not ``bad'': A distributional account of
+  negation.
+\newblock In {\em Proc. of the 2013 Workshop on Continuous Vector Space Models
+  and their Compositionality}.
+
+\bibitem[\protect\citeauthoryear{Hitzler, H{{\"o}}lldobler, and
+  Seda}{2004}]{hitzler2004logic}
+Hitzler, P.; H{{\"o}}lldobler, S.; and Seda, A.~K.
+\newblock 2004.
+\newblock Logic programs and connectionist networks.
+\newblock {\em Journal of Applied Logic} 2(3).
+
+\bibitem[\protect\citeauthoryear{H{\"o}lldobler, Kalinke, and
+  St{\"o}rr}{1999}]{holldobler1999approximating}
+H{\"o}lldobler, S.; Kalinke, Y.; and St{\"o}rr, H.-P.
+\newblock 1999.
+\newblock Approximating the semantics of logic programs by recurrent neural
+  networks.
+\newblock {\em Applied Intelligence} 11(1).
+
+\bibitem[\protect\citeauthoryear{Icard and Moss}{2013a}]{Icard:Moss:2013}
+Icard, T., and Moss, L.
+\newblock 2013a.
+\newblock A complete calculus of monotone and antitone higher-order functions.
+\newblock In {\em Proc. Topology, Algebra, and Categories in Logic}.
+
+\bibitem[\protect\citeauthoryear{Icard and Moss}{2013b}]{Icard:Moss:2013:LILT}
+Icard, T., and Moss, L.
+\newblock 2013b.
+\newblock Recent progress on monotonicity.
+\newblock {\em Linguistic Issues in Language Technology} 9(7).
+
+\bibitem[\protect\citeauthoryear{Katz}{1972}]{Katz72}
+Katz, J.~J.
+\newblock 1972.
+\newblock {\em Semantic Theory}.
+\newblock New York: Harper \& Row.
+
+\bibitem[\protect\citeauthoryear{Maas, Hannun, and Ng}{2013}]{maasrectifier}
+Maas, A.; Hannun, A.; and Ng, A.
+\newblock 2013.
+\newblock Rectifier nonlinearities improve neural network acoustic models.
+\newblock In {\em Proc. ICML}.
+
+\bibitem[\protect\citeauthoryear{MacCartney and
+  Manning}{2009}]{maccartney2009extended}
+MacCartney, B., and Manning, C.
+\newblock 2009.
+\newblock An extended model of natural logic.
+\newblock In {\em Proc. IWCS}.
+
+\bibitem[\protect\citeauthoryear{Miller}{1995}]{WordNet95}
+Miller, G.~A.
+\newblock 1995.
+\newblock Word{N}et: A lexical database for {E}nglish.
+\newblock {\em Communications of the {ACM}} 38(11):39--41.
+
+\bibitem[\protect\citeauthoryear{Pennington, Socher, and
+  Manning}{2014}]{pennington2014glove}
+Pennington, J.; Socher, R.; and Manning, C.
+\newblock 2014.
+\newblock Glo{V}e: {G}lobal vectors for word representation.
+\newblock In {\em Proc. EMNLP}.
+
+\bibitem[\protect\citeauthoryear{Rei and Briscoe}{2014}]{rei2014looking}
+Rei, M., and Briscoe, T.
+\newblock 2014.
+\newblock Looking for hyponyms in vector space.
+\newblock In {\em Proc. {CoNLL}}.
+
+\bibitem[\protect\citeauthoryear{Rockt{\"a}schel \bgroup et al\mbox.\egroup
+  }{2014}]{rocktaschellow}
+Rockt{\"a}schel, T.; Bosnjak, M.; Singh, S.; and Riedel, S.
+\newblock 2014.
+\newblock Low-dimensional embeddings of logic.
+\newblock In {\em Proc. the ACL 2014 Workshop on Semantic Parsing}.
+
+\bibitem[\protect\citeauthoryear{Socher \bgroup et al\mbox.\egroup
+  }{2011}]{socher2011semi}
+Socher, R.; Pennington, J.; Huang, E.; Ng, A.; and Manning, C.
+\newblock 2011.
+\newblock Semi-supervised recursive autoencoders for predicting sentiment
+  distributions.
+\newblock In {\em Proc. EMNLP}.
+
+\bibitem[\protect\citeauthoryear{Sutskever, Vinyals, and
+  Le}{2014}]{sutskever2014sequence}
+Sutskever, I.; Vinyals, O.; and Le, Q.
+\newblock 2014.
+\newblock Sequence to sequence learning with neural networks.
+\newblock In {\em Proc. {NIPS}}.
+
+\bibitem[\protect\citeauthoryear{van Benthem}{2008}]{vanBenthem08NATLOG}
+van Benthem, J.
+\newblock 2008.
+\newblock A brief history of natural logic.
+\newblock In Chakraborty, M.; L{\"o}we, B.; Nath~Mitra, M.; and Sarukki, S.,
+  eds., {\em Logic, Navya-Nyaya and Applications: Homage to {B}imal {M}atilal}.
+
+\end{thebibliography}
 \bibliographystyle{aaai}

### 2015-02-21 writing/F'14 paper/discussion.tex
Paper presentation tweaks; SNLI setup; misc updates

@@
 of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
-In addition,
-there remain subtle questions about how to fairly assess whether these
-models have truly generalized in the way we want them to. There is a
-constant tension between showing the models training data that gives
-them a chance to learn the target logical functions and revealing the
-answer to them in a way that leads to overfitting. The underlying
-logical theories provide only limited guidance on this point.
+In addition, interesting analytical questions remain about \ii{how} these models encode
+the underlying logics. Neither the underlying
+logical theories, nor any straightforward parameter inspection technique provides 
+much insight on this point, but we hope that further experiments may reveal 
+some structure in the representations.
 
@@
 of the logical complexity of natural language; in future experiments, we hope to test sentences
-with embedded quantifiers, multiple interacting quantifiers, relative
-clauses, and other kinds of recursive structure. Our SICK experiments 
+with more complex types of structure, as well as run similar experiments on non-tree structured models, such as LSTMs. Our SICK experiments 
 similarly only begin to reveal the potential of these models to learn to 

### 2015-02-21 writing/F'14 paper/intro.tex
Paper presentation tweaks; SNLI setup; misc updates

@@
 
-Supervised recursive neural network models (RNNs) for sentence meaning
+% TODO: TreeRNN?
+% TODO: More language about the representations being the object of study.
+
+Supervised recursive neural network models (RNNs, \newcite{goller1996learning}) for sentence meaning
 have been successful in an array of sophisticated language tasks,
@@
 \newcite{maccartney2009extended} as our formal model. This logic
-defines seven core relations of synonymy, entailment, contradiction,
+defines seven mutually-exclusive relations of synonymy, entailment, contradiction,
 and mutual consistency, as summarized in Table~\ref{b-table}, and it
@@
  moves}. Though the performance of the plain RNN model 
- is somewhat poor in our first experiment, we find that the stronger RNTN
-  model generalizes well in every case, suggesting that it has in fact
-  learned, or at least learned to simulate, our target logical
-  concepts.
+ is somewhat poor in our first experiment, we find that the stronger RNTN model
+  generalizes well in every case, suggesting that it has in fact learned, or at least
+ learned to simulate, our target logical concepts.
 

### 2015-02-21 writing/F'14 paper/join.tex
Paper presentation tweaks; SNLI setup; misc updates

@@
 We trained versions of both the RNN model and the RNTN model on these
-data sets. In both cases, the models are implemented as
+data sets\footnote{Since this task relies entirely on the learning of the two entity vectors, no obvious comparable baseline presents itself.}. In both cases, the models are implemented as
 described in Section~\ref{methods}, but since the items being compared
 are single terms rather than full tree structures, the composition
-layer is not involved, and the two models differ only in which
+layer is not involved, and the two models are not recursive and differ only in which
 layer function is used for the comparison layer. We simply present
@@
 
-      $a \nateq b$              & $b \nateq b$ \\
-      $a \natfor e$              & $b \natfor e$ \\
-      $a \natneg g$              & \strikeout{$e \nateq f$} \\
-      $d \natfor h$              & \strikeout{$g \natrev d$} \\
-      $e \natalt g$            & $h \natrev d$ \\
+      $a \nateq b$              & $b \natneg g$ \\
+      $a \natrev e$              & $b \natrev e$ \\
+      $d \natrev h$              & \strikeout{$e \nateq f$} \\
+       $e \natalt g$              & \strikeout{$g \natfor d$} \\
+      $g \natneg a$           & $h \natfor d$ \\
 

### 2015-02-21 writing/F'14 paper/quantifiers.tex
Paper presentation tweaks; SNLI setup; misc updates

@@
 %
-\paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic satisfactorily.
+\paragraph{Results} The results, reported in Table~\ref{qresultstable}, show that both models are able to learn to generalize the underlying logic satisfactorily.
 We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.
+
+% TODO: Baseline

### 2015-02-21 writing/F'14 paper/recursion.tex
Paper presentation tweaks; SNLI setup; misc updates

@@
       $a$               & $\natfor$ & $(a \plor b)$ \\
-      $a$               & $\natrev$ & $(a \pland b)$ \\
+      $(a \plor (b \plor c))$               & $\natrev$ & $(b \pland  \plneg c)$ \\
       %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
@@
     \end{tabular}
-    \caption{Examples of statements about relations between
-      well-formed formulae, defined in terms of sets of satisfying
+    \caption{Examples of the type of statements used for training and testing. These are relations between
+      well-formed formulae, computed in terms of sets of satisfying
       interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
@@
 each operator can be represented accurately by a single bit.
-In the propositional logic used here, the atomic symbols are variables over those values. 
-Statements of this logic thus describe conditions on
-unknown truth values of propositions. This logic distinguishes
-between $2^{6} = 64$ possible assignments of truth values, and
-expressions of this logic define arbitrary conditions on these
-possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
-possible statements that the intermediate vector representations need
-to be able to distinguish if we are to learn this logical system in full.
+In the more demainding task presented here, the atomic symbols are variables over binary values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions over valuations of
+these variables. Relational expressions between these conditions are essentially theorems of propositional logic, and to succeed, the model must learn to approximate the behavior of a theorem prover.
 
-In our experiments, we randomly generate a large set of unique pairs 
+In our experiments, we randomly generate unique pairs 
 of formulae and compute the relation that holds for each pair.
-We discard pairs in which either statement is a tautology or
-contradiction (a statement that is true of either all or no possible
-assignments), for which none of the seven relation labels in
-Table~\ref{b-table} can hold. The resulting set of formula pairs is
-then partitioned into bins based on the number of logical operators in
-the longer of the two formulae. We then randomly sample 20\% of each
+We discard pairs in which either statement is either a tautology or a
+contradiction, for which the seven relations in
+Table~\ref{b-table} are undefined. The resulting set of formula pairs is
+then partitioned into 13 bins according the number of logical operators in
+the longer of the two formulae. We then sample 20\% of each
 bin for a held-out test set.
@@
 referring to more than four of the six propositional variables.
+
+In order to test the model's generalization to unseen structures, we discard
+training examples with more than 4 connectives, yielding 60k short training examples,
+and 21k training examples containing up to 12 connectives.
+We evaluated both the RNN and RNTN\footnote{We did not run any strictly simpler baselines for this experiment because it is demonstrably impossible to perform substantially above chance on longer expressions without word order information.} on the resulting data, initializing the model parameters randomly,
+including the vector representations of the six variables.
 
@@
 
-\paragraph{Results} 
-We trained both the RNN and RNTN models on data with up to 4
-connectives (60k pairs) and tested it on a separate set of examples of up to 12
-connectives (21k pairs). We initialized the model parameters randomly,
-including the vector representations of the six variables.
-Figure~\ref{prop-results} shows test the relationship
+\paragraph{Results} Figure~\ref{prop-results} shows test the relationship
 between test accuracy and statement length. We found that both 

### 2015-02-21 writing/F'14 paper/sick.tex
Paper presentation tweaks; SNLI setup; misc updates

@@
   \caption{\label{examplesofsickdata}Examples of correct RNTN classifications on SICK.}
-\end{table*}
+\end{table*} % TODO: Update. Find incorrect examples?
 
@@
 We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.
-We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}), a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn.
+We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}, also used by the winning SICK submission), a corpus of noisy automatically labeled entailment examples over image captions, the same genre of text from which SICK was drawn.
 We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source.
@@
 
+% TODO: Update results
+
 Table \ref{examplesofsickdata} shows a few typical examples of correct classifications.

### 2015-02-21 writing/W'15 paper/newcorpus.tex
Paper presentation tweaks; SNLI setup; misc updates

\section{Discussion and conclusion}\label{sec:discussion}

data will be freely available

SICK presented an important challenge, we try to deliver on it.

aim to surpass sick in both size and label consistency

quote from SICK paper \cite{marelli2014sick}:

\begin{quote}
Not unreasonably, subjects found that, say, \ii{A woman is wearing an Egyptian headdress} does not contradict \ii{A woman is wearing an Indian headdress}, since one could easily imagine both sentences truthfully uttered to refer to a single scene where two different women are wearing different headdresses. In the future, a higher proportion of CONTRADICTION labels could be elicited by using grammatical and possibly visual cues (pictures) encouraging co-indexing of the entities in the two sentences.
\end{quote}

We use captions from the Flickr30k corpus \cite{hodoshimage} as the left hand side of each relation pair. That corpus ...

We randomly selected individual captions from that corpus to use in our own corpus, though in the interest of encouraging more complex source sentences without discarding too much source data, we downsample sentences with lengths of under 90 characters so that these sentences make up only half of our corpus, rather than the original 69\%.

... licencing ...

The instructions for this phase were integrated into the design of the data collection interface. The text of those instructions is shown in Figure~\ref{instructions-1}. We chose to use three classes, corresponding to ENTAILMENT, CONTRADICTION, and NEUTRAL classes used in SICK.

\begin{figure}
\footnotesize
\toprule
The Stanford University NLP Group is collecting data for use in research on computer understanding of English. We appreciate your help!

We will show you the caption for a photo. We will not show you the photo. Using only the caption and what you know about the world:
\begin{itemize}
\item Write one alternate caption that is \textbf{definitely} a \textbf{true} description of the photo. \ii{Example: For the caption "Two dogs are running through a field." you could write "There are animals outdoors."}
\item Write one alternate caption that \textbf{might be} a \textbf{true} description of the photo. \ii{Example: For the caption "Two dogs are running through a field." you could write "There are animals outdoors."}
\item Write one alternate caption that is \textbf{definitely} a \textbf{false} description of the photo. \ii{Example: For the caption "Two dogs are running through a field." you could write "There are animals outdoors." This is different from the maybe correct category because its impossible for the dogs to be both running and sitting.}
\end{itemize}
\textbf{Problems} (optional)   If something is wrong with the caption that makes it difficult to understand, do your best above and let us know here.
\bottomrule
\caption{\label{instructions-1}Examples of training data from the newly collected entailment corpus.}
\end{figure}

We posted 15.5k ImageFlickr captions. We showed each of the first 10k captions to five workers, yielding fifteen reponse sentences per source caption, with five assigned to each of the three labels. Then, to ensure maximal diversity in the test set, we showed the remaining 5.5k source captions to one worker each,  yielding only three responses per source caption.

We collected 143,023 pairs, after excluding less than 100 worker responses in which the worker reported that they could not understand the source caption, or in which a field was accidentally left blank. The average worker response sentence was 7.8 words long. 1100 workers participated in the task.

pre-verification distribution:

Recompute after removing blanks:
29305 contra
29304 neut
29305 ent

% ditto examples

\begin{table*}
  \centering\footnotesize
  \begin{tabular}{p{6.5cm}cp{6.5cm}}
  \toprule
A busy street with building lined up and people walking down the street outside and nighttime. &\ii{entailment}	&People lined up\\
\rule{0pt}{3ex}Two men stand in an electric outdoor lift. &\ii{neutral}	& The two men are inspectors for the FDA, waiting to inspect the second level of the pharmaceutical plant.\\
\rule{0pt}{3ex}A young redheaded girl, wearing a yellow shirt, black pants, and sneakers, jumping in a grassy field with blue skies and wispy clouds in the background. &\ii{entailment}	& A girl jumps in a grassy field\\
\rule{0pt}{3ex}A crowd of people, one with a guitar, are standing in the street under the 7th Avenue street sign. &\ii{neutral}	& The crowd listening to the man play guitar.\\
\rule{0pt}{3ex}A woman wearing bike shorts and a skirt is riding a bike and carrying a shoulder bag.  &\ii{contradiction}& A woman driving a car.\\
    \bottomrule
  \end{tabular}
  \caption{\label{examplesofscedata}The instructions presented to workers during data collection.}
\end{table*}

% Screenshot from phase 1

% Screenshot from phase 2

The data is freely available at (POST DATA), and is released under a CreativeCommons
Attribution-ShareAlike licence, which is the same licence used for the Flickr30k source captions.

\paragraph{validation}

The majority of the sentence pairs in the corpus have not been reviewed by anyone since their construction as part of the data collection Mechanical Turk task, except to remove blank responses and responses where the original author flagged the source caption for review. However, in order to measure the accuracy of our corpus, and in order to improve the quality of our test and development sets, we did perform an additional round of validation for TODOk examples.

This validation phase is not fundamentally different from the Mechanical Turk labeling task used for the SICK entailment data: we present workers with pairs of sentences (in our case, in batches of five per screenfull), and ask them to choose a single label for each pair. We supplied each pair to four annonators, yielding five labels per pair including the label used by the original composer of the pair.

TODO: Data filtering? Data re-labeling?

\begin{figure}
\footnotesize
\toprule
TODO: VALIDATION INSTRUCTIONS
\bottomrule\\

\caption{\label{instructions-2}The instructions presented to workers during data validation.}
\end{figure}

\paragraph{data partition}

\paragraph{simple baselines}

MFC

BLEU

Summing + NN

### 2015-02-21 RC/generate_neg_set_data.py

@@ -3,156 +3,167 @@ from itertools import *
 from collections import *
 import random
 
+
 def powerset(iterable):
     "From itertools: powerset([1,2,3]) --> () (1,) (2,) (3,) (1,2) (1,3) (2,3) (1,2,3)"
     s = list(iterable)
-    return chain.from_iterable(combinations(s, r) for r in range(len(s)+1))
+    return chain.from_iterable(combinations(s, r) for r in range(len(s) + 1))
+
 
 def get_candidate_worlds(num_vars):
-	return powerset(set(range(num_vars)))
+    return powerset(set(range(num_vars)))
+
 
 def get_satisfying_worlds_for_tree(tree, candidate_worlds):
-	if isinstance(tree, tuple):
-		if tree[0] == 'not':
-			child = get_satisfying_worlds_for_tree(tree[1], candidate_worlds)
-			return candidate_worlds.difference(child)
-	 	else:
-	 		left = get_satisfying_worlds_for_tree(tree[0], candidate_worlds)
-			right = get_satisfying_worlds_for_tree(tree[2], candidate_worlds)
-			if tree[1] == "and":
-				return left.intersection(right)
-			elif tree[1] == "or":
-	 			return left.union(right)
-	 		else:
-	 			print 'syntax error', tree
- 	else:
- 		result = []
- 		for world in candidate_worlds:
- 			if tree in world:
- 				result.append(world)
-		return set(result)
+    if isinstance(tree, tuple):
+        if tree[0] == 'not':
+            child = get_satisfying_worlds_for_tree(tree[1], candidate_worlds)
+            return candidate_worlds.difference(child)
+        else:
+            left = get_satisfying_worlds_for_tree(tree[0], candidate_worlds)
+            right = get_satisfying_worlds_for_tree(tree[2], candidate_worlds)
+            if tree[1] == "and":
+                return left.intersection(right)
+            elif tree[1] == "or":
+                return left.union(right)
+            else:
+                print 'syntax error', tree
+    else:
+        result = []
+        for world in candidate_worlds:
+            if tree in world:
+                result.append(world)
+        return set(result)
+
 
 def compute_relation(left, right, universe):
-	ne_intersection = left.intersection(right)
-	ne_just_left = left.difference(right)
-	ne_just_right = right.difference(left)
[…]

### 2015-02-21 RunF14Experiments/RunAndOrExperiments.sh

@@ -11,3 +11,33 @@ export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.001; dim = 25; 
 export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.001; dim = 25; td = 1; penult = 75; dropout = 1; tot = 1; name='taf3'; relu = 1; TrainModel(''\, 3\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh 
 export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.001; dim = 25; td = 1; penult = 75; dropout = 1; tot = 1; name='taf3'; relu = 1; TrainModel(''\, 4\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh 
 export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.001; dim = 25; td = 1; penult = 75; dropout = 1; tot = 1; name='taf3'; relu = 1; TrainModel(''\, 5\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh 
+
+export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.001; dim = 25; td = 3; penult = 75; dropout = 1; tot = 0; name='taf4'; relu = 1; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh 
+export MATLABCMD="cd quant; dataflag = 'and-or-deep'; lambda = 0.001; dim = 25; td = 3; penult = 75; dropout = 1; tot = 1; name='taf4'; relu = 1; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh 
[…]

### 2015-02-21 config/Defaults.m

@@ -19,6 +19,9 @@ hyperParams.tensorScale = 0.9;
 % How much of the output of the matrix parameters (in the range 0-1) should be initialized with an identity matrix.
 hyperParams.eyeScale = 0.5;
 
+% Use an older initialization scheme for comparability with older experiments.
+hyperParams.useCompatibilityInitialization = true;
+
 % The number of embedding transform layers. topDepth > 0 means NN layers will be
 % added above the embedding matrix. This is likely to only be useful when
 % learnWords is false, and so the embeddings do not exist in the same space
@@ -138,7 +141,9 @@ options.confusionFreq = 250;
 options.examplesFreq = 1000; 
 
 % How often (in steps) to save parameters to disk.
-options.checkpointFreq = 20000; 
+% Checkpoints saved at these intervals coexist with checkpoints created
+% automatically whenever a the best accuracy score is updated.
+options.checkpointFreq = 100000; 
 
 % The name assigned to the current call to TrainSGD. This can be used to
 % distinguish multiple phases of training in the same experiment.

### 2015-02-22 writing/F'14 paper/abstract.tex
Paper update

@@
-% \begin{abstract}
-% Recursive neural network models for sentence meaning have been successful in many tasks, but it remains an open question whether any such model can learn compositional semantic representations that support logical deduction.
-We pursue this question directly by evaluating whether two such models can correctly learn to identify logical relationships such as entailment and contradiction between sentences.
-We first describe a recursive neural tensor network model for entailment and demonstrate it on the SICK challenge.
-We then use artificial data from a logical grammar to test the model's ability to learn to handle basic relational reasoning, recursive functions, and quantification.
-The model performs competitively on the SICK data and generalizes well in all three experiments on simulated data, suggesting that it can learn suitable representations for logical inference in natural language.
-% \end{abstract}
-
-%%%% POTTS REVISION
 \begin{abstract} 
@@
   successful for many tasks, but it remains an open question whether
-  any such model can learn compositional semantic representations that
-  support logical deduction. We pursue this question by
-  evaluating whether two such models---RNNs and recursive neural
+  the fixed-length representations that they learn can support tasks as demanding as logical deduction.
+  We pursue this question by evaluating whether two such models---plain RNNs and recursive neural
   tensor networks (RNTNs)---can correctly learn to identify logical
-  relationships such as entailment and contradiction. % between sentences.
+  relationships such as entailment and contradiction using these representations. 
   In our first set of experiments, we generate artificial data from a

### 2015-02-22 writing/F'14 paper/intro.tex
Paper update

@@
 % TODO: More language about the representations being the object of study.
+% TODO: Is there any way to squeeze a useful example into the first page?
 
-Supervised recursive neural network models (RNNs, \newcite{goller1996learning}) for sentence meaning
+Supervised recursive neural network models (RNNs; \citealt{goller1996learning}) for sentence meaning
 have been successful in an array of sophisticated language tasks,
@@
 \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
-on the core semantic issues of quantification,
-entailment, and contradiction needed to identify the relationships
-between sentences like \ii{every animal walks}, \ii{every turtle
- moves}, and \ii{most reptiles don't move}.
+in their ability to model core semantic phenomena like quantification, entailment, and contradiction.
 
@@
 provides rules of semantic combination for projecting these relations
-from the lexicon up to complex phrases. The formal properties and
-inferential strength of this system are now well-understood
+from the lexicon up to complex phrases. The formal properties of this system are now well-understood
 \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
@@
 relationships correctly. In our first experiment
-(Section~\ref{sec:join}), we evaluate the ability of these models to
+(\S\ref{sec:join}), we evaluate the ability of these models to
 learn the core relational algebra of natural logic from data. Our
-second experiment (Section~\ref{sec:recursion}) extends this
+second experiment (\S\ref{sec:recursion}) extends this
 evaluation to relations between complex recursive structures like
-$(a \plor b)$ and $\plneg~(\plneg a \pland \plneg b)$, and our third
-experiment (Section~\ref{sec:quantifiers}) covers relations between
+$(a \plor b)$ and $\plneg (\plneg a \pland \plneg b)$, and our third
+experiment (\S\ref{sec:quantifiers}) covers relations between
 statements with quantification and negation like \ii{every reptile walks} and \ii{no turtle
@@
  is somewhat poor in our first experiment, we find that the stronger RNTN model
-  generalizes well in every case, suggesting that it has in fact learned, or at least
- learned to simulate, our target logical concepts.
+  generalizes well in every case, suggesting that it has learned to simulate our target logical concepts.
 

### 2015-02-22 writing/F'14 paper/join.tex
Paper update

@@
+\section{Reasoning about semantic relations}\label{sec:join}
+
+If a model is to learn the behavior of a relational logic like the one
+presented here from a finite amount of data, it must learn to deduce new
+relations from seen relations in a sound manner. The simplest such
+deductions involve atomic statements using the relations in
+Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
+\natrev c$, one can conclude that $a \natrev c$, by basic
+set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
+$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
+full set of sound such inferences is summarized in
+Table~\ref{tab:jointable}.
+
+% about the relations themselves that do not depend on the
+% internal structure of the things being compared. For example, given
+% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
+% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
+% understanding $a$, $b$, or $c$. These seven relations support more
+% than just transitivity: MacCartney and Manning's
+% \cite{maccartney2009extended} join table defines 32 valid inferences
+% that can be made on the basis of pairs of relations of the form $a R
+% b$ and $b R' c$, including several less intuitive ones such as that if
+% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
+
 \begin{table}[htp]
@@
 
-\section{Reasoning about semantic relations}\label{sec:join}
-
-If a model is to learn the behavior of a relational logic like the one
-presented here from a finite amount of data, it must learn to deduce new
-relations from seen relations in a sound manner. The simplest such
-deductions involve atomic statements using the relations in
-Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
-\natrev c$, one can conclude that $a \natrev c$, by basic
-set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
-full set of sound such inferences is summarized in
-Table~\ref{tab:jointable}.
-
-% about the relations themselves that do not depend on the
-% internal structure of the things being compared. For example, given
-% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
-% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven relations support more
-% than just transitivity: MacCartney and Manning's
-% \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of relations of the form $a R
-% b$ and $b R' c$, including several less intuitive ones such as that if
-% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-
 \paragraph{Experiments}
-To test our models' ability to learn these inferential patterns, we
+To generate data for these inferential patterns, we
 create small boolean structures for our logic in which terms denote
@@
 depicts a structure of this form. The lattice gives the full model,
-for which all the statements below it are valid. We divide these
-statements evenly into training and test sets, and remove from the
+and we generate a relational statement for each pair of labels in the model. 
+We divide these statements evenly into training and test sets, and remove from the
 test set statements which cannot be proven from the training
 statements.
-% using the logic depicted in Figure~\ref{lattice-figure}.
+In each experimental run, we create 80 randomly generated sets drawing from
+a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
+2960 relations.
 
-In each experimental run, we create 80 randomly generated sets drawing from
-a domain of seven elements. This yields a data set consisting of
-6400 statements about pairs of formulae. 3200 of these pairs are
-chosen as a test set, and that test set is further reduced to the 2960
-examples that can be provably derived from the training data.
-
-We trained versions of both the RNN model and the RNTN model on these
-data sets\footnote{Since this task relies entirely on the learning of the two entity vectors, no obvious comparable baseline presents itself.}. In both cases, the models are implemented as
+We trained models with both the RNN and RNTN comparison functions on these
+data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no strictly simpler model is likely viable as a baseline.} In both cases, the models are implemented as
 described in Section~\ref{methods}, but since the items being compared
 are single terms rather than full tree structures, the composition
-layer is not involved, and the two models are not recursive and differ only in which
-layer function is used for the comparison layer. We simply present
+layer is not used, and the two models are not recursive. We simply present
 the models with the (randomly initialized) embedding vectors for each
@@
 Table \ref{joinresultstable} shows the results. 
-The RNTN is able to accurately encode the relations 
+The NTN is able to accurately encode the relations 
 between the terms in the geometric relations between their vectors, 
 and is able to then use that information to recover relations that 
-are not overtly included in the training data. The RNN also generalizes fairly well, 
+are not overtly included in the training data. The NN also generalizes fairly well, 
 but makes enough errors that it remains an open question whether 
@@
 We cannot rule out the possibility that different optimization techniques or
-further hyperparameter tuning could lead an RNN model to succeed here.
+further hyperparameter tuning could lead an NN model to succeed here.

### 2015-02-22 writing/F'14 paper/methods.tex
Paper update

@@
 To apply these recursive models to our task, we propose the tree 
-pair model architecture depicted in Figure~\ref{sample-figure}, which is novel but
-closely modeled after past work on recursive NNs. 
+pair model architecture depicted in Figure~\ref{sample-figure}. 
 In it, the two phrases being compared are processed separately using a pair 
@@
 %
-\begin{gather} \label{rnn}
-\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}\,) \\
+\begin{gather} 
+\label{rnn}
+\vec{y}_{\textit{TreeRNN}} = f(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
 \label{rntn} 
-\vec{y}_{\textit{RNTN}} = \vec{y}_{\textit{RNN}} + f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
+\vec{y}_{\textit{TreeRNTN}} = \vec{y}_{\textit{TreeRNN}} + f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
 \end{gather} 
@@
 Because the classes are not balanced in the artificial data experiments, we report performance
-using both accuracy and macroaveraged F1, since the latter emphasizes
- performance on infrequent classes. We compute macroaveraged F1 
+using both accuracy and macroaveraged F1.\footnote{We compute macroaveraged F1 
 as the harmonic mean of average precision and average recall, both computed
 for all classes for which there is test data, setting precision to 0 
-where it is not defined. On the artificial data experiments, we report mean
+where it is not defined.} On the artificial data experiments, we report mean
 results over five fold cross-validation. Variance across runs is typically no

### 2015-02-22 writing/F'14 paper/recursion.tex
Paper update

@@
 \paragraph{Experiments}
-As in Section~\ref{sec:join}, we test this phenomenon within the
+As in \S\ref{sec:join}, we test this phenomenon within the
 framework of natural logic, but we now replace the unanalyzed symbols
@@
 a logic where the atomic symbols are simply the
-values $0$ and $1$. While learning the operators of that logic is not trivial, the outputs of
+values $\True$ and $\False$. While learning the operators of that logic is not trivial, the outputs of
 each operator can be represented accurately by a single bit.
-In the more demainding task presented here, the atomic symbols are variables over binary values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions over valuations of
-these variables. Relational expressions between these conditions are essentially theorems of propositional logic, and to succeed, the model must learn to approximate the behavior of a theorem prover.
+In the more demainding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.
+Relational expressions between these conditions are essentially theorems of propositional logic, and to succeed, the model must learn to approximate the behavior of a theorem prover.
 

### 2015-02-22 config/AndOr.m

@@ -93,7 +93,7 @@ options.confusionFreq = 1000;
 
 % How often to display which items are misclassified.
 % Should be a multiple of testFreq.
-options.examplesFreq = 10000; 
+options.examplesFreq = 25000; 
 
 
 end

### 2015-02-24 writing/F'14 paper/intro.tex
Experimental setup tweaks; add TODOs to paper

@@
 
-% TODO: TreeRNN?
+% TODO: TreeRNN.
 % TODO: More language about the representations being the object of study.
 % TODO: Is there any way to squeeze a useful example into the first page?
+% TODO: Clarify early that we study vector pair classification. Maybe move up fig. 1?
+% TODO: Clarify that we aren't implementing NatLog rules.
+% TODO: ``Something that does that'' epigram in intro.
+% TODO: Mention tree pair archichture.
 
@@
 \cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
-the ability of these models to learn compositional semantic grammars.
-However, it remains an open question whether they can achieve the
+the ability of these models to learn compositional semantic grammars, though many remain skeptical of the extent to which vector space representations can model complex compositional semantics, and it remains an open question whether learned can achieve the
 high-fidelity distributed representations of recent algebraic
-approaches \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
+approaches to vector space modeling \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
 and whether they can match the performance of grammars based in logical forms
-\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
-in their ability to model core semantic phenomena like quantification, entailment, and contradiction.
+in their ability to model core semantic phenomena like quantification, entailment, and contradiction \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}.
 
+% TODO: Condense `skeptical' sentence
+
+Recent work on the algebraic approach of
+\newcite{ClarkCoeckeSadrzadeh2011} has yielded rich frameworks 
+for computing the meanings of complex fragments of natural
+language compositionally from vector or tensor representations, but 
+has not yet yielded effective methods for learning these representations
+entirely from data in a typical machine learning settings. 
 To date, experimental investigations of these concepts using
@@
 model these phenomena in their full generality on complex linguistic
-structures. Recent work on the algebraic approach of
-\newcite{ClarkCoeckeSadrzadeh2011} has yielded rich frameworks 
-for computing the meanings of complex fragments of natural
-language compositionally from vector or tensor representations, but 
-has not yet yielded effective methods for learning these representations
-entirely from data in a typical machine learning setting.
+structures. 
+% TODO: Note about how their reps are learned instead.
 
@@
 need of intermediate representations in an artificial logical
-language. We use the natural logic developed by
+language. We use the powerful
+ natural logic developed by
 \newcite{maccartney2009extended} as our formal model. This logic
@@
 \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
+
+% TODO: Cut the below para. Replace with a more high-level summary of how we are paralleling NatLog, and how compositionality fits with that (Q: Does it?).
 
@@
  is somewhat poor in our first experiment, we find that the stronger RNTN model
-  generalizes well in every case, suggesting that it has learned to simulate our target logical concepts.
+  generalizes well in every case, suggesting that it has learned to simulate our target logical concepts. % TODO: Split into two sentences
 
@@
 the SICK entailment challenge data. The small size of this corpus puts
-data-intensive neural models like ours at a disadvantage, but we are
+data-hungry NN models like ours at a disadvantage, but we are
 nonetheless able to achieve competitive performance on it,
-surpassing several models with significant hand-engineered task-specific features.
+surpassing several submitted models with significant hand-engineered task-specific features.
 This suggests that the representational abilities that we observe in the previous sections
@@
 any clear level of inferential complexity for which other approaches work and NN models fail.
+
+% TODO: Contributions para.
 
@@
   \end{tabular}
-  \caption{\label{b-table}The seven natural logic relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}. 
-    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies where none of the other six do.} %, including when there 
+  \caption{\label{b-table}The seven natural logic relations of MacCartney and Manning \protect\shortcite{maccartney2009extended} are defined abstractly on pairs of sets drawing from the universe  $\mathcal{D}$, and can be applied to any pair of words, phrases, or sentences according to standard approaches to semantic interpretation.
+% TODO: Condense, leave in reference to denotation.
+The relation $\natind$ applies where none of the other six do.} %, including when there 
     %is insufficient knowledge to choose a label.}

### 2015-02-24 writing/F'14 paper/join.tex
Experimental setup tweaks; add TODOs to paper

@@
 
-If a model is to learn the behavior of a relational logic like the one
-presented here from a finite amount of data, it must learn to deduce new
+ % TODO: Why is it okay to study non-recursive models here.
+ % TODO: Run a baseline for shits and giggles?
+ % TODO: Mention that we do an easier version of this implicitly elsewhere.
+
+If a model is to learn the behavior of a relational logic like ours
+from a finite amount of data, it must learn to deduce new
 relations from seen relations in a sound manner. The simplest such
 deductions involve atomic statements using the relations in
-Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
-\natrev c$, one can conclude that $a \natrev c$, by basic
+Table~\ref{b-table}. 
+For instance, from the relation $p_1 \natrev p_2$ between two propositions, 
+one can conclude that $p_2 \natfor p_1$ using the definitions of the relations directly. 
+If one is also given the relation $p_1 \natrev p_3$, one can conclude that $p_1 \natrev p_2$, by basic
 set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
-full set of sound such inferences is summarized in
+$p_1 \natfor p_2$ and $p_2 \natneg p_3$, it follows that $p_1 \natalt p_3$.  The
+full set of sound such inferences on pairs of premise relations is depicted in
 Table~\ref{tab:jointable}.
@@
   \end{array}$
-  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.
+  \caption{Inference path from premises $p_1Rp_2$ (row) and $p_2Sp_3$ (column) to the relation that holds between $p_1$ and $p_3$, if any.
 These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
@@
 Cells containing a dot correspond to pairs
-of relations for which no valid inference can be drawn.}
+of relations for which no valid inference can be drawn.} % TODO: Move harder example in to replace caption.
   \label{tab:jointable}
@@
 2960 relations.
+% TODO: Explain what the lattice shows.
 
-We trained models with both the RNN and RNTN comparison functions on these
+We trained models with both the NN and NTN comparison functions on these
 data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no strictly simpler model is likely viable as a baseline.} In both cases, the models are implemented as
@@
     \begin{picture}(5,5.5)
-      \labelednode{2.50}{5}{}{1,2,3}
+      \labelednode{2.50}{5}{}{a,b,c}
 
@@
 
-      \labelednode{0.75}{3.5}{a,b}{1,2}
-      \labelednode{2.50}{3.5}{c}{1,3}
-      \labelednode{4.25}{3.5}{d}{2,3}
+      \labelednode{0.75}{3.5}{p_1,p_2}{a,b}
+      \labelednode{2.50}{3.5}{p_3}{a,c}
+      \labelednode{4.25}{3.5}{p_4}{b,c}
 
@@
 
-      \labelednode{0.75}{2}{e,f}{1}
-      \labelednode{2.50}{2}{}{2}
-      \labelednode{4.25}{2}{g,h}{3}
+      \labelednode{0.75}{2}{p_5,p_6}{a}
+      \labelednode{2.50}{2}{}{b}
+      \labelednode{4.25}{2}{p_7,p_8}{c}
 
@@
       \midrule
-
-      $a \nateq b$              & $b \natneg g$ \\
-      $a \natrev e$              & $b \natrev e$ \\
-      $d \natrev h$              & \strikeout{$e \nateq f$} \\
-       $e \natalt g$              & \strikeout{$g \natfor d$} \\
-      $g \natneg a$           & $h \natfor d$ \\
+      $p_1 \nateq p_2$              & $p_2 \natneg p_7$ \\
+      $p_1 \natrev p_5$              & $p_2 \natrev p_5$ \\
+      $p_4 \natrev p_8$              & \strikeout{$p_5 \nateq p_6$} \\
+      $p_5 \natalt p_7$              & \strikeout{$p_7 \natfor p_4$} \\
+      $p_7 \natneg p_1$           & $p_8 \natfor p_4$ \\
 
@@
 
-% Note: None of these test examples is derivable from the shown training data, 
-% but we suggest that there is additional training data, so we can cross lines
-% out willy-nilly without fear.
 \begin{table}[tp]
@@
 further hyperparameter tuning could lead an NN model to succeed here.
+
+In one run for example, both models successfully classified the test example $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$.
+On a somewhat similar example involving comparably frequent relations, the RNTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the RNN incorrectly assigned it $\natind$.
+
+% From train\test_1

### 2015-02-24 writing/F'14 paper/methods.tex
Experimental setup tweaks; add TODOs to paper

@@
 
-We study neural models that adhere to the linguistic \ii{principle of
+We limit the scope of our experiments in this paper neural network models 
+that adhere to the linguistic \ii{principle of
  compositionality}, which says that the meanings for complex
@@
 derived, it serves as a fixed-dimensional input for some subsequent function.
+
+% TODO: Say that there are two RNTNs in caption.
 
@@
 distribution over the seven relations represented in Table~\ref{b-table}.
+% TODO: Move 75 constant later.
 
@@
 To run the model forward and label a pair of phrases, the structure of
-the lower layers of the network is assembled so as to mirror the tree
-structures provided for each phrase. The word vectors are then looked
-up from the vocabulary matrix $V$ (one of the model parameters), and
+the lower layers of the network is assembled so as to match the tree
+structures provided for each phrase, which is either included in the source data or given by a parser.
+The word vectors are then looked up from the vocabulary matrix $V$ (one of the model parameters), and
 the composition and comparison functions are used to pass information
@@
 function, we use the negative log of the probability assigned to the
-correct label, plus L2 regularization.
+correct label with L2 regularization.
 
@@
 with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
-Training times on CPUs vary from hours to days across experiments, and
-the hyperparameters of AdaDelta, the dimensions of the representations,
-and the initialization strategy are all tuned manually.
+The hyperparameters of AdaDelta, the dimensions of the representations,
+and the initialization strategy are all tuned manually. Training times on CPUs vary from hours to days across experiments.
 Because the classes are not balanced in the artificial data experiments, we report performance

### 2015-02-24 writing/F'14 paper/quantifiers.tex
Paper update
Experimental setup tweaks; add TODOs to paper

@@
 
-We have seen that recursive NNs can learn an approximation of propositional
+We have seen that recursive models can learn an approximation of propositional
 logic.  However, natural languages can express functional meanings of
-considerably greater complexity than this.  As a step towards
-investigating whether our models can capture this complexity, we now
-directly measure the degree to which they are able to
+considerably greater complexity than this.  As a key test of whether our models 
+can capture this complexity, we now directly measure the degree to which they are able to
 develop suitable representations for the semantics of natural language
 quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantification 
-and negation are far from
-the only place in natural language where complex functional meanings
-are found, but they are natural starting point, since they can be tested
-in sentences whose structures are otherwise quite simple, and since they have
+and negation are far from the only place in natural language where complex functional meanings
+are found, but they are natural focus, since they have
 formed a standard case study in prior formal work on natural
-language inference \cite{Icard:Moss:2013:LILT}.
+language inference \cite{Icard:Moss:2013:LILT}..
 
@@
 
+In addition to the two models under study, we also train a summing neural network as a baseline, 
+which is identical to the RNN, except that instead of learning a parametric composition function,
+it simply sums the word vectors in a sentence to compose them. This baseline model does not
+use word order information, though the highly consistent sentence structure in this experiment means
+that it can potentially distinguish any two valid sentences except in case the two differ only in the placement of the word \ii{not}, potentially introducing errors into the 58.8\% of examples that contain a sentence with a single
+instance of \ii{not}.
+
 \begin{table}[tp]
-  \centering \small
+  \centering
   \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
     \toprule
-    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{25d RNN}  & \multicolumn{2}{c}{25d RNTN}\\
+    ~ & \multicolumn{2}{c}{25d SumNN} & \multicolumn{2}{c}{25d RNN}  & \multicolumn{2}{c}{25d RNTN}\\
     \midrule
-    Train & 35.4 &(7.5) &	\textbf{100}&		\textbf{(100)}&	99.5&	(99.3)\\	
-    Test & 35.4 &(7.5) &	\textbf{99.4}&\textbf{(99.5)}& \textbf{99.4} & (99.2)\\
+    Train &	96...&	(95...)&	\textbf{100}&		\textbf{(100)}&	99.5&	(99.3)\\	
+    Test & 90...&	(80...)&	\textbf{99.4}&\textbf{(99.5)}& \textbf{99.4} & (99.2)\\
     \bottomrule
   \end{tabular}
-  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}
+  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1. The most frequent class, $\natind$, appears in 35.4\% of examples.}
   \label{qresultstable}
 \end{table} 
+% TODO: Fill in baseline results
+
 %
@@
 %
-\paragraph{Results} The results, reported in Table~\ref{qresultstable}, show that both models are able to learn to generalize the underlying logic satisfactorily.
+\paragraph{Results} The results, reported in Table~\ref{qresultstable}, show that both recursive models are able to learn to generalize the underlying logic satisfactorily, and that the baseline model generalizes well but far from perfectly.
 We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

### 2015-02-24 writing/F'14 paper/recursion.tex
Experimental setup tweaks; add TODOs to paper

@@
 
-Complex embedding is a prominent feature of natural languages. 
-The principle of compositionality explains how the meanings of 
-such structures are derived systematically from their parts,
- enabling people to use novel phrases and sentences.
-Theoretical accounts of natural language syntax and semantics
-in turn rely heavily on recursive structures. This section examines the 
+% TODO: Check on citation for clark/coecke/sadr 2011
+
+% Run baseline anyway?
+
+% TODO: Clean up next sent.
+Recursive compositional structure is a prominent feature of natural language. 
+This section examines the 
 question of whether our models are able to learn a 
@@
       \midrule
-      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
+      $p_1$, $p_2$, $p_3$, $p_4$, $p_5$, $p_6$ & $\sem{x} \in \{\True, \False\}$ \\
       $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
@@
       \toprule
-      $\plneg a$        & $\natneg$ & $a$ \\
-      $\plneg \plneg a$ & $\nateq$  & $a$ \\
-      $a$               & $\natfor$ & $(a \plor b)$ \\
-      $(a \plor (b \plor c))$               & $\natrev$ & $(b \pland  \plneg c)$ \\
+      $\plneg p_3$        & $\natneg$ & $p_3$ \\
+      $\plneg \plneg p_6$ & $\nateq$  & $p_6$ \\
+      $p_3$               & $\natfor$ & $(p_3 \plor p_2)$ \\
+      $(p_1 \plor (p_2 \plor p_4))$               & $\natrev$ & $(p_2 \pland  \plneg p_4)$ \\
       %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
-      $\plneg\, (\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
+      $\plneg\, (\plneg p_1 \pland \plneg p_2)$ & $\nateq$ & $(p_1 \plor p_2)$ \\ 
       \bottomrule
@@
 each operator can be represented accurately by a single bit.
-In the more demainding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.
+In the more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.
 Relational expressions between these conditions are essentially theorems of propositional logic, and to succeed, the model must learn to approximate the behavior of a theorem prover.
+% TODO: Condense this para, and move bulk to footnote.
 
@@
 training examples with more than 4 connectives, yielding 60k short training examples,
-and 21k training examples containing up to 12 connectives.
+and 21k test examples containing up to 12 connectives.
 We evaluated both the RNN and RNTN\footnote{We did not run any strictly simpler baselines for this experiment because it is demonstrably impossible to perform substantially above chance on longer expressions without word order information.} on the resulting data, initializing the model parameters randomly,
@@
 
-\paragraph{Results} Figure~\ref{prop-results} shows test the relationship
+\paragraph{Results} Figure~\ref{prop-results} shows the relationship
 between test accuracy and statement length. We found that both 
@@
 increasing expression size, resulting in gradually dropping performance. 
+Though imperfect, model performance on the long examples is striking. 
+The RNTN classifies true examples of every class but $\nateq$ 
+(which occurs only once) correctly the majority of the time, where a 
+typical statement might be $( ( p_1  \pland ( p_2 \plor ( \plneg p_4 ) ) ) \plor p_3 ) \natcov ( \plneg ( ( \plneg ( ( \plneg ( p_4 \plor p_3 ) )  \plor ( ( \plneg p_3 ) \plor ( \plneg p_3 ) ) ) ) \pland ( \plneg ( ( \plneg ( ( \plneg p_3 ) \plor ( \plneg p_4 ) ) ) \plor ( \plneg p_2 ) ) ) ) )$.
+% TODO: Better present examples.
+
 It appears that the RNTN model overfit somewhat relative to the plain RNN,

### 2015-02-24 writing/F'14 paper/sick.tex
Paper update
Experimental setup tweaks; add TODOs to paper

@@
 
-The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data are not alone sufficient to address this issue.
+\mynote{NOTE: We might delete this section if we can't improve the numbers in time.}
+
+% TODO: Use ``model not guaranteed to do anything'' rhetoric earlier.
+The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and the diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data alone are not sufficient to address this issue.
 To investigate the ability of the model to learn on natural data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}.
@@
 
-Adapting to this task required us to make a few additions to the techniques discussed in Section \ref{methods}.
+Adapting to this task required us to make a few additions to the techniques discussed in \S\ref{methods}.
 In order to better handle rare words, we initialized our word embeddings using 200 dimensional vectors trained with
@@
     \toprule
-    ~&\multicolumn{1}{c}{\ii{neut.} only} & \multicolumn{1}{c}{30d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
+    ~&\multicolumn{1}{c}{25d SumNN} & \multicolumn{1}{c}{30d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
     \midrule
-    SICK Train &  56.7 & 95.4 &  \textbf{97.8}  \\
-    DG (Train) &  50.0 & 67.0 & \textbf{74.0}  \\
-    SICK Test & 56.7 & 74.9 & \textbf{76.9}  \\
+    SICK Train &  TODO & 95.4 &  \textbf{97.8}  \\
+    DG (Train) &  TODO & 67.0 & \textbf{74.0}  \\
+    SICK Test & TODO & 74.9 & \textbf{76.9}  \\
     \bottomrule
   \end{tabular}
-  \caption{Performance on collapsed SICK data.}
+  \caption{Performance on collapsed SICK data. The class NEUTRAL appears in 56.7\% of SICK examples and 50\% of DG examples.}
   \label{sresultstable}
@@
 
+% TODO: Better error analysis
+
 Table \ref{examplesofsickdata} shows a few typical examples of correct classifications.

### 2015-02-24 InitializeModelCompat.m

@@ -21,7 +21,7 @@ SCALE = 0.05;
 TSCALE = hyperParams.tensorScale * SCALE;
 
 % Randomly initialize softmax layer
-classifierParameters = [zeros(sum(hyperParams.numRelations) + SCALE, 1), ...
+classifierParameters = [zeros(sum(hyperParams.numRelations), 1), ...
                         rand(sum(hyperParams.numRelations), PENULT) .* (2 * SCALE) - SCALE];
 
 % Randomly initialize tensor parameters

### 2015-02-25 writing/F'14 paper/abstract.tex
Experimental setup tweaks; add TODOs to paper
Paper presentation overhaul.

@@
 \begin{abstract} 
-  Recursive neural networks (RNNs) for sentence meaning have been
+  Tree-structured recursive neural networks (TreeRNNs) for sentence meaning have been
   successful for many tasks, but it remains an open question whether
   the fixed-length representations that they learn can support tasks as demanding as logical deduction.
-  We pursue this question by evaluating whether two such models---plain RNNs and recursive neural
-  tensor networks (RNTNs)---can correctly learn to identify logical
+  We pursue this question by evaluating whether two such models---plain TreeRNNs and recursive neural
+  tensor networks (TreeRNTNs)---can correctly learn to identify logical
   relationships such as entailment and contradiction using these representations. 
@@
   inference in natural language.
+
+  % TODO: Rewrite latter part, w/ or w/o SICK.
 \end{abstract}

### 2015-02-25 writing/F'14 paper/discussion.tex
Experimental setup tweaks; add TODOs to paper
Paper presentation overhaul.

@@
 involving quantification and negation. We then show that the same models can learn to
-perform an entailment task on natural language image captions. The results suggest that RNTNs,
-and potentially also RNNs, have the capacity to learn to model these tasks with 
+perform an entailment task on natural language image captions. The results suggest that TreeRNTNs,
+and potentially also TreeRNNs, have the capacity to learn to model these tasks with 
 reasonably-sized training sets. These positive results are
@@
 
-Of course, challenges remain. Even
-the RNTN falls short of perfection in the recursion experiment, with
+Some questions about the abilities of these models remain open. Even
+the TreeRNTN falls short of perfection in the recursion experiment, with
 performance falling off steadily as the size of the expressions grows. It
 remains to be seen whether these deficiencies can be overcome with
-stronger models or optimization procedures. The space of possible stronger
-models is sizable, including extensions of the recursive models used here
-\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, recurrent
-models \cite{sutskever2014sequence}, and even learning-enabled variants 
-of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
-In addition, interesting analytical questions remain about \ii{how} these models encode
+stronger models or learning techniques. In addition, interesting 
+analytical questions remain about \ii{how} these models encode
 the underlying logics. Neither the underlying
@@
 
-Our artificial data experiments have only scratched the surface 
-of the logical complexity of natural language; in future experiments, we hope to test sentences
-with more complex types of structure, as well as run similar experiments on non-tree structured models, such as LSTMs. Our SICK experiments 
-similarly only begin to reveal the potential of these models to learn to 
-perform complex semantic inferences from corpora. Nonetheless, the
-rapid progress the field has made with these models in recent years
-provides ample reason to be optimistic that they can be trained to
+Our SICK experiments similarly only begin to reveal the potential of these models to learn to 
+perform complex semantic inferences from corpora, and there is ample room to develop our understanding
+using new sources of natural language data. Nonetheless, the rapid progress the field 
+has made with these models in recent years provides ample reason to be optimistic that 
+learned representation models can be trained to
 meet all the challenges of natural language semantics.
-
-% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators.
-This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic.
-The positive results that we have found, however, are extremely promising for the future of learned representation models in the applied modeling of meaning.
-We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators.
-We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.
-
-% There is ample room to build on these results.
-In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses.
-In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (-), or convolutional RNNs \cite{kalchbrenner2014convolutional}.

### 2015-02-25 writing/F'14 paper/intro.tex
Paper presentation overhaul.

@@
 
-% TODO: TreeRNN.
-% TODO: More language about the representations being the object of study.
 % TODO: Is there any way to squeeze a useful example into the first page?
-% TODO: Clarify early that we study vector pair classification. Maybe move up fig. 1?
-% TODO: Clarify that we aren't implementing NatLog rules.
-% TODO: ``Something that does that'' epigram in intro.
-% TODO: Mention tree pair archichture.
+% TODO: Can we fit a version of the ``find something that does that'' quote in intro somewhere?
 
-Supervised recursive neural network models (RNNs; \citealt{goller1996learning}) for sentence meaning
+% TODO: This para might be a bit repetitive.
+Tree-structured recursive neural network models (TreeRNNs; \citealt{goller1996learning}) for sentence meaning
 have been successful in an array of sophisticated language tasks,
@@
 \cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
-the ability of these models to learn compositional semantic grammars, though many remain skeptical of the extent to which vector space representations can model complex compositional semantics, and it remains an open question whether learned can achieve the
+the ability of these models to learn to produce and use strong 
+semantic representations for sentences,
+though many remain skeptical of the extent to which they can learn to represent
+the necessary complex compositional structures using fixed-size vectors. 
+In particular, it remains an open question whether learned models can achieve the
 high-fidelity distributed representations of recent algebraic
@@
 in their ability to model core semantic phenomena like quantification, entailment, and contradiction \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}.
-
-% TODO: Condense `skeptical' sentence
 
@@
 has not yet yielded effective methods for learning these representations
-entirely from data in a typical machine learning settings. 
+entirely from data in typical machine learning settings. 
 To date, experimental investigations of these concepts using
@@
 structures.
-% TODO: Note about how their reps are learned instead.
 
-Our research uses the task of \ii{natural language inference} 
+This paper describes four machine learning experiments
+that directly evaluate the abilities of these models to learn representations that 
+support specific semantic behaviors. These tasks follow the format of
+ \ii{natural language inference} 
 (also known as \ii{recognizing textual entailment};
 \citealt{dagan2006pascal}), in which the goal is to determine the core
-inferential relationship between two sentences. Much of the
-theoretical work on this task (and some successful implemented models;
+inferential relationship between two sentences. 
+We introduce a novel NN architecture for natural language inference which
+independently computes vector representations for each of two sentences using
+standard TreeRNN or TreeRNTN \cite{socher2013acl1} models, and
+produces a judgment for the pair using only those representations. This allows
+us to gauge the abilities of these two models to represent all of the necessary semantic
+information in the sentence vectors.
+
+Much of the
+theoretical work on natural language inference (and some successful implemented models;
 \citealt{maccartney2009extended,watanabe2012latent}) involves \ii{natural
@@
 need of intermediate representations in an artificial logical
-language. We use the powerful
- natural logic developed by
-\newcite{maccartney2009extended} as our formal model. This logic
+language. This logic
 defines seven mutually-exclusive relations of synonymy, entailment, contradiction,
@@
 provides rules of semantic combination for projecting these relations
-from the lexicon up to complex phrases. The formal properties of this system are now well-understood
-\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
+from the lexicon up to complex phrases. The formal properties of this system 
+are now well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
 
-% TODO: Cut the below para. Replace with a more high-level summary of how we are paralleling NatLog, and how compositionality fits with that (Q: Does it?).
-
-In our first set of experiments, we use this pre-specified logical
-grammar to generate controlled data sets encoding the semantic
-relationships between pairs of expressions and evaluate whether each
-of two classes of neural model---plain recursive NNs and recursive neural
-tensor networks (RNTNs; \citealt{socher2013acl1})---can learn those
-relationships correctly. In our first experiment
-(\S\ref{sec:join}), we evaluate the ability of these models to
-learn the core relational algebra of natural logic from data. Our
-second experiment (\S\ref{sec:recursion}) extends this
-evaluation to relations between complex recursive structures like
-$(a \plor b)$ and $\plneg (\plneg a \pland \plneg b)$, and our third
-experiment (\S\ref{sec:quantifiers}) covers relations between
-statements with quantification and negation like \ii{every reptile walks} and \ii{no turtle
- moves}. Though the performance of the plain RNN model 
- is somewhat poor in our first experiment, we find that the stronger RNTN model
-  generalizes well in every case, suggesting that it has learned to simulate our target logical concepts. % TODO: Split into two sentences
+While one of our experiments uses natural logic only as an approximate
+description of the desired model behavior, three experiments draw on this
+ tradition much more directly, and
+test models' abilities to learn to reproduce the precise behavior of the natural logic developed by
+\newcite{maccartney2009extended}: one
+covers learning to reason with the bare logical relations (\S\ref{sec:join}), a second covers reasoning with statements constructed compositionally from recursive functions (\S\ref{sec:recursion}),
+and a third covers the additional complexity that results from quantification (\S\ref{sec:quantifiers}).
+Though the performance of the plain TreeRNN model 
+ is somewhat poor in our first experiment, we find that the stronger TreeRNTN model
+  generalizes well in every case, suggesting that it has learned to simulate our target logical concepts.
 
@@
 naturalistic text. To address this question, we apply our models to
-the SICK entailment challenge data. The small size of this corpus puts
+the SICK entailment challenge data in \S\ref{sec:sick}. The small size of this corpus puts
 data-hungry NN models like ours at a disadvantage, but we are
@@
 This suggests that the representational abilities that we observe in the previous sections
-are not limited to carefully circumscribed tasks. We conclude that RNTN models
+are not limited to carefully circumscribed tasks. We conclude that TreeRNTN models
 are adequate for typical cases of natural logic inference, and that there is not yet

### 2015-02-25 writing/F'14 paper/join.tex
Paper presentation overhaul.

@@
-\section{Reasoning about semantic relations}\label{sec:join}
-
- % TODO: Why is it okay to study non-recursive models here.
- % TODO: Run a baseline for shits and giggles?
- % TODO: Mention that we do an easier version of this implicitly elsewhere.
-
-If a model is to learn the behavior of a relational logic like ours
-from a finite amount of data, it must learn to deduce new
-relations from seen relations in a sound manner. The simplest such
-deductions involve atomic statements using the relations in
-Table~\ref{b-table}. 
-For instance, from the relation $p_1 \natrev p_2$ between two propositions, 
-one can conclude that $p_2 \natfor p_1$ using the definitions of the relations directly. 
-If one is also given the relation $p_1 \natrev p_3$, one can conclude that $p_1 \natrev p_2$, by basic
-set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$p_1 \natfor p_2$ and $p_2 \natneg p_3$, it follows that $p_1 \natalt p_3$.  The
-full set of sound such inferences on pairs of premise relations is depicted in
-Table~\ref{tab:jointable}.
-
-% about the relations themselves that do not depend on the
-% internal structure of the things being compared. For example, given
-% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
-% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven relations support more
-% than just transitivity: MacCartney and Manning's
-% \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of relations of the form $a R
-% b$ and $b R' c$, including several less intuitive ones such as that if
-% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-
-\begin{table}[htp]
+\begin{table}[tp]
   \centering  \small
@@
   \end{array}$
-  \caption{Inference path from premises $p_1Rp_2$ (row) and $p_2Sp_3$ (column) to the relation that holds between $p_1$ and $p_3$, if any.
-These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
-We assess our models' ability to reproduce such inferential paths.
-Cells containing a dot correspond to pairs
-of relations for which no valid inference can be drawn.} % TODO: Move harder example in to replace caption.
+  \caption{In \S\ref{sec:join}, we assess our models' ability to learn to do inference over pairs of relations using the rules represented here.
+As an example, given that $p_1 \natfor p_2$ and $p_2 \natneg p_3$, the entry in the $\natfor$ row and the $\natneg$ column lets us conclude that $p_1 \natalt p_3$.
+These rules are derived from the definitions of the relations in Table~\ref{b-table}.  Cells containing a dot correspond to pairs
+of relations for which no valid inference can be drawn.} 
   \label{tab:jointable}
@@
 
+\section{Reasoning about semantic relations}\label{sec:join}
+
+The simplest kinds of deduction in natural logic involve atomic statements 
+using the relations in Table~\ref{b-table}. 
+For instance, from the relation $p_1 \natrev p_2$ between two propositions, 
+one can conclude that $p_2 \natfor p_1$ using the definitions of the relations directly. 
+If one is also given the relation $p_1 \natrev p_3$, one can conclude that $p_1 \natrev p_2$, by basic set-theoretic reasoning (transitivity of $\natrev$). The
+full set of sound such inferences on pairs of premise relations is depicted in
+Table~\ref{tab:jointable}. Though these basic inferences do not involve compositional
+sentence representations, any successful reasoning using compositional representations
+will rely on the ability to perform sound inferences of this kind, so our first experiment studies how well each model can learn to perform them them in isolation.
+
+% about the relations themselves that do not depend on the
+% internal structure of the things being compared. For example, given
+% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
+% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
+% understanding $a$, $b$, or $c$. These seven relations support more
+% than just transitivity: MacCartney and Manning's
+% \cite{maccartney2009extended} join table defines 32 valid inferences
+% that can be made on the basis of pairs of relations of the form $a R
+% b$ and $b R' c$, including several less intuitive ones such as that if
+% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
+
 \paragraph{Experiments}
-To generate data for these inferential patterns, we
-create small boolean structures for our logic in which terms denote
-sets of entities from a small domain.  Figure~\ref{lattice-figure}
-depicts a structure of this form. The lattice gives the full model,
-and we generate a relational statement for each pair of labels in the model. 
+We begin by creating a world model
+on which we will base the statements in the train and test sets.
+This takes the form of a small boolean structure in which terms denote
+sets of entities from a small domain.  Fig.~\ref{lattice-figure}a
+depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight terms ($p_{1\text{-}8}$). We then generate a 
+relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
 We divide these statements evenly into training and test sets, and remove from the
@@
 statements.
-In each experimental run, we create 80 randomly generated sets drawing from
-a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
+In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
 2960 relations.
-% TODO: Explain what the lattice shows.
 
 We trained models with both the NN and NTN comparison functions on these
-data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no strictly simpler model is likely viable as a baseline.} In both cases, the models are implemented as
-described in Section~\ref{methods}, but since the items being compared
+data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no simpler version of our model is a viable baseline.} In both cases, the models are implemented as
+described in \S\ref{methods}, but since the items being compared
 are single terms rather than full tree structures, the composition
@@
     \end{picture}}
-    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
+    \caption{Example boolean structure. The terms $p_{1\text{-}8}$ name the sets. Not all sets have names, and
     some sets have multiple names, so that learning $\nateq$ is non-trivial.}
   \end{subfigure}
-  \qquad
+  \qquad\small
     \begin{subfigure}[t]{0.43\textwidth}
@@
 \begin{table}[tp]
-  \centering \small
-  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
+  \centering\small
+  \begin{tabular}{ l r@{ \ }r r@{ \ }r }
     \toprule
-    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{15d RNN}  & \multicolumn{2}{c}{15d RNTN}\\
+    ~&\multicolumn{2}{c}{Train} & \multicolumn{2}{c}{Test}\\
     \midrule
-    Train &53.8 &(10.5)    & 99.8&	(99.0) & \textbf{100} & \textbf{(100)}\\
-    Test &52.0 &(10.3) &	94.0&(87.0)& \textbf{99.6} & \textbf{(95.5)}\\
+    $\natind$ only &53.8 &(10.5)    &53.8 &(10.5) \\
+    15d NN &				99.8&	(99.0) &94.0&(87.0) \\
+    15d NTN 				& \textbf{100} & \textbf{(100)} & \textbf{99.6} & \textbf{(95.5)}\\
     \bottomrule
   \end{tabular}
+
   \caption{Performance on the semantic relation experiments.
@@
 
-In one run for example, both models successfully classified the test example $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$.
-On a somewhat similar example involving comparably frequent relations, the RNTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the RNN incorrectly assigned it $\natind$.
+As an example, both models successfully classified the test item $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$.
+On another example involving comparably frequent relations, the NTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the NN incorrectly assigned it $\natind$.
 

### 2015-02-25 writing/F'14 paper/methods.tex
Paper presentation overhaul.

@@
 
-We limit the scope of our experiments in this paper neural network models 
+We limit the scope of our experiments in this paper to neural network models 
 that adhere to the linguistic \ii{principle of
@@
 via specific composition functions \cite{Partee84,Janssen97}. In our
-distributed setting, word meanings are vectors of dimension $n$. The
-composition function maps pairs of them to single vectors of dimension $n$, 
+distributed setting, word meanings are embedding vectors of dimension $n$. The
+composition function maps pairs of them to single phrase vectors of dimension $n$, 
 which can then be merged again to represent more complex
@@
 
-% TODO: Say that there are two RNTNs in caption.
-
 \begin{figure}[tp]
   \centering
- % \includegraphics[scale=0.35]{model.eps}
   \input{figure1}
-  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
-    The same structure is used for both the RNN and RNTN layer functions.} 
+  \caption{In our model, two separate tree-structured networks build up vector representations for each of two sentences using either NN or NTN layer functions.
+A comparison layer then uses the resulting vectors to produce features for a classifier.}
   \label{sample-figure}
@@
 To apply these recursive models to our task, we propose the tree 
-pair model architecture depicted in Figure~\ref{sample-figure}. 
+pair model architecture depicted in Fig.~\ref{sample-figure}. 
 In it, the two phrases being compared are processed separately using a pair 
-of recursive networks that share a single set of parameters. 
+of tree-structured networks that share a single set of parameters. 
 The resulting vectors are fed into a separate comparison
-layer that is meant to generate a feature vector (fixed at 75 dimensions) capturing the
+layer that is meant to generate a feature vector capturing the
 relation between the two phrases. The output of this layer is then
-given to a softmax classifier, which in turn produces a hypothesized
+given to a softmax classifier, which produces a
 distribution over the seven relations represented in Table~\ref{b-table}.
-% TODO: Move 75 constant later.
 
-For a composition layer, we evaluate models with both the plain neural
-network layer function \eqref{rnn} and the more powerful RNTN layer function
-\eqref{rntn} proposed in \newcite{chen2013learning}. The nonlinearity $f(x) = \tanh(x)$ is applied to the output of the layer function.
+For the sentence embedding portions of the network, we evaluate both TreeRNN models with
+ the standard NN layer function \eqref{TreeRNN} and those with the more powerful neural tensor 
+ network layer function
+\eqref{TreeRNTN} proposed in \newcite{chen2013learning}. The nonlinearity $f(x) = \tanh(x)$ is applied
+ elementwise to the output of either layer function.
 %
 \begin{gather} 
-\label{rnn}
+\label{TreeRNN}
 \vec{y}_{\textit{TreeRNN}} = f(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
-\label{rntn} 
+\label{TreeRNTN} 
 \vec{y}_{\textit{TreeRNTN}} = \vec{y}_{\textit{TreeRNN}} + f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
@@
 representations for the left and right children of the node, and
-$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
-them by an $n \times 2n$ matrix of learned weights, adds a bias $\vec{b}$, and applies the
-non-linearity element-wise to the resulting vector. The RNTN adds a learned third-order tensor 
+$\vec{y}$ is the node's output.  The TreeRNN concatenates them, multiplies
+them by an $n \times 2n$ matrix of learned weights, and adds a bias $\vec{b}$. 
+The TreeRNTN adds a learned third-order tensor 
 $\mathbf{T}$, of dimension $n \times n \times n$, modeling
 multiplicative interactions between the child vectors. 
+Other tree-structured networks have been successfully used in past work
+\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, but
+we believe that these two provide a valuable case study, and that positive results on 
+here are likely to generalize well to stronger models.
 
@@
 independently learned parameters and a separate nonlinearity function.
-Rather than use a $\tanh$ nonlinearity here, we found better results with a rectified linear function. In
-particular, we use the leaky rectified linear function
+Rather than use a $\tanh$ nonlinearity here, we found better results with the leaky rectified linear function
 \cite{maasrectifier}: $f(x)=\max(x, 0) +
@@
 structures provided for each phrase, which is either included in the source data or given by a parser.
-The word vectors are then looked up from the vocabulary matrix $V$ (one of the model parameters), and
+The word vectors are then looked up from the  vocabulary embedding matrix $V$ (one of the model parameters), and
 the composition and comparison functions are used to pass information
@@
 
-% Removed 'minibatch' -> For two of the experiments, we use minibatches
-% of size 1, which doesn't really count.
-We train the model using stochastic gradient descent (SGD)
+We initialize parameters uniformly, using the range $(-0.05, 0.05)$ for layer parameters
+and $(-0.01, 0.01)$ for embeddings, and train the model using stochastic gradient descent (SGD)
 with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
-The hyperparameters of AdaDelta, the dimensions of the representations,
-and the initialization strategy are all tuned manually. Training times on CPUs vary from hours to days across experiments.
-Because the classes are not balanced in the artificial data experiments, we report performance
+The classifier feature vector is fixed at 75 dimensions and 
+the dimensions of the recursive layers are tuned manually.
+Training times on CPUs vary from hours to days across experiments.
+On the first three experiments, which use artificial data, we report mean
+results over five fold cross-validation.
+Variance across runs is typically no
+more than two percentage points. In addition, because the classes are not necessarily balanced, 
+we report performance 
 using both accuracy and macroaveraged F1.\footnote{We compute macroaveraged F1 
@@
 for all classes for which there is test data, setting precision to 0 
-where it is not defined.} On the artificial data experiments, we report mean
-results over five fold cross-validation. Variance across runs is typically no
-more than two percentage points.
+where it is not defined.}
 Source code and generated data will be released after the review period.

### 2015-02-25 writing/F'14 paper/quantifiers.tex
Paper presentation overhaul.

@@
 \paragraph{Experiments}
-Our experimental data consist of pairs of sentences generated from a
-small artificial grammar. Each sentence contains a quantifier, a noun
+Our experimental data consist of pairs of sentences generated
+from a grammar for a simple English-like artificial language.
+Each sentence contains a quantifier, a noun
 which may be negated, and an intransitive verb which may be
@@
 each noun--noun pair, each verb--verb pair, and each
-quantifier--quantifier pair. The grammar accepts aligned pairs of
-sentences of this form and calculates the natural logic relationship
+quantifier--quantifier pair. The grammar generates pairs of
+sentences of this form and calculates the relations holding
 between them.  Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.
-  As in previous sections, the goal of
-learning is then to assign these relational labels accurately to
-unseen pairs of sentences.
 
@@
   \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
-\end{gather}\vspace{-0.8cm}
+\end{gather}
 
-In addition to the two models under study, we also train a summing neural network as a baseline, 
-which is identical to the RNN, except that instead of learning a parametric composition function,
-it simply sums the word vectors in a sentence to compose them. This baseline model does not
-use word order information, though the highly consistent sentence structure in this experiment means
-that it can potentially distinguish any two valid sentences except in case the two differ only in the placement of the word \ii{not}, potentially introducing errors into the 58.8\% of examples that contain a sentence with a single
+We use the same summing baseline as in the previous section.
+The highly consistent  sentence structure in this experiment means
+that that model can potentially distinguish two valid sentences unless they
+ differ only in the placement of the word \ii{not}, potentially introducing errors into 
+the 58.8\% of examples that contain a sentence with a single
 instance of \ii{not}.
@@
 \begin{table}[tp]
-  \centering
-  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
+  \centering\small
+    \begin{tabular}{ l r@{ \ }r r@{ \ }r }
+
     \toprule
-    ~ & \multicolumn{2}{c}{25d SumNN} & \multicolumn{2}{c}{25d RNN}  & \multicolumn{2}{c}{25d RNTN}\\
+    ~ & \multicolumn{2}{c}{Train} & \multicolumn{2}{c}{Test} \\
     \midrule
-    Train &	96...&	(95...)&	\textbf{100}&		\textbf{(100)}&	99.5&	(99.3)\\	
-    Test & 90...&	(80...)&	\textbf{99.4}&\textbf{(99.5)}& \textbf{99.4} & (99.2)\\
+    $\natind$ only &	35.4 & (7.5)	& 35.4	& (7.5)\\
+    25d SumNN	&	96...&	(95...)&	90...&	(80...)\\	
+    25d TreeRNN	&	\textbf{100}&	\textbf{(100)}&	99.5&	(99.3)\\
+    25d TreeRNTN	&	\textbf{99.4}&\textbf{(99.5)}&	\textbf{99.4} & (99.2)\\
     \bottomrule
   \end{tabular}
-  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1. The most frequent class, $\natind$, appears in 35.4\% of examples.}
+
+  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}
   \label{qresultstable}
@@
 %
-\paragraph{Results} The results, reported in Table~\ref{qresultstable}, show that both recursive models are able to learn to generalize the underlying logic satisfactorily, and that the baseline model generalizes well but far from perfectly.
+\paragraph{Results} The results, reported in Table~\ref{qresultstable}, show that both recursive models are able to learn to generalize the underlying logic satisfactorily, and that the baseline model generalizes well, but far from perfectly.
 We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

### 2015-02-25 writing/F'14 paper/recursion.tex
Paper presentation overhaul.

@@
 \section{Recursive structure}\label{sec:recursion}
-
-% TODO: Check on citation for clark/coecke/sadr 2011
-
-% Run baseline anyway?
+% TODO: Enter baseline results, use [[Tree]]RNN.
 
 % TODO: Clean up next sent.
-Recursive compositional structure is a prominent feature of natural language. 
-This section examines the 
-question of whether our models are able to learn a 
-compositional semantics over such recursive structures.
+Recursive structure is a prominent feature of natural language. 
+This section examines the degree to which our models are able to learn a 
+compositional semantics over recursive structures.
 In our evaluations, we exploit the fact that our logical language
@@
 \paragraph{Experiments}
-As in \S\ref{sec:join}, we test this phenomenon within the
-framework of natural logic, but we now replace the unanalyzed symbols
+As in \S\ref{sec:join}, we test this phenomenon on data generated from a logic,
+ but we now replace the unanalyzed symbols
 from that experiment with complex formulae. These formulae
@@
 operators are truth-functional ones.  Table~\ref{tab:pl} defines this
-logic, and Table~\ref{tab:plexs} gives some examples of relational
-statements that we can make in these terms. To compute these relations
+logic, and Table~\ref{tab:plexs} gives examples of the some of the shorter relational
+statements that we generate from it as train and test data. To compute these relations
 between statements, we exhaustively enumerate the sets of assignments
@@
 Table~\ref{b-table}. As a result, each relational statement represents
-a valid theorem of the propositional logic.
+a valid theorem of the propositional logic, and and to succeed, the models 
+must learn to reproduce the behavior of a theorem prover.\footnote{
+Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector TreeRNN
+model somewhat similar to our TreeRNTN can learn boolean logic, 
+a logic where the atomic symbols are simply the
+values $\True$ and $\False$. While learning the operators of that logic is not trivial, the outputs of
+each operator can be represented accurately by a single bit.
+In the more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.}
 
@@
 
-Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector RNN
-model somewhat similar to our RNTN can learn boolean logic, 
-a logic where the atomic symbols are simply the
-values $\True$ and $\False$. While learning the operators of that logic is not trivial, the outputs of
-each operator can be represented accurately by a single bit.
-In the more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.
-Relational expressions between these conditions are essentially theorems of propositional logic, and to succeed, the model must learn to approximate the behavior of a theorem prover.
-% TODO: Condense this para, and move bulk to footnote.
-
 In our experiments, we randomly generate unique pairs 
@@
 bin for a held-out test set.
+
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=5.75in]{decayfig.eps}
+%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
+  \caption{Model performance on propositional logic, by expression size.}
+%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
+%    representations. As measured by Euclidean distance, only the TreeRNTN
+%    achieves this for formulae containing more than a small number of
+%    connectives (\ii{and} in this example). The TreeRNN quickly collapses the representations,
+%    failing to capture the meaning contrasts.
+
+  \label{prop-results}
+
+\end{figure*}
 
@@
 and 21k test examples containing up to 12 connectives.
-We evaluated both the RNN and RNTN\footnote{We did not run any strictly simpler baselines for this experiment because it is demonstrably impossible to perform substantially above chance on longer expressions without word order information.} on the resulting data, initializing the model parameters randomly,
+We evaluated both the TreeRNN and TreeRNTN on the resulting data, initializing the model parameters randomly,
 including the vector representations of the six variables. 
+As a baseline, we also train a summing neural network
+which is largely identical to the TreeRNN, except that instead of learning a parametric composition function,
+it simply sums the term vectors in each expression to compose them. Unlike the two models under study,
+whose ability to accurately simulate the logic is an open question, this baseline does not use word order,
+and is as such guaranteed to ignore some information that it would need to succeed perfectly.
 
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=5.75in]{decayfig.eps}
-%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-  \caption{Model performance on propositional logic, by expression size.}
-%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
-%    representations. As measured by Euclidean distance, only the RNTN
-%    achieves this for formulae containing more than a small number of
-%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
-%    failing to capture the meaning contrasts.
-
-  \label{prop-results}
-
-\end{figure*}
-
-\paragraph{Results} Figure~\ref{prop-results} shows the relationship
-between test accuracy and statement length. We found that both 
+% TODO: Insert baseline result.
+\paragraph{Results} Fig.~\ref{prop-results} shows the relationship
+between test accuracy and statement length. BASELINE BASELINE
+ BASELINE BASELINE BASELINE BASELINE BASELINE BASELINE BASELINE   We found that both 
 models were able to perform well on unseen small test examples, 
-with RNN accuracy above
-98\% and RNTN accuracy above 99\% on formulae below length five.
+with TreeRNN accuracy above
+98\% and TreeRNTN accuracy above 99\% on formulae below length five.
 Starting at size four, performance gradually falls with increasing
-size. Training accuracy was 99.4\% for the RNN and 99.8\% for the RNTN.
+size. Training accuracy was 99.4\% for the TreeRNN and 99.8\% for the TreeRNTN.
 
@@
 increasing expression size, resulting in gradually dropping performance.
-Though imperfect, model performance on the long examples is striking. 
-The RNTN classifies true examples of every class but $\nateq$ 
-(which occurs only once) correctly the majority of the time, where a 
-typical statement might be $( ( p_1  \pland ( p_2 \plor ( \plneg p_4 ) ) ) \plor p_3 ) \natcov ( \plneg ( ( \plneg ( ( \plneg ( p_4 \plor p_3 ) )  \plor ( ( \plneg p_3 ) \plor ( \plneg p_3 ) ) ) ) \pland ( \plneg ( ( \plneg ( ( \plneg p_3 ) \plor ( \plneg p_4 ) ) ) \plor ( \plneg p_2 ) ) ) ) )$.
-% TODO: Better present examples.
 
-It appears that the RNTN model overfit somewhat relative to the plain RNN,
+It appears that the TreeRNTN model overfit somewhat relative to the plain TreeRNN,
 performing better on the training data and short examples, but worse on longer
 examples. Two factors are likely involved in this: even with the lower dimension,
-the RNTN composition function has about eight times as many parameters as the
-RNN, and thus more capacity to find a solution involving a large number of narrow 
+the TreeRNTN composition function has about eight times as many parameters as the
+TreeRNN, and thus more capacity to find a solution involving a large number of narrow 
 generalizations that work on the training data but don't generalize well. In addition,
-the RNTN used here has slightly weaker L2 regularization than the 
-RNN ($\lambda_{\textit{RNN}} = 0.001$, $\lambda_{\textit{RNTN}} = 0.0003$). This
-yielded better average performance, but may have facilitated this effect.
+the TreeRNTN used here has weaker L2 regularization than the 
+TreeRNN ($\lambda = 0.0003$ vs. $0.001$). This
+yielded better performance averaged across all bins, but may have facilitated this effect.
+
+Though imperfect, model performance on the most difficult set of expressions is nonetheless striking. 
+The weaker TreeRNTN classifies true examples of every class but $\nateq$ 
+(which is rare in long examples, and occurs only once here) correctly the majority of the time.
+% TODO: Better present examples.
 
@@
 %binary connectives, to assess how well they distinguish representations for semantically distinct
-%formulae. We found that only the RNTN is able to learn
+%formulae. We found that only the TreeRNTN is able to learn
 %substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
 % and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
-%The RNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed under three operators, and quickly approaches zero as depth increases.
-% The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}
+%The TreeRNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed under three operators, and quickly approaches zero as depth increases.
+% The TreeRNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}
 
@@
 % representations of the two sentences in each pair. Our findings are
-% shown in Figure~\ref{prop-falloff}, and show that while the RNTN can
-% distinguish the two sentences well at every size that we test, the RNN
+% shown in Figure~\ref{prop-falloff}, and show that while the TreeRNTN can
+% distinguish the two sentences well at every size that we test, the TreeRNN
 % fails after a depth of approximately six.

### 2015-02-25 writing/F'14 paper/sick.tex
Paper presentation overhaul.

@@
-\section{The SICK textual entailment challenge}
+\section{The SICK textual entailment challenge}\label{sec:sick}
 
-\mynote{NOTE: We might delete this section if we can't improve the numbers in time.}
-
-% TODO: Use ``model not guaranteed to do anything'' rhetoric earlier.
 The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and the diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data alone are not sufficient to address this issue.
@@
     \toprule
-A woman is not boiling shrimps.& \ii{contradiction}&	A woman is boiling shrimps.\\
-A wild deer is jumping a fence. &\ii{entailment}	&A deer is jumping a fence.\\
-A potato is being sliced by a woman. &\ii{entailment}	&A woman is slicing a potato.\\
-The plane, which is south African, is flying in a blue sky.& \ii{entailment}&	An airplane is flying through the air.\\
+Analysis\\
+Analysis\\
+Analysis\\
+Analysis\\
     \bottomrule
   \end{tabular}
-  \caption{\label{examplesofsickdata}Examples of correct RNTN classifications on SICK.}
+  \caption{\label{examplesofsickdata}Examples of correct TreeRNTN classifications on SICK.}
 \end{table*} % TODO: Update. Find incorrect examples?
@@
 GloVe \cite{pennington2014glove} on data from Wikipedia.
-Since 200 dimensional vectors are too large to be practical in an RNTN on a small dataset, a new embedding transformation layer is needed.
+Since 200 dimensional vectors are too large to be practical in an TreeRNTN on a small dataset, a new embedding transformation layer is needed.
 Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer.
@@
     \toprule
-    ~&\multicolumn{1}{c}{25d SumNN} & \multicolumn{1}{c}{30d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
+    ~&\multicolumn{1}{c}{SICK Train} & \multicolumn{1}{c}{DG Train}  & \multicolumn{1}{c}{SICK Test}\\
     \midrule
-    SICK Train &  TODO & 95.4 &  \textbf{97.8}  \\
-    DG (Train) &  TODO & 67.0 & \textbf{74.0}  \\
-    SICK Test & TODO & 74.9 & \textbf{76.9}  \\
+    NEUTRAL only &	56.7	&	50.0	  &	56.7\\
+    25d SumNN	&	TODO	&	TODO &	TODO	\\
+    30d TreeRNN	&	95.4		&	67.0	&	74.9		\\
+    50d TreeRNTN	&\textbf{97.8}	& \textbf{74.0}&\textbf{76.9}\\
     \bottomrule
   \end{tabular}
-  \caption{Performance on collapsed SICK data. The class NEUTRAL appears in 56.7\% of SICK examples and 50\% of DG examples.}
+  \caption{Classification accuracy for SICK.}
   \label{sresultstable}
@@
 Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment.
-Our RNTN did not reach the performance of the winning system (84.6\%), but it did exceed that of eight out of 18 submitted systems, including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
+Our TreeRNTN did not reach the performance of the winning system (84.6\%), but it did exceed that of eight out of 18 submitted systems, including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
 
@@
 
-% TODO: Better error analysis
+% TODO: Update analysis
 
@@
 The latter two show that the model can also identify the relationships between superficially different sentences, such as passive--active pairs and sentences with low lexical overlap.
+
+ANALYSIS\\
+ANALYSIS\\
+ANALYSIS\\
+ANALYSIS\\
+ANALYSIS\\
+ANALYSIS\\
+ANALYSIS\\
+ANALYSIS\\

### 2015-02-27 writing/F'14 paper/discussion.tex
Paper update.

@@
 involving quantification and negation. We then show that the same models can learn to
-perform an entailment task on natural language image captions. The results suggest that TreeRNTNs,
+perform an entailment task on natural language. The results suggest that TreeRNTNs,
 and potentially also TreeRNNs, have the capacity to learn to model these tasks with 
@@
 much insight on this point, but we hope that further experiments may reveal 
-some structure in the representations.
+structure in the learned parameters or the representations they produce.
 

### 2015-02-27 writing/F'14 paper/intro.tex
Paper update.

@@
 
-% TODO: Is there any way to squeeze a useful example into the first page?
-% TODO: Can we fit a version of the ``find something that does that'' quote in intro somewhere?
-
-% TODO: This para might be a bit repetitive.
 Tree-structured recursive neural network models (TreeRNNs; \citealt{goller1996learning}) for sentence meaning
@@
 the ability of these models to learn to produce and use strong 
-semantic representations for sentences,
-though many remain skeptical of the extent to which they can learn to represent
-the necessary complex compositional structures using fixed-size vectors. 
-In particular, it remains an open question whether learned models can achieve the
-high-fidelity distributed representations of recent algebraic
-approaches to vector space modeling \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
-and whether they can match the performance of grammars based in logical forms
+semantic representations for sentences, but it remains an 
+open question whether any such fully learned model can achieve the kind of
+high-fidelity distributed representations proposed in recent algebraic
+work on vector space modeling 
+\cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
+and whether any such model can match the performance of grammars based in logical forms
 in their ability to model core semantic phenomena like quantification, entailment, and contradiction \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}.
@@
 surpassing several submitted models with significant hand-engineered task-specific features.
+As with our simulated data, our concern is primarily with the quality of the learned representations and their ability to encode our target semantic relations.
+Thus, we are less focused on absolute performance than on patterns of accurate and inacccurate predictions, so even this small data set can provide powerful analytic insights into our core representational issues.
 This suggests that the representational abilities that we observe in the previous sections
@@
 any clear level of inferential complexity for which other approaches work and NN models fail.
-
-% TODO: Contributions para.
 

### 2015-02-27 writing/F'14 paper/join.tex
Paper update.

@@
   \end{array}$
-  \caption{In \S\ref{sec:join}, we assess our models' ability to learn to do inference over pairs of relations using the rules represented here.
+  \caption{In \S\ref{sec:join}, we assess our models' ability to learn to do inference over pairs of relations using the rules represented here, which are derived from the definitions of the relations in Table~\ref{b-table}.
 As an example, given that $p_1 \natfor p_2$ and $p_2 \natneg p_3$, the entry in the $\natfor$ row and the $\natneg$ column lets us conclude that $p_1 \natalt p_3$.
-These rules are derived from the definitions of the relations in Table~\ref{b-table}.  Cells containing a dot correspond to pairs
-of relations for which no valid inference can be drawn.} 
+Cells containing a dot correspond to situations for which no valid inference can be drawn.}
   \label{tab:jointable}
@@
 relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
-We divide these statements evenly into training and test sets, and remove from the
-test set statements which cannot be proven from the training
-statements.
+We divide these statements evenly into train and test sets, and delete the test set
+ examples which cannot be proven from the train examples, for which there is not enough information for even an ideal system to choose a correct label.
 In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
@@
 \paragraph{Results} 
-Table \ref{joinresultstable} shows the results. 
-The NTN is able to accurately encode the relations 
-between the terms in the geometric relations between their vectors, 
+The resuts (Table \ref{joinresultstable}) show that NTN is able to accurately encode the relations between the terms in the geometric relations between their vectors, 
 and is able to then use that information to recover relations that 

### 2015-02-27 writing/F'14 paper/methods.tex
Paper update.

@@
 multiplicative interactions between the child vectors. 
-Other tree-structured networks have been successfully used in past work
-\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, but
-we believe that these two provide a valuable case study, and that positive results on 
-here are likely to generalize well to stronger models.
-
 The comparison layer uses the same layer function as the
@@
 
-To run the model forward and label a pair of phrases, the structure of
-the lower layers of the network is assembled so as to match the tree
-structures provided for each phrase, which is either included in the source data or given by a parser.
+Other strong tree-structured networks have been proposed in past work
+\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, but
+we believe that these two provide a valuable case study, and that positive results on 
+here are likely to generalize well to stronger models.
+
+To run the model forward, we assemble the two tree networks so as to match the structures provided for each phrase, which is either included in the source data or given by a parser.
 The word vectors are then looked up from the vocabulary embedding matrix $V$ (one of the model parameters), and
 the composition and comparison functions are used to pass information
-up the tree and into the classifier at the top. For an objective
+up the tree and into the classifier. For an objective
 function, we use the negative log of the probability assigned to the

### 2015-02-27 writing/F'14 paper/quantifiers.tex
Paper update.

@@
 formed a standard case study in prior formal work on natural
-language inference \cite{Icard:Moss:2013:LILT}..
+language inference \cite{Icard:Moss:2013:LILT}.
 
 \paragraph{Experiments}
-Our experimental data consist of pairs of sentences generated
+Our data consist of pairs of sentences generated
 from a grammar for a simple English-like artificial language.
@@
 which may be negated, and an intransitive verb which may be
-negated. We use the basic quantifiers \ii{some}, \ii{most}, \ii{all},
+negated. We use the quantifiers \ii{some}, \ii{most}, \ii{all},
 \ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
-\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}. We also
+\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}, and also
 include five nouns, four intransitive verbs, and the negation symbol
 \ii{not}. In order to be able to define relations between sentences
-with differing lexical items, we define the lexical relations between
+with differing lexical items, we define the lexical relations for
 each noun--noun pair, each verb--verb pair, and each
-quantifier--quantifier pair. The grammar generates pairs of
-sentences of this form and calculates the relations holding
-between them.  Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.
+quantifier--quantifier pair. The grammar then generates pairs of
+sentences and calculates the relations
+between them.  
+Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.
 
@@
 
-We use the same summing baseline as in the previous section.
+We use the same summing baseline as in \S\ref{sec:recursion}.
 The highly consistent  sentence structure in this experiment means
 that that model can potentially distinguish two valid sentences unless they
- differ only in the placement of the word \ii{not}, potentially introducing errors into 
+ differ only in the placement of \ii{not}, potentially introducing errors on 
 the 58.8\% of examples that contain a sentence with a single
@@
     $\natind$ only &	35.4 & (7.5)	& 35.4	& (7.5)\\
-    25d SumNN	&	96...&	(95...)&	90...&	(80...)\\	
-    25d TreeRNN	&	\textbf{100}&	\textbf{(100)}&	99.5&	(99.3)\\
-    25d TreeRNTN	&	\textbf{99.4}&\textbf{(99.5)}&	\textbf{99.4} & (99.2)\\
+    25d SumNN	&	97.0&	(97.2)&	89.6&	(91.1)\\	
+    25d TreeRNN	&	\textbf{100}&	\textbf{(100)}&	\textbf{99.4}&	\textbf{(99.5)}\\
+    25d TreeRNTN	&	99.4&(99.5)&	99.4 & (99.2)\\
     \bottomrule
@@
 \end{table}
-% TODO: Fill in baseline results
 
@@
 %
-\paragraph{Results} The results, reported in Table~\ref{qresultstable}, show that both recursive models are able to learn to generalize the underlying logic satisfactorily, and that the baseline model generalizes well, but far from perfectly.
-We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.
-
-% TODO: Baseline
+\paragraph{Results} The results (Table~\ref{qresultstable}) show that both tree models are able to learn to generalize the underlying logic almost perfectly.
+The baseline summing model can largely memorize the training data, but does not generalize well.
+We do not find any consistent pattern in the handful of errors made by either tree model, and no errors were consistent across model restarts, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

### 2015-02-27 writing/F'14 paper/recursion.tex
Paper update.

@@
 \section{Recursive structure}\label{sec:recursion}
-% TODO: Enter baseline results, use [[Tree]]RNN.
 
-% TODO: Clean up next sent.
 Recursive structure is a prominent feature of natural language. 
-This section examines the degree to which our models are able to learn a 
+This section examines the degree to which our models can learn a 
 compositional semantics over recursive structures.
-In our evaluations, we exploit the fact that our logical language
+In our evaluations, we exploit the fact that our target language
 is infinite by testing on strings that are longer and more complex
@@
 \paragraph{Experiments}
-As in \S\ref{sec:join}, we test this phenomenon on data generated from a logic,
+As in \S\ref{sec:join}, we generate artificial data from a formal system,
  but we now replace the unanalyzed symbols
 from that experiment with complex formulae. These formulae
-represent a truth-functionally complete classical propositional logic:
+represent a complete classical propositional logic:
 each atomic symbol is a variable over the domain \{$\True$, $\False$\}, and the only
 operators are truth-functional ones.  Table~\ref{tab:pl} defines this
-logic, and Table~\ref{tab:plexs} gives examples of the some of the shorter relational
-statements that we generate from it as train and test data. To compute these relations
+logic, and Table~\ref{tab:plexs} gives some short examples relational statements from our data.
+ To compute these relations
 between statements, we exhaustively enumerate the sets of assignments
@@
 Table~\ref{b-table}. As a result, each relational statement represents
-a valid theorem of the propositional logic, and and to succeed, the models 
+a valid theorem of the propositional logic, and to succeed, the models 
 must learn to reproduce the behavior of a theorem prover.\footnote{
@@
 bin for a held-out test set.
-
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=5.75in]{decayfig.eps}
-%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-  \caption{Model performance on propositional logic, by expression size.}
-%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
-%    representations. As measured by Euclidean distance, only the TreeRNTN
-%    achieves this for formulae containing more than a small number of
-%    connectives (\ii{and} in this example). The TreeRNN quickly collapses the representations,
-%    failing to capture the meaning contrasts.
-
-  \label{prop-results}
-
-\end{figure*}
-
 If we do not implement any constraint that the two statements being
@@
 In order to test the model's generalization to unseen structures, we discard
-training examples with more than 4 connectives, yielding 60k short training examples,
-and 21k test examples containing up to 12 connectives.
-We evaluated both the TreeRNN and TreeRNTN on the resulting data, initializing the model parameters randomly,
-including the vector representations of the six variables. 
-As a baseline, we also train a summing neural network
-which is largely identical to the TreeRNN, except that instead of learning a parametric composition function,
-it simply sums the term vectors in each expression to compose them. Unlike the two models under study,
-whose ability to accurately simulate the logic is an open question, this baseline does not use word order,
+training examples with more than 4 logical operators, yielding 60k short training examples,
+and 21k test examples containing up to 12 operators.
+In addition to the two tree models, we also train a summing NN baseline
+which is largely identical to the TreeRNN, except that instead of using a learned composition function,
+it simply sums the term vectors in each expression to compose them. Unlike the two tree models, this baseline does not use word order,
 and is as such guaranteed to ignore some information that it would need to succeed perfectly.
 
-% TODO: Insert baseline result.
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=3.05in]{decayfig.pdf}
+%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
+  \caption{Results on recursive structure.
+The ``$\natind$ only'' line shows the frequency of the most frequent class, $\natind$, and the vertical dotted line marks the size of the longest training examples.}
+%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
+%    representations. As measured by Euclidean distance, only the TreeRNTN
+%    achieves this for formulae containing more than a small number of
+%    connectives (\ii{and} in this example). The TreeRNN quickly collapses the representations,
+%    failing to capture the meaning contrasts.
+
+  \label{prop-results} 
+\end{figure}
+
 \paragraph{Results} Fig.~\ref{prop-results} shows the relationship
-between test accuracy and statement length. BASELINE BASELINE
- BASELINE BASELINE BASELINE BASELINE BASELINE BASELINE BASELINE   We found that both 
+between test accuracy and statement length. While the summing baseline model performed poorly across the board, we found that both recursive
 models were able to perform well on unseen small test examples, 
 with TreeRNN accuracy above
-98\% and TreeRNTN accuracy above 99\% on formulae below length five.
-Starting at size four, performance gradually falls with increasing
-size. Training accuracy was 99.4\% for the TreeRNN and 99.8\% for the TreeRNTN.
+98\% and TreeRNTN accuracy above 99\% on formulae below length five, indicating
+that they learned correct approximations of the underlying
+logic. Training accuracy was 66.6\% for the SumNN, 99.4\% for the TreeRNN, and 99.8\% for the TreeRNTN.
 
-The performance of these models on small unseen test examples
-indicates that they learned correct approximations of the underlying
-logic. These approximations are accurate enough to yield
-correct answers when the composition layer is only applied a small
-number of times, but that the error in the approximation grows with
-increasing expression size, resulting in gradually dropping performance.
-
-It appears that the TreeRNTN model overfit somewhat relative to the plain TreeRNN,
-performing better on the training data and short examples, but worse on longer
-examples. Two factors are likely involved in this: even with the lower dimension,
+After the size four training cutoff, performance gradually decays with expression size for both tree models, suggesting that the learned approximations were accurate but lossy.
+Despite its stronger performance on short sentences, but the TreeRNTN's performance
+decayed more quickly than the TreeRNN's. 
+This suggests to us that it learned to interpret many specific fixed-size tree structures directly,
+allowing it to get away without learning strong generalizations about how to compose
+individual terms in a more general way.
+Two factors may have contributed to the learning of these narrower generalizations: 
+even with the lower dimension,
 the TreeRNTN composition function has about eight times as many parameters as the
-TreeRNN, and thus more capacity to find a solution involving a large number of narrow 
-generalizations that work on the training data but don't generalize well. In addition,
-the TreeRNTN used here has weaker L2 regularization than the 
-TreeRNN ($\lambda = 0.0003$ vs. $0.001$). This
-yielded better performance averaged across all bins, but may have facilitated this effect.
-
-Though imperfect, model performance on the most difficult set of expressions is nonetheless striking. 
-The weaker TreeRNTN classifies true examples of every class but $\nateq$ 
-(which is rare in long examples, and occurs only once here) correctly the majority of the time.
-% TODO: Better present examples.
+TreeRNN, and the TreeRNTN worked best with weaker L2 regularization than the 
+TreeRNN ($\lambda = 0.0003$ vs. $0.001$). Nonetheless, the performance of both 
+models on even the longest set of test sentences is good enough to indicate that
+ both can still perform at least least a rough approximation of the underlying theorem proving task.
 

### 2015-02-27 writing/F'14 paper/sick.tex
Paper update.

@@
 
-The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and the diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data alone are not sufficient to address this issue.
+The specific model architecture that we use is novel, and though the underlying tree structure approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and the diverse range of linguistic structures seen in typical natural language data.
 To investigate the ability of the model to learn on natural data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}.
 The corpus consists of about 10k natural language sentence pairs, labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}.
-At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of.
-
-\begin{table*}[htp]
-  \centering\small
-  \begin{tabular}{lcl}
-    \toprule
-Analysis\\
-Analysis\\
-Analysis\\
-Analysis\\
-    \bottomrule
-  \end{tabular}
-  \caption{\label{examplesofsickdata}Examples of correct TreeRNTN classifications on SICK.}
-\end{table*} % TODO: Update. Find incorrect examples?
+At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of, it is sufficient to provide some assurances as to the strength of these models.
 
@@
 This new layer aggregates any usable information from the embedding vectors into a more compact working representation.
+An identical layer is added to the SumNN between the word vectors and the comparison layer.
 
@@
   \centering \small
-  \begin{tabular}{ l r@{ \ } r@{ \ } r@{ \ } }
+    \begin{tabular}{ l@{~~} r@{~~~~} r@{~~~~} r@{~~~~} r@{~~~~} }
     \toprule
-    ~&\multicolumn{1}{c}{SICK Train} & \multicolumn{1}{c}{DG Train}  & \multicolumn{1}{c}{SICK Test}\\
+        ~&\ii{neut.}&	 30d  & 			30d & 50d\\
+    ~&only &SumNN  &TrRNN &TrRNTN\\ 
      \midrule
-    NEUTRAL only &	56.7	&	50.0	  &	56.7\\
-    25d SumNN	&	TODO	&	TODO &	TODO	\\
-    30d TreeRNN	&	95.4		&	67.0	&	74.9		\\
-    50d TreeRNTN	&\textbf{97.8}	& \textbf{74.0}&\textbf{76.9}\\
+    DG Train	& 50.0 & 61.0? & 67.0 & \textbf{74.0} \\
+    SICK Train	& 56.7 & 95.8? & 95.4 & \textbf{97.8} \\
+    SICK Test	& 56.7 & 74.1? & 74.9 & \textbf{76.9} \\
+    \midrule
+    \textsc{Pass} (4\%)	& 7 & 73? & \textbf{93} & 87 \\   
+    \textsc{Neg} (7\%)	& 5 & \textbf{100?} & 95 &75 \\
+    \textsc{Sub} (24\%)	& 35 & 73? & 73 & \textbf{76} \\
+    \textsc{Mult} (40\%)	& 76 & 66? & \textbf{78} &76 \\
+    \textsc{Diff} (25\%)	& \textbf{95} & 86? & 82 & 91 \\  
+    \midrule
+    \textsc{Lt10} (47\%)	& 50.0 & 75.2? & 77.0 &\textbf{77.9} \\    
     \bottomrule
   \end{tabular}
-  \caption{Classification accuracy for SICK.}
+  \caption{Classification accuracy, including a category breakdown for SICK test data. Categories are shown with their frequencies.}
   \label{sresultstable}
@@
 
-The results are shown in Table \ref{sresultstable}.
-Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment.
+% TODO: Update baseline results, apply bolding
+
+\begin{table*}[htp]
+  \centering\small
+  \begin{tabular}{lcl}
+    \toprule
+    A man is cutting a paper plate	& \ii{entailment} & A paper plate is being cut by the man (\textsc{Pass})\\
+    A little girl is playing the violin on a beach & \ii{contradiction} &	There is no girl playing the violin on a beach (\textsc{Neg})\\
+    A man is fixing a silencer to a gun & \ii{entailment} & A man is fitting a silencer to a weapon (\textsc{Sub})\\
+        A woman is breaking two eggs in a bowl & \ii{neutral} &A man is mixing a few ingredients in a bowl (\textsc{Multi})\\
+        Dough is being spread by a man & \ii{neutral} & A woman is slicing meat with a knife (\textsc{Diff})\\
+    \bottomrule
+  \end{tabular}
+  \caption{\label{examplesofsickdata}Examples of each category used in error analysis from the SICK test data. }
+\end{table*}
+
+\paragraph{Results} Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships, the results (Table~\ref{sresultstable}) show that it is possible to train our model to perform competitively on textual entailment.
 Our TreeRNTN did not reach the performance of the winning system (84.6\%), but it did exceed that of eight out of 18 submitted systems, including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
@@
 
-Table \ref{examplesofsickdata} shows a few typical examples of correct classifications.
-Our qualitative evaluation suggests that the model is able to successfully recognize negation and the insertion or deletion of words, as in the first two examples.
-The latter two show that the model can also identify the relationships between superficially different sentences, such as passive--active pairs and sentences with low lexical overlap.
+It is often the case for textual entailment (and is known to be the case for SICK; \citealt{marelli2014semeval}), it is possible to perform well without making any attempt to take advantage of compositional syntactic or semantic structure, and our summing baseline model is powerful enough to do this.
+Our tree models nonetheless perform substantially better than this model, and we remain confident that given sufficient data, it should be possible for the tree models, and not the summing model, to learn a truly high-quality solution.
 
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
+To better understand how the models differ, we manually annotated a fraction of the SICK test set, using mutually exclusive categories for passive/active alternation pairs (\textsc{Pass}), pairs differing only by the presence of negation (\textsc{Neg}), pairs differing by a single word or phrase substitution (\textsc{Sub}), pairs differing by multiple edits (\textsc{Mult}), and pairs with little or no content word overlap (\textsc{Diff}).
+Examples of each are in Table \ref{examplesofsickdata}.
+We annotated 100 random examples to judge the frequency of each category, and  continued selectively annotating until each categogy contained at least 25.
+We also use category (\textsc{Lt10}) for pairs in which neither sentence contains more than ten words.
+
+The results (Table \ref{examplesofsickdata}) show that tree models perform well in the two categories which most narrowly pick out a specific syntactic structure, \textsc{Pass} and \textsc{Neg}.
+The baseline model also performs well on \textsc{Neg}, where identifying the presence of negation in only one sentence is sufficient to assign \ii{contradiction}, but it does not do as well on P/A, for which distinguishing correct entailments from distractors requires identifying that the same phrase describes fills different syntactic roles in the two sentences.

### 2015-02-27 writing/F'14 paper/intro.tex
Paper update.

@@
 the ability of these models to learn to produce and use strong 
-semantic representations for sentences, but it remains an 
+semantic representations for sentences. However, it remains an 
 open question whether any such fully learned model can achieve the kind of
@@
 \newcite{ClarkCoeckeSadrzadeh2011} has yielded rich frameworks 
-for computing the meanings of complex fragments of natural
+for computing the meanings of fragments of natural
 language compositionally from vector or tensor representations, but 
 has not yet yielded effective methods for learning these representations
-entirely from data in typical machine learning settings. 
-To date, experimental investigations of these concepts using
+from data in typical machine learning settings. 
+Past experimental work on reasoning with
 distributed representations have been largely confined to short
@@
 need of intermediate representations in an artificial logical
-language. This logic
+language.
+In our first three experiments, we test our models' ability to learn the foundations of natural
+language inference by training them to reproduce the behavior of the natural logic of \newcite{maccartney2009extended} 
+on artificial data. This logic
 defines seven mutually-exclusive relations of synonymy, entailment, contradiction,
@@
 are now well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
-
-While one of our experiments uses natural logic only as an approximate
-description of the desired model behavior, three experiments draw on this
- tradition much more directly, and
-test models' abilities to learn to reproduce the precise behavior of the natural logic developed by
-\newcite{maccartney2009extended}: one
-covers learning to reason with the bare logical relations (\S\ref{sec:join}), a second covers reasoning with statements constructed compositionally from recursive functions (\S\ref{sec:recursion}),
-and a third covers the additional complexity that results from quantification (\S\ref{sec:quantifiers}).
+The first experiment using this logic 
+covers reasoning with the bare logical relations (\S\ref{sec:join}), the second extends this to reasoning with statements constructed compositionally from recursive functions (\S\ref{sec:recursion}),
+and the third covers the additional complexity that results from quantification (\S\ref{sec:quantifiers}).
 Though the performance of the plain TreeRNN model 
@@
 surpassing several submitted models with significant hand-engineered task-specific features.
-As with our simulated data, our concern is primarily with the quality of the learned representations and their ability to encode our target semantic relations.
-Thus, we are less focused on absolute performance than on patterns of accurate and inacccurate predictions, so even this small data set can provide powerful analytic insights into our core representational issues.
+Our concern is primarily with the quality of the learned representations and their ability to encode our target semantic relations.
+Thus, we are less focused on absolute performance than on patterns of accurate and inacccurate predictions, so even this small data set can offer insight into our core representational issues.
 This suggests that the representational abilities that we observe in the previous sections
@@
   \end{tabular}
-  \caption{\label{b-table}The seven natural logic relations of MacCartney and Manning \protect\shortcite{maccartney2009extended} are defined abstractly on pairs of sets drawing from the universe  $\mathcal{D}$, and can be applied to any pair of words, phrases, or sentences according to standard approaches to semantic interpretation.
-% TODO: Condense, leave in reference to denotation.
-The relation $\natind$ applies where none of the other six do.} %, including when there 
-    %is insufficient knowledge to choose a label.}
+  \caption{\label{b-table}The seven relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.} %-%
 

### 2015-02-27 writing/F'14 paper/join.tex
Paper update.

@@
 
-As an example, both models successfully classified the test item $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$.
+As an example from our test data, both models correctly labeled $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$.
 On another example involving comparably frequent relations, the NTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the NN incorrectly assigned it $\natind$.

### 2015-02-27 writing/F'14 paper/recursion.tex
Paper update.

@@
 
-Recursive structure is a prominent feature of natural language. 
-This section examines the degree to which our models can learn a 
-compositional semantics over recursive structures.
+A successful natural language inference system must reason 
+about relations not just over familiar
+atomic symbols, but also over novel structures built up 
+recursively from these symbols. 
+This section shows that our models can learn a 
+compositional semantics over such structures.
 In our evaluations, we exploit the fact that our target language

### 2015-02-27 writing/F'14 paper/sick.tex
Paper update.

--- 2015-02-25 writing/F'14 paper/sick.tex
@@
 
-The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and the diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data alone are not sufficient to address this issue.
-To investigate the ability of the model to learn on natural data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}.
+The specific model architecture that we use is novel, and though the underlying tree structure approach has been validated elsewhere, our experiments so far do not guarantee that it viable model for handling inference over real
+natural language data.
+To investigate our models' ability to handle the noisy labels and the diverse range of linguistic structures seen in typical natural language data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}.
 The corpus consists of about 10k natural language sentence pairs, labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}.
-At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of.
-
-\begin{table*}[htp]
-  \centering\small
-  \begin{tabular}{lcl}
-    \toprule
-Analysis\\
-Analysis\\
-Analysis\\
-Analysis\\
-    \bottomrule
-  \end{tabular}
-  \caption{\label{examplesofsickdata}Examples of correct TreeRNTN classifications on SICK.}
-\end{table*} % TODO: Update. Find incorrect examples?
+At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of, and our results
+nonetheless show that tree-structured NN models can learn to do inference in the real world.
 
@@
 We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source.
-We parsed the data from both sources with the Stanford PCFG Parser v.
-3.3.1 \cite{klein2003accurate}.
-We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word.
+We parsed the data from both sources with the Stanford PCFG Parser v.~3.3.1 \cite{klein2003accurate}.
+We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair by replacing them with a single head word.
 The training and test data on which we report performance are collapsed in this way, and both collapsed and uncollapsed copies of the training data are used in training.
@@
   \centering \small
-  \begin{tabular}{ l r@{ \ } r@{ \ } r@{ \ } }
+    \begin{tabular}{ l@{~~} r@{~~~~} r@{~~~~} r@{~~~~} r@{~~~~} }
     \toprule
-    ~&\multicolumn{1}{c}{SICK Train} & \multicolumn{1}{c}{DG Train}  & \multicolumn{1}{c}{SICK Test}\\
+        ~&\ii{neut.}&	 30d  & 			30d & 50d\\
+    ~&only &SumNN  &TrRNN &TrRNTN\\ 
      \midrule
-    NEUTRAL only &	56.7	&	50.0	  &	56.7\\
-    25d SumNN	&	TODO	&	TODO &	TODO	\\
-    30d TreeRNN	&	95.4		&	67.0	&	74.9		\\
-    50d TreeRNTN	&\textbf{97.8}	& \textbf{74.0}&\textbf{76.9}\\
+    DG Train	& 50.0 & 64.0 & 67.0 & \textbf{74.0} \\
+    SICK Train	& 56.7 & 95.2 & 95.4 & \textbf{97.8} \\
+    SICK Test	& 56.7 & 73.6 & 74.9 & \textbf{76.9} \\
+    \midrule
+    \textsc{Passive} (4\%)	& 0 		& 80  		& 68		&\textbf{88}\\   
+    \textsc{Neg} (7\%)	& 0 		& 88	 		& \textbf{100} & \textbf{100}\\
+    \textsc{Subst} (24\%)	& 28 		& 60  		& 64 		&  \textbf{72}\\
+    \textsc{MultiEd} (39\%)	&  \textbf{68} & 66  		&66 		& 64 \\
+    \textsc{Diff} (26\%)	& \textbf{96} &  	82		&79		& \textbf{96}\\  
+    \midrule
+    \textsc{Short} (47\%) & 50.0 & 73.9 & 75.9		& \textbf{77.3} \\    
     \bottomrule
   \end{tabular}
-  \caption{Classification accuracy for SICK.}
+  \caption{Classification accuracy, including a category breakdown for SICK test data. Categories are shown with their frequencies.}
   \label{sresultstable}
@@
 
-The results are shown in Table \ref{sresultstable}.
-Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment.
-Our TreeRNTN did not reach the performance of the winning system (84.6\%), but it did exceed that of eight out of 18 submitted systems, including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
+\begin{table*}[htp]
+  \centering\small
+  \begin{tabular}{lcl}
+    \toprule
+    A man is cutting a paper plate	& \ii{entailment} & A paper plate is being cut by the man (\textsc{Passive})\\
+    A little girl is playing the violin on a beach & \ii{contradiction} &	There is no girl playing the violin on a beach (\textsc{Neg})\\
+    A man is fixing a silencer to a gun & \ii{entailment} & A man is fitting a silencer to a weapon (\textsc{Subst})\\
+        A woman is breaking two eggs in a bowl & \ii{neutral} &A man is mixing a few ingredients in a bowl (\textsc{MultiEd})\\
+        Dough is being spread by a man & \ii{neutral} & A woman is slicing meat with a knife (\textsc{Diff})\\
+    \bottomrule
+  \end{tabular}
+  \caption{\label{examplesofsickdata}Examples of each category used in error analysis from the SICK test data. }
+\end{table*}
 
-% TODO: Update results
+\paragraph{Results} Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships, the results (Table~\ref{sresultstable}) show that our tree-structured models perform competitively on textual entailment, beating a strong baseline.
+Neither model reached the performance of the winning system (84.6\%), but the TreeRNTN did exceed that of eight out of 18 submitted systems, including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
 
-% TODO: Update analysis
+We know that it is possible to perform well on SICK (like other natural language inference corpora) without taking advantage of compositional syntactic or semantic structure \cite{marelli2014semeval}, and our summing baseline model is powerful enough to do this.
+Our tree models nonetheless perform substantially better, and we remain confident that given sufficient data, it should be possible for the tree models, and not the summing model, to learn a truly high-quality solution.
 
-Table \ref{examplesofsickdata} shows a few typical examples of correct classifications.
-Our qualitative evaluation suggests that the model is able to successfully recognize negation and the insertion or deletion of words, as in the first two examples.
-The latter two show that the model can also identify the relationships between superficially different sentences, such as passive--active pairs and sentences with low lexical overlap.
+To better understand how the models differ, we manually annotated a fraction of the SICK test set, using mutually exclusive categories for passive/active alternation pairs (\textsc{Passive}), pairs differing only by the presence of negation (\textsc{Neg}), pairs differing by a single word or phrase substitution (\textsc{Subst}), pairs differing by multiple edits (\textsc{Mult}), and pairs with little or no content word overlap (\textsc{Diff}).
+Examples of each are in Table \ref{examplesofsickdata}.
+We annotated 100 random examples to judge the frequency of each category, and  continued selectively annotating until each categogy contained at least 25.
+We also use category (\textsc{Short}) for pairs in which neither sentence contains more than ten words.
 
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
-ANALYSIS\\
+The results (Table \ref{examplesofsickdata}) show that the TreeRNTN performs especially strongly in the two categories which most narrowly pick out a specific syntactic configuration, \textsc{Passive} and \textsc{Neg}.
+It also beats the summing baseline on \textsc{Subst}, which most closely parallels the kind of monotonicity inferences studied in \S\ref{sec:quantifiers}.
+None of the three models perform better on the short \textsc{Short} pairs than on the rest of the data, suggesting that the decay phenomenon observed in \S\ref{sec:recursion} may not hurt the performance of models trained on typical natural language sentences.

### 2015-02-28 writing/F'14 paper/intro.tex
Paper update

@@
 The experiments with simulated data provide a convincing demonstration
-of the ability of neural networks to learn to encode semantic
-inferences in complex natural language sentences from reasonably-sized
+of the ability of neural networks to learn to build and use semantic representations for complex natural language sentences from reasonably-sized
 training sets. However, we are also interested in the more practical
@@
 nonetheless able to achieve competitive performance on it,
-surpassing several submitted models with significant hand-engineered task-specific features.
-Our concern is primarily with the quality of the learned representations and their ability to encode our target semantic relations.
-Thus, we are less focused on absolute performance than on patterns of accurate and inacccurate predictions, so even this small data set can offer insight into our core representational issues.
+surpassing several submitted models with significant hand-engineered task-specific features and our own NN baseline.
 This suggests that the representational abilities that we observe in the previous sections
 are not limited to carefully circumscribed tasks. We conclude that TreeRNTN models
-are adequate for typical cases of natural logic inference, and that there is not yet
+are adequate for typical cases of natural language inference, and that there is not yet
 any clear level of inferential complexity for which other approaches work and NN models fail.

### 2015-02-28 writing/F'14 paper/methods.tex
Paper update.
Paper update

@@
-\section{Recursive neural network models} \label{methods}
+\section{Tree-structured neural networks} \label{methods}
 
@@
 via specific composition functions \cite{Partee84,Janssen97}. In our
-distributed setting, word meanings are embedding vectors of dimension $n$. The
+distributed setting, word meanings are embedding vectors of dimension $n$. A learned
 composition function maps pairs of them to single phrase vectors of dimension $n$, 
 which can then be merged again to represent more complex
-phrases. Once the entire sentence-level representation has been
-derived, it serves as a fixed-dimensional input for some subsequent function.
+phrases, forming a tree structure. Once the entire sentence-level representation has been
+derived at the top of the tree, it serves as a fixed-dimensional input for some subsequent layer function.
 
@@
 
-Other strong tree-structured networks have been proposed in past work
+Other strong tree-structured models have been proposed in past work
 \cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, but
@@
 
-To run the model forward, we assemble the two tree networks so as to match the structures provided for each phrase, which is either included in the source data or given by a parser.
-The word vectors are then looked up from the vocabulary embedding matrix $V$ (one of the model parameters), and
+To run the model forward, we assemble the two tree-structured networks so as to match the structures provided for each phrase, which are either included in the source data or given by a parser.
+The word vectors are then looked up from the vocabulary embedding matrix $V$ (one of the learned model parameters), and
 the composition and comparison functions are used to pass information
@@
 Training times on CPUs vary from hours to days across experiments.
-On the first three experiments, which use artificial data, we report mean
-results over five fold cross-validation.
-Variance across runs is typically no
+On the experiments which use artificial data, we report mean
+results over five fold cross-validation, where
+variance across runs is typically no
 more than two percentage points. In addition, because the classes are not necessarily balanced, 
-we report performance 
-using both accuracy and macroaveraged F1.\footnote{We compute macroaveraged F1 
+we report both accuracy and macroaveraged F1.\footnote{We compute macroaveraged F1 
 as the harmonic mean of average precision and average recall, both computed

### 2015-02-28 writing/F'14 paper/quantifiers.tex
Paper update

@@
 considerably greater complexity than this.  As a key test of whether our models 
-can capture this complexity, we now directly measure the degree to which they are able to
+can capture this complexity, we now study the degree to which they are able to
 develop suitable representations for the semantics of natural language
-quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantification 
+quantifiers like \ii{most} and \ii{all} as they interact with negation and lexical entailments. Quantification 
 and negation are far from the only place in natural language where complex functional meanings
@@
   \\
-  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
+  \text{(most turtle) reptile} \natalt \text{(all turtle) (not animal)}\label{p3}
 \end{gather}
@@
 We use the same summing baseline as in \S\ref{sec:recursion}.
-The highly consistent  sentence structure in this experiment means
-that that model can potentially distinguish two valid sentences unless they
- differ only in the placement of \ii{not}, potentially introducing errors on 
-the 58.8\% of examples that contain a sentence with a single
-instance of \ii{not}.
+The highly consistent  sentence structure in this experiment means that this model
+is not as disadvantaged by the lack of word order information as it is in the previous experiment, 
+but the variable placement of \ii{not}  nonetheless introduces potential uncertainty in the 58.8\% 
+of examples that contain a sentence with a single token of it.
 

### 2015-02-28 writing/F'14 paper/sick.tex
Paper update

@@
 \section{The SICK textual entailment challenge}\label{sec:sick}
+
+% TODO: Something on how we can pack enough
 
@@
 natural language data.
-To investigate our models' ability to handle the noisy labels and the diverse range of linguistic structures seen in typical natural language data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}.
+To investigate our models' ability to handle the noisy labels and the diverse range of linguistic structures seen in typical natural language data, we use the SICK textual entailment challenge corpus \cite{marelli2014sick}.
 The corpus consists of about 10k natural language sentence pairs, labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}.
-At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of, and our results
-nonetheless show that tree-structured NN models can learn to do inference in the real world.
+At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of, and our results nonetheless show that tree-structured NN models can learn to do inference in the real world.
 
-Adapting to this task required us to make a few additions to the techniques discussed in \S\ref{methods}.
+Adapting to this task requires us to make a few additions to the techniques discussed in \S\ref{methods}.
 In order to better handle rare words, we initialized our word embeddings using 200 dimensional vectors trained with
@@
 
-We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.
+We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the 4,928 example SICK test set.
 We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}, also used by the winning SICK submission), a corpus of noisy automatically labeled entailment examples over image captions, the same genre of text from which SICK was drawn.
@@
   \centering \small
-    \begin{tabular}{ l@{~~} r@{~~~~} r@{~~~~} r@{~~~~} r@{~~~~} }
+    \begin{tabular}{ l@{\hspace{-0.25em}} r@{~~~~} r@{~~~~} r@{~~~~} r@{~~~~} }
     \toprule
-        ~&\ii{neut.}&	 30d  & 			30d & 50d\\
+        ~&\ii{neutral}&	 30d  & 			30d & 50d\\
     ~&only &SumNN  &TrRNN &TrRNTN\\ 
@@
   \centering\small
-  \begin{tabular}{lcl}
+  \begin{tabular}{l@{~~~}cl}
     \toprule
-    A man is cutting a paper plate	& \ii{entailment} & A paper plate is being cut by the man (\textsc{Passive})\\
+  The patient is being helped by the doctor	& \ii{entailment} & The doctor is helping the patient (\textsc{Passive})\\
     A little girl is playing the violin on a beach & \ii{contradiction} &	There is no girl playing the violin on a beach (\textsc{Neg})\\
-    A man is fixing a silencer to a gun & \ii{entailment} & A man is fitting a silencer to a weapon (\textsc{Subst})\\
+
+    The yellow dog is drinking water from a bottle& \ii{contradiction} &	The yellow dog is drinking water from a pot  (\textsc{Subst})\\
         A woman is breaking two eggs in a bowl & \ii{neutral} &A man is mixing a few ingredients in a bowl (\textsc{MultiEd})\\
@@
 
-We know that it is possible to perform well on SICK (like other natural language inference corpora) without taking advantage of compositional syntactic or semantic structure \cite{marelli2014semeval}, and our summing baseline model is powerful enough to do this.
+To better understand our results, we manually annotated a fraction of the SICK test set, using mutually exclusive categories for passive/active alternation pairs (\textsc{Passive}), pairs differing only by the presence of negation (\textsc{Neg}), pairs differing by a single word or phrase substitution (\textsc{Subst}), pairs differing by multiple edits (\textsc{MultiEd}), and pairs with little or no content word overlap (\textsc{Diff}).
+Examples of each are in Table \ref{examplesofsickdata}.
+We annotated 100 random examples to judge the frequency of each category, and  continued selectively annotating until each category contained at least 25.
+We also use the category \textsc{Short} for pairs in which neither sentence contains more than ten words.
+
+The results (Table \ref{examplesofsickdata}) show that the TreeRNTN performs especially strongly in the two categories which pick out specific syntactic configurations, \textsc{Passive} and \textsc{Neg}, suggesting that that model has learned to encode the relevant structures well.
+It also beats both other models on \textsc{Subst}, which most closely parallels the lexical entailment inferences addressed in \S\ref{sec:quantifiers}.
+In addition none of the models perform better on the \textsc{Short} pairs than on the rest of the data, suggesting that the performance decay observed in \S\ref{sec:recursion} may not impact models trained on typical natural language text.
+
+It is known that a model can perform well on SICK (like other natural language inference corpora) without taking advantage of compositional syntactic or semantic structure \cite{marelli2014semeval}, and our summing baseline model is powerful enough to do this.
 Our tree models nonetheless perform substantially better, and we remain confident that given sufficient data, it should be possible for the tree models, and not the summing model, to learn a truly high-quality solution.
-
-To better understand how the models differ, we manually annotated a fraction of the SICK test set, using mutually exclusive categories for passive/active alternation pairs (\textsc{Passive}), pairs differing only by the presence of negation (\textsc{Neg}), pairs differing by a single word or phrase substitution (\textsc{Subst}), pairs differing by multiple edits (\textsc{Mult}), and pairs with little or no content word overlap (\textsc{Diff}).
-Examples of each are in Table \ref{examplesofsickdata}.
-We annotated 100 random examples to judge the frequency of each category, and  continued selectively annotating until each categogy contained at least 25.
-We also use category (\textsc{Short}) for pairs in which neither sentence contains more than ten words.
-
-The results (Table \ref{examplesofsickdata}) show that the TreeRNTN performs especially strongly in the two categories which most narrowly pick out a specific syntactic configuration, \textsc{Passive} and \textsc{Neg}.
-It also beats the summing baseline on \textsc{Subst}, which most closely parallels the kind of monotonicity inferences studied in \S\ref{sec:quantifiers}.
-None of the three models perform better on the short \textsc{Short} pairs than on the rest of the data, suggesting that the decay phenomenon observed in \S\ref{sec:recursion} may not hurt the performance of models trained on typical natural language sentences.

### 2015-03-26 InitializeNNLayer.m
Forward pass through pyr with random matrices.

@@ -1,0 +1,53 @@
+function [matrix, bias] = InitializeNNLayer(indim, outdim, depth, initType)
+
+% Not currently set up for ReLU
+relu = 0;
+
+if initType == 0
+	if relu
+		scale = sqrt(2) / sqrt(outdim);
+	else
+		scale = 1 / sqrt(outdim);
+	end
+	matrix = rand(outdim, indim, depth) .* (2 * scale) - scale;
+	bias = zeros(outdim, depth);
+elseif initType == 1
+	if relu
+		scale = 2 / sqrt(indim);
+	else
+		scale = 1 / sqrt(indim);
+	end
+	matrix = rand(outdim, indim, depth) .* (2 * scale) - scale;
+	bias = zeros(outdim, depth);
+elseif initType == 2
+	scale = sqrt(6 / (outdim + indim));
+	matrix = rand(outdim, indim, depth) .* (2 * scale) - scale;
+	bias = zeros(outdim, depth);
+elseif initType == 3
+	matrix = zeros(outdim, indim, depth);
+	for d = 1:depth
+		for ind = 1:outdim
+		  indices = randi([1, indim], 1, 15);
+		  matrix(ind, indices, d) = normrnd(0, 1, 1, length(indices)) * 0.25;
+		end
+	end
+	if indim == outdim && depth > 0
+		ev = eig(matrix(:, :, 1));
+		matrix = 1.2 .* matrix ./ ev(1, 1);
+	elseif indim == 2 * outdim && depth > 0
+		ev = eig(matrix(1:outdim, 1:outdim, 1));
+		matrix = 1.2 .* matrix ./ ev(1, 1);
+	end
+	bias = zeros(outdim, depth);
+elseif initType > 3
+	matrix = zeros(outdim, indim, depth);
+	for d = 1:depth
+		for ind = 1:outdim
+		  indices = randi([1, indim], 1, 15);
+		  matrix(ind, indices, d) = normrnd(0, 1, 1, length(indices)) .* (1 / initType);
+		end
+	end
+	bias = zeros(outdim, depth);
+end
+
+end

### 2015-03-26 PyramidBatch.m

@@ -1,0 +1,109 @@
+% Want to distribute this code? Have other questions? -> sbowman@stanford.edu
+classdef PyramidBatch < handle
+    % Represents a single binary branching syntactic tree with three 
+    % representations at each node:
+    % - The index with which the feature vector can be looked up - if leaf
+    % - The text with which the tree can be displayed.
+    % - The features at the node.
+    
+    properties
+        NUMACTIONS = 3;
+        B = -1;
+        D = -1;
+        N = -1;
+        features = [];
+        wordCounts = [];
+        compositionInnerActivations = [];
+        % connectionInnerActivations = [];  % Maybe keep later? TODO.
+        connections = [];  % The above after softmax
+        connectionLabels = [];
+        % TODO: Dropout
+        % transformInnerActivations = []; % Stored activations for the embedding tranform layers. TODO.      
+    end
+    
+    methods(Static)
+        function pb = makePyramidBatch(pyramids, wordFeatures, hyperParams)
+            pb = PyramidBatch();
+            pb.B = length(pyramids);
+            pb.D = hyperParams.dim;
+            pb.N = 0;
+
+            % Find the length of the longest sequence. We use this to set the size of the main feature matrix,
+            % to this value has a large impact on the run time of the batch.
+            for b = 1:pb.B
+                pb.N = max([pb.N, pyramids.wordCount]);
+            end
+
+            pb.features = zeros(pb.N, pb.N * pb.D, pb.B);
+            pb.wordCounts = zeros(pb.B, 1);
+            pb.compositionInnerActivations = zeros(pb.N - 1, (pb.N - 1) * pb.D, pb.B);
+            % pb.connectionInnerActivations = zeros(pb.N - 1, pb.N - 1, pb.NUMACTIONS, pb.B);
+            pb.connections = zeros(pb.N - 1, pb.N - 1, pb.NUMACTIONS, pb.B);  % The above after softmax
+            pb.connectionLabels = zeros(pb.N - 1, pb.N - 1, pb.B);
+
+            for b = 1:pb.B
+                pb.wordCounts(b) = pyramids(b).wordCount;
+                for w = 1:pyramids(b).wordCount
+                    % We assume there is no embedding transform layer, so just use the word features. (TODO)
+                    pb.features(pb.N, (w - 1) * pb.D + 1:w * pb.D, b) = wordFeatures(pyramids(b).wordIndices(w), :)'; 
+                end
[…]

### 2015-03-26 RunExperiments/RunAndOrExperiments.sh

@@ -69,4 +69,44 @@ jude44.stanford.edu - 879583.scail.stanford.edu - cd quant; dataflag = 'and-or-d
 jude40.stanford.edu - 879586.scail.stanford.edu - cd quant; dataflag = 'and-or-deep-static'; lambda = 0.0000005; dim = 30; td = 1; penult = 90; dropout = 1; tot = 2; name='dimtest3'; relu = 1; TrainModel('', 1, @AndOr, name, dataflag, dim, penult, td, lambda, tot, relu, dropout, 16);
 jude42.stanford.edu - 879584.scail.stanford.edu - cd quant; dataflag = 'and-or-deep-static'; lambda = 0.000005; dim = 30; td = 1; penult = 90; dropout = 1; tot = 2; name='dimtest3'; relu = 1; TrainModel('', 1, @AndOr, name, dataflag, dim, penult, td, lambda, tot, relu, dropout, 4);
 
+# LSTM comparison
+export MATLABCMD="cd quant; dataflag = 'and-or-deep-static'; lambda = 0.0003; dim = 30; td = 1; penult = 90; tot = 1; name='/scr/sbowman/comptest'; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, 16);" ; qsub -v MATLABCMD quant/run.sh -l nodes=1:ppn=8
+export MATLABCMD="cd quant; dataflag = 'and-or-deep-3'; lambda = 0.0003; dim = 30; td = 1; penult = 90; tot = 1; name='/scr/sbowman/comptest'; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, 16);" ; qsub -v MATLABCMD quant/run.sh -l nodes=1:ppn=8
+export MATLABCMD="cd quant; dataflag = 'and-or-deep-6'; lambda = 0.0003; dim = 30; td = 1; penult = 90; tot = 1; name='/scr/sbowman/comptest'; TrainModel(''\, 1\, @AndOr\, name\, dataflag\, dim\, penult\, td\, lambda\, tot\, 16);" ; qsub -v MATLABCMD quant/run.sh -l nodes=1:ppn=8
 
[…]

### 2015-03-26 config/AndOr.m

@@ -1,13 +1,12 @@
-function [ hyperParams, options, wordMap, relationMap ] = AndOr(name, dataflag, dim, penult, top, lambda, composition, relu, maxNorm)
+function [ hyperParams, options, wordMap, relationMap ] = AndOr(name, dataflag, dim, penult, top, lambda, composition, mbs)
 % Configure the recursion experiments. 
 % NOTE: the {a-h} variables in the paper are actual multiletter names in the data used here.
 
 [hyperParams, options] = Defaults();
 
 
-hyperParams.name = [name, '-d', num2str(dim), '-pen', num2str(penult), '-top', num2str(top), ...
-				    '-comp', num2str(composition), '-relu', num2str(relu), '-l', num2str(lambda), ...
-				    '-mx', num2str(maxNorm)];
+hyperParams.name = [name, '-', dataflag, '-d', num2str(dim), '-pen', num2str(penult), '-top', num2str(top), ...
+				    '-comp', num2str(composition), '-mbs', num2str(mbs), '-l', num2str(lambda)];
 
 hyperParams.dim = dim;
 hyperParams.embeddingDim = dim;
@@ -44,19 +43,12 @@ hyperParams.topDropout = 1;
 
 hyperParams.topDepth = top;
 
-% Split longer test sets for crossvalidation without training on them.
-hyperParams.specialAndOrMode = 1;
-
-if relu
-  hyperParams.classNL = @LReLU;
-  hyperParams.classNLDeriv = @LReLUDeriv;
-end
+hyperParams.classNL = @LReLU;
+hyperParams.classNLDeriv = @LReLUDeriv;
 
 options.numPasses = 15000;
 
 options.miniBatchSize = mbs;
-
-options.maxGradientNorm = maxNorm;
 
 wordMap = InitializeMaps('/scr/sbowman/RC/longer2/wordlist.txt');
 
@@ -74,8 +66,25 @@ if strcmp(dataflag, 'and-or')
[…]

### 2015-03-31 writing/AAAI SSS '15/BowmanPottsManning/BowmanPottsManning.tex

@@
   \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{r r@{ }r r@{ }r r@{ }r r@{ }r r@{ }r} 
+  \begin{tabular}{rrrrrrr} 
     \toprule
-     Portion of & \multicolumn{4}{c}{NN} & \multicolumn{4}{c}{NTN} & \multicolumn{2}{c}{Baseline}\\
-     training data  & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} \\
+     Portion of & \multicolumn{2}{c}{25d NN} & \multicolumn{2}{c}{25d NTN} & \multicolumn{1}{c}{$\natfor$ only}\\
+     training data  & \multicolumn{1}{c}{w/ GloVe} & \multicolumn{1}{c}{w/o GloVe} & \multicolumn{1}{c}{w/ GloVe} & \multicolumn{1}{c}{w/o GloVe} \\
     \midrule
-     100\%   & 99.73 &(0.04) & 99.37$^\dagger$ &(0.14)    & 99.61 &(0.02) & \textbf{99.95}\nodagger &(0.03) & 37.05 &(--)\\
-     33\%    & 95.96 &(0.20) & 95.30\nodagger &(0.12)                & 95.83 &(0.35)          & 95.45$^\dagger$ &(0.31) & 37.05 &(--)\\
-     11\%    & 91.11 &(0.24) & 90.81$^\dagger$ &(0.20)    & 91.27 &(0.27)          & 90.90$^\dagger$ &(0.13) & 37.05 &(--)\\
-     % 2\%    & 86.0 &(100.0) & ... &(100.0)   & 86.2 &(100.0)           & 76.4 &(87.3) & 37.9 &(--)\\ % Accidental run with slightly different train/test split.
+     100\%   & 99.73 & 99.37   & 99.61  & \textbf{99.95}  & 37.05 \\
+     33\%    & 95.96 & 95.30    & 95.83  & 95.45  & 37.05 		\\
+     11\%    & 91.11 & 90.81   & 91.27  & 90.90  & 37.05 		\\
     \bottomrule

### 2015-03-31 writing/F'14 paper/arxiv/join.tex

--- 2015-02-27 writing/F'14 paper/join.tex
@@
 For instance, from the relation $p_1 \natrev p_2$ between two propositions, 
-one can conclude that $p_2 \natfor p_1$ using the definitions of the relations directly. 
-If one is also given the relation $p_1 \natrev p_3$, one can conclude that $p_1 \natrev p_2$, by basic set-theoretic reasoning (transitivity of $\natrev$). The
+one can infer the relation $p_2 \natfor p_1$ by applying the definitions of the relations directly. 
+If one is also given the relation $p_2 \natrev p_3$ one can conclude that $p_1 \natrev p_3$, by basic set-theoretic reasoning (transitivity of $\natrev$). The
 full set of sound such inferences on pairs of premise relations is depicted in
@@
 
-% about the relations themselves that do not depend on the
-% internal structure of the things being compared. For example, given
-% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
-% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven relations support more
-% than just transitivity: MacCartney and Manning's
-% \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of relations of the form $a R
-% b$ and $b R' c$, including several less intuitive ones such as that if
-% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-
 \paragraph{Experiments}
@@
 on which we will base the statements in the train and test sets.
-This takes the form of a small boolean structure in which terms denote
+This takes the form of a small Boolean structure in which terms denote
 sets of entities from a small domain.  Fig.~\ref{lattice-figure}a
-depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight terms ($p_{1\text{-}8}$). We then generate a 
+depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight proposition terms ($p_1$--$p_8$). We then generate a 
 relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
@@
  examples which cannot be proven from the train examples, for which there is not enough information for even an ideal system to choose a correct label.
-In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
-2960 relations.
+In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 examples and a test set of 
+2960 examples.
 
 We trained models with both the NN and NTN comparison functions on these
-data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no simpler version of our model is a viable baseline.} In both cases, the models are implemented as
+data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no simpler version of our model is a viable baseline.} %+%
+In both cases, the models are implemented as
 described in \S\ref{methods}, but since the items being compared
@@
     \end{picture}}
-    \caption{Example boolean structure. The terms $p_{1\text{-}8}$ name the sets. Not all sets have names, and
-    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
+    \caption{Example boolean structure. The terms $p_1$--$p_8$ name the sets. Not all sets have names, and  some sets have multiple names, so that learning $\nateq$ is non-trivial.}
   \end{subfigure}
@@
   \caption{Performance on the semantic relation experiments.
-These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
+These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.
+The ``$\natind$ only'' entries reflect the frequency of the most frequent class.}
   \label{joinresultstable}
@@
 it is capable of learning representations with these properties. 
-We cannot rule out the possibility that different optimization techniques or
+It is not possible for us to rule out the possibility that different optimization techniques or
 further hyperparameter tuning could lead an NN model to succeed here.
@@
 On another example involving comparably frequent relations, the NTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the NN incorrectly assigned it $\natind$.
-
-% From train\test_1

### 2015-04-04 writing/S'15 misc/newmodel.tex
Relation handling cleanup.

\section{Jointly learning parsing and semantic composition}

% TODO: Graph figure

\begin{figure}[tp]
  \centering\small
  \begin{tikzpicture}
    \def\dx{18pt}
    \def\dy{30pt}
    \newcounter{i}
    \tikzstyle{highlight}=[fill=dandelion!15]
    \tikzstyle{comment}=[fill=blue!15]

    \stepcounter{i}\node  (\arabic{i}) at (0*\dx,6*\dy) {$\vec{y}^{(1, 1)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-1*\dx,5*\dy) {$\vec{y}^{(2, 1)}$};
    \stepcounter{i}\node  (\arabic{i}) at (1*\dx,5*\dy) {$\vec{y}^{(2, 2)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-2*\dx,4*\dy) {$\vec{y}^{(3, 1)}$};
    \stepcounter{i}\node  (\arabic{i}) at (0*\dx,4*\dy) {$\vec{y}^{(3, 2)}$};
    \stepcounter{i}\node[highlight]  (\arabic{i}) at (2*\dx,4*\dy) {$\vec{y}^{(3, 3)}$};
    \stepcounter{i}\node[highlight]  (\arabic{i}) at (-3*\dx,3*\dy) {$\vec{y}^{(4, 1)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-1*\dx,3*\dy) {$\vec{y}^{(4, 2)}$};
    \stepcounter{i}\node  (\arabic{i}) at (1*\dx,3*\dy) {$\vec{y}^{(4, 3)}$};
    \stepcounter{i}\node  (\arabic{i}) at (3*\dx,3*\dy) {$\vec{y}^{(4, 4)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-4*\dx,2*\dy) {$\vec{y}^{(5, 1)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-2*\dx,2*\dy) {$\vec{y}^{(5, 2)}$};
    \stepcounter{i}\node  (\arabic{i}) at (0*\dx,2*\dy) {$\vec{y}^{(5, 3)}$};
    \stepcounter{i}\node  (\arabic{i}) at (2*\dx,2*\dy) {$\vec{y}^{(5, 4)}$};
    \stepcounter{i}\node[highlight]  (\arabic{i}) at (4*\dx,2*\dy) {$\vec{y}^{(5, 5)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-5*\dx,1*\dy) {My};
    \stepcounter{i}\node  (\arabic{i}) at (-3*\dx,1*\dy) {hovercraft};
    \stepcounter{i}\node  (\arabic{i}) at (-1*\dx,1*\dy) {is};
    \stepcounter{i}\node  (\arabic{i}) at (1*\dx,1*\dy) {full};
    \stepcounter{i}\node  (\arabic{i}) at (3*\dx,1*\dy) {of};
    \stepcounter{i}\node  (\arabic{i}) at (5*\dx,1*\dy) {eels};

    \stepcounter{i}\node[comment]  (\arabic{i}) at (4*\dx,4*\dy) {\textsc{compose}};
    \stepcounter{i}\node[comment]  (\arabic{i}) at (-5*\dx,3*\dy) {\textsc{copy-L}};
    \stepcounter{i}\node[comment]  (\arabic{i}) at (6*\dx,2*\dy) {\textsc{copy-R}};

    \tikzstyle{extra} = [->, draw=black, line width=.75pt]
    \tikzstyle{valid} = [->, draw=red, line width=1.25pt]

          \draw [valid] (2) -- (1);
          \draw [valid] (3) -- (1);

          \draw [valid] (4) -- (2);
          \draw [extra] (5) -- (2);
          \draw [valid] (5) -- (3);
          \draw [valid] (6) -- (3);

          \draw [valid] (7) -- (4);
          \draw [extra] (8) -- (4);
          \draw [valid] (8) -- (5);
          \draw [extra] (9) -- (5);
          \draw [valid] (9) -- (6);
          \draw [valid] (10) -- (6);

          \draw [valid] (11) -- (7);
          \draw [extra] (12) -- (7);
          \draw [valid] (12) -- (8);
          \draw [extra] (13) -- (8);
          \draw [valid] (13) -- (9);
          \draw [extra] (14) -- (9);
          \draw [valid] (14) -- (10);
          \draw [valid] (15) -- (10);

          \draw [valid] (16) -- (11);
          \draw [valid] (17) -- (11);
          \draw [extra] (17) -- (12);
          \draw [valid] (18) -- (12);
          \draw [extra] (18) -- (13);
          \draw [valid] (19) -- (13);
          \draw [extra] (19) -- (14);
          \draw [valid] (20) -- (14);
          \draw [extra] (20) -- (15);
          \draw [valid] (21) -- (15);
  \end{tikzpicture}

        \caption{The model structure used to compare \ii{turtle} and
          \ii{animal}.  Learned term representations are fed through
          either an NN or NTN comparison layer and then to a softmax
        classifier over the seven relations in Table~\ref{b-table}.}
  \label{sample-figure2}
\end{figure}

\begin{figure}[tp]
  \centering
    \footnotesize

	\newcommand{\labeledtreenode}[4][3]{\put(#2){\makebox(0,0){{\fcolorbox{black}{#4}{\makebox(#1,0.3){#3}}}}}}

	\newcommand{\textlabel}[4][3.5]{\put(#2){\makebox(0,0){{\makebox(#1,0.3){#3}}}}}

	\definecolor{lexcolor}{HTML}{F5F7C4}
	\definecolor{compositioncolor}{HTML}{BBEBFF}
	\definecolor{comparisoncolor}{HTML}{FFC895}
	\definecolor{softmaxcolor}{HTML}{A5FF8A}

	\setlength{\unitlength}{0.61cm}

	%\resizebox{3in}{!}{
	\begin{picture}(13.15, 5.5)

	  \labeledtreenode[2.4]{6.5,5}{$P(\sqsubset) = 0.8$}{softmaxcolor}  
	  \put(6.5,3.7){\vector(0,1){1}}  
	  \labeledtreenode[5]{6.5,3.4}{turtle \emph{vs.}~animal}{comparisoncolor}

	  \textlabel{2.75,5}{Softmax classifier}{black}
	  \textlabel{2.0,3.4}{\parbox{2cm}{Comparison N(T)N layer}}{black}

	  \textlabel{6.5,2.1}{Optional embedding}{black}
	  \textlabel{6.5,1.5}{transformation NN layers}{black}

	  \textlabel{6.5,0.5}{Learned term vectors}{black}

	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	  \put(1.5,0.75){\vector(0,1){0.85}}
	  \labeledtreenode{1.5,0.5}{turtle}{lexcolor}

	  \put(1.5,2.25){\vector(4,1){3.25}}
	  \labeledtreenode{1.5,1.9}{turtle}{compositioncolor}

	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	  \put(11.5,0.75){\vector(0,1){0.85}}
	  \labeledtreenode{11.5,0.5}{animal}{lexcolor}

	  \put(11.5,2.25){\vector(-4,1){3.25}}
	  \labeledtreenode{11.5,1.9}{animal}{compositioncolor}

	\end{picture}
	%}
        \caption{The model structure used to compare \ii{turtle} and
          \ii{animal}.  Learned term representations are fed through
          either an NN or NTN comparison layer and then to a softmax
        classifier over the seven relations in Table~\ref{b-table}.}
  \label{sample-figure}
\end{figure}

To compute the vector representation for a sentence, we populate the bottom row of the structure $\vec{y}^{(N,1...N)}$ with the embedddings of each of the words. We then compute the representations of each row starting at the second row from the bottom ($N - 1$), and moving from left to right. Computing one feature vector requires two steps. In the first step, a classifier produces a distribution $\vec{o}$ over the operations \{\textsc{copy-left, copy-right, compose}\}, and in the second step, that distribution is used to compute a feature vector $\vec{y}$.

To compute the distribution $\vec{o{}}$

Choose an operation

row i
col j
amount of context C

To compute the vector representation for a sentence

\begin{equation} 
\label{TreeRNN}
\vec{o}^{(i,j)} = \text{softmax}(\mathbf{M_o} \colvec{7}
	{\vec{y}^{(i + 1, j - C + 1)}}
	{...}
	{\vec{y}^{(i + 1, j)}}
	{\vec{y}^{(i + 1, j + 1)}}
	{...}
	{\vec{y}^{(i + 1, j + C)}}
	{\vec{o}^{(i, j - 1)}}
	 + \vec{b}_o\,) \\
\end{equation}

\begin{equation}
\vec{y}^{(i,j)} = 
o^{(i, j)}_{1} \circ \vec{y^{(i + 1, j)}} +\\
o^{(i, j)}_{2} \circ \vec{y^{(i + 1, j + 1)}} +\\
o^{(i, j)}_{3} \circ \tanh(\mathbf{M_y} \colvec{2}{\vec{y}^{(i + 1, j)}}{\vec{y}^{(i + 1, j + 1)}} + \vec{b}_y)
\end{equation}

For true relation $\rho$, and true operations $\omega^{(i,j)}$ for one example

\begin{equation}
loss = \lambda\theta^2-\ln(r_\rho) - \frac{1}{(N - 1)^2} \sum_{i = 1}^{N - 1} \sum_{j = 1}^{N - 1} \ln(o_{\omega^{(i,j)}})
\end{equation}

### 2015-04-04 run.sh

@@ -21,4 +21,4 @@ cd $PBS_O_WORKDIR
 echo `hostname`: $MATLABCMD 
 echo `hostname` - $PBS_JOBID - $MATLABCMD >> ~/machine_assignments.txt
 
-echo $MATLABCMD | /afs/cs/software/bin/matlab_r2012b -nodisplay
+echo $MATLABCMD | /afs/cs/software/bin/matlab_r2014b -nodisplay

### 2015-04-06 ComputeCostAndGrad.m
Support for delta clipping and cost scaling in pyr.

@@ -74,7 +74,7 @@ if nargout > 1 && (nargin < 6 || computeGradient)
     
     [ localSoftmaxGradient, softmaxDelta ] = ...
         ComputeSoftmaxClassificationGradients( ...
-          softmaxMatrix, relationProbs, max(trueRelation), ...
+          softmaxMatrix, relationProbs, trueRelation', ...
           extraInputs(:,hyperParams.topDepth), hyperParams);
     
     % Compute gradients for extra top layers

### 2015-04-06 InitializeModel.m
Support for delta clipping and cost scaling in pyr.

@@ -23,7 +23,7 @@ else
 end
 
 % Randomly initialize softmax layer
-softmaxMatrix = InitializeNNLayer(PENULT, sum(hyperParams.numRelations), 1, hyperParams.NNinitType);
+softmaxMatrix = InitializeNNLayer(PENULT, sum(hyperParams.numRelations), 1, 1);
 
 mergeMatrix = InitializeNNLayer(DIM * 2, PENULT, 1, hyperParams.NNinitType);
 

### 2015-04-06 layer-fns/ComputeSoftmaxClassificationGradients.m

@@ -33,6 +33,12 @@ if size(labels, 2) == 2
 		end
 		deltaDown(:, b) = matrix(relationRange, :)' * delta;
 	end
+
+	assert(sum(isinf(deltaDown(:))) == 0, ['Infs in classification delta.']); 
+	assert(sum(isnan(deltaDown(:))) == 0, ['NANs in classification delta.']); 
+
+	assert(sum(isinf(matrixGradients(:))) == 0, ['Infs in classification grad.']); 
+	assert(sum(isnan(matrixGradients(:))) == 0, ['NANs in classification grad.']); 
 else
 	delta = probs - targetprobs;
 	if nargin > 5
@@ -42,10 +48,16 @@ else
 
 	matrixGradients = delta * inPadded';
 	deltaDown = matrix' * delta;
-end
 
+	assert(sum(isinf(deltaDown(:))) == 0, ['Infs in classification delta.']); 
+	assert(sum(isnan(deltaDown(:))) == 0, ['NANs in classification delta.']); %%
+
+	assert(sum(isinf(matrixGradients(:))) == 0, ['Infs in classification grad.']); 
+	assert(sum(isnan(matrixGradients(:))) == 0, ['NANs in classification grad.']); 
+end
 
 % Remove bias deltas.
 deltaDown = deltaDown(2:end, :);
+
 
 end

### 2015-04-06 ComputeFullPyramidBatchCostAndGrad.m
Optimizations, mostly around NL gradients

@@ -121,7 +121,7 @@ if computeGrad
             MergeDeltaLeft, MergeDeltaRight ] = ...
           ComputeTensorLayerGradients(leftFeatures, rightFeatures, ...
               mergeMatrices, mergeMatrix, ...
-              extraDelta, hyperParams.classNLDeriv, tensorInnerOutput);
+              extraDelta, hyperParams.classNLDeriv, mergeOutput);
           localMergeMatricesGradients = sum(localMergeMatricesGradients, 4);
     else
          % Compute gradients for the merge NN layer
@@ -129,7 +129,7 @@ if computeGrad
          [localMergeMatrixGradients, ...
             MergeDeltaLeft, MergeDeltaRight] = ...
           ComputeRNNLayerGradients(leftFeatures, rightFeatures, ...
-              mergeMatrix, extraDelta, hyperParams.classNLDeriv, innerOutput);
+              mergeMatrix, extraDelta, hyperParams.classNLDeriv, mergeOutput);
 
           % Accumulate across batches.
     end

### 2015-04-06 PyramidBatch.m

@@ -9,7 +9,6 @@ classdef PyramidBatch < handle
         B = -1;  % Batch size.
         D = -1;  % Number of dimensions in the feature vector at each position.
         N = -1;  % The length of the longest sequence that can be handled within this pyramid.
-        R = -1;  % The number of rows in features and deltas.
         pyramids = [];  % The pyramid objects in the batch.
         wordIndices = [];  % The index into the embedding matrix for each word in the sequence.
         wordCounts = [];  % The number of words in each sequence.
@@ -20,8 +19,6 @@ classdef PyramidBatch < handle
                                          % the pre-nonlinearity activations from the embedding transform layer
                                          % if one is present.
         masks = [];  % Same structure as transformInnerActivations, but contains dropout masks for the embedding transform layer.
-        compositionInnerActivations = [];  % Same structure as features, but contains the pre-nonlinearity
-                                           % activations from the composition function, and has no bottom (word) layer.
         compositionActivations = [];  % Same structure as features, but contains the
                                       % activations from the composition function, and has no bottom (word) layer.
         connectionClassifierInputs = [];  % Inputs to the connection classifier, not computed separately, but stored to minimize runtime.
@@ -55,7 +52,6 @@ classdef PyramidBatch < handle
             pb.rawEmbeddings = zeros(hyperParams.embeddingDim, pb.B, hyperParams.embeddingTransformDepth * pb.N);
             pb.transformInnerActivations = zeros(pb.D, pb.B, hyperParams.embeddingTransformDepth * pb.N);
             pb.masks = zeros(pb.D, pb.B, hyperParams.embeddingTransformDepth * pb.N);
-            pb.compositionInnerActivations = zeros(pb.D, pb.B, pb.N - 1, pb.N - 1);
             pb.compositionActivations = zeros(pb.D, pb.B, pb.N - 1, pb.N - 1);
             pb.connectionClassifierInputs = zeros((2 * hyperParams.pyramidConnectionContextWidth) * pb.D + pb.NUMACTIONS, pb.B, pb.N - 1, pb.N - 1);
             pb.connections = zeros(pb.NUMACTIONS, pb.B, pb.N - 1, pb.N - 1);
@@ -118,8 +114,7 @@ classdef PyramidBatch < handle
 
                     % Build the composed representation
[…]

### 2015-04-06 Tree.m

@@ -11,7 +11,6 @@ classdef Tree < handle
         text = 'NO_TEXT';
         features = []; % DIM x 1 vector
         mask = []; % Used in dropout
-        featuresPreNL = [];
         wordIndex = -1; % -1 => Not a lexical item node.
         transformInnerActivations = []; % Stored activations for the embedding tranform layers.
         unknown = 0;
@@ -25,7 +24,6 @@ classdef Tree < handle
         function t = printAllProperties(obj)
             disp(obj.text)
             disp(obj.features)
-            disp(obj.featuresPreNL)
             disp(obj.wordIndex)
             disp(obj.transformInnerActivations)
             disp(obj.type)
@@ -184,11 +182,11 @@ classdef Tree < handle
                 end
                
                 if length(compMatrices) ~= 0
-                    [obj.features, obj.featuresPreNL] = ComputeTensorLayer(...
+                    obj.features = ComputeTensorLayer(...
                         lFeatures, rFeatures, compMatrices(:,:,:,typeInd),...
                         compMatrix(:,:,typeInd), compNL);
                 elseif length(compMatrix) ~= 0
-                    [obj.features, obj.featuresPreNL] = ComputeRNNLayer(lFeatures, rFeatures,...
+                    obj.features = ComputeRNNLayer(lFeatures, rFeatures,...
                         compMatrix(:,:,typeInd), compNL);
                 else
                     obj.features = ComputeSummingLayer(lFeatures, rFeatures);
@@ -259,7 +257,7 @@ classdef Tree < handle
                     ComputeTensorLayerGradients(lFeatures, rFeatures, ...
                           compMatrices(:,:,:,typeInd), ...
                           compMatrix(:,:,typeInd), delta, ...
-                          compNLDeriv, obj.featuresPreNL);
+                          compNLDeriv, obj.features);
 
                     upwardCompositionMatricesGradients(:,:,:,typeInd) = ...
                         tempCompositionMatricesGradients;
@@ -270,7 +268,7 @@ classdef Tree < handle
                         compDeltaRight] = ...
                     ComputeRNNLayerGradients(lFeatures, rFeatures, ...
                           compMatrix(:,:,typeInd), delta, ...
-                          compNLDeriv, obj.featuresPreNL);
+                          compNLDeriv, obj.features);
 
                     upwardCompositionMatrixGradients(:,:,typeInd) = ...
                         tempCompositionMatrixGradients;

### 2015-04-06 layer-fns/ComputeRNNLayerGradients.m

@@ -1,18 +1,19 @@
 % Want to distribute this code? Have other questions? -> sbowman@stanford.edu
 function [matrixGradients, deltaLeft, deltaRight] = ...
-      ComputeRNNLayerGradients(l, r, matrix, delta, ...
-                                  nonlinearityDeriv, innerOutput)
+      ComputeRNNLayerGradients(l, r, matrix, delta, nonlinearityDeriv, output)
 % Compute the gradients and deltas for an RNN layer for a given batch.
 
 in = [ones(1, size(l, 2)); l; r];
 
 % TODO: Make this happen less often.
-if nargin < 7
+if nargin < 6
     innerOutput = matrix * in;
+    NLDeriv = nonlinearityDeriv(innerOutput);
+else
+	NLDeriv = nonlinearityDeriv([], output);
 end
 
 % TODO: Efficient NLDeriv
-NLDeriv = nonlinearityDeriv(innerOutput);
 delta = NLDeriv .* delta;
 
 % Compute the matrix gradients

### 2015-04-06 layer-fns/ComputeSummingLayerGradients.m
Optimizations, mostly around NL gradients

@@ -1,6 +1,5 @@
 % Want to distribute this code? Have other questions? -> sbowman@stanford.edu
-function [deltaLeft, deltaRight] = ...
-      ComputeSummingLayerGradients(delta)
+function [deltaLeft, deltaRight] = ComputeSummingLayerGradients(delta)
 % Compute the gradients and deltas for a summing layer for a given example.
 
 deltaRight = delta;

### 2015-04-06 layer-fns/TanhDeriv.m

@@ -1,9 +1,13 @@
 % Want to distribute this code? Have other questions? -> sbowman@stanford.edu
-function outVec = TanhDeriv (inVec)
+function deriv = TanhDeriv (in, out)
 % Compute the gradient of the tanh nonlinearity.
 
 % TODO: Redefine in terms of the output vector -- faster.
- 
-outVec = sech(inVec).^2;
+
+if nargin > 1
+ 	deriv = 1 - out.^2;
+else
+	deriv = sech(in).^2;
+end
 
 end

### 2015-04-16 SequenceBatch.m
Pyramid -> Lattice

@@ -74,7 +74,7 @@ classdef SequenceBatch < handle
 
             for w = 1:sb.N
                 if ~isempty(embeddingTransformMatrix)
-                    % TODO: Adapt this zero-bias idea to PyramidBatch.
+                    % TODO: Adapt this zero-bias idea to LatticeBatch.
                     transformInputs = [ [w >= sb.N - sb.wordCounts + 1]'; sb.rawEmbeddings(:, :, w)];
                     [ sb.inputFeatures(:, :, w), sb.masks(:, :, w) ] = ...
                         Dropout(tanh(embeddingTransformMatrix * transformInputs), hyperParams.bottomDropout, trainingMode);

### 2015-04-16 writing/S'15 misc/newmodel.tex
Optimizations, mostly around NL gradients
Pyramid -> Lattice

@@
-\section{Jointly learning parsing and semantic composition}
+\section{Jointly learning syntactic parsing and semantic composition}
 
-% TODO: Graph figure
+\subsection{Why joint learning?}
+
+% tree and sequences have both been competitive, tree structures useful in some domains, but in many cases they are superceded by strong sequence models.
+
+% our hypothesis: (i) natural language has compositional structure that can be exploited by representation learning models (ii) there is no guarantee that standard syntactic formalisms are especially well suited for this purpose
+
+% new model, jointly learn syn and sem
+
+\subsection{Model definition}
 
@@
   \centering\small
-  \begin{tikzpicture}
-    \def\dx{18pt}
-    \def\dy{30pt}
-    \newcounter{i}
-    \tikzstyle{highlight}=[fill=dandelion!15]
-    \tikzstyle{comment}=[fill=blue!15]
+ 	\input{lattice_fig.tex}
 
-    \stepcounter{i}\node  (\arabic{i}) at (0*\dx,6*\dy) {$\vec{y}^{(1, 1)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (-1*\dx,5*\dy) {$\vec{y}^{(2, 1)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (1*\dx,5*\dy) {$\vec{y}^{(2, 2)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (-2*\dx,4*\dy) {$\vec{y}^{(3, 1)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (0*\dx,4*\dy) {$\vec{y}^{(3, 2)}$};
-    \stepcounter{i}\node[highlight]  (\arabic{i}) at (2*\dx,4*\dy) {$\vec{y}^{(3, 3)}$};
-    \stepcounter{i}\node[highlight]  (\arabic{i}) at (-3*\dx,3*\dy) {$\vec{y}^{(4, 1)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (-1*\dx,3*\dy) {$\vec{y}^{(4, 2)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (1*\dx,3*\dy) {$\vec{y}^{(4, 3)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (3*\dx,3*\dy) {$\vec{y}^{(4, 4)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (-4*\dx,2*\dy) {$\vec{y}^{(5, 1)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (-2*\dx,2*\dy) {$\vec{y}^{(5, 2)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (0*\dx,2*\dy) {$\vec{y}^{(5, 3)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (2*\dx,2*\dy) {$\vec{y}^{(5, 4)}$};
-    \stepcounter{i}\node[highlight]  (\arabic{i}) at (4*\dx,2*\dy) {$\vec{y}^{(5, 5)}$};
-    \stepcounter{i}\node  (\arabic{i}) at (-5*\dx,1*\dy) {My};
-    \stepcounter{i}\node  (\arabic{i}) at (-3*\dx,1*\dy) {hovercraft};
-    \stepcounter{i}\node  (\arabic{i}) at (-1*\dx,1*\dy) {is};
-    \stepcounter{i}\node  (\arabic{i}) at (1*\dx,1*\dy) {full};
-    \stepcounter{i}\node  (\arabic{i}) at (3*\dx,1*\dy) {of};
-    \stepcounter{i}\node  (\arabic{i}) at (5*\dx,1*\dy) {eels};
-
-    \stepcounter{i}\node[comment]  (\arabic{i}) at (4*\dx,4*\dy) {\textsc{compose}};
-    \stepcounter{i}\node[comment]  (\arabic{i}) at (-5*\dx,3*\dy) {\textsc{copy-L}};
-    \stepcounter{i}\node[comment]  (\arabic{i}) at (6*\dx,2*\dy) {\textsc{copy-R}};
-
-    \tikzstyle{extra} = [->, draw=black, line width=.75pt]
-    \tikzstyle{valid} = [->, draw=red, line width=1.25pt]
-
-          \draw [valid] (2) -- (1);
-          \draw [valid] (3) -- (1);
-
-          \draw [valid] (4) -- (2);
-          \draw [extra] (5) -- (2);
-          \draw [valid] (5) -- (3);
-          \draw [valid] (6) -- (3);
-
-          \draw [valid] (7) -- (4);
-          \draw [extra] (8) -- (4);
-          \draw [valid] (8) -- (5);
-          \draw [extra] (9) -- (5);
-          \draw [valid] (9) -- (6);
-          \draw [valid] (10) -- (6);
-
-          \draw [valid] (11) -- (7);
-          \draw [extra] (12) -- (7);
-          \draw [valid] (12) -- (8);
-          \draw [extra] (13) -- (8);
-          \draw [valid] (13) -- (9);
-          \draw [extra] (14) -- (9);
-          \draw [valid] (14) -- (10);
-          \draw [valid] (15) -- (10);
-
-          \draw [valid] (16) -- (11);
-          \draw [valid] (17) -- (11);
-          \draw [extra] (17) -- (12);
-          \draw [valid] (18) -- (12);
-          \draw [extra] (18) -- (13);
-          \draw [valid] (19) -- (13);
-          \draw [extra] (19) -- (14);
-          \draw [valid] (20) -- (14);
-          \draw [extra] (20) -- (15);
-          \draw [valid] (21) -- (15);
-  \end{tikzpicture}
-
-        \caption{The model structure used to compare \ii{turtle} and
-          \ii{animal}.  Learned term representations are fed through
-          either an NN or NTN comparison layer and then to a softmax
-        classifier over the seven relations in Table~\ref{b-table}.}
-  \label{sample-figure2}
-\end{figure}
-
-\begin{figure}[tp]
-  \centering
-    \footnotesize
-
-	\newcommand{\labeledtreenode}[4][3]{\put(#2){\makebox(0,0){{\fcolorbox{black}{#4}{\makebox(#1,0.3){#3}}}}}}
-
-	\newcommand{\textlabel}[4][3.5]{\put(#2){\makebox(0,0){{\makebox(#1,0.3){#3}}}}}
-
-	\definecolor{lexcolor}{HTML}{F5F7C4}
-	\definecolor{compositioncolor}{HTML}{BBEBFF}
-	\definecolor{comparisoncolor}{HTML}{FFC895}
-	\definecolor{softmaxcolor}{HTML}{A5FF8A}
-
-	\setlength{\unitlength}{0.61cm}
-
-	%\resizebox{3in}{!}{
-	\begin{picture}(13.15, 5.5)
-
-	  \labeledtreenode[2.4]{6.5,5}{$P(\sqsubset) = 0.8$}{softmaxcolor}  
-	  \put(6.5,3.7){\vector(0,1){1}}  
-	  \labeledtreenode[5]{6.5,3.4}{turtle \emph{vs.}~animal}{comparisoncolor}
-
-	  \textlabel{2.75,5}{Softmax classifier}{black}
-	  \textlabel{2.0,3.4}{\parbox{2cm}{Comparison N(T)N layer}}{black}
-
-	  \textlabel{6.5,2.1}{Optional embedding}{black}
-	  \textlabel{6.5,1.5}{transformation NN layers}{black}
-
-	  \textlabel{6.5,0.5}{Learned term vectors}{black}
-
-	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-	  \put(1.5,0.75){\vector(0,1){0.85}}
-	  \labeledtreenode{1.5,0.5}{turtle}{lexcolor}
-
-	  \put(1.5,2.25){\vector(4,1){3.25}}
-	  \labeledtreenode{1.5,1.9}{turtle}{compositioncolor}
-
-	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-	  \put(11.5,0.75){\vector(0,1){0.85}}
-	  \labeledtreenode{11.5,0.5}{animal}{lexcolor}
-
-	  \put(11.5,2.25){\vector(-4,1){3.25}}
-	  \labeledtreenode{11.5,1.9}{animal}{compositioncolor}
-
-	\end{picture}
-	%}
-        \caption{The model structure used to compare \ii{turtle} and
-          \ii{animal}.  Learned term representations are fed through
-          either an NN or NTN comparison layer and then to a softmax
-        classifier over the seven relations in Table~\ref{b-table}.}
-  \label{sample-figure}
+        \caption{The connection structure of a LatticeNN. The correct binary parse tree for the example sentence, projected onto a lattice with the left-first rule, is shown in the bolded arrows.}
+  \label{lattice-fig1}
 \end{figure}
@@
 Computing one feature vector requires two steps.
-In the first step, a classifier produces a distribution $\vec{o}$ over the operations \{\textsc{copy-left, copy-right, compose}\}, and in the second step, that distribution is used to compute a feature vector $\vec{y}$.
+In the first step, we produce a distribution $\vec{o}$ over the operations \{\textsc{copy-L, copy-R, compose}\}, and in the second step, that distribution is used to compute a feature vector $\vec{y}$.
 
-To compute the distribution $\vec{o{}}$
-
-Choose an operation
-
-row i
-col j
-amount of context C
-
-To compute the vector representation for a sentence
+At each row $i$, we begin by computing a scalar score $s$ for each node $(i, j)$ representing how appropriate it would be to compose the two child nodes $(i+1, j)$ and $(i+1, j+1)$ at $(i, j)$.
+This decision is made primarily on the basis of the two child nodes, and a $C$-word window of context around them that is meant to allow the model to do a limited amount of ambiguity resolution.
+When the context window extends off the edge of the lattice in either direction, a single learned padding vector is used.
+% TODO
+Including $j$ as an additional feature here is intended to allow the model to learn to use linear order in making its scoring decisions, allowing it to prefer to compose pairs of constituens that are closer to the left edge of the tree when all else is equal, and encouraging the model to conform to our left-first rule.
 
 \begin{equation}
-\label{TreeRNN}
-\vec{o}^{(i,j)} = \text{softmax}(\mathbf{M_o} \colvec{7}
+\label{scoringeqn}
+s^{(i,j)} = \vec{v}_s \colvec{7}
 	{\vec{y}^{(i + 1, j - C + 1)}}
@@
 	{\vec{y}^{(i + 1, j + C)}}
-	{\vec{o}^{(i, j - 1)}}
-	 + \vec{b}_o\,) \\
+	{j}
+	 + b_s \\
 \end{equation}
+%
+Since we wish for the model to compose only one pair of child nodes at each row $i$, we apply a softmax function to the full vector of scores computed at each row, and use the resulting values (which sum to one) to decide which nodes to compose.
+%
+\begin{equation} 
+o^{(i,\,...)}_{\textsc{compose}} = \text{softmax}(\colvec{4}
+	{s^{(i, 1)}}
+	{s^{(i, 2)}}
+	{...}
+	{s^{(i, i)}}	
+	) \\
+\end{equation}
+%
+At training time, optional secondary parsing supervision can be used in the form of a label $\kappa^{(i)}$ indicating the location of the merge at layer $i$ of the left-first projection of the reference parse.
 
+Then, for each node in the current row, we use these modified scores to compute another set of values summing to one, this time representing the fraction of the activation of the current node that should be drawn from each of the three operations:
+%
+\begin{gather}
+o^{(i, j)}_{\textsc{copy-L}} = \sum_{x = j + 1}^{i} o^{(i,x)}_{\textsc{compose}} \\
+o^{(i, j)}_{\textsc{copy-R}} = \sum_{x = 1}^{j - 1} o^{(i,x)}_{\textsc{compose}}
+\end{gather}
+%
+The use of the sum for the two \textsc{copy} weights reflects the intuition that if a pair of nodes will be composed to the left of the current node, the current node must \textsc{copy-R}, and vice versa, in order to preserve the left-first tree projection structure.
+
+% TODO: What guarantees can we make about how much of the activation from each layer will be propagated to the next layer?
+
+Finally, using these three computed weights, we compute the activations for the current node, using a standard RNN activation function for the \textsc{compose} operation:
+%
+
+\begin{eqnarray}
+&& \vec{y}^{(i,j)} =\\
+&&\nonumber  o^{(i, j)}_{\textsc{copy-L}} \circ \vec{y}^{(i + 1, j)} +\\
+&&\nonumber o^{(i, j)}_{\textsc{copy-R}} \circ \vec{y}^{(i + 1, j + 1)} +\\
+&&\nonumber o^{(i, j)}_{\textsc{compose}} \circ \tanh(\mathbf{M_y} \colvec{2}{\vec{y}^{(i + 1, j)}}{\vec{y}^{(i + 1, j + 1)}} + \vec{b}_y)
+\end{eqnarray}
+%
+Once all of the activations in the lattice have been computed, the top activation $\vec{y}^{(1,1)}$ can be treated as the embedding vector for the full sentence.
+For the sentiment classification task we study here, we pass it into a softmax classifier layer to predcit the probability of each sentiment label:
+%
 \begin{equation}
-\vec{y}^{(i,j)} = 
-o^{(i, j)}_{1} \circ \vec{y^{(i + 1, j)}} +\\
-o^{(i, j)}_{2} \circ \vec{y^{(i + 1, j + 1)}} +\\
-o^{(i, j)}_{3} \circ \tanh(\mathbf{M_y} \colvec{2}{\vec{y}^{(i + 1, j)}}{\vec{y}^{(i + 1, j + 1)}} + \vec{b}_y)
+\vec{l} = \text{softmax}(\mathbf{M_l}\,\vec{y}^{(1, 1)} + \vec{b}_l)
 \end{equation}
+%
+The loss for the example is then defined as the sum of three parts, a log likelihood term for the predicted sentiment class label, a sum of log likelihood terms for the correct composition location decisions (scaled by the number of ``middle'' layers in the lattice for which a composition location choice must be made), and an L2 regularization term.
+For sentiment relation label $\rho$, composition loctation label vector $\vec{\kappa}$, concatenated parameter vector $\theta$, and regularization strength hyperparameter $\lambda$ the loss is:
+%
 
-For true relation $\rho$, and true operations $\omega^{(i,j)}$ for one example
+\begin{eqnarray}
+loss & = & \lambda\theta^2 - \ln(l_\rho) -\\
+\nonumber&&\frac{1}{N - 1} \sum_{i = 1}^{N - 2} \ln(o^{(i,\kappa^{(i)})}_{\textsc{compose}})
+\end{eqnarray}
+%
+Parameter gradients can then be computed using standard backpropagation through structure \cite{goller1996learning}.
+Notably, when the composition location labels $\vec{\kappa}$, are used, the corresponding error signal at each row is used not only to train the scoring parameters $\vec{v}_s$, but also propagates error through the input vector used in eq.
+\ref{scoringeqn} and into the rest of the model.
 
-\begin{equation}
-loss = \lambda\theta^2-\ln(r_\rho) - \frac{1}{(N - 1)^2} \sum_{i = 1}^{N - 1} \sum_{j = 1}^{N - 1} \ln(o_{\omega^{(i,j)}})
-\end{equation}
+Batching
+
+% Possible supervision backoff proposal

### 2015-04-16 LoadEntailmentData.m

@@ -4,8 +4,8 @@ function data = LoadEntailmentData(filename, wordMap, relationMap, hyperParams, 
 
 if hyperParams.useTrees
     typeSig = '-trees';
-elseif hyperParams.usePyramids
-    typeSig = '-pyrs';
+elseif hyperParams.useLattices
+    typeSig = '-lats';
 else
     typeSig = ['-seqs-par' num2str(hyperParams.parensInSequences)];
 end
@@ -90,11 +90,11 @@ function [ data ] = ProcessAndSave(rawData, wordMap, lastSave, nextItemNo, filen
             data(dataInd).right = Tree.makeTree(rawData(dataInd).rightText, wordMap);
             data(dataInd).relation = rawData(dataInd).relation;
         end
-    elseif hyperParams.usePyramids
-        data = repmat(struct('relation', 0, 'left', Pyramid(), 'right', Pyramid()), numElements, 1);
+    elseif hyperParams.useLattices
+        data = repmat(struct('relation', 0, 'left', Lattice(), 'right', Lattice()), numElements, 1);
         parfor dataInd = 1:numElements
-            data(dataInd).left = Pyramid.makePyramid(rawData(dataInd).leftText, wordMap);
-            data(dataInd).right = Pyramid.makePyramid(rawData(dataInd).rightText, wordMap);
+            data(dataInd).left = Lattice.makeLattice(rawData(dataInd).leftText, wordMap);
+            data(dataInd).right = Lattice.makeLattice(rawData(dataInd).rightText, wordMap);
             data(dataInd).relation = rawData(dataInd).relation;
         end
     else

### 2015-04-16 config/SNLI.m

@@ -81,14 +81,14 @@ elseif composition == 3
   hyperParams.useThirdOrderMerge = 1;
   hyperParams.parensInSequences = 0;
 elseif composition == 4
-  hyperParams.usePyramids = 1;
+  hyperParams.useLattices = 1;
   hyperParams.lstm = 0;
   hyperParams.useTrees = 0;
   hyperParams.useThirdOrderComposition = 0;
   hyperParams.useThirdOrderMerge = 0;
   hyperParams.parensInSequences = 0;
 elseif composition == 5
-  hyperParams.usePyramids = 1;
+  hyperParams.useLattices = 1;
   hyperParams.lstm = 0;
   hyperParams.useTrees = 0;
   hyperParams.useThirdOrderComposition = 0;
@@ -141,7 +141,7 @@ elseif findstr(dataflag, 'snli095-only')
     hyperParams.testRelationIndices = [1];
     hyperParams.trainingMultipliers = [1];
 
-elseif findstr(dataflag, 'snli095HOL-only')
+elseif findstr(dataflag, 'snli095short-only')
     wordMap = InitializeMaps('./sick-data/sick-snli0.95_words.txt');
     hyperParams.vocabName = 'ss095'; 
 
@@ -151,9 +151,9 @@ elseif findstr(dataflag, 'snli095HOL-only')
     relationMap = cell(1, 1);
     relationMap{1} = containers.Map(hyperParams.relations{1}, 1:length(hyperParams.relations{1}));
 
-    hyperParams.trainFilenames = {'../data/snli_0.95_train_parsed_HOL.txt'};    
+    hyperParams.trainFilenames = {'../data/snli_0.95_train_words_parsed_short.txt'};    
     hyperParams.splitFilenames = {};    
-    hyperParams.testFilenames = {'../data/snli_0.95_dev_parsed_HOL.txt'};
+    hyperParams.testFilenames = {'../data/snli_0.95_dev_words_parsed_short.txt'};
 
     hyperParams.relationIndices = [1; 1; 1];
     hyperParams.testRelationIndices = [1];

### 2015-04-18 writing/S'15 misc/newmodel.tex
j => j/i
Writeup update.

@@
 
-% tree and sequences have both been competitive, tree structures useful in some domains, but in many cases they are superceded by strong sequence models.
+Sentence embedding methods like LSTMs and tree-structured RNNs have have rapidly become a standard tool in natural language processing tasks involving sentence meaning (TODO: Cite).
+These methods are generally built up out of standard neural network components, making it possible to learn them with minimal domain knowledge.
+They are also generally compositional, in that they build up their representations incrementally in a way that involves creating explicit representations for pieces of sentences.
+There is not yet any single model that is known to produce high-quality representations across domains, and one of the major open issues in the development of these models is the degree to which syntactic tree structures provide them with usable information that could not be efficiently extracted without them.
 
-% our hypothesis: (i) natural language has compositional structure that can be exploited by representation learning models (ii) there is no guarantee that standard syntactic formalisms are especially well suited for this purpose
-
-% new model, jointly learn syn and sem
-
-\subsection{Model definition}
+Our hypothesis on this issue is in two parts.
+(i) As is near-universally accepted in the language sciences, central components of natural language meaning are constructed compositionally according to regular tree structures in which each node in the tree structure for a sentence has an interpretable meaning derivable from the meanings of its children.
+Representation learning systems can learn more effectively by taking advantage of this fact.
+(ii) The syntactic structures used by typical software parsers do not reliably match the true tree structures underlying the construction of meaning for several reasons, even when those parsers are functioning correctly.
+In light of these two points, we propose a class of models which simultaneously learn to parse natural language sentences and to create compositional semantic representations using those parse structures, and can do this using supervision from both a downstream semantic task like sentiment classification and, when appropriate a syntactic parser.
+Such a model then serve both as a parser which is tuned so as to be optimally supportive of the construction of sentence meaning, as a function to construct representations of sentence meaning.
 
@@
 \end{figure}
+
+If our hypothesis is correct, and our models succeed at their goals, two important objectives will be served.
+The first is an engineering objective: our sentence embedding model will exploit compositional structure in the most effective way possible, producing better sentence representations that would have been possible otherwise.
+The second is a scientific objective: by learning a syntax optimal for semantic composition, we have a new source of information about humans might take advantage of syntactic structure in understanding natural language.
+This research is becoming possible only now, as it depends on both the tools (both software and hardware) for effectively training large neural networks that are only now becoming available, as well as on the availability of large volumes of natural language sentences annotated with explicit information about sentence meaning that are only now being collected.
+
+\paragraph{What's wrong with existing parsers?} One source of error is obvious: the parsers that are typically used to generate constituency or dependency structures for sentences are built around statistical machine learning systems trained on data that reflects only a fraction of the massive diversity of natural language, and a certain amount of error from both underfitting and overfitting is essentially unavoidable in this context.
+This is not, at least primarily, the kind of error that we aim to avoid.
+Our interest is in two other sources of error: cases of \textbf{syntax-semantics mismatch} where even the correct syntactic parse under the relevant formalism does not express the way in which meanings are constructed, and cases of \textbf{syntactic ambiguity} where there are multiple syntactically correct tree structures, and where choosing only one of them would mask a genuine ambiguity in meaning.
+
+% TODO: Example constructions for each!
+
+\subsection{Parsing in lattices}
+
+In this section, we propose a simple framework for syntactic parsing that lends itself joint learning.
+It is centered around a triangular lattice structure of the kind shown in Fig.~\ref{lattice-fig1}.
+In the lattice, every node in the bottom row contains a vector representation for a single word, the top node contains a vector representation of the whole sentence (of the same dimension), and every intermediate node contains a vector representation of some subset of the sentence.
+
+\begin{figure}[tp]
+  \centering\small
+ 	\input{lattice_fig_bad.tex}
+
+        \caption{The correct binary parse tree for the example sentence, projected onto a lattice \textit{without} the left-first rule.}
+  \label{lattice-fig1}
+\end{figure}
+
+\subsection{Model definition}
 
@@
 % TODO
-Including $j$ as an additional feature here is intended to allow the model to learn to use linear order in making its scoring decisions, allowing it to prefer to compose pairs of constituens that are closer to the left edge of the tree when all else is equal, and encouraging the model to conform to our left-first rule.
+Including $\frac{j}{i}$ as an additional feature here is intended to allow the model to learn to use linear order in making its scoring decisions, allowing it to prefer to compose pairs of constituens that are closer to the left edge of the tree when all else is equal, and encouraging the model to conform to our left-first rule.
 
@@
 	{\vec{y}^{(i + 1, j + C)}}
-	{j}
+	{\frac{j}{i}}
 	 + b_s \\
@@
 % Possible supervision backoff proposal
+
+\subsection{Related work}

### 2015-04-27 writing/S'15 misc/newmodel.tex
Paper update, start to TreeLSTM.

@@
 We then compute the representations of each row starting at the second row from the bottom ($N - 1$), and moving from left to right.
-Computing one feature vector requires two steps.
+Computing one feature vector requires two stages.
 In the first step, we produce a distribution $\vec{o}$ over the operations \{\textsc{copy-L, copy-R, compose}\}, and in the second step, that distribution is used to compute a feature vector $\vec{y}$.
 
-At each row $i$, we begin by computing a scalar score $s$ for each node $(i, j)$ representing how appropriate it would be to compose the two child nodes $(i+1, j)$ and $(i+1, j+1)$ at $(i, j)$.
+At each row $i$, we begin by computing a scoring feature vector $\vec{s}$ for each node $(i, j)$ which is meant to contain features that allow the model to judge how appropriate it would be to compose the two child nodes $(i+1, j)$ and $(i+1, j+1)$ at $(i, j)$.
 This decision is made primarily on the basis of the two child nodes, and a $C$-word window of context around them that is meant to allow the model to do a limited amount of ambiguity resolution.
@@
 % TODO
-Including $\frac{j}{i}$ as an additional feature here is intended to allow the model to learn to use linear order in making its scoring decisions, allowing it to prefer to compose pairs of constituens that are closer to the left edge of the tree when all else is equal, and encouraging the model to conform to our left-first rule.
+Including $\frac{j}{i}$ and $j$ as additional features here is intended to allow the model to learn to use linear order in making its scoring decisions, allowing it to prefer to compose pairs of constituens that are closer to the left edge of the tree when all else is equal, and encouraging the model to conform to our left-first rule.
 
@@
 \label{scoringeqn}
-s^{(i,j)} = \vec{v}_s \colvec{7}
+\vec{s}^{(i,j)} = \tanh(\mathbf{M}_s \colvec{8}
 	{\vec{y}^{(i + 1, j - C + 1)}}
@@
 	{\frac{j}{i}}
-	 + b_s \\
+	{j}
+	  + \vec{b_s}) \\
 \end{equation}
 %
+To convert these feature vectors to single scores, we take their inner product with a parameter vector $\vec{v}^T_s$.
 Since we wish for the model to compose only one pair of child nodes at each row $i$, we apply a softmax function to the full vector of scores computed at each row, and use the resulting values (which sum to one) to decide which nodes to compose.
@@
 o^{(i,\,...)}_{\textsc{compose}} = \text{softmax}(\colvec{4}
-	{s^{(i, 1)}}
-	{s^{(i, 2)}}
+	{\vec{v}^T_s \vec{s}^{(i, 1)} }
+	{\vec{v}^T_s \vec{s}^{(i, 2)}}
 	{...}
-	{s^{(i, i)}}	
+	{\vec{v}^T_s \vec{s}^{(i, i)}}	
 	) \\
@@
 % 
+\begin{equation}
+\vec{h}^{(i, j)} = \tanh(\mathbf{M_y} \colvec{2}{\vec{y}^{(i + 1, j)}}{\vec{y}^{(i + 1, j + 1)}} + \vec{b}_y)
+\end{equation}
 
 \begin{eqnarray}
-&& \vec{y}^{(i,j)} =\\
-&&\nonumber  o^{(i, j)}_{\textsc{copy-L}} \circ \vec{y}^{(i + 1, j)} +\\
+\vec{y}^{(i,j)} & = & o^{(i, j)}_{\textsc{copy-L}} \circ \vec{y}^{(i + 1, j)} +\\
 &&\nonumber o^{(i, j)}_{\textsc{copy-R}} \circ \vec{y}^{(i + 1, j + 1)} +\\
-&&\nonumber o^{(i, j)}_{\textsc{compose}} \circ \tanh(\mathbf{M_y} \colvec{2}{\vec{y}^{(i + 1, j)}}{\vec{y}^{(i + 1, j + 1)}} + \vec{b}_y)
+&&\nonumber o^{(i, j)}_{\textsc{compose}} \circ \vec{h}^{(i, j)}
 \end{eqnarray}
@@
 
+\subsection{LSTM composition}
+
+The LSTM variant of our model uses a TreeLSTM composition function.
+We largely model this function after the model of \newcite{tai2015improved}, however this model requires a clear distinction between raw inputs ($x_j$) and child hidden states ($h_{j1}$, $h_{j2}$).
+Since our model's intermediate representations $\vec{y}^{i,j}$ are linear combinations of composition function outputs and raw features, this distinction is not available.
+We follow \newcite{le2015compositional}'s TreeLSTM formulation in eliminating the input terms from the LSTM equations and treating inputs and child features identically (as in a standard tree-structured network), yielding the following:
+
+% TODO: Set up the alignments here properly.
+\begin{gather} 
+\vec{i}^{(i,j)} = \text{sigmoid}(\mathbf{M_i}\colvec{2}{\vec{y}^{(i+1,j)}}{\vec{y}^{(i+1,j+1)}} + \vec{b_i}) \\
+\vec{f}^{(i,j)}_l = \text{sigmoid}(\mathbf{M_{fl}}\colvec{2}{\vec{y}^{(i+1,j)}}{\vec{y}^{(i+1,j+1)}} + \vec{b_{fl}}) \\
+\vec{f}^{(i,j)}_r = \text{sigmoid}(\mathbf{M_{fr}}\colvec{2}{\vec{y}^{(i+1,j)}}{\vec{y}^{(i+1,j+1)}} + \vec{b_{fr}}) \\
+\vec{o}^{(i,j)} = \text{sigmoid}(\mathbf{M_o}\colvec{2}{\vec{y}^{(i+1,j)}}{\vec{y}^{(i+1,j+1)}} + \vec{b_o}) \\
+\vec{g}^{(i,j)} = \text{tanh}(\mathbf{M_g}\colvec{2}{\vec{y}^{(i+1,j)}}{\vec{y}^{(i+1,j+1)}} + \vec{b_g})\\
+\vec{c}^{(i,j)} = \vec{f}^{(i,j)}_l \circ \vec{c}^{(i+1,j)} + \vec{f}^{(i,j)}_r \circ \vec{c}^{(i+1,j+1)}+\\
+\nonumber ~~~~~~~~~~~~~ \vec{i}^{(i,j)} \circ \vec{g}^{(i,j)}\\
+\vec{h}^{(i,j)} = \vec{c}^{(i,j)} \circ \vec{o}^{(i,j)}
+\end{gather}
+
+The output $\vec{h}^{(i,j)}$ is then incorporated into $\vec{y}^{(i,j)}$ according to the operation weights, as in the basic model.
+
 % Possible supervision backoff proposal

### 2015-04-27 RunExperiments/RunQuantifierExperiments.sh
Paper update, start to TreeLSTM.

@@ -78,5 +78,15 @@ export MATLABCMD="cd quant; lambda = 0.0001; dim = 50; td = 1; penult = 75; drop
 export MATLABCMD="cd quant; lambda = 0.0001; dim = 20; td = 1; penult = 75; dropout = 1; tot = 0; name='latcomp-b'; relu = 0; TrainModel(''\, 1\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, dropout);" ; qsub -v MATLABCMD quant/run.sh
 export MATLABCMD="cd quant; lambda = 0.0001; dim = 20; td = 1; penult = 75; dropout = 1; tot = 1; name='latcomp-b'; relu = 0; TrainModel(''\, 1\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, dropout);" ; qsub -v MATLABCMD quant/run.sh
 
+export MATLABCMD="cd quant-naacl; lambda = 0.0003; dim = 25; td = 1; penult = 75; dropout = 1; tot = 0; name='/scr/sbowman/vintage-quant'; relu = 1; TrainModel(''\, 1\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, 1\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh
+export MATLABCMD="cd quant-naacl; lambda = 0.0003; dim = 25; td = 1; penult = 75; dropout = 1; tot = 1; name='/scr/sbowman/vintage-quant'; relu = 1; TrainModel(''\, 1\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, 1\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh
+export MATLABCMD="cd quant-naacl; lambda = 0.001; dim = 25; td = 1; penult = 75; dropout = 1; tot = 0; name='/scr/sbowman/vintage-quant'; relu = 1; TrainModel(''\, 1\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, 1\, dropout\, 32);" ; qsub -v MATLABCMD quant/run.sh
[…]

### 2015-04-27 config/SST.m

@@ -5,7 +5,7 @@ function [ hyperParams, options, wordMap, relationMap ] = SST(expName, dataflag,
 
 % Generate an experiment name that includes all of the hyperparameter values that
 % are being tuned.
-hyperParams.name = [expName, '-', dataflag, '-l', num2str(lambda), '-dim', num2str(dim),...
+hyperParams.name = [expName, '-', dataflag, '-l', num2str(lambda), '-p', num2str(penult), '-dim', num2str(dim),...
     '-ed', num2str(embDim), '-td', num2str(topDepth),...
     '-do', num2str(bottomDropout), '-', num2str(topDropout), '-co', num2str(collo),...
     '-comp', num2str(composition), '-lattEv', num2str(latte), '-curr', num2str(curr), ...
@@ -114,7 +114,13 @@ if strcmp(dataflag, 'sst-expanded')
 
     % Loading this data is fast, and the preprocessed file winds up huge.
     hyperParams.ignorePreprocessedFiles = true;
+
     hyperParams.relationCostMultipliers = [4.878182632, 2.433623131, 0.3014847996, 1.826731877, 3.980980277];
+
+    if penult == 0
+        hyperParams.relationCostMultipliers = (hyperParams.relationCostMultipliers + [1 1 1 1 1]) ./ 2
+    end
+
 elseif strcmp(dataflag, 'sst')
     wordMap = InitializeMaps('./sst-data/sst-words.txt');
     hyperParams.vocabName = 'sst'; 

### 2015-04-29 writing/F'14 paper/quantifiers.tex
Lattice edge tokens; better configs; optimizations

@@
 
-In each run, we sample 20\% of the 60k generated pairs to test on and 
-train on the remainder. In this setting, each model must learn a
+In each run, we partition the full set of possible sentences under the grammar into training and test sentences, and label all of the pairs from within each set to generate a training set of 27k pairs and a test set of 7k pairs.
+Because the model doesn't see the test sentences at training time, it cannot directly use the kind of reasoning described in \S\ref{sec:join}, each model must learn a
 complete reasoning system for the limited language and logic presented

### 2015-04-29 RunExperiments/RunSSTExperiments.sh

@@ -31,12 +31,21 @@ export MATLABCMD="cd quant; lambda = 0.000001; dim = 50; ed = 200; td = 2; penul
 
 																																																																								latte, curr, adad, ccs)
 % To run:
-export MATLABCMD="cd quant; lambda = 0.000001; dim = 25; ed = 200; td = 2; penult = 75; dropout = [0.5\, 0.5]; tot = 4; collo = 1; dataflag='sst-expanded'; name='/scr/sbowman/sst-expanded-3'; TrainModel(''\, 1\, @SST\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1\, 1\, 1\, 1);" ; qsub -v MATLABCMD quant/run.sh  
-export MATLABCMD="cd quant; lambda = 0.000001; dim = 25; ed = 200; td = 2; penult = 75; dropout = [0.5\, 0.5]; tot = 4; collo = 1; dataflag='sst-expanded'; name='/scr/sbowman/sst-expanded-3'; TrainModel(''\, 1\, @SST\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1\, 0\, 1\, 1);" ; qsub -v MATLABCMD quant/run.sh  
-export MATLABCMD="cd quant; lambda = 0.000001; dim = 25; ed = 200; td = 2; penult = 75; dropout = [0.5\, 0.5]; tot = 4; collo = 1; dataflag='sst-expanded'; name='/scr/sbowman/sst-expanded-3'; TrainModel(''\, 1\, @SST\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 0\, 1\, 1\, 1);" ; qsub -v MATLABCMD quant/run.sh  
[…]

### 2015-04-29 TestAndLog.m
Lattice edge tokens; better configs; optimizations

@@ -88,7 +88,9 @@ if mod(modelState.step, options.testFreq) == 0
 
     if conAcc(1) ~= -1
         Log(hyperParams.statlog, ['pass ', num2str(modelState.pass), ' step ', num2str(modelState.step), ...
-            ' train connection acc: ', num2str(conAcc), ' test connection acc: ',  num2str(testConAcc)]);
+            ' train connection acc: ', num2str(conAcc(1)), ' test connection acc: ',  num2str(testConAcc(1))]);
+        Log(hyperParams.examplelog, ['pass ', num2str(modelState.pass), ' step ', num2str(modelState.step), ...
+            ' train connection acc: ', num2str(conAcc'), ' test connection acc: ',  num2str(testConAcc')]);
     end
 
     FlushLogs(hyperParams);

### 2015-04-29 config/Sick.m

@@ -12,17 +12,6 @@ hyperParams.name = [expName, '-', dataflag, '-l', num2str(lambda), '-dim', num2s
     '-cdim', num2str(conDim)];
 
 
-%%
-
-hyperParams.latticeConnectionHiddenDim = conDim;
-
-if conDim == 0
-    hyperParams.latticeConnectionHiddenDim = 10;
-    hyperParams.latticeLocalCurriculum = false;
-end
-
-%%
-
 % The dimensionality of the word/phrase vectors. Currently fixed at 25 to match
 % the GloVe vectors.
 hyperParams.dim = dim;
@@ -66,42 +55,7 @@ hyperParams.topDropout = topDropout;
 
 hyperParams.useEyes = 1;
 
-if composition == -1
-  hyperParams.useTrees = 0;
-  hyperParams.useThirdOrderComposition = 0;
-  hyperParams.useThirdOrderMerge = 0;
-  hyperParams.useSumming = 1;
-elseif composition < 2
-  hyperParams.useThirdOrderComposition = composition;
-  hyperParams.useThirdOrderMerge = composition;
-elseif composition == 2
-  hyperParams.lstm = 1;
-  hyperParams.useTrees = 0;
-  hyperParams.eyeScale = 0;
-  hyperParams.useThirdOrderComposition = 0;
-  hyperParams.useThirdOrderMerge = 1;
-  hyperParams.parensInSequences = 0;
-elseif composition == 3
-  hyperParams.lstm = 0;
-  hyperParams.useTrees = 0;
-  hyperParams.useThirdOrderComposition = 0;
-  hyperParams.useThirdOrderMerge = 1;
-  hyperParams.parensInSequences = 0;
-elseif composition == 4
-  hyperParams.useLattices = 1;
-  hyperParams.lstm = 0;
-  hyperParams.useTrees = 0;
-  hyperParams.useThirdOrderComposition = 0;
-  hyperParams.useThirdOrderMerge = 0;
-  hyperParams.parensInSequences = 0;
-elseif composition == 5
-  hyperParams.useLattices = 1;
-  hyperParams.lstm = 0;
-  hyperParams.useTrees = 0;
-  hyperParams.useThirdOrderComposition = 0;
-  hyperParams.useThirdOrderMerge = 1;
-  hyperParams.parensInSequences = 0;
-end
+hyperParams = CompositionSetup(hyperParams, composition);
 
 hyperParams.loadWords = true;
[…]

### 2015-05-03 writing/F'14 paper/arxiv/abstract.tex
Sentence classification w/ TreeNNs

\begin{abstract} 
  Tree-structured recursive neural networks (TreeRNNs) for sentence meaning have been
  successful for many applications, but it remains an open question whether
  the fixed-length representations that they learn can support tasks as demanding as logical deduction.
  We pursue this question by evaluating whether two such models---plain TreeRNNs and tree-structured neural
  tensor networks (TreeRNTNs)---can correctly learn to identify logical
  labelships such as entailment and contradiction using these representations. 
  In our first set of experiments, we generate artificial data from a
  logical grammar and use it to evaluate the models' ability to learn
  to handle basic labelal reasoning, recursive structures, and
  quantification. We then evaluate the models on the more natural
  SICK challenge data. Both models perform competitively on the SICK
  data and generalize well in all three experiments on simulated data,
  suggesting that they can learn suitable representations for logical
  inference in natural language. 
\end{abstract}

### 2015-05-03 writing/F'14 paper/arxiv/discussion.tex
Sentence classification w/ TreeNNs

--- 2015-02-27 writing/F'14 paper/discussion.tex
@@
 
-This paper first evaluates two recursive models on a series of three increasingly
-challenging interpretive tasks involving natural language inference---the 
-core relational algebra of natural logic with entailment and
-exclusion; recursive propositional logic structures; and statements
-involving quantification and negation. We then show that the same models can learn to
+This paper first evaluates two recursive models on three natural language inference 
+tasks over clean artificial data, covering the 
+core labelal algebra of natural logic with entailment and
+exclusion, recursive structure, and quantification. 
+We then show that the same models can learn to
 perform an entailment task on natural language. The results suggest that TreeRNTNs,
-and potentially also TreeRNNs, have the capacity to learn to model these tasks with 
+and potentially also TreeRNNs, can learn to faithfully reproduce logical inference behaviors from
 reasonably-sized training sets. These positive results are
@@
 performance falling off steadily as the size of the expressions grows. It
-remains to be seen whether these deficiencies can be overcome with
+remains to be seen whether these deficiencies are limiting in practice,
+and whether they can be overcome with
 stronger models or learning techniques. In addition, interesting 
@@
 perform complex semantic inferences from corpora, and there is ample room to develop our understanding
-using new sources of natural language data. Nonetheless, the rapid progress the field 
+using new and larger sources of natural language data. Nonetheless, the rapid progress the field 
 has made with these models in recent years provides ample reason to be optimistic that 

### 2015-05-03 writing/F'14 paper/arxiv/intro.tex
Sentence classification w/ TreeNNs

--- 2015-02-28 writing/F'14 paper/intro.tex
@@
 \citealt{dagan2006pascal}), in which the goal is to determine the core
-inferential relationship between two sentences. 
+inferential labelship between two sentences. 
 We introduce a novel NN architecture for natural language inference which
@@
 on artificial data. This logic
-defines seven mutually-exclusive relations of synonymy, entailment, contradiction,
+defines seven mutually-exclusive labels of synonymy, entailment, contradiction,
 and mutual consistency, as summarized in Table~\ref{b-table}, and it
-provides rules of semantic combination for projecting these relations
+provides rules of semantic combination for projecting these labels
 from the lexicon up to complex phrases. The formal properties of this system 
@@
 The first experiment using this logic 
-covers reasoning with the bare logical relations (\S\ref{sec:join}), the second extends this to reasoning with statements constructed compositionally from recursive functions (\S\ref{sec:recursion}),
+covers reasoning with the bare logical labels (\S\ref{sec:join}), the second extends this to reasoning with statements constructed compositionally from recursive functions (\S\ref{sec:recursion}),
 and the third covers the additional complexity that results from quantification (\S\ref{sec:quantifiers}).
@@
   \end{tabular}
-  \caption{\label{b-table}The seven relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.} %-%
+  \caption{\label{b-table}The seven labels of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.} %-%
 
 \end{table*}
-
-% Citations to additional past work to be added.\\...\\...\\...\\...\\...
-
-% Deep learning methods in NLP which learn vector representations for words have seen successful uses in recent years on increasingly sophisticated tasks \cite{collobert2011natural, socher2011semi, socher2013acl1, chen2013learning}.
-Given the still modest performance of semantically rich NLP systems in many domains---question answering and machine translation, for instance---it is worth exploring the degree to which learned vectors can serve as general purpose semantic representations.
-Much of the work to date analyzing vector representations for words (see \cite{baroni2013frege}) has focused on lexical semantic behaviors---like the similarity between words like \ii{Paris} and \ii{France}.
-Good similarity functions are valuable for many NLP tasks, but there are real use cases for which it is necessary to know how words relate to one another or to some extrinsic representation, and to model the ways in which word meanings combine to form phrase, sentence, or document meanings.
-This paper explores the ability of linguistic representations developed using supervised deep learning techniques to support interpretation and reasoning.
-
-% There are two broad family of tasks that could be used to test the ability of a model to develop general purpose meaning representations.
-In an interpretation task, sentences are mapped onto some denotation, such as  \ii{true} or \ii{false} for statements, or a factual answer for questions.
-There has been preliminary work in developing distributed models for interpretation \cite{grefenstette2013towards, rocktaschellow}, but developing a representation of world knowledge that corresponds accurately to the content expressed in language introduces considerable design challenges.
-I approach the problem by way of an inference task instead.
-Inferring the truth of one sentence from another does not require any preexisting knowledge representations, but nonetheless requires a precise representation of sentence meaning.
-I borrow the structure of the task from MacCartney and Manning  \cite{maccartney2009extended}.
-In it, the model is presented with a pair of sentences, and made to label the logical relation between the sentences as equivalence, entailment, or any of five other classes, as here:
-
-%\begin{quote}
-%\begin{enumerate}\small
-%\item Many smartphone users avoid high bills overseas by disabling data service.
-%\item Not everyone uses their smartphones for email when traveling abroad.
-%\end{enumerate} 
-%$\Rightarrow$ Entailment
-%\end{quote}
-
-%In this paper, we test the ability of recursive models to on three simple tasks, each of which is meant to capture a property that is necessary in representing natural language meaning in the setting of inference.
-I begin with an overview of MacCartney and Manning's \cite{maccartney2009extended} framework for inference, and of the recursive neural networks that I study.
-by showing that these models can learn to correctly represent entailment representations between sentences.
-I then show that these models can capture the meanings of recursive structures accurately up to a sufficient depth.
-I finally close with a brief demonstration of the ability of these models to reason over short natural language sentences involving quantifiers.
-
-% \subsection{Natural language inference and natural logic}
-
-% In its simplest form, \ii{natural language inference} (also known as \ii{recognizing textual entailment}, as in \cite{dagan2006pascal}) is the task of determine whether one sentence entails another.
-Much of the theoretical work on this task (and some successful implemented models \cite{maccartney2009natural, watanabe2012latent}) involve \ii{natural logic}, formal systems that define sound rules of inference from one sentence of natural language to another without the need for intermediate representations in some other logic.
-The most powerful model that we are aware of for natural logic is due to MacCartney and Manning \cite{maccartney2009extended} and Icard \cite{icard2012inclusion}, and is centered around the definition of a set of seven logical relations which can hold between sentences, shown in Table \ref{b-table}.
-
-% This approach to natural logic can capture much of the complexity of natural language meaning within a well understood framework, and is also fairly straightforward to implement in a machine learning setting since it can be reduced to a seven-way classification problem on sentence pairs.
-Our goal in this paper is to learn recursive neural network models which are able to mimic key behaviors of this system.
-
-% \begin{table}
-% \begin{center}
-% \begin{tabular}{|c|c|c|c|} \hline
-% name & symbol & set-theoretic definition & example \\ \hline \hline
-% entailment & $x \sqsubset y$ & $x \subset y$ & \ii{crow, bird}  \\ \hline
-% reverse entailment & $x \sqsupset y$ & $x \supset y$ & \ii{Asian, Thai}  \\ \hline
-% equivalence & $x \equiv y$ & $x = y$ & \ii{couch, sofa} \\ \hline
-% alteration & $x$ $|$ $y$ & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{cat, dog} \\ \hline
-% negation & $x \natneg y$ & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{able, unable} \\ \hline
-% cover & $x \smallsmile y$ & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-ape} \\ \hline
-% independence & $x$ \# $y$ & (else) & \ii{hungry, hippo}\\ \hline
-% \end{tabular}
-% \caption{The entailment relations in  $\mathfrak{B}$.
-$\mathcal{D}$ is the universe of possible objects of the same type as those being compared, and the relation \# applies whenever none of the other six do, including when there is insufficient knowledge to choose a label.}
-% \label{b-table}
-% \end{center}
-% \end{table}
-
-% The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes.
-It should be noted that this is not a balanced classification problem.
-If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them.
-%  How much should we say about this?
-
-% The natural logic engine at the core of MacCartney and Manning's system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures.
-These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
-%for dozens of different cases of insertion, deletion and substitution of different types of lexical item.
-For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences.
-Here, the substitution of \ii{no} for \ii{some}  yields $\natneg$ and the substitution of \ii{dogs} for \ii{animals} yields $\sqsupset$.
-It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
-%\begin{quote}
-
-%1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
-%2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
-%3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}
-
-%\end{quote}
-
-% Work to date on inference in neural network models is quite limited.
-% \citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
-% In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations.
-They propose a deterministic inference engine (similar to MacCartney's) which is augmented by a probabilistic component that evaluates individual lexical substitution steps in the derivation using vector representations, though again these representations are not learned, and no evaluations of this system have been published to date.
-% \label{sec2}

### 2015-05-03 writing/F'14 paper/arxiv/join.tex
Sentence classification w/ TreeNNs

@@
   \end{array}$
-  \caption{In \S\ref{sec:join}, we assess our models' ability to learn to do inference over pairs of relations using the rules represented here, which are derived from the definitions of the relations in Table~\ref{b-table}.
+  \caption{In \S\ref{sec:join}, we assess our models' ability to learn to do inference over pairs of labels using the rules represented here, which are derived from the definitions of the labels in Table~\ref{b-table}.
 As an example, given that $p_1 \natfor p_2$ and $p_2 \natneg p_3$, the entry in the $\natfor$ row and the $\natneg$ column lets us conclude that $p_1 \natalt p_3$.
@@
 
-\section{Reasoning about semantic relations}\label{sec:join}
+\section{Reasoning about semantic labels}\label{sec:join}
 
 The simplest kinds of deduction in natural logic involve atomic statements 
-using the relations in Table~\ref{b-table}. 
-For instance, from the relation $p_1 \natrev p_2$ between two propositions, 
-one can infer the relation $p_2 \natfor p_1$ by applying the definitions of the relations directly. 
-If one is also given the relation $p_2 \natrev p_3$ one can conclude that $p_1 \natrev p_3$, by basic set-theoretic reasoning (transitivity of $\natrev$). The
-full set of sound such inferences on pairs of premise relations is depicted in
+using the labels in Table~\ref{b-table}. 
+For instance, from the label $p_1 \natrev p_2$ between two propositions, 
+one can infer the label $p_2 \natfor p_1$ by applying the definitions of the labels directly. 
+If one is also given the label $p_2 \natrev p_3$ one can conclude that $p_1 \natrev p_3$, by basic set-theoretic reasoning (transitivity of $\natrev$). The
+full set of sound such inferences on pairs of premise labels is depicted in
 Table~\ref{tab:jointable}. Though these basic inferences do not involve compositional
@@
 depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight proposition terms ($p_1$--$p_8$). We then generate a 
-relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
+labelal statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
 We divide these statements evenly into train and test sets, and delete the test set
@@
 of two terms, ensuring that the model has no information about the terms
-being compared except for the relations between them that appear in training.
+being compared except for the labels between them that appear in training.
 
@@
   \end{subfigure}  
-  \caption{Small example structure and data for learning relation composition.}
+  \caption{Small example structure and data for learning label composition.}
   \label{lattice-figure}
@@
 
-  \caption{Performance on the semantic relation experiments.
+  \caption{Performance on the semantic label experiments.
 These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.
@@
 \paragraph{Results} 
-The resuts (Table \ref{joinresultstable}) show that NTN is able to accurately encode the relations between the terms in the geometric relations between their vectors, 
-and is able to then use that information to recover relations that 
+The resuts (Table \ref{joinresultstable}) show that NTN is able to accurately encode the labels between the terms in the geometric labels between their vectors, 
+and is able to then use that information to recover labels that 
 are not overtly included in the training data. The NN also generalizes fairly well, 
@@
 As an example from our test data, both models correctly labeled $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$.
-On another example involving comparably frequent relations, the NTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the NN incorrectly assigned it $\natind$.
+On another example involving comparably frequent labels, the NTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the NN incorrectly assigned it $\natind$.

### 2015-05-03 writing/F'14 paper/arxiv/methods.tex
Sentence classification w/ TreeNNs

--- 2015-02-28 writing/F'14 paper/methods.tex
@@
 layer that is meant to generate a feature vector capturing the
-relation between the two phrases. The output of this layer is then
+label between the two phrases. The output of this layer is then
 given to a softmax classifier, which produces a
-distribution over the seven relations represented in Table~\ref{b-table}.
+distribution over the seven labels represented in Table~\ref{b-table}.
 

### 2015-05-03 writing/F'14 paper/arxiv/quantifiers.tex
Sentence classification w/ TreeNNs

--- 2015-02-28 writing/F'14 paper/quantifiers.tex
@@
 include five nouns, four intransitive verbs, and the negation symbol
-\ii{not}. In order to be able to define relations between sentences
-with differing lexical items, we define the lexical relations for
+\ii{not}. In order to be able to define labels between sentences
+with differing lexical items, we define the lexical labels for
 each noun--noun pair, each verb--verb pair, and each
 quantifier--quantifier pair. The grammar then generates pairs of
-sentences and calculates the relations
+sentences and calculates the labels
 between them.  
 Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.
-
-%nouns = ['warthogs', 'turtles', 'mammals', 'reptiles', 'pets']
-%verbs = ['walk', 'move', 'swim', 'growl']
-%dets = ['all', 'not_all', 'some', 'no', 'most', 'not_most', 'two', 'lt_two', 'three', 'lt_three']
-%adverbs = ['', 'not']
-
-% To assign relation labels to sentence pairs, we built a small
-% task-specific implemenation of MacCartney's logic that can
-% accurately label sentences of this restricted language. The logic is
-% not able to derive all intuitively true relations of this language,
-% and fails to derive a single unique relation for certain types of
-% statement, including De Morgean's laws (e.g. \ii{(all pets) growl
-% $\natneg$ (some pet) (not growl)}), and we simply discard these
-% examples. Exhaustively generating the valid sentences under this
-% grammar and choosing those to which a relation label can be assigned
-% yields 66k sentence pairs. Some examples of these data are provided
-% in Table~\ref{examplesofdata}.
 
@@
 in the training data, and must be able
-to recognize all of the lexical relations between the nouns, verbs,
+to recognize all of the lexical labels between the nouns, verbs,
 and quantifiers and how they interact. For instance, it might see
@@
 %
-% do not allow a blank line --- adds too much space
+
 %

### 2015-05-03 writing/F'14 paper/recursion.tex
Sentence classification w/ TreeNNs

@@
 A successful natural language inference system must reason 
-about relations not just over familiar
+about labels not just over familiar
 atomic symbols, but also over novel structures built up 
@@
 compositional semantics over such structures.
-In our evaluations, we exploit the fact that our target language
+In our evaluations, we exploit the fact that our logical language
 is infinite by testing on strings that are longer and more complex
@@
 operators are truth-functional ones.  Table~\ref{tab:pl} defines this
-logic, and Table~\ref{tab:plexs} gives some short examples relational statements from our data.
- To compute these relations
+logic, and Table~\ref{tab:plexs} gives some short examples labelal statements from our data.
+ To compute these labels
 between statements, we exhaustively enumerate the sets of assignments
 of truth values to propositional variables that would satisfy each of
-the statements, and then we convert the set-theoretic relation between
-those assignments into one of the seven relations in
-Table~\ref{b-table}. As a result, each relational statement represents
+the statements, and then we convert the set-theoretic label between
+those assignments into one of the seven labels in
+Table~\ref{b-table}. As a result, each labelal statement represents
 a valid theorem of the propositional logic, and to succeed, the models 
@@
 each operator can be represented accurately by a single bit.
-In the more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.}
+In the much more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.}
 
@@
     \end{tabular}
-    \caption{Examples of the type of statements used for training and testing. These are relations between
+    \caption{Examples of the type of statements used for training and testing. These are labels between
       well-formed formulae, computed in terms of sets of satisfying
@@
   \end{subtable}
-  \caption{Natural logic relations over sentences of propositional logic.}  
+  \caption{Natural logic labels over sentences of propositional logic.}  
   \label{prop-figure}
@@
 In our experiments, we randomly generate unique pairs 
-of formulae and compute the relation that holds for each pair.
+of formulae containing up to 12 instances of logical operators each and compute the label that holds for each pair.
 We discard pairs in which either statement is either a tautology or a
-contradiction, for which the seven relations in
+contradiction, for which the seven labels in
 Table~\ref{b-table} are undefined. The resulting set of formula pairs is
-then partitioned into 13 bins according the number of logical operators in
-the longer of the two formulae. We then sample 20\% of each
+then partitioned into 12 bins according the number of operators in
+the larger of the two formulae. We then sample 20\% of each
 bin for a held-out test set.
@@
 by statements in which the two formulae refer to largely separate
-subsets of the six variables, which means that the $\natind$ relation
+subsets of the six variables, which means that the $\natind$ label
 is almost always correct.  In an effort to balance the distribution of
-relation labels without departing from the basic task of modeling
+label labels without departing from the basic task of modeling
 propositional logic, we disallow individual pairs of statements from
@@
 training examples with more than 4 logical operators, yielding 60k short training examples,
-and 21k test examples containing up to 12 operators.
+and 21k test examples across all 12 bins.
 In addition to the two tree models, we also train a summing NN baseline
 which is largely identical to the TreeRNN, except that instead of using a learned composition function,
-it simply sums the term vectors in each expression to compose them. Unlike the two tree models, this baseline does not use word order,
-and is as such guaranteed to ignore some information that it would need to succeed perfectly.
+it simply sums the term vectors in each expression to compose them before passing them to the comparison layer. Unlike the two tree models, this baseline does not use word order,
+and is as such guaranteed to ignore some information that it would need in order to succeed perfectly.
 
@@
 %  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-  \caption{Results on recursive structure.
-The ``$\natind$ only'' line shows the frequency of the most frequent class, $\natind$, and the vertical dotted line marks the size of the longest training examples.}
+  \caption{Results on recursive structure. The vertical dotted line marks the size of the longest training examples.}
 %    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
@@
 
-\paragraph{Results} Fig.~\ref{prop-results} shows the relationship
-between test accuracy and statement length. While the summing baseline model performed poorly across the board, we found that both recursive
+\paragraph{Results} Fig.~\ref{prop-results} shows the labelship
+between test accuracy and statement size. While the summing baseline model performed poorly across the board, we found that both recursive
 models were able to perform well on unseen small test examples, 
@@
 After the size four training cutoff, performance gradually decays with expression size for both tree models, suggesting that the learned approximations were accurate but lossy.
-Despite its stronger performance on short sentences, but the TreeRNTN's performance
+Despite the TreeRNTN's stronger performance on short sentences, its performance
 decayed more quickly than the TreeRNN's. 
 This suggests to us that it learned to interpret many specific fixed-size tree structures directly,
-allowing it to get away without learning strong generalizations about how to compose
-individual terms in a more general way.
+allowing it to get away without learning as robust generalizations about how to compose
+terms in the general case.
 Two factors may have contributed to the learning of these narrower generalizations: 
@@
 TreeRNN, and the TreeRNTN worked best with weaker L2 regularization than the 
-TreeRNN ($\lambda = 0.0003$ vs. $0.001$). Nonetheless, the performance of both 
-models on even the longest set of test sentences is good enough to indicate that
- both can still perform at least least a rough approximation of the underlying theorem proving task.
+TreeRNN ($\lambda = 0.0003$ vs. $0.001$). 
+However, even in the most complex set of test examples, the TreeRNTN classifies true examples of every
+ class but $\nateq$ (which is rare in long examples, and occurs only once here) correctly 
+the majority of the time, and
+ the performance of both models on those examples indicates that both have learned
+  reasonable approximations of the underlying theorem proving task over recursive structure.
 

### 2015-05-03 writing/S'14 paper/intro.tex
Caltrain edits.
Removing a TODO
Sentence classification w/ TreeNNs

@@
 when it comes to the core semantic concepts of quantification,
-entailment, and contradiction needed to identify the relationships
+entailment, and contradiction needed to identify the labelships
 between sentences like \ii{every animal walks}, \ii{every turtle
@@
 \cite{dagan2006pascal}), in which the goal is to determine the core
-inferential relationship between two sentences. Much of the
+inferential labelship between two sentences. Much of the
 theoretical work on this task (and some successful implemented models
@@
 developed by \cite{maccartney2009extended} as our formal model. This
-logic defines seven core relations of synonymy, entailment,
+logic defines seven core labels of synonymy, entailment,
 contradiction, and mutual consistency, as summarized in
 Table~\ref{b-table}, and it provides rules of semantic combination for
-projecting these relations from the lexicon up to complex phrases. The
+projecting these labels from the lexicon up to complex phrases. The
 formal properties and inferential strength of this system are now
@@
 In our experiments, we use this pre-specified logical grammar to
-generate controlled data sets encoding the semantic relationships
+generate controlled data sets encoding the semantic labelships
 between pairs of expressions and evaluate whether each of two classes
 of neural model --- plain RNNs and recursive neural tensor networks
-(RNTNs, \cite{socher2013acl1}) --- can learn those relationships
+(RNTNs, \cite{socher2013acl1}) --- can learn those labelships
 correctly. In our first experiment (Section~\ref{sec:join}), we
-evaluate the ability of these models to learn the core relational
+evaluate the ability of these models to learn the core labelal
 algebra of natural logic from data. Our second experiment
-(Section~\ref{sec:recursion}) extends this evaluation to relations
+(Section~\ref{sec:recursion}) extends this evaluation to labels
 between complex recursive structures like $(a \plor b)$ and
 $\plneg(\plneg a \pland \plneg b)$, and our third experiment
-(Section~\ref{sec:quantifiers}) involves relations between quantified
+(Section~\ref{sec:quantifiers}) involves labels between quantified
 statements like \ii{every reptile walks} and \ii{no turtle moves}.
@@
   \end{tabular}
-  \caption{The entailment relations in $\mathfrak{B}$. 
+  \caption{The seven natural logic labels of \cite{maccartney2009extended}. 
     $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
+    and the label $\natind$ applies whenever none of the other six do.} %, including when there 
     %is insufficient knowledge to choose a label.}
@@
 I borrow the structure of the task from MacCartney and Manning  \cite{maccartney2009extended}.
-In it, the model is presented with a pair of sentences, and made to label the logical relation between the sentences as equivalence, entailment, or any of five other classes, as here:
+In it, the model is presented with a pair of sentences, and made to label the logical label between the sentences as equivalence, entailment, or any of five other classes, as here:
 
@@
 Much of the theoretical work on this task (and some successful implemented models \cite{maccartney2009natural, watanabe2012latent}) involve \ii{natural logic}, formal systems that define sound rules of inference from one sentence of natural language to another without the need for intermediate representations in some other logic.
-The most powerful model that we are aware of for natural logic is due to MacCartney and Manning \cite{maccartney2009extended} and Icard \cite{icard2012inclusion}, and is centered around the definition of a set of seven logical relations which can hold between sentences, shown in Table \ref{b-table}.
+The most powerful model that we are aware of for natural logic is due to MacCartney and Manning \cite{maccartney2009extended} and Icard \cite{icard2012inclusion}, and is centered around the definition of a set of seven logical labels which can hold between sentences, shown in Table \ref{b-table}.
 
@@
 % \end{tabular}
-% \caption{The entailment relations in  $\mathfrak{B}$.
-$\mathcal{D}$ is the universe of possible objects of the same type as those being compared, and the relation \# applies whenever none of the other six do, including when there is insufficient knowledge to choose a label.}
+% \caption{The entailment labels in  $\mathfrak{B}$.
+$\mathcal{D}$ is the universe of possible objects of the same type as those being compared, and the label \# applies whenever none of the other six do, including when there is insufficient knowledge to choose a label.}
 % \label{b-table}
@@
 It should be noted that this is not a balanced classification problem.
-If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them.
+If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# label will apply between them.
 % TODO: How much should we say about this?
@@
 % The natural logic engine at the core of MacCartney and Manning's system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures.
-These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
+These signatures are tables showing the entailment label that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
 %for dozens of different cases of insertion, deletion and substitution of different types of lexical item.
-For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences.
+For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the labels introduced by each of the two differences between the sentences.
 Here, the substitution of \ii{no} for \ii{some}  yields $\natneg$ and the substitution of \ii{dogs} for \ii{animals} yields $\sqsupset$.
-It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
+It would then use an additional store of knowledge about labels to resolve the resulting series of labels into one ($|$) that expresses the label between the two sentences being compared:
 %\begin{quote}

### 2015-05-03 writing/S'14 paper/join.tex
Caltrain edits.
Draft for arXiv.
Updates for arxiv
Sentence classification w/ TreeNNs

@@
-\section{Reasoning about semantic relations}\label{sec:join}
+\section{Reasoning about semantic labels}\label{sec:join}
 
-If a model is to learn the behavior of a relational logic like the one
+If a model is to learn the behavior of a labelal logic like the one
 presented here from a finite amount of data, it must learn to deduce new
-relations from seen relations in a sound manner. The simplest such
-deductions involve atomic statements using the relations in
+labels from seen labels in a sound manner. The simplest such
+deductions involve atomic statements using the labels in
 Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
@@
 Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
-of relations for which no valid inference can be drawn in our logic.
+of labels for which no valid inference can be drawn in our logic.
 
-% about the relations themselves that do not depend on the
+% about the labels themselves that do not depend on the
 % internal structure of the things being compared. For example, given
@@
 % $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven relations support more
+% understanding $a$, $b$, or $c$. These seven labels support more
 % than just transitivity: MacCartney and Manning's
 % \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of relations of the form $a R
+% that can be made on the basis of pairs of labels of the form $a R
 % b$ and $b R' c$, including several less intuitive ones such as that if
@@
   \end{array}$
-  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.
-These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
-We assess our model's ability to reproduce such inferential paths.}
+  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the label that holds between $a$ and $c$, if any.
+These inferences are based on basic set-theoretic truths about the meanings of the underlying labels as described in Table~\ref{b-table}.
+We assess our models' ability to reproduce such inferential paths.}
   \label{tab:jointable}
@@
 % using the logic depicted in Figure~\ref{lattice-figure}.
+
+% We test the model's ability to learn this behavior by creating
+% artificial data sets of terms which represent sets of numbers. Since
+% MacCartney and Manning's set of labels hold between sets as well as
+% between sentences, we can use the underlying set structure to generate
+% the all of the labels that hold between any pair of these terms, as
+% in Figure \ref{lattice-figure}. We train the model defined above on a
+% subset of these labels, but rather then presenting the model with a
+% pair of tree-structured sentences as inputs, simply present it with
+% two single terms, each of which corresponds to a single vector in the
+% (randomly initialized) vocabulary matrix $V$, ensuring that the model
+% has no information about the terms being compared except the labels
+% between them.
+
+In our experiments, we create 80 randomly generated sets drawing from
+a domain of seven elements. This yields a data set consisting of
+6400 statements about pairs of formulae. 3200 of these pairs are
+chosen as a test set, and that test set is further reduced to the 2960
+examples that can be provably derived from the training data.
+
+We trained versions of both the RNN model and the RNTN model on these
+data sets. In both cases, the models were implemented exactly as
+described in Section~\ref{methods}, but since the items being compared
+are single terms rather than full tree structures, the composition
+layer was not involved, and the two models differed only in which
+layer function was used for the comparison layer. We simply present
+the models with two single terms, each of which corresponds to a
+single vector in the (randomly initialized) vocabulary matrix $V$,
+thereby ensuring that the model has no information about the terms
+being compared except the labels between them.
+
+\paragraph{Results} 
+We found that the both models worked well with 15-dimensional vector
+representations for the 80 set symbols,
 
@@
 % out willy-nilly without fear.
-
-% We test the model's ability to learn this behavior by creating
-% artificial data sets of terms which represent sets of numbers. Since
-% MacCartney and Manning's set of relations hold between sets as well as
-% between sentences, we can use the underlying set structure to generate
-% the all of the relations that hold between any pair of these terms, as
-% in Figure \ref{lattice-figure}. We train the model defined above on a
-% subset of these relations, but rather then presenting the model with a
-% pair of tree-structured sentences as inputs, simply present it with
-% two single terms, each of which corresponds to a single vector in the
-% (randomly initialized) vocabulary matrix $V$, ensuring that the model
-% has no information about the terms being compared except the relations
-% between them.
-
-In our experiments, we create 80 randomly generated sets drawing from
-a domain of seven elements. This yields a data set consisting of
-6400 statements about pairs of formulae. 3200 of these pairs are
-chosen as a test set, and that test set is further reduced to the 2960
-examples that can be provably derived from the training data.
-
-We trained versions of both the RNN model and the RNTN model on these
-data sets. In both cases, the models were implemented exactly as
-described in Section~\ref{methods}, but since the items being compared
-are single terms rather than full tree structures, the composition
-layer was not involved, and the two models differed only in which
-layer function was used for the comparison layer. We simply present
-the models with two single terms, each of which corresponds to a
-single vector in the (randomly initialized) vocabulary matrix $V$,
-thereby ensuring that the model has no information about the terms
-being compared except the relations between them.
-
-\paragraph{Results} 
-We found that the RNTN model worked best with 11-dimensional vector
-representations for the 80 sets and a 90-dimensional feature vector
-for the classifier, though the performance of the model was not highly
-sensitive to either dimensionality parameter in any of the experiments
-discussed here. This model was able to correctly label 99.4\% (94.3
-macroaveraged F1) of the provably derivable test examples, and 95.4\%
-(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
-best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
-respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
-the RNN.
 
@@
 
-% and create a dataset consisting of the relations between every pair of
+% and create a dataset consisting of the labels between every pair of
 % sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
@@
 % MacCartney and Manning's join table (or by the symmetry of the
-% relations in about half of the cases) and the 240 that
+% labels in about half of the cases) and the 240 that
 % cannot.
@@
 
-These results are fairly straightforward to interpret. The RNTN model
-was able to accurately encode the relations between the terms in the
-geometric relations between their vectors, and was able to then use
-that information to recover relations that were not overtly included
+The RNTN model was able to accurately encode the labels between the terms in the
+geometric labels between their vectors, and was able to then use
+that information to recover labels that were not overtly included
 in the training data. In contrast, the RNN model was able to achieve

### 2015-05-03 writing/S'14 paper/methods.tex
Sentence classification w/ TreeNNs

@@
 layer that is meant to generate a feature vector capturing the
-relation between the two phrases. The output of this layer is then
+label between the two phrases. The output of this layer is then
 given to a softmax classifier, which in turn produces a hypothesized
-distribution over the seven relations represented in Table~\ref{b-table}.
+distribution over the seven labels represented in Table~\ref{b-table}.
 

### 2015-05-03 writing/S'14 paper/quantifiers.tex
Sentence classification w/ TreeNNs

@@
 include five nouns, four intransitive verbs, and the negation symbol
-\ii{not}. In order to be able to define relations between sentences
-with differing lexical items, we define the lexical relations between
+\ii{not}. In order to be able to define labels between sentences
+with differing lexical items, we define the lexical labels between
 each noun--noun pair, each verb--verb pair, and each
 quantifier--quantifier pair. The grammar accepts aligned pairs of
-sentences of this form and calculates the natural logic relationship
+sentences of this form and calculates the natural logic labelship
 between them.  Some examples of these data are provided in
 Table~\ref{examplesofdata}.  As in previous sections, the goal of
-learning is then to assign these relational labels accurately to
+learning is then to assign these labelal labels accurately to
 unseen pairs of sentences.
@@
 
-% To assign relation labels to sentence pairs, we built a small
+% To assign label labels to sentence pairs, we built a small
 % task-specific implemenation of MacCartney's logic that can
 % accurately label sentences of this restricted language. The logic is
-% not able to derive all intuitively true relations of this language,
-% and fails to derive a single unique relation for certain types of
+% not able to derive all intuitively true labels of this language,
+% and fails to derive a single unique label for certain types of
 % statement, including De Morgean's laws (e.g. \ii{(all pets) growl
@@
 % examples. Exhaustively generating the valid sentences under this
-% grammar and choosing those to which a relation label can be assigned
+% grammar and choosing those to which a label label can be assigned
 % yields 66k sentence pairs. Some examples of these data are provided
@@
     (most mammals) move          & $\natind$ &  (not-most (not turtles)) move\\
-    (most (not pets)) (not swim) & $\natrev$ & (not-most (not pets)) move 
-    \\[2ex]    
-    (no turtles) (not growl)     & $\natalt$ & (no turtles) (not swim)\\
-    (no warthogs) swim           & $\natrev$ & (no warthogs) move\\
-    (no warthogs) move           & $\natfor$ & (no (not reptiles)) swim\\
+    (most (not pets)) (not swim) & $\natrev$ & (not-most (not pets)) move\\ 
+    %\\[2ex]    
+    %(no turtles) (not growl)     & $\natalt$ & (no turtles) (not swim)\\
+    %(no warthogs) swim           & $\natrev$ & (no warthogs) move\\
+    %(no warthogs) move           & $\natfor$ & (no (not reptiles)) swim\\
     \bottomrule
@@
 on. Crucially, though, to succeed on this task, the model must be able
-to recognize all of the lexical relations between the nouns, verbs,
+to recognize all of the lexical labels between the nouns, verbs,
 and quantifiers and how they interact. For instance, it might see

### 2015-05-03 ComputeBatchSentenceClassificationCostAndGrad.m

@@ -1,0 +1,194 @@
+% Want to distribute this code? Have other questions? -> sbowman@stanford.edu
+function [ cost, grad, embGrad, acc, connectionAcc, confusion ] = ComputeBatchSentenceClassificationCostAndGrad(theta, decoder, data, separateWordFeatures, hyperParams, computeGrad)
+% Compute cost, gradient, accuracy, and confusions over a batch of examples for some parameters.
+% This is a well-behaved costGradFn, and can be @-passed to optimizers, including minFunc and TrainSGD.
+
+% NOTE: This is reasonably well optimized. The time complexity here lies almost entirely within the batch objects in normal cases.
+
+B = length(data);  % Batch size.
+
+if (nargin < 6 || computeGrad) && nargout > 1
+    computeGrad = 1;
+else
+    computeGrad = 0;
+    grad = [];
+    embGrad = [];
+end
+
+% Unpack theta
+[ ~, ~, ...
+    softmaxMatrix, trainedWordFeatures, connectionMatrix, ...
+    compositionMatrix, scoringVector, classifierExtraMatrix, embeddingTransformMatrix ] ...
+    = stack2param(theta, decoder);
+if hyperParams.trainWords && ~hyperParams.fastEmbed
+    wordFeatures = trainedWordFeatures;
+else
+    wordFeatures = separateWordFeatures;
+end
+
+% Create and batch objects and run them forward.
+if hyperParams.useLattices
+    batch = LatticeBatch.makeLatticeBatch([data(:).sentence], wordFeatures, hyperParams);
+else
+    batch = SequenceBatch.makeSequenceBatch([data(:).sentence], wordFeatures, hyperParams);
+end
+
+[ features, connectionCosts, connectionAcc ] = ...
+    batch.runForward(embeddingTransformMatrix, connectionMatrix, scoringVector, compositionMatrix, hyperParams, computeGrad);
+
+% Set up and run top dropout.
+[ features, mask ] = Dropout(features, hyperParams.topDropout, computeGrad);
+
+extraClassifierLayerInputs = zeros(hyperParams.penultDim, B, hyperParams.topDepth);
+extraClassifierLayerInnerOutputs = zeros(hyperParams.penultDim, B, hyperParams.topDepth - 1);
+extraClassifierLayerInputs(:, :, 1) = features;
+for layer = 1:(hyperParams.topDepth - 1) 
[…]

### 2015-05-03 layer-fns/ComputeBareSoftmaxGradients.m

@@ -4,7 +4,7 @@ function [ matrixGradients, deltasDown ] = ...
 % Compute the gradient for the softmax layer parameters using incoming
 % deltas rather than log loss and a class label vector.
 
-% TODO: (eventually maybe) add support for multiple relation classes, 
+% TODO: (eventually maybe) add support for multiple label classes, 
 % as in the other two Softmax functions.
 
 B = size(in, 2);

### 2015-05-03 writing/AAAI SSS '15/BowmanPottsManning.tex
Sentence classification w/ TreeNNs

@@
 \begin{quote}
-  Natural logic offers a powerful relational conception of meaning
+  Natural logic offers a powerful labelal conception of meaning
   that is a natural counterpart to distributed semantic
@@
   tensor networks. Our experiments evaluate the models' ability to
-  learn the basic algebra of natural logic relations from simulated
+  learn the basic algebra of natural logic labels from simulated
   data and from the WordNet noun graph.  The overall positive results
@@
 
-Natural logic offers a powerful \emph{relational} conception of
+Natural logic offers a powerful \emph{labelal} conception of
 semantics: the meanings for expressions are given, at least in part,
@@
   turtle}, and so forth. With generalized notions of entailment and
-contradiction, these relationships can be defined for all lexical
+contradiction, these labelships can be defined for all lexical
 categories as well as complex phrases, sentences, and even texts. The
 resulting theories of meaning offer valuable new analytic tools for
-tasks involving database inference, relation extraction, and textual
+tasks involving database inference, label extraction, and textual
 entailment.
@@
 Natural logic aligns well with distributed (e.g., vector)
-representations, which also naturally model meaning relationally.
+representations, which also naturally model meaning labelally.
 Distributed representations have been used successfully in a wide
@@
 and neural tensor networks (NTNs).  The natural logic is built from
-the seven relations defined in Table~\ref{b-table}. Its formal
+the seven labels defined in Table~\ref{b-table}. Its formal
 properties are now well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}, so it
@@
 In our experiments, we evaluate these models' ability to learn the
-basic algebra of natural logic relations from simulated data and from
+basic algebra of natural logic labels from simulated data and from
 the WordNet noun graph. The simulated data help us to achieve analytic
@@
   \end{tabular}
-  \protect\caption{\protect\label{b-table}The seven natural logic relations of \protect\cite{maccartney2009extended}. 
+  \protect\caption{\protect\label{b-table}The seven natural logic labels of \protect\cite{maccartney2009extended}. 
     $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the relation $\natind$ applies whenever none of the other six do.} 
+    and the label $\natind$ applies whenever none of the other six do.} 
 \end{table*}
 
-\section{Neural network models for relation classification} \label{methods}
+\section{Neural network models for label classification} \label{methods}
 
@@
 labeling a pair of words or sentences with one of a small set of
-logical relations. Unlike in a classical inference
+logical labels. Unlike in a classical inference
 setting, the model is not given specific premises to reason with at test time, but rather
 is expected to encode all of the information that it learns about each term (word) at training
-time in the term vectors, and to then use only the information encoded in the relevant pair of
+time in the term embedding vectors, and to then use only the information encoded in the relevant pair of
 term vectors at test time. 
-The architecture that we use, which is limited to
-only pairs of single terms, is depicted in
+The architecture that we use is depicted in
 Figure~\ref{sample-figure}. The model represents the two input terms
-as embeddings, which are fed into a comparison function based on one
-of two types of neural network layer functions to produce a
-representation for the relationship between the two terms. This
+as single embedding vectors, which are fed into a comparison function based on one
+of two types of neural network layer function to produce a
+representation for the labelship between the two terms. This
 representation is then fed into a softmax classifier, which outputs a
@@
         \caption{The model structure used to compare \ii{turtle} and
-          \ii{animal}.  Learned word representations are fed through
+          \ii{animal}.  Learned term representations are fed through
           either an NN or NTN comparison layer and then to a softmax
-        classifier over the seven relations in Table~\ref{b-table}.}
+        classifier over the seven labels in Table~\ref{b-table}.}
   \label{sample-figure}
@@
 
-\section{Reasoning about semantic relations}\label{sec:join}
+\section{Reasoning about semantic labels}\label{sec:join}
 
@@
 are 32 valid inferences that can be made on the basis of pairs of
-relations of the form $a R b$ and $b R' c$. Cells in the table
-containing a dot correspond to pairs of relations for which no valid
+labels of the form $a R b$ and $b R' c$. Cells in the table
+containing a dot correspond to pairs of labels for which no valid
 inference can be drawn in our logic.
 
-% about the relations themselves that do not depend on the
+% about the labels themselves that do not depend on the
 % internal structure of the things being compared. For example, given
@@
 % artificial data sets of terms which represent sets of numbers. Since
-% MacCartney and Manning's set of relations hold between sets as well as
+% MacCartney and Manning's set of labels hold between sets as well as
 % between sentences, we can use the underlying set structure to generate
-% the all of the relations that hold between any pair of these terms, as
+% the all of the labels that hold between any pair of these terms, as
 % in Figure \ref{lattice-figure}. We train the model defined above on a
-% subset of these relations, but rather then presenting the model with a
+% subset of these labels, but rather then presenting the model with a
 % pair of tree-structured sentences as inputs, simply present it with
@@
 % (randomly initialized) vocabulary matrix $V$, ensuring that the model
-% has no information about the terms being compared except the relations
+% has no information about the terms being compared except the labels
 % between them.
@@
 domain of seven elements. This yields 6400 statements about pairs of
-formulae, of which 3200 are chosen as a test set. For 240 of the test examples, 
-there is not sufficient evidence in the training data to support the choice of any 
-one relation, and we  exclude these from our primary test set. 
-We trained both the NN model and the 
-NTN model on these data sets.
+formulae, of which 3200 are chosen as a test set. 240 of the test examples 
+were excluded from the primary test set for lack of evidence in the training 
+data to support the choice of any one label. 
+We then trained and evaluated both the NN model and the NTN model on these data sets.
 
@@
       \end{array}$}
-    \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.
-These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}.
+    \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the label that holds between $a$ and $c$, if any.
+These inferences are based on basic set-theoretic truths about the meanings of the underlying labels as described in Table~\ref{b-table}.
 We assess our models' ability to reproduce such inferential paths.}
@@
   \end{subfigure}  
-  \caption{Experimental goal and set-up for reasoning about semantic relations.}
+  \caption{Experimental goal and set-up for reasoning about semantic labels.}
   \label{exp1}
@@
 These results are fairly straightforward to interpret. The NTN model
-was able to accurately encode the relations between the terms in the
-geometric relations between their vectors, and was able to then use
-that information to recover relations that were not overtly included
+was able to accurately encode the labels between the terms in the
+geometric labels between their vectors, and was able to then use
+that information to recover labels that were not overtly included
 in the training data. In contrast, the NN was able to achieve this
@@
 
-\section{Reasoning about lexical relations in WordNet}\label{sec:wordnet}
+\section{Reasoning about lexical labels in WordNet}\label{sec:wordnet}
 
@@
 there are no available resources labeling such a vocabulary with the
-relations from Table~\ref{b-table}. However, the relations in WordNet
+labels from Table~\ref{b-table}. However, the labels in WordNet
 \cite{fellbaum2010wordnet,WordNet95} come close and pose the same substantive
@@
 
-We extract three types of relation from WordNet. \ii{Hypernym} and
+We extract three types of label from WordNet. \ii{Hypernym} and
 \ii{hyponym} can be represented directly in the WordNet graph
 structure, and correspond closely to the $\natrev$ and $\natfor$
-relations from natural logic. As in natural logic, these relations are
+labels from natural logic. As in natural logic, these labels are
 mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
@@
 terms tend to exclude one another, thereby providing a loose approximation 
-of the natural logic exclusion relation $\natalt$. WordNet's antonym relation, however, is both too narrowly defined
-and too rare to use for this purpose. WordNet defines its relations over
+of the natural logic exclusion label $\natalt$. WordNet's antonym label, however, is both too narrowly defined
+and too rare to use for this purpose. WordNet defines its labels over
 sets of synonyms, rather than over individual terms, so we do not
-include a \ii{synonym} or \ii{equivalent} relation, but rather
+include a \ii{synonym} or \ii{equivalent} label, but rather
 consider only one member of each set of synonyms. Word pairs which do
-not fall into these three relations are not included in the data set.
+not fall into these three labels are not included in the data set.
 
 To limit the size of the vocabulary without otherwise simplifying the learning problem, we extract all of the
-instances of these three relations for single word nouns in WordNet that are hyponyms of the node 
+instances of these three labels for single word nouns in WordNet that are hyponyms of the node 
 \texttt{organism.n.01}. In order to balance the distribution of the classes, we slightly downsample instances 
-of the \ii{coordinate} relation, yielding a total of 36,772 relations among 3,217 terms.
+of the \ii{coordinate} label, yielding a total of 36,772 labels among 3,217 terms.
 We report results below using crossvalidation, choosing a disjoint 10\% test sample for each of five runs.
@@
 This paper evaluated two neural models on the task of learning 
-natural logic relations between distributed word representations. The
+natural logic labels between distributed word representations. The
 results suggest that at least the neural tensor network
@@
 to embed a vocabulary in a way that encodes a diverse
-set of relations, and to subsequently use those embeddings to infer new
-relations. In
+set of labels, and to subsequently use those embeddings to infer new
+labels. In
 \cite{Bowman:Potts:Manning:2014}, we extend these results to include

### 2015-05-03 writing/AAAI SSS '15/BowmanPottsManning/BowmanPottsManning.tex
Sentence classification w/ TreeNNs

@@
 labeling a pair of words or sentences with one of a small set of
-logical relations. Unlike in a classical inference
+logical labels. Unlike in a classical inference
 setting, the model is not given specific premises to reason with at test time, but rather
@@
 of two types of neural network layer function to produce a
-representation for the relationship between the two terms. This
+representation for the labelship between the two terms. This
 representation is then fed into a softmax classifier, which outputs a
@@
           either an NN or NTN comparison layer and then to a softmax
-        classifier over the seven relations in Table~\ref{b-table}.}
+        classifier over the seven labels in Table~\ref{b-table}.}
   \label{sample-figure}
@@
 were excluded from the primary test set for lack of evidence in the training 
-data to support the choice of any one relation. 
+data to support the choice of any one label. 
 We then trained and evaluated both the NN model and the NTN model on these data sets.

### 2015-05-03 writing/BayLearn abstract/long_abstract.tex
Sentence classification w/ TreeNNs

@@
   recursive neural tensor networks (RNTNs) --- can correctly learn
-  relationships such as entailment and contradiction between pairs of
+  labelships such as entailment and contradiction between pairs of
   sentences, where we have generated controlled data sets of sentences
   from a logical grammar.  Our first experiment evaluates whether
-  these models can learn the basic algebra of logical relations
+  these models can learn the basic algebra of logical labels
   involved. Our second and third experiments extend this evaluation to
@@
 In our experiments, we train pairs of recursive (tree structured) neural network models \cite{socher2013acl1} which are joined together with a shared top layer that generates features for a classifier.
-The classifier predicts the logical relation that holds between the sentences represented by the two trees.
+The classifier predicts the logical label that holds between the sentences represented by the two trees (entailment in the above; the table below reviews the full inventory of labels we predict).
 For an activation function, we use either a plain NN layer or a tensor combination layer.
@@
 
-\subsection*{Inference and the semantic relations}
+\subsection*{Inference and the semantic labels}
 
 Our aim is to evaluate the ability of learned recusive models to represent semantic structure.
-We pursue this through an inference task, where the models must learn to choose the logical relation that applies between a pair of statements.
-The possible relations are the seven below from \cite{maccartney2009extended}.
+We pursue this through an inference task, where the models must learn to choose the logical label that applies between a pair of statements.
+The possible labels are the seven below from \cite{maccartney2009extended}.
 This is distinct from the limited prior work on learning neural network models for formal semantics \cite{grefenstette2013towards, rocktaschellow}, which use an interpretation task in which the model evaluates sentences as \textit{true} or \textit{false} with respect to some representation of the world.
@@
 
-If any model is to learn the behavior of a relational logic like the one presented here from a finite amount of data, it must learn to deduce new relations from already seen relations.
+If any model is to learn the behavior of a labelal logic like the one presented here from a finite amount of data, it must learn to deduce new labels from already seen labels.
 Our first experiment evaluates the ability of our models to do this over pairs of atomic symbols.
-The model is trained on a randomly generated set of consistent statements like \{$a \natfor b$, $b \natneg c$, $c \nateq d$\}, and tested on novel examples that follow from the statements seen in training, like \{$b \natneg d$\}.
+The model is trained on a randomly generated set of consistent statements like \{$a \natfor b$, $b \natfor c$, $c \natneg d$\}, and tested on novel examples that follow from the statements seen in training, like \{$a \natfor c$\}.
 A tuned RNTN reached greater than 99\% test accuracy, but no plain RNN surpassed 89\%.
@@
 Our second experiment introduces compositional structure to our examples, replacing the atomic statements above with pairs of short statements of propositional logic, like $\plneg a\natalt((a~(\pland~ b))~(\pland~c))$.
-We train our models on only pairs of statements with up to four symbols, but observe that the RNTN performs reasonably both on pairs of that length and on much longer test pairs, as shown below.
+We train our models on only pairs of statements with up to four symbols (corresponding to test accuracy figures to the left of the dashed line below), but observe that the RNTN performs reasonably both on pairs of that length and on much longer test pairs.
 
@@
   \label{prop-results}
-\end{figure}
+\end{figure} \vspace{-.3cm}
 
@@
 RNTNs performed either well ($>85\%$ accuracy) or perfectly on all of them, while plain RNNs did not break 80\% in any setting.
+These experiments differentiate the increased power of RNTNs better than previous work and provide the most convincing demonstration to date of the ability of neural networks to model semantic inferences in complex natural language sentences.
 

### 2015-05-03 writing/F'14 paper/arxiv/sick.tex
Sentence classification w/ TreeNNs

--- 2015-02-28 writing/F'14 paper/sick.tex
@@
 \section{The SICK textual entailment challenge}\label{sec:sick}
-
-% TODO: Something on how we can pack enough
 
@@
      \midrule
-    DG Train	& 50.0 & 64.0 & 67.0 & \textbf{74.0} \\
-    SICK Train	& 56.7 & 95.2 & 95.4 & \textbf{97.8} \\
-    SICK Test	& 56.7 & 73.6 & 74.9 & \textbf{76.9} \\
+    DG Train	& 50.0 & 68.0 & 67.0 & \textbf{74.0} \\
+    SICK Train	& 56.7 & 96.6 & 95.4 & \textbf{97.8} \\
+    SICK Test	& 56.7 & 73.4 & 74.9 & \textbf{76.9} \\
     \midrule
-    \textsc{Passive} (4\%)	& 0 		& 80  		& 68		&\textbf{88}\\   
-    \textsc{Neg} (7\%)	& 0 		& 88	 		& \textbf{100} & \textbf{100}\\
-    \textsc{Subst} (24\%)	& 28 		& 60  		& 64 		&  \textbf{72}\\
-    \textsc{MultiEd} (39\%)	&  \textbf{68} & 66  		&66 		& 64 \\
-    \textsc{Diff} (26\%)	& \textbf{96} &  	82		&79		& \textbf{96}\\  
+    \textsc{Passive} (4\%)	& 0 		& 76  		& 68		&\textbf{88}\\   
+    \textsc{Neg} (7\%)		& 0 		& 96	 		& \textbf{100} & \textbf{100}\\
+    \textsc{Subst} (24\%)	& 28 		& \textbf{72}  		& 64 		&  \textbf{72}\\
+    \textsc{MultiEd} (39\%)	&  \textbf{68} & 61  		&66 		& 64 \\
+    \textsc{Diff} (26\%)		& \textbf{96} &  	68		&79		& \textbf{96}\\  
     \midrule
-    \textsc{Short} (47\%) & 50.0 & 73.9 & 75.9		& \textbf{77.3} \\    
+    \textsc{Short} (47\%) & 50.0 & 73.9 & 73.5		& \textbf{77.3} \\    
     \bottomrule
@@
 
-\paragraph{Results} Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships, the results (Table~\ref{sresultstable}) show that our tree-structured models perform competitively on textual entailment, beating a strong baseline.
+\paragraph{Results} Despite the small amount of high quality training data available and the lack of resources for learning lexical labelships, the results (Table~\ref{sresultstable}) show that our tree-structured models perform competitively on textual entailment, beating a strong baseline.
 Neither model reached the performance of the winning system (84.6\%), but the TreeRNTN did exceed that of eight out of 18 submitted systems, including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.
@@
 The results (Table \ref{examplesofsickdata}) show that the TreeRNTN performs especially strongly in the two categories which pick out specific syntactic configurations, \textsc{Passive} and \textsc{Neg}, suggesting that that model has learned to encode the relevant structures well.
-It also beats both other models on \textsc{Subst}, which most closely parallels the lexical entailment inferences addressed in \S\ref{sec:quantifiers}.
-In addition none of the models perform better on the \textsc{Short} pairs than on the rest of the data, suggesting that the performance decay observed in \S\ref{sec:recursion} may not impact models trained on typical natural language text.
+It also performs fairly on \textsc{Subst}, which most closely parallels the lexical entailment inferences addressed in \S\ref{sec:quantifiers}.
+In addition, none of the models perform dramatically better on the \textsc{Short} pairs than on the rest of the data, suggesting that the performance decay observed in \S\ref{sec:recursion} may not impact models trained on typical natural language text.
 

### 2015-05-03 writing/F'14 paper/intro.tex
Sentence classification w/ TreeNNs

@@
   \end{tabular}
-  \caption{\label{b-table}The seven relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.} %-%
+  \caption{\label{b-table}The seven labels of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.} %-%
 
@@
 It should be noted that this is not a balanced classification problem.
-If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them.
+If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# label will apply between them.
 %  How much should we say about this?

### 2015-05-03 writing/F'14 paper/join.tex
Sentence classification w/ TreeNNs

@@
 
-\section{Reasoning about semantic relations}\label{sec:join}
+\section{Reasoning about semantic labels}\label{sec:join}
 
 The simplest kinds of deduction in natural logic involve atomic statements 
-using the relations in Table~\ref{b-table}. 
-For instance, from the relation $p_1 \natrev p_2$ between two propositions, 
-one can conclude that $p_2 \natfor p_1$ using the definitions of the relations directly. 
-If one is also given the relation $p_1 \natrev p_3$, one can conclude that $p_1 \natrev p_2$, by basic set-theoretic reasoning (transitivity of $\natrev$). The
-full set of sound such inferences on pairs of premise relations is depicted in
+using the labels in Table~\ref{b-table}. 
+For instance, from the label $p_1 \natrev p_2$ between two propositions, 
+one can infer the label $p_2 \natfor p_1$ by applying the definitions of the labels directly. 
+If one is also given the label $p_2 \natrev p_3$ one can conclude that $p_1 \natrev p_3$, by basic set-theoretic reasoning (transitivity of $\natrev$). The
+full set of sound such inferences on pairs of premise labels is depicted in
 Table~\ref{tab:jointable}. Though these basic inferences do not involve compositional
@@
 
-% about the relations themselves that do not depend on the
+% about the labels themselves that do not depend on the
 % internal structure of the things being compared. For example, given
@@
 on which we will base the statements in the train and test sets.
-This takes the form of a small boolean structure in which terms denote
+This takes the form of a small Boolean structure in which terms denote
 sets of entities from a small domain.  Fig.~\ref{lattice-figure}a
-depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight terms ($p_{1\text{-}8}$). We then generate a 
-relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
+depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight proposition terms ($p_1$--$p_8$). We then generate a 
+labelal statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
 We divide these statements evenly into train and test sets, and delete the test set
  examples which cannot be proven from the train examples, for which there is not enough information for even an ideal system to choose a correct label.
-In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
-2960 relations.
+In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 examples and a test set of 
+2960 examples.
 
 We trained models with both the NN and NTN comparison functions on these
-data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no simpler version of our model is a viable baseline.} In both cases, the models are implemented as
+data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no simpler version of our model is a viable baseline.} %+%
+In both cases, the models are implemented as
 described in \S\ref{methods}, but since the items being compared
@@
 
-  \caption{Performance on the semantic relation experiments.
-These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
+  \caption{Performance on the semantic label experiments.
+These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.
+The ``$\natind$ only'' entries reflect the frequency of the most frequent class.}
   \label{joinresultstable}
@@
 As an example from our test data, both models correctly labeled $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$.
-On another example involving comparably frequent relations, the NTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the NN incorrectly assigned it $\natind$.
+On another example involving comparably frequent labels, the NTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the NN incorrectly assigned it $\natind$.
 

### 2015-05-03 writing/F'14 paper/quantifiers.tex
Sentence classification w/ TreeNNs

@@
 include five nouns, four intransitive verbs, and the negation symbol
-\ii{not}. In order to be able to define relations between sentences
-with differing lexical items, we define the lexical relations for
+\ii{not}. In order to be able to define labels between sentences
+with differing lexical items, we define the lexical labels for
 each noun--noun pair, each verb--verb pair, and each
 quantifier--quantifier pair. The grammar then generates pairs of
-sentences and calculates the relations
+sentences and calculates the labels
 between them.  

### 2015-05-03 writing/S'14 paper/abstract.tex
Sentence classification w/ TreeNNs

@@
   We find that the plain RNN achieves only mixed results on all three
-  experiments, whereas the stronger RNTN models generalizes well in
+  experiments, whereas the stronger RNTN model generalizes well in
   every setting and appears capable of learning suitable
@@
 
-% We address this question directly by for the first time evaluating whether each of two classes of neural model — plain RNNs and recursive neural tensor networks (RNTNs) — can correctly learn relationships such as entailment and contradiction between pairs of sentences, where we have generated controlled data sets of sentences from a logical grammar.
+% We address this question directly by for the first time evaluating whether each of two classes of neural model — plain RNNs and recursive neural tensor networks (RNTNs) — can correctly learn labelships such as entailment and contradiction between pairs of sentences, where we have generated controlled data sets of sentences from a logical grammar.

### 2015-05-03 writing/S'14 paper/discussion.tex
Sentence classification w/ TreeNNs

@@
 challenging interpretive tasks involving natural language inference:
-the core relational algebra of natural logic with entailment and
+the core labelal algebra of natural logic with entailment and
 exclusion; recursive propositional logic structures; and statements
@@
 
-% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators.
+% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical labels or the learning of recursive operators.
 This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic.
 The positive results that we have found, however, are extremely promising for the future of learned representation models in the applied modeling of meaning.
-We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators.
+We have seen that recursive neural tensor networks are able to encode lexical labels accurately and encode recursive operators.
 We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.

### 2015-05-03 writing/S'14 paper/recursion.tex
Sentence classification w/ TreeNNs

@@
 operators are truth-functional ones.  Table~\ref{tab:pl} defines this
-logic, and Table~\ref{tab:plexs} gives some examples of relational
-statements that we can make in these terms. To compute these relations
+logic, and Table~\ref{tab:plexs} gives some examples of labelal
+statements that we can make in these terms. To compute these labels
 between statements, we exhaustively enumerate the sets of assignments
 of truth values to propositional variables that would satisfy each of
-the statements, and then we convert the set-theoretic relation between
-those assignments into one of the seven relations in
-Table~\ref{b-table}. As a result, each relational statement represents
+the statements, and then we convert the set-theoretic label between
+those assignments into one of the seven labels in
+Table~\ref{b-table}. As a result, each labelal statement represents
 a valid theorem of the propositional logic.
@@
     \end{tabular}
-    \caption{Examples of statements about relations between
+    \caption{Examples of statements about labels between
       well-formed formulae, defined in terms of sets of satisfying
@@
   \end{subtable}
-  \caption{Propositional logic with natural logic relations.}  
+  \caption{Propositional logic with natural logic labels.}  
   \label{prop-figure}
@@
 For our experiments, we randomly generate a large set of  unique pairs 
-of formulae and compute the relation that holds for each pair.
+of formulae and compute the label that holds for each pair.
 We discard pairs in which either statement is a tautology or
 contradiction (a statement that is true of either all or no possible
-assignments), for which none of the seven relation labels in
+assignments), for which none of the seven label labels in
 Table~\ref{b-table} can hold. The resulting set of formula pairs is

### 2015-05-03 writing/S'15 misc/newmodel.tex
Sentence classification w/ TreeNNs

@@
 The loss for the example is then defined as the sum of three parts, a log likelihood term for the predicted sentiment class label, a sum of log likelihood terms for the correct composition location decisions (scaled by the number of ``middle'' layers in the lattice for which a composition location choice must be made), and an L2 regularization term.
-For sentiment relation label $\rho$, composition loctation label vector $\vec{\kappa}$, concatenated parameter vector $\theta$, and regularization strength hyperparameter $\lambda$ the loss is:
+For sentiment label label $\rho$, composition loctation label vector $\vec{\kappa}$, concatenated parameter vector $\theta$, and regularization strength hyperparameter $\lambda$ the loss is:
 %
@@
 Batching
+
+\subsection{Using pretrained word embeddings}
+
+In order to be able to take advantage of the information contained in high-dimensional pretrained word embeddings (we use the 200d vectors distributed with Glove) without the significant expense of large intermediate node representations, we use a projection neural network layer to map the pretrained embeddings  Like in (CITE: Irsoy, Cardie) this has the effect of allowing the intermediate node representations to live in a differently structured vector space from the words.
+This is defined as follows, where $N$ is the row index of the bottom row (equivalent to the length of the sentence), $\mathbf{M}_t$ and $\vec{b_t}$ are the parameters of the embedding transformation layer, $\mathbf{E}$ is the matrix of pretrained embeddings, $\delta$ is a one-hot vector with a 1 at index $w$, and $w$ is the index into that matrix for the word at position $j$:
+
+\begin{equation}
+\vec{y}^{(N,j)} = \tanh(\mathbf{M_t} \mathbf{E} \delta_{w} + \vec{b_t})
+\end{equation}
 

### 2015-05-03 config/Defaults.m

@@ -125,7 +125,7 @@ hyperParams.fastEmbed = false;
 
 hyperParams.clearActivations = false;
 
-% Applies to the unbatched ComputeCostAndGrad only.
+% Applies to unbatched computation only.
 % Full gradient vectors above this l2 norm will be rescaled down.
 hyperParams.clipGradients = true;
 hyperParams.maxGradNorm = 5;
@@ -137,10 +137,13 @@ hyperParams.connectionCostScale = 1;
 
 % One scalar for each label. Multiply the cost and gradient of examples with this label by this factor.
 % Currently doesn't support multiple label sets, but this could be fairly easily fixed.
-hyperParams.relationCostMultipliers = [];
+hyperParams.labelCostMultipliers = [];
 
 % Load at most this many lines of any one file. Useful in debugging.
 hyperParams.lineLimit = inf;
+
+% Treat parsing parens as tokens in sequence models.
+hyperParams.parensInSequences = false;
 
 %%% minFunc options: %%%
 

### 2015-05-03 config/GradCheck.m

@@ -1,7 +1,7 @@
-function [ hyperParams, options, wordMap, relationMap ] = GradCheck(transDepth, topDepth, composition, trainwords, fastemb, multipleClassSets, sentiment)
+function [ hyperParams, options, wordMap, labelMap ] = GradCheck(transDepth, topDepth, composition, trainwords, fastemb, multipleClassSets, sentiment)
 % Set up a gradient check for the main learned parameters.
 
-% TrainModel('', 1, @GradCheck, 1, 2, 7, 1, 0, 0, 1);
+% TrainModel('', 1, @GradCheck, 1, 2, 4, 1, 0, 0, 1);
 % NOTE: The LatticeLSTM doesn't do especially well on gradient checks at random initialization
 % but basically passes if tested on the model state after a few steps... keep an eye on this.
 
@@ -69,37 +69,37 @@ if sentiment
 	wordMap = InitializeMaps('./sst-data/gradcheckwords.txt');
 	hyperParams.vocabName = 'sst-gc'; 
 
-	hyperParams.numRelations = [5];
-	hyperParams.relationCostMultipliers = [1 0.0001 3 0.25 1];
+	hyperParams.numLabels = [5];
+	hyperParams.labelCostMultipliers = [1 0.0001 3 0.25 1];
 
-	hyperParams.relations = {{'0', '1', '2', '3', '4'}};
-	relationMap = cell(1, 1);
-	relationMap{1} = containers.Map(hyperParams.relations{1}, 1:length(hyperParams.relations{1}));
+	hyperParams.labels = {{'0', '1', '2', '3', '4'}};
+	labelMap = cell(1, 1);
+	labelMap{1} = containers.Map(hyperParams.labels{1}, 1:length(hyperParams.labels{1}));
 
 	hyperParams.trainFilenames = {};    
 	hyperParams.splitFilenames = {'./sst-data/gradcheckfile.txt'};    
 	hyperParams.testFilenames = {};
 else
 	if ~multipleClassSets
-		hyperParams.relations = {{'#', '=', '>', '<', '|', '^', 'v'}};
-		hyperParams.relationCostMultipliers = [1 0.0001 3 0.25 1 0.5 0.5];
-		hyperParams.numRelations = [7];
-		relationMap = cell(1, 1);
-		relationMap{1} = containers.Map(hyperParams.relations{1}, 1:length(hyperParams.relations{1}));
[…]

### 2015-05-03 writing/BayLearn abstract/long_abstract.tex
Writing replacement fix.

--- 2014-08-25 writing/BayLearn abstract/long_abstract.tex
@@
 In our experiments, we train pairs of recursive (tree structured) neural network models \cite{socher2013acl1} which are joined together with a shared top layer that generates features for a classifier.
-The classifier predicts the logical relation that holds between the sentences represented by the two trees.
+The classifier predicts the logical relation that holds between the sentences represented by the two trees (entailment in the above; the table below reviews the full inventory of relations we predict).
 For an activation function, we use either a plain NN layer or a tensor combination layer.
@@
 Our first experiment evaluates the ability of our models to do this over pairs of atomic symbols.
-The model is trained on a randomly generated set of consistent statements like \{$a \natfor b$, $b \natneg c$, $c \nateq d$\}, and tested on novel examples that follow from the statements seen in training, like \{$b \natneg d$\}.
+The model is trained on a randomly generated set of consistent statements like \{$a \natfor b$, $b \natfor c$, $c \natneg d$\}, and tested on novel examples that follow from the statements seen in training, like \{$a \natfor c$\}.
 A tuned RNTN reached greater than 99\% test accuracy, but no plain RNN surpassed 89\%.

### 2015-05-03 writing/F'14 paper/arxiv/intro.tex
Writing replacement fix.

@@
 \citealt{dagan2006pascal}), in which the goal is to determine the core
-inferential labelship between two sentences. 
+inferential relationship between two sentences. 
 We introduce a novel NN architecture for natural language inference which
@@
 on artificial data. This logic
-defines seven mutually-exclusive labels of synonymy, entailment, contradiction,
+defines seven mutually-exclusive relations of synonymy, entailment, contradiction,
 and mutual consistency, as summarized in Table~\ref{b-table}, and it
-provides rules of semantic combination for projecting these labels
+provides rules of semantic combination for projecting these relations
 from the lexicon up to complex phrases. The formal properties of this system 
@@
 The first experiment using this logic 
-covers reasoning with the bare logical labels (\S\ref{sec:join}), the second extends this to reasoning with statements constructed compositionally from recursive functions (\S\ref{sec:recursion}),
+covers reasoning with the bare logical relations (\S\ref{sec:join}), the second extends this to reasoning with statements constructed compositionally from recursive functions (\S\ref{sec:recursion}),
 and the third covers the additional complexity that results from quantification (\S\ref{sec:quantifiers}).
@@
   \end{tabular}
-  \caption{\label{b-table}The seven labels of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.} %-%
+  \caption{\label{b-table}The seven relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.} %-%
 

### 2015-05-03 writing/F'14 paper/arxiv/recursion.tex
Sentence classification w/ TreeNNs
Writing replacement fix.

--- 2015-02-27 writing/F'14 paper/recursion.tex
@@
 compositional semantics over such structures.
-In our evaluations, we exploit the fact that our target language
+In our evaluations, we exploit the fact that our logical language
 is infinite by testing on strings that are longer and more complex
 than any seen in training.
-
-% Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice
-%   said hello}, and \ii{Carl thinks that Bob said that Alice said
-%   hello}. Overt recursion of this kind is easy to find, and
-% theoretical accounts of natural language syntax and semantics rely
-% heavily on recursive structures. In order for a model to be able to
-% accurately learn natural language meanings, then, we expect that it
-% would need to be able to learn to represent the meanings of function
-% words in a such a way that they are able to behave correctly when
-% taking their own outputs as input. In evaluating the model, we take
-% advantage of the fact that recursive structures of this kind define
-% potentially infinite languages by testing the model on strings that
-% are longer and more complex than any seen in testing.
 
@@
 In our experiments, we randomly generate unique pairs 
-of formulae and compute the relation that holds for each pair.
+of formulae containing up to 12 instances of logical operators each and compute the relation that holds for each pair.
 We discard pairs in which either statement is either a tautology or a
@@
 Table~\ref{b-table} are undefined. The resulting set of formula pairs is
-then partitioned into 13 bins according the number of logical operators in
-the longer of the two formulae. We then sample 20\% of each
+then partitioned into 12 bins according the number of operators in
+the larger of the two formulae. We then sample 20\% of each
 bin for a held-out test set.
@@
   \includegraphics[width=3.05in]{decayfig.pdf}
-%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
-  \caption{Results on recursive structure.
-The ``$\natind$ only'' line shows the frequency of the most frequent class, $\natind$, and the vertical dotted line marks the size of the longest training examples.}
-%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
-%    representations. As measured by Euclidean distance, only the TreeRNTN
-%    achieves this for formulae containing more than a small number of
-%    connectives (\ii{and} in this example). The TreeRNN quickly collapses the representations,
-%    failing to capture the meaning contrasts.
+
+  \caption{Results on recursive structure. The vertical dotted line marks the size of the longest training examples.}
 
@@
 \paragraph{Results} Fig.~\ref{prop-results} shows the relationship
-between test accuracy and statement length. While the summing baseline model performed poorly across the board, we found that both recursive
+between test accuracy and statement size. While the summing baseline model performed poorly across the board, we found that both recursive
 models were able to perform well on unseen small test examples, 
@@
 TreeRNN, and the TreeRNTN worked best with weaker L2 regularization than the 
-TreeRNN ($\lambda = 0.0003$ vs. $0.001$). Nonetheless, the performance of both 
-models on even the longest set of test sentences is good enough to indicate that
- both can still perform at least least a rough approximation of the underlying theorem proving task.
-
-% \mynote{In an effort to better understand why model performance decays, 
-%we evaluated both models on pairs of long formulae involving
-%binary connectives, to assess how well they distinguish representations for semantically distinct
-%formulae. We found that only the TreeRNTN is able to learn
-%substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
-% and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
-%The TreeRNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed under three operators, and quickly approaches zero as depth increases.
-% The TreeRNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}
-
-% we discovered that this model looses information about in longer
-% statements in a particularly problematic way. It appears that that
-% model is unable to distinguish between two sentences when the only
-% difference between those sentences is embedded within a very deep
-% structure. We evaluated both models on sentences that differed in only
-% one term, but for which that one term was embedded under a large
-% number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and
-%   b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and
-%   b)))}. We then measured the Euclidean distance between the vector
-% representations of the two sentences in each pair. Our findings are
-% shown in Figure~\ref{prop-falloff}, and show that while the TreeRNTN can
-% distinguish the two sentences well at every size that we test, the TreeRNN
-% fails after a depth of approximately six.
+TreeRNN ($\lambda = 0.0003$ vs. $0.001$). 
+However, even in the most complex set of test examples, the TreeRNTN classifies true examples of every
+ class but $\nateq$ (which is rare in long examples, and occurs only once here) correctly 
+the majority of the time, and
+ the performance of both models on those examples indicates that both have learned
+  reasonable approximations of the underlying theorem proving task over recursive structure.

### 2015-05-03 writing/F'14 paper/arxiv/sick.tex
Writing replacement fix.

@@
 
-\paragraph{Results} Despite the small amount of high quality training data available and the lack of resources for learning lexical labelships, the results (Table~\ref{sresultstable}) show that our tree-structured models perform competitively on textual entailment, beating a strong baseline.
+\paragraph{Results} Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships, the results (Table~\ref{sresultstable}) show that our tree-structured models perform competitively on textual entailment, beating a strong baseline.
 Neither model reached the performance of the winning system (84.6\%), but the TreeRNTN did exceed that of eight out of 18 submitted systems, including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.

### 2015-05-03 writing/F'14 paper/discussion.tex
Writing replacement fix.

@@
 
-This paper first evaluates two recursive models on a series of three increasingly
-challenging interpretive tasks involving natural language inference---the 
+This paper first evaluates two recursive models on three natural language inference 
+tasks over clean artificial data, covering the 
 core relational algebra of natural logic with entailment and
-exclusion; recursive propositional logic structures; and statements
-involving quantification and negation. We then show that the same models can learn to
+exclusion, recursive structure, and quantification. 
+We then show that the same models can learn to
 perform an entailment task on natural language. The results suggest that TreeRNTNs,
-and potentially also TreeRNNs, have the capacity to learn to model these tasks with 
+and potentially also TreeRNNs, can learn to faithfully reproduce logical inference behaviors from
 reasonably-sized training sets. These positive results are

### 2015-05-03 writing/S'14 paper/join.tex
Writing replacement fix.

--- 2014-06-06 writing/S'14 paper/join.tex
@@
 % using the logic depicted in Figure~\ref{lattice-figure}.
+
+% We test the model's ability to learn this behavior by creating
+% artificial data sets of terms which represent sets of numbers. Since
+% MacCartney and Manning's set of relations hold between sets as well as
+% between sentences, we can use the underlying set structure to generate
+% the all of the relations that hold between any pair of these terms, as
+% in Figure \ref{lattice-figure}. We train the model defined above on a
+% subset of these relations, but rather then presenting the model with a
+% pair of tree-structured sentences as inputs, simply present it with
+% two single terms, each of which corresponds to a single vector in the
+% (randomly initialized) vocabulary matrix $V$, ensuring that the model
+% has no information about the terms being compared except the relations
+% between them.
+
+In our experiments, we create 80 randomly generated sets drawing from
+a domain of seven elements. This yields a data set consisting of
+6400 statements about pairs of formulae. 3200 of these pairs are
+chosen as a test set, and that test set is further reduced to the 2960
+examples that can be provably derived from the training data.
+
+We trained versions of both the RNN model and the RNTN model on these
+data sets. In both cases, the models were implemented exactly as
+described in Section~\ref{methods}, but since the items being compared
+are single terms rather than full tree structures, the composition
+layer was not involved, and the two models differed only in which
+layer function was used for the comparison layer. We simply present
+the models with two single terms, each of which corresponds to a
+single vector in the (randomly initialized) vocabulary matrix $V$,
+thereby ensuring that the model has no information about the terms
+being compared except the relations between them.
+
+\paragraph{Results} 
+We found that the both models worked well with 15-dimensional vector
+representations for the 80 set symbols,
 
@@
 
-These results are fairly straightforward to interpret. The RNTN model
-was able to accurately encode the relations between the terms in the
+The RNTN model was able to accurately encode the relations between the terms in the
 geometric relations between their vectors, and was able to then use

### 2015-05-03 writing/S'15 misc/abstract.tex
Writing replacement fix.

% \begin{abstract}
% Recursive neural network models for sentence meaning have been successful in many tasks, but it remains an open question whether any such model can learn compositional semantic representations that support logical deduction. We pursue this question directly by evaluating whether two such models can correctly learn to identify logical relationships such as entailment and contradiction between sentences. We first describe a recursive neural tensor network model for entailment and demonstrate it on the SICK challenge. We then use artificial data from a logical grammar to test the model's ability to learn to handle basic relational reasoning, recursive functions, and quantification. The model performs competitively on the SICK data and generalizes well in all three experiments on simulated data, suggesting that it can learn suitable representations for logical inference in natural language.
% \end{abstract}

%%%% POTTS REVISION
\begin{abstract} 
  Recursive neural networks (RNNs) for sentence meaning have been
  successful for many tasks, but it remains an open question whether
  they can learn compositional semantic representations that
  support logical deduction. We pursue this question by
  evaluating whether two such models---plain RNNs and recursive neural
  tensor networks (RNTNs)---can correctly learn to identify logical
  relationships such as entailment and contradiction.
  In our first set of experiments, we generate artificial data from a
  logical grammar and use it to evaluate the models' ability to learn
  to handle basic relational reasoning, recursive structures, and
  quantification. We then evaluate the models on the more natural
  SICK challenge data. Both models perform competitively on the SICK
  data and generalize well in all three experiments on simulated data,
  suggesting that they can learn suitable representations for logical
  inference in natural language. 
\end{abstract}

### 2015-05-03 writing/S'15 misc/abstract.tex
Paper cleanup.

@@
 % Recursive neural network models for sentence meaning have been successful in many tasks, but it remains an open question whether any such model can learn compositional semantic representations that support logical deduction.
-We pursue this question directly by evaluating whether two such models can correctly learn to identify logical relationships such as entailment and contradiction between sentences.
+We pursue this question directly by evaluating whether two such models can correctly learn to identify logical labelships such as entailment and contradiction between sentences.
 We first describe a recursive neural tensor network model for entailment and demonstrate it on the SICK challenge.
-We then use artificial data from a logical grammar to test the model's ability to learn to handle basic relational reasoning, recursive functions, and quantification.
+We then use artificial data from a logical grammar to test the model's ability to learn to handle basic labelal reasoning, recursive functions, and quantification.
 The model performs competitively on the SICK data and generalizes well in all three experiments on simulated data, suggesting that it can learn suitable representations for logical inference in natural language.
@@
   tensor networks (RNTNs)---can correctly learn to identify logical
-  relationships such as entailment and contradiction.
+  labelships such as entailment and contradiction.
   In our first set of experiments, we generate artificial data from a
   logical grammar and use it to evaluate the models' ability to learn
-  to handle basic relational reasoning, recursive structures, and
+  to handle basic labelal reasoning, recursive structures, and
   quantification. We then evaluate the models on the more natural

### 2015-05-03 writing/S'15 misc/newmodel.tex
Paper cleanup.

@@
 The loss for the example is then defined as the sum of three parts, a log likelihood term for the predicted sentiment class label, a sum of log likelihood terms for the correct composition location decisions (scaled by the number of ``middle'' layers in the lattice for which a composition location choice must be made), and an L2 regularization term.
-For sentiment label label $\rho$, composition loctation label vector $\vec{\kappa}$, concatenated parameter vector $\theta$, and regularization strength hyperparameter $\lambda$ the loss is:
+For sentiment label $\rho$, composition loctation label vector $\vec{\kappa}$, concatenated parameter vector $\theta$, and regularization strength hyperparameter $\lambda$ the loss is:
 %
@@
 loss & = & \lambda\theta^2 - \ln(l_\rho) -\\
-\nonumber&&\frac{1}{N - 1} \sum_{i = 1}^{N - 2} \ln(o^{(i,\kappa^{(i)})}_{\textsc{compose}})
+\nlonumber&&\frac{1}{N - 1} \sum_{i = 1}^{N - 2} \ln(o^{(i,\kappa^{(i)})}_{\textsc{compose}})
 \end{eqnarray}

### 2015-05-04 writing/SNLI paper/newcorpus.tex
Paper update

--- 2015-02-21 writing/W'15 paper/newcorpus.tex
@@
-\section{Discussion and conclusion}\label{sec:discussion}
+\section{A new corpus for natural language inference}\label{sec:discussion}
+
+need for new corpus
+
+Our ultimate aim in this work is to develop supervised models for sentence representation that can accurately capture natural language meaning.
+While sentiment tasks like SST have provided a useful testbed for sentence representation models, sentiment labeling only requires that models be able to encode a small piece of the full expressive capacity of language.
+We claim that the task of natural language inference (also called recognizing textual entailment, or RTE) is significantly more demanding, and that strong performance on this task is good evidence of a model's overall strength in sentence representation.
+
+what's new
+- size
+- grounding
 
@@
 
-aim to surpass sick in both size and label consistency
+aim to surpass sick in both size and r consistency
+
+Since our goal is to evaluate the quality of the single sentence representations in our model, we train an entailment classifier that has access only to the output of our sentence representation model, and not to any of the word or phrase representations that it used.
+We encourage future users of this corpus to evaluate their models in this way when possible.
+Our classifier is simply a stack of three 100d (NONLIN) layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
 
@@
 \footnotesize
-\toprule
+% \toprule
 The Stanford University NLP Group is collecting data for use in research on computer understanding of English. We appreciate your help!
@@
 \textbf{Problems} (optional)   If something is wrong with the caption that makes it difficult to understand, do your best above and let us know here.
-\bottomrule
+% \bottomrule
 \caption{\label{instructions-1}Examples of training data from the newly collected entailment corpus.}
@@
 % ditto examples
-
+%
 \begin{table*}
@@
 \footnotesize
-\toprule
+%\toprule
 TODO: VALIDATION INSTRUCTIONS
-\bottomrule\\
+%\bottomrule\\
 

### 2015-05-10 writing/Lattice paper/abstract.tex
Better memory management in SequenceBatch, etc.

\begin{abstract} 
In a tree-structured recursive neural network model, the meaning of a sentence is encoded using a neural network that is structured in a way that mirrors the syntactic structure of the sentence produced by a parser. This kind of syntactic structure is potentially valuable, but the reliance on an external parser is not ideal. There is no guarantee that the parsing formalisms used by freestanding syntact parsers are at all suitable for guiding semantic composition. We present a neural network model that is able to used a joint syntactic and semantic objective to learn to simultaneously discover semantically appropriate syntactic structures and to use those structures to produce semantic representations for some downstream task. We show that this model can reach state of the art performance on sentiment classification, and that it shows hints of being able to learn both syntactic and semantic representations that could be valuable in more demanding language understanding tasks.
\end{abstract}

### 2015-05-10 writing/Lattice paper/sst.tex
Better memory management in SequenceBatch, etc.

\section{Sentiment classification}

Our initial experiments are on the Stanford Sentiment Treebank, or SST \cite{socher2013acl1}, a benchmark for sentence embedding models. In particular, we learn a model for the fine-graned classification task, in which each sentence must be assigned an integer sentiment score in $\[1, 5\]$.

In addition to sentence-level labels, every sentence in SST is parsed, and every node in the parse tree is labeled with another score from the same label set. Since we do not require our model to use the same syntactic structures that are provided with the data, we cannot use supervision from these labels at the intermediate nodes of our parse trees. We can, however, take advantage of these labels by treating every labeled subtree of each sentence as its own sentence at training time yielding 327k labeled examples, many of them very short. We can then use the same single classifier to predict top-node labels for both sentences and their subtrees, shuffling .

### 2015-05-10 writing/S'15 misc/newcorpus.tex
Relation handling cleanup.
Better memory management in SequenceBatch, etc.

--- 2015-02-21 writing/W'15 paper/newcorpus.tex
@@
 aim to surpass sick in both size and label consistency
+
+Since our goal is to evaluate the quality of the single sentence representations in our model, we train an entailment classifier that has access only to the output of our sentence representation model, and not to any of the word or phrase representations that it used.
+We encourage future users of this corpus to evaluate their models in this way when possible.
+Our classifier is simply a stack of three 100d (NONLIN) layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
 
@@
 
-We use captions from the Flickr30k corpus \cite{hodoshimage} as the left hand side of each relation pair. That corpus ...
+We use captions from the Flickr30k corpus \cite{hodoshimage} as the left hand side of each label pair. That corpus ...
 

### 2015-05-10 config/AndOr.m

@@ -26,6 +26,8 @@ options.numPasses = 15000;
 
 options.miniBatchSize = mbs;
 
+options.updateFn = @AdaDeltaUpdate;
+
 wordMap = InitializeMaps('./propositionallogic/longer2/wordlist.txt');
 
 % The name assigned to the current vocabulary. Used in deciding whether to load a 

### 2015-05-10 config/Defaults.m

@@ -39,6 +39,8 @@ hyperParams.latticeConnectionHiddenDim = 8;
 % of the correct merge positions lower in the lattice to avoid training the composition model on bad inputs.
 hyperParams.latticeLocalCurriculum = false;
 
+hyperParams.latticeSlant = 0;
+
 % The number of embedding transform layers. topDepth = 1 means an NN layer will be
 % added above the embedding matrix. This is likely to only be useful when
 % learnWords is false, and so the embeddings do not exist in the same space

### 2015-05-10 layer-fns/ComputeRNNLayer.m

@@ -2,7 +2,8 @@
 function [ activations, activationsPreNL ] = ComputeRNNLayer(l, r, matrix, NL)
 % Run an RNN layer as in forward propagation.
 
-activationsPreNL = matrix * padarray([l; r], 1, 1, 'pre');
+in = [zeros([1, size(l, 2)], 'like', l); l; r];
+activationsPreNL = matrix * in;
 activations = NL(activationsPreNL);
 
 end

### 2015-05-10 layer-fns/ComputeRNNLayerGradients.m

@@ -3,7 +3,7 @@ function [matrixGradients, deltaLeft, deltaRight] = ...
       ComputeRNNLayerGradients(l, r, matrix, delta, nonlinearityDeriv, output)
 % Compute the gradients and deltas for an RNN layer for a given batch.
 
-in = padarray([l; r], 1, 1, 'pre');
+in = [zeros([1, size(l, 2)], 'like', l); l; r];
 
 if nargin < 6
     innerOutput = matrix * in;

### 2015-05-10 LatticeBatch.m
Better memory management in SequenceBatch, etc.

@@ -27,6 +27,7 @@ classdef LatticeBatch < handle
         activeNode = [];  % Triangular boolean matrix for each batch entry indicating whether each position 
                           % is within the lattice structure for that entry.
         supervisionWeights = [];  % Multiplicative weighting factors to apply to the cost/gradient for connections.
+        slantInputs = [];
 
         rawLeftEdgeEmbedding = [];   % The untransformed versions of the extra word embeddings that are fed to the 
         rawRightEdgeEmbedding = [];  % scorer at the edges of each sentence.
@@ -66,6 +67,7 @@ classdef LatticeBatch < handle
             lb.connectionLabels = fZeros([lb.B, lb.N - 1], hyperParams.gpu); %% deleted last
             lb.activeNode = fZeros([lb.B, lb.N, lb.N], hyperParams.gpu);
             lb.supervisionWeights = fOnes([lb.B, lb.N], hyperParams.gpu);
+            lb.slantInputs = fZeros([lb.N - 1, lb.B, lb.N - 1], hyperParams.gpu);
             lb.rawLeftEdgeEmbedding = wordFeatures(:, hyperParams.sentenceStartWordIndex);
             lb.rawRightEdgeEmbedding = wordFeatures(:, hyperParams.sentenceEndWordIndex);
             lb.leftEdgeEmbedding = fZeros([lb.D, 1], hyperParams.gpu);
@@ -157,6 +159,12 @@ classdef LatticeBatch < handle
                         % Dimensions: hiddenD x B
                         lb.scorerHiddenLayer(:, :, col, row) = hyperParams.compNL(connectionMatrix * scorerHiddenInputs);
                         lb.scores(col, :, row) = scoringVector * [ones(1, lb.B, 'like', lb.scorerHiddenLayer); lb.scorerHiddenLayer(:, :, col, row)];
+                    end
+
+                    if hyperParams.latticeSlant > 0
+                        % Use a slant layer to prefer left-side merges
+                        lb.slantInputs(1:row, :, row) = lb.scores(1:row, :, row);
+                        lb.scores(1:row, :, row) = ComputeSlantLayer(lb.scores(1:row, :, row), hyperParams.latticeSlant);
                     end
 
                     % Softmax the scores.
@@ -350,6 +358,11 @@ classdef LatticeBatch < handle
 
                     % Overwrite 0/0 deltas from inactive nodes.
                     deltasToScores(isnan(deltasToScores)) = 0;
+
[…]

### 2015-05-10 LoadSSTData.m

@@ -96,7 +96,7 @@ function [ data ] = ProcessAndSave(rawData, wordMap, lastSave, nextItemNo, filen
         end
     else
         data = repmat(struct('label', 0, 'sentence', Sequence()), numElements, 1);
-        parfor dataInd = 1:numElements
+        for dataInd = 1:numElements
             data(dataInd).sentence = Sequence.makeSequence(rawData(dataInd).sentenceText, wordMap, ...
                 hyperParams.parensInSequences, hyperParams.gpu && ~hyperParams.largeVocabMode);
             data(dataInd).label = rawData(dataInd).label;

### 2015-05-14 writing/SNLI paper/newcorpus.tex
Misc cleanup, new ALCIR feature.

@@
 
+\begin{table}
+\center
+  %\setlength{\tabcolsep}{15pt}
+  %\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{l l} 
+    \toprule
+Training pairs &  550,157\\
+Dev pairs &  10,000\\
+Test pairs &  10,000\\
+Total & 570,157\\
+\midrule
+L. mean token count & 14.1\\
+R. mean token count & 8.3 \\
+\midrule
+L. complete sentences & 70.2\%\\
+R. complete sentences & 75.4\%\\
+    \bottomrule
+  \end{tabular}
+\caption{\label{b-table}Key statistics for the raw sentence pairs. Since the left and right halves of each pair were collected separately, we report statistics for both.} 
+\end{table}
+
+\begin{table}
+\center
+  %\setlength{\tabcolsep}{15pt}
+  %\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{l l} 
+    \toprule
+Validated pairs \& 53,852
+Consensus label $\ne$ original label \& 8.5\%
+No consensus label $\ne$ \& 2.1\%
+Unanimous consensus \& 57.7\%
+
+    \bottomrule
+  \end{tabular}
+\caption{\label{b-table}Key statistics for the validated sentence pairs.} 
+\end{table}
+
+Total multiply annotated pairs: 53852
+Fraction where gold != original or no gold label: 0.0851593255589
+Fraction where gold != original: 0.0645844165491
+Fraction with no gold label: 0.0205749090099
+Fraction with unanimous agreement: 0.576561687588
+%% annotator agreement with gold labels: 0.889371871517
+%% annotator agreement with original labels: 0.855424084595
+
+\begin{table}
+\center
+  %\setlength{\tabcolsep}{15pt}
+  %\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{l lll} 
+    \toprule
+\textbf{Label} & \textbf{Train} & \textbf{Dev} & \textbf{Test}\\
+\midrule
+\ii{entailment} &33.3 & 33.3 & 33.0 \\
+\ii{neutral} & 33.2 & 32.7 & 32.5 \\
+\ii{contradiction} & 33.3 & 32.2 & 32.8 \\
+no gold label & 0.1 & 1.8 & 1.8 \\
+\bottomrule
+  \end{tabular}
+  % Dist. of each label in each data set.
+\caption{\label{b-table}Frequency (\%) of each label in the corpus. Note that unlabeled examples are omitted in standard evaluations.} 
+\end{table}
+
+No sentence overlap btw train and test and dev.
+
+Mention antonyms for motivation against unsup tasks
+
+% LHS length distribution: {1: 1, 2: 42, 3: 156, 4: 1095, 5: 3882, 6: 12120, 7: 26514, 8: 37434, 9: 44028, 10: 49245, 11: 50919, 12: 48363, 13: 43314, 14: 38121, 15: 33183, 16: 27621, 17: 23250, 18: 20247, 19: 18513, 20: 16386, 21: 13746, 22: 12066, 23: 9183, 24: 7131, 25: 6198, 26: 5007, 27: 3963, 28: 3438, 29: 2631, 30: 1959, 31: 1956, 32: 1434, 33: 1086, 34: 912, 35: 897, 36: 774, 37: 453, 38: 618, 39: 291, 40: 330, 41: 249, 42: 180, 43: 225, 44: 162, 45: 108, 46: 87, 47: 60, 48: 36, 49: 90, 50: 21, 51: 66, 52: 51, 53: 36, 54: 24, 55: 63, 56: 18, 57: 15, 58: 6, 59: 27, 60: 6, 61: 3, 62: 3, 63: 3, 64: 6, 65: 3, 66: 3, 67: 6, 68: 6, 69: 18, 70: 15, 71: 3, 72: 15, 73: 3, 75: 15, 79: 3, 82: 15}
+
+% RHS length distribution: {1: 39, 2: 1011, 3: 7980, 4: 29471, 5: 61196, 6: 74094, 7: 93600, 8: 85851, 9: 61359, 10: 46711, 11: 33241, 12: 22844, 13: 15994, 14: 11047, 15: 7601, 16: 5312, 17: 3732, 18: 2631, 19: 1878, 20: 1325, 21: 911, 22: 642, 23: 449, 24: 357, 25: 217, 26: 168, 27: 138, 28: 84, 29: 67, 30: 46, 31: 26, 32: 31, 33: 23, 34: 16, 35: 19, 36: 8, 37: 12, 38: 4, 39: 5, 40: 2, 41: 4, 42: 2, 43: 1, 44: 1, 48: 1, 51: 1, 55: 2, 56: 1, 60: 1, 62: 1}
+
 \paragraph{data partition}

### 2015-05-14 RunExperiments/RunSSTExperiments.sh

@@ -66,11 +66,11 @@ export MATLABCMD="cd quant; lambda = 0.000001; dim = 35; ed = 200; td = 2; penul
 export MATLABCMD="cd quant; lambda = 0.000001; dim = 35; ed = 200; td = 2; penult = 0; dropout = [0.9\, 0.9]; tot = 4; collo = 1; dataflag='sst-expanded'; name='/scr/sbowman/sst-expanded-5d'; TrainModel(''\, 1\, @SST\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1\, 1\, 0\, 1);" ; qsub -v MATLABCMD quant/run.sh -q john -l nodes=1:ppn=6
 export MATLABCMD="cd quant; lambda = 0.000001; dim = 75; ed = 200; td = 2; penult = 0; dropout = [0.9\, 0.9]; tot = 4; collo = 1; dataflag='sst-expanded'; name='/scr/sbowman/sst-expanded-5d'; TrainModel(''\, 1\, @SST\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1\, 1\, 0\, 1);" ; qsub -v MATLABCMD quant/run.sh -q john -l nodes=1:ppn=6
 
-export MATLABCMD="cd quant; lambda = 0.0001; dim = 40; ed = 300; td = 2; penult = 0; dropout = [1\, 1]; tot = 4; collo = 3; dataflag='sst-expanded'; name='/scr/sbowman/sst-expanded-5d'; TrainModel(''\, 1\, @SST\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 4\, 0\, 4\, .3);" ; qsub -v MATLABCMD quant/run.sh -q john -l nodes=1:ppn=6
[…]

### 2015-05-14 config/SST.m

@@ -97,7 +97,7 @@ options.testFreq = 250;
 options.detailedStatFreq = 250;
 
 if strcmp(dataflag, 'sst-expanded')
-    wordMap = InitializeMaps('./sst-data/sst-words.txt');
+    wordMap = LoadWordMap('./sst-data/sst-words.txt');
     hyperParams.vocabName = 'sst'; 
 
     hyperParams.numLabels = [5];
@@ -116,7 +116,7 @@ if strcmp(dataflag, 'sst-expanded')
     hyperParams.labelCostMultipliers = [4.878182632, 2.433623131, 0.3014847996, 1.826731877, 3.980980277];
 
 elseif strcmp(dataflag, 'sst')
-    wordMap = InitializeMaps('./sst-data/sst-words.txt');
+    wordMap = LoadWordMap('./sst-data/sst-words.txt');
     hyperParams.vocabName = 'sst'; 
 
     hyperParams.numLabels = [5];

### 2015-05-14 config/SynsetRelations.m

@@ -37,7 +37,7 @@ options.lr = lr;
 
 hyperParams.numLabels = [3]; 
 
-wordMap = InitializeMaps('synset-labels/longer_wordlist.txt');
+wordMap = LoadWordMap('synset-labels/longer_wordlist.txt');
 hyperParams.vocabName = 'synset'; 
 
 hyperParams.labels = {{'hypernym', 'hyponym', 'coordinate'}};

### 2015-05-14 config/Defaults.m
Misc cleanup, new ALCIR feature.

@@ -70,6 +70,10 @@ hyperParams.lambda = 0;
 hyperParams.bottomDropout = 1;
 hyperParams.topDropout = 1;
 
+% If this vector is populated with embedding indices, in ComputeSentenceClassificationExampleCostAndGrad,
+% the embeddings at those indices will be reinitialized before computing each example. Used for ALCIR.
+hyperParams.randomEmbeddingIndices = [];
+
 % L1 v. L2 regularization. If no regularization is needed, set
 % lambda to 0 and ignore this parameter.
 hyperParams.norm = 2;

### 2015-05-14 config/Quantifiers.m

@@ -21,7 +21,7 @@ hyperParams.topDepth = top;
 
 hyperParams.topDropout = tdrop;
 
-wordMap = InitializeMaps('./quantifiers/wordlist.tsv'); 
+wordMap = LoadWordMap('./quantifiers/wordlist.tsv'); 
 hyperParams.vocabName = 'quantifiers'
 
 hyperParams.labels = {{'#', '=', '>', '<', '|', '^', 'v'}};

### 2015-05-14 writing/SNLI paper/corpus.tex
Paper template

--- 2015-05-14 writing/SNLI paper/newcorpus.tex
@@
 \section{A new corpus for natural language inference}\label{sec:discussion}
+
+\subsection{Grounding with imagined images}
+
+\subsection{Data collection}
+
+\subsection{Data validation}
+
+\subsection{Statistics}
+
+Abstract (CP)
+
+Intro (CP)
+
+The corpus
+- Why we use our wacky sort of grounding (CP)
+- MTurk prompts, validation strategy, and other details about collection (Sam)
+- Statistics (Sam)
+
+Using the corpus as a platform for evaluation
+- Non-neural baselines: methods and results (Gabor)
+- RTE system(s): methods and results (Gabor)
+- An NN architecture for evaluating sentence embedding models through entailment (Sam)
+- NN model evaluations: methods and results (Sam)
+- Discussion (Sam)
 
@@
     \toprule
-Validated pairs \& 53,852
-Consensus label $\ne$ original label \& 8.5\%
-No consensus label $\ne$ \& 2.1\%
-Unanimous consensus \& 57.7\%
-
+Validated pairs \& 53,852\\
+Consensus label $\ne$ original label \& 8.5\%\\
+No consensus label $\ne$ \& 2.1\%\\
+Unanimous consensus \& 57.7\%\\
     \bottomrule

### 2015-05-14 writing/SNLI paper/entailment_eval.tex
Paper template

\section{SNLI as a platform for NLI research}

\subsection{Baselines}

\subsection{Existing NLI systems}

\subsection{Sentence embeddings and NLI}

\subsection{Results}

### 2015-05-17 writing/SNLI paper/corpus.tex
Paper push &c.

--- 2015-05-14 writing/SNLI paper/newcorpus.tex
@@
 We claim that the task of natural language inference (also called recognizing textual entailment, or RTE) is significantly more demanding, and that strong performance on this task is good evidence of a model's overall strength in sentence representation.
+
+\subsection{Grounding with imagined images}
+
+Quote from SICK paper \cite{marelli2014sick}:
+
+\begin{quote}
+Not unreasonably, subjects found that, say, \ii{A woman is wearing an Egyptian headdress} does not contradict \ii{A woman is wearing an Indian headdress}, since one could easily imagine both sentences truthfully uttered to refer to a single scene where two different women are wearing different headdresses.
+In the future, a higher proportion of CONTRADICTION labels could be elicited by using grammatical and possibly visual cues (pictures) encouraging co-indexing of the entities in the two sentences.
+\end{quote}
+
+\subsection{Data collection}
+
+\subsection{Data validation}
+
+\subsection{Statistics}
 
@@
 Our classifier is simply a stack of three 100d (NONLIN) layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
-
-quote from SICK paper \cite{marelli2014sick}:
-
-\begin{quote}
-Not unreasonably, subjects found that, say, \ii{A woman is wearing an Egyptian headdress} does not contradict \ii{A woman is wearing an Indian headdress}, since one could easily imagine both sentences truthfully uttered to refer to a single scene where two different women are wearing different headdresses.
-In the future, a higher proportion of CONTRADICTION labels could be elicited by using grammatical and possibly visual cues (pictures) encouraging co-indexing of the entities in the two sentences.
-\end{quote}
 
@@
 
-\paragraph{validation}
+\paragraph{Validation}
 
-The majority of the sentence pairs in the corpus have not been reviewed by anyone since their construction as part of the data collection Mechanical Turk task, except to remove blank responses and responses where the original author flagged the source caption for review.
-However, in order to measure the accuracy of our corpus, and in order to improve the quality of our test and development sets, we did perform an additional round of validation for TODOk examples.
+The majority of the sentence pairs in the corpus have not been reviewed by anyone since their construction as part of the data collection Mechanical Turk task, except to remove blank responses and responses some pairs which the original author flagged for review.
+However, in order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we did perform an additional round of validation for about 54k examples.
 
-This validation phase is not fundamentally different from the Mechanical Turk labeling task used for the SICK entailment data: we present workers with pairs of sentences (in our case, in batches of five per screenfull), and ask them to choose a single label for each pair.
+This validation phase is not fundamentally different from the Mechanical Turk labeling task used for the SICK entailment data: we present workers with pairs of sentences (in our case, in batches of five at a time), and ask them to choose a single label for each pair.
 We supplied each pair to four annonators, yielding five labels per pair including the label used by the original composer of the pair.
+
+The results of this validation process are summarized in Fig.~\ref{validation-stats}.
+For each pair that we validated, we chose a gold label.
+If any one of the three labels was chosen by at least three of the five annotators, that label is the gold label.
+If there was no such consensus, which occurred in about 0.2\% of cases, we chose the label `-'.
+While these unlabeled examples are included in the corpus distribution, they are unlikely to be helpful for the core classification task of NLI, and we
 
@@
 
+\begin{figure}
+\center
+\includegraphics[width=3.05in]{length_dist}
+    % From 1.0rc2
+\caption{\label{b-table}The distribution of sentence length.} 
+\end{figure}
+
 \begin{table}
@@
     \toprule
-Training pairs &  550,157\\
-Dev pairs &  10,000\\
-Test pairs &  10,000\\
-Total & 570,157\\
+Pairs in corpus & 570,157\\
+Training pairs &  550,159\\
+Dev pairs &  9,999\\
+Test pairs &  9,999\\
 \midrule
-L. mean token count & 14.1\\
-R. mean token count & 8.3 \\
+Premise mean token count & 14.1\\
+Hypothesis mean token count & 8.3 \\
 \midrule
-L. complete sentences & 70.2\%\\
-R. complete sentences & 75.4\%\\
+Premise complete sentences & 70.2 ??\%\\
+Hypothesis complete sentences & 75.4 ??\%\\
     \bottomrule
   \end{tabular}
-\caption{\label{b-table}Key statistics for the raw sentence pairs. Since the left and right halves of each pair were collected separately, we report statistics for both.} 
+    % From 1.0rc2
+\caption{\label{validation-stats}Key statistics for the raw sentence pairs. Since the two halves of each pair were collected separately, we report statistics for both.} 
 \end{table}
@@
     \toprule
-Validated pairs \& 53,852
-Consensus label $\ne$ original label \& 8.5\%
-No consensus label $\ne$ \& 2.1\%
-Unanimous consensus \& 57.7\%
-
+Validated pairs & 53,852\\
+gold label $\ne$ original label & 8.5\%\\
+Unanimous gold & 57.7\%\\
+No gold (no 3 labels match) & 2.1\%\\
+Label $=$ gold & 88.9\%\\
+Label $=$ original author's label & 85.5\%\\
     \bottomrule
   \end{tabular}
+    % From 1.0rc2
 \caption{\label{b-table}Key statistics for the validated sentence pairs.} 
 \end{table}
-
-Total multiply annotated pairs: 53852
-Fraction where gold != original or no gold label: 0.0851593255589
-Fraction where gold != original: 0.0645844165491
-Fraction with no gold label: 0.0205749090099
-Fraction with unanimous agreement: 0.576561687588
-%% annotator agreement with gold labels: 0.889371871517
-%% annotator agreement with original labels: 0.855424084595
 
@@
 \midrule
-\ii{entailment} &33.3 & 33.3 & 33.0 \\
-\ii{neutral} & 33.2 & 32.7 & 32.5 \\
-\ii{contradiction} & 33.3 & 32.2 & 32.8 \\
-no gold label & 0.1 & 1.8 & 1.8 \\
+\ii{entailment} &33.3 & 33.4 & 33.3 \\
+\ii{neutral} & 33.2 & 33.2 & 33.2 \\
+\ii{contradiction} & 33.3 & 33.2 & 33.3 \\
+- (no consensus) & 0.2 & 0.2 & 0.2 \\
 \bottomrule
   \end{tabular}
-  % Dist. of each label in each data set.
-\caption{\label{b-table}Frequency (\%) of each label in the corpus. Note that unlabeled examples are omitted in standard evaluations.} 
+  % From 1.0rc2
+\caption{\label{b-table}Frequency (\%) of each label in the corpus. Note that examples without gold labels are omitted in standard evaluations.} 
 \end{table}
@@
 
-\paragraph{data partition}
+\paragraph{Partition} We distribute the corpus with a pre-specified train/test/development split.
+The test and development sets contain 10k examples each.
+Each original ImageFlickr caption occurs in only one of the three sets, and all of the examples in the test and development sets have gone through validation.
 
-\paragraph{simple baselines}
-
-MFC
-
-BLEU
-
-Summing + NN
+\paragraph{General evaluation standards}
+While we hope that the corpus will be valuable in a variety of ways, we encourage researchers working on tools for semantic representation and inference to evaluate on our data in a uniform way: training on only the (parsed and/or unparsed) sentences included in the training set, and doing final evaluations on only the subset of the test set for which there are single gold labels.

### 2015-05-17 RunExperiments/RunSNLIExperiments.sh
Paper push &c.

@@ -130,16 +130,31 @@ export MATLABCMD="cd quant; lambda = 0.00001; dim = 40; ed = 200; td = 3; penult
 export MATLABCMD="cd quant; lambda = 0.00001; dim = 40; ed = 200; td = 3; penult = 25; dropout = [0.5\, 0.5]; tot = 2; collo = 1; dataflag='snli096-only'; name='/scr/sbowman/snli096'; TrainModel(''\, 1\, @SNLI\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1);" ; qsub -v MATLABCMD quant/run.sh -q john -l mem=31gb
 
 
-export MATLABCMD="cd quant; lambda = 0.0001; dim = 50; ed = 200; td = 3; penult = 100; dropout = [0.9\, 0.9]; tot = 2; collo = 1; dataflag='snlirc1-only'; name='/scr/sbowman/snlirc'; TrainModel(''\, 1\, @SNLI\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1);" ; qsub -v MATLABCMD quant/run.sh -q john -l mem=31gb
-export MATLABCMD="cd quant; lambda = 0.0000001; dim = 50; ed = 200; td = 3; penult = 100; dropout = [0.9\, 0.9]; tot = 2; collo = 1; dataflag='snlirc1-only'; name='/scr/sbowman/snlirc'; TrainModel(''\, 1\, @SNLI\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1);" ; qsub -v MATLABCMD quant/run.sh -q john -l mem=31gb
+export MATLABCMD="cd quant; lambda = 0.0001; dim = 50; ed = 200; td = 3; penult = 100; dropout = [0.9\, 0.9]; tot = 2; collo = 1; dataflag='snlirc2-only'; name='/scr/sbowman/snlirc'; TrainModel(''\, 1\, @SNLI\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1);" ; qsub -v MATLABCMD quant/run.sh -q john -l mem=31gb
[…]

### 2015-05-18 writing/SNLI paper/entailment_eval.tex
Beginnings of baseline classifiers section

@@
 \section{SNLI as a platform for NLI research}
+\todo{write me}
 
+%
+% Baselines
+%
 \subsection{Baselines}
+We evaluate a number of strong baselines on the SNLI corpus.
+The first class of baselines use models from the Excitement Open
+  Platform (EOP) 
+  \cite{pado2014design,magnini2014excitement}
+  -- an open source platform for RTE research which
+  is distributed alongside a number of RTE pipelines.
+We additionally evaluate against a both an unlexicalized and lexicalized
+  classifier with straightforward feature sets.
+\todo{some meaningful insight from \#s}
 
-\subsection{Existing NLI systems}
+%
+% EOP
+%
+\paragraph{Excitement Open Platform}
+% what is EOP
+The Excitement Open Platform is a tool to quickly develop RTE systems
+  while sharing components such as common lexical resources and 
+  evaluation sets.
+% 9 systems compared against
+A number of systems have been built using the platform, 9 of them
+  applicable to English are publicly distributed with version 1.2.1
+  of the software.
+% these fall into 2 classes
+These fall into two classes of algorithms: 2 are edit distance based,
+  whereas the remaining 7 make use of different features in a
+  maximum entropy classifier.
+
+% methodology
+We convert the 3-way classification task in SNLI into the RTE setting
+  by labeling both the \unknown\ and \contradiction\
+  labels as negative entailment, and treating the \entailment\ label as
+  the positive entailment.
+This creates a biased dataset of 66\% negative examples.
+% we report the best results from each class
+We run each of the 9 algorithms distributed with EOP on the 2-class
+  SNLI dataset, and report results for the best edit distance 
+  configuration and the best classifier based configuration, as
+  determined by performance on the development set.
+\todo{we may not have CPU time for this}.
+
+%
+% EOP RESULTS TABLE
+%
+% Some definitions
+\def\t#1{\small{#1}}
+\def\b#1{\t{\textbf{#1}}}
+\def\colspaceS{2.0mm}
+\def\colspaceM{3.0mm}
+\def\colspaceL{4.0mm}
+
+% The table
+\begin{table}
+\begin{center}
+\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceL}c@{\hskip \colspaceS}c}
+\hline
+\textbf{System} & \multicolumn{4}{c}{\b{SNLI}} & \multicolumn{2}{c}{\b{RTE-3}} \\
+ & \t{P} & \t{R} & \t{F$_1$} & \t{Acc.} & \t{F$_1$} & \t{Acc.} \\
+\hline
+\t{Edit Distance} & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} & 
+                    \t{??.?} & \t{??.?} \\
+\t{Classifier}    & \t{68.0} & \t{38.7} & \t{49.3} & \t{73.5} & 
+                    \t{64.2} & \t{65.3} \\
+\hline
+\end{tabular}
+\end{center}
+% The caption
+\caption{
+\label{tab:eopresults}
+2-class accuracy and precision/recall of existing competitive RTE
+systems from the Excitement Open Platform.
+Results are reported for an edit-distance based algorithm, and a
+  classifier-based system.
+For comparision, results are also reported on the RTE-3 dataset.
+\todo{numbers are approx.}
+}
+\end{table}
+%
+% END EOP RESULTS TABLE
+%
+
+% the best systems
+We report results in \reftab{eopresults}.
+The best edit distance algorithm tunes the the weight of the three 
+  case-insensitive edit distance operations on the training set, 
+  after removing stop words.
+The best classifier-based system \todo{does something}.
+\todo{some insight from results}.
+
+%
+% Lexicalized Classifier
+%
+\paragraph{Lexicalized Classifier}
+Unlike the RTE datasets, SNLI's size allows approaches which make use of
+  rich lexicalized features.
+We implmement a lexicalized classifier to provide a strong baseline for future work
+  on the dataset.
+Our classifier implements 6 features:
+\begin{enumerate}  % TODO(gabor) This takes up way too much space
+\setlength\itemsep{0em}
+  \item the BLEU score of the \hypothesis\ with respect
+  to the \premise, using an n-gram length between 1 and 4.
+
+  \item The length difference between the \hypothesis\ and the \premise, as a real-valued
+  feature.
+
+  \item The overlap between words in the \premise\ and \hypothesis. This is counted
+  both as an absolute count and a percentage of possible overlap, and is computed
+  both over all words and over just nouns, verbs, adjectives, and adverbs.
+
+  \item For every pair of words across the \premise\ and \hypothesis\ which share a 
+  POS tag, an indicator feature over the two words.
+
+  \item For every pair of bigrams across the \premise\ and \hypothesis\ which share a 
+  POS tag, an indicator feature over the two bigrams.
+
+  \item An indicator for every unigram and bigram in the \hypothesis.
+\end{enumerate}
+
+%
+% BOW RESULTS TABLE
+%
+% Some definitions
+\def\t#1{\small{#1}}
+\def\b#1{\t{\textbf{#1}}}
+\def\colspaceS{2.0mm}
+\def\colspaceM{3.0mm}
+\def\colspaceL{4.0mm}
+
+% The table
+\begin{table}
+\begin{center}
+\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceL}c}
+\hline
+\textbf{System} & \multicolumn{4}{c}{\b{2-Class}} & \b{3-Class} \\
+ & \t{P} & \t{R} & \t{F$_1$} & \t{Acc.} & \t{Acc.} \\
+\hline
+\t{Lexicalized}            & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} & \t{77.3} \\
+\t{Unigrams Only}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
+\t{Unlexicalized}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
+\hline
+\end{tabular}
+\end{center}
+% The caption
+\caption{
+\label{tab:bowresults}
+Accuracy on the 2-class and 3-class SNLI dataset, as well as precision/recall
+  on the 2-class task to allow for comparision with the RTE systems.
+Ablation studies are reported for removing bigram features and all lexical
+  features.
+\todo{numbers are approx.}
+}
+\end{table}
+%
+% END BOW RESULTS TABLE
+%
+
+\Fig{learning_curves_bow.pdf}{0.6}{bowlearncurve}{
+A learning curve for the lexicalized and unlexicalized baseline classifiers.
+Note the y axis starting at a random-chance accuracy of 33\%.
+\todo{numbers are approx.}
+}
+
+% Results
+We report results in \reftab{bowresults}, along with ablation studies for removing
+  the bigram features (leaving only the unigram entailment pair feature),
+  and for removing all lexicalized features.
+In addition, we report learning curves for the lexicalized and fully unlexicalized
+  classifiers in \reffig{bowlearncurve}.
+
+Note that \todo{something interesting happens}.
+
+%\subsection{Existing NLI systems}
 

### 2015-05-18 writing/F'14 paper/arxiv/intro.tex

--- 2015-02-28 writing/F'14 paper/intro.tex
@@
     \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
+    (strict) entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
+    (strict) reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
     equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
@@
   \end{tabular}
-  \caption{\label{b-table}The seven relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.} %-%
+  \caption{\label{b-table}The seven relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.
+The relations are defined so as to be mutually exclusive.} %-%
 

### 2015-05-18 writing/F'14 paper/arxiv/quantifiers.tex

--- 2015-02-28 writing/F'14 paper/quantifiers.tex
@@
 between them.  
-Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.
+For instance, our models might then see
+pairs like \eqref{p1} and \eqref{p2} in training and be required to 
+then label \eqref{p3}.
+
+\vspace{-0.6cm}
+\begin{gather}
+  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
+  \\
+  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
+  \\
+  \text{(most turtle) reptile} \natalt \text{(all turtle) (not animal)}\label{p3}
+\end{gather}
 
@@
 
-In each run, we sample 20\% of the 60k generated pairs to test on and 
-train on the remainder. In this setting, each model must learn a
-complete reasoning system for the limited language and logic presented
-in the training data, and must be able
-to recognize all of the lexical relations between the nouns, verbs,
-and quantifiers and how they interact. For instance, it might see
-pairs like \eqref{p1} and \eqref{p2} in training and be required to 
-then label \eqref{p3}.
-
-\vspace{-0.6cm}
-\begin{gather}
-  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
-  \\
-  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
-  \\
-  \text{(most turtle) reptile} \natalt \text{(all turtle) (not animal)}\label{p3}
-\end{gather}
+In each run, we randomly partition the set of valid \textit{single sentences} under the grammar into training and test, and then label all of the pairs from within each set to generate a training set of 27k pairs and a test set of 7k pairs.
+Because the model doesn't see the test sentences at training time, it cannot directly use the kind of reasoning described in \S\ref{sec:join}, and must instead both infer the word-level relations and learn a complete reasoning system over them for our logic.
 
@@
     $\natind$ only &	35.4 & (7.5)	& 35.4	& (7.5)\\
-    25d SumNN	&	97.0&	(97.2)&	89.6&	(91.1)\\	
-    25d TreeRNN	&	\textbf{100}&	\textbf{(100)}&	\textbf{99.4}&	\textbf{(99.5)}\\
-    25d TreeRNTN	&	99.4&(99.5)&	99.4 & (99.2)\\
+    25d SumNN	&	96.9&	(97.7)&	93.9&	(95.0)\\	
+    25d TreeRNN	&	99.6&	(99.6)&	99.2&	(99.3)\\
+    25d TreeRNTN	&	\textbf{100}&(\textbf{100})&	\textbf{99.7} & \textbf{(99.5)}\\
     \bottomrule
@@
 \paragraph{Results} The results (Table~\ref{qresultstable}) show that both tree models are able to learn to generalize the underlying logic almost perfectly.
-The baseline summing model can largely memorize the training data, but does not generalize well.
+The baseline summing model can largely memorize the training data, but does not generalize as well.
 We do not find any consistent pattern in the handful of errors made by either tree model, and no errors were consistent across model restarts, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

### 2015-05-18 writing/SNLI paper/corpus.tex
Paper update.

@@
 
-need for new corpus
+% TODO: Rewrite and expand motivation here (CP?)
 
@@
 \subsection{Grounding with imagined images}
+
+% TODO: Rewrite and expand (CP?)
 
@@
 
-\subsection{Data validation}
+We used Amazon Mechanical Turk for data collection.
+In each individual task, a worker was presented with premise scene descriptions from a preexisting corpus, and asked to supply hypotheses that fit each of our three labels (\ii{entailment}, \ii{neutral}, and \ii{contradiction}).
 
-\subsection{Statistics}
+For the premises, we used the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions from that corpus (corresponding to about 30k images) collected in an earlier crowdsourced effort.
+The captions were not authored by the photographers who took the source images, they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: ``Our trip to the Olympic Peninsula'').
+We presented each caption to at least one worker, and some to more than one.
+With the exception of a handful that were flagged by workers as excessively difficult, at least three labeled pairs are available for each caption.
 
-what's new
-- size
-- grounding
-
-data will be freely available
-
-SICK presented an important challenge, we try to deliver on it.
-
-aim to surpass sick in both size and r consistency
-
-Since our goal is to evaluate the quality of the single sentence representations in our model, we train an entailment classifier that has access only to the output of our sentence representation model, and not to any of the word or phrase representations that it used.
-We encourage future users of this corpus to evaluate their models in this way when possible.
-Our classifier is simply a stack of three 100d (NONLIN) layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
-
-We use captions from the Flickr30k corpus \cite{hodoshimage} as the left hand side of each relation pair. That corpus ...
-
-We randomly selected individual captions from that corpus to use in our own corpus, though in the interest of encouraging more complex source sentences without discarding too much source data, we downsample sentences with lengths of under 90 characters so that these sentences make up only half of our corpus, rather than the original 69\%.
-
-... licencing ...
-
-The instructions for this phase were integrated into the design of the data collection interface.
-The text of those instructions is shown in Figure~\ref{instructions-1}.
-We chose to use three classes, corresponding to ENTAILMENT, CONTRADICTION, and NEUTRAL classes used in SICK.
+The instructions that we provided to the workers are shown in Fig.~\ref{instructions-1}.
+Below the instructions were three fields for each of three requested sentences, corresponding to our \ii{entailment}, \ii{neutral}, and \ii{contradiction} labels, a fourth field (marked optional) for reporting problems, and a link to an FAQ page.
+That FAQ grew over the course of data collection, but discussed disallowed techniques (e.g., reusing the same sentence for many different prompts, which we saw in a few cases), sentence length and complexity (we did not enforce a minimum length, and we allowed bare NPs as well as full sentences), and logistical issues around payment timing.
 
@@
 \end{itemize}
-\textbf{Problems} (optional)   If something is wrong with the caption that makes it difficult to understand, do your best above and let us know here.
-% \bottomrule
-\caption{\label{instructions-1}Examples of training data from the newly collected entailment corpus.}
+\caption{\label{instructions-1}The instructions used on Mechanical Turk for data collection.}
 \end{figure}
 
-We posted 15.5k ImageFlickr captions.
-We showed each of the first 10k captions to five workers, yielding fifteen reponse sentences per source caption, with five assigned to each of the three labels.
-Then, to ensure maximal diversity in the test set, we showed the remaining 5.5k source captions to one worker each,  yielding only three responses per source caption.
+% TODO: Do we need statistics about the number/distribution of workers?
+% TODO: Do we need statistics about how much of ImageFlickr we rejected/double-annotated?
 
-We collected 143,023 pairs, after excluding less than 100 worker responses in which the worker reported that they could not understand the source caption, or in which a field was accidentally left blank.
-The average worker response sentence was 7.8 words long.
-1100 workers participated in the task.
-
-pre-verification distribution:
-
-Recompute after removing blanks:
-29305 contra
-29304 neut
-29305 ent
-
-% ditto examples
-%
-\begin{table*}
-  \centering\footnotesize
-  \begin{tabular}{p{6.5cm}cp{6.5cm}}
-  \toprule
-A busy street with building lined up and people walking down the street outside and nighttime. &\ii{entailment}	&People lined up\\
-\rule{0pt}{3ex}Two men stand in an electric outdoor lift. &\ii{neutral}	& The two men are inspectors for the FDA, waiting to inspect the second level of the pharmaceutical plant.\\
-\rule{0pt}{3ex}A young redheaded girl, wearing a yellow shirt, black pants, and sneakers, jumping in a grassy field with blue skies and wispy clouds in the background.
-&\ii{entailment}	& A girl jumps in a grassy field\\
-\rule{0pt}{3ex}A crowd of people, one with a guitar, are standing in the street under the 7th Avenue street sign. &\ii{neutral}	& The crowd listening to the man play guitar.\\
-\rule{0pt}{3ex}A woman wearing bike shorts and a skirt is riding a bike and carrying a shoulder bag.  &\ii{contradiction}& A woman driving a car.\\
-    \bottomrule
-  \end{tabular}
-  \caption{\label{examplesofscedata}The instructions presented to workers during data collection.}
-\end{table*}
-
-% Screenshot from phase 1
-
-% Screenshot from phase 2
-
-The data is freely available at (POST DATA), and is released under a CreativeCommons
-Attribution-ShareAlike licence, which is the same licence used for the Flickr30k source captions.
-
-\paragraph{Validation}
+\subsection{Data validation}
 
 The majority of the sentence pairs in the corpus have not been reviewed by anyone since their construction as part of the data collection Mechanical Turk task, except to remove blank responses and responses some pairs which the original author flagged for review.
-However, in order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we did perform an additional round of validation for about 54k examples.
+However, in order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we did perform an additional round of validation for about 54k (TODO) examples.
 
@@
 We supplied each pair to four annonators, yielding five labels per pair including the label used by the original composer of the pair.
+The instructions were broadly similar to the instructions for initial data collection shown in Fig.~\ref{instructions-1}, and linked to a similar FAQ.
 
@@
 
-\begin{figure}
-\footnotesize
-%\toprule
-TODO: VALIDATION INSTRUCTIONS
-%\bottomrule\\
+\begin{table*}
+  \centering\footnotesize
+  \begin{tabular}{p{7.5cm}lp{5.5cm}}
+  \toprule
+A young man in a striped hoodie looks at an artistic Levi's advertisement that says "We are all workers." & \ii{entailment} & There is a man in a hoodie looking at an advertisement.\\
+\rule{0pt}{3ex}Two men on laptops in a dark room. & \ii{entailment} & Two men are on their laptops.\\
+\rule{0pt}{3ex}A naked baby and toddler smear each other in paint. & \ii{neutral} & The kids are in the bathtub.\\
+\rule{0pt}{3ex}A woman is focusing on her game of curling. & \ii{neutral} & There is a women playing a game indoors\\
+\rule{0pt}{3ex}A group of people dressed in green and white are dancing in a circle outside of a white building. & \ii{contradiction} & A group performs indoors.\\
+\rule{0pt}{3ex}A balding professor giving a lecture to students on the topic of "opportunity". & \ii{neutral} & A female professor with hair giving a lecture to students.\\
+% TODO: Replace with *validated* examples from the post-rc2 training set, include all 5 labels
+    \bottomrule
+% From 1.0rc2
+  \end{tabular}
+  \caption{\label{snli-examples}Randomly chosen examples from SNLI.}
+\end{table*}
 
-\caption{\label{instructions-2}The instructions presented to workers during data validation.}
-\end{figure}
+\subsection{The distributed corpus}
+
+% Screenshot from phase 1
+
+% Screenshot from phase 2
+
+The data is freely available at (POST DATA), and is released under a CreativeCommons
+Attribution-ShareAlike licence, which is the same licence used for the Flickr30k source captions.
 
@@
 
-No sentence overlap btw train and test and dev.
-
-Mention antonyms for motivation against unsup tasks
-
-% LHS length distribution: {1: 1, 2: 42, 3: 156, 4: 1095, 5: 3882, 6: 12120, 7: 26514, 8: 37434, 9: 44028, 10: 49245, 11: 50919, 12: 48363, 13: 43314, 14: 38121, 15: 33183, 16: 27621, 17: 23250, 18: 20247, 19: 18513, 20: 16386, 21: 13746, 22: 12066, 23: 9183, 24: 7131, 25: 6198, 26: 5007, 27: 3963, 28: 3438, 29: 2631, 30: 1959, 31: 1956, 32: 1434, 33: 1086, 34: 912, 35: 897, 36: 774, 37: 453, 38: 618, 39: 291, 40: 330, 41: 249, 42: 180, 43: 225, 44: 162, 45: 108, 46: 87, 47: 60, 48: 36, 49: 90, 50: 21, 51: 66, 52: 51, 53: 36, 54: 24, 55: 63, 56: 18, 57: 15, 58: 6, 59: 27, 60: 6, 61: 3, 62: 3, 63: 3, 64: 6, 65: 3, 66: 3, 67: 6, 68: 6, 69: 18, 70: 15, 71: 3, 72: 15, 73: 3, 75: 15, 79: 3, 82: 15}
-
-% RHS length distribution: {1: 39, 2: 1011, 3: 7980, 4: 29471, 5: 61196, 6: 74094, 7: 93600, 8: 85851, 9: 61359, 10: 46711, 11: 33241, 12: 22844, 13: 15994, 14: 11047, 15: 7601, 16: 5312, 17: 3732, 18: 2631, 19: 1878, 20: 1325, 21: 911, 22: 642, 23: 449, 24: 357, 25: 217, 26: 168, 27: 138, 28: 84, 29: 67, 30: 46, 31: 26, 32: 31, 33: 23, 34: 16, 35: 19, 36: 8, 37: 12, 38: 4, 39: 5, 40: 2, 41: 4, 42: 2, 43: 1, 44: 1, 48: 1, 51: 1, 55: 2, 56: 1, 60: 1, 62: 1}
-
 \paragraph{Partition} We distribute the corpus with a pre-specified train/test/development split.
@@
 
+\paragraph{Parses}
+
+The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser \cite{klein2003accurate}, both in the standard PTB format and in the binarized unlabeled format used by tree-structured neural network models, as in \S\ref{sentence-embedding}.
+
+% TODO (CM?)
+
 \paragraph{General evaluation standards}

### 2015-05-18 writing/SNLI paper/entailment_eval.tex
Paper update.

--- 2015-05-14 writing/SNLI paper/entailment_eval.tex
@@
 \section{SNLI as a platform for NLI research}
+
+Since our goal is to evaluate the quality of the single sentence representations in our model, we train an entailment classifier that has access only to the output of our sentence representation model, and not to any of the word or phrase representations that it used.
+We encourage future users of this corpus to evaluate their models in this way when possible.
+Our classifier is simply a stack of three 100d (NONLIN) layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
 
@@
 
-\subsection{Sentence embeddings and NLI}
+\subsection{Sentence embeddings and NLI}\label{sentence-embedding}
+\input{model_structure_fig.tex}
 

### 2015-05-18 writing/SNLI paper/model_structure_fig.tex
Paper update.

--- 2015-05-03 writing/AAAI SSS '15/BowmanPottsManning/BowmanPottsManning.tex
@@
-\def\year{2015}
-%File: formatting-instruction.tex
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-
-\usepackage[leqno, fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{qtree}
-\usepackage[numbers]{natbib}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{colortbl}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{xcolor}
-
-\frenchspacing
-\setlength{\pdfpagewidth}{8.5in}
-\setlength{\pdfpageheight}{11in}
-\pdfinfo{
-/Title Learning Distributed Word Representations for Natural Logic Reasoning
-/Author Samuel R. Bowman, Christopher Potts, Christopher D. Manning
-/Keywords distributional semantics, vector space models, natural logic, natural language processing, neural networks}
-\setcounter{secnumdepth}{1}
-
-\newcommand{\nateq}{\equiv}
-\newcommand{\natind}{\mathbin{\#}}
-%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
-\newcommand{\natneg}{\mathbin{^{\wedge}}}
-\newcommand{\natfor}{\sqsubset}
-\newcommand{\natrev}{\sqsupset}
-\newcommand{\natalt}{\mathbin{|}}
-\newcommand{\natcov}{\mathbin{\smallsmile}}
-
-\newcommand{\plneg}{\mathop{\textit{not}}}
-\newcommand{\pland}{\mathbin{\textit{and}}}
-\newcommand{\plor}{\mathbin{\textit{or}}}
-
-% Strikeout
-\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
-\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}
-
-\newcommand{\word}[1]{\emph{#1}}
-\def\ii#1{\textit{#1}}
-
- \begin{document}
-% The file aaai.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-\title{Learning Distributed Word Representations for Natural Logic Reasoning}
-\author{Samuel R. Bowman, Christopher Potts, and Christopher D. Manning\\
-  Stanford University\\
-  450 Serra Mall \\
-  Stanford, CA 94305-2150\\
-  \texttt{\{sbowman,cgpotts,manning\}@stanford.edu}} %TODO: Three affiliations?
-\maketitle
-\begin{abstract}
-\begin{quote}
-  Natural logic offers a powerful labelal conception of meaning
-  that is a natural counterpart to distributed semantic
-  representations, which have proven valuable in a wide range of
-  sophisticated language tasks. However, it remains an open question
-  whether it is possible to train distributed representations to
-  support the rich, diverse logical reasoning captured by natural
-  logic. We address this question using two neural network-based
-  models for learning embeddings: plain neural networks and neural
-  tensor networks. Our experiments evaluate the models' ability to
-  learn the basic algebra of natural logic labels from simulated
-  data and from the WordNet noun graph.  The overall positive results
-  are promising for the future of learned distributed representations in
-  the applied modeling of logical semantics.
-\end{quote}
-\end{abstract}
-
-Natural logic offers a powerful \emph{labelal} conception of
-semantics: the meanings for expressions are given, at least in part,
-by their inferential connections with other expressions
-\cite{Katz72,maccartney2009extended,vanBenthem08NATLOG}. For instance,
-\word{turtle} is analyzed, not primarily by its extension in the
-world, but rather by its lexical network: it entails \word{reptile},
-excludes \word{chair}, is entailed by \word{sea
-  turtle}, and so forth. With generalized notions of entailment and
-contradiction, these labelships can be defined for all lexical
-categories as well as complex phrases, sentences, and even texts. The
-resulting theories of meaning offer valuable new analytic tools for
-tasks involving database inference, label extraction, and textual
-entailment.
-
-Natural logic aligns well with distributed (e.g., vector)
-representations, which also naturally model meaning labelally.
-Distributed representations have been used successfully in a wide
-array of sophisticated language tasks, including part of speach tagging,
-\cite{collobert2011natural}, semantic role labeling \cite{collobert2011natural},
-sentiment analysis \cite{socher2011semi}, and translation \cite{sutskever2014sequence} 
-among many others.
-However, it remains an open question whether it is possible to train
-such representations to support the rich, diverse logical reasoning
-captured by natural logic; while they excel at synonymy (similarity),
-the results are more mixed for entailment, contradiction, and mutual
-consistency.  Using the natural logic of \cite{maccartney2009extended}
-as our formal model, we address this open question using two neural
-network-based models for learning embeddings: plain neural networks
-and neural tensor networks (NTNs).  The natural logic is built from
-the seven labels defined in Table~\ref{b-table}. Its formal
-properties are now well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}, so it
-provides a rigorous set of goals for our neural models. To keep the
-discussion manageable, we limit attention to experiments involving the
-lexicon; for a more extended treatment of complex expressions
-involving logical connectives and quantifiers, see
-\cite{Bowman:Potts:Manning:2014}. % CITET
-
-In our experiments, we evaluate these models' ability to learn the
-basic algebra of natural logic labels from simulated data and from
-the WordNet noun graph. The simulated data help us to achieve analytic
-insights into what the models learn, and the WordNet data show how they
-fare with a real natural language vocabulary.  We find that only the NTN is able to fully
-learn the underlying algebra, but that both models excel in the 
-WordNet experiment.
-
-\begin{table*}[tp]
-  \centering%\small
-  \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{l c l l} 
-    \toprule
-    Name & Symbol & Set-theoretic definition & Example \\ 
-    \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle $\natfor$ reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile $\natrev$ turtle}  \\ 
-    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch $\nateq$ sofa} \\ 
-    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle $\natalt$ warthog} \\ 
-    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able $\natneg$ unable} \\
-    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal $\natcov$ non-turtle} \\ 
-    independence       & $x \natind y$   & (else) & \ii{turtle $\natind$ pet}\\
-    \bottomrule
-  \end{tabular}
-  \protect\caption{\protect\label{b-table}The seven natural logic labels of \protect\cite{maccartney2009extended}. 
-    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the label $\natind$ applies whenever none of the other six do.} 
-\end{table*}
-
-\section{Neural network models for label classification} \label{methods}
-
-We build embedding-based models using the method of
-\cite{Bowman:Potts:Manning:2014}, which is centered on the task of
-labeling a pair of words or sentences with one of a small set of
-logical labels. Unlike in a classical inference
-setting, the model is not given specific premises to reason with at test time, but rather
-is expected to encode all of the information that it learns about each term (word) at training
-time in the term embedding vectors, and to then use only the information encoded in the relevant pair of
-term vectors at test time. 
-The architecture that we use is depicted in
-Figure~\ref{sample-figure}. The model represents the two input terms
-as single embedding vectors, which are fed into a comparison function based on one
-of two types of neural network layer function to produce a
-representation for the labelship between the two terms. This
-representation is then fed into a softmax classifier, which outputs a
-distribution over possible labels. The entire network, including the
-embeddings, is trained using backpropagation with AdaGrad \cite{duchi2011adaptive}
-and L2 regularization.
-
-The simpler version of the comparison concatenates the two
-input vectors before feeding them into a LReLU \cite{maasrectifier} neural network (NN)
-layer.  The more powerful neural tensor network (NTN) version adds an
-additional third-order tensor parameter to allow for multiplicative
-interactions between the two inputs \cite{chen2013learning}. For more
-details on the implementation and training of the layer functions, see
-\cite{Bowman:Potts:Manning:2014}.
-
 \begin{figure}[tp]
@@
 
-	  \labeledtreenode[2.4]{6.5,5}{$P(\sqsubset) = 0.8$}{softmaxcolor}  
-	  \put(6.5,3.7){\vector(0,1){1}}  
-	  \labeledtreenode[5]{6.5,3.4}{turtle \emph{vs.}~animal}{comparisoncolor}
-
-	  \textlabel{2.75,5}{Softmax classifier}{black}
-	  \textlabel{2.0,3.4}{\parbox{2cm}{Comparison N(T)N layer}}{black}
-
-	  \textlabel{6.5,2.1}{Optional embedding}{black}
-	  \textlabel{6.5,1.5}{transformation NN layers}{black}
-
-	  \textlabel{6.5,0.5}{Learned term vectors}{black}
+	  \labeledtreenode[5]{6.5,6.5}{softmax classifier}{softmaxcolor}  
+	  \put(6.5,5.25){\vector(0,1){1}}  
+	  \labeledtreenode[5]{6.5,5.25}{$\tanh$ layer}{comparisoncolor}
+	  \put(6.5,4){\vector(0,1){1}}  
+	  \labeledtreenode[5]{6.5,4}{$\tanh$ layer}{comparisoncolor}	 
+	  \put(6.5,2.75){\vector(0,1){1}}  
+           \labeledtreenode[5]{6.5,2.75}{$\tanh$ layer}{comparisoncolor}
 
@@
 
-	  \put(1.5,0.75){\vector(0,1){0.85}}
-	  \labeledtreenode{1.5,0.5}{turtle}{lexcolor}
-
-	  \put(1.5,2.25){\vector(4,1){3.25}}
-	  \labeledtreenode{1.5,1.9}{turtle}{compositioncolor}
+	  \put(3,1.5){\vector(2,1){0.85}}
+	  \labeledtreenode{1.5,1.5}{turtle}{lexcolor}
 
@@
 
-	  \put(11.5,0.75){\vector(0,1){0.85}}
-	  \labeledtreenode{11.5,0.5}{animal}{lexcolor}
-
-	  \put(11.5,2.25){\vector(-4,1){3.25}}
-	  \labeledtreenode{11.5,1.9}{animal}{compositioncolor}
-
 	\end{picture}
 	%}
-        \caption{The model structure used to compare \ii{turtle} and
-          \ii{animal}.  Learned term representations are fed through
-          either an NN or NTN comparison layer and then to a softmax
-        classifier over the seven labels in Table~\ref{b-table}.}
+        \caption{The standard model structure used for entailment evaluation.}
   \label{sample-figure}
 \end{figure}
-
-This model used here differs from the one described in that work in
-two ways. First, because the inputs are single terms, we do not use
-a composition function here. Second, for our experiment on WordNet data, 
-we introduce an additional $\tanh$ neural network layer between the embedding 
-input and the comparison function, which facilitates initializing the 
-embeddings themselves from pretrained vectors and was found to help 
-performance in that setting.
-
-%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?
-
-\section{Reasoning about semantic labels}\label{sec:join}
-
-In this experiment, we test our model's ability to learn and use the
-natural logic inference rules schematized in
-Figure~\ref{tab:jointable}. For instance, given that $a \natrev b$ and
-$b \natrev c$, one can conclude that $a \natrev c$, by basic
-set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  There
-are 32 valid inferences that can be made on the basis of pairs of
-labels of the form $a R b$ and $b R' c$. Cells in the table
-containing a dot correspond to pairs of labels for which no valid
-inference can be drawn in our logic.
-
-% about the labels themselves that do not depend on the
-% internal structure of the things being compared. For example, given
-% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
-% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven labels support more
-% than just transitivity: MacCartney and Manning's
-% \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of labels of the form $a R
-% b$ and $b R' c$, including several less intuitive ones such as that if
-% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-
-\paragraph{Experiments}
-To test our models' ability to learn these inferential patterns, we
-create artificial boolean structures in which terms denote sets of entities from
-a small domain (e.g., Figure~\ref{lattice-figure}), employ logical
-deduction to identify the valid statements, divide those into train
-and test sets, and remove from the test set those statements which
-cannot be proven from the training statements
-(Figure~\ref{unprovable}). 
-% using the logic depicted in Figure~\ref{lattice-figure}.
-%
-%
-% We test the model's ability to learn this behavior by creating
-% artificial data sets of terms which represent sets of numbers. Since
-% MacCartney and Manning's set of labels hold between sets as well as
-% between sentences, we can use the underlying set structure to generate
-% the all of the labels that hold between any pair of these terms, as
-% in Figure \ref{lattice-figure}. We train the model defined above on a
-% subset of these labels, but rather then presenting the model with a
-% pair of tree-structured sentences as inputs, simply present it with
-% two single terms, each of which corresponds to a single vector in the
-% (randomly initialized) vocabulary matrix $V$, ensuring that the model
-% has no information about the terms being compared except the labels
-% between them.
-%
-In our experiments, we create 80 randomly generated sets drawn from a
-domain of seven elements. This yields 6400 statements about pairs of
-formulae, of which 3200 are chosen as a test set. 240 of the test examples 
-were excluded from the primary test set for lack of evidence in the training 
-data to support the choice of any one label. 
-We then trained and evaluated both the NN model and the NTN model on these data sets.
-
-\paragraph{Results} 
-We found that the NTN model worked best with 11-dimensional vector
-representations for the 80 sets and a 90-dimensional feature vector
-for the classifier, though the performance of the model was not highly
-dependant on either dimensionality setting. 
-Averaging over five randomly generated data sets, the model was able to correctly label 98.1\% (standard error $0.67\%$) of the provably derivable test examples, and 87.7\%
-($\textit{SE} = 3.59\%$) of the remaining test examples. The simpler NN worked
-best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 84.8\% ($\textit{SE} = 0.84\%$) and 68.7\% ($\textit{SE} = 1.58\%$),
-respectively. Training accuracy was 99.8\% ($\textit{SE} = 0.04\%$) for the NTN and 94.7\% ($\textit{SE} = 0.89\%$) for
-the NN.
-
-\begin{figure*}[tp]
-  \centering
-  \begin{subfigure}[b]{0.42\textwidth}
-    \centering  
-    \setlength{\arraycolsep}{8pt}
-    \renewcommand{\arraystretch}{1.1}
-    \newcommand{\UNK}{\cdot}  
-    \resizebox{2.4in}{!}{$\begin{array}[c]{c@{ \ }|*{7}{c}|}
-        % \hline
-        \multicolumn{1}{c}{}
-        & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
-        \cline{2-8}
-        \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
-        \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
-        \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
-        \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
-        \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
-        \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
-        \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
-        \cline{2-8}
-      \end{array}$}
-    \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the label that holds between $a$ and $c$, if any.
-These inferences are based on basic set-theoretic truths about the meanings of the underlying labels as described in Table~\ref{b-table}.
-We assess our models' ability to reproduce such inferential paths.}
-    \label{tab:jointable}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
-    \setlength{\unitlength}{1cm}
-    \resizebox{1.5in}{!}{\begin{picture}(5,5.5)
-      \labelednode{2.50}{5}{}{1,2,3}
-
-      \put(0.75,4){\line(3,1){1.5}}
-      \put(2.5,4){\line(0,1){0.5}}
-      \put(4.25,4){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{3.5}{a,b}{1,2}
-      \labelednode{2.50}{3.5}{c}{1,3}
-      \labelednode{4.25}{3.5}{d}{2,3}
-
-      \put(0.75,2.5){\line(0,1){0.5}}
-      \put(0.75,2.5){\line(3,1){1.5}}
-
-      \put(2.5,2.5){\line(-3,1){1.5}}
-      \put(2.5,2.5){\line(3,1){1.5}}
-
-      \put(4.25,2.5){\line(0,1){0.5}}
-      \put(4.25,2.5){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{2}{e,f}{1}
-      \labelednode{2.50}{2}{}{2}
-      \labelednode{4.25}{2}{g,h}{3}
-
-      \put(2.5,1){\line(-3,1){1.5}}
-      \put(2.5,1){\line(0,1){0.5}}
-      \put(2.5,1){\line(3,1){1.5}}
-
-      \labelednode{2.5}{0.5}{}{}
-    \end{picture}}
-    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
-    some sets have multiple names, so that learning $\nateq$ is non-trivial.}\label{lattice-figure}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.2\textwidth}
-    \centering
-    \setlength{\tabcolsep}{12pt}
-    \begin{tabular}[b]{c  c}
-      \toprule
-      Train & Test \\
-      \midrule
-
-      $a \nateq b$              & $b \nateq b$ \\
-      $a \natfor e$              & $b \natfor e$ \\
-      $a \natneg g$              & \strikeout{$e \nateq f$} \\
-      $d \natfor h$              & \strikeout{$g \natrev d$} \\
-      $e \natalt g$            & $h \natrev d$ \\ 
-      $\vdots$		& $\vdots$	\\
-      \bottomrule
-    \end{tabular}
-
-    \caption{A train/test split of the atomic statements about the
-      model.  Test statements not provable from the training data are
-      crossed out.}\label{unprovable}
-  \end{subfigure}  
-  \caption{Experimental goal and set-up for reasoning about semantic labels.}
-  \label{exp1}
-\end{figure*} 
-% Note: None of these test examples is derivable from the shown training data, 
-% but we suggest that there is additional training data, so we can cross lines
-% out willy-nilly without fear.
-
-% RNTN log: tue-j-11-6x80-hip.txt
-% Train PER: 0.0021875
-
-% RNN log: tue-j-11-6x80-r-75.txt
-% Train PER: 0.047188
-
-% and create a dataset consisting of the labels between every pair of
-% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
-% test dataset, and that test dataset was further split into the 2960
-% examples that can be provably derived from the test data using
-% MacCartney and Manning's join table (or by the symmetry of the
-% labels in about half of the cases) and the 240 that
-% cannot.
-
-% We tested a version of both the RNN model and the RNTN model on these
-% data. In both cases, the models were implemented exactly as described
-% in Section~\ref{methods}, but since the items being compared are
-% single terms rather than full tree structures, the composition layer
-% was not used, and the two models differed only in which layer function
-% was used for the comparison layer. We found that the RNTN model worked
-% best with 11 dimensional vector representations for the 80 sets and a
-% 90 dimensional feature vector for the classifier. This model was able
-% to correctly label 99.3\% of the derivable test examples, and 99.1\%
-% of the remaining examples. The simpler RNN model worked best with 11
-% and 75 dimensions, respectively, but was able to achieve accuracies of
-% only 90.0\% and 87.\%, respectively.
-
-These results are fairly straightforward to interpret. The NTN model
-was able to accurately encode the labels between the terms in the
-geometric labels between their vectors, and was able to then use
-that information to recover labels that were not overtly included
-in the training data. In contrast, the NN was able to achieve this
-behavior only incompletely. It is possible but not likely that it
-could be made to find a good solution with further optimization on
-different learning algorithms, or that it would do better on a larger
-universe of sets, for which there would be a larger set of training
-data to learn from, but the NTN is readily able to achieve these
-effects in the setting discussed here.
-
-It's noteworthy that both models performed well above chance on the
-unprovable examples. This indicates that both were able to identify
-statistical cues to the properties of the underlying boolean
-structure, thereby going beyond what one could obtain from the natural
-logic. Information of this kind is not sufficient to yield the
-near-perfect accuracy that we saw on the provable data, but it does
-highlight one of the ways in which trained inference models can turn
-out to be more powerful that rigidly defined proof systems.
-
-\newcommand{\nodagger}{\phantom{$^\dagger$}}
-
-\begin{table*}
-\centering%\resizebox{5.5in}{!}
-{
-  \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{rrrrrrr} 
-    \toprule
-     Portion of & \multicolumn{2}{c}{25d NN} & \multicolumn{2}{c}{25d NTN} & \multicolumn{1}{c}{$\natfor$ only}\\
-     training data  & \multicolumn{1}{c}{w/ GloVe} & \multicolumn{1}{c}{w/o GloVe} & \multicolumn{1}{c}{w/ GloVe} & \multicolumn{1}{c}{w/o GloVe} \\
-    \midrule
-     100\%   & 99.73 & 99.37   & 99.61  & \textbf{99.95}  & 37.05 \\
-     33\%    & 95.96 & 95.30    & 95.83  & 95.45  & 37.05 		\\
-     11\%    & 91.11 & 90.81   & 91.27  & 90.90  & 37.05 		\\
-    \bottomrule
-  \end{tabular}}
- \caption{Mean test \% accuracy scores (with standard error) on the WordNet data over five-fold crossvalidation.
-The baseline figure is the frequency of the most frequent class, \ii{hypernym}.\label{wn-table}}
-\end{table*}
-
-\section{Reasoning about lexical labels in WordNet}\label{sec:wordnet}
-
-Using simulated data as above is reassuring about what the models
-learn and why, but we also want to know how they perform with a real
-natural language vocabulary. Unfortunately, as far as we are aware,
-there are no available resources labeling such a vocabulary with the
-labels from Table~\ref{b-table}. However, the labels in WordNet
-\cite{fellbaum2010wordnet,WordNet95} come close and pose the same substantive
-challenges within a somewhat easier classification problem.
-
-% WordNet \cite{fellbaum2010wordnet} provides a different type of
-% lattice structure which is compatible with our learning paradigm.
-
-We extract three types of label from WordNet. \ii{Hypernym} and
-\ii{hyponym} can be represented directly in the WordNet graph
-structure, and correspond closely to the $\natrev$ and $\natfor$
-labels from natural logic. As in natural logic, these labels are
-mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
-(perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then
-\ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
-terms, which share a direct hypernym, like \ii{dalmatian}, \ii{pug},
-and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
-terms tend to exclude one another, thereby providing a loose approximation 
-of the natural logic exclusion label $\natalt$. WordNet's antonym label, however, is both too narrowly defined
-and too rare to use for this purpose. WordNet defines its labels over
-sets of synonyms, rather than over individual terms, so we do not
-include a \ii{synonym} or \ii{equivalent} label, but rather
-consider only one member of each set of synonyms. Word pairs which do
-not fall into these three labels are not included in the data set.
-
-To limit the size of the vocabulary without otherwise simplifying the learning problem, we extract all of the
-instances of these three labels for single word nouns in WordNet that are hyponyms of the node 
-\texttt{organism.n.01}. In order to balance the distribution of the classes, we slightly downsample instances 
-of the \ii{coordinate} label, yielding a total of 36,772 labels among 3,217 terms.
-We report results below using crossvalidation, choosing a disjoint 10\% test sample for each of five runs.
-Unlike in the previous experiment,
-it is not straightforward here to determine in advance
-how much data should be required to train an accurate model, so we performed training runs with 
-various fractions of the remaining data. Embeddings were fixed at 25 dimensions and were initialized 
-randomly or using distributional vectors from GloVe \cite{pennington2014glove}. The feature vector 
-produced by the comparison layer was fixed at 80 dimensions.
-
-\paragraph{Results} 
-The structured WordNet data contains substantially more redundancy than the random artificial data above, and we find that the NTN performs perfectly with random initialization.
-The plain NN performs almost as well, providing
-a point of contrast with the results of Section~\ref{sec:join}. We also find that initialization with GloVe is helpful in allowing the models to maintain fair performance
-with smaller amounts of training data. Some of the randomly initialized model runs failed to learn
-usable representations at all and labeled all examples with the most frequent labels.
-We excluded these runs from the statistics, but marked settings for which this occurred with the symbol $\dagger$.
-For all of the remaining runs, training accuracy was consistently above 99\%.
-
-\section{Conclusion}\label{sec:discussion}
-
-This investigation builds upon extensive prior research on
-constructing compositional word meanings in vector spaces
-\cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
-extracting entailment information from distributional vector space
-models \cite{baroni2012entailment,rei2014looking}, and transferring
-information between logic programs and neural networks in constrained
-settings \cite{hitzler2004logic,holldobler1999approximating}. Our
-primary contribution is to establish a tight connection between these
-approaches and formal developments within natural logic.  More
-specifically, we showed that neural network models, optimized using
-established techniques, can learn distributed representations that
-closely approximate the behavior one would expect from a natural logic
-inference algorithm.
-
-% Ours is not the first investigation into logical behavior in vector space models: 
-%  there is extensive prior work on constructing compositional word meanings in vector spaces \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
-%  extracting entailment information from distributional vector space models \cite{baroni2012entailment,rei2014looking},
-%  and transferring information between logic programs and neural networks in 
-%  constrained settings \cite{holldobler1999approximating,hitzler2004logic}. However, this investigation
-%   is the first that we are aware of on logical behavior in representations learned within conventionally
-%    optimized neural network models.
-
-This paper evaluated two neural models on the task of learning 
-natural logic labels between distributed word representations. The
-results suggest that at least the neural tensor network
-has the capacity to
-meet this challenge with reasonably-sized training sets, learning both 
-to embed a vocabulary in a way that encodes a diverse
-set of labels, and to subsequently use those embeddings to infer new
-labels. In
-\cite{Bowman:Potts:Manning:2014}, we extend these results to include
-complex expressions involving logical connectives and quantifiers, with similar
-conclusions about (tree-structured recursive versions of) these models. These findings
-are promising for the future of learned distributed representations in the
-applied modeling of logical semantics.
-
-\section{Acknowledgments}
-
-We thank Jeffrey Pennington, Richard Socher, and audiences at CSLI, Nuance, and BayLearn.
-
-\begin{thebibliography}{}
-
-\bibitem[\protect\citeauthoryear{Baroni \bgroup et al\mbox.\egroup
-  }{2012}]{baroni2012entailment}
-Baroni, M.; Bernardi, R.; Do, N.-Q.; and Shan, C.-c.
-\newblock 2012.
-\newblock Entailment above the word level in distributional semantics.
-\newblock In {\em Proc. EACL}.
-
-\bibitem[\protect\citeauthoryear{Bowman, Potts, and
-  Manning}{2014}]{Bowman:Potts:Manning:2014}
-Bowman, S.; Potts, C.; and Manning, C.
-\newblock 2014.
-\newblock Recursive neural networks can learn logical semantics.
-\newblock {arXiv} manuscript 1406.1827.
-
-\bibitem[\protect\citeauthoryear{Chen \bgroup et al\mbox.\egroup
-  }{2013}]{chen2013learning}
-Chen, D.; Socher, R.; Manning, C.; and Ng, A.
-\newblock 2013.
-\newblock Learning new facts from knowledge bases with neural tensor networks
-  and semantic word vectors.
-\newblock In {\em Proc. {ICLR}}.
-
-\bibitem[\protect\citeauthoryear{Clark, Coecke, and
-  Sadrzadeh}{2011}]{ClarkCoeckeSadrzadeh2011}
-Clark, S.; Coecke, B.; and Sadrzadeh, M.
-\newblock 2011.
-\newblock Mathematical foundations for a compositional distributed model of
-  meaning.
-\newblock {\em Linguistic Analysis} 36(1--4):345--384.
-
-\bibitem[\protect\citeauthoryear{Collobert \bgroup et al\mbox.\egroup
-  }{2011}]{collobert2011natural}
-Collobert, R.; Weston, J.; Bottou, L.; Karlen, M.; Kavukcuoglu, K.; and Kuksa,
-  P.
-\newblock 2011.
-\newblock Natural language processing (almost) from scratch.
-\newblock {\em JLMR} 12.
-
-\bibitem[\protect\citeauthoryear{Duchi, Hazan, and
-  Singer}{2011}]{duchi2011adaptive}
-Duchi, J.; Hazan, E.; and Singer, Y.
-\newblock 2011.
-\newblock Adaptive subgradient methods for online learning and stochastic
-  optimization.
-\newblock {\em JMLR}.
-
-\bibitem[\protect\citeauthoryear{Fellbaum}{2010}]{fellbaum2010wordnet}
-Fellbaum, C.
-\newblock 2010.
-\newblock Word{N}et.
-\newblock In {\em Theory and Applications of Ontology: Computer Applications}.
-  Springer.
-
-\bibitem[\protect\citeauthoryear{Grefenstette}{2013}]{grefenstette2013towards}
-Grefenstette, E.
-\newblock 2013.
-\newblock Towards a formal distributional semantics: Simulating logical calculi
-  with tensors.
-\newblock arXiv manuscript 1304.5823.
-
-\bibitem[\protect\citeauthoryear{Hermann, Grefenstette, and
-  Blunsom}{2013}]{Hermann-etal:2013}
-Hermann, K.; Grefenstette, E.; and Blunsom, P.
-\newblock 2013.
-\newblock ``{N}ot not bad'' is not ``bad'': A distributional account of
-  negation.
-\newblock In {\em Proc. of the 2013 Workshop on Continuous Vector Space Models
-  and their Compositionality}.
-
-\bibitem[\protect\citeauthoryear{Hitzler, H{{\"o}}lldobler, and
-  Seda}{2004}]{hitzler2004logic}
-Hitzler, P.; H{{\"o}}lldobler, S.; and Seda, A.~K.
-\newblock 2004.
-\newblock Logic programs and connectionist networks.
-\newblock {\em Journal of Applied Logic} 2(3).
-
-\bibitem[\protect\citeauthoryear{H{\"o}lldobler, Kalinke, and
-  St{\"o}rr}{1999}]{holldobler1999approximating}
-H{\"o}lldobler, S.; Kalinke, Y.; and St{\"o}rr, H.-P.
-\newblock 1999.
-\newblock Approximating the semantics of logic programs by recurrent neural
-  networks.
-\newblock {\em Applied Intelligence} 11(1).
-
-\bibitem[\protect\citeauthoryear{Icard and Moss}{2013a}]{Icard:Moss:2013}
-Icard, T., and Moss, L.
-\newblock 2013a.
-\newblock A complete calculus of monotone and antitone higher-order functions.
-\newblock In {\em Proc. Topology, Algebra, and Categories in Logic}.
-
-\bibitem[\protect\citeauthoryear{Icard and Moss}{2013b}]{Icard:Moss:2013:LILT}
-Icard, T., and Moss, L.
-\newblock 2013b.
-\newblock Recent progress on monotonicity.
-\newblock {\em Linguistic Issues in Language Technology} 9(7).
-
-\bibitem[\protect\citeauthoryear{Katz}{1972}]{Katz72}
-Katz, J.~J.
-\newblock 1972.
-\newblock {\em Semantic Theory}.
-\newblock New York: Harper \& Row.
-
-\bibitem[\protect\citeauthoryear{Maas, Hannun, and Ng}{2013}]{maasrectifier}
-Maas, A.; Hannun, A.; and Ng, A.
-\newblock 2013.
-\newblock Rectifier nonlinearities improve neural network acoustic models.
-\newblock In {\em Proc. ICML}.
-
-\bibitem[\protect\citeauthoryear{MacCartney and
-  Manning}{2009}]{maccartney2009extended}
-MacCartney, B., and Manning, C.
-\newblock 2009.
-\newblock An extended model of natural logic.
-\newblock In {\em Proc. IWCS}.
-
-\bibitem[\protect\citeauthoryear{Miller}{1995}]{WordNet95}
-Miller, G.~A.
-\newblock 1995.
-\newblock Word{N}et: A lexical database for {E}nglish.
-\newblock {\em Communications of the {ACM}} 38(11):39--41.
-
-\bibitem[\protect\citeauthoryear{Pennington, Socher, and
-  Manning}{2014}]{pennington2014glove}
-Pennington, J.; Socher, R.; and Manning, C.
-\newblock 2014.
-\newblock Glo{V}e: {G}lobal vectors for word representation.
-\newblock In {\em Proc. EMNLP}.
-
-\bibitem[\protect\citeauthoryear{Rei and Briscoe}{2014}]{rei2014looking}
-Rei, M., and Briscoe, T.
-\newblock 2014.
-\newblock Looking for hyponyms in vector space.
-\newblock In {\em Proc. {CoNLL}}.
-
-\bibitem[\protect\citeauthoryear{Rockt{\"a}schel \bgroup et al\mbox.\egroup
-  }{2014}]{rocktaschellow}
-Rockt{\"a}schel, T.; Bosnjak, M.; Singh, S.; and Riedel, S.
-\newblock 2014.
-\newblock Low-dimensional embeddings of logic.
-\newblock In {\em Proc. the ACL 2014 Workshop on Semantic Parsing}.
-
-\bibitem[\protect\citeauthoryear{Socher \bgroup et al\mbox.\egroup
-  }{2011}]{socher2011semi}
-Socher, R.; Pennington, J.; Huang, E.; Ng, A.; and Manning, C.
-\newblock 2011.
-\newblock Semi-supervised recursive autoencoders for predicting sentiment
-  distributions.
-\newblock In {\em Proc. EMNLP}.
-
-\bibitem[\protect\citeauthoryear{Sutskever, Vinyals, and
-  Le}{2014}]{sutskever2014sequence}
-Sutskever, I.; Vinyals, O.; and Le, Q.
-\newblock 2014.
-\newblock Sequence to sequence learning with neural networks.
-\newblock In {\em Proc. {NIPS}}.
-
-\bibitem[\protect\citeauthoryear{van Benthem}{2008}]{vanBenthem08NATLOG}
-van Benthem, J.
-\newblock 2008.
-\newblock A brief history of natural logic.
-\newblock In Chakraborty, M.; L{\"o}we, B.; Nath~Mitra, M.; and Sarukki, S.,
-  eds., {\em Logic, Navya-Nyaya and Applications: Homage to {B}imal {M}atilal}.
-
-\end{thebibliography}
-\bibliographystyle{aaai}
-
-\end{document}

### 2015-05-18 writing/SNLI paper/entailment_eval.tex

--- 2015-05-18 writing/SNLI paper/entailment_eval.tex
@@
 
-%
-% Baselines
-%
+Since our goal is to evaluate the quality of the single sentence representations in our model, we train an entailment classifier that has access only to the output of our sentence representation model, and not to any of the word or phrase representations that it used.
+We encourage future users of this corpus to evaluate their models in this way when possible.
+Our classifier is simply a stack of three 100d (NONLIN) layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
+
 \subsection{Baselines}

### 2015-05-18 writing/SNLI paper/macros.tex

\def\entailment{\textit{entailment}}
\def\unknown{\textit{unknown}}
\def\contradiction{\textit{contradiction}}
\def\hypothesis{hypothesis}
\def\premise{premise}

\usepackage{color}
\newcommand{\hl}[2]{\colorbox{#2}{#1}}
\newcommand{\hly}[1]{\hl{yellow}{#1}}
\def\todo#1{\hl{{\bf TODO:} #1}{yellow}}
\def\needcite{\hl{{$^{\tt\small[citation\ needed]}$}}{blue}}
\def\needfig{\hl{Figure X}{green}}
\def\needtab{\hl{Table Y}{green}}

\newcommand\fig[1]{\begin{center} \includegraphics{#1} \end{center}}
\newcommand\Fig[4]{\begin{figure}[tb] \begin{center} \includegraphics[scale=#2]{#1} \end{center} \caption{\label{fig:#3} #4} \end{figure}}
\newcommand\FigTop[4]{\begin{figure}[t] \begin{center} \includegraphics[scale=#2]{#1} \end{center} \caption{\label{fig:#3} #4} \end{figure}}
\newcommand\FigStar[4]{\begin{figure*}[tb] \begin{center} \includegraphics[scale=#2]{#1} \end{center} \caption{\label{fig:#3} #4} \end{figure*}}

\newcommand\refeqn[1]{(\ref{eqn:#1})}
\newcommand\refeqns[2]{(\ref{eqn:#1}) and (\ref{eqn:#2})}
\newcommand\refchp[1]{Chapter~\ref{chp:#1}}
\newcommand\refsec[1]{Section~\ref{sec:#1}}
\newcommand\refsecs[2]{Sections~\ref{sec:#1} and~\ref{sec:#2}}
\newcommand\reffig[1]{Figure~\ref{fig:#1}}
\newcommand\reffigs[2]{Figures~\ref{fig:#1} and~\ref{fig:#2}}
\newcommand\reffigss[3]{Figures~\ref{fig:#1},~\ref{fig:#2}, and~\ref{fig:#3}}
\newcommand\reffigsss[4]{Figures~\ref{fig:#1},~\ref{fig:#2},~\ref{fig:#3}, and~\ref{fig:#4}}
\newcommand\reftab[1]{Table~\ref{tab:#1}}
\newcommand\refapp[1]{Appendix~\ref{sec:#1}}
\newcommand\refthm[1]{Theorem~\ref{thm:#1}}
\newcommand\refthms[2]{Theorems~\ref{thm:#1} and~\ref{thm:#2}}
\newcommand\reflem[1]{Lemma~\ref{lem:#1}}
\newcommand\reflems[2]{Lemmas~\ref{lem:#1} and~\ref{lem:#2}}
\newcommand\refprop[1]{Proposition~\ref{prop:#1}}
\newcommand\refdef[1]{Definition~\ref{def:#1}}
\newcommand\refcor[1]{Corollary~\ref{cor:#1}}
\newcommand\refalg[1]{Algorithm~\ref{alg:#1}}

### 2015-05-19 writing/SNLI paper/model_structure_fig.tex
Transfer update; new figure.

@@
   \centering
-    \footnotesize
 
-	\newcommand{\labeledtreenode}[4][3]{\put(#2){\makebox(0,0){{\fcolorbox{black}{#4}{\makebox(#1,0.3){#3}}}}}}
+ \begin{tikzpicture}
+    \def\dx{19pt}
+    \def\dy{29pt}
 
-	\newcommand{\textlabel}[4][3.5]{\put(#2){\makebox(0,0){{\makebox(#1,0.3){#3}}}}}
+    \tikzstyle{label}=[text width=40mm,align=center]    
+    \tikzstyle{softmax}=[fill=red!40,text width=40mm,align=center]
+    \tikzstyle{preclass}=[fill=orange!40,text width=50mm,align=center]
+    \tikzstyle{e}=[fill=green!40,text width=25mm,align=center]
+    \tikzstyle{m}=[fill=blue!40,text width=34mm,align=center]
 
-	\definecolor{lexcolor}{HTML}{F5F7C4}
-	\definecolor{compositioncolor}{HTML}{BBEBFF}
-	\definecolor{comparisoncolor}{HTML}{FFC895}
-	\definecolor{softmaxcolor}{HTML}{A5FF8A}
+    \node[softmax]  (softmax) at (0*\dx,6*\dy) {3-way softmax classifier};
+    \node[preclass]  (pc3) at (0*\dx,5*\dy) {100d $\tanh$ layer};
+    \node[preclass]  (pc2) at (0*\dx,4*\dy) {100d $\tanh$ layer};
+    \node[preclass]  (pc1) at (0*\dx,3*\dy) {100d $\tanh$ layer};
+    \node[e]  (pe) at (-3*\dx,1.85*\dy) {50d premise};
+    \node[e]  (he) at (3*\dx,1.85*\dy) {50d hypothesis};
+    \node[m]  (pem) at (-3*\dx,0.5*\dy) {embedding model\\ with premise input};
+    \node[m]  (hem) at (3*\dx,0.5*\dy) {embedding model\\ with hypothesis input};
 
-	\setlength{\unitlength}{0.61cm}
+    \pgfsetarrowsend{latex}
+    \tikzstyle{fwd} = [draw=black, line width=1pt]
 
-	%\resizebox{3in}{!}{
-	\begin{picture}(13.15, 5.5)
+          \draw [fwd] (pc3) -- (softmax);
+          \draw [fwd] (pc2) -- (pc3);
+          \draw [fwd] (pc1) -- (pc2);
+          \draw [fwd] (pe) -- (pc1);
+          \draw [fwd] (he) -- (pc1);
+          \draw [fwd] (hem) -- (he);
+          \draw [fwd] (pem) -- (pe);
 
-	  \labeledtreenode[5]{6.5,6.5}{softmax classifier}{softmaxcolor}  
-	  \put(6.5,5.25){\vector(0,1){1}}  
-	  \labeledtreenode[5]{6.5,5.25}{$\tanh$ layer}{comparisoncolor}
-	  \put(6.5,4){\vector(0,1){1}}  
-	  \labeledtreenode[5]{6.5,4}{$\tanh$ layer}{comparisoncolor}	 
-	  \put(6.5,2.75){\vector(0,1){1}}  
-           \labeledtreenode[5]{6.5,2.75}{$\tanh$ layer}{comparisoncolor}
+  \end{tikzpicture}
 
-	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-	  \put(3,1.5){\vector(2,1){0.85}}
-	  \labeledtreenode{1.5,1.5}{turtle}{lexcolor}
-
-	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-	\end{picture}
-	%}
         \caption{The standard model structure used for entailment evaluation.}

### 2015-05-19 ComputeSentenceClassificationExampleCostAndGrad.m

@@ -24,8 +24,14 @@ end
 DIM = hyperParams.dim;
 EMBDIM = hyperParams.embeddingDim;
 
+if isfield(hyperParams, 'smallVecs') && hyperParams.smallVecs
+    sigma = 0.25;
+else
+    sigma = 1;
+end
+
 wordFeatures(:, hyperParams.randomEmbeddingIndices) = ...
-    fNormrnd(0, 1, [EMBDIM, length(hyperParams.randomEmbeddingIndices)], ...
+    fNormrnd(0, sigma, [EMBDIM, length(hyperParams.randomEmbeddingIndices)], ...
              hyperParams.gpu, hyperParams.gpu && hyperParams.largeVocabMode);
 wordFeatures(1, hyperParams.randomEmbeddingIndices) = 1;
 

### 2015-05-19 RunExperiments/RunSickExperiments.sh

@@ -192,12 +192,13 @@ export MATLABCMD="cd quant; conCost = 20; lambda = 0.0005; dim = 30; ed = 200; t
 export MATLABCMD="cd quant; conCost = 5; lambda = 0.0005; dim = 50; ed = 200; td = 2; penult = 75; comp = 4; dropout = [0.75\, 0.75]; collo = 1; dataflag='dg-only-sst'; name='/scr/sbowman/updaterule4b'; TrainModel(''\, 1\, @Sick\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, comp\, dropout(1)\, dropout(2)\, collo\, 1\, conCost);" ; qsub -v MATLABCMD quant/run.sh -q john
 export MATLABCMD="cd quant; conCost = 5; lambda = 0.0005; dim = 50; ed = 200; td = 2; penult = 75; comp = 6; dropout = [0.75\, 0.75]; collo = 1; dataflag='dg-only-sst'; name='/scr/sbowman/updaterule4b'; TrainModel(''\, 1\, @Sick\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, comp\, dropout(1)\, dropout(2)\, collo\, 1\, conCost);" ; qsub -v MATLABCMD quant/run.sh -q john
 
-export MATLABCMD="cd quant; lambda = 0.0001; dim = 50; ed = 200; td = 3; penult = 100; tot = 1; dropout = [0.9\, 0.9]; collo = 1; dataflag='sick-only-dev'; name='transfer'; TrainModel('snlirc-snlirc2-only-l1e-07-dim50-ed200-td3-pen100-do0.9-0.9-co1-comp2-dp1/ckpt-best-tr150517185056@353375.mat'\, 1\, @Sick\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 0\, 1);" ; qsub -v MATLABCMD quant/run.sh -q john -l nodes=1:ppn=12
[…]

### 2015-05-19 config/SNLI.m

@@ -1,4 +1,4 @@
-function [ hyperParams, options, wordMap, labelMap ] = SNLI(expName, dataflag, embDim, dim, topDepth, penult, lambda, composition, bottomDropout, topDropout, collo, dp)
+function [ hyperParams, options, wordMap, labelMap ] = SNLI(expName, dataflag, embDim, dim, topDepth, penult, lambda, composition, bottomDropout, topDropout, collo, dp, ad)
 % Configuration for experiments involving the SemEval SICK challenge and ImageFlickr 30k. 
 
 [hyperParams, options] = Defaults();
@@ -9,7 +9,7 @@ hyperParams.name = [expName, '-', dataflag, '-l', num2str(lambda), '-dim', num2s
     '-ed', num2str(embDim), '-td', num2str(topDepth),...
     '-pen', num2str(penult), '-do', num2str(bottomDropout), '-', num2str(topDropout), '-co', num2str(collo),...
     '-comp', num2str(composition), ...
-    '-dp', num2str(dp)];
+    '-dp', num2str(dp), '-ad', num2str(ad)];
 
 hyperParams.parensInSequences = 0;
 
@@ -47,8 +47,9 @@ hyperParams.penultDim = penult;
 % Regularization coefficient.
 hyperParams.lambda = lambda; % 0.002 works?;
 
-% How many examples to run before taking a parameter update step on the accumulated gradients.
-options.miniBatchSize = 32;
+if ad
+    options.updateFn = @AdaDeltaUpdate;
+end
 
 % TODO:
 % hyperParams.maxDeltaNorm = dp;

### 2015-05-19 config/ALCIR.m
Transfer update; new figure.

@@ -11,7 +11,11 @@ hyperParams.sentenceClassificationMode = 1;
 
 if slant == -2
     hyperParams.randomEmbeddingIndices = [1, 2];
+elseif slant == -3
+    hyperParams.randomEmbeddingIndices = [1, 2];
+    hyperParams.smallVecs = 1;
 end
+
 
 hyperParams.dim = dim;
 hyperParams.embeddingDim = dim;
@@ -37,7 +41,7 @@ elseif slant == 4
 elseif slant == 5
     hyperParams.latticeFirstPastThreshold = 0.9;
     hyperParams.latticeFirstPastHardMax = true;
-elseif slant == 6
+elseif slant == 6 % Here and above works.
     hyperParams.latticeFirstPastThreshold = 0.95;
     hyperParams.latticeFirstPastHardMax = true;
 elseif slant == 7

### 2015-05-19 config/Sick.m

@@ -28,6 +28,8 @@ else
     hyperParams.vocabPath = ['../data/wordsource.scaled.' num2str(embDim) 'd.txt'];    
 end
 
+options.updateFn = @AdaDeltaUpdate;
+
 % The number of embedding transform layers. topDepth > 0 means NN layers will be
 % added above the embedding matrix. This is likely to only be useful when
 % learnWords is false, and so the embeddings do not exist in the same space
@@ -68,9 +70,9 @@ options.miniBatchSize = 32;
 % Amount to upsample SICK data.
 datamult = 8;
 
-if findstr(dataflag, 'sick-only-dev')
-    wordMap = LoadWordMap('sick-data/combined_words.txt');
-    hyperParams.vocabName = 'sick_all'; 
+if strcmp(dataflag, 'sick-only-dev')
+    wordMap = LoadWordMap('sick-data/sick_basic_words.txt');
+    hyperParams.vocabName = 'sick_all';
 
     hyperParams.numLabels = [3];
    	hyperParams.labels = {{'ENTAILMENT', 'CONTRADICTION', 'NEUTRAL'}};
@@ -84,7 +86,7 @@ if findstr(dataflag, 'sick-only-dev')
     				 './sick-data/SICK_trial_parsed_18plusparens.txt', ...
     				 './sick-data/SICK_trial_parsed_lt18_parens.txt'};
     hyperParams.splitFilenames = {};
-elseif findstr(dataflag, 'sick-only')
+elseif strcmp(dataflag, 'sick-only')
     % The number of labels.
     hyperParams.numLabels = [3];
 
@@ -92,15 +94,30 @@ elseif findstr(dataflag, 'sick-only')
     labelMap = cell(1, 1);
     labelMap{1} = containers.Map(hyperParams.labels{1}, 1:length(hyperParams.labels{1}));
 
-    wordMap = LoadWordMap('sick-data/combined_words.txt');
+    wordMap = LoadWordMap('sick-data/sick_basic_words.txt');
     hyperParams.vocabName = 'sick_all';
 
-    hyperParams.trainFilenames = {'./sick-data/SICK_train_parsed_exactAlign.txt', ...
-                     './sick-data/SICK_train_parsed.txt', ...
-                     './sick-data/SICK_trial_parsed_exactAlign.txt', ...
-                     './sick-data/SICK_trial_parsed.txt'};
[…]

### 2015-05-19 writing/SNLI paper/abstract.tex
Pre-lunch paper update.

--- 2015-05-03 writing/AAAI SSS '15/BowmanPottsManning/BowmanPottsManning.tex
@@
-\def\year{2015}
-%File: formatting-instruction.tex
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
+\begin{abstract} 
+\todo{Write (CP?)}
 
-\usepackage[leqno, fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{qtree}
-\usepackage[numbers]{natbib}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{colortbl}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{xcolor}
-
-\frenchspacing
-\setlength{\pdfpagewidth}{8.5in}
-\setlength{\pdfpageheight}{11in}
-\pdfinfo{
-/Title Learning Distributed Word Representations for Natural Logic Reasoning
-/Author Samuel R. Bowman, Christopher Potts, Christopher D. Manning
-/Keywords distributional semantics, vector space models, natural logic, natural language processing, neural networks}
-\setcounter{secnumdepth}{1}
-
-\newcommand{\nateq}{\equiv}
-\newcommand{\natind}{\mathbin{\#}}
-%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
-\newcommand{\natneg}{\mathbin{^{\wedge}}}
-\newcommand{\natfor}{\sqsubset}
-\newcommand{\natrev}{\sqsupset}
-\newcommand{\natalt}{\mathbin{|}}
-\newcommand{\natcov}{\mathbin{\smallsmile}}
-
-\newcommand{\plneg}{\mathop{\textit{not}}}
-\newcommand{\pland}{\mathbin{\textit{and}}}
-\newcommand{\plor}{\mathbin{\textit{or}}}
-
-% Strikeout
-\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
-\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}
-
-\newcommand{\word}[1]{\emph{#1}}
-\def\ii#1{\textit{#1}}
-
- \begin{document}
-% The file aaai.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-\title{Learning Distributed Word Representations for Natural Logic Reasoning}
-\author{Samuel R. Bowman, Christopher Potts, and Christopher D. Manning\\
-  Stanford University\\
-  450 Serra Mall \\
-  Stanford, CA 94305-2150\\
-  \texttt{\{sbowman,cgpotts,manning\}@stanford.edu}} %TODO: Three affiliations?
-\maketitle
-\begin{abstract}
-\begin{quote}
-  Natural logic offers a powerful labelal conception of meaning
-  that is a natural counterpart to distributed semantic
-  representations, which have proven valuable in a wide range of
-  sophisticated language tasks. However, it remains an open question
-  whether it is possible to train distributed representations to
-  support the rich, diverse logical reasoning captured by natural
-  logic. We address this question using two neural network-based
-  models for learning embeddings: plain neural networks and neural
-  tensor networks. Our experiments evaluate the models' ability to
-  learn the basic algebra of natural logic labels from simulated
-  data and from the WordNet noun graph.  The overall positive results
-  are promising for the future of learned distributed representations in
-  the applied modeling of logical semantics.
-\end{quote}
 \end{abstract}
-
-Natural logic offers a powerful \emph{labelal} conception of
-semantics: the meanings for expressions are given, at least in part,
-by their inferential connections with other expressions
-\cite{Katz72,maccartney2009extended,vanBenthem08NATLOG}. For instance,
-\word{turtle} is analyzed, not primarily by its extension in the
-world, but rather by its lexical network: it entails \word{reptile},
-excludes \word{chair}, is entailed by \word{sea
-  turtle}, and so forth. With generalized notions of entailment and
-contradiction, these labelships can be defined for all lexical
-categories as well as complex phrases, sentences, and even texts. The
-resulting theories of meaning offer valuable new analytic tools for
-tasks involving database inference, label extraction, and textual
-entailment.
-
-Natural logic aligns well with distributed (e.g., vector)
-representations, which also naturally model meaning labelally.
-Distributed representations have been used successfully in a wide
-array of sophisticated language tasks, including part of speach tagging,
-\cite{collobert2011natural}, semantic role labeling \cite{collobert2011natural},
-sentiment analysis \cite{socher2011semi}, and translation \cite{sutskever2014sequence} 
-among many others.
-However, it remains an open question whether it is possible to train
-such representations to support the rich, diverse logical reasoning
-captured by natural logic; while they excel at synonymy (similarity),
-the results are more mixed for entailment, contradiction, and mutual
-consistency.  Using the natural logic of \cite{maccartney2009extended}
-as our formal model, we address this open question using two neural
-network-based models for learning embeddings: plain neural networks
-and neural tensor networks (NTNs).  The natural logic is built from
-the seven labels defined in Table~\ref{b-table}. Its formal
-properties are now well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}, so it
-provides a rigorous set of goals for our neural models. To keep the
-discussion manageable, we limit attention to experiments involving the
-lexicon; for a more extended treatment of complex expressions
-involving logical connectives and quantifiers, see
-\cite{Bowman:Potts:Manning:2014}. % CITET
-
-In our experiments, we evaluate these models' ability to learn the
-basic algebra of natural logic labels from simulated data and from
-the WordNet noun graph. The simulated data help us to achieve analytic
-insights into what the models learn, and the WordNet data show how they
-fare with a real natural language vocabulary.  We find that only the NTN is able to fully
-learn the underlying algebra, but that both models excel in the 
-WordNet experiment.
-
-\begin{table*}[tp]
-  \centering%\small
-  \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{l c l l} 
-    \toprule
-    Name & Symbol & Set-theoretic definition & Example \\ 
-    \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle $\natfor$ reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile $\natrev$ turtle}  \\ 
-    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch $\nateq$ sofa} \\ 
-    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle $\natalt$ warthog} \\ 
-    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able $\natneg$ unable} \\
-    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal $\natcov$ non-turtle} \\ 
-    independence       & $x \natind y$   & (else) & \ii{turtle $\natind$ pet}\\
-    \bottomrule
-  \end{tabular}
-  \protect\caption{\protect\label{b-table}The seven natural logic labels of \protect\cite{maccartney2009extended}. 
-    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
-    and the label $\natind$ applies whenever none of the other six do.} 
-\end{table*}
-
-\section{Neural network models for label classification} \label{methods}
-
-We build embedding-based models using the method of
-\cite{Bowman:Potts:Manning:2014}, which is centered on the task of
-labeling a pair of words or sentences with one of a small set of
-logical labels. Unlike in a classical inference
-setting, the model is not given specific premises to reason with at test time, but rather
-is expected to encode all of the information that it learns about each term (word) at training
-time in the term embedding vectors, and to then use only the information encoded in the relevant pair of
-term vectors at test time. 
-The architecture that we use is depicted in
-Figure~\ref{sample-figure}. The model represents the two input terms
-as single embedding vectors, which are fed into a comparison function based on one
-of two types of neural network layer function to produce a
-representation for the labelship between the two terms. This
-representation is then fed into a softmax classifier, which outputs a
-distribution over possible labels. The entire network, including the
-embeddings, is trained using backpropagation with AdaGrad \cite{duchi2011adaptive}
-and L2 regularization.
-
-The simpler version of the comparison concatenates the two
-input vectors before feeding them into a LReLU \cite{maasrectifier} neural network (NN)
-layer.  The more powerful neural tensor network (NTN) version adds an
-additional third-order tensor parameter to allow for multiplicative
-interactions between the two inputs \cite{chen2013learning}. For more
-details on the implementation and training of the layer functions, see
-\cite{Bowman:Potts:Manning:2014}.
-
-\begin{figure}[tp]
-  \centering
-    \footnotesize
-
-	\newcommand{\labeledtreenode}[4][3]{\put(#2){\makebox(0,0){{\fcolorbox{black}{#4}{\makebox(#1,0.3){#3}}}}}}
-
-	\newcommand{\textlabel}[4][3.5]{\put(#2){\makebox(0,0){{\makebox(#1,0.3){#3}}}}}
-
-	\definecolor{lexcolor}{HTML}{F5F7C4}
-	\definecolor{compositioncolor}{HTML}{BBEBFF}
-	\definecolor{comparisoncolor}{HTML}{FFC895}
-	\definecolor{softmaxcolor}{HTML}{A5FF8A}
-
-	\setlength{\unitlength}{0.61cm}
-
-	%\resizebox{3in}{!}{
-	\begin{picture}(13.15, 5.5)
-
-	  \labeledtreenode[2.4]{6.5,5}{$P(\sqsubset) = 0.8$}{softmaxcolor}  
-	  \put(6.5,3.7){\vector(0,1){1}}  
-	  \labeledtreenode[5]{6.5,3.4}{turtle \emph{vs.}~animal}{comparisoncolor}
-
-	  \textlabel{2.75,5}{Softmax classifier}{black}
-	  \textlabel{2.0,3.4}{\parbox{2cm}{Comparison N(T)N layer}}{black}
-
-	  \textlabel{6.5,2.1}{Optional embedding}{black}
-	  \textlabel{6.5,1.5}{transformation NN layers}{black}
-
-	  \textlabel{6.5,0.5}{Learned term vectors}{black}
-
-	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-	  \put(1.5,0.75){\vector(0,1){0.85}}
-	  \labeledtreenode{1.5,0.5}{turtle}{lexcolor}
-
-	  \put(1.5,2.25){\vector(4,1){3.25}}
-	  \labeledtreenode{1.5,1.9}{turtle}{compositioncolor}
-
-	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-	  \put(11.5,0.75){\vector(0,1){0.85}}
-	  \labeledtreenode{11.5,0.5}{animal}{lexcolor}
-
-	  \put(11.5,2.25){\vector(-4,1){3.25}}
-	  \labeledtreenode{11.5,1.9}{animal}{compositioncolor}
-
-	\end{picture}
-	%}
-        \caption{The model structure used to compare \ii{turtle} and
-          \ii{animal}.  Learned term representations are fed through
-          either an NN or NTN comparison layer and then to a softmax
-        classifier over the seven labels in Table~\ref{b-table}.}
-  \label{sample-figure}
-\end{figure}
-
-This model used here differs from the one described in that work in
-two ways. First, because the inputs are single terms, we do not use
-a composition function here. Second, for our experiment on WordNet data, 
-we introduce an additional $\tanh$ neural network layer between the embedding 
-input and the comparison function, which facilitates initializing the 
-embeddings themselves from pretrained vectors and was found to help 
-performance in that setting.
-
-%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?
-
-\section{Reasoning about semantic labels}\label{sec:join}
-
-In this experiment, we test our model's ability to learn and use the
-natural logic inference rules schematized in
-Figure~\ref{tab:jointable}. For instance, given that $a \natrev b$ and
-$b \natrev c$, one can conclude that $a \natrev c$, by basic
-set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
-$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  There
-are 32 valid inferences that can be made on the basis of pairs of
-labels of the form $a R b$ and $b R' c$. Cells in the table
-containing a dot correspond to pairs of labels for which no valid
-inference can be drawn in our logic.
-
-% about the labels themselves that do not depend on the
-% internal structure of the things being compared. For example, given
-% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
-% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
-% understanding $a$, $b$, or $c$. These seven labels support more
-% than just transitivity: MacCartney and Manning's
-% \cite{maccartney2009extended} join table defines 32 valid inferences
-% that can be made on the basis of pairs of labels of the form $a R
-% b$ and $b R' c$, including several less intuitive ones such as that if
-% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-
-\paragraph{Experiments}
-To test our models' ability to learn these inferential patterns, we
-create artificial boolean structures in which terms denote sets of entities from
-a small domain (e.g., Figure~\ref{lattice-figure}), employ logical
-deduction to identify the valid statements, divide those into train
-and test sets, and remove from the test set those statements which
-cannot be proven from the training statements
-(Figure~\ref{unprovable}). 
-% using the logic depicted in Figure~\ref{lattice-figure}.
-%
-%
-% We test the model's ability to learn this behavior by creating
-% artificial data sets of terms which represent sets of numbers. Since
-% MacCartney and Manning's set of labels hold between sets as well as
-% between sentences, we can use the underlying set structure to generate
-% the all of the labels that hold between any pair of these terms, as
-% in Figure \ref{lattice-figure}. We train the model defined above on a
-% subset of these labels, but rather then presenting the model with a
-% pair of tree-structured sentences as inputs, simply present it with
-% two single terms, each of which corresponds to a single vector in the
-% (randomly initialized) vocabulary matrix $V$, ensuring that the model
-% has no information about the terms being compared except the labels
-% between them.
-%
-In our experiments, we create 80 randomly generated sets drawn from a
-domain of seven elements. This yields 6400 statements about pairs of
-formulae, of which 3200 are chosen as a test set. 240 of the test examples 
-were excluded from the primary test set for lack of evidence in the training 
-data to support the choice of any one label. 
-We then trained and evaluated both the NN model and the NTN model on these data sets.
-
-\paragraph{Results} 
-We found that the NTN model worked best with 11-dimensional vector
-representations for the 80 sets and a 90-dimensional feature vector
-for the classifier, though the performance of the model was not highly
-dependant on either dimensionality setting. 
-Averaging over five randomly generated data sets, the model was able to correctly label 98.1\% (standard error $0.67\%$) of the provably derivable test examples, and 87.7\%
-($\textit{SE} = 3.59\%$) of the remaining test examples. The simpler NN worked
-best with 11 and 75 dimensions, respectively, but was able to achieve
-accuracies of only 84.8\% ($\textit{SE} = 0.84\%$) and 68.7\% ($\textit{SE} = 1.58\%$),
-respectively. Training accuracy was 99.8\% ($\textit{SE} = 0.04\%$) for the NTN and 94.7\% ($\textit{SE} = 0.89\%$) for
-the NN.
-
-\begin{figure*}[tp]
-  \centering
-  \begin{subfigure}[b]{0.42\textwidth}
-    \centering  
-    \setlength{\arraycolsep}{8pt}
-    \renewcommand{\arraystretch}{1.1}
-    \newcommand{\UNK}{\cdot}  
-    \resizebox{2.4in}{!}{$\begin{array}[c]{c@{ \ }|*{7}{c}|}
-        % \hline
-        \multicolumn{1}{c}{}
-        & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
-        \cline{2-8}
-        \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
-        \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
-        \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
-        \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
-        \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
-        \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
-        \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
-        \cline{2-8}
-      \end{array}$}
-    \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the label that holds between $a$ and $c$, if any.
-These inferences are based on basic set-theoretic truths about the meanings of the underlying labels as described in Table~\ref{b-table}.
-We assess our models' ability to reproduce such inferential paths.}
-    \label{tab:jointable}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
-    \setlength{\unitlength}{1cm}
-    \resizebox{1.5in}{!}{\begin{picture}(5,5.5)
-      \labelednode{2.50}{5}{}{1,2,3}
-
-      \put(0.75,4){\line(3,1){1.5}}
-      \put(2.5,4){\line(0,1){0.5}}
-      \put(4.25,4){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{3.5}{a,b}{1,2}
-      \labelednode{2.50}{3.5}{c}{1,3}
-      \labelednode{4.25}{3.5}{d}{2,3}
-
-      \put(0.75,2.5){\line(0,1){0.5}}
-      \put(0.75,2.5){\line(3,1){1.5}}
-
-      \put(2.5,2.5){\line(-3,1){1.5}}
-      \put(2.5,2.5){\line(3,1){1.5}}
-
-      \put(4.25,2.5){\line(0,1){0.5}}
-      \put(4.25,2.5){\line(-3,1){1.5}}
-
-      \labelednode{0.75}{2}{e,f}{1}
-      \labelednode{2.50}{2}{}{2}
-      \labelednode{4.25}{2}{g,h}{3}
-
-      \put(2.5,1){\line(-3,1){1.5}}
-      \put(2.5,1){\line(0,1){0.5}}
-      \put(2.5,1){\line(3,1){1.5}}
-
-      \labelednode{2.5}{0.5}{}{}
-    \end{picture}}
-    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
-    some sets have multiple names, so that learning $\nateq$ is non-trivial.}\label{lattice-figure}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.2\textwidth}
-    \centering
-    \setlength{\tabcolsep}{12pt}
-    \begin{tabular}[b]{c  c}
-      \toprule
-      Train & Test \\
-      \midrule
-
-      $a \nateq b$              & $b \nateq b$ \\
-      $a \natfor e$              & $b \natfor e$ \\
-      $a \natneg g$              & \strikeout{$e \nateq f$} \\
-      $d \natfor h$              & \strikeout{$g \natrev d$} \\
-      $e \natalt g$            & $h \natrev d$ \\ 
-      $\vdots$		& $\vdots$	\\
-      \bottomrule
-    \end{tabular}
-
-    \caption{A train/test split of the atomic statements about the
-      model.  Test statements not provable from the training data are
-      crossed out.}\label{unprovable}
-  \end{subfigure}  
-  \caption{Experimental goal and set-up for reasoning about semantic labels.}
-  \label{exp1}
-\end{figure*} 
-% Note: None of these test examples is derivable from the shown training data, 
-% but we suggest that there is additional training data, so we can cross lines
-% out willy-nilly without fear.
-
-% RNTN log: tue-j-11-6x80-hip.txt
-% Train PER: 0.0021875
-
-% RNN log: tue-j-11-6x80-r-75.txt
-% Train PER: 0.047188
-
-% and create a dataset consisting of the labels between every pair of
-% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
-% test dataset, and that test dataset was further split into the 2960
-% examples that can be provably derived from the test data using
-% MacCartney and Manning's join table (or by the symmetry of the
-% labels in about half of the cases) and the 240 that
-% cannot.
-
-% We tested a version of both the RNN model and the RNTN model on these
-% data. In both cases, the models were implemented exactly as described
-% in Section~\ref{methods}, but since the items being compared are
-% single terms rather than full tree structures, the composition layer
-% was not used, and the two models differed only in which layer function
-% was used for the comparison layer. We found that the RNTN model worked
-% best with 11 dimensional vector representations for the 80 sets and a
-% 90 dimensional feature vector for the classifier. This model was able
-% to correctly label 99.3\% of the derivable test examples, and 99.1\%
-% of the remaining examples. The simpler RNN model worked best with 11
-% and 75 dimensions, respectively, but was able to achieve accuracies of
-% only 90.0\% and 87.\%, respectively.
-
-These results are fairly straightforward to interpret. The NTN model
-was able to accurately encode the labels between the terms in the
-geometric labels between their vectors, and was able to then use
-that information to recover labels that were not overtly included
-in the training data. In contrast, the NN was able to achieve this
-behavior only incompletely. It is possible but not likely that it
-could be made to find a good solution with further optimization on
-different learning algorithms, or that it would do better on a larger
-universe of sets, for which there would be a larger set of training
-data to learn from, but the NTN is readily able to achieve these
-effects in the setting discussed here.
-
-It's noteworthy that both models performed well above chance on the
-unprovable examples. This indicates that both were able to identify
-statistical cues to the properties of the underlying boolean
-structure, thereby going beyond what one could obtain from the natural
-logic. Information of this kind is not sufficient to yield the
-near-perfect accuracy that we saw on the provable data, but it does
-highlight one of the ways in which trained inference models can turn
-out to be more powerful that rigidly defined proof systems.
-
-\newcommand{\nodagger}{\phantom{$^\dagger$}}
-
-\begin{table*}
-\centering%\resizebox{5.5in}{!}
-{
-  \setlength{\tabcolsep}{15pt}
-  \renewcommand{\arraystretch}{1.1}
-  \begin{tabular}{rrrrrrr} 
-    \toprule
-     Portion of & \multicolumn{2}{c}{25d NN} & \multicolumn{2}{c}{25d NTN} & \multicolumn{1}{c}{$\natfor$ only}\\
-     training data  & \multicolumn{1}{c}{w/ GloVe} & \multicolumn{1}{c}{w/o GloVe} & \multicolumn{1}{c}{w/ GloVe} & \multicolumn{1}{c}{w/o GloVe} \\
-    \midrule
-     100\%   & 99.73 & 99.37   & 99.61  & \textbf{99.95}  & 37.05 \\
-     33\%    & 95.96 & 95.30    & 95.83  & 95.45  & 37.05 		\\
-     11\%    & 91.11 & 90.81   & 91.27  & 90.90  & 37.05 		\\
-    \bottomrule
-  \end{tabular}}
- \caption{Mean test \% accuracy scores (with standard error) on the WordNet data over five-fold crossvalidation.
-The baseline figure is the frequency of the most frequent class, \ii{hypernym}.\label{wn-table}}
-\end{table*}
-
-\section{Reasoning about lexical labels in WordNet}\label{sec:wordnet}
-
-Using simulated data as above is reassuring about what the models
-learn and why, but we also want to know how they perform with a real
-natural language vocabulary. Unfortunately, as far as we are aware,
-there are no available resources labeling such a vocabulary with the
-labels from Table~\ref{b-table}. However, the labels in WordNet
-\cite{fellbaum2010wordnet,WordNet95} come close and pose the same substantive
-challenges within a somewhat easier classification problem.
-
-% WordNet \cite{fellbaum2010wordnet} provides a different type of
-% lattice structure which is compatible with our learning paradigm.
-
-We extract three types of label from WordNet. \ii{Hypernym} and
-\ii{hyponym} can be represented directly in the WordNet graph
-structure, and correspond closely to the $\natrev$ and $\natfor$
-labels from natural logic. As in natural logic, these labels are
-mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
-(perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then
-\ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
-terms, which share a direct hypernym, like \ii{dalmatian}, \ii{pug},
-and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
-terms tend to exclude one another, thereby providing a loose approximation 
-of the natural logic exclusion label $\natalt$. WordNet's antonym label, however, is both too narrowly defined
-and too rare to use for this purpose. WordNet defines its labels over
-sets of synonyms, rather than over individual terms, so we do not
-include a \ii{synonym} or \ii{equivalent} label, but rather
-consider only one member of each set of synonyms. Word pairs which do
-not fall into these three labels are not included in the data set.
-
-To limit the size of the vocabulary without otherwise simplifying the learning problem, we extract all of the
-instances of these three labels for single word nouns in WordNet that are hyponyms of the node 
-\texttt{organism.n.01}. In order to balance the distribution of the classes, we slightly downsample instances 
-of the \ii{coordinate} label, yielding a total of 36,772 labels among 3,217 terms.
-We report results below using crossvalidation, choosing a disjoint 10\% test sample for each of five runs.
-Unlike in the previous experiment,
-it is not straightforward here to determine in advance
-how much data should be required to train an accurate model, so we performed training runs with 
-various fractions of the remaining data. Embeddings were fixed at 25 dimensions and were initialized 
-randomly or using distributional vectors from GloVe \cite{pennington2014glove}. The feature vector 
-produced by the comparison layer was fixed at 80 dimensions.
-
-\paragraph{Results} 
-The structured WordNet data contains substantially more redundancy than the random artificial data above, and we find that the NTN performs perfectly with random initialization.
-The plain NN performs almost as well, providing
-a point of contrast with the results of Section~\ref{sec:join}. We also find that initialization with GloVe is helpful in allowing the models to maintain fair performance
-with smaller amounts of training data. Some of the randomly initialized model runs failed to learn
-usable representations at all and labeled all examples with the most frequent labels.
-We excluded these runs from the statistics, but marked settings for which this occurred with the symbol $\dagger$.
-For all of the remaining runs, training accuracy was consistently above 99\%.
-
-\section{Conclusion}\label{sec:discussion}
-
-This investigation builds upon extensive prior research on
-constructing compositional word meanings in vector spaces
-\cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
-extracting entailment information from distributional vector space
-models \cite{baroni2012entailment,rei2014looking}, and transferring
-information between logic programs and neural networks in constrained
-settings \cite{hitzler2004logic,holldobler1999approximating}. Our
-primary contribution is to establish a tight connection between these
-approaches and formal developments within natural logic.  More
-specifically, we showed that neural network models, optimized using
-established techniques, can learn distributed representations that
-closely approximate the behavior one would expect from a natural logic
-inference algorithm.
-
-% Ours is not the first investigation into logical behavior in vector space models: 
-%  there is extensive prior work on constructing compositional word meanings in vector spaces \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
-%  extracting entailment information from distributional vector space models \cite{baroni2012entailment,rei2014looking},
-%  and transferring information between logic programs and neural networks in 
-%  constrained settings \cite{holldobler1999approximating,hitzler2004logic}. However, this investigation
-%   is the first that we are aware of on logical behavior in representations learned within conventionally
-%    optimized neural network models.
-
-This paper evaluated two neural models on the task of learning 
-natural logic labels between distributed word representations. The
-results suggest that at least the neural tensor network
-has the capacity to
-meet this challenge with reasonably-sized training sets, learning both 
-to embed a vocabulary in a way that encodes a diverse
-set of labels, and to subsequently use those embeddings to infer new
-labels. In
-\cite{Bowman:Potts:Manning:2014}, we extend these results to include
-complex expressions involving logical connectives and quantifiers, with similar
-conclusions about (tree-structured recursive versions of) these models. These findings
-are promising for the future of learned distributed representations in the
-applied modeling of logical semantics.
-
-\section{Acknowledgments}
-
-We thank Jeffrey Pennington, Richard Socher, and audiences at CSLI, Nuance, and BayLearn.
-
-\begin{thebibliography}{}
-
-\bibitem[\protect\citeauthoryear{Baroni \bgroup et al\mbox.\egroup
-  }{2012}]{baroni2012entailment}
-Baroni, M.; Bernardi, R.; Do, N.-Q.; and Shan, C.-c.
-\newblock 2012.
-\newblock Entailment above the word level in distributional semantics.
-\newblock In {\em Proc. EACL}.
-
-\bibitem[\protect\citeauthoryear{Bowman, Potts, and
-  Manning}{2014}]{Bowman:Potts:Manning:2014}
-Bowman, S.; Potts, C.; and Manning, C.
-\newblock 2014.
-\newblock Recursive neural networks can learn logical semantics.
-\newblock {arXiv} manuscript 1406.1827.
-
-\bibitem[\protect\citeauthoryear{Chen \bgroup et al\mbox.\egroup
-  }{2013}]{chen2013learning}
-Chen, D.; Socher, R.; Manning, C.; and Ng, A.
-\newblock 2013.
-\newblock Learning new facts from knowledge bases with neural tensor networks
-  and semantic word vectors.
-\newblock In {\em Proc. {ICLR}}.
-
-\bibitem[\protect\citeauthoryear{Clark, Coecke, and
-  Sadrzadeh}{2011}]{ClarkCoeckeSadrzadeh2011}
-Clark, S.; Coecke, B.; and Sadrzadeh, M.
-\newblock 2011.
-\newblock Mathematical foundations for a compositional distributed model of
-  meaning.
-\newblock {\em Linguistic Analysis} 36(1--4):345--384.
-
-\bibitem[\protect\citeauthoryear{Collobert \bgroup et al\mbox.\egroup
-  }{2011}]{collobert2011natural}
-Collobert, R.; Weston, J.; Bottou, L.; Karlen, M.; Kavukcuoglu, K.; and Kuksa,
-  P.
-\newblock 2011.
-\newblock Natural language processing (almost) from scratch.
-\newblock {\em JLMR} 12.
-
-\bibitem[\protect\citeauthoryear{Duchi, Hazan, and
-  Singer}{2011}]{duchi2011adaptive}
-Duchi, J.; Hazan, E.; and Singer, Y.
-\newblock 2011.
-\newblock Adaptive subgradient methods for online learning and stochastic
-  optimization.
-\newblock {\em JMLR}.
-
-\bibitem[\protect\citeauthoryear{Fellbaum}{2010}]{fellbaum2010wordnet}
-Fellbaum, C.
-\newblock 2010.
-\newblock Word{N}et.
-\newblock In {\em Theory and Applications of Ontology: Computer Applications}.
-  Springer.
-
-\bibitem[\protect\citeauthoryear{Grefenstette}{2013}]{grefenstette2013towards}
-Grefenstette, E.
-\newblock 2013.
-\newblock Towards a formal distributional semantics: Simulating logical calculi
-  with tensors.
-\newblock arXiv manuscript 1304.5823.
-
-\bibitem[\protect\citeauthoryear{Hermann, Grefenstette, and
-  Blunsom}{2013}]{Hermann-etal:2013}
-Hermann, K.; Grefenstette, E.; and Blunsom, P.
-\newblock 2013.
-\newblock ``{N}ot not bad'' is not ``bad'': A distributional account of
-  negation.
-\newblock In {\em Proc. of the 2013 Workshop on Continuous Vector Space Models
-  and their Compositionality}.
-
-\bibitem[\protect\citeauthoryear{Hitzler, H{{\"o}}lldobler, and
-  Seda}{2004}]{hitzler2004logic}
-Hitzler, P.; H{{\"o}}lldobler, S.; and Seda, A.~K.
-\newblock 2004.
-\newblock Logic programs and connectionist networks.
-\newblock {\em Journal of Applied Logic} 2(3).
-
-\bibitem[\protect\citeauthoryear{H{\"o}lldobler, Kalinke, and
-  St{\"o}rr}{1999}]{holldobler1999approximating}
-H{\"o}lldobler, S.; Kalinke, Y.; and St{\"o}rr, H.-P.
-\newblock 1999.
-\newblock Approximating the semantics of logic programs by recurrent neural
-  networks.
-\newblock {\em Applied Intelligence} 11(1).
-
-\bibitem[\protect\citeauthoryear{Icard and Moss}{2013a}]{Icard:Moss:2013}
-Icard, T., and Moss, L.
-\newblock 2013a.
-\newblock A complete calculus of monotone and antitone higher-order functions.
-\newblock In {\em Proc. Topology, Algebra, and Categories in Logic}.
-
-\bibitem[\protect\citeauthoryear{Icard and Moss}{2013b}]{Icard:Moss:2013:LILT}
-Icard, T., and Moss, L.
-\newblock 2013b.
-\newblock Recent progress on monotonicity.
-\newblock {\em Linguistic Issues in Language Technology} 9(7).
-
-\bibitem[\protect\citeauthoryear{Katz}{1972}]{Katz72}
-Katz, J.~J.
-\newblock 1972.
-\newblock {\em Semantic Theory}.
-\newblock New York: Harper \& Row.
-
-\bibitem[\protect\citeauthoryear{Maas, Hannun, and Ng}{2013}]{maasrectifier}
-Maas, A.; Hannun, A.; and Ng, A.
-\newblock 2013.
-\newblock Rectifier nonlinearities improve neural network acoustic models.
-\newblock In {\em Proc. ICML}.
-
-\bibitem[\protect\citeauthoryear{MacCartney and
-  Manning}{2009}]{maccartney2009extended}
-MacCartney, B., and Manning, C.
-\newblock 2009.
-\newblock An extended model of natural logic.
-\newblock In {\em Proc. IWCS}.
-
-\bibitem[\protect\citeauthoryear{Miller}{1995}]{WordNet95}
-Miller, G.~A.
-\newblock 1995.
-\newblock Word{N}et: A lexical database for {E}nglish.
-\newblock {\em Communications of the {ACM}} 38(11):39--41.
-
-\bibitem[\protect\citeauthoryear{Pennington, Socher, and
-  Manning}{2014}]{pennington2014glove}
-Pennington, J.; Socher, R.; and Manning, C.
-\newblock 2014.
-\newblock Glo{V}e: {G}lobal vectors for word representation.
-\newblock In {\em Proc. EMNLP}.
-
-\bibitem[\protect\citeauthoryear{Rei and Briscoe}{2014}]{rei2014looking}
-Rei, M., and Briscoe, T.
-\newblock 2014.
-\newblock Looking for hyponyms in vector space.
-\newblock In {\em Proc. {CoNLL}}.
-
-\bibitem[\protect\citeauthoryear{Rockt{\"a}schel \bgroup et al\mbox.\egroup
-  }{2014}]{rocktaschellow}
-Rockt{\"a}schel, T.; Bosnjak, M.; Singh, S.; and Riedel, S.
-\newblock 2014.
-\newblock Low-dimensional embeddings of logic.
-\newblock In {\em Proc. the ACL 2014 Workshop on Semantic Parsing}.
-
-\bibitem[\protect\citeauthoryear{Socher \bgroup et al\mbox.\egroup
-  }{2011}]{socher2011semi}
-Socher, R.; Pennington, J.; Huang, E.; Ng, A.; and Manning, C.
-\newblock 2011.
-\newblock Semi-supervised recursive autoencoders for predicting sentiment
-  distributions.
-\newblock In {\em Proc. EMNLP}.
-
-\bibitem[\protect\citeauthoryear{Sutskever, Vinyals, and
-  Le}{2014}]{sutskever2014sequence}
-Sutskever, I.; Vinyals, O.; and Le, Q.
-\newblock 2014.
-\newblock Sequence to sequence learning with neural networks.
-\newblock In {\em Proc. {NIPS}}.
-
-\bibitem[\protect\citeauthoryear{van Benthem}{2008}]{vanBenthem08NATLOG}
-van Benthem, J.
-\newblock 2008.
-\newblock A brief history of natural logic.
-\newblock In Chakraborty, M.; L{\"o}we, B.; Nath~Mitra, M.; and Sarukki, S.,
-  eds., {\em Logic, Navya-Nyaya and Applications: Homage to {B}imal {M}atilal}.
-
-\end{thebibliography}
-\bibliographystyle{aaai}
-
-\end{document}

### 2015-05-19 writing/SNLI paper/entailment_eval.tex
Transfer update; new figure.
Pre-lunch paper update.

@@
 \section{SNLI as a platform for NLI research}
-\todo{write me}
 
-Since our goal is to evaluate the quality of the single sentence representations in our model, we train an entailment classifier that has access only to the output of our sentence representation model, and not to any of the word or phrase representations that it used.
-We encourage future users of this corpus to evaluate their models in this way when possible.
-Our classifier is simply a stack of three 100d (NONLIN) layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
+The most immediate application for the SNLI corpus is in developing models for the task of NLI.
+In particular, since SNLI is dramatically larger than any existing corpus of comparable quality, we expect it to make possible the use of parameter-rich models like neural networks for this task.
+In this section, we explore the performance of a range of standard and novel models trained on the corpus.
 
-\subsection{Baselines}
+\subsection{Standard entailment models}
 We evaluate a number of strong baselines on the SNLI corpus.
 The first class of baselines use models from the Excitement Open
-  Platform (EOP) 
-  \cite{pado2014design,magnini2014excitement}
+  Platform (EOP,
+  \citealt{pado2014design,magnini2014excitement})
   -- an open source platform for RTE research which
@@
 
-\Fig{learning_curves_bow.pdf}{0.6}{bowlearncurve}{
-A learning curve for the lexicalized and unlexicalized baseline classifiers.
-Note the y axis starting at a random-chance accuracy of 33\%.
-\todo{numbers are approx.}
-}
+% Commenting out until PDF available. -S
+%\Fig{learning_curves_bow.pdf}{0.6}{bowlearncurve}{
+%A learning curve for the lexicalized and unlexicalized baseline classifiers.
+%Note the y axis starting at a random-chance accuracy of 33\%.
+%\todo{numbers are approx.}
+%}
 
@@
 
+Since our goal is to evaluate the quality of the single sentence representations in our model, we train an entailment classifier that has access only to the output of our sentence representation model, and not to any of the word or phrase representations that it used.
+We encourage future users of this corpus to evaluate their models in this way when possible.
+Our classifier is simply a stack of three 100d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
+\hline
+\textbf{Sentence model} & \b{Train (3-Class)}  & \b{Test (3-Class)}\\
+ & \t{Acc.} & \t{Acc.} \\
+\hline
+\t{Sum of words}            & \t{} & \t{} \\
+\t{RNN}            & \t{} & \t{} \\
+\t{LSTM RNN}            & \t{75?} & \t{72?} \\
+\t{TreeRNN}            & \t{} & \t{} \\
+\t{TreeRNTN}            & \t{58?} & \t{56?} \\
+\t{LSTM TreeRNN}            & \t{72?} & \t{69?} \\
+\hline
+\end{tabular}
+\end{center}
+% The caption
+\caption{
+\label{tab:nnresults}
+Accuracy on the 2-class and 3-class SNLI dataset, as well as precision/recall
+  on the 2-class task to allow for comparision with the RTE systems.
+Ablation studies are reported for removing bigram features and all lexical
+  features.
+\todo{numbers are approx.}
+}
+\end{table}
+
 \subsection{Results}

### 2015-05-19 writing/SNLI paper/transfer.tex
Pre-lunch paper update.

\subsection{SNLI and transfer learning}

\begin{table}
\begin{center}
\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c}
\hline
\textbf{Corpus} & \multicolumn{2}{c}{\b{Train Acc.}} &\multicolumn{2}{c}{ \b{Test Acc.}} \\
 & \t{cold start} & \t{transfer} & \t{cold start} & \t{transfer}  \\
\hline
\t{SICK}            & \t{98} & \t{95} & \t{65?} & \t{68?}  \\
\t{SST}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
\t{SUBJ}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
\t{FactBank?}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
\hline
\end{tabular}
\end{center}
% The caption
\caption{
\label{tab:transferresults}
LSTM
\todo{numbers are approx.}
}
\end{table}

### 2015-05-19 writing/SNLI paper/corpus.tex
Pre-lunch paper update.

@@
 
-% TODO: Rewrite and expand motivation here (CP?)
+\todo{Rewrite and expand motivation here (CP?)}
 
@@
 
-% TODO: Rewrite and expand (CP?)
+\todo{Rewrite and expand (CP?)}
 
@@
 
-% TODO: Do we need statistics about the number/distribution of workers?
-% TODO: Do we need statistics about how much of ImageFlickr we rejected/double-annotated?
+\todo{Do we need statistics about the number/distribution of workers?}
+\todo{Do we need statistics about how much of ImageFlickr we rejected/double-annotated?}
 
@@
 If there was no such consensus, which occurred in about 0.2\% of cases, we chose the label `-'.
-While these unlabeled examples are included in the corpus distribution, they are unlikely to be helpful for the core classification task of NLI, and we
-
-TODO: Data filtering? Data re-labeling?
+While these unlabeled examples are included in the corpus distribution, they are unlikely to be helpful for the core classification task of NLI, and we do not include them when training and evaluating the models discussed in later sections.
 
@@
 
-% Screenshot from phase 1
-
-% Screenshot from phase 2
-
-The data is freely available at (POST DATA), and is released under a CreativeCommons
-Attribution-ShareAlike licence, which is the same licence used for the Flickr30k source captions.
+The data will be released upon publication under a CreativeCommons
+Attribution-ShareAlike licence, the same licence used for the Flickr30k source captions.
 
@@
 
-% TODO (CM?)
+\todo{Update and expand (CM, SB)}
 

### 2015-05-19 writing/SNLI paper/entailment_eval.tex
Paper update.

@@
 \subsection{Sentence embeddings and NLI}\label{sentence-embedding}
+
+SNLI is uniquely suitable for training neural network models that produce vector representations of sentence meaning. In this section, we compare the performance of six such models on SNLI. 
+In order to best evaluate the strengths of these six models at producing informative representations, we use sentence embedding as an intermediate step in the NLI classification task: each model most produce vector representation of each of the two sentences without using any context from the other sentence, and the two resulting vectors are then passed to a neural network classifier which predicts the label for the pair.
+This choice allows us to focus on evaluating the ability of existing models to produce representations that fully capture sentence meaning, at the possible cost of excluding from consideration neural models for NLI that directly compare the two inputs at the word or phrase level.
+
 \input{model_structure_fig.tex}
 
-Since our goal is to evaluate the quality of the single sentence representations in our model, we train an entailment classifier that has access only to the output of our sentence representation model, and not to any of the word or phrase representations that it used.
-We encourage future users of this corpus to evaluate their models in this way when possible.
-Our classifier is simply a stack of three 100d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
+Our neural network classifier, depicted in Fig.~\ref{modelstructure} is simply a stack of three 100d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
+
+We test six sentence representation models, each set to use 50d word and phrase embeddings.
+Our baseline model in this experiment simply sums the embeddings of the word in each sentence.
+Next, we experiment with two recurrent models: a plain RNN, and a one-layer LSTM RNN \cite{hochreiter1997long}.
+Finally, we experiment with three tree-structured models: a plain TreeRNN \cite{socher2011semi}, a tensor-parameterized TreeRNTN \cite{socher2013acl1}, and a TreeLSTM \cite{tai2015improved}.
+
+The word embeddings for all of the moedels are initialized the 200d reference GloVe vectors \cite{pennington2014glove} and fine tuned.
+In addition, all of the models use an additional $\tanh$ neural network layer to map these 200d embeddings into the 50d working space.
+All of the models are randomly initialized using standard techniques and trained using AdaDelta \cite{zeiler2012adadelta} minibatch SGD until performance on the development set stops improving.
+We applied L2 regularization to all models, manually tuning the strength coefficient $\lambda$ for each.
+All models were reimplemented in a common framework for this paper, and the implementations will be made available at publication time.
 
@@
 \hline
-\textbf{Sentence model} & \b{Train (3-Class)}  & \b{Test (3-Class)}\\
- & \t{Acc.} & \t{Acc.} \\
+\textbf{Sentence model} & \b{Train}  & \b{Test}\\
 \hline
@@
 \label{tab:nnresults}
-Accuracy on the 2-class and 3-class SNLI dataset, as well as precision/recall
-  on the 2-class task to allow for comparision with the RTE systems.
-Ablation studies are reported for removing bigram features and all lexical
-  features.
-\todo{numbers are approx.}
+Accuracy in 3-class classification on the SNLI training and test sets for each model.
+\todo{Replace with final numbers}
 }
@@
 
-\subsection{Results}
+Note that \todo{something interesting happens}.

### 2015-05-19 writing/SNLI paper/model_structure_fig.tex
Pre-lunch paper update.
Paper update.

@@
   \centering
-
+\scalebox{0.85}{
  \begin{tikzpicture}
-    \def\dx{19pt}
+    \def\dx{21pt}
     \def\dy{29pt}
@@
     \tikzstyle{e}=[fill=green!40,text width=25mm,align=center]
-    \tikzstyle{m}=[fill=blue!40,text width=34mm,align=center]
+    \tikzstyle{m}=[draw=black,text width=38mm,align=center]
 
@@
 
-  \end{tikzpicture}
+  \end{tikzpicture}}
 
-        \caption{The standard model structure used for entailment evaluation.}
-  \label{sample-figure}
+        \caption{The common classification architecture used in all NLI evaluations: for each model evaluated in Table~\ref{tab:nnresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 50d inputs shown here.}
+  \label{modelstructure}
 \end{figure}

### 2015-05-19 writing/SNLI paper/transfer.tex
Paper update.

@@
-\subsection{SNLI and transfer learning}
+\section{SNLI and transfer learning}
+
+We argue that for a sentence representation to provide the information necessary to reliably perform NLI classification, it must contain a faithful and thorough description of the sentence's meaning.
+This claim has a readily testable consequence: the knowledge that a representation learning system learns when it is trained on SNLI should transfer well to any other task that involves understanding sentence meaning.
+To test this, we perform transfer learning experiments on four other tasks, in which models are initialized using the parameters of a model trained on  on SNLI (as in \S\ref{sentence-embedding}) and then trained to convergence on the target tasks.
 

### 2015-05-19 writing/SNLI paper/corpus.tex
Paper update.

@@
 
+\subsection{The distributed corpus}
+\todo{Update all statistics to reflect the latest relase candidate.}
+
 \begin{table*}
@@
 \rule{0pt}{3ex}A balding professor giving a lecture to students on the topic of "opportunity". & \ii{neutral} & A female professor with hair giving a lecture to students.\\
-% TODO: Replace with *validated* examples from the post-rc2 training set, include all 5 labels
     \bottomrule
-% From 1.0rc2
+% From 1.0rc2`
   \end{tabular}
-  \caption{\label{snli-examples}Randomly chosen examples from SNLI.}
+  \caption{\label{snli-examples}Randomly chosen examples from SNLI.\\
+\todo{Replace with *validated* examples from the post-rc2 training set, include all 5 labels.}}
 \end{table*}
 
-\subsection{The distributed corpus}
+Table~\ref{snli-examples} shows a set of randomly chosen validated examples from the development sent with their labels.
+Qualitatively, we find the data that we collected to draw fairly extensively on commonsense knowledge, suggesting that superficial word alignment models are unlikely to be able to succeed at classification.
+We find the sentences that we collected be largely fluent, correctly spelled English, with a mix of full sentences and bare noun phrases of a style that is appropriate for the underlying caption task.
+Punctution and capitalization are only intermittently present.
 

### 2015-05-19 writing/SNLI paper/transfer.tex
Paper update.

@@
 This claim has a readily testable consequence: the knowledge that a representation learning system learns when it is trained on SNLI should transfer well to any other task that involves understanding sentence meaning.
-To test this, we perform transfer learning experiments on four other tasks, in which models are initialized using the parameters of a model trained on  on SNLI (as in \S\ref{sentence-embedding}) and then trained to convergence on the target tasks.
+To test this, we perform transfer learning experiments on four other tasks, in which models are initialized using the parameters of an LSTM RNN-based model trained on  on SNLI (as in \S\ref{sentence-embedding}) and then trained to convergence on the target tasks.
+
+We first evaluate on SICK (\citealt{marelli2014sick}), the smaller entailment corpus SNLI was modeled on, training on the standard 4.5k training sentence pairs.
+We then evaluate on SUBJ (\citealt{pang2004sentimental}), a two-way sentence classification task, using 5-fold cross-validation over the 10k sentences.
+We next evaluate on the SST (\citealt{socher2013acl1}), a five-way sentiment classification task with labels for both phrases and sentences.
+We train on both types of label within the 8.5k sentence training set and test on only the sentence-level labels in the test set.
+We finally evaluate on FactBank.
+\todo{FactBank} ....
+
+ For SICK, we use pretrained parameters for all of the LSTM parameters, pretrained parameters for the $\tanh$ layers in the classifier, and pretrained embeddings for all of the words that occurred in both corpora, randomly initializing only the softmax layer and the remaining word embeddings.
+We use a similar strategy for the other three tasks, but do not transfer the initialization for the $\tanh$ layers, since the switch between sentence pair classification (in SNLI) and single sentence classification means that the input sizes for the bottom $\tanh$ do not match.
+In both cases we re-initialize AdaDelta as for a new model run.
+
+\todo{Do we want to use all four target tasks?}
 
@@
 \t{SICK}            & \t{98} & \t{95} & \t{65?} & \t{68?}  \\
+\t{SUBJ}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
 \t{SST}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
-\t{SUBJ}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
 \t{FactBank?}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
@@
 % The caption
-\caption{
-\label{tab:transferresults}
-LSTM
-\todo{numbers are approx.}
-}
+\caption{\label{tab:transferresults}
+The performance of an LSTM RNN classification model on four tasks under both a random initialization and an initialization copied from a model trained on SNLI.
+\todo{Fill in numbers.}}
 \end{table}
+
+The results are shown in Table~\ref{tab:transferresults}.
+For each task, the transfer model is compared with a model that was initialized randomly as normal.
+Note that \todo{something interesting happens}.
+Our transfer model shows \todo{something about the state of the art on the tasks.}

### 2015-05-20 writing/SNLI paper/corpus.tex
Paper update.

@@
 
+\todo{SICK is referenced below, so we should set it up as a point of reference either here or in the intro, including mentioning that it had both entailment and similarity labels.}
+
 \subsection{Data collection}
@@
 We used Amazon Mechanical Turk for data collection.
-In each individual task, a worker was presented with premise scene descriptions from a preexisting corpus, and asked to supply hypotheses that fit each of our three labels (\ii{entailment}, \ii{neutral}, and \ii{contradiction}).
+In each individual task (each HIT), a worker was presented with premise scene descriptions from a preexisting corpus, and asked to supply hypotheses that fit each of our three labels (\ii{entailment}, \ii{neutral}, and \ii{contradiction}), forcing the resulting data to be balanced among these classes.
 
 For the premises, we used the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions from that corpus (corresponding to about 30k images) collected in an earlier crowdsourced effort.
-The captions were not authored by the photographers who took the source images, they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: ``Our trip to the Olympic Peninsula'').
+The captions were not authored by the photographers who took the source images, they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \textit{``Our trip to the Olympic Peninsula''}).
 We presented each caption to at least one worker, and some to more than one.
@@
 \footnotesize
-% \toprule
-The Stanford University NLP Group is collecting data for use in research on computer understanding of English. We appreciate your help!
-
 We will show you the caption for a photo. We will not show you the photo. Using only the caption and what you know about the world:
@@
 \end{itemize}
+\todo{Fix the broken examples here (SB)}
 \caption{\label{instructions-1}The instructions used on Mechanical Turk for data collection.}
@@
 \todo{Do we need statistics about the number/distribution of workers?}
-\todo{Do we need statistics about how much of ImageFlickr we rejected/double-annotated?}
 
@@
 
-The majority of the sentence pairs in the corpus have not been reviewed by anyone since their construction as part of the data collection Mechanical Turk task, except to remove blank responses and responses some pairs which the original author flagged for review.
-However, in order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we did perform an additional round of validation for about 54k (TODO) examples.
+The majority of the sentence pairs in the corpus have not been reviewed by anyone since their construction as part of the data collection Mechanical Turk task, except to remove a small number of responses where either a response field was empty or the original author flagged the source caption for review.
+In order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we did perform an additional round of validation for about 10\% of our data.
 
-This validation phase is not fundamentally different from the Mechanical Turk labeling task used for the SICK entailment data: we present workers with pairs of sentences (in our case, in batches of five at a time), and ask them to choose a single label for each pair.
+This validation phase is not fundamentally different from the Mechanical Turk labeling task used to label the SICK entailment data: we present workers with pairs of sentences (in our case, in batches of five at a time), and ask them to choose a single label for each pair.
 We supplied each pair to four annonators, yielding five labels per pair including the label used by the original composer of the pair.
@@
 
-The results of this validation process are summarized in Fig.~\ref{validation-stats}.
 For each pair that we validated, we chose a gold label.
@@
 While these unlabeled examples are included in the corpus distribution, they are unlikely to be helpful for the core classification task of NLI, and we do not include them when training and evaluating the models discussed in later sections.
+The results of this validation process are summarized in Tables \ref{validation-stats} and \ref{validation-freq}, and suggest that while fundamental limits of both our crowdsourcing methods and the task itself impose an upper bound on the consistency of our labels, the data is sufficiently high quality to pose a challenging but realistic machine learning task.
 
@@
     \toprule
-Pairs in corpus & 570,157\\
 Training pairs &  550,159\\
-Dev pairs &  9,999\\
+Development pairs &  9,999\\
 Test pairs &  9,999\\
+\textbf{Total sentence pairs} & 570,157\\
 \midrule
@@
     % From 1.0rc2
-\caption{\label{validation-stats}Key statistics for the raw sentence pairs. Since the two halves of each pair were collected separately, we report statistics for both.} 
+\caption{\label{collection-stats}Key statistics for the raw sentence pairs. Since the two halves of each pair were collected separately, we report statistics for both.} 
 \end{table}
@@
     % From 1.0rc2
-\caption{\label{b-table}Key statistics for the validated sentence pairs.} 
+\caption{\label{validation-stats}Key statistics for the validated sentence pairs.} 
 \end{table}
@@
   % From 1.0rc2
-\caption{\label{b-table}Frequency (\%) of each label in the corpus. Note that examples without gold labels are omitted in standard evaluations.} 
+\caption{\label{validation-freq}Frequency (\%) of each label in the corpus. Note that examples without gold labels are omitted in standard evaluations.} 
 \end{table}

### 2015-05-20 utils/make_wordlist.py

@@ -11,19 +11,18 @@ from collections import defaultdict
 
 # TODO: Ensure that there are no duplicates.
 
-BASENAME = "sick_basic"
-TRAINING_FILES = ["sick-data/SICK_train_parsed.txt"]
-DEV_FILES = ["sick-data/SICK_trial_parsed.txt"]
-TEST_FILES = ["sick-data/SICK_test_annotated_rearranged_parsed.txt"]
+BASENAME = "sst-snlirc2-transfer"
+TRAINING_FILES = ["sst-data/train.txt"]
+DEV_FILES = ["sst-data/dev.txt"]
+TEST_FILES = ["sst-data/test.txt"]
 
-# TRANSFER_SOURCE_WORDLIST = "../data/snlirc2_words.txt"
-TRANSFER_SOURCE_WORDLIST = ""
+TRANSFER_SOURCE_WORDLIST = "../data/snlirc2_words.txt"
 VECTOR_WORDLIST = "utils/glove.6B.wordlist.txt"
 
 EXCLUSIONS = set(['(', '(0', '(1', '(2', '(3', '(4', ')', '', ' ', '\n', '\r'])
 THRESHOLD = 50
 
-ENTAILMENT_MODE = True
+ENTAILMENT_MODE = False
 
 
 def count_words(filenames):

### 2015-05-22 writing/SNLI paper/corpus.tex
Misc, including rc3 stats.

@@
 That FAQ grew over the course of data collection, but discussed disallowed techniques (e.g., reusing the same sentence for many different prompts, which we saw in a few cases), sentence length and complexity (we did not enforce a minimum length, and we allowed bare NPs as well as full sentences), and logistical issues around payment timing.
+About 2500 workers contributed data.
 
@@
 
-\todo{Do we need statistics about the number/distribution of workers?}
-
 \subsection{Data validation}
@@
 The instructions were broadly similar to the instructions for initial data collection shown in Fig.~\ref{instructions-1}, and linked to a similar FAQ.
+Though we used a very restrictive qualification (based on past approval rate) to select workers, we nonetheless discovered and deleted some instances of random guessing during an early round of validation, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
 
@@
     \bottomrule
-% From 1.0rc2`
+% From 1.0rc2
   \end{tabular}
@@
     \toprule
-Training pairs &  550,159\\
-Development pairs &  9,999\\
-Test pairs &  9,999\\
-\textbf{Total sentence pairs} & 570,157\\
+Training pairs &  550,153\\
+Development pairs &  10,000\\
+Test pairs & 10,000\\
 \midrule
@@
 \midrule
-Premise complete sentences & 70.2 ??\%\\
-Hypothesis complete sentences & 75.4 ??\%\\
+Premise `S'-rooted parses & 74.0\%\\
+Hypothesis `S'-rooted parses & 88.9\%\\
     \bottomrule
   \end{tabular}
-    % From 1.0rc2
+% From 1.0rc3
 \caption{\label{collection-stats}Key statistics for the raw sentence pairs. Since the two halves of each pair were collected separately, we report statistics for both.} 
@@
     \toprule
-Validated pairs & 53,852\\
-gold label $\ne$ original label & 8.5\%\\
-Unanimous gold & 57.7\%\\
-No gold (no 3 labels match) & 2.1\%\\
-Label $=$ gold & 88.9\%\\
-Label $=$ original author's label & 85.5\%\\
+Validated pairs & 59,445\\
+gold label $\ne$ original label & 8.4\%\\
+Unanimous gold & 58.2\%\\
+No gold (no 3 labels match) & 1.9\%\\
+Label $=$ gold & 89.0\%\\
+Label $=$ original author's label & 85.7\%\\
     \bottomrule
   \end{tabular}
-    % From 1.0rc2
+    % From 1.0rc3
 \caption{\label{validation-stats}Key statistics for the validated sentence pairs.} 
@@
 \midrule
-\ii{entailment} &33.3 & 33.4 & 33.3 \\
-\ii{neutral} & 33.2 & 33.2 & 33.2 \\
-\ii{contradiction} & 33.3 & 33.2 & 33.3 \\
-- (no consensus) & 0.2 & 0.2 & 0.2 \\
+\ii{entailment} &33.3 & 33.3 & 33.7 \\
+\ii{neutral} & 33.2 & 32.4 & 32.2 \\
+\ii{contradiction} &  33.3 & 32.7 & 32.4  \\
+- (no consensus) & 0.1 & 1.6 & 1.8 \\
 \bottomrule
   \end{tabular}	
-  % From 1.0rc2
+% From 1.0rc3
 \caption{\label{validation-freq}Frequency (\%) of each label in the corpus. Note that examples without gold labels are omitted in standard evaluations.} 

### 2015-05-22 writing/SNLI paper/intro.tex

\section{Introduction}
\todo{Write (CP)}

One point of comparison that might be good to set up (maybe not in these words, this is Sam's sketch):

Translation can also be used to train/evaluate NNs, and also demands some degree of sensitivity to compositional syntactic and semantic structure. Plus, it's easier to get good data for that task. But NLI is the better benchmark for developing NNs for language understanding, because:

(i) Typical translation tasks require natural language generation, which is a separate difficult problem that must be learned in parallel with the semantic encoding task of interest, making results harder to interpret. We can just a vanilla well-understood classifier on top of our sentence model.

(ii) Contradiction vs. entailment decisions in particular specifically target the abilities of NN models to learn lexical and phrasal representations (like alternation) that don't resemble similarity, either in their correlation with distributional information or their transitivity behavior. MT doesn't seem to have a good parallel to this. Since modeling similarity is almost the only aspect of NN behavior in NLP that's reasonably well understood and basically known to work, using a benchmark that explicitly demands something more sophisticated than this is likely to pay off by better exposing the weaknesses of current standard models.

It might be also worth making an explicit comparison with sentiment as a benchmark, but that's low-hanging fruit.

### 2015-05-22 writing/SNLI paper/transfer.tex
Paper update.

@@
 \begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c}
+\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c}
 \hline
-\textbf{Corpus} & \multicolumn{2}{c}{\b{Train Acc.}} &\multicolumn{2}{c}{ \b{Test Acc.}} \\
- & \t{cold start} & \t{transfer} & \t{cold start} & \t{transfer}  \\
+\textbf{Corpus} & \multicolumn{2}{c}{\b{Train Acc.}} &\multicolumn{3}{c}{ \b{Test Acc.}} \\
+ & \t{cold start} & \t{transfer} & \t{cold start} & \t{transfer} & \t{SotA} \\
 \hline
-\t{SICK}            & \t{98} & \t{95} & \t{65?} & \t{68?}  \\
-\t{SUBJ}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
-\t{SST}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
-\t{FactBank?}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
+\t{SICK}            & \t{98} & \t{95} & \t{63?} & \t{69?} & \t{84.5} \\
+\t{SUBJ}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?}& \t{??.?} \\
+\t{SST}          & \t{??.?} & \t{??.?} & \t{43?} & \t{43?} & \t{50.6}\\
+\t{FactBank?}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?}& \t{??.?} \\
 \hline
@@
 \end{center}
-% The caption
+
 \caption{\label{tab:transferresults}
-The performance of an LSTM RNN classification model on four tasks under both a random initialization and an initialization copied from a model trained on SNLI.
-\todo{Fill in numbers.}}
+The performance of an LSTM RNN classification model on four tasks under both a random initialization and an initialization copied from a model trained on SNLI. \todo{Final numbers.}}
+
 \end{table}
+..
+The final column shows the state of the art results from \cite{lai2014illinois}, TODO, and \cite{tai2015improved}, respectively
+...
 

### 2015-05-22 writing/SNLI paper/corpus.tex
Assorted updates, new paper stats

@@
     \toprule
-Training pairs &  550,153\\
+Training pairs &  550,152\\
 Development pairs &  10,000\\
@@
     \toprule
-Validated pairs & 59,445\\
-gold label $\ne$ original label & 8.4\%\\
-Unanimous gold & 58.2\%\\
-No gold (no 3 labels match) & 1.9\%\\
+Validated pairs & 56,941\\
+gold label $\ne$ original label & 6.8\%\\
+Unanimous gold & 58.3\%\\
+No gold (no 3 labels match) & 2.0\%\\
 Label $=$ gold & 89.0\%\\
-Label $=$ original author's label & 85.7\%\\
+Label $=$ original author's label & 85.8\%\\
     \bottomrule

### 2015-05-22 config/PragBank.m

@@ -18,7 +18,7 @@ if collo == 1
     hyperParams.vocabPath = ['/scr/nlp/data/glove_vecs/glove.6B.' num2str(embDim) 'd.txt'];
 elseif collo == 2
     hyperParams.vocabPath = '/u/nlp/data/senna_embeddings/combined.txt';  
-    assert(embDim == 50, 'The Collobert and Weston-sourced vectors only come in dim 50.'); 
+    assert(embDim == 50, 'The Collobert and Weston-sourced vectors only come in 50d.'); 
 elseif collo == 3
     hyperParams.vocabPath = ['/scr/nlp/data/glove_vecs/glove.840B.' num2str(embDim) 'd.txt'];
 end
@@ -53,9 +53,6 @@ hyperParams = CompositionSetup(hyperParams, composition);
 hyperParams.loadWords = true;
 hyperParams.trainWords = true;
 
-wordMap = LoadWordMap('../data/subj_words.txt');
-hyperParams.vocabName = 'subj'; 
-
 hyperParams.numLabels = [ 7 ];
 
 hyperParams.labels = {{'CT_plus', 'CT_minus', 'PR_plus', 'PR_minus', 'PS_plus', 'PS_minus', 'Uu'}};
@@ -65,5 +62,15 @@ labelMap{1} = containers.Map(hyperParams.labels{1}, 1:length(hyperParams.labels{
 hyperParams.trainFilenames = {'pragbank-data/train.txt'};    
 hyperParams.splitFilenames = {};    
 hyperParams.testFilenames = {'pragbank-data/test.txt'};
+
+
+if strcmp(dataflag, 'pragbank')
+	wordMap = LoadWordMap('pragbank-data/pragbank_words.txt');
+	hyperParams.vocabName = 'subj'; 
+elseif strcmp(dataflag, 'pragbank-transfer')
+	wordMap = LoadWordMap('pragbank-data/pragbank-snlirc2-transfer_words.txt');
+	hyperParams.vocabName = 'subj';
+	hyperParams.sourceWordMap = LoadWordMap('../data/snlirc2_words.txt');
+end
 
 end

### 2015-05-22 RunExperiments/RunSUBJExperiments.sh
Assorted updates, new paper stats

@@ -13,6 +13,16 @@ export MATLABCMD="cd quant; lambda = 0.001; dim = 50; ed = 200; td = 2; penult =
 export MATLABCMD="cd quant; lambda = 0.00001; dim = 50; ed = 200; td = 2; penult = 75; dropout = [0.5\, 0.5]; tot = 4; collo = 1; dataflag='sst-expanded'; name='/scr/sbowman/subj-init-b'; TrainModel(''\, 1\, @SUBJ\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 0\, 0\, 0\, 0.1);" ; qsub -v MATLABCMD quant/run.sh  
 export MATLABCMD="cd quant; lambda = 0.0000001; dim = 50; ed = 200; td = 2; penult = 75; dropout = [0.5\, 0.5]; tot = 4; collo = 1; dataflag='sst-expanded'; name='/scr/sbowman/subj-init-b'; TrainModel(''\, 1\, @SUBJ\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 0\, 0\, 0\, 0.1);" ; qsub -v MATLABCMD quant/run.sh  
 
+% Transfer
+export MATLABCMD="cd quant; lambda = 0.0001; dim = 50; ed = 200; td = 3; penult = 0; dropout = [0.9\, 0.9]; tot = 2; collo = 1; dataflag='subj-transfer'; name='/scr/sbowman/subjt-'; TrainModel('snlirc-snlirc2-only-l1e-07-dim50-ed200-td3-pen100-do0.9-0.9-co1-comp2-dp1/ckpt-best-tr150517185056@353375.mat'\, 1\, @SUBJ\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo);" ; qsub -v MATLABCMD quant/run.sh 
[…]

### 2015-05-22 config/Defaults.m

@@ -74,6 +74,8 @@ hyperParams.topDropout = 1;
 % the embeddings at those indices will be reinitialized before computing each example. Used for ALCIR.
 hyperParams.randomEmbeddingIndices = [];
 
+hyperParams.restartUpdateRuleInTransfer = false;
+
 % L1 v. L2 regularization. If no regularization is needed, set
 % lambda to 0 and ignore this parameter.
 hyperParams.norm = 2;

### 2015-05-22 utils/make_wordlist.py

@@ -11,10 +11,10 @@ from collections import defaultdict
 
 # TODO: Ensure that there are no duplicates.
 
-BASENAME = "pragbank"
-TRAINING_FILES = ["pragbank-data/train.txt"]
-DEV_FILES = []
-TEST_FILES = ["pragbank-data/test.txt"]
+BASENAME = "snlirc3"
+TRAINING_FILES = ["../data/snli_1.0rc3_train.txt"]
+DEV_FILES = ["../data/snli_1.0rc3_dev.txt"]
+TEST_FILES = ["../data/snli_1.0rc3_test.txt"]
 
 # TRANSFER_SOURCE_WORDLIST = "../data/snlirc2_words.txt"
 TRANSFER_SOURCE_WORDLIST = ""
@@ -23,7 +23,7 @@ VECTOR_WORDLIST = "utils/glove.6B.wordlist.txt"
 EXCLUSIONS = set(['(', '(0', '(1', '(2', '(3', '(4', ')', '', ' ', '\n', '\r'])
 THRESHOLD = 50
 
-ENTAILMENT_MODE = False
+ENTAILMENT_MODE = True
 
 
 def count_words(filenames):

### 2015-05-23 writing/SNLI paper/corpus.tex
Updating paper to match details of rc3.

@@
   \centering\footnotesize
-  \begin{tabular}{p{7.5cm}lp{5.5cm}}
+  \begin{tabular}{p{6cm}p{0.75cm}p{1.4cm}p{6cm}}
   \toprule
-A young man in a striped hoodie looks at an artistic Levi's advertisement that says "We are all workers." & \ii{entailment} & There is a man in a hoodie looking at an advertisement.\\
-\rule{0pt}{3ex}Two men on laptops in a dark room. & \ii{entailment} & Two men are on their laptops.\\
-\rule{0pt}{3ex}A naked baby and toddler smear each other in paint. & \ii{neutral} & The kids are in the bathtub.\\
-\rule{0pt}{3ex}A woman is focusing on her game of curling. & \ii{neutral} & There is a women playing a game indoors\\
-\rule{0pt}{3ex}A group of people dressed in green and white are dancing in a circle outside of a white building. & \ii{contradiction} & A group performs indoors.\\
-\rule{0pt}{3ex}A balding professor giving a lecture to students on the topic of "opportunity". & \ii{neutral} & A female professor with hair giving a lecture to students.\\
+A man inspects the uniform of a figure in some East Asian country. & \b{contr.}& \t{$\{$c c c c c$\}$} & The man is sleeping\\
+\rule{0pt}{3ex}An older and younger man smiling. & \b{neut.}& \t{$\{$n n e n n$\}$} & Two men are smiling and laughing at the cats playing on the floor.\\
+\rule{0pt}{3ex}A black race car starts up in front of a crowd of people. & \b{contr.}& \t{$\{$c c c c c$\}$} & A man is driving down a lonely road.\\
+\rule{0pt}{3ex}A soccer game with multiple males playing. & \b{ent.}& \t{$\{$e e e e e$\}$} & Some men are playing a sport.\\
+\rule{0pt}{3ex}A smiling costumed woman is holding an umbrella. & \b{neut.}& \t{$\{$n n e c n$\}$} & A happy woman in a fairy costume holds an umbrella.\\
     \bottomrule
-% From 1.0rc2
+% From 1.0rc3
   \end{tabular}
-  \caption{\label{snli-examples}Randomly chosen examples from SNLI.\\
-\todo{Replace with *validated* examples from the post-rc2 training set, include all 5 labels.}}
+  \caption{\label{snli-examples}Randomly chosen examples from the SNLI development set, shown with the gold labels and the full set of annotator labels (abbreviated).}
 \end{table*}
@@
 We find the sentences that we collected be largely fluent, correctly spelled English, with a mix of full sentences and bare noun phrases of a style that is appropriate for the underlying caption task.
-Punctution and capitalization are only intermittently present.
+Punctution and capitalization are only intermittently presE
 
@@
 \midrule
-\ii{entailment} &33.3 & 33.3 & 33.7 \\
+\ii{entailment} &		33.3 & 	33.3 & 	33.7 \\
 \ii{neutral} & 		33.2 & 	32.4 & 	32.2 \\

### 2015-05-23 RunExperiments/RunSNLIExperiments.sh
Updating paper to match details of rc3.

@@ -220,6 +220,21 @@ export MATLABCMD="cd quant; lambda = 0.00003; dim = 50; ed = 300; td = 3; penult
 export MATLABCMD="cd quant; lambda = 0.00003; dim = 50; ed = 300; td = 3; penult = 100; dropout = [1\, 1]; tot = 2; collo = 3; dataflag='snlirc3-only-short'; name='/scr/sbowman/snlirc3'; TrainModel(''\, 1\, @SNLI\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1\, 0\, 4);" ; qsub -v MATLABCMD quant/run.sh -l mem=6gb
 export MATLABCMD="cd quant; lambda = 0.00003; dim = 50; ed = 300; td = 3; penult = 100; dropout = [1\, 1]; tot = 2; collo = 3; dataflag='snlirc3-only-short'; name='/scr/sbowman/snlirc3'; TrainModel(''\, 1\, @SNLI\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1\, 0\, 5);" ; qsub -v MATLABCMD quant/run.sh -l mem=6gb
 
+% Next round
+export MATLABCMD="cd quant; lambda = 0.000001; dim = 50; ed = 300; td = 3; penult = 100; dropout = [1\, 1]; tot = 7; collo = 3; dataflag='snlirc3-only-short'; name='/scr/sbowman/snlirc3b'; TrainModel(''\, 1\, @SNLI\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1\, 0.5\, 1);" ; qsub -v MATLABCMD quant/run.sh -q john -l mem=6gb
+export MATLABCMD="cd quant; lambda = 0.000001; dim = 50; ed = 300; td = 3; penult = 100; dropout = [1\, 1]; tot = 7; collo = 3; dataflag='snlirc3-only-short'; name='/scr/sbowman/snlirc3b'; TrainModel(''\, 1\, @SNLI\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1\, 5\, 1);" ; qsub -v MATLABCMD quant/run.sh -q john -l mem=6gb
[…]

### 2015-05-25 writing/SNLI paper/abstract.tex
Draft abstract

@@
 \begin{abstract} 
-\todo{Write (CP?)}
-
+  Entailment and contradiction are fundamental to natural language
+  meaning and lie at the heart of tasks like summarization, textual
+  entailment, and commonsense reasoning.
+  %
+  However, computational modeling of these concepts is currently
+  hampered by the limited nature of existing annotated resources.
+  %
+  To address this, we present a new corpus of sentence pairs labeled
+  for entailment, contradiction, and semantic independence. At 550K
+  pairs, it is an order of magnitude larger than all other resources
+  of its type.
+  %
+  And, in contrast to many such resources, all of the sentences were
+  written by humans in a grounded, naturalistic context, and 54K of
+  the labels have been independently validated.
+  %
+  In this paper, we use this corpus to evaluate a wide variety of
+  models for natural language inference, finding that Tree-Structured
+  Long Short-Term Memory networks (TreeLSTMs) achieve the best
+  performance.
+  %
+  In addition, to enhance the case for TreeLSTMs, and deepen the
+  argument for the centrality of entailment and contradiction in
+  semantics, we show that the TreeLSTM's representatations, trained on
+  our corpus, perform well in a disparate set of additional semantic
+  tasks.  
 \end{abstract}

### 2015-05-25 writing/SNLI paper/intro.tex
Rough draft intro

@@
-\section{Introduction}
-\todo{Write (CP)}
+\section{Introduction}\label{sec:introduction}
 
-One point of comparison that might be good to set up (maybe not in these words, this is Sam's sketch):
+The semantic concepts of entailment and contradiction are central to
+all aspects of natural language meaning
+\cite{Katz72,vanBenthem08NATLOG}, from the lexicon to the content of
+entire texts. Thus, \emph{natural language
+  inference} (NLI) --- characterizing and using these relations in
+computational systems
+\cite{dagan2006pascal,MacCartney09,maccartney2009extended} --- is
+essential in tasks ranging from information retrieval to semantic
+parsing to commonsense reasoning.
 
-Translation can also be used to train/evaluate NNs, and also demands some degree of sensitivity to compositional syntactic and semantic structure.
+NLI has been addressed using a wide variety of techniques, including
+those grounded in syntactic structures, knowledge bases, and symbolic
+logic. In recent years, it has become an important testing ground for
+approaches employing \emph{distributed} word and phrase
+representations. Distributed representations excel at capturing
+relations based in similarity, but it is less clear that they can be
+trained to support the full range of logical inferences required for
+NLI. In the SemEval 2014 task aimed at evaluating distributed
+representations for NLI, the best-performing systems relied heavily on
+additional features and reasoning capabilities
+\cite{marelli2014semeval}.
+
+Our primary objective in this paper is to provide a new empirical
+evaluation of a wide range of models for distributed semantic
+representations in the context of NLI. However, in our view, the
+existing corpus resources in this area do not permit such an
+assessment. They are generally too small for training modern
+data-intensive, wide-coverage models, and they are often beset with
+indeterminacies of event and entity coreference that significantly
+impact annotation quality.
+
+To address this, we present a new corpus of sentence pairs labeled for
+entailment, contradiction, and semantic independence. At 550,152
+sentence-pairs, this corpus is orders of magnitude larger than all
+other resources of its type. And, in contrast to many such resources,
+all of the sentences were written by humans in a grounded,
+naturalistic context. In a separate validation phase, we collected
+four additional judgments for each label for 56,941 of the examples,
+and 98\% of cases emerge with a gold label, which attests to the high
+quality of the data.
+
+In this paper, we use this corpus to evaluate a wide variety of models
+for natural language inference, including rule-based systems, simple
+linear classifiers, and compositional neural networks. We find that
+Tree-Structured Long Short-Term Memory networks (TreeLSTMs;
+\citealt{tai2015improved,le2015compositional}) achieve the best
+performance. In addition, we enhance the case for TreeLSTMs by showing
+that its representations, trained on our corpus, perform well in a
+disparate set of additional semantic tasks.  The success of these
+transfer-learning experiments is also a testament to the centrality
+of NLI in semantics.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% One point of comparison that might be good to set up (maybe not in these words, this is Sam's sketch):
+
+% Translation can also be used to train/evaluate NNs, and also demands some degree of sensitivity to compositional syntactic and semantic structure.
 Plus, it's easier to get good data for that task.
@@
 
-(i) Typical translation tasks require natural language generation, which is a separate difficult problem that must be learned in parallel with the semantic encoding task of interest, making results harder to interpret.
+% (i) Typical translation tasks require natural language generation, which is a separate difficult problem that must be learned in parallel with the semantic encoding task of interest, making results harder to interpret.
 We can just a vanilla well-understood classifier on top of our sentence model.
 
-(ii) Contradiction vs.
+% (ii) Contradiction vs.
 entailment decisions in particular specifically target the abilities of NN models to learn lexical and phrasal representations (like alternation) that don't resemble similarity, either in their correlation with distributional information or their transitivity behavior.
@@
 
-It might be also worth making an explicit comparison with sentiment as a benchmark, but that's low-hanging fruit.
+% It might be also worth making an explicit comparison with sentiment as a benchmark, but that's low-hanging fruit.

### 2015-05-25 writing/SNLI paper/intro.tex
A few tweaks to the draft

@@
 representations. Distributed representations excel at capturing
-relations based in similarity, but it is less clear that they can be
+relations based in similarity, and they have proven effective at
+modeling simple dimensions of meaning like evaluative sentiment
+\cite{socher2013acl1}, but it is less clear that they can be
 trained to support the full range of logical inferences required for

### 2015-05-25 writing/SNLI paper/corpus.tex
Initial draft of framing sections

@@
 
-\todo{Rewrite and expand motivation here (CP?)}
+\subsection{Motivation}
 
-Our ultimate aim in this work is to develop supervised models for sentence representation that can accurately capture natural language meaning.
+To date, the primary sources of annotated NLI corpora have been the
+Recognizing Textual Entailment (RTE)
+tasks.\footnote{\url{http://aclweb.org/aclwiki/index.php?title=Textual_Entailment_Resource_Pool}}
+These are generally high-quality, hand-labeled data sets, and they
+have stimulated innovative logical and statistical models of natural
+language reasoning, but their small size limits their utility as a
+testbed for distributed representations. The recent SemEval 2014 data
+set called Sentences Involving Compositional Knowledge (SICK) is a
+step up in terms of size, but only to about 10,000 examples, and its
+partly automatic construction introduced some spurious patterns into
+the data (\citealt{marelli2014semeval}, $\S$6). Finally, the
+Denotation Graph data set \cite{hodoshimage} contains millions of
+examples, but it was constructed using fully automatic methods, and is
+noisy enough that it is probably suitable only as a source of
+supplementary training data.
+
+Existing resources suffer from a subtler issue that impacts even
+projects using only human-provided annotations: indeterminacies of
+event and entity coreference lead to insurmountable indeterminacy
+concerning the correct semantic label \cite{marelli2014sick}. For
+instance, the sentence pair \word{A boat sank in the Pacific
+  ocean} and \word{A boat sank in the ocean} could reasonably in
+numerous ways depending on how the objects and events are grounded. If
+the two boats are different, then the sentences are semantically
+independent. If the boats are the same and the oceans are the same,
+then the first sentence entails the second (in the relevant
+commonsense way). If the boats are the same and the oceans are
+different, then the two events are (commonsense) contradictory. This
+kind of indeterminacy can be resolved only once the questions of
+coreference are resolved.
+
+With our corpus, we sought to address the issues of size, quality, and
+indeterminacy. To do this, we employed a crowdsourcing framework with
+the following crucial innovations. First, the examples were grounded
+in specific scenarios, which helps greatly in controlling event and
+entity coreference that beset existing corpora. Second, the prompt
+gave participants the freedom to produce entirely novel sentences
+within the task setting, which led to richer examples than we see with
+the more proscribed string-editing techniques of earlier approaches,
+without sacrificing consistency. Third, a subset of the resulting
+sentences were sent to a validation task aimed at flagging low-quality
+data and identifying areas of inferential uncertainty.
+
+% Our ultimate aim in this work is to develop supervised models for sentence representation that can accurately capture natural language meaning.
 While sentiment tasks like SST have provided a useful testbed for sentence representation models, sentiment labeling only requires that models be able to encode a small piece of the full expressive capacity of language.
@@
 
-\subsection{Grounding with imagined images}
+% \subsection{Grounding with imagined images}
 
-\todo{Rewrite and expand (CP?)}
+% \todo{Rewrite and expand (CP?)} --- I incorporated this above.
 
-Quote from SICK paper \cite{marelli2014sick}:
+% Quote from SICK paper \cite{marelli2014sick}:
 
-\begin{quote}
-Not unreasonably, subjects found that, say, \ii{A woman is wearing an Egyptian headdress} does not contradict \ii{A woman is wearing an Indian headdress}, since one could easily imagine both sentences truthfully uttered to refer to a single scene where two different women are wearing different headdresses.
+% \begin{quote}
+% Not unreasonably, subjects found that, say, \ii{A woman is wearing an Egyptian headdress} does not contradict \ii{A woman is wearing an Indian headdress}, since one could easily imagine both sentences truthfully uttered to refer to a single scene where two different women are wearing different headdresses.
 In the future, a higher proportion of CONTRADICTION labels could be elicited by using grammatical and possibly visual cues (pictures) encouraging co-indexing of the entities in the two sentences.
-\end{quote}
+% \end{quote}
 
-\todo{SICK is referenced below, so we should set it up as a point of reference either here or in the intro, including mentioning that it had both entailment and similarity labels.}
+%\todo{SICK is referenced below, so we should set it up as a point of reference either here or in the intro, including mentioning that it had both entailment and similarity labels.}
 

### 2015-05-26 writing/SNLI paper/corpus.tex
Lots of paper cleanup.

@@
 \section{A new corpus for natural language inference}\label{sec:discussion}
-
-\subsection{Motivation}
 
@@
 have stimulated innovative logical and statistical models of natural
-language reasoning, but their small size limits their utility as a
-testbed for distributed representations. The recent SemEval 2014 data
-set called Sentences Involving Compositional Knowledge (SICK) is a
-step up in terms of size, but only to about 10,000 examples, and its
+language reasoning, but their small size of a few hundred examples each
+limits their utility as a testbed for learned distributed representations. 
+The data for the SemEval 2014 task called Sentences Involving Compositional Knowledge (SICK) is a
+step up in terms of size, but only to 4,500 training examples, and its
 partly automatic construction introduced some spurious patterns into
 the data (\citealt{marelli2014semeval}, $\S$6). Finally, the
-Denotation Graph data set \cite{hodoshimage} contains millions of
-examples, but it was constructed using fully automatic methods, and is
+Denotation Graph entailment set \cite{hodoshimage} contains millions of
+examples of entailments between sentences and artificially constructed
+short phrases, but it was labeled using fully automatic methods, and is
 noisy enough that it is probably suitable only as a source of
 supplementary training data.
+\todo{also cf. PPDB?}
 
@@
 instance, the sentence pair \word{A boat sank in the Pacific
-  ocean} and \word{A boat sank in the ocean} could reasonably in
+  ocean} and \word{A boat sank in the ocean} could reasonably be labeled in
 numerous ways depending on how the objects and events are grounded. If
-the two boats are different, then the sentences are semantically
-independent. If the boats are the same and the oceans are the same,
+the two boats are assumed to be different, then the sentences are semantically
+independent. If the boats and the oceans are assumed to be the same,
 then the first sentence entails the second (in the relevant
-commonsense way). If the boats are the same and the oceans are
-different, then the two events are (commonsense) contradictory. This
+commonsense way). If the boats are assumed to be the same and 
+\ii{the ocean} in the second sentence is interpreted with a specific referent 
+that is not the Pacific, then the two events are (commonsense) contradictory. This
 kind of indeterminacy can be resolved only once the questions of
@@
 
-For the premises, we used the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions from that corpus (corresponding to about 30k images) collected in an earlier crowdsourced effort.
-The captions were not authored by the photographers who took the source images, they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \textit{``Our trip to the Olympic Peninsula''}).
+For the premises, we used the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions  (corresponding to about 30k images) collected in an earlier crowdsourced effort.
+The captions were not authored by the photographers who took the source images, they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \textit{Our trip to the Olympic Peninsula}).
 We presented each caption to at least one worker, and some to more than one.
@@
 \begin{figure}
-\footnotesize
+\begin{framed}
+\small
 We will show you the caption for a photo. We will not show you the photo. Using only the caption and what you know about the world:
@@
 \item Write one alternate caption that is \textbf{definitely} a \textbf{true} description of the photo.
-\ii{Example: For the caption "Two dogs are running through a field." you could write "There are animals outdoors."}
+\ii{Example: For the caption ``\ii{Two dogs are running through a field.}'' you could write ``\ii{There are animals outdoors.}"}
 \item Write one alternate caption that \textbf{might be} a \textbf{true} description of the photo.
-\ii{Example: For the caption "Two dogs are running through a field." you could write "There are animals outdoors."}
+\ii{Example: For the caption ``\ii{Two dogs are running through a field.}" you could write ``\ii{Some puppies are running to catch a stick.}"}
 \item Write one alternate caption that is \textbf{definitely} a \textbf{false} description of the photo.
-\ii{Example: For the caption "Two dogs are running through a field." you could write "There are animals outdoors." This is different from the maybe correct category because its impossible for the dogs to be both running and sitting.}
+\ii{Example: For the caption ``\ii{Two dogs are running through a field.}" you could write ``\ii{The pets are sitting on a couch.}" This is different from the maybe correct category because it's impossible for the dogs to be both running and sitting.}
 \end{itemize}
-\todo{Fix the broken examples here (SB)}
+\end{framed}
+
 \caption{\label{instructions-1}The instructions used on Mechanical Turk for data collection.}
@@
 Though we used a very restrictive qualification (based on past approval rate) to select workers, we nonetheless discovered and deleted some instances of random guessing during an early round of validation, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
+\todo{Leave in mention of qual?}
 
@@
 If there was no such consensus, which occurred in about 0.2\% of cases, we chose the label `-'.
-While these unlabeled examples are included in the corpus distribution, they are unlikely to be helpful for the core classification task of NLI, and we do not include them when training and evaluating the models discussed in later sections.
+While these unlabeled examples are included in the corpus distribution, they are unlikely to be helpful for the standard NLI classification task, and we do not include them when training and evaluating the models discussed in later sections.
 The results of this validation process are summarized in Tables \ref{validation-stats} and \ref{validation-freq}, and suggest that while fundamental limits of both our crowdsourcing methods and the task itself impose an upper bound on the consistency of our labels, the data is sufficiently high quality to pose a challenging but realistic machine learning task.
@@
 \subsection{The distributed corpus}
-\todo{Update all statistics to reflect the latest relase candidate.}
 
@@
   \end{tabular}
-  \caption{\label{snli-examples}Randomly chosen examples from the SNLI development set, shown with the gold labels and the full set of annotator labels (abbreviated).}
+  \caption{\label{snli-examples}Randomly chosen examples from the SNLI development set, shown with the gold labels and the full set of annotator labels (abbreviated).
+Unlike in many existing NLI corpora, the premise and hypothesis sentences in SNLI are often too dissimilar in structure to support approaches to NLI that rely on building sentence-to-sentence alignments.}
 \end{table*}
@@
 Qualitatively, we find the data that we collected to draw fairly extensively on commonsense knowledge, suggesting that superficial word alignment models are unlikely to be able to succeed at classification.
-We find the sentences that we collected be largely fluent, correctly spelled English, with a mix of full sentences and bare noun phrases of a style that is appropriate for the underlying caption task.
-Punctution and capitalization are only intermittently presE
+We find the sentences that we collected be largely fluent, correctly spelled English, with a mix of full sentences and caption-style noun phrase fragments.
+Punctution and capitalization are only intermittently present.
 
@@
 
-The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser \cite{klein2003accurate}, both in the standard PTB format and in the binarized unlabeled format used by tree-structured neural network models, as in \S\ref{sentence-embedding}.
+The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser \cite{klein2003accurate} (trained on the standard training set as well as on the Brown Corpus; \citealt{francis1979brown}), both in the standard PTB format and in the binarized unlabeled format used by tree-structured neural network models, as in \S\ref{sentence-embedding}.
 
-\todo{Update and expand (CM, SB)}
+\todo{Update as appropriate (CM)}
 

### 2015-05-26 writing/SNLI paper/entailment_eval.tex
Paper update.
Add PDF of learning curve
Misc, including rc3 stats.
Add numbers for the lexical baseline classifier
Assorted updates, new paper stats
Paper update
Updating paper to match details of rc3.
Lots of paper cleanup.

@@
 For comparision, results are also reported on the RTE-3 dataset.
-\todo{numbers are approx.}
+\todo{numbers are still dev.}
 }
@@
 %
-% Some definitions
-\def\t#1{\small{#1}}
-\def\b#1{\t{\textbf{#1}}}
-\def\colspaceS{2.0mm}
-\def\colspaceM{3.0mm}
-\def\colspaceL{4.0mm}
 
@@
 \begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceL}c}
+\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceM}c}
 \hline
-\textbf{System} & \multicolumn{4}{c}{\b{2-Class}} & \b{3-Class} \\
- & \t{P} & \t{R} & \t{F$_1$} & \t{Acc.} & \t{Acc.} \\
+\textbf{System} & \b{Dev} & \multicolumn{4}{c}{\b{Test}} \\
+ & \t{Acc.} & \t{P} & \t{R} & \t{F$_1$} & \t{Acc.} \\
 \hline
-\t{Lexicalized}            & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} & \t{77.3} \\
-\t{Unigrams Only}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
-\t{Unlexicalized}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} \\
+\t{Lexicalized}            & \t{??.?} & \t{79.4} & \t{82.3} & \t{80.9} & \t{78.2} \\
+\t{Unigrams Only}          & \t{??.?} & \t{71.1} & \t{77.7} & \t{74.3} & \t{71.6} \\
+\t{Unlexicalized}          & \t{??.?} & \t{56.0} & \t{64.4} & \t{59.9} & \t{50.39} \\
 \hline
@@
   on the 2-class task to allow for comparision with the RTE systems.
-Ablation studies are reported for removing bigram features and all lexical
+Ablation results are reported for models lacking bigram features and lacking all lexical
   features.
-\todo{numbers are approx.}
+\todo{fill in dev numbers}
 }
@@
 % Commenting out until PDF available. -S
-%\Fig{learning_curves_bow.pdf}{0.6}{bowlearncurve}{
-%A learning curve for the lexicalized and unlexicalized baseline classifiers.
-%Note the y axis starting at a random-chance accuracy of 33\%.
-%\todo{numbers are approx.}
-%}
+% Running 'make' should generate PDF from gnuplot
+\Fig{learning_curves_bow.pdf}{0.6}{bowlearncurve}{
+A learning curve for the lexicalized and unlexicalized baseline classifiers,
+plotted on a log scale.
+Note the y axis starting at a random-chance accuracy of 33\%.
+\todo{numbers are approx.}
+}
 
@@
 
-The word embeddings for all of the moedels are initialized the 200d reference GloVe vectors \cite{pennington2014glove} and fine tuned.
-In addition, all of the models use an additional $\tanh$ neural network layer to map these 200d embeddings into the 50d working space.
+The word embeddings for all of the moedels are initialized the 300d reference GloVe vectors \cite{pennington2014glove} and fine tuned.
+In addition, all of the models use an additional $\tanh$ neural network layer to map these 300d embeddings into the lower-dimensianal phrase and sentence embedding space.
 All of the models are randomly initialized using standard techniques and trained using AdaDelta \cite{zeiler2012adadelta} minibatch SGD until performance on the development set stops improving.
@@
 \hline
-\t{Sum of words}            & \t{} & \t{} \\
-\t{RNN}            & \t{} & \t{} \\
-\t{LSTM RNN}            & \t{75?} & \t{72?} \\
-\t{TreeRNN}            & \t{} & \t{} \\
-\t{TreeRNTN}            & \t{58?} & \t{56?} \\
-\t{LSTM TreeRNN}            & \t{72?} & \t{69?} \\
+\t{100d Sum of words}            & \t{79?} & \t{76?} \\
+\t{100d RNN}            & \t{?} & \t{68} \\	
+\t{100d LSTM RNN}            & \t{75?} & \t{76?} \\
+\t{100d TreeRNN}            & \t{69?} & \t{69?} \\
+\t{50d TreeRNTN}            & \t{58?} & \t{56?} \\
+\t{100d LSTM TreeRNN}            & \t{75?} & \t{73?} \\
 \hline

### 2015-05-26 writing/SNLI paper/intro.tex
Lots of paper cleanup.

@@
 \cite{socher2013acl1}, but it is less clear that they can be
-trained to support the full range of logical inferences required for
-NLI. In the SemEval 2014 task aimed at evaluating distributed
+trained to support the full range of logical and commonsense
+ inferences required for NLI. In the SemEval 2014 task aimed at evaluating distributed
 representations for NLI, the best-performing systems relied heavily on
@@
 To address this, we present a new corpus of sentence pairs labeled for
-entailment, contradiction, and semantic independence. At 550,152
+entailment, contradiction, and semantic independence. At 570,152
 sentence-pairs, this corpus is orders of magnitude larger than all
@@
 naturalistic context. In a separate validation phase, we collected
-four additional judgments for each label for 56,941 of the examples,
-and 98\% of cases emerge with a gold label, which attests to the high
-quality of the data.
+four additional judgments for each label for 56,941 of the examples.
+Of these, 98\% of cases emerge with a three annotator consensus, 
+and 58\% see a unanimous consensus from all five annotators, 
+which attests to the high quality of the data.
 
@@
 of NLI in semantics.
+
+\todo{Plain LSTMs strong, used in transfer.}
 

### 2015-05-26 writing/SNLI paper/model_structure_fig.tex
Lots of paper cleanup.

@@
 
-        \caption{The common classification architecture used in all NLI evaluations: for each model evaluated in Table~\ref{tab:nnresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 50d inputs shown here.}
+        \caption{The classification architecture used in all NLI evaluations: for each model evaluated in Table~\ref{tab:nnresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 50d inputs shown here.
+In some transfer learning experiments (\S\ref{sec:transfer}), the premise embedding can be replaced with a zero vector for single-sentence tasks.}
   \label{modelstructure}

### 2015-05-26 writing/SNLI paper/transfer.tex
Lots of paper cleanup.

@@
-\section{SNLI and transfer learning}
+\section{SNLI and transfer learning}\label{sec:transfer}
 
@@
 We train on both types of label within the 8.5k sentence training set and test on only the sentence-level labels in the test set.
-We finally evaluate on FactBank.
-\todo{FactBank} ....
+We finally evaluate on PragBank, an 800 example 6-way sentence classification task focused on veridicality.
+\todo{Cite} ....
 
- For SICK, we use pretrained parameters for all of the LSTM parameters, pretrained parameters for the $\tanh$ layers in the classifier, and pretrained embeddings for all of the words that occurred in both corpora, randomly initializing only the softmax layer and the remaining word embeddings.
-We use a similar strategy for the other three tasks, but do not transfer the initialization for the $\tanh$ layers, since the switch between sentence pair classification (in SNLI) and single sentence classification means that the input sizes for the bottom $\tanh$ do not match.
-In both cases we re-initialize AdaDelta as for a new model run.
+We use pretrained parameters for all of the LSTM parameters, pretrained parameters for the $\tanh$ layers in the classifier stack, and pretrained embeddings for all of the words that occurred in both corpora, randomly initializing only the softmax layer and the remaining word embeddings.
+In order to use this starategy in sentence classification tasks, we pass the classifier the embedding of the sentence to be classified in the hypothesis input position, as well as a zero vector in the premise input position.
+After transfer, we re-initialize the accumulators in AdaDelta to re-set the learning rates to high values.
+We additionally use L2 regularization with each task, separately tuning the regularization strength on the transfer model and the corresponding baseline model without transfer.
 
+\todo{Do we still want to re-initialize LRs?}
 \todo{Do we want to use all four target tasks?}
@@
 \hline
-\t{SICK}            & \t{98} & \t{95} & \t{63?} & \t{69?} & \t{84.5} \\
+\t{SICK}            & \t{98} & \t{95} & \t{63?} & \t{80.0?} & \t{84.5} \\
 \t{SUBJ}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?}& \t{??.?} \\
 \t{SST}          & \t{??.?} & \t{??.?} & \t{43?} & \t{43?} & \t{50.6}\\
-\t{FactBank?}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?}& \t{??.?} \\
+\t{PragBank}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?}& \t{??.?} \\
 \hline
@@
 \caption{\label{tab:transferresults}
-The performance of an LSTM RNN classification model on four tasks under both a random initialization and an initialization copied from a model trained on SNLI. \todo{Final numbers.}}
+The performance of an LSTM RNN model on four tasks under both a random initialization and a transfer initialization copied from a model trained on SNLI.
+All figures are classification accuracy, except for PragBank, where report macroaveraged F1 to enable comparison with existing results on that task.
+\todo{Final numbers.}}
 
 \end{table}
-..
-The final column shows the state of the art results from \cite{lai2014illinois}, TODO, and \cite{tai2015improved}, respectively
-...
+
+The final column shows the state of the art results from \cite{lai2014illinois}, TODO, \cite{tai2015improved}, TODO, respectively.
+\todo{Does Tai have the SotA on SST?}
 
@@
 Note that \todo{something interesting happens}.
-Our transfer model shows \todo{something about the state of the art on the tasks.}
+Within the domain of NLI, our transfer model shows the best number yet reported on SICK for a fully-learned model, and approaches the overall state of the art (and to the 84\% level of interannotator agreement, a likely performance cieling), while the randomly initialized model performs poorly.
+This suggests that the introduction of SNLI makes it newly possible to train representation learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on semantically sophisticated tasks.
+
+\todo{something about the state of the art on the other tasks.}

### 2015-05-26 writing/SNLI paper/intro.tex
Tweaks concerning theory contributions

@@
 modeling simple dimensions of meaning like evaluative sentiment
-\cite{socher2013acl1}, but it is less clear that they can be
+(e.g., \citealt{socher2013acl1}), but it is less clear that they can be
 trained to support the full range of logical and commonsense
@@
 evaluation of a wide range of models for distributed semantic
-representations in the context of NLI. However, in our view, the
-existing corpus resources in this area do not permit such an
-assessment. They are generally too small for training modern
-data-intensive, wide-coverage models, and they are often beset with
+representations in the context of NLI, thereby deepening the case for
+NLI as a powerful evaluation of domain-general approaches to semantic
+representation. However, in our view, existing NLI corpora do not
+permit such an assessment. They are generally too small for training
+modern data-intensive, wide-coverage models, many contain sentences
+that were algorithmically generated, and they are often beset with
 indeterminacies of event and entity coreference that significantly

### 2015-05-26 writing/SNLI paper/corpus.tex
Tweaks  and now ref to PPDB

@@
 have stimulated innovative logical and statistical models of natural
-language reasoning, but their small size of a few hundred examples each
+language reasoning, but their small size (a few hundred examples each)
 limits their utility as a testbed for learned distributed representations. 
@@
 noisy enough that it is probably suitable only as a source of
-supplementary training data.
-\todo{also cf. PPDB?}
+supplementary training data. The Paraphase Database of \cite{ganitkevitch2013ppdb}
+addresses these issues of size and quality, but only for the limited
+fragment of the NLI problem that is captured by paraphase.
 
@@
 commonsense way). If the boats are assumed to be the same and 
-\ii{the ocean} in the second sentence is interpreted with a specific referent 
+\word{the ocean} in the second sentence is interpreted with a specific referent 
 that is not the Pacific, then the two events are (commonsense) contradictory. This
@@
 
-For the premises, we used the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions  (corresponding to about 30k images) collected in an earlier crowdsourced effort.
-The captions were not authored by the photographers who took the source images, they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \textit{Our trip to the Olympic Peninsula}).
+For the premises, we used the Flickr30k corpus \cite{hodoshimage}, a database of approximately 160k captions  (corresponding to about 30k images) collected in an earlier crowdsourced effort.
+The captions were not authored by the photographers who took the source images, and they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \word{Our trip to the Olympic Peninsula}).
 We presented each caption to at least one worker, and some to more than one.
-With the exception of a handful that were flagged by workers as excessively difficult, at least three labeled pairs are available for each caption.
+With the exception of a handful that were flagged by workers as excessively difficult, our corpus contains at least three labeled pairs for each caption.
 
@@
 Below the instructions were three fields for each of three requested sentences, corresponding to our \ii{entailment}, \ii{neutral}, and \ii{contradiction} labels, a fourth field (marked optional) for reporting problems, and a link to an FAQ page.
-That FAQ grew over the course of data collection, but discussed disallowed techniques (e.g., reusing the same sentence for many different prompts, which we saw in a few cases), sentence length and complexity (we did not enforce a minimum length, and we allowed bare NPs as well as full sentences), and logistical issues around payment timing.
-About 2500 workers contributed data.
+That FAQ grew over the course of data collection.
+It warned about disallowed techniques (e.g., reusing the same sentence for many different prompts, which we saw in a few cases), provided guidance concerning sentence length and complexity (we did not enforce a minimum length, and we allowed bare NPs as well as full sentences), and reviewed logistical issues around payment timing.
+About 2,500 workers contributed data.
 
@@
 
-The majority of the sentence pairs in the corpus have not been reviewed by anyone since their construction as part of the data collection Mechanical Turk task, except to remove a small number of responses where either a response field was empty or the original author flagged the source caption for review.
-In order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we did perform an additional round of validation for about 10\% of our data.
-
-This validation phase is not fundamentally different from the Mechanical Turk labeling task used to label the SICK entailment data: we present workers with pairs of sentences (in our case, in batches of five at a time), and ask them to choose a single label for each pair.
-We supplied each pair to four annonators, yielding five labels per pair including the label used by the original composer of the pair.
+% I suggest removing this sentence. It sounds like a negative, but
+% there is no practical alternative except what we did do.
+%
+% The majority of the sentence pairs in the corpus have not been reviewed by anyone since their construction as part of the data collection Mechanical Turk task, except to remove a small number of cases where either a response field was empty or the original author flagged the source caption for review.
+In order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we performed an additional round of validation for about 10\% of our data.
+This validation phase was not fundamentally different from the Mechanical Turk labeling task used to label the SICK entailment data: we presented workers with pairs of sentences (in our case, in batches of five at a time), and ask them to choose a single label for each pair.
+We supplied each pair to four annonators, yielding five labels per pair including the label used by the original author.
 The instructions were broadly similar to the instructions for initial data collection shown in Fig.~\ref{instructions-1}, and linked to a similar FAQ.
 Though we used a very restrictive qualification (based on past approval rate) to select workers, we nonetheless discovered and deleted some instances of random guessing during an early round of validation, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
-\todo{Leave in mention of qual?}
+\todo{Leave in mention of qual? I would--CP}
 
@@
 While these unlabeled examples are included in the corpus distribution, they are unlikely to be helpful for the standard NLI classification task, and we do not include them when training and evaluating the models discussed in later sections.
-The results of this validation process are summarized in Tables \ref{validation-stats} and \ref{validation-freq}, and suggest that while fundamental limits of both our crowdsourcing methods and the task itself impose an upper bound on the consistency of our labels, the data is sufficiently high quality to pose a challenging but realistic machine learning task.
+The results of this validation process are summarized in Tables \ref{validation-stats}, \ref{kappas}, and \ref{validation-freq}, and suggest that while fundamental limits of both our crowdsourcing methods and the task itself impose an upper bound on the consistency of our labels, the data is sufficiently high quality to pose a challenging but realistic machine learning task.
 
@@
 
-Table~\ref{snli-examples} shows a set of randomly chosen validated examples from the development sent with their labels.
+Table~\ref{snli-examples} shows a set of randomly chosen validated examples from the development set with their labels.
 Qualitatively, we find the data that we collected to draw fairly extensively on commonsense knowledge, suggesting that superficial word alignment models are unlikely to be able to succeed at classification.
 We find the sentences that we collected be largely fluent, correctly spelled English, with a mix of full sentences and caption-style noun phrase fragments.
-Punctution and capitalization are only intermittently present.
+Punctuation and capitalization are only intermittently present.
+
+% \marginpar{{\tiny Does this mean that expected punctuation is missing, 
+% or that the sentences required little punctuation even by 
+% prescriptive standards?}}
 
@@
     % From 1.0rc3
-\caption{\label{validation-stats}Key statistics for the validated sentence pairs.} 
+\caption{\label{validation-stats}Key statistics for the validated pairs.} 
+\end{table}
+
+\begin{table}
+\center
+  %\setlength{\tabcolsep}{15pt}
+  %\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{l r} 
+    \toprule
+    contradiction & 0.77 \\
+    entailment & 0.72 \\
+    neutral & 0.60 \\[1ex]
+    overall & 0.70 \\
+    \bottomrule
+  \end{tabular}
+    % From 1.0rc3
+\caption{\label{kappas}Fleiss $\kappa$ statistics.} 
 \end{table}

### 2015-05-26 writing/SNLI paper/entailment_eval.tex
Update the baselines section

@@
   \citealt{pado2014design,magnini2014excitement})
-  -- an open source platform for RTE research which
-  is distributed alongside a number of RTE pipelines.
-We additionally evaluate against a both an unlexicalized and lexicalized
-  classifier with straightforward feature sets.
-\todo{some meaningful insight from \#s}
+  -- an open source platform for RTE research.
+%  which
+%  is distributed alongside a number of RTE pipelines.
+We additionally evaluate against a strong classifier-based baseline based on
+  both  unlexicalized and lexicalized features.
 
@@
 
-% methodology
-We convert the 3-way classification task in SNLI into the RTE setting
-  by labeling both the \unknown\ and \contradiction\
-  labels as negative entailment, and treating the \entailment\ label as
-  the positive entailment.
-This creates a biased dataset of 66\% negative examples.
+%% methodology
+%We convert the 3-way classification task in SNLI into the RTE setting
+%  by labeling both the \unknown\ and \contradiction\
+%  labels as negative entailment, and treating the \entailment\ label as
+%  the positive entailment.
+%This creates a biased dataset of 66\% negative examples.
 % we report the best results from each class
-We run each of the 9 algorithms distributed with EOP on the 2-class
-  SNLI dataset, and report results for the best edit distance 
-  configuration and the best classifier based configuration, as
+We run the top performing edit-distance based algorithm and the top
+  performing classifier-based algorithm on the SNLI test set, as
   determined by performance on the development set.
-\todo{we may not have CPU time for this}.
+Note that these models were run using the default configuration
+  with minimal tuning.
+The results should therefore be taken as a strong baseline for
+  RTE-style approaches to the problem, rather than necessarily
+  representing the state-of-the-art system's performance on the
+  task.
+%We run each of the 9 algorithms distributed with EOP on the 2-class
+%  SNLI dataset, and report results for the best edit distance 
+%  configuration and the best classifier based configuration, as
+%  determined by performance on the development set.
 
@@
                     \t{??.?} & \t{??.?} \\
-\t{Classifier}    & \t{68.0} & \t{38.7} & \t{49.3} & \t{73.5} & 
+\t{Classifier}    & \t{67.9} & \t{38.7} & \t{49.3} & \t{7X.X} & 
                     \t{64.2} & \t{65.3} \\
@@
 \label{tab:eopresults}
-2-class accuracy and precision/recall of existing competitive RTE
+3-class accuracy and precision/recall of existing competitive RTE
 systems from the Excitement Open Platform.
@@
   classifier-based system.
-For comparision, results are also reported on the RTE-3 dataset.
+In order to calibrate the difficulty of the dataset against the 
+  classical RTE problems, results are also reported on the (2-class) 
+  RTE-3 dataset.
 \todo{numbers are still dev.}
@@
   after removing stop words.
-The best classifier-based system \todo{does something}.
-\todo{some insight from results}.
+The best classifier-based system makes use of information from
+  WordNet \cite{miller1995wordnet} and VerbOcean
+  \cite{chklovski2004verbocean}, and makes use of features
+  based on tree patterns and dependency tree skeletons
+  \cite{wang2007recognizing}.
+Unsurprisingly, the classification-based approach outperforms simple
+  edit distance metrics, and performs quite well despite relatively
+  little lexicalization.
+
+% TODO(Sam): Can you add these to the Bibtex?
+%
+%@article{miller1995wordnet,
+%  title={{W}ord{N}et: a lexical database for English},
+%  author={Miller, George A},
+%  journal={Communications of the ACM},
+%  volume={38},
+%  number={11},
+%  pages={39--41},
+%  year={1995},
+%  publisher={ACM}
+%}
+%
+%@inproceedings{chklovski2004verbocean,
+%  title={VerbOcean: Mining the Web for Fine-Grained Semantic Verb Relations.},
+%  author={Chklovski, Timothy and Pantel, Patrick},
+%  booktitle={EMNLP},
+%  year={2004}
+%}
+%
+%@inproceedings{wang2007recognizing,
+%  title={Recognizing textual entailment using sentence similarity based on dependency tree skeletons},
+%  author={Wang, Rui and Neumann, G{\"u}nter},
+%  booktitle={ACL-PASCAL Workshop on Textual Entailment and Paraphrasing},
+%  year={2007},
+%}
 
@@
   rich lexicalized features.
-We implmement a lexicalized classifier to provide a strong baseline for future work
-  on the dataset.
+We implmement such a lexicalized classifier as a strong baseline 
+  for the task.
 Our classifier implements 6 features:
-\begin{enumerate}  % TODO(gabor) This takes up way too much space
-\setlength\itemsep{0em}
+\begin{enumerate}
+\setlength\itemsep{-0.25em}
   \item the BLEU score of the \hypothesis\ with respect
@@
 
-  \item The overlap between words in the \premise\ and \hypothesis. This is counted
-  both as an absolute count and a percentage of possible overlap, and is computed
-  both over all words and over just nouns, verbs, adjectives, and adverbs.
+  \item The overlap between words in the \premise\ and \hypothesis. We include
+  both as an absolute count and a percentage of possible overlap, and compute
+  the count both over all words and over just nouns, verbs, adjectives, 
+  and adverbs.
 
-  \item For every pair of words across the \premise\ and \hypothesis\ which share a 
+  \item\label{lst:ngram} An indicator for every unigram and bigram in the \hypothesis.
+
+  \item\label{lst:unigram} Cross-unigrams: for every pair of words across the \premise\ and \hypothesis\ which share a 
   POS tag, an indicator feature over the two words.
 
-  \item For every pair of bigrams across the \premise\ and \hypothesis\ which share a 
+  \item\label{lst:bigram} Cross-bigrams: for every pair of bigrams across the \premise\ and \hypothesis\ which share a 
   POS tag, an indicator feature over the two bigrams.
-
-  \item An indicator for every unigram and bigram in the \hypothesis.
 \end{enumerate}
@@
 \label{tab:bowresults}
-Accuracy on the 2-class and 3-class SNLI dataset, as well as precision/recall
-  on the 2-class task to allow for comparision with the RTE systems.
-Ablation results are reported for models lacking bigram features and lacking all lexical
-  features.
+Accuracy on the SNLI dataset, including precision/recall for the
+  \textit{entailed} class on the test set.
+Ablation results are reported for models lacking cross-bigram features 
+  (Feature \ref{lst:bigram}), and lacking all lexical
+  features (Features \ref{lst:ngram}--\ref{lst:bigram}).
 \todo{fill in dev numbers}
@@
 % Running 'make' should generate PDF from gnuplot
-\Fig{learning_curves_bow.pdf}{0.6}{bowlearncurve}{
+\Fig{learning_curves_bow.pdf}{0.45}{bowlearncurve}{
 A learning curve for the lexicalized and unlexicalized baseline classifiers,
@@
 Note the y axis starting at a random-chance accuracy of 33\%.
-\todo{numbers are approx.}
 }
@@
 
-Note that \todo{something interesting happens}.
+% Insights 1: lexicalization helps a bunch
+It is interesting to note the substantial jump in accuracy from using
+  lexicalized features, and even from using the relatively sparse
+  cross-bigram features.
+In fact, the addition of these features alone allow the classifier to
+  outperform the RTE baselines -- this is in some sense unsurprising:
+  the RTE systems have been tuned on relatively small corpora
+  (~1600 examples), whereas a classifier trained on SNLI can make use of
+  over two orders of magnitude more data.
+
+% Insights 2: the learning curve for the unlexicalized classifier is sad
+More surprising is how quickly the lexicalized classifier outperforms its
+  unlexicalized counterpart (see \reffig{bowlearncurve}).
+With only 100 training examples, the cross-bigram classifier is already
+  outperforming its unlexicalized counterpart.
+Empirically we notice that the top weighted features for the classifier
+  trained on 100 examples tend to be high precision entailments;
+  e.g.,
+  \textit{playing} $\rightarrow$ \textit{outside}
+  (most scenes are outdoors), or \textit{a banana} $\rightarrow$
+  \textit{person eating}.
+If relatively few spurious entailments get high weight -- as it appears
+  is the case -- then it makes sense that when these do fire they
+  provide some boost in accuracy.
 

### 2015-05-26 writing/SNLI paper/abstract.tex
Rough draft intro
Lots of paper cleanup.

@@
 \begin{abstract} 
+  % Entailment and contradiction are fundamental to natural language
+  % meaning and lie at the heart of tasks like summarization, textual
+  % entailment, and commonsense reasoning.
   Entailment and contradiction are fundamental to natural language
-  meaning and lie at the heart of tasks like summarization, textual
-  entailment, and commonsense reasoning.
+  meaning and so have emerged (in the form of tasks involving
+  reasoning, inference, and summarization) as an important testing
+  ground for semantic representations.
   %
-  However, computational modeling of these concepts is currently
-  hampered by the limited nature of existing annotated resources.
+  %However, computational modeling of these concepts is currently
+  %hampered by the limited nature of existing annotated resources.
+  However, progess is currently hampered by the limitations of
+  existing annotated resources.
   %
   To address this, we present a new corpus of sentence pairs labeled
-  for entailment, contradiction, and semantic independence. At 550K
-  pairs, it is an order of magnitude larger than all other resources
+  for entailment, contradiction, and semantic independence. At 570K
+  pairs, it is orders of magnitude larger than all other resources
   of its type.
@@
   tasks.  
+  \todo{Structure more around results (vs. data).}
 \end{abstract}

### 2015-05-26 writing/SNLI paper/corpus.tex
Better connections between prose and tables

@@
 In order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we performed an additional round of validation for about 10\% of our data.
-This validation phase was not fundamentally different from the Mechanical Turk labeling task used to label the SICK entailment data: we presented workers with pairs of sentences (in our case, in batches of five at a time), and ask them to choose a single label for each pair.
+This validation phase was not fundamentally different from the Mechanical Turk labeling task used to label the SICK entailment data: we presented workers with pairs of sentences (in our case, in batches of five), and ask them to choose a single label for each pair.
 We supplied each pair to four annonators, yielding five labels per pair including the label used by the original author.
@@
 
-For each pair that we validated, we chose a gold label.
-If any one of the three labels was chosen by at least three of the five annotators, that label is the gold label.
-If there was no such consensus, which occurred in about 0.2\% of cases, we chose the label `-'.
-While these unlabeled examples are included in the corpus distribution, they are unlikely to be helpful for the standard NLI classification task, and we do not include them when training and evaluating the models discussed in later sections.
-The results of this validation process are summarized in Tables \ref{validation-stats}, \ref{kappas}, and \ref{validation-freq}, and suggest that while fundamental limits of both our crowdsourcing methods and the task itself impose an upper bound on the consistency of our labels, the data is sufficiently high quality to pose a challenging but realistic machine learning task.
+For each pair that we validated, we chose a gold label. If any one of
+the three labels was chosen by at least three of the five annotators,
+that label is the gold label. If there was no such consensus, which
+occurred in about 0.2\% of cases, we chose the label `-'. While these
+unlabeled examples are included in the corpus distribution, they are
+unlikely to be helpful for the standard NLI classification task, and
+we do not include them when training and evaluating the models
+discussed in later sections. The results of this validation process
+are summarized in Tables \ref{validation-stats}, \ref{kappas}, and
+\ref{validation-freq}. Nearly all of the examples received a majority
+label, indicating broad consensus about the nature of the data and
+categories. The gold-labeled examples are essentially evenly
+distributed across the three categories in our train, development, and
+testing (Table~\ref{validation-freq}). The Fleiss $\kappa$ scores in
+Table~\ref{kappas} are likely to be conservative given our large and
+unevenly distributed pool of workers, but they still provide insights
+about the levels of disagreement across the three semantic
+classes. This disagreement likely reflects not just the limitations of
+large crowdsourcing efforts but also the uncertainty inherent in NLI.
+Regardless, the overall rates of agreement is extremely high,
+suggesting that the corpus is sufficiently high quality to pose a
+challenging but realistic machine learning task.
 
@@
 
-% \marginpar{{\tiny Does this mean that expected punctuation is missing, 
-% or that the sentences required little punctuation even by 
-% prescriptive standards?}}
+% Does this mean that expected punctuation is missing, or that the
+% sentences required little punctuation even by prescriptive
+% standards?
 
@@
 \begin{table}
-\center
-  %\setlength{\tabcolsep}{15pt}
-  %\renewcommand{\arraystretch}{1.2}
+  \centering
   \begin{tabular}{l r} 
@@
 
+%%%%% This is the first table I would cut to save space.
+%%%%% I summarized its content quickly in the prose. CP
+%
 \begin{table}

### 2015-05-26 writing/SNLI paper/corpus.tex
Experiment updates, and nicer stats tables.

--- 2015-05-26 writing/SNLI paper/corpus.tex
@@
 While these unlabeled examples are included in the corpus distribution, they are unlikely to be helpful for the standard NLI classification task, and we do not include them when training and evaluating the models discussed in later sections.
-The results of this validation process are summarized in Tables \ref{validation-stats}, \ref{kappas}, and \ref{validation-freq}, and suggest that while fundamental limits of both our crowdsourcing methods and the task itself impose an upper bound on the consistency of our labels, the data is sufficiently high quality to pose a challenging but realistic machine learning task.
+The results of this validation process are summarized in Tables \ref{validation-stats}, and \ref{validation-freq}, and suggest that while fundamental limits of both our crowdsourcing methods and the task itself impose an upper bound on the consistency of our labels, the data is sufficiently high quality to pose a challenging but realistic machine learning task.
 
@@
     % From 1.0rc2
+\todo{Replace w/ mean+SD if low on space.}
 \caption{\label{b-table}The distribution of sentence length.} 
@@
     \toprule
+\multicolumn{2}{l}{\textbf{Data set sizes:}}\\
 Training pairs &  550,152\\
@@
 \midrule
+\multicolumn{2}{l}{\textbf{Sentence length:}}\\
 Premise mean token count & 14.1\\
@@
 \midrule
+\multicolumn{2}{l}{\textbf{Frequency of complete sentences:}}\\
 Premise `S'-rooted parses & 74.0\%\\
@@
 % From 1.0rc3
-\caption{\label{collection-stats}Key statistics for the raw sentence pairs. Since the two halves of each pair were collected separately, we report statistics for both.} 
+\caption{\label{collection-stats}Key statistics for the raw sentence pairs. Since the two halves of each pair were collected separately, we report some statistics for both.} 
 \end{table}
@@
     \toprule
+\multicolumn{2}{l}{\textbf{General:}}\\
 Validated pairs & 56,941\\
-gold label $\ne$ original label & 6.8\%\\
-Unanimous gold & 58.3\%\\
-No gold (no 3 labels match) & 2.0\%\\
-Label $=$ gold & 89.0\%\\
-Label $=$ original author's label & 85.8\%\\
+Pairs w/ unanimous gold label & 58.3\%\\
+\midrule
+\multicolumn{2}{l}{\textbf{Individual annotator label agreement:}}\\
+Annotator label $=$ gold & 89.0\%\\
+Annotator label $=$ author's label & 85.8\%\\
+\midrule
+\multicolumn{2}{l}{\textbf{Gold label/author's label agreement:}}\\
+Gold label $=$ author's label & 91.2\%\\
+Gold label $\ne$ author's label & 6.8\% \\
+No gold label (no 3 labels match) & 2.0\%\\
+\midrule
+\multicolumn{2}{l}{\textbf{Fleiss $\kappa$:}}\\
+    \ii{contradiction} & 0.77 \\
+    \ii{entailment} & 0.72 \\
+    \ii{neutral} & 0.60 \\
+    Overall & 0.70 \\
     \bottomrule
@@
     % From 1.0rc3
-\caption{\label{validation-stats}Key statistics for the validated pairs.} 
-\end{table}
-
-\begin{table}
-\center
-  %\setlength{\tabcolsep}{15pt}
-  %\renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{l r} 
-    \toprule
-    contradiction & 0.77 \\
-    entailment & 0.72 \\
-    neutral & 0.60 \\[1ex]
-    overall & 0.70 \\
-    \bottomrule
-  \end{tabular}
-    % From 1.0rc3
-\caption{\label{kappas}Fleiss $\kappa$ statistics.} 
+\caption{\label{validation-stats}Key statistics for the validated pairs.
+The \ii{author's label} is the label that was used by the worker who wrote the premise to create the sentence pair.
+A \ii{gold label} reflects a consensus of three votes among the author and the four annotators.}
 \end{table}

### 2015-05-26 config/Defaults.m

@@ -26,7 +26,7 @@ hyperParams.eyeScale = 0.5;
 % Which initialization scheme to use
 hyperParams.NNinitType = 1;
 hyperParams.NTNinitType = 1;
-hyperParams.LSTMinitType = 2;
+hyperParams.LSTMinitType = 5;
 
 % How far *in each direction* should the connection classifier in a Lattice model look.
 % Setting this to 1 means to only use the immediate left and right composition inputs with no
@@ -74,7 +74,7 @@ hyperParams.topDropout = 1;
 % the embeddings at those indices will be reinitialized before computing each example. Used for ALCIR.
 hyperParams.randomEmbeddingIndices = [];
 
-hyperParams.restartUpdateRuleInTransfer = true;
+hyperParams.restartUpdateRuleInTransfer = false;
 
 % L1 v. L2 regularization. If no regularization is needed, set
 % lambda to 0 and ignore this parameter.

### 2015-05-26 config/PragBank.m

@@ -9,7 +9,8 @@ hyperParams.name = [expName, '-', dataflag, '-l', num2str(lambda), '-dim', num2s
     '-do', num2str(bottomDropout), '-', num2str(topDropout), '-co', num2str(collo),...
     '-comp', num2str(composition)];
 
-hyperParams.sentenceClassificationMode = 1;
+% Treat this as an entailment problem, with the target word as the hypothesis.
+hyperParams.sentenceClassificationMode = 0;
 
 hyperParams.dim = dim;
 hyperParams.embeddingDim = embDim;
@@ -59,9 +60,9 @@ hyperParams.labels = {{'CT_plus', 'CT_minus', 'PR_plus', 'PR_minus', 'PS_plus', 
 labelMap = cell(1, 1);
 labelMap{1} = containers.Map(hyperParams.labels{1}, 1:length(hyperParams.labels{1}));
 
-hyperParams.trainFilenames = {'pragbank-data/train.txt'};    
+hyperParams.trainFilenames = {'pragbank-data/train-parsed.txt'};    
 hyperParams.splitFilenames = {};    
-hyperParams.testFilenames = {'pragbank-data/test.txt'};
+hyperParams.testFilenames = {'pragbank-data/test-parsed.txt'};
 
 
 if strcmp(dataflag, 'pragbank')

### 2015-05-26 run.sh
Experiment updates, and nicer stats tables.

@@ -2,7 +2,7 @@
 
 ### Generic job script for all experiments.
 
-#PBS -l nodes=1:ppn=6 	### Request at least 4 cores
+#PBS -l nodes=1:ppn=6 	### Request at least 6 cores
 #PBS -l walltime=60:00:00	### Die after 60h
 #PBS -l mem=6000MB
 #PBS -q nlp

### 2015-05-26 utils/ParseThreeColPartial.java

@@ -47,9 +47,9 @@ class ParseThreeColPartial {
 		try {
 			BufferedReader br = new BufferedReader(new FileReader(args[0]));  
 			String line = null;  
-			String[] columnDetail = new String[4];
-			String[] binarized = new String[2];
-			String[] parsed = new String[2];
+			String[] columnDetail = new String[5];
+			String[] binarized = new String[1];
+			String[] parsed = new String[1];
 
 			int num_done = 0;
 			while ((line = br.readLine()) != null) {
@@ -62,7 +62,7 @@ class ParseThreeColPartial {
 						parsed[i - 1] = parse.toString();	
 						binarized[i - 1] = unlabeledPrint(collapsed);	
 					}
-					System.out.println(columnDetail[0] + "\t" + binarized[0] + "\t" + "\t" + parsed[0] + "\t" + "\t" + columnDetail[1] + "\t" + "\t");
+					System.out.println(columnDetail[0] + "\t" + binarized[0] + "\t" + "\t" + parsed[0] + "\t" + "\t" + columnDetail[1] + "\t" + columnDetail[2]  + "\t");
 					num_done++;
 					if (num_done % 1000 == 0) {
 						System.err.println("Finished " + num_done + ".");

### 2015-05-26 writing/SNLI paper/corpus.tex

@@
 
-For each pair that we validated, we chose a gold label.
-If any one of the three labels was chosen by at least three of the five annotators, that label is the gold label.
-If there was no such consensus, which occurred in about 0.2\% of cases, we chose the label `-'.
-While these unlabeled examples are included in the corpus distribution, they are unlikely to be helpful for the standard NLI classification task, and we do not include them when training and evaluating the models discussed in later sections.
-The results of this validation process are summarized in Tables \ref{validation-stats}, and \ref{validation-freq}, and suggest that while fundamental limits of both our crowdsourcing methods and the task itself impose an upper bound on the consistency of our labels, the data is sufficiently high quality to pose a challenging but realistic machine learning task.
+For each pair that we validated, we chose a gold label. If any one of
+the three labels was chosen by at least three of the five annotators,
+that label is the gold label. If there was no such consensus, which
+occurred in about 0.2\% of cases, we chose the label `-'. While these
+unlabeled examples are included in the corpus distribution, they are
+unlikely to be helpful for the standard NLI classification task, and
+we do not include them when training and evaluating the models
+discussed in later sections. The results of this validation process
+are summarized in Tables \ref{validation-stats} and
+\ref{validation-freq}. Nearly all of the examples received a majority
+label, indicating broad consensus about the nature of the data and
+categories. The gold-labeled examples are essentially evenly
+distributed across the three categories in our train, development, and
+testing (Table~\ref{validation-freq}). The Fleiss $\kappa$ scores in
+Table~\ref{kappas} are likely to be conservative given our large and
+unevenly distributed pool of workers, but they still provide insights
+about the levels of disagreement across the three semantic
+classes. This disagreement likely reflects not just the limitations of
+large crowdsourcing efforts but also the uncertainty inherent in NLI.
+Regardless, the overall rates of agreement is extremely high,
+suggesting that the corpus is sufficiently high quality to pose a
+challenging but realistic machine learning task.
 

### 2015-05-26 writing/SNLI paper/corpus.tex
Merge remote-tracking branch 'origin/master'
Conflicts:
	writing/SNLI paper/corpus.tex
Corpus tweaks.
Nix label freq table.

@@
 
-% I suggest removing this sentence. It sounds like a negative, but
-% there is no practical alternative except what we did do.
-%
-% The majority of the sentence pairs in the corpus have not been reviewed by anyone since their construction as part of the data collection Mechanical Turk task, except to remove a small number of cases where either a response field was empty or the original author flagged the source caption for review.
 In order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we performed an additional round of validation for about 10\% of our data.
@@
 Though we used a very restrictive qualification (based on past approval rate) to select workers, we nonetheless discovered and deleted some instances of random guessing during an early round of validation, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
-\todo{Leave in mention of qual? I would--CP}
+%\todo{Leave in mention of qual? I would--CP Settled -SB}
 
@@
 discussed in later sections. The results of this validation process
-are summarized in Tables \ref{validation-stats} and
-\ref{validation-freq}. Nearly all of the examples received a majority
+are summarized in Table \ref{validation-stats}. 
+Nearly all of the examples received a majority
 label, indicating broad consensus about the nature of the data and
@@
 distributed across the three categories in our train, development, and
-testing (Table~\ref{validation-freq}). The Fleiss $\kappa$ scores in
+testing. The Fleiss $\kappa$ scores in
 Table~\ref{kappas} are likely to be conservative given our large and
@@
 We find the sentences that we collected be largely fluent, correctly spelled English, with a mix of full sentences and caption-style noun phrase fragments.
-Punctuation and capitalization are only intermittently present.
-
-% Does this mean that expected punctuation is missing, or that the
-% sentences required little punctuation even by prescriptive
-% standards?
+Punctuation and capitalization are often omitted.
 
@@
 \includegraphics[width=3.05in]{length_dist}
-    % From 1.0rc2
+    % From 1.0rc2, though still accurate as of rc3
 \todo{Replace w/ mean+SD if low on space.}
@@
 
-%%%%% This is the first table I would cut to save space.
-%%%%% I summarized its content quickly in the prose. CP
-%
-\begin{table}
-\center
-  %\setlength{\tabcolsep}{15pt}
-  %\renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{l lll} 
-    \toprule
-\textbf{Label} & \textbf{Train} & \textbf{Dev} & \textbf{Test}\\
-\midrule
-\ii{entailment} &		33.3 & 	33.3 & 	33.7 \\
-\ii{neutral} & 		33.2 & 	32.4 & 	32.2 \\
-\ii{contradiction} &  	33.3 & 	32.7 & 	32.4  \\
-- (no consensus) & 	0.1 & 	1.6 & 	1.8 \\
-\bottomrule
-  \end{tabular}	
-% From 1.0rc3
-\caption{\label{validation-freq}Frequency (\%) of each label in the corpus. Note that examples without gold labels are omitted in standard evaluations.} 
-\end{table}
-
 \paragraph{Partition} We distribute the corpus with a pre-specified train/test/development split.

### 2015-05-27 writing/SNLI paper/conclusion.tex

\subsection{Conclusion}\label{sec:conclusion}

Natural languages are powerful vehicles for complex, flexible
commonsense reasoning, and nearly all questions about meaningfulness
in language can be reduced to questions of contextual and entailment
and contradiction. This suggests that NLI is an ideal testing ground
for theories of semantic representation, and that training for NLI
tasks can provide rich domain-general semantic representations.  To
date, however, it has not been possible to fully realize this
potential due to the limited nature of existing NLI resources.  This
paper sought to remedy this with a new, large-scale, naturalistic
corpus of sentence pairs labeled for entailment, contradiction, and
independence. We used this corpus to evaluate a wide range of models,
and we found that not only do neural models perform best, but also
that their learned representations transfer well to new domains.
These results are powerful new evidence that neural representations
can be trained to accurately encode rich, multifaceted semantic
information.

### 2015-05-27 writing/SNLI paper/transfer.tex
Pre workshopping cleanup.

@@
 \todo{Cite} ....
+\noindent\todo{Pick best 2-3 tasks.}
 
@@
 
-\todo{Do we still want to re-initialize LRs?}
-\todo{Do we want to use all four target tasks?}
+\noindent\todo{Do we still want to re-initialize LRs?}
 
@@
 \hline
-\t{SICK}            & \t{98} & \t{95} & \t{63?} & \t{80.0?} & \t{84.5} \\
-\t{SUBJ}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?}& \t{??.?} \\
-\t{SST}          & \t{??.?} & \t{??.?} & \t{43?} & \t{43?} & \t{50.6}\\
-\t{PragBank}          & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?}& \t{??.?} \\
+\t{SICK}            & \t{100} & \t{99} & \t{62.9?} & \t{79.8?} & \t{84.5} \\
+\t{SUBJ}          & \t{??.?} & \t{??.?} & \t{90?} & \t{91?}& \t{95.5} \\
+\t{SST}          & \t{??.?} & \t{??.?} & \t{43?} & \t{42?} & \t{50.6}\\
+\t{PragBank}          & \t{??.?} & \t{??.?} & \t{50} & \t{56}& \t{~75} \\
 \hline
@@
 
-The final column shows the state of the art results from \cite{lai2014illinois}, TODO, \cite{tai2015improved}, TODO, respectively.
-\todo{Does Tai have the SotA on SST?}
+The final column shows the state of the art results from \cite{lai2014illinois}, \cite{zhao2015self}, \cite{tai2015improved}, TODO, respectively.
+\noindent\todo{Does Tai have the SotA on SST?}
 
@@
 
-\todo{something about the state of the art on the other tasks.}
+\noindent\todo{something about SotA on the other tasks.}

### 2015-05-27 writing/SNLI paper/corpus.tex
Pre workshopping cleanup.

@@
 distributed across the three categories in our train, development, and
-testing. The Fleiss $\kappa$ scores in
-Table~\ref{kappas} are likely to be conservative given our large and
+testing. The Fleiss $\kappa$ scores 
+(computed over every example with a full five annotations)
+are likely to be conservative given our large and
 unevenly distributed pool of workers, but they still provide insights
@@
 
-\todo{Update as appropriate (CM)}
+\noindent\todo{Update as appropriate (CM)}
 

### 2015-05-27 writing/SNLI paper/entailment_eval.tex
Update (still prelim) results
Update bibtex
Pre workshopping cleanup.

@@
 
-% TODO(Sam): Can you add these to the Bibtex?
-%
-%@article{miller1995wordnet,
-%  title={{W}ord{N}et: a lexical database for English},
-%  author={Miller, George A},
-%  journal={Communications of the ACM},
-%  volume={38},
-%  number={11},
-%  pages={39--41},
-%  year={1995},
-%  publisher={ACM}
-%}
-%
-%@inproceedings{chklovski2004verbocean,
-%  title={VerbOcean: Mining the Web for Fine-Grained Semantic Verb Relations.},
-%  author={Chklovski, Timothy and Pantel, Patrick},
-%  booktitle={EMNLP},
-%  year={2004}
-%}
-%
-%@inproceedings{wang2007recognizing,
-%  title={Recognizing textual entailment using sentence similarity based on dependency tree skeletons},
-%  author={Wang, Rui and Neumann, G{\"u}nter},
-%  booktitle={ACL-PASCAL Workshop on Textual Entailment and Paraphrasing},
-%  year={2007},
-%}
-
 %
@@
   the RTE systems have been tuned on relatively small corpora
-  (~1600 examples), whereas a classifier trained on SNLI can make use of
+  ($\sim$1600 examples), whereas a classifier trained on SNLI can make use of
   over two orders of magnitude more data.
@@
 \t{100d Sum of words}            & \t{79?} & \t{76?} \\
-\t{100d RNN}            & \t{?} & \t{68} \\	
+\t{100d RNN}            & \t{67?} & \t{69?} \\	
 \t{100d LSTM RNN}            & \t{75?} & \t{76?} \\

### 2015-05-27 writing/SNLI paper/abstract.tex
Pre workshopping cleanup.
Comments from KST and Ignacio.

@@
   %hampered by the limited nature of existing annotated resources.
-  However, progess is currently hampered by the limitations of
+  However, progress is currently hampered by the limitations of
   existing annotated resources.
@@
   And, in contrast to many such resources, all of the sentences were
-  written by humans in a grounded, naturalistic context, and 54K of
-  the labels have been independently validated.
+  written by humans in a grounded, naturalistic context.
   %
@@
   argument for the centrality of entailment and contradiction in
-  semantics, we show that the TreeLSTM's representatations, trained on
+  semantics, we show that the TreeLSTM's representations, trained on
   our corpus, perform well in a disparate set of additional semantic

### 2015-05-27 writing/SNLI paper/corpus.tex
Comments from KST and Ignacio.

@@
 coreference are resolved.
+
+\noindent\todo{Introduce name of corpus.}
 
@@
 With the exception of a handful that were flagged by workers as excessively difficult, our corpus contains at least three labeled pairs for each caption.
+
+\noindent\todo{Explain our 5x->1x annotation switch...}
 
@@
 
+\noindent\todo{More on *why* `-' isn't a valid test label}
+
 \subsection{The distributed corpus}
@@
 \midrule
-\multicolumn{2}{l}{\textbf{Frequency of complete sentences:}}\\
+\multicolumn{2}{l}{\textbf{Parser output:}}\\
 Premise `S'-rooted parses & 74.0\%\\
 Hypothesis `S'-rooted parses & 88.9\%\\
+Unique word tokens (ignoring case) & ...\\
     \bottomrule
@@
 % From 1.0rc3
+\todo{Unique word count.}
 \caption{\label{collection-stats}Key statistics for the raw sentence pairs. Since the two halves of each pair were collected separately, we report some statistics for both.} 
@@
 
-\noindent\todo{Update as appropriate (CM)}
+\noindent\todo{Use version number for 4/20/15 model}
 

### 2015-05-27 writing/SNLI paper/transfer.tex
Bit more cleanup.
Comments from KST and Ignacio.

@@
 To test this, we perform transfer learning experiments on four other tasks, in which models are initialized using the parameters of an LSTM RNN-based model trained on  on SNLI (as in \S\ref{sentence-embedding}) and then trained to convergence on the target tasks.
+
+\noindent\todo{Mention domain limitation to scenes}
 
@@
 
-The final column shows the state of the art results from \cite{lai2014illinois}, \cite{zhao2015self}, \cite{tai2015improved}, TODO, respectively.
+The final column shows the state of the art results from \citealt{lai2014illinois}, \citealt{zhao2015self}, \citealt{tai2015improved}, and \todo{CITE}.
 \noindent\todo{Does Tai have the SotA on SST?}
@@
 Note that \todo{something interesting happens}.
-Within the domain of NLI, our transfer model shows the best number yet reported on SICK for a fully-learned model, and approaches the overall state of the art (and to the 84\% level of interannotator agreement, a likely performance cieling), while the randomly initialized model performs poorly.
+Within the domain of NLI, our transfer model shows the best number yet reported on SICK for a fully-learned model, and approaches the overall state of the art (and to the 84\% level of interannotator agreement, a likely performance ceiling), while the randomly initialized model performs poorly.
 This suggests that the introduction of SNLI makes it newly possible to train representation learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on semantically sophisticated tasks.

### 2015-05-27 writing/SNLI paper/entailment_eval.tex
Bit more cleanup.
Comments from KST and Ignacio.

@@
   RTE-3 dataset.
-\todo{numbers are still dev.}
+\todo{Final numbers}
 }
@@
 We report results in \reftab{eopresults}.
-The best edit distance algorithm tunes the the weight of the three 
+The best edit distance algorithm tunes the weight of the three 
   case-insensitive edit distance operations on the training set, 
@@
   rich lexicalized features.
-We implmement such a lexicalized classifier as a strong baseline 
+We implement such a lexicalized classifier as a strong baseline 
   for the task.
@@
   features (Features \ref{lst:ngram}--\ref{lst:bigram}).
-\todo{fill in dev numbers}
+\todo{Final numbers}
 }
@@
 
-The word embeddings for all of the moedels are initialized the 300d reference GloVe vectors \cite{pennington2014glove} and fine tuned.
-In addition, all of the models use an additional $\tanh$ neural network layer to map these 300d embeddings into the lower-dimensianal phrase and sentence embedding space.
+The word embeddings for all of the moedels are initialized the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned.
+In addition, all of the models use an additional $\tanh$ neural network layer to map these 300d embeddings into the lower-dimensional phrase and sentence embedding space.
 All of the models are randomly initialized using standard techniques and trained using AdaDelta \cite{zeiler2012adadelta} minibatch SGD until performance on the development set stops improving.
@@
 Accuracy in 3-class classification on the SNLI training and test sets for each model.
-\todo{Replace with final numbers}
+\todo{Replace with final numbers/dims}
 }

### 2015-05-27 writing/SNLI paper/entailment_eval.tex
Misc updates.

@@
 
-The word embeddings for all of the moedels are initialized the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned.
+The word embeddings for all of the models are initialized the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned.
 In addition, all of the models use an additional $\tanh$ neural network layer to map these 300d embeddings into the lower-dimensional phrase and sentence embedding space.

### 2015-05-27 config/SynsetRelations.m

@@ -11,7 +11,7 @@ hyperParams.dim = 25;
 % added above the embedding matrix. This is likely to only be useful when
 % learnWords is false, and so the embeddings do not exist in the same space
 % the rest of the constituents do.
-hyperParams.embeddingTransformDepth = transDepth;
+hyperParams.useEmbeddingTransform = transDepth;
 
 % The dimensionality of the comparison layer(s).
 hyperParams.penultDim = penult;

### 2015-05-27 writing/SNLI paper/corpus.tex
Misc updates.

@@
 
-For the premises, we used the Flickr30k corpus \cite{hodoshimage}, a database of approximately 160k captions  (corresponding to about 30k images) collected in an earlier crowdsourced effort.
+For the premises, we used captions from the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions (corresponding to about 30k images) collected in an earlier crowdsourced effort.
 The captions were not authored by the photographers who took the source images, and they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \word{Our trip to the Olympic Peninsula}).
-We presented each caption to at least one worker, and some to more than one.
-With the exception of a handful that were flagged by workers as excessively difficult, our corpus contains at least three labeled pairs for each caption.
-
-\noindent\todo{Explain our 5x->1x annotation switch...}
 

### 2015-05-27 RunExperiments/RunSSTExperiments.sh

@@ -101,6 +101,7 @@ export MATLABCMD="cd quant; lambda = 3e-6; dim = 50; ed = 200; td = 3; penult = 
 export MATLABCMD="cd quant; lambda = 0.0003; dim = 50; ed = 200; td = 3; penult = 100; dropout = [1\, 1]; tot = 2; collo = 1; dataflag='sst-expanded-transfer'; name='/scr/sbowman/sst-expanded-7t'; TrainModel('/scr/sbowman/snlirc2-snlirc2-only-l1e-05-dim50-ed200-td3-pen100-do1-1-co1-comp2-dp1-gc0/ckpt-best-tr150522185602@435250.mat'\, 1\, @SST\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 4\, 0\, 0\, .3);" ; qsub -v MATLABCMD quant/run.sh 
 export MATLABCMD="cd quant; lambda = 0.00003; dim = 50; ed = 200; td = 3; penult = 100; dropout = [1\, 1]; tot = 2; collo = 1; dataflag='sst-expanded-transfer'; name='/scr/sbowman/sst-expanded-7t'; TrainModel('/scr/sbowman/snlirc2-snlirc2-only-l1e-05-dim50-ed200-td3-pen100-do1-1-co1-comp2-dp1-gc0/ckpt-best-tr150522185602@435250.mat'\, 1\, @SST\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 4\, 0\, 0\, .3);" ; qsub -v MATLABCMD quant/run.sh 
 
+export MATLABCMD="cd quant; lambda = 0.0003; dim = 50; ed = 200; td = 3; penult = 100; dropout = [1\, 1]; tot = 2; collo = 1; dataflag='sst-expanded-transfer'; name='/scr/sbowman/sst-expanded-7tswap'; TrainModel('/scr/sbowman/snlirc2-snlirc2-only-l1e-05-dim50-ed200-td3-pen100-do1-1-co1-comp2-dp1-gc0/ckpt-best-tr150522185602@435250.mat'\, 1\, @SST\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 4\, 0\, 0\, .3);" ; qsub -v MATLABCMD quant/run.sh 
 
[…]

### 2015-05-27 config/PDTB.m

@@ -1,4 +1,5 @@
-function [ hyperParams, options, wordMap, labelMap ] = TREC(expName, dataflag, embDim, dim, topDepth, penult, lambda, composition, bottomDropout, topDropout, collo, adi)
+function [ hyperParams, options, wordMap, labelMap ] = SNLI(expName, dataflag, embDim, dim, topDepth, penult, lambda, composition, bottomDropout, topDropout, collo, dp, gc, adi)
+% Configuration for experiments involving the SemEval SICK challenge and ImageFlickr 30k. 
 
 [hyperParams, options] = Defaults();
 
@@ -6,14 +7,16 @@ function [ hyperParams, options, wordMap, labelMap ] = TREC(expName, dataflag, e
 % are being tuned.
 hyperParams.name = [expName, '-', dataflag, '-l', num2str(lambda), '-dim', num2str(dim),...
     '-ed', num2str(embDim), '-td', num2str(topDepth),...
-    '-do', num2str(bottomDropout), '-', num2str(topDropout), '-co', num2str(collo),...
-    '-comp', num2str(composition), '-adi', num2str(adi)  ];
+    '-pen', num2str(penult), '-do', num2str(bottomDropout), '-', num2str(topDropout), '-co', num2str(collo),...
+    '-comp', num2str(composition), ...
+    '-dp', num2str(dp), '-gc', num2str(gc),  '-adi', num2str(adi)];
 
 hyperParams.restartUpdateRuleInTransfer = adi;
 
-hyperParams.sentenceClassificationMode = 1;
 
-hyperParams.testFraction = 0.1;
+hyperParams.parensInSequences = 0;
+
+hyperParams.dataPortion = dp;
 
 hyperParams.dim = dim;
 hyperParams.embeddingDim = embDim;
@@ -41,14 +44,16 @@ hyperParams.topDepth = topDepth;
 % to the parameters that are in use at each step.
 hyperParams.largeVocabMode = true;
 
-% The dimensionality of the classifier extra layers.
+% The dimensionality of the comparison layer(s).
 hyperParams.penultDim = penult;
 
 % Regularization coefficient.
 hyperParams.lambda = lambda; % 0.002 works?;
 
-% How many examples to run before taking a parameter update step on the accumulated gradients.
-options.miniBatchSize = 32;
+if gc > 0
[…]

### 2015-05-28 writing/SNLI paper/abstract.tex
Implementing CM's suggestions.

@@
   Entailment and contradiction are fundamental to natural language
-  meaning and so have emerged (in the form of tasks involving
-  reasoning, inference, and summarization) as an important testing
+  meaning and so are an important testing
   ground for semantic representations.
@@
   However, progress is currently hampered by the limitations of
-  existing annotated resources.
+  existing annotated resources for training.
   %
@@
   tasks.  
-  \todo{Structure more around results (vs. data).}
 \end{abstract}

### 2015-05-28 writing/SNLI paper/corpus.tex
Implementing CM's suggestions.

@@
 have stimulated innovative logical and statistical models of natural
-language reasoning, but their small size (a few hundred examples each)
+language reasoning, but their small size (fewer than a thousand examples each)
 limits their utility as a testbed for learned distributed representations. 
@@
 coreference are resolved.
-
-\noindent\todo{Introduce name of corpus.}
+\todo{Cite Marneffe `08 \S4.3 for coref issue.}
 
@@
 that label is the gold label. If there was no such consensus, which
-occurred in about 0.2\% of cases, we chose the label `-'. While these
+occurred in about 0.2\% of cases, we chose the label `-'. While rare these
 unlabeled examples are included in the corpus distribution, they are
@@
 
-\noindent\todo{More on *why* `-' isn't a valid test label}
-
 \subsection{The distributed corpus}
@@
   \end{tabular}
-  \caption{\label{snli-examples}Randomly chosen examples from the SNLI development set, shown with the gold labels and the full set of annotator labels (abbreviated).
-Unlike in many existing NLI corpora, the premise and hypothesis sentences in SNLI are often too dissimilar in structure to support approaches to NLI that rely on building sentence-to-sentence alignments.}
+  \caption{\label{snli-examples}Randomly chosen examples from our development set, shown with the gold labels and the full set of annotator labels (abbreviated).
+Unlike in many existing NLI corpora, the premise and hypothesis sentences in our corpus are often too dissimilar in structure to support approaches to NLI that rely on building sentence-to-sentence alignments.}
 \end{table*}
@@
 Qualitatively, we find the data that we collected to draw fairly extensively on commonsense knowledge, suggesting that superficial word alignment models are unlikely to be able to succeed at classification.
-We find the sentences that we collected be largely fluent, correctly spelled English, with a mix of full sentences and caption-style noun phrase fragments.
+We find the sentences that we collected to be largely fluent, correctly spelled English, with a mix of full sentences and caption-style noun phrase fragments.
 Punctuation and capitalization are often omitted.
@@
 Hypothesis `S'-rooted parses & 88.9\%\\
-Unique word tokens (ignoring case) & ...\\
+Distinct words (ignoring case) & 37,026\\
     \bottomrule
@@
 
-The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser \cite{klein2003accurate} (trained on the standard training set as well as on the Brown Corpus; \citealt{francis1979brown}), both in the standard PTB format and in the binarized unlabeled format used by tree-structured neural network models, as in \S\ref{sentence-embedding}.
+The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser \cite{klein2003accurate}, trained on the standard training set as well as on the Brown Corpus \cite{francis1979brown}, which we found to improve the parse quality of the descriptive sentences and noun phrases found in the descriptions).
+Parses are available in both the standard PTB format and in the binarized unlabeled format used by tree-structured neural network models, as in \S\ref{sentence-embedding}.
 

### 2015-05-28 writing/SNLI paper/entailment_eval.tex
Implementing CM's suggestions.

@@
-\section{SNLI as a platform for NLI research}
+\section{Our data as a platform for NLI research}
 
-The most immediate application for the SNLI corpus is in developing models for the task of NLI.
-In particular, since SNLI is dramatically larger than any existing corpus of comparable quality, we expect it to make possible the use of parameter-rich models like neural networks for this task.
+The most immediate application for our corpus is in developing models for the task of NLI.
+In particular, since it is dramatically larger than any existing corpus of comparable quality, we expect it to be suitable for training  parameter-rich models like neural networks for this task.
 In this section, we explore the performance of a range of standard and novel models trained on the corpus.
@@
 \subsection{Standard entailment models}
-We evaluate a number of strong baselines on the SNLI corpus.
+We evaluate a number of strong baselines on our corpus.
 The first class of baselines use models from the Excitement Open
@@
 We run the top performing edit-distance based algorithm and the top
-  performing classifier-based algorithm on the SNLI test set, as
+  performing classifier-based algorithm on our test set, as
   determined by performance on the development set.
@@
 \hline
-\textbf{System} & \multicolumn{4}{c}{\b{SNLI}} & \multicolumn{2}{c}{\b{RTE-3}} \\
+\textbf{System} & \multicolumn{4}{c}{\b{Our corpus}} & \multicolumn{2}{c}{\b{RTE-3}} \\
  & \t{P} & \t{R} & \t{F$_1$} & \t{Acc.} & \t{F$_1$} & \t{Acc.} \\
@@
 \paragraph{Lexicalized Classifier}
-Unlike the RTE datasets, SNLI's size allows approaches which make use of
+Unlike the RTE datasets, our corpus's size allows approaches which make use of
   rich lexicalized features.
@@
 \label{tab:bowresults}
-Accuracy on the SNLI dataset, including precision/recall for the
+Accuracy on the our dataset, including precision/recall for the
   \textit{entailed} class on the test set.
@@
   the RTE systems have been tuned on relatively small corpora
-  ($\sim$1600 examples), whereas a classifier trained on SNLI can make use of
+  ($\sim$1600 examples), whereas a classifier trained on our corpus can make use of
   over two orders of magnitude more data.
+  \todo{Use curve to justify large size of corpus.}
 
@@
 
-SNLI is uniquely suitable for training neural network models that produce vector representations of sentence meaning. In this section, we compare the performance of six such models on SNLI. 
-In order to best evaluate the strengths of these six models at producing informative representations, we use sentence embedding as an intermediate step in the NLI classification task: each model most produce vector representation of each of the two sentences without using any context from the other sentence, and the two resulting vectors are then passed to a neural network classifier which predicts the label for the pair.
+Our corpus is uniquely suitable for training neural network models that produce vector representations of sentence meaning.
+In this section, we compare the performance of six such models on the corpus.
+In order to best evaluate the strengths of these six models at producing informative representations, we use sentence embedding as an intermediate step in the NLI classification task: each model must produce a vector representation of each of the two sentences without using any context from the other sentence, and the two resulting vectors are then passed to a neural network classifier which predicts the label for the pair.
 This choice allows us to focus on evaluating the ability of existing models to produce representations that fully capture sentence meaning, at the possible cost of excluding from consideration neural models for NLI that directly compare the two inputs at the word or phrase level.
@@
 
-The word embeddings for all of the models are initialized the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned.
+The word embeddings for all of the models are initialized with the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned.
 In addition, all of the models use an additional $\tanh$ neural network layer to map these 300d embeddings into the lower-dimensional phrase and sentence embedding space.
@@
 \label{tab:nnresults}
-Accuracy in 3-class classification on the SNLI training and test sets for each model.
+Accuracy in 3-class classification on the our training and test sets for each model.
 \todo{Replace with final numbers/dims}

### 2015-05-28 writing/SNLI paper/transfer.tex
Implementing CM's suggestions.

@@
-\section{SNLI and transfer learning}\label{sec:transfer}
+\section{NLI data and transfer learning}\label{sec:transfer}
 
 We argue that for a sentence representation to provide the information necessary to reliably perform NLI classification, it must contain a faithful and thorough description of the sentence's meaning.
-This claim has a readily testable consequence: the knowledge that a representation learning system learns when it is trained on SNLI should transfer well to any other task that involves understanding sentence meaning.
-To test this, we perform transfer learning experiments on four other tasks, in which models are initialized using the parameters of an LSTM RNN-based model trained on  on SNLI (as in \S\ref{sentence-embedding}) and then trained to convergence on the target tasks.
+This claim has a readily testable consequence: the knowledge that a representation learning system learns when it is trained on our corpus should transfer well to any other task that involves understanding sentence meaning, at least to the extent that the style of language used in that task is broadly similar to the style of our corpus.
+To test this, we perform transfer learning experiments on four other tasks, in which models are initialized using the parameters of an LSTM RNN-based model trained on our corpus (as in \S\ref{sentence-embedding}) and then trained to convergence on the target tasks.
 
-\noindent\todo{Mention domain limitation to scenes}
-
-We first evaluate on SICK (\citealt{marelli2014sick}), the smaller entailment corpus SNLI was modeled on, training on the standard 4.5k training sentence pairs.
+We first evaluate on SICK (\citealt{marelli2014sick}), the smaller entailment corpus ours was modeled on, training on the standard 4.5k training sentence pairs.
 We then evaluate on SUBJ (\citealt{pang2004sentimental}), a two-way sentence classification task, using 5-fold cross-validation over the 10k sentences.
-We next evaluate on the SST (\citealt{socher2013acl1}), a five-way sentiment classification task with labels for both phrases and sentences.
+We next evaluate on the Stanford Sentiment Treebank (SST; \citealt{socher2013acl1}), a five-way sentiment classification task with labels for both phrases and sentences.
 We train on both types of label within the 8.5k sentence training set and test on only the sentence-level labels in the test set.
@@
 \caption{\label{tab:transferresults}
-The performance of an LSTM RNN model on four tasks under both a random initialization and a transfer initialization copied from a model trained on SNLI.
+The performance of an LSTM RNN model on four tasks under both a random initialization and a transfer initialization copied from a model trained on our corpus.
 All figures are classification accuracy, except for PragBank, where report macroaveraged F1 to enable comparison with existing results on that task.
@@
 The final column shows the state of the art results from \citealt{lai2014illinois}, \citealt{zhao2015self}, \citealt{tai2015improved}, and \todo{CITE}.
-\noindent\todo{Does Tai have the SotA on SST?}
 
@@
 Within the domain of NLI, our transfer model shows the best number yet reported on SICK for a fully-learned model, and approaches the overall state of the art (and to the 84\% level of interannotator agreement, a likely performance ceiling), while the randomly initialized model performs poorly.
-This suggests that the introduction of SNLI makes it newly possible to train representation learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on semantically sophisticated tasks.
+This suggests that the introduction of a large high-quality corpus makes it newly possible to train representation learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on semantically sophisticated tasks.
 

### 2015-05-28 utils/make_wordlist.py

@@ -11,13 +11,14 @@ from collections import defaultdict
 
 # TODO: Ensure that there are no duplicates.
 
-BASENAME = "pdtb-rc3"
-TRAINING_FILES = ["../data/pdtb2-train.tsv"]
-DEV_FILES = ["../data/pdtb2-train.tsv"]
-TEST_FILES = ["../data/pdtb2-train.tsv"]
+BASENAME = "temp"
+TRAINING_FILES = ["../Distributions/snli_1.0rc3/snli_1.0rc3_dev.txt",
+                  "../Distributions/snli_1.0rc3/snli_1.0rc3_test.txt", "../Distributions/snli_1.0rc3/snli_1.0rc3_train.txt"]
+DEV_FILES = []
+TEST_FILES = []
 
-TRANSFER_SOURCE_WORDLIST = "../data/snlirc3_words.txt"
-# TRANSFER_SOURCE_WORDLIST = ""
+# TRANSFER_SOURCE_WORDLIST = "../data/snlirc3_words.txt"
+TRANSFER_SOURCE_WORDLIST = ""
 VECTOR_WORDLIST = "utils/glove.840B.wordlist.txt"
 
 EXCLUSIONS = set(['(', '(0', '(1', '(2', '(3', '(4', ')', '', ' ', '\n', '\r'])
@@ -73,10 +74,12 @@ else:
     transfer_source_words = set()
 
 training_words = count_words(TRAINING_FILES)
+print str(len(training_words)) + " words in training."
 # for word in training_words:
 #    print str(training_words[word]) + '\t' + word
 
 dev_test_words = count_words(TEST_FILES + DEV_FILES)
+print str(len(dev_test_words)) + " words in dev/test."
 
 wordlist = create_wordlist(
     training_words, dev_test_words, vector_words.union(transfer_source_words))

### 2015-05-28 Tree.m
Optimizations.

@@ -162,12 +162,14 @@ classdef Tree < handle
                                 compositionMatrix, embeddingTransformMatrix, compNL, trainingMode, hyperParams)
             % Recomputes features using fresh parameters.
 
-            if (~isempty(obj.daughters))
-                for daughterIndex = 1:length(obj.daughters)
-                    obj.daughters(daughterIndex).updateFeatures(...
-                        wordFeatures, compositionMatrices, compositionMatrix, embeddingTransformMatrix, ...
-                        compNL, trainingMode, hyperParams);
-                end
+            if ~isempty(obj.daughters)
+                obj.daughters(1).updateFeatures(...
+                    wordFeatures, compositionMatrices, compositionMatrix, embeddingTransformMatrix, ...
+                    compNL, trainingMode, hyperParams);
+
+                obj.daughters(2).updateFeatures(...
+                    wordFeatures, compositionMatrices, compositionMatrix, embeddingTransformMatrix, ...
+                    compNL, trainingMode, hyperParams);
                 
                 lFeatures = obj.daughters(1).features;
                 rFeatures = obj.daughters(2).features;
@@ -222,7 +224,6 @@ classdef Tree < handle
                         compNLDeriv, hyperParams)
             
             DIM = length(delta);
-            EMBDIM = size(embeddingTransformMatrix, 2) - 1;
             NUMTRANS = size(embeddingTransformMatrix, 3) .* (length(embeddingTransformMatrix) > 0);
 
             if size(compositionMatrix, 3) == 1 % Using tied composition parameters
@@ -236,7 +237,7 @@ classdef Tree < handle
             upwardWordGradients = sparse([], [], [], ...
                 size(wordFeatures, 1), size(wordFeatures, 2), 10);            
             
-            if size(compositionMatrices, 1) == 0
+            if isempty(compositionMatrices)
                 upwardCompositionMatricesGradients = [];
             else
                 upwardCompositionMatricesGradients = zeros(size(compositionMatrices));
@@ -303,7 +304,7 @@ classdef Tree < handle
                   incomingCompositionMatricesGradients, ...
                   incomingCompositionMatrixGradients, ...
                   incomingEmbeddingTransformMatrixGradients ] = ...
-                  obj.getLeftDaughter.getGradient( ...
+                  obj.daughters(1).getGradient( ...
                                 compDeltaLeft, [], wordFeatures, ...
                                 compositionMatrices, compositionMatrix, embeddingTransformMatrix, ...
[…]

### 2015-05-28 config/SST.m

@@ -180,10 +180,4 @@ elseif strcmp(dataflag, 'sst')
     hyperParams.testFilenames = {'./sst-data/dev.txt'};
 end
 
-% Temp testing method.
-if ccs == -100
-    hyperParams.lineLimit = 500;
-    hyperParams.loadWords = 0;
-end
-
 end

### 2015-05-28 writing/SNLI paper/corpus.tex
Paper updates; config updates.

@@
 event and entity coreference lead to insurmountable indeterminacy
-concerning the correct semantic label \cite{marelli2014sick}. For
+concerning the correct semantic label (\citealt{de2008finding} $\S4.3$, \citealt{marelli2014sick}). For
 instance, the sentence pair \word{A boat sank in the Pacific
@@
 coreference are resolved.
-\todo{Cite Marneffe `08 \S4.3 for coref issue.}
 
@@
 
-The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser \cite{klein2003accurate}, trained on the standard training set as well as on the Brown Corpus \cite{francis1979brown}, which we found to improve the parse quality of the descriptive sentences and noun phrases found in the descriptions).
+The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser 3.5.2 \cite{klein2003accurate}, trained on the standard training set as well as on the Brown Corpus \cite{francis1979brown}, which we found to improve the parse quality of the descriptive sentences and noun phrases found in the descriptions).
 Parses are available in both the standard PTB format and in the binarized unlabeled format used by tree-structured neural network models, as in \S\ref{sentence-embedding}.
-
-\noindent\todo{Use version number for 4/20/15 model}
 

### 2015-05-28 RunExperiments/RunRTE3Experiments.sh

@@ -1,0 +1,2 @@
+export MATLABCMD="cd quant; lambda = 0.00003; dim = 100; ed = 300; td = 3; penult = 200; dropout = [0.95\, 0.95]; tot = 2; collo = 3; dataflag='rte3'; name='/scr/sbowman/rte3-t2'; TrainModel(''\, 1\, @RTE3\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1\, 5\, 5);" ; qsub -v MATLABCMD quant/run.sh
+export MATLABCMD="cd quant; lambda = 0.00003; dim = 100; ed = 300; td = 3; penult = 200; dropout = [0.95\, 0.95]; tot = 2; collo = 3; dataflag='rte3-transfer'; name='/scr/sbowman/rte3-t2'; TrainModel('snlirc3d-snlirc3-only-l3e-05-dim100-ed300-td3-pen200-do0.95-0.95-co3-comp2-dp1-gc5-lstminit5/ckpt-best-tr150527095225@147125.mat'\, 1\, @RTE3\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1\, 5\, 5);" ; qsub -v MATLABCMD quant/run.sh

### 2015-05-28 config/SNLI.m

@@ -62,10 +62,6 @@ hyperParams.topDropout = topDropout;
 hyperParams = CompositionSetup(hyperParams, composition);
 hyperParams.useThirdOrderMerge = false;
 
-if (composition == 7) || (composition < 2)
-    options.miniBatchSize = 16;
-end
-
 hyperParams.loadWords = true;
 hyperParams.trainWords = true;
 

### 2015-05-29 config/SUBJ.m
Bib updates, params.

@@ -79,6 +79,6 @@ hyperParams.testFilenames = {};
 
 if strcmp(dataflag, 'subj')
 elseif strcmp(dataflag, 'subj-transfer')
-	hyperParams.sourceWordMap = LoadWordMap('../data/snlirc2_words.txt');
+	hyperParams.sourceWordMap = LoadWordMap('../data/snlirc3_words.txt');
 end
 end

### 2015-05-29 writing/SNLI paper/intro.tex
Pre workshopping cleanup.
Comments from KST and Ignacio.
Implementing CM's suggestions.
Bib updates, params.

@@
 parsing to commonsense reasoning.
+\todo{Non-MacCartney NLI citation.}
 
@@
 modeling simple dimensions of meaning like evaluative sentiment
-(e.g., \citealt{socher2013acl1}), but it is less clear that they can be
+(e.g., \citealt{socher2013recursive}), but it is less clear that they can be
 trained to support the full range of logical and commonsense
@@
 entailment, contradiction, and semantic independence. At 570,152
-sentence-pairs, this corpus is orders of magnitude larger than all
+sentence-pairs, this corpus is two orders of magnitude larger than all
 other resources of its type. And, in contrast to many such resources,
@@
 
-\todo{Plain LSTMs strong, used in transfer.}
+\todo{TreeLSTM vs. LSTM result TBD.}
 

### 2015-05-29 TransferInitialization.m

@@ -13,7 +13,7 @@ function freshModelState = TransferInitialization(freshModelState, pretrainedMod
 
 % Unpack pretrained model state.
 [ mergeMatrices, mergeMatrix, ...
-    ~, pretrainedWordFeatures, connectionMatrix, ...
+    pretrainedSoftmaxMatrix, pretrainedWordFeatures, connectionMatrix, ...
     compositionMatrix, scoringVector, classifierExtraMatrix, embeddingTransformMatrix] ...
     = stack2param(pretrainedModelState.theta, pretrainedModelState.thetaDecoder);
 [ GmergeMatrices, GmergeMatrix, ...
@@ -24,6 +24,14 @@ function freshModelState = TransferInitialization(freshModelState, pretrainedMod
     ~, DpretrainedWordFeatures, DconnectionMatrix, ...
     DcompositionMatrix, DscoringVector, DclassifierExtraMatrix, DembeddingTransformMatrix] ...
     = stack2param(pretrainedModelState.sumSqDelta, pretrainedModelState.thetaDecoder);
+
+if hyperParams.transferSoftmax && (size(softmaxMatrix, 1) == 3)
+	Log(hyperParams.statlog, ['Copying entailment softmax layer.']);
+	softmaxMatrix = pretrainedSoftmaxMatrix;
+elseif hyperParams.transferSoftmax && (size(softmaxMatrix, 1) == 2)
+	Log(hyperParams.statlog, ['Copying entailment softmax layer from three-way entailment to two-way.']);
+	softmaxMatrix = pretrainedSoftmaxMatrix([1 3], :);
+end
 
 if ~isempty(wordFeatures)
 	wordFeaturesInStack = true;

### 2015-05-29 RunExperiments/RunSUBJExperiments.sh
Numbers; configs.

@@ -63,7 +63,9 @@ export MATLABCMD="cd quant; lambda = 0.00001; dim = 50; ed = 200; td = 3; penult
 % 0001 optimal for notrans
 % 001-00001 optimal for transfer ??
 
-export MATLABCMD="cd quant; lambda = 0.0001; dim = 50; ed = 200; td = 3; penult = 100; dropout = [1\, 1]; tot = 2; collo = 1; dataflag='subj-transfer'; name='/scr/sbowman/subj6swap'; TrainModel('/scr/sbowman/snlirc2-snlirc2-only-l1e-05-dim50-ed200-td3-pen100-do1-1-co1-comp2-dp1-gc0/ckpt-best-tr150522185602@435250.mat'\, 1\, @SUBJ\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1);" ; qsub -v MATLABCMD quant/run.sh 
-export MATLABCMD="cd quant; lambda = 0.0001; dim = 50; ed = 200; td = 3; penult = 100; dropout = [1\, 1]; tot = 2; collo = 1; dataflag='subj'; name='/scr/sbowman/subj6swap'; TrainModel(''\, 1\, @SUBJ\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 0);" ; qsub -v MATLABCMD quant/run.sh
+export MATLABCMD="cd quant; lambda = 0.001; dim = 100; ed = 300; td = 3; penult = 200; dropout = [0.95\, 0.95]; tot = 2; collo = 3; dataflag='subj-transfer'; name='/scr/sbowman/subj7'; TrainModel('snlirc3d-snlirc3-only-l3e-05-dim100-ed300-td3-pen200-do0.95-0.95-co3-comp2-dp1-gc5-lstminit5/ckpt-best-tr150527095225@147125.mat'\, 1\, @SUBJ\, name\, dataflag\, ed\, dim\, td\, penult\, lambda\, tot\, dropout(1)\, dropout(2)\, collo\, 1);" ; qsub -v MATLABCMD quant/run.sh 
[…]

### 2015-05-29 writing/SNLI paper/entailment_eval.tex
Optimizations.
Numbers; configs.
Removing transfer section. :P

@@
 % what is EOP
-The Excitement Open Platform is a tool to quickly develop RTE systems
+The Excitement Open Platform is a tool to quickly develop NLI systems
   while sharing components such as common lexical resources and 
@@
 The results should therefore be taken as a strong baseline for
-  RTE-style approaches to the problem, rather than necessarily
+  NLI-style approaches to the problem, rather than necessarily
   representing the state-of-the-art system's performance on the
@@
   RTE-3 dataset.
+\todo{Cite RTE-3 here.}
 \todo{Final numbers}
@@
 \t{100d Sum of words}            & \t{79?} & \t{76?} \\
-\t{100d RNN}            & \t{67?} & \t{69?} \\	
-\t{100d LSTM RNN}            & \t{75?} & \t{76?} \\
+\t{100d RNN}            & \t{75?} & \t{72?} \\	
+\t{100d LSTM RNN}            & \t{81.4?} & \t{78.2?} \\
 \t{100d TreeRNN}            & \t{69?} & \t{69?} \\
-\t{50d TreeRNTN}            & \t{58?} & \t{56?} \\
-\t{100d LSTM TreeRNN}            & \t{75?} & \t{73?} \\
+\t{50d TreeRNTN}            & \t{61?} & \t{60?} \\
+\t{100d LSTM TreeRNN}            & \t{72?} & \t{73?} \\
 \hline
@@
 
-Note that \todo{something interesting happens}.
+\subsubsection{Evaluating on other inference corpora}
+
+To the extent that successfully training a sentence model on our corpus forces that model to encode broadly accurate representations of English scene descriptions, we should expect it to be possible to adapt these representations in a different setting.
+In this section, we experiment with transfer learning on SICK, the corpus for the most recent NLI shared task.
+
+To perform transfer learning, we take the parameters of strong LSTM RNN model trained on our corpus and use them to initialize new model which trained from that point only on the training portion of SICK.
+The only newly initialized parameters are the embeddings for words that appear in the new training set, but not in our training set, and those parameters are initialized using GloVe embeddings, as before.
+We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength which is re-tuned.
+We additionally transfer the accumulators that are used by AdaDelta to set the learning rates with the intention of ensuring that the model does not learn too quickly in its first few epochs and destroy the knowledge accumulated in the pre-transfer phase of training.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
+\hline
+\textbf{Sentence model} & \b{Train}  & \b{Test}\\
+\hline
+\t{100d LSTM RNN (random init.)}            & \t{100} & \t{71.3} \\
+% From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
+\t{100d LSTM RNN (transfer init.)}            & \t{100?} & \t{80?} \\
+% From: 
+\hline
+\end{tabular}
+\end{center}
+
+\caption{\label{tab:transferresults}
+The performance (\% accuracy) of an LSTM RNN model on SICK under both a random initialization and a transfer initialization copied from a model trained on our corpus.  \todo{Final numbers.}} 
+\end{table}
+
+\todo{Why not RTE3?}
+
+We compare the performance of an LSTM model with both the transfer initialization and a normal random initialization.
+We tune the regularization parameter separately for each case.
+Table \ref{tab:transferresults} shows these results.
+
+Our transfer model shows the best performance yet reported on SICK for an unaugmented neural network model, and approaches the overall state of the art at 84.6\% \cite{lai2014illinois}, as well as the 84\% level of interannotator agreement, which represents a likely performance ceiling.
+The randomly initialized model performs relatively poorly.
+This suggests that the introduction of a large high-quality corpus makes it newly possible to train representation learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on inference tasks.

### 2015-05-29 writing/SNLI paper/abstract.tex
Paper updates; config updates.
Removing transfer section. :P

@@
 \begin{abstract} 
-  % Entailment and contradiction are fundamental to natural language
-  % meaning and lie at the heart of tasks like summarization, textual
-  % entailment, and commonsense reasoning.
-  Entailment and contradiction are fundamental to natural language
-  meaning and so are an important testing
-  ground for semantic representations.
-  %
-  %However, computational modeling of these concepts is currently
-  %hampered by the limited nature of existing annotated resources.
-  However, progress is currently hampered by the limitations of
-  existing annotated resources for training.
-  %
-  To address this, we present a new corpus of sentence pairs labeled
-  for entailment, contradiction, and semantic independence. At 570K
-  pairs, it is orders of magnitude larger than all other resources
-  of its type.
-  %
-  And, in contrast to many such resources, all of the sentences were
-  written by humans in a grounded, naturalistic context.
-  %
-  In this paper, we use this corpus to evaluate a wide variety of
-  models for natural language inference, finding that Tree-Structured
-  Long Short-Term Memory networks (TreeLSTMs) achieve the best
-  performance.
-  %
-  In addition, to enhance the case for TreeLSTMs, and deepen the
-  argument for the centrality of entailment and contradiction in
-  semantics, we show that the TreeLSTM's representations, trained on
-  our corpus, perform well in a disparate set of additional semantic
-  tasks.  
+Understanding entailment and contradiction is  fundamental to understanding natural language, and inference about entailment and contradiction is a valuable testing ground for the development of semantic representations.
+However, research in this area has been dramatically limited by the lack of large-scale resources. 
+To address this, we introduce a new freely available corpus of labeled sentence pairs, written by humans in a novel grounded task framing based on image captioning. 
+At 570K pairs, it is two orders of magnitude larger than all other resources of its type. 
+In this paper, we use this corpus to evaluate a wide variety of natural language inference models, finding that Tree-Structured Long Short-Term Memory networks (TreeLSTMs) achieve the best performance.
+In addition, we show that models trained on our corpus perform competitively on an existing NLI test set.
 \end{abstract}

### 2015-05-30 writing/SNLI paper/abstract.tex
Incremental paper update.

@@
 At 570K pairs, it is two orders of magnitude larger than all other resources of its type. 
-In this paper, we use this corpus to evaluate a wide variety of natural language inference models, finding that Tree-Structured Long Short-Term Memory networks (TreeLSTMs) achieve the best performance.
-In addition, we show that models trained on our corpus perform competitively on an existing NLI test set.
+We also use this corpus to evaluate several natural language inference models.
+We find that a simple Long Short-Term Memory (LSTM) based model trained on our corpus not only performs competitively on our own evaluation set, but also achieves the best reported performance for a neural network model on an existing natural language inference benchmark.
 \end{abstract}

### 2015-05-30 writing/SNLI paper/entailment_eval.tex
Incremental paper update.

@@
   RTE-3 dataset.
-\todo{Cite RTE-3 here.}
 \todo{Final numbers}
@@
 % the best systems
-We report results in \reftab{eopresults}.
+We report results in \reftab{eopresults}, covering both our own test set 
+and the standard RTE-3 test set \cite{giampiccolo2007third}.
 The best edit distance algorithm tunes the weight of the three 
@@
   little lexicalization.
+\todo{Is the RTE3 model trained on RTE3?}
 
@@
 
-% Commenting out until PDF available. -S
-% Running 'make' should generate PDF from gnuplot
-\Fig{learning_curves_bow.pdf}{0.45}{bowlearncurve}{
-A learning curve for the lexicalized and unlexicalized baseline classifiers,
-plotted on a log scale.
-Note the y axis starting at a random-chance accuracy of 33\%.
-}
-
 % Results
@@
   and for removing all lexicalized features.
-In addition, we report learning curves for the lexicalized and fully unlexicalized
-  classifiers in \reffig{bowlearncurve}.
 
@@
 In fact, the addition of these features alone allow the classifier to
-  outperform the RTE baselines -- this is in some sense unsurprising:
-  the RTE systems have been tuned on relatively small corpora
+  outperform the EOP models -- this is in some sense unsurprising:
+  the EOP systems have been tuned on relatively small corpora
   ($\sim$1600 examples), whereas a classifier trained on our corpus can make use of
@@
   \todo{Use curve to justify large size of corpus.}
-
-% Insights 2: the learning curve for the unlexicalized classifier is sad
-More surprising is how quickly the lexicalized classifier outperforms its
-  unlexicalized counterpart (see \reffig{bowlearncurve}).
-With only 100 training examples, the cross-bigram classifier is already
-  outperforming its unlexicalized counterpart.
-Empirically we notice that the top weighted features for the classifier
-  trained on 100 examples tend to be high precision entailments;
-  e.g.,
-  \textit{playing} $\rightarrow$ \textit{outside}
-  (most scenes are outdoors), or \textit{a banana} $\rightarrow$
-  \textit{person eating}.
-If relatively few spurious entailments get high weight -- as it appears
-  is the case -- then it makes sense that when these do fire they
-  provide some boost in accuracy.
-
-%\subsection{Existing NLI systems}
 
@@
 
-Our neural network classifier, depicted in Fig.~\ref{modelstructure} is simply a stack of three 100d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
+Our neural network classifier, depicted in Fig.~\ref{modelstructure} is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
 
-We test six sentence representation models, each set to use 50d word and phrase embeddings.
+We test three sentence representation models, each set to use 100d word and phrase embeddings.
 Our baseline model in this experiment simply sums the embeddings of the word in each sentence.
-Next, we experiment with two recurrent models: a plain RNN, and a one-layer LSTM RNN \cite{hochreiter1997long}.
-Finally, we experiment with three tree-structured models: a plain TreeRNN \cite{socher2011semi}, a tensor-parameterized TreeRNTN \cite{socher2013acl1}, and a TreeLSTM \cite{tai2015improved}.
+In addition, we experiment with two simple sequence embedding models models: a plain RNN and an LSTM RNN \cite{hochreiter1997long}.
+Both use the simple conventional layout with one layer per input token.
 
@@
 We applied L2 regularization to all models, manually tuning the strength coefficient $\lambda$ for each.
-All models were reimplemented in a common framework for this paper, and the implementations will be made available at publication time.
+All models were implemented in a common framework for this paper, and the implementations will be made available at publication time.
 
@@
 \t{100d RNN}            & \t{75?} & \t{72?} \\	
-\t{100d LSTM RNN}            & \t{81.4?} & \t{78.2?} \\
-\t{100d TreeRNN}            & \t{69?} & \t{69?} \\
-\t{50d TreeRNTN}            & \t{61?} & \t{60?} \\
-\t{100d LSTM TreeRNN}            & \t{72?} & \t{73?} \\
+\t{100d LSTM RNN}            & \t{81.4?} & \t{78.0?} \\
+% \t{100d TreeRNN}            & \t{69?} & \t{69?} \\
+% \t{50d TreeRNTN}            & \t{61?} & \t{60?} \\
+% \t{100d LSTM TreeRNN}            & \t{72?} & \t{73?} \\
 \hline
@@
 \label{tab:nnresults}
-Accuracy in 3-class classification on the our training and test sets for each model.
+Accuracy in 3-class classification on our training and test sets for each model.
 \todo{Replace with final numbers/dims}
@@
 % From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
-\t{100d LSTM RNN (transfer init.)}            & \t{100?} & \t{80?} \\
-% From: 
+\t{100d LSTM RNN (transfer init.)}            & \t{100?} & \t{80.4?} \\
+% From: ...
 \hline
@@
 
-\todo{Why not RTE3?}
-
 We compare the performance of an LSTM model with both the transfer initialization and a normal random initialization.
@@
 This suggests that the introduction of a large high-quality corpus makes it newly possible to train representation learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on inference tasks.
+
+We attempted to apply this same transfer evaluation technique to the RTE-3 challenge dataset, but found that the small size of the training set (800 examples) did not allow our model to adapt to the
+
+\subsection{Further analysis}
+
+Fig.~\reffig{bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
+Here, we observe further confirmation that supervised neural network models like our LSTM-based model are only competitive
+% describe intersection
+% promising with more data
+% useful resource in the development of more sophisticated SE models
+
+\Fig{learning_curves_bow.pdf}{0.45}{bowlearncurve}{
+A learning curve for the lexicalized and unlexicalized baseline classifiers and the LSTM,
+plotted on a log scale.
+The minibatch-based training method that we chose for the LSTM requires at least 64 examples to function.
+Note that the y axis starts at a random-chance accuracy of 33\%.
+\todo{One more LSTM data point at ~80}}
+
+% Insights 2: the learning curve for the unlexicalized classifier is sad
+Setting aside the neural network model, were surprised at how quickly
+the lexicalized classifier outperforms its
+  unlexicalized counterpart (see \reffig{bowlearncurve}).
+With only 100 training examples, the cross-bigram classifier is already
+  outperforming its unlexicalized counterpart.
+Empirically we notice that the top weighted features for the classifier
+  trained on 100 examples tend to be high precision entailments;
+  e.g.,
+  \textit{playing} $\rightarrow$ \textit{outside}
+  (most scenes are outdoors), or \textit{a banana} $\rightarrow$
+  \textit{person eating}.
+If relatively few spurious entailments get high weight -- as it appears
+  is the case -- then it makes sense that when these do fire they
+  provide some boost in accuracy.
+
+  \todo{Insert example-specific error analysis}

### 2015-05-30 writing/SNLI paper/intro.tex
Incremental paper update.

@@
+\begin{table*}[t]
+  \centering\footnotesize
+  \begin{tabular}{p{6cm}p{0.75cm}p{1.4cm}p{6cm}}
+  \toprule
+A man inspects the uniform of a figure in some East Asian country. & \b{contr.}& \t{$\{$c c c c c$\}$} & The man is sleeping\\
+\rule{0pt}{3ex}An older and younger man smiling. & \b{neut.}& \t{$\{$n n e n n$\}$} & Two men are smiling and laughing at the cats playing on the floor.\\
+\rule{0pt}{3ex}A black race car starts up in front of a crowd of people. & \b{contr.}& \t{$\{$c c c c c$\}$} & A man is driving down a lonely road.\\
+\rule{0pt}{3ex}A soccer game with multiple males playing. & \b{ent.}& \t{$\{$e e e e e$\}$} & Some men are playing a sport.\\
+\rule{0pt}{3ex}A smiling costumed woman is holding an umbrella. & \b{neut.}& \t{$\{$n n e c n$\}$} & A happy woman in a fairy costume holds an umbrella.\\
+    \bottomrule
+% From 1.0rc3
+  \end{tabular}
+  \caption{\label{snli-examples}Randomly chosen examples from the development section of our new corpus, shown with both the gold labels and the full set of annotator labels (abbreviated).}
+\end{table*}
+
 \section{Introduction}\label{sec:introduction}
@@
 linear classifiers, and compositional neural networks. We find that
-Tree-Structured Long Short-Term Memory networks (TreeLSTMs;
-\citealt{tai2015improved,le2015compositional}) achieve the best
-performance. In addition, we enhance the case for TreeLSTMs by showing
-that its representations, trained on our corpus, perform well in a
-disparate set of additional semantic tasks.  The success of these
-transfer-learning experiments is also a testament to the centrality
-of NLI in semantics.
-
-\todo{TreeLSTM vs. LSTM result TBD.}
+a neural network model based on a Long Short-Term Memory network (LSTM; 
+\citealt{hochreiter1997long}) for sentence embedding achieves the best
+performance, and we validate that strong performance by showing that
+a model trained on our data can be successfully adapted to an existing
+NLI challenge task, yielding the best reported performance for a neural network model.
 

### 2015-05-30 writing/SNLI paper/corpus.tex
Incremental paper update.

@@
+% Note: The examples have been moved to the previous section in order to appear on p. 2.
+
 \section{A new corpus for natural language inference}\label{sec:discussion}
@@
 
-We used Amazon Mechanical Turk for data collection.
-In each individual task (each HIT), a worker was presented with premise scene descriptions from a preexisting corpus, and asked to supply hypotheses that fit each of our three labels (\ii{entailment}, \ii{neutral}, and \ii{contradiction}), forcing the resulting data to be balanced among these classes.
-
-For the premises, we used captions from the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions (corresponding to about 30k images) collected in an earlier crowdsourced effort.
-The captions were not authored by the photographers who took the source images, and they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \word{Our trip to the Olympic Peninsula}).
-
-The instructions that we provided to the workers are shown in Fig.~\ref{instructions-1}.
-Below the instructions were three fields for each of three requested sentences, corresponding to our \ii{entailment}, \ii{neutral}, and \ii{contradiction} labels, a fourth field (marked optional) for reporting problems, and a link to an FAQ page.
-That FAQ grew over the course of data collection.
-It warned about disallowed techniques (e.g., reusing the same sentence for many different prompts, which we saw in a few cases), provided guidance concerning sentence length and complexity (we did not enforce a minimum length, and we allowed bare NPs as well as full sentences), and reviewed logistical issues around payment timing.
-About 2,500 workers contributed data.
-
 \begin{figure}
@@
 
+We used Amazon Mechanical Turk for data collection.
+In each individual task (each HIT), a worker was presented with premise scene descriptions from a preexisting corpus, and asked to supply hypotheses that fit each of our three labels (\ii{entailment}, \ii{neutral}, and \ii{contradiction}), forcing the resulting data to be balanced among these classes.
+
+For the premises, we used captions from the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions (corresponding to about 30k images) collected in an earlier crowdsourced effort.
+The captions were not authored by the photographers who took the source images, and they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \word{Our trip to the Olympic Peninsula}).
+
+The instructions that we provided to the workers are shown in Fig.~\ref{instructions-1}.
+Below the instructions were three fields for each of three requested sentences, corresponding to our \ii{entailment}, \ii{neutral}, and \ii{contradiction} labels, a fourth field (marked optional) for reporting problems, and a link to an FAQ page.
+That FAQ grew over the course of data collection.
+It warned about disallowed techniques (e.g., reusing the same sentence for many different prompts, which we saw in a few cases), provided guidance concerning sentence length and complexity (we did not enforce a minimum length, and we allowed bare NPs as well as full sentences), and reviewed logistical issues around payment timing.
+About 2,500 workers contributed data.
+
+Table \ref{collection-stats} reports some key statistics about the collected corpus, and Fig.~\ref{length-dist} shows the distributions of sentence lengths for both our source hypotheses and our newly collected premises.
+We observed that while premise sentences varied considerably in length, hypothesis sentences tended to be as short as possible while still providing enough information to yield a clear judgment, clustering at around seven words.
+
+\begin{table}
+\center
+  %\setlength{\tabcolsep}{15pt}
+  %\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{l l} 
+    \toprule
+\multicolumn{2}{l}{\textbf{Data set sizes:}}\\
+Training pairs &  550,152\\
+Development pairs &  10,000\\
+Test pairs & 10,000\\
+\midrule
+\multicolumn{2}{l}{\textbf{Sentence length:}}\\
+Premise mean token count & 14.1\\
+Hypothesis mean token count & 8.3 \\
+\midrule
+\multicolumn{2}{l}{\textbf{Parser output:}}\\
+Premise `S'-rooted parses & 74.0\%\\
+Hypothesis `S'-rooted parses & 88.9\%\\
+Distinct words (ignoring case) & 37,026\\
+    \bottomrule
+  \end{tabular}
+% From 1.0rc3
+\caption{\label{collection-stats}Key statistics for the raw sentence pairs. Since the two halves of each pair were collected separately, we report some statistics for both.} 
+\end{table}
+
+\begin{figure}
+\center
+\includegraphics[width=3.05in]{length_dist}
+    % From 1.0rc2, though still accurate as of rc3
+\todo{Replace w/ mean+SD if low on space.}
+\caption{\label{length-dist}The distribution of sentence length.} 
+\end{figure}
+
 \subsection{Data validation}
@@
 Though we used a very restrictive qualification (based on past approval rate) to select workers, we nonetheless discovered and deleted some instances of random guessing during an early round of validation, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
-%\todo{Leave in mention of qual? I would--CP Settled -SB}
 
@@
 challenging but realistic machine learning task.
-
-\subsection{The distributed corpus}
-
-\begin{table*}
-  \centering\footnotesize
-  \begin{tabular}{p{6cm}p{0.75cm}p{1.4cm}p{6cm}}
-  \toprule
-A man inspects the uniform of a figure in some East Asian country. & \b{contr.}& \t{$\{$c c c c c$\}$} & The man is sleeping\\
-\rule{0pt}{3ex}An older and younger man smiling. & \b{neut.}& \t{$\{$n n e n n$\}$} & Two men are smiling and laughing at the cats playing on the floor.\\
-\rule{0pt}{3ex}A black race car starts up in front of a crowd of people. & \b{contr.}& \t{$\{$c c c c c$\}$} & A man is driving down a lonely road.\\
-\rule{0pt}{3ex}A soccer game with multiple males playing. & \b{ent.}& \t{$\{$e e e e e$\}$} & Some men are playing a sport.\\
-\rule{0pt}{3ex}A smiling costumed woman is holding an umbrella. & \b{neut.}& \t{$\{$n n e c n$\}$} & A happy woman in a fairy costume holds an umbrella.\\
-    \bottomrule
-% From 1.0rc3
-  \end{tabular}
-  \caption{\label{snli-examples}Randomly chosen examples from our development set, shown with the gold labels and the full set of annotator labels (abbreviated).
-Unlike in many existing NLI corpora, the premise and hypothesis sentences in our corpus are often too dissimilar in structure to support approaches to NLI that rely on building sentence-to-sentence alignments.}
-\end{table*}
-
-Table~\ref{snli-examples} shows a set of randomly chosen validated examples from the development set with their labels.
-Qualitatively, we find the data that we collected to draw fairly extensively on commonsense knowledge, suggesting that superficial word alignment models are unlikely to be able to succeed at classification.
-We find the sentences that we collected to be largely fluent, correctly spelled English, with a mix of full sentences and caption-style noun phrase fragments.
-Punctuation and capitalization are often omitted.
-
-The data will be released upon publication under a CreativeCommons
-Attribution-ShareAlike licence, the same licence used for the Flickr30k source captions.
-
-\begin{figure}
-\center
-\includegraphics[width=3.05in]{length_dist}
-    % From 1.0rc2, though still accurate as of rc3
-\todo{Replace w/ mean+SD if low on space.}
-\caption{\label{b-table}The distribution of sentence length.} 
-\end{figure}
-
-\begin{table}
-\center
-  %\setlength{\tabcolsep}{15pt}
-  %\renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{l l} 
-    \toprule
-\multicolumn{2}{l}{\textbf{Data set sizes:}}\\
-Training pairs &  550,152\\
-Development pairs &  10,000\\
-Test pairs & 10,000\\
-\midrule
-\multicolumn{2}{l}{\textbf{Sentence length:}}\\
-Premise mean token count & 14.1\\
-Hypothesis mean token count & 8.3 \\
-\midrule
-\multicolumn{2}{l}{\textbf{Parser output:}}\\
-Premise `S'-rooted parses & 74.0\%\\
-Hypothesis `S'-rooted parses & 88.9\%\\
-Distinct words (ignoring case) & 37,026\\
-    \bottomrule
-  \end{tabular}
-% From 1.0rc3
-\todo{Unique word count.}
-\caption{\label{collection-stats}Key statistics for the raw sentence pairs. Since the two halves of each pair were collected separately, we report some statistics for both.} 
-\end{table}
 
@@
 
+\subsection{The distributed corpus}
+
+Table~\ref{snli-examples} shows a set of randomly chosen validated examples from the development set with their labels.
+Qualitatively, we find the data that we collected to draw fairly extensively on commonsense knowledge, and that hypothesis and premise sentences often differ structurally in significant ways, suggesting that superficial word alignment models are unlikely to be able to succeed at classification.
+We also find the sentences that we collected to be largely fluent, correctly spelled English, with a mix of full sentences and caption-style noun phrase fragments, though punctuation and capitalization are often omitted.
+
+The data will be released upon publication under a CreativeCommons
+Attribution-ShareAlike licence, the same licence used for the Flickr30k source captions.
+
 \paragraph{Partition} We distribute the corpus with a pre-specified train/test/development split.
@@
 The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser 3.5.2 \cite{klein2003accurate}, trained on the standard training set as well as on the Brown Corpus \cite{francis1979brown}, which we found to improve the parse quality of the descriptive sentences and noun phrases found in the descriptions).
-Parses are available in both the standard PTB format and in the binarized unlabeled format used by tree-structured neural network models, as in \S\ref{sentence-embedding}.
+Parses are available in both the standard PTB format and in the binarized unlabeled format used by tree-structured neural network models.
 

### 2015-05-30 writing/SNLI paper/entailment_eval.tex
Paper update.

@@
   over two orders of magnitude more data.
-  \todo{Use curve to justify large size of corpus.}
 
@@
 
+The results are shown in Table~\ref{tab:nnresults}.
+The sum of words model performed slightly worse than the fundamentially similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
+Of the two RNN models, our observations find that the more robust ability to learn long-term dependencies serves it well, giving it the strongest performance of any model we tested.
+
 \subsubsection{Evaluating on other inference corpora}
@@
 
-We attempted to apply this same transfer evaluation technique to the RTE-3 challenge dataset, but found that the small size of the training set (800 examples) did not allow our model to adapt to the
+\todo{Comparison with other models on 2?3?-way SICK}
+
+We attempted to apply this same transfer evaluation technique to the RTE-3 challenge dataset, but found that the small size of the training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text use in that corpus, such that neither our transfer model nor our randomly initialized model reached competitive performance.
 
@@
 Fig.~\reffig{bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
-Here, we observe further confirmation that supervised neural network models like our LSTM-based model are only competitive
-% describe intersection
-% promising with more data
+This learning curve shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance.
+It is also worth noting that though the LSTM and the lexicalized model show similar performance when trained on the data we present here, the steeper slope for the LSTM suggests that its ability to learn arbitrarily structured representations of sentence meaning would give it an advantage over the more constrained lexicalized model on still larger datasets.
+
 % useful resource in the development of more sophisticated SE models
@@
 % Insights 2: the learning curve for the unlexicalized classifier is sad
-Setting aside the neural network model, were surprised at how quickly
-the lexicalized classifier outperforms its
-  unlexicalized counterpart (see \reffig{bowlearncurve}).
-With only 100 training examples, the cross-bigram classifier is already
-  outperforming its unlexicalized counterpart.
+We were struck by how quickly the lexicalized classifier outperforms its unlexicalized counterpart.
+With only 100 training examples, the cross-bigram classifier is already outperforming its unlexicalized counterpart.
 Empirically we notice that the top weighted features for the classifier

### 2015-05-31 writing/SNLI paper/conclusion.tex
Implementing CM's suggestions.
Lots of paper tidying.

@@
 commonsense reasoning, and nearly all questions about meaningfulness
-in language can be reduced to questions of contextual and entailment
-and contradiction. This suggests that NLI is an ideal testing ground
+in language can be reduced to questions of entailment
+and contradiction in context. This suggests that NLI is an ideal testing ground
 for theories of semantic representation, and that training for NLI
@@
 independence. We used this corpus to evaluate a wide range of models,
-and we found that not only do neural models perform best, but also
-that their learned representations transfer well to new domains.
+and we found that not only do neural models perform best, both on our own
+task and on a standard challenge dataset.
 These results are powerful new evidence that neural representations

### 2015-05-31 writing/SNLI paper/corpus.tex
Paper update.
Lots of paper tidying.

@@
 Recognizing Textual Entailment (RTE)
-tasks.\footnote{\url{http://aclweb.org/aclwiki/index.php?title=Textual_Entailment_Resource_Pool}}
+challenge tasks.\footnote{\url{http://aclweb.org/aclwiki/index.php?title=Textual_Entailment_Resource_Pool}}
 These are generally high-quality, hand-labeled data sets, and they
@@
 noisy enough that it is probably suitable only as a source of
-supplementary training data. The Paraphase Database of \cite{ganitkevitch2013ppdb}
+supplementary training data. The Paraphase Database of \citealt{ganitkevitch2013ppdb}
 addresses these issues of size and quality, but only for the limited
@@
 event and entity coreference lead to insurmountable indeterminacy
-concerning the correct semantic label (\citealt{de2008finding} $\S4.3$, \citealt{marelli2014sick}). For
+concerning the correct semantic label (\citealt{de2008finding} $\S4.3$; \citealt{marelli2014sick}). For
 instance, the sentence pair \word{A boat sank in the Pacific
@@
 We used Amazon Mechanical Turk for data collection.
-In each individual task (each HIT), a worker was presented with premise scene descriptions from a preexisting corpus, and asked to supply hypotheses that fit each of our three labels (\ii{entailment}, \ii{neutral}, and \ii{contradiction}), forcing the resulting data to be balanced among these classes.
-
-For the premises, we used captions from the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions (corresponding to about 30k images) collected in an earlier crowdsourced effort.
-The captions were not authored by the photographers who took the source images, and they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \word{Our trip to the Olympic Peninsula}).
+In each individual task (each HIT), a worker was presented with premise scene descriptions from a preexisting corpus, and asked to supply hypotheses that fit each of our three labels---\ii{entailment}, \ii{neutral}, and \ii{contradiction}---forcing the resulting data to be balanced among these classes.
 
@@
 About 2,500 workers contributed data.
+
+For the premises, we used captions from the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions (corresponding to about 30k images) collected in an earlier crowdsourced effort.
+The captions were not authored by the photographers who took the source images, and they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \word{Our trip to the Olympic Peninsula}).
 
@@
     % From 1.0rc2, though still accurate as of rc3
-\todo{Replace w/ mean+SD if low on space.}
+% \todo{Replace w/ mean+SD if low on space.}
 \caption{\label{length-dist}The distribution of sentence length.} 
@@
 In order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we performed an additional round of validation for about 10\% of our data.
-This validation phase was not fundamentally different from the Mechanical Turk labeling task used to label the SICK entailment data: we presented workers with pairs of sentences (in our case, in batches of five), and ask them to choose a single label for each pair.
-We supplied each pair to four annonators, yielding five labels per pair including the label used by the original author.
-The instructions were broadly similar to the instructions for initial data collection shown in Fig.~\ref{instructions-1}, and linked to a similar FAQ.
-Though we used a very restrictive qualification (based on past approval rate) to select workers, we nonetheless discovered and deleted some instances of random guessing during an early round of validation, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
+This validation phase followed the same basic form as the Mechanical Turk labeling task used to label the SICK entailment data: we presented workers with pairs of sentences in batches of five, and ask them to choose a single label for each pair.
+We supplied each pair to four annotators, yielding five labels per pair including the label used by the original author.
+The instructions were similar to the instructions for initial data collection shown in Fig.~\ref{instructions-1}, and linked to a similar FAQ.
+Though we initially used a very restrictive qualification (based on past approval rate) to select workers for the validation task, we nonetheless discovered (and excluded from the corpus) some instances of random guessing in an early batch of work, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
 
-For each pair that we validated, we chose a gold label. If any one of
-the three labels was chosen by at least three of the five annotators,
-that label is the gold label. If there was no such consensus, which
-occurred in about 0.2\% of cases, we chose the label `-'. While rare these
-unlabeled examples are included in the corpus distribution, they are
+For each pair that we validated, we assigned a gold label. If any one of
+the three labels was chosen by at least three of the five annotators, it was 
+chosen as the gold label. If there was no such consensus, which
+occurred in about 2\% of cases, we assigned the placeholder label `-'. 
+While these unlabeled examples are included in the corpus distribution, they are
 unlikely to be helpful for the standard NLI classification task, and
-we do not include them when training and evaluating the models
-discussed in later sections. The results of this validation process
+we do not include them in either training or evaluation in the experiments that we 
+discuss in this paper.
+
+The results of this validation process
 are summarized in Table \ref{validation-stats}. 
@@
     % From 1.0rc3
-\caption{\label{validation-stats}Key statistics for the validated pairs.
-The \ii{author's label} is the label that was used by the worker who wrote the premise to create the sentence pair.
-A \ii{gold label} reflects a consensus of three votes among the author and the four annotators.}
+\caption{\label{validation-stats}Statistics for the validated pairs.
+The \ii{author's label} is the label used by the worker who wrote the premise to create the sentence pair.
+A \ii{gold label} reflects a consensus of three votes from among the author and the four annotators.}
 \end{table}
@@
 
-The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser 3.5.2 \cite{klein2003accurate}, trained on the standard training set as well as on the Brown Corpus \cite{francis1979brown}, which we found to improve the parse quality of the descriptive sentences and noun phrases found in the descriptions).
-Parses are available in both the standard PTB format and in the binarized unlabeled format used by tree-structured neural network models.
+The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser 3.5.2 \cite{klein2003accurate}, trained on the standard training set as well as on the Brown Corpus (\citealt{francis1979brown}, which we found to improve the parse quality of the descriptive sentences and noun phrases found in the descriptions).
 
-\paragraph{General evaluation standards}
-While we hope that the corpus will be valuable in a variety of ways, we encourage researchers working on tools for semantic representation and inference to evaluate on our data in a uniform way: training on only the (parsed and/or unparsed) sentences included in the training set, and doing final evaluations on only the subset of the test set for which there are single gold labels.
+% Parses are available in both the standard PTB format and in the binarized unlabeled format used by tree-structured neural network models.
+
+%\paragraph{General evaluation standards}
+%While we hope that the corpus will be valuable in a variety of ways, we encourage researchers working on tools for semantic representation and inference to evaluate on our data in a uniform way: training on only the (parsed and/or unparsed) sentences included in the training set, and doing final evaluations on only the subset of the test set for which there are single gold labels.

### 2015-05-31 writing/SNLI paper/model_structure_fig.tex
Paper updates; config updates.
Lots of paper tidying.

--- 2015-05-19 writing/SNLI paper/model_structure_fig.tex
@@
     \tikzstyle{preclass}=[fill=orange!40,text width=50mm,align=center]
-    \tikzstyle{e}=[fill=green!40,text width=25mm,align=center]
+    \tikzstyle{e}=[fill=green!40,text width=26mm,align=center]
     \tikzstyle{m}=[draw=black,text width=38mm,align=center]
@@
     \node[softmax]  (softmax) at (0*\dx,6*\dy) {3-way softmax classifier};
-    \node[preclass]  (pc3) at (0*\dx,5*\dy) {100d $\tanh$ layer};
-    \node[preclass]  (pc2) at (0*\dx,4*\dy) {100d $\tanh$ layer};
-    \node[preclass]  (pc1) at (0*\dx,3*\dy) {100d $\tanh$ layer};
-    \node[e]  (pe) at (-3*\dx,1.85*\dy) {50d premise};
-    \node[e]  (he) at (3*\dx,1.85*\dy) {50d hypothesis};
+    \node[preclass]  (pc3) at (0*\dx,5*\dy) {200d $\tanh$ layer};
+    \node[preclass]  (pc2) at (0*\dx,4*\dy) {200d $\tanh$ layer};
+    \node[preclass]  (pc1) at (0*\dx,3*\dy) {200d $\tanh$ layer};
+    \node[e]  (pe) at (-3*\dx,1.85*\dy) {100d premise};
+    \node[e]  (he) at (3*\dx,1.85*\dy) {100d hypothesis};
     \node[m]  (pem) at (-3*\dx,0.5*\dy) {embedding model\\ with premise input};
@@
 
-        \caption{The common classification architecture used in all NLI evaluations: for each model evaluated in Table~\ref{tab:nnresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 50d inputs shown here.}
+        \caption{The classification architecture used in all NLI evaluations: for each sentence embedding model evaluated in Tables \ref{tab:nnresults} and \ref{tab:transferresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 100d inputs shown here.}
   \label{modelstructure}

### 2015-05-31 writing/SNLI paper/abstract.tex
Lots of paper tidying.

@@
 We also use this corpus to evaluate several natural language inference models.
-We find that a simple Long Short-Term Memory (LSTM) based model trained on our corpus not only performs competitively on our own evaluation set, but also achieves the best reported performance for a neural network model on an existing natural language inference benchmark.
+We find that a simple Long Short-Term Memory (LSTM) based neural network model trained on our corpus not only performs best on our own evaluation set, but also achieves the best reported performance for a neural network model on an existing natural language inference benchmark.
 \end{abstract}

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
Lots of paper tidying.

@@
 \subsection{Standard entailment models}
-We evaluate a number of strong baselines on our corpus.
-The first class of baselines use models from the Excitement Open
+
+The first class of models from the Excitement Open
   Platform (EOP,
-  \citealt{pado2014design,magnini2014excitement})
-  -- an open source platform for RTE research.
+  \citealt{pado2014design,magnini2014excitement})---an open source platform for RTE research.
 %  which
@@
   classifier-based system.
-In order to calibrate the difficulty of the dataset against the 
+In order to calibrate the difficulty of our corpus against the 
   classical RTE problems, results are also reported on the (2-class) 
-  RTE-3 dataset.
+  RTE-3 challenge data.
 \todo{Final numbers}
@@
 
-  \item The overlap between words in the \premise\ and \hypothesis. We include
-  both as an absolute count and a percentage of possible overlap, and compute
-  the count both over all words and over just nouns, verbs, adjectives, 
+  \item The overlap between words in the \premise\ and \hypothesis,
+  both as an absolute count and a percentage of possible overlap, and both over 
+  all words and over just nouns, verbs, adjectives, 
   and adverbs.
@@
 Accuracy on the our dataset, including precision/recall for the
-  \textit{entailed} class on the test set.
+  \textit{entailment} class on the test set.
 Ablation results are reported for models lacking cross-bigram features 
@@
 \todo{Final numbers}
+\todo{Do we want P/R for NNs?}
 }
@@
 In fact, the addition of these features alone allow the classifier to
-  outperform the EOP models -- this is in some sense unsurprising:
+  outperform the EOP models---this is in some sense unsurprising:
   the EOP systems have been tuned on relatively small corpora
@@
 
-Our corpus is uniquely suitable for training neural network models that produce vector representations of sentence meaning.
-In this section, we compare the performance of six such models on the corpus.
-In order to best evaluate the strengths of these six models at producing informative representations, we use sentence embedding as an intermediate step in the NLI classification task: each model must produce a vector representation of each of the two sentences without using any context from the other sentence, and the two resulting vectors are then passed to a neural network classifier which predicts the label for the pair.
-This choice allows us to focus on evaluating the ability of existing models to produce representations that fully capture sentence meaning, at the possible cost of excluding from consideration neural models for NLI that directly compare the two inputs at the word or phrase level.
+Our corpus is uniquely suitable for training neural network models that produce distributed representations of sentence meaning.
+In this section, we compare the performance of three such models on the corpus.
+In order to best evaluate the strengths of these models at producing informative sentence representations, we use sentence embedding as an intermediate step in the NLI classification task: each model must produce a vector representation of each of the two sentences without using any context from the other sentence, and the two resulting vectors are then passed to a neural network classifier which predicts the label for the pair.
+This choice allows us to focus on existing models for sentence embedding, and it allows us to evaluate the ability of those models to learn useful representations of meaning (which may be independently useful for subsequent tasks), at the cost of excluding from consideration possible strong neural models for NLI that directly compare the two inputs at the word or phrase level.
 
@@
 
-Our neural network classifier, depicted in Fig.~\ref{modelstructure} is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the representation model itself.
+Our neural network classifier, depicted in Fig.~\ref{modelstructure} is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
 
-We test three sentence representation models, each set to use 100d word and phrase embeddings.
-Our baseline model in this experiment simply sums the embeddings of the word in each sentence.
+We test three sentence embedding models, each set to use 100d phrase and sentence embeddings.
+Our baseline sentence embedding simply sums the embeddings of the words in each sentence.
 In addition, we experiment with two simple sequence embedding models models: a plain RNN and an LSTM RNN \cite{hochreiter1997long}.
@@
 Accuracy in 3-class classification on our training and test sets for each model.
-\todo{Replace with final numbers/dims}
+\todo{Replace with final numbers}
 }
@@
 The sum of words model performed slightly worse than the fundamentially similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
-Of the two RNN models, our observations find that the more robust ability to learn long-term dependencies serves it well, giving it the strongest performance of any model we tested.
+Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it the strongest performance of any model we tested.
 
-\subsubsection{Evaluating on other inference corpora}
+\subsubsection{Evaluating on SICK}
 
-To the extent that successfully training a sentence model on our corpus forces that model to encode broadly accurate representations of English scene descriptions, we should expect it to be possible to adapt these representations in a different setting.
-In this section, we experiment with transfer learning on SICK, the corpus for the most recent NLI shared task.
+To the extent that successfully training a neural network model on our corpus forces that model to encode broadly accurate representations of English scene descriptions, and to build an entailment classifier over those relations, we should it to be possible to trained models for use on other NLI tasks.
+In this section, we evaluate on the SICK task using a simple transfer learning method, and achieve competitive results.
 
-To perform transfer learning, we take the parameters of strong LSTM RNN model trained on our corpus and use them to initialize new model which trained from that point only on the training portion of SICK.
-The only newly initialized parameters are the embeddings for words that appear in the new training set, but not in our training set, and those parameters are initialized using GloVe embeddings, as before.
+To perform transfer, we take the parameters of strong LSTM RNN model trained on our corpus and use them to initialize new model, which trained from that point only on the training portion of SICK.
+The only newly initialized parameters are the embeddings for words that appear in the SICK training set, but not in our training set, and those parameters are initialized using GloVe embeddings as above.
 We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength which is re-tuned.
-We additionally transfer the accumulators that are used by AdaDelta to set the learning rates with the intention of ensuring that the model does not learn too quickly in its first few epochs and destroy the knowledge accumulated in the pre-transfer phase of training.
+We additionally transfer the accumulators that are used by AdaDelta to set the learning rates.
+This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
 
@@
 % From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
-\t{100d LSTM RNN (transfer init.)}            & \t{100?} & \t{80.4?} \\
+\t{100d LSTM RNN (transfer init.)}            & \t{100?} & \t{80.8?} \\
 % From: ...
@@
 
-We attempted to apply this same transfer evaluation technique to the RTE-3 challenge dataset, but found that the small size of the training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text use in that corpus, such that neither our transfer model nor our randomly initialized model reached competitive performance.
+We attempted to apply this same transfer evaluation technique to the RTE-3 challenge, but found that the small size of the training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text used in that corpus, such that neither our transfer model nor our randomly initialized model reached competitive performance.
 
@@
 
-Fig.~\reffig{bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
-This learning curve shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance.
+Fig.~\ref{fig:bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
+This learning curve shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance for both.
 It is also worth noting that though the LSTM and the lexicalized model show similar performance when trained on the data we present here, the steeper slope for the LSTM suggests that its ability to learn arbitrarily structured representations of sentence meaning would give it an advantage over the more constrained lexicalized model on still larger datasets.
@@
   \textit{person eating}.
-If relatively few spurious entailments get high weight -- as it appears
-  is the case -- then it makes sense that when these do fire they
+If relatively few spurious entailments get high weight---as it appears
+  is the case---then it makes sense that when these do fire they
   provide some boost in accuracy.

### 2015-05-31 writing/SNLI paper/intro.tex
Lots of paper tidying.

@@
   \centering\footnotesize
-  \begin{tabular}{p{6cm}p{0.75cm}p{1.4cm}p{6cm}}
+  \begin{tabular}{p{6.4cm}p{1.7cm}p{6.4cm}}
   \toprule
-A man inspects the uniform of a figure in some East Asian country. & \b{contr.}& \t{$\{$c c c c c$\}$} & The man is sleeping\\
-\rule{0pt}{3ex}An older and younger man smiling. & \b{neut.}& \t{$\{$n n e n n$\}$} & Two men are smiling and laughing at the cats playing on the floor.\\
-\rule{0pt}{3ex}A black race car starts up in front of a crowd of people. & \b{contr.}& \t{$\{$c c c c c$\}$} & A man is driving down a lonely road.\\
-\rule{0pt}{3ex}A soccer game with multiple males playing. & \b{ent.}& \t{$\{$e e e e e$\}$} & Some men are playing a sport.\\
-\rule{0pt}{3ex}A smiling costumed woman is holding an umbrella. & \b{neut.}& \t{$\{$n n e c n$\}$} & A happy woman in a fairy costume holds an umbrella.\\
+A man inspects the uniform of a figure in some East Asian country. & \fulllabel{contradiction}{c c c c c} & The man is sleeping\\
+\rule{0pt}{3ex}An older and younger man smiling. & \fulllabel{neutral}{n n e n n} & Two men are smiling and laughing at the cats playing on the floor.\\
+\rule{0pt}{3ex}A black race car starts up in front of a crowd of people. & \fulllabel{contradiction}{c c c c c} & A man is driving down a lonely road.\\
+\rule{0pt}{3ex}A soccer game with multiple males playing. & \fulllabel{entailment}{e e e e e} & Some men are playing a sport.\\
+\rule{0pt}{3ex}A smiling costumed woman is holding an umbrella. & \fulllabel{neutral}{n n e c n} & A happy woman in a fairy costume holds an umbrella.\\
     \bottomrule
@@
 
-NLI has been addressed using a wide variety of techniques, including
-those grounded in syntactic structures, knowledge bases, and symbolic
+NLI has been addressed using a variety of techniques, including
+those based on sequence neural network models, knowledge bases, and symbolic
 logic. In recent years, it has become an important testing ground for
@@
 linear classifiers, and compositional neural networks. We find that
-a neural network model based on a Long Short-Term Memory network (LSTM; 
-\citealt{hochreiter1997long}) for sentence embedding achieves the best
+a neural network model which uses a Long Short-Term Memory network (LSTM; 
+\citealt{hochreiter1997long}) to represent sentence meaning achieves the best
 performance, and we validate that strong performance by showing that

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
Dropout mention, numbers

@@
 All of the models are randomly initialized using standard techniques and trained using AdaDelta \cite{zeiler2012adadelta} minibatch SGD until performance on the development set stops improving.
-We applied L2 regularization to all models, manually tuning the strength coefficient $\lambda$ for each.
+We applied L2 regularization to all models, manually tuning the strength coefficient $\lambda$ for each, and additionally applied dropout \cite{srivastava2014dropout} to the inputs and outputs of the sentence embedding models (though not to its internal connections) with a fixed dropout rate.
 All models were implemented in a common framework for this paper, and the implementations will be made available at publication time.

### 2015-05-31 writing/SNLI paper/abstract.tex
NN eval numbers.

--- 2015-05-30 writing/SNLI paper/abstract.tex
@@
 We also use this corpus to evaluate several natural language inference models.
-We find that a simple Long Short-Term Memory (LSTM) based model trained on our corpus not only performs competitively on our own evaluation set, but also achieves the best reported performance for a neural network model on an existing natural language inference benchmark.
+We find that a simple Long Short-Term Memory (LSTM) based neural network model trained on our corpus performs well on our evaluation set and achieves the best reported performance for a neural network model on an existing natural language inference benchmark.
 \end{abstract}

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
NN eval numbers.

@@
 \hline
-\t{100d Sum of words}            & \t{79?} & \t{76?} \\
-\t{100d RNN}            & \t{75?} & \t{72?} \\	
-\t{100d LSTM RNN}            & \t{81.4?} & \t{78.0?} \\
+\t{100d Sum of words}            & \t{79.3} & \t{75.3} \\
+% scr/snlirc3d-snlirc3-only-l0.0001-dim100-ed300-td3-pen200-do0.9-0.9-co3-comp-1-dp1-gc5-lstminit5/stat_log
+% 290500, dev 76.4, mostly converged, but check
+
+\t{100d RNN}            & \t{73.4} & \t{71.8} \\	
+% scr/snlirc3d-snlirc3-only-l0.0001-dim100-ed300-td3-pen200-do0.9-0.9-co3-comp3-dp1-gc5-lstminit5/stat_log
+% 307500, dev 71.744,
+
+\t{100d LSTM RNN}            & \t{84.8} & \b{78.0} \\
+% scr/snlirc3d-snlirc3-only-l3e-05-dim100-ed300-td3-pen200-do0.95-0.95-co3-comp2-dp1-gc5-lstminit5/stat_log
+% 372000, dev 79.039
+
 % \t{100d TreeRNN}            & \t{69?} & \t{69?} \\
@@
 Accuracy in 3-class classification on our training and test sets for each model.
-\todo{Replace with final numbers}
 }
@@
 % From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
-\t{100d LSTM RNN (transfer init.)}            & \t{100?} & \t{80.8?} \\
-% From: ...
+
+\t{100d LSTM RNN (transfer init.)}            & \t{99.9} & \b{80.8} \\
+% From: ~/quant/transfer5-sick-only-transfer-l0.0001-dim100-ed300-td3-pen200-do0.95-0.95-ws3-adi0-comp2-cdim1/stat_log
 \hline
@@
 \caption{\label{tab:transferresults}
-The performance (\% accuracy) of an LSTM RNN model on SICK under both a random initialization and a transfer initialization copied from a model trained on our corpus.  \todo{Final numbers.}} 
+The performance (\% accuracy) of an LSTM RNN model on SICK under both a random initialization and a transfer initialization copied from a model trained on our corpus.} 
 \end{table}

### 2015-05-31 writing/SNLI paper/intro.tex
NN eval numbers.

@@
 for natural language inference, including rule-based systems, simple
-linear classifiers, and compositional neural networks. We find that
-a neural network model which uses a Long Short-Term Memory network (LSTM; 
-\citealt{hochreiter1997long}) to represent sentence meaning achieves the best
-performance, and we validate that strong performance by showing that
-a model trained on our data can be successfully adapted to an existing
+linear classifiers, and compositional neural networks. 
+We find that two models achieve comparable performance: a feature-rich
+classifier model and a neural network model which uses a Long Short-Term Memory network (LSTM; 
+\citealt{hochreiter1997long}) to represent sentence meaning. We further evaluate the LSTM model
+by taking advantage of its ready support for transfer learning, and show that it can be adapted to an existing
 NLI challenge task, yielding the best reported performance for a neural network model.

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
Update results; curves

--- 2015-05-31 writing/SNLI paper/entailment_eval.tex
@@
   evaluation sets.
-% 9 systems compared against
-A number of systems have been built using the platform, 9 of them
-  applicable to English are publicly distributed with version 1.2.1
-  of the software.
-% these fall into 2 classes
-These fall into two classes of algorithms: 2 are edit distance based,
-  whereas the remaining 7 make use of different features in a
-  maximum entropy classifier.
+We evaluate against three algorithms included in the distribution,
+  and also evaluated in \newcite{magnini2014excitement}.
+These are: 
+  (1) a simple edit-distance based algorithm;
+  (2) a classifier-based algorithm; and
+  (3) an algorithm based around tree transformations.
 
-%% methodology
-%We convert the 3-way classification task in SNLI into the RTE setting
-%  by labeling both the \unknown\ and \contradiction\
-%  labels as negative entailment, and treating the \entailment\ label as
-%  the positive entailment.
-%This creates a biased dataset of 66\% negative examples.
-% we report the best results from each class
-We run the top performing edit-distance based algorithm and the top
-  performing classifier-based algorithm on our test set, as
-  determined by performance on the development set.
-Note that these models were run using the default configuration
-  with minimal tuning.
-The results should therefore be taken as a strong baseline for
-  NLI-style approaches to the problem, rather than necessarily
-  representing the state-of-the-art system's performance on the
-  task.
-%We run each of the 9 algorithms distributed with EOP on the 2-class
-%  SNLI dataset, and report results for the best edit distance 
-%  configuration and the best classifier based configuration, as
+%% 9 systems compared against
+%A number of systems have been built using the platform, 9 of them
+%  applicable to English are publicly distributed with version 1.2.1
+%  of the software.
+%% these fall into 2 classes
+%These fall into two classes of algorithms: 2 are edit distance based,
+%  whereas the remaining 7 make use of different features in a
+%  maximum entropy classifier.
+%
+%%% methodology
+%%We convert the 3-way classification task in SNLI into the RTE setting
+%%  by labeling both the \unknown\ and \contradiction\
+%%  labels as negative entailment, and treating the \entailment\ label as
+%%  the positive entailment.
+%%This creates a biased dataset of 66\% negative examples.
+%% we report the best results from each class
+%We run the top performing edit-distance based algorithm and the top
+%  performing classifier-based algorithm on our test set, as
 %  determined by performance on the development set.
+%Note that these models were run using the default configuration
+%  with minimal tuning.
+%The results should therefore be taken as a strong baseline for
+%  NLI-style approaches to the problem, rather than necessarily
+%  representing the state-of-the-art system's performance on the
+%  task.
+%%We run each of the 9 algorithms distributed with EOP on the 2-class
+%%  SNLI dataset, and report results for the best edit distance 
+%%  configuration and the best classifier based configuration, as
+%%  determined by performance on the development set.
 
@@
 \begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceL}c@{\hskip \colspaceS}c}
+\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
 \hline
-\textbf{System} & \multicolumn{4}{c}{\b{Our corpus}} & \multicolumn{2}{c}{\b{RTE-3}} \\
- & \t{P} & \t{R} & \t{F$_1$} & \t{Acc.} & \t{F$_1$} & \t{Acc.} \\
+\textbf{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
 \hline
-\t{Edit Distance} & \t{??.?} & \t{??.?} & \t{??.?} & \t{??.?} & 
-                    \t{??.?} & \t{??.?} \\
-\t{Classifier}    & \t{67.9} & \t{38.7} & \t{49.3} & \t{7X.X} & 
-                    \t{64.2} & \t{65.3} \\
+\t{Edit Distance Based}  & \t{71.9} & \t{65.4} & \t{64.4} \\
+\t{Classifier Based}     & \t{72.2} & \t{74.3} & \t{65.3} \\
+\hline
+\t{Transformation Based} & \t{??.?} & \t{76.7} & \t{67.1} \\
 \hline
@@
 \label{tab:eopresults}
-3-class accuracy and precision/recall of existing competitive RTE
-systems from the Excitement Open Platform.
-Results are reported for an edit-distance based algorithm, and a
-  classifier-based system.
-In order to calibrate the difficulty of our corpus against the 
-  classical RTE problems, results are also reported on the (2-class) 
-  RTE-3 challenge data.
-\todo{Final numbers}
+Accuracies of various entailment algorithms included in the
+  Excitement Open Platform.
+The edit distance and classifier based accuracies are are on a 2-class
+  classification task; on our corpus and SICK this results in a 66\%
+  majority class baseline.
+All RTE results are 2-class accuracy.
+The transformation based model allows for 3-class predictions on our
+  corpus and SICK; these are reported in the last row.
 }
@@
 % the best systems
-We report results in \reftab{eopresults}, covering both our own test set 
-and the standard RTE-3 test set \cite{giampiccolo2007third}.
-The best edit distance algorithm tunes the weight of the three 
-  case-insensitive edit distance operations on the training set, 
-  after removing stop words.
-The best classifier-based system makes use of information from
-  WordNet \cite{miller1995wordnet} and VerbOcean
-  \cite{chklovski2004verbocean}, and makes use of features
-  based on tree patterns and dependency tree skeletons
-  \cite{wang2007recognizing}.
-Unsurprisingly, the classification-based approach outperforms simple
-  edit distance metrics, and performs quite well despite relatively
-  little lexicalization.
-\todo{Is the RTE3 model trained on RTE3?}
+We report results in \reftab{eopresults}, covering both our own test set,
+  the SICK test data, and the standard RTE-3 test set \cite{giampiccolo2007third}.
+Each of the models was trained on its own training set.
+The results for RTE-3 are taken from \newcite{magnini2014excitement}.
+Note that the edit distance and classifier based models are evaluated only on
+  2-class entailment; all RTE results are likewise reported for 2-class entailment.
+To convert 3-class problems like SICK and our corpus to this setting, all instances
+  of \contradiction\ and \unknown\ are converted to nonentailment.
+This yields a random-chance baseline of 66\%.
+
+%The best edit distance algorithm tunes the weight of the three 
+%  case-insensitive edit distance operations on the training set, 
+%  after removing stop words.
+%The best classifier-based system makes use of information from
+%  WordNet \cite{miller1995wordnet} and VerbOcean
+%  \cite{chklovski2004verbocean}, and makes use of features
+%  based on tree patterns and dependency tree skeletons
+%  \cite{wang2007recognizing}.
+%Unsurprisingly, the classification-based approach outperforms simple
+%  edit distance metrics, and performs quite well despite relatively
+%  little lexicalization.
+%\todo{Is the RTE3 model trained on RTE3?}
 
@@
 \begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceS}c@{\hskip \colspaceM}c}
+\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceS}c}
 \hline
-\textbf{System} & \b{Dev} & \multicolumn{4}{c}{\b{Test}} \\
- & \t{Acc.} & \t{P} & \t{R} & \t{F$_1$} & \t{Acc.} \\
+\b{System} & \b{Train} & \b{Dev} & \b{Test} \\
 \hline
-\t{Lexicalized}            & \t{??.?} & \t{79.4} & \t{82.3} & \t{80.9} & \t{78.2} \\
-\t{Unigrams Only}          & \t{??.?} & \t{71.1} & \t{77.7} & \t{74.3} & \t{71.6} \\
-\t{Unlexicalized}          & \t{??.?} & \t{56.0} & \t{64.4} & \t{59.9} & \t{50.39} \\
+\t{Lexicalized}            & \t{99.7} & \t{78.7} & \t{78.2} \\
+\t{Unigrams Only}          & \t{93.1} & \t{72.2} & \t{71.6} \\
+\t{Unlexicalized}          & \t{49.4} & \t{50.1} & \t{50.39} \\
 \hline
@@
   features (Features \ref{lst:ngram}--\ref{lst:bigram}).
-\todo{Final numbers}
-\todo{Do we want P/R for NNs?}
 }

### 2015-05-31 writing/SNLI paper/conclusion.tex

@@
 independence. We used this corpus to evaluate a wide range of models,
-and we found that not only do neural models perform best, both on our own
+and we found that neural models perform competitively, both on our own
 task and on a standard challenge dataset.

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
Error analysis stuff

--- 2015-05-31 writing/SNLI paper/entailment_eval.tex
@@
 We were struck by how quickly the lexicalized classifier outperforms its unlexicalized counterpart.
-With only 100 training examples, the cross-bigram classifier is already outperforming its unlexicalized counterpart.
-Empirically we notice that the top weighted features for the classifier
+With only 100 training examples, the cross-bigram classifier already outperforms its unlexicalized counterpart.
+Empirically, we find that the top weighted features for the classifier
   trained on 100 examples tend to be high precision entailments;
@@
 If relatively few spurious entailments get high weight---as it appears
-  is the case---then it makes sense that when these do fire they
-  provide some boost in accuracy.
+is the case---then it makes sense that, when these do fire, they
+boost accuracy.
 
-  \todo{Insert example-specific error analysis}
+%\todo{Insert example-specific error analysis}
+
+There are revealing patterns in the errors common to all the models
+considered here. Despite the large size of the training corpus and the
+distributional information captured by GloVe initialization, many
+lexical relationships are still misanalyzed, leading to incorrect
+predictions of \ii{independent}, even for pairs that are common in the
+training corpus like \word{beach}/\word{surf} and
+\word{sprinter}/\word{runner}. Semantic mistakes at the phrasal level
+(e.g., predicting contradiction for \word{A shopper buys cat food at a
+  Walmart}/\word{A person shops for their pet at a stor}) indicate
+that additional attention to compositional semantics would pay off.
+%
+% Others that could replace the above:
+% \word{Two teen girls relax on a black futon}/\word{Two young girls are sitting inside}
+% \word{A male is placing an order in a deli}/\word{A man buying a sandwich at a deli}
+%
+However, many of the persistent problems run deeper, to inferences
+that depend on world knowledge and context-specific inferences, as in
+the entailment pair \word{A race car driver leaps from a burning
+  car}/\word{A race car driver escaping danger}, for which all our
+models predict \ii{independent}. Some of these examples have qualities
+reminiscent of Winograd schema \cite{Winograd:1972,Levesque:2013}. For
+example, it seems likely that all the models wrongly predict
+entailment for \word{A young girl throws sand toward the
+  ocean}/\word{A girl can't stand the ocean}, presumably because of
+distributional associations between \word{throws} and \word{can't
+  stand}.
+
+% A woman prepares ingredients for a bowl of soup.	A soup bowl prepares a woman.

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
Fix class bias in SICK

--- 2015-05-31 writing/SNLI paper/entailment_eval.tex
@@
 The edit distance and classifier based accuracies are are on a 2-class
-  classification task; on our corpus and SICK this results in a 66\%
-  majority class baseline.
+  classification task; this results in a majority-class baseline
+  of 66\% on our class and 71\% on SICK.
 All RTE results are 2-class accuracy.

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
Incremental paper update.

@@
 
-Our neural network classifier, depicted in Fig.~\ref{modelstructure} is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
+Our neural network classifier, depicted in Figure\ref{modelstructure} is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
 
@@
 The sum of words model performed slightly worse than the fundamentially similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
-Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it the strongest performance of any model we tested.
+Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it a substantial advantage over the plain RNN, and allowing it to roughly tie in performance with the lexicalized classifier.
 
@@
 
-To the extent that successfully training a neural network model on our corpus forces that model to encode broadly accurate representations of English scene descriptions, and to build an entailment classifier over those relations, we should it to be possible to trained models for use on other NLI tasks.
-In this section, we evaluate on the SICK task using a simple transfer learning method, and achieve competitive results.
+To the extent that successfully training a neural network model like our LSTM on our corpus forces that model to encode broadly accurate representations of English scene descriptions, and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
+In this section, we evaluate on the SICK entailment task using a simple transfer learning method, and achieve competitive results.
 
-To perform transfer, we take the parameters of strong LSTM RNN model trained on our corpus and use them to initialize new model, which trained from that point only on the training portion of SICK.
+To perform transfer, we take the parameters of the LSTM RNN model trained on our corpus and use them to initialize new model, which trained from that point only on the training portion of SICK.
 The only newly initialized parameters are the embeddings for words that appear in the SICK training set, but not in our training set, and those parameters are initialized using GloVe embeddings as above.
@@
 
-Our transfer model shows the best performance yet reported on SICK for an unaugmented neural network model, and approaches the overall state of the art at 84.6\% \cite{lai2014illinois}, as well as the 84\% level of interannotator agreement, which represents a likely performance ceiling.
+Our transfer model shows the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which represents a likely performance ceiling.
 The randomly initialized model performs relatively poorly.
 This suggests that the introduction of a large high-quality corpus makes it newly possible to train representation learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on inference tasks.
-
-\todo{Comparison with other models on 2?3?-way SICK}
 
@@
 
-Fig.~\ref{fig:bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
+Figure\ref{fig:bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
 This learning curve shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance for both.
 It is also worth noting that though the LSTM and the lexicalized model show similar performance when trained on the data we present here, the steeper slope for the LSTM suggests that its ability to learn arbitrarily structured representations of sentence meaning would give it an advantage over the more constrained lexicalized model on still larger datasets.
+Since the LSTM is trained using SGD with minibatches of 64 examples, and was tuned in that setting, our ability to meaningfully evaluate it on extremely small fractions of the data was limited.
 
@@
 A learning curve for the lexicalized and unlexicalized baseline classifiers and the LSTM,
-plotted on a log scale.
-The minibatch-based training method that we chose for the LSTM requires at least 64 examples to function.
-Note that the y axis starts at a random-chance accuracy of 33\%.
-\todo{One more LSTM data point at ~80}}
+plotted on a log scale. Note that the y axis starts at a random-chance accuracy of 33\%.}
 

### 2015-05-31 writing/SNLI paper/intro.tex
Incremental paper update.

@@
 computational systems
-\cite{dagan2006pascal,MacCartney09,maccartney2009extended} --- is
+\cite{Fyodorov-etal:2000,Condoravdi-etAl:2003,BosMar:2005,dagan2006pascal,maccartney2009extended} --- is
 essential in tasks ranging from information retrieval to semantic
 parsing to commonsense reasoning.
-\todo{Non-MacCartney NLI citation.}
 
 NLI has been addressed using a variety of techniques, including
-those based on sequence neural network models, knowledge bases, and symbolic
-logic. In recent years, it has become an important testing ground for
+those based on symbolic logic, knowledge bases, and neural networks. 
+In recent years, it has become an important testing ground for
 approaches employing \emph{distributed} word and phrase
@@
 trained to support the full range of logical and commonsense
-inferences required for NLI. In the SemEval 2014 task aimed at evaluating distributed
+inferences required for NLI \cite{Weston-etal:2015,Weston-etal:2015Q}. 
+In the SemEval 2014 entailment task, which was aimed at evaluating distributed
 representations for NLI, the best-performing systems relied heavily on
@@
 
-Our primary objective in this paper is to provide a new empirical
-evaluation of a wide range of models for distributed semantic
-representations in the context of NLI, thereby deepening the case for
-NLI as a powerful evaluation of domain-general approaches to semantic
-representation. However, in our view, existing NLI corpora do not
+Our ultimate objective in this paper is to provide an empirical
+evaluation of learning centered approaches to NLI with a focus on
+systems for representation learning,
+thereby deepening the case for NLI as a powerful evaluation of 
+domain-general approaches to semantic representation. 
+However, in our view, existing NLI corpora do not
 permit such an assessment. They are generally too small for training
@@
 
-To address this, we present a new corpus of sentence pairs labeled for
+To address this, this paper introduces a new corpus of sentence pairs labeled for
 entailment, contradiction, and semantic independence. At 570,152
@@
 four additional judgments for each label for 56,941 of the examples.
-Of these, 98\% of cases emerge with a three annotator consensus, 
-and 58\% see a unanimous consensus from all five annotators, 
-which attests to the high quality of the data.
+Of these, 98\% of cases emerge with a three-annotator consensus, 
+and 58\% see a unanimous consensus from all five annotators.
 
-In this paper, we use this corpus to evaluate a wide variety of models
+In this paper, we use this corpus to evaluate a variety of models
 for natural language inference, including rule-based systems, simple
-linear classifiers, and compositional neural networks. 
+linear classifiers, and neural network-based models. 
 We find that two models achieve comparable performance: a feature-rich
@@
 by taking advantage of its ready support for transfer learning, and show that it can be adapted to an existing
-NLI challenge task, yielding the best reported performance for a neural network model.
+NLI challenge task, yielding the best reported performance by a neural network model and approaching the overall state of the art.
 

### 2015-05-31 writing/SNLI paper/abstract.tex
Incremental paper update.

@@
 Understanding entailment and contradiction is  fundamental to understanding natural language, and inference about entailment and contradiction is a valuable testing ground for the development of semantic representations.
-However, research in this area has been dramatically limited by the lack of large-scale resources. 
-To address this, we introduce a new freely available corpus of labeled sentence pairs, written by humans in a novel grounded task framing based on image captioning. 
+However, machine learning research in this area has been dramatically limited by the lack of large-scale resources. 
+To address this, we introduce a new freely available corpus of labeled sentence pairs, written by humans doing a novel grounded task based on image captioning.
 At 570K pairs, it is two orders of magnitude larger than all other resources of its type. 
-We also use this corpus to evaluate several natural language inference models.
-We find that a simple Long Short-Term Memory (LSTM) based neural network model trained on our corpus performs well on our evaluation set and achieves the best reported performance for a neural network model on an existing natural language inference benchmark.
+We find that this increase in scale allows lexicalized classifier is able to outperform some sophisticated existing entailment models, and that it allows a neural network-based model to perform competitively on natural language inference benchmarks for the first time.
 \end{abstract}

### 2015-05-31 writing/SNLI paper/corpus.tex
Dropout mention, numbers
Incremental paper update.

@@
 instance, the sentence pair \word{A boat sank in the Pacific
-  ocean} and \word{A boat sank in the ocean} could reasonably be labeled in
+  ocean} and \word{A boat sank in an ocean} could reasonably be labeled in
 numerous ways depending on how the objects and events are grounded. If
@@
 commonsense way). If the boats are assumed to be the same and 
-\word{the ocean} in the second sentence is interpreted with a specific referent 
+\word{an ocean} in the second sentence is interpreted with a specific referent 
 that is not the Pacific, then the two events are (commonsense) contradictory. This
@@
 the following crucial innovations. First, the examples were grounded
-in specific scenarios, which helps greatly in controlling event and
-entity coreference that beset existing corpora. Second, the prompt
+in specific scenarios, and the premise and hypothesis sentences in each example 
+are constrained to describe that scenario from the same perspective, 
+which helps greatly in controlling event and entity coreference. Second, the prompt
 gave participants the freedom to produce entirely novel sentences
@@
 without sacrificing consistency. Third, a subset of the resulting
-sentences were sent to a validation task aimed at flagging low-quality
-data and identifying areas of inferential uncertainty.
+sentences were sent to a validation task aimed at providing highly reliable set of annotations over the same data, and at identifying areas of inferential uncertainty.
 
@@
 We used Amazon Mechanical Turk for data collection.
-In each individual task (each HIT), a worker was presented with premise scene descriptions from a preexisting corpus, and asked to supply hypotheses that fit each of our three labels---\ii{entailment}, \ii{neutral}, and \ii{contradiction}---forcing the resulting data to be balanced among these classes.
+In each individual task (each HIT), a worker was presented with premise scene descriptions from a preexisting corpus, and asked to supply hypotheses for each of our three labels---\ii{entailment}, \ii{neutral}, and \ii{contradiction}---forcing the data to be balanced among these classes.
 
-The instructions that we provided to the workers are shown in Fig.~\ref{instructions-1}.
+The instructions that we provided to the workers are shown in Figure\ref{instructions-1}.
 Below the instructions were three fields for each of three requested sentences, corresponding to our \ii{entailment}, \ii{neutral}, and \ii{contradiction} labels, a fourth field (marked optional) for reporting problems, and a link to an FAQ page.
@@
 It warned about disallowed techniques (e.g., reusing the same sentence for many different prompts, which we saw in a few cases), provided guidance concerning sentence length and complexity (we did not enforce a minimum length, and we allowed bare NPs as well as full sentences), and reviewed logistical issues around payment timing.
-About 2,500 workers contributed data.
+About 2,500 workers contributed.
 
@@
 
-Table \ref{collection-stats} reports some key statistics about the collected corpus, and Fig.~\ref{length-dist} shows the distributions of sentence lengths for both our source hypotheses and our newly collected premises.
+Table \ref{collection-stats} reports some key statistics about the collected corpus, and Figure\ref{length-dist} shows the distributions of sentence lengths for both our source hypotheses and our newly collected premises.
 We observed that while premise sentences varied considerably in length, hypothesis sentences tended to be as short as possible while still providing enough information to yield a clear judgment, clustering at around seven words.
@@
 We supplied each pair to four annotators, yielding five labels per pair including the label used by the original author.
-The instructions were similar to the instructions for initial data collection shown in Fig.~\ref{instructions-1}, and linked to a similar FAQ.
+The instructions were similar to the instructions for initial data collection shown in Figure\ref{instructions-1}, and linked to a similar FAQ.
 Though we initially used a very restrictive qualification (based on past approval rate) to select workers for the validation task, we nonetheless discovered (and excluded from the corpus) some instances of random guessing in an early batch of work, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
@@
 large crowdsourcing efforts but also the uncertainty inherent in NLI.
-Regardless, the overall rates of agreement is extremely high,
+Regardless, the overall rate of agreement is extremely high,
 suggesting that the corpus is sufficiently high quality to pose a
@@
 \multicolumn{2}{l}{\textbf{General:}}\\
-Validated pairs & 56,941\\
+Validated pairs & 56,951\\
 Pairs w/ unanimous gold label & 58.3\%\\
@@
 Table~\ref{snli-examples} shows a set of randomly chosen validated examples from the development set with their labels.
-Qualitatively, we find the data that we collected to draw fairly extensively on commonsense knowledge, and that hypothesis and premise sentences often differ structurally in significant ways, suggesting that superficial word alignment models are unlikely to be able to succeed at classification.
+Qualitatively, we find the data that we collected draws fairly extensively on commonsense knowledge, and that hypothesis and premise sentences often differ structurally in significant ways, suggesting that superficial word alignment models are unlikely to be able to succeed at classification.
 We also find the sentences that we collected to be largely fluent, correctly spelled English, with a mix of full sentences and caption-style noun phrase fragments, though punctuation and capitalization are often omitted.
@@
 The test and development sets contain 10k examples each.
-Each original ImageFlickr caption occurs in only one of the three sets, and all of the examples in the test and development sets have gone through validation.
+Each original ImageFlickr caption occurs in only one of the three sets, and all of the examples in the test and development sets have been validated.
 
@@
 
-The distributed corpus includes parses for every sentence produced by the Stanford PCFG Parser 3.5.2 \cite{klein2003accurate}, trained on the standard training set as well as on the Brown Corpus (\citealt{francis1979brown}, which we found to improve the parse quality of the descriptive sentences and noun phrases found in the descriptions).
+The distributed corpus includes parses produced by the Stanford PCFG Parser 3.5.2 \cite{klein2003accurate}, trained on the standard training set as well as on the Brown Corpus (\citealt{francis1979brown}), which we found to improve the parse quality of the descriptive sentences and noun phrases found in the descriptions.
 

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
CM/CP edits.

@@
 
-The first class of models from the Excitement Open
+The first class of models is from the Excitement Open
   Platform (EOP,
@@
 %  is distributed alongside a number of RTE pipelines.
-We additionally evaluate against a strong classifier-based baseline based on
+We additionally evaluate against a strong classifier-based baseline using
   both unlexicalized and lexicalized features.
@@
 \begin{center}
+\def\t#1{\small{#1}}
 \begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
 \hline
-\textbf{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
+\b{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
 \hline
@@
 The results for RTE-3 are taken from \newcite{magnini2014excitement}.
-Note that the edit distance and classifier based models are evaluated only on
+The edit distance and classifier based models are evaluated only on
   2-class entailment; all RTE results are likewise reported for 2-class entailment.
@@
   of \contradiction\ and \unknown\ are converted to nonentailment.
-This yields a random-chance baseline of 66\%.
+This yields a chance baseline of 66\%.
 
@@
 \setlength\itemsep{-0.25em}
-  \item the BLEU score of the \hypothesis\ with respect
+  \item The BLEU score of the \hypothesis\ with respect
   to the \premise, using an n-gram length between 1 and 4.
@@
 % Insights 1: lexicalization helps a bunch
-It is interesting to note the substantial jump in accuracy from using
+There is a substantial jump in accuracy from using
   lexicalized features, and even from using the relatively sparse
@@
 In fact, the addition of these features alone allow the classifier to
-  outperform the EOP models---this is in some sense unsurprising:
+  outperform the EOP models.
+  This seems initially surprising, but it makes sense given how our corpus differs from existing ones:
   the EOP systems have been tuned on relatively small corpora
@@
 In this section, we compare the performance of three such models on the corpus.
-In order to best evaluate the strengths of these models at producing informative sentence representations, we use sentence embedding as an intermediate step in the NLI classification task: each model must produce a vector representation of each of the two sentences without using any context from the other sentence, and the two resulting vectors are then passed to a neural network classifier which predicts the label for the pair.
+To focus specifically on the strengths of these models at producing informative sentence representations, we use sentence embedding as an intermediate step in the NLI classification task: each model must produce a vector representation of each of the two sentences without using any context from the other sentence, and the two resulting vectors are then passed to a neural network classifier which predicts the label for the pair.
 This choice allows us to focus on existing models for sentence embedding, and it allows us to evaluate the ability of those models to learn useful representations of meaning (which may be independently useful for subsequent tasks), at the cost of excluding from consideration possible strong neural models for NLI that directly compare the two inputs at the word or phrase level.
@@
 
-Our neural network classifier, depicted in Figure\ref{modelstructure} is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
+Our neural network classifier, depicted in Figure\ref{modelstructure}, is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
 
@@
 
-The word embeddings for all of the models are initialized with the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned.
+The word embeddings for all of the models are initialized with the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned as part of training.
 In addition, all of the models use an additional $\tanh$ neural network layer to map these 300d embeddings into the lower-dimensional phrase and sentence embedding space.
@@
 
-\subsubsection{Evaluating on SICK}
+\subsection{Analysis of performance}
 
-To the extent that successfully training a neural network model like our LSTM on our corpus forces that model to encode broadly accurate representations of English scene descriptions, and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
-In this section, we evaluate on the SICK entailment task using a simple transfer learning method, and achieve competitive results.
-
-To perform transfer, we take the parameters of the LSTM RNN model trained on our corpus and use them to initialize new model, which trained from that point only on the training portion of SICK.
-The only newly initialized parameters are the embeddings for words that appear in the SICK training set, but not in our training set, and those parameters are initialized using GloVe embeddings as above.
-We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength which is re-tuned.
-We additionally transfer the accumulators that are used by AdaDelta to set the learning rates.
-This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
-\hline
-\textbf{Sentence model} & \b{Train}  & \b{Test}\\
-\hline
-\t{100d LSTM RNN (random init.)}            & \t{100} & \t{71.3} \\
-% From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
-
-\t{100d LSTM RNN (transfer init.)}            & \t{99.9} & \b{80.8} \\
-% From: ~/quant/transfer5-sick-only-transfer-l0.0001-dim100-ed300-td3-pen200-do0.95-0.95-ws3-adi0-comp2-cdim1/stat_log
-\hline
-\end{tabular}
-\end{center}
-
-\caption{\label{tab:transferresults}
-The performance (\% accuracy) of an LSTM RNN model on SICK under both a random initialization and a transfer initialization copied from a model trained on our corpus.} 
-\end{table}
-
-We compare the performance of an LSTM model with both the transfer initialization and a normal random initialization.
-We tune the regularization parameter separately for each case.
-Table \ref{tab:transferresults} shows these results.
-
-Our transfer model shows the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which represents a likely performance ceiling.
-The randomly initialized model performs relatively poorly.
-This suggests that the introduction of a large high-quality corpus makes it newly possible to train representation learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on inference tasks.
-
-We attempted to apply this same transfer evaluation technique to the RTE-3 challenge, but found that the small size of the training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text used in that corpus, such that neither our transfer model nor our randomly initialized model reached competitive performance.
-
-\subsection{Further analysis}
-
-Figure\ref{fig:bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
+Figure \ref{fig:bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
 This learning curve shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance for both.
-It is also worth noting that though the LSTM and the lexicalized model show similar performance when trained on the data we present here, the steeper slope for the LSTM suggests that its ability to learn arbitrarily structured representations of sentence meaning would give it an advantage over the more constrained lexicalized model on still larger datasets.
+It is also worth noting that, though the LSTM and the lexicalized model show similar performance when trained on the data we present here, the steeper slope for the LSTM suggests that its ability to learn arbitrarily structured representations of sentence meaning would give it an advantage over the more constrained lexicalized model on still larger datasets.
 Since the LSTM is trained using SGD with minibatches of 64 examples, and was tuned in that setting, our ability to meaningfully evaluate it on extremely small fractions of the data was limited.
@@
 A learning curve for the lexicalized and unlexicalized baseline classifiers and the LSTM,
-plotted on a log scale. Note that the y axis starts at a random-chance accuracy of 33\%.}
+plotted on a log scale. The y axis starts at a random-chance accuracy of 33\%.}
 
@@
 We were struck by how quickly the lexicalized classifier outperforms its unlexicalized counterpart.
-With only 100 training examples, the cross-bigram classifier already outperforms its unlexicalized counterpart.
+With only 100 training examples, the cross-bigram classifier already outperformed its unlexicalized counterpart.
 Empirically, we find that the top weighted features for the classifier
@@
   \textit{playing} $\rightarrow$ \textit{outside}
-  (most scenes are outdoors), or \textit{a banana} $\rightarrow$
+  (most scenes are outdoors), \textit{a banana} $\rightarrow$
   \textit{person eating}.
@@
 (e.g., predicting contradiction for \word{A shopper buys cat food at a
-  Walmart}/\word{A person shops for their pet at a stor}) indicate
+  Walmart}/\word{A person shops for their pet at a store}) indicate
 that additional attention to compositional semantics would pay off.
@@
 reminiscent of Winograd schema \cite{Winograd:1972,Levesque:2013}. For
-example, it seems likely that all the models wrongly predict
+example, all the models wrongly predict
 entailment for \word{A young girl throws sand toward the
@@
 
+\section{Transfer learning with SICK}
+
+To the extent that successfully training a neural network model like our LSTM on our corpus forces that model to encode broadly accurate representations of English scene descriptions, and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
+In this section, we evaluate on the SICK entailment task using a simple transfer learning method, and achieve competitive results.
+
+To perform transfer, we take the parameters of the LSTM RNN model trained on our corpus and use them to initialize a new model, which is trained from that point only on the training portion of SICK.
+The only newly initialized parameters are the embeddings for words that appear in the SICK training set, but not in our training set, and those parameters are initialized using GloVe embeddings as above.
+We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength, which is re-tuned.
+We additionally transfer the accumulators that are used by AdaDelta to set the learning rates.
+This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
+\hline
+\textbf{Sentence model} & \b{Train}  & \b{Test}\\
+\hline
+\t{LSTM RNN (random init.)}            & \t{100} & \t{71.3} \\
+% From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
+
+\t{LSTM RNN (transfer init.)}            & \t{99.9} & \b{80.8} \\
+% From: ~/quant/transfer5-sick-only-transfer-l0.0001-dim100-ed300-td3-pen200-do0.95-0.95-ws3-adi0-comp2-cdim1/stat_log
+\hline
+\end{tabular}
+\end{center}
+
+\caption{\label{tab:transferresults}
+The performance (\% accuracy) of a 100d LSTM RNN model on SICK under both a random initialization and a transfer initialization copied from a model trained on our corpus.} 
+\end{table}
+
+We compare the performance of an LSTM model with both the transfer initialization and a normal random initialization.
+We tune the regularization parameter separately for each case.
+Table~\ref{tab:transferresults} shows these results.
+
+Our transfer model shows the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which represents a likely performance ceiling.
+The randomly initialized model performs relatively poorly.
+This suggests that the introduction of a large high-quality corpus makes it newly possible to train representation-learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on inference tasks.
+
+We attempted to apply this same transfer evaluation technique to the RTE-3 challenge, but found that the small size of the training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text used in that corpus, such that neither our transfer model nor our randomly initialized model reached competitive performance.
+Further research on effective transfer learning on small data sets with neural models might facilitate improvements here.
+
 % A woman prepares ingredients for a bowl of soup.	A soup bowl prepares a woman.

### 2015-05-31 writing/SNLI paper/abstract.tex
Typo correction

@@
 At 570K pairs, it is two orders of magnitude larger than all other resources of its type. 
-We find that this increase in scale allows lexicalized classifier is able to outperform some sophisticated existing entailment models, and that it allows a neural network-based model to perform competitively on natural language inference benchmarks for the first time.
+We find that this increase in scale allows lexicalized classifiers to outperform some sophisticated existing entailment models, and that it allows a neural network-based model to perform competitively on natural language inference benchmarks for the first time.
 \end{abstract}

### 2015-05-31 writing/SNLI paper/intro.tex
Missing hyphen

@@
 Our ultimate objective in this paper is to provide an empirical
-evaluation of learning centered approaches to NLI with a focus on
+evaluation of learning-centered approaches to NLI with a focus on
 systems for representation learning,

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
New para on topic/focus and coref

@@
 
+Analysis of the models' predictions also yields insights into the
+extent to which they grapple with event and entity coreference. For
+the most part, the original image prompts contained a focal element
+that the caption writer identified with a syntactic subject, following
+information structuring conventions associating subjects and topics in
+English \cite{Ward04}. Our annotators generally followed suit, writing
+sentences with similar topic/focus (theme/rheme) structure that
+promotes a coherence, situation-specific construal of each sentence
+pair. This is information that our models can easily take advantage
+of, but it can lead them astray. For instance, all of them stumble
+with the amusingly simple case \emph{A woman prepares ingredients for
+  a bowl of soup}/\emph{A soup bowl prepares a woman}, in which prior
+expectations about parallelism are not met. Another headline example
+of this type is \emph{A man wearing padded arm protection is being
+  bitten by a German shepherd dog}/\emph{A man bit a dog}, which all
+the models wrongly diagnose as \ii{entailment}, though the sentences
+report two very different stories.
+
 \section{Transfer learning with SICK}

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
booktabs style throughout
Bolding in results; fixed examples; style

@@
 \hline
-\t{Transformation Based} & \t{??.?} & \t{76.7} & \t{67.1} \\
+\t{Transformation Based} & \b{??.?} & \b{76.7} & \b{67.1} \\
 \hline
@@
 \begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceS}c}
+\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
 \hline
-\b{System} & \b{Train} & \b{Dev} & \b{Test} \\
+\b{System} & \b{Train}  & \b{Test} \\ % \t{Dev}
 \hline
-\t{Lexicalized}            & \t{99.7} & \t{78.7} & \t{78.2} \\
-\t{Unigrams Only}          & \t{93.1} & \t{72.2} & \t{71.6} \\
-\t{Unlexicalized}          & \t{49.4} & \t{50.1} & \t{50.39} \\
+\t{Lexicalized}            & \t{99.7} 	& \b{78.2} \\ % & \t{78.7} 
+\t{Unigrams Only}          & \t{93.1} & \t{71.6} \\ % & \t{72.2} 
+\t{Unlexicalized}          & \t{49.4} 	 & \t{50.39} \\ % & \t{50.1} 
 \hline
@@
 The results are shown in Table~\ref{tab:nnresults}.
-The sum of words model performed slightly worse than the fundamentially similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
-Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it a substantial advantage over the plain RNN, and allowing it to roughly tie in performance with the lexicalized classifier.
+The sum of words model performed slightly worse than the fundamentally similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
+Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it a substantial advantage over the plain RNN, and  resulting in performance that is essentially equivalent to the lexicalized classifier.
 
@@
 This learning curve shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance for both.
-It is also worth noting that, though the LSTM and the lexicalized model show similar performance when trained on the data we present here, the steeper slope for the LSTM suggests that its ability to learn arbitrarily structured representations of sentence meaning would give it an advantage over the more constrained lexicalized model on still larger datasets.
-Since the LSTM is trained using SGD with minibatches of 64 examples, and was tuned in that setting, our ability to meaningfully evaluate it on extremely small fractions of the data was limited.
+In addition, though the LSTM and the lexicalized model show similar performance when trained on the current full corpus, the steeper slope for the LSTM suggests that its ability to learn arbitrarily structured representations of sentence meaning would give it an advantage over the more constrained lexicalized model on still larger datasets.
 
@@
 A learning curve for the lexicalized and unlexicalized baseline classifiers and the LSTM,
-plotted on a log scale. The y axis starts at a random-chance accuracy of 33\%.}
+plotted on a log scale.
+The y axis starts at a random-chance accuracy of 33\%.
+Since the LSTM is trained using SGD with minibatches of 64 examples, and was tuned in that setting, our ability to meaningfully evaluate it on extremely small fractions of the data was limited.}
 
 % Insights 2: the learning curve for the unlexicalized classifier is sad
-We were struck by how quickly the lexicalized classifier outperforms its unlexicalized counterpart.
+We were struck by the speed with which the lexicalized classifier outperforms its unlexicalized counterpart.
 With only 100 training examples, the cross-bigram classifier already outperformed its unlexicalized counterpart.
@@
 
-%\todo{Insert example-specific error analysis}
-
 There are revealing patterns in the errors common to all the models
@@
 \word{sprinter}/\word{runner}. Semantic mistakes at the phrasal level
-(e.g., predicting contradiction for \word{A shopper buys cat food at a
-  Walmart}/\word{A person shops for their pet at a store}) indicate
+(e.g., predicting contradiction for word{A male is placing an order in a 
+deli}/\word{A man buying a sandwich at a deli}) indicate
 that additional attention to compositional semantics would pay off.
@@
 % \word{A male is placing an order in a deli}/\word{A man buying a sandwich at a deli}
-%
+% \word{A shopper buys cat food at a Walmart}/\word{A person shops for their pet at a store}
 However, many of the persistent problems run deeper, to inferences
@@
 the entailment pair \word{A race car driver leaps from a burning
-  car}/\word{A race car driver escaping danger}, for which all our
-models predict \ii{independent}. Some of these examples have qualities
+  car}/\word{A race car driver escaping danger}, for which both
+the lexicalized classifier and the LSTM predict \ii{neutral}. 
+In other cases, models' attempts to shortcut this kind of inference 
+through lexical cues can lead them astray. 
+Some of these examples have qualities
 reminiscent of Winograd schema \cite{Winograd:1972,Levesque:2013}. For
@@
 sentences with similar topic/focus (theme/rheme) structure that
-promotes a coherence, situation-specific construal of each sentence
+promotes a coherent, situation-specific construal of each sentence
 pair. This is information that our models can easily take advantage

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
Merge error repair.
Remaining merge repair
Macros for figs and tables

--- 2015-05-31 writing/SNLI paper/entailment_eval.tex
@@
 \begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
-\hline
+\toprule
 \b{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
-\hline
+\midrule
 \t{Edit Distance Based}  & \t{71.9} & \t{65.4} & \t{64.4} \\
 \t{Classifier Based}     & \t{72.2} & \t{74.3} & \t{65.3} \\
-\hline
+%\addlinespace[1ex]
 \t{Transformation Based} & \t{??.?} & \t{76.7} & \t{67.1} \\
-\hline
+\bottomrule
 \end{tabular}
@@
 % the best systems
-We report results in \reftab{eopresults}, covering both our own test set,
+We report results in \reftab{tab:eopresults}, covering both our own test set,
   the SICK test data, and the standard RTE-3 test set \cite{giampiccolo2007third}.
@@
 \begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceS}c}
-\hline
+\toprule
 \b{System} & \b{Train} & \b{Dev} & \b{Test} \\
-\hline
+\midrule
 \t{Lexicalized}            & \t{99.7} & \t{78.7} & \t{78.2} \\
 \t{Unigrams Only}          & \t{93.1} & \t{72.2} & \t{71.6} \\
-\t{Unlexicalized}          & \t{49.4} & \t{50.1} & \t{50.39} \\
-\hline
+\t{Unlexicalized}          & \t{49.4} & \t{50.1} & \t{50.4} \\ % Last value was 50.39; rounded for consistency.
+\bottomrule
 \end{tabular}
@@
 % Results
-We report results in \reftab{bowresults}, along with ablation studies for removing
+We report results in \reftab{tab:bowresults}, along with ablation studies for removing
   the bigram features (leaving only the unigram entailment pair feature),
@@
 
-Our neural network classifier, depicted in Figure\ref{modelstructure}, is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
+Our neural network classifier, depicted in \reffig{modelstructure}, is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
 
@@
 \begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
-\hline
+\toprule
 \textbf{Sentence model} & \b{Train}  & \b{Test}\\
-\hline
+\midrule
 \t{100d Sum of words}            & \t{79.3} & \t{75.3} \\
@@
 % \t{100d LSTM TreeRNN}            & \t{72?} & \t{73?} \\
-\hline
+\bottomrule
 \end{tabular}
@@
 
-The results are shown in Table~\ref{tab:nnresults}.
+The results are shown in \reftab{tab:nnresults}.
 The sum of words model performed slightly worse than the fundamentially similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
@@
 
-Figure \ref{fig:bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
+\Reffig{fig:bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
 This learning curve shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance for both.
@@
 \begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
-\hline
+\toprule
 \textbf{Sentence model} & \b{Train}  & \b{Test}\\
-\hline
+\midrule
 \t{LSTM RNN (random init.)}            & \t{100} & \t{71.3} \\
@@
 % From: ~/quant/transfer5-sick-only-transfer-l0.0001-dim100-ed300-td3-pen200-do0.95-0.95-ws3-adi0-comp2-cdim1/stat_log
-\hline
+\bottomrule
 \end{tabular}
@@
 We tune the regularization parameter separately for each case.
-Table~\ref{tab:transferresults} shows these results.
+\Reftab{tab:transferresults} shows these results.
 

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
Tweaks for page limit.

--- 2015-05-31 writing/SNLI paper/entailment_eval.tex
@@
-\section{Our data as a platform for NLI research}
+\section{Our data as a platform for evaluation}
 
@@
 \begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
-\hline
+\toprule
 \b{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
-\hline
+\midrule
 \t{Edit Distance Based}  & \t{71.9} & \t{65.4} & \t{64.4} \\
 \t{Classifier Based}     & \t{72.2} & \t{74.3} & \t{65.3} \\
-\hline
+\midrule
 \t{Transformation Based} & \b{??.?} & \b{76.7} & \b{67.1} \\
-\hline
+\bottomrule
 \end{tabular}
@@
 \begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
-\hline
-\b{System} & \b{Train}  & \b{Test} \\ % \t{Dev}
-\hline
+\toprule
+\b{System} & \b{Train} & \b{Test} \\
+\midrule
 \t{Lexicalized}            & \t{99.7}  & \b{78.2} \\ % & \t{78.7}
 \t{Unigrams Only}          & \t{93.1} & \t{71.6} \\ % & \t{72.2}
-\t{Unlexicalized}          & \t{49.4} 	 & \t{50.39} \\ % & \t{50.1} 
-\hline
+\t{Unlexicalized}          & \t{49.4} & \t{50.4} \\ % & \t{50.1} % Last value was 50.39; rounded for consistency.
+\bottomrule
 \end{tabular}
@@
 A learning curve for the lexicalized and unlexicalized baseline classifiers and the LSTM,
-plotted on a log scale.
-The y axis starts at a random-chance accuracy of 33\%.
-Since the LSTM is trained using SGD with minibatches of 64 examples, and was tuned in that setting, our ability to meaningfully evaluate it on extremely small fractions of the data was limited.}
+plotted on a log scale. The y axis starts at a random-chance accuracy of 33\%. The minibatch size of 64 that we used to tune the LSTM sets a lower bound on data for that model.}
 
@@
 Some of these examples have qualities
-reminiscent of Winograd schema \cite{Winograd:1972,Levesque:2013}. For
+reminiscent of Winograd schemas \cite{Winograd:1972,Levesque:2013}. For
 example, all the models wrongly predict

### 2015-05-31 writing/SNLI paper/entailment_eval.tex

--- 2015-05-31 writing/SNLI paper/entailment_eval.tex
@@
 
-The results are shown in Table~\ref{tab:nnresults}.
+The results are shown in \reftab{tab:nnresults}.
 The sum of words model performed slightly worse than the fundamentally similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.

### 2015-05-31 writing/SNLI paper/entailment_eval.tex
Numbers

--- 2015-05-31 writing/SNLI paper/entailment_eval.tex
@@
-\section{Our data as a platform for evaluation}
+\section{Our data as a platform for NLI research}
 
@@
 
-The first class of models is from the Excitement Open
+The first class of models from the Excitement Open
   Platform (EOP,
@@
 %  is distributed alongside a number of RTE pipelines.
-We additionally evaluate against a strong classifier-based baseline using
+We additionally evaluate against a strong classifier-based baseline based on
   both unlexicalized and lexicalized features.
@@
 %
+% Some definitions
+\def\t#1{\small{#1}}
+\def\b#1{\t{\textbf{#1}}}
+\def\colspaceS{2.0mm}
+\def\colspaceM{3.0mm}
+\def\colspaceL{4.0mm}
 
@@
 \begin{center}
-\def\t#1{\small{#1}}
 \begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
-\toprule
-\b{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
-\midrule
-\t{Edit Distance Based}  & \t{71.9} & \t{65.4} & \t{64.4} \\
-\t{Classifier Based}     & \t{72.2} & \t{74.3} & \t{65.3} \\
-\midrule
-\t{Transformation Based} & \b{??.?} & \b{76.7} & \b{67.1} \\
-\bottomrule
+\hline
+\textbf{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
+\hline
+\t{Edit Distance Based}  & \t{71.9} & \t{65.4} & \t{61.9} \\
+\t{Classifier Based}     & \t{72.2} & \t{74.3} & \t{61.5} \\
+\hline
+\t{Transformation Based} & \t{36.0} & \t{76.7} & \t{56.4} \\
+\hline
 \end{tabular}
@@
 The results for RTE-3 are taken from \newcite{magnini2014excitement}.
-The edit distance and classifier based models are evaluated only on
+Note that the edit distance and classifier based models are evaluated only on
   2-class entailment; all RTE results are likewise reported for 2-class entailment.
@@
   of \contradiction\ and \unknown\ are converted to nonentailment.
-This yields a chance baseline of 66\%.
+This yields a random-chance baseline of 66\%.
 
@@
 \setlength\itemsep{-0.25em}
-  \item The BLEU score of the \hypothesis\ with respect
+  \item the BLEU score of the \hypothesis\ with respect
   to the \premise, using an n-gram length between 1 and 4.
@@
 \begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
-\toprule
-\b{System} & \b{Train} & \b{Test} \\
-\midrule
-\t{Lexicalized}            & \t{99.7}  & \b{78.2} \\ % & \t{78.7}
-\t{Unigrams Only}          & \t{93.1} & \t{71.6} \\ % & \t{72.2}
-\t{Unlexicalized}          & \t{49.4} & \t{50.4} \\ % & \t{50.1} % Last value was 50.39; rounded for consistency.
-\bottomrule
+\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c@{\hskip \colspaceS}c}
+\hline
+\b{System} & \b{Train} & \b{Dev} & \b{Test} \\
+\hline
+\t{Lexicalized}            & \t{99.7} & \t{78.7} & \t{78.2} \\
+\t{Unigrams Only}          & \t{93.1} & \t{72.2} & \t{71.6} \\
+\t{Unlexicalized}          & \t{49.4} & \t{50.1} & \t{50.39} \\
+\hline
 \end{tabular}
@@
 % Insights 1: lexicalization helps a bunch
-There is a substantial jump in accuracy from using
+It is interesting to note the substantial jump in accuracy from using
   lexicalized features, and even from using the relatively sparse
@@
 In fact, the addition of these features alone allow the classifier to
-  outperform the EOP models.
-  This seems initially surprising, but it makes sense given how our corpus differs from existing ones:
+  outperform the EOP models---this is in some sense unsurprising:
   the EOP systems have been tuned on relatively small corpora
@@
 In this section, we compare the performance of three such models on the corpus.
-To focus specifically on the strengths of these models at producing informative sentence representations, we use sentence embedding as an intermediate step in the NLI classification task: each model must produce a vector representation of each of the two sentences without using any context from the other sentence, and the two resulting vectors are then passed to a neural network classifier which predicts the label for the pair.
+In order to best evaluate the strengths of these models at producing informative sentence representations, we use sentence embedding as an intermediate step in the NLI classification task: each model must produce a vector representation of each of the two sentences without using any context from the other sentence, and the two resulting vectors are then passed to a neural network classifier which predicts the label for the pair.
 This choice allows us to focus on existing models for sentence embedding, and it allows us to evaluate the ability of those models to learn useful representations of meaning (which may be independently useful for subsequent tasks), at the cost of excluding from consideration possible strong neural models for NLI that directly compare the two inputs at the word or phrase level.
@@
 
-Our neural network classifier, depicted in Figure\ref{modelstructure}, is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
+Our neural network classifier, depicted in Fig.~\ref{modelstructure} is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
 
@@
 
-The word embeddings for all of the models are initialized with the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned as part of training.
+The word embeddings for all of the models are initialized with the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned.
 In addition, all of the models use an additional $\tanh$ neural network layer to map these 300d embeddings into the lower-dimensional phrase and sentence embedding space.
@@
 \begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
-\toprule
+\hline
 \textbf{Sentence model} & \b{Train}  & \b{Test}\\
-\midrule
+\hline
 \t{100d Sum of words}            & \t{79.3} & \t{75.3} \\
@@
 % \t{100d LSTM TreeRNN}            & \t{72?} & \t{73?} \\
-\bottomrule
+\hline
 \end{tabular}
@@
 The results are shown in Table~\ref{tab:nnresults}.
-The sum of words model performed slightly worse than the fundamentally similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
-Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it a substantial advantage over the plain RNN, and  resulting in performance that is essentially equivalent to the lexicalized classifier.
+The sum of words model performed slightly worse than the fundamentially similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
+Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it the strongest performance of any model we tested.
 
-\subsection{Analysis of performance}
+\subsubsection{Evaluating on SICK}
 
-Figure \ref{fig:bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
+To the extent that successfully training a neural network model on our corpus forces that model to encode broadly accurate representations of English scene descriptions, and to build an entailment classifier over those relations, we should it to be possible to trained models for use on other NLI tasks.
+In this section, we evaluate on the SICK task using a simple transfer learning method, and achieve competitive results.
+
+To perform transfer, we take the parameters of strong LSTM RNN model trained on our corpus and use them to initialize new model, which trained from that point only on the training portion of SICK.
+The only newly initialized parameters are the embeddings for words that appear in the SICK training set, but not in our training set, and those parameters are initialized using GloVe embeddings as above.
+We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength which is re-tuned.
+We additionally transfer the accumulators that are used by AdaDelta to set the learning rates.
+This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
+\hline
+\textbf{Sentence model} & \b{Train}  & \b{Test}\\
+\hline
+\t{100d LSTM RNN (random init.)}            & \t{100} & \t{71.3} \\
+% From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
+
+\t{100d LSTM RNN (transfer init.)}            & \t{99.9} & \b{80.8} \\
+% From: ~/quant/transfer5-sick-only-transfer-l0.0001-dim100-ed300-td3-pen200-do0.95-0.95-ws3-adi0-comp2-cdim1/stat_log
+\hline
+\end{tabular}
+\end{center}
+
+\caption{\label{tab:transferresults}
+The performance (\% accuracy) of an LSTM RNN model on SICK under both a random initialization and a transfer initialization copied from a model trained on our corpus.} 
+\end{table}
+
+We compare the performance of an LSTM model with both the transfer initialization and a normal random initialization.
+We tune the regularization parameter separately for each case.
+Table \ref{tab:transferresults} shows these results.
+
+Our transfer model shows the best performance yet reported on SICK for an unaugmented neural network model, and approaches the overall state of the art at 84.6\% \cite{lai2014illinois}, as well as the 84\% level of interannotator agreement, which represents a likely performance ceiling.
+The randomly initialized model performs relatively poorly.
+This suggests that the introduction of a large high-quality corpus makes it newly possible to train representation learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on inference tasks.
+
+\todo{Comparison with other models on 2?3?-way SICK}
+
+We attempted to apply this same transfer evaluation technique to the RTE-3 challenge, but found that the small size of the training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text used in that corpus, such that neither our transfer model nor our randomly initialized model reached competitive performance.
+
+\subsection{Further analysis}
+
+Fig.~\ref{fig:bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
 This learning curve shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance for both.
-In addition, though the LSTM and the lexicalized model show similar performance when trained on the current full corpus, the steeper slope for the LSTM suggests that its ability to learn arbitrarily structured representations of sentence meaning would give it an advantage over the more constrained lexicalized model on still larger datasets.
+It is also worth noting that though the LSTM and the lexicalized model show similar performance when trained on the data we present here, the steeper slope for the LSTM suggests that its ability to learn arbitrarily structured representations of sentence meaning would give it an advantage over the more constrained lexicalized model on still larger datasets.
 
@@
 A learning curve for the lexicalized and unlexicalized baseline classifiers and the LSTM,
-plotted on a log scale. The y axis starts at a random-chance accuracy of 33\%. The minibatch size of 64 that we used to tune the LSTM sets a lower bound on data for that model.}
+plotted on a log scale.
+The minibatch-based training method that we chose for the LSTM requires at least 64 examples to function.
+Note that the y axis starts at a random-chance accuracy of 33\%.
+\todo{One more LSTM data point at ~80}}
 
 % Insights 2: the learning curve for the unlexicalized classifier is sad
-We were struck by the speed with which the lexicalized classifier outperforms its unlexicalized counterpart.
-With only 100 training examples, the cross-bigram classifier already outperformed its unlexicalized counterpart.
+We were struck by how quickly the lexicalized classifier outperforms its unlexicalized counterpart.
+With only 100 training examples, the cross-bigram classifier already outperforms its unlexicalized counterpart.
 Empirically, we find that the top weighted features for the classifier
@@
   \textit{playing} $\rightarrow$ \textit{outside}
-  (most scenes are outdoors), \textit{a banana} $\rightarrow$
+  (most scenes are outdoors), or \textit{a banana} $\rightarrow$
   \textit{person eating}.
@@
 boost accuracy.
+
+%\todo{Insert example-specific error analysis}
 
@@
 \word{sprinter}/\word{runner}. Semantic mistakes at the phrasal level
-(e.g., predicting contradiction for word{A male is placing an order in a 
-deli}/\word{A man buying a sandwich at a deli}) indicate
+(e.g., predicting contradiction for \word{A shopper buys cat food at a
+  Walmart}/\word{A person shops for their pet at a stor}) indicate
 that additional attention to compositional semantics would pay off.
@@
 % \word{A male is placing an order in a deli}/\word{A man buying a sandwich at a deli}
-% \word{A shopper buys cat food at a Walmart}/\word{A person shops for their pet at a store}
+%
 However, many of the persistent problems run deeper, to inferences
@@
 the entailment pair \word{A race car driver leaps from a burning
-  car}/\word{A race car driver escaping danger}, for which both
-the lexicalized classifier and the LSTM predict \ii{neutral}. 
-In other cases, models' attempts to shortcut this kind of inference 
-through lexical cues can lead them astray. 
-Some of these examples have qualities
-reminiscent of Winograd schemas \cite{Winograd:1972,Levesque:2013}. For
-example, all the models wrongly predict
+  car}/\word{A race car driver escaping danger}, for which all our
+models predict \ii{independent}. Some of these examples have qualities
+reminiscent of Winograd schema \cite{Winograd:1972,Levesque:2013}. For
+example, it seems likely that all the models wrongly predict
 entailment for \word{A young girl throws sand toward the
@@
 
-Analysis of the models' predictions also yields insights into the
-extent to which they grapple with event and entity coreference. For
-the most part, the original image prompts contained a focal element
-that the caption writer identified with a syntactic subject, following
-information structuring conventions associating subjects and topics in
-English \cite{Ward04}. Our annotators generally followed suit, writing
-sentences with similar topic/focus (theme/rheme) structure that
-promotes a coherent, situation-specific construal of each sentence
-pair. This is information that our models can easily take advantage
-of, but it can lead them astray. For instance, all of them stumble
-with the amusingly simple case \emph{A woman prepares ingredients for
-  a bowl of soup}/\emph{A soup bowl prepares a woman}, in which prior
-expectations about parallelism are not met. Another headline example
-of this type is \emph{A man wearing padded arm protection is being
-  bitten by a German shepherd dog}/\emph{A man bit a dog}, which all
-the models wrongly diagnose as \ii{entailment}, though the sentences
-report two very different stories.
-
-\section{Transfer learning with SICK}
-
-To the extent that successfully training a neural network model like our LSTM on our corpus forces that model to encode broadly accurate representations of English scene descriptions, and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
-In this section, we evaluate on the SICK entailment task using a simple transfer learning method, and achieve competitive results.
-
-To perform transfer, we take the parameters of the LSTM RNN model trained on our corpus and use them to initialize a new model, which is trained from that point only on the training portion of SICK.
-The only newly initialized parameters are the embeddings for words that appear in the SICK training set, but not in our training set, and those parameters are initialized using GloVe embeddings as above.
-We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength, which is re-tuned.
-We additionally transfer the accumulators that are used by AdaDelta to set the learning rates.
-This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
-\toprule
-\textbf{Sentence model} & \b{Train}  & \b{Test}\\
-\midrule
-\t{LSTM RNN (random init.)}            & \t{100} & \t{71.3} \\
-% From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
-
-\t{LSTM RNN (transfer init.)}            & \t{99.9} & \b{80.8} \\
-% From: ~/quant/transfer5-sick-only-transfer-l0.0001-dim100-ed300-td3-pen200-do0.95-0.95-ws3-adi0-comp2-cdim1/stat_log
-\bottomrule
-\end{tabular}
-\end{center}
-
-\caption{\label{tab:transferresults}
-The performance (\% accuracy) of a 100d LSTM RNN model on SICK under both a random initialization and a transfer initialization copied from a model trained on our corpus.} 
-\end{table}
-
-We compare the performance of an LSTM model with both the transfer initialization and a normal random initialization.
-We tune the regularization parameter separately for each case.
-Table~\ref{tab:transferresults} shows these results.
-
-Our transfer model shows the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which represents a likely performance ceiling.
-The randomly initialized model performs relatively poorly.
-This suggests that the introduction of a large high-quality corpus makes it newly possible to train representation-learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on inference tasks.
-
-We attempted to apply this same transfer evaluation technique to the RTE-3 challenge, but found that the small size of the training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text used in that corpus, such that neither our transfer model nor our randomly initialized model reached competitive performance.
-Further research on effective transfer learning on small data sets with neural models might facilitate improvements here.
-
 % A woman prepares ingredients for a bowl of soup.	A soup bowl prepares a woman.

### 2015-05-31 writing/SNLI paper/conclusion.tex

@@
-\subsection{Conclusion}\label{sec:conclusion}
+\section{Conclusion}\label{sec:conclusion}
 
@@
 corpus of sentence pairs labeled for entailment, contradiction, and
-independence. We used this corpus to evaluate a wide range of models,
-and we found that neural models perform competitively, both on our own
-task and on a standard challenge dataset.
+independence. We used this corpus to evaluate a range of models,
+and we found that simple lexicalized models and neural network
+models perform competitively, and that the representations learned by
+a neural network model from our corpus can be used in transfer learning
+to dramatically improve performance on a standard challenge dataset.
 These results are powerful new evidence that neural representations

### 2015-05-31 writing/SNLI paper/corpus.tex
CM/CP edits.

@@
 
-\section{A new corpus for natural language inference}\label{sec:discussion}
+\section{A new corpus for NLI}\label{sec:discussion}
 
@@
 
-The instructions that we provided to the workers are shown in Figure\ref{instructions-1}.
+The instructions that we provided to the workers are shown in \reffig{instructions-1}.
 Below the instructions were three fields for each of three requested sentences, corresponding to our \ii{entailment}, \ii{neutral}, and \ii{contradiction} labels, a fourth field (marked optional) for reporting problems, and a link to an FAQ page.
@@
 
-Table \ref{collection-stats} reports some key statistics about the collected corpus, and Figure\ref{length-dist} shows the distributions of sentence lengths for both our source hypotheses and our newly collected premises.
+\Reftab{collection-stats} reports some key statistics about the collected corpus, and \reffig{length-dist} shows the distributions of sentence lengths for both our source hypotheses and our newly collected premises.
 We observed that while premise sentences varied considerably in length, hypothesis sentences tended to be as short as possible while still providing enough information to yield a clear judgment, clustering at around seven words.
@@
 We supplied each pair to four annotators, yielding five labels per pair including the label used by the original author.
-The instructions were similar to the instructions for initial data collection shown in Figure\ref{instructions-1}, and linked to a similar FAQ.
+The instructions were similar to the instructions for initial data collection shown in \reffig{instructions-1}, and linked to a similar FAQ.
 Though we initially used a very restrictive qualification (based on past approval rate) to select workers for the validation task, we nonetheless discovered (and excluded from the corpus) some instances of random guessing in an early batch of work, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
@@
 The results of this validation process
-are summarized in Table \ref{validation-stats}. 
+are summarized in \reftab{validation-stats}. 
 Nearly all of the examples received a majority
@@
 
-Table~\ref{snli-examples} shows a set of randomly chosen validated examples from the development set with their labels.
+\Reftab{snli-examples} shows a set of randomly chosen validated examples from the development set with their labels.
 Qualitatively, we find the data that we collected draws fairly extensively on commonsense knowledge, and that hypothesis and premise sentences often differ structurally in significant ways, suggesting that superficial word alignment models are unlikely to be able to succeed at classification.

### 2015-05-31 writing/SNLI paper/model_structure_fig.tex

@@
 
-        \caption{The classification architecture used in all NLI evaluations: for each sentence embedding model evaluated in Tables \ref{tab:nnresults} and \ref{tab:transferresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 100d inputs shown here.}
+        \caption{The classification architecture used in all NLI evaluations: for each sentence embedding model evaluated in \reftabs{tab:nnresults}{tab:transferresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 100d inputs shown here.}
   \label{modelstructure}

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Expand the EOP and classifier sections

--- 2015-05-31 writing/SNLI paper/entailment_eval.tex
@@
 These are: 
-  (1) a simple edit-distance based algorithm;
-  (2) a classifier-based algorithm; and
-  (3) an algorithm based around tree transformations.
+  (1) a simple edit-distance based algorithm; and
+  (2) a classifier-based algorithm.
+%  (3) an algorithm based around tree transformations.
 
@@
 \toprule
-\b{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
+\textbf{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
 \midrule
-\t{Edit Distance Based}  & \t{71.9} & \t{65.4} & \t{64.4} \\
-\t{Classifier Based}     & \t{72.2} & \t{74.3} & \t{65.3} \\
-\midrule
-\t{Transformation Based} & \b{??.?} & \b{76.7} & \b{67.1} \\
+\t{Edit Distance Based}  & \t{71.9} & \t{65.4} & \t{61.9} \\
+\t{Classifier Based}     & \t{72.2} & \t{74.3} & \t{61.5} \\
+%\midrule
+%\t{Transformation Based} & \t{36.0} & \t{76.7} & \t{56.4} \\
 \bottomrule
@@
 \label{tab:eopresults}
-Accuracies of various entailment algorithms included in the
+Accuracies of two entailment algorithms included in the
   Excitement Open Platform.
-The edit distance and classifier based accuracies are are on a 2-class
+Note that all accuracies are are on a 2-class
   classification task; this results in a majority-class baseline
   of 66\% on our class and 71\% on SICK.
-All RTE results are 2-class accuracy.
-The transformation based model allows for 3-class predictions on our
-  corpus and SICK; these are reported in the last row.
+%All RTE results are 2-class accuracy.
+%The transformation based model allows for 3-class predictions on our
+%  corpus and SICK; these are reported in the last row.
 }
@@
 Each of the models was trained on its own training set.
-The results for RTE-3 are taken from \newcite{magnini2014excitement}.
+%The results for RTE-3 are taken from \newcite{magnini2014excitement}.
 The edit distance and classifier based models are evaluated only on
@@
   of \contradiction\ and \unknown\ are converted to nonentailment.
-This yields a chance baseline of 66\%.
+This yields a chance baseline of 66\% on our corpus, and 71\% on SICK.
+
+Note that all results are reported without using any of the included lexical
+  resources.
+We do not attempt to necessarily compare against the state-of-the-art
+  RTE systems as much as give a general sense for the relative
+  difficulties of the datasets.
 
@@
   \item\label{lst:bigram} Cross-bigrams: for every pair of bigrams across the \premise\ and \hypothesis\ which share a 
-  POS tag, an indicator feature over the two bigrams.
+  POS tag on the second word, an indicator feature over the two bigrams.
 \end{enumerate}
@@
 We report results in \reftab{tab:bowresults}, along with ablation studies for removing
-  the bigram features (leaving only the unigram entailment pair feature),
+  the cross-bigram features (leaving only the cross-unigram feature),
   and for removing all lexicalized features.
-
 % Insights 1: lexicalization helps a bunch
@@
   cross-bigram features.
-In fact, the addition of these features alone allow the classifier to
-  outperform the EOP models.
-  This seems initially surprising, but it makes sense given how our corpus differs from existing ones:
-  the EOP systems have been tuned on relatively small corpora
-  ($\sim$1600 examples), whereas a classifier trained on our corpus can make use of
-  over two orders of magnitude more data.
+The latter of these was initially surprising, but could be attributed to
+  the classifier automatically learning explicit negations or adjective
+  modification.
+A similar for bigram features in sentiment analysis was shown in
+  \newcite{sidaw12simple}.
+
+% Insight 2: do well without alignments
+Another surprising fact is that the classifier performs as well as it
+  does without any notion of alignment or tree transformations.
+Although likely richer models would continue to improve performance,
+  the results suggest that given enough data
+  the more noisy part-of-speech overlap constraint can produce an effective
+  model.
+
+%In fact, the addition of these features alone allow the classifier to
+%  outperform many of the .
+%  This seems initially surprising, but it makes sense given how our corpus differs from existing ones:
+%  the EOP systems have been tuned on relatively small corpora
+%  ($\sim$1600 examples), whereas a classifier trained on our corpus can make use of
+%  over two orders of magnitude more data.
 

### 2015-06-01 writing/SNLI paper/conclusion.tex
CP comments.

@@
 to dramatically improve performance on a standard challenge dataset.
-These results are powerful new evidence that neural representations
-can be trained to accurately encode rich, multifaceted semantic
-information.
+This result serves as powerful new evidence that neural representations can, 
+when given a sufficiently extensive and challenging task, 
+be trained to accurately encode rich, multifaceted semantic information, and it invites
+further research into the problem of learning maximally faithful and expressive
+ representations.

### 2015-06-01 writing/SNLI paper/corpus.tex
CP comments.

@@
 noisy enough that it is probably suitable only as a source of
-supplementary training data. The Paraphase Database of \citealt{ganitkevitch2013ppdb}
+supplementary training data. The Paraphase Database of \newcite{ganitkevitch2013ppdb}
 addresses these issues of size and quality, but only for the limited
@@
 in specific scenarios, and the premise and hypothesis sentences in each example 
-are constrained to describe that scenario from the same perspective, 
+were constrained to describe that scenario from the same perspective, 
 which helps greatly in controlling event and entity coreference. Second, the prompt
@@
 without sacrificing consistency. Third, a subset of the resulting
-sentences were sent to a validation task aimed at providing highly reliable set of annotations over the same data, and at identifying areas of inferential uncertainty.
+sentences were sent to a validation task aimed at providing a highly 
+reliable set of annotations over the same data, and at identifying areas of inferential uncertainty.
 
@@
 In order to measure the quality of our corpus, and in order to construct maximally useful testing and development sets, we performed an additional round of validation for about 10\% of our data.
-This validation phase followed the same basic form as the Mechanical Turk labeling task used to label the SICK entailment data: we presented workers with pairs of sentences in batches of five, and ask them to choose a single label for each pair.
+This validation phase followed the same basic form as the Mechanical Turk labeling task used to label the SICK entailment data: we presented workers with pairs of sentences in batches of five, and asked them to choose a single label for each pair.
 We supplied each pair to four annotators, yielding five labels per pair including the label used by the original author.
@@
 label, indicating broad consensus about the nature of the data and
-categories. The gold-labeled examples are essentially evenly
-distributed across the three categories in our train, development, and
-testing. The Fleiss $\kappa$ scores 
+categories. The gold-labeled examples are very nearly evenly
+distributed across the three labels. The Fleiss $\kappa$ scores 
 (computed over every example with a full five annotations)
@@
 The data will be released upon publication under a CreativeCommons
-Attribution-ShareAlike licence, the same licence used for the Flickr30k source captions.
+Attribution-ShareAlike license, the same licence used for the Flickr30k source captions.
 

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Updated classifier numbers
Tentative 3-class results
CP comments.

@@
 
-The most immediate application for our corpus is in developing models for the task of NLI.
+% The most immediate application for our corpus is in developing models for the task of NLI.
 In particular, since it is dramatically larger than any existing corpus of comparable quality, we expect it to be suitable for training  parameter-rich models like neural networks for this task.
@@
 \t{Edit Distance Based}        & \t{71.9} & \t{65.4} & \t{61.9} \\
-\t{Classifier Based}     & \t{72.2} & \t{74.3} & \t{61.5} \\
+\t{Classifier Based}           & \t{72.2} & \t{71.4} & \t{61.5} \\
 %\midrule
+%\t{Classifier Based (3-class)} & \t{??.?} & \t{65.6} & \t{} \\
 %\t{Transformation Based} & \t{36.0} & \t{76.7} & \t{56.4} \\
@@
 \label{tab:eopresults}
-Accuracies of two entailment algorithms included in the
+Accuracies for two entailment algorithms included in the
   Excitement Open Platform.
@@
 %All RTE results are 2-class accuracy.
-%The transformation based model allows for 3-class predictions on our
+%The transformation-based model allows for 3-class predictions on our
 %  corpus and SICK; these are reported in the last row.
@@
 %  the EOP systems have been tuned on relatively small corpora
-%  ($\sim$1600 examples), whereas a classifier trained on our corpus can make use of
+%  ($\approx$1600 examples), whereas a classifier trained on our corpus can make use of
 %  over two orders of magnitude more data.
@@
 
-Our corpus is uniquely suitable for training neural network models that produce distributed representations of sentence meaning.
+Our corpus is suitably large and diverse to make it possible to train neural network models that produce distributed representations of sentence meaning.
 In this section, we compare the performance of three such models on the corpus.
@@
 
-Our neural network classifier, depicted in \reffig{modelstructure}, is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
+Our neural network classifier, depicted in \reffig{modelstructure} (and based on a one-layer model in \citealt{Bowman:Potts:Manning:2014}), is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
 
@@
 Our baseline sentence embedding simply sums the embeddings of the words in each sentence.
-In addition, we experiment with two simple sequence embedding models models: a plain RNN and an LSTM RNN \cite{hochreiter1997long}.
+In addition, we experiment with two simple sequence embedding models: a plain RNN and an LSTM RNN \cite{hochreiter1997long}.
 Both use the simple conventional layout with one layer per input token.
 
-The word embeddings for all of the models are initialized with the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned as part of training.
+The word embeddings for all of the models are initialized with the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine-tuned as part of training.
 In addition, all of the models use an additional $\tanh$ neural network layer to map these 300d embeddings into the lower-dimensional phrase and sentence embedding space.
@@
 A learning curve for the lexicalized and unlexicalized baseline classifiers and the LSTM,
-plotted on a log scale. The y axis starts at a random-chance accuracy of 33\%. The minibatch size of 64 that we used to tune the LSTM sets a lower bound on data for that model.}
+plotted on a log scale. The y-axis starts at a random-chance accuracy of 33\%. The minibatch size of 64 that we used to tune the LSTM sets a lower bound on data for that model.}
 
@@
 \word{sprinter}/\word{runner}. Semantic mistakes at the phrasal level
-(e.g., predicting contradiction for word{A male is placing an order in a 
+(e.g., predicting contradiction for \word{A male is placing an order in a 
 deli}/\word{A man buying a sandwich at a deli}) indicate
@@
 the lexicalized classifier and the LSTM predict \ii{neutral}. 
-In other cases, models' attempts to shortcut this kind of inference 
+In other cases, the models' attempts to shortcut this kind of inference 
 through lexical cues can lead them astray. 
@@
 English \cite{Ward04}. Our annotators generally followed suit, writing
-sentences with similar topic/focus (theme/rheme) structure that
-promotes a coherent, situation-specific construal of each sentence
+sentences that, while structurally diverse, share topic/focus (theme/rheme)
+structure with their premises.
+This promotes a coherent, situation-specific construal of each sentence
 pair. This is information that our models can easily take advantage

### 2015-06-01 writing/SNLI paper/intro.tex
CP comments.

@@
 \begin{table*}[t]
-  \centering\footnotesize
+  \centering\small
   \begin{tabular}{p{6.4cm}p{1.7cm}p{6.4cm}}
@@
 those based on symbolic logic, knowledge bases, and neural networks. 
-In recent years, it has become an important testing ground for
+In recent years, it has become an imp`ortant testing ground for
 approaches employing \emph{distributed} word and phrase
@@
 trained to support the full range of logical and commonsense
-inferences required for NLI \cite{Weston-etal:2015,Weston-etal:2015Q}. 
+inferences required for NLI \cite{Bowman:Potts:Manning:2014,Weston-etal:2015,Weston-etal:2015Q}. 
 In the SemEval 2014 entailment task, which was aimed at evaluating distributed
@@
 entailment, contradiction, and semantic independence. At 570,152
-sentence-pairs, this corpus is two orders of magnitude larger than all
+sentence pairs, this corpus is two orders of magnitude larger than all
 other resources of its type. And, in contrast to many such resources,

### 2015-06-01 writing/SNLI paper/intro.tex
Finished paper?

--- 2015-05-31 writing/SNLI paper/intro.tex
@@
 representations. Distributed representations excel at capturing
-relations based in similarity, and they have proven effective at
+relations based in similarity, and have proven effective at
 modeling simple dimensions of meaning like evaluative sentiment
@@
 trained to support the full range of logical and commonsense
-inferences required for NLI \cite{Weston-etal:2015,Weston-etal:2015Q}. 
-In the SemEval 2014 entailment task, which was aimed at evaluating distributed
+inferences required for NLI \cite{Bowman:Potts:Manning:2014,Weston-etal:2015,Weston-etal:2015Q}. 
+In a SemEval 2014 task aimed at evaluating distributed
 representations for NLI, the best-performing systems relied heavily on
@@
 systems for representation learning,
-thereby deepening the case for NLI as a powerful evaluation of 
+advancing the case for NLI as a tool for the evaluation of 
 domain-general approaches to semantic representation. 
@@
 entailment, contradiction, and semantic independence. At 570,152
-sentence-pairs, this corpus is two orders of magnitude larger than all
+sentence pairs, this corpus is two orders of magnitude larger than all
 other resources of its type. And, in contrast to many such resources,
-all of the sentences were written by humans in a grounded,
+all of its sentences and labels were written by humans in a grounded,
 naturalistic context. In a separate validation phase, we collected
@@
 We find that two models achieve comparable performance: a feature-rich
-classifier model and a neural network model which uses a Long Short-Term Memory network (LSTM; 
-\citealt{hochreiter1997long}) to represent sentence meaning. We further evaluate the LSTM model
+classifier model and a neural network model centered around a Long Short-Term Memory network (LSTM; 
+\citealt{hochreiter1997long}). We further evaluate the LSTM model
 by taking advantage of its ready support for transfer learning, and show that it can be adapted to an existing

### 2015-06-01 writing/SNLI paper/conclusion.tex
Finished paper?

--- 2015-05-31 writing/SNLI paper/conclusion.tex
@@
-\subsection{Conclusion}\label{sec:conclusion}
+\section{Conclusion}\label{sec:conclusion}
 
-Natural languages are powerful vehicles for complex, flexible
-commonsense reasoning, and nearly all questions about meaningfulness
+Natural languages are powerful vehicles for reasoning, 
+and nearly all questions about meaningfulness
 in language can be reduced to questions of entailment
@@
 corpus of sentence pairs labeled for entailment, contradiction, and
-independence. We used this corpus to evaluate a wide range of models,
-and we found that neural models perform competitively, both on our own
-task and on a standard challenge dataset.
-These results are powerful new evidence that neural representations
-can be trained to accurately encode rich, multifaceted semantic
+independence. We used this corpus to evaluate a range of models,
+and found that both simple lexicalized models and neural network
+models perform well, and that the representations learned by
+a neural network model on our corpus can be used to dramatically 
+improve performance on a standard challenge dataset. These results 
+serve as powerful new evidence that neural representations can, 
+when given a sufficiently extensive and challenging task, 
+be trained to accurately encode rich, multifaceted semantic
 information.

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Finished paper?

@@
 %  is distributed alongside a number of RTE pipelines.
-We additionally evaluate against a strong classifier-based baseline using
+We additionally evaluate against a strong but simple classifier-based model with
   both unlexicalized and lexicalized features.
@@
 \midrule
-\t{Edit Distance Based}        & \t{71.9} & \t{65.4} & \t{61.9} \\
-\t{Classifier Based}           & \t{72.2} & \t{71.4} & \t{61.5} \\
+\t{Edit Distance Based}        & \t{71.9} & \t{65.4} & \b{61.9} \\
+\t{Classifier Based}           & \b{72.2} & \b{71.4} & \t{61.5} \\
 %\midrule
@@
 \label{tab:eopresults}
-Accuracies for two entailment algorithms included in the
+2-class accuracy for two entailment algorithms included in the
   Excitement Open Platform.
-Note that all accuracies are are on a 2-class
-  classification task; this results in a majority-class baseline
-  of 66\% on our class and 71\% on SICK.
 %All RTE results are 2-class accuracy.
@@
   of \contradiction\ and \unknown\ are converted to nonentailment.
-This yields a chance baseline of 66\% on our corpus, and 71\% on SICK.
-
-Note that all results are reported without using any of the included lexical
-  resources.
-We do not attempt to necessarily compare against the state-of-the-art
-  RTE systems as much as give a general sense for the relative
+This yields a chance baseline accuracy of 66\% on our corpus, and 71\% on SICK.
+We train these models without using any of the included lexical
+  resources, so these results do not represent state-of-the-art
+  RTE systems, but rather give an indication of the relative
   difficulties of the datasets.
@@
 \paragraph{Lexicalized Classifier}
-Unlike the RTE datasets, our corpus's size allows approaches which make use of
-  rich lexicalized features.
-We implement such a lexicalized classifier as a strong baseline 
-  for the task.
-Our classifier implements 6 features:
+Unlike the RTE datasets, our corpus's size supports approaches which make use of rich lexicalized features.
+We evaluate a simple lexicalized classifier to explore the ability of non-specialized models to exploit these features in lieu of more involved language understanding.
+Our classifier implements 6 feature types:
 \begin{enumerate}
@@
 \label{tab:bowresults}
-Accuracy on the our dataset, including precision/recall for the
-  \textit{entailment} class on the test set.
+3-class accuracy on our dataset.
 Ablation results are reported for models lacking cross-bigram features 
@@
   cross-bigram features.
-The latter of these was initially surprising, but could be attributed to
-  the classifier automatically learning explicit negations or adjective
-  modification.
-A similar for bigram features in sentiment analysis was shown in
-  \newcite{sidaw12simple}.
+The latter of these was initially surprising, but can be explained if
+  the classifier automatically learns to recognize explicit negations and adjective
+  modification---such a result was shown in
+  \newcite{sidaw12simple} for bigram features in sentiment analysis.
 
@@
   does without any notion of alignment or tree transformations.
-Although likely richer models would continue to improve performance,
-  the results suggest that given enough data
-  the more noisy part-of-speech overlap constraint can produce an effective
-  model.
+Although we expect that richer models would perform better,
+  the results suggest that given enough data, cross bigrams with the noisy 
+  part-of-speech overlap constraint can produce an effective model.
 
@@
 We test three sentence embedding models, each set to use 100d phrase and sentence embeddings.
-Our baseline sentence embedding simply sums the embeddings of the words in each sentence.
+Our baseline sentence embedding model simply sums the embeddings of the words in each sentence.
 In addition, we experiment with two simple sequence embedding models: a plain RNN and an LSTM RNN \cite{hochreiter1997long}.
@@
 % scr/snlirc3d-snlirc3-only-l0.0001-dim100-ed300-td3-pen200-do0.9-0.9-co3-comp-1-dp1-gc5-lstminit5/stat_log
-% 290500, dev 76.4, mostly converged, but check
+% 290500, dev 76.4, converged
 
-\t{100d RNN}            & \t{73.4} & \t{71.8} \\	
+\t{100d RNN}            & \t{73.1} & \t{72.2} \\	
 % scr/snlirc3d-snlirc3-only-l0.0001-dim100-ed300-td3-pen200-do0.9-0.9-co3-comp3-dp1-gc5-lstminit5/stat_log
-% 307500, dev 71.744,
+% 338500, dev 72.5, mostly converged
 
-\t{100d LSTM RNN}            & \t{84.8} & \b{78.0} \\
+\t{100d LSTM RNN}            & \t{84.8} & \b{77.6} \\
 % scr/snlirc3d-snlirc3-only-l3e-05-dim100-ed300-td3-pen200-do0.95-0.95-co3-comp2-dp1-gc5-lstminit5/stat_log
-% 372000, dev 79.039
+% 372000, dev 79.11, mostly converged
 
@@
 The sum of words model performed slightly worse than the fundamentally similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
-Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it a substantial advantage over the plain RNN, and  resulting in performance that is essentially equivalent to the lexicalized classifier.
+Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it a substantial advantage over the plain RNN, and  resulting in performance that is essentially equivalent to the lexicalized classifier on the test set (LSTM performance near the stopping iteration varies by up to 0.5\% between evaluation steps).
+While the lexicalized model fits the test set perfectly, the gap between train and test set accuracy is relatively small for all three neural network models, suggesting that research into significantly higher capacity versions of these models could be productive.
 
-\subsection{Analysis of performance}
+\subsection{Analysis and discussion}
 
@@
 We were struck by the speed with which the lexicalized classifier outperforms its unlexicalized counterpart.
-With only 100 training examples, the cross-bigram classifier already outperformed its unlexicalized counterpart.
+With only 100 training examples, the cross-bigram classifier is already performing better.
 Empirically, we find that the top weighted features for the classifier
@@
 is the case---then it makes sense that, when these do fire, they
-boost accuracy.
+boost accuracy in identifying entailments.
 
@@
 
-To the extent that successfully training a neural network model like our LSTM on our corpus forces that model to encode broadly accurate representations of English scene descriptions, and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
-In this section, we evaluate on the SICK entailment task using a simple transfer learning method, and achieve competitive results.
-
-To perform transfer, we take the parameters of the LSTM RNN model trained on our corpus and use them to initialize a new model, which is trained from that point only on the training portion of SICK.
-The only newly initialized parameters are the embeddings for words that appear in the SICK training set, but not in our training set, and those parameters are initialized using GloVe embeddings as above.
-We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength, which is re-tuned.
-We additionally transfer the accumulators that are used by AdaDelta to set the learning rates.
-This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
+To the extent that successfully training a neural network model like our LSTM on our corpus forces that model to encode broadly accurate representations of English scene descriptions and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
+In this section, we evaluate on the SICK entailment task using a simple transfer learning method and achieve competitive results.
 
@@
 \toprule
-\textbf{Sentence model} & \b{Train}  & \b{Test}\\
+\textbf{Training sets} & \b{Train}  & \b{Test}\\
 \midrule
-\t{LSTM RNN (random init.)}            & \t{100} & \t{71.3} \\
+\t{Our data only}            & \t{42.0} & \t{46.7} \\
+% scr/transfer5-sick-only-transfer-l0.0001-dim100-ed300-td3-pen200-do0.95-0.95-ws3-adi1-comp2-cdim1/stat_log
+% Step 0, no learning intended after transfer
+\t{SICK only}            & \t{100} & \t{71.3} \\
 % From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
-
-\t{LSTM RNN (transfer init.)}            & \t{99.9} & \b{80.8} \\
+\t{Our data and SICK (transfer)}            & \t{99.9} & \b{80.8} \\
 % From: ~/quant/transfer5-sick-only-transfer-l0.0001-dim100-ed300-td3-pen200-do0.95-0.95-ws3-adi0-comp2-cdim1/stat_log
@@
 
+training sets
+
 \caption{\label{tab:transferresults}
-The performance (\% accuracy) of a 100d LSTM RNN model on SICK under both a random initialization and a transfer initialization copied from a model trained on our corpus.} 
+LSTM 3-class accuracy on the SICK train and test sets under three training regimes.} 
 \end{table}
 
-We compare the performance of an LSTM model with both the transfer initialization and a normal random initialization.
-We tune the regularization parameter separately for each case.
-\Reftab{tab:transferresults} shows these results.
+To perform transfer, we take the parameters of the LSTM RNN model trained on our corpus and use them to initialize a new model, which is trained from that point only on the training portion of SICK.
+The only newly initialized parameters are softmax layer parameters and the embeddings for words that appear in SICK, but not in our corpus (which are populated with GloVe embeddings as above).
+We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength, which is re-tuned.
+We additionally transfer the accumulators that are used by AdaDelta to set the learning rates.
+This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
 
-Our transfer model shows the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which represents a likely performance ceiling.
+The results in \reftab{tab:transferresults} compare LSTM models trained each corpus separately with one trained on both using transfer. We tune the regularization parameter separately for each case.
+
+While training on either corpora alone yields poor performance, transfer yields the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which like represents an approximate performance ceiling.
 The randomly initialized model performs relatively poorly.

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
One more clarification.

@@
 This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
+We tune the regularization parameter separately for each case.
 
-The results in \reftab{tab:transferresults} compare LSTM models trained each corpus separately with one trained on both using transfer. We tune the regularization parameter separately for each case.
-
-While training on either corpora alone yields poor performance, transfer yields the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which like represents an approximate performance ceiling.
-The randomly initialized model performs relatively poorly.
-This suggests that the introduction of a large high-quality corpus makes it newly possible to train representation-learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on inference tasks.
+The results are shown in \reftab{tab:transferresults}.
+Training on SICK alone yields poor performance, and the model trained on our data alone fails dramatically by labeling more \ii{neutral} examples as \ii{contradiction}s than correctly.
+Transferring representations, in contrast, yields the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which likely represents an approximate performance ceiling.
+This suggests that the introduction of a large high-quality corpus makes it  possible to train representation-learning models for sentence meaning that are competitive with the best hand-engineered models on inference tasks.
 

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
A tiny bit more storytelling.

@@
 The results are shown in \reftab{tab:transferresults}.
-Training on SICK alone yields poor performance, and the model trained on our data alone fails dramatically by labeling more \ii{neutral} examples as \ii{contradiction}s than correctly.
+Training on SICK alone yields poor performance, and the model trained on our data misunderstands the subtly different meanings of the (identically named) SICK labels, labeling more \ii{neutral} examples as \ii{contradiction}s than correctly.
 Transferring representations, in contrast, yields the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which likely represents an approximate performance ceiling.
@@
 
-We attempted to apply this same transfer evaluation technique to the RTE-3 challenge, but found that the small size of the training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text used in that corpus, such that neither our transfer model nor our randomly initialized model reached competitive performance.
+We attempted to apply this same transfer evaluation technique to the RTE-3 challenge, but found that the small training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text used in that corpus, such that no training configuration yielded competitive performance.
 Further research on effective transfer learning on small data sets with neural models might facilitate improvements here.

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Added SICK classifier results.

@@
 \toprule
-\textbf{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
+\b{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
 \midrule
@@
 \begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
+\def\t#1{\small{#1}}
+\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceL}c}
 \toprule
-\b{System} & \b{Train} & \b{Test} \\
+\b{System}	 & \multicolumn{2}{c}{\b{Our corpus}} & \b{SICK}\vspace{-0.5mm}\\
+ & \b{Train} & \b{Test} & \b{Test} \vspace{-0.5mm}\\
 \midrule
-\t{Lexicalized}            & \t{99.7}  & \b{78.2} \\ % & \t{78.7}
-\t{Unigrams Only}          & \t{93.1} & \t{71.6} \\ % & \t{72.2}
-\t{Unlexicalized}          & \t{49.4} & \t{50.4} \\ % & \t{50.1} % Last value was 50.39; rounded for consistency.
+\t{Lexicalized}            & \t{99.7}  & \b{78.2} & \b{77.3} \\ % & \t{78.7}
+\t{Unigrams Only}          & \t{93.1} & \t{71.6} & \t{74.0} \\ % & \t{72.2}
+\t{Unlexicalized}          & \t{49.4} & \t{50.4} & \t{69.4}\\ % & \t{50.1} % Last value was 50.39; rounded for consistency.
 \bottomrule

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Making room.

@@
 \label{tab:bowresults}
-3-class accuracy on our dataset.
-Ablation results are reported for models lacking cross-bigram features 
+3-class accuracy, training and testing on either our data or SICK, including models lacking cross-bigram features 
   (Feature \ref{lst:bigram}), and lacking all lexical

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Commented 3-class numbers

--- 2015-06-01 writing/SNLI paper/entailment_eval.tex
@@
 
-% The most immediate application for our corpus is in developing models for the task of NLI.
+The most immediate application for our corpus is in developing models for the task of NLI.
 In particular, since it is dramatically larger than any existing corpus of comparable quality, we expect it to be suitable for training  parameter-rich models like neural networks for this task.
@@
 %  is distributed alongside a number of RTE pipelines.
-We additionally evaluate against a strong but simple classifier-based model with
+We additionally evaluate against a strong classifier-based baseline using
   both unlexicalized and lexicalized features.
@@
 \midrule
-\t{Edit Distance Based}        & \t{71.9} & \t{65.4} & \b{61.9} \\
-\t{Classifier Based}           & \b{72.2} & \b{71.4} & \t{61.5} \\
+\t{Edit Distance Based}        & \t{71.9} & \t{65.4} & \t{61.9} \\
+\t{Classifier Based}           & \t{72.2} & \t{71.4} & \t{61.5} \\
 %\midrule
 %\t{Classifier Based (3-class)} & \t{??.?} & \t{65.6} & \t{} \\
-%\t{Transformation Based} & \t{36.0} & \t{76.7} & \t{56.4} \\
+%\t{Transformation Based} & \t{36.0} & \t{76.7} & \t{56.4} \\  % broken! At least, on SNLI...
 \bottomrule
@@
 \label{tab:eopresults}
-2-class accuracy for two entailment algorithms included in the
+Accuracies of two entailment algorithms included in the
   Excitement Open Platform.
+Note that all accuracies are are on a 2-class
+  classification task; this results in a majority-class baseline
+  of 66\% on our class and 71\% on SICK.
 %All RTE results are 2-class accuracy.
-%The transformation-based model allows for 3-class predictions on our
+%The transformation based model allows for 3-class predictions on our
 %  corpus and SICK; these are reported in the last row.
@@
   of \contradiction\ and \unknown\ are converted to nonentailment.
-This yields a chance baseline accuracy of 66\% on our corpus, and 71\% on SICK.
-We train these models without using any of the included lexical
-  resources, so these results do not represent state-of-the-art
-  RTE systems, but rather give an indication of the relative
+This yields a chance baseline of 66\% on our corpus, and 71\% on SICK.
+
+Note that all results are reported without using any of the included lexical
+  resources.
+We do not attempt to necessarily compare against the state-of-the-art
+  RTE systems as much as give a general sense for the relative
   difficulties of the datasets.
@@
 \paragraph{Lexicalized Classifier}
-Unlike the RTE datasets, our corpus's size supports approaches which make use of rich lexicalized features.
-We evaluate a simple lexicalized classifier to explore the ability of non-specialized models to exploit these features in lieu of more involved language understanding.
-Our classifier implements 6 feature types:
+Unlike the RTE datasets, our corpus's size allows approaches which make use of
+  rich lexicalized features.
+We implement such a lexicalized classifier as a strong baseline 
+  for the task.
+Our classifier implements 6 features:
 \begin{enumerate}
@@
 \label{tab:bowresults}
-3-class accuracy on our dataset.
+Accuracy on the our dataset, including precision/recall for the
+  \textit{entailment} class on the test set.
 Ablation results are reported for models lacking cross-bigram features 
@@
   cross-bigram features.
-The latter of these was initially surprising, but can be explained if
-  the classifier automatically learns to recognize explicit negations and adjective
-  modification---such a result was shown in
-  \newcite{sidaw12simple} for bigram features in sentiment analysis.
+The latter of these was initially surprising, but could be attributed to
+  the classifier automatically learning explicit negations or adjective
+  modification.
+A similar for bigram features in sentiment analysis was shown in
+  \newcite{sidaw12simple}.
 
@@
   does without any notion of alignment or tree transformations.
-Although we expect that richer models would perform better,
-  the results suggest that given enough data, cross bigrams with the noisy 
-  part-of-speech overlap constraint can produce an effective model.
+Although likely richer models would continue to improve performance,
+  the results suggest that given enough data
+  the more noisy part-of-speech overlap constraint can produce an effective
+  model.
 
@@
 %  the EOP systems have been tuned on relatively small corpora
-%  ($\approx$1600 examples), whereas a classifier trained on our corpus can make use of
+%  ($\sim$1600 examples), whereas a classifier trained on our corpus can make use of
 %  over two orders of magnitude more data.
@@
 
-Our corpus is suitably large and diverse to make it possible to train neural network models that produce distributed representations of sentence meaning.
+Our corpus is uniquely suitable for training neural network models that produce distributed representations of sentence meaning.
 In this section, we compare the performance of three such models on the corpus.
@@
 
-Our neural network classifier, depicted in \reffig{modelstructure} (and based on a one-layer model in \citealt{Bowman:Potts:Manning:2014}), is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
+Our neural network classifier, depicted in \reffig{modelstructure}, is simply a stack of three 200d $\tanh$ layers, with the bottom layer taking the concatenated sentence representations as input and the top layer feeding a softmax classifier, all trained jointly with the sentence embedding model itself.
 
 We test three sentence embedding models, each set to use 100d phrase and sentence embeddings.
-Our baseline sentence embedding model simply sums the embeddings of the words in each sentence.
-In addition, we experiment with two simple sequence embedding models: a plain RNN and an LSTM RNN \cite{hochreiter1997long}.
+Our baseline sentence embedding simply sums the embeddings of the words in each sentence.
+In addition, we experiment with two simple sequence embedding models models: a plain RNN and an LSTM RNN \cite{hochreiter1997long}.
 Both use the simple conventional layout with one layer per input token.
 
-The word embeddings for all of the models are initialized with the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine-tuned as part of training.
+The word embeddings for all of the models are initialized with the 300d reference GloVe vectors (840B token version, \citealt{pennington2014glove}) and fine tuned as part of training.
 In addition, all of the models use an additional $\tanh$ neural network layer to map these 300d embeddings into the lower-dimensional phrase and sentence embedding space.
@@
 % scr/snlirc3d-snlirc3-only-l0.0001-dim100-ed300-td3-pen200-do0.9-0.9-co3-comp-1-dp1-gc5-lstminit5/stat_log
-% 290500, dev 76.4, converged
+% 290500, dev 76.4, mostly converged, but check
 
-\t{100d RNN}            & \t{73.1} & \t{72.2} \\	
+\t{100d RNN}            & \t{73.4} & \t{71.8} \\	
 % scr/snlirc3d-snlirc3-only-l0.0001-dim100-ed300-td3-pen200-do0.9-0.9-co3-comp3-dp1-gc5-lstminit5/stat_log
-% 338500, dev 72.5, mostly converged
+% 307500, dev 71.744,
 
-\t{100d LSTM RNN}            & \t{84.8} & \b{77.6} \\
+\t{100d LSTM RNN}            & \t{84.8} & \b{78.0} \\
 % scr/snlirc3d-snlirc3-only-l3e-05-dim100-ed300-td3-pen200-do0.95-0.95-co3-comp2-dp1-gc5-lstminit5/stat_log
-% 372000, dev 79.11, mostly converged
+% 372000, dev 79.039
 
@@
 The sum of words model performed slightly worse than the fundamentally similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
-Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it a substantial advantage over the plain RNN, and  resulting in performance that is essentially equivalent to the lexicalized classifier on the test set (LSTM performance near the stopping iteration varies by up to 0.5\% between evaluation steps).
-While the lexicalized model fits the test set perfectly, the gap between train and test set accuracy is relatively small for all three neural network models, suggesting that research into significantly higher capacity versions of these models could be productive.
+Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it a substantial advantage over the plain RNN, and  resulting in performance that is essentially equivalent to the lexicalized classifier.
 
-\subsection{Analysis and discussion}
+\subsection{Analysis of performance}
 
@@
 A learning curve for the lexicalized and unlexicalized baseline classifiers and the LSTM,
-plotted on a log scale. The y-axis starts at a random-chance accuracy of 33\%. The minibatch size of 64 that we used to tune the LSTM sets a lower bound on data for that model.}
+plotted on a log scale. The y axis starts at a random-chance accuracy of 33\%. The minibatch size of 64 that we used to tune the LSTM sets a lower bound on data for that model.}
 
@@
 We were struck by the speed with which the lexicalized classifier outperforms its unlexicalized counterpart.
-With only 100 training examples, the cross-bigram classifier is already performing better.
+With only 100 training examples, the cross-bigram classifier already outperformed its unlexicalized counterpart.
 Empirically, we find that the top weighted features for the classifier
@@
 is the case---then it makes sense that, when these do fire, they
-boost accuracy in identifying entailments.
+boost accuracy.
 
@@
 \word{sprinter}/\word{runner}. Semantic mistakes at the phrasal level
-(e.g., predicting contradiction for \word{A male is placing an order in a 
+(e.g., predicting contradiction for word{A male is placing an order in a 
 deli}/\word{A man buying a sandwich at a deli}) indicate
@@
 the lexicalized classifier and the LSTM predict \ii{neutral}. 
-In other cases, the models' attempts to shortcut this kind of inference 
+In other cases, models' attempts to shortcut this kind of inference 
 through lexical cues can lead them astray. 
@@
 English \cite{Ward04}. Our annotators generally followed suit, writing
-sentences that, while structurally diverse, share topic/focus (theme/rheme)
-structure with their premises.
-This promotes a coherent, situation-specific construal of each sentence
+sentences with similar topic/focus (theme/rheme) structure that
+promotes a coherent, situation-specific construal of each sentence
 pair. This is information that our models can easily take advantage
@@
 
-To the extent that successfully training a neural network model like our LSTM on our corpus forces that model to encode broadly accurate representations of English scene descriptions and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
-In this section, we evaluate on the SICK entailment task using a simple transfer learning method and achieve competitive results.
+To the extent that successfully training a neural network model like our LSTM on our corpus forces that model to encode broadly accurate representations of English scene descriptions, and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
+In this section, we evaluate on the SICK entailment task using a simple transfer learning method, and achieve competitive results.
+
+To perform transfer, we take the parameters of the LSTM RNN model trained on our corpus and use them to initialize a new model, which is trained from that point only on the training portion of SICK.
+The only newly initialized parameters are the embeddings for words that appear in the SICK training set, but not in our training set, and those parameters are initialized using GloVe embeddings as above.
+We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength, which is re-tuned.
+We additionally transfer the accumulators that are used by AdaDelta to set the learning rates.
+This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
 
@@
 \toprule
-\textbf{Training sets} & \b{Train}  & \b{Test}\\
+\textbf{Sentence model} & \b{Train}  & \b{Test}\\
 \midrule
-\t{Our data only}            & \t{42.0} & \t{46.7} \\
-% scr/transfer5-sick-only-transfer-l0.0001-dim100-ed300-td3-pen200-do0.95-0.95-ws3-adi1-comp2-cdim1/stat_log
-% Step 0, no learning intended after transfer
-\t{SICK only}            & \t{100} & \t{71.3} \\
+\t{LSTM RNN (random init.)}            & \t{100} & \t{71.3} \\
 % From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
-\t{Our data and SICK (transfer)}            & \t{99.9} & \b{80.8} \\
+
+\t{LSTM RNN (transfer init.)}            & \t{99.9} & \b{80.8} \\
 % From: ~/quant/transfer5-sick-only-transfer-l0.0001-dim100-ed300-td3-pen200-do0.95-0.95-ws3-adi0-comp2-cdim1/stat_log
@@
 
-training sets
-
 \caption{\label{tab:transferresults}
-LSTM 3-class accuracy on the SICK train and test sets under three training regimes.} 
+The performance (\% accuracy) of a 100d LSTM RNN model on SICK under both a random initialization and a transfer initialization copied from a model trained on our corpus.} 
 \end{table}
 
-To perform transfer, we take the parameters of the LSTM RNN model trained on our corpus and use them to initialize a new model, which is trained from that point only on the training portion of SICK.
-The only newly initialized parameters are softmax layer parameters and the embeddings for words that appear in SICK, but not in our corpus (which are populated with GloVe embeddings as above).
-We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength, which is re-tuned.
-We additionally transfer the accumulators that are used by AdaDelta to set the learning rates.
-This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
+We compare the performance of an LSTM model with both the transfer initialization and a normal random initialization.
 We tune the regularization parameter separately for each case.
+\Reftab{tab:transferresults} shows these results.
 
-The results are shown in \reftab{tab:transferresults}.
-Training on SICK alone yields poor performance, and the model trained on our data misunderstands the subtly different meanings of the (identically named) SICK labels, labeling more \ii{neutral} examples as \ii{contradiction}s than correctly.
-Transferring representations, in contrast, yields the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which likely represents an approximate performance ceiling.
-This suggests that the introduction of a large high-quality corpus makes it  possible to train representation-learning models for sentence meaning that are competitive with the best hand-engineered models on inference tasks.
+Our transfer model shows the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which represents a likely performance ceiling.
+The randomly initialized model performs relatively poorly.
+This suggests that the introduction of a large high-quality corpus makes it newly possible to train representation-learning models for sentence meaning of a level of quality that is competitive with the best hand-engineered models on inference tasks.
 
-We attempted to apply this same transfer evaluation technique to the RTE-3 challenge, but found that the small training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text used in that corpus, such that no training configuration yielded competitive performance.
+We attempted to apply this same transfer evaluation technique to the RTE-3 challenge, but found that the small size of the training set (800 examples) did not allow the model to adapt to the unfamiliar genre of text used in that corpus, such that neither our transfer model nor our randomly initialized model reached competitive performance.
 Further research on effective transfer learning on small data sets with neural models might facilitate improvements here.

### 2015-06-01 writing/SNLI paper/corpus.tex

@@
 addresses these issues of size and quality, but only for the limited
-fragment of the NLI problem that is captured by paraphase.
+fragment of the NLI problem that is captured by paraphrase.
 
@@
 \item Write one alternate caption that is \textbf{definitely} a \textbf{false} description of the photo.
-\ii{Example: For the caption ``\ii{Two dogs are running through a field.}" you could write ``\ii{The pets are sitting on a couch.}" This is different from the maybe correct category because it's impossible for the dogs to be both running and sitting.}
+\ii{Example: For the caption ``\ii{Two dogs are running through a field.}" you could write ``\ii{The pets are sitting on a couch.}" This is different from the} maybe correct \ii{category because it's impossible for the dogs to be both running and sitting.}
 \end{itemize}
@@
 The instructions were similar to the instructions for initial data collection shown in \reffig{instructions-1}, and linked to a similar FAQ.
-Though we initially used a very restrictive qualification (based on past approval rate) to select workers for the validation task, we nonetheless discovered (and excluded from the corpus) some instances of random guessing in an early batch of work, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
+Though we initially used a very restrictive qualification (based on past approval rate) to select workers for the validation task, we nonetheless discovered (and deleted) some instances of random guessing in an early batch of work, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
 
@@
 are likely to be conservative given our large and
-unevenly distributed pool of workers, but they still provide insights
+unevenly distributed pool of annotators, but they still provide insights
 about the levels of disagreement across the three semantic
 classes. This disagreement likely reflects not just the limitations of
-large crowdsourcing efforts but also the uncertainty inherent in NLI.
+large crowdsourcing efforts but also the uncertainty inherent in naturalistic NLI.
 Regardless, the overall rate of agreement is extremely high,
@@
 \multicolumn{2}{l}{\textbf{Individual annotator label agreement:}}\\
-Annotator label $=$ gold & 89.0\%\\
-Annotator label $=$ author's label & 85.8\%\\
+Individual label $=$ gold label & 89.0\%\\
+Individual label $=$ author's label & 85.8\%\\
 \midrule
@@
 The data will be released upon publication under a CreativeCommons
-Attribution-ShareAlike license, the same licence used for the Flickr30k source captions.
+Attribution-ShareAlike license, the same license used for the Flickr30k source captions.
 
-\paragraph{Partition} We distribute the corpus with a pre-specified train/test/development split.
+\paragraph{Partition}
+
+We distribute the corpus with a pre-specified train/test/development split.
 The test and development sets contain 10k examples each.

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Tabular hacks
Prose changes around new results.

--- 2015-06-01 writing/SNLI paper/entailment_eval.tex
@@
 %\t{Classifier Based (3-class)} & \t{??.?} & \t{65.6} & \t{} \\
-%\t{Transformation Based} & \t{36.0} & \t{76.7} & \t{56.4} \\
+%\t{Transformation Based} & \t{36.0} & \t{76.7} & \t{56.4} \\  % broken! At least, on SNLI...
 \bottomrule
@@
 \toprule
-\b{System}	 & \multicolumn{2}{c}{\b{Our corpus}} & \b{SICK}\vspace{-0.5mm}\\
- & \b{Train} & \b{Test} & \b{Test} \vspace{-0.5mm}\\
+\b{System}	 & \multicolumn{2}{c}{\hspace{-1.2em}\b{Our corpus}} & \b{SICK}\vspace{-0.5mm}\\
+ & \t{Train} & \t{Test} & \t{Test} \vspace{-0.5mm}\\
 \midrule
@@
 % Insights 1: lexicalization helps a bunch
-There is a substantial jump in accuracy from using
-  lexicalized features, and even from using the relatively sparse
+On our large corpus in particular, there is a substantial jump in accuracy from using
+  lexicalized features, and another from using the very sparse
   cross-bigram features.
-The latter of these was initially surprising, but can be explained if
+The latter  result can be explained if
   the classifier automatically learns to recognize explicit negations and adjective
-  modification---such a result was shown in
+  modification, and a similar result was shown in
   \newcite{sidaw12simple} for bigram features in sentiment analysis.

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
More caption munging to fix paper length

@@
 \label{tab:eopresults}
-2-class accuracy for two entailment algorithms included in the
+2-class test accuracy for two systems included in the
   Excitement Open Platform.
@@
 \begin{center}
-\def\t#1{\small{#1}}
 \begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceL}c}
 \toprule
-\b{System}	 & \multicolumn{2}{c}{\hspace{-1.2em}\b{Our corpus}} & \b{SICK}\vspace{-0.5mm}\\
- & \t{Train} & \t{Test} & \t{Test} \vspace{-0.5mm}\\
+\b{System}	 & \multicolumn{2}{c}{\hspace{-1.2em}\b{Our corpus}} & \b{SICK}\\
+ & \t{Train} & \t{Test} & \t{Test}\\
 \midrule
@@
 
-training sets
-
 \caption{\label{tab:transferresults}

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Clarification (I think)

@@
 \label{tab:bowresults}
-3-class accuracy, training and testing on either our data or SICK, including models lacking cross-bigram features 
+3-class accuracy, training on either our data or SICK, including models lacking cross-bigram features 
   (Feature \ref{lst:bigram}), and lacking all lexical

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Last few style edits.

@@
 We report results in \reftab{tab:bowresults}, along with ablation studies for removing
-  the cross-bigram features (leaving only the cross-unigram feature),
+  the cross-bigram features (leaving only the cross-unigram feature)
   and for removing all lexicalized features.
@@
 The results are shown in \reftab{tab:transferresults}.
-Training on SICK alone yields poor performance, and the model trained on our data misunderstands the subtly different meanings of the (identically named) SICK labels, labeling more \ii{neutral} examples as \ii{contradiction}s than correctly.
+Training on SICK alone yields poor performance, and the model trained on our data alone misunderstands the subtly different meanings of the (identically named) SICK labels, labeling more \ii{neutral} examples as \ii{contradiction}s than correctly.
 Transferring representations, in contrast, yields the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which likely represents an approximate performance ceiling.

### 2015-06-01 writing/SNLI paper/corpus.tex
Minor changes

@@
 \Reftab{snli-examples} shows a set of randomly chosen validated examples from the development set with their labels.
-Qualitatively, we find the data that we collected draws fairly extensively on commonsense knowledge, and that hypothesis and premise sentences often differ structurally in significant ways, suggesting that superficial word alignment models are unlikely to be able to succeed at classification.
+Qualitatively, we find the data that we collected draws fairly extensively on commonsense knowledge, and that hypothesis and premise sentences often differ structurally in significant ways, suggesting that there is room for improvement beyond superficial word alignment models.
 We also find the sentences that we collected to be largely fluent, correctly spelled English, with a mix of full sentences and caption-style noun phrase fragments, though punctuation and capitalization are often omitted.

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Minor changes

--- 2015-06-01 writing/SNLI paper/entailment_eval.tex
@@
   evaluation sets.
-We evaluate against three algorithms included in the distribution,
+We evaluate against two algorithms included in the distribution,
   and also evaluated in \newcite{magnini2014excitement}.
 These are: 
-  (1) a simple edit-distance based algorithm; and
-  (2) a classifier-based algorithm.
+  a simple edit-distance based algorithm; and
+  a classifier-based algorithm.
 %  (3) an algorithm based around tree transformations.
@@
 \label{tab:eopresults}
-2-class test accuracy for two systems included in the
+2-class accuracy for two entailment algorithms included in the
   Excitement Open Platform.
@@
 \begin{center}
+\def\t#1{\small{#1}}
 \begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceL}c}
 \toprule
-\b{System}	 & \multicolumn{2}{c}{\hspace{-1.2em}\b{Our corpus}} & \b{SICK}\\
- & \t{Train} & \t{Test} & \t{Test}\\
+\b{System}	 & \multicolumn{2}{c}{\hspace{-1.2em}\b{Our corpus}} & \b{SICK}\vspace{-0.5mm}\\
+ & \t{Train} & \t{Test} & \t{Test} \vspace{-0.5mm}\\
 \midrule
@@
 % Insights 1: lexicalization helps a bunch
-On our large corpus in particular, there is a substantial jump in accuracy from using
-  lexicalized features, and another from using the very sparse
+There is a substantial jump in accuracy from using
+  lexicalized features, and even from using the relatively sparse
   cross-bigram features.
-The latter  result can be explained if
+The latter of these was initially surprising, but can be explained if
   the classifier automatically learns to recognize explicit negations and adjective
-  modification, and a similar result was shown in
+  modification---such a result was shown in
   \newcite{sidaw12simple} for bigram features in sentiment analysis.
@@
 
+training sets
+
 \caption{\label{tab:transferresults}

### 2015-06-01 writing/SNLI paper/entailment_eval.tex
Gah; another typo

@@
   evaluation sets.
-We evaluate against two algorithms included in the distribution,
-  and also evaluated in \newcite{magnini2014excitement}.
-These are: 
-  a simple edit-distance based algorithm; and
+We evaluate against two algorithms included in the distribution:
+%  and also evaluated in \newcite{magnini2014excitement}.
+%These are: 
+  a simple edit-distance based algorithm and
   a classifier-based algorithm.

### 2015-06-10 writing/F'14 paper/cameraready/join.tex

--- 2015-05-03 writing/F'14 paper/join.tex
@@
 \paragraph{Results} 
-The resuts (Table \ref{joinresultstable}) show that NTN is able to accurately encode the relations between the terms in the geometric relations between their vectors, 
+The results (Table \ref{joinresultstable}) show that NTN is able to accurately encode the relations between the terms in the geometric relations between their vectors, 
 and is able to then use that information to recover relations that 

### 2015-06-10 writing/F'14 paper/cameraready/methods.tex

--- 2015-02-28 writing/F'14 paper/methods.tex
@@
 them by an $n \times 2n$ matrix of learned weights, and adds a bias $\vec{b}$. 
-The TreeRNTN adds a learned third-order tensor 
+The TreeRNTN adds a learned full rank third-order tensor 
 $\mathbf{T}$, of dimension $n \times n \times n$, modeling
@@
 Other strong tree-structured models have been proposed in past work
-\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, but
+\cite{sochergrounded,irsoydeep,tai2015improved}, but
 we believe that these two provide a valuable case study, and that positive results on 
@@
 up the tree and into the classifier. For an objective
-function, we use the negative log of the probability assigned to the
-correct label with L2 regularization.
+function, we use the negative log likelihood of the
+correct label with tuned L2 regularization.
 

### 2015-06-10 writing/F'14 paper/cameraready/quantifiers.tex

--- 2015-05-18 writing/F'14 paper/arxiv/quantifiers.tex
@@
 
-In each run, we randomly partition the set of valid \textit{single sentences} under the grammar into training and test, and then label all of the pairs from within each set to generate a training set of 27k pairs and a test set of 7k pairs.
-Because the model doesn't see the test sentences at training time, it cannot directly use the kind of reasoning described in \S\ref{sec:join}, and must instead both infer the word-level relations and learn a complete reasoning system over them for our logic.
+In each run, we randomly partition the set of valid \textit{single sentences} into train and test, and then label all of the pairs from within each set to generate a training set of 27k pairs and a test set of 7k pairs.
+Because the model doesn't see the test sentences at training time, it cannot directly use the kind of reasoning described in \S\ref{sec:join} (treating sentences as unanalyzed symbols), and must instead infer the word-level relations and learn a complete reasoning system over them for our logic.
 

### 2015-06-10 writing/F'14 paper/cameraready/recursion.tex

--- 2015-05-03 writing/F'14 paper/recursion.tex
@@
 operators are truth-functional ones.  Table~\ref{tab:pl} defines this
-logic, and Table~\ref{tab:plexs} gives some short examples relational statements from our data.
+logic, and Table~\ref{tab:plexs} gives some short examples of relational statements from our data.
  To compute these relations
@@
 each operator can be represented accurately by a single bit.
-In the much more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.}
+In the much more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to {\tiny $2^{2^{6}}$} distinct conditions on valuations.}
 

### 2015-06-10 writing/F'14 paper/cameraready/sick.tex

--- 2015-05-03 writing/F'14 paper/arxiv/sick.tex
@@
 
-The results (Table \ref{examplesofsickdata}) show that the TreeRNTN performs especially strongly in the two categories which pick out specific syntactic configurations, \textsc{Passive} and \textsc{Neg}, suggesting that that model has learned to encode the relevant structures well.
+The results (Table \ref{sresultstable}) show that the TreeRNTN performs especially strongly in the two categories which pick out specific syntactic configurations, \textsc{Passive} and \textsc{Neg}, suggesting that that model has learned to encode the relevant structures well.
 It also performs fairly on \textsc{Subst}, which most closely parallels the lexical entailment inferences addressed in \S\ref{sec:quantifiers}.

### 2015-06-10 writing/F'14 paper/cameraready/abstract.tex

--- 2015-05-03 writing/F'14 paper/arxiv/abstract.tex
@@
   tensor networks (TreeRNTNs)---can correctly learn to identify logical
-  labelships such as entailment and contradiction using these representations. 
+  relationships such as entailment and contradiction using these representations. 
   In our first set of experiments, we generate artificial data from a
   logical grammar and use it to evaluate the models' ability to learn
-  to handle basic labelal reasoning, recursive structures, and
+  to handle basic relational reasoning, recursive structures, and
   quantification. We then evaluate the models on the more natural

### 2015-06-13 writing/S'15 LSTM/abstract.tex
Paper push.

\begin{abstract}
Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but these models largely have not outperformed simpler sequence-based models by substantial margins. We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured are built around, mitigating the advantage of the tree-structured models. We investigate this possibility by evaluating both models on an artificial talk for which recursive compositional structure is crucial, and find that the sequence model is able to identify and exploit the underlying structure, though that it is less efficient at learning than the tree-structured models, only succeeding after exposure to a larger and richer set of training data.
\end{abstract}

### 2015-06-13 writing/S'15 LSTM/data.tex
Paper push.

--- 2015-05-03 writing/F'14 paper/arxiv/intro.tex
@@
-\section{Introduction}\label{sec:intro}
+\subsection{Natural language inference as an ev}
+
+borrow data
+
+framing is entailment - true NLI has nice property, this generalizes that
+
+data is from simulated language meant to highlight recursion
+
+\paragraph{The task} The
+
+\todo{Rework into a 1-2 paragraph introduction to NLI and NatLog.}
 
@@
 
-\begin{table*}[tp]
+\begin{table}[tp]
   \centering\small
-  \setlength{\tabcolsep}{15pt}
   \renewcommand{\arraystretch}{1}
@@
     \toprule
-    Name & Symbol & Set-theoretic definition & Example \\ 
+    Name & Symb. & Set-theoretic definition \\ 
     \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
-    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
-    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
-    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
-    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
-    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
+strict entailment         & $\natfor$   & $x \subset y$  \\ 
+    strict rev. entailment & $\natrev$   & $x \supset y$  \\ 
+    equivalence        & $\nateq$    & $x = y$   \\ 
+    alternation        & $\natalt$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ \\ 
+    negation           & $\natneg$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$   \\
+    cover              & $\natcov$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ \\ 
+    independence       & $\natind$   & (else)\\
     \bottomrule
   \end{tabular}
-  \caption{\label{b-table}The seven relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.} %-%
+  \caption{\label{b-table}The seven relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.
+The relations are defined so as to be mutually exclusive.
+(adapted from \protect\citealt{Bowman:Potts:Manning:2014})} %-%
 
-\end{table*}
+\end{table}
+
+A successful natural language inference system must reason 
+about relations not just over familiar
+atomic symbols, but also over novel structures built up 
+recursively from these symbols. 
+This section shows that our models can learn a 
+compositional semantics over such structures.
+In our evaluations, we exploit the fact that our logical language
+is infinite by testing on strings that are longer and more complex
+than any seen in training.
+
+% Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice
+%   said hello}, and \ii{Carl thinks that Bob said that Alice said
+%   hello}. Overt recursion of this kind is easy to find, and
+% theoretical accounts of natural language syntax and semantics rely
+% heavily on recursive structures. In order for a model to be able to
+% accurately learn natural language meanings, then, we expect that it
+% would need to be able to learn to represent the meanings of function
+% words in a such a way that they are able to behave correctly when
+% taking their own outputs as input. In evaluating the model, we take
+% advantage of the fact that recursive structures of this kind define
+% potentially infinite languages by testing the model on strings that
+% are longer and more complex than any seen in testing.
+
+\paragraph{Experiments}
+As in \S\ref{sec:join}, we generate artificial data from a formal system,
+ but we now replace the unanalyzed symbols
+from that experiment with complex formulae. These formulae
+represent a complete classical propositional logic:
+each atomic symbol is a variable over the domain \{$\True$, $\False$\}, and the only
+operators are truth-functional ones.  Table~\ref{tab:pl} defines this
+logic, and Table~\ref{tab:plexs} gives some short examples of relational statements from our data.
+ To compute these relations
+between statements, we exhaustively enumerate the sets of assignments
+of truth values to propositional variables that would satisfy each of
+the statements, and then we convert the set-theoretic relation between
+those assignments into one of the seven relations in
+Table~\ref{b-table}. As a result, each relational statement represents
+a valid theorem of the propositional logic, and to succeed, the models 
+must learn to reproduce the behavior of a theorem prover.\footnote{
+Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector TreeRNN
+model somewhat similar to our TreeRNTN can learn boolean logic, 
+a logic where the atomic symbols are simply the
+values $\True$ and $\False$. While learning the operators of that logic is not trivial, the outputs of
+each operator can be represented accurately by a single bit.
+In the much more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to {\tiny $2^{2^{6}}$} distinct conditions on valuations.}
+
+\begin{table}[tp]
+  \centering\small
+%  \begin{subtable}[t]{0.45\textwidth}
+%    \centering
+%    \begin{tabular}[t]{l l}
+%      \toprule
+%      Formula     & Interpretation \\
+%      \midrule
+%      $p_1$, $p_2$, $p_3$, $p_4$, $p_5$, $p_6$ & $\sem{x} \in \{\True, \False\}$ \\
+%      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
+%      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
+%      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
+%      \bottomrule
+%    \end{tabular}    
+%    \caption{Well-formed formulae. $\varphi$ and $\psi$
+%      range over all well-formed formulae, and $\sem{\cdot}$ is
+%      the interpretation function mapping formulae into $\{\True,
+%      \False\}$.}\label{tab:pl}
+%  \end{subtable}
+    \begin{tabular}[t]{r c l}
+      \toprule
+      $\plneg p_3$        & $\natneg$ & $p_3$ \\
+      $\plneg \plneg p_6$ & $\nateq$  & $p_6$ \\
+      $p_3$               & $\natfor$ & $(p_3 \plor p_2)$ \\
+      $(p_1 \plor (p_2 \plor p_4))$               & $\natrev$ & $(p_2 \pland  \plneg p_4)$ \\
+      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
+      $\plneg\, (\plneg p_1 \pland \plneg p_2)$ & $\nateq$ & $(p_1 \plor p_2)$ \\ 
+      $(p_3 \pland \plneg p_1 ) \plor \plneg p_3$    & $\natrev$& $\plneg\, (p_3 \plor p_2)$ \\
+      %<	( not ( c ( or b ) ) )	( ( c ( and ( not a ) ) ) ( or ( not c ) ) )
+      \bottomrule
+    \end{tabular}
+    \caption{Examples of short expressions from the artificial data introduced by \protect\citealt{Bowman:Potts:Manning:2014}.}\label{tab:plexs}
+\end{table}
+
+\todo{Note about extra parens}
+
+In our experiments, we randomly generate unique pairs 
+of formulae containing up to 12 instances of logical operators each and compute the relation that holds for each pair.
+We discard pairs in which either statement is either a tautology or a
+contradiction, for which the seven relations in
+Table~\ref{b-table} are undefined. The resulting set of formula pairs is
+then partitioned into 12 bins according the number of operators in
+the larger of the two formulae. We then sample 20\% of each
+bin for a held-out test set.
+If we do not implement any constraint that the two statements being
+compared are similar in any way, then the generated data are dominated
+by statements in which the two formulae refer to largely separate
+subsets of the six variables, which means that the $\natind$ relation
+is almost always correct.  In an effort to balance the distribution of
+relation labels without departing from the basic task of modeling
+propositional logic, we disallow individual pairs of statements from
+referring to more than four of the six propositional variables.
+
+In order to test the model's generalization to unseen structures, we discard
+training examples with more than 4 logical operators, yielding 60k short training examples,
+and 21k test examples across all 12 bins.
+In addition to the two tree models, we also train a summing NN baseline
+which is largely identical to the TreeRNN, except that instead of using a learned composition function,
+it simply sums the term vectors in each expression to compose them before passing them to the comparison layer. Unlike the two tree models, this baseline does not use word order,
+and is as such guaranteed to ignore some information that it would need in order to succeed perfectly.
+
+\paragraph{Model tuning} We stipulate dimensionalities of 50 and 100 for the sentence embeddings and the classifier layers, respectively. We initialize

### 2015-06-13 writing/S'15 LSTM/figure1.tex
Paper push.

\begin{subfigure}[t]{0.45\textwidth}
\centering
\scalebox{0.75}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{29pt}

    \tikzstyle{label}=[text width=40mm,align=center]    
    \tikzstyle{softmax}=[fill=red!40,text width=40mm,align=center]
    \tikzstyle{preclass}=[fill=orange!40,text width=50mm,align=center]
    \tikzstyle{e}=[fill=green!40,text width=26mm,align=center]
    \tikzstyle{m}=[draw=black,text width=38mm,align=center]

    \node[softmax]  (softmax) at (0*\dx,6*\dy) {7-way softmax classifier};
    \node[preclass]  (pc3) at (0*\dx,5*\dy) {100d $\tanh$ NN layer};
    \node[preclass]  (pc2) at (0*\dx,4*\dy) {100d $\tanh$ NN layer};
    \node[preclass,fill=yellow!40]  (pc1) at (0*\dx,3*\dy) {100d $\tanh$ NTN layer};
    \node[e]  (pe) at (-3*\dx,1.75*\dy) {50d premise};
    \node[e]  (he) at (3*\dx,1.75*\dy) {50d hypothesis};
    \node[m]  (pem) at (-3*\dx,0.5*\dy) {sentence model\\ with premise input};
    \node[m]  (hem) at (3*\dx,0.5*\dy) {sentence model\\ with hypothesis input};

    \pgfsetarrowsend{latex}
    \tikzstyle{fwd} = [draw=black, line width=1pt]

          \draw [fwd] (pc3) -- (softmax);
          \draw [fwd] (pc2) -- (pc3);
          \draw [fwd] (pc1) -- (pc2);
          \draw [fwd] (pe) -- (pc1);
          \draw [fwd] (he) -- (pc1);
          \draw [fwd] (hem) -- (he);
          \draw [fwd] (pem) -- (pe);

  \end{tikzpicture}}

 \caption{The general architecture shared across models.}\label{fig:model:top}

\end{subfigure}

\begin{subfigure}[t]{0.45\textwidth}
  \centering
\scalebox{0.75}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{29pt}

    \tikzstyle{word}=[fill=purple!40,text width=16mm,text height=2mm,align=center]
    \tikzstyle{node}=[fill=blue!40,text width=16mm,text height=2mm,align=center]
    \tikzstyle{empty}=[fill=blue!0,text width=8mm,text height=2mm,align=center]

    \node[empty]  (null) at (0*\dx,7*\dy) {...};
    \node[node]  (aorb) at (0*\dx,6*\dy) {a or b};
    \node[word]  (a) at (-2.5*\dx,5*\dy) {a};
    \node[node]  (orb) at (2.5*\dx,5*\dy) {or b};
    \node[word]  (or) at (0*\dx,4*\dy) {or};
    \node[word]  (b) at (5*\dx,4*\dy) {b};

    \pgfsetarrowsend{latex}
    \tikzstyle{fwd} = [draw=black, line width=1pt]

          \draw [fwd] (or) -- (orb);
          \draw [fwd] (b) -- (orb);
          \draw [fwd] (a) -- (aorb);
          \draw [fwd] (orb) -- (aorb);
          \draw [fwd] (aorb) -- (null);
  \end{tikzpicture}}

     \caption{The architecture for the TreeRNN and TreeRNTN sentence models. Terminal nodes are learned embeddings and nonterminal nodes are either NN or NTN layers with $\tanh$ nonlinearities.}\label{fig:model:tree}

  \end{subfigure}
 \vspace{0.5cm}

\begin{subfigure}[t]{0.45\textwidth}
  \centering
\scalebox{0.75}{
 \begin{tikzpicture}
    \def\dx{35pt}
    \def\dy{29pt}

    \tikzstyle{word}=[fill=purple!40,text width=16mm,text height=2mm,align=center]
    \tikzstyle{node}=[fill=blue!40,text width=16mm,text height=2mm,align=center]
    \tikzstyle{empty}=[fill=blue!0,text width=8mm,text height=2mm,align=center]

    \node[word]  (a) at (-3*\dx,1*\dy) {a};
    \node[node]  (aN) at (-3*\dx,2*\dy) {a};

    \node[word]  (or) at (-1*\dx,1*\dy) {or};
    \node[node]  (orN) at (-1*\dx,2*\dy) {a or};

    \node[word]  (b) at (1*\dx,1*\dy) {b};
    \node[node]  (bN) at (1*\dx,2*\dy) {a or b};

    \node[empty]  (nullN) at (2.75*\dx,2*\dy) {...};

    \pgfsetarrowsend{latex}
    \tikzstyle{fwd} = [draw=black, line width=1pt]

          \draw [fwd] (or) -- (orN);
          \draw [fwd] (b) -- (bN);
          \draw [fwd] (a) -- (aN);
          \draw [fwd] (aN) -- (orN);
          \draw [fwd] (orN) -- (bN);
          \draw [fwd] (bN) -- (nullN);

  \end{tikzpicture}}

   \caption{The architecture for the LSTM sentence model. Nodes in the lower row are learned embeddings and nodes on the upper row are LSTM units.}\label{fig:model:seq}

    \end{subfigure}

### 2015-06-13 writing/S'15 LSTM/methods.tex
Paper push.

\section{Sentence embedding models for NLI} \label{methods}
\todo{Rewrite in a shorter form, add LSTMs}

\begin{figure}[t!]
  \centering
  \input{figure1}
  \caption{In our model, two separate sentence models---based on either tree (b) or sequence (c) models---build up vector representations for each of two sentences. A multilayer classifier component (a) then uses the resulting vectors to predict a label that reflects the logical relationship between the two sentences.} 
  \label{sample-figure}
\end{figure}

We use the model architecture depicted in Figure \ref{fig:model:top}. This approach is loosely based on the model used in the experiments by \newcite{Bowman:Potts:Manning:2014}. The model architecture uses two copies of a single sentence embedding network---a TreeRNN, TreeRNTN, or LSTM RNN, depending on the precise experiment---to embed the premise and hypothesis sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations. Since the embeddings are computed separately, the embedding models are forced encode every element of the sentence meaning that the downstream model will need. Though this prevents us from exploiting cross-sentence alignment information that could allow for the learning of more effective entailment models, it allows us to focus our evaluation on models that capture the meanings of single sentences.

\paragraph{Classifier}
The classifier component of the model consists of a combinging layer which takes the two sentence representations as inputs, followed by two neural network layers, then a softmax classifier.
For the combining layer, we use a (recursive) neural tensor network ((R)NTN, \citealt{chen2013learning}) layer, which sums the output of a plain recursive/recurrent neural network layer with a vector computed using two multiplications with a learned (full rank) third-order tensor parameter:
\begin{gather} 
\label{TreeRNN}
\vec{y}_{\textit{RNN}} = \tanh(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
\label{TreeRNTN} 
\vec{y}_{\textit{RNTN}} = \vec{y}_{\textit{RNN}} + \tanh(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather} 
Our model is largely identical to the Bowman et al. model, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence embedding models are compared in as similar a setting as possible.

\paragraph{Sentence embedding}
The sentence embedding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector. We experiment with two tree-structured models (Figure \ref{fig:model:tree}), one with an RNTN activation function or a plain RNN activation function. Both are structured using the parses that are provided with the data. In addition, we use a sequence model (Figure \ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} with the precise formulation from \newcite{zaremba2015recurrent}.

We additionally experimented with a plain RNN---the sequence equivalent of the TreeRNN---but were not able to coerce that weaker model into learning to model even the training data well, so they are omitted below. LSTM, in contrast are the most widely used of a class of models that attempt to avoid the extreme sensitivity to initialization and gradient magnitude that make RNN training so unreliable.

\paragraph{Training} We randomly initialize all embeddings and layer parameters, and then train all parameters using minibatch stochastic gradient descent with AdaDelta \cite{zeiler2012adadelta} learning rates. Our objective is the standard negative log likelihood classification objective with L2 regularization. All models were stopped at 100 epochs (TODO), it which point all had largely converged without significantly declining from their peak performances.

### 2015-06-13 writing/S'15 LSTM/recursion.tex
Paper push.

\section{The task}\label{sec:recursion}
\paragraph{Reasoning about entailment} 
The data that we use define a version of the recognizing textual entailment task, in which the goal is to determine what kind of logical consequence relation holds between two sentences, drawing on a small fixed vocabulary of relations such as entailment, contradiction, or synonymy. This task is well suited to evaluating neural network models for sentence interpretation: models must develop comprehensive represenations of the meanings of each sentence to do well at the task, but the data does not force these representations to take a specific form, allowing the model to learn whatever kind of representations it can use most effectively.

The data we use are labeled with the seven relation system of \newcite{maccartney2009extended}, depicted in Table \ref{b-table}.

\begin{table}[tp]
  \centering\small
  \renewcommand{\arraystretch}{1}
  \begin{tabular}{l c l l} 
    \toprule
    Name & Symb. & Set-theoretic definition \\ 
    \midrule
strict entailment         & $\natfor$   & $x \subset y$  \\ 
    strict rev. entailment & $\natrev$   & $x \supset y$  \\ 
    equivalence        & $\nateq$    & $x = y$   \\ 
    alternation        & $\natalt$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ \\ 
    negation           & $\natneg$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$   \\
    cover              & $\natcov$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ \\ 
    independence       & $\natind$   & (else)\\
    \bottomrule
  \end{tabular}
  \caption{\label{b-table}The seven relations are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences. The relations are defined so as to be mutually exclusive. (table adapted from \protect\citealt{Bowman:Potts:Manning:2014})} %-%
\end{table}

\paragraph{The artificial language} The language described in Bowman et al. (\S4) is extremely impoverished: its vocabulary consists only of six unanalyzed word types ($p_1, p_2, p_3, p_4, p_5, p_6$), \word{and}, \word{or}, and \word{not}. Sentences of the language can be straightforwardly interpreted as statements of propositional logic (where the six unanalyzed words are variables), and labeled sentence pairs can be interpreted as theorems of that logic. Some example pairs are provided in Table \ref{tab:plexs}.

Crucially, the language is defined such that any sentence can be embedded under negation or conjunction to create a new sentence, allowing for arbitrary-depth recursion, and the scope of negation and conjunction are determined only by bracketing with parentheses (rather than bare word order). The compositional structure of each sentence can thus be an arbitrary tree structure, and interpreting a sentence correctly requires using that structure.

The data comes with parentheses representing complete binary bracketing. Our two models use this information differently. For the tree structured model, we assumble the tree structure for each sentence according to the parentheses, but do not treat them as word tokens. For the sequence model, we treat the parentheses as words, and learn their embeddings.

\begin{table}[tp]
  \centering\small
%  \begin{subtable}[t]{0.45\textwidth}
%    \centering
%    \begin{tabular}[t]{l l}
%      \toprule
%      Formula     & Interpretation \\
%      \midrule
%      $p_1$, $p_2$, $p_3$, $p_4$, $p_5$, $p_6$ & $\sem{x} \in \{\True, \False\}$ \\
%      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
%      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
%      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
%      \bottomrule
%    \end{tabular}    
%    \caption{Well-formed formulae. $\varphi$ and $\psi$
%      range over all well-formed formulae, and $\sem{\cdot}$ is
%      the interpretation function mapping formulae into $\{\True,
%      \False\}$.}\label{tab:pl}
%  \end{subtable}
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg p_3$        & $\natneg$ & $p_3$ \\
      $\plneg \plneg p_6$ & $\nateq$  & $p_6$ \\
      $p_3$               & $\natfor$ & $(p_3 \plor p_2)$ \\
      $(p_1 \plor (p_2 \plor p_4))$               & $\natrev$ & $(p_2 \pland  \plneg p_4)$ \\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg\, (\plneg p_1 \pland \plneg p_2)$ & $\nateq$ & $(p_1 \plor p_2)$ \\ 
      $(p_3 \pland \plneg p_1 ) \plor \plneg p_3$    & $\natrev$& $\plneg\, (p_3 \plor p_2)$ \\
      %<	( not ( c ( or b ) ) )	( ( c ( and ( not a ) ) ) ( or ( not c ) ) )
      \bottomrule
    \end{tabular}
    \caption{Examples of short expressions from the artificial data introduced by \protect\citealt{Bowman:Potts:Manning:2014}. For brevity, we only show the parentheses that are needed to disambiguate the sentences rather than the full binary bracketings.}\label{tab:plexs}
\end{table}

\paragraph{Corpus size and structure}
The sentence pairs in the corpus are divided into thirteen bins according to the number of logical connectives (\word{and, or, not}) in the longer of the two sentences in the pair (the data contain no sentences with more than 12 connectives). We test the model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how its performance depends on the complexity of the sentences. In three experiments, we train our models on the training portions of bins 0--3 (62k examples), bins 0--4 (90k), and bins 0--6 (160k). Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures. If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, the models will inevitably will start to make errors on sentences above a some length, and the (necessary) use of L2 regularization exacerbates this for sentences which are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-06-13 config/CompositionSetup.m

@@ -9,7 +9,7 @@ if composition == -1
 	hyperParams.useSumming = 1;
 elseif composition < 2
 	hyperParams.useThirdOrderComposition = composition;
-	hyperParams.useThirdOrderMerge = 0;
+	hyperParams.useThirdOrderMerge = 1;
 elseif composition == 2
 	hyperParams.lstm = 1;
 	hyperParams.useTrees = 0;
@@ -45,7 +45,7 @@ elseif composition == 7
 	hyperParams.useTrees = 1;
 	hyperParams.eyeScale = 0;
 	hyperParams.useThirdOrderComposition = 0;
-	hyperParams.useThirdOrderMerge = 0;
+	hyperParams.useThirdOrderMerge = 1;
 end
 
 end

### 2015-06-13 layer-fns/ComputeInnerTensorLayer.m

@@ -2,8 +2,8 @@
 function innerTensorLayerOutput = ComputeInnerTensorLayer(l, r, matrices, matrix)
 % Run an NTN layer as in forward propagation, not including the nonlinearity.
 
-outDim = size(matrix, 1)
-B = size(l, 2)
+outDim = size(matrix, 1);
+B = size(l, 2);
 
 % Apply third-order tensor
 % NOTE: Sadly, there doesn't seem to be a good MATLAB primitive for this.

### 2015-06-13 writing/S'15 LSTM/discussion.tex
Paper push.

\section{Results and discussion}\label{sec:discussion}

\begin{figure*}[t!]
  \centering
\begin{subfigure}[t]{0.29\textwidth}
  \includegraphics[height=1.6in]{fig3.pdf}
  \caption{Training expression size $\le$3.}
  \end{subfigure}~~
\begin{subfigure}[t]{0.29\textwidth}
    \includegraphics[height=1.6in]{fig4.pdf}
  \caption{Training expression size $\le$4.}
  \end{subfigure}~~
\begin{subfigure}[t]{0.4\textwidth}
      \includegraphics[height=1.6in]{fig6.pdf}
  \caption{Training expression size $\le$6.}
\end{subfigure}
  \caption{Results on three experiments with increasingly rich training sets. The horizontal axis on each graph divides the test set data points (expression pairs) into bins by the number of logical operators in the more complex of the two expressions in the pair.}
  \label{prop-results} 
\end{figure*}

The results are shown in Figure \ref{prop-results}. All three models were able to largely memorize the training data (TODO, TODO, and TODO for $\le$6). With sufficient training data (in the $\le$6 setting), the two tree models were able to perform at above 99\% (TODO) on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets.

The LSTM performs fairly poorly in the $\le3$ setting, with performance at 3 a mediorce TODO\%, and an abrubt drop to TODO\% at 4, suggesting that the model did not acquire the needed generalizations. With more ample training data in the $\le6$ condition, though, the LSTM shows fairly good accuracy on short examples, and a smooth decay with no abrupt drop.

\paragraph{Conclusion}

We find that

### 2015-06-13 writing/S'15 LSTM/intro.tex
Paper push.

\section{Introduction}\label{sec:intro}

Neural network models that encode sentences as vectors have been successfully used on a wide array of NLP tasks, including \todo{citations}. Many of the most successful of these models fall into one of two categories: sequence models based on recurrent neural networks build representations incrementally from left to right *CITE, while tree-structured models based on \ii{recursive} neural networks *CITE build representations incrementally according the hierarchical sturcture of linguistic phrases.

Tree-structured models are presented as the more principled choice, since they directly exploit the principle of compositionality---fundamental to linguistic semantic theory \cite{Partee84,Janssen97}---that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.

While both classes of model can perform objectively well on many tasks, and both are under active development, tree-structured models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect. Even when the engineering hurdles around building tree-structured models are dealt with, head-to-head comparisons with sequence models show either modest improvements \todo{Cite TreeLSTM paper, and ...} or none at all \todo{cite Jiwei}.

\todo{We use focused artificial data set to effectively demonstrate...}
We argue that this question is best studied through focused artificial data experiments, rather than through the inspection of existing learned models. There is no theory of distributed representation that is sufficiently detailed to allow us to look for the presence of tree-structured representations in model parameters or activations, and performing in-domain error analysis on trained models is unlikely to provide a sufficiently clear picture. An artificial data experiment, in contrast, allows us to single out the behavior that we want to see and evaluate a model's ability to learn it under ideal circumstances with ample clean data, such that if the is able to perform that behavior in any setting, it will do so in our experiment.

In particular, we aim to single out the behavior of learning to interpret sentences that use recursive structure. We specifically wish to test the recursion aspect of interpretation by training the model on sentences whose length and recursion depth are limited, and testing it on longer and more complex sentences, such that only a model that recognizes and exploits the recursive structure will be able to succeed. An existing experimental design meets this criterion nicely: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two types of tree-structured neural network to learn to interpret (by supplying entailment judgments) statements of a simple propositional logic-based language, with a focus on evaluating the models on expressions more complex than those used in training. We propose a way to adapt the experiment presented in that work to sequence models and use it to compare an LSTM-based model with the tree-structured models in that paper, and we extend the experiment by testing multiple different sizes of training set in order to better understand how efficient each model is at learning the underlying generalizations.

As in \newcite{Bowman:Potts:Manning:2014}, we find that both TreeRNNs and TreeRNTNs are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size. Perhaps unsurprisingly, we also find that extending the training set to include larger structures mitigates this decay. More importantly, though, we find that a sequence model centered on a single-layer LSTM RNN is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree-structured models. We take these results as evidence that sequence models are capable of learning tree-structured representations when trained on a task that demands them. \todo{higher level conclusions based on the two interpretations from above}

\subsection{Related work}

TODO CITE JIWEI performs head-to-head comparisons of tree and sequence models on five NLP tasks over small to moderately-sized data sets and find that tree-structured models robustly outperform sequences on only one, concluding that sequence models are largely adequate for NLP. This paper aims to explore that result, specifically focusing on the question of whether compositional structure (implicit or explicit) can be dismissed as unnecessary in research on representation learning for standard NLP tasks.

Recent results on parsing by \todo{Cite Grammar as a Foreign Lg, Dyer} hints the outcome that we find. The simplest model presented in TODO CITE GaaFL encodes each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) for the sentence using only the information present in the vector. This suggests that the intermediate vector representation contains a tree-structured representation, and that that representation was constructed incrementally by the sequence model. However, the massive size of the training set that was necessary to train that model---250M tokens---makes an alternate explanation possible: the model could have primarily learned to simply recognize and generate only tree-structures that it had already seen, representing those structures as simple hashes rather than structured objects. If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without exploiting the underlying tree structure of the linearized parses in a way that would generalize to substantially new structures.

### 2015-06-13 writing/S'15 LSTM/intro.tex
Paper push.

@@
 
-While both classes of model can perform objectively well on many tasks, and both are under active development, tree-structured models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect.
-Even when the engineering hurdles around building tree-structured models are dealt with, head-to-head comparisons with sequence models show either modest improvements \todo{Cite TreeLSTM paper, and ...} or none at all \todo{cite Jiwei}.
+While both classes of model can perform objectively well on many tasks, and both are under active development, tree-structured models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \todo{Cite TreeLSTM paper, and ...} or none at all \todo{cite Jiwei}.
 
-\todo{We use focused artificial data set to effectively demonstrate...}
-We argue that this question is best studied through focused artificial data experiments, rather than through the inspection of existing learned models.
-There is no theory of distributed representation that is sufficiently detailed to allow us to look for the presence of tree-structured representations in model parameters or activations, and performing in-domain error analysis on trained models is unlikely to provide a sufficiently clear picture.
-An artificial data experiment, in contrast, allows us to single out the behavior that we want to see and evaluate a model's ability to learn it under ideal circumstances with ample clean data, such that if the is able to perform that behavior in any setting, it will do so in our experiment.
+In this paper, we use focused artificial data experiments to effectively expose a possible explanation for these underwhelming results: standard sequence models can learn to implicitly exploit recursive compositional structure.
 
-In particular, we aim to single out the behavior of learning to interpret sentences that use recursive structure.
-We specifically wish to test the recursion aspect of interpretation by training the model on sentences whose length and recursion depth are limited, and testing it on longer and more complex sentences, such that only a model that recognizes and exploits the recursive structure will be able to succeed.
-An existing experimental design meets this criterion nicely: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two types of tree-structured neural network to learn to interpret (by supplying entailment judgments) statements of a simple propositional logic-based language, with a focus on evaluating the models on expressions more complex than those used in training.
+In particular, our aim is to test standard tree and sequence models on their handling of recursive structure by training each model on sentences whose length and recursion depth are limited, and then testing each on longer and more complex sentences, such that only a model that recognizes and exploits the recursive structure will be able to succeed.
+An existing experimental design meets this criterion nicely:\newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two types of tree-structured neural network to learn to interpret (by supplying entailment judgments) statements of a simple propositional logic-based language, with a focus on evaluating the models on expressions more complex than those used in training.
 We propose a way to adapt the experiment presented in that work to sequence models and use it to compare an LSTM-based model with the tree-structured models in that paper, and we extend the experiment by testing multiple different sizes of training set in order to better understand how efficient each model is at learning the underlying generalizations.
@@
 More importantly, though, we find that a sequence model centered on a single-layer LSTM RNN is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree-structured models.
-We take these results as evidence that sequence models are capable of learning tree-structured representations when trained on a task that demands them.
-\todo{higher level conclusions based on the two interpretations from above}
+We take these results as evidence that the similar performance of tree and sequence models can be attributed to their exploitation of similar structures, and that there is value in attempting to further develop this ability in sequence models without fully sacrificing the flexibility that makes these models succeed.
 
@@
 TODO CITE JIWEI performs head-to-head comparisons of tree and sequence models on five NLP tasks over small to moderately-sized data sets and find that tree-structured models robustly outperform sequences on only one, concluding that sequence models are largely adequate for NLP.
-This paper aims to explore that result, specifically focusing on the question of whether compositional structure (implicit or explicit) can be dismissed as unnecessary in research on representation learning for standard NLP tasks.
+This paper aims to explore that result, specifically focusing on the question of whether compositional structure (implicit or explicit) can be dismissed as unessential in research on representation learning for standard NLP tasks.
 

### 2015-06-13 writing/S'15 LSTM/methods.tex
Cleanup, citations, etc.

@@
 \section{Sentence embedding models for NLI} \label{methods}
-\todo{Rewrite in a shorter form, add LSTMs}
 
@@
 \end{figure}
+
+\begin{figure*}[t]
+  \centering
+\begin{subfigure}[t]{0.29\textwidth}
+  \includegraphics[height=1.6in]{fig3.pdf}
+  \caption{Training expression size $\le$3.}
+  \end{subfigure}~~
+\begin{subfigure}[t]{0.29\textwidth}
+    \includegraphics[height=1.6in]{fig4.pdf}
+  \caption{Training expression size $\le$4.}
+  \end{subfigure}~~
+\begin{subfigure}[t]{0.4\textwidth}
+      \includegraphics[height=1.6in]{fig6.pdf}
+  \caption{Training expression size $\le$6.}
+\end{subfigure}
+  \caption{Results on three experiments with increasingly rich training sets.
+The horizontal axis on each graph divides the test set data points (expression pairs) into bins by the number of logical operators in the more complex of the two expressions in the pair.}
+  \label{prop-results} 
+\end{figure*}
 

### 2015-06-13 writing/S'15 LSTM/intro.tex
Cleanup, citations, etc.

@@
 
-Neural network models that encode sentences as vectors have been successfully used on a wide array of NLP tasks, including \todo{citations}.
-Many of the most successful of these models fall into one of two categories: sequence models based on recurrent neural networks build representations incrementally from left to right *CITE, while tree-structured models based on \ii{recursive} neural networks *CITE build representations incrementally according the hierarchical sturcture of linguistic phrases.
+Neural network models that encode sentences as vectors have been successfully used on a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}.
+Many of the most successful of these models fall into one of two categories: sequence models based on recurrent neural networks build representations incrementally from left to right \cite{sutskever2014sequence} \todo{Who to cite as early work on text RNNs}, while tree-structured models based on \ii{recursive} neural networks \cite{socher2011semi,irsoydeep,tai2015improved} build representations incrementally according the hierarchical sturcture of linguistic phrases.
 
-Tree-structured models are presented as the more principled choice, since they directly exploit the principle of compositionality---fundamental to linguistic semantic theory \cite{Partee84,Janssen97}---that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.
+Tree-structured models are presented as the more principled choice, since they directly exploit the principle of compositionality---fundamental to linguistic semantic theory \cite{Partee84,Janssen97}---which states that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.
 
-While both classes of model can perform objectively well on many tasks, and both are under active development, tree-structured models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \todo{Cite TreeLSTM paper, and ...} or none at all \todo{cite Jiwei}.
-
+While both classes of model can perform objectively well on many tasks, and both are under active development, tree-structured models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}.
 In this paper, we use focused artificial data experiments to effectively expose a possible explanation for these underwhelming results: standard sequence models can learn to implicitly exploit recursive compositional structure.
@@
 In particular, our aim is to test standard tree and sequence models on their handling of recursive structure by training each model on sentences whose length and recursion depth are limited, and then testing each on longer and more complex sentences, such that only a model that recognizes and exploits the recursive structure will be able to succeed.
-An existing experimental design meets this criterion nicely:\newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two types of tree-structured neural network to learn to interpret (by supplying entailment judgments) statements of a simple propositional logic-based language, with a focus on evaluating the models on expressions more complex than those used in training.
-We propose a way to adapt the experiment presented in that work to sequence models and use it to compare an LSTM-based model with the tree-structured models in that paper, and we extend the experiment by testing multiple different sizes of training set in order to better understand how efficient each model is at learning the underlying generalizations.
+An existing experimental design meets this criterion nicely: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two types of tree-structured neural network to learn to interpret (by supplying entailment judgments) statements of a simple artificial language.
+We propose a way to adapt the experiment presented in that work to sequence models and use it to compare an LSTM-based model with the tree-structured models in that paper, and we extend the experiment by testing on several structured subsets of the training set in order to better understand how efficient each model is at learning the underlying generalizations.
 
 As in \newcite{Bowman:Potts:Manning:2014}, we find that both TreeRNNs and TreeRNTNs are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size.
-Perhaps unsurprisingly, we also find that extending the training set to include larger structures mitigates this decay.
-More importantly, though, we find that a sequence model centered on a single-layer LSTM RNN is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree-structured models.
+We also find that extending the training set to include larger structures mitigates this decay.
+We find that a sequence model centered on a single-layer LSTM RNN is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree-structured models.
 We take these results as evidence that the similar performance of tree and sequence models can be attributed to their exploitation of similar structures, and that there is value in attempting to further develop this ability in sequence models without fully sacrificing the flexibility that makes these models succeed.
 
-\subsection{Related work}
+\paragraph{Related work}
+\newcite{li2015tree} performs head-to-head comparisons of tree and sequence models on five NLP tasks over small to moderately-sized data sets and find that tree-structured models robustly outperform sequences on only one, concluding that sequence models are largely adequate for NLP.
 
-TODO CITE JIWEI performs head-to-head comparisons of tree and sequence models on five NLP tasks over small to moderately-sized data sets and find that tree-structured models robustly outperform sequences on only one, concluding that sequence models are largely adequate for NLP.
-This paper aims to explore that result, specifically focusing on the question of whether compositional structure (implicit or explicit) can be dismissed as unessential in research on representation learning for standard NLP tasks.
-
-Recent results on parsing by \todo{Cite Grammar as a Foreign Lg, Dyer} hints the outcome that we find.
-The simplest model presented in TODO CITE GaaFL encodes each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) for the sentence using only the information present in the vector.
-This suggests that the intermediate vector representation contains a tree-structured representation, and that that representation was constructed incrementally by the sequence model.
-However, the massive size of the training set that was necessary to train that model---250M tokens---makes an alternate explanation possible: the model could have primarily learned to simply recognize and generate only tree-structures that it had already seen, representing those structures as simple hashes rather than structured objects.
+Recent successes on parsing by \newcite{vinyals2014grammar} and \cite{dyer2015transition} hint that sequence models may be useful in manipulating tree structures.
+The simplest model presented in \newcite{vinyals2014grammar} uses an LSTM sequence model to each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) using only the information present in the vector, suggesting that the LSTM was able to encode the tree.
+However, the massive size of the training set that was necessary to train that model---250M tokens---makes it possible that the model could have primarily learned to recognize and generate only tree-structures that it had already seen, representing those structures as simple hashes rather than structured objects.
 If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without exploiting the underlying tree structure of the linearized parses in a way that would generalize to substantially new structures.

### 2015-06-13 writing/S'15 LSTM/intro.tex
Paper push.

@@
 
-Neural network models that encode sentences as vectors have been successfully used on a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}.
-Many of the most successful of these models fall into one of two categories: sequence models based on recurrent neural networks build representations incrementally from left to right \cite{sutskever2014sequence} \todo{Who to cite as early work on text RNNs}, while tree-structured models based on \ii{recursive} neural networks \cite{socher2011semi,irsoydeep,tai2015improved} build representations incrementally according the hierarchical sturcture of linguistic phrases.
+Neural network models that encode sentences as vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}.
+Many of the most successful of these models fall into one of two categories: sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree-structured models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical sturcture of linguistic phrases.
 
@@
 In particular, our aim is to test standard tree and sequence models on their handling of recursive structure by training each model on sentences whose length and recursion depth are limited, and then testing each on longer and more complex sentences, such that only a model that recognizes and exploits the recursive structure will be able to succeed.
-An existing experimental design meets this criterion nicely: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two types of tree-structured neural network to learn to interpret (by supplying entailment judgments) statements of a simple artificial language.
+An existing experimental design meets this criterion nicely: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two types of tree-structured neural network to learn to interpret statements of a simple artificial language.
 We propose a way to adapt the experiment presented in that work to sequence models and use it to compare an LSTM-based model with the tree-structured models in that paper, and we extend the experiment by testing on several structured subsets of the training set in order to better understand how efficient each model is at learning the underlying generalizations.
@@
 We find that a sequence model centered on a single-layer LSTM RNN is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree-structured models.
-We take these results as evidence that the similar performance of tree and sequence models can be attributed to their exploitation of similar structures, and that there is value in attempting to further develop this ability in sequence models without fully sacrificing the flexibility that makes these models succeed.
 
 \paragraph{Related work}
-\newcite{li2015tree} performs head-to-head comparisons of tree and sequence models on five NLP tasks over small to moderately-sized data sets and find that tree-structured models robustly outperform sequences on only one, concluding that sequence models are largely adequate for NLP.
+\newcite{li2015tree} performs head-to-head comparisons of tree and sequence models on five NLP tasks over small to moderately-sized data sets and find that tree-structured models robustly outperform sequences on only one, concluding that ``recurrent models are sufficient to capture the compositional semantics required for many NLP tasks.''
 
-Recent successes on parsing by \newcite{vinyals2014grammar} and \cite{dyer2015transition} hint that sequence models may be useful in manipulating tree structures.
-The simplest model presented in \newcite{vinyals2014grammar} uses an LSTM sequence model to each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) using only the information present in the vector, suggesting that the LSTM was able to encode the tree.
-However, the massive size of the training set that was necessary to train that model---250M tokens---makes it possible that the model could have primarily learned to recognize and generate only tree-structures that it had already seen, representing those structures as simple hashes rather than structured objects.
-If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without exploiting the underlying tree structure of the linearized parses in a way that would generalize to substantially new structures.
+Recent successes on parsing by \newcite{vinyals2014grammar} and \newcite{dyer2015transition} hint that sequence models may be useful in manipulating tree structures.
+The simplest model presented in \newcite{vinyals2014grammar} uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) using only the information present in the vector, suggesting that the LSTM was able to encode the tree.
+However, the massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model could have primarily learned to recognize and generate only tree-structures that it had already seen, representing those structures as simple hashes rather than structured objects.
+If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without showing demonstrating an ability to represent recursive structure, at least in a way that would generalize to substantially new structures.

### 2015-06-13 TestAndLog.m

@@ -33,7 +33,7 @@ if mod(modelState.step, options.testFreq) == 0
 
         if mod(modelState.step, options.examplesFreq) == 0 && modelState.step > 0
             hyperParams.showExamples = true;
-            Log(hyperParams.examplelog, 'Training data:')
+            Log(hyperParams.examplelog, ['pass ', num2str(modelState.pass), ' step ', num2str(modelState.step), ' training data:'])
         else
             hyperParams.showExamples = false;
         end
@@ -61,7 +61,7 @@ if mod(modelState.step, options.testFreq) == 0
         end
 
         if (mod(modelState.step, options.examplesFreq) == 0 || mod(modelState.step, options.detailedStatFreq) == 0) && modelState.step > 0
-            Log(hyperParams.examplelog, 'Test data:');
+            Log(hyperParams.examplelog, ['pass ', num2str(modelState.pass), ' step ', num2str(modelState.step), ' test data:']);
         end
         [testAcc, testMf1, ~, testConAcc] = TestModel(CostGradFunc, modelState.theta, modelState.thetaDecoder, testDatasets, modelState.separateWordFeatures, hyperParams);
         modelState.bestTestAcc = max(testAcc, modelState.bestTestAcc);

### 2015-06-13 writing/S'15 LSTM/abstract.tex
Paper push.

@@
 We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured are built around, mitigating the advantage of the tree-structured models.
-We investigate this possibility by evaluating both models on an artificial talk for which recursive compositional structure is crucial, and find that the sequence model is able to identify and exploit the underlying structure, though that it is less efficient at learning than the tree-structured models, only succeeding after exposure to a larger and richer set of training data.
+We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to identify and exploit the underlying structure, though that it is less efficient at learning than the tree-structured models, only succeeding after exposure to a larger and richer set of training data.
 \end{abstract}

### 2015-06-13 writing/S'15 LSTM/discussion.tex
Cleanup, citations, etc.
Paper push.

@@
 \section{Results and discussion}\label{sec:discussion}
-
-\begin{figure*}[t!]
-  \centering
-\begin{subfigure}[t]{0.29\textwidth}
-  \includegraphics[height=1.6in]{fig3.pdf}
-  \caption{Training expression size $\le$3.}
-  \end{subfigure}~~
-\begin{subfigure}[t]{0.29\textwidth}
-    \includegraphics[height=1.6in]{fig4.pdf}
-  \caption{Training expression size $\le$4.}
-  \end{subfigure}~~
-\begin{subfigure}[t]{0.4\textwidth}
-      \includegraphics[height=1.6in]{fig6.pdf}
-  \caption{Training expression size $\le$6.}
-\end{subfigure}
-  \caption{Results on three experiments with increasingly rich training sets.
-The horizontal axis on each graph divides the test set data points (expression pairs) into bins by the number of logical operators in the more complex of the two expressions in the pair.}
-  \label{prop-results} 
-\end{figure*}
 
@@
 With sufficient training data (in the $\le$6 setting), the two tree models were able to perform at above 99\% (TODO) on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets.
+These robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.
 
@@
 
-We find that
+We find that all three \todo{Still using three?} models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree structured model's biases allow them to learn to do this more effectively from less data.
+We interpret these results as evidence that the similar performance of tree and sequence models can be attributed to their fundamentally similar representational systems, with both models recognizing compositional structure when it is present, the former doing so more reliably, and the latter benefiting elsewhere---either because of architectural choices that make backpropagation more effective, or because interpreting sentences from left to right allows the model to better simulate other aspects of human language comprehension.
+We finally suggest that there is value in attempting to further develop this ability in sequence models without fully sacrificing the flexibility that makes these models succeed.

### 2015-06-13 writing/S'15 LSTM/figure1.tex
Paper push.

@@
     \def\dx{21pt}
-    \def\dy{29pt}
+    \def\dy{27pt}
 
-    \tikzstyle{label}=[text width=40mm,align=center]    
-    \tikzstyle{softmax}=[fill=red!40,text width=40mm,align=center]
-    \tikzstyle{preclass}=[fill=orange!40,text width=50mm,align=center]
-    \tikzstyle{e}=[fill=green!40,text width=26mm,align=center]
-    \tikzstyle{m}=[draw=black,text width=38mm,align=center]
+    \tikzstyle{label}=[text width=40mm,align=center,text height=2mm]    
+    \tikzstyle{softmax}=[fill=red!40,text width=40mm,align=center,text height=2mm]
+    \tikzstyle{preclass}=[fill=orange!40,text width=50mm,align=center,text height=2mm]
+    \tikzstyle{e}=[fill=green!40,text width=26mm,align=center,text height=2mm]
+    \tikzstyle{m}=[draw=black,text width=38mm,align=center,text height=2mm]
 
@@
     \def\dx{21pt}
-    \def\dy{29pt}
+    \def\dy{27pt}
 
@@
 
-     \caption{The architecture for the TreeRNN and TreeRNTN sentence models.
-Terminal nodes are learned embeddings and nonterminal nodes are either NN or NTN layers with $\tanh$ nonlinearities.}\label{fig:model:tree}
+     \caption{The architecture for the tree-structured sentence models. Terminal nodes are learned embeddings and nonterminal nodes are NN or NTN layers.}\label{fig:model:tree}
 
@@
     \def\dx{35pt}
-    \def\dy{29pt}
+    \def\dy{27pt}
 

### 2015-06-13 writing/S'15 LSTM/methods.tex
Paper push.

@@
 
-\begin{figure}[t!]
+\begin{figure}[t]
   \centering
   \input{figure1}
-  \caption{In our model, two separate sentence models---based on either tree (b) or sequence (c) models---build up vector representations for each of two sentences.
+  \caption{In our model, two copies of a sentence model---based on either tree (b) or sequence (c) models---encode the two sentences.
 A multilayer classifier component (a) then uses the resulting vectors to predict a label that reflects the logical relationship between the two sentences.}
@@
   \centering
-\begin{subfigure}[t]{0.29\textwidth}
-  \includegraphics[height=1.6in]{fig3.pdf}
-  \caption{Training expression size $\le$3.}
+  \begin{subfigure}[t]{0.04\textwidth}
+      \includegraphics[height=1.45in]{scale.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+  \includegraphics[height=1.45in]{fig3c.pdf}
+  \caption{Training on sz. $\le$3.}
+  \end{subfigure}~~~
+\begin{subfigure}[t]{0.24\textwidth}
+    \includegraphics[height=1.45in]{fig4c.pdf}
+  \caption{Training on sz. $\le$4.}
+  \end{subfigure}~~~~
+\begin{subfigure}[t]{0.24\textwidth}
+      \includegraphics[height=1.45in]{fig6c.pdf}
+  \caption{Training on sz. $\le$6.}
 \end{subfigure}~~
-\begin{subfigure}[t]{0.29\textwidth}
-    \includegraphics[height=1.6in]{fig4.pdf}
-  \caption{Training expression size $\le$4.}
-  \end{subfigure}~~
-\begin{subfigure}[t]{0.4\textwidth}
-      \includegraphics[height=1.6in]{fig6.pdf}
-  \caption{Training expression size $\le$6.}
+\begin{subfigure}[t]{0.08\textwidth}
+      \includegraphics[height=1.45in]{leg.pdf}
 \end{subfigure}
@@
 We use the model architecture depicted in Figure \ref{fig:model:top}.
-This approach is loosely based on the model used in the experiments by \newcite{Bowman:Potts:Manning:2014}.
+This approach builds on the model used in \newcite{Bowman:Potts:Manning:2014}.
 The model architecture uses two copies of a single sentence embedding network---a TreeRNN, TreeRNTN, or LSTM RNN, depending on the precise experiment---to embed the premise and hypothesis sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations.
-Since the embeddings are computed separately, the embedding models are forced encode every element of the sentence meaning that the downstream model will need.
-Though this prevents us from exploiting cross-sentence alignment information that could allow for the learning of more effective entailment models, it allows us to focus our evaluation on models that capture the meanings of single sentences.
+Since the embeddings are computed separately, the embedding models are forced to encode every element of the sentence meaning that the downstream model will need.
 
@@
 \end{gather}
-Our model is largely identical to the Bowman et al.
-model, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence embedding models are compared in as similar a setting as possible.
+
+Our model is largely identical to the Bowman et al.~model, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence embedding models are compared in as similar a setting as possible.
 
@@
 Both are structured using the parses that are provided with the data.
-In addition, we use a sequence model (Figure \ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} with the precise formulation from \newcite{zaremba2015recurrent}.
-
-We additionally experimented with a plain RNN---the sequence equivalent of the TreeRNN---but were not able to coerce that weaker model into learning to model even the training data well, so they are omitted below.
-LSTM, in contrast are the most widely used of a class of models that attempt to avoid the extreme sensitivity to initialization and gradient magnitude that make RNN training so unreliable.
+In addition, we use a sequence model (Figure \ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}.\todo{Okay to omit negative result on sequence RNN?}
 
@@
 Our objective is the standard negative log likelihood classification objective with L2 regularization.
-All models were stopped at 100 epochs (TODO), it which point all had largely converged without significantly declining from their peak performances.
+All models were stopped at 100 epochs (TODO), at which point all had largely converged without significantly declining from their peak performances.

### 2015-06-13 writing/S'15 LSTM/data.tex
Paper push.

--- 2015-05-03 writing/F'14 paper/arxiv/intro.tex
@@
-\section{Introduction}\label{sec:intro}
+\section{Recursive structure in artificial data}\label{sec:recursion}
+\paragraph{Reasoning about entailment} 
+The data that we use define a version of the recognizing textual entailment task, in which the goal is to determine what kind of logical consequence relation holds between two sentences, drawing on a small fixed vocabulary of relations such as entailment, contradiction, or synonymy.
+This task is well suited to evaluating neural network models for sentence interpretation: models must develop comprehensive represenations of the meanings of each sentence to do well at the task, but the data does not force these representations to take a specific form, allowing the model to learn whatever kind of representations it can use most effectively.
 
-Tree-structured recursive neural network models (TreeRNNs; \citealt{goller1996learning}) for sentence meaning
-have been successful in an array of sophisticated language tasks,
-including sentiment analysis \cite{socher2011semi,irsoydeep},
-image description \cite{sochergrounded}, and paraphrase detection
-\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
-the ability of these models to learn to produce and use strong 
-semantic representations for sentences. However, it remains an 
-open question whether any such fully learned model can achieve the kind of
-high-fidelity distributed representations proposed in recent algebraic
-work on vector space modeling 
-\cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
-and whether any such model can match the performance of grammars based in logical forms
-in their ability to model core semantic phenomena like quantification, entailment, and contradiction \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}.
+The data we use are labeled with the seven relation system of \newcite{maccartney2009extended}, depicted in Table \ref{b-table}.
 
-Recent work on the algebraic approach of
-\newcite{ClarkCoeckeSadrzadeh2011} has yielded rich frameworks 
-for computing the meanings of fragments of natural
-language compositionally from vector or tensor representations, but 
-has not yet yielded effective methods for learning these representations
-from data in typical machine learning settings. 
-Past experimental work on reasoning with
-distributed representations have been largely confined to short
-phrases \cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment}.
-However, for robust natural language understanding, it is essential to
-model these phenomena in their full generality on complex linguistic
-structures.
-
-This paper describes four machine learning experiments
-that directly evaluate the abilities of these models to learn representations that 
-support specific semantic behaviors. These tasks follow the format of
- \ii{natural language inference} 
-(also known as \ii{recognizing textual entailment};
-\citealt{dagan2006pascal}), in which the goal is to determine the core
-inferential relationship between two sentences. 
-We introduce a novel NN architecture for natural language inference which
-independently computes vector representations for each of two sentences using
-standard TreeRNN or TreeRNTN \cite{socher2013acl1} models, and
-produces a judgment for the pair using only those representations. This allows
-us to gauge the abilities of these two models to represent all of the necessary semantic
-information in the sentence vectors.
-
-Much of the
-theoretical work on natural language inference (and some successful implemented models;
-\citealt{maccartney2009extended,watanabe2012latent}) involves \ii{natural
-  logics}, which are formal systems that define rules of inference
-between natural language words, phrases, and sentences without the
-need of intermediate representations in an artificial logical
-language.
-In our first three experiments, we test our models' ability to learn the foundations of natural
-language inference by training them to reproduce the behavior of the natural logic of \newcite{maccartney2009extended} 
-on artificial data. This logic
-defines seven mutually-exclusive relations of synonymy, entailment, contradiction,
-and mutual consistency, as summarized in Table~\ref{b-table}, and it
-provides rules of semantic combination for projecting these relations
-from the lexicon up to complex phrases. The formal properties of this system 
-are now well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.
-The first experiment using this logic 
-covers reasoning with the bare logical relations (\S\ref{sec:join}), the second extends this to reasoning with statements constructed compositionally from recursive functions (\S\ref{sec:recursion}),
-and the third covers the additional complexity that results from quantification (\S\ref{sec:quantifiers}).
-Though the performance of the plain TreeRNN model 
- is somewhat poor in our first experiment, we find that the stronger TreeRNTN model
-  generalizes well in every case, suggesting that it has learned to simulate our target logical concepts.
-
-The experiments with simulated data provide a convincing demonstration
-of the ability of neural networks to learn to build and use semantic representations for complex natural language sentences from reasonably-sized
-training sets. However, we are also interested in the more practical
-question of whether they can learn these representations from
-naturalistic text. To address this question, we apply our models to
-the SICK entailment challenge data in \S\ref{sec:sick}. The small size of this corpus puts
-data-hungry NN models like ours at a disadvantage, but we are
-nonetheless able to achieve competitive performance on it,
-surpassing several submitted models with significant hand-engineered task-specific features and our own NN baseline.
-This suggests that the representational abilities that we observe in the previous sections
-are not limited to carefully circumscribed tasks. We conclude that TreeRNTN models
-are adequate for typical cases of natural language inference, and that there is not yet
-any clear level of inferential complexity for which other approaches work and NN models fail.
-
-\begin{table*}[tp]
+\begin{table}[tp]
   \centering\small
-  \setlength{\tabcolsep}{15pt}
   \renewcommand{\arraystretch}{1}
@@
     \toprule
-    Name & Symbol & Set-theoretic definition & Example \\ 
+    Name & Symb. & Set-theoretic definition \\ 
     \midrule
-    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
-    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
-    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
-    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
-    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
-    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
-    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
+strict entailment         & $\natfor$   & $x \subset y$  \\ 
+    strict rev. entailment & $\natrev$   & $x \supset y$  \\ 
+    equivalence        & $\nateq$    & $x = y$   \\ 
+    alternation        & $\natalt$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ \\ 
+    negation           & $\natneg$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$   \\
+    cover              & $\natcov$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ \\ 
+    independence       & $\natind$   & (else)\\
     \bottomrule
   \end{tabular}
-  \caption{\label{b-table}The seven relations of MacCartney and Manning \protect\shortcite{maccartney2009extended}'s logic are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.} %-%
+  \caption{\label{b-table}MacCartney's seven mutually exclusive relations are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.
+  } %-%
+\end{table}
 
-\end{table*}
+\begin{table}[tp]
+  \centering\small
+%  \begin{subtable}[t]{0.45\textwidth}
+%    \centering
+%    \begin{tabular}[t]{l l}
+%      \toprule
+%      Formula     & Interpretation \\
+%      \midrule
+%      $p_1$, $p_2$, $p_3$, $p_4$, $p_5$, $p_6$ & $\sem{x} \in \{\True, \False\}$ \\
+%      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
+%      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
+%      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
+%      \bottomrule
+%    \end{tabular}    
+%    \caption{Well-formed formulae. $\varphi$ and $\psi$
+%      range over all well-formed formulae, and $\sem{\cdot}$ is
+%      the interpretation function mapping formulae into $\{\True,
+%      \False\}$.}\label{tab:pl}
+%  \end{subtable}
+    \begin{tabular}[t]{r c l}
+      \toprule
+      $\plneg p_3$        & $\natneg$ & $p_3$ \\
+%      $\plneg \plneg p_6$ & $\nateq$  & $p_6$ \\
+      $p_3$               & $\natfor$ & $(p_3 \plor p_2)$ \\
+      $(p_1 \plor (p_2 \plor p_4))$               & $\natrev$ & $(p_2 \pland  \plneg p_4)$ \\
+      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
+      $\plneg\, (\plneg p_1 \pland \plneg p_2)$ & $\nateq$ & $(p_1 \plor p_2)$ \\ 
+      $(p_3 \pland \plneg p_1 ) \plor \plneg p_3$    & $\natrev$& $\plneg\, (p_3 \plor p_2)$ \\
+      %<	( not ( c ( or b ) ) )	( ( c ( and ( not a ) ) ) ( or ( not c ) ) )
+      \bottomrule
+    \end{tabular}
+    \caption{Examples of short expressions from the artificial data introduced by \protect\citealt{Bowman:Potts:Manning:2014}.
+For brevity, we only show the parentheses that are needed to disambiguate the sentences rather than the full binary bracketings.}\label{tab:plexs}
+\end{table}
+
+\paragraph{The artificial language} The language described in Bowman et al.~(\S4) is designed to highlight the use of recursive structure with minimal additional complexity.
+It's vocabulary consists only of six unanalyzed word types ($p_1, p_2, p_3, p_4, p_5, p_6$), \word{and}, \word{or}, and \word{not}.
+Sentences of the language can be straightforwardly interpreted as statements of propositional logic (where the six unanalyzed words are variables), and labeled sentence pairs can be interpreted as theorems of that logic.
+Some example pairs are provided in Table \ref{tab:plexs}.
+
+Crucially, the language is defined such that any sentence can be embedded under negation or conjunction to create a new sentence, allowing for arbitrary-depth recursion, and the scope of negation and conjunction are determined only by bracketing with parentheses (rather than bare word order).
+The compositional structure of each sentence can thus be an arbitrary tree structure, and interpreting a sentence correctly requires using that structure.
+
+The data comes with parentheses representing complete binary bracketing.
+Our two models use this information differently.
+For the tree structured model, we assumble the tree structure for each sentence according to the parentheses, but do not treat them as word tokens.
+For the sequence model, we treat the parentheses as words, and learn their embeddings.
+This provides the two models with equivalent data, so that their ability to handle unseen structures can be fairly compared.
+
+\paragraph{The corpus}
+The sentence pairs in the corpus are divided into thirteen bins according to the number of logical connectives (\word{and, or, not}) in the longer of the two sentences in the pair.
+We test the model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how its performance depends on the complexity of the sentences.
+In three experiments, we train our models on the training portions of bins 0--3 (62k examples), bins 0--4 (90k), and bins 0--6 (160k).
+Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures.
+If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, the models will inevitably will start to make errors on sentences above some length, and the (necessary) use of L2 regularization exacerbates this for sentences which are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-06-13 writing/S'15 LSTM/abstract.tex
Paper trim and proof, new figures.

@@
 We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured are built around, mitigating the advantage of the tree-structured models.
-We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to identify and exploit the underlying structure, though that it is less efficient at learning than the tree-structured models, only succeeding after exposure to a larger and richer set of training data.
+We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to identify and exploit the underlying structure, though it is less efficient at learning than the tree-structured models, only succeeding after exposure to a larger and richer set of training data.
 \end{abstract}

### 2015-06-13 writing/S'15 LSTM/data.tex
Paper trim and proof, new figures.

@@
 
-The data we use are labeled with the seven relation system of \newcite{maccartney2009extended}, depicted in Table \ref{b-table}.
+The data we use are labeled with the seven mutually exclusive logical reations of \newcite{maccartney2009extended}, which distinguish entailment in two directions ($\natfor$, $\natrev$), equivalence ($\nateq$), two types of contradiction ($\natneg$, $\natalt$), and two types of semantic independence ($\natind$, $\natcov$).
 
-\begin{table}[tp]
-  \centering\small
-  \renewcommand{\arraystretch}{1}
-  \begin{tabular}{l c l l} 
-    \toprule
-    Name & Symb. & Set-theoretic definition \\ 
-    \midrule
-strict entailment         & $\natfor$   & $x \subset y$  \\ 
-    strict rev. entailment & $\natrev$   & $x \supset y$  \\ 
-    equivalence        & $\nateq$    & $x = y$   \\ 
-    alternation        & $\natalt$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ \\ 
-    negation           & $\natneg$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$   \\
-    cover              & $\natcov$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ \\ 
-    independence       & $\natind$   & (else)\\
-    \bottomrule
-  \end{tabular}
-  \caption{\label{b-table}MacCartney's seven mutually exclusive relations are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.
-  } %-%
-\end{table}
+%\begin{table}[tp]
+%  \centering\small
+%  \renewcommand{\arraystretch}{1}
+%  \begin{tabular}{l c l l} 
+%    \toprule
+%    Name & Symb. & Set-theoretic definition \\ 
+%    \midrule
+%strict entailment         & $\natfor$   & $x \subset y$  \\ 
+%    strict rev. entailment & $\natrev$   & $x \supset y$  \\ 
+%    equivalence        & $\nateq$    & $x = y$   \\ 
+%    alternation        & $\natalt$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ \\ 
+%    negation           & $\natneg$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$   \\
+%    cover              & $\natcov$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ \\ 
+%    independence       & $\natind$   & (else)\\
+%    \bottomrule
+%  \end{tabular}
+%  \caption{\label{b-table}MacCartney's seven mutually exclusive relations are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.
+%  } %-%
+%\end{table}
 
@@
 
-The data comes with parentheses representing complete binary bracketing.
+The data come with parentheses representing complete binary bracketing.
 Our two models use this information differently.
@@
 We test the model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how its performance depends on the complexity of the sentences.
-In three experiments, we train our models on the training portions of bins 0--3 (62k examples), bins 0--4 (90k), and bins 0--6 (160k).
+In three experiments, we train our models on the training portions of bins 0--3 (62k examples), 0--4 (90k), and 0--6 (160k).
 Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures.

### 2015-06-13 writing/S'15 LSTM/discussion.tex
Paper trim and proof, new figures.

@@
 
-The results are shown in Figure \ref{prop-results}.
-All three models were able to largely memorize the training data (TODO, TODO, and TODO for $\le$6).
-With sufficient training data (in the $\le$6 setting), the two tree models were able to perform at above 99\% (TODO) on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets.
-These robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.
+The results are shown in Figure \ref{prop-results}\todo{Next version of figures will have vertical dividing lines at 3,4,6}.
+All four models were able to largely memorize the training data (TODO, TODO, TODO, and TODO for $\le$6).
+In the $\le$6 setting, the three tree models were able to perform at above TODO\% on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets.
+All four robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.
 
-The LSTM performs fairly poorly in the $\le3$ setting, with performance at 3 a mediorce TODO\%, and an abrubt drop to TODO\% at 4, suggesting that the model did not acquire the needed generalizations.
+The LSTM performs fairly poorly in the $\le3$ setting, with performance at 3 a mediocre TODO\%, and an abrupt drop to TODO\% at 4, suggesting that the model did not acquire the needed generalizations.
 With more ample training data in the $\le6$ condition, though, the LSTM shows fairly good accuracy on short examples, and a smooth decay with no abrupt drop.
+
+[Something about the learning curve that doesn't involve a figure. Something about the learning curve that doesn't involve a figure. Something about the learning curve that doesn't involve a figure.]
 
@@
 
-We find that all three \todo{Still using three?} models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree structured model's biases allow them to learn to do this more effectively from less data.
+We find that all four models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree structured model's biases allow them to learn to do this more effectively from less data.
 We interpret these results as evidence that the similar performance of tree and sequence models can be attributed to their fundamentally similar representational systems, with both models recognizing compositional structure when it is present, the former doing so more reliably, and the latter benefiting elsewhere---either because of architectural choices that make backpropagation more effective, or because interpreting sentences from left to right allows the model to better simulate other aspects of human language comprehension.
-We finally suggest that there is value in attempting to further develop this ability in sequence models without fully sacrificing the flexibility that makes these models succeed.
+We finally suggest that, because of the well supported linguistic claim that the kind of recursive structure that we study here is key to language understanding, there is likely to be value in developing sequence models that can more efficiently exploit this structure without fully sacrificing the flexibility that makes these models succeed.

### 2015-06-13 writing/S'15 LSTM/methods.tex
Paper trim and proof, new figures.

@@
 
-We use the model architecture depicted in Figure \ref{fig:model:top}.
-This approach builds on the model used in \newcite{Bowman:Potts:Manning:2014}.
-The model architecture uses two copies of a single sentence embedding network---a TreeRNN, TreeRNTN, or LSTM RNN, depending on the precise experiment---to embed the premise and hypothesis sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations.
+We use the model architecture depicted in Figure \ref{fig:model:top}, which builds on the model used in \newcite{Bowman:Potts:Manning:2014}.
+The model architecture uses two copies of a single sentence embedding network (a tree or sequence model) to embed the premise and hypothesis (left and right side) sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations.
 Since the embeddings are computed separately, the embedding models are forced to encode every element of the sentence meaning that the downstream model will need.
@@
 The classifier component of the model consists of a combinging layer which takes the two sentence representations as inputs, followed by two neural network layers, then a softmax classifier.
-For the combining layer, we use a (recursive) neural tensor network ((R)NTN, \citealt{chen2013learning}) layer, which sums the output of a plain recursive/recurrent neural network layer with a vector computed using two multiplications with a learned (full rank) third-order tensor parameter:
+For the combining layer, we use a neural tensor network (NTN, \citealt{chen2013learning}) layer, which sums the output of a plain recursive/recurrent neural network layer with a vector computed using two multiplications with a learned (full rank) third-order tensor parameter:
 \begin{gather} 
 \label{TreeRNN}
-\vec{y}_{\textit{RNN}} = \tanh(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
+\vec{y}_{\textit{NN}} = \tanh(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
 \label{TreeRNTN} 
-\vec{y}_{\textit{RNTN}} = \vec{y}_{\textit{RNN}} + \tanh(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
+\vec{y}_{\textit{NTN}} = \vec{y}_{\textit{RNN}} + \tanh(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
 \end{gather}
@@
 The sentence embedding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector.
-We experiment with two tree-structured models (Figure \ref{fig:model:tree}), one with an RNTN activation function or a plain RNN activation function.
-Both are structured using the parses that are provided with the data.
-In addition, we use a sequence model (Figure \ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}.\todo{Okay to omit negative result on sequence RNN?}
+We experiment with tree-structured models (Figure \ref{fig:model:tree}) with TreeRNN (eqn.~\ref{TreeRNN}), TreeRNTN (eqn.~\ref{TreeRNTN}), and TreeLSTM \cite{tai2015improved} activation functions.
+In addition, we use a sequence model (Figure \ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}.
+Experiments with simpler non-LSTM RNN sequence models tended to underfit the training data, and are not included here.
 

### 2015-06-13 writing/S'15 LSTM/intro.tex
Paper trim and proof, new figures.

@@
 
-Tree-structured models are presented as the more principled choice, since they directly exploit the principle of compositionality---fundamental to linguistic semantic theory \cite{Partee84,Janssen97}---which states that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.
+Tree-structured models are presented as the more principled choice, since they directly exploit the principle of compositionality---a fundamental idea in linguistic semantics \cite{Partee84,Janssen97}---which states that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.
 
 While both classes of model can perform objectively well on many tasks, and both are under active development, tree-structured models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}.
-In this paper, we use focused artificial data experiments to effectively expose a possible explanation for these underwhelming results: standard sequence models can learn to implicitly exploit recursive compositional structure.
+In this paper, we use focused artificial data experiments to effectively expose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive compositional structure.
 
-In particular, our aim is to test standard tree and sequence models on their handling of recursive structure by training each model on sentences whose length and recursion depth are limited, and then testing each on longer and more complex sentences, such that only a model that recognizes and exploits the recursive structure will be able to succeed.
-An existing experimental design meets this criterion nicely: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two types of tree-structured neural network to learn to interpret statements of a simple artificial language.
-We propose a way to adapt the experiment presented in that work to sequence models and use it to compare an LSTM-based model with the tree-structured models in that paper, and we extend the experiment by testing on several structured subsets of the training set in order to better understand how efficient each model is at learning the underlying generalizations.
+In particular, our aim is to test standard tree and sequence models on their handling of recursive structure by training each model on sentences whose length and recursion depth are limited, and then testing each on longer and more complex sentences, such that only models that recognize and exploit the recursive structure will be able to succeed.
+An existing experimental design meets this criterion nicely: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two tree-structured models to learn to interpret statements of a simple artificial language.
+We propose a way to adapt than experiment to sequence models and use it to compare an LSTM-based model with the three tree-structured models and we extend the experiment by testing on several structured subsets of the training set in order to better understand how efficiently each model learns the underlying generalizations.
 
-As in \newcite{Bowman:Potts:Manning:2014}, we find that both TreeRNNs and TreeRNTNs are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size.
+As in \newcite{Bowman:Potts:Manning:2014}, we find that standard tree-structured neural networks are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size.
 We also find that extending the training set to include larger structures mitigates this decay.
-We find that a sequence model centered on a single-layer LSTM RNN is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree-structured models.
+In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree-structured models.
 
@@
 
-Recent successes on parsing by \newcite{vinyals2014grammar} and \newcite{dyer2015transition} hint that sequence models may be useful in manipulating tree structures.
-The simplest model presented in \newcite{vinyals2014grammar} uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) using only the information present in the vector, suggesting that the LSTM was able to encode the tree.
-However, the massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model could have primarily learned to recognize and generate only tree-structures that it had already seen, representing those structures as simple hashes rather than structured objects.
+Recent successes on parsing by \newcite{vinyals2014grammar} and \newcite{dyer2015transition} hint that sequence models may be able to exploit recursive structure.
+The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) using only the information present in the vector, suggesting that the LSTM was able to encode the tree.
+However, the massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model primarily learned to recognize and generate only tree-structures that it had already seen, representing those structures as simple hashes rather than structured objects.
 If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without showing demonstrating an ability to represent recursive structure, at least in a way that would generalize to substantially new structures.

### 2015-06-14 writing/S'15 LSTM/discussion.tex
Text summary of learning curve

@@
 The results are shown in Figure \ref{prop-results}\todo{Next version of figures will have vertical dividing lines at 3,4,6}.
-All four models were able to largely memorize the training data (TODO, TODO, TODO, and TODO for $\le$6).
+All four models were able to largely memorize the training data (TODO, TODO, TODO, and TODO for $\le$6, pooling across bins).
 In the $\le$6 setting, the three tree models were able to perform at above TODO\% on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets.
@@
 
-[Something about the learning curve that doesn't involve a figure. Something about the learning curve that doesn't involve a figure. Something about the learning curve that doesn't involve a figure.]
+Learning curve experiments suggest that none of these models are saturated, and that still larger data sets could help even on this simple task.
+The LSTM in the $\le6$ strikingly moves from TODO\% accuracy with a random 33\% of the available training data to TODO\% on all of it.
+The plain TreeRNN shows the smallest improvement, moving from TODO\% to TODO\%.
 

### 2015-06-14 writing/S'15 LSTM/data.tex
Delete SAT instances; logging; paper push.

@@
 The data that we use define a version of the recognizing textual entailment task, in which the goal is to determine what kind of logical consequence relation holds between two sentences, drawing on a small fixed vocabulary of relations such as entailment, contradiction, or synonymy.
-This task is well suited to evaluating neural network models for sentence interpretation: models must develop comprehensive represenations of the meanings of each sentence to do well at the task, but the data does not force these representations to take a specific form, allowing the model to learn whatever kind of representations it can use most effectively.
+This task is well suited to evaluating neural network models for sentence interpretation: models must develop comprehensive represenations of the meanings of each sentence to do well at the task, but the data do not force these representations to take a specific form, allowing the model to learn whatever kind of representations it can use most effectively.
 
-The data we use are labeled with the seven mutually exclusive logical reations of \newcite{maccartney2009extended}, which distinguish entailment in two directions ($\natfor$, $\natrev$), equivalence ($\nateq$), two types of contradiction ($\natneg$, $\natalt$), and two types of semantic independence ($\natind$, $\natcov$).
+The data we use are labeled with the seven mutually exclusive logical reations of \newcite{maccartney2009extended}, which distinguish entailment in two directions ($\natfor$, $\natrev$), equivalence ($\nateq$), exhaustive and non-exhaustive contradiction ($\natneg$, $\natalt$), and two types of semantic independence ($\natind$, $\natcov$).
 
@@
 \paragraph{The artificial language} The language described in Bowman et al.~(\S4) is designed to highlight the use of recursive structure with minimal additional complexity.
-It's vocabulary consists only of six unanalyzed word types ($p_1, p_2, p_3, p_4, p_5, p_6$), \word{and}, \word{or}, and \word{not}.
+Its vocabulary consists only of six unanalyzed word types ($p_1, p_2, p_3, p_4, p_5, p_6$), \word{and}, \word{or}, and \word{not}.
 Sentences of the language can be straightforwardly interpreted as statements of propositional logic (where the six unanalyzed words are variables), and labeled sentence pairs can be interpreted as theorems of that logic.
@@
 Our two models use this information differently.
-For the tree structured model, we assumble the tree structure for each sentence according to the parentheses, but do not treat them as word tokens.
+For the tree model, we assumble the tree structure for each sentence according to the parentheses, but do not treat them as word tokens.
 For the sequence model, we treat the parentheses as words, and learn their embeddings.
@@
 Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures.
-If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, the models will inevitably will start to make errors on sentences above some length, and the (necessary) use of L2 regularization exacerbates this for sentences which are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.
+If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, the models will inevitably will start to make errors on sentences above some length, and the use of L2 regularization (which helps overall performance) exacerbates this for sentences which are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-06-14 writing/S'15 LSTM/discussion.tex
Delete SAT instances; logging; paper push.

@@
 
-The results are shown in Figure \ref{prop-results}\todo{Next version of figures will have vertical dividing lines at 3,4,6}.
+The results are shown in Figure~\ref{prop-results}\todo{Next version of figures will have vertical dividing lines at 3,4,6}.
 All four models were able to largely memorize the training data (TODO, TODO, TODO, and TODO for $\le$6, pooling across bins).
@@
 
-\paragraph{Conclusion}
+\section{Conclusion}
 
-We find that all four models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree structured model's biases allow them to learn to do this more effectively from less data.
+We find that all four models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree models' biases allow them to learn to do this more effectively from less data.
 We interpret these results as evidence that the similar performance of tree and sequence models can be attributed to their fundamentally similar representational systems, with both models recognizing compositional structure when it is present, the former doing so more reliably, and the latter benefiting elsewhere---either because of architectural choices that make backpropagation more effective, or because interpreting sentences from left to right allows the model to better simulate other aspects of human language comprehension.

### 2015-06-14 writing/S'15 LSTM/abstract.tex
Delete SAT instances; logging; paper push.

@@
 Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but these models largely have not outperformed simpler sequence-based models by substantial margins.
-We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured are built around, mitigating the advantage of the tree-structured models.
-We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to identify and exploit the underlying structure, though it is less efficient at learning than the tree-structured models, only succeeding after exposure to a larger and richer set of training data.
+We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured ones are built around---at least in cases where there are clear cues to that structure in the data---mitigating the advantage of the tree-structured models.
+We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to exploit the underlying structure, though it is less efficient at learning than the tree models, only succeeding after exposure to a larger and richer set of training data.
 \end{abstract}

### 2015-06-14 writing/S'15 LSTM/intro.tex
Delete SAT instances; logging; paper push.

@@
 Neural network models that encode sentences as vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}.
-Many of the most successful of these models fall into one of two categories: sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree-structured models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical sturcture of linguistic phrases.
+Many of the most successful of these models are based on either tree models or sequence models.
+sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical sturcture of linguistic phrases.
 
-Tree-structured models are presented as the more principled choice, since they directly exploit the principle of compositionality---a fundamental idea in linguistic semantics \cite{Partee84,Janssen97}---which states that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.
+Tree models are presented as the more principled choice, since they directly exploit the principle of compositionality---a fundamental idea in linguistic semantics \cite{Partee84,Janssen97}---which states that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.
 
-While both classes of model can perform objectively well on many tasks, and both are under active development, tree-structured models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}.
-In this paper, we use focused artificial data experiments to effectively expose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive compositional structure.
+While both classes of model can perform objectively well on many tasks, and both are under active development, tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}.
+In this paper, we use focused artificial data experiments to expose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive compositional structure.
 
-In particular, our aim is to test standard tree and sequence models on their handling of recursive structure by training each model on sentences whose length and recursion depth are limited, and then testing each on longer and more complex sentences, such that only models that recognize and exploit the recursive structure will be able to succeed.
-An existing experimental design meets this criterion nicely: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two tree-structured models to learn to interpret statements of a simple artificial language.
-We propose a way to adapt than experiment to sequence models and use it to compare an LSTM-based model with the three tree-structured models and we extend the experiment by testing on several structured subsets of the training set in order to better understand how efficiently each model learns the underlying generalizations.
+In particular, we aim test standard tree and sequence models on their handling of recursive structure by training the models on sentences whose length and recursion depth are limited, and then testing them on longer and more complex sentences, such that only models that exploit the recursive structure will be able to succeed.
+An existing experimental design meets this criterion: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two tree models to learn to interpret statements of a simple artificial language.
+We propose a way to adapt than experiment to sequence models involving decorating the statements with an explicit bracketing, and use the experiment to compare an LSTM sequence model with the three treee models, with a focus on what data each model needs in order to learn the needed generalizations.
 
-As in \newcite{Bowman:Potts:Manning:2014}, we find that standard tree-structured neural networks are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size.
+As in \newcite{Bowman:Potts:Manning:2014}, we find that standard tree neural networks are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size.
 We also find that extending the training set to include larger structures mitigates this decay.
-In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree-structured models.
+In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree models.
 
 \paragraph{Related work}
-\newcite{li2015tree} performs head-to-head comparisons of tree and sequence models on five NLP tasks over small to moderately-sized data sets and find that tree-structured models robustly outperform sequences on only one, concluding that ``recurrent models are sufficient to capture the compositional semantics required for many NLP tasks.''
+\newcite{li2015tree} performs head-to-head comparisons of tree and sequence models on five NLP tasks over small to moderately-sized data sets and find that tree models robustly outperform sequences on only one, concluding that ``recurrent models are sufficient to capture the compositional semantics required for many NLP tasks.''
 
@@
 The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) using only the information present in the vector, suggesting that the LSTM was able to encode the tree.
-However, the massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model primarily learned to recognize and generate only tree-structures that it had already seen, representing those structures as simple hashes rather than structured objects.
+However, the massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model primarily learned to recognize and generate only tree structures that it had already seen, representing those structures as simple hashes rather than structured objects.
 If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without showing demonstrating an ability to represent recursive structure, at least in a way that would generalize to substantially new structures.
+Our paper aims to understand whether this ability to generalize is likely or possible.

### 2015-06-14 writing/S'15 LSTM/methods.tex
Delete SAT instances; logging; paper push.

@@
 
-We use the model architecture depicted in Figure \ref{fig:model:top}, which builds on the model used in \newcite{Bowman:Potts:Manning:2014}.
+We use the model architecture depicted in Figure~\ref{fig:model:top}, which builds on the model used in \newcite{Bowman:Potts:Manning:2014}.
 The model architecture uses two copies of a single sentence embedding network (a tree or sequence model) to embed the premise and hypothesis (left and right side) sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations.
@@
 The sentence embedding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector.
-We experiment with tree-structured models (Figure \ref{fig:model:tree}) with TreeRNN (eqn.~\ref{TreeRNN}), TreeRNTN (eqn.~\ref{TreeRNTN}), and TreeLSTM \cite{tai2015improved} activation functions.
-In addition, we use a sequence model (Figure \ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}.
+We experiment with tree models (Figure~\ref{fig:model:tree}) with TreeRNN (eqn.~\ref{TreeRNN}), TreeRNTN (eqn.~\ref{TreeRNTN}), and TreeLSTM \cite{tai2015improved} activation functions.
+In addition, we use a sequence model (Figure~\ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}.
 Experiments with simpler non-LSTM RNN sequence models tended to underfit the training data, and are not included here.

### 2015-06-14 writing/S'15 LSTM/abstract.tex
Paper push.

@@
 We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured ones are built around---at least in cases where there are clear cues to that structure in the data---mitigating the advantage of the tree-structured models.
-We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to exploit the underlying structure, though it is less efficient at learning than the tree models, only succeeding after exposure to a larger and richer set of training data.
+We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to exploit the underlying structure, though it is less efficient at learning than the tree models, only succeeding after exposure to a larger and richer training set.
 \end{abstract}

### 2015-06-14 writing/S'15 LSTM/methods.tex
Paper push.

@@
 Our objective is the standard negative log likelihood classification objective with L2 regularization.
-All models were stopped at 100 epochs (TODO), at which point all had largely converged without significantly declining from their peak performances.
+All models were stopped after TODO gradient steps, at which point all had largely converged without significantly declining from their peak performances.

### 2015-06-14 writing/S'15 LSTM/intro.tex
Paper push.

@@
 sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical sturcture of linguistic phrases.
-
-Tree models are presented as the more principled choice, since they directly exploit the principle of compositionality---a fundamental idea in linguistic semantics \cite{Partee84,Janssen97}---which states that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.
+Tree models are presented as the more principled choice, since they align with standard linguistic assumptions about constituent structure and the compositional derivation of complex meanings.
 
 While both classes of model can perform objectively well on many tasks, and both are under active development, tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}.
-In this paper, we use focused artificial data experiments to expose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive compositional structure.
+In this paper, we use focused artificial data experiments to expose a possible explanation for these results: when trained on strings that contain indications of recursive syntactic structure, standard sequence models implicitly learn tree-like constituent structures even though they do not process their inputs as trees.
+That is, to the extent that constituent structure it useful, they can learn approximations of it without being required to do so by their architecture.
 
@@
 
-As in \newcite{Bowman:Potts:Manning:2014}, we find that standard tree neural networks are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size.
+As in Bowman et al., we find that standard tree neural networks are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size.
 We also find that extending the training set to include larger structures mitigates this decay.

### 2015-06-14 writing/S'15 LSTM/abstract.tex
Abstract update.

@@
 We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured ones are built around---at least in cases where there are clear cues to that structure in the data---mitigating the advantage of the tree-structured models.
-We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to exploit the underlying structure, though it is less efficient at learning than the tree models, only succeeding after exposure to a larger and richer training set.
+We investigate this possibility using artificial data task for which recursive compositional structure is crucial, and find that the sequence model is able to exploit the underlying structure, though it is less efficient at learning than the tree models, only succeeding after exposure to a larger and richer set of training data.
 \end{abstract}

### 2015-06-14 writing/S'15 LSTM/intro.tex
Abstract update.

@@
 Many of the most successful of these models are based on either tree models or sequence models.
-sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical sturcture of linguistic phrases.
+Sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical sturcture of linguistic phrases.
 Tree models are presented as the more principled choice, since they align with standard linguistic assumptions about constituent structure and the compositional derivation of complex meanings.

### 2015-06-14 writing/S'15 LSTM/intro.tex
Semi-major intro reframing.

@@
 While both classes of model can perform objectively well on many tasks, and both are under active development, tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}.
-In this paper, we use focused artificial data experiments to expose a possible explanation for these results: when trained on strings that contain indications of recursive syntactic structure, standard sequence models implicitly learn tree-like constituent structures even though they do not process their inputs as trees.
-That is, to the extent that constituent structure it useful, they can learn approximations of it without being required to do so by their architecture.
+We propose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive syntactic structure in generating sentence representations, thereby demonstrating the behavior that tree-structured models are explicitly designed around.
+This would require both that sequence models are able to identify syntactic structure in natural language---which we believe is plausible, on the basis of other recent research---and that they are able to use that structure to guide the way in which they interpret sentences.
+In this paper, we  evaluate this second ability, focusing on cases where the relevant syntactic structure is clearly indicated in the data.
 
-In particular, we aim test standard tree and sequence models on their handling of recursive structure by training the models on sentences whose length and recursion depth are limited, and then testing them on longer and more complex sentences, such that only models that exploit the recursive structure will be able to succeed.
-An existing experimental design meets this criterion: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two tree models to learn to interpret statements of a simple artificial language.
-We propose a way to adapt than experiment to sequence models involving decorating the statements with an explicit bracketing, and use the experiment to compare an LSTM sequence model with the three treee models, with a focus on what data each model needs in order to learn the needed generalizations.
+We compare standard tree and sequence models on their handling of recursive structure by training the models on sentences whose length and recursion depth are limited, and then testing them on longer and more complex sentences, such that only models that exploit the recursive structure will be able to generalize in a way that yields correct interpretations for these test sentences.
+Our methods are based on those of \newcite{Bowman:Potts:Manning:2014}, who describe an experiment and corresponding artificial dataset which tests this ability in two tree models.
+We propose a way to adapt than experiment to sequence models by decorating the statements with an explicit bracketing, and use the experiment to compare an LSTM sequence model with the three tree models, with a focus on what data each model needs in order to learn the needed generalizations.
 
-As in Bowman et al., we find that standard tree neural networks are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size.
+As in Bowman et al., we find that standard tree neural networks are able to make the necessary generalizations, with their performance decaying gradually as the structures in the test set grow in size.
 We also find that extending the training set to include larger structures mitigates this decay.
@@
 \paragraph{Related work}
-\newcite{li2015tree} performs head-to-head comparisons of tree and sequence models on five NLP tasks over small to moderately-sized data sets and find that tree models robustly outperform sequences on only one, concluding that ``recurrent models are sufficient to capture the compositional semantics required for many NLP tasks.''
-
-Recent successes on parsing by \newcite{vinyals2014grammar} and \newcite{dyer2015transition} hint that sequence models may be able to exploit recursive structure.
-The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) using only the information present in the vector, suggesting that the LSTM was able to encode the tree.
-However, the massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model primarily learned to recognize and generate only tree structures that it had already seen, representing those structures as simple hashes rather than structured objects.
+Recent successes on parsing by \newcite{vinyals2014grammar} and \newcite{dyer2015transition} provide evidence that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks.
+The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) with high accuracy using only the information present in the vector.
+This shows that the LSTM was able to identify the correct syntactic structures and also hints that it was able to develop a generalizable method for encoding these structures in vectors.
+The massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model primarily learned to generate only tree structures that it had already seen, representing those structures as simple hashes rather than structured objects.
 If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without showing demonstrating an ability to represent recursive structure, at least in a way that would generalize to substantially new structures.
-Our paper aims to understand whether this ability to generalize is likely or possible.

### 2015-06-14 writing/S'15 LSTM/abstract.tex
Abstract and intro revisions

@@
 \begin{abstract}
-Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but these models largely have not outperformed simpler sequence-based models by substantial margins.
-We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured ones are built around---at least in cases where there are clear cues to that structure in the data---mitigating the advantage of the tree-structured models.
-We investigate this possibility using artificial data task for which recursive compositional structure is crucial, and find that the sequence model is able to exploit the underlying structure, though it is less efficient at learning than the tree models, only succeeding after exposure to a larger and richer set of training data.
+%Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but 
+Tree-structured neural networks aim to deliver robust sentential meaning representations, but 
+these models largely have not %outperformed simpler sequence-based models by substantial margins. 
+significantly outperformed simpler sequence-based models.
+We hypothesize that sequence models like LSTMs are able to discover and implicitly use %the same kinds of 
+recursive compositional structure, %that the tree-structured ones are built around---
+at least in cases where there are clear cues to that structure in the data, mitigating the advantage of a tree-structured architecture.
+We investigate this possibility using an artificial data task for which recursive compositional structure is crucial, and find that the sequence model is able to exploit the underlying structure, though it is less efficient at learning than the tree models, succeeding only after exposure to a larger and richer training set.
 \end{abstract}

### 2015-06-14 writing/S'15 LSTM/intro.tex
Abstract and intro revisions

@@
 Neural network models that encode sentences as vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}.
-Many of the most successful of these models are based on either tree models or sequence models.
-Sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical sturcture of linguistic phrases.
-Tree models are presented as the more principled choice, since they align with standard linguistic assumptions about constituent structure and the compositional derivation of complex meanings.
+Many of the most successful of these models are based on either tree or sequence architectures.
+Sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical structure of linguistic phrases.
+Tree models are often presented as the more principled choice, since they align with standard linguistic assumptions about constituent structure and the compositional derivation of complex meanings.
 
-While both classes of model can perform objectively well on many tasks, and both are under active development, tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}.
+While both classes of model perform %objectively 
+well on many tasks, and both are under active development, tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}.
 We propose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive syntactic structure in generating sentence representations, thereby demonstrating the behavior that tree-structured models are explicitly designed around.
-This would require both that sequence models are able to identify syntactic structure in natural language---which we believe is plausible, on the basis of other recent research---and that they are able to use that structure to guide the way in which they interpret sentences.
+If correct, this means that
+%This would require both that 
+sequence models are able to identify syntactic structure in natural language---which we believe is plausible, on the basis of other recent research---and that they are able to use that structure to guide interpretation.
+%the way in which they interpret sentences.
 In this paper, we  evaluate this second ability, focusing on cases where the relevant syntactic structure is clearly indicated in the data.
@@
 Our methods are based on those of \newcite{Bowman:Potts:Manning:2014}, who describe an experiment and corresponding artificial dataset which tests this ability in two tree models.
-We propose a way to adapt than experiment to sequence models by decorating the statements with an explicit bracketing, and use the experiment to compare an LSTM sequence model with the three tree models, with a focus on what data each model needs in order to learn the needed generalizations.
+We adapt that experiment to sequence models by decorating the statements with an explicit bracketing, and we use this design to compare an LSTM sequence model with three tree models, with a focus on what data each model needs in order to learn the needed generalizations.
 
@@
 We also find that extending the training set to include larger structures mitigates this decay.
-In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree models.
+In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it does this only when trained on a larger and more complex training set than is needed by the tree models.
 
-\paragraph{Related work}
-Recent successes on parsing by \newcite{vinyals2014grammar} and \newcite{dyer2015transition} provide evidence that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks.
+%\paragraph{Related work}
+%Recent successes on parsing by 
+Out results engage with those of \newcite{vinyals2014grammar} and \newcite{dyer2015transition}, who find that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks.
 The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) with high accuracy using only the information present in the vector.
 This shows that the LSTM was able to identify the correct syntactic structures and also hints that it was able to develop a generalizable method for encoding these structures in vectors.
-The massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model primarily learned to generate only tree structures that it had already seen, representing those structures as simple hashes rather than structured objects.
-If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without showing demonstrating an ability to represent recursive structure, at least in a way that would generalize to substantially new structures.
+However, the massive size of the data set needed to train the model---250M tokens---leaves open the possibility that it primarily learned to generate only tree structures that it had already seen, representing them as simple hashes rather than structured objects.
+If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without actually representing recursive structure, at least in a way that would generalize to substantially new structures.
+Our experiments suggest that LSMTs can generalize in the requisite ways, but at the cost of very large training sets of the right type.

### 2015-06-14 writing/S'15 LSTM/data.tex
Data tweaks

@@
 \paragraph{Reasoning about entailment} 
-The data that we use define a version of the recognizing textual entailment task, in which the goal is to determine what kind of logical consequence relation holds between two sentences, drawing on a small fixed vocabulary of relations such as entailment, contradiction, or synonymy.
-This task is well suited to evaluating neural network models for sentence interpretation: models must develop comprehensive represenations of the meanings of each sentence to do well at the task, but the data do not force these representations to take a specific form, allowing the model to learn whatever kind of representations it can use most effectively.
+The data that we use define a version of the \emph{recognizing textual entailment} task, in which the goal is to determine what kind of logical consequence relation holds between two sentences, drawing on a small fixed vocabulary of relations such as entailment, contradiction, and synonymy.
+This task is well suited to evaluating neural network models for sentence interpretation: models must develop comprehensive representations of the meanings of each sentence to do well at the task, but the data do not force these representations to take a specific form, allowing the model to learn whatever kind of representations it can use most effectively.
 
-The data we use are labeled with the seven mutually exclusive logical reations of \newcite{maccartney2009extended}, which distinguish entailment in two directions ($\natfor$, $\natrev$), equivalence ($\nateq$), exhaustive and non-exhaustive contradiction ($\natneg$, $\natalt$), and two types of semantic independence ($\natind$, $\natcov$).
+The data we use are labeled with the seven mutually exclusive logical relations of \newcite{maccartney2009extended}, which distinguish entailment in two directions ($\natfor$, $\natrev$), equivalence ($\nateq$), exhaustive and non-exhaustive contradiction ($\natneg$, $\natalt$), and two types of semantic independence ($\natind$, $\natcov$).
 
@@
 Sentences of the language can be straightforwardly interpreted as statements of propositional logic (where the six unanalyzed words are variables), and labeled sentence pairs can be interpreted as theorems of that logic.
-Some example pairs are provided in Table \ref{tab:plexs}.
+Some example pairs are provided in Table~\ref{tab:plexs}.
 
 Crucially, the language is defined such that any sentence can be embedded under negation or conjunction to create a new sentence, allowing for arbitrary-depth recursion, and the scope of negation and conjunction are determined only by bracketing with parentheses (rather than bare word order).
-The compositional structure of each sentence can thus be an arbitrary tree structure, and interpreting a sentence correctly requires using that structure.
+The compositional structure of each sentence can thus be an arbitrary tree, and interpreting a sentence correctly requires using that structure.
 
@@
 Our two models use this information differently.
-For the tree model, we assumble the tree structure for each sentence according to the parentheses, but do not treat them as word tokens.
-For the sequence model, we treat the parentheses as words, and learn their embeddings.
-This provides the two models with equivalent data, so that their ability to handle unseen structures can be fairly compared.
+For the tree model, the parentheses are not word tokens, but rather used in the expected way to build the tree.
+For the sequence model, the parentheses are words with associated learned embeddings.
+This provides the two models with equivalent data, so their ability to handle unseen structures can be fairly compared.
 

### 2015-06-14 writing/S'15 LSTM/data.tex
Tweaks

@@
 Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures.
-If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, the models will inevitably will start to make errors on sentences above some length, and the use of L2 regularization (which helps overall performance) exacerbates this for sentences which are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.
+If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and the use of L2 regularization (which helps overall performance) exacerbates this for sentences that are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-06-14 writing/S'15 LSTM/methods.tex
Tweaks

@@
 
-We use the model architecture depicted in Figure~\ref{fig:model:top}, which builds on the model used in \newcite{Bowman:Potts:Manning:2014}.
+We use the architecture depicted in Figure~\ref{fig:model:top}, which builds on the one used in \newcite{Bowman:Potts:Manning:2014}.
 The model architecture uses two copies of a single sentence embedding network (a tree or sequence model) to embed the premise and hypothesis (left and right side) sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations.
@@
 \paragraph{Classifier}
-The classifier component of the model consists of a combinging layer which takes the two sentence representations as inputs, followed by two neural network layers, then a softmax classifier.
+The classifier component of the model consists of a combining layer which takes the two sentence representations as inputs, followed by two neural network layers, then a softmax classifier.
 For the combining layer, we use a neural tensor network (NTN, \citealt{chen2013learning}) layer, which sums the output of a plain recursive/recurrent neural network layer with a vector computed using two multiplications with a learned (full rank) third-order tensor parameter:

### 2015-06-15 writing/S'15 LSTM/intro.tex
Google parsing paper interpretation tweak.

@@
 %Recent successes on parsing by 
-Out results engage with those of \newcite{vinyals2014grammar} and \newcite{dyer2015transition}, who find that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks.
+Our results engage with those of \newcite{vinyals2014grammar} and \newcite{dyer2015transition}, who find that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks.
 The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) with high accuracy using only the information present in the vector.
@@
 If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without actually representing recursive structure, at least in a way that would generalize to substantially new structures.
-Our experiments suggest that LSMTs can generalize in the requisite ways, but at the cost of very large training sets of the right type.
+Our experiments show that LSTMs can manipulate tree-structured data, suggesting that there are are no fundamental obstacles to this kind of generalization.

### 2015-06-15 writing/S'15 LSTM/abstract.tex
Chris turn at abstract.

@@
 %Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but 
-Tree-structured neural networks aim to deliver robust sentential meaning representations, but 
-these models largely have not %outperformed simpler sequence-based models by substantial margins. 
-significantly outperformed simpler sequence-based models.
-We hypothesize that sequence models like LSTMs are able to discover and implicitly use %the same kinds of 
+%Tree-structured neural networks aim to deliver robust sentential meaning representations, but 
+Tree-structured neural networks encode a particular tree geometry for a sentence in the network design.
+However, these models have at best only slightly outperformed simpler sequence-based models.
+% these models largely have not %outperformed simpler sequence-based models by substantial margins. 
+% significantly outperformed simpler sequence-based models.
+We hypothesize that neural sequence models like LSTMs are in fact able to discover and implicitly use %the same kinds of 
 recursive compositional structure, %that the tree-structured ones are built around---
-at least in cases where there are clear cues to that structure in the data, mitigating the advantage of a tree-structured architecture.
-We investigate this possibility using an artificial data task for which recursive compositional structure is crucial, and find that the sequence model is able to exploit the underlying structure, though it is less efficient at learning than the tree models, succeeding only after exposure to a larger and richer training set.
+at least for tasks with clear cues to that structure in the data. 
+% , mitigating the advantage of a tree-structured architecture. 
+We demonstrate this possibility using an artificial data task for which recursive compositional structure is crucial, and find that the sequence model can learn the underlying patterning.
+The sequence model is better in that it learns the value of tree structure from the data in an emergent way, while the tree-structured model is better in being able to learn with greater statistical efficiency due to its informative prior model structure.
+% structure, though it is less efficient at learning than the tree models, succeeding only after exposure to a larger and richer training set.
 \end{abstract}

### 2015-06-15 writing/S'15 LSTM/intro.tex
Some intro work

@@
 
-Neural network models that encode sentences as vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}.
-Many of the most successful of these models are based on either tree or sequence architectures.
-Sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical structure of linguistic phrases.
-Tree models are often presented as the more principled choice, since they align with standard linguistic assumptions about constituent structure and the compositional derivation of complex meanings.
+Neural network models that encode sentences as real-valued vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}.
+% Many of the most successful of these models are based on either tree or sequence architectures. 
+These models for language may be either 
+sequence models based on recurrent neural networks, which build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, or tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi}, which build representations incrementally according the hierarchical structure of linguistic phrases.
 
-While both classes of model perform %objectively 
-well on many tasks, and both are under active development, tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}.
-We propose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive syntactic structure in generating sentence representations, thereby demonstrating the behavior that tree-structured models are explicitly designed around.
-If correct, this means that
+While both model classes perform %objectively 
+well on many tasks, and both are under active development,
+tree models are often presented as the more principled choice, since they align with standard linguistic assumptions about constituent structure and the compositional derivation of complex meanings.
+Nevertheless,
+tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}.
+
+We propose a possible explanation for these results: standard sequence models can learn to
+%  implicitly 
+exploit recursive syntactic structure in generating sentence meaning representations, thereby 
+% demonstrating 
+learning to use the 
+% behavior 
+structure that tree-structured models are explicitly designed around. This first requires that
 %This would require both that 
-sequence models are able to identify syntactic structure in natural language---which we believe is plausible, on the basis of other recent research---and that they are able to use that structure to guide interpretation.
+sequence models are able to identify syntactic structure in natural language.
+We believe this is plausible, on the basis of other recent research \cite{Karpathy2015vaurn}.
 %the way in which they interpret sentences.
-In this paper, we  evaluate this second ability, focusing on cases where the relevant syntactic structure is clearly indicated in the data.
+In this paper, we  evaluate whether LSTM models are able to use that structure to guide interpretation, 
+focusing on cases where the relevant syntactic structure is clearly indicated in the data.
 

### 2015-06-15 writing/S'15 LSTM/intro.tex
Slight argument weakening.

@@
 % Many of the most successful of these models are based on either tree or sequence architectures. 
-These models for language may be either 
+These models are generally either 
 sequence models based on recurrent neural networks, which build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, or tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi}, which build representations incrementally according the hierarchical structure of linguistic phrases.

### 2015-06-15 writing/S'15 LSTM/discussion.tex
Paper push.
Learning curve

@@
+\begin{figure}[t]
+  \centering
+      \includegraphics[height=1.1in]{lcc.pdf}
+  \caption{Learning curve for the $\le6$ experiment.}
+  \label{fig:lc} 
+\end{figure}
+
 \section{Results and discussion}\label{sec:discussion}
 
-The results are shown in Figure~\ref{prop-results}\todo{Next version of figures will have vertical dividing lines at 3,4,6}.
-All four models were able to largely memorize the training data (TODO, TODO, TODO, and TODO for $\le$6, pooling across bins).
-In the $\le$6 setting, the three tree models were able to perform at above TODO\% on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets.
+The results are shown in Figure~\ref{prop-results}.
+All four models were able to largely memorize the training data (??.?, ??.?, ??.?, and ??.?
+for $\le$6, pooling across bins).
+In the $\le$6 setting, the three tree models were able to perform at above ??\% on examples of size $\le$2 with a steady decay in performance continuing through ??.?\% at size 12, and similar, if steeper, decays with smaller training sets.
 All four robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.
 
-The LSTM performs fairly poorly in the $\le3$ setting, with performance at 3 a mediocre TODO\%, and an abrupt drop to TODO\% at 4, suggesting that the model did not acquire the needed generalizations.
+The LSTM performs fairly poorly in the $\le3$ setting, with performance at 3 a mediocre ??.?\%, and an abrupt drop to \% at 4, suggesting that the model did not acquire the needed generalizations.
 With more ample training data in the $\le6$ condition, though, the LSTM shows fairly good accuracy on short examples, and a smooth decay with no abrupt drop.
 
-Learning curve experiments suggest that none of these models are saturated, and that still larger data sets could help even on this simple task.
-The LSTM in the $\le6$ strikingly moves from TODO\% accuracy with a random 33\% of the available training data to TODO\% on all of it.
-The plain TreeRNN shows the smallest improvement, moving from TODO\% to TODO\%.
+The learning curve (Figure~\ref{fig:lc}) suggests that additional data is unlikely to change these basic results, though the disparity in performance at the 11\% level is striking: the LSTM sequence model and the TreeRNTN (the model with the most parameters by far) lag far behind the other two.
 

### 2015-06-15 writing/S'15 LSTM/data.tex
LC and partial proofread.

@@
 
-The data come with parentheses representing complete binary bracketing.
-Our two models use this information differently.
-For the tree model, the parentheses are not word tokens, but rather used in the expected way to build the tree.
+The data come with parentheses representing a complete binary bracketing.
+Our models use this information in two ways.
+For the tree models, the parentheses are not word tokens, but rather used in the expected way to build the tree.
 For the sequence model, the parentheses are words with associated learned embeddings.
-This provides the two models with equivalent data, so their ability to handle unseen structures can be fairly compared.
+This provides the models with equivalent data, so their ability to handle unseen structures can be fairly compared.
 
@@
 Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures.
-If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and the use of L2 regularization (which helps overall performance) exacerbates this for sentences that are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.
+If their performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and the use of L2 regularization (which helps overall performance) exacerbates this for sentences that are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-06-15 writing/S'15 LSTM/intro.tex
Learning curve
LC and partial proofread.

@@
 These models are generally either 
-sequence models based on recurrent neural networks, which build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, or tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi}, which build representations incrementally according the hierarchical structure of linguistic phrases.
+sequence models based on recurrent neural networks, which build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, or tree models based on \ii{recursive} neural networks, which build representations incrementally according the hierarchical structure of linguistic phrases \cite{goller1996learning,socher2011semi}.
 
@@
 sequence models are able to identify syntactic structure in natural language.
-We believe this is plausible, on the basis of other recent research \cite{Karpathy2015vaurn}.
+We believe this is plausible, on the basis of other recent research \cite{vinyals2014grammar,Karpathy2015vaurn}.
 %the way in which they interpret sentences.
@@
 If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without actually representing recursive structure, at least in a way that would generalize to substantially new structures.
-Our experiments show that LSTMs can manipulate tree-structured data, suggesting that there are are no fundamental obstacles to this kind of generalization.
+However, our experiments show that LSTMs can manipulate tree-structured data, suggesting that there are are no fundamental obstacles to this kind of generalization.

### 2015-06-15 writing/S'15 LSTM/methods.tex
LC and partial proofread.

@@
-\section{Sentence embedding models for NLI} \label{methods}
+\section{Testing sentence models on entailment} \label{methods}
 
@@
 We use the architecture depicted in Figure~\ref{fig:model:top}, which builds on the one used in \newcite{Bowman:Potts:Manning:2014}.
-The model architecture uses two copies of a single sentence embedding network (a tree or sequence model) to embed the premise and hypothesis (left and right side) sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations.
-Since the embeddings are computed separately, the embedding models are forced to encode every element of the sentence meaning that the downstream model will need.
+The model architecture uses two copies of a single sentence model (a tree or sequence model) to encode the premise and hypothesis (left and right side) sentences, and then uses those encodings as the features for a multilayer classifier which predicts one of the seven relations.
+Since the encodings are computed separately, the sentence models are forced to encode every element of the sentence meaning that the downstream model will need.
 
@@
 \label{TreeRNTN} 
-\vec{y}_{\textit{NTN}} = \vec{y}_{\textit{RNN}} + \tanh(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
+\vec{y}_{\textit{NTN}} = \vec{y}_{\textit{NN}} + \tanh(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
 \end{gather}
 
-Our model is largely identical to the Bowman et al.~model, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence embedding models are compared in as similar a setting as possible.
+Our model is largely identical to the Bowman et al.~model, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence models are compared in as similar a setting as possible.
 
-\paragraph{Sentence embedding}
-The sentence embedding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector.
+\paragraph{Sentence models}
+The sentence encoding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector.
 We experiment with tree models (Figure~\ref{fig:model:tree}) with TreeRNN (eqn.~\ref{TreeRNN}), TreeRNTN (eqn.~\ref{TreeRNTN}), and TreeLSTM \cite{tai2015improved} activation functions.
@@
 Our objective is the standard negative log likelihood classification objective with L2 regularization.
-All models were stopped after TODO gradient steps, at which point all had largely converged without significantly declining from their peak performances.
+All models were stopped after ???
+gradient steps, at which point all had largely converged without significantly declining from their peak performances.

### 2015-06-15 writing/S'15 LSTM/discussion.tex
LC and partial proofread.
Discussion cleanup.

@@
 The results are shown in Figure~\ref{prop-results}.
-All four models were able to largely memorize the training data (??.?, ??.?, ??.?, and ??.?
-for $\le$6, pooling across bins).
-In the $\le$6 setting, the three tree models were able to perform at above ??\% on examples of size $\le$2 with a steady decay in performance continuing through ??.?\% at size 12, and similar, if steeper, decays with smaller training sets.
-All four robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.
+All four models were able to largely memorize the training data, reaching ??.?, ??.?, ??.?, and ??.?
+overall accuracy for the $\le$6 setting.
+In that setting, the three tree models were able to perform at above ??.?\% on examples of size $\le$2 with a steady decay in performance continuing through ??.?\% at 12, and similar, if steeper, decays with smaller training sets.
+The LSTM performs fairly poorly in the $\le3$ setting, with performance at 3 a mediocre ??.?\%, and an abrupt drop to ??.?\% at 4, suggesting that the model did not acquire the needed generalizations.
+With more ample training data in the $\le6$ condition, though, the LSTM shows fairly good accuracy on short examples, and a smooth decay with no abrupt drop.
 
-The LSTM performs fairly poorly in the $\le3$ setting, with performance at 3 a mediocre ??.?\%, and an abrupt drop to \% at 4, suggesting that the model did not acquire the needed generalizations.
-With more ample training data in the $\le6$ condition, though, the LSTM shows fairly good accuracy on short examples, and a smooth decay with no abrupt drop.
+All four models robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a neural bag of words model does reasonably on the shortest examples but falls below 60\% by bin~4.
 

### 2015-06-15 writing/S'15 LSTM/discussion.tex
Typo corrections

@@
 We interpret these results as evidence that the similar performance of tree and sequence models can be attributed to their fundamentally similar representational systems, with both models recognizing compositional structure when it is present, the former doing so more reliably, and the latter benefiting elsewhere---either because of architectural choices that make backpropagation more effective, or because interpreting sentences from left to right allows the model to better simulate other aspects of human language comprehension.
-We finally suggest that, because of the well supported linguistic claim that the kind of recursive structure that we study here is key to language understanding, there is likely to be value in developing sequence models that can more efficiently exploit this structure without fully sacrificing the flexibility that makes these models succeed.
+Finally, we suggest that, because of the well-supported linguistic claim that the kind of recursive structure that we study here is key to language understanding, there is likely to be value in developing sequence models that can more efficiently exploit this structure without fully sacrificing the flexibility that makes these models succeed.

### 2015-06-15 writing/S'15 LSTM/intro.tex
Typo corrections

@@
 
-Neural network models that encode sentences as real-valued vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}.
+Neural networks that encode sentences as real-valued vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}.
 % Many of the most successful of these models are based on either tree or sequence architectures. 
 These models are generally either 
-sequence models based on recurrent neural networks, which build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, or tree models based on \ii{recursive} neural networks, which build representations incrementally according the hierarchical structure of linguistic phrases \cite{goller1996learning,socher2011semi}.
+sequence models based on recurrent neural networks, which build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, or tree models based on \ii{recursive} neural networks, which build representations incrementally
+according to the hierarchical structure of linguistic phrases \cite{goller1996learning,socher2011semi}.
 
@@
 This shows that the LSTM was able to identify the correct syntactic structures and also hints that it was able to develop a generalizable method for encoding these structures in vectors.
-However, the massive size of the data set needed to train the model---250M tokens---leaves open the possibility that it primarily learned to generate only tree structures that it had already seen, representing them as simple hashes rather than structured objects.
+However, the massive size of the dataset needed to train the model---250M tokens---leaves open the possibility that it primarily learned to generate only tree structures that it had already seen, representing them as simple hashes rather than structured objects.
 If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without actually representing recursive structure, at least in a way that would generalize to substantially new structures.
-However, our experiments show that LSTMs can manipulate tree-structured data, suggesting that there are are no fundamental obstacles to this kind of generalization.
+However, our experiments show that LSTMs can manipulate tree-structured data, suggesting that there are no fundamental obstacles to this kind of generalization.

### 2015-06-15 writing/S'15 LSTM/intro.tex
Tiny wording changes

@@
 As in Bowman et al., we find that standard tree neural networks are able to make the necessary generalizations, with their performance decaying gradually as the structures in the test set grow in size.
-We also find that extending the training set to include larger structures mitigates this decay.
-In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it does this only when trained on a larger and more complex training set than is needed by the tree models.
+We additionally find that extending the training set to include larger structures mitigates this decay.
+Then considering sequence models, we find that a single-layer LSTM is also able to generalize to unseen large structures, but that it does this only when trained on a larger and more complex training set than is needed by the tree models.
 
@@
 However, the massive size of the dataset needed to train the model---250M tokens---leaves open the possibility that it primarily learned to generate only tree structures that it had already seen, representing them as simple hashes rather than structured objects.
+% CDM: I wonder about just leaving this next sentence out
 If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without actually representing recursive structure, at least in a way that would generalize to substantially new structures.

### 2015-06-15 writing/S'15 LSTM/methods.tex
Tiny wording changes

@@
   \caption{Results on three experiments with increasingly rich training sets.
-The horizontal axis on each graph divides the test set data points (expression pairs) into bins by the number of logical operators in the more complex of the two expressions in the pair.}
+The horizontal axis on each graph divides the test set data points (expression pairs) into bins by the number of logical operators in the more complex of the two expressions in the pair.
+The dotted line shows the size we trained on up to.}
   \label{prop-results} 

### 2015-06-15 writing/S'15 LSTM/intro.tex
Comment replies.

@@
 This shows that the LSTM was able to identify the correct syntactic structures and also hints that it was able to develop a generalizable method for encoding these structures in vectors.
-However, the massive size of the dataset needed to train the model---250M tokens---leaves open the possibility that it primarily learned to generate only tree structures that it had already seen, representing them as simple hashes rather than structured objects.
+However, the massive size of the dataset needed to train the model---250M tokens---leaves open the possibility that it primarily learned to generate only tree structures that it had already seen, representing them as simple hashes---which would not capture unseen tree structures---rather than as structured objects.
 % CDM: I wonder about just leaving this next sentence out
-If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without actually representing recursive structure, at least in a way that would generalize to substantially new structures.
-However, our experiments show that LSTMs can manipulate tree-structured data, suggesting that there are no fundamental obstacles to this kind of generalization.
+% If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without actually representing recursive structure, at least in a way that would generalize to substantially new structures.
+Our experiments, though, show that LSTMs can manipulate tree-structured data, suggesting that there is no fundamental obstacle to this kind of generalization.

### 2015-06-15 writing/S'15 LSTM/methods.tex
Comment replies.

@@
 \end{subfigure}
-  \caption{Results on three experiments with increasingly rich training sets.
-The horizontal axis on each graph divides the test set data points (expression pairs) into bins by the number of logical operators in the more complex of the two expressions in the pair.
+  \caption{Test accuracy on three experiments with increasingly rich training sets.
+The horizontal axis on each graph divides the test set expression pairs into bins by the number of logical operators in the more complex of the two expressions in the pair.
 The dotted line shows the size we trained on up to.}

### 2015-06-15 writing/S'15 LSTM/data.tex
Comment replies.

@@
 We test the model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how its performance depends on the complexity of the sentences.
-In three experiments, we train our models on the training portions of bins 0--3 (62k examples), 0--4 (90k), and 0--6 (160k).
-Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures.
-If their performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and the use of L2 regularization (which helps overall performance) exacerbates this for sentences that are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.
+In three experiments, we train our models on the training portions of bins 0--3 (62k examples), 0--4 (90k), and 0--6 (160k), and test on every bin but the trivial bin 0.
+Capping the size of the training sentences allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures.
+If their performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and L2 regularization (which helps overall performance) exacerbates this by preventing the model from using the kind of numerically precise, nonlinearity-saturating functions that would generalize best.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-06-16 writing/S'15 LSTM/discussion.tex
Paper submission candidate.

--- 2015-06-15 writing/S'15 LSTM/discussion.tex
@@
   \centering
-      \includegraphics[height=1.1in]{lcc.pdf}
-  \caption{Learning curve for the $\le6$ experiment.}
+      \includegraphics[height=1.3in]{lcc.pdf}
+  \caption{Learning curve for the $\le$6 experiment.}
   \label{fig:lc} 
@@
 The results are shown in Figure~\ref{prop-results}. 
-All four models were able to largely memorize the training data (??.?, ??.?, ??.?, and ??.?
-for $\le$6, pooling across bins).
-In the $\le$6 setting, the three tree models were able to perform at above ??\% on examples of size $\le$2 with a steady decay in performance continuing through ??.?\% at size 12, and similar, if steeper, decays with smaller training sets.
-All four robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.
+The tree models fit the training data well, reaching 99.6, 98.5, and 97.7\% overall accuracy respectively in the $\le$6 setting, with the LSTM underfitting slightly at 94.4\%. 
+In that setting, all models generalized well to structures of familiar length, with the tree models all surpassing 97\% on examples of size 4, and the LSTM reaching 94.8\%.
+On the longer test sentences, the tree models decayed smoothly in performance across the board, while the LSTM decayed more quickly and more abruptly, with a striking difference in the $\le$4 setting, where LSTM performance falls 10\% from 4 to 5, compared to 4.4\% for the next worse model.
+However, the LSTM improves considerably with more ample training data in the $\le$6 condition, showing only a 3\% drop and generalization results better than the best model's in the $\le$3 setting.
+% TODO: These stats?
 
-The LSTM performs fairly poorly in the $\le3$ setting, with performance at 3 a mediocre ??.?\%, and an abrupt drop to \% at 4, suggesting that the model did not acquire the needed generalizations.
-With more ample training data in the $\le6$ condition, though, the LSTM shows fairly good accuracy on short examples, and a smooth decay with no abrupt drop.
+All four models robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time and a neural bag of words model does reasonably on the shortest examples but falls below 60\% by bin~4.
 
@@
 
-We find that all four models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree models' biases allow them to learn to do this more effectively from less data.
-We interpret these results as evidence that the similar performance of tree and sequence models can be attributed to their fundamentally similar representational systems, with both models recognizing compositional structure when it is present, the former doing so more reliably, and the latter benefiting elsewhere---either because of architectural choices that make backpropagation more effective, or because interpreting sentences from left to right allows the model to better simulate other aspects of human language comprehension.
-We finally suggest that, because of the well supported linguistic claim that the kind of recursive structure that we study here is key to language understanding, there is likely to be value in developing sequence models that can more efficiently exploit this structure without fully sacrificing the flexibility that makes these models succeed.
+We find that all four models can exploit a recursively defined language to interpret sentences with complex unseen structures, and that tree models' biases allow them to learn to do this more effectively from less data.
+We interpret these results as evidence that both tree and sequence architectures can play valuable roles in the construction of sentence models over data with recursive syntactic structure: tree architectures provide an explicit bias that makes it possible to learn from relatively small and impoverished data sets, while sequence-based architectures can solve the same problems---given sufficient data---without being constrained by this bias.
+Finally, we suggest that, because of the well-supported linguistic claim that the kind of recursive structure that we study here is key to the understanding of real natural languages, there is likely to be value in developing sequence models that can more efficiently exploit this structure without fully sacrificing the flexibility that makes them succeed.

### 2015-06-16 writing/S'15 LSTM/methods.tex
Paper submission candidate.

@@
   \begin{subfigure}[t]{0.04\textwidth}
-      \includegraphics[height=1.45in]{scale.pdf}
+      \includegraphics[height=1.5in]{scale.pdf}
 \end{subfigure}
 \begin{subfigure}[t]{0.24\textwidth}
-  \includegraphics[height=1.45in]{fig3c.pdf}
+  \includegraphics[height=1.5in]{fig3c.pdf}
   \caption{Training on sz. $\le$3.}
-  \end{subfigure}~~~
+  \end{subfigure}~~~~
 \begin{subfigure}[t]{0.24\textwidth}
-    \includegraphics[height=1.45in]{fig4c.pdf}
+    \includegraphics[height=1.5in]{fig4c.pdf}
   \caption{Training on sz. $\le$4.}
@@
 \begin{subfigure}[t]{0.24\textwidth}
-      \includegraphics[height=1.45in]{fig6c.pdf}
+      \includegraphics[height=1.5in]{fig6c.pdf}
   \caption{Training on sz. $\le$6.}
@@
 \begin{subfigure}[t]{0.08\textwidth}
-      \includegraphics[height=1.45in]{leg.pdf}
+      \includegraphics[height=1.5in]{leg.pdf}
 \end{subfigure}
@@
 
-\paragraph{Training} We randomly initialize all embeddings and layer parameters, and then train all parameters using minibatch stochastic gradient descent with AdaDelta \cite{zeiler2012adadelta} learning rates.
-Our objective is the standard negative log likelihood classification objective with L2 regularization.
-All models were stopped after ???
-gradient steps, at which point all had largely converged without significantly declining from their peak performances.
+\paragraph{Training} We randomly initialize all embeddings and layer parameters, and train them using minibatch stochastic gradient descent with AdaDelta \cite{zeiler2012adadelta} learning rates.
+Our objective is the standard negative log likelihood classification objective with L2 regularization (tuned on a separate train/test split).
+All models were trained for 72 epochs, after which all had largely converged without significantly declining from their peak performances.

### 2015-06-16 writing/S'15 LSTM/data.tex
Abstract formatting for pasting, final tweaks.

@@
 Capping the size of the training sentences allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures.
-If their performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and L2 regularization (which helps overall performance) exacerbates this by preventing the model from using the kind of numerically precise, nonlinearity-saturating functions that would generalize best.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.
+If their performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and L2 regularization (which helps overall performance) exacerbates this by discouraging the model from using the kind of numerically precise, nonlinearity-saturating functions that would generalize best.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-06-16 writing/S'15 LSTM/abstract.tex
Abstract formatting for pasting, final tweaks.

@@
 \begin{abstract} 
-%Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but 
-%Tree-structured neural networks aim to deliver robust sentential meaning representations, but 
 Tree-structured neural networks encode a particular tree geometry for a sentence in the network design.
 However, these models have at best only slightly outperformed simpler sequence-based models.
-% these models largely have not %outperformed simpler sequence-based models by substantial margins. 
-% significantly outperformed simpler sequence-based models.
-We hypothesize that neural sequence models like LSTMs are in fact able to discover and implicitly use %the same kinds of 
-recursive compositional structure, %that the tree-structured ones are built around---
-at least for tasks with clear cues to that structure in the data. 
-% , mitigating the advantage of a tree-structured architecture. 
+We hypothesize that neural sequence models like LSTMs are in fact able to discover and implicitly use recursive compositional structure, at least for tasks with clear cues to that structure in the data.
 We demonstrate this possibility using an artificial data task for which recursive compositional structure is crucial, and find that the sequence model can learn the underlying patterning.
 The sequence model is better in that it learns the value of tree structure from the data in an emergent way, while the tree-structured model is better in being able to learn with greater statistical efficiency due to its informative prior model structure.
-% structure, though it is less efficient at learning than the tree models, succeeding only after exposure to a larger and richer training set.
 \end{abstract}

### 2015-06-16 writing/S'15 LSTM/intro.tex
Abstract formatting for pasting, final tweaks.

@@
 Neural networks that encode sentences as real-valued vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}.
-% Many of the most successful of these models are based on either tree or sequence architectures. 
 These models are generally either 
@@
 
-While both model classes perform %objectively 
+While both model classes perform
 well on many tasks, and both are under active development,
@@
 We propose a possible explanation for these results: standard sequence models can learn to
-%  implicitly 
 exploit recursive syntactic structure in generating sentence meaning representations, thereby 
-% demonstrating 
 learning to use the 
-% behavior 
 structure that tree-structured models are explicitly designed around. This first requires that
-%This would require both that 
-sequence models are able to identify syntactic structure in natural language.
-We believe this is plausible, on the basis of other recent research \cite{vinyals2014grammar,Karpathy2015vaurn}.
-%the way in which they interpret sentences.
+sequence models are able to identify syntactic structure in natural language. We believe this is plausible, on the basis of other recent research \cite{vinyals2014grammar,Karpathy2015vaurn}.
 In this paper, we  evaluate whether LSTM models are able to use that structure to guide interpretation, 
@@
 
-%\paragraph{Related work}
-%Recent successes on parsing by 
 Our results engage with those of \newcite{vinyals2014grammar} and \newcite{dyer2015transition}, who find that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks.
@@
 However, the massive size of the dataset needed to train the model---250M tokens---leaves open the possibility that it primarily learned to generate only tree structures that it had already seen, representing them as simple hashes---which would not capture unseen tree structures---rather than as structured objects.
-% CDM: I wonder about just leaving this next sentence out
-% If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without actually representing recursive structure, at least in a way that would generalize to substantially new structures.
-Our experiments, though, show that LSTMs can manipulate tree-structured data, suggesting that there is no fundamental obstacle to this kind of generalization.
+Our experiments, though, show that LSTMs can learn to understand tree structures when given enough data, suggesting that there is no fundamental obstacle to this kind of generalization.

### 2015-06-16 writing/S'15 LSTM/intro.tex
Final clarifications.

@@
 The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) with high accuracy using only the information present in the vector.
-This shows that the LSTM was able to identify the correct syntactic structures and also hints that it was able to develop a generalizable method for encoding these structures in vectors.
-However, the massive size of the dataset needed to train the model---250M tokens---leaves open the possibility that it primarily learned to generate only tree structures that it had already seen, representing them as simple hashes---which would not capture unseen tree structures---rather than as structured objects.
-Our experiments, though, show that LSTMs can learn to understand tree structures when given enough data, suggesting that there is no fundamental obstacle to this kind of generalization.
+This shows that the LSTM is able to identify the correct syntactic structures and also hints that it is able to develop a generalizable method for encoding these structures in vectors.
+However, the massive size of the dataset needed to train the model---250M tokens---leaves open the possibility that it primarily learns to generate only tree structures that it has already seen, representing them as simple hashes---which would not capture unseen tree structures---rather than as structured objects.
+Our experiments, though, show that LSTMs can learn to understand tree structures when given enough data, suggesting that there is no fundamental obstacle to learning this kind of structured representation.

### 2015-06-16 writing/S'15 LSTM/methods.tex
Final clarifications.

@@
 The horizontal axis on each graph divides the test set expression pairs into bins by the number of logical operators in the more complex of the two expressions in the pair.
-The dotted line shows the size we trained on up to.}
+The dotted line shows the maximum size in each training set.}
   \label{prop-results} 
@@
 In addition, we use a sequence model (Figure~\ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}.
-Experiments with simpler non-LSTM RNN sequence models tended to underfit the training data, and are not included here.
+Experiments with simpler non-LSTM RNN sequence models tended to badly underfit the training data, and are not included here.
 

### 2015-06-16 writing/S'15 LSTM/data.tex
Longer example.
Final clarifications.

--- 2015-06-15 writing/S'15 LSTM/data.tex
@@
 %      $\plneg \plneg p_6$ & $\nateq$  & $p_6$ \\
-      $p_3$               & $\natfor$ & $(p_3 \plor p_2)$ \\
-      $(p_1 \plor (p_2 \plor p_4))$               & $\natrev$ & $(p_2 \pland  \plneg p_4)$ \\
+      $p_3$               & $\natfor$ & $p_3 \plor p_2$ \\
+%      $p_1 \plor (p_2 \plor p_4)$               & $\natrev$ & $p_2 \pland  \plneg p_4$ \\
+      $(\plneg p_2) \pland p_6  $ & $\natalt$ & $     \plneg ( p_6 \plor ( p_5 \plor p_3 ) ) $\\
       %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
-      $\plneg\, (\plneg p_1 \pland \plneg p_2)$ & $\nateq$ & $(p_1 \plor p_2)$ \\ 
-      $(p_3 \pland \plneg p_1 ) \plor \plneg p_3$    & $\natrev$& $\plneg\, (p_3 \plor p_2)$ \\
+ $    p_4 \plor ( \plneg ( ( p_1   $ & $ \natfor $ & $  \plneg ( ( ( ( \plneg p_6 ) \plor ( \plneg p_4 ) )   $\\      
+$ \plor p_6 )  \plor p_4 ) )$ &&$ \pland ( \plneg p_5 ) )\pland ( p_6 \pland p_6 ) ) $\\
+% $(p_3 \pland \plneg p_1 ) \plor \plneg p_3$    & $\natrev$& $\plneg\, (p_3 \plor p_2)$ \\
       %<	( not ( c ( or b ) ) )	( ( c ( and ( not a ) ) ) ( or ( not c ) ) )
@@
     \end{tabular}
-    \caption{Examples of short expressions from the artificial data introduced by \protect\citealt{Bowman:Potts:Manning:2014}.
+    \caption{Examples of short to moderate length pairs from the artificial data introduced by \protect\citealt{Bowman:Potts:Manning:2014}.
 For brevity, we only show the parentheses that are needed to disambiguate the sentences rather than the full binary bracketings.}\label{tab:plexs}
@@
 Capping the size of the training sentences allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures.
-If their performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and L2 regularization (which helps overall performance) exacerbates this by preventing the model from using the kind of numerically precise, nonlinearity-saturating functions that would generalize best.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.
+If their performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and L2 regularization (which helps overall performance) exacerbates this by discouraging the model from using the kind of numerically precise, nonlinearity-saturating functions that generalize best.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-06-16 writing/S'15 LSTM/figure1.tex
Delete SAT instances; logging; paper push.
Learning curve
Paper submission candidate.
Final clarifications.

@@
 \centering
-\scalebox{0.75}{
+\scalebox{0.72}{
  \begin{tikzpicture}
@@
   \centering
-\scalebox{0.75}{
+\scalebox{0.72}{
  \begin{tikzpicture}
@@
 
-     \caption{The architecture for the tree-structured sentence models. Terminal nodes are learned embeddings and nonterminal nodes are NN or NTN layers.}\label{fig:model:tree}
+     \caption{The architecture for the tree-structured sentence models. Terminal nodes are learned embeddings and nonterminal nodes are NN, NTN, or TreeLSTM layers.}\label{fig:model:tree}
 
@@
 
-   \caption{The architecture for the LSTM sentence model. Nodes in the lower row are learned embeddings and nodes on the upper row are LSTM units.}\label{fig:model:seq}
+   \caption{The architecture for the sequence sentence model. Nodes in the lower row are learned embeddings and nodes on the upper row are LSTM layers.}\label{fig:model:seq}
 

### 2015-06-16 writing/S'15 LSTM/arxiv-package/intro.tex
Anon and non-anon sharable versions.

--- 2015-06-16 writing/S'15 LSTM/intro.tex
@@
 We compare standard tree and sequence models on their handling of recursive structure by training the models on sentences whose length and recursion depth are limited, and then testing them on longer and more complex sentences, such that only models that exploit the recursive structure will be able to generalize in a way that yields correct interpretations for these test sentences.
-Our methods are based on those of \newcite{Bowman:Potts:Manning:2014}, who describe an experiment and corresponding artificial dataset which tests this ability in two tree models.
-We adapt that experiment to sequence models by decorating the statements with an explicit bracketing, and we use this design to compare an LSTM sequence model with three tree models, with a focus on what data each model needs in order to learn the needed generalizations.
+Our methods are based on  \newcite{Bowman:Potts:Manning:2014}, in which we describe an experiment and corresponding artificial dataset which tests this ability in two tree models.
+We adapt that experiment here to sequence models by decorating the statements with an explicit bracketing, and we use this design to compare an LSTM sequence model with three tree models, with a focus on what data each model needs in order to learn the needed generalizations.
 

### 2015-06-16 writing/S'15 LSTM/arxiv-package/acl2015.tex
Anon and non-anon sharable versions.

%
% File acl2015.tex
%
% Contact: car@ir.hit.edu.cn, gdzhou@suda.edu.cn
%%
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl2015}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}

%\setlength\titlebox{5cm}

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\title{Instructions for ACL-2015 Proceedings}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  This document contains the instructions for preparing a camera-ready
  manuscript for the proceedings of ACL-2015. The document itself
  conforms to its own specifications, and is therefore an example of
  what your manuscript should look like. These instructions should be
  used for both papers submitted for review and for final versions of
  accepted papers.  Authors are asked to conform to all the directions
  reported in this document.
\end{abstract}

\section{Credits}

This document has been adapted from the instructions for earlier ACL
proceedings, including those for ACL-2012 by Maggie Li and Michael
White, those from ACL-2010 by Jing-Shing Chang and Philipp Koehn,
those for ACL-2008 by Johanna D. Moore, Simone Teufel, James Allan,
and Sadaoki Furui, those for ACL-2005 by Hwee Tou Ng and Kemal
Oflazer, those for ACL-2002 by Eugene Charniak and Dekang Lin, and
earlier ACL and EACL formats. Those versions were written by several
people, including John Chen, Henry S. Thompson and Donald
Walker. Additional elements were taken from the formatting
instructions of the {\em International Joint Conference on Artificial
  Intelligence}.

\section{Introduction}

The following instructions are directed to authors of papers submitted
to ACL-2015 or accepted for publication in its proceedings. All
authors are required to adhere to these specifications. Authors are
required to provide a Portable Document Format (PDF) version of their
papers. \textbf{The proceedings are designed for printing on A4
paper.}

We will make more detailed instructions available at
\url{http://acl2015.org/publication.html}. Please check this website 
regularly.

\section{General Instructions}

Manuscripts must be in two-column format.  Exceptions to the
two-column format include the title, authors' names and complete
addresses, which must be centered at the top of the first page, and
any full-width figures or tables (see the guidelines in
Subsection~\ref{ssec:first}). {\bf Type single-spaced.}  Start all
pages directly under the top margin. See the guidelines later
regarding formatting the first page.  The manuscript should be
printed single-sided and its length
should not exceed the maximum page limit described in Section~\ref{sec:length}.
Do not number the pages.

\subsection{Electronically-available resources}

We strongly prefer that you prepare your PDF files using \LaTeX\ with
the official ACL 2015 style file (acl2015.sty) and bibliography style
(acl.bst). These files are available at
\url{http://acl2015.org}. You will also find the document
you are currently reading (acl2015.pdf) and its \LaTeX\ source code
(acl2015.tex) on this website.

You can alternatively use Microsoft Word to produce your PDF file. In
this case, we strongly recommend the use of the Word template file
(acl2015.dot) on the ACL 2015 website (\url{http://acl2015.org}). 
If you have an option, we recommend that you use the \LaTeX2e version. 
If you will be using the Microsoft Word template, we suggest that you 
anonymize your source file so that the pdf produced does not retain your
identity.  This can be done by removing any personal information
from your source document properties.

\subsection{Format of Electronic Manuscript}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's
Portable Document Format (PDF). PDF files are usually produced from
\LaTeX\ using the \textit{pdflatex} command. If your version of
\LaTeX\ produces Postscript files, you can convert these into PDF
using \textit{ps2pdf} or \textit{dvipdf}. On Windows, you can also use
Adobe Distiller to generate PDF.

Please make sure that your PDF file includes all the necessary fonts
(especially tree diagrams, symbols, and fonts with Asian
characters). When you print or create the PDF file, there is usually
an option in your printer setup to include none, all or just
non-standard fonts.  Please make sure that you select the option of
including ALL the fonts. \textbf{Before sending it, test your PDF by
  printing it from a computer different from the one where it was
  created.} Moreover, some word processors may generate very large PDF
files, where each page is rendered as an image. Such images may
reproduce poorly. In this case, try alternative ways to obtain the
PDF. One way on some systems is to install a driver for a postscript
printer, send your document to the printer specifying ``Output to a
file'', then convert the file to PDF.

It is of utmost importance to specify the \textbf{A4 format} (21 cm
x 29.7 cm) when formatting the paper. When working with
{\tt dvips}, for instance, one should specify {\tt -t a4}.
Or using the command \verb|\special{papersize=210mm,297mm}| in the latex
preamble (directly below the \verb|\usepackage| commands). Then using 
{\tt dvipdf} and/or {\tt pdflatex} which would make it easier for some.

Print-outs of the PDF file on A4 paper should be identical to the
hardcopy version. If you cannot meet the above requirements about the
production of your electronic submission, please contact the
publication chairs as soon as possible.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts two columns to a page, in the manner these
instructions are formatted. The exact dimensions for a page on A4
paper are:

\begin{itemize}
\item Left and right margins: 2.5 cm
\item Top margin: 2.5 cm
\item Bottom margin: 2.5 cm
\item Column width: 7.7 cm
\item Column height: 24.7 cm
\item Gap between columns: 0.6 cm
\end{itemize}

\noindent Papers should not be submitted on any other paper size.
 If you cannot meet the above requirements about the production of 
 your electronic submission, please contact the publication chairs 
 above as soon as possible.

\subsection{Fonts}

For reasons of uniformity, Adobe's {\bf Times Roman} font should be
used. In \LaTeX2e{} this is accomplished by putting

\begin{quote}
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}
in the preamble. If Times Roman is unavailable, use {\bf Computer
  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
  10\% less dense than Adobe's Times Roman font.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|rl|}
\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
paper title & 15 pt & bold \\
author names & 12 pt & bold \\
author affiliation & 12 pt & \\
the word ``Abstract'' & 12 pt & bold \\
section titles & 12 pt & bold \\
document text & 11 pt  &\\
captions & 11 pt & \\
abstract text & 10 pt & \\
bibliography & 10 pt & \\
footnotes & 9 pt & \\
\hline
\end{tabular}
\end{center}
\caption{\label{font-table} Font guide. }
\end{table}

\subsection{The First Page}
\label{ssec:first}

Center the title, author's name(s) and affiliation(s) across both
columns. Do not use footnotes for affiliations. Do not include the
paper ID number assigned during the submission process. Use the
two-column format only when you begin the abstract.

{\bf Title}: Place the title centered at the top of the first page, in
a 15-point bold font. (For a complete guide to font sizes and styles,
see Table~\ref{font-table}) Long titles should be typed on two lines
without a blank line intervening. Approximately, put the title at 2.5
cm from the top of the page, followed by a blank line, then the
author's names(s), and the affiliation on the following line. Do not
use only initials for given names (middle initials are allowed). Do
not format surnames in all capitals (e.g., use ``Schlangen'' not
``SCHLANGEN'').  Do not format title and section headings in all
capitals as well except for proper names (such as ``BLEU'') that are
conventionally in all capitals.  The affiliation should contain the
author's complete address, and if possible, an electronic mail
address. Start the body of the first page 7.5 cm from the top of the
page.

The title, author names and addresses should be completely identical
to those entered to the electronical paper submission website in order
to maintain the consistency of author information among all
publications of the conference. If they are different, the publication
chairs may resolve the difference without consulting with you; so it
is in your own interest to double-check that the information is
consistent.

{\bf Abstract}: Type the abstract at the beginning of the first
column. The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by about
0.6 cm on each side. Center the word {\bf Abstract} in a 12 point bold
font above the body of the abstract. The abstract should be a concise
summary of the general thesis and conclusions of the paper. It should
be no longer than 200 words. The abstract text should be in 10 point font.

{\bf Text}: Begin typing the main body of the text immediately after
the abstract, observing the two-column format as shown in 
the present document. Do not include page numbers.

{\bf Indent} when starting a new paragraph. Use 11 points for text and 
subsection headings, 12 points for section headings and 15 points for
the title.

\subsection{Sections}

{\bf Headings}: Type and label section and subsection headings in the
style shown on the present document.  Use numbered sections (Arabic
numerals) in order to facilitate cross references. Number subsections
with the section number and the subsection number separated by a dot,
in Arabic numerals. Do not number subsubsections.

{\bf Citations}: Citations within the text appear in parentheses
as~\cite{Gusfield:97} or, if the author's name appears in the text
itself, as Gusfield~\shortcite{Gusfield:97}.  Append lowercase letters
to the year in cases of ambiguity.  Treat double authors as
in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two
authors are involved. Collapse multiple citations as
in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations
as sentence constituents. We suggest that instead of
\begin{quote}
  ``\cite{Gusfield:97} showed that ...''
\end{quote}
you use
\begin{quote}
``Gusfield \shortcite{Gusfield:97}   showed that ...''
\end{quote}

If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
can use the command \verb|\newcite| to get ``author (year)'' citations.

As reviewing will be double-blind, the submitted version of the papers
should not include the authors' names and affiliations. Furthermore,
self-references that reveal the author's identity, e.g.,
\begin{quote}
``We previously showed \cite{Gusfield:97} ...''  
\end{quote}
should be avoided. Instead, use citations such as 
\begin{quote}
``Gusfield \shortcite{Gusfield:97}
previously showed ... ''
\end{quote}

\textbf{Please do not use anonymous citations} and do not include
acknowledgements when submitting your papers. Papers that do not
conform to these requirements may be rejected without review.

\textbf{References}: Gather the full set of references together under
the heading {\bf References}; place the section before any Appendices,
unless they contain references. Arrange the references alphabetically
by first author, rather than by order of occurrence in the text.
Provide as complete a citation as possible, using a consistent format,
such as the one for {\em Computational Linguistics\/} or the one in the 
{\em Publication Manual of the American 
Psychological Association\/}~\cite{APA:83}.  Use of full names for
authors rather than initials is preferred.  A list of abbreviations
for common computer science journals can be found in the ACM 
{\em Computing Reviews\/}~\cite{ACM:83}.

The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
American Psychological Association format, allowing regular citations, 
short citations and multiple citations as described above.

{\bf Appendices}: Appendices, if any, directly follow the text and the
references (but see above).  Letter them in sequence and provide an
informative title: {\bf Appendix A. Title of Appendix}.

\subsection{Footnotes}

{\bf Footnotes}: Put footnotes at the bottom of the page and use 9
points text. They may be numbered or referred to by asterisks or other
symbols.\footnote{This is how a footnote should appear.} Footnotes
should be separated from the text by a line.\footnote{Note the line
separating the footnotes from the text.}

\subsection{Graphics}

{\bf Illustrations}: Place figures, tables, and photographs in the
paper near where they are first discussed, rather than at the end, if
possible.  Wide illustrations may run across both columns.  Color
illustrations are discouraged, unless you have verified that  
they will be understandable when printed in black ink.

{\bf Captions}: Provide a caption for every illustration; number each one
sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
Caption of the Table.''  Type the captions of the figures and 
tables below the body, using 11 point text.

\section{XML conversion and supported \LaTeX\ packages}

Following ACL 2014 we will also we will attempt to automatically convert 
your \LaTeX\ source files to publish papers in machine-readable 
XML with semantic markup in the ACL Anthology, in addition to the 
traditional PDF format.  This will allow us to create, over the next 
few years, a growing corpus of scientific text for our own future research, 
and picks up on recent initiatives on converting ACL papers from earlier 
years to XML.

We encourage you to submit a ZIP file of your \LaTeX\ sources along
with the camera-ready version of your paper. We will then convert them
to XML automatically, using the LaTeXML tool
(\url{http://dlmf.nist.gov/LaTeXML}). LaTeXML has \emph{bindings} for
a number of \LaTeX\ packages, including the ACL 2015 stylefile. These
bindings allow LaTeXML to render the commands from these packages
correctly in XML. For best results, we encourage you to use the
packages that are officially supported by LaTeXML, listed at
\url{http://dlmf.nist.gov/LaTeXML/manual/included.bindings}

\section{Translation of non-English Terms}

It is also advised to supplement non-English characters and terms
with appropriate transliterations and/or translations
since not all readers understand all such characters and terms.
Inline transliteration or translation can be represented in
the order of: original-form transliteration ``translation''.

\section{Length of Submission}
\label{sec:length}

Long papers may consist of up to 8 pages of content, plus two extra
pages for references. Short papers may consist of up to 4 pages of
content, plus two extra pages for references.  Papers that do not
conform to the specified length and formatting requirements may be
rejected without review.

\section*{Acknowledgments}

The acknowledgments should go immediately before the references.  Do
not number the acknowledgments section. Do not include this section
when submitting your paper for review.

% include your own bib file like this:
%\bibliographystyle{acl}
%\bibliography{acl2015}

\begin{thebibliography}{}

\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
Alfred~V. Aho and Jeffrey~D. Ullman.
\newblock 1972.
\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
\newblock Prentice-{Hall}, Englewood Cliffs, NJ.

\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
{American Psychological Association}.
\newblock 1983.
\newblock {\em Publications Manual}.
\newblock American Psychological Association, Washington, DC.

\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
{Association for Computing Machinery}.
\newblock 1983.
\newblock {\em Computing Reviews}, 24(11):503--512.

\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
Ashok~K. Chandra, Dexter~C. Kozen, and Larry~J. Stockmeyer.
\newblock 1981.
\newblock Alternation.
\newblock {\em Journal of the Association for Computing Machinery},
  28(1):114--133.

\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
Dan Gusfield.
\newblock 1997.
\newblock {\em Algorithms on Strings, Trees and Sequences}.
\newblock Cambridge University Press, Cambridge, UK.

\end{thebibliography}

\end{document}

### 2015-06-18 writing/F'14 paper/cameraready/join.tex
Cleanup for workshop camera ready.

@@
 sentence representations, any successful reasoning using compositional representations
-will rely on the ability to perform sound inferences of this kind, so our first experiment studies how well each model can learn to perform them them in isolation.
+will rely on the ability to perform sound inferences of this kind in order to be able to use unseen relational facts within larger derivations.
+Our first experiment studies how well each model can learn to perform them them in isolation.
 
@@
 % $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.
-
-\paragraph{Experiments}
-We begin by creating a world model
-on which we will base the statements in the train and test sets.
-This takes the form of a small Boolean structure in which terms denote
-sets of entities from a small domain.  Fig.~\ref{lattice-figure}a
-depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight proposition terms ($p_1$--$p_8$). We then generate a 
-relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
-We divide these statements evenly into train and test sets, and delete the test set
- examples which cannot be proven from the train examples, for which there is not enough information for even an ideal system to choose a correct label.
-In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 examples and a test set of 
-2960 examples.
-
-We trained models with both the NN and NTN comparison functions on these
-data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no simpler version of our model is a viable baseline.} %+%
-In both cases, the models are implemented as
-described in \S\ref{methods}, but since the items being compared
-are single terms rather than full tree structures, the composition
-layer is not used, and the two models are not recursive. We simply present
-the models with the (randomly initialized) embedding vectors for each
-of two terms, ensuring that the model has no information about the terms
-being compared except for the relations between them that appear in training.
 
@@
     \end{picture}}
-    \caption{Example boolean structure. The terms $p_1$--$p_8$ name the sets. Not all sets have names, and  some sets have multiple names, so that learning $\nateq$ is non-trivial.}
+    \caption{Example boolean structure, shown with edges idicating inclusion.
+The terms $p_1$--$p_8$ name the sets.
+Not all sets have names, and  some sets have multiple names, so that learning $\nateq$ is non-trivial.}
   \end{subfigure}
@@
     \caption{A few examples of atomic statements about the
-      model.  Test statements that are not provable from the training data shown are
+      model depicted above.  Test statements that are not provable from the training data shown are
       crossed out.}
@@
 
+\paragraph{Experiments}
+We begin by creating a world model
+on which we will base the statements in the train and test sets.
+This takes the form of a small Boolean structure in which terms denote
+sets of entities from a small domain.  Fig.~\ref{lattice-figure}a
+depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight proposition terms ($p_1$--$p_8$). We then generate a 
+relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
+We divide these statements evenly into train and test sets, and delete the test set
+ examples which cannot be proven from the train examples, for which there is not enough information for even an ideal system to choose a correct label.
+In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 examples and a test set of 
+2960 examples.
+
+We trained models with both the NN and NTN comparison functions on these
+data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no simpler version of our model is a viable baseline.} %+%
+In both cases, the models are implemented as
+described in \S\ref{methods}, but since the items being compared
+are single terms rather than full tree structures, the composition
+layer is not used, and the two models are not recursive. We simply present
+the models with the (randomly initialized) embedding vectors for each
+of two terms, ensuring that the model has no information about the terms
+being compared except for the relations between them that appear in training.
+
 \paragraph{Results} 
@@
 It is not possible for us to rule out the possibility that different optimization techniques or
-further hyperparameter tuning could lead an NN model to succeed here.
+finer-grained hyperparameter tuning could lead an NN model to succeed.
 

### 2015-06-18 writing/F'14 paper/cameraready/methods.tex
Cleanup for workshop camera ready.

@@
 via specific composition functions \cite{Partee84,Janssen97}. In our
-distributed setting, word meanings are embedding vectors of dimension $n$. A learned
-composition function maps pairs of them to single phrase vectors of dimension $n$, 
+distributed setting, word meanings are embedding vectors of dimension $N$. A learned
+composition function maps pairs of them to single phrase vectors of dimension $N$, 
 which can then be merged again to represent more complex
@@
 $\vec{y}$ is the node's output.  The TreeRNN concatenates them, multiplies
-them by an $n \times 2n$ matrix of learned weights, and adds a bias $\vec{b}$. 
+them by an $N \times 2N$ matrix of learned weights, and adds a bias $\vec{b}$. 
 The TreeRNTN adds a learned full rank third-order tensor 
-$\mathbf{T}$, of dimension $n \times n \times n$, modeling
+$\mathbf{T}$, of dimension $N \times N \times N$, modeling
 multiplicative interactions between the child vectors. 
@@
 where it is not defined.}
-Source code and generated data will be released after the review period.
+Source code and generated data can be downloaded from \url{http://stanford.edu/~sbowman/}.

### 2015-06-18 writing/F'14 paper/cameraready/sick.tex
Cleanup for workshop camera ready.

@@
 The corpus consists of about 10k natural language sentence pairs, labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}.
-At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of, and our results nonetheless show that tree-structured NN models can learn to do inference in the real world.
+At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of, and our results nonetheless show that tree-structured NN models can learn to approximate natural logic-style inference in the real world.
 
@@
 Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer.
-This new layer aggregates any usable information from the embedding vectors into a more compact working representation.
+This new layer allows the model to choose which aspects of the 200 dimensional representations from the unsupervised source it most values, rather than relying on GloVe---which is has no knowledge of the task---to do so, as would be the case were GloVe asked to directly produce vectors of the lower dimensionality.
 An identical layer is added to the SumNN between the word vectors and the comparison layer.
 
+\begin{table*}[tp]
+  \centering\small
+  \begin{tabular}{l@{~~~}cl}
+    \toprule
+  The patient is being helped by the doctor	& \ii{entailment} & The doctor is helping the patient (\textsc{Passive})\\
+    A little girl is playing the violin on a beach & \ii{contradiction} &	There is no girl playing the violin on a beach (\textsc{Neg})\\
+
+    The yellow dog is drinking water from a bottle& \ii{contradiction} &	The yellow dog is drinking water from a pot  (\textsc{Subst})\\
+        A woman is breaking two eggs in a bowl & \ii{neutral} &A man is mixing a few ingredients in a bowl (\textsc{MultiEd})\\
+        Dough is being spread by a man & \ii{neutral} & A woman is slicing meat with a knife (\textsc{Diff})\\
+    \bottomrule
+  \end{tabular}
+  \caption{\label{examplesofsickdata}Examples of each category used in error analysis from the SICK test data. }
+\end{table*}
+
 We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the 4,928 example SICK test set.
-We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}, also used by the winning SICK submission), a corpus of noisy automatically labeled entailment examples over image captions, the same genre of text from which SICK was drawn.
-We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source.
+We also report training accuracy on a small sample from each data source.} (4500 examples) with 600k examples of approximate entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}, also used by the winning SICK submission), a corpus of noisy automatically labeled entailment examples over image captions, the same genre of text from which SICK was drawn.
+We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source, and forced the model to sample SICK examples and DG examples about equally often during training.
+
 We parsed the data from both sources with the Stanford PCFG Parser v.~3.3.1 \cite{klein2003accurate}.
@@
 
-\begin{table*}[htp]
-  \centering\small
-  \begin{tabular}{l@{~~~}cl}
-    \toprule
-  The patient is being helped by the doctor	& \ii{entailment} & The doctor is helping the patient (\textsc{Passive})\\
-    A little girl is playing the violin on a beach & \ii{contradiction} &	There is no girl playing the violin on a beach (\textsc{Neg})\\
-
-    The yellow dog is drinking water from a bottle& \ii{contradiction} &	The yellow dog is drinking water from a pot  (\textsc{Subst})\\
-        A woman is breaking two eggs in a bowl & \ii{neutral} &A man is mixing a few ingredients in a bowl (\textsc{MultiEd})\\
-        Dough is being spread by a man & \ii{neutral} & A woman is slicing meat with a knife (\textsc{Diff})\\
-    \bottomrule
-  \end{tabular}
-  \caption{\label{examplesofsickdata}Examples of each category used in error analysis from the SICK test data. }
-\end{table*}
-
 \paragraph{Results} Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships, the results (Table~\ref{sresultstable}) show that our tree-structured models perform competitively on textual entailment, beating a strong baseline.

### 2015-06-18 writing/F'14 paper/cameraready/intro.tex
Cleanup for workshop camera ready.

--- 2015-05-18 writing/F'14 paper/arxiv/intro.tex
@@
 
-Tree-structured recursive neural network models (TreeRNNs; \citealt{goller1996learning}) for sentence meaning
+Tree-structured recursive neural network models (TreeRNNs; \citealt{goller1996learning,socher2011semi}) for sentence meaning
 have been successful in an array of sophisticated language tasks,

### 2015-06-18 writing/F'14 paper/cameraready/quantifiers.tex
Cleanup for workshop camera ready.

@@
 In each run, we randomly partition the set of valid \textit{single sentences} into train and test, and then label all of the pairs from within each set to generate a training set of 27k pairs and a test set of 7k pairs.
-Because the model doesn't see the test sentences at training time, it cannot directly use the kind of reasoning described in \S\ref{sec:join} (treating sentences as unanalyzed symbols), and must instead infer the word-level relations and learn a complete reasoning system over them for our logic.
+Because the model doesn't see the test sentences at training time, it cannot directly use the kind of reasoning described in \S\ref{sec:join} at the sentence level (by treating sentences as unanalyzed symbols), and must instead jointly learn the word-level relations and a complete reasoning system over them for our logic.
 

### 2015-07-14 writing/S'15 LSTM/arxiv-package/discussion.tex
Anon and non-anon sharable versions.
Updated results for LSTM paper.

--- 2015-06-16 writing/S'15 LSTM/discussion.tex
@@
 However, the LSTM improves considerably with more ample training data in the $\le$6 condition, showing only a 3\% drop and generalization results better than the best model's in the $\le$3 setting.
-% TODO: These stats?
 
@@
 
-The learning curve (Figure~\ref{fig:lc}) suggests that additional data is unlikely to change these basic results, though the disparity in performance at the 11\% level is striking: the LSTM sequence model and the TreeRNTN (the model with the most parameters by far) lag far behind the other two.
+The learning curve (Figure~\ref{fig:lc}) suggests that additional data is unlikely to change these basic results.
+The learning curve (Figure~\ref{fig:lc}) suggests that additional data is unlikely to change these basic results.
+The LSTM lags behind the tree models across the curve, but appears to gain accuracy at a similar rate.
 

### 2015-07-14 writing/S'15 LSTM/discussion.tex
Updated results for LSTM paper.

@@
 
-The learning curve (Figure~\ref{fig:lc}) suggests that additional data is unlikely to change these basic results, though the disparity in performance at the 11\% level is striking: the LSTM sequence model and the TreeRNTN (the model with the most parameters by far) lag far behind the other two.
+The learning curve (Figure~\ref{fig:lc}) suggests that additional data is unlikely to change these basic results.
+The LSTM lags behind the tree models across the curve, but appears to gain accuracy at a similar rate.
 

### 2015-07-14 writing/S'15 LSTM/methods.tex
Updated results for LSTM paper.

@@
   \begin{subfigure}[t]{0.04\textwidth}
-      \includegraphics[height=1.5in]{scale.pdf}
+      \includegraphics[height=1.4in]{scale.pdf}
 \end{subfigure}
 \begin{subfigure}[t]{0.24\textwidth}
-  \includegraphics[height=1.5in]{fig3c.pdf}
+  \includegraphics[height=1.4in]{fig3c.pdf}
   \caption{Training on sz. $\le$3.}
@@
 \begin{subfigure}[t]{0.24\textwidth}
-    \includegraphics[height=1.5in]{fig4c.pdf}
+    \includegraphics[height=1.4in]{fig4c.pdf}
   \caption{Training on sz. $\le$4.}
@@
 \begin{subfigure}[t]{0.24\textwidth}
-      \includegraphics[height=1.5in]{fig6c.pdf}
+      \includegraphics[height=1.4in]{fig6c.pdf}
   \caption{Training on sz. $\le$6.}
@@
 \begin{subfigure}[t]{0.08\textwidth}
-      \includegraphics[height=1.5in]{leg.pdf}
+      \includegraphics[height=1.4in]{leg.pdf}
 \end{subfigure}
@@
 Our objective is the standard negative log likelihood classification objective with L2 regularization (tuned on a separate train/test split).
-All models were trained for 72 epochs, after which all had largely converged without significantly declining from their peak performances.
+All models were trained for 100 epochs, after which all had largely converged without significantly declining from their peak performances.\todo{Replot learning curve, resummarize numbers.}

### 2015-08-07 writing/SNLI paper/abstract.tex
Remove newlines from abstract for ease of copying.
First pass of SNLI camera ready revisions.

@@
 However, machine learning research in this area has been dramatically limited by the lack of large-scale resources.
-To address this, we introduce a new freely available corpus of labeled sentence pairs, written by humans doing a novel grounded task based on image captioning.
+To address this, we introduce the Stanford Natural Language Inference corpus, a new freely available collection of labeled sentence pairs, written by humans doing a novel grounded task based on image captioning.
 At 570K pairs, it is two orders of magnitude larger than all other resources of its type.

### 2015-08-07 writing/SNLI paper/corpus.tex
First pass of SNLI camera ready revisions.

@@
 limits their utility as a testbed for learned distributed representations. 
-The data for the SemEval 2014 task called Sentences Involving Compositional Knowledge (SICK) is a
-step up in terms of size, but only to 4,500 training examples, and its
+The data for the SemEval 2014 task called Sentences Involving Compositional Knowledge (SICK) is a step up in terms of size, but only to 4,500 training examples, and its
 partly automatic construction introduced some spurious patterns into
-the data (\citealt{marelli2014semeval}, $\S$6). Finally, the
+the data (\citealt{marelli2014semeval}, $\S$6). The
 Denotation Graph entailment set \cite{hodoshimage} contains millions of
@@
 noisy enough that it is probably suitable only as a source of
-supplementary training data. The Paraphase Database of \newcite{ganitkevitch2013ppdb}
-addresses these issues of size and quality, but only for the limited
-fragment of the NLI problem that is captured by paraphrase.
+supplementary training data. 
+Outside the domain of sentence-level entailment, \newcite{levy2014focused} introduce
+a large corpus of semi-automatically annotated entailment examples 
+between subject--verb--object relation triples, and the second release of the 
+Paraphase Database \cite{ganitkevitch2ppdb} includes automatically generated entailment annotations
+over a large corpus of pairs of words and short phrases.
 
@@
 
-With our corpus, we sought to address the issues of size, quality, and
+With SNLI, we sought to address the issues of size, quality, and
 indeterminacy. To do this, we employed a crowdsourcing framework with
@@
 We observed that while premise sentences varied considerably in length, hypothesis sentences tended to be as short as possible while still providing enough information to yield a clear judgment, clustering at around seven words.
+We also observed that the bulk of the sentences from both sources were syntactically complete rather than fragments, and the frequency with which the parser produces a parse rooted with an `S' (sentence) node attests to this.
 
@@
 % From 1.0rc3
-\caption{\label{collection-stats}Key statistics for the raw sentence pairs. Since the two halves of each pair were collected separately, we report some statistics for both.} 
+\caption{\label{collection-stats}Key statistics for the raw sentence pairs in SNLI. Since the two halves of each pair were collected separately, we report some statistics for both.} 
 \end{table}
@@
 
+\todo{Upload SNLI and add a URL}
+
 \paragraph{Partition}

### 2015-08-07 writing/SNLI paper/entailment_eval.tex
First pass of SNLI camera ready revisions.

@@
 
-% The most immediate application for our corpus is in developing models for the task of NLI.
-In particular, since it is dramatically larger than any existing corpus of comparable quality, we expect it to be suitable for training  parameter-rich models like neural networks for this task.
+The most immediate application for our corpus is in developing models for the task of NLI.
+In particular, since it is dramatically larger than any existing corpus of comparable quality, we expect it to be suitable for training parameter-rich models like neural networks, which have not previously been competetive at this task.
 In this section, we explore the performance of a range of standard and novel models trained on the corpus.
@@
 \toprule
-\b{System} & \b{Our corpus} & \b{SICK} & \b{RTE-3} \\
+\b{System} & \b{SNLI} & \b{SICK} & \b{RTE-3} \\
 \midrule
@@
   2-class entailment; all RTE results are likewise reported for 2-class entailment.
-To convert 3-class problems like SICK and our corpus to this setting, all instances
+To convert 3-class problems like SICK and SNLI to this setting, all instances
   of \contradiction\ and \unknown\ are converted to nonentailment.
-This yields a chance baseline accuracy of 66\% on our corpus, and 71\% on SICK.
+This yields a chance baseline accuracy of 66\% on SNLI, and 71\% on SICK.
 We train these models without using any of the included lexical
@@
 \paragraph{Lexicalized Classifier}
-Unlike the RTE datasets, our corpus's size supports approaches which make use of rich lexicalized features.
+Unlike the RTE datasets, SNLI's size supports approaches which make use of rich lexicalized features.
 We evaluate a simple lexicalized classifier to explore the ability of non-specialized models to exploit these features in lieu of more involved language understanding.
@@
 \toprule
-\b{System}	 & \multicolumn{2}{c}{\hspace{-1.2em}\b{Our corpus}} & \b{SICK}\\
+\b{System}	 & \multicolumn{2}{c}{\hspace{-1.2em}\b{SNLI}} & \b{SICK}\\
  & \t{Train} & \t{Test} & \t{Test}\\
@@
   (Feature \ref{lst:bigram}), and lacking all lexical
-  features (Features \ref{lst:ngram}--\ref{lst:bigram}).
+  features (Features \ref{lst:ngram}--\ref{lst:bigram}). For the model trained on SNLI, we report results both on the test set and the training set to judge overfitting.
 }
@@
 
-Our corpus is suitably large and diverse to make it possible to train neural network models that produce distributed representations of sentence meaning.
+SNLI is suitably large and diverse to make it possible to train neural network models that produce distributed representations of sentence meaning.
 In this section, we compare the performance of three such models on the corpus.
@@
 This learning curve shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance for both.
-In addition, though the LSTM and the lexicalized model show similar performance when trained on the current full corpus, the steeper slope for the LSTM suggests that its ability to learn arbitrarily structured representations of sentence meaning would give it an advantage over the more constrained lexicalized model on still larger datasets.
+In addition, though the LSTM and the lexicalized model show similar performance when trained on the current full corpus, the somewhat steeper slope for the LSTM hints that its ability to learn arbitrarily structured representations of sentence meaning may give it an advantage over the more constrained lexicalized model on still larger datasets.
 
@@
 
-To the extent that successfully training a neural network model like our LSTM on our corpus forces that model to encode broadly accurate representations of English scene descriptions and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
+To the extent that successfully training a neural network model like our LSTM on SNLI forces that model to encode broadly accurate representations of English scene descriptions and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
 In this section, we evaluate on the SICK entailment task using a simple transfer learning method and achieve competitive results.
@@
 LSTM 3-class accuracy on the SICK train and test sets under three training regimes.} 
+% TODO (Gabor): Report training accuracy for the model trained on SICK.
 \end{table}
 
-To perform transfer, we take the parameters of the LSTM RNN model trained on our corpus and use them to initialize a new model, which is trained from that point only on the training portion of SICK.
-The only newly initialized parameters are softmax layer parameters and the embeddings for words that appear in SICK, but not in our corpus (which are populated with GloVe embeddings as above).
+To perform transfer, we take the parameters of the LSTM RNN model trained on SNLI and use them to initialize a new model, which is trained from that point only on the training portion of SICK.
+The only newly initialized parameters are softmax layer parameters and the embeddings for words that appear in SICK, but not in SNLI (which are populated with GloVe embeddings as above).
 We use the same model hyperparameters that were used to train the original model, with the exception of the L2 regularization strength, which is re-tuned.
@@
 The results are shown in \reftab{tab:transferresults}.
-Training on SICK alone yields poor performance, and the model trained on our data alone misunderstands the subtly different meanings of the (identically named) SICK labels, labeling more \ii{neutral} examples as \ii{contradiction}s than correctly.
+Training on SICK alone yields poor performance, and the model trained on SNLI fails when tested on SICK data, labeling more \ii{neutral} examples as \ii{contradiction}s than correctly, possibly as a result of subtle differences in the label definitions.
 Transferring representations, in contrast, yields the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which likely represents an approximate performance ceiling.

### 2015-08-07 writing/SNLI paper/intro.tex
First pass of SNLI camera ready revisions.

@@
 
-To address this, this paper introduces a new corpus of sentence pairs labeled for
+To address this, this paper introduces the Stanford Natural Language
+Inference (SNLI) corpus, a collection of sentence pairs labeled for
 entailment, contradiction, and semantic independence. At 570,152
-sentence pairs, this corpus is two orders of magnitude larger than all
+sentence pairs, SNLI is two orders of magnitude larger than all
 other resources of its type. And, in contrast to many such resources,

### 2015-08-07 writing/SNLI paper/transfer.tex
Bib updates, params.
Numbers; configs.
First pass of SNLI camera ready revisions.

@@
 We argue that for a sentence representation to provide the information necessary to reliably perform NLI classification, it must contain a faithful and thorough description of the sentence's meaning.
-This claim has a readily testable consequence: the knowledge that a representation learning system learns when it is trained on our corpus should transfer well to any other task that involves understanding sentence meaning, at least to the extent that the style of language used in that task is broadly similar to the style of our corpus.
-To test this, we perform transfer learning experiments on four other tasks, in which models are initialized using the parameters of an LSTM RNN-based model trained on our corpus (as in \S\ref{sentence-embedding}) and then trained to convergence on the target tasks.
+This claim has a readily testable consequence: the knowledge that a representation learning system learns when it is trained on SNLI should transfer well to any other task that involves understanding sentence meaning, at least to the extent that the style of language used in that task is broadly similar to the style of SNLI.
+To test this, we perform transfer learning experiments on four other tasks, in which models are initialized using the parameters of an LSTM RNN-based model trained on SNLI (as in \S\ref{sentence-embedding}) and then trained to convergence on the target tasks.
 
@@
 \hline
-\t{SICK}            & \t{100} & \t{99} & \t{62.9?} & \t{79.8?} & \t{84.5} \\
+\t{SICK}            & \t{100} & \t{99} & \t{71.3?} & \t{80.1?} & \t{84.5} \\
 \t{SUBJ}          & \t{??.?} & \t{??.?} & \t{90?} & \t{91?}& \t{95.5} \\
@@
 \caption{\label{tab:transferresults}
-The performance of an LSTM RNN model on four tasks under both a random initialization and a transfer initialization copied from a model trained on our corpus.
+The performance of an LSTM RNN model on four tasks under both a random initialization and a transfer initialization copied from a model trained on SNLI.
 All figures are classification accuracy, except for PragBank, where report macroaveraged F1 to enable comparison with existing results on that task.
@@
 
-The results are shown in Table~\ref{tab:transferresults}.
+The results are shown in \tabref{tab:transferresults}.
 For each task, the transfer model is compared with a model that was initialized randomly as normal.

### 2015-08-12 writing/SNLI paper/corpus.tex
Add data link; clarification.

@@
 
-The data will be released upon publication under a CreativeCommons
+The corpus is available under a CreativeCommons
 Attribution-ShareAlike license, the same license used for the Flickr30k source captions.
-
-\todo{Upload SNLI and add a URL}
+It can be downloaded at:\\\href{http://nlp.stanford.edu/projects/snli/}{\texttt{nlp.stanford.edu/projects/snli/}}
 

### 2015-08-12 writing/SNLI paper/intro.tex
Add data link; clarification.

@@
   \end{tabular}
-  \caption{\label{snli-examples}Randomly chosen examples from the development section of our new corpus, shown with both the gold labels and the full set of annotator labels (abbreviated).}
+  \caption{\label{snli-examples}Randomly chosen examples from the development section of our new corpus, shown with both the selected gold labels and the full set of labels (abbreviated) from the individual annotators, including (in the first position) the label used by the initial author of the pair.}
 \end{table*}

### 2015-08-12 writing/SNLI paper/corpus.tex
Revision to coreference para

@@
 event and entity coreference lead to insurmountable indeterminacy
-concerning the correct semantic label (\citealt{de2008finding} $\S4.3$; \citealt{marelli2014sick}). For
-instance, the sentence pair \word{A boat sank in the Pacific
-  ocean} and \word{A boat sank in an ocean} could reasonably be labeled in
-numerous ways depending on how the objects and events are grounded. If
-the two boats are assumed to be different, then the sentences are semantically
-independent. If the boats and the oceans are assumed to be the same,
-then the first sentence entails the second (in the relevant
-commonsense way). If the boats are assumed to be the same and 
-\word{an ocean} in the second sentence is interpreted with a specific referent 
-that is not the Pacific, then the two events are (commonsense) contradictory. This
-kind of indeterminacy can be resolved only once the questions of
-coreference are resolved.
+concerning the correct semantic label (\citealt{de2008finding} $\S4.3$; \citealt{marelli2014sick}). 
+For instance, the sentence pair \word{A boat sank in the Pacific Ocean} and \word{A boat sank in the Atlantic Ocean} could be labeled as a contradiction if the annotator assumes that the two events under discussion are coreferent, but could also be reasonably labeled as neutral if that assumption is not made.
+For a similar case with potentially different inutitions, \ii{Ruth Bader Ginsburg was appointed to the US Supreme Court} and \ii{I had a sandwich for lunch today} could be labeled as neutral if the annotator allows that they need not be refer to the same event, but is necessarily a contradiction if the annotator takes coreference to be required.
+Entity coreference presents a similar kind of uncertainty, as in the pair \word{A tourist visited New York} and \word{A tourist visited the city}.
+Assuming coreference between \ii{New York} and \ii{the city} justifies labeling the pair as an entailment, but without that assumption \ii{the city} could be taken to refer to a specific unknown city, leaving the pair neutral.
+
+% If
+% the two boats are assumed to be different, then the sentences are semantically
+% independent. If the boats and the oceans are assumed to be the same,
+%then the first sentence entails the second (in the relevant
+%commonsense way). If the boats are assumed to be the same and 
+%\word{an ocean} in the second sentence is interpreted with a specific referent 
+%that is not the Pacific, then the two events are (commonsense) contradictory. This
+%kind of indeterminacy can be resolved only once the questions of
+%coreference are resolved.
 

### 2015-08-12 writing/SNLI paper/corpus.tex
Clarify coref para

@@
 For instance, the sentence pair \word{A boat sank in the Pacific Ocean} and \word{A boat sank in the Atlantic Ocean} could be labeled as a contradiction if the annotator assumes that the two events under discussion are coreferent, but could also be reasonably labeled as neutral if that assumption is not made.
-For a similar case with potentially different inutitions, \ii{Ruth Bader Ginsburg was appointed to the US Supreme Court} and \ii{I had a sandwich for lunch today} could be labeled as neutral if the annotator allows that they need not be refer to the same event, but is necessarily a contradiction if the annotator takes coreference to be required.
+In order to ensure that our labeling scheme is consistent across annotators, we need to select one of these approaches, but there are pitfalls to both choices.
+If we instruct annotators not to assume that entities are coreferent, than we will only ever see contradictions between sentences that make broad universal assertions, but if we instruct them to always assume coreference, other problems emerge.
+For a similar example with potentially different inutitions, \ii{Ruth Bader Ginsburg was appointed to the US Supreme Court} and \ii{I had a sandwich for lunch today} would be labeled as neutral if the annotator allows that they need not be refer to the same event, but is unintuitively a contradiction if we require that events corefer.
+
 Entity coreference presents a similar kind of uncertainty, as in the pair \word{A tourist visited New York} and \word{A tourist visited the city}.
 Assuming coreference between \ii{New York} and \ii{the city} justifies labeling the pair as an entailment, but without that assumption \ii{the city} could be taken to refer to a specific unknown city, leaving the pair neutral.
+This kind of indeterminacy can be resolved only once the questions of coreference are resolved.
 

### 2015-08-14 writing/SNLI paper/corpus.tex
More work on coref para.

--- 2015-08-12 writing/SNLI paper/corpus.tex
@@
 event and entity coreference lead to insurmountable indeterminacy
-concerning the correct semantic label (\citealt{de2008finding} $\S4.3$; \citealt{marelli2014sick}). For
-instance, the sentence pair \word{A boat sank in the Pacific
-  ocean} and \word{A boat sank in an ocean} could reasonably be labeled in
-numerous ways depending on how the objects and events are grounded. If
-the two boats are assumed to be different, then the sentences are semantically
-independent. If the boats and the oceans are assumed to be the same,
-then the first sentence entails the second (in the relevant
-commonsense way). If the boats are assumed to be the same and 
-\word{an ocean} in the second sentence is interpreted with a specific referent 
-that is not the Pacific, then the two events are (commonsense) contradictory. This
-kind of indeterminacy can be resolved only once the questions of
-coreference are resolved.
+concerning the correct semantic label (\citealt{de2008finding} $\S4.3$; \citealt{marelli2014sick}). 
+For an example of the pitfalls surrounding entity coreference, consider the sentence pair \word{A boat sank in the Pacific Ocean} and \word{A boat sank in the Atlantic Ocean}.
+The pair could be labeled as a contradiction if one assumes that the two sentences refer to the same single event, but could also be reasonably labeled as neutral if that assumption is not made.
+In order to ensure that our labeling scheme assigns a single correct label to every pair, we must select one of these approaches across the board, but both choices present problems.
+If we opt not to assume that events are coreferent, then we will only ever find contradictions between sentences that make broad universal assertions, but if we opt to assume coreference, new counterintuitive predictions emerge.
+For example, \ii{Ruth Bader Ginsburg was appointed to the US Supreme Court} and \ii{I had a sandwich for lunch today} would unintuitively be labeled as a contradiction, rather than neutral, under this assumption.
+
+Entity coreference presents a similar kind of indeterminacy, as in the pair \word{A tourist visited New York} and \word{A tourist visited the city}.
+Assuming coreference between \ii{New York} and \ii{the city} justifies labeling the pair as an entailment, but without that assumption \ii{the city} could be taken to refer to a specific unknown city, leaving the pair neutral.
+This kind of indeterminacy of label can be resolved only once the questions of coreference are resolved.
 

### 2015-08-14 writing/SNLI paper/transfer.tex
Transfer citation.

@@
 This claim has a readily testable consequence: the knowledge that a representation learning system learns when it is trained on SNLI should transfer well to any other task that involves understanding sentence meaning, at least to the extent that the style of language used in that task is broadly similar to the style of SNLI.
-To test this, we perform transfer learning experiments on four other tasks, in which models are initialized using the parameters of an LSTM RNN-based model trained on SNLI (as in \S\ref{sentence-embedding}) and then trained to convergence on the target tasks.
+To test this, we perform transfer learning \cite{pratt1991direct} experiments on four other tasks, in which models are initialized using the parameters of an LSTM RNN-based model trained on SNLI (as in \S\ref{sentence-embedding}) and then trained to convergence on the target tasks.
 

### 2015-08-14 writing/SNLI paper/entailment_eval.tex
Transfer citation.
Minor edits

@@
 The most immediate application for our corpus is in developing models for the task of NLI.
-In particular, since it is dramatically larger than any existing corpus of comparable quality, we expect it to be suitable for training parameter-rich models like neural networks, which have not previously been competetive at this task.
+In particular, since it is dramatically larger than any existing corpus of comparable quality, we expect it to be suitable for training parameter-rich models like neural networks, which have not previously been competitive at this task.
 In this section, we explore the performance of a range of standard and novel models trained on the corpus.
@@
 %  is distributed alongside a number of RTE pipelines.
-We additionally evaluate against a strong but simple classifier-based model with
-  both unlexicalized and lexicalized features.
+We additionally evaluate against a strong but simple classifier-based model 
+  making use of both unlexicalized and lexicalized features.
 
@@
   the SICK test data, and the standard RTE-3 test set \cite{giampiccolo2007third}.
-Each of the models was trained on its own training set.
+Each of the models was re-trained on the training set of each corpus.
 %The results for RTE-3 are taken from \newcite{magnini2014excitement}.

### 2015-08-14 writing/SNLI paper/entailment_eval.tex
Hedging of EOP models

@@
 \label{tab:eopresults}
-2-class test accuracy for two systems included in the
+2-class test accuracy for two simple baseline systems included in the
   Excitement Open Platform.
@@
 This yields a chance baseline accuracy of 66\% on SNLI, and 71\% on SICK.
-We train these models without using any of the included lexical
-  resources, so these results do not represent state-of-the-art
-  RTE systems, but rather give an indication of the relative
-  difficulties of the datasets.
+This is intended to demonstrate the difficulty of the task, rather than necessarily
+  the performance of a state-of-the-art RTE system.
+In particular, we omit the models which rely on richer lexical resources or more
+  sophisticated algorithms, as the training for these algorithms did not finish
+  after multiple days.
+%We train these models without using any of the included lexical
+%  resources, so these results do not represent state-of-the-art
+%  RTE systems, but rather give an indication of the relative
+%  difficulties of the datasets.
 
@@
 We evaluate a simple lexicalized classifier to explore the ability of non-specialized models to exploit these features in lieu of more involved language understanding.
-Our classifier implements 6 feature types:
+Our classifier implements 6 feature types; 3 unlexicalized and 3 lexicalized:
 \begin{enumerate}

### 2015-08-14 writing/SNLI paper/entailment_eval.tex
Hedging tweak; VG citation space; clarification.

@@
 In particular, we omit the models which rely on richer lexical resources or more
-  sophisticated algorithms, as the training for these algorithms did not finish
-  after multiple days.
+  sophisticated algorithms. None of the strongest Excitement Open Platform models
+  are designed to scale to large training corpora without modification, and none of them
+  succeded at training on the full SNLI training set, with no model showing any measurable    
+  progress after several days.
+
 %We train these models without using any of the included lexical
@@
 To the extent that successfully training a neural network model like our LSTM on SNLI forces that model to encode broadly accurate representations of English scene descriptions and to build an entailment classifier over those relations, we should expect it to be readily possible to adapt the trained model for use on other NLI tasks.
-In this section, we evaluate on the SICK entailment task using a simple transfer learning method and achieve competitive results.
+In this section, we evaluate on the SICK entailment task using a simple transfer learning method \cite{pratt1991direct} and achieve competitive results.
 

### 2015-08-14 writing/SNLI paper/entailment_eval.tex
New learning curve.

@@
 
-\Fig{learning_curves_bow.pdf}{0.45}{bowlearncurve}{
-A learning curve for the lexicalized and unlexicalized baseline classifiers and the LSTM,
-plotted on a log scale. The y-axis starts at a random-chance accuracy of 33\%. The minibatch size of 64 that we used to tune the LSTM sets a lower bound on data for that model.}
+\Fig{new_lcc.pdf}{0.83}{bowlearncurve}{
+A learning curve showing how the baseline classifiers and the LSTM perform when trained to convergence on varied amounts of training data.
+The y-axis starts near a random-chance accuracy of 33\%.
+The minibatch size of 64 that we used to tune the LSTM sets a lower bound on data for that model.}
 

### 2015-08-14 writing/SNLI paper/corpus.tex

@@
 
-For the premises, we used captions from the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions (corresponding to about 30k images) collected in an earlier crowdsourced effort.
+For the premises, we used captions from the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions (corresponding to about 30k images) collected in an earlier crowdsourced effort.\footnote{
+We additionally include about 4k sentence pairs from a pilot study in which the premise sentences were instead drawn from the VisualGenome corpus (CITE CITE CITE CITE CITE).
+These examples appear only in the training set, and have pair identifiers prefixed with \texttt{vg} in the corpus.
+ } 
 The captions were not authored by the photographers who took the source images, and they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \word{Our trip to the Olympic Peninsula}).
+In order to ensure that the label for each sentence pair can be recovered solely based on the available text, we did not use the images at all during corpus collection.
 
@@
 
-\begin{figure}
-\center
-\includegraphics[width=3.05in]{length_dist}
-    % From 1.0rc2, though still accurate as of rc3
-% \todo{Replace w/ mean+SD if low on space.}
-\caption{\label{length-dist}The distribution of sentence length.} 
-\end{figure}
-
 \subsection{Data validation}
@@
 discuss in this paper.
+
+\begin{figure}
+\center
+\includegraphics[width=3.05in]{length_dist}
+    % From 1.0rc2, though still accurate as of rc3
+% \todo{Replace w/ mean+SD if low on space.}
+\caption{\label{length-dist}The distribution of sentence length.} 
+\end{figure}
 

### 2015-08-14 writing/SNLI paper/entailment_eval.tex
better EOP model
SICK and RTE results

@@
 \t{Classifier Based}               & \b{72.2} & \b{71.4} & \t{61.5} \\
+\t{Classifier + Lexical Resources} & \b{--}   & \b{78.8} & \t{63.6} \\
 %\midrule
@@
 2-class test accuracy for two simple baseline systems included in the
-  Excitement Open Platform.
+  Excitement Open Platform, as well as SICK and RTE results for
+  a model making use of more sophisticated lexical resources.
 %All RTE results are 2-class accuracy.
@@
 In particular, we omit the models which rely on richer lexical resources or more
-  sophisticated algorithms. None of the strongest Excitement Open Platform models
-  are designed to scale to large training corpora without modification, and none of them
-  succeded at training on the full SNLI training set, with no model showing any measurable    
-  progress after several days.
+  sophisticated algorithms. 
+None of the strongest Excitement Open Platform models
+  are not designed to scale to large training corpora without modification, 
+  and none of them succeeded at training on the full SNLI training set, with no 
+  model showing any measurable progress after several days.
+We include the performance of one such model on the SICK and RTE-3 datasets to
+  gauge its relative performance.
 

### 2015-08-14 writing/SNLI paper/entailment_eval.tex
Fix bolding
Train numbers on SICK

@@
 \midrule
-\t{Edit Distance Based}            & \t{71.9} & \t{65.4} & \b{61.9} \\
-\t{Classifier Based}               & \b{72.2} & \b{71.4} & \t{61.5} \\
-\t{Classifier + Lexical Resources} & \b{--}   & \b{78.8} & \t{63.6} \\
+\t{Edit Distance Based}            & \t{71.9} & \t{65.4} & \t{61.9} \\
+\t{Classifier Based}               & \b{72.2} & \t{71.4} & \t{61.5} \\
+\t{Classifier + Lexical Resources} & \t{--}   & \b{78.8} & \b{63.6} \\
 %\midrule
@@
 \begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceL}c}
+\begin{tabular}{l@{\hskip \colspaceL}c@{\hskip \colspaceS}c@{\hskip \colspaceL}c@{\hskip \colspaceS}c}
 \toprule
-\b{System}	 & \multicolumn{2}{c}{\hspace{-1.2em}\b{SNLI}} & \b{SICK}\\
- & \t{Train} & \t{Test} & \t{Test}\\
+\b{System}	 & \multicolumn{2}{c}{\hspace{-1.2em}\b{SNLI}} & \multicolumn{2}{c}{\b{SICK}}\\
+ & \t{Train} & \t{Test} & \t{Train} & \t{Test}\\
 \midrule
-\t{Lexicalized}            & \t{99.7}  & \b{78.2} & \b{77.3} \\ % & \t{78.7}
-\t{Unigrams Only}          & \t{93.1} & \t{71.6} & \t{74.0} \\ % & \t{72.2}
-\t{Unlexicalized}          & \t{49.4} & \t{50.4} & \t{69.4}\\ % & \t{50.1} % Last value was 50.39; rounded for consistency.
+\t{Lexicalized}            & \t{99.7}  & \b{78.2} & \t{90.4} & \b{77.8} \\ % & \t{78.7}
+\t{Unigrams Only}          & \t{93.1} & \t{71.6}  & \t{88.1} & \t{77.0} \\ % & \t{72.2}
+\t{Unlexicalized}          & \t{49.4} & \t{50.4}  & \t{69.9} & \t{69.6} \\ % & \t{50.1} % Last value was 50.39; rounded for consistency.
 \bottomrule
@@
   (Feature \ref{lst:bigram}), and lacking all lexical
-  features (Features \ref{lst:ngram}--\ref{lst:bigram}). For the model trained on SNLI, we report results both on the test set and the training set to judge overfitting.
+  features (Features \ref{lst:ngram}--\ref{lst:bigram}).
+We report results both on the test set and the training set to judge overfitting.
 }

### 2015-08-14 writing/SNLI paper/conclusion.tex
Some of CM's comments

@@
 a neural network model on our corpus can be used to dramatically 
-improve performance on a standard challenge dataset. These results 
-serve as powerful new evidence that neural representations can, 
-when given a sufficiently extensive and challenging task, 
-be trained to accurately encode rich, multifaceted semantic
-information.
+improve performance on a standard challenge dataset. We believe that
+the model presents a valuable, challenging testbed for the continued 
+development of machine learning to semantic representation.

### 2015-08-14 writing/SNLI paper/corpus.tex
Some of CM's comments

--- 2015-08-14 writing/SNLI paper/corpus.tex
@@
 were constrained to describe that scenario from the same perspective, 
-which helps greatly in controlling event and entity coreference. Second, the prompt
+which helps greatly in controlling event and entity coreference.\footnote{
+For example, with the premise sentence \word{A dog is lying in the grass}, a worker could safely assume that the dog is the most prominent thing in the photo, and very likely the only dog, and  build contradicting sentences that violate these assumptions.
+} 
+Second, the prompt
 gave participants the freedom to produce entirely novel sentences
@@
 
-% \todo{Rewrite and expand (CP?)} --- I incorporated this above.
-
 % Quote from SICK paper \cite{marelli2014sick}:
@@
 % \end{quote}
-
-%\todo{SICK is referenced below, so we should set it up as a point of reference either here or in the intro, including mentioning that it had both entailment and similarity labels.}
 
@@
 We used Amazon Mechanical Turk for data collection.
-In each individual task (each HIT), a worker was presented with premise scene descriptions from a preexisting corpus, and asked to supply hypotheses for each of our three labels---\ii{entailment}, \ii{neutral}, and \ii{contradiction}---forcing the data to be balanced among these classes.
+In each individual task (each HIT), a worker was presented with premise scene descriptions from a pre-existing corpus, and asked to supply hypotheses for each of our three labels---\ii{entailment}, \ii{neutral}, and \ii{contradiction}---forcing the data to be balanced among these classes.
 
@@
 
-For the premises, we used captions from the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions (corresponding to about 30k images) collected in an earlier crowdsourced effort.
+For the premises, we used captions from the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions (corresponding to about 30k images) collected in an earlier crowdsourced effort.\footnote{
+We additionally include about 4k sentence pairs from a pilot study in which the premise sentences were instead drawn from the VisualGenome corpus (under construction).
+These examples appear only in the training set, and have pair identifiers prefixed with \texttt{vg} in the corpus.
+ } % TODO: Insert citation.
 The captions were not authored by the photographers who took the source images, and they tend to contain relatively literal scene descriptions that are suited to our approach, rather than those typically associated with personal photographs (as in their example: \word{Our trip to the Olympic Peninsula}).
+In order to ensure that the label for each sentence pair can be recovered solely based on the available text, we did not use the images at all during corpus collection.
 
@@
   %\renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{l l} 
+  \begin{tabular}{l r} 
     \toprule
@@
 discuss in this paper.
+
+\begin{figure}
+\center
+\includegraphics[width=3.05in]{length_dist}
+    % From 1.0rc2, though still accurate as of rc3
+\caption{\label{length-dist}The distribution of sentence length.} 
+\end{figure}
 

### 2015-08-14 writing/SNLI paper/entailment_eval.tex
Some of CM's comments

@@
   of \contradiction\ and \unknown\ are converted to nonentailment.
-This yields a chance baseline accuracy of 66\% on SNLI, and 71\% on SICK.
+This yields a most-frequent-class baseline accuracy of 66\% on SNLI, and 71\% on SICK.
 This is intended to demonstrate the difficulty of the task, rather than necessarily
   the performance of a state-of-the-art RTE system.
-In particular, we omit the models which rely on richer lexical resources or more
+In particular, we could not evaluate models which rely on richer lexical resources or more
   sophisticated algorithms. 
 None of the strongest Excitement Open Platform models
-  are not designed to scale to large training corpora without modification, 
+  are designed to scale to large training corpora without modification, 
   and none of them succeeded at training on the full SNLI training set, with no 
@@
 %  little lexicalization.
-%\todo{Is the RTE3 model trained on RTE3?}
 
@@
   cross-bigram features.
-The latter  result can be explained if
-  the classifier automatically learns to recognize explicit negations and adjective
-  modification, and a similar result was shown in
+The latter  result suggests that there is value in letting
+  the classifier automatically learn to recognize structures like explicit negations and adjective
+  modification. A similar result was shown in
   \newcite{sidaw12simple} for bigram features in sentiment analysis.
@@
 % Insight 2: do well without alignments
-Another surprising fact is that the classifier performs as well as it
+It is surprising that the classifier performs as well as it
   does without any notion of alignment or tree transformations.
@@
 In addition, we experiment with two simple sequence embedding models: a plain RNN and an LSTM RNN \cite{hochreiter1997long}.
-Both use the simple conventional layout with one layer per input token.
 
@@
 Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it a substantial advantage over the plain RNN, and  resulting in performance that is essentially equivalent to the lexicalized classifier on the test set (LSTM performance near the stopping iteration varies by up to 0.5\% between evaluation steps).
-While the lexicalized model fits the test set perfectly, the gap between train and test set accuracy is relatively small for all three neural network models, suggesting that research into significantly higher capacity versions of these models could be productive.
+While the lexicalized model fits the training set almost perfectly, the gap between train and test set accuracy is relatively small for all three neural network models, suggesting that research into significantly higher capacity versions of these models would be productive.
 
@@
 the models wrongly diagnose as \ii{entailment}, though the sentences
-report two very different stories.
+report two very different stories. We would expect a model with access
+to explicit information about syntactic or semantic structure to perform
+better on cases like these.
+% CP is a fan of ``headline'' -- newspaper theme -- so didn't delete.
 
@@
 \begin{center}
-\begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}c@{\hskip \colspaceL}c}
+\begin{tabular}{l@{\hskip \colspaceL}@{\hskip \colspaceL}r@{\hskip \colspaceL}r}
 \toprule
@@
 % Step 0, no learning intended after transfer
-\t{SICK only}            & \t{100} & \t{71.3} \\
+\t{SICK only}            & \t{100.0} & \t{71.3} \\
 % From: ~/quant/transfer-sick-only-l0.0001-dim50-ed200-td3-pen100-do0.9-0.9-ws1-par0-comp2-cdim1/
@@
 This lowers the starting learning rates, and is intended to ensure that the model does not learn too quickly in its first few epochs after transfer and destroy the knowledge accumulated in the pre-transfer phase of training.
-We tune the regularization parameter separately for each case.
 
 The results are shown in \reftab{tab:transferresults}.
-Training on SICK alone yields poor performance, and the model trained on SNLI fails when tested on SICK data, labeling more \ii{neutral} examples as \ii{contradiction}s than correctly, possibly as a result of subtle differences in the label definitions.
+Training on SICK alone yields poor performance, and the model trained on SNLI fails when tested on SICK data, labeling more \ii{neutral} examples as \ii{contradiction}s than correctly, possibly as a result of subtle differences in how the labeling task was presented.
 Transferring representations, in contrast, yields the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which likely represents an approximate performance ceiling.

### 2015-08-15 writing/SNLI paper/corpus.tex
VG citation.

@@
 For the premises, we used captions from the Flickr30k corpus \cite{hodoshimage}, a collection of approximately 160k captions (corresponding to about 30k images) collected in an earlier crowdsourced effort.\footnote{
-We additionally include about 4k sentence pairs from a pilot study in which the premise sentences were instead drawn from the VisualGenome corpus (under construction).
-These examples appear only in the training set, and have pair identifiers prefixed with \texttt{vg} in the corpus.
+We additionally include about 4k sentence pairs from a pilot study in which the premise sentences were instead drawn from the VisualGenome corpus (under construction; \texttt{\href{https://visualgenome.org/}{visualgenome.org}}).
+These examples appear only in the training set, and have pair identifiers prefixed with \texttt{vg} in our corpus.
  } % TODO: Insert citation.

### 2015-08-15 writing/SNLI paper/corpus.tex
Make some of my changes.

@@
 which helps greatly in controlling event and entity coreference.\footnote{
-For example, with the premise sentence \word{A dog is lying in the grass}, a worker could safely assume that the dog is the most prominent thing in the photo, and very likely the only dog, and  build contradicting sentences that violate these assumptions.
+Issues of coreference are not completely solved, but greatly mitigated.
+For example, with the premise sentence \word{A dog is lying in the grass}, a worker could safely assume that the dog is the most prominent thing in the photo, and very likely the only dog, and  build contradicting sentences assuming reference to the same dog.
 } 
@@
 % From 1.0rc3
-\caption{\label{collection-stats}Key statistics for the raw sentence pairs in SNLI. Since the two halves of each pair were collected separately, we report some statistics for both.} 
+\caption{\label{collection-stats}Key statistics for the raw sentence pairs in SNLI\@. Since the two halves of each pair were collected separately, we report some statistics for both.} 
 \end{table}
@@
 We supplied each pair to four annotators, yielding five labels per pair including the label used by the original author.
-The instructions were similar to the instructions for initial data collection shown in \reffig{instructions-1}, and linked to a similar FAQ.
+The instructions were similar to the instructions for initial data collection shown in \reffig{instructions-1}, and linked to a similar FAQ\@.
 Though we initially used a very restrictive qualification (based on past approval rate) to select workers for the validation task, we nonetheless discovered (and deleted) some instances of random guessing in an early batch of work, and subsequently instituted a fully closed qualification restricted to about 30 trusted workers.
@@
 classes. This disagreement likely reflects not just the limitations of
-large crowdsourcing efforts but also the uncertainty inherent in naturalistic NLI.
+large crowdsourcing efforts but also the uncertainty inherent in naturalistic NLI\@.
 Regardless, the overall rate of agreement is extremely high,

### 2015-08-15 writing/SNLI paper/entailment_eval.tex
Make some of my changes.

@@
 
-The most immediate application for our corpus is in developing models for the task of NLI.
+The most immediate application for our corpus is in developing models for the task of NLI\@.
 In particular, since it is dramatically larger than any existing corpus of comparable quality, we expect it to be suitable for training parameter-rich models like neural networks, which have not previously been competitive at this task.
-In this section, we explore the performance of a range of standard and novel models trained on the corpus.
+In this section, we explore the performance of three classes of models: (i)~models from a well-known NLI system, the Excitement Open Platform; (ii)~variants of a strong but simple feature-based classifier model, which makes use of both unlexicalized and lexicalized features, and (iii)~distributed representation models, including a baseline model and neural network sequence models.
 
-\subsection{Standard entailment models}
+%
+% EOP
+%
+
+\subsection{Excitement Open Platform models}
 
@@
 %  is distributed alongside a number of RTE pipelines.
-We additionally evaluate against a strong but simple classifier-based model 
-  making use of both unlexicalized and lexicalized features.
-
 %
-% EOP
-%
-\paragraph{Excitement Open Platform}
 % what is EOP
-The Excitement Open Platform is a tool to quickly develop NLI systems
+EOP is a tool for quickly developing NLI systems
   while sharing components such as common lexical resources and 
   evaluation sets.
-We evaluate against two algorithms included in the distribution:
+We evaluate on two algorithms included in the distribution:
 %  and also evaluated in \newcite{magnini2014excitement}.
@@
 % the best systems
-We report results in \reftab{tab:eopresults}, covering both our own test set,
+Our initial goal was to better understand the difficulty of the task of classifying SNLI corpus inferences, rather than necessarily the performance of a state-of-the-art RTE system.
+We approached this by running the same system on several data sets:
+our own test set,
 the SICK test data, and the standard RTE-3 test set \cite{giampiccolo2007third}.
-Each of the models was re-trained on the training set of each corpus.
+We report results in \reftab{tab:eopresults}.
+Each of the models was separately trained on the training set of each corpus.
 %The results for RTE-3 are taken from \newcite{magnini2014excitement}.
@@
   of \contradiction\ and \unknown\ are converted to nonentailment.
-This yields a most-frequent-class baseline accuracy of 66\% on SNLI, and 71\% on SICK.
-This is intended to demonstrate the difficulty of the task, rather than necessarily
-  the performance of a state-of-the-art RTE system.
-In particular, we could not evaluate models which rely on richer lexical resources or more
+This yields a most-frequent-class baseline accuracy of 66\% on SNLI, and 71\% on SICK\@.
+We were unable to run EOP models which rely on richer lexical resources or more
   sophisticated algorithms. 
@@
 %
-\paragraph{Lexicalized Classifier}
+\subsection{Lexicalized Classifier}
 Unlike the RTE datasets, SNLI's size supports approaches which make use of rich lexicalized features.

### 2015-08-15 writing/SNLI paper/corpus.tex
Spellcheck

--- 2015-08-15 writing/SNLI paper/corpus.tex
@@
 between subject--verb--object relation triples, and the second release of the 
-Paraphase Database \cite{ganitkevitch2ppdb} includes automatically generated entailment annotations
+Paraphrase Database \cite{ganitkevitch2ppdb} includes automatically generated entailment annotations
 over a large corpus of pairs of words and short phrases.

### 2015-08-15 writing/SNLI paper/transfer.tex
Spellcheck

@@
 We use pretrained parameters for all of the LSTM parameters, pretrained parameters for the $\tanh$ layers in the classifier stack, and pretrained embeddings for all of the words that occurred in both corpora, randomly initializing only the softmax layer and the remaining word embeddings.
-In order to use this starategy in sentence classification tasks, we pass the classifier the embedding of the sentence to be classified in the hypothesis input position, as well as a zero vector in the premise input position.
+In order to use this strategy in sentence classification tasks, we pass the classifier the embedding of the sentence to be classified in the hypothesis input position, as well as a zero vector in the premise input position.
 After transfer, we re-initialize the accumulators in AdaDelta to re-set the learning rates to high values.

### 2015-08-15 writing/SNLI paper/entailment_eval.tex
Rest of Chris edits.

@@
 The results are shown in \reftab{tab:nnresults}.
-The sum of words model performed slightly worse than the fundamentally similar lexicalized classifier---while the sum of words model is able to use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
+The sum of words model performed slightly worse than the fundamentally similar lexicalized classifier---while the sum of words model can use pretrained word embeddings to better handle rare words, it lacks even the rudimentary sensitivity to word order that the lexicalized model's bigram features provide.
 Of the two RNN models, the LSTM's more robust ability to learn long-term dependencies serves it well, giving it a substantial advantage over the plain RNN, and  resulting in performance that is essentially equivalent to the lexicalized classifier on the test set (LSTM performance near the stopping iteration varies by up to 0.5\% between evaluation steps).
@@
 \Reffig{fig:bowlearncurve} shows a learning curve for the LSTM and the lexicalized and unlexicalized feature-based models.
-This learning curve shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance for both.
+It shows that the large size of the corpus is crucial to both the LSTM and the lexicalized model, and suggests that additional data would yield still better performance for both.
 In addition, though the LSTM and the lexicalized model show similar performance when trained on the current full corpus, the somewhat steeper slope for the LSTM hints that its ability to learn arbitrarily structured representations of sentence meaning may give it an advantage over the more constrained lexicalized model on still larger datasets.
@@
 the models wrongly diagnose as \ii{entailment}, though the sentences
-report two very different stories. We would expect a model with access
-to explicit information about syntactic or semantic structure to perform
+report two very different stories. A model with access
+to explicit information about syntactic or semantic structure should perform
 better on cases like these.
@@
 Training on SICK alone yields poor performance, and the model trained on SNLI fails when tested on SICK data, labeling more \ii{neutral} examples as \ii{contradiction}s than correctly, possibly as a result of subtle differences in how the labeling task was presented.
-Transferring representations, in contrast, yields the best performance yet reported on SICK for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which likely represents an approximate performance ceiling.
+In contrast, transferring SNLI representations to SICK yields the best performance yet reported for an unaugmented neural network model, surpasses the available EOP models, and approaches both the overall state of the art at 84.6\% \cite{lai2014illinois} and the 84\% level of interannotator agreement, which likely represents an approximate performance ceiling.
 This suggests that the introduction of a large high-quality corpus makes it  possible to train representation-learning models for sentence meaning that are competitive with the best hand-engineered models on inference tasks.

### 2015-08-18 writing/SNLI paper/entailment_eval.tex
Add in EOP results

--- 2015-08-14 writing/SNLI paper/entailment_eval.tex
@@
 \t{Edit Distance Based}            & \t{71.9} & \t{65.4} & \t{61.9} \\
-\t{Classifier Based}               & \b{72.2} & \t{71.4} & \t{61.5} \\
-\t{Classifier + Lexical Resources} & \t{--}   & \b{78.8} & \b{63.6} \\
+\t{Classifier Based}               & \t{72.2} & \t{71.4} & \t{61.5} \\
+\t{$~~~$ + Lexical Resources} & \b{75.0} & \b{78.8} & \b{63.6} \\
 %\midrule
@@
 %The results for RTE-3 are taken from \newcite{magnini2014excitement}.
-The edit distance and classifier based models are evaluated only on
+These models are evaluated only on
   2-class entailment; all RTE results are likewise reported for 2-class entailment.
@@
 This yields a most-frequent-class baseline accuracy of 66\% on SNLI, and 71\% on SICK.
-This is intended to demonstrate the difficulty of the task, rather than necessarily
+This is intended primarily to demonstrate the difficulty of the task, rather than necessarily
   the performance of a state-of-the-art RTE system.
-In particular, we could not evaluate models which rely on richer lexical resources or more
-  sophisticated algorithms. 
-None of the strongest Excitement Open Platform models
-  are designed to scale to large training corpora without modification, 
-  and none of them succeeded at training on the full SNLI training set, with no 
-  model showing any measurable progress after several days.
-We include the performance of one such model on the SICK and RTE-3 datasets to
-  gauge its relative performance.
+The edit distance algorithm tunes the weight of the three 
+  case-insensitive edit distance operations on the training set, 
+  after removing stop words.
+In addition to the base classifier-based system distributed with the platform, we
+  train a variant which includes information from
+  WordNet \cite{miller1995wordnet} and VerbOcean
+  \cite{chklovski2004verbocean}, and makes use of features
+  based on tree patterns and dependency tree skeletons
+  \cite{wang2007recognizing}.
+
+%In particular, we could not evaluate models which rely on richer lexical resources or more
+%  sophisticated algorithms. 
+%None of the strongest Excitement Open Platform models
+%  are designed to scale to large training corpora without modification, 
+%  and none of them succeeded at training on the full SNLI training set, with no 
+%  model showing any measurable progress after several days.
+%We include the performance of one such model on the SICK and RTE-3 datasets to
+%  gauge its relative performance.
 

### 2015-08-18 writing/SNLI paper/entailment_eval.tex

--- 2015-08-15 writing/SNLI paper/entailment_eval.tex
@@
 This yields a most-frequent-class baseline accuracy of 66\% on SNLI, and 71\% on SICK\@.
-We were unable to run EOP models which rely on richer lexical resources or more
-  sophisticated algorithms. 
-None of the strongest Excitement Open Platform models
-  are designed to scale to large training corpora without modification, 
-  and none of them succeeded at training on the full SNLI training set, with no 
-  model showing any measurable progress after several days.
-We include the performance of one such model on the SICK and RTE-3 datasets to
-  gauge its relative performance.
+This is intended primarily to demonstrate the difficulty of the task, rather than necessarily
+  the performance of a state-of-the-art RTE system.
+The edit distance algorithm tunes the weight of the three 
+  case-insensitive edit distance operations on the training set, 
+  after removing stop words.
+In addition to the base classifier-based system distributed with the platform, we
+  train a variant which includes information from
+  WordNet \cite{miller1995wordnet} and VerbOcean
+  \cite{chklovski2004verbocean}, and makes use of features
+  based on tree patterns and dependency tree skeletons
+  \cite{wang2007recognizing}.
+
+%In particular, we could not evaluate models which rely on richer lexical resources or more
+%  sophisticated algorithms. 
+%None of the strongest Excitement Open Platform models
+%  are designed to scale to large training corpora without modification, 
+%  and none of them succeeded at training on the full SNLI training set, with no 
+%  model showing any measurable progress after several days.
+%We include the performance of one such model on the SICK and RTE-3 datasets to
+%  gauge its relative performance.
 

### 2015-08-18 writing/S'15 LSTM/methods.tex
Prepping LSTM paper.

@@
 Our objective is the standard negative log likelihood classification objective with L2 regularization (tuned on a separate train/test split).
-All models were trained for 100 epochs, after which all had largely converged without significantly declining from their peak performances.\todo{Replot learning curve, resummarize numbers.}
+All models were trained for 100 epochs, after which all had largely converged without significantly declining from their peak performances.

### 2015-08-18 writing/S'15 LSTM/discussion.tex
More results updating, new arxiv package.

@@
 The results are shown in Figure~\ref{prop-results}. 
-The tree models fit the training data well, reaching 99.6, 98.5, and 97.7\% overall accuracy respectively in the $\le$6 setting, with the LSTM underfitting slightly at 94.4\%. 
+The tree models fit the training data well, reaching 98.9, 98.8, and 98.4\% overall accuracy respectively in the $\le$6 setting, with the LSTM underfitting slightly at 94.8\%. 
 In that setting, all models generalized well to structures of familiar length, with the tree models all surpassing 97\% on examples of size 4, and the LSTM reaching 94.8\%.

### 2015-10-05 writing/S'15 LSTM/data.tex
Paper work

@@
     \end{tabular}
-    \caption{Examples of short to moderate length pairs from the artificial data introduced by \protect\citealt{Bowman:Potts:Manning:2014}.
-For brevity, we only show the parentheses that are needed to disambiguate the sentences rather than the full binary bracketings.}\label{tab:plexs}
+    \caption{Examples of short to moderate length pairs from the artificial data introduced in \protect\citealt{Bowman:Potts:Manning:2014}.
+We only show the parentheses that are needed to disambiguate the sentences rather than the full binary bracketings.}\label{tab:plexs}
 \end{table}

### 2015-10-05 writing/S'15 LSTM/nips2015.tex
Paper work

\documentclass{article} % For LaTeX2e
\usepackage{nips15submit_e,times}
\usepackage{hyperref}
\usepackage{url}
%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09

\title{Formatting Instructions for NIPS 2015}

\author{
David S.~Hippocampus\thanks{ Use footnote for providing further information
about author (webpage, alternative address)---\emph{not} for acknowledging
funding agencies.} \\
Department of Computer Science\\
Cranberry-Lemon University\\
Pittsburgh, PA 15213 \\
\texttt{hippo@cs.cranberry-lemon.edu} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\AND
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
(if needed)\\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
The abstract paragraph should be indented 1/2~inch (3~picas) on both left and
right-hand margins. Use 10~point type, with a vertical spacing of 11~points.
The word \textbf{Abstract} must be centered, bold, and in point size 12. Two
line spaces precede the abstract. The abstract must be limited to one
paragraph.
\end{abstract}

\section{Submission of papers to NIPS 2015}

NIPS requires electronic submissions.  The electronic submission site is  
\begin{center}
   \url{http://papers.nips.cc}
\end{center}

Please read carefully the
instructions below, and follow them faithfully.
\subsection{Style}

Papers to be submitted to NIPS 2015 must be prepared according to the
instructions presented here. Papers may be only up to eight pages long,
including figures. Since 2009 an additional ninth page \textit{containing only
cited references} is allowed. Papers that exceed nine pages will not be
reviewed, or in any other way considered for presentation at the conference.
%This is a strict upper bound.

Please note that this year we have introduced automatic line number generation
into the style file (for \LaTeXe and Word versions). This is to help reviewers
refer to specific lines of the paper when they make their comments. Please do
NOT refer to these line numbers in your paper as they will be removed from the
style file for the final version of accepted papers.

The margins in 2015 are the same as since 2007, which allow for $\approx 15\%$
more words in the paper compared to earlier years. We are also again using 
double-blind reviewing. Both of these require the use of new style files.

Authors are required to use the NIPS \LaTeX{} style files obtainable at the
NIPS website as indicated below. Please make sure you use the current files and
not previous versions. Tweaking the style files may be grounds for rejection.

%% \subsection{Double-blind reviewing}

%% This year we are doing double-blind reviewing: the reviewers will not know 
%% who the authors of the paper are. For submission, the NIPS style file will 
%% automatically anonymize the author list at the beginning of the paper.

%% Please write your paper in such a way to preserve anonymity. Refer to
%% previous work by the author(s) in the third person, rather than first
%% person. Do not provide Web links to supporting material at an identifiable
%% web site.

%%\subsection{Electronic submission}
%%
%% \textbf{THE SUBMISSION DEADLINE IS June 5, 2015. SUBMISSIONS MUST BE LOGGED BY
%% 23:00, June 5, 2015, UNIVERSAL TIME}

%% You must enter your submission in the electronic submission form available at
%% the NIPS website listed above. You will be asked to enter paper title, name of
%% all authors, keyword(s), and data about the contact
%% author (name, full address, telephone, fax, and email). You will need to
%% upload an electronic (postscript or pdf) version of your paper.

%% You can upload more than one version of your paper, until the
%% submission deadline. We strongly recommended uploading your paper in
%% advance of the deadline, so you can avoid last-minute server congestion.
%%
%% Note that your submission is only valid if you get an e-mail
%% confirmation from the server. If you do not get such an e-mail, please
%% try uploading again.

\subsection{Retrieval of style files}

The style files for NIPS and other conference information are available on the World Wide Web at
\begin{center}
   \url{http://www.nips.cc/}
\end{center}
The file \verb+nips2015.pdf+ contains these 
instructions and illustrates the
various formatting requirements your NIPS paper must satisfy. \LaTeX{}
users can choose between two style files:
\verb+nips15submit_09.sty+ (to be used with \LaTeX{} version 2.09) and
\verb+nips15submit_e.sty+ (to be used with \LaTeX{}2e). The file
\verb+nips2015.tex+ may be used as a ``shell'' for writing your paper. All you
have to do is replace the author, title, abstract, and text of the paper with
your own. The file
\verb+nips2015.rtf+ is provided as a shell for MS Word users.

The formatting instructions contained in these style files are summarized in
sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

%% \subsection{Keywords for paper submission}
%% Your NIPS paper can be submitted with any of the following keywords (more than one keyword is possible for each paper):

%% \begin{verbatim}
%% Bioinformatics
%% Biological Vision
%% Brain Imaging and Brain Computer Interfacing
%% Clustering
%% Cognitive Science
%% Control and Reinforcement Learning
%% Dimensionality Reduction and Manifolds
%% Feature Selection
%% Gaussian Processes
%% Graphical Models
%% Hardware Technologies
%% Kernels
%% Learning Theory
%% Machine Vision
%% Margins and Boosting
%% Neural Networks
%% Neuroscience
%% Other Algorithms and Architectures
%% Other Applications
%% Semi-supervised Learning
%% Speech and Signal Processing
%% Text and Language Applications

%% \end{verbatim}

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).
Use 10~point type with a vertical spacing of 11~points. Times New Roman is the
preferred typeface throughout. Paragraphs are separated by 1/2~line space,
with no indentation.

Paper title is 17~point, initial caps/lower case, bold, centered between
2~horizontal rules. Top rule is 4~points thick and bottom rule is 1~point
thick. Allow 1/4~inch space above and below title to rules. All pages should
start at 1~inch (6~picas) from the top of the page.

%The version of the paper submitted for review should have ``Anonymous Author(s)'' as the author of the paper.

For the final version, authors' names are
set in boldface, and each name is centered above the corresponding
address. The lead author's name is to be listed first (left-most), and
the co-authors' names (if different address) are set to follow. If
there is only one co-author, list both author and co-author side by side.

Please pay special attention to the instructions in section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

First level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 12. One line space before the first level
heading and 1/2~line space after the first level heading.

\subsection{Headings: second level}

Second level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 10. One line space before the second level
heading and 1/2~line space after the second level heading.

\subsubsection{Headings: third level}

Third level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 10. One line space before the third level
heading and 1/2~line space after the third level heading.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone, regardless of the formatter being used.

\subsection{Citations within the text}

Citations within the text should be numbered consecutively. The corresponding
number is to appear enclosed in square brackets, such as [1] or [2]-[5]. The
corresponding references are to be listed in the same order at the end of the
paper, in the \textbf{References} section. (Note: the standard
\textsc{Bib\TeX} style \texttt{unsrt} produces this.) As to the format of the
references themselves, any style is acceptable as long as it is used
consistently.

As submission is double blind, refer to your own published work in the 
third person. That is, use ``In the previous work of Jones et al.\ [4]'',
not ``In our previous work [4]''. If you cite your other papers that
are not widely available (e.g.\ a journal paper under review), use
anonymous author names in the citation, e.g.\ an author of the
form ``A.\ Anonymous''.

\subsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first footnote} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches
(12~picas).\footnote{Sample of the second footnote}

\subsection{Figures}

All artwork must be neat, clean, and legible. Lines should be dark
enough for purposes of reproduction; art work should not be
hand-drawn. The figure number and caption always appear after the
figure. Place one line space before the figure caption, and one line
space after the figure. The figure caption is lower case (except for
first word and proper nouns); figures are numbered consecutively.

Make sure the figure caption does not get separated from the figure.
Leave sufficient space to avoid splitting the figure and figure caption.

You may use color figures. 
However, it is best for the
figure captions and the paper body to make sense if the paper is printed
either in black/white or in color.
\begin{figure}[h]
\begin{center}
%\framebox[4.0in]{$\;$}
\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Sample figure caption.}
\end{figure}

\subsection{Tables}

All tables must be centered, neat, clean and legible. Do not use hand-drawn
tables. The table number and title always appear before the table. See
Table~\ref{sample-table}.

Place one line space before the table title, one line space after the table
title, and one line space after the table. The table title must be lower case
(except for first word and proper nouns); tables are numbered consecutively.

\begin{table}[t]
\caption{Sample table title}
\label{sample-table}
\begin{center}
\begin{tabular}{ll}
\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
\\ \hline \\
Dendrite         &Input terminal \\
Axon             &Output terminal \\
Soma             &Cell body (contains cell nucleus) \\
\end{tabular}
\end{center}
\end{table}

\section{Final instructions}
Do not change any aspects of the formatting parameters in the style files.
In particular, do not modify the width or length of the rectangle the text
should fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PostScript or PDF files}

Please prepare PostScript or PDF files with paper size ``US Letter'', and
not, for example, ``A4''. The -t
letter option on dvips will produce US Letter files.

Fonts were the main cause of problems in the past years. Your PDF file must
only contain Type 1 or Embedded TrueType fonts. Here are a few instructions
to achieve this.

\begin{itemize}

\item You can check which fonts a PDF files uses.  In Acrobat Reader,
select the menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
available out-of-the-box on most Linux machines.

\item The IEEE has recommendations for generating PDF files whose fonts
are also acceptable for NIPS. Please see
\url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item LaTeX users:

\begin{itemize}

\item Consider directly generating PDF files using \verb+pdflatex+
(especially if you are a MiKTeX user). 
PDF figures must be substituted for EPS figures, however.

\item Otherwise, please generate your PostScript and PDF files with the following commands:
\begin{verbatim} 
dvips mypaper.dvi -t letter -Ppdf -G0 -o mypaper.ps
ps2pdf mypaper.ps mypaper.pdf
\end{verbatim}

Check that the PDF files only contains Type 1 fonts. 
%For the final version, please send us both the Postscript file and
%the PDF file.

\item xfig "patterned" shapes are implemented with 
bitmap fonts.  Use "solid" shapes instead. 
\item The \verb+\bbold+ package almost always uses bitmap
fonts.  You can try the equivalent AMS Fonts with command
\begin{verbatim}
\usepackage[psamsfonts]{amssymb}
\end{verbatim}
 or use the following workaround for reals, natural and complex: 
\begin{verbatim}
\newcommand{\RR}{I\!\!R} %real numbers
\newcommand{\Nat}{I\!\!N} %natural numbers 
\newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}

\item Sometimes the problematic fonts are used in figures
included in LaTeX files. The ghostscript program \verb+eps2eps+ is the simplest
way to clean such figures. For black and white figures, slightly better
results can be achieved with program \verb+potrace+.
\end{itemize}
\item MSWord and Windows users (via PDF file):
\begin{itemize}
\item Install the Microsoft Save as PDF Office 2007 Add-in from
\url{http://www.microsoft.com/downloads/details.aspx?displaylang=en\&familyid=4d951911-3e7e-4ae6-b059-a2e79ed87041}
\item Select ``Save or Publish to PDF'' from the Office or File menu
\end{itemize}
\item MSWord and Mac OS X users (via PDF file):
\begin{itemize}
\item From the print menu, click the PDF drop-down box, and select ``Save
as PDF...''
\end{itemize}
\item MSWord and Windows users (via PS file):
\begin{itemize}
\item To create a new printer
on your computer, install the AdobePS printer driver and the Adobe Distiller PPD file from
\url{http://www.adobe.com/support/downloads/detail.jsp?ftpID=204} {\it Note:} You must reboot your PC after installing the
AdobePS driver for it to take effect.
\item To produce the ps file, select ``Print'' from the MS app, choose
the installed AdobePS printer, click on ``Properties'', click on ``Advanced.''
\item Set ``TrueType Font'' to be ``Download as Softfont''
\item Open the ``PostScript Options'' folder
\item Select ``PostScript Output Option'' to be ``Optimize for Portability''
\item Select ``TrueType Font Download Option'' to be ``Outline''
\item Select ``Send PostScript Error Handler'' to be ``No''
\item Click ``OK'' three times, print your file.
\item Now, use Adobe Acrobat Distiller or ps2pdf to create a PDF file from
the PS file. In Acrobat, check the option ``Embed all fonts'' if
applicable.
\end{itemize}

\end{itemize}
If your file contains Type 3 fonts or non embedded TrueType fonts, we will
ask you to fix it.

\subsection{Margins in LaTeX}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+
from the graphicx package. Always specify the figure width as a multiple of
the line width as in the example below using .eps graphics
\begin{verbatim}
   \usepackage[dvips]{graphicx} ... 
   \includegraphics[width=0.8\linewidth]{myfile.eps} 
\end{verbatim}
or % Apr 2009 addition
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ... 
   \includegraphics[width=0.8\linewidth]{myfile.pdf} 
\end{verbatim}
for .pdf graphics. 
See section 4.4 in the graphics bundle documentation (\url{http://www.ctan.org/tex-archive/macros/latex/required/graphics/grfguide.ps})

A number of width problems arise when LaTeX cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command.

\subsubsection*{Acknowledgments}

Use unnumbered third level headings for the acknowledgments. All
acknowledgments go at the end of the paper. Do not include 
acknowledgments in the anonymized submission, only in the 
final paper.

\subsubsection*{References}

References follow the acknowledgments. Use unnumbered third level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to `small' (9-point) 
when listing the references. {\bf Remember that this year you can use
a ninth page as long as it contains \emph{only} cited references.}

\small{
[1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
and T.K. Leen (eds.), {\it Advances in Neural Information Processing
Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.

[2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
Realistic Neural Models with the GEneral NEural SImulation System.}
New York: TELOS/Springer-Verlag.

[3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
and recall at excitatory recurrent synapses and cholinergic modulation
in rat hippocampal region CA3. {\it Journal of Neuroscience}
{\bf 15}(7):5249-5262.
}

\end{document}

### 2015-10-05 writing/SNLI paper/arxiv_package/abstract.tex
Paper work

--- 2015-05-31 writing/SNLI paper/abstract.tex
@@
 However, machine learning research in this area has been dramatically limited by the lack of large-scale resources.
-To address this, we introduce a new freely available corpus of labeled sentence pairs, written by humans doing a novel grounded task based on image captioning.
+To address this, we introduce the Stanford Natural Language Inference corpus, a new, freely available collection of labeled sentence pairs, written by humans doing a novel grounded task based on image captioning.
 At 570K pairs, it is two orders of magnitude larger than all other resources of its type.
-We find that this increase in scale allows lexicalized classifiers to outperform some sophisticated existing entailment models, and that it allows a neural network-based model to perform competitively on natural language inference benchmarks for the first time.
+This increase in scale allows lexicalized classifiers to outperform some sophisticated existing entailment models, and it allows a neural network-based model to perform competitively on natural language inference benchmarks for the first time.
 \end{abstract}

### 2015-10-05 writing/SNLI paper/arxiv_package/entailment_eval.tex
Paper work

--- 2015-08-18 writing/SNLI paper/entailment_eval.tex
@@
 In particular, since it is dramatically larger than any existing corpus of comparable quality, we expect it to be suitable for training parameter-rich models like neural networks, which have not previously been competitive at this task.
-In this section, we explore the performance of three classes of models: (i)~models from a well-known NLI system, the Excitement Open Platform; (ii)~variants of a strong but simple feature-based classifier model, which makes use of both unlexicalized and lexicalized features, and (iii)~distributed representation models, including a baseline model and neural network sequence models.
+Our ability to evaluate standard classifier-base NLI models, however, was limited to those which were designed to scale to SNLI's size without modification, so a more complete comparison of approaches will have to wait for future work.
+In this section, we explore the performance of three classes of models which could scale readily: (i)~models from a well-known NLI system, the Excitement Open Platform; (ii)~variants of a strong but simple feature-based classifier model, which makes use of both unlexicalized and lexicalized features, and (iii)~distributed representation models, including a baseline model and neural network sequence models.
 
@@
 We evaluate on two algorithms included in the distribution:
-%  and also evaluated in \newcite{magnini2014excitement}.
-%These are: 
   a simple edit-distance based algorithm and
-  a classifier-based algorithm.
+  a classifier-based algorithm, the latter both in a bare form and augmented 
+  with EOP's full suite of lexical resources.
 %  (3) an algorithm based around tree transformations.
@@
 %The results for RTE-3 are taken from \newcite{magnini2014excitement}.
-The edit distance and classifier based models are evaluated only on
-2-class entailment; all RTE results are likewise reported for 2-class entailment.
+All models are evaluated only on 2-class entailment.
 To convert 3-class problems like SICK and SNLI to this setting, all instances

### 2015-10-05 writing/SNLI paper/arxiv_package/intro.tex
Paper work

--- 2015-08-12 writing/SNLI paper/intro.tex
@@
 
-Our ultimate objective in this paper is to provide an empirical
-evaluation of learning-centered approaches to NLI with a focus on
-systems for representation learning,
+Our ultimate objective is to provide an empirical
+evaluation of learning-centered approaches to NLI,
 advancing the case for NLI as a tool for the evaluation of 

### 2015-10-05 writing/SNLI paper/arxiv_package/model_structure_fig.tex
Paper work

--- 2015-05-31 writing/SNLI paper/model_structure_fig.tex
@@
     \node[e]  (he) at (3*\dx,1.85*\dy) {100d hypothesis};
-    \node[m]  (pem) at (-3*\dx,0.5*\dy) {embedding model\\ with premise input};
-    \node[m]  (hem) at (3*\dx,0.5*\dy) {embedding model\\ with hypothesis input};
+    \node[m]  (pem) at (-3*\dx,0.5*\dy) {sentence model\\ with premise input};
+    \node[m]  (hem) at (3*\dx,0.5*\dy) {sentence model\\ with hypothesis input};
 
@@
 
-        \caption{The classification architecture used in all NLI evaluations: for each sentence embedding model evaluated in \reftabs{tab:nnresults}{tab:transferresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 100d inputs shown here.}
+        \caption{The neural network classification architecture: for each sentence embedding model evaluated in \reftabs{tab:nnresults}{tab:transferresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 100d inputs shown here.}
   \label{modelstructure}

### 2015-10-05 writing/SNLI paper/cusp.tex
Paper work

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[12pt]{article} % use larger type; default would be 10pt

\usepackage[round]{natbib}
\bibpunct[,~]{(}{)}{,}{a}{,}{,}

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{letterpaper} % or letterpaper (US) or a5paper or....
\geometry{margin=1in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options
\usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black]{hyperref}

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...

%%% END Article customizations

%%% The "real" document content comes below...

\def\ii#1{\textit{#1}}
\newcommand{\word}[1]{\emph{#1}}
\newcommand{\fulllabel}[2]{\textbf{#1}\newline\textsc{#2}}

\begin{document}
{\centering \textbf{A large annotated corpus of entailments and contradictions}\\
 Samuel R. Bowman, Gabor Angeli, Christopher Potts, and Christopher D. Manning\\
 Stanford University\\~\\
 }

Corpus methods have the potential to offer a range of fast reproducible experimantal designs to researchers in semantics. However, these methods have been largely impractical due to a lack of large corpora with any kind of semantic grounding. While there are serious obstacles to collecting such data, we argue that focusing on inferences of entailment and contradiction as a form of grounding mitigates these, and we present a large new corpus of utterance pairs labeled for the  entailment and contradiction, the Stanford Natural Language Inference corpus \citep[introduced earlier this year in][]{snli:emnlp2015}.

An ideal corpus for model-theoretic semantics would pair utterances in context with either a truth value or, even better, a set of truth conditions expressed in a logical form of some kind. Doing this requires specifying a representational system to describe the contexts of utterances, and in the latter case, the contents of those utterances as well. How best to represent this kind of information, though, is very much an open problem, and in fact it is the very problem that such a corpus would be meant to help solve.

Following work in natural logic \citep{Moss09} and focusing on grounding the meanings of utterances in inferences that they licence, rather than in their truth conditions, makes it possible to collect a representation-agnostic corpus. In this framing, natural language is treated as its own representation language, and the semantic annotation for a sentence consists simply of examples of sentences with respect to which it is entailing, contradictory, or semantically neutral. High quality corpora of this kind exist \citep{dagan2006pascal, marelli2014sick}, but at a few thousand examples they have so far been too small for corpus methods to offer much insight. Our new corpus of 570k examples aims to remedy this.

In our data collection effort, annotators rercuited through Amazon Mechanical Turk were presented with short scene descriptions (either sentences or complex NPs), and asked to redescribe each scene first faithfully (yielding a pair with the label \ii{entailment}), then deliberately unfaithfully (\ii{contradiction}), and then with the introduction of uncertain details (\ii{neutral}). We collected 570k such pairs of utterances, and had additional annotators relabel 57k of these, including those shown in Table~\ref{examples}.  We find the resulting pairs reliably evince valid semantic relationships between sentences with diverse surface forms, and we hope that they can serve as a valuable resource in exploratory and experimental research on meaning. They are available for download at: \href{http://nlp.stanford.edu/projects/snli/}{\texttt{nlp.stanford.edu/projects/snli/}}

\begin{table}[bh!]
  \centering\small
  \begin{tabular}{p{6.0cm}p{2.3cm}p{6.0cm}}
  \toprule
\rule{0pt}{3ex}A black race car starts up in front of a crowd of people. & \fulllabel{contradiction}{c c c c c} & A man is driving down a lonely road.\\
\rule{0pt}{3ex}A soccer game with multiple males playing. & \fulllabel{entailment}{e e e e e} & Some men are playing a sport.\\
\rule{0pt}{3ex}A smiling costumed woman is holding an umbrella. & \fulllabel{neutral}{n n e c n} & A happy woman in a fairy costume holds an umbrella.\\
    \bottomrule
% From 1.0rc3
  \end{tabular}
\caption{Randomly chosen examples from our corpus, shown with both labels from individual annotators and consensus labels.\label{examples}}
\end{table}

\bibliographystyle{plainnat}
\bibliography{MLSemantics}

\end{document}

### 2015-10-05 writing/SNLI paper/test.tex
Paper work

%sample file for QITL 2013 Abstract

\documentclass[11pt,a4paper]{article}
\usepackage{graphicx}
% uncomment according to your operating system:
% ------------------------------------------------
\usepackage[latin1]{inputenc}    %% european characters can be used (Windows, old Linux)
%\usepackage[utf8]{inputenc}     %% european characters can be used (Linux)
%\usepackage[applemac]{inputenc} %% european characters can be used (Mac OS)
% ------------------------------------------------
\usepackage[T1]{fontenc}   %% get hyphenation and accented letters right
\usepackage{mathptmx}      %% use fitting times fonts also in formulas
\usepackage{linguex}       %% for linguistics glosses
\usepackage[round,colon]{natbib} %% for authornames with years for in-text references 
\bibliographystyle{author1} 
% do not change these lines:
\pagestyle{empty}                %% no page numbers!
\usepackage[left=35mm, right=35mm, top=15mm, bottom=20mm, noheadfoot]{geometry}
%% please don't change geometry settings!

% begin the document
\begin{document}
\thispagestyle{empty}

\title{\textbf{{\LaTeX} template for the QITL5 abstract \\ Leuven, Sept. 12-14, 2013}}
\author{Kris Heylen\quad Dirk Speelman\\
University of Leuven\\
\{kris.heylen,dirk.speelman\}@arts.kuleuven.be}

\date{} % <--- leave date empty
\maketitle\thispagestyle{empty} %% <-- you need this for the first page

\section{General information}
The abstract should be written in English (with consistent British or American spelling), be approx. 500 to 1000 words long (not including references, figures or tables), and  be prepared as an A4 sized  (210$\times$297 mm) PDF file. Please use the Times font, 18 pt for the title, 13 pt for the authors' names and affiliations, 14pt for section titles, and 11 pt for the text. The left and right margins should be 3.5 cm, the top margin 3 cm and the bottom margin 2 cm.  We advise you to use the {\LaTeX} and MS Word templates provided on the conference website in order to ensure the correct formatting.

Paragraphs directly following another paragraph have an indent of 0.5cm. Paragraphs following a section title, a figure, a table, or an example should have no indents. Text should be justified throughout and  single spaced.

\section{Tables, figures, examples, formulae and references}

Tables, like table \ref{tab:corpussizes}, should be centered and numbered with captions in bold face below the table. Tables should have outer borders, but the choice of inner bordering is free. Please, use a Times font no smaller than 9 pt.

\begin{table}[htbp]
\centering
\small
\begin{tabular}{|r|cccc|l|}
\hline
    & Usenet     & Popular news & Quality news & Legalese    & Total\\
\hline
BE    & 22 million & 905 million        & 373 million        & 70 million  & 1.4 billion\\
NL    & 26 million & 126 million        & 161 million        & 115 million & 428 million\\
\hline
Total & 48 million & 1 billion          & 499 million        & 185 million & 1.8 billion\\
\hline
\end{tabular}
{\bf \caption{Corpus sizes}
\label{tab:corpussizes}}
\end{table}

\noindent
Examples should be numbered with arabic numerals. There is an extra 8 pt spacing before and after example blocks  A simple example should look like:

\ex.  This is$_{[COPULA]}$ an example in English

\noindent
Examples with glosses should look like this:

\exg. *Taro-ga   Hanako-ni   ringo-o    age-ta\\
  Taro-\textsc{nom} Hanako-\textsc{dat}  apple-\textsc{acc} give-\textsc{pst}\\ 
  `Taro gave an apple to Hanako.'

\noindent
Also possible are repeated or contrasted examples with an additional numbering level, using small Latin letters:

\ex.  
\ag.  Taro-ga   Hanako-ni   ringo-o    age-ta\\
      Taro-\textsc{nom} Hanako-\textsc{dat}  apple-\textsc{acc} 
      give-\textsc{pst}\\
      `Taro gave an apple to Hanako.'
\bg.  **Taro-ga   Hanako-ni   ringo-o    tabe-ta\\
      Taro-\textsc{nom} Hanako-\textsc{dat}  ringo-\textsc{acc} 
      eat-\textsc{pst}\\
      `Taro ate an apple to Hanako.'

\noindent
You can include figures, like figure \ref{fig:varieties}. Figures should be centered with captions below in bold face. There is no fixed size for figures, but they should be readable and stay within the margins.

\begin{figure}[h]
%uncomment next line to include a graphic file
%\centerline{\includegraphics[width=10cm]{fig1.eps}}
%and comment out next line
\centerline{\framebox[6cm]{\rule{0cm}{3.5cm} figure example}}
{\bf \caption{Example figure} 
\label{fig:varieties}}
\end{figure}

\noindent
Formulae, like formula \ref{eq:cosine}, should be centered and numbered on the right.

\begin{equation}
cos(V_1,V_2) = cos(\vec{x}, \vec{y}) = 1 - \frac{\vec{x} \cdot \vec{y}}{|\vec{x}| |\vec{y}|} = \frac{\sum_{i=1}^{n} x_i y_i}{\sum_{i=1}^{n} x_i^2 \sum_{i=1}^{n}y_i^2}
\label{eq:cosine}
\end{equation}

\noindent
In-text references should have author names and years. Author names integrated in the text should have the publication year in round brackets behind the name with optional page numbers separated by a comma. Non-integrated references between brackets should have author names and year, separated by a comma. References should have maximally 2 author names. More than 2 author names are abbreviated with {\em et al.} Multiple references between brackets are separated by a semicolon.

Here are some example uses of references: \citet{HeylenSpeelmanGeeraerts2012} present a visualization technique for semantic vector spaces. \citet[p.~47-100]{geeraerts10theories} gives an historical overview of structuralist lexical theories. Some studies have stressed the importance of concept characteristics \citep{speelman09concept,SpeelmanGrondelaersGeeraerts2010}.

Please, also use the bibliography style for books, journal articles, proceedings and book chapters as exemplified in the references section below.

\begin{thebibliography}{00}
\addcontentsline{toc}{chapter}{References}

\bibitem[\protect\citename{Geeraerts}, 2010]{geeraerts10theories}
Geeraerts, Dirk. 2010. {\em Theories of Lexical Semantics.} Berlin: Mouton De Gruyter.

\bibitem[\protect\citename{Heylen \bgroup et al.\egroup}, 2012]{HeylenSpeelmanGeeraerts2012}
Heylen, Kris, Dirk Speelman and Dirk Geeraerts. 2012. Looking at word meaning. An interactive visualization of Semantic Vector Spaces for Dutch synsets. In: {\em Proceedings of the EACL-2012 joint workshop of LINGVIS \& UNCLH: Visualization of Language Patters and Uncovering Language History from Multilingual Resources}, 16--24.

\bibitem[\protect\citename{Speelman and Geeraerts}, 2008]{speelman09concept}
Speelman, Dirk, and Dirk Geeraerts. 2008. The Role of Concept Characteristics in Lexical Dialectometry. {\em International Journal of Humanities and Arts Computing} 2 (1-2): 221--242.

\bibitem[\protect\citename{Speelman \bgroup et al.\egroup}, 2010]{SpeelmanGrondelaersGeeraerts2010}	
Speelman, Dirk, Stefan Grondelaers,  and Dirk Geeraerts. 2006. A profile-based calculation of region and register variation: the synchronic and diachronic status of the two main national varieties of Dutch. In: Wilson, Archer and Rayson (eds.), {\em Corpus Linguistics around the World}, 195--202. Amsterdam: Rodopi.

\end{thebibliography}

\end{document}

### 2015-10-05 writing/S'15 LSTM/abstract.tex
Revise LSTM v. Tree paper.

@@
 We hypothesize that neural sequence models like LSTMs are in fact able to discover and implicitly use recursive compositional structure, at least for tasks with clear cues to that structure in the data.
-We demonstrate this possibility using an artificial data task for which recursive compositional structure is crucial, and find that the sequence model can learn the underlying patterning.
+We demonstrate this possibility using an artificial data task for which recursive compositional structure is crucial, and find that while tree-based models reliably outperform an LSTM-based sequence model, the sequence model can learn the underlying patterning.
 The sequence model is better in that it learns the value of tree structure from the data in an emergent way, while the tree-structured model is better in being able to learn with greater statistical efficiency due to its informative prior model structure.

### 2015-10-05 writing/S'15 LSTM/data.tex
Revise LSTM v. Tree paper.

@@
     \end{tabular}
-    \caption{Examples of short to moderate length pairs from the artificial data introduced in \protect\citealt{Bowman:Potts:Manning:2014}.
+    \caption{Examples of short to moderate length pairs from the artificial data introduced in \protect\cite{Bowman:Potts:Manning:2014}.
 We only show the parentheses that are needed to disambiguate the sentences rather than the full binary bracketings.}\label{tab:plexs}
@@
 
-\paragraph{The artificial language} The language described in Bowman et al.~(\S4) is designed to highlight the use of recursive structure with minimal additional complexity.
+\paragraph{The artificial language} The language described in \newcite{Bowman:Potts:Manning:2014} (\S4) is designed to highlight the use of recursive structure with minimal additional complexity.
 Its vocabulary consists only of six unanalyzed word types ($p_1, p_2, p_3, p_4, p_5, p_6$), \word{and}, \word{or}, and \word{not}.

### 2015-10-05 writing/S'15 LSTM/discussion.tex
Revise LSTM v. Tree paper.

@@
   \centering
-      \includegraphics[height=1.3in]{lcc.pdf}
+      \includegraphics[height=1.3in]{lc.pdf}
   \caption{Learning curve for the $\le$6 experiment.}
@@
 
-All four models robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time and a neural bag of words model does reasonably on the shortest examples but falls below 60\% by bin~4.
+All four models robustly beat the simple baselines reported in \newcite{Bowman:Potts:Manning:2014}: the most frequent class occurs just over 50\% of the time and a neural bag of words model does reasonably on the shortest examples but falls below 60\% by bin~4.
 
@@
 
-We find that all four models can exploit a recursively defined language to interpret sentences with complex unseen structures, and that tree models' biases allow them to learn to do this more effectively from less data.
+We found that all four models were able to effectively exploit a recursively defined language to interpret sentences with complex unseen structures.
+We find that tree models' biases allow them to do this with greater efficiency, outperforming sequence-based models in every experiment.
+However, our sequence model was nonetheless able to generalize smoothly to from seen sentence structures to unseen ones, showing that its lack of explicit recursive structure did not prevent it from recognizing recursive structure in our artificial language.
 We interpret these results as evidence that both tree and sequence architectures can play valuable roles in the construction of sentence models over data with recursive syntactic structure: tree architectures provide an explicit bias that makes it possible to learn from relatively small and impoverished data sets, while sequence-based architectures can solve the same problems---given sufficient data---without being constrained by this bias.

### 2015-10-05 writing/S'15 LSTM/intro.tex
Revise LSTM v. Tree paper.

@@
 
-As in Bowman et al., we find that standard tree neural networks are able to make the necessary generalizations, with their performance decaying gradually as the structures in the test set grow in size.
+As in \newcite{Bowman:Potts:Manning:2014}, we find that standard tree neural networks are able to make the necessary generalizations, with their performance decaying gradually as the structures in the test set grow in size.
 We additionally find that extending the training set to include larger structures mitigates this decay.
-Then considering sequence models, we find that a single-layer LSTM is also able to generalize to unseen large structures, but that it does this only when trained on a larger and more complex training set than is needed by the tree models.
+Then considering sequence models, we find that a single-layer LSTM is also able to generalize to unseen large structures, but that it does this only when trained on a larger and more complex training set than is needed by the tree models to reach the same generalization performance.
 
-Our results engage with those of \newcite{vinyals2014grammar} and \newcite{dyer2015transition}, who find that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks.
-The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) with high accuracy using only the information present in the vector.
+Our results engage with those of \cite{vinyals2014grammar} and \cite{dyer2015transition}, who find that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks.
+The simplest model presented in \cite{vinyals2014grammar} uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) with high accuracy using only the information present in the vector.
 This shows that the LSTM is able to identify the correct syntactic structures and also hints that it is able to develop a generalizable method for encoding these structures in vectors.

### 2015-10-05 writing/S'15 LSTM/methods.tex
Paper work
Revise LSTM v. Tree paper.

@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
   \centering
@@
   \label{sample-figure}
-\end{figure}
+\end{figure*}
 
@@
   \begin{subfigure}[t]{0.04\textwidth}
-      \includegraphics[height=1.4in]{scale.pdf}
+      \includegraphics[height=1.2in]{scale.pdf}
 \end{subfigure}
 \begin{subfigure}[t]{0.24\textwidth}
-  \includegraphics[height=1.4in]{fig3c.pdf}
+  \includegraphics[height=1.2in]{fig3c.pdf}
   \caption{Training on sz. $\le$3.}
-  \end{subfigure}~~~~
+  \end{subfigure}~~~
 \begin{subfigure}[t]{0.24\textwidth}
-    \includegraphics[height=1.4in]{fig4c.pdf}
+    \includegraphics[height=1.2in]{fig4c.pdf}
   \caption{Training on sz. $\le$4.}
-  \end{subfigure}~~~~
+  \end{subfigure}~~~
 \begin{subfigure}[t]{0.24\textwidth}
-      \includegraphics[height=1.4in]{fig6c.pdf}
+      \includegraphics[height=1.2in]{fig6c.pdf}
   \caption{Training on sz. $\le$6.}
@@
 \begin{subfigure}[t]{0.08\textwidth}
-      \includegraphics[height=1.4in]{leg.pdf}
+      \includegraphics[height=1.2in]{leg.pdf}
 \end{subfigure}
@@
 The classifier component of the model consists of a combining layer which takes the two sentence representations as inputs, followed by two neural network layers, then a softmax classifier.
-For the combining layer, we use a neural tensor network (NTN, \citealt{chen2013learning}) layer, which sums the output of a plain recursive/recurrent neural network layer with a vector computed using two multiplications with a learned (full rank) third-order tensor parameter:
+For the combining layer, we use a neural tensor network (NTN, \cite{chen2013learning}) layer, which sums the output of a plain recursive/recurrent neural network layer with a vector computed using two multiplications with a learned (full rank) third-order tensor parameter:
 \begin{gather} 
@@
 
-Our model is largely identical to the Bowman et al.~model, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence models are compared in as similar a setting as possible.
+Our model is largely identical to the model from \newcite{Bowman:Potts:Manning:2014}, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence models are compared in as similar a setting as possible.
+
+We only study models that encode entire sentences in fixed length vectors, rather than models with attention \cite{bahdanau2014neural}, which gives the downstream model (here, the classifier) the potential to access access each input token individually through a soft content addressing system.
+While attention simplifies the problem of learning complex correspondances between input and output, there is no reason to believe that it should improve or harm a model's ability to track structural information like a given token's position in a tree.
+As such, we expect our results to reflect the same basic behaviors that would be seen in attention-based models.
 

### 2015-10-06 writing/S'15 LSTM/abstract.tex
Reframing the results.

--- 2015-06-16 writing/S'15 LSTM/abstract.tex
@@
 We hypothesize that neural sequence models like LSTMs are in fact able to discover and implicitly use recursive compositional structure, at least for tasks with clear cues to that structure in the data.
-We demonstrate this possibility using an artificial data task for which recursive compositional structure is crucial, and find that the sequence model can learn the underlying patterning.
-The sequence model is better in that it learns the value of tree structure from the data in an emergent way, while the tree-structured model is better in being able to learn with greater statistical efficiency due to its informative prior model structure.
+We demonstrate this possibility using an artificial data task for which recursive compositional structure is crucial, and find an LSTM-based sequence model can learn to exploit the underlying tree structure.
+However, its performance consistently lags behind that of tree-structured models, even on large training sets, suggesting that tree-structured models retain an advantage in tasks for which it is necessary to exploit recursive structure.
 \end{abstract}

### 2015-10-06 writing/S'15 LSTM/intro.tex
Reframing the results.

@@
 We compare standard tree and sequence models on their handling of recursive structure by training the models on sentences whose length and recursion depth are limited, and then testing them on longer and more complex sentences, such that only models that exploit the recursive structure will be able to generalize in a way that yields correct interpretations for these test sentences.
-Our methods are based on those of \newcite{Bowman:Potts:Manning:2014}, who describe an experiment and corresponding artificial dataset which tests this ability in two tree models.
+Our methods are based on those of \cite{Bowman:Potts:Manning:2014}, who describe an experiment and corresponding artificial dataset which tests this ability in two tree models.
 We adapt that experiment to sequence models by decorating the statements with an explicit bracketing, and we use this design to compare an LSTM sequence model with three tree models, with a focus on what data each model needs in order to learn the needed generalizations.
 
-As in \newcite{Bowman:Potts:Manning:2014}, we find that standard tree neural networks are able to make the necessary generalizations, with their performance decaying gradually as the structures in the test set grow in size.
+As in \cite{Bowman:Potts:Manning:2014}, we find that standard tree neural networks are able to make the necessary generalizations, with their performance decaying gradually as the structures in the test set grow in size.
 We additionally find that extending the training set to include larger structures mitigates this decay.
@@
 Our experiments, though, show that LSTMs can learn to understand tree structures when given enough data, suggesting that there is no fundamental obstacle to learning this kind of structured representation.
+We also find, though, that sequence models lag behind tree models across the board, even on training corpora that are quite large relative to the complexity of the underlying grammar, suggesting that tree models can play a valuable role in tasks that require recursive interpretation.

### 2015-10-08 writing/S'15 LSTM/abstract.tex
LSTM paper prose refinement.

--- 2015-06-16 writing/S'15 LSTM/abstract.tex
@@
 We hypothesize that neural sequence models like LSTMs are in fact able to discover and implicitly use recursive compositional structure, at least for tasks with clear cues to that structure in the data.
-We demonstrate this possibility using an artificial data task for which recursive compositional structure is crucial, and find that the sequence model can learn the underlying patterning.
-The sequence model is better in that it learns the value of tree structure from the data in an emergent way, while the tree-structured model is better in being able to learn with greater statistical efficiency due to its informative prior model structure.
+We demonstrate this possibility using an artificial data task for which recursive compositional structure is crucial, and find an LSTM-based sequence model can indeed learn to exploit the underlying tree structure.
+However, its performance consistently lags behind that of tree-structured models, even on large training sets, suggesting that tree-structured models are more effective at exploiting recursive structure.
 \end{abstract}

### 2015-10-08 writing/S'15 LSTM/data.tex
LSTM paper prose refinement.

@@
       %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
- $    p_4 \plor ( \plneg ( ( p_1   $ & $ \natfor $ & $  \plneg ( ( ( ( \plneg p_6 ) \plor ( \plneg p_4 ) )   $\\      
-$ \plor p_6 )  \plor p_4 ) )$ &&$ \pland ( \plneg p_5 ) )\pland ( p_6 \pland p_6 ) ) $\\
+ $    p_4 \plor ( \plneg ( ( p_1  \plor p_6 )  \plor p_4 ) ) $ & $ \natfor $ & $  \plneg ( ( ( ( \plneg p_6 ) \plor ( \plneg p_4 ) )  \pland ( \plneg p_5 ) )\pland ( p_6 \pland p_6 ) ) $\\      
 % $(p_3 \pland \plneg p_1 ) \plor \plneg p_3$    & $\natrev$& $\plneg\, (p_3 \plor p_2)$ \\
@@
 Its vocabulary consists only of six unanalyzed word types ($p_1, p_2, p_3, p_4, p_5, p_6$), \word{and}, \word{or}, and \word{not}.
-Sentences of the language can be straightforwardly interpreted as statements of propositional logic (where the six unanalyzed words are variables), and labeled sentence pairs can be interpreted as theorems of that logic.
+Sentences of the language can be straightforwardly interpreted as statements of propositional logic (where the six unanalyzed words types are variable names), and labeled sentence pairs can be interpreted as theorems of that logic.
 Some example pairs are provided in Table~\ref{tab:plexs}.
 
-Crucially, the language is defined such that any sentence can be embedded under negation or conjunction to create a new sentence, allowing for arbitrary-depth recursion, and the scope of negation and conjunction are determined only by bracketing with parentheses (rather than bare word order).
+Crucially, the language is defined such that any sentence can be embedded under negation or conjunction to create a new sentence, allowing for arbitrary-depth recursion, and such that the scope of negation and conjunction are determined only by bracketing with parentheses (rather than bare word order).
 The compositional structure of each sentence can thus be an arbitrary tree, and interpreting a sentence correctly requires using that structure.
@@
 Our models use this information in two ways.
-For the tree models, the parentheses are not word tokens, but rather used in the expected way to build the tree.
-For the sequence model, the parentheses are words with associated learned embeddings.
-This provides the models with equivalent data, so their ability to handle unseen structures can be fairly compared.
+For the tree models, the parentheses are not word tokens, but rather are used in the expected way to build the tree.
+For the sequence model, the parentheses are word tokens with associated learned embeddings.
+This approach provides the models with equivalent data, so their ability to handle unseen structures can be reasonably compared.
 

### 2015-10-08 writing/S'15 LSTM/discussion.tex
LSTM paper prose refinement.

@@
-\begin{figure}[t]
-  \centering
-      \includegraphics[height=1.3in]{lc.pdf}
-  \caption{Learning curve for the $\le$6 experiment.}
-  \label{fig:lc} 
-\end{figure}
-
 \section{Results and discussion}\label{sec:discussion}
@@
 
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[t]{0.04\textwidth}
+      \includegraphics[height=1.25in]{scale.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+  \includegraphics[height=1.25in]{fig3c.pdf}
+  \caption{Training on sz. $\le$3.}
+  \end{subfigure}~~~
+\begin{subfigure}[t]{0.24\textwidth}
+    \includegraphics[height=1.25in]{fig4c.pdf}
+  \caption{Training on sz. $\le$4.}
+  \end{subfigure}~~~
+\begin{subfigure}[t]{0.24\textwidth}
+      \includegraphics[height=1.25in]{fig6c.pdf}
+  \caption{Training on sz. $\le$6.}
+\end{subfigure}~~
+\begin{subfigure}[t]{0.08\textwidth}
+      \includegraphics[height=1.25in]{leg.pdf}
+\end{subfigure}
+  \caption{Test accuracy on three experiments with increasingly rich training sets.
+The horizontal axis on each graph divides the test set expression pairs into bins by the number of logical operators in the more complex of the two expressions in the pair.
+The dotted line shows the size of the largest examples in the training set in each experiment.}
+  \label{prop-results} 
+\end{figure*}
+
+\begin{figure}[t]
+  \centering
+      \includegraphics[height=1.1in]{lc.pdf}
+  \caption{Learning curve for the $\le$6 experiment.}
+  \label{fig:lc} 
+\end{figure}
+
 \section{Conclusion}
@@
 We found that all four models were able to effectively exploit a recursively defined language to interpret sentences with complex unseen structures.
-We find that tree models' biases allow them to do this with greater efficiency, outperforming sequence-based models in every experiment.
+We find that tree models' biases allow them to do this with greater efficiency, outperforming sequence-based models substantially in every experiment.
 However, our sequence model was nonetheless able to generalize smoothly to from seen sentence structures to unseen ones, showing that its lack of explicit recursive structure did not prevent it from recognizing recursive structure in our artificial language.
-We interpret these results as evidence that both tree and sequence architectures can play valuable roles in the construction of sentence models over data with recursive syntactic structure: tree architectures provide an explicit bias that makes it possible to learn from relatively small and impoverished data sets, while sequence-based architectures can solve the same problems---given sufficient data---without being constrained by this bias.
+
+We interpret these results as evidence that both tree and sequence architectures can play valuable roles in the construction of sentence models over data with recursive syntactic structure.
+Tree architectures provide an explicit bias that makes it possible to efficiently learn to compositional interpretation, which is difficult for sequence models.
+Sequence models, on the other hand, lack this bias, but have other advantages.
+Since they use a consistent graph structure across examples, it is easy to accelerate minibatch training in ways that yield substantially faster training times than are possible with tree models, especially with GPUs.
+In addition, when sequence models integrate each word into a partial sentence representation, they have access to the entire sentence representation up to that point, which may provide valuable cues for the resolution of lexical ambiguity, which is not present in our artificial language, but is a serious concern in natural language text.
+
 Finally, we suggest that, because of the well-supported linguistic claim that the kind of recursive structure that we study here is key to the understanding of real natural languages, there is likely to be value in developing sequence models that can more efficiently exploit this structure without fully sacrificing the flexibility that makes them succeed.

### 2015-10-08 writing/S'15 LSTM/figure1.tex
Paper work
LSTM paper prose refinement.

@@
 \centering
-\scalebox{0.72}{
+\scalebox{0.65}{
  \begin{tikzpicture}
@@
     \tikzstyle{preclass}=[fill=orange!40,text width=50mm,align=center,text height=2mm]
-    \tikzstyle{e}=[fill=green!40,text width=26mm,align=center,text height=2mm]
+    \tikzstyle{e}=[fill=cyan!40,text width=26mm,align=center,text height=2mm]
     \tikzstyle{m}=[draw=black,text width=38mm,align=center,text height=2mm]
@@
     \node[softmax]  (softmax) at (0*\dx,6*\dy) {7-way softmax classifier};
-    \node[preclass]  (pc3) at (0*\dx,5*\dy) {100d $\tanh$ NN layer};
-    \node[preclass]  (pc2) at (0*\dx,4*\dy) {100d $\tanh$ NN layer};
-    \node[preclass,fill=yellow!40]  (pc1) at (0*\dx,3*\dy) {100d $\tanh$ NTN layer};
+    \node[preclass,fill=orange!40]  (pc3) at (0*\dx,5*\dy) {100d $\tanh$ NN layer};
+    \node[preclass,fill=yellow!40]  (pc2) at (0*\dx,4*\dy) {100d $\tanh$ NN layer};
+    \node[preclass,fill=green!40]  (pc1) at (0*\dx,3*\dy) {100d $\tanh$ NTN layer};
     \node[e]  (pe) at (-3*\dx,1.75*\dy) {50d premise};
@@
 
-\end{subfigure}
-
+\end{subfigure}\\
 \begin{subfigure}[t]{0.45\textwidth}
   \centering
-\scalebox{0.72}{
+\scalebox{0.65}{
  \begin{tikzpicture}
@@
 
-  \end{subfigure}
- \vspace{0.5cm}
-
+  \end{subfigure}\qquad
 \begin{subfigure}[t]{0.45\textwidth}
   \centering
-\scalebox{0.72}{
+\scalebox{0.65}{
  \begin{tikzpicture}

### 2015-10-08 writing/S'15 LSTM/intro.tex
LSTM paper prose refinement.

@@
 These models are generally either 
-sequence models based on recurrent neural networks, which build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, or tree models based on \ii{recursive} neural networks, which build representations incrementally
+sequence models based on recurrent neural networks, which build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, or tree-structured models based on \ii{recursive} neural networks, which build representations incrementally
 according to the hierarchical structure of linguistic phrases \cite{goller1996learning,socher2011semi}.
@@
 learning to use the 
-structure that tree-structured models are explicitly designed around. This first requires that
+structure that tree models are explicitly designed around. This first requires that
 sequence models are able to identify syntactic structure in natural language. We believe this is plausible, on the basis of other recent research \cite{vinyals2014grammar,Karpathy2015vaurn}.
@@
 We compare standard tree and sequence models on their handling of recursive structure by training the models on sentences whose length and recursion depth are limited, and then testing them on longer and more complex sentences, such that only models that exploit the recursive structure will be able to generalize in a way that yields correct interpretations for these test sentences.
-Our methods are based on those of \cite{Bowman:Potts:Manning:2014}, who describe an experiment and corresponding artificial dataset which tests this ability in two tree models.
+Our methods extend those of our earlier work in \cite{Bowman:Potts:Manning:2014}, which introduces an experiment and corresponding artificial dataset to test this ability in two tree models.
 We adapt that experiment to sequence models by decorating the statements with an explicit bracketing, and we use this design to compare an LSTM sequence model with three tree models, with a focus on what data each model needs in order to learn the needed generalizations.

### 2015-10-08 writing/S'15 LSTM/methods.tex
LSTM paper prose refinement.

@@
 \section{Testing sentence models on entailment} \label{methods}
+
+We use the architecture depicted in Figure~\ref{fig:model:top}, which builds on the one used in \newcite{Bowman:Potts:Manning:2014}.
+The model architecture uses two copies of a single sentence model (a tree or sequence model) to encode the premise and hypothesis (left and right side) expressions, and then uses those encodings as the features for a multilayer classifier which predicts one of the seven relations.
+Since the encodings are computed separately, the sentence models are forced to encode every element of the sentence meaning that the downstream model will need.
 
@@
 \end{figure*}
-
-\begin{figure*}[t]
-  \centering
-  \begin{subfigure}[t]{0.04\textwidth}
-      \includegraphics[height=1.2in]{scale.pdf}
-\end{subfigure}
-\begin{subfigure}[t]{0.24\textwidth}
-  \includegraphics[height=1.2in]{fig3c.pdf}
-  \caption{Training on sz. $\le$3.}
-  \end{subfigure}~~~
-\begin{subfigure}[t]{0.24\textwidth}
-    \includegraphics[height=1.2in]{fig4c.pdf}
-  \caption{Training on sz. $\le$4.}
-  \end{subfigure}~~~
-\begin{subfigure}[t]{0.24\textwidth}
-      \includegraphics[height=1.2in]{fig6c.pdf}
-  \caption{Training on sz. $\le$6.}
-\end{subfigure}~~
-\begin{subfigure}[t]{0.08\textwidth}
-      \includegraphics[height=1.2in]{leg.pdf}
-\end{subfigure}
-  \caption{Test accuracy on three experiments with increasingly rich training sets.
-The horizontal axis on each graph divides the test set expression pairs into bins by the number of logical operators in the more complex of the two expressions in the pair.
-The dotted line shows the maximum size in each training set.}
-  \label{prop-results} 
-\end{figure*}
-
-We use the architecture depicted in Figure~\ref{fig:model:top}, which builds on the one used in \newcite{Bowman:Potts:Manning:2014}.
-The model architecture uses two copies of a single sentence model (a tree or sequence model) to encode the premise and hypothesis (left and right side) sentences, and then uses those encodings as the features for a multilayer classifier which predicts one of the seven relations.
-Since the encodings are computed separately, the sentence models are forced to encode every element of the sentence meaning that the downstream model will need.
 
@@
 
-We only study models that encode entire sentences in fixed length vectors, rather than models with attention \cite{bahdanau2014neural}, which gives the downstream model (here, the classifier) the potential to access access each input token individually through a soft content addressing system.
-While attention simplifies the problem of learning complex correspondances between input and output, there is no reason to believe that it should improve or harm a model's ability to track structural information like a given token's position in a tree.
+We only study models that encode entire sentences in fixed length vectors, and we set aside models with attention \cite{bahdanau2014neural}, a technique which gives the downstream model (here, the classifier) the potential to access each input token individually through a soft content addressing system.
+While attention simplifies the problem of learning complex correspondences between input and output, there is no apparent reason to believe that it should improve or harm a model's ability to track structural information like a given token's position in a tree.
 As such, we expect our results to reflect the same basic behaviors that would be seen in attention-based models.
@@
 The sentence encoding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector.
-We experiment with tree models (Figure~\ref{fig:model:tree}) with TreeRNN (eqn.~\ref{TreeRNN}), TreeRNTN (eqn.~\ref{TreeRNTN}), and TreeLSTM \cite{tai2015improved} activation functions.
+We experiment with tree-structured models (Figure~\ref{fig:model:tree}) with TreeRNN (eqn.~\ref{TreeRNN}), TreeRNTN (eqn.~\ref{TreeRNTN}), and TreeLSTM \cite{tai2015improved} activation functions.
 In addition, we use a sequence model (Figure~\ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}.
-Experiments with simpler non-LSTM RNN sequence models tended to badly underfit the training data, and are not included here.
+In experiments with a simpler non-LSTM RNN sequence model, the model tended to badly underfit the training data, and those results are not included here.
 

### 2015-11-09 writing/S'15 LSTM/abstract.tex
Long-delayed sync.

@@
 We demonstrate this possibility using an artificial data task for which recursive compositional structure is crucial, and find an LSTM-based sequence model can indeed learn to exploit the underlying tree structure.
-However, its performance consistently lags behind that of tree-structured models, even on large training sets, suggesting that tree-structured models are more effective at exploiting recursive structure.
+However, its performance consistently lags behind that of tree models, even on large training sets, suggesting that tree-structured models are more effective at exploiting recursive structure.
 \end{abstract}

### 2015-11-09 writing/SNLI paper/entailment_eval.tex
Minor tweak to wording
Long-delayed sync.

@@
 We applied L2 regularization to all models, manually tuning the strength coefficient $\lambda$ for each, and additionally applied dropout \cite{srivastava2014dropout} to the inputs and outputs of the sentence embedding models (though not to its internal connections) with a fixed dropout rate.
-All models were implemented in a common framework for this paper, and the implementations will be made available at publication time.
+All models were implemented in a common framework for this paper.
 

### 2015-11-09 run.sh

@@ -8,7 +8,7 @@
 #PBS -q nlp
 
 # Usage example:
-# export MATLABCMD="cd quant; lambda = 0.0001; dim = 25; td = 1; penult = 75; dropout = 1; tot = 0; name='tq'; relu = 1; TrainModel(''\, 1\, @Quantifiers\, name\, dim\, penult\, td\, lambda\, tot\, relu\, dropout\, 32);"; qsub -v MATLABCMD quant/run.sh
+# export REMBED_FLAGS="--learning_rate 0.1"; qsub -v REMBED_FLAGS run.sh
 
 # Change to the submission directory. (Usually ~ for me... I include a 'cd some/dir;' 
 # at the start of MATLABCMD to get to my real working directory.)

### 2015-11-09 writing/S'15 LSTM/data.tex
Camera ready for CoCo.

@@
     \caption{Examples of short to moderate length pairs from the artificial data introduced in \protect\cite{Bowman:Potts:Manning:2014}.
-We only show the parentheses that are needed to disambiguate the sentences rather than the full binary bracketings.}\label{tab:plexs}
+We only show the parentheses that are needed to disambiguate the sentences rather than the full binary bracketings that the models use.}\label{tab:plexs}
 \end{table}
@@
 
-\paragraph{The corpus}
-The sentence pairs in the corpus are divided into thirteen bins according to the number of logical connectives (\word{and, or, not}) in the longer of the two sentences in the pair.
-We test the model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how its performance depends on the complexity of the sentences.
+\paragraph{The data}
+Our sentence pairs are divided into thirteen bins according to the number of logical connectives (\word{and, or, not}) in the longer of the two sentences in each pair.
+We test each model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how each model's performance depends on the complexity of the sentences.
 In three experiments, we train our models on the training portions of bins 0--3 (62k examples), 0--4 (90k), and 0--6 (160k), and test on every bin but the trivial bin 0.
-Capping the size of the training sentences allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures.
-If their performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and L2 regularization (which helps overall performance) exacerbates this by discouraging the model from using the kind of numerically precise, nonlinearity-saturating functions that generalize best.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.
+Capping the size of the training sentences allows us to evaluate how the models interpret the sentences: if a model's performance falls off abruptly above the cutoff, it is reasonable to conclude that it relies heavily on specific sentence structures and cannot generalize to new structures.
+If a model's performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and L2 regularization (which helps overall performance) exacerbates this by discouraging the model from using the kind of numerically precise, nonlinearity-saturating functions that generalize best.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-11-09 writing/S'15 LSTM/discussion.tex
Camera ready for CoCo.

@@
 The tree models fit the training data well, reaching 98.9, 98.8, and 98.4\% overall accuracy respectively in the $\le$6 setting, with the LSTM underfitting slightly at 94.8\%. 
-In that setting, all models generalized well to structures of familiar length, with the tree models all surpassing 97\% on examples of size 4, and the LSTM reaching 94.8\%.
-On the longer test sentences, the tree models decayed smoothly in performance across the board, while the LSTM decayed more quickly and more abruptly, with a striking difference in the $\le$4 setting, where LSTM performance falls 10\% from 4 to 5, compared to 4.4\% for the next worse model.
+In that setting, all models generalized well to structures of familiar length, with the tree models all surpassing 97\% on examples in bin 4, and the LSTM reaching 94.8\%.
+On the longer test sentences, the tree models decay smoothly in performance across the board, while the LSTM decays more quickly and more abruptly, with a striking difference in the $\le$4 setting, where LSTM performance falls 10\% from bin 4 to bin 5, compared to 4.4\% for the next worse model.
 However, the LSTM improves considerably with more ample training data in the $\le$6 condition, showing only a 3\% drop and generalization results better than the best model's in the $\le$3 setting.
@@
 
-We found that all four models were able to effectively exploit a recursively defined language to interpret sentences with complex unseen structures.
+We find that all four models are able to effectively exploit a recursively defined language to interpret sentences with complex unseen structures.
 We find that tree models' biases allow them to do this with greater efficiency, outperforming sequence-based models substantially in every experiment.
-However, our sequence model was nonetheless able to generalize smoothly to from seen sentence structures to unseen ones, showing that its lack of explicit recursive structure did not prevent it from recognizing recursive structure in our artificial language.
+However, our sequence model is nonetheless able to generalize smoothly from seen sentence structures to unseen ones, showing that its lack of explicit recursive structure does not prevent it from recognizing recursive structure in our artificial language.
 

### 2015-11-09 writing/S'15 LSTM/intro.tex
Camera ready for CoCo.

@@
 We propose a possible explanation for these results: standard sequence models can learn to
-exploit recursive syntactic structure in generating sentence meaning representations, thereby 
+exploit recursive syntactic structure in generating representations of sentence meaning, thereby 
 learning to use the 
-structure that tree models are explicitly designed around. This first requires that
-sequence models are able to identify syntactic structure in natural language. We believe this is plausible, on the basis of other recent research \cite{vinyals2014grammar,Karpathy2015vaurn}.
-In this paper, we  evaluate whether LSTM models are able to use that structure to guide interpretation, 
-focusing on cases where the relevant syntactic structure is clearly indicated in the data.
+structure that tree models are explicitly designed around. This requires that
+sequence models be able to identify syntactic structure in natural language. We believe this is plausible on the basis of other recent research \cite{vinyals2014grammar,Karpathy2015vaurn}.
+In this paper, we  evaluate whether LSTM sequence models are able to use such structure to guide interpretation, 
+focusing on cases where syntactic structure is clearly indicated in the data.
 
@@
 Our methods extend those of our earlier work in \cite{Bowman:Potts:Manning:2014}, which introduces an experiment and corresponding artificial dataset to test this ability in two tree models.
-We adapt that experiment to sequence models by decorating the statements with an explicit bracketing, and we use this design to compare an LSTM sequence model with three tree models, with a focus on what data each model needs in order to learn the needed generalizations.
+We adapt that experiment to sequence models by decorating the statements with an explicit bracketing, and we use this design to compare an LSTM sequence model with three tree models, with a focus on what data each model needs in order to generalize well.
 
@@
 This shows that the LSTM is able to identify the correct syntactic structures and also hints that it is able to develop a generalizable method for encoding these structures in vectors.
-However, the massive size of the dataset needed to train the model---250M tokens---leaves open the possibility that it primarily learns to generate only tree structures that it has already seen, representing them as simple hashes---which would not capture unseen tree structures---rather than as structured objects.
+However, the massive size of the dataset needed to train that model, 250M tokens, leaves open the possibility that it primarily learns to generate only tree structures that it has already seen, representing them as simple hashes---which would not capture unseen tree structures---rather than as structured objects.
 Our experiments, though, show that LSTMs can learn to understand tree structures when given enough data, suggesting that there is no fundamental obstacle to learning this kind of structured representation.

### 2015-11-09 writing/S'15 LSTM/methods.tex
Camera ready for CoCo.

@@
 The model architecture uses two copies of a single sentence model (a tree or sequence model) to encode the premise and hypothesis (left and right side) expressions, and then uses those encodings as the features for a multilayer classifier which predicts one of the seven relations.
-Since the encodings are computed separately, the sentence models are forced to encode every element of the sentence meaning that the downstream model will need.
+Since the encodings are computed separately, the sentence models must encode complete representations of the meanings of the two sentences for the downstream model to be able to succeed.
 
@@
   \input{figure1}
-  \caption{In our model, two copies of a sentence model---based on either tree (b) or sequence (c) models---encode the two sentences.
+  \caption{In our model, two copies of a sentence model---based on either tree (b) or sequence (c) models---encode the two input sentences.
 A multilayer classifier component (a) then uses the resulting vectors to predict a label that reflects the logical relationship between the two sentences.}
@@
 
-Our model is largely identical to the model from \newcite{Bowman:Potts:Manning:2014}, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence models are compared in as similar a setting as possible.
+Our model is largely identical to the model from \newcite{Bowman:Potts:Manning:2014}, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all four models, rather than just the TreeRNTN model, so as to ensure that the sentence models are compared in as similar a setting as possible.
 
@@
 \paragraph{Sentence models}
-The sentence encoding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector.
+The sentence encoding component of the model transforms the (learned) embeddings of the input words for each sentence into a single vector representing that sentence.
 We experiment with tree-structured models (Figure~\ref{fig:model:tree}) with TreeRNN (eqn.~\ref{TreeRNN}), TreeRNTN (eqn.~\ref{TreeRNTN}), and TreeLSTM \cite{tai2015improved} activation functions.
